### 2018-03-15 PI_theory_notes.Rmd
Initialise all of the old files from the dungfly repository

---
title: 'Biodiversity and ecosystem function in ephemeral patches: some theory ideas'
author: "Brad Duthie"
bibliography: references.bib
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: evolution.csl
biblio-style: apalike
---

> "If increasing clutch size leads to losses through sib competition, it becomes optimal to compensate by reducing the time spent gathering resources for the clutch of eggs, leading to a smaller clutch (egg size times number) and also by laying larger (and hence even fewer) eggs" [@Parker1986].

[Parker's citations](https://scholar.google.co.uk/scholar?cites=12394960503460892627&as_sdt=2005&sciodt=0,5&hl=en)

Brief introduction to the model and parameters
--------------------------------------------------------------------------------

Define the relationship between total investment budget $M$, parental investment (PI) $m_{i}$ (i.e., the investment that each offspring receives), and total offspring produced $n_{i}$, as follows,

$$n_{i} = \frac{M_{i}}{m_{i}}.$$

Define the effect of competition of the offspring of $j$ on the offspring of $i$ as follows,

$$\alpha_{ij}(m_j, n_j) = \frac{m_{j}n_{j}}{R}.$$

In the above, $R$ defines the total available resources on a patch. The fitness of an individual $i$ ($W_{i}$), defined by number of *viable* offspring produced, will be the product of offspring produced $n_{j}$ and the viability of each offspring, as affected by parental investment $m_{i}$ and effects of competition $\alpha_{ij}(m_j, n_j)$ summed over all competitors $j$<sup><a name="note_1">note_1</a></sup>,

$$W_{i} = \left(1 - e^{-c\left(m_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j})\right)}\right).$$

In the above, $m_{min}$ is a minimum amount of investment required to produce a viable offspring. The parameter $c$ defines the shape of the curve relating parental investment to viability (i.e., how 'diminishing' are the returns in fitness $W_{i}$ as $m_{i}$ increases). From the above equation, it is possible to figure out how much parental investment is optimal ($m^{*}_{i}$) for an individual $i$.

- **[note_1](#note_1):** I [originally had this](https://github.com/bradduthie/dungfly/commit/0b5e10f6fe57911c2d701fc647e26c3baeda55d8) as just $m_{j}R^{-1}$, but this confuses what $N$ really is, as I think it should be the competitors arriving at a patch, not the offspring that they produce; hence, more competitors, more offspring, and this needs to be reflected in the equation. Also note, of course, $m_{j}n_{j} = M_{j}$. @Parker1986 go a different route here, choosing to look at survival given competition as a function of only $n_{j}$ and ignoring effects of increased $m_{j}$ on competition, reasoning that competition should typically "become important only after a considerable amount of growth beyond the egg stage has occurred". I am including both $n_{j}$ and $m_{j}$ to increase the generality, and because I think that in many systems, this assumption might be violated.

Also, it is worth acknowledging that I've [used this mathematical trick before](https://bradduthie.github.io/DuthieEtAl2016c.pdf) in the context of inbreeding [@Duthie2016b], so I'm mostly building off of the hope that a lot of the same logic can apply to decreases in fitness from competitive effects #overlyhonestmethods.

Tedious maths to get optimal parental investment
--------------------------------------------------------------------------------

**This section can be skipped** for anyone who is not interested in how to derive $m^{*}_{i}$ from the equation for $W_{i}$, but I need the equations and code for the concrete example.

```{r, echo = TRUE}
alpha_ij <- function(mi, R) sum(mi/R); # Note: using 'cc' to avoid confusing 'c()'.
W_i      <- function(mi, cc, mmin, mj, R) (1 - exp(-cc*(mi - mmin - alpha_ij(mj,R))));
fWi      <- expression((1-exp(-cc*(mi-mmin-Alpha))),'mi');
dfWi     <- D(fWi, 'mi');
```

For simplicity, we can define $m_{min} = 1$, so unity is the minimum amount of parental investment required to produce a viable offspring in the absence of competition.  For simplicity, we can also set $c = 1$. In the absence of competition (as $R \to \infty$), we have the following slope.

```{r, echo = TRUE}
m_vals <- seq(from = 0, to = 6, length = 1000);
viable <- W_i(c = 1, mi = m_vals, mmin = 1, mj = 1, R = Inf);
par(mar = c(5, 5, 2, 1));
plot(x = m_vals, y = viable, type = "l", lwd = 3, 
     xlab = expression(paste("Parental Investment (",italic(m),")")),
     ylab = expression(paste("Individual fitness (",italic(W[i]),")")),
     cex.lab = 1.5, cex.axis = 1.5, xaxs="i", yaxs="i", ylim = c(0,1));
```

The optimal parental investment $m_{i}^{*}$ is obtained by finding the line tangent to the curve that passes through the origin; the point where the curve and this line meet defines $m_{i}^{*}$ because it maximises the rate at which viable offspring are produced. To get the tangent line, we first need to get the derivative of $W_{i}$ with respect to $m_{i}$,

$$\frac{\partial W_{i}}{\partial {m_{i}}} = ce^{-c\left(m_i - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j}) \right)}.$$

A function for the above is below,

```{r, echo = TRUE}
dW_i <- function(mi, cc, mmin, mj, R) cc * exp(-cc * (mi - mmin - alpha_ij(mj, R)));
```

In general, the equation for any line that is tangent to a function $f$ (with derivative $f'$) at the point $a$ is,

$$y = f'(a)(x - a) + f(a).$$

Because we want to pass through the origin where $x = 0$ and $y = 0$, we use the following to describe the tangent line above,

$$0 = ce^{-c(m_i - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{i}))}(0 - m_{i}) + \left(1 - e^{-c\left(m_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j})\right)}\right).$$

We can just solve this numerically for the parameters set in finding `viable` above. Below is a quick function to get $m_{i}^{*}$.

```{r, echo = TRUE}
get_mi <- function(low_guess, high_guess, cc = 1, mmin = 1, mj = 1, R = Inf){
    fm <- function(mi, cc = cc, mmin = mmin, mj = mj, R = R){
        W_val <- W_i(mi = mi, cc = cc, mmin = mmin, mj = mj, R = R);
        W_der <- dW_i(mi = mi, cc = cc, mmin = mmin, mj = mj, R = R);
        return( W_der*(0-mi) + W_val );
    }
    lg  <- fm(mi = low_guess, cc = cc, mmin = mmin, mj = mj, R = R);
    hg  <- fm(mi = high_guess, cc = cc, mmin = mmin, mj = mj, R = R);
    if(lg > 0){
      u <- low_guess;
      l <- high_guess;
    }else{
      u <- high_guess;
      l <- low_guess;
    }
    fml <- fm(mi = l, cc = cc, mmin = mmin, mj = mj, R = R);
    fmu <- fm(mi = u, cc = cc, mmin = mmin, mj = mj, R = R);
    if((fml > 0 & fmu > 0) | (fml < 0 & fmu < 0)){
      return("Value of m is outside the range");
    }else{
      check  <- 1;
      mguess <- 0.5 * (l+u);
      i      <- 0;
      while(abs(check) > 0.000001 & i < 1000000){
          check <- fm(mi = mguess, cc = cc, mmin = mmin, mj = mj, R = R);
          if(check > 0){
            u      <- mguess;
            mguess <- 0.5*(l + mguess); 
          }else{
            l      <- mguess;
            mguess <- 0.5*(u + mguess);
          }
          i <- i+1;
      }
    return(mguess);
  }
} # Running the below returns the estimate
m_star <- get_mi(low_guess = 0, high_guess = 6, cc = 1, mmin = 1, mj = 1, R = Inf);
```

Running the above code gets an $m_{i}^{*}$ <a name="m_star-no_comp">value</a> of `r m_star`. We can plot the tangent line on the function using the code below.

```{r, echo = TRUE}
m_vals <- seq(from = 0, to = 6, length = 1000);
viable <- W_i(c = 1, mi = m_vals, mmin = 1, mj = 1, R = Inf);
par(mar = c(5, 5, 2, 1));
plot(x = m_vals, y = viable, type = "l", lwd = 3, 
     xlab = expression(paste("Parental Investment (",italic(m),")")),
     ylab = expression(paste("Individual fitness (",italic(W[i]),")")),
     cex.lab = 1.5, cex.axis = 1.5, xaxs="i", yaxs="i", ylim = c(0,1));
slope_mstar <- W_i(mi = m_star, cc = 1, mmin = 1, mj = 1, R = Inf) / m_star;
abline(a = 0, b = slope_mstar, lty = "dotted", lwd = 2);
```

Where am (was) I going with all of this?
--------------------------------------------------------------------------------

I've removed all competition in the tedious maths bits above by setting $R = \infty$ so that resources are limitless and the offspring of parents feed without restriction. Note that there is still an optimum parental investment *above the minimum* because the viability of the offspring is affected by parental investment *independent of competition* -- this could be something like the probability of being able to fight off pathogens or persevere through abiotic stress (i.e., some component of fitness that is independent of density). Now, I want to **see what happens when competition decreases viability**. Given finite resources, assume that parental investment $m_{i}$ influences both competitive ability and viability independent of competition (e.g., something like body size or feeding rate). **It seems intuitive that if competition is high, then it might increase fitness to increase investment to produce fewer competitive offspring**. Note that this also assumes a situation in which an individual moving to a different patch to oviposit is either not possible, or is costly (i.e., at least more costly than it would be to stay and invest optimally on the current patch).

The [corrected version](https://github.com/bradduthie/dungfly/commit/e81c673a78e60e3c95038976129d190cfd5956fd
) of $\alpha_{ij}(m_j, n_j)$ to include $n_{j}$ should make things easier by recognising that *total* competition should depend on the total investment budgets of the individuals arriving to a patch to oviposit. In other words, total competition does not depend on the parental investment decisions of other individuals -- at least not yet. This type of recursion (in which, e.g., competition changed with *realised* viability given different parental investment) is just going to be too complicated of a starting point, so it is better to first consider **how optimal parental investment, and therefore offspring number, is modified by existing competitors on a patch**, without thinking about feedbacks that follow. In a more formal analysis, it might be interesting to consider the PI decisions for an initial coloniser of a patch, followed by the decision of the second arriver, and a third, and so forth. This would link with the established literature optimal foraging and oviposition per patch somehow, which has been studied in ovipositing wasps, though not (to my knowledge), in the context of parental investment where individuals might adjust allocation of resources to eggs or modify behaviour (e.g., guarding).

A simple, concrete example to get started
--------------------------------------------------------------------------------

I will begin with an example of a single ovipositing female on a patch of resources where $R = 100$. The female has a total investment budget of $M_{i} = 10$. Conveniently, the way $\alpha_{ij}(m_j, n_j) = m_{j}n_{j}{R^{-1}}$ is defined allows to just substitute $M_{j} = m_{j}n_{j} = 10$, so $\alpha_{ij}(m_j, n_j) = M_{j}R^{-1} = 10R^{-1} = 0.1$. We can start working from the general equation,

$$W_{i} = \left(1 - e^{-c\left(m_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j})\right)}\right).$$

And we can substitute in the relevant values for competition; assume $c = 1$ and $m_{min} = 1$ for simplicity,

$$W_{i} = \left(1 - e^{-\left(m_{i} - 1 - 0.1\right)}\right).$$

If a focal female's fitness $W_{i}$ is **not** decreased by competition among her offspring, then we had a optimal PI $m^{*}_{i}$ of `r m_star`, [as above](#m_star-no_comp). Now that we have added some competition among her offspring, optimal PI per offsrping is calculated with the code below.

```{r, echo = TRUE}
m_star_1 <- get_mi(low_guess = 0, high_guess = 6, cc = 1, mmin = 1, mj = 10, R = 100);
```

The value of `m_star_1`, and optimal PI $m^{*}_{i}$ given competition amongst offspring is therefore `r m_star_1`, a value that is **higher than optimal PI in the absence of competition**.

Okay, now let's consider what happens when the next competitor arrives *immediately after the first* to oviposit (i.e., ignore any priority effects for now). Assume that she notices that the first competitor is there and can adjust her PI accordingly. If her own total investment budget is also 10, then $\sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j}) = (10 + 10)/100 = 0.2$. We can do the same exercise as before to figure out *her* optimal parental investment.

```{r, echo = TRUE}
m_star_2 <- get_mi(low_guess = 0, high_guess = 6, cc = 1, mmin = 1, mj = c(10, 10), R = 100);
```

The value of `m_star_2` given this increased competition is now `r m_star_2`. Continuing with this logic, we can see how optimal parental investment changes with up to 100 arriving competitors.

```{r, echo = TRUE}
m_stars  <- NULL;
tot_comp <- 1:100;
for(competitors in tot_comp){
    mjs     <- rep(x = 10, times = competitors)
    m_star  <- get_mi(low_guess = 0, high_guess = 20, cc = 1, 
                      mmin = 1, mj = m_stars, R = 100);
    m_stars <- c(m_stars, m_star);
}
par(mar = c(5, 5, 2, 1));
plot(x = tot_comp, y = m_stars, type = "l", pch = 20, lwd = 3, 
     xlab = "Number of competitors on a patch",
     ylab = expression(paste("Optimal parental investment (","m*"[i],")")),
     cex.lab = 1.5, cex.axis = 1.5);
```

This seems like a substantial change in optimal parental investment when competition affects offspring survival, but can be mitigated with increasing PI. Note that total offspring produced per competitor also changes in accordance with $n_{i} = M_{i}m_{i}^{-1}$.

```{r, echo = TRUE}
n_prod <- 10 / m_stars;
par(mar = c(5, 5, 2, 1));
plot(x = tot_comp, y = n_prod, type = "l", pch = 20, lwd = 3, 
     xlab = "Number of competitors on a patch",
     ylab = expression(paste("Offspring per competitor (","n"[i],")")),
     cex.lab = 1.5, cex.axis = 1.5);
```

Given the assumptions underlying this model, it does not seem unreasonable to expect fairly substantial decreases in how an individual arriving to a patch chooses to allocate from their total investment budget $M_{i}$.

Key thoughts before looking at eco-evo feebacks
--------------------------------------------------------------------------------

**It is important to recognise that decreased offspring production as a consequence of shifting parental investment *is* competition in a community ecology sense.** If, for example, a focal individual ultimately *increases* their fitness by producing fewer viable offspring, competition has still occurred. Further, if a focal individual decreases the number of offspring that they produce due to increasd density of other individuals, then competition has occurred **even if those other competitors fail to produce offspring** (e.g., due to sterility). This is because competition is simply defined as an interaction in which the density of a species $i$ decreases the growth rate of a species $j$ (including where $i = j$). Using absolute competition coefficients $\alpha_{ij}$ describing the total effect of species $j$ on species $i$, this is described by the following equation [@Chesson2000b],

$$\frac{dN_{i}}{dt} = r_{i}N_{i}\left(1 - \sum \alpha_{ij}N_{j}\right).$$

In the above, $N$ defines species densities and $r$ is the intrinsic rate of growth in the absence of competition. Competition is merely defined by the presence of a species $j$ ($N_{j}$) on the growth rate of species $i$ (and vice versa); no particular mechanism is assumed. If more $j$ causes focal individuals to shift towards producing fewer $i$, then competition has occurred, though this is counterintuitive. The term **"intimidation competition"** seems appropriate, and apparently [has not been coined in the literature](https://scholar.google.co.uk/scholar?hl=en&as_sdt=0%2C5&q=ecology+%22intimidation+competition%22&btnG=). I'll need to dig a bit more to see if the general concept has been pointed out already.

Different competition functions
--------------------------------------------------------------------------------

I have only looked at a simple competition function,

$$\alpha_{ij}(m_j, n_j) = \frac{m_{j}n_{j}}{R}.$$

But this might be an unrealistic assumption. For example, allocating twice the parental investment to half the number of offspring will result in an identical $\alpha_{ij}(m_j, n_j)$, but it might be that competition is not linearly related to the product of $m_j$ and $n_j$. Rather, twice as much PI in half as many offspring might increase (or decrease) the total value of $\alpha_{ij}(m_j, n_j)$. We can modify the effect of each $m_j$ and $n_j$ on competition using the exponents $a$ and $b$,

$$\alpha_{ij}(m_j, n_j) = \frac{m_{j}^{a}n_{j}^{b}}{R}.$$

If $a = 1$ and $b = 1$, then $\alpha_{ij}(m_j, n_j)$ remains constant as $m_{j}^{a}$ varies (because $M_{j} = m_{j}n_{j}$; see solid line below). If, e.g., $a = 2$ and $b = 1/2$, then $\alpha_{ij}(m_j, n_j)$ increases with increasing parental investment $m_{j}$ of competitors (see dashed line below). If, e.g., $a = 1/2$ and $b = 2$, then $\alpha_{ij}(m_j, n_j)$ decreases with increasing parental investment $m_{j}$ of competitors (see dotted line below).

```{r, echo = TRUE}
M_j     <- 10;
aij_exp <- function(mj, nj, a, b, R = 100) ((mj^a) * (nj^b)) / R;
mj_vals <- seq(from = 0, to = 6, length = 1000);
comp_1  <- aij_exp(mj = mj_vals, nj = M_j/mj_vals, a = 1,   b = 1,   R = 100);
comp_2  <- aij_exp(mj = mj_vals, nj = M_j/mj_vals, a = 2,   b = 0.5, R = 100);
comp_3  <- aij_exp(mj = mj_vals, nj = M_j/mj_vals, a = 0.5, b = 2,   R = 100);
par(mar = c(5, 5, 2, 1));
plot(x = mj_vals, y = comp_1, type = "l", lwd = 2, ylim = c(0, 1),
     xlab = "Parental investment of competitors", ylab = "Total competition on offspring",
     cex.axis = 1.5, cex.lab = 1.5);
points(x = mj_vals, y = comp_2, type = "l", lty = "dashed", lwd = 2);
points(x = mj_vals, y = comp_3, type = "l", lty = "dotted", lwd = 2);
```

While the general result that optimal parental investment increases with increasing competition should not change, just how much competition is predicted to change as a result of changes in parental investment might therefore depend on how competition is affected differently by changes to $m_{j}$ versus $n_{j}$. **This might have interesting consequences for community ecology** and ecological and evolutionary feedbacks more generally, in that changes in PI could increase or decrease population growth, which could in turn affect changes in PI through changing species densities.

Eco-evolutionary approach to species coexistence
--------------------------------------------------------------------------------

To advance theory further, I would like to determine what effect the modulation of parental investment (and therefore offspring number) has on population dynamics and therefore species coexistence. I will use the recently developed approach of @Patel2018 to investigate how feedbacks affect community stability. We can consider a simple eco-evo case of two species, $i$, and $j$ (note that I'm indexing *species* now, not individuals as before), in which species densities $N_{i}$ and $N_{j}$ change and species $i$ parental investment trait $m_{i}$ evolves. **Note that** in contrast to previous models such as @Parker1986, it is not assumed tha PI is adjusted with phenotypic plasticity, but evolves as a quantitative trait. The total effect of ecology and evolution is defined by the Jacobian matrix $\boldsymbol{J}$,

$$\boldsymbol{J} = \begin{bmatrix}
    \boldsymbol{A}  & \boldsymbol{B}  \\
    \boldsymbol{C}  & \boldsymbol{D}  \\  \end{bmatrix}
    = \begin{bmatrix}
    \frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \frac{\partial N_{i}W_{i}}{\partial N_{j}} & \frac{\partial N_{i}W_{i}}{\partial m_{i}} & \frac{\partial N_{i}W_{i}}{\partial m_{j}}  \\
    \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \frac{\partial N_{j}W_{j}}{\partial N_{j}} & \frac{\partial N_{j}W_{j}}{\partial m_{i}} & \frac{\partial N_{j}W_{j}}{\partial m_{j}} \\
    \varepsilon\frac{\partial m_{i}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{i}}{\partial N_{j}} & \varepsilon\frac{\partial m_{i}}{\partial m_{i}} & \varepsilon\frac{\partial m_{i}}{\partial m_{j}}  \\
    \varepsilon\frac{\partial m_{j}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{j}}{\partial N_{j}} & \varepsilon\frac{\partial m_{j}}{\partial m_{i}} & \varepsilon\frac{\partial m_{j}}{\partial m_{j}}  \\
\end{bmatrix}.$$

As in @Patel2018, we can consider the sign of $\boldsymbol{J}$ elements to infer how evolution and ecology interact to affect overall stability. Note also that the value $\varepsilon$ defines the rate of evolution relative to ecology [@Patel2018]. Also note that I am including both traits $m$ and $n$ into the matrix explicitly rather than simply using $m$; this is important more for maths reasons than anything else (we need to be able to invert a matrix later, and having two negatively correlated traits allows us to do this). I am going to start by assuming a simple competition function and showing why that it cannot affect eco-evolutionary dynamics,

$$\alpha_{ij}(m_j, n_j) = \frac{m_{j}n_{j}}{R}.$$

The above function of competition means that there should be **no direct trait change effects on ecology** (e.g., $\partial N_{i}W_{i} / \partial m_{i}$ because the total growth rate *in terms of expected viable offspring* $N_{i}W_{i}$ will not change with $m_j$ because it will be exactly balanced by a change in $n_j$ to define competitive effects. If this is true, then there will be no ultimate effect of competition on the stability of the system overall. Conceptually, this is because evolutionary traits do not feed back at all to affect the ecology of the system -- i.e., everything is driven by the stability of the ecology. Mathematically, this is due to multiplying all values of the resulting $\boldsymbol{B}$ (which is either a matrix or vector of zeroes) to obtain the stability criteria with feedbacks. Hence, eco-evolutionary feedbacks should *only* be important for the more general case where tradeoff in parental investment and offspring number does not produce equivalent results for competition,

$$\alpha_{ij}(m_j, n_j) = \frac{m_{j}^{a}n_{j}^{b}}{R}.$$

For evolution to *have a direct effect on ecology*, at least one of the exponents $a$ or $b$ needs to be inequal to 1. Below I consider four separate cases that include relatively slow and fast evolution of traits given different effects of $m$ and $n$ on competition.

General derivation of PI affected by competition
--------------------------------------------------------------------------------

**This section can be skipped** for anyone who does not need convincing that optimal parental investment $m^{*}_{i}$ will increase with an increase in competition more generally. To demonstrate this is the case, we need to show that $\partial m^{*}_{i} / \partial{\alpha_{ij}(m_j, n_j)}$ for a wide range of conditions. Assume that $\alpha_{ij}(m_j, n_j)$ is positive. We can then start with the general equation for which $m_{i} = m^{*}_{i}$,

$$0 = ce^{-c(m_i - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{i}))}(0 - m_{i}) + \left(1 - e^{-c\left(m_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j})\right)}\right).$$

Substituting $m^{*}_{i}$ for $m_{i}$, we have the following,

$$0 = -m^{*}_{i}ce^{-c(m^{*}_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m^{*}_{i}))} + 1 - e^{-c\left(m^{*}_{i} - m_{min} - \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j})\right)}.$$

```{r, echo = FALSE, eval = FALSE}
library("Ryacas"); # The Ryacas library can simplify
yacas("Solve(0 == -mi * cc * Exp(-cc*(mi - mmin - aij)) + 1 - Exp(-cc*(mi - mmin - aij)), aij)");
```

Just to keep the notation simpler for a moment, allow,

$$\alpha_{ij}(m_{j}, n_{j}) \equiv \sum_{j = 1}^{N}\alpha_{ij}(m_{j}, n_{j}).$$

We can isolate $\alpha_{ij}(m_{j}, n_{j})$ algebraically from the above (the package [Ryacas](https://cran.r-project.org/package=Ryacas) is useful here).

$$\alpha_{ij}(m_{j}, n_{j}) = m^{*}_{i} - m_{min} + \frac{1}{c}\log\left(\frac{1}{m^{*}_{i}c+1}\right).$$

From here, we can differentiate $\alpha_{ij}(m_{i})$ with respect to $m_{i}^{*}$.

```{r, echo = TRUE}
fWi_opt     <- expression((mi - mmin + (1/cc)*log(1/(mi * cc + 1))), 'mi');
dfWi_opt    <- D(fWi_opt, 'mi');
```

The result is as follows,

$$\frac{\partial \alpha_{ij}(m_{j}, n_{j})}{\partial m_{i}^{*}} = 1 - \frac{c(m_{i}^{*}c + 1)}{c(m_{i}^{*}c + 1)^2}.$$

This actualy simplifies nicely to the following (double-checked with R and [mathomatic](http://mathomatic.orgserve.de/math/)),

$$\frac{\partial \alpha_{ij}(m_{j}, n_{j})}{\partial m_{i}^{*}} = \frac{m_{i}^{*}c}{m_{i}^{*}c + 1}.$$

And via the [chain rule](https://en.wikipedia.org/wiki/Chain_rule) of calculus, we can do the following,

$$\frac{\partial m_{i}^{*}}{\partial \alpha_{ij}(m_{j}, n_{j})} = \frac{m_{i}^{*}c + 1}{m_{i}^{*}c}.$$

Overall, for any $m_{i}^{*} > 0$ and $c > 0$ (parental investment increases offspring viability), optimal parental investment will increase with increasing competition. We can therefore conclude that this relationship is general in that the specific form of the competition function does not matter as long as competition decreases offspring viability, nor does the value of $m_min$. **It is potentially interesting to note** that if $\alpha_{ij}(m_{j}, n_{j}) < 0$ (mutualism or facilitation), then optimal parental investment should actually decrease with increasing density of other individuals, ultimately producing more offspring.

It is worth noting that the above corresponds to equation 4 in @Parker1986.

Empirical predictions
--------------------------------------------------------------------------------

One potentially interesting way to test this could be by looking for the *intimidation competition* that I mentioned earlier. As an example, we can imagine a focal female arriving at a patch to oviposit (though any individual that reproduces in an ephemeral patch should work -- potentially including plants in temporary ponds). She has the ability to somehow modulate how many offspring she produces and how much investment goes into each offspring (note again for emphasis -- 'parental investment' is not synonymous with 'parental care', and investment might take the form of, e.g., increased egg provisioning or egg guarding). If she observes the density of competitors to be low, then she should make relatively low PI and produce more offspring. If she observes the density of competiters to instead by high, her PI should be elevated and conseqent offspring reduction reduced. It might be possible to test for this by comparing a focal female's offspring production across a density gradient of competitors who are fertile versus sterile. Consider, e.g., the simple case:

1. A focal female ovipositing on a patch at low density of competitors.
2. A focal female ovipositing on a patch at high density of competitors.
3. A focal female ovipositing on a patch at high density of **sterile** competitors.
4. A focal female ovipositing on a patch at high density of competitors *that cannot physically interfere.*

If females cannot or do not modulate PI with competition, then we would predict no significant difference between a focal female's offspring production in 1 and 3. But if females can modulate PI with competition, then we would predict offspring produced in 3 to be *lower* than offspring produced in 1 due to the effect of intimidation from competitors, but perhaps for these offspring to have traits more associated with higher viability (e.g., larger body size, higher feeding rate). We should also predict offspring production to be much lower in 2 than either 1 or 3 due to effects of competition among larvae. The addition of 4 might be needed to control for effects of competitors physically interfering with the focal female (even if they are sterile). The idea might be to have a mesh separating the focal female physically but still give the illusion of high density in the surrounding area.

Four cases of evolution and competition (ignore -- wrong approach)
--------------------------------------------------------------------------------

**EVERYTHING BELOW IS BAD -- it's an original approach before trying the random matrix stuff**.

**I. Slow evolution. Competition increases with increased offspring production**

Allow us to start by considering the very simple case where $a = 0$ and $b = 1$, which effectively makes competition only a function of offspring production (because $m_{j}^{0} = 1$), as assumed in @Parker1986. This implies that $\partial NW / \partial m > 0$ for all combinations of $i$ and $j$; that is, there will be a *direct positive effect of evolution on ecology* because the growth rate of both species will increase with increasing $m_{i}$ and $m_{j}$ due to decreased competition. Note, we cannot model decrease in $n_i$ and $n_j$ (rather, we don't need to) because of complete degeneracy for perfectly correlated traits [@Patel2018] -- i.e., it's effectively the same trait.

Additionally, there will be a *direct positive effect of ecology on PI evolution* because higher densities cause increased parental investment and therefore higher values of $m_{i}$ and $m_{j}$. Finally, we need to assume that the trace of $\boldsymbol{D}$ is negative to make the system evolutionarily stable. I'll need to think about how this is a reasonable or unreasonable assumption.

$$\boldsymbol{J} = \begin{bmatrix}
    -  & - & + & + \\
    -  & - & + & + \\
    \varepsilon(+)  & \varepsilon(+) & - & 0 \\
    \varepsilon(+)  & \varepsilon(+) & 0 & - \\
\end{bmatrix}.$$

I believe that the above is correct for this simplified case. Small positive values ($\varepsilon$ is small) in the lower left subset of the matrix imply that an increase in density (of either species) increases parental investment (of either species) and decreases offspring number, as demonstrated above.

To start, I am going to **set intraspecific competition as slightly higher than interspecific competition** and see how the feedback of density on traits affects coexistence through assessing the stability criteria of @Patel2018. Under these conditions, ecological dynamics should be stable, leading to extinction of species because. This can be modelled with the submatrix $\boldsymbol{A}$,

$$\boldsymbol{A} = \begin{bmatrix}
    -1      & -0.9  \\
    -0.9    & -1    \\
\end{bmatrix}.$$

```{r, echo = TRUE}
A <- matrix(data = c(-1, -0.9, -0.9, -1), nrow = 2, ncol = 2);
```

Stable coexistence of species occurs when the real component of the leading eigenvalue of the Jacobian matrix is negative. In the case of $\boldsymbol{A}$ above, the leading eigenvalue is negative, which implies an stable system.

 ```{r, echo = TRUE}
s_A <- eigen(A)$values; print(s_A);
```

Now the point is to look at the stability criteria of the feedback between species density and trait change and see if it drives the system to be unstable. The stability criteria given slowly-evolving traits requires that the ecology be stable, and for  [@Patel2018],

$$s\left(\boldsymbol{D} + \boldsymbol{C}\boldsymbol{A}^{-1}(-\boldsymbol{B})\right) < 0.$$

Where $\boldsymbol{A}$, $\boldsymbol{B}$, $\boldsymbol{C}$, and $\boldsymbol{D}$, are defined within $\boldsymbol{J}$ above. Verbally, we need to do element-wise addition of $\boldsymbol{D}$ to the matrix product of $\boldsymbol{C}$ and inverse of $\boldsymbol{A}$, then multiply by $\boldsymbol{B}$ where all elements are multiplied by $-1$. If the resulting matrix has a real component of the leading eigenvalue that is less than zero, then the feedbacks between ecology and evolution will be stabilising. Assume $\varepsilon = 0.001$, so,

$$\boldsymbol{B} = \begin{bmatrix}
    1       & 1   \\
    1       & 1  \\
\end{bmatrix},$$

$$\boldsymbol{C} = \begin{bmatrix}
    1      & 1  \\
    1      & 1  \\
\end{bmatrix},$$ and

$$\boldsymbol{D} = \begin{bmatrix}
    -1   & 0  \\
    0   & -1  \\
\end{bmatrix}.$$

Note that values of $\varepsilon$ are not themselves incorporated into the matrices -- the effect of high versus low $\varepsilon$ is incorporated into stability criteria.

```{r, echo = TRUE}
B      <- matrix(data = 1, nrow = 2, ncol = 2);
C      <- matrix(data = 1, nrow = 2, ncol = 2);
D      <- matrix(data = c(-1, 0, 0, -1), nrow = 2, ncol = 2);
AA    <- D + C %*% solve(A) %*% (-B);
s_AA  <- eigen(AA)$values; print(s_AA);
```

**Conclusion for I:** The **feedback between ecology and evolution therefore destabilises the whole system**, meaning that population densities and traits will not return to their original state if perturbed.

**II. Slow evolution. Competition increases with increased parental investment**

**Now let's try things when the exponent $a = 1$ but $b = 0$, which effectively makes competition *only* a result of parental investment per offspring** (admittedly, it's hard to imagine why this would ever be, but let's give it a try first). This results in the following Jacobian matrix,

$$\boldsymbol{J} = \begin{bmatrix}
    -  & - & - & - \\
    -  & - & - & - \\
    \varepsilon(+)  & \varepsilon(+) & - & 0 \\
    \varepsilon(+)  & \varepsilon(+) & 0 & - \\
\end{bmatrix}.$$

Note that the only thing has changed is $\boldsymbol{B}$; now an increase in $m$ *decreases* population growth because $m_{i}$ and $m_{j}$ both increase copetition. The values in $\boldsymbol{C}$ do not change because evolution should, else being equal, still cause increased parental investment with increased population size,

$$\boldsymbol{B} = \begin{bmatrix}
    -1       & -1  \\
    -1       & -1  \\
\end{bmatrix}$$

We have the same stability criteria $s\left(\boldsymbol{D} + \boldsymbol{C}\boldsymbol{A}^{-1}(-\boldsymbol{B})\right) < 0$ as above for the case where competition increased with offspring number $n$. We only need to change $\boldsymbol{B}$ to assess it.

```{r, echo = TRUE}
B      <- matrix(data = -1, nrow = 2, ncol = 2);
AA    <- D + C %*% solve(A) %*% (-B);
s_AA  <- eigen(AA)$values; print(s_AA);
```

**Conclusion II.** In the above, the leading eigenvalue is negative, so **the whole system is stable. Hence, when competition increases with increasing parental investment as opposed to increasing offspring number, the whole system is also stable.** Note that under these conditions (slow evolution), the whole system cannot be stable if $\boldsymbol{A}$ by itself (the ecology) is not stable.

**III. Fast evolution. Competition increases with increased offspring production**

Here we need to go back to our positive values in $\boldsymbol{B}$ for $m$ and negative values for $n$,

$$\boldsymbol{B} = \begin{bmatrix}
    1       & 1   \\
    1       & 1   \\
\end{bmatrix}.$$

```{r, echo = TRUE}
B      <- matrix(data = 1, nrow = 2, ncol = 2);
```

But a new stability criteria is needed for when evolution happens quickly relative to ecology (high $\varepsilon$), as shown by @Patel2018. We might also consider this kind of situation as occurring when individuals can plastically respond to their environment to adjust parental investment as the situation requires to maximise their inclusive fitness. **Note that, as mentioned above, the evolution of the system is neutrally stable**; there are only complete negative correlations between a single species $m$ and $n$, and the eigenvalues of $\boldsymbol{D}$ are all unity. Consequently, **the stability caused by feedbacks between ecology and evolution solely determine the stability of the system as a whole**. The stability criteria under these circumstances are as follows,

$$s\left(\boldsymbol{A} + \boldsymbol{B}\boldsymbol{D}^{-1}(-\boldsymbol{C})\right) < 0$$

```{r, echo =TRUE}
AA    <- A + B %*% solve(D) %*% (-C);
s_AA  <- eigen(AA)$values; print(s_AA);
```

**Conclusion III.** The leading eigenvalue of the above is positive, meaning that **when individuals can quickly adjust their parental investment and offspring production (or traits quickly evolve), and competition is determined by offspring number**, the whole system is unstable, even if the ecology by itself *is* stable.

**IV. Fast evolution. Competition increases with increased parental investment**

To determine the stability when evolution is fast, but competition increases with increased parental investment, we must start by first putting $\boldsymbol{B}$ back to what it was.

$$\boldsymbol{B} = \begin{bmatrix}
    -1       & -1   \\
    -1       & -1   \\
\end{bmatrix}.$$

```{r, echo = TRUE}
B      <- matrix(data = -1, nrow = 2, ncol = 2);
```

We then test the stability criteria.

```{r, echo = TRUE}
B      <- matrix(data = -1, nrow = 2, ncol = 2);
AA     <- A + B %*% solve(D) %*% (-C);
s_AA   <- eigen(AA)$values; print(s_AA);
```

**Conclusion IV.** Less interesting than it sounds; **when competition is increased by parental investment, the rapid change in traits stabilises the whole system.** (recall that we have defined the system to be evolutionarily stable a priori). Unfortunately, it **very** rarely stabilises a system that is not already stable. The code below finds one case:

```{r, echo = TRUE}
vals <- NULL;
for(i in 1:10000){
    u      <- runif(n = 4, min = -10, max = -0.0001);
    v      <- runif(n = 8, min = 0, max = 10);
    A      <- matrix(data = c(u[1], u[2], u[3], u[4]), nrow = 2, ncol = 2);
    if(max(Re(eigen(A)$values)) > 0){
        B      <- matrix(data = c(v[1], v[2], v[3], v[4]), nrow = 2, ncol = 2);
        C      <- matrix(data = c(v[5], v[6], v[7], v[8]), nrow = 2, ncol = 2);
        AA     <- A + B %*% solve(D) %*% (-C);
        s_AA   <- eigen(AA)$values; 
        vals   <- c(vals, max(Re(s_AA)));
        if(max(Re(s_AA)) < 0){
            break;
        }
    }
}
print(A);
print(B);
print(C);
print(D);
print(eigen(A)$values);
print(eigen(AA)$values);
print(paste("Stabilising evolution finally found after ", i, "tries"));
```

Essentially, how quickly everything happens needs to differe between species.

References
--------------------------------------------------------------------------------

### 2018-03-15 random_matrix_stability.Rmd
Initialise all of the old files from the dungfly repository

---
title: "Stability concerns in random matrices"
author: "Brad Duthie"
bibliography: references.bib
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: evolution.csl
biblio-style: apalike
---

Background on the key question
--------------------------------------------------------------------------------

Recently, @Patel2018 demonstrated the stability criteria for eco-evolutionary systems. A key result was that an unstable (stable) ecological community could be stabilised (destabilised) by fast eco-evolutionary feedback, assuming the evolutionary system was itself stable. Likewise, @Patel2018 showed that an unstable (stable) evolutionary system could be stabilised (destabilised) by fast ecological feedback, assuming the ecological community was itself stable. **Here I am going to show why the stability criteria for eco-evolutionary systems is such that stable ecological systems are *increasingly* likely to be destablised, and unstable ecological systems are *decreasingly* likely to be stabilised, at higher numbers of species traits**. I will use a random matrix approach similar to that of Allesina and colleagues [@Allesina2011; @Allesina2012; @Allesina2015a]. The mathematical concept of what I describe below is not new, and is a similar point to what @May1971 explained nearly half a century ago (that increasing species number decreases stability). Still, this point was not, that I can find, noticed or acknowledged by @Patel2018 when coming up with their stability criteria.

Additionally, I want to look at random ecological matrices from the context of varying generation time, as partly inspired by the use of $\varepsilon$ by @Patel2018 and partly by @Jirsa2004 to separate generation time of different species in a community. Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of the other. Yet I've not been able to find anything on this in the literature in a random matrix context. I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**. Hence, the goals here are twofold.

1. Demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.

2. Determine if variance in generation time stabilises or destabilises random matrices.

I'm not sure if either would make a publication, or perhaps just a blog post, and whether they would need to be completely separated. I will present one a time.

The destabilising effect of rapid evolution
--------------------------------------------------------------------------------

Assume a community of $n$ species, each with the same evolving trait (e.g., body size). Each species $i$ therefore has its own associated trait $m_{i}$ (originally I was thinking about parental investment -- I thought that I had come up with something really interesting, then realised that @Parker1986 got there first). We can describe the dynamics of a system with two species $i$ and $j$ using the Jacobian matrix $\boldsymbol{J}$ below [@Patel2018],

$$\boldsymbol{J} = \begin{bmatrix}
    \boldsymbol{A}  & \boldsymbol{B}  \\
    \boldsymbol{C}  & \boldsymbol{D}  \\  \end{bmatrix}
    = \begin{bmatrix}
    \frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \frac{\partial N_{i}W_{i}}{\partial N_{j}} & \frac{\partial N_{i}W_{i}}{\partial m_{i}} & \frac{\partial N_{i}W_{i}}{\partial m_{j}}  \\
    \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \frac{\partial N_{j}W_{j}}{\partial N_{j}} & \frac{\partial N_{j}W_{j}}{\partial m_{i}} & \frac{\partial N_{j}W_{j}}{\partial m_{j}} \\
    \varepsilon\frac{\partial m_{i}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{i}}{\partial N_{j}} & \varepsilon\frac{\partial m_{i}}{\partial m_{i}} & \varepsilon\frac{\partial m_{i}}{\partial m_{j}}  \\
    \varepsilon\frac{\partial m_{j}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{j}}{\partial N_{j}} & \varepsilon\frac{\partial m_{j}}{\partial m_{i}} & \varepsilon\frac{\partial m_{j}}{\partial m_{j}}  \\
\end{bmatrix}.$$

Note that the matrix $\boldsymbol{J}$ can be separated into four separate submatrices. The submatrix $\boldsymbol{A}$ describes the effects that the density of one species has on its own population growth (diagonal) or that of another species (off-diagonal). The submatrix $\boldsymbol{B}$ describes the effect the trait $m$ of each species has on the population growth of itself or another species. The submatrix $\boldsymbol{C}$ describes the effect that the density of a species has on a species' changing $m$, and $\boldsymbol{D}$ describes the effect that traits have on traits. The value of $\varepsilon$ defines the rate of evolutionary change relative to ecological change; here I will just assume a high value of $\varepsilon$, but the conclusions regarding an increasing number of traits' effect on a relatively slowly changing ecoogical system will be equally applicable to an incresing number of species in relativley slowly evolving system. I will therefore only assume a high value of $\varepsilon$ for now.

For the whole system (ecology and evolution) to be stable when traits change quickly relative to population densities, evolution by itself $\boldsymbol{D}$ must be stable. This is the case when all of the eigenvalues of $\boldsymbol{D}$ are negative. Additionally, for the system $\boldsymbol{J}$ as a whole to be stable, we need to consider the feedback between ecology and evolution. @Patel2018 showed that the system as a whole is stable when the following is satisfied,

$$s\left(\boldsymbol{A} + \boldsymbol{B}\boldsymbol{D}^{-1}(-\boldsymbol{C})\right) < 0.$$

I am specifically interested in the case in which $\boldsymbol{D}$ is known to be stable. Using random matrices containing element values that are real numbers uniformly distributed between $-4$ and $4$, I will show that rapid evolution is almost never stabilising and almost always destabilising when over six species and their traits $m$ are considered. To do this, I will use the following code, which finds a stable random matrix $\boldsymbol{D}$ and assigns random element values (possibly, but not necessarily stable) to $\boldsymbol{A}$, $\boldsymbol{B}$, and $\boldsymbol{C}$. I will do this for 20000 replicates of species numbers from $n = 2$ to $n = 8$.

```{r, echo = TRUE}
random_comm <- function(dm = 2, iter = 10000){
    stabilised_fastevo <- NULL;
    unstabled_fastevo  <- NULL;
    nn           <- dm * dm;
    tot_stabled  <- 0;
    tot_ustabled <- 0;
    tot_eco_stab <- 0;
    stabilised_fastevo <- NULL;
    unstabled_fastevo  <- NULL;
    while(iter > 0){
        stabilised   <- 0;
        unstabilised <- 0;
        Adat <- runif(n = nn, min = -4, max = 4);
        Bdat <- runif(n = nn, min = -4, max = 4);
        Cdat <- runif(n = nn, min = -4, max = 4);
        A    <- matrix(data = Adat, nrow = dm, ncol = dm);
        B    <- matrix(data = Bdat, nrow = dm, ncol = dm);
        C    <- matrix(data = Cdat, nrow = dm, ncol = dm);
        egD  <- 1;
        while(egD >= 0){
             Ddat <- runif(n = nn, min = -4, max = 4);
             D    <- matrix(data = Ddat, nrow = dm, ncol = dm);
             egD  <- max(Re(eigen(D)$values))
        }
        eco_stable       <- max(Re(eigen(A)$values)) < 0;
        slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
        stable_fast_evo  <- slow_evo_maxE < 0;
        if(eco_stable == FALSE & stable_fast_evo == TRUE){
            list_ele <- length(stabilised_fastevo$A) + 1;
            stabilised_fastevo$A[[list_ele]] <- A;
            stabilised_fastevo$B[[list_ele]] <- B;
            stabilised_fastevo$C[[list_ele]] <- C;
            stabilised_fastevo$D[[list_ele]] <- D;
            stabilised   <- 1;
        }
        if(eco_stable == TRUE & stable_fast_evo == FALSE){
            list_ele <- length(unstabled_fastevo$A) + 1;
            unstabled_fastevo$A[[list_ele]] <- A;
            unstabled_fastevo$B[[list_ele]] <- B;
            unstabled_fastevo$C[[list_ele]] <- C;
            unstabled_fastevo$D[[list_ele]] <- D;
            unstabilised   <- 1;
        }
        if(stabilised == 1){
            tot_stabled <- tot_stabled + 1;
        }
        if(unstabilised == 1){
            tot_ustabled <- tot_ustabled + 1;
        }
        if(eco_stable == TRUE){
            tot_eco_stab <- tot_eco_stab + 1;
        }
        iter <- iter - 1;
    }
    results <- list(ecologically_stable = tot_eco_stab, 
                    stabilised = tot_stabled, destabilised = tot_ustabled,
                    stabilised_all   = stabilised_fastevo, 
                    destabilised_all = unstabled_fastevo);
    return(results);
}
```

Note that the code above takes a while to run for a large number of species, and this is mostly caused by having to find a random matrix $\boldsymbol{D}$ that is stable. The code first checks whether $\boldsymbol{A}$ is stable, and if it is stable, whether or not eco-evolutionary dynamics destabilise the stable ecological community. If it is not stable, then the code checks whether eco-evolutionary dynamics stabilise the unstable ecological community. As expected [@May1971], ecological stability decreases with an increasing number of species.

```{r, echo = FALSE}
dat       <- read.table(file = "rand_sims.txt", header = TRUE);
species   <- dat[,1];
pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion stable (n)",
     cex.axis = 1.5, cex.lab = 1.5);
```

More interesting, the plot below shows eco-evolutionary results for ecological communities that **are** stable. The x-axis shows the number of species in the community (recall that each species has the same evolving trait potentially affecting ecology and evolution), and the y-axis shows the proportion of of these communities that are *destabilised* by eco-evolutionary dynamics.

```{r, echo = FALSE}
pr_destabilised <- dat[,3] / dat[,2];
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion destabilised",
     cex.axis = 1.5, cex.lab = 1.5);
```

Note that the proportion of communities that are destabilised by eco-evolutionary effects increases with increasing species number. The plot below shows the eco-evolutionary results for ecological communities that are unstable. The y-axis here shows the proportion of communities that are *stabilised* by eco-evolutionary dynamics. This decreases with increasing species number.

```{r, echo = FALSE}
pr_stabilised <- dat[,5] / dat[,4];
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion stabilised",
     cex.axis = 1.5, cex.lab = 1.5);
```

Hence, overall, **as species number and therefore evolving species traits increases, the probability that stable ecological communities become destabilised increases rapidly and the probability that unstable ecological communities are stabilised becomes vanishingly small**. I am not sure if I can prove this mathematically, but I don't think I need to for the point to be made effectively. What I want to figure out is whether or not this point can or should be published in the peer-reviwed literature -- else it's a blog post.

The effect of variation in generation time on ecological stability
--------------------------------------------------------------------------------

Here I will only consider $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time. I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently. I assume that the response of a particular species' growth rate (rows in $\boldsymbol{A}$) is consistently affected by a value $\varepsilon_{i}$; higher values of $\varepsilon_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times. The matrix below $\boldsymbol{A}$ for two species $i$ and $j$ with generation times $\varepsilon_{i}$ and $\varepsilon_{j}$, respectively.

$$\boldsymbol{A} = \begin{bmatrix}
    \varepsilon_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \varepsilon_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
    \varepsilon_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \varepsilon_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
\end{bmatrix}.$$

I will allow random values of $\varepsilon_{i}$ to vary from $1$ to $100$, modelling two orders of magnitude of generation times. I will compare the stability of two sets of random community matrices -- one in which all $\varepsilon = 1$ and another in which $\varepsilon$ varies.

```{r, echo = TRUE, eval = FALSE}
nn     <- 8;
A1_stt <- 0;
A2_stt <- 0;
iter   <- 100000;
while(iter > 0){
    A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
    A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
    A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
    A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
    epsil   <- runif(n = nn, min = 1, max = 100);
    eps_dat <- rep(x = epsil, times = nn);
    eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
    A2      <- A2 * eps_mat;
    A1_stb  <- max(Re(eigen(A1)$values)) < 0;
    A2_stb  <- max(Re(eigen(A2)$values)) < 0;
    if(A1_stb == TRUE){
      A1_stt <- A1_stt + 1;
    }
    if(A2_stb == TRUE){
      A2_stt <- A2_stt + 1;
    }
    iter    <- iter - 1;
}
```

The number of stable random matrices when $\varepsilon$ does not vary was $13$, and the number of stable random matrices when it does was $48$ (several re-runs of the above code found this result to be roughly the same for $n = 8$, with smaller differences for lower $n$, and presumably larger differences for higher $n$). The effect of varying $\varepsilon$ appears to be more extreme with an increasing number of species. Hence, overall, **variation in generation time appears to be stabilising for random ecological communities**. Again, I am not sure if this is something that I can prove mathematically, and I cannot decide whether the result is interesting enough to try to publish by itself or in tandem with the above. If not, it might be another blog post.

Combining the two (work in progress)
--------------------------------------------------------------------------------

```{r, echo = FALSE}
dat       <- read.table(file = "rand_sims_var_gen.txt", header = TRUE);
species   <- dat[,1];
pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion stable (n)",
     cex.axis = 1.5, cex.lab = 1.5);
```

```{r, echo = FALSE}
pr_destabilised <- dat[,3] / dat[,2];
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion destabilised",
     cex.axis = 1.5, cex.lab = 1.5);
```

```{r, echo = FALSE}
pr_stabilised <- dat[,5] / dat[,4];
par(mar = c(5, 5, 2, 1));
plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
     xlab = "Species number", ylab = "Proportion stabilised",
     cex.axis = 1.5, cex.lab = 1.5);
```

References
--------------------------------------------------------------------------------

### 2018-03-15 random_matrix_PI.R

@@ -1,0 +1,216 @@
+
+
+
+
+random_comm <- function(dm = 2, iter = 10000){
+    stabilised_fastevo <- NULL;
+    unstabled_fastevo  <- NULL;
+    nn           <- dm * dm;
+    tot_stabled  <- 0;
+    tot_ustabled <- 0;
+    tot_eco_stab <- 0;
+    stabilised_fastevo <- NULL;
+    unstabled_fastevo  <- NULL;
+    while(iter > 0){
+        stabilised   <- 0;
+        unstabilised <- 0;
+        Adat <- runif(n = nn, min = -4, max = 4);
+        Bdat <- runif(n = nn, min = -4, max = 4);
+        Cdat <- runif(n = nn, min = -4, max = 4);
+        A    <- matrix(data = Adat, nrow = dm, ncol = dm);
+        B    <- matrix(data = Bdat, nrow = dm, ncol = dm);
+        C    <- matrix(data = Cdat, nrow = dm, ncol = dm);
+        egD  <- 1;
+        while(egD >= 0){
+             Ddat <- runif(n = nn, min = -4, max = 4);
+             D    <- matrix(data = Ddat, nrow = dm, ncol = dm);
+             egD  <- max(Re(eigen(D)$values))
+        }
+        eco_stable       <- max(Re(eigen(A)$values)) < 0;
+        slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
+        stable_fast_evo  <- slow_evo_maxE < 0;
+        if(eco_stable == FALSE & stable_fast_evo == TRUE){
+            list_ele <- length(stabilised_fastevo$A) + 1;
+            stabilised_fastevo$A[[list_ele]] <- A;
+            stabilised_fastevo$B[[list_ele]] <- B;
+            stabilised_fastevo$C[[list_ele]] <- C;
+            stabilised_fastevo$D[[list_ele]] <- D;
+            stabilised   <- 1;
+        }
+        if(eco_stable == TRUE & stable_fast_evo == FALSE){
+            list_ele <- length(unstabled_fastevo$A) + 1;
+            unstabled_fastevo$A[[list_ele]] <- A;
+            unstabled_fastevo$B[[list_ele]] <- B;
+            unstabled_fastevo$C[[list_ele]] <- C;
+            unstabled_fastevo$D[[list_ele]] <- D;
[…]

### 2018-03-15 random_matrix_stability.Rmd
Organised notes of plotted results

@@
 
-The number of stable random matrices when $\varepsilon$ does not vary was $13$, and the number of stable random matrices when it does was $48$ (several re-runs of the above code found this result to be roughly the same for $n = 8$, with smaller differences for lower $n$, and presumably larger differences for higher $n$).
-The effect of varying $\varepsilon$ appears to be more extreme with an increasing number of species.
-Hence, overall, **variation in generation time appears to be stabilising for random ecological communities**.
-Again, I am not sure if this is something that I can prove mathematically, and I cannot decide whether the result is interesting enough to try to publish by itself or in tandem with the above.
+The number of stable random matrices when $\varepsilon$ does not vary was $13$, and the number of stable random matrices when it does was $48$.
+To further investigate, it is worth looking at **how the amount of variation in generation time in a community affects stability**.
+Below is some code to look at stability in random communities across different magnitudes of variance in generation time for communities with eight species.
+
+```{r, echo = TRUE, eval = FALSE}
+res  <- NULL;
+mags <- 10^(0:5);
+for(i in mags){
+    nn     <- 8;
+    A1_stt <- 0;
+    A2_stt <- 0;
+    iter   <- 1000000;
+    while(iter > 0){
+        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
+        epsil   <- runif(n = nn, min = 1, max = i);
+        eps_dat <- rep(x = epsil, times = nn);
+        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
+        A2      <- A2 * eps_mat;
+        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
+        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
+        if(A1_stb == TRUE){
+          A1_stt <- A1_stt + 1;
+        }
+        if(A2_stb == TRUE){
+          A2_stt <- A2_stt + 1;
+        }
+        iter    <- iter - 1;
+    }
+    res <- rbind(res, c(A1_stt, A2_stt));
+    print(i);
+}
+```
+
+The code below produces the following results showing the proportion of random communities that are stable as the variation in generation time increases in orders of magnitude (note that I'm using eight species, so the y-axis shows a very low proportion range).
+
+```{r, echo = TRUE, eval = TRUE}
+mag_var <- read.table(file = "var_gen_mag.csv", header = TRUE, sep = ",");
+stabi_3 <- mag_var[,3] / 1000000;
+par(mar = c(5, 5, 2, 1));
+plot(x = log10(mag_var[,1]), y = stabi_3, type = "b", lwd = 2, ylim = c(0, 0.0005),
+     xlab = "LOG Generation time variation among species", ylab = "Proportion stable",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
+
+**It looks like once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 20 for identical generation times versus magnitudes of generation time variation of 1000.
+
+```{r, echo = TRUE, eval = FALSE}
+res <- NULL;
+for(i in 2:8){
+    nn     <- i;
+    A1_stt <- 0;
+    A2_stt <- 0;
+    iter   <- 1000000;
+    while(iter > 0){
+        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
+        epsil   <- runif(n = nn, min = 1, max = 1000);
+        eps_dat <- rep(x = epsil, times = nn);
+        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
+        A2      <- A2 * eps_mat;
+        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
+        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
+        if(A1_stb == TRUE){
+          A1_stt <- A1_stt + 1;
+        }
+        if(A2_stb == TRUE){
+          A2_stt <- A2_stt + 1;
+        }
+        iter    <- iter - 1;
+    }
+    res <- rbind(res, c(A1_stt, A2_stt));
+    print(i);
+}
+```
+
+The plot below shows the result for the proportion of stable communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
+
+```{r, echo = TRUE, eval = TRUE}
+gen_var <- read.table(file = "dif_sp_gen_var.csv", header = TRUE, sep = ",");
+stabi_1 <- gen_var[,2] / 1000000;
+stabi_2 <- gen_var[,3] / 1000000;
+par(mar = c(5, 5, 2, 1));
+plot(x = gen_var[,1], y = stabi_1, type = "b", lwd = 2, ylim = c(0, 1),
+     xlab = "Species number", ylab = "Proportion stable",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
+points(x = gen_var[,1], y = stabi_2, type = "b", lwd = 2, ylim = c(0, 1),
+       lty = "solid", pch = 20);
+legend(x = 2, y = 1, lwd = 2,
+       legend = c("Generation time varies", "Generation time does not vary"),
+       lty = c("solid", "dotted"));
+```
+
+It is difficult to see the difference between these two numbers, so the figure below shows the ratio of the proportion of stable communities given generation variation relative to no generation variation (No_gen_var:Gen_var).
+
+```{r, echo = TRUE, eval = TRUE}
+par(mar = c(5, 5, 2, 1));
+plot(x = gen_var[,1], y = stabi_2 / stabi_1, type = "b", lwd = 2, ylim = c(0, 5),
+     xlab = "Species number", ylab = "Stable ratio (var to no var)",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
+
+This is interesting.
+**As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
+Assuming that I have done this correctly, this could be a fairly high impact result -- variation in generation time is important for community stability.
+I have not seen this take-home message in the literatur -- at least, not presented in this way.
+Again, I am not sure if any of this is something that I can prove mathematically, and I cannot decide whether the result is interesting enough to try to publish by itself or in tandem with the above.
 If not, it might be another blog post.

### 2018-03-16 random_matrix_PI.R
Update to simulate more

@@ -194,6 +194,9 @@ random_comm <- function(dm = 2, iter = 10000){
       tot_eco_stab <- tot_eco_stab + 1;
     }
     iter <- iter - 1;
+    if(iter %% 1000000 == 0){
+        print(c(dm, iter));
+    }
   }
   results <- list(ecologically_stable = tot_eco_stab, 
                   stabilised = tot_stabled, destabilised = tot_ustabled,
@@ -202,9 +205,9 @@ random_comm <- function(dm = 2, iter = 10000){
   return(results);
 }
 
-NN       <- 8;
+NN       <- 20;
 res_tabl <- matrix(data = 0, nrow = (NN-1), ncol = 5);
-its      <- 10000;
+its      <- 100000000;
 for(i in 2:NN){
   res              <- random_comm(dm = i, iter = its); 
   res_tabl[i-1, 1] <- i;                               # Species and traits

### 2018-03-16 random_matrix_stability.Rmd
More simulations included
More thoughts on stability

@@
 To do this, I will use the following code, which finds a stable random matrix $\boldsymbol{D}$ and assigns random element values (possibly, but not necessarily stable) to $\boldsymbol{A}$, $\boldsymbol{B}$, and $\boldsymbol{C}$.
-I will do this for 20000 replicates of species numbers from $n = 2$ to $n = 8$.
+I will do this for 100000 replicates of species numbers from $n = 2$ to $n = 8$.
 
@@
 I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
-I assume that the response of a particular species' growth rate (rows in $\boldsymbol{A}$) is consistently affected by a value $\varepsilon_{i}$; higher values of $\varepsilon_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
-The matrix below $\boldsymbol{A}$ for two species $i$ and $j$ with generation times $\varepsilon_{i}$ and $\varepsilon_{j}$, respectively.
+I assume that the response of a particular species' growth rate (rows in $\boldsymbol{A}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
+Higher values of $\gamma_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
+The matrix below $\boldsymbol{A}$ for two species $i$ and $j$ with generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
 
 $$\boldsymbol{A} = \begin{bmatrix}
-    \varepsilon_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \varepsilon_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
-    \varepsilon_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \varepsilon_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
+    \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
+    \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
 \end{bmatrix}.$$
 
-I will allow random values of $\varepsilon_{i}$ to vary from $1$ to $100$, modelling two orders of magnitude of generation times.
-I will compare the stability of two sets of random community matrices -- one in which all $\varepsilon = 1$ and another in which $\varepsilon$ varies.
+I will first allow random values of $\gamma_{i}$ to vary from $1$ to $100$, modelling two orders of magnitude of generation times.
+I will compare the stability of two sets of random community matrices -- one in which all $\gamma = 1$ and another in which $\gamma$ varies.
 
@@
     A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
-    epsil   <- runif(n = nn, min = 1, max = 100);
-    eps_dat <- rep(x = epsil, times = nn);
-    eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
+    gamval  <- runif(n = nn, min = 1, max = 100);
+    eps_dat <- rep(x = gamval, times = nn);
+    eps_mat <- matrix(data = gamval, nrow = nn, ncol = nn, byrow = FALSE);
     A2      <- A2 * eps_mat;
@@
 
-The number of stable random matrices when $\varepsilon$ does not vary was $13$, and the number of stable random matrices when it does was $48$.
-To further investigate, it is worth looking at **how the amount of variation in generation time in a community affects stability**.
+The number of stable random matrices when $\gamma$ does not vary was $13$, and the number of stable random matrices when it does was $48$.
+To further investigate, it is worth looking at **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
 Below is some code to look at stability in random communities across different magnitudes of variance in generation time for communities with eight species.
@@
     A2_stt <- 0;
-    iter   <- 1000000;
+    iter   <- 10000000;
     while(iter > 0){
@@
         A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
-        epsil   <- runif(n = nn, min = 1, max = i);
-        eps_dat <- rep(x = epsil, times = nn);
-        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
-        A2      <- A2 * eps_mat;
+        gamval  <- runif(n = nn, min = 1, max = i);
+        gam_dat <- rep(x = gamval, times = nn);
+        gam_mat <- matrix(data = gam_dat, nrow = nn, ncol = nn, byrow = FALSE);
+        A2      <- A2 * gam_mat;
         A1_stb  <- max(Re(eigen(A1)$values)) < 0;
@@
 
+Code to print the above is below.
+
+```{r, eval = FALSE, echo = TRUE}
+random_comm <- function(dm = 2, iter = 10000){
+  stabilised_fastevo <- NULL;
+  unstabled_fastevo  <- NULL;
+  nn           <- dm * dm;
+  tot_stabled  <- 0;
+  tot_ustabled <- 0;
+  tot_eco_stab <- 0;
+  stabilised_fastevo <- NULL;
+  unstabled_fastevo  <- NULL;
+  while(iter > 0){
+    epsil   <- runif(n = dm, min = 1, max = 1000);
+    eps_dat <- rep(x = epsil, times = dm);
+    eps_mat <- matrix(data = c(epsil, epsil), nrow = dm, ncol = dm, byrow = FALSE);
+    stabilised   <- 0;
+    unstabilised <- 0;
+    Adat <- runif(n = nn, min = -4, max = 4);
+    Bdat <- runif(n = nn, min = -4, max = 4);
+    Cdat <- runif(n = nn, min = -4, max = 4);
+    A    <- matrix(data = Adat, nrow = dm, ncol = dm) * eps_mat;
+    B    <- matrix(data = Bdat, nrow = dm, ncol = dm) * eps_mat;
+    C    <- matrix(data = Cdat, nrow = dm, ncol = dm) * eps_mat;
+    egD  <- 1;
+    while(egD >= 0){
+      Ddat <- runif(n = nn, min = -4, max = 4);
+      D    <- matrix(data = Ddat, nrow = dm, ncol = dm) * eps_mat;
+      egD  <- max(Re(eigen(D)$values))
+    }
+    eco_stable       <- max(Re(eigen(A)$values)) < 0;
+    slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
+    stable_fast_evo  <- slow_evo_maxE < 0;
+    if(eco_stable == FALSE & stable_fast_evo == TRUE){
+      list_ele <- length(stabilised_fastevo$A) + 1;
+      stabilised_fastevo$A[[list_ele]] <- A;
+      stabilised_fastevo$B[[list_ele]] <- B;
+      stabilised_fastevo$C[[list_ele]] <- C;
+      stabilised_fastevo$D[[list_ele]] <- D;
+      stabilised   <- 1;
+    }
+    if(eco_stable == TRUE & stable_fast_evo == FALSE){
+      list_ele <- length(unstabled_fastevo$A) + 1;
+      unstabled_fastevo$A[[list_ele]] <- A;
+      unstabled_fastevo$B[[list_ele]] <- B;
+      unstabled_fastevo$C[[list_ele]] <- C;
+      unstabled_fastevo$D[[list_ele]] <- D;
+      unstabilised   <- 1;
+    }
+    if(stabilised == 1){
+      tot_stabled <- tot_stabled + 1;
+    }
+    if(unstabilised == 1){
+      tot_ustabled <- tot_ustabled + 1;
+    }
+    if(eco_stable == TRUE){
+      tot_eco_stab <- tot_eco_stab + 1;
+    }
+    iter <- iter - 1;
+    if(iter %% 1000000 == 0){
+        print(c(dm, iter));
+    }
+  }
+  results <- list(ecologically_stable = tot_eco_stab, 
+                  stabilised = tot_stabled, destabilised = tot_ustabled,
+                  stabilised_all   = stabilised_fastevo, 
+                  destabilised_all = unstabled_fastevo);
+  return(results);
+}
+
+NN       <- 20;
+res_tabl <- matrix(data = 0, nrow = (NN-1), ncol = 5);
+its      <- 100000000;
+for(i in 2:NN){
+  res              <- random_comm(dm = i, iter = its); 
+  res_tabl[i-1, 1] <- i;                               # Species and traits
+  res_tabl[i-1, 2] <- res$ecologically_stable;         # Total stable
+  res_tabl[i-1, 3] <- res$destabilised;                # Total destabilised
+  res_tabl[i-1, 4] <- its - res_tabl[i-1, 2];          # Total unstable
+  res_tabl[i-1, 5] <- res$stabilised;                  # Total stabilised
+  print(i);
+}
+```
+
 References

### 2018-03-17 random_matrix_stability.Rmd
Add introductory paragraph

@@
 ---
+
+********************************************************************************
+
+**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
+But in ecological communities, the magnitude of interaction effects will be highly dependent upon species generation time.
+The per capita effect of individuals from species with relatively short generation times will be lower than that of individuals from species with longer generation times.
+Interaction effects on are therefore expected to be non-random within species, with some species having consistently weaker or stronger effects on other species per capita.
+Here we show that by accounting for variation in generation time among species across three orders of magnitude, community stability is substantially increased.
+Further, the influence of variation in generation time becomes exponentially more important as community size increases.
+We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
+
+********************************************************************************
+
+The above is a draft introductory paragraph or abstract.
+I believe that we could get [an article on the above](https://www.nature.com/natecolevol/about/content) clearly in the space of 3500, or maybe even 1500 words.
+As far as I've been able to find, such an article would be novel and interesting (and mathematically correct).
 
@@
 mag_var <- read.table(file = "var_gen_mag.csv", header = TRUE, sep = ",");
-stabi_3 <- mag_var[,3] / 1000000;
+stabi_3 <- mag_var[,3] / 10000000;
 par(mar = c(5, 5, 2, 1));

### 2018-03-17 random_matrix_stability.Rmd
fix abstract

@@
 **Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
-But in ecological communities, the magnitude of interaction effects will be highly dependent upon species generation time.
-The per capita effect of individuals from species with relatively short generation times will be lower than that of individuals from species with longer generation times.
-Interaction effects on are therefore expected to be non-random within species, with some species having consistently weaker or stronger effects on other species per capita.
-Here we show that by accounting for variation in generation time among species across three orders of magnitude, community stability is substantially increased.
+But in ecological communities, the rate of change of species density and the magnitude of interaction effects depend on species generation time.
+Per capita effects of individuals from species with short generation times will be lower than individuals from species with long generation times, and densities of species with shorter generation times will change more rapidly.
+Interaction effects are therefore expected to be correlated within species, with some species having consistently weaker or stronger effects on others per capita.
+Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
 Further, the influence of variation in generation time becomes exponentially more important as community size increases.

### 2018-03-17 random_matrix_stability.Rmd
Re-orient notes

@@
 
+The below is a draft introductory paragraph or abstract.
+I believe that we could get [an article on the above](https://www.nature.com/natecolevol/about/content) clearly in the space of 3500, or maybe even 1500 words.
+As far as I've been able to find, such an article would be novel and interesting (and mathematically correct).
+These notes are divided into two parts at the moment.
+
+1. [The effect of variation in generation time on community stability.](#gen_time_stability)
+2. [The effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel)
+
+I suspect that part 2 will need to branch off soon, but I don't want to forget about it.
+
 ********************************************************************************
@@
 
-The above is a draft introductory paragraph or abstract.
-I believe that we could get [an article on the above](https://www.nature.com/natecolevol/about/content) clearly in the space of 3500, or maybe even 1500 words.
-As far as I've been able to find, such an article would be novel and interesting (and mathematically correct).
-
-Background on the key question
+<a name="gen_time_stability">The effect of variation in generation time on community stability</a>
 --------------------------------------------------------------------------------
 
-Recently, @Patel2018 demonstrated the stability criteria for eco-evolutionary systems.
-A key result was that an unstable (stable) ecological community could be stabilised (destabilised) by fast eco-evolutionary feedback, assuming the evolutionary system was itself stable.
-Likewise, @Patel2018 showed that an unstable (stable) evolutionary system could be stabilised (destabilised) by fast ecological feedback, assuming the ecological community was itself stable.
-**Here I am going to show why the stability criteria for eco-evolutionary systems is such that stable ecological systems are *increasingly* likely to be destablised, and unstable ecological systems are *decreasingly* likely to be stabilised, at higher numbers of species traits**.
-I will use a random matrix approach similar to that of Allesina and colleagues [@Allesina2011; @Allesina2012; @Allesina2015a].
-The mathematical concept of what I describe below is not new, and is a similar point to what @May1971 explained nearly half a century ago (that increasing species number decreases stability).
-Still, this point was not, that I can find, noticed or acknowledged by @Patel2018 when coming up with their stability criteria.
-
-Additionally, I want to look at random ecological matrices from the context of varying generation time, as partly inspired by the use of $\varepsilon$ by @Patel2018 and partly by @Jirsa2004 to separate generation time of different species in a community.
 Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of the other.
@@
 I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**.
-Hence, the goals here are twofold.
+This work is partly inspired by the use of $\varepsilon$ by @Patel2018 and partly by @Jirsa2004 to separate generation time of different species in a community.
 
-1. Demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.
-
-2. Determine if variance in generation time stabilises or destabilises random matrices.
-
-I'm not sure if either would make a publication, or perhaps just a blog post, and whether they would need to be completely separated. I will present one a time.
-
-The destabilising effect of rapid evolution
---------------------------------------------------------------------------------
-
-Assume a community of $n$ species, each with the same evolving trait (e.g., body size).
-Each species $i$ therefore has its own associated trait $m_{i}$ (originally I was thinking about parental investment -- I thought that I had come up with something really interesting, then realised that @Parker1986 got there first).
-We can describe the dynamics of a system with two species $i$ and $j$ using the Jacobian matrix $\boldsymbol{J}$ below [@Patel2018],
-
-$$\boldsymbol{J} = \begin{bmatrix}
-    \boldsymbol{A}  & \boldsymbol{B}  \\
-    \boldsymbol{C}  & \boldsymbol{D}  \\  \end{bmatrix}
-    = \begin{bmatrix}
-    \frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \frac{\partial N_{i}W_{i}}{\partial N_{j}} & \frac{\partial N_{i}W_{i}}{\partial m_{i}} & \frac{\partial N_{i}W_{i}}{\partial m_{j}}  \\
-    \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \frac{\partial N_{j}W_{j}}{\partial N_{j}} & \frac{\partial N_{j}W_{j}}{\partial m_{i}} & \frac{\partial N_{j}W_{j}}{\partial m_{j}} \\
-    \varepsilon\frac{\partial m_{i}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{i}}{\partial N_{j}} & \varepsilon\frac{\partial m_{i}}{\partial m_{i}} & \varepsilon\frac{\partial m_{i}}{\partial m_{j}}  \\
-    \varepsilon\frac{\partial m_{j}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{j}}{\partial N_{j}} & \varepsilon\frac{\partial m_{j}}{\partial m_{i}} & \varepsilon\frac{\partial m_{j}}{\partial m_{j}}  \\
-\end{bmatrix}.$$
-
-Note that the matrix $\boldsymbol{J}$ can be separated into four separate submatrices.
-The submatrix $\boldsymbol{A}$ describes the effects that the density of one species has on its own population growth (diagonal) or that of another species (off-diagonal).
-The submatrix $\boldsymbol{B}$ describes the effect the trait $m$ of each species has on the population growth of itself or another species.
-The submatrix $\boldsymbol{C}$ describes the effect that the density of a species has on a species' changing $m$, and $\boldsymbol{D}$ describes the effect that traits have on traits.
-The value of $\varepsilon$ defines the rate of evolutionary change relative to ecological change; here I will just assume a high value of $\varepsilon$, but the conclusions regarding an increasing number of traits' effect on a relatively slowly changing ecoogical system will be equally applicable to an incresing number of species in relativley slowly evolving system.
-I will therefore only assume a high value of $\varepsilon$ for now.
-
-For the whole system (ecology and evolution) to be stable when traits change quickly relative to population densities, evolution by itself $\boldsymbol{D}$ must be stable.
-This is the case when all of the eigenvalues of $\boldsymbol{D}$ are negative.
-Additionally, for the system $\boldsymbol{J}$ as a whole to be stable, we need to consider the feedback between ecology and evolution.
-@Patel2018 showed that the system as a whole is stable when the following is satisfied,
-
-$$s\left(\boldsymbol{A} + \boldsymbol{B}\boldsymbol{D}^{-1}(-\boldsymbol{C})\right) < 0.$$
-
-I am specifically interested in the case in which $\boldsymbol{D}$ is known to be stable.
-Using random matrices containing element values that are real numbers uniformly distributed between $-4$ and $4$, I will show that rapid evolution is almost never stabilising and almost always destabilising when over six species and their traits $m$ are considered.
-To do this, I will use the following code, which finds a stable random matrix $\boldsymbol{D}$ and assigns random element values (possibly, but not necessarily stable) to $\boldsymbol{A}$, $\boldsymbol{B}$, and $\boldsymbol{C}$.
-I will do this for 100000 replicates of species numbers from $n = 2$ to $n = 8$.
-
-```{r, echo = TRUE}
-random_comm <- function(dm = 2, iter = 10000){
-    stabilised_fastevo <- NULL;
-    unstabled_fastevo  <- NULL;
-    nn           <- dm * dm;
-    tot_stabled  <- 0;
-    tot_ustabled <- 0;
-    tot_eco_stab <- 0;
-    stabilised_fastevo <- NULL;
-    unstabled_fastevo  <- NULL;
-    while(iter > 0){
-        stabilised   <- 0;
-        unstabilised <- 0;
-        Adat <- runif(n = nn, min = -4, max = 4);
-        Bdat <- runif(n = nn, min = -4, max = 4);
-        Cdat <- runif(n = nn, min = -4, max = 4);
-        A    <- matrix(data = Adat, nrow = dm, ncol = dm);
-        B    <- matrix(data = Bdat, nrow = dm, ncol = dm);
-        C    <- matrix(data = Cdat, nrow = dm, ncol = dm);
-        egD  <- 1;
-        while(egD >= 0){
-             Ddat <- runif(n = nn, min = -4, max = 4);
-             D    <- matrix(data = Ddat, nrow = dm, ncol = dm);
-             egD  <- max(Re(eigen(D)$values))
-        }
-        eco_stable       <- max(Re(eigen(A)$values)) < 0;
-        slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
-        stable_fast_evo  <- slow_evo_maxE < 0;
-        if(eco_stable == FALSE & stable_fast_evo == TRUE){
-            list_ele <- length(stabilised_fastevo$A) + 1;
-            stabilised_fastevo$A[[list_ele]] <- A;
-            stabilised_fastevo$B[[list_ele]] <- B;
-            stabilised_fastevo$C[[list_ele]] <- C;
-            stabilised_fastevo$D[[list_ele]] <- D;
-            stabilised   <- 1;
-        }
-        if(eco_stable == TRUE & stable_fast_evo == FALSE){
-            list_ele <- length(unstabled_fastevo$A) + 1;
-            unstabled_fastevo$A[[list_ele]] <- A;
-            unstabled_fastevo$B[[list_ele]] <- B;
-            unstabled_fastevo$C[[list_ele]] <- C;
-            unstabled_fastevo$D[[list_ele]] <- D;
-            unstabilised   <- 1;
-        }
-        if(stabilised == 1){
-            tot_stabled <- tot_stabled + 1;
-        }
-        if(unstabilised == 1){
-            tot_ustabled <- tot_ustabled + 1;
-        }
-        if(eco_stable == TRUE){
-            tot_eco_stab <- tot_eco_stab + 1;
-        }
-        iter <- iter - 1;
-    }
-    results <- list(ecologically_stable = tot_eco_stab, 
-                    stabilised = tot_stabled, destabilised = tot_ustabled,
-                    stabilised_all   = stabilised_fastevo, 
-                    destabilised_all = unstabled_fastevo);
-    return(results);
-}
-```
-
-Note that the code above takes a while to run for a large number of species, and this is mostly caused by having to find a random matrix $\boldsymbol{D}$ that is stable.
-The code first checks whether $\boldsymbol{A}$ is stable, and if it is stable, whether or not eco-evolutionary dynamics destabilise the stable ecological community.
-If it is not stable, then the code checks whether eco-evolutionary dynamics stabilise the unstable ecological community.
-As expected [@May1971], ecological stability decreases with an increasing number of species.
-
-```{r, echo = FALSE}
-dat       <- read.table(file = "rand_sims.txt", header = TRUE);
-species   <- dat[,1];
-pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stable (n)",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-More interesting, the plot below shows eco-evolutionary results for ecological communities that **are** stable.
-The x-axis shows the number of species in the community (recall that each species has the same evolving trait potentially affecting ecology and evolution), and the y-axis shows the proportion of of these communities that are *destabilised* by eco-evolutionary dynamics.
-
-```{r, echo = FALSE}
-pr_destabilised <- dat[,3] / dat[,2];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion destabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-Note that the proportion of communities that are destabilised by eco-evolutionary effects increases with increasing species number.
-The plot below shows the eco-evolutionary results for ecological communities that are unstable.
-The y-axis here shows the proportion of communities that are *stabilised* by eco-evolutionary dynamics.
-This decreases with increasing species number.
-
-```{r, echo = FALSE}
-pr_stabilised <- dat[,5] / dat[,4];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-Hence, overall, **as species number and therefore evolving species traits increases, the probability that stable ecological communities become destabilised increases rapidly and the probability that unstable ecological communities are stabilised becomes vanishingly small**.
-I am not sure if I can prove this mathematically, but I don't think I need to for the point to be made effectively.
-What I want to figure out is whether or not this point can or should be published in the peer-reviwed literature -- else it's a blog post.
-
-The effect of variation in generation time on ecological stability
---------------------------------------------------------------------------------
-
-Here I will only consider $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time.
+Here I will consider a Jacobian matrix of interacting species $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time.
 I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
@@
 
+<a name="comment_on_Patel">The effect of increasing numbers of evolutionary traits on eco-evolutionary stability</a>
+--------------------------------------------------------------------------------
+
+Recently, @Patel2018 demonstrated the stability criteria for eco-evolutionary systems.
+A key result was that an unstable (stable) ecological community could be stabilised (destabilised) by fast eco-evolutionary feedback, assuming the evolutionary system was itself stable.
+Likewise, @Patel2018 showed that an unstable (stable) evolutionary system could be stabilised (destabilised) by fast ecological feedback, assuming the ecological community was itself stable.
+**Here I am going to show why the stability criteria for eco-evolutionary systems is such that stable ecological systems are *increasingly* likely to be destablised, and unstable ecological systems are *decreasingly* likely to be stabilised, at higher numbers of species traits**.
+I will use a random matrix approach similar to that of Allesina and colleagues [@Allesina2011; @Allesina2012; @Allesina2015a].
+The mathematical concept of what I describe below is not new, and is a similar point to what @May1971 explained nearly half a century ago (that increasing species number decreases stability).
+Still, this point was not, that I can find, noticed or acknowledged by @Patel2018 when coming up with their stability criteria.
+
+Hence, the goal here is to demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.
+
+The destabilising effect of rapid evolution
+--------------------------------------------------------------------------------
+
+Assume a community of $n$ species, each with the same evolving trait (e.g., body size).
+Each species $i$ therefore has its own associated trait $m_{i}$ (originally I was thinking about parental investment -- I thought that I had come up with something really interesting, then realised that @Parker1986 got there first).
+We can describe the dynamics of a system with two species $i$ and $j$ using the Jacobian matrix $\boldsymbol{J}$ below [@Patel2018],
+
+$$\boldsymbol{J} = \begin{bmatrix}
+    \boldsymbol{A}  & \boldsymbol{B}  \\
+    \boldsymbol{C}  & \boldsymbol{D}  \\  \end{bmatrix}
+    = \begin{bmatrix}
+    \frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \frac{\partial N_{i}W_{i}}{\partial N_{j}} & \frac{\partial N_{i}W_{i}}{\partial m_{i}} & \frac{\partial N_{i}W_{i}}{\partial m_{j}}  \\
+    \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \frac{\partial N_{j}W_{j}}{\partial N_{j}} & \frac{\partial N_{j}W_{j}}{\partial m_{i}} & \frac{\partial N_{j}W_{j}}{\partial m_{j}} \\
+    \varepsilon\frac{\partial m_{i}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{i}}{\partial N_{j}} & \varepsilon\frac{\partial m_{i}}{\partial m_{i}} & \varepsilon\frac{\partial m_{i}}{\partial m_{j}}  \\
+    \varepsilon\frac{\partial m_{j}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{j}}{\partial N_{j}} & \varepsilon\frac{\partial m_{j}}{\partial m_{i}} & \varepsilon\frac{\partial m_{j}}{\partial m_{j}}  \\
+\end{bmatrix}.$$
+
+Note that the matrix $\boldsymbol{J}$ can be separated into four separate submatrices.
+The submatrix $\boldsymbol{A}$ describes the effects that the density of one species has on its own population growth (diagonal) or that of another species (off-diagonal).
+The submatrix $\boldsymbol{B}$ describes the effect the trait $m$ of each species has on the population growth of itself or another species.
+The submatrix $\boldsymbol{C}$ describes the effect that the density of a species has on a species' changing $m$, and $\boldsymbol{D}$ describes the effect that traits have on traits.
+The value of $\varepsilon$ defines the rate of evolutionary change relative to ecological change; here I will just assume a high value of $\varepsilon$, but the conclusions regarding an increasing number of traits' effect on a relatively slowly changing ecoogical system will be equally applicable to an incresing number of species in relativley slowly evolving system.
+I will therefore only assume a high value of $\varepsilon$ for now.
+
+For the whole system (ecology and evolution) to be stable when traits change quickly relative to population densities, evolution by itself $\boldsymbol{D}$ must be stable.
+This is the case when all of the eigenvalues of $\boldsymbol{D}$ are negative.
+Additionally, for the system $\boldsymbol{J}$ as a whole to be stable, we need to consider the feedback between ecology and evolution.
+@Patel2018 showed that the system as a whole is stable when the following is satisfied,
+
+$$s\left(\boldsymbol{A} + \boldsymbol{B}\boldsymbol{D}^{-1}(-\boldsymbol{C})\right) < 0.$$
+
+I am specifically interested in the case in which $\boldsymbol{D}$ is known to be stable.
+Using random matrices containing element values that are real numbers uniformly distributed between $-4$ and $4$, I will show that rapid evolution is almost never stabilising and almost always destabilising when over six species and their traits $m$ are considered.
+To do this, I will use the following code, which finds a stable random matrix $\boldsymbol{D}$ and assigns random element values (possibly, but not necessarily stable) to $\boldsymbol{A}$, $\boldsymbol{B}$, and $\boldsymbol{C}$.
+I will do this for 100000 replicates of species numbers from $n = 2$ to $n = 8$.
+
+```{r, echo = TRUE}
+random_comm <- function(dm = 2, iter = 10000){
+    stabilised_fastevo <- NULL;
+    unstabled_fastevo  <- NULL;
+    nn           <- dm * dm;
+    tot_stabled  <- 0;
+    tot_ustabled <- 0;
+    tot_eco_stab <- 0;
+    stabilised_fastevo <- NULL;
+    unstabled_fastevo  <- NULL;
+    while(iter > 0){
+        stabilised   <- 0;
+        unstabilised <- 0;
+        Adat <- runif(n = nn, min = -4, max = 4);
+        Bdat <- runif(n = nn, min = -4, max = 4);
+        Cdat <- runif(n = nn, min = -4, max = 4);
+        A    <- matrix(data = Adat, nrow = dm, ncol = dm);
+        B    <- matrix(data = Bdat, nrow = dm, ncol = dm);
+        C    <- matrix(data = Cdat, nrow = dm, ncol = dm);
+        egD  <- 1;
+        while(egD >= 0){
+             Ddat <- runif(n = nn, min = -4, max = 4);
+             D    <- matrix(data = Ddat, nrow = dm, ncol = dm);
+             egD  <- max(Re(eigen(D)$values))
+        }
+        eco_stable       <- max(Re(eigen(A)$values)) < 0;
+        slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
+        stable_fast_evo  <- slow_evo_maxE < 0;
+        if(eco_stable == FALSE & stable_fast_evo == TRUE){
+            list_ele <- length(stabilised_fastevo$A) + 1;
+            stabilised_fastevo$A[[list_ele]] <- A;
+            stabilised_fastevo$B[[list_ele]] <- B;
+            stabilised_fastevo$C[[list_ele]] <- C;
+            stabilised_fastevo$D[[list_ele]] <- D;
+            stabilised   <- 1;
+        }
+        if(eco_stable == TRUE & stable_fast_evo == FALSE){
+            list_ele <- length(unstabled_fastevo$A) + 1;
+            unstabled_fastevo$A[[list_ele]] <- A;
+            unstabled_fastevo$B[[list_ele]] <- B;
+            unstabled_fastevo$C[[list_ele]] <- C;
+            unstabled_fastevo$D[[list_ele]] <- D;
+            unstabilised   <- 1;
+        }
+        if(stabilised == 1){
+            tot_stabled <- tot_stabled + 1;
+        }
+        if(unstabilised == 1){
+            tot_ustabled <- tot_ustabled + 1;
+        }
+        if(eco_stable == TRUE){
+            tot_eco_stab <- tot_eco_stab + 1;
+        }
+        iter <- iter - 1;
+    }
+    results <- list(ecologically_stable = tot_eco_stab, 
+                    stabilised = tot_stabled, destabilised = tot_ustabled,
+                    stabilised_all   = stabilised_fastevo, 
+                    destabilised_all = unstabled_fastevo);
+    return(results);
+}
+```
+
+Note that the code above takes a while to run for a large number of species, and this is mostly caused by having to find a random matrix $\boldsymbol{D}$ that is stable.
+The code first checks whether $\boldsymbol{A}$ is stable, and if it is stable, whether or not eco-evolutionary dynamics destabilise the stable ecological community.
+If it is not stable, then the code checks whether eco-evolutionary dynamics stabilise the unstable ecological community.
+As expected [@May1971], ecological stability decreases with an increasing number of species.
+
+```{r, echo = FALSE}
+dat       <- read.table(file = "rand_sims.txt", header = TRUE);
+species   <- dat[,1];
+pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
+par(mar = c(5, 5, 2, 1));
+plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
+     xlab = "Species number", ylab = "Proportion stable (n)",
+     cex.axis = 1.5, cex.lab = 1.5);
+```
+
+More interesting, the plot below shows eco-evolutionary results for ecological communities that **are** stable.
+The x-axis shows the number of species in the community (recall that each species has the same evolving trait potentially affecting ecology and evolution), and the y-axis shows the proportion of of these communities that are *destabilised* by eco-evolutionary dynamics.
+
+```{r, echo = FALSE}
+pr_destabilised <- dat[,3] / dat[,2];
+par(mar = c(5, 5, 2, 1));
+plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
+     xlab = "Species number", ylab = "Proportion destabilised",
+     cex.axis = 1.5, cex.lab = 1.5);
+```
+
+Note that the proportion of communities that are destabilised by eco-evolutionary effects increases with increasing species number.
+The plot below shows the eco-evolutionary results for ecological communities that are unstable.
+The y-axis here shows the proportion of communities that are *stabilised* by eco-evolutionary dynamics.
+This decreases with increasing species number.
+
+```{r, echo = FALSE}
+pr_stabilised <- dat[,5] / dat[,4];
+par(mar = c(5, 5, 2, 1));
+plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
+     xlab = "Species number", ylab = "Proportion stabilised",
+     cex.axis = 1.5, cex.lab = 1.5);
+```
+
+Hence, overall, **as species number and therefore evolving species traits increases, the probability that stable ecological communities become destabilised increases rapidly and the probability that unstable ecological communities are stabilised becomes vanishingly small**.
+I am not sure if I can prove this mathematically, but I don't think I need to for the point to be made effectively.
+What I want to figure out is whether or not this point can or should be published in the peer-reviwed literature -- else it's a blog post.
+
 Combining the two (work in progress)

### 2018-03-17 random_matrix_stability.Rmd
Rearrange and clean up abstract

@@
 
+<a name="gen_time_stability">The effect of variation in generation time on community stability</a>
+--------------------------------------------------------------------------------
+
 ********************************************************************************
@@
 **Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
-But in ecological communities, the rate of change of species density and the magnitude of interaction effects depend on species generation time.
-Per capita effects of individuals from species with short generation times will be lower than individuals from species with long generation times, and densities of species with shorter generation times will change more rapidly.
-Interaction effects are therefore expected to be correlated within species, with some species having consistently weaker or stronger effects on others per capita.
+But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
+Densities of species with relatively short generation times will change more rapidly than densities of species with longer generation times.
+Interaction effects are therefore expected to be correlated within species, with some species densities changing more rapidly than others in a community.
 Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
-Further, the influence of variation in generation time becomes exponentially more important as community size increases.
+The influence of variation in generation time becomes exponentially more important as community size increases.
+And communities are further stabilised when the interaction effects of long-lived species are especially strong.
 We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
@@
 ********************************************************************************
-
-<a name="gen_time_stability">The effect of variation in generation time on community stability</a>
---------------------------------------------------------------------------------
 
@@
 
-Hence, the goal here is to demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.
-
-The destabilising effect of rapid evolution
---------------------------------------------------------------------------------
+**Hence, the goal here is to demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.**
 
@@
 
-Combining the two (work in progress)
---------------------------------------------------------------------------------
-
-```{r, echo = FALSE}
-dat       <- read.table(file = "rand_sims_var_gen.txt", header = TRUE);
-species   <- dat[,1];
-pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stable (n)",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-```{r, echo = FALSE}
-pr_destabilised <- dat[,3] / dat[,2];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion destabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-```{r, echo = FALSE}
-pr_stabilised <- dat[,5] / dat[,4];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-Code to print the above is below.
-
-```{r, eval = FALSE, echo = TRUE}
-random_comm <- function(dm = 2, iter = 10000){
-  stabilised_fastevo <- NULL;
-  unstabled_fastevo  <- NULL;
-  nn           <- dm * dm;
-  tot_stabled  <- 0;
-  tot_ustabled <- 0;
-  tot_eco_stab <- 0;
-  stabilised_fastevo <- NULL;
-  unstabled_fastevo  <- NULL;
-  while(iter > 0){
-    epsil   <- runif(n = dm, min = 1, max = 1000);
-    eps_dat <- rep(x = epsil, times = dm);
-    eps_mat <- matrix(data = c(epsil, epsil), nrow = dm, ncol = dm, byrow = FALSE);
-    stabilised   <- 0;
-    unstabilised <- 0;
-    Adat <- runif(n = nn, min = -4, max = 4);
-    Bdat <- runif(n = nn, min = -4, max = 4);
-    Cdat <- runif(n = nn, min = -4, max = 4);
-    A    <- matrix(data = Adat, nrow = dm, ncol = dm) * eps_mat;
-    B    <- matrix(data = Bdat, nrow = dm, ncol = dm) * eps_mat;
-    C    <- matrix(data = Cdat, nrow = dm, ncol = dm) * eps_mat;
-    egD  <- 1;
-    while(egD >= 0){
-      Ddat <- runif(n = nn, min = -4, max = 4);
-      D    <- matrix(data = Ddat, nrow = dm, ncol = dm) * eps_mat;
-      egD  <- max(Re(eigen(D)$values))
-    }
-    eco_stable       <- max(Re(eigen(A)$values)) < 0;
-    slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
-    stable_fast_evo  <- slow_evo_maxE < 0;
-    if(eco_stable == FALSE & stable_fast_evo == TRUE){
-      list_ele <- length(stabilised_fastevo$A) + 1;
-      stabilised_fastevo$A[[list_ele]] <- A;
-      stabilised_fastevo$B[[list_ele]] <- B;
-      stabilised_fastevo$C[[list_ele]] <- C;
-      stabilised_fastevo$D[[list_ele]] <- D;
-      stabilised   <- 1;
-    }
-    if(eco_stable == TRUE & stable_fast_evo == FALSE){
-      list_ele <- length(unstabled_fastevo$A) + 1;
-      unstabled_fastevo$A[[list_ele]] <- A;
-      unstabled_fastevo$B[[list_ele]] <- B;
-      unstabled_fastevo$C[[list_ele]] <- C;
-      unstabled_fastevo$D[[list_ele]] <- D;
-      unstabilised   <- 1;
-    }
-    if(stabilised == 1){
-      tot_stabled <- tot_stabled + 1;
-    }
-    if(unstabilised == 1){
-      tot_ustabled <- tot_ustabled + 1;
-    }
-    if(eco_stable == TRUE){
-      tot_eco_stab <- tot_eco_stab + 1;
-    }
-    iter <- iter - 1;
-    if(iter %% 1000000 == 0){
-        print(c(dm, iter));
-    }
-  }
-  results <- list(ecologically_stable = tot_eco_stab, 
-                  stabilised = tot_stabled, destabilised = tot_ustabled,
-                  stabilised_all   = stabilised_fastevo, 
-                  destabilised_all = unstabled_fastevo);
-  return(results);
-}
-
-NN       <- 20;
-res_tabl <- matrix(data = 0, nrow = (NN-1), ncol = 5);
-its      <- 100000000;
-for(i in 2:NN){
-  res              <- random_comm(dm = i, iter = its); 
-  res_tabl[i-1, 1] <- i;                               # Species and traits
-  res_tabl[i-1, 2] <- res$ecologically_stable;         # Total stable
-  res_tabl[i-1, 3] <- res$destabilised;                # Total destabilised
-  res_tabl[i-1, 4] <- its - res_tabl[i-1, 2];          # Total unstable
-  res_tabl[i-1, 5] <- res$stabilised;                  # Total stabilised
-  print(i);
-}
-```
-
 References

### 2018-03-17 random_matrix_stability.Rmd
Updates and tweak wording

@@
 output:
-  html_document: default
-  pdf_document:
-    fig_caption: yes
   word_document:
@@
     reference_docx: docx_template.docx
+  pdf_document:
+    fig_caption: yes
+  html_document: default
 header-includes:
@@
 I believe that we could get [an article on the above](https://www.nature.com/natecolevol/about/content) clearly in the space of 3500, or maybe even 1500 words.
-As far as I've been able to find, such an article would be novel and interesting (and mathematically correct).
+As far as I've been able to find, such an article would be novel and interesting (and, I hope, mathematically correct).
 These notes are divided into two parts at the moment.
@@
 Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
-The influence of variation in generation time becomes exponentially more important as community size increases.
-And communities are further stabilised when the interaction effects of long-lived species are especially strong.
+Further, the influence of variation in generation time becomes exponentially more important as community size increases.
 We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
@@
 
-Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of the other.
-Yet I've not been able to find anything on this in the literature in a random matrix context.
+Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
+Yet I've not been able to find anything on this in the literature in a random matrix context (or any other context).
 I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**.
-This work is partly inspired by the use of $\varepsilon$ by @Patel2018 and partly by @Jirsa2004 to separate generation time of different species in a community.
+This work is partly inspired by the use of $\varepsilon$ by @Patel2018 to separate the time scale of ecological versus evolutionary dynamics, and partly inspired by @Jirsa2004 to separate generation time of different species in a community.
 
@@
 I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
-I assume that the response of a particular species' growth rate (rows in $\boldsymbol{A}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
+I assume that the effect that the density of any species $j$ ($N_{j}$) has on the growth rate of a focal species $i$ ($\partial N_{i}W_{i}/\partial N_{j}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
 Higher values of $\gamma_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
-The matrix below $\boldsymbol{A}$ for two species $i$ and $j$ with generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
+The matrix below $\boldsymbol{A}$ describes the dynamics for two species $i$ and $j$ with effects of generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
 
@@
 
-I will first allow random values of $\gamma_{i}$ to vary from $1$ to $100$, modelling two orders of magnitude of generation times.
-I will compare the stability of two sets of random community matrices -- one in which all $\gamma = 1$ and another in which $\gamma$ varies.
-
-```{r, echo = TRUE, eval = FALSE}
-nn     <- 8;
-A1_stt <- 0;
-A2_stt <- 0;
-iter   <- 100000;
-while(iter > 0){
-    A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
-    A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-    A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
-    A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
-    gamval  <- runif(n = nn, min = 1, max = 100);
-    eps_dat <- rep(x = gamval, times = nn);
-    eps_mat <- matrix(data = gamval, nrow = nn, ncol = nn, byrow = FALSE);
-    A2      <- A2 * eps_mat;
-    A1_stb  <- max(Re(eigen(A1)$values)) < 0;
-    A2_stb  <- max(Re(eigen(A2)$values)) < 0;
-    if(A1_stb == TRUE){
-      A1_stt <- A1_stt + 1;
-    }
-    if(A2_stb == TRUE){
-      A2_stt <- A2_stt + 1;
-    }
-    iter    <- iter - 1;
-}
-```
-
-The number of stable random matrices when $\gamma$ does not vary was $13$, and the number of stable random matrices when it does was $48$.
-To further investigate, it is worth looking at **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
-Below is some code to look at stability in random communities across different magnitudes of variance in generation time for communities with eight species.
+I will investigate **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
+Below is some code to look at stability in random communities with eight species across different magnitudes of variance in generation time.
+Random values of $\gamma_{i}$ are chosen for each species $i$ from a uniform distribution $\mathcal{U}\left(1, \max(\gamma_{i})\right)$.
 
@@
 
-The code below produces the following results showing the proportion of random communities that are stable as the variation in generation time increases in orders of magnitude (note that I'm using eight species, so the y-axis shows a very low proportion range).
+The code above produces the following results showing the proportion of random communities that are stable as $\max(\gamma_{i})$ increases in orders of magnitude (note that I'm using eight species, so the y-axis shows a very low proportion range).
 
@@
 plot(x = log10(mag_var[,1]), y = stabi_3, type = "b", lwd = 2, ylim = c(0, 0.0005),
-     xlab = "LOG Generation time variation among species", ylab = "Proportion stable",
+     xlab = expression(paste("LOG max(",gamma[i],")")), ylab = "Proportion stable",
      cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
@@
 
-**It looks like once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 20 for identical generation times versus magnitudes of generation time variation of 1000.
+**Once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 8 for identical generation times versus magnitudes of generation time variation of 1000.
 
@@
 
-The plot below shows the result for the proportion of stable communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
+The plot below shows the result for the proportion of stable random communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
 
@@
 par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = stabi_1, type = "b", lwd = 2, ylim = c(0, 1),
+plot(x = gen_var[,1], y = stabi_1, type = "b", lwd = 2, ylim = c(0, 0.3),
      xlab = "Species number", ylab = "Proportion stable",
      cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
-points(x = gen_var[,1], y = stabi_2, type = "b", lwd = 2, ylim = c(0, 1),
+points(x = gen_var[,1], y = stabi_2, type = "b", lwd = 2, ylim = c(0, 0.3),
        lty = "solid", pch = 20);
-legend(x = 2, y = 1, lwd = 2,
+legend(x = 4, y = 0.3, lwd = 2,
        legend = c("Generation time varies", "Generation time does not vary"),
@@
 
-This is interesting.
 **As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
-Assuming that I have done this correctly, this could be a fairly high impact result -- variation in generation time is important for community stability.
-I have not seen this take-home message in the literatur -- at least, not presented in this way.
-Again, I am not sure if any of this is something that I can prove mathematically, and I cannot decide whether the result is interesting enough to try to publish by itself or in tandem with the above.
-If not, it might be another blog post.
+In other words, all else being equal, variation in generation time becomes more important for stabilising system dynamics as the size of the system increases.
+
+********************************************************************************
 

### 2018-03-18 random_matrix_stability.Rmd
Tweak the method for getting data and graphs

@@
 
+<!---
+
+Switch code so that the same matrix is checked for stability before and after generation change is added.
+Then record number stabilised versus unstabilised.
+Show a plot with the proportion stabilised by varying generation time.
+
+```{r, echo = TRUE, eval = FALSE}
+tot_res <- NULL;
+res     <- NULL;
+for(i in 2:8){
+    nn             <- i;
+    A1_stt         <- 0;
+    A2_stt         <- 0;
+    iter           <- 1000000;
+    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+    while(iter > 0){
+        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        epsil   <- runif(n = nn, min = 1, max = 1000);
+        eps_dat <- rep(x = epsil, times = nn);
+        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
+        A2      <- A1 * eps_mat;
+        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
+        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
+        if(A1_stb == TRUE){
+          A1_stt                  <- A1_stt + 1;
+          tot_res[[i-1]][iter, 1] <- 1;
+        }
+        if(A2_stb == TRUE){
+          A2_stt                  <- A2_stt + 1;
+          tot_res[[i-1]][iter, 2] <- 1;
+        }
+        iter    <- iter - 1;
+    }
+    res <- rbind(res, c(A1_stt, A2_stt));
+    print(i);
+}
+```
+--->
+
 ********************************************************************************

### 2018-03-18 random_matrix_stability.Rmd
New figure and way of showing stuff

--- 2018-03-17 random_matrix_stability.Rmd
@@
 
+**A better way of doing this** is actually to check *the same* random matrices before and after the variation in generation time is added.
+Below is some code to do this.
+Unlike the above, the same matrix is initialised, and its stability is checked before and after the variation in generation time is added.
+Following 1000000 simulated matrices for each number of species, matrices are categorised in one of four ways: (1) Unstable before and after the addition of variance in generation time, (2) stable before and after the addition of variance in generation time, (3) stable before but not after the addition of variation in generation time (destabilised), and (4) unstable before but not after the addition of variation in generation time (stabilised).
+
+```{r, echo = TRUE, eval = FALSE}
+tot_res <- NULL;
+res     <- NULL;
+for(i in 2:8){
+    nn             <- i;
+    A1_stt         <- 0;
+    A2_stt         <- 0;
+    iter           <- 1000000;
+    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+    while(iter > 0){
+        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
+        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        epsil   <- runif(n = nn, min = 1, max = 1000);
+        eps_dat <- rep(x = epsil, times = nn);
+        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
+        A2      <- A1 * eps_mat;
+        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
+        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
+        if(A1_stb == TRUE){
+          A1_stt                  <- A1_stt + 1;
+          tot_res[[i-1]][iter, 1] <- 1;
+        }
+        if(A2_stb == TRUE){
+          A2_stt                  <- A2_stt + 1;
+          tot_res[[i-1]][iter, 2] <- 1;
+        }
+        iter    <- iter - 1;
+    }
+    res <- rbind(res, c(A1_stt, A2_stt));
+    print(i);
+}
+
+all_res <- matrix(data = 0, nrow = 7, ncol = 5);
+for(i in 1:7){
+    unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
+    stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
+    unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
+    stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
+    all_res[i, 1] <- i;
+    all_res[i, 2] <- sum(unstables);
+    all_res[i, 3] <- sum(stables);
+    all_res[i, 4] <- sum(unstabled);
+    all_res[i, 5] <- sum(stabled);
+}
+```
+
+Rather than show the ratio of stable to unstable matrices given variation in generation time, it is perhaps more intuitive to compare the proportion of random matrices stabilised versus the proportion of random matrices destabilised by varying generation time for different species numbers.
+The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
+
+```{r, echo = TRUE, eval = TRUE}
+gen_var     <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
+unstable    <- gen_var[,2] / 1000000;
+stable      <- gen_var[,3] / 1000000;
+destablised <- gen_var[,4] / 1000000;
+stabilised  <- gen_var[,5] / 1000000;
+par(mar = c(5, 5, 2, 1));
+plot(x = gen_var[,1], y = destablised, type = "b", lwd = 2, ylim = c(0, 0.02),
+     xlab = "Species number", ylab = "Proportion of communities",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
+polygon(x = c(gen_var[,1], rev(gen_var[,1]), -1), y = c(stable, rep(-1, 8)), 
+        col = "grey", border = NA);
+box();
+points(x = gen_var[,1], y = destablised, type = "b", lwd = 2, lty = "dashed");
+points(x = gen_var[,1], y = stabilised, type = "b", lwd = 2, ylim = c(0, 0.02),
+       lty = "solid", pch = 20);
+legend(x = 5.25, y = 0.02, lwd = 2, cex = 1.25,
+       legend = c("Stabilised", "Destabilised"),
+       lty = c("solid", "dashed"));
+```
+
 ********************************************************************************

### 2018-03-18 random_matrix_stability.Rmd
Better method add

@@
 output:
+  html_document: default
+  pdf_document:
+    fig_caption: yes
   word_document:
@@
     reference_docx: docx_template.docx
-  pdf_document:
-    fig_caption: yes
-  html_document: default
 header-includes:
@@
 The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
+Shading shows the proportion of communities that are expected to be stable in the absence of variation in generation time.
 
@@
      cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
-polygon(x = c(gen_var[,1], rev(gen_var[,1]), -1), y = c(stable, rep(-1, 8)), 
+polygon(x = c(gen_var[,1], rev(gen_var[,1]), -1), 
+        y = c(stable + destablised, rep(-1, 8)), 
         col = "grey", border = NA);
@@
 ```
+
+We can also look see this a bit more clearly when we specifically look at how variation in generation time changes random matrices across different species numbers.
+Specifically, the figure below shows only results for stable matrices.
+Points on the line show the proportion of random matrices that are stable *because of variation in generation time*.
+In other words, given that a random matrix is stable, what is the probability that it would be unstable if not for the effect of variation in generation time.
+
+```{r, echo = TRUE, eval = TRUE}
+par(mar = c(5, 5, 2, 1));
+y1 <- stabilised / (stable + stabilised);
+plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
+     xlab = "Species number", ylab = "Pr stable due to Var(gen)",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
+
+As the number of species increases, the probability that the whole community is stable because of variation in generation time (i.e., it would not be otherwise) increases.
+Likewise, we can look at the probability that if a whole community is unstable, it is unstable because of variation in generation time.
+*Note the y-axis difference*.
+
+```{r, echo = TRUE, eval = TRUE}
+par(mar = c(5, 5, 2, 1));
+y1 <- destablised / (unstable + destablised);
+plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 0.02),
+     xlab = "Species number", ylab = "Pr unstable due to Var(gen)",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
+
+Unstable communities are rarely unstable due to variation in generation time, and as the number of species increases, the probability that unstability will be caused by such variation becomes vanishingly small.
+**Hence, for large communities, variation in generation time is highly likely to be a critical reason for community stability**.
 

### 2018-03-19 random_matrix_stability.Rmd
Add the 20 species plots

@@
 gen_var     <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-unstable    <- gen_var[,2] / 1000000;
-stable      <- gen_var[,3] / 1000000;
-destablised <- gen_var[,4] / 1000000;
-stabilised  <- gen_var[,5] / 1000000;
+tot_sim     <- sum(gen_var[1,2:5]);
+unstable    <- gen_var[,2] / tot_sim;
+stable      <- gen_var[,3] / tot_sim;
+destablised <- gen_var[,4] / tot_sim;
+stabilised  <- gen_var[,5] / tot_sim;
 par(mar = c(5, 5, 2, 1));
@@
      cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
-polygon(x = c(gen_var[,1], rev(gen_var[,1]), -1), 
-        y = c(stable + destablised, rep(-1, 8)), 
+polygon(x = c(gen_var[,1], rev(gen_var[,1]), -4), 
+        y = c(stable + destablised, rep(-1, 20)), 
         col = "grey", border = NA);
@@
        lty = "solid", pch = 20);
-legend(x = 5.25, y = 0.02, lwd = 2, cex = 1.25,
+legend(x = 12, y = 0.02, lwd = 2, cex = 1.5,
        legend = c("Stabilised", "Destabilised"),

### 2018-03-20 random_matrix_stability.Rmd
Update figures to remove some species

--- 2018-03-18 random_matrix_stability.Rmd
@@
 gen_var     <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-unstable    <- gen_var[,2] / 1000000;
-stable      <- gen_var[,3] / 1000000;
-destablised <- gen_var[,4] / 1000000;
-stabilised  <- gen_var[,5] / 1000000;
+gen_var     <- gen_var[1:13,];
+tot_sim     <- sum(gen_var[1,2:5]);
+unstable    <- gen_var[,2] / tot_sim;
+stable      <- gen_var[,3] / tot_sim;
+destablised <- gen_var[,4] / tot_sim;
+stabilised  <- gen_var[,5] / tot_sim;
 par(mar = c(5, 5, 2, 1));
@@
      cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
-polygon(x = c(gen_var[,1], rev(gen_var[,1]), -1), 
-        y = c(stable + destablised, rep(-1, 8)), 
+polygon(x = c(gen_var[,1], rev(gen_var[,1]), -4), 
+        y = c(stable + destablised, rep(-1, dim(gen_var)[1] + 1)), 
         col = "grey", border = NA);
@@
        lty = "solid", pch = 20);
-legend(x = 5.25, y = 0.02, lwd = 2, cex = 1.25,
+legend(x = 9, y = 0.02, lwd = 2, cex = 1.5,
        legend = c("Stabilised", "Destabilised"),

### 2018-03-20 random_matrix_stability.Rmd
Consolidate some of the results

--- 2018-03-18 random_matrix_stability.Rmd
@@
 
-**Once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 8 for identical generation times versus magnitudes of generation time variation of 1000.
+**Once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 14 for identical generation times versus magnitudes of generation time variation of 1000.
 
-```{r, echo = TRUE, eval = FALSE}
-res <- NULL;
-for(i in 2:8){
-    nn     <- i;
-    A1_stt <- 0;
-    A2_stt <- 0;
-    iter   <- 1000000;
-    while(iter > 0){
-        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-        A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
-        epsil   <- runif(n = nn, min = 1, max = 1000);
-        eps_dat <- rep(x = epsil, times = nn);
-        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
-        A2      <- A2 * eps_mat;
-        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
-        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
-        if(A1_stb == TRUE){
-          A1_stt <- A1_stt + 1;
-        }
-        if(A2_stb == TRUE){
-          A2_stt <- A2_stt + 1;
-        }
-        iter    <- iter - 1;
-    }
-    res <- rbind(res, c(A1_stt, A2_stt));
-    print(i);
-}
-```
-
-The plot below shows the result for the proportion of stable random communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
-
-```{r, echo = TRUE, eval = TRUE}
-gen_var <- read.table(file = "dif_sp_gen_var.csv", header = TRUE, sep = ",");
-stabi_1 <- gen_var[,2] / 1000000;
-stabi_2 <- gen_var[,3] / 1000000;
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = stabi_1, type = "b", lwd = 2, ylim = c(0, 0.3),
-     xlab = "Species number", ylab = "Proportion stable",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
-points(x = gen_var[,1], y = stabi_2, type = "b", lwd = 2, ylim = c(0, 0.3),
-       lty = "solid", pch = 20);
-legend(x = 4, y = 0.3, lwd = 2,
-       legend = c("Generation time varies", "Generation time does not vary"),
-       lty = c("solid", "dotted"));
-```
-
-It is difficult to see the difference between these two numbers, so the figure below shows the ratio of the proportion of stable communities given generation variation relative to no generation variation (No_gen_var:Gen_var).
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = stabi_2 / stabi_1, type = "b", lwd = 2, ylim = c(0, 5),
-     xlab = "Species number", ylab = "Stable ratio (var to no var)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-```
-
-**As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
-In other words, all else being equal, variation in generation time becomes more important for stabilising system dynamics as the size of the system increases.
-
-**A better way of doing this** is actually to check *the same* random matrices before and after the variation in generation time is added.
-Below is some code to do this.
-Unlike the above, the same matrix is initialised, and its stability is checked before and after the variation in generation time is added.
+To investigate the effect of variation in generation time, I will check the stability of random matrices before and after the variation in generation time is added.
 Following 1000000 simulated matrices for each number of species, matrices are categorised in one of four ways: (1) Unstable before and after the addition of variance in generation time, (2) stable before and after the addition of variance in generation time, (3) stable before but not after the addition of variation in generation time (destabilised), and (4) unstable before but not after the addition of variation in generation time (stabilised).
@@
 res     <- NULL;
-for(i in 2:8){
+for(i in 2:14){
     nn             <- i;
@@
 
-all_res <- matrix(data = 0, nrow = 7, ncol = 5);
-for(i in 1:7){
+all_res <- matrix(data = 0, nrow = 13, ncol = 5);
+for(i in 1:19){
     unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
@@
 
-Rather than show the ratio of stable to unstable matrices given variation in generation time, it is perhaps more intuitive to compare the proportion of random matrices stabilised versus the proportion of random matrices destabilised by varying generation time for different species numbers.
-The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
+We can first show the stability of random matrices before and after variation in generation time is added.
+The plot below shows the result for the proportion of stable random communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
+
+```{r, echo = TRUE, eval = TRUE}
+gen_var  <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
+gen_var  <- gen_var[1:13,];
+tot_sim  <- sum(gen_var[1,2:5]);
+before_v <- gen_var[,3] / tot_sim;
+after_v  <- ((gen_var[,3] - gen_var[,4]) + gen_var[,5]) / tot_sim;
+par(mar = c(5, 5, 2, 1));
+plot(x = gen_var[,1], y = before_v, type = "b", lwd = 2, ylim = c(0, 0.3),
+     xlab = "Species number", ylab = "Proportion stable",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
+points(x = gen_var[,1], y = after_v, type = "b", lwd = 2, ylim = c(0, 0.3),
+       lty = "solid", pch = 20);
+legend(x = 4, y = 0.3, lwd = 2,
+       legend = c("Generation time varies", "Generation time does not vary"),
+       lty = c("solid", "dotted"));
+```
+
+It is difficult to see the difference between these two numbers, so the figure below shows the ratio of the proportion of stable communities given generation variation relative to no generation variation (Gen_var : No_gen_var).
+
+```{r, echo = TRUE, eval = TRUE}
+par(mar = c(5, 5, 2, 1));
+plot(x = gen_var[,1], y = after_v / before_v, type = "b", lwd = 2, ylim = c(0, 20),
+     xlab = "Species number", ylab = "Stable ratio (var to no var)",
+     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
+
+**As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
+In other words, all else being equal, variation in generation time becomes more important for stabilising system dynamics as the size of the system increases.
+
+Rather than show the ratio of stable to unstable matrices given variation in generation time, **it is perhaps more intuitive to compare the proportion of random matrices stabilised versus the proportion of random matrices destabilised by varying generation time for different species numbers.** The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
 Shading shows the proportion of communities that are expected to be stable in the absence of variation in generation time.
@@
 ```{r, echo = TRUE, eval = TRUE}
-gen_var     <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-unstable    <- gen_var[,2] / 1000000;
-stable      <- gen_var[,3] / 1000000;
-destablised <- gen_var[,4] / 1000000;
-stabilised  <- gen_var[,5] / 1000000;
+unstable    <- gen_var[,2] / tot_sim;
+stable      <- gen_var[,3] / tot_sim;
+destablised <- gen_var[,4] / tot_sim;
+stabilised  <- gen_var[,5] / tot_sim;
 par(mar = c(5, 5, 2, 1));
@@
 Points on the line show the proportion of random matrices that are stable *because of variation in generation time*.
-In other words, given that a random matrix is stable, what is the probability that it would be unstable if not for the effect of variation in generation time.
+In other words, given that a random matrix is stable, what is the probability that it would be unstable if not for the effect of variation in generation time?
 

### 2018-03-20 random_matrix_stability.Rmd
Obvious code fix

@@
 res     <- NULL;
-for(i in 2:14){
+for(i in 2:15){
     nn             <- i;
@@
 
-all_res <- matrix(data = 0, nrow = 13, ncol = 5);
-for(i in 1:19){
+all_res <- matrix(data = 0, nrow = 14, ncol = 5);
+for(i in 1:14){
     unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;

### 2018-03-22 random_matrix_stability.Rmd
Small tweaks and recompile

@@
 
-The below is a draft introductory paragraph or abstract.
-I believe that we could get [an article on the above](https://www.nature.com/natecolevol/about/content) clearly in the space of 3500, or maybe even 1500 words.
-As far as I've been able to find, such an article would be novel and interesting (and, I hope, mathematically correct).
-These notes are divided into two parts at the moment.
+These notes are divided into three parts at the moment.
+Part 1 is an initial start to a [brief communcation](https://www.nature.com/natecolevol/about/content) for Nature Ecology and Evolution.
+My goal, at the moment, is to write something short and to the point; the point is something fairly simple, and I think can be expressed in 1500 words.
+As far as I've been able to find, such an letter would be novel and interesting (and, I hope, mathematically correct).
+The most likely group that would have discovered this already would be the [Allesina lab](https://allesinalab.uchicago.edu/), which works a lot with random matrices, but I've not seen anything about generation time variation published by them.
 
-1. [The effect of variation in generation time on community stability.](#gen_time_stability)
-2. [The effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel)
+1. [Variation in generation time drives stability in large random communities](#working_draft)
+2. [The effect of variation in generation time on community stability.](#gen_time_stability)
+3. [The effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel)
 
-I suspect that part 2 will need to branch off soon, but I don't want to forget about it.
+Part 2 includes a lot of the messy notes used to get part 1, and might eventually be developed into supporting information.
+I suspect that part 3 will need to branch off soon, as the take-home message of it is quite different, but I don't want to forget about it.
+
+<a name = "working_draft">Variation in generation time drives stability in large random communities</a>
+
+********************************************************************************
+
+**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
+But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
+Here we show that variation in generation time becomes increasingly important for community stabilisation with increasing community size, and is therefore a potentially key contributor to the stability of large ecosystems.**
+
+********************************************************************************
+
+Coexistence of species rich ecological communities in nature remains a long-standing puzzle in ecology [@Hutchinson1961; @May1972; @Allesina2012; @Grilli2017].
+
+Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
 
@@
 par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = after_v / before_v, type = "b", lwd = 2, ylim = c(0, 20),
+plot(x = gen_var[,1], y = after_v / before_v, type = "b", lwd = 2, ylim = c(0, 40),
      xlab = "Species number", ylab = "Stable ratio (var to no var)",

### 2018-03-22 random_matrix_stability.Rmd
New first paragraph

@@
 
-Coexistence of species rich ecological communities in nature remains a long-standing puzzle in ecology [@Hutchinson1961; @May1972; @Allesina2012; @Grilli2017].
-
-Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
+Understanding the stability and therefore persistence of species rich ecological communities in nature remains a long-standing puzzle in ecology [@Hutchinson1961; @May1972; @Allesina2012; @Grilli2017].
+The general conditions under which such communities return to equilibrium when perturbed (i.e., are stabilised) are well-established, as is the knowledge that stabilisation becomes increasingly unlikely as community size increases [@May1972].
+Recently, analyses of random communities have investigated the effects of community structure, finding, e.g., that predator-prey interactions [@Allesina2012], increasing numbers of species interaction partners [@Allesina2015], correlations between interaction strengths [@Tang2014a], and higher order species interactions can be stabilising [@Allesina2011; @Levine2017].
+All such investigations assume that the rates at which interacting species affect one another are uniform.
+But generation times can vary by orders of magnitude between interacting species, causing response time in species dynamics to vary following perturbation.
 

### 2018-03-23 random_matrix_stability.Rmd
Updating table

@@
 
+********************************************************************************
+
 <a name = "working_draft">Variation in generation time drives stability in large random communities</a>
+================================================================================
+
+```{r, echo = FALSE}
+library(knitr);
+dat  <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
+dat  <- cbind(dat, (dat[,3] - dat[,4]) + dat[,5]);
+rownames(dat) <- NULL;
+colnames(dat) <- c("Species ($\\{S}$)", "Unstable", "Stable", "Stabilised", 
+                   "Destabilised", "Stable ($\\{Var(gamma)}$)")
+```
 
@@
 Understanding the stability and therefore persistence of species rich ecological communities in nature remains a long-standing puzzle in ecology [@Hutchinson1961; @May1972; @Allesina2012; @Grilli2017].
-The general conditions under which such communities return to equilibrium when perturbed (i.e., are stabilised) are well-established, as is the knowledge that stabilisation becomes increasingly unlikely as community size increases [@May1972].
+The general conditions under which such communities return to equilibrium densities when perturbed (i.e., are stabilised) are well-established, as is the knowledge that stabilisation becomes increasingly unlikely as community size increases [@May1972].
 Recently, analyses of random communities have investigated the effects of community structure, finding, e.g., that predator-prey interactions [@Allesina2012], increasing numbers of species interaction partners [@Allesina2015], correlations between interaction strengths [@Tang2014a], and higher order species interactions can be stabilising [@Allesina2011; @Levine2017].
@@
 But generation times can vary by orders of magnitude between interacting species, causing response time in species dynamics to vary following perturbation.
+We find that this variation becomes increasingly more likely to stabilise rather than destabilise simulated communities as species number ($S$) increases, and that the probability that community stability is solely caused by variation in generation time is near unity for randomly assembled communities of $S > 10$.
+
+In communities of interacting species, the density of any species $j$ ($N_{j}$) can affect species $i$'s mean fitness ($W_{i}$), and therefore its population growth rate [$\partial N_{i}W_{i}/\partial N_{j}$; @Patel2018].
+
+**References**
+
+<div id="refs"></div>
+
+********************************************************************************
+
+```{r, echo = FALSE}
+kable(dat);
+```
+
+**Table 1:** Stability results from 100 million simulations of random interaction matrices for 2-15 species ($S$).
+Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\gamma = 1$).
+Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $Var(\gamma)$, respectively.
+Column 6 reports the total number of communities that are stable given $Var(\gamma)$ (compare to column 3).
+
+| Species ($S$) | Unstable  | Stable   | Stabilised   | Destabilised | Stable ($Var(\gamma)$) |
+|---------------|-----------|----------|--------------|--------------|------------------------|
+| 2             | 73788328  | 23789690 | 1210228      | 1211754      | 23788164               |
+| 3             | 87850079  | 9449619  | 1613996      | 1086306      | 9977309                |
+| 4             | 94978045  | 3153446  | 1224826      | 643683       | 3734589                |
+| 5             | 98173012  | 878074   | 662411       | 286503       | 1253982                |
+| 6             | 99415760  | 204708   | 279491       | 100041       | 384158                 |
+| 7             | 99837320  | 39806    | 95472        | 27402        | 107876                 |
+| 8             | 99959849  | 6356     | 27660        | 6135         | 27881                  |
+| 9             | 99991252  | 887      | 6693         | 1168         | 6412                   |
+| 10            | 99998284  | 106      | 1439         | 171          | 1374                   |
+| 11            | 99999719  | 10       | 255          | 16           | 249                    |
+| 12            | 99999951  | 0        | 46           | 3            | 43                     |
+| 13            | 99999994  | 0        | 5            | 1            | 4                      |
+| 14            | 99999998  | 0        | 2            | 0            | 2                      |
+| 15            | 100000000 | 0        | 0            | 0            | 0                      |
+
+********************************************************************************
+
+**Figure 1:**
+
+********************************************************************************
 
@@
     A2_stt         <- 0;
-    iter           <- 1000000;
+    iter           <- 100000000;
     tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
@@
 --------------------------------------------------------------------------------
+
+<div id="refs"></div>

### 2018-03-23 random_matrix_stability.Rmd
Work on figure

@@
 library(knitr);
-dat  <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-dat  <- cbind(dat, (dat[,3] - dat[,4]) + dat[,5]);
+dat  <- read.table(file = "mat_dat.csv", header = TRUE, sep = ",");
+dat  <- cbind(dat, (dat[,3] - dat[,5]) + dat[,4]);
 rownames(dat) <- NULL;
-colnames(dat) <- c("Species ($\\{S}$)", "Unstable", "Stable", "Stabilised", 
-                   "Destabilised", "Stable ($\\{Var(gamma)}$)")
+colnames(dat) <- c("Species (S)", "Unstable", "Stable", "Stabilised", 
+                   "Destabilised", "Stable ($\\ Var(\u03b3)$ )")
 ```
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, caption = "Stability results from 100 million simulations of random interaction matrices for 2-15 species (S).
+Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\\u03b3 = 1$).
+Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $\\ Var(\u03b3)$, respectively.
+Column 6 reports the total number of communities that are stable given $\\ Var(\u03b3)$ (compare to column 3)."}
 kable(dat);
@@
 
-| Species ($S$) | Unstable  | Stable   | Stabilised   | Destabilised | Stable ($Var(\gamma)$) |
-|---------------|-----------|----------|--------------|--------------|------------------------|
-| 2             | 73788328  | 23789690 | 1210228      | 1211754      | 23788164               |
-| 3             | 87850079  | 9449619  | 1613996      | 1086306      | 9977309                |
-| 4             | 94978045  | 3153446  | 1224826      | 643683       | 3734589                |
-| 5             | 98173012  | 878074   | 662411       | 286503       | 1253982                |
-| 6             | 99415760  | 204708   | 279491       | 100041       | 384158                 |
-| 7             | 99837320  | 39806    | 95472        | 27402        | 107876                 |
-| 8             | 99959849  | 6356     | 27660        | 6135         | 27881                  |
-| 9             | 99991252  | 887      | 6693         | 1168         | 6412                   |
-| 10            | 99998284  | 106      | 1439         | 171          | 1374                   |
-| 11            | 99999719  | 10       | 255          | 16           | 249                    |
-| 12            | 99999951  | 0        | 46           | 3            | 43                     |
-| 13            | 99999994  | 0        | 5            | 1            | 4                      |
-| 14            | 99999998  | 0        | 2            | 0            | 2                      |
-| 15            | 100000000 | 0        | 0            | 0            | 0                      |
+********************************************************************************
 
-********************************************************************************
+```{r, echo = FALSE}
+bar_dat                      <- t(cbind(dat[,3], dat[,6]));
+log_bar_dat                  <- log(bar_dat);
+log_bar_dat[log_bar_dat < 0] <- 0;
+barplot(log_bar_dat, beside = TRUE, col = c("grey60", "grey80"),
+        names.arg = dat[,1], ylim = c(0, 25), xlab = "Species number (S)",
+        ylab = "Ln number of stable communities", cex.lab = 1.25, 
+        cex.axis = 1.25, xlim = c(1, 42));
+legend("topleft", c("\u03b3 = 1", "\u03b3 = [1, 1000]"), pch=15, 
+       col=c("grey60","grey80"), cex = 1.5);
+box(lwd = 2);
+par(new = TRUE);
+y1     <- dat[,4] / (dat[,3] + dat[,4]);
+x1     <- 2:15;
+x1_adj <- c(0.27, 0.25, 0.23, 0.21, 0.21, 0.20, 0.19, 0.18, 0.17, 0.16, 0.15,
+            0.14, 0.13, 0.12);
+plot(x = x1 + x1_adj, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+     xlab = "", ylab = "", type = "b", xlim = c(2, 15));
+
+#plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
+#     ylab = "Pr. stable due to Var(\u03b3)",
+#     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+```
 
@@
      cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+print(stabilised);
+print(stable);
+print(y1);
 ```

### 2018-03-24 random_matrix_stability.Rmd
Get Fig 1 right

@@
 log_bar_dat[log_bar_dat < 0] <- 0;
+par(oma = c(4, 4, 1, 4), mar = c(0.5, 0.5, 0.5, 0.5), lwd = 1);
 barplot(log_bar_dat, beside = TRUE, col = c("grey60", "grey80"),
@@
         ylab = "Ln number of stable communities", cex.lab = 1.25, 
-        cex.axis = 1.25, xlim = c(1, 42));
+        cex.axis = 1.25, xlim = c(1, 42), cex.names = 1.25);
 legend("topleft", c("\u03b3 = 1", "\u03b3 = [1, 1000]"), pch=15, 
@@
 y1     <- dat[,4] / (dat[,3] + dat[,4]);
-x1     <- 2:15;
-x1_adj <- c(0.27, 0.25, 0.23, 0.21, 0.21, 0.20, 0.19, 0.18, 0.17, 0.16, 0.15,
-            0.14, 0.13, 0.12);
-plot(x = x1 + x1_adj, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
-     xlab = "", ylab = "", type = "b", xlim = c(2, 15));
-
-#plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
-#     ylab = "Pr. stable due to Var(\u03b3)",
-#     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+x1     <- seq(from = 2.315, length = 14, by = 0.95)
+plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+     xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, cex = 1.25);
+axis(side = 4, at = seq(from = 0, to = 1, by = 0.2), cex.axis = 1.25);
+mtext(expression(paste("Species number (S)")), outer = TRUE, side = 1,
+      line = 2, cex = 1.25);
+mtext(expression(paste("Ln number of stable communities")), outer = TRUE, 
+      side = 2, line = 2, cex = 1.25);
+mtext(expression(paste("Pr. stable due to Var(",gamma,")")), outer = TRUE, 
+      side = 4, line = 2.2, cex = 1.25);
 ```

### 2018-03-24 random_matrix_stability.Rmd
Another paragraph

--- 2018-03-23 random_matrix_stability.Rmd
@@
 Recently, analyses of random communities have investigated the effects of community structure, finding, e.g., that predator-prey interactions [@Allesina2012], increasing numbers of species interaction partners [@Allesina2015], correlations between interaction strengths [@Tang2014a], and higher order species interactions can be stabilising [@Allesina2011; @Levine2017].
-All such investigations assume that the rates at which interacting species affect one another are uniform.
+These investigations assume that the rates at which interacting species affect one another are uniform.
 But generation times can vary by orders of magnitude between interacting species, causing response time in species dynamics to vary following perturbation.
@@
 
-In communities of interacting species, the density of any species $j$ ($N_{j}$) can affect species $i$'s mean fitness ($W_{i}$), and therefore its population growth rate [$\partial N_{i}W_{i}/\partial N_{j}$; @Patel2018].
+In communities of interacting species, the density of any species $j$ ($N_{j}$) can affect species $i$'s mean fitness ($W_{i}$), and therefore its population growth rate [$N_{i}W_{i}$; @Patel2018].
+But the time scale on which species $i$'s growth rate changes ($\gamma_{i}$) in response to the density of another species will vary depending on species $i$'s life history.
+Short-lived species can experience tens to hundreds of generations during the same time scale that long-lived species experience a single generation.
+Such timescale differences can have profound consequences for ecological system dynamics [@Hastings2004].
+In ecological communities, species with short generation times can be expected to respond more quickly to changes in species densities (higher $\gamma_{i}$), while species with long generation times will respond consistently more slowly (lower $\gamma_{i}$).
+The relative speed at which the growth rate of species $i$ is affected by a change in the density species of $j$ can therefore be modulated by $\gamma_{i}$ and expressed as $\gamma_{i} \times \partial N_{i}W_{i}/\partial N_{j}$.
+For large communities, it is convenient to represent these interactions for all $i$ and $j$ pairs using a Jacobian matrix,
+
+$$\boldsymbol{A} = \begin{bmatrix}
+    \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \cdots & \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{S}}  \\
+    \vdots & \ddots & \vdots \\
+    \gamma_{S} \frac{\partial N_{S}W_{S}}{\partial N_{i}}  & \cdots & \gamma_{S} \frac{\partial N_{S}W_{S}}{\partial N_{S}}  \\
+\end{bmatrix}.$$
+
+A community is stable if the real components of all eigenvalues of $\boldsymbol{A}$ are negative.
+
+We examined the stability of 100 million random communities in which $S = [2, 15]$ before and after the addition of variatince in $\gamma_{i}$.
 
@@
         ylab = "Ln number of stable communities", cex.lab = 1.25, 
-        cex.axis = 1.25, xlim = c(1, 42));
+        cex.axis = 1.25, xlim = c(1, 42), cex.names = 1);
 legend("topleft", c("\u03b3 = 1", "\u03b3 = [1, 1000]"), pch=15, 
@@
 y1     <- dat[,4] / (dat[,3] + dat[,4]);
-x1     <- 2:15;
-x1_adj <- c(0.27, 0.25, 0.23, 0.21, 0.21, 0.20, 0.19, 0.18, 0.17, 0.16, 0.15,
-            0.14, 0.13, 0.12);
-plot(x = x1 + x1_adj, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
-     xlab = "", ylab = "", type = "b", xlim = c(2, 15));
-
-#plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
-#     ylab = "Pr. stable due to Var(\u03b3)",
-#     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
+x1     <- seq(from = 2.315, length = 14, by = 0.95)
+plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+     xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, cex = 1.25);
+axis(side = 4, at = seq(from = 0, to = 1, by = 0.2), cex.axis = 1.25);
+mtext(expression(paste("Species number (S)")), outer = TRUE, side = 1,
+      line = 2, cex = 1.25);
+mtext(expression(paste("Ln number of stable communities")), outer = TRUE, 
+      side = 2, line = 2, cex = 1.25);
+mtext(expression(paste("Pr. stable due to Var(",gamma,")")), outer = TRUE, 
+      side = 4, line = 2.2, cex = 1.25);
 ```
 
-**Figure 1:**
+**Figure 1:** Stability of random communities when generation time ($\gamma$) varies and when it does not for 2-15 species (x-axis).
+Bars compare the $\ln$ total number of communities found to be stable out of 100 million before (dark grey) and after (light grey) $Var(\gamma)$ is added to a community (left y-axis).
+Points show the proportion of these random communities in which stability is a direct consequence of $Var(\gamma)$.
 

### 2018-03-25 random_matrix_stability.Rmd
Another paragraph of results

@@
 
-We examined the stability of 100 million random communities in which $S = [2, 15]$ before and after the addition of variatince in $\gamma_{i}$.
+We examined the stability of 100 million random communities in which $S$ ranged between 2-15 before and after the addition of variation in $\gamma_{i}$.
+Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in \gamma_{i} was generated with a random uniform distribution of $\mathcal{U}\left(1, 1000\right)$ (see Supporting Information S1).
+Random communities were categorised in one of four ways (Table 1): (1) Unstable before and after the addition of variance in $\gamma_{i}$ (unstable), (2) stable before and after the addition of variance in $\gamma_{i}$ (stable), (3) unstable before but not after the addition of variation in $\gamma_{i}$ (stabilised), and (4) stable before but not after the addition of variation in $\gamma_{i}$ (destabilised; Table 1).
+
+As $S$ increases, so does the effect of variation in $\gamma_{i}$ on increasing system stability (Figure 1).
+At high $S$, nearly all stable communities were only stable due to the effect of variation in $\gamma_{i}$, which rarely destabilised species rich communities (Supporting Information S1).
+All else being equal, variation in generation time therefore becomes more influential in stabilising random communities as community size increases, meaning that variation in generation time might be an important contributor to stability in large communities.
 

### 2018-03-25 Patel_comment.Rmd
Patel comments initialise in second file

--- 2018-03-24 random_matrix_stability.Rmd
@@
 ---
-
-These notes are divided into three parts at the moment.
-Part 1 is an initial start to a [brief communcation](https://www.nature.com/natecolevol/about/content) for Nature Ecology and Evolution.
-My goal, at the moment, is to write something short and to the point; the point is something fairly simple, and I think can be expressed in 1500 words.
-As far as I've been able to find, such an letter would be novel and interesting (and, I hope, mathematically correct).
-The most likely group that would have discovered this already would be the [Allesina lab](https://allesinalab.uchicago.edu/), which works a lot with random matrices, but I've not seen anything about generation time variation published by them.
-
-1. [Variation in generation time drives stability in large random communities](#working_draft)
-2. [The effect of variation in generation time on community stability.](#gen_time_stability)
-3. [The effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel)
-
-Part 2 includes a lot of the messy notes used to get part 1, and might eventually be developed into supporting information.
-I suspect that part 3 will need to branch off soon, as the take-home message of it is quite different, but I don't want to forget about it.
-
-********************************************************************************
-
-<a name = "working_draft">Variation in generation time drives stability in large random communities</a>
-================================================================================
-
-```{r, echo = FALSE}
-library(knitr);
-dat  <- read.table(file = "mat_dat.csv", header = TRUE, sep = ",");
-dat  <- cbind(dat, (dat[,3] - dat[,5]) + dat[,4]);
-rownames(dat) <- NULL;
-colnames(dat) <- c("Species (S)", "Unstable", "Stable", "Stabilised", 
-                   "Destabilised", "Stable ($\\ Var(\u03b3)$ )")
-```
-
-********************************************************************************
-
-**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
-But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
-Here we show that variation in generation time becomes increasingly important for community stabilisation with increasing community size, and is therefore a potentially key contributor to the stability of large ecosystems.**
-
-********************************************************************************
-
-Understanding the stability and therefore persistence of species rich ecological communities in nature remains a long-standing puzzle in ecology [@Hutchinson1961; @May1972; @Allesina2012; @Grilli2017].
-The general conditions under which such communities return to equilibrium densities when perturbed (i.e., are stabilised) are well-established, as is the knowledge that stabilisation becomes increasingly unlikely as community size increases [@May1972].
-Recently, analyses of random communities have investigated the effects of community structure, finding, e.g., that predator-prey interactions [@Allesina2012], increasing numbers of species interaction partners [@Allesina2015], correlations between interaction strengths [@Tang2014a], and higher order species interactions can be stabilising [@Allesina2011; @Levine2017].
-All such investigations assume that the rates at which interacting species affect one another are uniform.
-But generation times can vary by orders of magnitude between interacting species, causing response time in species dynamics to vary following perturbation.
-We find that this variation becomes increasingly more likely to stabilise rather than destabilise simulated communities as species number ($S$) increases, and that the probability that community stability is solely caused by variation in generation time is near unity for randomly assembled communities of $S > 10$.
-
-In communities of interacting species, the density of any species $j$ ($N_{j}$) can affect species $i$'s mean fitness ($W_{i}$), and therefore its population growth rate [$\partial N_{i}W_{i}/\partial N_{j}$; @Patel2018].
-
-**References**
-
-<div id="refs"></div>
-
-********************************************************************************
-
-```{r, echo = FALSE, caption = "Stability results from 100 million simulations of random interaction matrices for 2-15 species (S).
-Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\\u03b3 = 1$).
-Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $\\ Var(\u03b3)$, respectively.
-Column 6 reports the total number of communities that are stable given $\\ Var(\u03b3)$ (compare to column 3)."}
-kable(dat);
-```
-
-**Table 1:** Stability results from 100 million simulations of random interaction matrices for 2-15 species ($S$).
-Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\gamma = 1$).
-Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $Var(\gamma)$, respectively.
-Column 6 reports the total number of communities that are stable given $Var(\gamma)$ (compare to column 3).
-
-********************************************************************************
-
-```{r, echo = FALSE}
-bar_dat                      <- t(cbind(dat[,3], dat[,6]));
-log_bar_dat                  <- log(bar_dat);
-log_bar_dat[log_bar_dat < 0] <- 0;
-par(oma = c(4, 4, 1, 4), mar = c(0.5, 0.5, 0.5, 0.5), lwd = 1);
-barplot(log_bar_dat, beside = TRUE, col = c("grey60", "grey80"),
-        names.arg = dat[,1], ylim = c(0, 25), xlab = "Species number (S)",
-        ylab = "Ln number of stable communities", cex.lab = 1.25, 
-        cex.axis = 1.25, xlim = c(1, 42), cex.names = 1.25);
-legend("topleft", c("\u03b3 = 1", "\u03b3 = [1, 1000]"), pch=15, 
-       col=c("grey60","grey80"), cex = 1.5);
-box(lwd = 2);
-par(new = TRUE);
-y1     <- dat[,4] / (dat[,3] + dat[,4]);
-x1     <- seq(from = 2.315, length = 14, by = 0.95)
-plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
-     xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, cex = 1.25);
-axis(side = 4, at = seq(from = 0, to = 1, by = 0.2), cex.axis = 1.25);
-mtext(expression(paste("Species number (S)")), outer = TRUE, side = 1,
-      line = 2, cex = 1.25);
-mtext(expression(paste("Ln number of stable communities")), outer = TRUE, 
-      side = 2, line = 2, cex = 1.25);
-mtext(expression(paste("Pr. stable due to Var(",gamma,")")), outer = TRUE, 
-      side = 4, line = 2.2, cex = 1.25);
-```
-
-**Figure 1:**
-
-********************************************************************************
-
-<a name="gen_time_stability">The effect of variation in generation time on community stability</a>
---------------------------------------------------------------------------------
-
-********************************************************************************
-
-**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
-But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
-Densities of species with relatively short generation times will change more rapidly than densities of species with longer generation times.
-Interaction effects are therefore expected to be correlated within species, with some species densities changing more rapidly than others in a community.
-Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
-Further, the influence of variation in generation time becomes exponentially more important as community size increases.
-We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
-
-********************************************************************************
-
-Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
-Yet I've not been able to find anything on this in the literature in a random matrix context (or any other context).
-I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**.
-This work is partly inspired by the use of $\varepsilon$ by @Patel2018 to separate the time scale of ecological versus evolutionary dynamics, and partly inspired by @Jirsa2004 to separate generation time of different species in a community.
-
-Here I will consider a Jacobian matrix of interacting species $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time.
-I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
-I assume that the effect that the density of any species $j$ ($N_{j}$) has on the growth rate of a focal species $i$ ($\partial N_{i}W_{i}/\partial N_{j}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
-Higher values of $\gamma_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
-The matrix below $\boldsymbol{A}$ describes the dynamics for two species $i$ and $j$ with effects of generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
-
-$$\boldsymbol{A} = \begin{bmatrix}
-    \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
-    \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
-\end{bmatrix}.$$
-
-I will investigate **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
-Below is some code to look at stability in random communities with eight species across different magnitudes of variance in generation time.
-Random values of $\gamma_{i}$ are chosen for each species $i$ from a uniform distribution $\mathcal{U}\left(1, \max(\gamma_{i})\right)$.
-
-```{r, echo = TRUE, eval = FALSE}
-res  <- NULL;
-mags <- 10^(0:5);
-for(i in mags){
-    nn     <- 8;
-    A1_stt <- 0;
-    A2_stt <- 0;
-    iter   <- 10000000;
-    while(iter > 0){
-        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-        A2_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A2      <- matrix(data = A2_dat, nrow = nn, ncol = nn);
-        gamval  <- runif(n = nn, min = 1, max = i);
-        gam_dat <- rep(x = gamval, times = nn);
-        gam_mat <- matrix(data = gam_dat, nrow = nn, ncol = nn, byrow = FALSE);
-        A2      <- A2 * gam_mat;
-        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
-        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
-        if(A1_stb == TRUE){
-          A1_stt <- A1_stt + 1;
-        }
-        if(A2_stb == TRUE){
-          A2_stt <- A2_stt + 1;
-        }
-        iter    <- iter - 1;
-    }
-    res <- rbind(res, c(A1_stt, A2_stt));
-    print(i);
-}
-```
-
-The code above produces the following results showing the proportion of random communities that are stable as $\max(\gamma_{i})$ increases in orders of magnitude (note that I'm using eight species, so the y-axis shows a very low proportion range).
-
-```{r, echo = TRUE, eval = TRUE}
-mag_var <- read.table(file = "var_gen_mag.csv", header = TRUE, sep = ",");
-stabi_3 <- mag_var[,3] / 10000000;
-par(mar = c(5, 5, 2, 1));
-plot(x = log10(mag_var[,1]), y = stabi_3, type = "b", lwd = 2, ylim = c(0, 0.0005),
-     xlab = expression(paste("LOG max(",gamma[i],")")), ylab = "Proportion stable",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-```
-
-**Once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 14 for identical generation times versus magnitudes of generation time variation of 1000.
-
-To investigate the effect of variation in generation time, I will check the stability of random matrices before and after the variation in generation time is added.
-Following 1000000 simulated matrices for each number of species, matrices are categorised in one of four ways: (1) Unstable before and after the addition of variance in generation time, (2) stable before and after the addition of variance in generation time, (3) stable before but not after the addition of variation in generation time (destabilised), and (4) unstable before but not after the addition of variation in generation time (stabilised).
-
-```{r, echo = TRUE, eval = FALSE}
-tot_res <- NULL;
-res     <- NULL;
-for(i in 2:15){
-    nn             <- i;
-    A1_stt         <- 0;
-    A2_stt         <- 0;
-    iter           <- 100000000;
-    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
-    while(iter > 0){
-        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-        epsil   <- runif(n = nn, min = 1, max = 1000);
-        eps_dat <- rep(x = epsil, times = nn);
-        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
-        A2      <- A1 * eps_mat;
-        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
-        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
-        if(A1_stb == TRUE){
-          A1_stt                  <- A1_stt + 1;
-          tot_res[[i-1]][iter, 1] <- 1;
-        }
-        if(A2_stb == TRUE){
-          A2_stt                  <- A2_stt + 1;
-          tot_res[[i-1]][iter, 2] <- 1;
-        }
-        iter    <- iter - 1;
-    }
-    res <- rbind(res, c(A1_stt, A2_stt));
-    print(i);
-}
-
-all_res <- matrix(data = 0, nrow = 14, ncol = 5);
-for(i in 1:14){
-    unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
-    stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
-    unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
-    stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
-    all_res[i, 1] <- i;
-    all_res[i, 2] <- sum(unstables);
-    all_res[i, 3] <- sum(stables);
-    all_res[i, 4] <- sum(unstabled);
-    all_res[i, 5] <- sum(stabled);
-}
-```
-
-We can first show the stability of random matrices before and after variation in generation time is added.
-The plot below shows the result for the proportion of stable random communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
-
-```{r, echo = TRUE, eval = TRUE}
-gen_var  <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-gen_var  <- gen_var[1:13,];
-tot_sim  <- sum(gen_var[1,2:5]);
-before_v <- gen_var[,3] / tot_sim;
-after_v  <- ((gen_var[,3] - gen_var[,4]) + gen_var[,5]) / tot_sim;
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = before_v, type = "b", lwd = 2, ylim = c(0, 0.3),
-     xlab = "Species number", ylab = "Proportion stable",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
-points(x = gen_var[,1], y = after_v, type = "b", lwd = 2, ylim = c(0, 0.3),
-       lty = "solid", pch = 20);
-legend(x = 4, y = 0.3, lwd = 2,
-       legend = c("Generation time varies", "Generation time does not vary"),
-       lty = c("solid", "dotted"));
-```
-
-It is difficult to see the difference between these two numbers, so the figure below shows the ratio of the proportion of stable communities given generation variation relative to no generation variation (Gen_var : No_gen_var).
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = after_v / before_v, type = "b", lwd = 2, ylim = c(0, 40),
-     xlab = "Species number", ylab = "Stable ratio (var to no var)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-```
-
-**As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
-In other words, all else being equal, variation in generation time becomes more important for stabilising system dynamics as the size of the system increases.
-
-Rather than show the ratio of stable to unstable matrices given variation in generation time, **it is perhaps more intuitive to compare the proportion of random matrices stabilised versus the proportion of random matrices destabilised by varying generation time for different species numbers.** The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
-Shading shows the proportion of communities that are expected to be stable in the absence of variation in generation time.
-
-```{r, echo = TRUE, eval = TRUE}
-unstable    <- gen_var[,2] / tot_sim;
-stable      <- gen_var[,3] / tot_sim;
-destablised <- gen_var[,4] / tot_sim;
-stabilised  <- gen_var[,5] / tot_sim;
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = destablised, type = "b", lwd = 2, ylim = c(0, 0.02),
-     xlab = "Species number", ylab = "Proportion of communities",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
-polygon(x = c(gen_var[,1], rev(gen_var[,1]), -4), 
-        y = c(stable + destablised, rep(-1, dim(gen_var)[1] + 1)), 
-        col = "grey", border = NA);
-box();
-points(x = gen_var[,1], y = destablised, type = "b", lwd = 2, lty = "dashed");
-points(x = gen_var[,1], y = stabilised, type = "b", lwd = 2, ylim = c(0, 0.02),
-       lty = "solid", pch = 20);
-legend(x = 9, y = 0.02, lwd = 2, cex = 1.5,
-       legend = c("Stabilised", "Destabilised"),
-       lty = c("solid", "dashed"));
-```
-
-We can also look see this a bit more clearly when we specifically look at how variation in generation time changes random matrices across different species numbers.
-Specifically, the figure below shows only results for stable matrices.
-Points on the line show the proportion of random matrices that are stable *because of variation in generation time*.
-In other words, given that a random matrix is stable, what is the probability that it would be unstable if not for the effect of variation in generation time?
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-y1 <- stabilised / (stable + stabilised);
-plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Pr stable due to Var(gen)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-print(stabilised);
-print(stable);
-print(y1);
-```
-
-As the number of species increases, the probability that the whole community is stable because of variation in generation time (i.e., it would not be otherwise) increases.
-Likewise, we can look at the probability that if a whole community is unstable, it is unstable because of variation in generation time.
-*Note the y-axis difference*.
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-y1 <- destablised / (unstable + destablised);
-plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 0.02),
-     xlab = "Species number", ylab = "Pr unstable due to Var(gen)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-```
-
-Unstable communities are rarely unstable due to variation in generation time, and as the number of species increases, the probability that unstability will be caused by such variation becomes vanishingly small.
-**Hence, for large communities, variation in generation time is highly likely to be a critical reason for community stability**.
-
-********************************************************************************
 
@@
 --------------------------------------------------------------------------------
-
-<div id="refs"></div>

### 2018-03-25 random_matrix_stability.Rmd
Clean file and split Patel comments

@@
 
-These notes are divided into three parts at the moment.
-Part 1 is an initial start to a [brief communcation](https://www.nature.com/natecolevol/about/content) for Nature Ecology and Evolution.
-My goal, at the moment, is to write something short and to the point; the point is something fairly simple, and I think can be expressed in 1500 words.
-As far as I've been able to find, such an letter would be novel and interesting (and, I hope, mathematically correct).
+**AUTHOR'S NOTE:** This is an initial draft of a potential [brief communcation](https://www.nature.com/natecolevol/about/content) for *Nature Ecology and Evolution*.
+I think that the take-home message can be expressed in 1500 words.
+As far as I've been able to find, such an letter would be novel and interesting.
 The most likely group that would have discovered this already would be the [Allesina lab](https://allesinalab.uchicago.edu/), which works a lot with random matrices, but I've not seen anything about generation time variation published by them.
-
-1. [Variation in generation time drives stability in large random communities](#working_draft)
-2. [The effect of variation in generation time on community stability.](#gen_time_stability)
-3. [The effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel)
-
-Part 2 includes a lot of the messy notes used to get part 1, and might eventually be developed into supporting information.
-I suspect that part 3 will need to branch off soon, as the take-home message of it is quite different, but I don't want to forget about it.
+In developing this theory, I also came up with some brief comments on [the effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel), making the simple point after reading @Patel2018 that ecological systems are increasingly likely to be destablised, and unstable ecological systems are decreasingly likely to be stabilised, at higher numbers of evolving species traits.
 
@@
 We examined the stability of 100 million random communities in which $S$ ranged between 2-15 before and after the addition of variation in $\gamma_{i}$.
-Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in \gamma_{i} was generated with a random uniform distribution of $\mathcal{U}\left(1, 1000\right)$ (see Supporting Information S1).
+Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in \gamma_{i} was generated with a random uniform distribution of $\mathcal{U}\left(1, 1000\right)$ (see [Supporting Information S1](#SI_1)).
 Random communities were categorised in one of four ways (Table 1): (1) Unstable before and after the addition of variance in $\gamma_{i}$ (unstable), (2) stable before and after the addition of variance in $\gamma_{i}$ (stable), (3) unstable before but not after the addition of variation in $\gamma_{i}$ (stabilised), and (4) stable before but not after the addition of variation in $\gamma_{i}$ (destabilised; Table 1).
@@
 
-<a name="gen_time_stability">The effect of variation in generation time on community stability</a>
+<a name="gen_time_stability">Variation in generation time drives stability in large random communities: Supporting Information S1</a>
 --------------------------------------------------------------------------------
@@
 
-**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
-But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
-Densities of species with relatively short generation times will change more rapidly than densities of species with longer generation times.
-Interaction effects are therefore expected to be correlated within species, with some species densities changing more rapidly than others in a community.
-Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
-Further, the influence of variation in generation time becomes exponentially more important as community size increases.
-We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
+1. [**Effect of different variation in $\gamma$**](#SI_1)
+2. [**Destabilisation with increasing $S$**](#SI_2)
+3. [**Code underlying random matrix simulation**](#SI_3)
 
@@
 
-Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
-Yet I've not been able to find anything on this in the literature in a random matrix context (or any other context).
-I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**.
-This work is partly inspired by the use of $\varepsilon$ by @Patel2018 to separate the time scale of ecological versus evolutionary dynamics, and partly inspired by @Jirsa2004 to separate generation time of different species in a community.
+<a name="SI_1">**Effect of different variation in $\gamma$**</a>
 
-Here I will consider a Jacobian matrix of interacting species $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time.
-I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
-I assume that the effect that the density of any species $j$ ($N_{j}$) has on the growth rate of a focal species $i$ ($\partial N_{i}W_{i}/\partial N_{j}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
-Higher values of $\gamma_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
-The matrix below $\boldsymbol{A}$ describes the dynamics for two species $i$ and $j$ with effects of generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
+Below shows how the amount of variation in generation time (range of $\gamma_{i}$) in a community affects stability.
+Random communities of $S=8$ species are simulated 1 million times across different magnitudes of variance in $\gamma_{i}$.
+Random values of $\gamma_{i}$ are chosen for each species $i$ from a uniform distribution $\mathcal{U}\left(1, \max(\gamma_{i})\right)$, where $\max(\gamma_{i})$ ranges from 1 to 100000.
 
-$$\boldsymbol{A} = \begin{bmatrix}
-    \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
-    \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
-\end{bmatrix}.$$
-
-I will investigate **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
-Below is some code to look at stability in random communities with eight species across different magnitudes of variance in generation time.
-Random values of $\gamma_{i}$ are chosen for each species $i$ from a uniform distribution $\mathcal{U}\left(1, \max(\gamma_{i})\right)$.
-
-```{r, echo = TRUE, eval = FALSE}
+```{r, echo = FALSE, eval = FALSE}
 res  <- NULL;
@@
 ```
-
-The code above produces the following results showing the proportion of random communities that are stable as $\max(\gamma_{i})$ increases in orders of magnitude (note that I'm using eight species, so the y-axis shows a very low proportion range).
-
-```{r, echo = TRUE, eval = TRUE}
+```{r, echo = FALSE, eval = TRUE}
 mag_var <- read.table(file = "var_gen_mag.csv", header = TRUE, sep = ",");
@@
 
-**Once the generation time varies by an order of magnitude of around 1000, there is little return in terms of stability.** To continue, we can simulate different sizes of communities with species numbers between 2 and 14 for identical generation times versus magnitudes of generation time variation of 1000.
+Figure S1-1: The proportion of random $S=8$ communities that are stable as $\max(\gamma_{i})$ increases over 5 orders of magnitude.
+Each point represents a proportion calculated from 1 million random simulated matrices.
 
-To investigate the effect of variation in generation time, I will check the stability of random matrices before and after the variation in generation time is added.
-Following 1000000 simulated matrices for each number of species, matrices are categorised in one of four ways: (1) Unstable before and after the addition of variance in generation time, (2) stable before and after the addition of variance in generation time, (3) stable before but not after the addition of variation in generation time (destabilised), and (4) unstable before but not after the addition of variation in generation time (stabilised).
+Once the generation time varies by an order of magnitude of around 1000, there is little increase in terms of stability.
+Magnitudes higher than $\gamma_{i} = 1000$ are also likely to be biologically unrealistic for most communities.
 
-```{r, echo = TRUE, eval = FALSE}
-tot_res <- NULL;
-res     <- NULL;
-for(i in 2:15){
-    nn             <- i;
-    A1_stt         <- 0;
-    A2_stt         <- 0;
-    iter           <- 100000000;
-    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
-    while(iter > 0){
-        A1_dat  <- runif(n = nn * nn, min = -4, max = 4);
-        A1      <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-        epsil   <- runif(n = nn, min = 1, max = 1000);
-        eps_dat <- rep(x = epsil, times = nn);
-        eps_mat <- matrix(data = epsil, nrow = nn, ncol = nn, byrow = FALSE);
-        A2      <- A1 * eps_mat;
-        A1_stb  <- max(Re(eigen(A1)$values)) < 0;
-        A2_stb  <- max(Re(eigen(A2)$values)) < 0;
-        if(A1_stb == TRUE){
-          A1_stt                  <- A1_stt + 1;
-          tot_res[[i-1]][iter, 1] <- 1;
-        }
-        if(A2_stb == TRUE){
-          A2_stt                  <- A2_stt + 1;
-          tot_res[[i-1]][iter, 2] <- 1;
-        }
-        iter    <- iter - 1;
-    }
-    res <- rbind(res, c(A1_stt, A2_stt));
-    print(i);
-}
+********************************************************************************
 
-all_res <- matrix(data = 0, nrow = 14, ncol = 5);
-for(i in 1:14){
-    unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
-    stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
-    unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
-    stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
-    all_res[i, 1] <- i;
-    all_res[i, 2] <- sum(unstables);
-    all_res[i, 3] <- sum(stables);
-    all_res[i, 4] <- sum(unstabled);
-    all_res[i, 5] <- sum(stabled);
-}
-```
+<a name="SI_2">**Destabilisation with increasing $S$**</a>
 
-We can first show the stability of random matrices before and after variation in generation time is added.
-The plot below shows the result for the proportion of stable random communities when generation time varies (solid line, closed points) versus when generation time varies over two orders of magnitude (dotted line, open points).
+Here we use the data from Table 1 in the main text to show that the proportion of already stable communities that are destabilised by the addition of variance in $\gamma_{i}$ decreases with increasing $S$.
 
-```{r, echo = TRUE, eval = TRUE}
-gen_var  <- read.table(file = "full_data.csv", header = TRUE, sep = ",");
-gen_var  <- gen_var[1:13,];
-tot_sim  <- sum(gen_var[1,2:5]);
-before_v <- gen_var[,3] / tot_sim;
-after_v  <- ((gen_var[,3] - gen_var[,4]) + gen_var[,5]) / tot_sim;
+```{r, echo = FALSE, eval = TRUE}
 par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = before_v, type = "b", lwd = 2, ylim = c(0, 0.3),
-     xlab = "Species number", ylab = "Proportion stable",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "dotted");
-points(x = gen_var[,1], y = after_v, type = "b", lwd = 2, ylim = c(0, 0.3),
-       lty = "solid", pch = 20);
-legend(x = 4, y = 0.3, lwd = 2,
-       legend = c("Generation time varies", "Generation time does not vary"),
-       lty = c("solid", "dotted"));
-```
-
-It is difficult to see the difference between these two numbers, so the figure below shows the ratio of the proportion of stable communities given generation variation relative to no generation variation (Gen_var : No_gen_var).
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = after_v / before_v, type = "b", lwd = 2, ylim = c(0, 40),
-     xlab = "Species number", ylab = "Stable ratio (var to no var)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-```
-
-**As species number increases, variation in generation time appears to have a stronger influence in stabilising random communities**.
-In other words, all else being equal, variation in generation time becomes more important for stabilising system dynamics as the size of the system increases.
-
-Rather than show the ratio of stable to unstable matrices given variation in generation time, **it is perhaps more intuitive to compare the proportion of random matrices stabilised versus the proportion of random matrices destabilised by varying generation time for different species numbers.** The figure below shows the proportion of matrices stabilised (solid lines) and destabilised (dashed lines) by variation in generation time.
-Shading shows the proportion of communities that are expected to be stable in the absence of variation in generation time.
-
-```{r, echo = TRUE, eval = TRUE}
-unstable    <- gen_var[,2] / tot_sim;
-stable      <- gen_var[,3] / tot_sim;
-destablised <- gen_var[,4] / tot_sim;
-stabilised  <- gen_var[,5] / tot_sim;
-par(mar = c(5, 5, 2, 1));
-plot(x = gen_var[,1], y = destablised, type = "b", lwd = 2, ylim = c(0, 0.02),
-     xlab = "Species number", ylab = "Proportion of communities",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "dashed");
-polygon(x = c(gen_var[,1], rev(gen_var[,1]), -4), 
-        y = c(stable + destablised, rep(-1, dim(gen_var)[1] + 1)), 
-        col = "grey", border = NA);
-box();
-points(x = gen_var[,1], y = destablised, type = "b", lwd = 2, lty = "dashed");
-points(x = gen_var[,1], y = stabilised, type = "b", lwd = 2, ylim = c(0, 0.02),
-       lty = "solid", pch = 20);
-legend(x = 9, y = 0.02, lwd = 2, cex = 1.5,
-       legend = c("Stabilised", "Destabilised"),
-       lty = c("solid", "dashed"));
-```
-
-We can also look see this a bit more clearly when we specifically look at how variation in generation time changes random matrices across different species numbers.
-Specifically, the figure below shows only results for stable matrices.
-Points on the line show the proportion of random matrices that are stable *because of variation in generation time*.
-In other words, given that a random matrix is stable, what is the probability that it would be unstable if not for the effect of variation in generation time?
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-y1 <- stabilised / (stable + stabilised);
-plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Pr stable due to Var(gen)",
-     cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
-print(stabilised);
-print(stable);
-print(y1);
-```
-
-As the number of species increases, the probability that the whole community is stable because of variation in generation time (i.e., it would not be otherwise) increases.
-Likewise, we can look at the probability that if a whole community is unstable, it is unstable because of variation in generation time.
-*Note the y-axis difference*.
-
-```{r, echo = TRUE, eval = TRUE}
-par(mar = c(5, 5, 2, 1));
-y1 <- destablised / (unstable + destablised);
-plot(x = gen_var[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 0.02),
-     xlab = "Species number", ylab = "Pr unstable due to Var(gen)",
+y1 <- dat[,5] / (dat[,2] + dat[,5]);
+plot(x = dat[,1], y = y1, type = "b", lwd = 2, ylim = c(0, 0.02),
+     xlab = "Species number (S)", 
+     ylab = expression(paste("Pr. unstable due to Var(",gamma,")")),
      cex.axis = 1.5, cex.lab = 1.5, lty = "solid");
@@
 Unstable communities are rarely unstable due to variation in generation time, and as the number of species increases, the probability that unstability will be caused by such variation becomes vanishingly small.
-**Hence, for large communities, variation in generation time is highly likely to be a critical reason for community stability**.
+Hence, for large communities, variation in generation time might be critical for community stability.
 
@@
 
-<a name="comment_on_Patel">The effect of increasing numbers of evolutionary traits on eco-evolutionary stability</a>
---------------------------------------------------------------------------------
+<a name="SI_3">**Code underlying random matrix simulation**</a>
 
-Recently, @Patel2018 demonstrated the stability criteria for eco-evolutionary systems.
-A key result was that an unstable (stable) ecological community could be stabilised (destabilised) by fast eco-evolutionary feedback, assuming the evolutionary system was itself stable.
-Likewise, @Patel2018 showed that an unstable (stable) evolutionary system could be stabilised (destabilised) by fast ecological feedback, assuming the ecological community was itself stable.
-**Here I am going to show why the stability criteria for eco-evolutionary systems is such that stable ecological systems are *increasingly* likely to be destablised, and unstable ecological systems are *decreasingly* likely to be stabilised, at higher numbers of species traits**.
-I will use a random matrix approach similar to that of Allesina and colleagues [@Allesina2011; @Allesina2012; @Allesina2015a].
-The mathematical concept of what I describe below is not new, and is a similar point to what @May1971 explained nearly half a century ago (that increasing species number decreases stability).
-Still, this point was not, that I can find, noticed or acknowledged by @Patel2018 when coming up with their stability criteria.
-
-**Hence, the goal here is to demonstrate the inherent destabilising effect of rapid evolution, which can also be interpreted as plasticity [@Patel2018], on ecological communities.**
-
-Assume a community of $n$ species, each with the same evolving trait (e.g., body size).
-Each species $i$ therefore has its own associated trait $m_{i}$ (originally I was thinking about parental investment -- I thought that I had come up with something really interesting, then realised that @Parker1986 got there first).
-We can describe the dynamics of a system with two species $i$ and $j$ using the Jacobian matrix $\boldsymbol{J}$ below [@Patel2018],
-
-$$\boldsymbol{J} = \begin{bmatrix}
-    \boldsymbol{A}  & \boldsymbol{B}  \\
-    \boldsymbol{C}  & \boldsymbol{D}  \\  \end{bmatrix}
-    = \begin{bmatrix}
-    \frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \frac{\partial N_{i}W_{i}}{\partial N_{j}} & \frac{\partial N_{i}W_{i}}{\partial m_{i}} & \frac{\partial N_{i}W_{i}}{\partial m_{j}}  \\
-    \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \frac{\partial N_{j}W_{j}}{\partial N_{j}} & \frac{\partial N_{j}W_{j}}{\partial m_{i}} & \frac{\partial N_{j}W_{j}}{\partial m_{j}} \\
-    \varepsilon\frac{\partial m_{i}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{i}}{\partial N_{j}} & \varepsilon\frac{\partial m_{i}}{\partial m_{i}} & \varepsilon\frac{\partial m_{i}}{\partial m_{j}}  \\
-    \varepsilon\frac{\partial m_{j}}{\partial N_{i}}  & \varepsilon\frac{\partial m_{j}}{\partial N_{j}} & \varepsilon\frac{\partial m_{j}}{\partial m_{i}} & \varepsilon\frac{\partial m_{j}}{\partial m_{j}}  \\
-\end{bmatrix}.$$
-
-Note that the matrix $\boldsymbol{J}$ can be separated into four separate submatrices.
-The submatrix $\boldsymbol{A}$ describes the effects that the density of one species has on its own population growth (diagonal) or that of another species (off-diagonal).
-The submatrix $\boldsymbol{B}$ describes the effect the trait $m$ of each species has on the population growth of itself or another species.
-The submatrix $\boldsymbol{C}$ describes the effect that the density of a species has on a species' changing $m$, and $\boldsymbol{D}$ describes the effect that traits have on traits.
-The value of $\varepsilon$ defines the rate of evolutionary change relative to ecological change; here I will just assume a high value of $\varepsilon$, but the conclusions regarding an increasing number of traits' effect on a relatively slowly changing ecoogical system will be equally applicable to an incresing number of species in relativley slowly evolving system.
-I will therefore only assume a high value of $\varepsilon$ for now.
-
-For the whole system (ecology and evolution) to be stable when traits change quickly relative to population densities, evolution by itself $\boldsymbol{D}$ must be stable.
-This is the case when all of the eigenvalues of $\boldsymbol{D}$ are negative.
-Additionally, for the system $\boldsymbol{J}$ as a whole to be stable, we need to consider the feedback between ecology and evolution.
-@Patel2018 showed that the system as a whole is stable when the following is satisfied,
-
-$$s\left(\boldsymbol{A} + \boldsymbol{B}\boldsymbol{D}^{-1}(-\boldsymbol{C})\right) < 0.$$
-
-I am specifically interested in the case in which $\boldsymbol{D}$ is known to be stable.
-Using random matrices containing element values that are real numbers uniformly distributed between $-4$ and $4$, I will show that rapid evolution is almost never stabilising and almost always destabilising when over six species and their traits $m$ are considered.
-To do this, I will use the following code, which finds a stable random matrix $\boldsymbol{D}$ and assigns random element values (possibly, but not necessarily stable) to $\boldsymbol{A}$, $\boldsymbol{B}$, and $\boldsymbol{C}$.
-I will do this for 100000 replicates of species numbers from $n = 2$ to $n = 8$.
-
-```{r, echo = TRUE}
-random_comm <- function(dm = 2, iter = 10000){
-    stabilised_fastevo <- NULL;
-    unstabled_fastevo  <- NULL;
-    nn           <- dm * dm;
-    tot_stabled  <- 0;
-    tot_ustabled <- 0;
-    tot_eco_stab <- 0;
-    stabilised_fastevo <- NULL;
-    unstabled_fastevo  <- NULL;
-    while(iter > 0){
-        stabilised   <- 0;
-        unstabilised <- 0;
-        Adat <- runif(n = nn, min = -4, max = 4);
-        Bdat <- runif(n = nn, min = -4, max = 4);
-        Cdat <- runif(n = nn, min = -4, max = 4);
-        A    <- matrix(data = Adat, nrow = dm, ncol = dm);
-        B    <- matrix(data = Bdat, nrow = dm, ncol = dm);
-        C    <- matrix(data = Cdat, nrow = dm, ncol = dm);
-        egD  <- 1;
-        while(egD >= 0){
-             Ddat <- runif(n = nn, min = -4, max = 4);
-             D    <- matrix(data = Ddat, nrow = dm, ncol = dm);
-             egD  <- max(Re(eigen(D)$values))
-        }
-        eco_stable       <- max(Re(eigen(A)$values)) < 0;
-        slow_evo_maxE    <- max(Re(eigen(A + B %*% solve(D) %*% (-C))$values));
-        stable_fast_evo  <- slow_evo_maxE < 0;
-        if(eco_stable == FALSE & stable_fast_evo == TRUE){
-            list_ele <- length(stabilised_fastevo$A) + 1;
-            stabilised_fastevo$A[[list_ele]] <- A;
-            stabilised_fastevo$B[[list_ele]] <- B;
-            stabilised_fastevo$C[[list_ele]] <- C;
-            stabilised_fastevo$D[[list_ele]] <- D;
-            stabilised   <- 1;
-        }
-        if(eco_stable == TRUE & stable_fast_evo == FALSE){
-            list_ele <- length(unstabled_fastevo$A) + 1;
-            unstabled_fastevo$A[[list_ele]] <- A;
-            unstabled_fastevo$B[[list_ele]] <- B;
-            unstabled_fastevo$C[[list_ele]] <- C;
-            unstabled_fastevo$D[[list_ele]] <- D;
-            unstabilised   <- 1;
-        }
-        if(stabilised == 1){
-            tot_stabled <- tot_stabled + 1;
-        }
-        if(unstabilised == 1){
-            tot_ustabled <- tot_ustabled + 1;
-        }
-        if(eco_stable == TRUE){
-            tot_eco_stab <- tot_eco_stab + 1;
-        }
-        iter <- iter - 1;
-    }
-    results <- list(ecologically_stable = tot_eco_stab, 
-                    stabilised = tot_stabled, destabilised = tot_ustabled,
-                    stabilised_all   = stabilised_fastevo, 
-                    destabilised_all = unstabled_fastevo);
-    return(results);
-}
+```{r}
+# Will add code here.
 ```
 
-Note that the code above takes a while to run for a large number of species, and this is mostly caused by having to find a random matrix $\boldsymbol{D}$ that is stable.
-The code first checks whether $\boldsymbol{A}$ is stable, and if it is stable, whether or not eco-evolutionary dynamics destabilise the stable ecological community.
-If it is not stable, then the code checks whether eco-evolutionary dynamics stabilise the unstable ecological community.
-As expected [@May1971], ecological stability decreases with an increasing number of species.
-
-```{r, echo = FALSE}
-dat       <- read.table(file = "rand_sims.txt", header = TRUE);
-species   <- dat[,1];
-pr_stable <- dat[,2] / (dat[,2] + dat[,4]);
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stable, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stable (n)",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-More interesting, the plot below shows eco-evolutionary results for ecological communities that **are** stable.
-The x-axis shows the number of species in the community (recall that each species has the same evolving trait potentially affecting ecology and evolution), and the y-axis shows the proportion of of these communities that are *destabilised* by eco-evolutionary dynamics.
-
-```{r, echo = FALSE}
-pr_destabilised <- dat[,3] / dat[,2];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_destabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion destabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-Note that the proportion of communities that are destabilised by eco-evolutionary effects increases with increasing species number.
-The plot below shows the eco-evolutionary results for ecological communities that are unstable.
-The y-axis here shows the proportion of communities that are *stabilised* by eco-evolutionary dynamics.
-This decreases with increasing species number.
-
-```{r, echo = FALSE}
-pr_stabilised <- dat[,5] / dat[,4];
-par(mar = c(5, 5, 2, 1));
-plot(x = species, y = pr_stabilised, type = "b", lwd = 2, ylim = c(0, 1),
-     xlab = "Species number", ylab = "Proportion stabilised",
-     cex.axis = 1.5, cex.lab = 1.5);
-```
-
-Hence, overall, **as species number and therefore evolving species traits increases, the probability that stable ecological communities become destabilised increases rapidly and the probability that unstable ecological communities are stabilised becomes vanishingly small**.
-I am not sure if I can prove this mathematically, but I don't think I need to for the point to be made effectively.
-What I want to figure out is whether or not this point can or should be published in the peer-reviwed literature -- else it's a blog post.
-
-References
---------------------------------------------------------------------------------
-
-<div id="refs"></div>
+********************************************************************************

### 2018-03-25 random_matrix_stability.Rmd
Minor corrections

@@
 We examined the stability of 100 million random communities in which $S$ ranged between 2-15 before and after the addition of variation in $\gamma_{i}$.
-Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in \gamma_{i} was generated with a random uniform distribution of $\mathcal{U}\left(1, 1000\right)$ (see [Supporting Information S1](#SI_1)).
+Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in $\gamma_{i}$ was generated with a random uniform distribution such that $\gamma_{i} = \mathcal{U}\left(1, 1000\right)$ (see [Supporting Information S1](#SI_1)).
 Random communities were categorised in one of four ways (Table 1): (1) Unstable before and after the addition of variance in $\gamma_{i}$ (unstable), (2) stable before and after the addition of variance in $\gamma_{i}$ (stable), (3) unstable before but not after the addition of variation in $\gamma_{i}$ (stabilised), and (4) stable before but not after the addition of variation in $\gamma_{i}$ (destabilised; Table 1).
 
-As $S$ increases, so does the effect of variation in $\gamma_{i}$ on increasing system stability (Figure 1).
-At high $S$, nearly all stable communities were only stable due to the effect of variation in $\gamma_{i}$, which rarely destabilised species rich communities (Supporting Information S1).
+We found that as $S$ increases, so does the effect of variation in $\gamma_{i}$ on increasing system stability (Figure 1).
+At high $S$, nearly all stable communities were only stable due to the effect of variation in $\gamma_{i}$, which rarely destabilised species rich communities ([Supporting Information S1](#SI_2)).
 All else being equal, variation in generation time therefore becomes more influential in stabilising random communities as community size increases, meaning that variation in generation time might be an important contributor to stability in large communities.

### 2018-03-25 random_matrix_stability.Rmd
More on author's note

@@
 In developing this theory, I also came up with some brief comments on [the effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel), making the simple point after reading @Patel2018 that ecological systems are increasingly likely to be destablised, and unstable ecological systems are decreasingly likely to be stabilised, at higher numbers of evolving species traits.
+
+For the final paper, I am strongly considering changing the simulations a bit to consider 3-5 scenarios.
+This isn't going to change anything about the general conclusions, I don't think.
+What it will do is make the simulations more consistent with what others have tried in the past:
+
+1.
+Random network but with elements drawn from a normal distribution with N(0, 0.4) and the diagonal elements of the matrix being forced at -1 (helping out stability a bit -- since in all stable matrices, diagonal elements must be negative).
+2. Competitive network with elements drawn from N(0, 0.4) and diagonals at -1, but when sampling form the normal distribution returns positive values, the sign is reversed so all values are negative.
+3. Mutualist network created as above, but with all signs forced to be positive (again, except the diagonal).
+4. Predator-prey network created as above, but forcing the sign of a_ij to be the opposite of a_ji.
+
+If the below conclusions hold for all of the above, then I think the results can be made even stronger.
+The best to present would probably be 1, with 2-4 mentioned and then shown in the supporting information.
 
@@
 Elements of $\boldsymbol{A}$ were selected from a random uniform distribution $\mathcal{U}\left(-4, 4\right)$, and variation in $\gamma_{i}$ was generated with a random uniform distribution such that $\gamma_{i} = \mathcal{U}\left(1, 1000\right)$ (see [Supporting Information S1](#SI_1)).
+For simplicity, we assumed complete connectance of simulated communities (i.e., elements of $\boldsymbol{A}$ were allowed to be nonzero with a probability of $C = 1$), but this did not effect our general conclusions.
 Random communities were categorised in one of four ways (Table 1): (1) Unstable before and after the addition of variance in $\gamma_{i}$ (unstable), (2) stable before and after the addition of variance in $\gamma_{i}$ (stable), (3) unstable before but not after the addition of variation in $\gamma_{i}$ (stabilised), and (4) stable before but not after the addition of variation in $\gamma_{i}$ (destabilised; Table 1).
@@
 All else being equal, variation in generation time therefore becomes more influential in stabilising random communities as community size increases, meaning that variation in generation time might be an important contributor to stability in large communities.
+
+More broadly, individual processes within complex dynamic systems might respond relatively slowly or quickly to perturbations in the system.
+We show that accounting for this variation becomes more important with increasing system size.
 
@@
 # Will add code here.
+# For consistency with Dougoud, maybe run some simulations under their parameter
+# values:
+# C = 1, elements = N(0, 0.4), and theta = -1 (diag).
+# Can then force competitive (folded normal to all negative values), or
+# mutualistic networks -- alongside the random ones. Could do a predator-prey
+# one as well forcing matrix signs to differ.
 ```

### 2018-03-25 random_matrix_stability.Rmd
Author's note on feasibility criteria

@@
 The best to present would probably be 1, with 2-4 mentioned and then shown in the supporting information.
+It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable.
+Feasibility criteria are as follows,
+
+$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.
+The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive).
+The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have).
+The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0).
+And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off diagonal effecting species interactions,
+
+$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
+
+To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
+So really, we're just multiplying $(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
+This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
+
+Hence, the below is going to be refined quite a bit, but I highly doubt any of the major conclusions will change.
 

### 2018-03-27 random_matrix_stability.Rmd
Fix sign
New plots

@@
 To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
-So really, we're just multiplying $(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
+So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
 This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
 
-Hence, the below is going to be refined quite a bit, but I highly doubt any of the major conclusions will change.
+A lot is going to change.
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+random      <- read.csv(file = "sim_results/random_community.csv");
+competition <- read.csv(file = "sim_results/competitor_community.csv");
+mutualism   <- read.csv(file = "sim_results/mutualist_community.csv");
+pred_prey   <- read.csv(file = "sim_results/predator_prey_community.csv");
+# ==============================================================================
+par(mfrow = c(2, 2));
+# ==============================================================================
+dat      <- random;
+Sp       <- dat[,1];
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible");
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+legend(x = 40, y = 0.9, col = c("black", "black", "red", "red"), cex = 1.2,
+       legend = c(expression(paste("Stable, no Var(",gamma,")")),
+                             expression(paste("Stable, Var(",gamma,")")),
+                             expression(paste("Feasible, no Var(",gamma,")")),
+                             expression(paste("Feasible, Var(",gamma,")"))),
+       lty = c("solid", "dashed", "solid", "dashed"), lwd = 2);
+text(x = 60, y = 0.98, labels = "Random", cex = 2);
+# ==============================================================================
+dat      <- competition;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible");
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 60, y = 0.98, labels = "Competitive", cex = 2);
+# ==============================================================================
+dat      <- mutualism;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible");
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 60, y = 0.98, labels = "Mutualist", cex = 2);
+# ==============================================================================
+dat      <- pred_prey;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible");
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 60, y = 0.98, labels = "Predator-prey", cex = 2);
+# ==============================================================================
+```
 

### 2018-03-27 test_mat.R
New plot and data read into

@@ -78,7 +78,7 @@ summarise_randmat <- function(tot_res, fea_res){
         feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
         unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
         feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
-        all_res[i, 1] <- i;
+        all_res[i, 1] <- i + 1;
         all_res[i, 2] <- sum(unstables);
         all_res[i, 3] <- sum(stables);
         all_res[i, 4] <- sum(unstabled);
@@ -90,6 +90,107 @@ summarise_randmat <- function(tot_res, fea_res){
     }
     return(all_res);
 }
+
+
+# Get:
+# (1) proportion stable made unstable
+# (2) proportion unstable made stable
+# (3) proportion feasible made infeasible
+# (4) proportion infeasible made feasible
+
+
+random      <- read.csv(file = "sim_results/random_community.csv");
+competition <- read.csv(file = "sim_results/competitor_community.csv");
+mutualism   <- read.csv(file = "sim_results/mutualist_community.csv");
+pred_prey   <- read.csv(file = "sim_results/predator_prey_community.csv");
+
+
+
+dat <- pred_prey;
+
+pr_destab <- dat[,4] / (dat[,4] + dat[,3]);
+pr_stabld <- dat[,5] / (dat[,5] + dat[,2]);
+pr_defeas <- dat[,8] / (dat[,8] + dat[,7]);
+pr_feased <- dat[,9] / (dat[,9] + dat[,6]);
+Sp        <- dat[,1];
+
+plot(dat[,1], pr_destab, type = "l", lwd = 2, ylim = c(0, 1));
+points(dat[,1], pr_stabld, type = "l", lwd = 2, col = "red");
+points(dat[,1], pr_defeas, type = "l", lwd = 2, col = "black", lty = "dashed");
+points(dat[,1], pr_feased, type = "l", lwd = 2, col = "red", lty = "dashed");
+
+
+
+# ==============================================================================
+par(mfrow = c(2, 2), mar = c(4, 4, 1, 1));
+# ==============================================================================
+dat      <- random;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
[…]

### 2018-03-28 test_mat.R
Add code to show eigen analysis

@@ -1,4 +1,4 @@
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 1000){
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 999){
     tot_res <- NULL;
     fea_res <- NULL;
     for(i in 2:max_sp){
@@ -20,7 +20,10 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 1000){
             eps_dat  <- rep(x = epsil, times = nn);
             eps_mat  <- matrix(data = epsil, nrow = nn, ncol = nn, 
                                byrow = FALSE);
+            avg_mat  <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                               byrow = FALSE);
             A2       <- A1 * eps_mat;
+            A1       <- A1 * avg_mat;
             A1_stb   <- max(Re(eigen(A1)$values)) < 0;
             A2_stb   <- max(Re(eigen(A2)$values)) < 0;
             A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
@@ -191,7 +194,160 @@ text(x = 60, y = 0.98, labels = "Predator-prey", cex = 2);
 
 
 
+mat <- matrix(data = rnorm(n = 4, mean = 0, sd = 0.4), nrow = 2);
+diag(mat) <- -1;
+plot(mat, xlim = c(-0.5, 0.5), ylim = c(-1.25, 1.25), asp = 1, pch = 20)
+points(Re(eigen(mat)$vector));
+abline(h = 0, lty = "dotted", lwd = 0.8);
+abline(v = 0, lty = "dotted", lwd = 0.8);
 
+
+
+
+
+
+plot(x = 0, y = 0, type = "n", xlim = c(-0.3, 0.3), ylim = c(-2.5, 2.5),
+     xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+abline(h = 0, lty = "dotted", lwd = 0.8);
+abline(v = 0, lty = "dotted", lwd = 0.8);
+
+iter     <- 10000;
+while(iter > 0){
+    mat1       <- matrix(data = rnorm(n = 16, mean = 0, sd = 0.4), nrow = 4);
+    diag(mat1) <- -1;
+    mat2       <- mat1;
+    mat2[1,]   <- mat2[1,] * runif(n = 1, min = 1, max = 1000);
[…]

### 2018-03-28 test_mat.R
More code for analysis

@@ -345,11 +345,68 @@ plot_eig_set <- function(eigres){
     text(x = x0 + (x1 - x0)*0.2, y = 1500, "Stabilised", cex = 1.5);
 }
 
+get_rand_feas <- function(int_type = 0, iter = 1000){
+    nn             <- 20;
+    eigres         <- matrix(data = 0, ncol = 42, nrow = iter);
+    eigres[,1]     <- int_type;
+    while(iter > 0){
+        r_vec     <- runif(n = nn, min = 1, max = rmx);
+        A1_dat    <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+        A1        <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A1        <- species_interactions(mat = A1, type = int_type);
+        diag(A1)  <- -1;
+        avg_mat   <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        A1        <- A1 * avg_mat;
+        mat1_e    <- eigen(A1)$values;
+        real_mat1 <- Re(mat1_e);
+        imag_mat1 <- Im(mat1_e);
+        stab1     <- max(real_mat1) < 0;
+        A1_fea    <- min(-1*solve(A1) %*% r_vec) > 0;
+        if(stab1 == TRUE & A1_fea == FALSE){
+            eigres[iter,2] <- 1;
+        }
+        if(stab1 == TRUE & A1_fea == TRUE){
+            eigres[iter,2] <- 2;
+        }
+        eigres[iter,3:22]  <- real_mat1;
+        eigres[iter,23:42] <- imag_mat1;
+        iter    <- iter - 1;
+    }
+    return(eigres);
+}
 
-eres <- get_rand_eigs(int_type = 1);
+
+plot_feas_set <- function(eigres){
+    par(mfrow = c(1,2), mar = c(5, 5, 1, 1));
+    x0 <- min(eigres[,3:22]);
+    x1 <- max(eigres[,3:22]);
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(1 %in% eigres[,2]){
[…]

### 2018-03-28 random_matrix_stability.Rmd
Code examples up

@@
 
-**AUTHOR'S NOTE:** This is an initial draft of a potential [brief communcation](https://www.nature.com/natecolevol/about/content) for *Nature Ecology and Evolution*.
-I think that the take-home message can be expressed in 1500 words.
-As far as I've been able to find, such an letter would be novel and interesting.
-The most likely group that would have discovered this already would be the [Allesina lab](https://allesinalab.uchicago.edu/), which works a lot with random matrices, but I've not seen anything about generation time variation published by them.
-In developing this theory, I also came up with some brief comments on [the effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel), making the simple point after reading @Patel2018 that ecological systems are increasingly likely to be destablised, and unstable ecological systems are decreasingly likely to be stabilised, at higher numbers of evolving species traits.
+********************************************************************************
 
-For the final paper, I am strongly considering changing the simulations a bit to consider 3-5 scenarios.
-This isn't going to change anything about the general conclusions, I don't think.
-What it will do is make the simulations more consistent with what others have tried in the past:
+**NEW AUTHORS NOTE:** After analysing the new results, I am thinking that a different angle is needed for a strong publication. To recap what has happened:
 
-1.
-Random network but with elements drawn from a normal distribution with N(0, 0.4) and the diagonal elements of the matrix being forced at -1 (helping out stability a bit -- since in all stable matrices, diagonal elements must be negative).
-2. Competitive network with elements drawn from N(0, 0.4) and diagonals at -1, but when sampling form the normal distribution returns positive values, the sign is reversed so all values are negative.
-3. Mutualist network created as above, but with all signs forced to be positive (again, except the diagonal).
-4. Predator-prey network created as above, but forcing the sign of a_ij to be the opposite of a_ji.
+- I started analysing random communities of simulated species using matrices and a mathematical trick similar to that used in @Patel2018 in attempt to see if variation in generation time among species will stabilise or destabilise large communities.
+- I initially found variation in generation time to be highly stabilising, and, in particular, as species richness increased, the overwhelming number of stable random communties were stable *due to variation in generation time* (and would otherwise be unstable).
+Results are summarised in [this figure](#old_fig_1) and [this table](#old_tab_1).
+- I then read the very recent paper by @Dougoud2018, who showed that the classic stability criteria used by @May1972 and @Allesina2012 (among others) have been flawed.
+Some "stable" communities in these studies are stable (i.e., return to equilibrium values if perturbed) at *negative* species densities.
+@Dougoud2018 demonstrate how to check whether or not random communities are "feasible", meaning that they are stable and have species densities that are all positive (and hence, sensible ecological communities).
+- I applied the criteria of @Dougoud2018 to my random communities under the same parameter values that they and others have used; I also looked at communities characterised only by competitive, mutualistic, and predator-prey interactions.
+- Strangely, variation in generation time has the exact *opposite* effect in random, competitive, and predator-prey communities; variation in generation time therefore *increases* stability but *decreases* feasibility (i.e., variation in generation time shifts communities toward being "stable" with negative species densities and away from being "stable" with positive species densities).
+The overall effect is therefore to make real ecological communities less stable.
+- I am still working out what this means conceptually by multiplying each row of the matrix by a single number, the shape of the matrix is somehow being changed to decrease the eigenvalues.
+- The exception to this is mutualistic communities -- variation in generation time had zero effect on mutualist community stability.
+- **I am tempted to broaden these results to make a more general point about complex dynamic systems**: "Variation in process time increases stability but decreases feasibility in complex systems", with a focus on biological systems.
+There might be some types of systems where negative values are actually meaningful (e.g., phenotypes in evolving systems, gene regulatory networks?
+Suggestions very welcome), just not in ecological ones where species densities obviously cannot be negative.
 
-If the below conclusions hold for all of the above, then I think the results can be made even stronger.
-The best to present would probably be 1, with 2-4 mentioned and then shown in the supporting information.
-It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable.
-Feasibility criteria are as follows,
-
-$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
-The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.
-The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive).
-The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have).
-The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0).
-And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off diagonal effecting species interactions,
-
-$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
-
-To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
-So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
-This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
-
-A lot is going to change.
+The plot below shows the story for the effect of variation in generation time on stability and feasibility.
 
@@
 
+And here's the same plot with the axes shifted a bit.
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+random      <- read.csv(file = "sim_results/random_community.csv");
+competition <- read.csv(file = "sim_results/competitor_community.csv");
+mutualism   <- read.csv(file = "sim_results/mutualist_community.csv");
+pred_prey   <- read.csv(file = "sim_results/predator_prey_community.csv");
+# ==============================================================================
+par(mfrow = c(2, 2));
+# ==============================================================================
+dat      <- random;
+Sp       <- dat[,1];
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible", xlim = c(0, 30));
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+legend(x = 12, y = 0.9, col = c("black", "black", "red", "red"), cex = 1.2,
+       legend = c(expression(paste("Stable, no Var(",gamma,")")),
+                             expression(paste("Stable, Var(",gamma,")")),
+                             expression(paste("Feasible, no Var(",gamma,")")),
+                             expression(paste("Feasible, Var(",gamma,")"))),
+       lty = c("solid", "dashed", "solid", "dashed"), lwd = 2);
+text(x = 17, y = 0.98, labels = "Random", cex = 2);
+# ==============================================================================
+dat      <- competition;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible", xlim = c(0, 30));
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 15, y = 0.98, labels = "Competitive", cex = 2);
+# ==============================================================================
+dat      <- mutualism;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible", xlim = c(0, 10));
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 6, y = 0.98, labels = "Mutualist", cex = 2);
+# ==============================================================================
+dat      <- pred_prey;
+stab_nov <- dat[,3] + dat[,4];
+stab_var <- dat[,3] + dat[,5];
+feas_nov <- dat[,7] + dat[,8];
+feas_var <- dat[,7] + dat[,9];
+plot(x = Sp, y = stab_nov / 100000, type = "l", lwd = 2, ylim = c(0,1),
+     cex.lab = 1.5, cex.axis = 1.5, xlab = "Species Number (S)",
+     ylab = "Proportion stable or feasible");
+points(x = Sp, y = stab_var / 100000, type = "l", lwd = 2, lty = "dashed");
+points(x = Sp, y = feas_nov / 100000, type = "l", lwd = 2, col = "red");
+points(x = Sp, y = feas_var / 100000, type = "l", lwd = 2, col = "red",
+       lty = "dashed");
+text(x = 60, y = 0.98, labels = "Predator-prey", cex = 2);
+# ==============================================================================
+```
+
+We can delve a bit deeper into this mess by looking at the distribution of eigenvalues in sample communities that are unstable, stable, destabilised, or stabilised.
+The plot below shows these examples for communities where species number $S = 20$.
+This gives the same type of plots as shown in @Allesina2012.
+Black points show eigenvalues before variation in generation time is added, and red points show eigenvalues after variation in generation time is added.
+The shift is not typically drastic, but sometimes causes the sign of the real components (x-axis) to flip, thereby potentially affecting stability.
+Immediately below shows plots for a random community; where not points are plot in any of the below, no examples exist.
+
+```{r, echo = FALSE}
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 999){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    for(i in 2:max_sp){
+        nn             <- i;
+        A1_stt         <- 0;
+        A2_stt         <- 0;
+        A1_fet         <- 0;
+        A2_fet         <- 0;
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        while(iter > 0){
+            r_vec    <- runif(n = nn, min = 1, max = rmx);
+            A1_dat   <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+            A1       <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+            A1       <- species_interactions(mat = A1, type = int_type);
+            diag(A1) <- -1;
+            epsil    <- runif(n = nn, min = 1, max = eps_max);
+            eps_dat  <- rep(x = epsil, times = nn);
+            eps_mat  <- matrix(data = epsil, nrow = nn, ncol = nn, 
+                               byrow = FALSE);
+            avg_mat  <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                               byrow = FALSE);
+            A2       <- A1 * eps_mat;
+            A1       <- A1 * avg_mat;
+            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
+            A2_stb   <- max(Re(eigen(A2)$values)) < 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
+            A2_fea   <- min(-1*solve(A2) %*% r_vec) > 0;
+            if(A1_stb == TRUE){
+                tot_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A2_stb == TRUE){
+                tot_res[[i-1]][iter, 2] <- 1;
+            }
+            if(A1_fea == TRUE){
+                fea_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A2_fea == TRUE){
+                fea_res[[i-1]][iter, 2] <- 1;
+            }
+            iter    <- iter - 1;
+        }
+        print(i);
+    }
+    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
+    return(all_res);
+}
+
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+
+summarise_randmat <- function(tot_res, fea_res){
+    sims    <- length(tot_res);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 9);
+    for(i in 1:sims){
+        unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
+        stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
+        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
+        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
+        non_feas  <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == FALSE;
+        feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
+        unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
+        feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
+        all_res[i, 1] <- i + 1;
+        all_res[i, 2] <- sum(unstables);
+        all_res[i, 3] <- sum(stables);
+        all_res[i, 4] <- sum(unstabled);
+        all_res[i, 5] <- sum(stabled);
+        all_res[i, 6] <- sum(non_feas);
+        all_res[i, 7] <- sum(feasibl);
+        all_res[i, 8] <- sum(unfeased);
+        all_res[i, 9] <- sum(feased);
+    }
+    return(all_res);
+}
+
+get_rand_eigs <- function(int_type = 0, iter = 1000){
+    nn             <- 20;
+    eigres         <- matrix(data = 0, ncol = 82, nrow = iter);
+    eigres[,1]     <- int_type;
+    while(iter > 0){
+        r_vec     <- runif(n = nn, min = 1, max = 1000);
+        A1_dat    <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+        A1        <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A1        <- species_interactions(mat = A1, type = int_type);
+        diag(A1)  <- -1;
+        epsil     <- runif(n = nn, min = 1, max = 999);
+        eps_dat   <- rep(x = epsil, times = nn);
+        eps_mat   <- matrix(data = epsil, nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        avg_mat   <- matrix(data = mean(1:999), nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        A2        <- A1 * eps_mat;
+        A1        <- A1 * avg_mat;
+        mat1_e    <- eigen(A1)$values;
+        mat2_e    <- eigen(A2)$values;
+        real_mat1 <- Re(mat1_e);
+        imag_mat1 <- Im(mat1_e);
+        real_mat2 <- Re(mat2_e);
+        imag_mat2 <- Im(mat2_e);
+        stab1     <- max(real_mat1) < 0;
+        stab2     <- max(real_mat2) < 0;
+        if(stab1 == FALSE & stab2 == FALSE){
+            eigres[iter,2] <- 1;
+        }
+        if(stab1 == TRUE & stab2 == TRUE){
+            eigres[iter,2] <- 2;
+        }
+        if(stab1 == TRUE & stab2 == FALSE){
+            eigres[iter,2] <- 3;
+        }
+        if(stab1 == FALSE & stab2 == TRUE){
+            eigres[iter,2] <- 4;
+        }
+        eigres[iter,3:22]  <- real_mat1;
+        eigres[iter,23:42] <- imag_mat1;
+        eigres[iter,43:62] <- real_mat2;
+        eigres[iter,63:82] <- imag_mat2;
+        iter    <- iter - 1;
+    }
+    return(eigres);
+}
+
+plot_eig_set <- function(eigres){
+    par(mfrow = c(2,2), mar = c(5, 5, 1, 1));
+    x0 <- min(cbind(eigres[,3:22], eigres[,43:62]));
+    x1 <- max(cbind(eigres[,3:22], eigres[,43:62]));
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(1 %in% eigres[,2]){
+        row <- which(eigres[,2] == 1)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        m2r <- eigres[row,43:62];
+        m2i <- eigres[row,63:82];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+        points(x = m2r, y = m2i, pch = 20, cex = 1.5, col = "red");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Unstable", cex = 1.5);
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(2 %in% eigres[,2]){
+        row <- which(eigres[,2] == 2)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        m2r <- eigres[row,43:62];
+        m2i <- eigres[row,63:82];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+        points(x = m2r, y = m2i, pch = 20, cex = 1.5, col = "red");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Stable", cex = 1.5);
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(3 %in% eigres[,2]){
+        row <- which(eigres[,2] == 3)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        m2r <- eigres[row,43:62];
+        m2i <- eigres[row,63:82];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+        points(x = m2r, y = m2i, pch = 20, cex = 1.5, col = "red");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Destabilised", cex = 1.5);
+    plot(x = 0, y = 0, type = "n", xlim =c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(4 %in% eigres[,2]){
+        row <- which(eigres[,2] == 4)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        m2r <- eigres[row,43:62];
+        m2i <- eigres[row,63:82];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+        points(x = m2r, y = m2i, pch = 20, cex = 1.5, col = "red");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Stabilised", cex = 1.5);
+}
+```
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+eres <- get_rand_eigs(int_type = 0, iter = 10000);
+plot_eig_set(eres);
+```
+
+Below shows plots for competitive communities.
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+eres <- get_rand_eigs(int_type = 1, iter = 10000);
+plot_eig_set(eres);
+```
+
+Below shows plots for mutualist communities.
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+eres <- get_rand_eigs(int_type = 2, iter = 10000);
+plot_eig_set(eres);
+```
+
+Below shows plots for predator-prey communities.
+
+```{r, echo = FALSE, fig.height = 10, fig.width = 10}
+eres <- get_rand_eigs(int_type = 3, iter = 10000);
+plot_eig_set(eres);
+```
+
+What I would like to explore is some sort of pattern in these points that links to the previous plot showing probability of stabilisation versus destabilisation of generation time, but I'm stuck for now.
+What I can do is look at stable versus feasible systems; the latter seem pretty much the same, except that real eigenvalues in feasible systems are all lower values.
+Below shows one example for a predator-prey community of $S = 20$.
+
+```{r, echo = FALSE, fig.height = 6, fig.width = 10}
+get_rand_feas <- function(int_type = 0, iter = 1000){
+    nn             <- 20;
+    eigres         <- matrix(data = 0, ncol = 42, nrow = iter);
+    eigres[,1]     <- int_type;
+    while(iter > 0){
+        r_vec     <- runif(n = nn, min = 1, max = 10);
+        A1_dat    <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+        A1        <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A1        <- species_interactions(mat = A1, type = int_type);
+        diag(A1)  <- -1;
+        avg_mat   <- matrix(data = mean(1:999), nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        A1        <- A1 * avg_mat;
+        mat1_e    <- eigen(A1)$values;
+        real_mat1 <- Re(mat1_e);
+        imag_mat1 <- Im(mat1_e);
+        stab1     <- max(real_mat1) < 0;
+        A1_fea    <- min(-1*solve(A1) %*% r_vec) > 0;
+        if(stab1 == TRUE & A1_fea == FALSE){
+            eigres[iter,2] <- 1;
+        }
+        if(stab1 == TRUE & A1_fea == TRUE){
+            eigres[iter,2] <- 2;
+        }
+        eigres[iter,3:22]  <- real_mat1;
+        eigres[iter,23:42] <- imag_mat1;
+        iter    <- iter - 1;
+    }
+    return(eigres);
+}
+
+plot_feas_set <- function(eigres){
+    par(mfrow = c(1,2), mar = c(5, 5, 1, 1));
+    x0 <- min(eigres[,3:22]);
+    x1 <- max(eigres[,3:22]);
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(1 %in% eigres[,2]){
+        row <- which(eigres[,2] == 1)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Stable", cex = 1.5);
+    plot(x = 0, y = 0, type = "n", xlim = c(x0, x1), ylim = c(-1500, 1500),
+         xlab = "Real component", ylab = "Imaginary component", cex.lab = 1.25);
+    abline(h = 0, lty = "dotted", lwd = 0.8);
+    abline(v = 0, lty = "dotted", lwd = 0.8);
+    if(2 %in% eigres[,2]){
+        row <- which(eigres[,2] == 2)[1];
+        m1r <- eigres[row,3:22];
+        m1i <- eigres[row,23:42];
+        points(x = m1r, y = m1i, pch = 20, cex = 1.5, col = "black");
+    }
+    text(x = x0 + (x1 - x0)*0.2, y = 1500, "Feasible", cex = 1.5);
+}
+fea_test <- get_rand_feas(iter = 10000, int_type = 3);
+plot_feas_set(fea_test)
+```
+
+********************************************************************************
+
+**AUTHOR'S NOTE:** This is an initial draft of a potential [brief communcation](https://www.nature.com/natecolevol/about/content) for *Nature Ecology and Evolution*.
+I think that the take-home message can be expressed in 1500 words.
+As far as I've been able to find, such an letter would be novel and interesting.
+The most likely group that would have discovered this already would be the [Allesina lab](https://allesinalab.uchicago.edu/), which works a lot with random matrices, but I've not seen anything about generation time variation published by them.
+In developing this theory, I also came up with some brief comments on [the effect of increasing numbers of evolutionary traits on eco-evolutionary stability.](#comment_on_Patel), making the simple point after reading @Patel2018 that ecological systems are increasingly likely to be destablised, and unstable ecological systems are decreasingly likely to be stabilised, at higher numbers of evolving species traits.
+
+For the final paper, I am strongly considering changing the simulations a bit to consider 3-5 scenarios.
+This isn't going to change anything about the general conclusions, I don't think.
+What it will do is make the simulations more consistent with what others have tried in the past:
+
+1.
+Random network but with elements drawn from a normal distribution with N(0, 0.4) and the diagonal elements of the matrix being forced at -1 (helping out stability a bit -- since in all stable matrices, diagonal elements must be negative).
+2. Competitive network with elements drawn from N(0, 0.4) and diagonals at -1, but when sampling form the normal distribution returns positive values, the sign is reversed so all values are negative.
+3. Mutualist network created as above, but with all signs forced to be positive (again, except the diagonal).
+4. Predator-prey network created as above, but forcing the sign of a_ij to be the opposite of a_ji.
+
+If the below conclusions hold for all of the above, then I think the results can be made even stronger.
+The best to present would probably be 1, with 2-4 mentioned and then shown in the supporting information.
+It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable.
+Feasibility criteria are as follows,
+
+$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.
+The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive).
+The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have).
+The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0).
+And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off diagonal effecting species interactions,
+
+$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
+
+To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
+So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
+This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
+
 ********************************************************************************
@@
 
+<a name = "old_tab_1">**Table 1:**</a> Stability results from 100 million simulations of random interaction matrices for 2-15 species ($S$).
+Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\gamma = 1$).
+Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $Var(\gamma)$, respectively.
+Column 6 reports the total number of communities that are stable given $Var(\gamma)$ (compare to column 3).
+
 ```{r, echo = FALSE, caption = "Stability results from 100 million simulations of random interaction matrices for 2-15 species (S).
@@
 
-**Table 1:** Stability results from 100 million simulations of random interaction matrices for 2-15 species ($S$).
-Columns 2 and 3 report the number of stable and unstable communities in the absence of variation in species generation time (all $\gamma = 1$).
-Columns 4 and 5 report the number of unstable communities stabilised or stable communities destabilised by $Var(\gamma)$, respectively.
-Column 6 reports the total number of communities that are stable given $Var(\gamma)$ (compare to column 3).
+********************************************************************************
 
-********************************************************************************
+<a name = "old_fig_1">**Figure 1:**</a> Stability of random communities when generation time ($\gamma$) varies and when it does not for 2-15 species (x-axis).
+Bars compare the $\ln$ total number of communities found to be stable out of 100 million before (dark grey) and after (light grey) $Var(\gamma)$ is added to a community (left y-axis).
+Points show the proportion of these random communities in which stability is a direct consequence of $Var(\gamma)$.
 
@@
 ```
-
-**Figure 1:** Stability of random communities when generation time ($\gamma$) varies and when it does not for 2-15 species (x-axis).
-Bars compare the $\ln$ total number of communities found to be stable out of 100 million before (dark grey) and after (light grey) $Var(\gamma)$ is added to a community (left y-axis).
-Points show the proportion of these random communities in which stability is a direct consequence of $Var(\gamma)$.
 

### 2018-03-28 random_matrix_stability.Rmd
Fix bullet

@@
 The overall effect is therefore to make real ecological communities less stable.
-- I am still working out what this means conceptually by multiplying each row of the matrix by a single number, the shape of the matrix is somehow being changed to decrease the eigenvalues.
-- The exception to this is mutualistic communities -- variation in generation time had zero effect on mutualist community stability.
+- I am still working out what this means conceptually (trying to do this by plotting the eigenvalues and comparing different random systems).
+- The exception to all of this is mutualistic communities -- variation in generation time had zero effect on mutualist community stability.
 - **I am tempted to broaden these results to make a more general point about complex dynamic systems**: "Variation in process time increases stability but decreases feasibility in complex systems", with a focus on biological systems.
@@
 The shift is not typically drastic, but sometimes causes the sign of the real components (x-axis) to flip, thereby potentially affecting stability.
-Immediately below shows plots for a random community; where not points are plot in any of the below, no examples exist.
+Immediately below shows plots for a random community; where no points are plotted in any of the figures below, no examples exist.
 

### 2018-03-29 test_mat.R
More code for eigenalysis of transforming matrix

@@ -406,7 +406,122 @@ plot_feas_set <- function(eigres){
 }
 
 
-eres <- get_rand_eigs(int_type = 2, iter = 10000);
+eres <- get_rand_eigs(int_type = 0, iter = 10000);
 plot_eig_set(eres);
+
+
+
+
+
+
+get_var_change <- function(int_type = 0, iter = 1000){
+    m_changes      <- NULL;
+    nn             <- 20;
+    eigres         <- matrix(data = 0, ncol = 82, nrow = iter);
+    eigtests       <- matrix(data = 0, nrow = iter, ncol = 21);
+    eigres[,1]     <- int_type;
+    while(iter > 0){
+        r_vec     <- runif(n = nn, min = 1, max = rmx);
+        A1_dat    <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+        A1        <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A1        <- species_interactions(mat = A1, type = int_type);
+        diag(A1)  <- -1;
+        epsil     <- runif(n = nn, min = 1, max = eps_max);
+        eps_dat   <- rep(x = epsil, times = nn);
+        eps_mat   <- matrix(data = epsil, nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        avg_mat   <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                            byrow = FALSE);
+        A2        <- A1 * eps_mat;
+        A1        <- A1 * avg_mat;
+        mat1_e    <- eigen(A1)$values;
+        mat2_e    <- eigen(A2)$values;
+        real_mat1 <- Re(mat1_e);
+        imag_mat1 <- Im(mat1_e);
+        real_mat2 <- Re(mat2_e);
+        imag_mat2 <- Im(mat2_e);
+        stab1     <- max(real_mat1) < 0;
+        stab2     <- max(real_mat2) < 0;
+        mchange   <- solve(t(A1) %*% A1) %*% t(A1) %*% A2;
+        m_changes[[iter]] <- mchange;
+        if(stab1 == FALSE & stab2 == FALSE){
+            eigres[iter,2]    <- 1;
+            eigtests[iter,1]  <- 1;
+        }
+        if(stab1 == TRUE & stab2 == TRUE){
+            eigres[iter,2] <- 2;
[…]

### 2018-03-29 random_matrix_stability.Rmd
Notes on isolating transforming matrix, and it's lack of fruit

@@
 
+I also looked at the matrices that decribe the *changes* between the community matrix with no variation in generation time ($A_{1}$) versus after the addition of generation time ($A_{2}$), which can be described by the transforming matrix ($B$), where,
+
+$$A_{2} = A_{1}B$$
+$$A_{1}^{t} A_{2} = A_{1}^{t} A_{1}B$$
+$$(A^{t}_{1}A_{1})^{-1}A_{1}^{t} A_{2} = (A^{t}_{1}A_{1})^{-1}A_{1}^{t} A_{1}B$$
+
+$$B = (A^{t}_{1}A_{1})^{-1}A_{1}^{t} A_{2}$$
+The transforming matrix B does not appear to have any clear properties when $A_{2}$ stabilises versus destabilises -- its values appear the same, as do the eigenvalues after decomposition. Weird.
+
 ********************************************************************************

### 2018-03-30 test_mat.R
Some more playing around

@@ -1,6 +1,7 @@
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 999){
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, eps_max = 999){
     tot_res <- NULL;
     fea_res <- NULL;
+    nea_res <- NULL;
     for(i in 2:max_sp){
         nn             <- i;
         A1_stt         <- 0;
@@ -10,8 +11,9 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 999){
         iter           <- iters;
         tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
         fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        nea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
         while(iter > 0){
-            r_vec    <- runif(n = nn, min = 1, max = rmx);
+            r_vec    <- rnorm(n = nn, mean = 0, sd = rmx);
             A1_dat   <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
             A1       <- matrix(data = A1_dat, nrow = nn, ncol = nn);
             A1       <- species_interactions(mat = A1, type = int_type);
@@ -24,10 +26,13 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 10, eps_max = 999){
                                byrow = FALSE);
             A2       <- A1 * eps_mat;
             A1       <- A1 * avg_mat;
+            #r_vec    <- r_vec * epsil / 100;
             A1_stb   <- max(Re(eigen(A1)$values)) < 0;
             A2_stb   <- max(Re(eigen(A2)$values)) < 0;
             A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
             A2_fea   <- min(-1*solve(A2) %*% r_vec) > 0;
+            A1_nea   <- max(-1*solve(A1) %*% r_vec) < 0;
+            A2_nea   <- max(-1*solve(A2) %*% r_vec) < 0;
             if(A1_stb == TRUE){
                 tot_res[[i-1]][iter, 1] <- 1;
             }
[…]

### 2018-03-30 random_matrix_stability.Rmd
Add note

@@
 ********************************************************************************
+
+**UPDATE 30 MAR**:
+
+It appears that one of the assumptions that I was making was rather badly thought out.
+In checking for feasibility, I simulated random intrinsic species growth rates that were sampled from a uniform distribution $\mathcal{U}\left(0, 1\right)$.
+I thought this would be reasonable at first because species intrinsic population growth rates surely would need to be positive for a species to avoid extinction; this of course is not the case.
+A simple counter-example is that of an obligate mutualist.
+In the absence of a mutualist partner, the obligate mutualist *by definition* should have a negative intrinsic growth rate.
+Hence, the more appropriate sampling for elements of the vector $\textbf{r}$ should be something like $\mathcal{U}\left(-1, 1\right)$.
+When doing it this way, variation in generation time appears to have no effect, rather than a negative effect, on feasibility.
+More simulations of this are needed to confirm.
+To answer the obvious question, sampling from something like $\mathcal{U}\left(-1, 0\right)$ to get elements for $\textbf{r}$ *does not* appear to increase feasibility at first glance, but I'm simulating more communities to double-check.
+
+**Note**: @Song2018, rehashing @May1972, point out that "stable communities will only emerge as long as a derived measure of com- munity complexity is below a certain threshold" defined by $\sigma\sqrt{S \times C}$, "where $\sigma$ represents the standard deviation of the distribution of interaction strengths, S corresponds to the number of species in the community, and C corresponds to the connectance in the community".
+I believe that $Var(\gamma_{i})$ would be independent of all of these parameters (not modifying the *strength* of interactions or the *connectance*, but the speed of system components).
+
+Nevertheless, @Song2018 state that as long as $\sigma\sqrt{S \times C} < 1$, the system will be stable regardless of other parameters in the system.
+I believe that the simulations with $Var(\gamma_{i})$ now show otherwise, but I need to make sure.
+In the simulations that I have run $S \times C = S$ because $C = 1$ (I've not tried anything else).
+Is the standard deviation of interaction strengths changing as a consequence of $\sqrt{Var(\gamma_{i})}$?
 

### 2018-04-07 random_matrix_stability.Rmd
New comments

@@
 ********************************************************************************
+
+**UPDATE 7 APR**:
+
+Currently, I am having 1 million simulations run for four community types (random, competitor, mutualist, predator-prey).
+But the manuscript itself is likely to focus genearally on **stability of complex systems as a consequence of variation in the response rate of individual components** (that will not be the title).
+It will be in the form of a [general 1500 word letter](https://www.nature.com/nature/for-authors/formatting-guide#a1), with an appropriate [introductory summary paragraph](http://s3-service-broker-live-19ea8b98-4d41-4cb4-be4c-d68f4963b7dd.s3.amazonaws.com/uploads/ckeditor/attachments/7808/2c_Summary_para.pdf).
+At least two key results will be emphasised:
+
+1.
+**System stability increases with variation in component response rate.
+And stable large complex systems are almost certain to be dependent upon this variation for preserving their stability** As sub-points, this is generally true for random, competitor, and predator-prey, but not mutualist systems.
+Further, feasibility is not increased by this variation, but it can be argued that no ecological system is isolated from other processes (evolutionary, social, economic).
+The importance of the bold result needs to be emphasised while also acknowledging the nuance, and further acknowledging that *most* stable systems will be actually *destabilised* by variation in component response rate.
+Paradoxically, initial analysis appears to reveal that *addition* of such variation is more likely to destabilise already stable complex systems than it is to stabilise existing unstable complex systems.
+But overall, **complex systems with variation in component response rate are more likely to be stable than systems with no such variation**.
+The logic behind this needs to be made clear.
+2.
+**System stability increases when component response rate is correlated with component interaction type**.
+For example, if components that typically respond positively to increases in other components (e.g., a generalist top level predator in a community) tend to move respond more or less quickly than other types of components, the whole system is highly likely to be stabilised (more so than just given general variance as discussed in 1 above).
+Simulations for this are still running, but if this is even more stabilising, then it would be worth including in the manuscript.
+
+Other key results will come if I can think of other interesting ways to vary component response rate.
+
+Key figures of this paper will include:
+
+1. Before and after side-by-side histograms of system stability proportion (or count), with overlayed line showing the percentage of systems that are stable *due to* variation in component time.
+2. A heat map of eigenvalues before and after variation over many simulations.
 
@@
 ```{r, echo = FALSE, fig.height = 10, fig.width = 10}
-random      <- read.csv(file = "sim_results/random_community.csv");
-competition <- read.csv(file = "sim_results/competitor_community.csv");
-mutualism   <- read.csv(file = "sim_results/mutualist_community.csv");
-pred_prey   <- read.csv(file = "sim_results/predator_prey_community.csv");
+random      <- read.csv(file = "sim_results_old/random_community.csv");
+competition <- read.csv(file = "sim_results_old/competitor_community.csv");
+mutualism   <- read.csv(file = "sim_results_old/mutualist_community.csv");
+pred_prey   <- read.csv(file = "sim_results_old/predator_prey_community.csv");
 # ==============================================================================

### 2018-04-07 test_mat.R
Code to test component correlations

@@ -1,3 +1,58 @@
+
+plot_eigen_c <- function(eigen_list, xlim = c(-2000, 500), ylim = c(-800, 800)){
+    coords1 <- expand.grid(real = xlim[1]:xlim[2], imag = ylim[1]:ylim[2]);
+    coords2 <- expand.grid(real = xlim[1]:xlim[2], imag = ylim[1]:ylim[2]);
+    heat1   <- rep(0, length = dim(coords1)[1]);
+    heat2   <- rep(0, length = dim(coords2)[1]);
+    el1     <- eigen_list$eig_1;
+    el2     <- eigen_list$eig_2;
+    for(i in 1:dim(coords1)[1]){
+        rightRe1  <- el1[,1] == coords1[i,1];
+        rightIm1  <- el1[,2] == coords1[i,2];
+        heat1[i]  <- sum(rightRe1 == TRUE & rightIm1 == TRUE);
+        rightRe2  <- el2[,1] == coords2[i,1];
+        rightIm2  <- el2[,2] == coords2[i,2];
+        heat2[i]  <- sum(rightRe2 == TRUE & rightIm2 == TRUE);
+        if(i %% 5000 == 0){
+            pct <- round((i / dim(coords1)[1]) * 100, digits = 2);
+            print(paste(pct, "percent complete"));
+        }
+    }
+    pl1 <- cbind(coords1, heat1);
+    pl2 <- cbind(coords2, heat2);
+    return(list(A1 = pl1, A2 = pl2));
+}
+
+eigen_cloud <- function(nn, iters, int_type = 0, rmx = 0.4, eps_max = 999){
+    e_coord1 <- NULL;
+    e_coord2 <- NULL;
+    iter     <- iters;
+    while(iter > 0){
+        A1_dat   <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+        A1       <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+        A1       <- species_interactions(mat = A1, type = int_type);
+        diag(A1) <- -1;
+        epsil    <- runif(n = nn, min = 1, max = eps_max);
+        eps_dat  <- rep(x = epsil, times = nn);
+        eps_mat  <- matrix(data = epsil, nrow = nn, ncol = nn, 
+                          byrow = FALSE);
+        avg_mat  <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                           byrow = FALSE);
[…]

### 2018-04-08 random_matrix_stability.Rmd
Add an outline, and some old stuff, back into the working file

@@
 
+**UPDATE 8 APR**:
+
+Here's a working outline of the manuscript:
+
+1.
+Introductory [summary paragraph](http://s3-service-broker-live-19ea8b98-4d41-4cb4-be4c-d68f4963b7dd.s3.amazonaws.com/uploads/ckeditor/attachments/7808/2c_Summary_para.pdf), which will be broadly focused on complex system stability.
+2.
+Go back to @May1972 and note how the original ecological result has been broadened for systems such as evolutionary networks, gene regulatory networks, economics, etc.
+(references needed).
+Perhaps make the broader point that even local systems have different elements of different types of systems.
+3.
+Briefly point out the advances that have been made in terms of network structures (e.g., random versus competitive interactions), but then hit the point that different components of complex systems are expected to respond on different time scales [e.g.
+@Patel2018], and no one has looked at how this might affect stability of systems more broadly.
+The current paper is going to show that variation in component response rate increases stability.
+4. A paragraph with some broad explanation of the idea of matrix stability, along the lines of @Allesina2012. This will effectively explain the idea of random matrix theory to the reader.
+5.
+A paragraph introducing, mathematically, how I am adding varying rates of component response -- multipling each row $i$ by a scalar $\gamma_{i}$.
+First this will be done with some simple function giving $Var(\gamma_{i})$, then in a more structured way such that random components that are more likely to be positively influenced by other components having faster or slower rates (higher or lower $\gamma_{i}$) of response.
+A sentence or two is needed to suggest what this might mean biologically (e.g., generalist predators have longer generation times than their prey).
+6.
+A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in supporting information.
+The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.
+I could then argue that $Var(\gamma_{i})$ has a minimum such correlation of zero, and can only increase in stabilisation by having an absolute value of correlation being greater than zero).
+At least two [key figures](#fig_ideas) will be included, likely with four panels (1.
+without variation, 2.
+with unstructured variation, and 3.
+and 4.
+with structured variation).
+7.
+Some sort of mention is needed for a couple odd results; first, that in mutualist networks, stability is not affected -- perhaps because of the lack of variance in component response rate given number of interaction types (note, I think, competitive networks can have indirect positives due to intransitive effects).
+Second, *feasibility* is not increased, so pure ecological networks might not be affected by this (but emphasise that *pure* ecological networks don't really exist in nature -- they are incorporated into evolutionary, economic, social, systems).
+8.
+Some bold statement is needed here -- something to the effect of the argument that nearly all large stable complex systems will be stable *due to* the variation simulated in this paper.
+The odds that it would be stable without such variation becomes vanishingly small as system size increases.
+
 **UPDATE 7 APR**:
@@
 
-Key figures of this paper will include:
+<a name="fig_ideas">Key figures</a> of this paper will include:
 
@@
 2. A heat map of eigenvalues before and after variation over many simulations.
+
+The manuscript overall is going to need to be very general, not just ecology focused.
+I will need to talk about complex systems such as evolutionary systems, social systems, and financial systems (the last one here is a very clear case where negative values are allowed, so feasibility is not as big of a deal).
+
+Relevant papers [here](https://www.nature.com/articles/nature16948), and [here](http://science.sciencemag.org/content/337/6092/349)?
+I also need to add some code to skip feasibility analyses when matrices have determinant of zero (are singular), as part of the feasibility analysis requires matrix inversion.
 
@@
 
+**OLD STUFF**
+
+I'm adding some stuff from [earlier](https://raw.githubusercontent.com/bradduthie/random_matrix_stability/e1314dfc13353aa301f34782827c092b0c5c7b52/random_matrix_stability.Rmd?token=AFSzia2ycJ9XXiZYYMrWG9PrWVHCXRheks5a03EvwA%3D%3D) back in now, as it might be used in the writing again.
+Note that the language underlying all of this needs to be adjusted to be about complex systems more generally.
+
+********************************************************************************
+
+**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size.
+But in ecological communities, the rate that species densities change as a consequence of species interactions will depend on relative species generation times.
+Densities of species with relatively short generation times will change more rapidly than densities of species with longer generation times.
+Interaction effects are therefore expected to be correlated within species, with some species densities changing more rapidly than others in a community.
+Here we show that by accounting for variation in generation time among species, community stability is substantially increased.
+Further, the influence of variation in generation time becomes exponentially more important as community size increases.
+We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
+
+********************************************************************************
+
+Species generation time varies by orders of magnitude within communities, potentially affecting the rate at which the density of one species affects the population growth rate of another.
+Yet I've not been able to find anything on this in the literature in a random matrix context (or any other context).
+I want to **determine if the effects of varying generation time among community members is stabilising or destabilising in an ecological system**.
+This work is partly inspired by the use of $\varepsilon$ by @Patel2018 to separate the time scale of ecological versus evolutionary dynamics, and partly inspired by @Jirsa2004 to separate generation time of different species in a community.
+
+Here I will consider a Jacobian matrix of interacting species $\boldsymbol{A}$ and focus on the stability of ecological interactions given variation in generation time.
+I will use the same technique of @Patel2018 for varying ecological and evolutionary time scales, but a bit differently.
+I assume that the effect that the density of any species $j$ ($N_{j}$) has on the growth rate of a focal species $i$ ($\partial N_{i}W_{i}/\partial N_{j}$) is consistently affected by a value $\gamma_{i}$ (I am using $\gamma_{i}$ instead of $\varepsilon_{i}$ to avoid confusing generation time variance with ecological versus evolutionary rate difference).
+Higher values of $\gamma_{i}$ correspond to shorter absolute generation times and lower values correspond to longer generation times.
+The matrix below $\boldsymbol{A}$ describes the dynamics for two species $i$ and $j$ with effects of generation times $\gamma_{i}$ and $\gamma_{j}$, respectively.
+
+$$\boldsymbol{A} = \begin{bmatrix}
+    \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{i}}  & \gamma_{i}\frac{\partial N_{i}W_{i}}{\partial N_{j}}  \\
+    \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{i}}  & \gamma_{j} \frac{\partial N_{j}W_{j}}{\partial N_{j}}  \\
+\end{bmatrix}.$$
+
+I will investigate **how the amount of variation in generation time (range of $\gamma$) in a community affects stability**.
+
 ********************************************************************************

### 2018-04-08 random_matrix_stability.Rmd
Draft of summary paragraph

@@
 ---
+
+********************************************************************************
+
+[Working draft of the manuscript](#ms)
 
@@
 ********************************************************************************
+
+<a name="ms">The manuscript draft</a>
+
+********************************************************************************
+
+**The stability of a complex system generally decreases with increasing system size, as is demonstrated by random matrix theory [@May1972; @Allesina2012].
+This counter-intuitive result, first shown by May [@May1972], is broadly relevant for understanding the dynamics and persistence of systems such as chemical, biological, and social networks.
+Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Allesina2012].
+But more broadly, stabilising mechanisms in complex systems remain under-explored, and the effect of variation in the response rate of individual system components remains unexplored.
+Here I show that when components of a complex system respond to system dynamics at different rates, system stability is strongly increased.
+This increase in stability is driven by the correlation between a focal component's response rate and the number of other components in the system that have a negative or positive effect on it.
+When variation in component response rate ($\gamma$) is included in random interaction matrices, the effect of such emerging correlations necessarily increases stability.
+Further, these effects become increasingly more important as system size increases.
+My results therefore reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
+Future research in complex systems should therefore account for the varying response rates of individual system components when assessing whole system stability.**

### 2018-04-08 random_matrix_stability.Rmd
New results and manuscript start

@@
 
+**New result -- a correlation between the number of positive or negative values in a row of the Jacobian matrix (e.g., corresponding to a species that is generally positively affected by increases in other species -- generalist predator, or generally negatively affected by increases in other species -- a producer or resource), and the response rate of that row (e.g., species generation), is *highly* stabilising, and I believe that I can dig into the matrices and show that *this* is the mechanism that is tweaked when $Var(\gamma_{i})$ is included.
+In other words, variation in generation time inevitably increases or decreases the correlation (from zero) between sum of row signs and *types* of interactions and generation time -- which I will now generalise to something like reaction rate.** I will present these results in figures soon.
+I will also show that 1.
+general variation in $\gamma_{i}$ changes the correlation between sign counts and generation rate and 2.
+when this correlation is higher or lower, eigen values are more likely to all be negative.
+
 Here's a working outline of the manuscript:
@@
 The odds that it would be stable without such variation becomes vanishingly small as system size increases.
+
+**Here's the correlation codes:**
+
+```{r, eval = FALSE}
+SA1      <- sign(A1);
+Sadj     <- apply(X = SA1, MARGIN = 1, FUN = sum);
+mnadj    <- 1 - min(Sadj);
+Sadj     <- exp(Sadj + mnadj);
+```
+
+The above is saved as `random_pos.csv`.
+The rate of change $\gamma_{i}$ in row $i$ equals the sum values in the row, plus the minimum number of values (e.g., four positive and three negative value equals one, plus whatever is the minimum -- if -3, then the end value should be four).
+
+```{r, eval = FALSE}
+SA1      <- sign(A1);
+Sadj     <- apply(X = -SA1, MARGIN = 1, FUN = sum);
+mnadj    <- 1 - min(Sadj);
+Sadj     <- exp(Sadj + mnadj);
+```
+
+The above is saved as `random_neg.csv`. The rate of change $\gamma_{i}$ in row $i$ equals the opposite of the sum values in the row, plus the minimum number of values.
 

### 2018-04-14 random_matrix_stability.Rmd
New links and fixes

@@
 
-[Working draft of the manuscript](#ms)
+[Working draft of the manuscript](https://bradduthie.github.io/random_matrix_stability/ms.html)
 
@@
 ```{r, echo = FALSE, fig.height = 10, fig.width = 10}
-random      <- read.csv(file = "sim_results_old/random_community.csv");
-competition <- read.csv(file = "sim_results_old/competitor_community.csv");
-mutualism   <- read.csv(file = "sim_results_old/mutualist_community.csv");
-pred_prey   <- read.csv(file = "sim_results_old/predator_prey_community.csv");
+random      <- read.csv(file = "sim_results_old/oldest/random_community.csv");
+competition <- read.csv(file = "sim_results_old/oldest/competitor_community.csv");
+mutualism   <- read.csv(file = "sim_results_old/oldest/mutualist_community.csv");
+pred_prey   <- read.csv(file = "sim_results_old/oldest/predator_prey_community.csv");
 # ==============================================================================

### 2018-04-14 ms.Rmd
New manuscript file initialised

---
title: "Stability concerns in random matrices"
author: "Brad Duthie"
bibliography: references.bib
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

********************************************************************************

**The stability of a complex system generally decreases with increasing system size, as is demonstrated by random matrix theory [@May1972; @Allesina2012]. This counter-intuitive result, first shown by May [@May1972], is broadly relevant for understanding the dynamics and persistence of systems such as chemical, biological, and social networks. Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Haldane2011; @Allesina2012; @Mougi2012; @Allesina2015a; @Gao2016; @Grilli2017; @Patel2018]. But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains unexplored. Here I show that when components of a complex system respond to system dynamics at different rates, system stability is markedly increased. This increase in stability is especially strong when variation in component response rate ($\gamma$) is high and has a skewed distribution. Further, these effects become increasingly more important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for variation in $\gamma$. My results therefore reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems. Future research in complex systems should therefore account for the varying response rates of individual system components when assessing whole system stability.**

**References**

<div id="refs"></div>

********************************************************************************

**Figure 1**

```{r, eval = TRUE, echo = FALSE, fig.height = 8}
make_gammas <- function(nn = 10, distribution = 1, mn = 10, sdd = 1){
    if(distribution == 0){
        dat          <- rep(x = mn, times = nn);
    }
    if(distribution == 1){
        dat          <- rnorm(n = nn, mean = mn, sd = sdd);
        dat[dat < 0] <- -dat[dat < 0];
    }
    if(distribution == 2){
        dat          <- runif(n = nn, min = 0, max = 20);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 3){
        dat          <- rexp(n = nn);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 4){
        dat          <- 1 - (sdd * rexp(n = nn));
        dat          <- dat - min(dat);
        dat          <- dat - mean(dat) + mn;
        dat[dat < 0] <- -dat[dat < 0];
    }
    if(distribution == 5){
        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 6){
        dat          <- rgamma(n = nn, shape = 2, scale = 2);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    return(dat);
}
show_gammas <- function(nn = 1000000){
    y1 <- make_gammas(nn = nn, distribution = 1);
    y2 <- make_gammas(nn = nn, distribution = 2);
    y3 <- make_gammas(nn = nn, distribution = 3);
    y4 <- make_gammas(nn = nn, distribution = 4);
    y5 <- make_gammas(nn = nn, distribution = 5);
    y6 <- make_gammas(nn = nn, distribution = 6);
    par(mfrow = c(3, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y5, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y6, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    mtext(expression(paste("Component ",gamma[i]," value")),
          outer=TRUE,side=1,line=3.0,cex=2);
    mtext(expression(paste("Relative frequency")),
          outer=TRUE,side=2,line=2.5,cex=2);
}
show_gammas();
```

********************************************************************************

### 2018-04-15 ms.Rmd
Figure legend and sub-panel letters

@@
 
-**Figure 1**
+**Figure 1: Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.
+The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\boldsymbol{\gamma = 10}$ across different system sizes ($\boldsymbol{S}$; i.e., component numbers) and $\boldsymbol{\gamma}$ standard deviations ($\boldsymbol{\sigma_{\gamma} = \{1, 2, 4, 8\}}$).
+All distributions shown below are mean-centred at $\boldsymbol{\mu = 10}$.
+Distributions are as follows: (a) normal, (b) uniform, (c) exponential, (d) reversed exponential (built from 1 minus an exponential distribution, which is then mean-centred and standardised), beta ($\boldsymbol{\alpha = 0.5}$ and $\boldsymbol{\beta = 0.5}$), and gamma ($\boldsymbol{k = 2}$ and $\boldsymbol{\theta = 2}$).
+Each panel shows 1 million randomly generated \boldsymbol{$\gamma$} values.**
 
@@
     par(mfrow = c(3, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
-    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h1 <- hist(y1, breaks = 1000, plot = FALSE);
+    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h1$counts)*1.2));
+    mtext("a", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h2 <- hist(y2, breaks = 1000, plot = FALSE);
+    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h2$counts)*1.2));
+    mtext("b", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h3 <- hist(y3, breaks = 1000, plot = FALSE);
+    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h3$counts)*1.2));
+    mtext("c", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h4 <- hist(y4, breaks = 1000, plot = FALSE);
+    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h4$counts)*1.2));
+    mtext("d", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    hist(y5, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h5 <- hist(y5, breaks = 1000, plot = FALSE);
+    hist(y5, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h5$counts)*1.2));
+    mtext("e", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    hist(y6, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
+    h6 <- hist(y6, breaks = 1000, plot = FALSE);
+    hist(y6, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h6$counts)*1.2));
+    mtext("f", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
     box();
-    mtext(expression(paste("Component ",gamma[i]," value")),
+    mtext(expression(paste("Component ",gamma," value")),
           outer=TRUE,side=1,line=3.0,cex=2);

### 2018-04-16 ms.Rmd
New MS figures

--- 2018-04-14 ms.Rmd
@@
 
-**Figure 1**
+**Figure 1: Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.
+The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\boldsymbol{\gamma = 100}$ across different system sizes ($\boldsymbol{S}$; i.e., component numbers) and $\boldsymbol{\gamma}$ standard deviations ($\boldsymbol{\sigma_{\gamma} = \{1, 4, 8, 12\}}$).
+All distributions shown below are mean-centred at $\boldsymbol{\mu = 100}$.
+Distributions are as follows: (a) normal, (b) uniform, (c) exponential, (d) reversed exponential (built from 1 minus an exponential distribution, which is then mean-centred and standardised), beta ($\boldsymbol{\alpha = 0.5}$ and $\boldsymbol{\beta = 0.5}$), and gamma ($\boldsymbol{k = 2}$ and $\boldsymbol{\theta = 2}$).
+Each panel shows 1 million randomly generated \boldsymbol{$\gamma$} values.**
 
@@
 }
-show_gammas();
+show_gammas(mn = 100);
 ```

### 2018-04-16 test_mat.R
Fix some distributions

@@ -133,34 +133,22 @@ make_gammas <- function(nn = 10, distribution = 1, mn = 10, sdd = 1){
         dat          <- rep(x = mn, times = nn);
     }
     if(distribution == 1){
-        dat          <- rnorm(n = nn, mean = mn, sd = sdd);
-        dat[dat < 0] <- -dat[dat < 0];
+        dat             <- runif(n = nn, min = 0, max = 2*mn);
     }
     if(distribution == 2){
-        dat          <- runif(n = nn, min = 0, max = 20);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
-    }
-    if(distribution == 3){
         dat          <- rexp(n = nn);
         dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
+        dat          <- dat - min(dat) + min(abs(dat));
     }
-    if(distribution == 4){
-        dat          <- 1 - (sdd * rexp(n = nn));
-        dat          <- dat - min(dat);
-        dat          <- dat - mean(dat) + mn;
-        dat[dat < 0] <- -dat[dat < 0];
-    }
-    if(distribution == 5){
+    if(distribution == 3){
         dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
         dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
+        dat          <- dat - min(dat) + min(abs(dat));
     }
-    if(distribution == 6){
+    if(distribution == 4){
         dat          <- rgamma(n = nn, shape = 2, scale = 2);
         dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
+        dat          <- dat - min(dat) + min(abs(dat));
     }
     return(dat);
 }
@@ -172,30 +160,20 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1,
     tot_res <- NULL;
     fea_res <- NULL;
     for(i in 2:max_sp){
-        nn             <- i;
         iter           <- iters;
         tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
         while(iter > 0){
[…]

### 2018-04-16 test_mat.R
Rework distributions and output based on them

@@ -4,7 +4,7 @@
 
 summarise_randmat <- function(tot_res, fea_res){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 53);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 37);
     for(i in 1:sims){
         all_res[i, 1]  <- i + 1;
         # Stable and unstable
@@ -18,91 +18,63 @@ summarise_randmat <- function(tot_res, fea_res){
         all_res[i, 9]  <- sum(tot_res[[i]][,4] == TRUE);
         all_res[i, 10] <- sum(tot_res[[i]][,5] == FALSE);
         all_res[i, 11] <- sum(tot_res[[i]][,5] == TRUE);
-        all_res[i, 12] <- sum(tot_res[[i]][,6] == FALSE);
-        all_res[i, 13] <- sum(tot_res[[i]][,6] == TRUE);
-        all_res[i, 14] <- sum(tot_res[[i]][,7] == FALSE);
-        all_res[i, 15] <- sum(tot_res[[i]][,7] == TRUE);
         # Stabilised and destabilised
-        all_res[i, 16] <- sum(tot_res[[i]][,1] == FALSE & 
+        all_res[i, 12] <- sum(tot_res[[i]][,1] == FALSE & 
                                   tot_res[[i]][,2] == TRUE);
-        all_res[i, 17] <- sum(tot_res[[i]][,1] == FALSE & 
+        all_res[i, 13] <- sum(tot_res[[i]][,1] == FALSE & 
                                   tot_res[[i]][,3] == TRUE);
-        all_res[i, 18] <- sum(tot_res[[i]][,1] == FALSE & 
+        all_res[i, 14] <- sum(tot_res[[i]][,1] == FALSE & 
                                   tot_res[[i]][,4] == TRUE);
-        all_res[i, 19] <- sum(tot_res[[i]][,1] == FALSE & 
+        all_res[i, 15] <- sum(tot_res[[i]][,1] == FALSE & 
                                   tot_res[[i]][,5] == TRUE);
-        all_res[i, 20] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,6] == TRUE);
-        all_res[i, 21] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,7] == TRUE);
[…]

### 2018-04-19 test_mat.R
Add all the old stuff

@@ -100,12 +100,13 @@ species_interactions <- function(mat, type = 0){
 }
 
 
-make_gammas <- function(nn = 10, distribution = 1, mn = 10, sdd = 1){
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
     if(distribution == 0){
         dat          <- rep(x = mn, times = nn);
     }
     if(distribution == 1){
-        dat             <- runif(n = nn, min = 0, max = 2*mn);
+        mval         <- sdd * sqrt(12);
+        dat          <- runif(n = nn, min = 0, max = mval);
     }
     if(distribution == 2){
         dat          <- rexp(n = nn);
@@ -244,45 +245,52 @@ show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
 ################################################################################
 ################################################################################
 
-
-image(-1*egi$hmat, col = heat.colors(2000));
-
-make_e_image <- function(ei_ret){
-     hmat   <- ei_ret$hmat;
-     xcoord <- ei_ret$xcoord;
-     ycoord <- ei_ret$ycoord;
-     adj_ei <- -1 * log(hmat + 0.001);
-     xaxis  <- which(xcoord == 0) / length(xcoord);
-     yaxis  <- which(ycoord == 0) / length(ycoord);
-     coomat <- matrix(data = 0, nrow = dim(hmat)[1], ncol = dim(hmat)[2]);
-     coomat[,xaxis] <- 1;
-     coomat[yaxis,] <- 1;
-     image(adj_ei, col = heat.colors(2000), axes = FALSE);
-     abline(h = xaxis, lwd = 2);
-     abline(v = yaxis, lwd = 2);
-     box();
+eig_plot <- function(ecl){
+     xlims <- c(min(ecl$d1[,1]), max(ecl$d1[,1]));
+     ylims <- c(min(ecl$d1[,2]), max(ecl$d1[,2]));
+     tpink <- "#FFC0CB14"
+     par(mfrow = c(2, 2), mar = c(1, 1, 1, 1), oma = c(5, 5, 1, 1));
+     # Uniform distribution (a)
+     plot(x = ecl$d0[,1], y = ecl$d0[,2], pch = 4, col = "pink", cex = 0.3, 
[…]

### 2018-04-19 R/evolve_gammas.R
Separate R files

@@ -1,0 +1,130 @@
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, eps_max = 999){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    nea_res <- NULL;
+    for(i in 2:max_sp){
+        nn             <- i;
+        A1_stt         <- 0;
+        A2_stt         <- 0;
+        A1_fet         <- 0;
+        A2_fet         <- 0;
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        while(iter > 0){
+            r_vec    <- rnorm(n = nn, mean = 0, sd = rmx);
+            A1_dat   <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
+            A1       <- matrix(data = A1_dat, nrow = nn, ncol = nn);
+            A1       <- species_interactions(mat = A1, type = int_type);
+            diag(A1) <- -1;
+            epsil    <- runif(n = nn, min = 1, max = eps_max);
+            eps_dat  <- rep(x = epsil, times = nn);
+            eps_mat  <- matrix(data = epsil, nrow = nn, ncol = nn, 
+                               byrow = FALSE);
+            avg_mat  <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
+                               byrow = FALSE);
+            A2       <- A1 * eps_mat;
+            A3_stb   <- rand_mat_ga(A1);
+            A1       <- A1 * avg_mat;
+            
+            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
+            A2_stb   <- max(Re(eigen(A2)$values)) < 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
+            A2_fea   <- min(-1*solve(A2) %*% r_vec) > 0;
+            
+            if(A1_stb == TRUE){
+                tot_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A2_stb == TRUE){
+                tot_res[[i-1]][iter, 2] <- 1;
+            }
+            if(A3_stb == 1){
+                tot_res[[i-1]][iter, 3] <- 1;
+            }
[…]

### 2018-04-19 R/plot_figs.R

@@ -1,0 +1,158 @@
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- sdd * sqrt(12);
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
+
+show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
+    y1 <- make_gammas(nn = nn, distribution = 1, sdd = sdd, mn = mn);
+    y2 <- make_gammas(nn = nn, distribution = 2, sdd = sdd, mn = mn);
+    y3 <- make_gammas(nn = nn, distribution = 3, sdd = sdd, mn = mn);
+    y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
+    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
+    h1 <- hist(y1, breaks = 1000, plot = FALSE);
+    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h1$counts)*1.2));
+    mtext("a", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
[…]

### 2018-04-19 R/sim_mat.R

@@ -1,0 +1,196 @@
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- sdd * sqrt(12);
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
+
+summarise_randmat <- function(tot_res, fea_res){
+    sims    <- length(tot_res);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 37);
+    for(i in 1:sims){
+        all_res[i, 1]  <- i + 1;
+        # Stable and unstable
+        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
+        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
+        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
+        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
[…]

### 2018-04-19 R/test_mat.R

@@ -1,503 +1,3 @@
-################################################################################
-################################################################################
-################################################################################
-
-summarise_randmat <- function(tot_res, fea_res){
-    sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 37);
-    for(i in 1:sims){
-        all_res[i, 1]  <- i + 1;
-        # Stable and unstable
-        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
-        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
-        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
-        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
-        all_res[i, 6]  <- sum(tot_res[[i]][,3] == FALSE);
-        all_res[i, 7]  <- sum(tot_res[[i]][,3] == TRUE);
-        all_res[i, 8]  <- sum(tot_res[[i]][,4] == FALSE);
-        all_res[i, 9]  <- sum(tot_res[[i]][,4] == TRUE);
-        all_res[i, 10] <- sum(tot_res[[i]][,5] == FALSE);
-        all_res[i, 11] <- sum(tot_res[[i]][,5] == TRUE);
-        # Stabilised and destabilised
-        all_res[i, 12] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,2] == TRUE);
-        all_res[i, 13] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,3] == TRUE);
-        all_res[i, 14] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,4] == TRUE);
-        all_res[i, 15] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,5] == TRUE);
-        all_res[i, 16] <- sum(tot_res[[i]][,1] == TRUE & 
-                                  tot_res[[i]][,2] == FALSE);
-        all_res[i, 17] <- sum(tot_res[[i]][,1] == TRUE & 
-                                  tot_res[[i]][,3] == FALSE);
-        all_res[i, 18] <- sum(tot_res[[i]][,1] == TRUE & 
[…]

### 2018-04-19 ms.Rmd
Add Figure 2

--- 2018-04-14 ms.Rmd
@@
 
+```{r, echo = FALSE}
+source(file = "R/sim_mat.R");
+source(file = "R/plot_figs.R");
+```
+
 ********************************************************************************
@@
 
-**Figure 1**
+**Figure 1: Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$).
+Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$).
+Each panel shows 1 million randomly generated $\gamma$ values.
 
 ```{r, eval = TRUE, echo = FALSE, fig.height = 8}
-make_gammas <- function(nn = 10, distribution = 1, mn = 10, sdd = 1){
-    if(distribution == 0){
-        dat          <- rep(x = mn, times = nn);
-    }
-    if(distribution == 1){
-        dat          <- rnorm(n = nn, mean = mn, sd = sdd);
-        dat[dat < 0] <- -dat[dat < 0];
-    }
-    if(distribution == 2){
-        dat          <- runif(n = nn, min = 0, max = 20);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
-    }
-    if(distribution == 3){
-        dat          <- rexp(n = nn);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
-    }
-    if(distribution == 4){
-        dat          <- 1 - (sdd * rexp(n = nn));
-        dat          <- dat - min(dat);
-        dat          <- dat - mean(dat) + mn;
-        dat[dat < 0] <- -dat[dat < 0];
-    }
-    if(distribution == 5){
-        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
-    }
-    if(distribution == 6){
-        dat          <- rgamma(n = nn, shape = 2, scale = 2);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat + mn;
-    }
-    return(dat);
-}
-show_gammas <- function(nn = 1000000){
-    y1 <- make_gammas(nn = nn, distribution = 1);
-    y2 <- make_gammas(nn = nn, distribution = 2);
-    y3 <- make_gammas(nn = nn, distribution = 3);
-    y4 <- make_gammas(nn = nn, distribution = 4);
-    y5 <- make_gammas(nn = nn, distribution = 5);
-    y6 <- make_gammas(nn = nn, distribution = 6);
-    par(mfrow = c(3, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
-    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    hist(y5, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    hist(y6, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
-    box();
-    mtext(expression(paste("Component ",gamma[i]," value")),
-          outer=TRUE,side=1,line=3.0,cex=2);
-    mtext(expression(paste("Relative frequency")),
-          outer=TRUE,side=2,line=2.5,cex=2);
-}
 show_gammas();
@@
 ********************************************************************************
+
+```{r, eval = TRUE, echo = FALSE}
+eig_cloud <- eigen_cloud(sp = 20, iters = 1000);
+Pr_D0     <- sum(eig_cloud$d0[,1] < 0) / length(eig_cloud$d0[,1]);
+Pr_D1     <- sum(eig_cloud$d1[,1] < 0) / length(eig_cloud$d1[,1]);
+Pr_D2     <- sum(eig_cloud$d2[,1] < 0) / length(eig_cloud$d2[,1]);
+Pr_D3     <- sum(eig_cloud$d3[,1] < 0) / length(eig_cloud$d3[,1]);
+Pr_D4     <- sum(eig_cloud$d4[,1] < 0) / length(eig_cloud$d4[,1]);
+```
+
+**Figure 2: Distributions of eigenvalues across different distributions of component response rate  ($\boldsymbol{\gamma}$) values.** Panels show 100000 random communities for $S = 20$ and $C = 1$, where off-diagonal elements are randomly sampled from a standard normal distribution $N(0, 1)$, diagonal elements are $-1$, and each row of elements is multiplied by a random scalar value $\gamma$ sampled from one of four distributions (black points in all sub-panels):  **a**, uniform $\Pr(\gamma) = 1/\sqrt{12}$ for $x \in [0, \sqrt{12}]$.
+**b**, exponential $\Pr(\gamma) = e^{-\lambda}$ for $\lambda = 1$.
+**c**, beta $\Pr(\gamma) = \frac{x^{\alpha = 1}(1-x)^{\beta - 1}}{B(\alpha, \beta)}$, where $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$ for $\alpha = 0.5$ and $\beta = 0.5$.
+**d**, gamma $\Pr(\gamma) = \frac{1}{\Gamma(k)\theta^{k}}x^{k-1}e^{\frac{-x}{\theta}}$ for $k = 2$ and $\theta = 2$.
+Eigenvalues associated with each $\gamma$ distribution are compared with otherwise identical communities in which $\gamma = 1$ (pink points in all sub-panels).
+In all distributions, $SD(\gamma) = 1$; note that $\gamma$ distributions cannot be mean standardised without changing the sign of some community matrix elements, but the sign of eigenvalues (and therefore community stability) is unaffected by multiplying elements by any scalar value.
+Proportions of real components of eigenvalues (x-axis) below 0 are (a) `r Pr_D1`, (b) `r Pr_D2`, (c) `r Pr_D3`, (d) `r Pr_D4`, as compared to `r Pr_D0` when $\gamma = 1$.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 8}
+eig_plot(ecl = eig_cloud);
+```
+
+********************************************************************************

### 2018-04-19 ms.Rmd
Re-run with more iters

@@
 ---
-title: "Stability concerns in random matrices"
+title: "Component response rate variation drives stability in large complex systems"
 author: "Brad Duthie"
@@
 ```{r, eval = TRUE, echo = FALSE}
-eig_cloud <- eigen_cloud(sp = 20, iters = 1000);
+eig_cloud <- eigen_cloud(sp = 20, iters = 10000);
 Pr_D0     <- sum(eig_cloud$d0[,1] < 0) / length(eig_cloud$d0[,1]);

### 2018-04-19 ms.Rmd
Fix notation

@@
 This increase in stability is especially strong when variation in component response rate ($\gamma$) is high and has a skewed distribution.
-Further, these effects become increasingly more important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for variation in $\gamma$.
+Further, these effects become increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for variation in $\gamma$.
 My results therefore reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
@@
 
-**Figure 2: Distributions of eigenvalues across different distributions of component response rate  ($\boldsymbol{\gamma}$) values.** Panels show 100000 random communities for $S = 20$ and $C = 1$, where off-diagonal elements are randomly sampled from a standard normal distribution $N(0, 1)$, diagonal elements are $-1$, and each row of elements is multiplied by a random scalar value $\gamma$ sampled from one of four distributions (black points in all sub-panels):  **a**, uniform $\Pr(\gamma) = 1/\sqrt{12}$ for $x \in [0, \sqrt{12}]$.
-**b**, exponential $\Pr(\gamma) = e^{-\lambda}$ for $\lambda = 1$.
-**c**, beta $\Pr(\gamma) = \frac{x^{\alpha = 1}(1-x)^{\beta - 1}}{B(\alpha, \beta)}$, where $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$ for $\alpha = 0.5$ and $\beta = 0.5$.
-**d**, gamma $\Pr(\gamma) = \frac{1}{\Gamma(k)\theta^{k}}x^{k-1}e^{\frac{-x}{\theta}}$ for $k = 2$ and $\theta = 2$.
+**Figure 2: Distributions of eigenvalues across different distributions of component response rate  ($\boldsymbol{\gamma}$) values.** Panels show 100000 random communities for $S = 20$ and $C = 1$, where off-diagonal elements are randomly sampled from a standard normal distribution $N(0, 1)$, diagonal elements are $-1$, and each row of elements is multiplied by a random scalar value $\gamma$ sampled from one of four distributions (black points in all sub-panels):  **a**, uniform $\Pr(\gamma = x) = 1/\sqrt{12}$ where $x \in [0, \sqrt{12}]$.
+**b**, exponential $\Pr(\gamma = x) = e^{-x}$ where $x \geq 0$.
+**c**, beta $\Pr(\gamma = x) = \frac{x^{\alpha - 1}(1-x)^{\beta - 1}}{B(\alpha, \beta)}$ where $x \in [0, 1]$, and where $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$ for $\alpha = 0.5$ and $\beta = 0.5$.
+**d**, gamma $\Pr(\gamma = x) = \frac{1}{\Gamma(k)\theta^{k}}x^{k-1}e^{\frac{-x}{\theta}}$ for $k = 2$ and $\theta = 2$, where $x \in [0, 1]$.
 Eigenvalues associated with each $\gamma$ distribution are compared with otherwise identical communities in which $\gamma = 1$ (pink points in all sub-panels).

### 2018-04-20 R/plot_figs.R
New plot of big results

@@ -269,10 +269,90 @@ pr_destabilised_plot <- function(dat){
           line = 3.5);
 }
 
-
-
-
-
+stable_N_plot <- function(dat, S = 32){
+    Ns          <- 1:S;
+    par(mfrow = c(2, 2), oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("red", "black"),
+            names.arg = dat[Ns,1], ylim = c(0, 15), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1.25, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1,
+            xaxt = "n", yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:32,12] / (dat[1:32,3] + dat[1:32,12]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = 32);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1.25, col = "blue");
+    text(x = x1[31], y = 0.93, labels = "a", cex = 4);
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,7]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("red", "black"),
+            names.arg = dat[Ns,1], ylim = c(0, 15), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1.25, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1,
+            xaxt = "n", yaxt = "n");
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:32,13] / (dat[1:32,3] + dat[1:32,13]);
[…]

### 2018-04-20 R/test_mat.R

@@ -1544,8 +1544,11 @@ abline(v = 0, lwd = 2, col = "orange");
 
 
 
+12 16
 
 
+dat[,12] / (dat[,2] + dat[,12]);
+dat[,16] / (dat[,3] + dat[,1]);
 
 
 

### 2018-04-20 R/evolve_gammas.R
New code

@@ -1,7 +1,6 @@
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, eps_max = 999){
+Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
     tot_res <- NULL;
     fea_res <- NULL;
-    nea_res <- NULL;
     for(i in 2:max_sp){
         nn             <- i;
         A1_stt         <- 0;
@@ -12,39 +11,32 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, eps_max = 999){
         tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
         fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
         while(iter > 0){
-            r_vec    <- rnorm(n = nn, mean = 0, sd = rmx);
-            A1_dat   <- rnorm(n = nn * nn, mean = 0, sd = 0.4);
-            A1       <- matrix(data = A1_dat, nrow = nn, ncol = nn);
-            A1       <- species_interactions(mat = A1, type = int_type);
-            diag(A1) <- -1;
-            epsil    <- runif(n = nn, min = 1, max = eps_max);
-            eps_dat  <- rep(x = epsil, times = nn);
-            eps_mat  <- matrix(data = epsil, nrow = nn, ncol = nn, 
-                               byrow = FALSE);
-            avg_mat  <- matrix(data = mean(1:eps_max), nrow = nn, ncol = nn, 
-                               byrow = FALSE);
-            A2       <- A1 * eps_mat;
-            A3_stb   <- rand_mat_ga(A1);
-            A1       <- A1 * avg_mat;
-            
-            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
-            A2_stb   <- max(Re(eigen(A2)$values)) < 0;
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
+            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            A0       <- A0 * C_mat;
[…]

### 2018-04-20 R/plot_figs.R

@@ -269,8 +269,8 @@ pr_destabilised_plot <- function(dat){
           line = 3.5);
 }
 
-stable_N_plot <- function(dat, S = 32){
-    Ns          <- 1:S;
+stable_N_plot <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
     par(mfrow = c(2, 2), oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
     #=================================
     bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
@@ -286,10 +286,11 @@ stable_N_plot <- function(dat, S = 32){
     par(new = TRUE);
     y1     <- dat[1:32,12] / (dat[1:32,3] + dat[1:32,12]);
     x1     <- seq(from = 2.132, to = 15.1112, length = 32);
-    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
          xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
-         cex = 1.25, col = "blue");
-    text(x = x1[31], y = 0.93, labels = "a", cex = 4);
+         cex = 1, col = "blue", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "blue");
+    text(x = x1[30], y = 0.92, labels = "a", cex = 3);
     #=================================
     bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,7]));
     log_bar_dat                  <- log(bar_dat);
@@ -303,11 +304,12 @@ stable_N_plot <- function(dat, S = 32){
     par(new = TRUE);
     y1     <- dat[1:32,13] / (dat[1:32,3] + dat[1:32,13]);
     x1     <- seq(from = 2.132, to = 15.1112, length = 32);
-    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1),
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
          xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
-         cex = 1.25, col = "blue");
[…]

### 2018-04-20 ms.Rmd
Figure 3 stuff

@@
 
+********************************************************************************
+
 **References**
@@
 ********************************************************************************
+
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** Panels show the $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+For each $S$, 1 million complex systems are randomly generated and tested for stability before and after variation in $\gamma$ is included at four different distributions: **a**, uniform, **b**, exponential, **c**, beta, and **d**, gamma.
+Red and black bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
+Blue lines show the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 8}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+stable_N_plot(dat);
+```

### 2018-04-21 ms.Rmd
Some more references and how to use them

@@
 Future research in complex systems should therefore account for the varying response rates of individual system components when assessing whole system stability.**
+
+<!---
+
+@Koch1999 Suggests that the results of @May1972 can apply to neuroscience
+
+@May2008 Suggests that random matrix theory can apply to banking
+--->
+
+<!---
+
+2.
+Go back to @May1972 and note how the original ecological result has been broadened for systems such as evolutionary networks, gene regulatory networks, economics, etc.
+(references needed).
+Perhaps make the broader point that even local systems have different elements of different types of systems.
+3.
+Briefly point out the advances that have been made in terms of network structures (e.g., random versus competitive interactions), but then hit the point that different components of complex systems are expected to respond on different time scales [e.g.
+@Patel2018], and no one has looked at how this might affect stability of systems more broadly.
+The current paper is going to show that variation in component response rate increases stability.
+4. A paragraph with some broad explanation of the idea of matrix stability, along the lines of @Allesina2012. This will effectively explain the idea of random matrix theory to the reader.
+5.
+A paragraph introducing, mathematically, how I am adding varying rates of component response -- multipling each row $i$ by a scalar $\gamma_{i}$.
+First this will be done with some simple function giving $Var(\gamma_{i})$, then in a more structured way such that random components that are more likely to be positively influenced by other components having faster or slower rates (higher or lower $\gamma_{i}$) of response.
+A sentence or two is needed to suggest what this might mean biologically (e.g., generalist predators have longer generation times than their prey).
+6.
+A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in supporting information.
+The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.
+I could then argue that $Var(\gamma_{i})$ has a minimum such correlation of zero, and can only increase in stabilisation by having an absolute value of correlation being greater than zero).
+At least two [key figures](#fig_ideas) will be included, likely with four panels (1.
+without variation, 2.
+with unstructured variation, and 3.
+and 4.
+with structured variation).
+7.
+Some sort of mention is needed for a couple odd results; first, that in mutualist networks, stability is not affected -- perhaps because of the lack of variance in component response rate given number of interaction types (note, I think, competitive networks can have indirect positives due to intransitive effects).
+Second, *feasibility* is not increased, so pure ecological networks might not be affected by this (but emphasise that *pure* ecological networks don't really exist in nature -- they are incorporated into evolutionary, economic, social, systems).
+8.
+Some bold statement is needed here -- something to the effect of the argument that nearly all large stable complex systems will be stable *due to* the variation simulated in this paper.
+The odds that it would be stable without such variation becomes vanishingly small as system size increases.
+
+--->
 

### 2018-04-22 ms.Rmd
Add mroe comments

@@
 @May2008 Suggests that random matrix theory can apply to banking
+
+Download: Jansen, V. A., & Kokkoris, G. D. (2003). Complexity and stability revisited. Ecology Letters, 6(6), 498-502.
+
+Hastings, H. M. (1982). The May-Wigner stability theorem. Journal of Theoretical Biology, 97(2), 155-166.
+
+Maybe see how variance in gamma modifies the general findings of May with respect to connectivity and interaction strength?
+-- Link to variance in off diagonals, which Tang and Allesina (frontiers 2014) finds to be important.
+
+Gray, R. T., & Robinson, P. A. (2008). Stability and synchronization of random brain networks with a distribution of connection strengths. Neurocomputing, 71(7-9), 1373-1387.
+
+Gray & Robinson 2009 Stability of random brain networks with excitatory and inhibitory connections. Neurocomputing
+
+Need to consider that the off-diagonal elements size relative to the diagonal size will affect stability -- higher relative values make things *less* stable.
+I think this is okay because we're multiplying diagonals and off diagonals by the same values -- this is why if we multiply the whole matrix by a scalar it only affects the magnitude and not the sign of eigenvalues.
+
+**Idea**: Since the *variance* of the off-diagonal elements (relative to the diagonals?
+Perhaps the diagonal values?) affects stability -- increased variance increases stability, the relevance of $\gamma$ increases with increasing system size because the variance also increases -- more off-diagonal elements relative to diagonals, and we're essentially adding a new source of variation, $\gamma$
+
+**Idea**: Look at what happens to feasibility when r < 0 for all species.
+Could even potentially make the argument that this is *expected* for real communities, because in the absence of other components of the system (producers, nutrients, whatever), the growth rate of a population *must* be negative?
+
+Note further, that Allesina report on the criteria of May that \sigma \sqrt{SC} < 1, meaning that \sigma needs to be low for the random matrix to be stable.
+Hence, the multiplying rows by $\gamma$ is increasing the stability overall by decreasing the variance?
+This doesn't seem to make sense though -- really it's adding an element of variation, Var(inherent) + Var(due to \gamma).
+But, also maybe should think about gamma in relation to element values.
+If, e.g., components with similar gammas tend to be more likely to interact, then this might be stabilising?
+**VARIANCE APPEARS HIGHER GIVEN VARIATION IN gamma**
+
+**Potential solution** -- The increase in variation caused by gamma creates a situation of sometimes increasing but sometimes decreasing the correlation between pairs of off diagonal elements, and sometimes increasing but sometimes decreasing total off-diagonal variation.
+This increase creates more extreme total variations, some of which cause leading eigenvalues to all be negative.
+Note that the probability of stability versus instability would be affected by creating disproportionately more extreme variation (high or low) communities that are sometimes stable?
+**THIS DOES NOT APPEAR TO BE THE CASE BASED ON SIMULATION**
 --->

### 2018-04-22 R/scratch.R
More scratch to figure out details

@@ -1,15 +1,16 @@
 
 rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1, 
                          gamma_mn = 1, C = 1){
-  tot_res <- NULL;
-  fea_res <- NULL;
-  vars_t  <- NULL;
+  tot_res  <- NULL;
+  fea_res  <- NULL;
+  vars_t   <- NULL;
+  off_cors <- NULL;
   for(i in 2:max_sp){
     iter           <- iters;
     tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
     fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
     while(iter > 0){
-      r_vec    <- runif(n = i, min = -1, max = 0); #rnorm(n = i, mean = 0, sd = rmx);
+      r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
       A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
       A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
       A0       <- species_interactions(mat = A0, type = int_type);
@@ -29,8 +30,8 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1,
       A1       <- A0 * gam1;
       A2       <- A0 * gam1;
       A3       <- A0 * gam0;
-      diag(A2) <- gam0;
-      diag(A3) <- gam1;
+      diag(A2) <- -1 * gam0;
+      diag(A3) <- -1 * gam1;
       
 
       A4       <- A0 * gam4;
@@ -78,18 +79,31 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1,
       if(A4_fea == TRUE){
         fea_res[[i-1]][iter, 5] <- 1;
       }
-      v0      <- get_var_A(A0);
-      v1      <- get_var_A(A1);
-      v2      <- get_var_A(A2);
-      v3      <- get_var_A(A3);
-      v4      <- get_var_A(A4);
-      vars_t  <- rbind(vars_t, c(i, v0, v1, v2, v3, v4));
+      
+      stbs     <- c(A0_stb, A1_stb, A2_stb, A3_stb, A4_stb);
+      
+      v0       <- get_var_A(A0);
+      v1       <- get_var_A(A1);
+      v2       <- get_var_A(A2);
+      v3       <- get_var_A(A3);
[…]

### 2018-04-24 R/scratch.R
Restructure for the new simplified presentation

@@ -642,10 +642,16 @@ points(x = x0, y = y0, type = "l", col = "red", lwd = 3)
 
 
 
+
+
+
 A_comp <- NULL;
 A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
 A_mat  <- matrix(data = A_dat, nrow = 1000);
-gammas <- runif(n = 1000, min = 0, max = 1);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+gammas <- runif(n = 1000, min = 0, max = 2);
 mu_gam <- mean(gammas);
 diag(A_mat) <- -1;
 A1     <- gammas * A_mat;
@@ -667,20 +673,20 @@ A1vec       <- as.vector(A1_vm);
 A1vec       <- A1vec[is.na(A1vec) == FALSE];
 
 par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-8.75, 8.25), ylim = c(-8.5,8.5), pch = 4, cex = 0.7,
+plot(A0_r, A0_i, xlim = c(-16.75, 16.25), ylim = c(-16.5,16.5), pch = 4, cex = 0.7,
      xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
 vl <- seq(from = 0, to = 2*pi, by = 0.001);
 x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
 y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
 x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
 y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
-text(x = -9.5, y = 8, labels = "a", cex = 2);
+text(x = -19, y = 14, labels = "a", cex = 2);
 points(x = x0, y = y0, type = "l", lwd = 3);
 points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-plot(A1_r, A1_i, xlim = c(-8.75, 8.25), ylim = c(-8.5,8.5), pch = 4, cex = 0.7,
[…]

### 2018-04-24 R/sim_mat.R

@@ -17,35 +17,9 @@ species_interactions <- function(mat, type = 0){
     return(mat);
 }
 
-make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
-    if(distribution == 0){
-        dat          <- rep(x = mn, times = nn);
-    }
-    if(distribution == 1){
-        mval         <- sdd * sqrt(12);
-        dat          <- runif(n = nn, min = 0, max = mval);
-    }
-    if(distribution == 2){
-        dat          <- rexp(n = nn);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 3){
-        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 4){
-        dat          <- rgamma(n = nn, shape = 2, scale = 2);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    return(dat);
-}
-
 summarise_randmat <- function(tot_res, fea_res){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 37);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
     for(i in 1:sims){
         all_res[i, 1]  <- i + 1;
         # Stable and unstable
@@ -53,75 +27,31 @@ summarise_randmat <- function(tot_res, fea_res){
         all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
         all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
         all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
-        all_res[i, 6]  <- sum(tot_res[[i]][,3] == FALSE);
-        all_res[i, 7]  <- sum(tot_res[[i]][,3] == TRUE);
-        all_res[i, 8]  <- sum(tot_res[[i]][,4] == FALSE);
-        all_res[i, 9]  <- sum(tot_res[[i]][,4] == TRUE);
-        all_res[i, 10] <- sum(tot_res[[i]][,5] == FALSE);
-        all_res[i, 11] <- sum(tot_res[[i]][,5] == TRUE);
[…]

### 2018-04-24 ms.Rmd
Rework the abstract and figures

@@
 **The stability of a complex system generally decreases with increasing system size, as is demonstrated by random matrix theory [@May1972; @Allesina2012].
-This counter-intuitive result, first shown by May [@May1972], is broadly relevant for understanding the dynamics and persistence of systems such as chemical, biological, and social networks.
-Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Haldane2011; @Allesina2012; @Mougi2012; @Allesina2015a; @Gao2016; @Grilli2017; @Patel2018].
+This counter-intuitive result, first shown by May [@May1972], is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010] and socio-economic [@Haldane2011; @Suweis2014; @Bardoscia2017] networks.
+Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Allesina2012; @Mougi2012; @Allesina2015a; @Gao2016; @Grilli2017; @Patel2018].
 But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains unexplored.
-Here I show that when components of a complex system respond to system dynamics at different rates, system stability is markedly increased.
-This increase in stability is especially strong when variation in component response rate ($\gamma$) is high and has a skewed distribution.
-Further, these effects become increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for variation in $\gamma$.
+Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), system stability is markedly increased.
+Stability increases despite higher variation among interaction strengths ($\boldsymbol{\sigma^{2}}$) due to the clustering of eigenvalues at the centre of eigenvalue distributions.
+This effect of variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for $\boldsymbol{Var(\gamma)}$.
 My results therefore reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
 Future research in complex systems should therefore account for the varying response rates of individual system components when assessing whole system stability.**
+
+<!---
+
+2. Go back to @May1972 and explain the basic idea of random matrices as representing communities.
+
+3.
+A paragraph with some broad explanation of the idea of matrix stability, along the lines of @Allesina2012.
+This will effectively explain the idea of random matrix theory to the reader.
+Explain the general conclusion that stability decreases whenever $\sigma\sqrt{SC} > 1$.
+Note how the original ecological result has been broadened for different types of networks (e.g., nestedness, predator-prey, mutualist, etc.),
+
+4.
+But none of these studies have considered that different components operate on different time scales (but see [e.g.
+@Patel2018]).
+Then explain that this, intuitively, should decrease stability because it introduces another source of variation and therefore increases $\sigma$.
+But then explain that it also changes the distribution of eignvalues, bringing them towards the centre of the plot (figure 1).
+I show that this actually increases the stability of complex systems.
+
+5.
+A paragraph introducing, mathematically, how I am adding varying rates of component response -- multipling each row $i$ by a scalar $\gamma_{i}$.
+First this will be done with some simple function giving $Var(\gamma_{i})$, then in a more structured way such that random components that are more likely to be positively influenced by other components having faster or slower rates (higher or lower $\gamma_{i}$) of response.
+A sentence or two is needed to suggest what this might mean biologically (e.g., generalist predators have longer generation times than their prey).
+
+6.
+A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in supporting information.
+The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.
+I could then argue that $Var(\gamma_{i})$ has a minimum such correlation of zero, and can only increase in stabilisation by having an absolute value of correlation being greater than zero).
+At least two [key figures](#fig_ideas) will be included, likely with four panels (1.
+without variation, 2.
+with unstructured variation, and 3.
+and 4.
+with structured variation).
+
+7.
+Some sort of mention is needed for a couple odd results; first, that in mutualist networks, stability is not affected -- perhaps because of the lack of variance in component response rate given number of interaction types (note, I think, competitive networks can have indirect positives due to intransitive effects).
+Second, *feasibility* is not increased, so pure ecological networks might not be affected by this (but emphasise that *pure* ecological networks don't really exist in nature -- they are incorporated into evolutionary, economic, social, systems).
+
+8.
+Some bold statement is needed here -- something to the effect of the argument that nearly all large stable complex systems will be stable *due to* the variation simulated in this paper.
+The odds that it would be stable without such variation becomes vanishingly small as system size increases.
+
+--->
 
@@
 
-**Idea**: Since the *variance* of the off-diagonal elements (relative to the diagonals?
-Perhaps the diagonal values?) affects stability -- increased variance increases stability, the relevance of $\gamma$ increases with increasing system size because the variance also increases -- more off-diagonal elements relative to diagonals, and we're essentially adding a new source of variation, $\gamma$
-
-**Idea**: Look at what happens to feasibility when r < 0 for all species.
-Could even potentially make the argument that this is *expected* for real communities, because in the absence of other components of the system (producers, nutrients, whatever), the growth rate of a population *must* be negative?
-
 Note further, that Allesina report on the criteria of May that \sigma \sqrt{SC} < 1, meaning that \sigma needs to be low for the random matrix to be stable.
@@
 
-<!---
-
-2.
-Go back to @May1972 and note how the original ecological result has been broadened for systems such as evolutionary networks, gene regulatory networks, economics, etc.
-(references needed).
-Perhaps make the broader point that even local systems have different elements of different types of systems.
-3.
-Briefly point out the advances that have been made in terms of network structures (e.g., random versus competitive interactions), but then hit the point that different components of complex systems are expected to respond on different time scales [e.g.
-@Patel2018], and no one has looked at how this might affect stability of systems more broadly.
-The current paper is going to show that variation in component response rate increases stability.
-4. A paragraph with some broad explanation of the idea of matrix stability, along the lines of @Allesina2012. This will effectively explain the idea of random matrix theory to the reader.
-5.
-A paragraph introducing, mathematically, how I am adding varying rates of component response -- multipling each row $i$ by a scalar $\gamma_{i}$.
-First this will be done with some simple function giving $Var(\gamma_{i})$, then in a more structured way such that random components that are more likely to be positively influenced by other components having faster or slower rates (higher or lower $\gamma_{i}$) of response.
-A sentence or two is needed to suggest what this might mean biologically (e.g., generalist predators have longer generation times than their prey).
-6.
-A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in supporting information.
-The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.
-I could then argue that $Var(\gamma_{i})$ has a minimum such correlation of zero, and can only increase in stabilisation by having an absolute value of correlation being greater than zero).
-At least two [key figures](#fig_ideas) will be included, likely with four panels (1.
-without variation, 2.
-with unstructured variation, and 3.
-and 4.
-with structured variation).
-7.
-Some sort of mention is needed for a couple odd results; first, that in mutualist networks, stability is not affected -- perhaps because of the lack of variance in component response rate given number of interaction types (note, I think, competitive networks can have indirect positives due to intransitive effects).
-Second, *feasibility* is not increased, so pure ecological networks might not be affected by this (but emphasise that *pure* ecological networks don't really exist in nature -- they are incorporated into evolutionary, economic, social, systems).
-8.
-Some bold statement is needed here -- something to the effect of the argument that nearly all large stable complex systems will be stable *due to* the variation simulated in this paper.
-The odds that it would be stable without such variation becomes vanishingly small as system size increases.
-
---->
-
 ********************************************************************************
@@
 
-**Figure 1: Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$).
-Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$).
-Each panel shows 1 million randomly generated $\gamma$ values.
+**Figure 1: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = d = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
+Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
 
-```{r, eval = TRUE, echo = FALSE, fig.height = 8}
-show_gammas();
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+gammas <- runif(n = 1000, min = 0, max = 2);
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -15.5, y = 19, labels = "a", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+text(x = -15.5, y = 19, labels = "b", cex = 2);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
 ```
@@
 
-```{r, eval = TRUE, echo = FALSE}
-eig_cloud <- eigen_cloud(sp = 20, iters = 10000);
-Pr_D0     <- sum(eig_cloud$d0[,1] < 0) / length(eig_cloud$d0[,1]);
-Pr_D1     <- sum(eig_cloud$d1[,1] < 0) / length(eig_cloud$d1[,1]);
-Pr_D2     <- sum(eig_cloud$d2[,1] < 0) / length(eig_cloud$d2[,1]);
-Pr_D3     <- sum(eig_cloud$d3[,1] < 0) / length(eig_cloud$d3[,1]);
-Pr_D4     <- sum(eig_cloud$d4[,1] < 0) / length(eig_cloud$d4[,1]);
-```
-
-**Figure 2: Distributions of eigenvalues across different distributions of component response rate  ($\boldsymbol{\gamma}$) values.** Panels show 100000 random communities for $S = 20$ and $C = 1$, where off-diagonal elements are randomly sampled from a standard normal distribution $N(0, 1)$, diagonal elements are $-1$, and each row of elements is multiplied by a random scalar value $\gamma$ sampled from one of four distributions (black points in all sub-panels):  **a**, uniform $\Pr(\gamma = x) = 1/\sqrt{12}$ where $x \in [0, \sqrt{12}]$.
-**b**, exponential $\Pr(\gamma = x) = e^{-x}$ where $x \geq 0$.
-**c**, beta $\Pr(\gamma = x) = \frac{x^{\alpha - 1}(1-x)^{\beta - 1}}{B(\alpha, \beta)}$ where $x \in [0, 1]$, and where $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$ for $\alpha = 0.5$ and $\beta = 0.5$.
-**d**, gamma $\Pr(\gamma = x) = \frac{1}{\Gamma(k)\theta^{k}}x^{k-1}e^{\frac{-x}{\theta}}$ for $k = 2$ and $\theta = 2$, where $x \in [0, 1]$.
-Eigenvalues associated with each $\gamma$ distribution are compared with otherwise identical communities in which $\gamma = 1$ (pink points in all sub-panels).
-In all distributions, $SD(\gamma) = 1$; note that $\gamma$ distributions cannot be mean standardised without changing the sign of some community matrix elements, but the sign of eigenvalues (and therefore community stability) is unaffected by multiplying elements by any scalar value.
-Proportions of real components of eigenvalues (x-axis) below 0 are (a) `r Pr_D1`, (b) `r Pr_D2`, (c) `r Pr_D3`, (d) `r Pr_D4`, as compared to `r Pr_D0` when $\gamma = 1$.
+**Figure 2: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+For each $S$, 1 million complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
+Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
+Red and black bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
+Blue lines show the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
 
 ```{r, eval = TRUE, echo = FALSE, fig.height = 8}
-eig_plot(ecl = eig_cloud);
+#dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+#stable_N_plot(dat);
 ```
@@
 
-**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** Panels show the $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
-For each $S$, 1 million complex systems are randomly generated and tested for stability before and after variation in $\gamma$ is included at four different distributions: **a**, uniform, **b**, exponential, **c**, beta, and **d**, gamma.
-Red and black bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
-Blue lines show the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 8}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-stable_N_plot(dat);
-```
+**Figure 3: Distributions of variation in component response time evolved to generate stable systems**

### 2018-04-25 R/plot_figs.R
New plotting function

@@ -360,6 +360,37 @@ stable_N_plot <- function(dat, S_s = 32){
 
 
 
-
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("grey80", "grey40"),
+            names.arg = dat[Ns,1], ylim = c(0, 15), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c("E[\u03b3]", "Var(\u03b3)"), pch=15, 
+           col=c("grey80","grey40"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
 
 

### 2018-04-25 R/plot_figs.R
Small plot change

@@ -368,10 +368,10 @@ plot_stables <- function(dat, S_s = 32){
     log_bar_dat                  <- log(bar_dat);
     log_bar_dat[log_bar_dat < 0] <- 0; 
     barplot(log_bar_dat, beside = TRUE, col = c("grey80", "grey40"),
-            names.arg = dat[Ns,1], ylim = c(0, 15), xlab = "",
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
             ylab = "Ln number of stable communities", cex.lab = 1, 
             cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
-    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
     box(lwd = 2);
     par(new = TRUE);
     y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);

### 2018-04-25 ms.Rmd
New Figure
Got a second paragraph

@@
 
+In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations.
+Complexity in this case is defined by the size of the system (i.e., the number of interacting components; $S$), its inter-connectivity (i.e., the probability that one component will affect another; $C$), and the variance of interaction strengths ($\sigma^{2}$)[@Allesina2012].
+May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ has profound consequences across multiple disciplines, raising the question of how complex systems in, e.g., ecology[@Mougi2012; @Allesina2012; @Allesina2015; @Grilli2017] or banking [@May2008; @Haldane2011; @Bardoscia2017] are predicted to persist or change.
+
 <!---
+
+Part of the answer might be that communities aren't actually random ....
 
@@
 
-**Figure 1: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = d = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
-Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+```{r, echo = FALSE}
 A_comp <- NULL;
@@
 A1vec       <- A1vec[is.na(A1vec) == FALSE];
+```
 
+**Figure 1: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = d = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
+Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
+Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
 par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
@@
 
-```{r, eval = TRUE, echo = FALSE, fig.height = 8}
-#dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-#stable_N_plot(dat);
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+plot_stables(dat);
 ```

### 2018-04-26 ms.Rmd
Cite Allesina as saying that variance in diagonals is understudied -- based on the 2015 SI

@@
 Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Allesina2012; @Mougi2012; @Allesina2015a; @Gao2016; @Grilli2017; @Patel2018].
-But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains unexplored.
+But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem [@Allesina2015].
 Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), system stability is markedly increased.

### 2018-04-28 R/scratch.R
Change some script in a small way

@@ -650,7 +650,9 @@ A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
 A_mat  <- matrix(data = A_dat, nrow = 1000);
 C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
 C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
-A_mat     <- A_mat * C_mat;
+A_mat  <- A_mat * C_mat;
+#gammas <- c(20, rep(1, 999));
+#gammas <- c(rep(1.5, 500), rep(0.5, 500))
 gammas <- runif(n = 1000, min = 0, max = 2);
 mu_gam <- mean(gammas);
 diag(A_mat) <- -1;

### 2018-04-28 R/sim_mat.R

@@ -68,6 +68,9 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
             A0       <- A0 * C_mat;
             diag(A0) <- -1;
             gam1     <- runif(n = i, min = 0, max = 2);
+            #if(i %% 2 == 0){
+            #  gam1 <- c(rep(1.5, i/2), rep(0.5, i/2));
+            #}
             A1       <- A0 * gam1;
             A0       <- A0 * mean(gam1);
             A0_stb   <- max(Re(eigen(A0)$values)) < 0;

### 2018-04-28 random_matrix_stability.Rmd
New notes

@@
 ********************************************************************************
+
+**UPDATE 28 APR**:
+
+I have made some more work on [the manuscript](https://bradduthie.github.io/random_matrix_stability/ms.html).
+The different distributions of $\gamma$ are now not the central focus.
+Instead, I want to focus more on the distribution of eigenvalues and allow a uniform distribution $\gamma \sim \mathcal{U}(0, 2)$.
+I want to show the eigenvalues plotted out and go into a bit more detail of what is actually driving these distributions.
+The figure below shows one plot **a** where $\gamma = E[\gamma] \approx 1$ and one plot **b** where $\gamma \sim \mathcal{U}(0, 2)$.
+The sum of the eigenvalues needs to equal the sum of the rows [@Allesina2012], so the sum of the eigenvalues is identical in each.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+gammas <- runif(n = 1000, min = 0, max = 2);
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -15.5, y = 19, labels = "a", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+text(x = -15.5, y = 19, labels = "b", cex = 2);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+It's worth noting that the **b** in the figure above resembles the power law graph of @Allesina2015, which also has variation in the distribution of values due to the structure of connectivity ($C$).
+It's also worth looking at a what happens when only 1 $\gamma$ out of the 1000 total is different.
+The plot on below in subpanel **a** shows what happens when one $\gamma = 18$ and the rest are $\gamma = 1 - 17/999$ (causing $E[\gamma] = 1$).
+The plot in subpanel **b** shows what happens when one $\gamma = 1/18$ and the rest are $\gamma = 1 + (17/18)/999$ (also causing $E[\gamma] = 1$).
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat  <- A_mat * C_mat;
+#gamma1 <- c(4,   rep(1 - 1/333,  999));
+#gamma2 <- c(1/4, rep(1 + 1/1332, 999));
+gamma1 <- c(18,   rep(1 - 17/999,  999));
+gamma2 <- c(1/18, rep(1 + (17/18)/999, 999));
+#gamma1 <- c(40,   rep(1 - 13/333,  999));
+#gamma2 <- c(1/40, rep(1 + (39/40)/999, 999));
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gamma1 * A_mat;
+A0     <- gamma2 * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-20.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -15.5, y = 16, labels = "a", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-20.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+text(x = -15.5, y = 16, labels = "b", cex = 2);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+This does not add a whole lot of insight into what is happening, although it does show a couple eigenvalues leaving the circle in **b**.
+In **a**, the lowest real eigenvalue is `r min(A0_r)` and the highest is `r max(A0_r)`, while in **b**, the lowest real eigenvalue is `r min(A1_r)` and the highest is `r max(A1_r)`.
+More revealing is to consider a situation where only two values of $\gamma$ exist: 0.05 and 1.95.
+By selecting half $\gamma$ to be each, it becomes a bit clearer to see what is going on.
+In the
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat  <- A_mat * C_mat;
+gammas <- c(rep(0.05, 500), rep(1.95, 500));
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-19.5, 15.5), ylim = c(-19.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -16.5, y = 14, labels = "a", cex = 2);
+abline(v = 0, col = "blue");
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-19.5, 15.5), ylim = c(-19.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+text(x = -16.5, y = 14, labels = "b", cex = 2);
+abline(v = 0, col = "blue");
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+I have added the blue vertical line where the x-axis equals zero.
+In **a**, the lowest real eigenvalue is `r min(A0_r)` and the highest is `r max(A0_r)`, while in **b**, the lowest real eigenvalue is `r min(A1_r)` and the highest is `r max(A1_r)`.
+But note the clump in **b**.
+Instead of being uniform, there are low values and high values.
+Overall, there are `r sum(A0_r < 0)` eigenvalues in **a** where the real component is less than zero, and there are `r sum(A1_r < 0)` eigenvalues in **b** where the real component is less than zero.
+Overall, because the mean is forced such that $E[\Re{(\lambda_{1})}] = 1$ (the mean on the x-axis is exactly -1 in both panels), the smaller circle on the inside created where the subset of $\gamma$ values is very low has a small probability of pulling real eigenvalues to the left of 0, and in smaller systems (lower $S$), these eigenvalues can be the leading one, thereby stabilising the system.
 
@@
 library(knitr);
-dat  <- read.table(file = "mat_dat.csv", header = TRUE, sep = ",");
+dat  <- read.table(file = "sim_results_old/oldest/mat_dat.csv", 
+                   header = TRUE, sep = ",");
 dat  <- cbind(dat, (dat[,3] - dat[,5]) + dat[,4]);
@@
 ```{r, echo = FALSE, eval = TRUE}
-mag_var <- read.table(file = "var_gen_mag.csv", header = TRUE, sep = ",");
+mag_var <- read.table(file = "sim_results_old/oldest/var_gen_mag.csv", 
+                      header = TRUE, sep = ",");
 stabi_3 <- mag_var[,3] / 10000000;

### 2018-04-28 random_matrix_stability.Rmd
Add a bunch of stuff, and more notes

@@
 diag(A_mat) <- -1;
-A1     <- gamma1 * A_mat;
-A0     <- gamma2 * A_mat;
+A0     <- gamma1 * A_mat;
+A1     <- gamma2 * A_mat;
 A0_e   <- eigen(A0)$values;
@@
 Overall, because the mean is forced such that $E[\Re{(\lambda_{1})}] = 1$ (the mean on the x-axis is exactly -1 in both panels), the smaller circle on the inside created where the subset of $\gamma$ values is very low has a small probability of pulling real eigenvalues to the left of 0, and in smaller systems (lower $S$), these eigenvalues can be the leading one, thereby stabilising the system.
+
+Note that the sum of the real components of all eigenvalues needs to equal the sum of the diagonal of a matrix [@Allesina2015].
+So we can break the two haves of the matrix represented by **b** above into the upper half where the sum equals `r sum(diag(A1[1:500, 1:500]))`, and the lower half where the sum equals `r sum(diag(A1[501:1000, 501:1000]))`.
+We can then partition the eigenvalues in each half, $\lambda_{upper}$ and $\lambda_{lower}$.
+
+We can also note that the values of the off diagonals in the upper half of the matrix are multiplied by 1.95, so $\sigma_{upper} = \sigma(1.95) = 1.95$ and $\sigma_{lower} = \sigma(0.05) = 0.05$.
+This effectively creates two circles, one with a radius of $1.95\sqrt{SC}$; this is shown explicitly below with the red circles.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat  <- A_mat * C_mat;
+gammas <- c(rep(0.05, 500), rep(1.95, 500));
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-19.5, 16.5), ylim = c(-19.5,16.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -16.5, y = 14, labels = "a", cex = 2);
+abline(v = 0, col = "blue");
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-19.5, 16.5), ylim = c(-19.5,16.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vecU      <- as.vector(A1_vm[1:500,]);
+A1vecU      <- A1vecU[is.na(A1vecU) == FALSE];
+A1vecL      <- as.vector(A1_vm[501:1000,]);
+A1vecL      <- A1vecL[is.na(A1vecL) == FALSE];
+x0 <- sqrt(500) * sd(A1vecU) * cos(vl) + mean(diag(A1[1:500,1:500]));
+y0 <- sqrt(500) * sd(A1vecU) * sin(vl);
+x1 <- sqrt(500) * sd(A1vecL) * cos(vl) + mean(diag(A1[501:1000,501:1000]));
+y1 <- sqrt(500) * sd(A1vecL ) * sin(vl);
+text(x = -16.5, y = 14, labels = "b", cex = 2);
+abline(v = 0, col = "blue");
+points(x = x1, y = y1, type = "l", col = "orange", lwd = 2)
+points(x = x0, y = y0, type = "l", lwd = 2, lty = "solid", col = "orange");
+
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+x3 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y3 <- sqrt(1000) * sd(A1vec) * sin(vl);
+points(x = x3, y = y3, type = "l", col = "red", lwd = 3, lty = "dashed");
+
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+Note that the large orange is pushed leftward (red dashed circles are identical), though I am not entirely sure why yet.
+It should be noted that the upper-left and lower-right squares of the matrix should be partitionable.
+The remaining squares have off-diagonals pulled from matching distributions, so we should really be able to think of the whole thing as a combination of the upper and lower rows, I think.
+
+Also note how the smaller circle is pulled rightwards (centre = `r mean(x0)`) while the larger circle is pulled leftwards (centre = `r mean(x1)`), while the overall centre remains the same (red dashed circle, centre = `r mean(x3)`).
 

### 2018-05-17 ms.Rmd
Add new Figure

@@
 ```{r, echo = FALSE}
+A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
+A1 <- as.matrix(A1[,-1]);
+
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(t(A0_vm));
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(t(A1_vm));
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+fhalf       <- 1:(0.5*length(A1vec));
+shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st       <- pr_st[,-1];
+```
+
+**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
+**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
+In **a**, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in **b** $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in **b** but not **a** is stable.
+In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
+Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
+A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
+text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
+points(A0_r, A0_i, pch = 4, cex = 0.7);
+
+plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "black", yaxt = "n");
+
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
+A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
+A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
+points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+
+points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);
+
+text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+********************************************************************************
+
+```{r, echo = FALSE}
 A_comp <- NULL;
@@
 
-**Figure 1: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
 **a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = d = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
@@
 
-**Figure 2: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
@@
 
-**Figure 3: Distributions of variation in component response time evolved to generate stable systems**
+**Figure 4: Distributions of variation in component response time evolved to generate stable systems**

### 2018-05-17 R/find_bi_gamma.R
Add function and plotting for bimodal gamma systems

@@ -1,3 +1,38 @@
+stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
+    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
+    A0_count <- 0;
+    A1_count <- 0;
+    while(iters > 0){
+        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
+        A_mat  <- matrix(data = A_dat, nrow = S);
+        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+        A_mat  <- A_mat * C_mat;
+        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
+        mu_gam <- mean(gammas);
+        diag(A_mat) <- -1;
+        A1     <- gammas * A_mat;
+        A0     <- mu_gam * A_mat;
+        A0_e   <- eigen(A0)$values;
+        A0_r   <- Re(A0_e);
+        A0_i   <- Im(A0_e);
+        A1_e   <- eigen(A1)$values;
+        A1_r   <- Re(A1_e);
+        A1_i   <- Im(A1_e);
+        if(max(A0_r) < 0){
+            ress[iters, 1] <- 1;
+            A0_count       <- A0_count + 1;
+            }
+        if(max(A1_r) < 0){
+            ress[iters, 2] <- 1;
+            A1_count       <- A1_count + 1;
+        }
+        print(c(iters, A0_count, A1_count));
+        iters <- iters - 1;
+    }
+    return(ress);
+}
+
 
 find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
     while(iters > 0){
@@ -67,6 +102,7 @@ plot_eigs <- function(A0, A1){
     mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
     mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
 }
+
 
 
 

### 2018-05-17 R/plot_figs.R

@@ -1,4 +1,4 @@
-make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+ make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
     if(distribution == 0){
         dat          <- rep(x = mn, times = nn);
     }

### 2018-05-17 ms.Rmd
Minor change to Figure 1 caption

@@
 **b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
-In **a**, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in **b** $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in **b** but not **a** is stable.
+In **a**, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in **b** $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in **b** but not **a** is stable because in **b** $\max\left(\Re(\lambda)\right) < 0$.
 In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.

### 2018-05-18 ms.Rmd
More writing

@@
 Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), system stability is markedly increased.
-Stability increases despite higher variation among interaction strengths ($\boldsymbol{\sigma^{2}}$) due to the clustering of eigenvalues at the centre of eigenvalue distributions.
+Stability increases due to the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher variation among interaction strengths ($\boldsymbol{\sigma^{2}}$).
 This effect of variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for $\boldsymbol{Var(\gamma)}$.
@@
 
+Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., species[@Allesina2012] or banks[@Haldane2011]).
+One element of such a matrix $M_{ij}$ defines how component $j$ affects component $i$ in the system at a point of equilibrium[@Allesina2012].
+Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
+Traditionally, values of off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$).
+In a large system (high $S$), eigenvalues are distributed uniformly[@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$[@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma^{2}$ increase.
+
+The above stability criteria assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+In ecological communities, for example, the rate at which population density changes following perterbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
+Species with short generation times will respond quickly (high $\gamma$) to perturbations relative to species with long generation times (low $\gamma$).
+Similarly, the speed at which individual banks respond to perturbations in financial networks, or individual institutions respond to perturbations in complex social-ecological systems, is likely to vary.
+The effect that this variance in the rate at which individual system components respond to perturbations has not been investigated in complex systems theory.
+Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
+Here I show why, despite higher $\sigma$, complex systems in which $\gamma$ varies actually tend to be more stable, especially when $S$ is high.
+
+```{r echo = FALSE}
+S     <- 200;
+C     <- 0.05
+sigma <- 0.4;
+```
+
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a scalar value $\gamma_{i}$[@Patel2018].
+The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
+This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
+I assume $S = 200$, $C = 0.05$, $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
+Eigenvalues of $M$ are distributed uniformly within a circle centred at ($0, -1$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+1a).
+Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
+But when
+
+<!--- Say that the amnat paper did this with multiplying a row --->
+
 <!---
@@
 2. Go back to @May1972 and explain the basic idea of random matrices as representing communities.
+
+===== ACTUALLY, use paragraph three to shift to considering component time scales not having been done (skip 3, sort of).
 

### 2018-05-18 ms.Rmd
More writing

@@
 sigma <- 0.4;
+pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st <- pr_st[,-1];
 ```
@@
 I assume $S = 200$, $C = 0.05$, $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
-Eigenvalues of $M$ are distributed uniformly within a circle centred at ($0, -1$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+Eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
 1a).
 Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
-But when
+But when $\gamma_{i}$ values are separated into two groups, eigenvalues are no longer uniformly distributed (Fig.
+1b).
+Instead, two distinct clusters of eigenvalues appear (grey cicles in Fig.
+1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
+The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$) and all $\Re({\lambda}) < 0$.
+The latter cluster has real components that have shifted to the right, but has a smaller radius.
+Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,2])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+
+Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
+I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions).
 

### 2018-05-18 ms.Rmd
Fix Fig 3 legend

@@
 Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Red and black bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
-Blue lines show the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
+Light and dark grey bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
+The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
 

### 2018-05-18 ms.Rmd
New paragraphs

@@
 S     <- 200;
-C     <- 0.05
+C     <- 0.05;
 sigma <- 0.4;
@@
 
+```{r echo = FALSE}
+S     <- 1000;
+C     <- 1;
+sigma <- 0.4;
+```
+
 Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions).
+I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions which give similar results).
+As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
+1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
+Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered tightly around the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (red circle Fig 2b).
+This larger radius occurs because the addition of $Var(\gamma)$ increases the realised $\sigma$ of $M$.
+The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
+But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold of $\sigma\sqrt{SC} > 1$.
 
-<!--- Say that the amnat paper did this with multiplying a row --->
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supporting Information for different values of $\sigma$ and $C$).
+One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
+Under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$.
+
+<!--- Remember in Fig 3 that the line is showing the A1s that are stable *due to* var(gamma) -- sometimes A0s were stable, but they were destabilised by var(gamma) --->
 
@@
 **b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
-In **a**, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in **b** $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in **b** but not **a** is stable because in **b** $\max\left(\Re(\lambda)\right) < 0$.
+In a, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
 In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
@@
 **Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = d = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
 Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.

### 2018-05-18 R/plot_figs.R
Minor edit

@@ -381,7 +381,7 @@ plot_stables <- function(dat, S_s = 32){
          cex = 1, col = "black", yaxs="i");
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c("E[\u03b3]", "Var(\u03b3)"), pch=15, 
+    legend("topleft", c("\u03b3 = 1", "Var(\u03b3)"), pch=15, 
            col=c("grey80","grey40"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 

### 2018-05-18 SI.Rmd
Initialise and use SI

---
title: "Component response rate variation drives stability in large complex systems"
author: "Brad Duthie"
bibliography: references.bib
output:
  html_document: default
  pdf_document:
    fig_caption: yes
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
subtitle: Supporting information
biblio-style: apalike
---

```{r, echo = FALSE}
source(file = "R/sim_mat.R");
source(file = "R/plot_figs.R");
```

Code and simulations underlying Fig. 1
===============================================================================

The sample $M$ used for the eigenvalue distributions in Fig. 1 of the text is available on [GitHub](), and was produced with by running the following function.

```{r}
find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
    while(iters > 0){
        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
        A_mat  <- matrix(data = A_dat, nrow = S);
        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
        A_mat  <- A_mat * C_mat;
        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
        mu_gam <- mean(gammas);
        diag(A_mat) <- -1;
        A1     <- gammas * A_mat;
        A0     <- mu_gam * A_mat;
        A0_e   <- eigen(A0)$values;
        A0_r   <- Re(A0_e);
        A0_i   <- Im(A0_e);
        A1_e   <- eigen(A1)$values;
        A1_r   <- Re(A1_e);
        A1_i   <- Im(A1_e);
        if(max(A0_r) >= 0 & max(A1_r) < 0){
            return(list(A0 = A0, A1 = A1));
            break;
        }
        print(iters);
        iters <- iters - 1;
    }
}
```

The above function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$. The function is used to illustrate the concept of how a fast versus slow component responses can cause a system to become stable. Simulations were run for `iter = 1000000`, but terminated once an acceptable `A0` and `A1` were found. The code below plots the eigenvalue distributions of `A0` and `A1` in panels **a** and **b**, respectively.

```{r, echo = FALSE}
A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
A0 <- as.matrix(A0[,-1]);
A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
A1 <- as.matrix(A1[,-1]);

A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(t(A0_vm));
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(t(A1_vm));
A1vec       <- A1vec[is.na(A1vec) == FALSE];
fhalf       <- 1:(0.5*length(A1vec));
shalf       <- (0.5*length(A1vec)+1):length(A1vec);
```

The plot itself can be recreated with the code below.

```{r, fig.height = 6, fig.width = 9}
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
vl <- seq(from = 0, to = 2*pi, by = 0.001);
A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
text(x = -3.5, y = 2.25, labels = "a", cex = 2);
points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
points(A0_r, A0_i, pch = 4, cex = 0.7);

plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
     col = "black", yaxt = "n");

vl <- seq(from = 0, to = 2*pi, by = 0.001);
A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");

points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);

text(x = -3.5, y = 2.25, labels = "b", cex = 2);
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
```

To find out how frequently $M$ was stable given that all $\gamma = 1$ versus $\gamma = \{1.95, 0.05\}$, the function below was created.

```{r}
stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
    A0_count <- 0;
    A1_count <- 0;
    while(iters > 0){
        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
        A_mat  <- matrix(data = A_dat, nrow = S);
        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
        A_mat  <- A_mat * C_mat;
        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
        mu_gam <- mean(gammas);
        diag(A_mat) <- -1;
        A1     <- gammas * A_mat;
        A0     <- mu_gam * A_mat;
        A0_e   <- eigen(A0)$values;
        A0_r   <- Re(A0_e);
        A0_i   <- Im(A0_e);
        A1_e   <- eigen(A1)$values;
        A1_r   <- Re(A1_e);
        A1_i   <- Im(A1_e);
        if(max(A0_r) < 0){
            ress[iters, 1] <- 1;
            A0_count       <- A0_count + 1;
            }
        if(max(A1_r) < 0){
            ress[iters, 2] <- 1;
            A1_count       <- A1_count + 1;
        }
        print(c(iters, A0_count, A1_count));
        iters <- iters - 1;
    }
    return(ress);
}
```

```{r, echo = FALSE}
pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
pr_st       <- pr_st[,-1];
```

The function above was run for `iters = 1000000`, and the resulting matrix `ress` was returned. Each row of `ress` represents a single $M$ given $\gamma = 1$ (column 1) versus $\gamma = \{1.95, 0.05\}$ (column 2). Values of 0 indicate that $M$ was found to be unstable (at least one real component of its eigenvalues greater than or equal to zero), whereas values of 1 indicate that $M$ was found to be stable (all real components of eigenvalues are negative). The frequencies of stable $M$ were `r sum(pr_st[,1])` given $\gamma = 1$ and `r sum(pr_st[,2])` given $\gamma = \{1.95, 0.05\}$, as reported in the main text and legend of Fig. 1.

Code and simulations underlying Fig. 2
===============================================================================

Figure 2 of the main text shows how eigenvalue distributions in a system where $S = 1000$, $C = 1$, and $\sigma = 0.4$. Eigenvalues can be reproduced using the code below for when $\gamma = 1$ (panel a) and $\gamma \sim \mathcal{U}(0, 2)$ (panel b).

```{r}
A_comp <- NULL;
A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
A_mat  <- matrix(data = A_dat, nrow = 1000);
C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
A_mat     <- A_mat * C_mat;
gammas <- runif(n = 1000, min = 0, max = 2);
mu_gam <- mean(gammas);
diag(A_mat) <- -1;
A1     <- gammas * A_mat;
A0     <- mu_gam * A_mat;
A0_e   <- eigen(A0)$values;
A0_r   <- Re(A0_e);
A0_i   <- Im(A0_e);
A1_e   <- eigen(A1)$values;
A1_r   <- Re(A1_e);
A1_i   <- Im(A1_e);

A0_vm       <- A0;
diag(A0_vm) <- NA;
A0vec       <- as.vector(A0_vm);
A0vec       <- A0vec[is.na(A0vec) == FALSE];
A1_vm       <- A1;
diag(A1_vm) <- NA;
A1vec       <- as.vector(A1_vm);
A1vec       <- A1vec[is.na(A1vec) == FALSE];
```

The code below reproduces the figure itself.

```{r, fig.height = 6, fig.width = 9}
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
vl <- seq(from = 0, to = 2*pi, by = 0.001);
x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
text(x = -15.5, y = 19, labels = "a", cex = 2);
points(x = x0, y = y0, type = "l", lwd = 3);
points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
     yaxt = "n");
text(x = -15.5, y = 19, labels = "b", cex = 2);
points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
```

### 2018-05-18 SI.Rmd
Add more SI

@@
 ```
+
+Stability across increasing $S$
+===============================================================================
+
+```{r, echo = TRUE}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+```
+
+The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S$ from 2 to 32.
+In this table, the `A0` refers to matrices when $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
+Each row summarises data for a given $S$ over 1 million randomly simulated $M$ (`A0` and `A1`).
+The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 1 million).
+Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
+The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
+
+```{r, echo = FALSE}
+library(knitr);
+d_tab <- dat[,1:7];
+kable(d_tab);
+```
+
+The results underlying this table were produced with the `rand_gen_var` function below.
+
+```{r}
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    for(i in 2:max_sp){
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+        while(iter > 0){
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
+            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            A0       <- A0 * C_mat;
+            diag(A0) <- -1;
+            gam1     <- runif(n = i, min = 0, max = 2);
+            A1       <- A0 * gam1;
+            A0       <- A0 * mean(gam1);
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
+            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
+            if(A0_stb == TRUE){
+                tot_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_stb == TRUE){
+                tot_res[[i-1]][iter, 2] <- 1;
+            }
+            if(A0_fea == TRUE){
+                fea_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_fea == TRUE){
+                fea_res[[i-1]][iter, 2] <- 1;
+            }
+            iter    <- iter - 1;
+        }
+        print(i);
+    }
+    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
+    return(all_res);
+}
+```
+
+The above function calls the two functions `species_interactions` and `summarise_randmat`, which are provided below.
+
+```{r}
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+
+summarise_randmat <- function(tot_res, fea_res){
+    sims    <- length(tot_res);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
+    for(i in 1:sims){
+        all_res[i, 1]  <- i + 1;
+        # Stable and unstable
+        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
+        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
+        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
+        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
+        # Stabilised and destabilised
+        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
+                                  tot_res[[i]][,2] == TRUE);
+        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
+                                  tot_res[[i]][,2] == FALSE);
+        # Feasible and infeasible
+        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
+        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
+        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
+        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
+        # Feased and defeased
+        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
+                                  fea_res[[i]][,2] == TRUE);
+        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
+                                  fea_res[[i]][,2] == FALSE);
+    }
+    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
+                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
+                "A0_feasible", "A1_infeasible", "A1_feasible", 
+                "A1_made_feasible", "A1_made_infeasible");
+    colnames(all_res) <- cnames;
+    return(all_res);
+}
+````
+
+Note that feasibility results were ommited for the table above, but are shown below.

### 2018-05-18 R/sim_mat.R
Minor

@@ -68,9 +68,6 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
             A0       <- A0 * C_mat;
             diag(A0) <- -1;
             gam1     <- runif(n = i, min = 0, max = 2);
-            #if(i %% 2 == 0){
-            #  gam1 <- c(rep(1.5, i/2), rep(0.5, i/2));
-            #}
             A1       <- A0 * gam1;
             A0       <- A0 * mean(gam1);
             A0_stb   <- max(Re(eigen(A0)$values)) < 0;

### 2018-05-19 ms.Rmd
More writing
Minor edits and more writing

@@
 But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem [@Allesina2015].
-Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), system stability is markedly increased.
+Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), the potential for system stability is markedly increased.
 Stability increases due to the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher variation among interaction strengths ($\boldsymbol{\sigma^{2}}$).
@@
 Traditionally, values of off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$).
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$)[@May1972; @Allesina2012].
 In a large system (high $S$), eigenvalues are distributed uniformly[@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$[@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
@@
 
-The above stability criteria assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
-In ecological communities, for example, the rate at which population density changes following perterbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
+The above stability criterion assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
 Species with short generation times will respond quickly (high $\gamma$) to perturbations relative to species with long generation times (low $\gamma$).
-Similarly, the speed at which individual banks respond to perturbations in financial networks, or individual institutions respond to perturbations in complex social-ecological systems, is likely to vary.
-The effect that this variance in the rate at which individual system components respond to perturbations has not been investigated in complex systems theory.
+Similarly, the speed at which individual banks respond to perturbations in financial networks, or individuals or institutions respond to perturbations in complex social networks, is likely to vary.
+The effect of such variance has not been investigated in complex systems theory.
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
@@
 This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
-I assume $S = 200$, $C = 0.05$, $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
-Eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
+Rows are then multiplied by $y_{i}$ to generate $M$.
+When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
 1a).
@@
 1b).
-Instead, two distinct clusters of eigenvalues appear (grey cicles in Fig.
+Instead, two distinct clusters of eigenvalues appear (grey circles in Fig.
 1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
@@
 Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions which give similar results).
+I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions, which give similar results).
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
@@
 
+```{r, echo = FALSE}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+```
+
 To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supporting Information for different values of $\sigma$ and $C$).
-One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
-Under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$.
+One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$).
+I found that number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results).
+This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
 

### 2018-05-19 ms.Rmd
More tweaks

@@
 One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$).
-I found that number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results).
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results).
 This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
+
+Some care is needed when interpreting these results.
+First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supporting Information).
+Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
+This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
+1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
+<!--- Mention specific ecological systems here?
+--->
+
+<!--- Grob the genetic algorithm results and show that by changing the component response time, we can guide a system to much more stability.
+State that mutualism is not stabilised (SI), but this is unsurprising as mutualisms are unstable (cite Allesina).
+Finally, feasibility of ecological systems was not changed by $Var(\gamma)$.
+Conclude that more investigation is needed.
+Then need a strong concluding paragraph.
+--->
 

### 2018-05-19 ms.tex
More tweaks

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Component response rate variation drives stability in large complex systems},
            pdfauthor={Brad Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Component response rate variation drives stability in large complex
systems}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Brad Duthie}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\linenumbers
\bibliographystyle{amnatnat}

\begin{document}
\maketitle

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{The stability of a complex system generally decreases with
increasing system size, as is demonstrated by random matrix
theory\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
This counter-intuitive result, first shown by
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}, is broadly
relevant for understanding the dynamics and persistence of systems such
as
ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}},
neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{3},\protect\hyperlink{ref-Gray2009}{4}},
biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{5},\protect\hyperlink{ref-MacArthur2010}{6}}
and
socio-economic\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}--\protect\hyperlink{ref-Bardoscia2017}{9}}
networks. Much attention has especially been given to the stability of
ecological communities such as food webs or mutualist networks, with
recent work investigating how different community structures affect
stability\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10}--\protect\hyperlink{ref-Patel2018}{14}}.
But more broadly, stabilising mechanisms in complex systems remain
under-developed, and the effect of variation in the response rate of
individual system components remains an open
problem\textsuperscript{\protect\hyperlink{ref-Allesina2015}{15}}. Here
I show that when components of a complex system respond to system
dynamics at different rates (\(\boldsymbol{\omega}\)), the potential for
system stability is markedly increased. Stability increases due to the
clustering of some eigenvalues toward the centre of eigenvalue
distributions despite the destabilising effect of higher variation among
interaction strengths (\(\boldsymbol{\sigma^{2}}\)). This effect of
variation in \(\boldsymbol{\omega}\) becomes increasingly important as
system size increases, to the extent that the largest stable complex
systems would otherwise be unstable if not for
\(\boldsymbol{Var(\omega)}\). My results therefore reveal a previously
unconsidered driver of system stability that is likely to be pervasive
across all complex systems. Future research in complex systems should
therefore account for the varying response rates of individual system
components when assessing whole system stability.}

In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
demonstrated that randomly assembled systems of sufficient complexity
are almost inevitably unstable given infinitesimally small
perturbations. Complexity in this case is defined by the size of the
system (i.e., the number of interacting components; \(S\)), its
inter-connectivity (i.e., the probability that one component will affect
another; \(C\)), and the variance of interaction strengths
(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
May's finding that the probability of local stability falls to near zero
given a sufficiently high threshold of \(\sigma\sqrt{SC}\) has profound
consequences across multiple disciplines, raising the question of how
complex systems in, e.g.,
ecology\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10},\protect\hyperlink{ref-Grilli2017}{13},\protect\hyperlink{ref-Allesina2015}{15}}
or
banking\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7},\protect\hyperlink{ref-Bardoscia2017}{9},\protect\hyperlink{ref-May2008}{16}}
are predicted to persist or change.

Randomly assembled complex systems can be represented as large square
matrices (\(M\)) with \(S\) components (e.g.,
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
element of such a matrix \(M_{ij}\) defines how component \(j\) affects
component \(i\) in the system at a point of
equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
Off-diagonal elements (\(i \neq j\)) therefore define interactions
between components, while diagonal elements (\(i = j\)) define component
self-regulation (e.g., carrying capacity in ecological communities).
Traditionally, values of off-diagonal elements are assigned non-zero
values with a probability \(C\), which are sampled from a distribution
with variance \(\sigma^{2}\); diagonal elements are set to
-1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}.
Local system stability is assessed using eigenanalysis, with the system
being stable if the real parts of all eigenvalues (\(\lambda\)) of \(M\)
are negative
(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
In a large system (high \(S\)), eigenvalues are distributed
uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}} within a
circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
\(\Im = 0\), with a radius of
\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}
(Figs 1a and 2a). Local stability of randomly assembled systems
therefore becomes increasingly unlikely as \(S\), \(C\), and
\(\sigma^{2}\) increase.

The above stability criterion assumes that individual components respond
to perturbations of the system at the same rate (\(\omega\)), but this
is highly unlikely in any complex system. In ecological communities, for
example, the rate at which population density changes following
perturbation will depend on the generation time of individuals, which
might vary by orders of magnitude among species. Species with short
generation times will respond quickly (high \(\omega\)) to perturbations
relative to species with long generation times (low \(\omega\)).
Similarly, the speed at which individual banks respond to perturbations
in financial networks, or individuals or institutions respond to
perturbations in complex social networks, is likely to vary. The effect
of such variance has not been investigated in complex systems theory.
Intuitively, variation in \(\omega\) might be expected to decrease
system stability by introducing a new source of variation into the
system and thereby increasing \(\sigma\). Here I show why, despite
higher \(\sigma\), complex systems in which \(\omega\) varies actually
tend to be more stable, especially when \(S\) is high.

Rows in \(M\) define how a given component \(i\) is affected by other
components of the system, meaning that the rate of component response
time can be modelled by multiplying all row elements by a scalar value
\(\omega_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
The distribution of \(\omega\) over \(S\) components thereby models the
distribution of component response rates. An instructive example
compares one \(M\) where \(\omega_{i} = 1\) for all \(i\) in \(S\) to
the same \(M\) when half of \(\omega_{i} = 1.95\) and half of
\(\omega_{i} = 0.05\). This models one system in which \(\omega\) is
invariant and one in which \(\omega\) varies, but systems are otherwise
identical (note \(E[\omega_{i}] = 1\) in both cases). I assume
\(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
set to \(-1\) and non-zero off-diagonal elements are drawn from
\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by \(y_{i}\) to
generate \(M\). When \(\omega_{i} = 1\), eigenvalues of \(M\) are
distributed uniformly within a circle centred at (\(-1, 0\)) with a
radius of 1.265 (Fig. 1a). Hence, the real components of eigenvalues are
highly unlikely to all be negative when all \(\omega_{i} = 1\). But when
\(\omega_{i}\) values are separated into two groups, eigenvalues are no
longer uniformly distributed (Fig. 1b). Instead, two distinct clusters
of eigenvalues appear (grey circles in Fig. 1b), one centred at
(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
large radius, but the real components have shifted to the left (in
comparison to when \(\omega = 1\)) and all \(\Re({\lambda}) < 0\). The
latter cluster has real components that have shifted to the right, but
has a smaller radius. Overall, for 1 million randomly assembled \(M\),
this division between slow and fast component response rates results in
more stable systems: 0 stable given \(\omega = 1\) versus 0 stable given
\(\omega = \{1.95, 0.5\}\).

Higher stability in systems with variation in \(\omega\) can be observed
by sampling \(\omega_{i}\) values from various distributions. I now
focus on a uniform distribution where \(\omega \sim \mathcal{U}(0, 2)\)
(see Supporting Information for other distributions, which give similar
results). As with the case of \(\omega = \{1.95, 0.5\}\) (Fig. 1b),
\(E[\omega] = 1\) when \(\omega \sim \mathcal{U}(0, 2)\), allowing
comparison of \(M\) before and after variation in component response
rate. Figure 2 shows a comparison of eigenvalue distributions given
\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
\(\omega = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
Uniform variation in \(\omega\) leads to a non-uniform distribution of
eigenvalues, some of which are clustered tightly around the centre of
the distribution, but others of which are spread outside the former
radius of 12.649 (red circle Fig 2b). This larger radius occurs because
the addition of \(Var(\omega)\) increases the realised \(\sigma\) of
\(M\). The clustering and spreading of eigenvalues introduced by
\(Var(\omega)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\omega = 1\), the effect of \(Var(\omega)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold of \(\sigma\sqrt{SC} > 1\).

To investigate the effect of \(Var(\omega)\) on system stability, I
simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
\(S\) ranging from \(2-32\) (see Supporting Information for different
values of \(\sigma\) and \(C\)). One million \(M\) were simulated for
each \(S\), and the stability of \(M\) was assessed given \(\omega = 1\)
versus \(\omega \sim \mathcal{U}(0, 2)\) (note that under these
conditions, \(\sigma\sqrt{SC} = 1\) given \(S = 25\) when
\(\omega = 1\)). I found that the number of stable random systems was
consistently higher given \(Var(\omega)\) than when \(\omega = 1\) (Fig.
3), and that the difference between the probabilities of observing a
stable system increased with an increase in \(S\); i.e., the potential
for \(Var(\omega)\) to drive stability increased with system complexity.
For the highest values of \(S\), nearly all systems that were stable
given \(Var(\omega)\) would not have been stable given \(\omega = 1\),
and the maximum observed \(S\) for which a system was stable was \(31\)
given \(Var(\omega)\) versus \(27\) given \(\omega = 1\) (see Supporting
Information for full results). This suggests that the stability of large
systems might be dependent upon variation in the response rate of their
individual components, meaning that factors such as generation time (in
ecological networks), transaction speed (in economic networks), or
communication speed (in social networks) needs to be considered when
investigating the stability of complex systems.

Some care is needed when interpreting these results. First, I emphasise
that \(Var(\omega)\) is not stabilising per se; that is, adding
variation in \(\omega\) to a particular system \(M\) does not
necessarily increase the probability that the system will be stable (see
Supporting Information). Rather, systems that are observed to be stable
are more likely to vary in \(\omega\), and for this \(Var(\omega)\) to
be critical to their stability. This is caused by the shift in the
distribution of eigenvalues that occurs by introducing \(Var(\omega)\)
(Fig. 1b, 2b), which can sometimes result in all \(\Re(\lambda) < 0\)
but might also increase \(\Re(\lambda)\) values.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{References}

\hypertarget{refs}{}
\hypertarget{ref-May1972}{}
1. May, R. M. Will a large complex system be stable? \emph{Nature}
\textbf{238,} 413--414 (1972).

\hypertarget{ref-Allesina2012}{}
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\hypertarget{ref-Gray2008}{}
3. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
random brain networks with a distribution of connection strengths.
\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).

\hypertarget{ref-Gray2009}{}
4. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
with excitatory and inhibitory connections. \emph{Neurocomputing}
\textbf{72,} 1849--1858 (2009).

\hypertarget{ref-Rosenfeld2009}{}
5. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
high-dimensional biochemical networks. \emph{Gene Regulation and Systems
Biology} \textbf{3,} 1--10 (2009).

\hypertarget{ref-MacArthur2010}{}
6. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
and criticality of adaptive regulatory networks. \emph{Physics Review
Letters} \textbf{104,} 168701 (2010).

\hypertarget{ref-Haldane2011}{}
7. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
\emph{Nature} \textbf{469,} 351--355 (2011).

\hypertarget{ref-Suweis2014}{}
8. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
networks. \emph{PLoS ONE} \textbf{9,} (2014).

\hypertarget{ref-Bardoscia2017}{}
9. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G. Pathways
towards instability in financial networks. \emph{Nature Communications}
\textbf{8,} 1--7 (2017).

\hypertarget{ref-Mougi2012}{}
10. Mougi, A. \& Kondoh, M. Diversity of interaction types and
ecological community stability. \emph{Science} \textbf{337,} 349--351
(2012).

\hypertarget{ref-Allesina2015a}{}
11. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\hypertarget{ref-Gao2016}{}
12. Gao, J., Barzel, B. \& Barabási, A. L. Universal resilience patterns
in complex networks. \emph{Nature} \textbf{530,} 307--312 (2016).

\hypertarget{ref-Grilli2017}{}
13. Grilli, J. \emph{et al.} Feasibility and coexistence of large
ecological communities. \emph{Nature Communications} \textbf{8,} (2017).

\hypertarget{ref-Patel2018}{}
14. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
effects of eco-evolutionary feedbacks on community stability.
\emph{American Naturalist} \textbf{191,} 1--29 (2018).

\hypertarget{ref-Allesina2015}{}
15. Allesina, S. \emph{et al.} Predicting the stability of large
structured food webs. \emph{Nature Communications} \textbf{6,} 7842
(2015).

\hypertarget{ref-May2008}{}
16. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).

\hypertarget{ref-Tao2010}{}
17. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
(2010).

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 1: Example distribution of eigenvalues before (a) and
after (b) separating a randomly generated complex system into fast
(\(\boldsymbol{\omega} = 1.95\)) and slow
(\(\boldsymbol{\omega} = 0.05\)) component response rates.} Each panel
shows the same system where \(S = 200\), \(C = 0.05\), and
\(\sigma = 0.4\), and in each case \(E[\omega] = 1\) (i.e., only the
distribution of \(\omega\) differs between panels). \textbf{a.}
Eigenvalues plotted when all \(\omega = 1\); distributions of points are
uniformly distributed within the grey circle with a radius of
\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
Eigenvalues plotted when half \(\omega = 1.95\) and half
\(\omega = 0.05\); distributions of points can be partitioned into one
large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
\(\omega = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
centred at \(\omega = -0.05\). In a, the maximum real eigenvalaue
\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
complex system in b but not a is stable because in b
\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
complex systems under the same parameter values, 0 were stable when
\(\omega = 1\) while 0 were stable when \(\omega = \{1.95, 0.05\}\).
Overall, complex systems that are separated into fast versus slow
components tend to be more stable than otherwise identical systems with
identical component response rates.

\includegraphics{ms_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\omega}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
Eigenvalues plotted in the absence of \(Var(\omega)\) where
\(E[\omega] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\omega \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
the distribution's centre (-1, 0). Black and red elipses in both panels
show the circle centred on the distribution in panels a and b,
respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
of \(\Re(\lambda) < 0\) are 0.55 and 0.554 for a and b, respectively.

\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 3: Stability of large complex systems with and without
variation in component response rate (\(\boldsymbol{\omega}\)).} The
\(\ln\) number of systems that are stable across different system sizes
(\(S\)) given \(C = 1\), and the proportion of systems in which
variation in \(\omega\) is critical for system stability. For each
\(S\), 1 million complex systems are randomly generated. Stability of
each complex system is tested given variation in \(\omega\) by randomly
sampling \(\omega \sim \mathcal{U}(0, 2)\). Stability given
\(Var(\omega)\) is then compared to stability in an otherwise identical
system in which \(\omega = E[\mathcal{U}(0, 2)]\) for all components.
Light and dark grey bars show the number of stable systems in the
absence and presence of variance in \(\omega\), respectively. The black
line shows the proportion of systems that are stable when
\(Var(\omega) > 0\), but would be unstable if \(Var(\omega) = 0\).

\begin{verbatim}
## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
## conversion failure on 'γ = 1' in 'mbcsToSbcs': dot substituted for <ce>
\end{verbatim}

\begin{verbatim}
## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
## conversion failure on 'γ = 1' in 'mbcsToSbcs': dot substituted for <b3>
\end{verbatim}

\begin{verbatim}
## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
## conversion failure on 'Var(γ)' in 'mbcsToSbcs': dot substituted for <ce>
\end{verbatim}

\begin{verbatim}
## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
## conversion failure on 'Var(γ)' in 'mbcsToSbcs': dot substituted for <b3>
\end{verbatim}

\begin{verbatim}
## Warning in text.default(x, y, ...): conversion failure on 'γ = 1' in
## 'mbcsToSbcs': dot substituted for <ce>
\end{verbatim}

\begin{verbatim}
## Warning in text.default(x, y, ...): conversion failure on 'γ = 1' in
## 'mbcsToSbcs': dot substituted for <b3>
\end{verbatim}

\begin{verbatim}
## Warning in text.default(x, y, ...): conversion failure on 'Var(γ)' in
## 'mbcsToSbcs': dot substituted for <ce>
\end{verbatim}

\begin{verbatim}
## Warning in text.default(x, y, ...): conversion failure on 'Var(γ)' in
## 'mbcsToSbcs': dot substituted for <b3>
\end{verbatim}

\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 4: Distributions of variation in component response time
evolved to generate stable systems}

\end{document}

### 2018-05-20 ms.Rmd
Update small tweaks to ms, mainly instructions for making a TEX file to submit to arxiv

@@
 output:
-  html_document: default
   pdf_document:
     fig_caption: yes
+    keep_tex: yes
+  html_document: default
   word_document:
@@
 ---
+
+<!---
+"This will generate the intermediate files that are actually used to generate the PDF.
+You need only go to the directory where your .Rmd file is, gather the .tex and image directory (called filename_files) with"
+
+arxiv comand: tar cvfz filename.tgz filename.tex filename_files
+--->
 

### 2018-05-20 ms.tex
Update small tweaks to ms, mainly instructions for making a TEX file to submit to arxiv

@@
 I show that when components of a complex system respond to system
-dynamics at different rates (\(\boldsymbol{\omega}\)), the potential for
+dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
 system stability is markedly increased. Stability increases due to the
@@
 interaction strengths (\(\boldsymbol{\sigma^{2}}\)). This effect of
-variation in \(\boldsymbol{\omega}\) becomes increasingly important as
+variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
 system size increases, to the extent that the largest stable complex
 systems would otherwise be unstable if not for
-\(\boldsymbol{Var(\omega)}\). My results therefore reveal a previously
+\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
 unconsidered driver of system stability that is likely to be pervasive
@@
 The above stability criterion assumes that individual components respond
-to perturbations of the system at the same rate (\(\omega\)), but this
+to perturbations of the system at the same rate (\(\gamma\)), but this
 is highly unlikely in any complex system. In ecological communities, for
@@
 might vary by orders of magnitude among species. Species with short
-generation times will respond quickly (high \(\omega\)) to perturbations
-relative to species with long generation times (low \(\omega\)).
+generation times will respond quickly (high \(\gamma\)) to perturbations
+relative to species with long generation times (low \(\gamma\)).
 Similarly, the speed at which individual banks respond to perturbations
@@
 of such variance has not been investigated in complex systems theory.
-Intuitively, variation in \(\omega\) might be expected to decrease
+Intuitively, variation in \(\gamma\) might be expected to decrease
 system stability by introducing a new source of variation into the
 system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\omega\) varies actually
+higher \(\sigma\), complex systems in which \(\gamma\) varies actually
 tend to be more stable, especially when \(S\) is high.
@@
 time can be modelled by multiplying all row elements by a scalar value
-\(\omega_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
-The distribution of \(\omega\) over \(S\) components thereby models the
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
+The distribution of \(\gamma\) over \(S\) components thereby models the
 distribution of component response rates. An instructive example
-compares one \(M\) where \(\omega_{i} = 1\) for all \(i\) in \(S\) to
-the same \(M\) when half of \(\omega_{i} = 1.95\) and half of
-\(\omega_{i} = 0.05\). This models one system in which \(\omega\) is
-invariant and one in which \(\omega\) varies, but systems are otherwise
-identical (note \(E[\omega_{i}] = 1\) in both cases). I assume
+compares one \(M\) where \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
+the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
+\(\gamma_{i} = 0.05\). This models one system in which \(\gamma\) is
+invariant and one in which \(\gamma\) varies, but systems are otherwise
+identical (note \(E[\gamma_{i}] = 1\) in both cases). I assume
 \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
@@
 \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by \(y_{i}\) to
-generate \(M\). When \(\omega_{i} = 1\), eigenvalues of \(M\) are
+generate \(M\). When \(\gamma_{i} = 1\), eigenvalues of \(M\) are
 distributed uniformly within a circle centred at (\(-1, 0\)) with a
 radius of 1.265 (Fig. 1a). Hence, the real components of eigenvalues are
-highly unlikely to all be negative when all \(\omega_{i} = 1\). But when
-\(\omega_{i}\) values are separated into two groups, eigenvalues are no
+highly unlikely to all be negative when all \(\gamma_{i} = 1\). But when
+\(\gamma_{i}\) values are separated into two groups, eigenvalues are no
 longer uniformly distributed (Fig. 1b). Instead, two distinct clusters
@@
 large radius, but the real components have shifted to the left (in
-comparison to when \(\omega = 1\)) and all \(\Re({\lambda}) < 0\). The
+comparison to when \(\gamma = 1\)) and all \(\Re({\lambda}) < 0\). The
 latter cluster has real components that have shifted to the right, but
@@
 this division between slow and fast component response rates results in
-more stable systems: 0 stable given \(\omega = 1\) versus 0 stable given
-\(\omega = \{1.95, 0.5\}\).
+more stable systems: 0 stable given \(\gamma = 1\) versus 0 stable given
+\(\gamma = \{1.95, 0.5\}\).
 
-Higher stability in systems with variation in \(\omega\) can be observed
-by sampling \(\omega_{i}\) values from various distributions. I now
-focus on a uniform distribution where \(\omega \sim \mathcal{U}(0, 2)\)
+Higher stability in systems with variation in \(\gamma\) can be observed
+by sampling \(\gamma_{i}\) values from various distributions. I now
+focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
 (see Supporting Information for other distributions, which give similar
-results). As with the case of \(\omega = \{1.95, 0.5\}\) (Fig. 1b),
-\(E[\omega] = 1\) when \(\omega \sim \mathcal{U}(0, 2)\), allowing
+results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b),
+\(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
 comparison of \(M\) before and after variation in component response
@@
 expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
-\(\omega = 1\), eigenvalues are distributed uniformly in a circle
+\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
 centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
-Uniform variation in \(\omega\) leads to a non-uniform distribution of
+Uniform variation in \(\gamma\) leads to a non-uniform distribution of
 eigenvalues, some of which are clustered tightly around the centre of
@@
 radius of 12.649 (red circle Fig 2b). This larger radius occurs because
-the addition of \(Var(\omega)\) increases the realised \(\sigma\) of
+the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
 \(M\). The clustering and spreading of eigenvalues introduced by
-\(Var(\omega)\) can destabilise previously stable systems or stabilise
+\(Var(\gamma)\) can destabilise previously stable systems or stabilise
 systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\omega = 1\), the effect of \(Var(\omega)\)
+complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
 can often lead to stability above
@@
 
-To investigate the effect of \(Var(\omega)\) on system stability, I
+To investigate the effect of \(Var(\gamma)\) on system stability, I
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
@@
 values of \(\sigma\) and \(C\)). One million \(M\) were simulated for
-each \(S\), and the stability of \(M\) was assessed given \(\omega = 1\)
-versus \(\omega \sim \mathcal{U}(0, 2)\) (note that under these
+each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
+versus \(\gamma \sim \mathcal{U}(0, 2)\) (note that under these
 conditions, \(\sigma\sqrt{SC} = 1\) given \(S = 25\) when
-\(\omega = 1\)). I found that the number of stable random systems was
-consistently higher given \(Var(\omega)\) than when \(\omega = 1\) (Fig.
+\(\gamma = 1\)). I found that the number of stable random systems was
+consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
 3), and that the difference between the probabilities of observing a
 stable system increased with an increase in \(S\); i.e., the potential
-for \(Var(\omega)\) to drive stability increased with system complexity.
+for \(Var(\gamma)\) to drive stability increased with system complexity.
 For the highest values of \(S\), nearly all systems that were stable
-given \(Var(\omega)\) would not have been stable given \(\omega = 1\),
+given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\),
 and the maximum observed \(S\) for which a system was stable was \(31\)
-given \(Var(\omega)\) versus \(27\) given \(\omega = 1\) (see Supporting
+given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see Supporting
 Information for full results). This suggests that the stability of large
@@
 Some care is needed when interpreting these results. First, I emphasise
-that \(Var(\omega)\) is not stabilising per se; that is, adding
-variation in \(\omega\) to a particular system \(M\) does not
+that \(Var(\gamma)\) is not stabilising per se; that is, adding
+variation in \(\gamma\) to a particular system \(M\) does not
 necessarily increase the probability that the system will be stable (see
 Supporting Information). Rather, systems that are observed to be stable
-are more likely to vary in \(\omega\), and for this \(Var(\omega)\) to
+are more likely to vary in \(\gamma\), and for this \(Var(\gamma)\) to
 be critical to their stability. This is caused by the shift in the
-distribution of eigenvalues that occurs by introducing \(Var(\omega)\)
+distribution of eigenvalues that occurs by introducing \(Var(\gamma)\)
 (Fig. 1b, 2b), which can sometimes result in all \(\Re(\lambda) < 0\)
@@
 after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\omega} = 1.95\)) and slow
-(\(\boldsymbol{\omega} = 0.05\)) component response rates.} Each panel
+(\(\boldsymbol{\gamma} = 1.95\)) and slow
+(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
 shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\omega] = 1\) (i.e., only the
-distribution of \(\omega\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\omega = 1\); distributions of points are
+\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
+distribution of \(\gamma\) differs between panels). \textbf{a.}
+Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
 uniformly distributed within the grey circle with a radius of
 \(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\omega = 1.95\) and half
-\(\omega = 0.05\); distributions of points can be partitioned into one
+Eigenvalues plotted when half \(\gamma = 1.95\) and half
+\(\gamma = 0.05\); distributions of points can be partitioned into one
 large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\omega = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
-centred at \(\omega = -0.05\). In a, the maximum real eigenvalaue
+\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
+centred at \(\gamma = -0.05\). In a, the maximum real eigenvalaue
 \(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
@@
 complex systems under the same parameter values, 0 were stable when
-\(\omega = 1\) while 0 were stable when \(\omega = \{1.95, 0.05\}\).
+\(\gamma = 1\) while 0 were stable when \(\gamma = \{1.95, 0.05\}\).
 Overall, complex systems that are separated into fast versus slow
@@
 introducing variation in component response rate
-(\(\boldsymbol{\omega}\)) in complex systems.} Each panel show the same
+(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
 system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
-Eigenvalues plotted in the absence of \(Var(\omega)\) where
-\(E[\omega] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\omega \sim \mathcal{U}(0, 2)\), which increases the variance of
+Eigenvalues plotted in the absence of \(Var(\gamma)\) where
+\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
+\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
 interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.55 and 0.554 for a and b, respectively.
+of \(\Re(\lambda) < 0\) are 0.555 and 0.549 for a and b, respectively.
 
@@
 \textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate (\(\boldsymbol{\omega}\)).} The
+variation in component response rate (\(\boldsymbol{\gamma}\)).} The
 \(\ln\) number of systems that are stable across different system sizes
 (\(S\)) given \(C = 1\), and the proportion of systems in which
-variation in \(\omega\) is critical for system stability. For each
+variation in \(\gamma\) is critical for system stability. For each
 \(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\omega\) by randomly
-sampling \(\omega \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\omega)\) is then compared to stability in an otherwise identical
-system in which \(\omega = E[\mathcal{U}(0, 2)]\) for all components.
+each complex system is tested given variation in \(\gamma\) by randomly
+sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
+\(Var(\gamma)\) is then compared to stability in an otherwise identical
+system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
 Light and dark grey bars show the number of stable systems in the
-absence and presence of variance in \(\omega\), respectively. The black
+absence and presence of variance in \(\gamma\), respectively. The black
 line shows the proportion of systems that are stable when
-\(Var(\omega) > 0\), but would be unstable if \(Var(\omega) = 0\).
+\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
 

### 2018-05-21 ms.Rmd
Add website
More writing

@@
 arxiv comand: tar cvfz filename.tgz filename.tex filename_files
+
+https://gist.github.com/JJ/ef9d3d8f1142df064bd5
 --->
@@
 Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supporting Information for other distributions, which give similar results).
+I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
@@
 
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supporting Information for different values of $\sigma$ and $C$).
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supplementary Information for different values of $\sigma$ and $C$).
 One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$).
@@
 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supporting Information for full results).
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supplementary Information for full results).
 This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
@@
 Some care is needed when interpreting these results.
-First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supporting Information).
+First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
 Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-<!--- Mention specific ecological systems here?
---->
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm because the space of possible $\gamma$ values was too large to searh exhaustively[@Hamblin2013] (see Supplementary Information).
+For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
+Eigenanalysis of each set of $\gamma$ values was performed, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase was satisfied.
+Across $S = \{2, 3, ..., 49, 50\}$, sets of $\gamma$ values were frequently found result in stable systems, suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
+<!--- Fix these last few sentences --->
+
+I have focused broadly on random complex systems, but
+
+My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
 
@@
 6.
-A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in supporting information.
+A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in Supplementary Information.
 The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.

### 2018-05-21 ms.Rmd
Minor tweaks

@@
 
-I have focused broadly on random complex systems, but
+I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina].
+These include systems in which all interactions (i.e., all off-diagonal elements of $M$) are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
+In general, competitive and mutualist networks tend to be stabilising, and predator-prey network tend to be stabilising[@Allesina2011].
+When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist communities does not (see Supplementary Information).
+Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
 

### 2018-05-21 SI.Rmd
More supporting information

@@
 
-Code and simulations underlying Fig. 1
-===============================================================================
+Contents
+================================================================================
+
+- [Code and simulations underlying Fig. 1](#Fig1) 
+- [Code and simulations underlying Fig. 2](#Fig2) 
+- [Stability across increasing $S$](#Incr_S}) 
+- [Stability of ecological networks](#ecological) 
+    - [Competitor networks](#competition) 
+    - [Mutualist networks](#mutualism) 
+    - [Predator-prey networks](#pred-prey) 
+- [Feasilbility of complex systems](#Feasibility)
+
+Code and simulations underlying Fig. 1 {#Fig1}
+================================================================================
 
@@
 
-Code and simulations underlying Fig. 2
+Code and simulations underlying Fig. 2 {#Fig2}
 ===============================================================================
@@
 
-Stability across increasing $S$
+Stability across increasing $S$ {#Incr_S}
 ===============================================================================
@@
 }
-````
+```
 
 Note that feasibility results were ommited for the table above, but are shown below.
+
+Stability of ecological networks {#ecological}
+================================================================================
+
+While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
+In these networks, the matrix $M$ is interpreted as a community matrix and each row and column is interpreted as a single species.
+The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $M$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015a; @Tang2014b].
+While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $M$ are negative), mutualist networks (all off-diagonal elements of $M$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2011; @Allesina2012; @Allesina2015a; @Tang2014b].
+In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.
+
+I investigate competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012].
+To create these networks, I first generated a random matrix $M$, then changed the elements of $M$ accordingly.
+If $M$ was a competitive network, then the sign of any positive off-diagonal elements was reversed to be negative.
+If $M$ was a mutualist network, then the sign of any positive off-diagonal elements was reversed to be positive.
+And if $M$ was a predator-prey network, then all $i$ and $j$ pairs of elements were checked; any pairs of the same sign were changed so that one was negative and the other was positive.
+The `species_interaction` function used to do this is below.
+
+```{r}
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+```
+
+This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#Incr_S) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
+The following tables for restricted ecological communities can therefore be compared with the random $M$ results above.
+
+**Competition** {#competition}
+
+Results for competitor interaction networks are shown below
+
+```{r, echo = TRUE}
+cdat   <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
+cdat   <- cdat[,-1];
+cd_tab <- cdat[,1:7];
+kable(cd_tab);
+```
+
+**Mutualism** {#mutualism}
+
+Results for mutualist interaction networks are shown below
+
+```{r, echo = TRUE}
+mdat   <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
+mdat   <- mdat[,-1];
+md_tab <- mdat[,1:7];
+kable(md_tab);
+```
+
+**Predator-prey** {#pred-prey}
+
+Results for predator-prey interaction networks are shown below
+
+```{r, echo = TRUE}
+pdat   <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
+pdat   <- pdat[,-1];
+pd_tab <- pdat[,1:7];
+kable(pd_tab);
+```
+
+Feasilbility of complex systems {#Feasibility}
+================================================================================
+
+It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable.
+Feasibility criteria are as follows,
+
+$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.
+The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive).
+The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have).
+The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0).
+And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off diagonal effecting species interactions,
+
+$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
+
+To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
+So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
+This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
+
+References
+================================================================================

### 2018-05-21 SI.Rmd
Fix subtitles

@@
 - [Feasilbility of complex systems](#Feasibility) 
+- [Genetic algorithm](#ga)
 
@@
 
-**Competition** {#competition}
+<a name="competition">**Competition**</a>
 
@@
 
-**Mutualism** {#mutualism}
+<a name="mutualism">**Mutualism**</a>
 
@@
 
-**Predator-prey** {#pred-prey}
+<a name="pred-prey">**Predator-prey**</a>
 
@@
 
+Genetic algorithm {#ga}
+================================================================================
+
 References

### 2018-05-22 ms.Rmd
Restructure manuscript

@@
 output:
+  html_document: default
   pdf_document:
@@
     keep_tex: yes
-  html_document: default
   word_document:
@@
 
+********************************************************************************
+
+```{r, echo = FALSE}
+A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
+A1 <- as.matrix(A1[,-1]);
+
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(t(A0_vm));
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(t(A1_vm));
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+fhalf       <- 1:(0.5*length(A1vec));
+shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st       <- pr_st[,-1];
+```
+
+**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
+**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
+In a, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
+In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
+Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
+A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
+text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
+points(A0_r, A0_i, pch = 4, cex = 0.7);
+
+plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "black", yaxt = "n");
+
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
+A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
+A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
+points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+
+points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);
+
+text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+********************************************************************************
+
 Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a scalar value $\gamma_{i}$[@Patel2018].
@@
 The latter cluster has real components that have shifted to the right, but has a smaller radius.
-Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,2])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
 
@@
 ```
+
+********************************************************************************
+
+```{r, echo = FALSE}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+gammas <- runif(n = 1000, min = 0, max = 2);
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+```
+
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
+Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
+Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+text(x = -15.5, y = 19, labels = "a", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
+     yaxt = "n");
+text(x = -15.5, y = 19, labels = "b", cex = 2);
+points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+********************************************************************************
 
@@
 ```
+
+********************************************************************************
+
+**Figure 3: Stability of large complex systems with and without variation in component response rate($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+For each $S$, 1 million complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
+Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
+Light and dark grey bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
+The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+plot_stables(dat);
+```
+
+********************************************************************************
 
@@
 
-I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina].
+I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
 These include systems in which all interactions (i.e., all off-diagonal elements of $M$) are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
@@
 <div id="refs"></div>
-
-********************************************************************************
-
-```{r, echo = FALSE}
-A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
-A0 <- as.matrix(A0[,-1]);
-A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
-A1 <- as.matrix(A1[,-1]);
-
-A0_e   <- eigen(A0)$values;
-A0_r   <- Re(A0_e);
-A0_i   <- Im(A0_e);
-A1_e   <- eigen(A1)$values;
-A1_r   <- Re(A1_e);
-A1_i   <- Im(A1_e);
-
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(t(A0_vm));
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(t(A1_vm));
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
-fhalf       <- 1:(0.5*length(A1vec));
-shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st       <- pr_st[,-1];
-```
-
-**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
-In a, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
-In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
-Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
-A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
-text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
-points(A0_r, A0_i, pch = 4, cex = 0.7);
-
-plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "black", yaxt = "n");
-
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
-A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
-points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
-A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
-A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
-points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
-
-points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);
-
-text(x = -3.5, y = 2.25, labels = "b", cex = 2);
-mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-```
-
-********************************************************************************
-
-```{r, echo = FALSE}
-A_comp <- NULL;
-A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
-A_mat  <- matrix(data = A_dat, nrow = 1000);
-C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
-C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
-A_mat     <- A_mat * C_mat;
-gammas <- runif(n = 1000, min = 0, max = 2);
-mu_gam <- mean(gammas);
-diag(A_mat) <- -1;
-A1     <- gammas * A_mat;
-A0     <- mu_gam * A_mat;
-A0_e   <- eigen(A0)$values;
-A0_r   <- Re(A0_e);
-A0_i   <- Im(A0_e);
-A1_e   <- eigen(A1)$values;
-A1_r   <- Re(A1_e);
-A1_i   <- Im(A1_e);
-
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(A0_vm);
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(A1_vm);
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
-```
-
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
-Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
-Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
-y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
-text(x = -15.5, y = 19, labels = "a", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
-     yaxt = "n");
-text(x = -15.5, y = 19, labels = "b", cex = 2);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
-mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-```
-
-********************************************************************************
-
-**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
-For each $S$, 1 million complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
-Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Light and dark grey bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
-The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
-plot_stables(dat);
-```
-
-********************************************************************************
-
-**Figure 4: Distributions of variation in component response time evolved to generate stable systems**

### 2018-05-22 ms.tex
Restructure manuscript

@@
 
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Figure 1: Example distribution of eigenvalues before (a) and
+after (b) separating a randomly generated complex system into fast
+(\(\boldsymbol{\gamma} = 1.95\)) and slow
+(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
+shows the same system where \(S = 200\), \(C = 0.05\), and
+\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
+distribution of \(\gamma\) differs between panels). \textbf{a.}
+Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
+uniformly distributed within the grey circle with a radius of
+\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
+Eigenvalues plotted when half \(\gamma = 1.95\) and half
+\(\gamma = 0.05\); distributions of points can be partitioned into one
+large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
+\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
+centred at \(\gamma = -0.05\). In a, the maximum real eigenvalaue
+\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
+\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
+complex system in b but not a is stable because in b
+\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
+complex systems under the same parameter values, 1 were stable when
+\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
+Overall, complex systems that are separated into fast versus slow
+components tend to be more stable than otherwise identical systems with
+identical component response rates.
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-4-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
 Rows in \(M\) define how a given component \(i\) is affected by other
@@
 this division between slow and fast component response rates results in
-more stable systems: 0 stable given \(\gamma = 1\) versus 0 stable given
-\(\gamma = \{1.95, 0.5\}\).
+more stable systems: 1 stable given \(\gamma = 1\) versus 32 stable
+given \(\gamma = \{1.95, 0.5\}\).
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
+introducing variation in component response rate
+(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
+system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
+Eigenvalues plotted in the absence of \(Var(\gamma)\) where
+\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
+\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
+interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
+the distribution's centre (-1, 0). Black and red elipses in both panels
+show the circle centred on the distribution in panels a and b,
+respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
+of \(\Re(\lambda) < 0\) are 0.549 and 0.553 for a and b, respectively.
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
-(see Supporting Information for other distributions, which give similar
-results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b),
-\(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
+(see Supplementary Information for other distributions, which give
+similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
+1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
 comparison of \(M\) before and after variation in component response
@@
 
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Figure 3: Stability of large complex systems with and without
+variation in component response rate(\(\boldsymbol{\gamma}\)).} The
+\(\ln\) number of systems that are stable across different system sizes
+(\(S\)) given \(C = 1\), and the proportion of systems in which
+variation in \(\gamma\) is critical for system stability. For each
+\(S\), 1 million complex systems are randomly generated. Stability of
+each complex system is tested given variation in \(\gamma\) by randomly
+sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
+\(Var(\gamma)\) is then compared to stability in an otherwise identical
+system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
+Light and dark grey bars show the number of stable systems in the
+absence and presence of variance in \(\gamma\), respectively. The black
+line shows the proportion of systems that are stable when
+\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
 To investigate the effect of \(Var(\gamma)\) on system stability, I
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from \(2-32\) (see Supporting Information for different
+\(S\) ranging from \(2-32\) (see Supplementary Information for different
 values of \(\sigma\) and \(C\)). One million \(M\) were simulated for
@@
 and the maximum observed \(S\) for which a system was stable was \(31\)
-given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see Supporting
-Information for full results). This suggests that the stability of large
-systems might be dependent upon variation in the response rate of their
-individual components, meaning that factors such as generation time (in
-ecological networks), transaction speed (in economic networks), or
-communication speed (in social networks) needs to be considered when
-investigating the stability of complex systems.
+given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see
+Supplementary Information for full results). This suggests that the
+stability of large systems might be dependent upon variation in the
+response rate of their individual components, meaning that factors such
+as generation time (in ecological networks), transaction speed (in
+economic networks), or communication speed (in social networks) needs to
+be considered when investigating the stability of complex systems.
 
@@
 necessarily increase the probability that the system will be stable (see
-Supporting Information). Rather, systems that are observed to be stable
-are more likely to vary in \(\gamma\), and for this \(Var(\gamma)\) to
-be critical to their stability. This is caused by the shift in the
-distribution of eigenvalues that occurs by introducing \(Var(\gamma)\)
-(Fig. 1b, 2b), which can sometimes result in all \(\Re(\lambda) < 0\)
-but might also increase \(\Re(\lambda)\) values.
+Supplementary Information). Rather, systems that are observed to be
+stable are more likely to vary in \(\gamma\), and for this
+\(Var(\gamma)\) to be critical to their stability. This is caused by the
+shift in the distribution of eigenvalues that occurs by introducing
+\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
+\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
+further investigate the potential of \(Var(\gamma)\) to be stabilising,
+I used a simple genetic algorithm because the space of possible
+\(\gamma\) values was too large to searh
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}}
+(see Supplementary Information). For each of 10000 random \(M\), the
+genetic algorithm initialised 1000 different sets of
+\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis of
+each set of \(\gamma\) values was performed, and the 20 sets with the
+lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
+subsequent mutation and crossover between the resulting new population
+of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
+\(M\) was found, 20 generations occurred, or a converge criteria of
+minimum fitness increase was satisfied. Across
+\(S = \{2, 3, ..., 49, 50\}\), sets of \(\gamma\) values were frequently
+found result in stable systems, suggesting that varying component
+response time might by itself be sufficient to stabilise complex
+systems.
+
+I have focused broadly on random complex systems, but it is also
+worthwhile to consider more restricted interactions such as those of
+specific ecological
+networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
+include systems in which all interactions (i.e., all off-diagonal
+elements of \(M\)) are negative (e.g., competitive networks), positive
+(e.g., mutualist networks), or \(i\) and \(j\) pairs have opposing signs
+(e.g., predator-prey networks). In general, competitive and mutualist
+networks tend to be stabilising, and predator-prey network tend to be
+stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
+When \(Var(\gamma)\) is applied to each, the proportion of stable
+competitive and predator-prey networks increases, but the proportion of
+stable mutualist communities does not (see Supplementary Information).
+Additionally, when each component of \(M\) is interpreted as a unique
+species and given a random intrinsic growth
+rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
+feasibility is not increased by \(Var(\gamma)\), suggesting that
+variation in species generation time might be unlikely to drive
+stability in purely multi-species networks (see Supplementary
+Information).
+
+My results show that complex systems are more likely to be stable when
+the response rates of system components vary. These results are broadly
+applicable to complex biological and social networks.
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\hypertarget{ref-Hamblin2013}{}
+18. Hamblin, S. On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+184--194 (2013).
 
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the grey circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
-centred at \(\gamma = -0.05\). In a, the maximum real eigenvalaue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 0 were stable when
-\(\gamma = 1\) while 0 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
+\hypertarget{ref-Allesina2011}{}
+19. Allesina, S. \& Levine, J. M. A competitive network theory of
+species diversity. \emph{Proceedings of the National Academy of Sciences
+of the United States of America} \textbf{108,} 5638--5642 (2011).
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
-Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
-the distribution's centre (-1, 0). Black and red elipses in both panels
-show the circle centred on the distribution in panels a and b,
-respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.555 and 0.549 for a and b, respectively.
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate (\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\)) given \(C = 1\), and the proportion of systems in which
-variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Light and dark grey bars show the number of stable systems in the
-absence and presence of variance in \(\gamma\), respectively. The black
-line shows the proportion of systems that are stable when
-\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
-
-\begin{verbatim}
-## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
-## conversion failure on 'γ = 1' in 'mbcsToSbcs': dot substituted for <ce>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
-## conversion failure on 'γ = 1' in 'mbcsToSbcs': dot substituted for <b3>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
-## conversion failure on 'Var(γ)' in 'mbcsToSbcs': dot substituted for <ce>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in strwidth(legend, units = "user", cex = cex, font = text.font):
-## conversion failure on 'Var(γ)' in 'mbcsToSbcs': dot substituted for <b3>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in text.default(x, y, ...): conversion failure on 'γ = 1' in
-## 'mbcsToSbcs': dot substituted for <ce>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in text.default(x, y, ...): conversion failure on 'γ = 1' in
-## 'mbcsToSbcs': dot substituted for <b3>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in text.default(x, y, ...): conversion failure on 'Var(γ)' in
-## 'mbcsToSbcs': dot substituted for <ce>
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in text.default(x, y, ...): conversion failure on 'Var(γ)' in
-## 'mbcsToSbcs': dot substituted for <b3>
-\end{verbatim}
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 4: Distributions of variation in component response time
-evolved to generate stable systems}
+\hypertarget{ref-Dougoud2018}{}
+20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+The feasibility of equilibria in large ecosystems: a primary but
+neglected concept in the complexity-stability debate. \emph{PLOS
+Computational Biology} \textbf{14,} e1005988 (2018).
 

### 2018-05-22 R/plot_figs.R
Minor changes, one fixes the compile by removing unicode from gamma

@@ -381,8 +381,9 @@ plot_stables <- function(dat, S_s = 32){
          cex = 1, col = "black", yaxs="i");
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c("\u03b3 = 1", "Var(\u03b3)"), pch=15, 
-           col=c("grey80","grey40"), cex = 1.5, horiz = TRUE);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("grey80","grey40"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
           line = 3.0);

### 2018-05-22 R/scratch.R

@@ -1,4 +1,11 @@
 
+
+eout <- read.csv("Evo_out.txt");
+
+dat <- eout[eout[,1] == 40,];
+
+
+
 rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1, 
                          gamma_mn = 1, C = 1){
   tot_res  <- NULL;

### 2018-05-22 ms.Rmd
Real draft of the ms made

@@
 output:
-  html_document: default
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
   word_document:
@@
     reference_docx: docx_template.docx
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+  html_document: default
 header-includes:
@@
 Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), the potential for system stability is markedly increased.
-Stability increases due to the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher variation among interaction strengths ($\boldsymbol{\sigma^{2}}$).
+Stability is caused by the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher interaction strength variation ($\boldsymbol{\sigma^{2}}$).
 This effect of variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for $\boldsymbol{Var(\gamma)}$.
@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, complex systems in which $\gamma$ varies actually tend to be more stable, especially when $S$ is high.
+Here I show why, despite higher $\sigma$, complex systems in which $\gamma$ varies are actually more likely to be stable, especially when $S$ is high.
 
@@
 In a, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
-In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` were stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
+In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` was stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
 Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
@@
 
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a scalar value $\gamma_{i}$[@Patel2018].
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018].
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
@@
 I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
-Rows are then multiplied by $y_{i}$ to generate $M$.
+Rows are then multiplied by $\gamma_{i}$ to generate $M$.
 When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
@@
 Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Light and dark grey bars show the number of stable systems in the absence and presence of variance in $\gamma$, respectively.
+Light and dark grey bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
 The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm because the space of possible $\gamma$ values was too large to searh exhaustively[@Hamblin2013] (see Supplementary Information).
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm (the space of possible $\gamma$ values was too large to searh exhaustively[@Hamblin2013]; see Supplementary Information).
 For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
 Eigenanalysis of each set of $\gamma$ values was performed, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase was satisfied.
-Across $S = \{2, 3, ..., 49, 50\}$, sets of $\gamma$ values were frequently found result in stable systems, suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
-<!--- Fix these last few sentences --->
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase between generations was satisfied.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were frequently found that resulted in stable systems (the highest at $S = 38$), suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
 
@@
 These include systems in which all interactions (i.e., all off-diagonal elements of $M$) are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
-In general, competitive and mutualist networks tend to be stabilising, and predator-prey network tend to be stabilising[@Allesina2011].
+In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
 When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist communities does not (see Supplementary Information).
@@
 My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
-
-<!--- Grob the genetic algorithm results and show that by changing the component response time, we can guide a system to much more stability.
-State that mutualism is not stabilised (SI), but this is unsurprising as mutualisms are unstable (cite Allesina).
-Finally, feasibility of ecological systems was not changed by $Var(\gamma)$.
-Conclude that more investigation is needed.
-Then need a strong concluding paragraph.
---->
-
-<!--- Remember in Fig 3 that the line is showing the A1s that are stable *due to* var(gamma) -- sometimes A0s were stable, but they were destabilised by var(gamma) --->
-
-<!---
-
-Part of the answer might be that communities aren't actually random ....
-
-2. Go back to @May1972 and explain the basic idea of random matrices as representing communities.
-
-===== ACTUALLY, use paragraph three to shift to considering component time scales not having been done (skip 3, sort of).
-
-3.
-A paragraph with some broad explanation of the idea of matrix stability, along the lines of @Allesina2012.
-This will effectively explain the idea of random matrix theory to the reader.
-Explain the general conclusion that stability decreases whenever $\sigma\sqrt{SC} > 1$.
-Note how the original ecological result has been broadened for different types of networks (e.g., nestedness, predator-prey, mutualist, etc.),
-
-4.
-But none of these studies have considered that different components operate on different time scales (but see [e.g.
-@Patel2018]).
-Then explain that this, intuitively, should decrease stability because it introduces another source of variation and therefore increases $\sigma$.
-But then explain that it also changes the distribution of eignvalues, bringing them towards the centre of the plot (figure 1).
-I show that this actually increases the stability of complex systems.
-
-5.
-A paragraph introducing, mathematically, how I am adding varying rates of component response -- multipling each row $i$ by a scalar $\gamma_{i}$.
-First this will be done with some simple function giving $Var(\gamma_{i})$, then in a more structured way such that random components that are more likely to be positively influenced by other components having faster or slower rates (higher or lower $\gamma_{i}$) of response.
-A sentence or two is needed to suggest what this might mean biologically (e.g., generalist predators have longer generation times than their prey).
-
-6.
-A paragraph explaining the results of the random matrices, with brief mention of other types of systems (e.g., predator-prey, mutualist, competitor) presented in Supplementary Information.
-The obvious increase in stability with different variation of $\gamma_{i}$ will be presented, along with an intuitive explanation of why this is the case (still waiting on simulations -- it would be great if correlation between positive or negative processes and response times led to increased stability, as appears to be the case.
-I could then argue that $Var(\gamma_{i})$ has a minimum such correlation of zero, and can only increase in stabilisation by having an absolute value of correlation being greater than zero).
-At least two [key figures](#fig_ideas) will be included, likely with four panels (1.
-without variation, 2.
-with unstructured variation, and 3.
-and 4.
-with structured variation).
-
-7.
-Some sort of mention is needed for a couple odd results; first, that in mutualist networks, stability is not affected -- perhaps because of the lack of variance in component response rate given number of interaction types (note, I think, competitive networks can have indirect positives due to intransitive effects).
-Second, *feasibility* is not increased, so pure ecological networks might not be affected by this (but emphasise that *pure* ecological networks don't really exist in nature -- they are incorporated into evolutionary, economic, social, systems).
-
-8.
-Some bold statement is needed here -- something to the effect of the argument that nearly all large stable complex systems will be stable *due to* the variation simulated in this paper.
-The odds that it would be stable without such variation becomes vanishingly small as system size increases.
-
---->
-
-<!---
-
-@Koch1999 Suggests that the results of @May1972 can apply to neuroscience
-
-@May2008 Suggests that random matrix theory can apply to banking
-
-Download: Jansen, V. A., & Kokkoris, G. D. (2003). Complexity and stability revisited. Ecology Letters, 6(6), 498-502.
-
-Hastings, H. M. (1982). The May-Wigner stability theorem. Journal of Theoretical Biology, 97(2), 155-166.
-
-Maybe see how variance in gamma modifies the general findings of May with respect to connectivity and interaction strength?
--- Link to variance in off diagonals, which Tang and Allesina (frontiers 2014) finds to be important.
-
-Gray, R. T., & Robinson, P. A. (2008). Stability and synchronization of random brain networks with a distribution of connection strengths. Neurocomputing, 71(7-9), 1373-1387.
-
-Gray & Robinson 2009 Stability of random brain networks with excitatory and inhibitory connections. Neurocomputing
-
-Need to consider that the off-diagonal elements size relative to the diagonal size will affect stability -- higher relative values make things *less* stable.
-I think this is okay because we're multiplying diagonals and off diagonals by the same values -- this is why if we multiply the whole matrix by a scalar it only affects the magnitude and not the sign of eigenvalues.
-
-Note further, that Allesina report on the criteria of May that \sigma \sqrt{SC} < 1, meaning that \sigma needs to be low for the random matrix to be stable.
-Hence, the multiplying rows by $\gamma$ is increasing the stability overall by decreasing the variance?
-This doesn't seem to make sense though -- really it's adding an element of variation, Var(inherent) + Var(due to \gamma).
-But, also maybe should think about gamma in relation to element values.
-If, e.g., components with similar gammas tend to be more likely to interact, then this might be stabilising?
-**VARIANCE APPEARS HIGHER GIVEN VARIATION IN gamma**
-
-**Potential solution** -- The increase in variation caused by gamma creates a situation of sometimes increasing but sometimes decreasing the correlation between pairs of off diagonal elements, and sometimes increasing but sometimes decreasing total off-diagonal variation.
-This increase creates more extreme total variations, some of which cause leading eigenvalues to all be negative.
-Note that the probability of stability versus instability would be affected by creating disproportionately more extreme variation (high or low) communities that are sometimes stable?
-**THIS DOES NOT APPEAR TO BE THE CASE BASED ON SIMULATION**
---->
 

### 2018-05-22 R/evolve_gammas.R
Evolve gammas updated

@@ -1,7 +1,10 @@
-Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
+Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
     tot_res <- NULL;
     fea_res <- NULL;
-    for(i in 2:max_sp){
+    if(from >= to){
+        stop("Argument 'from' must be less than argument 'to'");
+    }
+    for(i in from:to){
         nn             <- i;
         A1_stt         <- 0;
         A2_stt         <- 0;
@@ -19,10 +22,8 @@ Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
             C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
             A0       <- A0 * C_mat;
             diag(A0) <- -1;
-            gam0     <- make_gammas(nn = i, distribution = 0, sdd = 1);
-            gam1     <- make_gammas(nn = i, distribution = 1, sdd = 1);
+            gam1     <- runif(n = i, min = 0, max = 2);
             A1       <- A0 * gam1;
-            A0       <- A0 * gam0;
             A0_stb   <- max(Re(eigen(A0)$values)) < 0;
             A1_stb   <- rand_mat_ga(A1);
             A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
@@ -99,7 +100,7 @@ rand_mat_ga <- function(A1, max_it = 20, converg = 0.01){
         mu_dat   <- rbinom(n = nn*1000, size = 1, prob = 0.2);
         mu_dat2  <- rnorm(n = nn*1000, mean = 0, sd = 0.02);
         mu_dat2[mu_dat2 < 0] <- -mu_dat2[mu_dat2 < 0];
-        mu_dat2[mu_dat2 > 1] <- 1;
+        mu_dat2[mu_dat2 > 2] <- 2;
         mu_dat3  <- mu_dat * mu_dat2;
         mu_mat   <- matrix(data = mu_dat3, nrow = 1000, ncol = nn);
         new_gen  <- new_gen + mu_mat;
@@ -114,12 +115,11 @@ rand_mat_ga <- function(A1, max_it = 20, converg = 0.01){
     if(find_st == 1){
         s_row <- which(isst == 1)[1];
         writt <- c(nn, inds[s_row,]);
-        cat(writt, file = "Evo_out.txt", append = TRUE);
[…]

### 2018-05-22 SI.Rmd
Update SI

@@
 - [Feasilbility of complex systems](#Feasibility) 
+- [Different distributions of $\gamma$](#gam_dist)
 - [Genetic algorithm](#ga)
@@
 
+Different distributions of $\gamma$ {#gam_dist}
+================================================================================
+
 Genetic algorithm {#ga}

### 2018-05-22 SI.Rmd
genetic algorithm stuff

@@
 output:
-  html_document: default
   pdf_document:
     fig_caption: yes
+  html_document: default
   word_document:
@@
 
+**Note: All code and data are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability)**
+
 - [Code and simulations underlying Fig. 1](#Fig1) 
@@
     - [Predator-prey networks](#pred-prey) 
+- [Different connectance (C) values](#connectance)
+    - [C = 0.3](#connect3)
+    - [C = 0.5](#connect5)
+    - [C = 0.7](#connect7)
+    - [C = 0.9](#connect9)
+- [Different distributions of $\gamma$](#gam_dist)
 - [Feasilbility of complex systems](#Feasibility) 
-- [Different distributions of $\gamma$](#gam_dist)
 - [Genetic algorithm](#ga)
@@
 
-The sample $M$ used for the eigenvalue distributions in Fig. 1 of the text is available on [GitHub](), and was produced with by running the following function.
+The sample $M$ used for the eigenvalue distributions in Fig.
+1 of the text is available on [GitHub](https://github.com/bradduthie/random_matrix_stability/tree/master/sim_results/bi_gamma), and was produced by running the following function.
 
@@
 The above function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$.
-The function is used to illustrate the concept of how a fast versus slow component responses can cause a system to become stable.
+The function is used to illustrate the concept of how fast versus slow component responses can cause a system to become stable.
 Simulations were run for `iter = 1000000`, but terminated once an acceptable `A0` and `A1` were found.
@@
 The frequencies of stable $M$ were `r sum(pr_st[,1])` given $\gamma = 1$ and `r sum(pr_st[,2])` given $\gamma = \{1.95, 0.05\}$, as reported in the main text and legend of Fig.
-1.
+1 (raw data are [available on GitHub](https://github.com/bradduthie/random_matrix_stability/blob/master/sim_results/bi_gamma/bi_pr_st.csv)).
 
@@
 
-Figure 2 of the main text shows how eigenvalue distributions in a system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+Figure 2 of the main text shows eigenvalue distributions in a system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
 Eigenvalues can be reproduced using the code below for when $\gamma = 1$ (panel a) and $\gamma \sim \mathcal{U}(0, 2)$ (panel b).
@@
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 dat <- read.csv(file = "sim_results/C_1/random_all.csv");
@@
 
-The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S$ from 2 to 32.
-In this table, the `A0` refers to matrices when $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
+The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
+In this table, the `A0` refers to matrices where $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
 Each row summarises data for a given $S$ over 1 million randomly simulated $M$ (`A0` and `A1`).
@@
 
-Note that feasibility results were ommited for the table above, but are shown below.
+Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).
 
@@
     return(mat);
-}
+} # Note: -1 values are added in the diagonal later
 ```
@@
 This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#Incr_S) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
-The following tables for restricted ecological communities can therefore be compared with the random $M$ results above.
+The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#Incr_S).
+As with the [results above](#Incr_S), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added.
+The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
+Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
+The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
 
@@
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 cdat   <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
@@
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 mdat   <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
@@
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 pdat   <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
@@
 
+Overall, as expected[@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable.
+But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
+This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
+
+Different connectance (C) values {#connectance}
+================================================================================
+
+In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $M$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
+Here I present four tables showing the number of stable communities given $C = \{0.3, 0.
+5, 0.7, 0.9 \}$.
+In all cases, uniform variation in component response time ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
+In contrast to the main text, 100000 rather than 1 million $M$ were simulated.
+As with the results on [stability with increasing $S$](#Incr_S) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added.
+The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
+Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
+The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
+
+<a name="connect3">**Connectance $C = 0.3$**</a>
+
+```{r, echo = FALSE}
+C3dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
+C3dat   <- C3dat[,-1];
+C3d_tab <- C3dat[,1:7];
+kable(C3d_tab);
+```
+
+<a name="connect5">**Connectance $C = 0.5$**</a>
+
+```{r, echo = FALSE}
+C5dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
+C5dat   <- C5dat[,-1];
+C5d_tab <- C5dat[,1:7];
+kable(C5d_tab);
+```
+
+<a name="connect7">**Connectance $C = 0.7$**</a>
+
+```{r, echo = FALSE}
+C7dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
+C7dat   <- C7dat[,-1];
+C7d_tab <- C7dat[,1:7];
+kable(C7d_tab);
+```
+
+<a name="connect9">**Connectance $C = 0.9$**</a>
+
+```{r, echo = FALSE}
+C9dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
+C9dat   <- C9dat[,-1];
+C9d_tab <- C9dat[,1:7];
+kable(C9d_tab);
+```
+
+Different distributions of $\gamma$ {#gam_dist}
+================================================================================
+
+In the main text, I considered a uniform distribution of component response rates $\gamma \sim \mathcal{U}(0, 2)$.
+The number of unstable and stable $M$ matrices are reported in [a table above](#Incr_S) across different values of $S$.
+Here I show complementary results for three different distributions including an exponential, beta, and gamma distribution of $\gamma$ values.
+The shape of these distributions is shown in the figure below.
+
+********************************************************************************
+
+**Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$) for b-d.
+Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$).
+Each panel shows 1 million randomly generated $\gamma$ values.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 7, fig.width = 6}
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- 2;
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
+
+show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
+    y1 <- make_gammas(nn = nn, distribution = 1, sdd = sdd, mn = mn);
+    y2 <- make_gammas(nn = nn, distribution = 2, sdd = sdd, mn = mn);
+    y3 <- make_gammas(nn = nn, distribution = 3, sdd = sdd, mn = mn);
+    y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
+    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
+    h1 <- hist(y1, breaks = 1000, plot = FALSE);
+    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h1$counts)*1.2));
+    mtext("a", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
+    h2 <- hist(y2, breaks = 1000, plot = FALSE);
+    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h2$counts)*1.2));
+    mtext("b", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
+    h3 <- hist(y3, breaks = 1000, plot = FALSE);
+    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h3$counts)*1.2));
+    mtext("c", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
+    h4 <- hist(y4, breaks = 1000, plot = FALSE);
+    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h4$counts)*1.2));
+    mtext("d", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
+    mtext(expression(paste("Component ",gamma," value")),
+          outer=TRUE,side=1,line=3.0,cex=2);
+    mtext(expression(paste("Relative frequency")),
+          outer=TRUE,side=2,line=2.5,cex=2);
+}
+
+show_gammas();
+```
+
+********************************************************************************
+
+The same 100000 $M$ matrices were used to investigate stability when applying each of these different distributions of $\gamma$ values.
+The table below shows the number of $M$ that were unstable (_unst) and stable (_stbl) for the exponential (Exp), beta, and gamma distributions.
+
+```{r, echo = TRUE}
+fordist   <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
+kable(fordist);
+```
+
+In comparison to the uniform distribution (a), proportionally fewer random systems are found with the exponential distribution (b), while more are found with the beta (c) and gamma (d) distributions.
+
+Genetic algorithm {#ga}
+================================================================================
+
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluate for each unique $M$.
+This is technically impossible because any $\gamma_{i}$ can take any real value between $0-2$, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
+In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
+The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below 20 generations pass or the fitness increase from one generation to the next is below a certain criteria).
+The genetic algorithm relies on five functions.
+The first outer function `Evo_rand_gen_var` runs all of the simulations (`from` and `to` refer to $S$ values, and `iters` refers to the number of $M$ to try for each $S$).
+
+```{r}
+Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    if(from >= to){
+        stop("Argument 'from' must be less than argument 'to'");
+    }
+    for(i in from:to){
+        nn             <- i;
+        A1_stt         <- 0;
+        A2_stt         <- 0;
+        A1_fet         <- 0;
+        A2_fet         <- 0;
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        while(iter > 0){
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
+            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            A0       <- A0 * C_mat;
+            diag(A0) <- -1;
+            gam1     <- runif(n = i, min = 0, max = 2);
+            A1       <- A0 * gam1;
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            A1_stb   <- rand_mat_ga(A1);
+            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
+            if(A0_stb == TRUE){
+                tot_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_stb == TRUE){
+                tot_res[[i-1]][iter, 2] <- 1;
+            }
+            if(A0_fea == TRUE){
+                fea_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_fea == TRUE){
+                fea_res[[i-1]][iter, 2] <- 1;
+            }
+            iter    <- iter - 1;
+        }
+        print(i);
+    }
+    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
+    return(all_res);
+}
+```
+
+Note that `Evo_rand_gen_var` calls three custom sub-functions, `species_interactions`, `rand_mat_ga`, and `summarise_randmat`.
+The first simply allows for non-random interactions between components (e.g., modelling [ecological interactions](#ecological) of random, competition, mutualism, or predator-prey).
+
+```{r}
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+```
+
+The sub-function `rand_mat_ga` does the work of the genetic algorithm, searching for $\gamma$ vectors that are stabilising.
+
+```{r}
+rand_mat_ga <- function(A1, max_it = 20, converg = 0.01){
+    nn       <- dim(A1)[1];
+    rind     <- runif(n = nn*1000, min = 0, max = 1);
+    inds     <- matrix(data = rind, nrow = 1000, ncol = nn);
+    lastf    <- -10;
+    ccrit    <- 10;
+    find_st  <- 0;
+    iter     <- max_it;
+    while(iter > 0 & find_st < 1 & ccrit > converg){
+        ivar  <- rep(x = 0, length = dim(inds)[1]);
+        ifit  <- rep(x = 0, length = dim(inds)[1]);
+        isst  <- rep(x = 0, length = dim(inds)[1]);
+        for(i in 1:dim(inds)[1]){
+            ifit[i] <- -1*max(Re(eigen(inds[i,]*A1)$values));
+            ivar[i] <- var(inds[i,]);
+            isst[i] <- max(Re(eigen(inds[i,]*A1)$values)) < 0;
+        }
+        most_fit <- order(ifit, decreasing = TRUE)[1:20];
+        parents  <- inds[most_fit,];
+        new_gen  <- matrix(data = t(parents), nrow = 1000, ncol = nn, 
+                           byrow = TRUE);
+        mu_dat   <- rbinom(n = nn*1000, size = 1, prob = 0.2);
+        mu_dat2  <- rnorm(n = nn*1000, mean = 0, sd = 0.02);
+        mu_dat2[mu_dat2 < 0] <- -mu_dat2[mu_dat2 < 0];
+        mu_dat2[mu_dat2 > 2] <- 2;
+        mu_dat3  <- mu_dat * mu_dat2;
+        mu_mat   <- matrix(data = mu_dat3, nrow = 1000, ncol = nn);
+        new_gen  <- new_gen + mu_mat;
+        new_gen  <- crossover(inds = new_gen, pr = 0.1);
+        inds     <- new_gen;
+        find_st  <- max(isst);
+        newf     <- mean(ifit);
+        ccrit    <- newf - lastf;
+        lastf    <- newf;
+        iter     <- iter - 1;
+    }
+    if(find_st == 1){
+        s_row <- which(isst == 1)[1];
+        writt <- c(nn, inds[s_row,]);
+        cat(writt, file = "evo_out.txt", append = TRUE);
+        cat("\n", file = "evo_out.txt", append = TRUE);
+    }
+    return(find_st);
+}
+```
+
+The while loop continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed.
+Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and there is selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
+In mutation, any values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
+Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
+
+Crossover occurs in the `crossover` function below.
+
+```{r} 
+crossover <- function(inds, pr = 0.1){
+    crossed <- floor(dim(inds)[1] * pr);
+    cross1  <- sample(x = 1:dim(inds)[1], size = crossed);
+    cross2  <- sample(x = 1:dim(inds)[1], size = crossed);
+    for(i in 1:length(cross1)){
+        fromv   <- sample(x = 1:dim(inds)[2], size = 1);
+        tov     <- sample(x = 1:dim(inds)[2], size = 1);
+        temp                   <- inds[cross1[i],fromv:tov];
+        inds[cross1[i],fromv:tov] <- inds[cross2[i],fromv:tov];
+        inds[cross2[i],fromv:tov] <- temp;
+    }
+    return(inds);
+}
+```
+
+After all $M$ are simulated in `Evo_rand_gen_var`, the `summarise_randmat` formats the data into a table.
+
+```{r}
+summarise_randmat <- function(tot_res, fea_res){
+    sims    <- length(tot_res);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 10);
+    for(i in 1:sims){
+        unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
+        stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
+        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
+        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
+        non_feas  <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == FALSE;
+        feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
+        unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
+        feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
+        foundd    <- tot_res[[i]][,3] == TRUE;
+        all_res[i, 1]  <- i + 1;
+        all_res[i, 2]  <- sum(unstables);
+        all_res[i, 3]  <- sum(stables);
+        all_res[i, 4]  <- sum(unstabled);
+        all_res[i, 5]  <- sum(stabled);
+        all_res[i, 6]  <- sum(non_feas);
+        all_res[i, 7]  <- sum(feasibl);
+        all_res[i, 8]  <- sum(unfeased);
+        all_res[i, 9]  <- sum(feased);
+        all_res[i, 10] <- sum(foundd);
+    }
+    return(all_res);
+}
+```
+
+Stability results from this table are shown below.
+
+```{r, echo = FALSE}
+evo_res   <- read.csv(file = "sim_results/evolved/evo_results.csv");
+kable(evo_res);
+```
+
+The distribution of one of the $\gamma$ vectors at $S = 39$ is shown below (all are available on [GitHub]()).
+
+**Work in progress**
+
 Feasilbility of complex systems {#Feasibility}
 ================================================================================
+
+**Work in progress**
 
@@
 And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off diagonal effecting species interactions,
+Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions,
 
@@
 
-Different distributions of $\gamma$ {#gam_dist}
-================================================================================
-
-Genetic algorithm {#ga}
-================================================================================
-
 References

### 2018-05-25 notebook/SI.Rmd
Move several R files

--- 2018-05-22 SI.Rmd
@@
 output:
+  html_document: default
   pdf_document:
     fig_caption: yes
-  html_document: default
   word_document:

### 2018-05-25 notebook/ms.tex
Move several R files

--- 2018-05-22 ms.tex
@@
 dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
-system stability is markedly increased. Stability increases due to the
+system stability is markedly increased. Stability is caused by the
 clustering of some eigenvalues toward the centre of eigenvalue
-distributions despite the destabilising effect of higher variation among
-interaction strengths (\(\boldsymbol{\sigma^{2}}\)). This effect of
+distributions despite the destabilising effect of higher interaction
+strength variation (\(\boldsymbol{\sigma^{2}}\)). This effect of
 variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
@@
 system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\gamma\) varies actually
-tend to be more stable, especially when \(S\) is high.
+higher \(\sigma\), complex systems in which \(\gamma\) varies are
+actually more likely to be stable, especially when \(S\) is high.
 
@@
 \(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 were stable when
+complex systems under the same parameter values, 1 was stable when
 \(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
@@
 components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a scalar value
+time can be modelled by multiplying all row elements by a real scalar
+value
 \(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
@@
 set to \(-1\) and non-zero off-diagonal elements are drawn from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by \(y_{i}\) to
-generate \(M\). When \(\gamma_{i} = 1\), eigenvalues of \(M\) are
-distributed uniformly within a circle centred at (\(-1, 0\)) with a
-radius of 1.265 (Fig. 1a). Hence, the real components of eigenvalues are
-highly unlikely to all be negative when all \(\gamma_{i} = 1\). But when
-\(\gamma_{i}\) values are separated into two groups, eigenvalues are no
-longer uniformly distributed (Fig. 1b). Instead, two distinct clusters
-of eigenvalues appear (grey circles in Fig. 1b), one centred at
-(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
-large radius, but the real components have shifted to the left (in
-comparison to when \(\gamma = 1\)) and all \(\Re({\lambda}) < 0\). The
-latter cluster has real components that have shifted to the right, but
-has a smaller radius. Overall, for 1 million randomly assembled \(M\),
-this division between slow and fast component response rates results in
-more stable systems: 1 stable given \(\gamma = 1\) versus 32 stable
-given \(\gamma = \{1.95, 0.5\}\).
+\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
+\(\gamma_{i}\) to generate \(M\). When \(\gamma_{i} = 1\), eigenvalues
+of \(M\) are distributed uniformly within a circle centred at
+(\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real components
+of eigenvalues are highly unlikely to all be negative when all
+\(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated into
+two groups, eigenvalues are no longer uniformly distributed (Fig. 1b).
+Instead, two distinct clusters of eigenvalues appear (grey circles in
+Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
+(\(-0.05, 0\)). The former has a large radius, but the real components
+have shifted to the left (in comparison to when \(\gamma = 1\)) and all
+\(\Re({\lambda}) < 0\). The latter cluster has real components that have
+shifted to the right, but has a smaller radius. Overall, for 1 million
+randomly assembled \(M\), this division between slow and fast component
+response rates results in more stable systems: 1 stable given
+\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
 
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.549 and 0.553 for a and b, respectively.
+of \(\Re(\lambda) < 0\) are 0.547 and 0.557 for a and b, respectively.
 
@@
 Light and dark grey bars show the number of stable systems in the
-absence and presence of variance in \(\gamma\), respectively. The black
-line shows the proportion of systems that are stable when
+absence and presence of \(Var(\gamma)\), respectively. The black line
+shows the proportion of systems that are stable when
 \(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
@@
 further investigate the potential of \(Var(\gamma)\) to be stabilising,
-I used a simple genetic algorithm because the space of possible
-\(\gamma\) values was too large to searh
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}}
-(see Supplementary Information). For each of 10000 random \(M\), the
+I used a simple genetic algorithm (the space of possible \(\gamma\)
+values was too large to searh
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
+see Supplementary Information). For each of 10000 random \(M\), the
 genetic algorithm initialised 1000 different sets of
@@
 \(M\) was found, 20 generations occurred, or a converge criteria of
-minimum fitness increase was satisfied. Across
-\(S = \{2, 3, ..., 49, 50\}\), sets of \(\gamma\) values were frequently
-found result in stable systems, suggesting that varying component
-response time might by itself be sufficient to stabilise complex
-systems.
+minimum fitness increase between generations was satisfied. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were frequently
+found that resulted in stable systems (the highest at \(S = 38\)),
+suggesting that varying component response time might by itself be
+sufficient to stabilise complex systems.
 
@@
 (e.g., predator-prey networks). In general, competitive and mutualist
-networks tend to be stabilising, and predator-prey network tend to be
+networks tend to be destabilising, and predator-prey network tend to be
 stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.

### 2018-05-25 notebook/SI.Rmd
Edit SI

@@
 The code below plots the eigenvalue distributions of `A0` and `A1` in panels **a** and **b**, respectively.
+The plot itself can be recreated with the function and code below.
 
 ```{r, echo = FALSE}
-A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
-A0 <- as.matrix(A0[,-1]);
-A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
-A1 <- as.matrix(A1[,-1]);
-
+load(bg_data);
+plot_Fig_1 <- function(A0, A1){
+    S_val       <- dim(A0)[1];
     A0_e        <- eigen(A0)$values;
@@
     A1_i        <- Im(A1_e);
-
     A0_vm       <- A0;
@@
     shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-```
-
-The plot itself can be recreated with the code below.
-
-```{r, fig.height = 6, fig.width = 9}
     par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
@@
     vl   <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
-A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
+    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
     text(x = -3.5, y = 2.25, labels = "a", cex = 2);
@@
     points(A0_r, A0_i, pch = 4, cex = 0.7);
-
     plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
@@
          col = "black", yaxt = "n");
-
     vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
-A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
+    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:(0.5*S_val)]);
+    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
     points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
-A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
-A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
+    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
+                  mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
+    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
     points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
-
-points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);
-
+    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7);   
     text(x = -3.5, y = 2.25, labels = "b", cex = 2);
@@
     mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+plot_Fig_1(A0 = A0, A1 = A1);
 ```
@@
 
-```{r, echo = FALSE}
-pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st       <- pr_st[,-1];
+```{r, echo = TRUE}
+load(bi_pr_st);
+pr_st  <- bi_pr_st[,-1];
 ```
@@
 Eigenvalues can be reproduced using the code below for when $\gamma = 1$ (panel a) and $\gamma \sim \mathcal{U}(0, 2)$ (panel b).
+The function below reproduces the figure.
 
-```{r}
+```{r, fig.height = 6, fig.width = 9}
+plot_Fig_2 <- function(){
     A_comp <- NULL;
@@
     A1_i   <- Im(A1_e);
-
     A0_vm       <- A0;
@@
     A1vec       <- A1vec[is.na(A1vec) == FALSE];
-```
-
-The code below reproduces the figure itself.
-
-```{r, fig.height = 6, fig.width = 9}
     par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1);
     vl <- seq(from = 0, to = 2*pi, by = 0.001);
@@
     mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+plot_Fig_2();
 ```

### 2018-05-25 R/Figure_1.R
Relevant R package files

@@ -1,0 +1,155 @@
+#' Find a stabilised system
+#' 
+#' This function finds a random matrix (M) for which the complex system it
+#' represents is unstable before the addition of variation in component
+#' response rate, but stable after its addition.
+#'
+#'@return If successful, a list that includes a matrix (A0) that is unstable in 
+#'the absence of component response rate variation and a matrix (A1) that is
+#'is identical to A0 but includes variation in component response rate and is
+#'stable.
+#'@param S Size of the complex system
+#'@param C Connectedness of complex system
+#'@param Osd Standard deviation of interaction strength
+#'@param iters Number of random matrices to simulate
+#'@examples
+#'find_bgamma(S = 200, C = 0.05, Osd = 0.4, iters = 2);
+#'@useDynLib random_matrix_stability
+#'@export
+find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
+    while(iters > 0){
+        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
+        A_mat  <- matrix(data = A_dat, nrow = S);
+        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+        A_mat  <- A_mat * C_mat;
+        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
+        mu_gam <- mean(gammas);
+        diag(A_mat) <- -1;
+        A1     <- gammas * A_mat;
+        A0     <- mu_gam * A_mat;
+        A0_e   <- eigen(A0)$values;
+        A0_r   <- Re(A0_e);
+        A0_i   <- Im(A0_e);
+        A1_e   <- eigen(A1)$values;
+        A1_r   <- Re(A1_e);
+        A1_i   <- Im(A1_e);
+        if(max(A0_r) >= 0 & max(A1_r) < 0){
+            return(list(A0 = A0, A1 = A1));
+            break;
+        }
+        print(iters);
+        iters <- iters - 1;
+    }
+}
+
+#' Find stable bimodal systems
+#' 
+#' Produces a two panel plot for a matrix (M) before and after the addition of 
[…]

### 2018-05-25 R/Figure_2.R

@@ -1,0 +1,59 @@
+#' Plot eigenvalues of a large system
+#' 
+#' Produces a two panel plot for a matrix (M) before and after the addition of 
+#' uniform variation in component response rate (gamma).
+#'
+#'@return A plot as in Figure 2  of the manuscript, showing distribution of 
+#'eigenvalues before and after the addition of uniform variation in component 
+#'response rate
+#'@examples
+#'plot_Fig_2();
+#'@useDynLib random_matrix_stability
+#'@export
+plot_Fig_2 <- function(){
+    A_comp <- NULL;
+    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+    A_mat  <- matrix(data = A_dat, nrow = 1000);
+    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+    A_mat     <- A_mat * C_mat;
+    gammas <- runif(n = 1000, min = 0, max = 2);
+    mu_gam <- mean(gammas);
+    diag(A_mat) <- -1;
+    A1     <- gammas * A_mat;
+    A0     <- mu_gam * A_mat;
+    A0_e   <- eigen(A0)$values;
+    A0_r   <- Re(A0_e);
+    A0_i   <- Im(A0_e);
+    A1_e   <- eigen(A1)$values;
+    A1_r   <- Re(A1_e);
+    A1_i   <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(A0_vm);
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(A1_vm);
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1);
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
[…]

### 2018-05-25 R/Figure_1.R
Restructure everything

@@ -14,7 +14,6 @@
 #'@param iters Number of random matrices to simulate
 #'@examples
 #'find_bgamma(S = 200, C = 0.05, Osd = 0.4, iters = 2);
-#'@useDynLib random_matrix_stability
 #'@export
 find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
     while(iters > 0){
@@ -56,7 +55,6 @@ find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
 #'@examples
 #'load(bi_pr_st);
 #'plot_Fig_1(A0 = A0, A1 = A1);
-#'@useDynLib random_matrix_stability
 #'@export
 plot_Fig_1 <- function(A0, A1){
     S_val       <- dim(A0)[1];
@@ -117,7 +115,6 @@ plot_Fig_1 <- function(A0, A1){
 #'@param iters Number of random matrices to simulate
 #'@examples
 #'dat <- stab_bgamma(iters = 4);
-#'@useDynLib random_matrix_stability
 #'@export
 stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
     ress     <- matrix(data = 0, nrow = iters, ncol = 2);

### 2018-05-25 R/Figure_2.R

@@ -8,7 +8,6 @@
 #'response rate
 #'@examples
 #'plot_Fig_2();
-#'@useDynLib random_matrix_stability
 #'@export
 plot_Fig_2 <- function(){
     A_comp <- NULL;

### 2018-05-25 R/random_mats.R
More adds

@@ -1,0 +1,140 @@
+#' Find a stabilised system
+#' 
+#' Compares random matrices in which variation in component response rate does
+#' not vary to random matrices in which this variation for 2 to max_sp 
+#' components.
+#'
+#'@return A table of stability results, where rows summarise for each component
+#'number (S) the number of stable or unstable (also, feasible and infeasible)
+#'random matrices produced.
+#'@param max_sp Maximum number of components to randomise
+#'@param iters Number of iterations (i.e., random matrices) per component
+#'@param int_type Type of interaction between components (0 is random)
+#'@param rmx Standard deviation of non-zero matrix element components
+#'@param C Connectedness of matrices (i.e., probability of non-zero matrix 
+#'element components.
+#'@examples
+#'rand_gen_var(max_sp = 2, iters = 4);
+#'@export
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    for(i in 2:max_sp){
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+        while(iter > 0){
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
+            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            A0       <- A0 * C_mat;
+            diag(A0) <- -1;
+            gam1     <- runif(n = i, min = 0, max = 2);
+            A1       <- A0 * gam1;
+            A0       <- A0 * mean(gam1);
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
+            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
[…]

### 2018-05-25 notebook/SI.Rmd
Add more to rework the project

@@
 
+```{r}
+data(ecology);
+```
+
 ```{r, echo = FALSE}
-cdat   <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
-cdat   <- cdat[,-1];
 cd_tab <- cdat[,1:7];
@@
 ```{r, echo = FALSE}
-mdat   <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
-mdat   <- mdat[,-1];
 md_tab <- mdat[,1:7];
@@
 ```{r, echo = FALSE}
-pdat   <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
-pdat   <- pdat[,-1];
 pd_tab <- pdat[,1:7];

### 2018-05-26 notebook/SI.Rmd
Update SI

@@
 ```{r, echo = FALSE}
-source(file = "R/sim_mat.R");
-source(file = "R/plot_figs.R");
+source(file = "R_old/sim_mat.R");
+source(file = "R_old/plot_figs.R");
 ```
+
+**Note: All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability) as part of the RandomMatrixStability package**.
+The text below explains how all results from the main manuscript can be recreated using the RandomMatrixStability package available on [GitHub](https://github.com/bradduthie/random_matrix_stability).
+To download this package, which includes all functions and tools for recreating the text, this supplemental information, and running all code, the [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
+
+```{r, eval = FALSE}
+install.packages("devtools");
+library(devtools);
+```
+
+The code below installs the RandomMatrixStability package using devtools.
+
+```{r, eval = FALSE}
+install_github("bradduthie/random_matrix_stability")
+```
+
+While downloading this package is recommended, all relevant code is also reproduced below with explanation.
 
@@
 ================================================================================
-
-**Note: All code and data are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability)**
 
@@
 
+The above functions produced the `bi_pr_st` data.
+
 ```{r, echo = TRUE}
-load(bi_pr_st);
+data(bi_pr_st);
 pr_st  <- bi_pr_st[,-1];
@@
 
-```{r, echo = FALSE}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
+Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations. The data used to make this figure is read into R below.
+
+```{r, echo = TRUE}
+dat <- data(random_all);
+dat <- dat[,-1]; # Extra row indicating column removed
 ```
@@
 This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#Incr_S) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
+This produced the data set below.
+
+```{r}
+data(ecology);
+```
+
 The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#Incr_S).
@@
 Results for competitor interaction networks are shown below
-
-```{r}
-data(ecology);
-```
 
@@
 
+All data reported below for various values of $C$ are accessible using the below.
+
+```{r}
+data(connectedness);
+```
+
+This includes the objects `C3dat`, `C5dat`, `C7dat`, and `C9dat`, which includes the results for $C = 0.3$, $C = 0.5$, $C = 0.7$, and $C = 0.9$, respectively.
+
 <a name="connect3">**Connectance $C = 0.3$**</a>
@@
 ```{r, echo = FALSE}
-C3dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
 C3dat   <- C3dat[,-1];
@@
 ```{r, echo = FALSE}
-C5dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
 C5dat   <- C5dat[,-1];
@@
 ```{r, echo = FALSE}
-C7dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
 C7dat   <- C7dat[,-1];
@@
 ```{r, echo = FALSE}
-C9dat   <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
 C9dat   <- C9dat[,-1];

### 2018-05-26 R/gammas.R
Functions associated with distributions of gamma initialised

@@ -1,0 +1,89 @@
+#' Make a distribution of gamma values
+#' 
+#' Makes one of five distributions of gamma values, including a constant value
+#' (0), a uniform distribution (1), an exponential distribution (2), a beta
+#' distribution (3), or a gamma distribution (4).
+#'
+#'@return A vector with elements sampled froma set distribution
+#'@param nn The number of elements in the returned vector
+#'@param distribution The type of distribution to return
+#'@param mn The mean of the vector to be returned
+#'@param sdd The standard devation of the vector to be returned (where 
+#'applicable)
+#'@examples
+#'gammas <- make_gammas(nn = 8);
+#'@export
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- 2;
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
+
+#' Produce figure of gamma distributions
+#' 
+#' Produces a four panel histogram of gamma distributions from the make_gammas
+#' funciton, including a uniform distribution, exponential distribution,
+#' beta distribution, and gamma distribution
+#'
+#'@return A four panel plot showing histograms of distributions
+#'@param nn The number of samples from distributions to make the histogram
[…]

### 2018-05-26 notebook/SI.Rmd
New results for evolutionary output included in SI

@@
 ```{r, echo = TRUE}
-fordist   <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
-kable(fordist);
+data(fourdists)
+kable(fourdists);
 ```
@@
 ```{r}
-summarise_randmat <- function(tot_res, fea_res){
+summarise_randmat_ga <- function(tot_res, fea_res){
     sims    <- length(tot_res);
@@
 ```{r, echo = FALSE}
-evo_res   <- read.csv(file = "sim_results/evolved/evo_results.csv");
+data(evo_res);
 kable(evo_res);
@@
 The distribution of one of the $\gamma$ vectors at $S = 39$ is shown below (all are available on [GitHub]()).
+Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
+<!--- NOTE: This needs to be updated with the new data --->
+
+```{r}
+data(evo_out);
+plot_evo_out(evo_out);
+```
+
+The distribution of $\gamma$ values found by the genetic algorithm is uniform.
 

### 2018-05-26 R/ga.R

@@ -1,0 +1,260 @@
+#' Run genetic algorithms to find stability
+#' 
+#' Outer function for investigating the potential for a complex system to be
+#' stabilised by component response rates. For a stretch of values of component
+#' number (i.e., system size, S), the function runs some number of iterations, 
+#' and in each iteration, if a system is not initially found to be stable with
+#' a set of component response times, then a genetic algorithm is called in 
+#' attempt to find a stable solution.
+#'
+#'@return A table showing frequency of stable complex systems found
+#'@param from The lowest system size, S
+#'@param to The highest system size, S
+#'@param iters The number of iterations (complex systems) tried for each S
+#'@param int_type The type of interaction, including random (0), competitive 
+#'(1), mutualist (2), or predator-prey (3)
+#'@param rmx The standard devation of interaction (off-diagonal) elements in a
+#'complex system
+#'@param C The connectedness of a complex system (i.e., the probability that an
+#'off-diagonal element does not equal zero)
+#'@examples
+#'sim <- Evo_rand_gen_var(from = 2, to = 2, iters = 1);
+#'@export
+Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    if(from >= to){
+        stop("Argument 'from' must be less than argument 'to'");
+    }
+    for(i in from:to){
+        nn             <- i;
+        A1_stt         <- 0;
+        A2_stt         <- 0;
+        A1_fet         <- 0;
+        A2_fet         <- 0;
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        while(iter > 0){
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
[…]

### 2018-05-28 notebook/SI.Rmd
Update notebook files

@@
 ```{r, echo = FALSE}
-load(bg_data);
+data(bg_data);
 plot_Fig_1 <- function(A0, A1){

### 2018-05-28 notebook/SI.Rmd
Progress

@@
 
-```{r, echo = FALSE}
-data(bg_data);
+```{r, echo = FALSE, fig.height = 6, fig.width = 9}
+A0 <- read.csv("sim_results/bi_gamma/S200_A0.csv");
+A1 <- read.csv("sim_results/bi_gamma/S200_A1.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- as.matrix(A1[,-1]);
 plot_Fig_1 <- function(A0, A1){
@@
 ```{r, echo = TRUE}
-data(bi_pr_st);
+bi_pr_st <- read.csv("sim_results/bi_gamma/bi_pr_st.csv");
 pr_st    <- bi_pr_st[,-1];
@@
 ```{r, echo = TRUE}
-dat <- data(random_all);
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
 dat <- dat[,-1]; # Extra row indicating column removed
@@
 ```{r}
-data(ecology);
+cdat <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
+mdat <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
+pdat <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
 ```
@@
 ```{r}
-data(connectedness);
+C3dat <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
+C5dat <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
+C7dat <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
+C9dat <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
 ```
@@
 ```{r, echo = TRUE}
-data(fourdists)
+fourdists <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
 kable(fourdists);
@@
 ```{r, echo = FALSE}
-data(evo_res);
+plot_evo_out <- function(evo_out){
+    gammas <- get_top_evo_out(evo_out, size = 6);
+    par(mfrow = c(3, 2), mar = c(0.25, 0.25, 0.25, 0.25), oma = c(5, 5, 1, 1));
+    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1));
+    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1));
+    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1));
+    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1));
+    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xlim = c(0, 1));
+    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1));
+    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
+          text = expression(paste("Component response rate (",gamma,")")));
+    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
+}
+get_top_evo_out <- function(evo_out, size){
+    highest <- max(evo_out);
+    gammas  <- NULL;
+    while(size > 0){
+        pos <- which(evo_out == highest);
+        len <- length(pos);
+        nli <- NULL;
+        for(i in 1:len){
+            start          <- pos[i] + 1;
+            end            <- pos[i] + highest;
+            gammas[[size]] <- evo_out[start:end];
+            size           <- size - 1;
+            if(size == 0){
+                break;
+            }
+        }
+        highest <- highest - 1;
+    }
+    return(gammas);
+}
+evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
 kable(evo_res);
@@
 ```{r}
-data(evo_out);
+evo_out <- scan(file = "sim_results/evolved/Evo_out.txt");
 plot_evo_out(evo_out);

### 2018-05-28 notebook/ms.Rmd
Progress

--- 2018-05-22 ms.Rmd
@@
 title: "Component response rate variation drives stability in large complex systems"
-author: "Brad Duthie"
-bibliography: references.bib
+author: "A. Bradley Duthie"
+date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
+  FK9 4LA alexander.duthie@stir.ac.uk
 output:
+  html_document: default
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
   word_document:
@@
     reference_docx: docx_template.docx
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-  html_document: default
+bibliography: references.bib
 header-includes:
@@
 <!---
+Current word count: 1683
+--->
+
+<!---
 "This will generate the intermediate files that are actually used to generate the PDF.
@@
 ```{r, echo = FALSE}
-source(file = "R/sim_mat.R");
-source(file = "R/plot_figs.R");
+source(file = "R_old/sim_mat.R");
+source(file = "R_old/plot_figs.R");
 ```
@@
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were frequently found that resulted in stable systems (the highest at $S = 38$), suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were frequently found that resulted in stable systems (the highest at $S = 39$), suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
 

### 2018-05-28 notebook/ms.tex
Progress

@@
             pdftitle={Component response rate variation drives stability in large complex systems},
-            pdfauthor={Brad Duthie},
+            pdfauthor={A. Bradley Duthie},
             colorlinks=true,
@@
   \posttitle{\par}
-  \author{Brad Duthie}
+  \author{A. Bradley Duthie}
   \preauthor{\centering\large\emph}
   \postauthor{\par}
-  \date{}
-  \predate{}\postdate{}
+  \predate{\centering\large\emph}
+  \postdate{\par}
+  \date{Biological and Environmental Sciences, University of Stirling, Stirling,
+UK, FK9 4LA
+\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}
 
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.547 and 0.557 for a and b, respectively.
+of \(\Re(\lambda) < 0\) are 0.551 and 0.553 for a and b, respectively.
 
@@
 \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were frequently
-found that resulted in stable systems (the highest at \(S = 38\)),
+found that resulted in stable systems (the highest at \(S = 39\)),
 suggesting that varying component response time might by itself be

### 2018-05-28 R/ga.R
Add all else

@@ -8,8 +8,7 @@
 #' attempt to find a stable solution.
 #'
 #'@return A table showing frequency of stable complex systems found
-#'@param from The lowest system size, S
-#'@param to The highest system size, S
+#'@param max_sp The maximum system component size (S) to simulate
 #'@param iters The number of iterations (complex systems) tried for each S
 #'@param int_type The type of interaction, including random (0), competitive 
 #'(1), mutualist (2), or predator-prey (3)
@@ -18,15 +17,12 @@
 #'@param C The connectedness of a complex system (i.e., the probability that an
 #'off-diagonal element does not equal zero)
 #'@examples
-#'sim <- Evo_rand_gen_var(from = 2, to = 2, iters = 1);
+#'sim <- Evo_rand_gen_var(max_sp = 2, iters = 1);
 #'@export
-Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
+Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
     tot_res <- NULL;
     fea_res <- NULL;
-    if(from >= to){
-        stop("Argument 'from' must be less than argument 'to'");
-    }
-    for(i in from:to){
+    for(i in 2:max_sp){
         nn             <- i;
         A1_stt         <- 0;
         A2_stt         <- 0;
@@ -232,7 +228,7 @@ get_top_evo_out <- function(evo_out, size){
 #'@return A list of vectors of component response rates (gammas)
 #'@param evo_out Output of component response values
 #'@export
-plot_evo_out(evo_out){
+plot_evo_out <- function(evo_out){
     gammas <- get_top_evo_out(evo_out, size = 6);
     par(mfrow = c(3, 2), mar = c(0.25, 0.25, 0.25, 0.25), oma = c(5, 5, 1, 1));
     hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),

### 2018-05-29 R/ga.R
Try to change figure xlim values

@@ -230,19 +230,19 @@ get_top_evo_out <- function(evo_out, size){
 #'@export
 plot_evo_out <- function(evo_out){
     gammas <- get_top_evo_out(evo_out, size = 6);
-    par(mfrow = c(3, 2), mar = c(0.25, 0.25, 0.25, 0.25), oma = c(5, 5, 1, 1));
+    par(mfrow = c(3, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
     hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1));
+         xaxt = "n", xlim = c(0, 1.3));
     hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.3));
     hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1));
+         xaxt = "n", xlim = c(0, 1.3));
     hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.3));
     hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xlim = c(0, 1));
+         xlim = c(0, 3));
     hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1));
+         yaxt = "n", xlim = c(0, 1.3));
     mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
           text = expression(paste("Component response rate (",gamma,")")));
     mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);

### 2018-06-01 notebook/ms.Rmd
Minor edits, particularly emphasising the genetic algorithm results

@@
 output:
-  html_document: default
   pdf_document:
@@
     keep_tex: yes
+  html_document: default
   word_document:
@@
 This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
-I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn from $\mathcal{N}(0, \sigma^{2})$.
+I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
 Rows are then multiplied by $\gamma_{i}$ to generate $M$.
@@
 
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from $2-32$ (see Supplementary Information for different values of $\sigma$ and $C$).
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32 (see Supplementary Information for different values of $\sigma$ and $C$).
 One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$).
@@
 
-Some care is needed when interpreting these results.
+Some caution is needed when interpreting these results.
 First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm (the space of possible $\gamma$ values was too large to searh exhaustively[@Hamblin2013]; see Supplementary Information).
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
 For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
@@
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were frequently found that resulted in stable systems (the highest at $S = 39$), suggesting that varying component response time might by itself be sufficient to stabilise complex systems.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supporting Information).
+Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems.
 
 I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
-These include systems in which all interactions (i.e., all off-diagonal elements of $M$) are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
+These include systems in which all interactions are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
 In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
-When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist communities does not (see Supplementary Information).
+When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
 Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).

### 2018-06-01 notebook/ms.tex
Minor edits, particularly emphasising the genetic algorithm results

@@
 \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
-set to \(-1\) and non-zero off-diagonal elements are drawn from
+set to \(-1\) and non-zero off-diagonal elements are drawn randomly from
 \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.551 and 0.553 for a and b, respectively.
+of \(\Re(\lambda) < 0\) are 0.549 and 0.557 for a and b, respectively.
 
@@
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from \(2-32\) (see Supplementary Information for different
+\(S\) ranging from 2-32 (see Supplementary Information for different
 values of \(\sigma\) and \(C\)). One million \(M\) were simulated for
@@
 
-Some care is needed when interpreting these results. First, I emphasise
-that \(Var(\gamma)\) is not stabilising per se; that is, adding
-variation in \(\gamma\) to a particular system \(M\) does not
+Some caution is needed when interpreting these results. First, I
+emphasise that \(Var(\gamma)\) is not stabilising per se; that is,
+adding variation in \(\gamma\) to a particular system \(M\) does not
 necessarily increase the probability that the system will be stable (see
@@
 I used a simple genetic algorithm (the space of possible \(\gamma\)
-values was too large to searh
+values was too large to search
 exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
@@
 minimum fitness increase between generations was satisfied. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were frequently
-found that resulted in stable systems (the highest at \(S = 39\)),
-suggesting that varying component response time might by itself be
-sufficient to stabilise complex systems.
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were orders of
+magnitude higher than when \(\gamma = 1\) (see Supporting Information).
+Hence, managing the response rates of system components in a targetted
+way can potentially facilitate the stabilisation of complex systems.
 
@@
 networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions (i.e., all off-diagonal
-elements of \(M\)) are negative (e.g., competitive networks), positive
-(e.g., mutualist networks), or \(i\) and \(j\) pairs have opposing signs
-(e.g., predator-prey networks). In general, competitive and mutualist
-networks tend to be destabilising, and predator-prey network tend to be
+include systems in which all interactions are negative (e.g.,
+competitive networks), positive (e.g., mutualist networks), or \(i\) and
+\(j\) pairs have opposing signs (e.g., predator-prey networks). In
+general, competitive and mutualist networks tend to be destabilising,
+and predator-prey network tend to be
 stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
@@
 competitive and predator-prey networks increases, but the proportion of
-stable mutualist communities does not (see Supplementary Information).
+stable mutualist networks does not (see Supplementary Information).
 Additionally, when each component of \(M\) is interpreted as a unique

### 2018-06-01 R/ga.R
Small changes to plotting function

@@ -229,20 +229,26 @@ get_top_evo_out <- function(evo_out, size){
 #'@param evo_out Output of component response values
 #'@export
 plot_evo_out <- function(evo_out){
-    gammas <- get_top_evo_out(evo_out, size = 6);
-    par(mfrow = c(3, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
+    gammas <- get_top_evo_out(evo_out, size = 9);
+    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
     hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.3));
+         xaxt = "n", xlim = c(0, 1.1));
     hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.3));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
     hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.3));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
     hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.3));
+         xaxt = "n", xlim = c(0, 1.1));
     hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xlim = c(0, 3));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
     hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1.3));
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xlim = c(0, 1.1));
+    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
[…]

### 2018-06-01 notebook/SI.Rmd
Attempt to increase figure size
Edits and complete the genetic algorithm sectoin

--- 2018-05-28 notebook/SI.Rmd
@@
 title: "Component response rate variation drives stability in large complex systems"
+subtitle: "Supplemental Information"
 author: "Brad Duthie"
+date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
+  FK9 4LA alexander.duthie@stir.ac.uk
 bibliography: references.bib
 output:
-  html_document: default
   pdf_document:
     fig_caption: yes
+  html_document: default
   word_document:
@@
 csl: nature.csl
-subtitle: Supporting information
 biblio-style: apalike
@@
 
-**Note: All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability) as part of the RandomMatrixStability package**.
-The text below explains how all results from the main manuscript can be recreated using the RandomMatrixStability package available on [GitHub](https://github.com/bradduthie/random_matrix_stability).
-To download this package, which includes all functions and tools for recreating the text, this supplemental information, and running all code, the [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
+********************************************************************************
+
+**This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with all of the code required to recreate the analysis in the main text, and with additional analyses to support its conclusions.
+All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability) as part of the RandomMatrixStability package.**
+
+The [RandomMatrixStability package](https://github.com/bradduthie/random_matrix_stability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for functions as part of the package.
+The RandomMatrixStability package package is available on [GitHub](https://github.com/bradduthie/random_matrix_stability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
 
@@
 
-While downloading this package is recommended, all relevant code is also reproduced below with explanation.
+While downloading this package is recommended, all relevant code is also reproduced below with explanation, so it is possible to recreate all analyses using only this supplemental information.
 
-Contents
+********************************************************************************
+
+Supplemental Information table of contents
 ================================================================================
@@
 - [Code and simulations underlying Fig. 2](#Fig2) 
-- [Stability across increasing $S$](#Incr_S}) 
+- [Stability across increasing $S$](#IncrS) 
 - [Stability of ecological networks](#ecological) 
@@
 - [Genetic algorithm](#ga)
+
+********************************************************************************
 
@@
 
-The above function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$.
+The above `find_bgamma` function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$.
 The function is used to illustrate the concept of how fast versus slow component responses can cause a system to become stable.
@@
 
-```{r, echo = FALSE}
-data(bg_data);
+```{r, echo = FALSE, fig.height = 5, fig.width = 8}
+A0 <- read.csv("sim_results/bi_gamma/S200_A0.csv");
+A1 <- read.csv("sim_results/bi_gamma/S200_A1.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- as.matrix(A1[,-1]);
 plot_Fig_1 <- function(A0, A1){
@@
 
-The function above was run for `iters = 1000000`, and the resulting matrix `ress` was returned.
+The function `stab_bgamma` was run for `iters = 1000000`, and the resulting matrix `ress` was returned.
 Each row of `ress` represents a single $M$ given $\gamma = 1$ (column 1) versus $\gamma = \{1.95, 0.05\}$ (column 2).
@@
 
-Stability across increasing $S$ {#Incr_S}
+Stability across increasing $S$ {#IncrS}
 ===============================================================================
 
-Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations. The data used to make this figure is read into R below.
+Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations. The data used to make this figure are read into R below.
 
 ```{r, echo = TRUE}
-dat <- data(random_all);
-dat <- dat[,-1]; # Extra row indicating column removed
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1]; # Extra row-indicating column removed
 ```
@@
 
+Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
+Hence, more randomly created complex systems ($M$) are generated given variation in $\gamma$ than when $\gamma = 1$.
 The results underlying this table were produced with the `rand_gen_var` function below.
@@
 
-I investigate competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012].
+I investigated competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012].
 To create these networks, I first generated a random matrix $M$, then changed the elements of $M$ accordingly.
@@
 
-This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#Incr_S) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
+This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
 This produced the data set below.
@@
 ```{r}
-data(ecology);
+cdat <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
+mdat <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
+pdat <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
 ```
 
-The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#Incr_S).
-As with the [results above](#Incr_S), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added.
+The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#IncrS) (but note that counts from systems with comparable probabilities of stability will be an order of magnitude lower in the tables below due to the smaller number of $M$ matrices generated).
+As with the [results above](#IncrS), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added.
 The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
@@
 
-Different connectance (C) values {#connectance}
+Different inter-connectivity (C) values {#connectance}
 ================================================================================
 
-In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $M$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
+In the main text, for simplicity, I assumed inter-connectivity values of $C = 1$, meaning that all off-diagonal elements of a matrix $M$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
 Here I present four tables showing the number of stable communities given $C = \{0.3, 0.
@@
 In contrast to the main text, 100000 rather than 1 million $M$ were simulated.
-As with the results on [stability with increasing $S$](#Incr_S) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added.
+As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added.
 The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
@@
 ```{r}
-data(connectedness);
+C3dat <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
+C5dat <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
+C7dat <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
+C9dat <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
 ```
 
-This includes the objects `C3dat`, `C5dat`, `C7dat`, and `C9dat`, which includes the results for $C = 0.3$, $C = 0.5$, $C = 0.7$, and $C = 0.9$, respectively.
+These objects `C3dat`, `C5dat`, `C7dat`, and `C9dat` include the results for $C = 0.3$, $C = 0.5$, $C = 0.7$, and $C = 0.9$, respectively.
 
@@
 In the main text, I considered a uniform distribution of component response rates $\gamma \sim \mathcal{U}(0, 2)$.
-The number of unstable and stable $M$ matrices are reported in [a table above](#Incr_S) across different values of $S$.
+The number of unstable and stable $M$ matrices are reported in [a table above](#IncrS) across different values of $S$.
 Here I show complementary results for three different distributions including an exponential, beta, and gamma distribution of $\gamma$ values.
@@
 
-Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluate for each unique $M$.
-This is technically impossible because any $\gamma_{i}$ can take any real value between $0-2$, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $M$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
 Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
 In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
-The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below 20 generations pass or the fitness increase from one generation to the next is below a certain criteria).
+The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
 The genetic algorithm relies on five functions.
-The first outer function `Evo_rand_gen_var` runs all of the simulations (`from` and `to` refer to $S$ values, and `iters` refers to the number of $M$ to try for each $S$).
+The first outer function `Evo_rand_gen_var` runs all of the simulations (`max_sp` refers to the maximum $S$ value simulated, and `iters` refers to the number of $M$ to try for each $S$).
 
 ```{r}
-Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
+Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
     tot_res <- NULL;
     fea_res <- NULL;
-    if(from >= to){
-        stop("Argument 'from' must be less than argument 'to'");
-    }
-    for(i in from:to){
+    for(i in 2:max_sp){
         nn             <- i;
@@
 
-The while loop continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed.
-Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and there is selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
-In mutation, any values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
+The while loop in `rand_mat_ga` continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed (by default, `converg = 0.01`).
+Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
+In mutation, any $\gamma_{i}$ values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
 Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
@@
 
-Stability results from this table are shown below.
+Some stability results from this table are shown below. Each histogram shows a different distribution of $\gamma$ that was found to be stabilising.
 
 ```{r, echo = FALSE}
-data(evo_res);
+plot_evo_out <- function(evo_out){
+    gammas <- get_top_evo_out(evo_out, size = 9);
+    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
+    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xlim = c(0, 1.1));
+    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1.1));
+    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
+          text = expression(paste("Component response rate (",gamma,")")));
+    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
+}
+get_top_evo_out <- function(evo_out, size){
+    highest <- max(evo_out);
+    gammas  <- NULL;
+    while(size > 0){
+        pos <- which(evo_out == highest);
+        len <- length(pos);
+        nli <- NULL;
+        for(i in 1:len){
+            start          <- pos[i] + 1;
+            end            <- pos[i] + highest;
+            gammas[[size]] <- evo_out[start:end];
+            size           <- size - 1;
+            if(size == 0){
+                break;
+            }
+        }
+        highest <- highest - 1;
+    }
+    return(gammas);
+}
+evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
 kable(evo_res);
@@
 
-The distribution of one of the $\gamma$ vectors at $S = 39$ is shown below (all are available on [GitHub]()).
+The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
+Recall that 1 million random matrices were generated for the less computationally intense task of [comparing](#IncrS) $M$ when $\gamma = 1$ versus when $\gamma \sim \mathcal{U}(0, 2)$, so it is more informative to compare stability in column 5 above with column 3 above.
+**This comparison shows the high number of stable $M$ that can be produced through a targetted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
+Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.**
+
 Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
-<!--- NOTE: This needs to be updated with the new data --->
 
-```{r}
-data(evo_out);
+```{r, fig.height = 6, fig.width = 7}
+evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
 plot_evo_out(evo_out);
@@
 The distribution of $\gamma$ values found by the genetic algorithm is uniform.
-
-**Work in progress**
+A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $M$.
 

### 2018-06-01 notebook/SI.Rmd
Update feasibility part

@@
 
-Feasilbility of complex systems {#Feasibility}
+Feasibility of complex systems {#Feasibility}
 ================================================================================
 
-**Work in progress**
+For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
+Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
+Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
+While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
+This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
 
-It also would be useful to look at feasibility criteria established by @Dougoud2018, who very recently made the point that some of May and Allesina's criteria allows for negative species densities when stable.
-Feasibility criteria are as follows,
+Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
 
 $$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+
 The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.

### 2018-06-01 notebook/SI.Rmd
First complete draft

@@
 ```{r, echo = FALSE}
+get_top_evo_out <- function(evo_out, size){
+    highest <- max(evo_out);
+    gammas  <- NULL;
+    while(size > 0){
+        pos <- which(evo_out == highest);
+        len <- length(pos);
+        nli <- NULL;
+        for(i in 1:len){
+            start          <- pos[i] + 1;
+            end            <- pos[i] + highest;
+            gammas[[size]] <- evo_out[start:end];
+            size           <- size - 1;
+            if(size == 0){
+                break;
+            }
+        }
+        highest <- highest - 1;
+    }
+    return(gammas);
+}
 plot_evo_out <- function(evo_out){
@@
     mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
-}
-get_top_evo_out <- function(evo_out, size){
-    highest <- max(evo_out);
-    gammas  <- NULL;
-    while(size > 0){
-        pos <- which(evo_out == highest);
-        len <- length(pos);
-        nli <- NULL;
-        for(i in 1:len){
-            start          <- pos[i] + 1;
-            end            <- pos[i] + highest;
-            gammas[[size]] <- evo_out[start:end];
-            size           <- size - 1;
-            if(size == 0){
-                break;
-            }
-        }
-        highest <- highest - 1;
-    }
-    return(gammas);
 }
@@
 
-The above is not nearly as nasty as it looks, especially because it is entirely reasonable to simply use convenient values to parameterise it.
-The variable $x^{*}$ is just the vector of species abundances at equilibrium (we need all of them to be positive).
-The matrix $I$ is just the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is just the strength of intraspecific competition -- we can just set this to $\theta = -1$ (as others have).
-The variable $C$ is just the connectance of the community, which we will also set to 1 for convenience (lower values would mean that some species don't interact with one another, corresponding to an off-diagonal matrix element of 0).
-And $\delta$ just affects the strength of interactions -- which we can set to $\delta = 0$ for strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so we're just adding the diagonal matrix of -1s ($\theta I$) to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions,
+In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive).
+The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements).
+As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
+The variable $C$ is just the interconnectedness of the community, which was set to $C = 1$ throughout the manuscript and supporing information, except [where otherwise noted](#connectance).
+The variable $\delta$ just affects the strength of interactions, which are held in $A$.
+In the main text, $\delta = 0$ underlying strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions.
+The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information.
 
@@
 
-To check the feasibility criteria, all that needs to be done is to invert $(\theta I + A)$ and multiply the matrix by the vector of growth rates $r$, which we can also just set to 1.
-So really, we're just multiplying $-(\theta I + A)$ by a vector of ones and checking to make sure that all the values are positive.
-This isn't much extra work, but it will probably go a long way toward satisfying any reviewers familiar with @Dougoud2018.
+To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$.
+Feasibility is satisfied if all of the elements of the resulting vector are positive.
+
+The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$, as shown in the `rand_gen_var` function in [the section](#IncrS) on "Stability across increasing $S$" above.
+Hence, each component $i$ of the complex system $M$ is assumed to be a species with a growth rate of $r_{i}$.
+Note that negative intrinsic growth rates are not unrealistic, and will occur in obligate mutualists in the absence of a partner.
+
+When feasibility was evaluated with and without variation in $\gamma$, there was no increase in stability for $M$ where $\gamma$ varied as compared to where $\gamma = 1$.
+Results below illustrate this result, which was general to all other simulations performed.
+
+```{r, echo = FALSE}
+dat      <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat      <- dat[,-1]; # Extra row-indicating column removed
+feas_tab <- cbind(dat[,1], dat[,8:13]);
+colnames(feas_tab)[1] <- "S";
+kable(feas_tab);
+```
+
+Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks.
 

### 2018-06-01 notebook/SI.Rmd
Add more notebook files; final draft edits

@@
 title: "Component response rate variation drives stability in large complex systems"
-subtitle: "Supplemental Information"
 author: "Brad Duthie"
@@
   FK9 4LA alexander.duthie@stir.ac.uk
-bibliography: references.bib
 output:
@@
     reference_docx: docx_template.docx
+bibliography: references.bib
 header-includes:
@@
 csl: nature.csl
+subtitle: Supplemental Information
 biblio-style: apalike
@@
 - [Different distributions of $\gamma$](#gam_dist)
+- [Genetic algorithm](#ga) 
 - [Feasilbility of complex systems](#Feasibility)
-- [Genetic algorithm](#ga)
 

### 2018-06-01 notebook/SI.Rmd
Fix spelling

@@
 - [Genetic algorithm](#ga) 
-- [Feasilbility of complex systems](#Feasibility)
+- [Feasibility of complex systems](#Feasibility)
 

### 2018-06-01 notebook/SI.Rmd
Minor edits to feasibility; name correction

@@
 title: "Component response rate variation drives stability in large complex systems"
-author: "Brad Duthie"
+author: "A. Bradley Duthie"
 date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
@@
 As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
-The variable $C$ is just the interconnectedness of the community, which was set to $C = 1$ throughout the manuscript and supporing information, except [where otherwise noted](#connectance).
-The variable $\delta$ just affects the strength of interactions, which are held in $A$.
-In the main text, $\delta = 0$ underlying strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal effecting species interactions.
-The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information.
+The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supporing information, except [where otherwise noted](#connectance).
+The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
+In the main text, implicitly, $\delta = 0$ underlying strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
+The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
 

### 2018-06-02 notebook/SI.Rmd
Prep everything for arXiv

@@
     fig_caption: yes
+    keep_tex: yes
   html_document: default
@@
 As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
-The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supporing information, except [where otherwise noted](#connectance).
+The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
 The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.

### 2018-06-02 notebook/arxiv/ms.tex
Prep everything for arXiv

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Component response rate variation drives stability in large complex systems},
            pdfauthor={A. Bradley Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Component response rate variation drives stability in large complex systems},
            pdfauthor={A. Bradley Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

\setlength{\droptitle}{-2em}
  \title{Component response rate variation drives stability in large complex
systems}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{A. Bradley Duthie}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\bibliographystyle{amnatnat}

\begin{document}
\maketitle

\textbf{The stability of a complex system generally decreases with
increasing system size, as is demonstrated by random matrix
theory\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
This counter-intuitive result, first shown by
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}, is broadly
relevant for understanding the dynamics and persistence of systems such
as
ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}},
neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{3},\protect\hyperlink{ref-Gray2009}{4}},
biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{5},\protect\hyperlink{ref-MacArthur2010}{6}}
and
socio-economic\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}--\protect\hyperlink{ref-Bardoscia2017}{9}}
networks. Much attention has especially been given to the stability of
ecological communities such as food webs or mutualist networks, with
recent work investigating how different community structures affect
stability\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10}--\protect\hyperlink{ref-Patel2018}{14}}.
But more broadly, stabilising mechanisms in complex systems remain
under-developed, and the effect of variation in the response rate of
individual system components remains an open
problem\textsuperscript{\protect\hyperlink{ref-Allesina2015}{15}}. Here
I show that when components of a complex system respond to system
dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
system stability is markedly increased. Stability is caused by the
clustering of some eigenvalues toward the centre of eigenvalue
distributions despite the destabilising effect of higher interaction
strength variation (\(\boldsymbol{\sigma^{2}}\)). This effect of
variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
system size increases, to the extent that the largest stable complex
systems would otherwise be unstable if not for
\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
unconsidered driver of system stability that is likely to be pervasive
across all complex systems. Future research in complex systems should
therefore account for the varying response rates of individual system
components when assessing whole system stability.}

In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
demonstrated that randomly assembled systems of sufficient complexity
are almost inevitably unstable given infinitesimally small
perturbations. Complexity in this case is defined by the size of the
system (i.e., the number of interacting components; \(S\)), its
inter-connectivity (i.e., the probability that one component will affect
another; \(C\)), and the variance of interaction strengths
(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
May's finding that the probability of local stability falls to near zero
given a sufficiently high threshold of \(\sigma\sqrt{SC}\) has profound
consequences across multiple disciplines, raising the question of how
complex systems in, e.g.,
ecology\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10},\protect\hyperlink{ref-Grilli2017}{13},\protect\hyperlink{ref-Allesina2015}{15}}
or
banking\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7},\protect\hyperlink{ref-Bardoscia2017}{9},\protect\hyperlink{ref-May2008}{16}}
are predicted to persist or change.

Randomly assembled complex systems can be represented as large square
matrices (\(M\)) with \(S\) components (e.g.,
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
element of such a matrix \(M_{ij}\) defines how component \(j\) affects
component \(i\) in the system at a point of
equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
Off-diagonal elements (\(i \neq j\)) therefore define interactions
between components, while diagonal elements (\(i = j\)) define component
self-regulation (e.g., carrying capacity in ecological communities).
Traditionally, values of off-diagonal elements are assigned non-zero
values with a probability \(C\), which are sampled from a distribution
with variance \(\sigma^{2}\); diagonal elements are set to
-1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}.
Local system stability is assessed using eigenanalysis, with the system
being stable if the real parts of all eigenvalues (\(\lambda\)) of \(M\)
are negative
(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
In a large system (high \(S\)), eigenvalues are distributed
uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}} within a
circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
\(\Im = 0\), with a radius of
\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}
(Figs 1a and 2a). Local stability of randomly assembled systems
therefore becomes increasingly unlikely as \(S\), \(C\), and
\(\sigma^{2}\) increase.

The above stability criterion assumes that individual components respond
to perturbations of the system at the same rate (\(\gamma\)), but this
is highly unlikely in any complex system. In ecological communities, for
example, the rate at which population density changes following
perturbation will depend on the generation time of individuals, which
might vary by orders of magnitude among species. Species with short
generation times will respond quickly (high \(\gamma\)) to perturbations
relative to species with long generation times (low \(\gamma\)).
Similarly, the speed at which individual banks respond to perturbations
in financial networks, or individuals or institutions respond to
perturbations in complex social networks, is likely to vary. The effect
of such variance has not been investigated in complex systems theory.
Intuitively, variation in \(\gamma\) might be expected to decrease
system stability by introducing a new source of variation into the
system and thereby increasing \(\sigma\). Here I show why, despite
higher \(\sigma\), complex systems in which \(\gamma\) varies are
actually more likely to be stable, especially when \(S\) is high.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 1: Example distribution of eigenvalues before (a) and
after (b) separating a randomly generated complex system into fast
(\(\boldsymbol{\gamma} = 1.95\)) and slow
(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
shows the same system where \(S = 200\), \(C = 0.05\), and
\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
distribution of \(\gamma\) differs between panels). \textbf{a.}
Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
uniformly distributed within the grey circle with a radius of
\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
Eigenvalues plotted when half \(\gamma = 1.95\) and half
\(\gamma = 0.05\); distributions of points can be partitioned into one
large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
centred at \(\gamma = -0.05\). In a, the maximum real eigenvalaue
\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
complex system in b but not a is stable because in b
\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
complex systems under the same parameter values, 1 was stable when
\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
Overall, complex systems that are separated into fast versus slow
components tend to be more stable than otherwise identical systems with
identical component response rates.

\includegraphics{unnamed-chunk-4-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

Rows in \(M\) define how a given component \(i\) is affected by other
components of the system, meaning that the rate of component response
time can be modelled by multiplying all row elements by a real scalar
value
\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
The distribution of \(\gamma\) over \(S\) components thereby models the
distribution of component response rates. An instructive example
compares one \(M\) where \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
\(\gamma_{i} = 0.05\). This models one system in which \(\gamma\) is
invariant and one in which \(\gamma\) varies, but systems are otherwise
identical (note \(E[\gamma_{i}] = 1\) in both cases). I assume
\(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
set to \(-1\) and non-zero off-diagonal elements are drawn randomly from
\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
\(\gamma_{i}\) to generate \(M\). When \(\gamma_{i} = 1\), eigenvalues
of \(M\) are distributed uniformly within a circle centred at
(\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real components
of eigenvalues are highly unlikely to all be negative when all
\(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated into
two groups, eigenvalues are no longer uniformly distributed (Fig. 1b).
Instead, two distinct clusters of eigenvalues appear (grey circles in
Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
(\(-0.05, 0\)). The former has a large radius, but the real components
have shifted to the left (in comparison to when \(\gamma = 1\)) and all
\(\Re({\lambda}) < 0\). The latter cluster has real components that have
shifted to the right, but has a smaller radius. Overall, for 1 million
randomly assembled \(M\), this division between slow and fast component
response rates results in more stable systems: 1 stable given
\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
introducing variation in component response rate
(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
Eigenvalues plotted in the absence of \(Var(\gamma)\) where
\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
the distribution's centre (-1, 0). Black and red elipses in both panels
show the circle centred on the distribution in panels a and b,
respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
of \(\Re(\lambda) < 0\) are 0.548 and 0.552 for a and b, respectively.

\includegraphics{unnamed-chunk-7-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

Higher stability in systems with variation in \(\gamma\) can be observed
by sampling \(\gamma_{i}\) values from various distributions. I now
focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
(see \hyperlink{SIstart}{Supplementary Information} for other distributions, 
which give
similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
comparison of \(M\) before and after variation in component response
rate. Figure 2 shows a comparison of eigenvalue distributions given
\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
Uniform variation in \(\gamma\) leads to a non-uniform distribution of
eigenvalues, some of which are clustered tightly around the centre of
the distribution, but others of which are spread outside the former
radius of 12.649 (red circle Fig 2b). This larger radius occurs because
the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
\(M\). The clustering and spreading of eigenvalues introduced by
\(Var(\gamma)\) can destabilise previously stable systems or stabilise
systems that are otherwise unstable. But where systems are otherwise too
complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
can often lead to stability above
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
threshold of \(\sigma\sqrt{SC} > 1\).

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Figure 3: Stability of large complex systems with and without
variation in component response rate(\(\boldsymbol{\gamma}\)).} The
\(\ln\) number of systems that are stable across different system sizes
(\(S\)) given \(C = 1\), and the proportion of systems in which
variation in \(\gamma\) is critical for system stability. For each
\(S\), 1 million complex systems are randomly generated. Stability of
each complex system is tested given variation in \(\gamma\) by randomly
sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
\(Var(\gamma)\) is then compared to stability in an otherwise identical
system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
Light and dark grey bars show the number of stable systems in the
absence and presence of \(Var(\gamma)\), respectively. The black line
shows the proportion of systems that are stable when
\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).

\includegraphics{unnamed-chunk-9-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

To investigate the effect of \(Var(\gamma)\) on system stability, I
simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
\(S\) ranging from 2-32 (see \hyperlink{SIstart}{Supplementary Information}
for different values of \(\sigma\) and \(C\)). One million \(M\) were simulated 
for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
versus \(\gamma \sim \mathcal{U}(0, 2)\) (note that under these
conditions, \(\sigma\sqrt{SC} = 1\) given \(S = 25\) when
\(\gamma = 1\)). I found that the number of stable random systems was
consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
3), and that the difference between the probabilities of observing a
stable system increased with an increase in \(S\); i.e., the potential
for \(Var(\gamma)\) to drive stability increased with system complexity.
For the highest values of \(S\), nearly all systems that were stable
given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\),
and the maximum observed \(S\) for which a system was stable was \(31\)
given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see
\hyperlink{SIstart}{Supplementary Information} for full results). This 
suggests that the
stability of large systems might be dependent upon variation in the
response rate of their individual components, meaning that factors such
as generation time (in ecological networks), transaction speed (in
economic networks), or communication speed (in social networks) needs to
be considered when investigating the stability of complex systems.

Some caution is needed when interpreting these results. First, I
emphasise that \(Var(\gamma)\) is not stabilising per se; that is,
adding variation in \(\gamma\) to a particular system \(M\) does not
necessarily increase the probability that the system will be stable (see
\hyperlink{SIstart}{Supplementary Information}). Rather, systems that are 
observed to be stable are more likely to vary in \(\gamma\), and for this
\(Var(\gamma)\) to be critical to their stability. This is caused by the
shift in the distribution of eigenvalues that occurs by introducing
\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
further investigate the potential of \(Var(\gamma)\) to be stabilising,
I used a simple genetic algorithm (the space of possible \(\gamma\)
values was too large to search
exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
random \(M\), the
genetic algorithm initialised 1000 different sets of
\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis of
each set of \(\gamma\) values was performed, and the 20 sets with the
lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
subsequent mutation and crossover between the resulting new population
of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
\(M\) was found, 20 generations occurred, or a converge criteria of
minimum fitness increase between generations was satisfied. Across
\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
resulted in stable systems with probabilities that were orders of
magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary Information}).
Hence, managing the response rates of system components in a targetted
way can potentially facilitate the stabilisation of complex systems.

I have focused broadly on random complex systems, but it is also
worthwhile to consider more restricted interactions such as those of
specific ecological
networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
include systems in which all interactions are negative (e.g.,
competitive networks), positive (e.g., mutualist networks), or \(i\) and
\(j\) pairs have opposing signs (e.g., predator-prey networks). In
general, competitive and mutualist networks tend to be destabilising,
and predator-prey network tend to be
stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
When \(Var(\gamma)\) is applied to each, the proportion of stable
competitive and predator-prey networks increases, but the proportion of
stable mutualist networks does not (see 
\hyperlink{SIstart}{Supplementary Information}).
Additionally, when each component of \(M\) is interpreted as a unique
species and given a random intrinsic growth
rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
feasibility is not increased by \(Var(\gamma)\), suggesting that
variation in species generation time might be unlikely to drive
stability in purely multi-species networks (see \hyperlink{SIstart}{Supplementary
Information}).

My results show that complex systems are more likely to be stable when
the response rates of system components vary. These results are broadly
applicable to complex biological and social networks.

\textbf{Acknowledgements}

I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
Conversations with L. Bussi\`ere and N. Bunnefeld improved the quality of
this work.

\textbf{References}

\hypertarget{refs}{}
\hypertarget{ref-May1972}{}
1. May, R. M. Will a large complex system be stable? \emph{Nature}
\textbf{238,} 413--414 (1972).

\hypertarget{ref-Allesina2012}{}
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\hypertarget{ref-Gray2008}{}
3. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
random brain networks with a distribution of connection strengths.
\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).

\hypertarget{ref-Gray2009}{}
4. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
with excitatory and inhibitory connections. \emph{Neurocomputing}
\textbf{72,} 1849--1858 (2009).

\hypertarget{ref-Rosenfeld2009}{}
5. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
high-dimensional biochemical networks. \emph{Gene Regulation and Systems
Biology} \textbf{3,} 1--10 (2009).

\hypertarget{ref-MacArthur2010}{}
6. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
and criticality of adaptive regulatory networks. \emph{Physics Review
Letters} \textbf{104,} 168701 (2010).

\hypertarget{ref-Haldane2011}{}
7. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
\emph{Nature} \textbf{469,} 351--355 (2011).

\hypertarget{ref-Suweis2014}{}
8. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
networks. \emph{PLoS ONE} \textbf{9,} (2014).

\hypertarget{ref-Bardoscia2017}{}
9. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G. Pathways
towards instability in financial networks. \emph{Nature Communications}
\textbf{8,} 1--7 (2017).

\hypertarget{ref-Mougi2012}{}
10. Mougi, A. \& Kondoh, M. Diversity of interaction types and
ecological community stability. \emph{Science} \textbf{337,} 349--351
(2012).

\hypertarget{ref-Allesina2015a}{}
11. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\hypertarget{ref-Gao2016}{}
12. Gao, J., Barzel, B. \& Barabási, A. L. Universal resilience patterns
in complex networks. \emph{Nature} \textbf{530,} 307--312 (2016).

\hypertarget{ref-Grilli2017}{}
13. Grilli, J. \emph{et al.} Feasibility and coexistence of large
ecological communities. \emph{Nature Communications} \textbf{8,} (2017).

\hypertarget{ref-Patel2018}{}
14. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
effects of eco-evolutionary feedbacks on community stability.
\emph{American Naturalist} \textbf{191,} 1--29 (2018).

\hypertarget{ref-Allesina2015}{}
15. Allesina, S. \emph{et al.} Predicting the stability of large
structured food webs. \emph{Nature Communications} \textbf{6,} 7842
(2015).

\hypertarget{ref-May2008}{}
16. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).

\hypertarget{ref-Tao2010}{}
17. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
(2010).

\hypertarget{ref-Hamblin2013}{}
18. Hamblin, S. On the practical usage of genetic algorithms in ecology
and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
184--194 (2013).

\hypertarget{ref-Allesina2011}{}
19. Allesina, S. \& Levine, J. M. A competitive network theory of
species diversity. \emph{Proceedings of the National Academy of Sciences
of the United States of America} \textbf{108,} 5638--5642 (2011).

\hypertarget{ref-Dougoud2018}{}
20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
The feasibility of equilibria in large ecosystems: a primary but
neglected concept in the complexity-stability debate. \emph{PLOS
Computational Biology} \textbf{14,} e1005988 (2018).

\clearpage

\begin{center}
\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
\end{center}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{This supplementary information supports the manuscript
``Component response rate variation drives stability in large complex
systems'' with all of the code required to recreate the analysis in the
main text, and with additional analyses to support its conclusions. All
text, code, and data underlying this manuscript are publicly available
on \href{https://github.com/bradduthie/random_matrix_stability}{GitHub}
as part of the RandomMatrixStability package.}

The
\href{https://github.com/bradduthie/random_matrix_stability}{RandomMatrixStability
package} includes all functions and tools for recreating the text, this
supplemental information, and running all code; additional documentation
is also provided for functions as part of the package. The
RandomMatrixStability package package is available on
\href{https://github.com/bradduthie/random_matrix_stability}{GitHub}; to
download it, the
\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
library} is needed.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
\KeywordTok{library}\NormalTok{(devtools);}
\end{Highlighting}
\end{Shaded}

The code below installs the RandomMatrixStability package using
devtools.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/random_matrix_stability"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

While downloading this package is recommended, all relevant code is also
reproduced below with explanation, so it is possible to recreate all
analyses using only this Supplementary Information.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\section{Supplementary Information table of
contents}\label{supplemental-information-table-of-contents}

\begin{itemize}
\tightlist
\item
  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
\item
  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
\item
  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
\item
  \protect\hyperlink{ecological}{Stability of ecological networks}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{competition}{Competitor networks}
  \item
    \protect\hyperlink{mutualism}{Mutualist networks}
  \item
    \protect\hyperlink{pred-prey}{Predator-prey networks}
  \end{itemize}
\item
  \protect\hyperlink{connectance}{Different connectance (C) values}

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{connect3}{C = 0.3}
  \item
    \protect\hyperlink{connect5}{C = 0.5}
  \item
    \protect\hyperlink{connect7}{C = 0.7}
  \item
    \protect\hyperlink{connect9}{C = 0.9}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
\item
  \protect\hyperlink{ga}{Genetic algorithm}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\end{itemize}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
1}\label{Fig1}}

The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
text is available on
\href{https://github.com/bradduthie/random_matrix_stability/tree/master/sim_results/bi_gamma}{GitHub},
and was produced by running the following function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
            \NormalTok{break;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(iters);}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
is found that is not stable when all component response rates are set to
\(\gamma = 1\), but is stable when half of component response rates are
\(1.95\) and half are \(0.05\). The function is used to illustrate the
concept of how fast versus slow component responses can cause a system
to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
but terminated once an acceptable \texttt{A0} and \texttt{A1} were
found. The code below plots the eigenvalue distributions of \texttt{A0}
and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
plot itself can be recreated with the function and code below.

\includegraphics{unnamed-chunk-4-1.pdf}

To find out how frequently \(M\) was stable given that all
\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
was created.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(ress);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above functions produced the \texttt{bi\_pr\_st} data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
\end{Highlighting}
\end{Shaded}

The function \texttt{stab\_bgamma} was run for
\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
returned. Each row of \texttt{ress} represents a single \(M\) given
\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
Values of 0 indicate that \(M\) was found to be unstable (at least one
real component of its eigenvalues greater than or equal to zero),
whereas values of 1 indicate that \(M\) was found to be stable (all real
components of eigenvalues are negative). The frequencies of stable \(M\)
were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
reported in the main text and legend of Fig. 1 (raw data are
\href{https://github.com/bradduthie/random_matrix_stability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
on GitHub}).

\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
2}\label{Fig2}}

Figure 2 of the main text shows eigenvalue distributions in a system
where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
reproduced using the code below for when \(\gamma = 1\) (panel a) and
\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
reproduces the figure.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
\NormalTok{\}}
\KeywordTok{plot_Fig_2}\NormalTok{();}
\end{Highlighting}
\end{Shaded}

\includegraphics{unnamed-chunk-8-1.pdf}

\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
\(S\)}{Stability across increasing S}}\label{IncrS}}

Figure 3 of the main text reports the number of stable random complex
systems found over 1 million iterations. The data used to make this
figure are read into R below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_1/random_all.csv"}\NormalTok{);}
\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[,-}\DecValTok{1}\NormalTok{]; }\CommentTok{# Extra row-indicating column removed}
\end{Highlighting}
\end{Shaded}

The table below shows the results for all simulations of random \(M\)
matrices at \(\sigma = 0.4\) and \(C = 1\) given a range of
\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
Each row summarises data for a given \(S\) over 1 million randomly
simulated \(M\) (\texttt{A0} and \texttt{A1}). The column
\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
unstable, and the column \texttt{A0\_stable} shows the number of
\texttt{A0} matrices that are stable (these two columns sum to 1
million). Similarly, the column \texttt{A1\_unstable} shows the number
of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
the number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 293 & 999707 & 293 & 999707 & 0 & 0\tabularnewline
3 & 3602 & 996398 & 3609 & 996391 & 0 & 7\tabularnewline
4 & 14937 & 985063 & 15008 & 984992 & 0 & 71\tabularnewline
5 & 39289 & 960711 & 39783 & 960217 & 36 & 530\tabularnewline
6 & 78845 & 921155 & 80207 & 919793 & 389 & 1751\tabularnewline
7 & 133764 & 866236 & 136904 & 863096 & 1679 & 4819\tabularnewline
8 & 204112 & 795888 & 208241 & 791759 & 5391 & 9520\tabularnewline
9 & 288041 & 711959 & 291775 & 708225 & 12619 & 16353\tabularnewline
10 & 384024 & 615976 & 384931 & 615069 & 23153 & 24060\tabularnewline
11 & 485975 & 514025 & 481019 & 518981 & 35681 & 30725\tabularnewline
12 & 590453 & 409547 & 577439 & 422561 & 48302 & 35288\tabularnewline
13 & 689643 & 310357 & 669440 & 330560 & 57194 & 36991\tabularnewline
14 & 777496 & 222504 & 751433 & 248567 & 60959 & 34896\tabularnewline
15 & 850159 & 149841 & 821613 & 178387 & 58567 & 30021\tabularnewline
16 & 905057 & 94943 & 877481 & 122519 & 51255 & 23679\tabularnewline
17 & 943192 & 56808 & 919536 & 80464 & 40854 & 17198\tabularnewline
18 & 969018 & 30982 & 949944 & 50056 & 30102 & 11028\tabularnewline
19 & 984301 & 15699 & 970703 & 29297 & 20065 & 6467\tabularnewline
20 & 992601 & 7399 & 983507 & 16493 & 12587 & 3493\tabularnewline
21 & 996765 & 3235 & 991532 & 8468 & 7030 & 1797\tabularnewline
22 & 998693 & 1307 & 995567 & 4433 & 3884 & 758\tabularnewline
23 & 999503 & 497 & 997941 & 2059 & 1883 & 321\tabularnewline
24 & 999861 & 139 & 999059 & 941 & 899 & 97\tabularnewline
25 & 999964 & 36 & 999617 & 383 & 380 & 33\tabularnewline
26 & 999993 & 7 & 999878 & 122 & 121 & 6\tabularnewline
27 & 999995 & 5 & 999946 & 54 & 53 & 4\tabularnewline
28 & 1000000 & 0 & 999975 & 25 & 25 & 0\tabularnewline
29 & 1000000 & 0 & 999997 & 3 & 3 & 0\tabularnewline
30 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
31 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
matrices found is greater than 1 (compare column 5 to column 3), and
this ratio increases with increasing \(S\) (column 1). Hence, more
randomly created complex systems (\(M\)) are generated given variation
in \(\gamma\) than when \(\gamma = 1\). The results underlying this
table were produced with the \texttt{rand\_gen\_var} function below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(i);}
    \NormalTok{\}}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The above function calls the two functions
\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
are provided below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\}}

\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \CommentTok{# Stable and unstable}
        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \CommentTok{# Stabilised and destabilised}
        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \CommentTok{# Feasible and infeasible}
        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \CommentTok{# Feased and defeased}
        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
    \NormalTok{\}}
    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Note that feasibility results were ommited for the table above, but are
\protect\hyperlink{Feasibility}{reported below}.

\hypertarget{ecological}{\section{Stability of ecological
networks}\label{ecological}}

While the foundational work of
May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
to complex networks, much attention has been given specifically to
ecological networks of interacting species. In these networks, the
matrix \(M\) is interpreted as a community matrix and each row and
column is interpreted as a single species. The effect that the density
of any species \(i\) has on the population dynamics of species \(j\) is
found in \(M_{ij}\), meaning that \(M\) holds the effects of pair-wise
interactions between \(S\)
species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Tang2014b}{4}}.
While May's original
work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
randomly assembled communities, recent work has specifically looked at
more restricted ecological communities including competitive networks
(all off-diagonal elements of \(M\) are negative), mutualist networks
(all off-diagonal elements of \(M\) are positive), and predator-prey
networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
is negative and \(j\) on \(i\) is positive, or vice
versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Allesina2011}{5}}.
In general, competitor and mutualist networks tend to be unstable, while
predator-prey networks tend to be highly stabilising.

I investigated competitor, mutualist, and predator-prey networks
following Allesina et
al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
these networks, I first generated a random matrix \(M\), then changed
the elements of \(M\) accordingly. If \(M\) was a competitive network,
then the sign of any positive off-diagonal elements was reversed to be
negative. If \(M\) was a mutualist network, then the sign of any
positive off-diagonal elements was reversed to be positive. And if \(M\)
was a predator-prey network, then all \(i\) and \(j\) pairs of elements
were checked; any pairs of the same sign were changed so that one was
negative and the other was positive. The \texttt{species\_interaction}
function used to do this is below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
\end{Highlighting}
\end{Shaded}

This function was applied to all created matrices \(M\), then the number
of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
as it was} in the main text for random matrices for values of \(S\) from
2 to 50 (100 in the case of the relatively more stable predator-prey
interactions), except that only 100000 random \(M\) were generated
instead of 1 million. This produced the data set below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

The following tables for restricted ecological communities can therefore
be compared with the random \(M\) \protect\hyperlink{IncrS}{results
above} (but note that counts from systems with comparable probabilities
of stability will be an order of magnitude lower in the tables below due
to the smaller number of \(M\) matrices generated). As with the
\protect\hyperlink{IncrS}{results above}, in the tables below,
\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
refers to matrices after \(Var(\gamma)\) is added. The column
\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
unstable, and the column \texttt{A0\_stable} shows the number of
\texttt{A0} matrices that are stable (these two columns sum to 100000).
Similarly, the column \texttt{A1\_unstable} shows the number of
\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

\textbf{Competition}

Results for competitor interaction networks are shown below

\begin{longtable}[]{@{}lllllll@{}}
\toprule
X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
1 & 2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
2 & 3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
3 & 4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
4 & 5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
5 & 6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
6 & 7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
7 & 8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
8 & 9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
9 & 10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
10 & 11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
11 & 12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
12 & 13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
13 & 14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
14 & 15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
15 & 16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
16 & 17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
17 & 18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
18 & 19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
19 & 20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
20 & 21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
21 & 22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
22 & 23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
23 & 24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
24 & 25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
25 & 26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
26 & 27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
27 & 28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
28 & 29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
29 & 30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
30 & 31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
31 & 32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
32 & 33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
33 & 34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Mutualism}

Results for mutualist interaction networks are shown below

\begin{longtable}[]{@{}lllllll@{}}
\toprule
X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
1 & 2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
2 & 3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
3 & 4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
4 & 5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
5 & 6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
6 & 7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
7 & 8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
8 & 9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
9 & 10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
10 & 11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
11 & 12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Predator-prey}

Results for predator-prey interaction networks are shown below

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised\tabularnewline
\midrule
\endhead
1 & 2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
2 & 3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
3 & 4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
4 & 5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
5 & 6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
6 & 7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
7 & 8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
8 & 9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
9 & 10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
10 & 11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
11 & 12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
12 & 13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
13 & 14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
14 & 15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
15 & 16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
16 & 17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
17 & 18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
18 & 19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
19 & 20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
20 & 21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
21 & 22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
22 & 23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
23 & 24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
24 & 25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
25 & 26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
26 & 27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
27 & 28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
28 & 29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
29 & 30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
30 & 31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
31 & 32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
32 & 33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
33 & 34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
34 & 35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
35 & 36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
36 & 37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
37 & 38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
38 & 39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
39 & 40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
40 & 41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
41 & 42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
42 & 43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
43 & 44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
44 & 45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
45 & 46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
46 & 47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
47 & 48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
48 & 49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
49 & 50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
50 & 51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
51 & 52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
52 & 53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
53 & 54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
54 & 55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
55 & 56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
56 & 57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
57 & 58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
58 & 59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
59 & 60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
60 & 61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
61 & 62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
62 & 63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
63 & 64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
64 & 65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
65 & 66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
66 & 67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
67 & 68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
68 & 69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
69 & 70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
70 & 71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
71 & 72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
72 & 73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
73 & 74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
74 & 75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
75 & 76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
76 & 77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
77 & 78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
78 & 79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
79 & 80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
80 & 81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
81 & 82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
82 & 83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
83 & 84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
84 & 85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
85 & 86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
86 & 87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
87 & 88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
88 & 89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
89 & 90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
90 & 91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
91 & 92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
92 & 93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
93 & 94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
94 & 95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
95 & 96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
96 & 97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
97 & 98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
98 & 99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
99 & 100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
\bottomrule
\end{longtable}

Overall, as
expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
predator-prey communities are relatively stable while mutualist
communties are highly unstable. But interestingly, while \(Var(\gamma)\)
stabilises predator-prey and competitor communities, it does not
stabilise mutualist communities. This is unsurprising because purely
mutualist communities are characterised by a very
positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
alone will shift all real parts of eigenvalues to negative values.

\hypertarget{connectance}{\section{Different inter-connectivity (C)
values}\label{connectance}}

In the main text, for simplicity, I assumed inter-connectivity values of
\(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
potentially nonzero and sampled from a normal distribution
\(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here I present
four tables showing the number of stable communities given
\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
component response time (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
higher number of stable communities than when \(\gamma\) did not vary
(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
million \(M\) were simulated. As with the results on
\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
that are unstable, and the column \texttt{A0\_stable} shows the number
of \texttt{A0} matrices that are stable (these two columns sum to
100000). Similarly, the column \texttt{A1\_unstable} shows the number of
\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
number that are stable. The columns \texttt{A1\_stabilised} and
\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
stabilised or destabilised, respectively, by \(Var(\gamma)\).

All data reported below for various values of \(C\) are accessible using
the below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{C3dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt3.csv"}\NormalTok{);}
\NormalTok{C5dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt5.csv"}\NormalTok{);}
\NormalTok{C7dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt7.csv"}\NormalTok{);}
\NormalTok{C9dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt9.csv"}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

These objects \texttt{C3dat}, \texttt{C5dat}, \texttt{C7dat}, and
\texttt{C9dat} include the results for \(C = 0.3\), \(C = 0.5\),
\(C = 0.7\), and \(C = 0.9\), respectively.

\textbf{Connectance \(C = 0.3\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.5\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.7\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\textbf{Connectance \(C = 0.9\)}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{gam_dist}{\section{\texorpdfstring{Different distributions
of
\(\gamma\)}{Different distributions of \textbackslash{}gamma}}\label{gam_dist}}

In the main text, I considered a uniform distribution of component
response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
and stable \(M\) matrices are reported in \protect\hyperlink{IncrS}{a
table above} across different values of \(S\). Here I show complementary
results for three different distributions including an exponential,
beta, and gamma distribution of \(\gamma\) values. The shape of these
distributions is shown in the figure below.

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\textbf{Distributions of component response rate
(\(\boldsymbol{\gamma}\)) values in complex systems.} The stabilities of
simulated complex systems with these \(\gamma\) distributions are
compared to otherwise identical complex systems with a fixed component
response rate of \(\gamma = 1\) across different system sizes (\(S\);
i.e., component numbers) given a unit \(\gamma\) standard deviation
(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
\(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
panel shows 1 million randomly generated \(\gamma\) values.

\includegraphics{unnamed-chunk-23-1.pdf}

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

The same 100000 \(M\) matrices were used to investigate stability when
applying each of these different distributions of \(\gamma\) values. The
table below shows the number of \(M\) that were unstable (\_unst) and
stable (\_stbl) for the exponential (Exp), beta, and gamma
distributions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
\KeywordTok{kable}\NormalTok{(fourdists);}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lllllll@{}}
\toprule
S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
gamma\_stbl\tabularnewline
\midrule
\endhead
2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
\ldots{}\tabularnewline
50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
\bottomrule
\end{longtable}

In comparison to the uniform distribution (a), proportionally fewer
random systems are found with the exponential distribution (b), while
more are found with the beta (c) and gamma (d) distributions.

\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}

Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
the proportion of stable complex systems, the search space of all
possible \(\gamma\) vectors would be evaluated for each unique \(M\).
This is technically impossible because \(\gamma_{i}\) can take any real
value between 0-2, but even rounding \(\gamma\) to reasonable values
would result in a search space too large to practically explore. Under
these conditions, genetic algorithms are highly useful tools for finding
practical solutions by mimicking the process of biological
evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{6}}. In
this case, the practical solution is finding vectors of \(\gamma\) that
decrease the most positive real eigenvalue of \(M\). The genetic
algorithm below achieves this by initialising a large population of 1000
different potential \(\gamma\) vectors and allowing this population to
evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
values between vectors), selection, and reproduction until either a
\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
``giving up'' critiera is met (in the below, this ``giving up''"
criteria is met when 20 generations pass, or if the fitness increase
from one generation to the next is below a certain criteria). The
genetic algorithm relies on five functions. The first outer function
\texttt{Evo\_rand\_gen\_var} runs all of the simulations
(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
\texttt{iters} refers to the number of \(M\) to try for each \(S\)).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
            \NormalTok{\}}
            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{\}}
        \KeywordTok{print}\NormalTok{(i);}
    \NormalTok{\}}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
\texttt{summarise\_randmat}. The first simply allows for non-random
interactions between components (e.g., modelling
\protect\hyperlink{ecological}{ecological interactions} of random,
competition, mutualism, or predator-prey).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
    \NormalTok{\}}
    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
                \NormalTok{\}}
            \NormalTok{\}}
        \NormalTok{\}}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(mat);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
algorithm, searching for \(\gamma\) vectors that are stabilising.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
        \NormalTok{\}}
        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{\}}
    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(find_st);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

The while loop in \texttt{rand\_mat\_ga} continues until either
\texttt{iter} generations have occured, a solution \(\gamma\) vector is
found that results in all \(\Re(\lambda) < 0\), or some criteria of
minimum fitness increase is observed (by default,
\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
values are mutated, crossover occurs between \(\gamma\) vectors, and
selection occurs in each generation such that the 20 \(\gamma\) vectors
that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
zero are multiplied by \(-1\), and any values that mutate above 2 are
set to 2. Note also that if a solution is found, then one such
\(\gamma\) vector causing stability is printed to a file.

Crossover occurs in the \texttt{crossover} function below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(inds);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
\texttt{summarise\_randmat} formats the data into a table.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{summarise_randmat_ga <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{10}\NormalTok{);}
    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
        \NormalTok{unstables <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{stables   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{non_feas  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{feasibl   <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{unfeased  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
        \NormalTok{feased    <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{foundd    <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{3}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstables);}
        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stables);}
        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(non_feas);}
        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feasibl);}
        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unfeased);}
        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feased);}
        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(foundd);}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(all_res);}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Some stability results from this table are shown below. Each histogram
shows a different distribution of \(\gamma\) that was found to be
stabilising.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
A1\_stabilised & A1\_destabilised\tabularnewline
\midrule
\endhead
2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

The distributions of nine \(\gamma\) vectors from the highest \(S\)
values are shown below. Recall that 1 million random matrices were
generated for the less computationally intense task of
\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
compare stability in column 5 above with column 3 above. \textbf{This
comparison shows the high number of stable \(M\) that can be produced
through a targetted search of \(\gamma\) values, and suggests that many
otherwise unstable systems could potentially be stabilised by an
informed manipulation of their component response times. Such a
possibility might conceivably reduce the dimensionality of problems
involving stability in social-ecological or economic systems.}

Distributions of \(\gamma\) values in vectors for the highest values of
\(S\) are shown below.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
\end{Highlighting}
\end{Shaded}

\includegraphics{unnamed-chunk-31-1.pdf}

The distribution of \(\gamma\) values found by the genetic algorithm is
uniform. A uniform distribution was used to initialise \(\gamma\)
values, so there is therefore no evidence that a particular distribution
of \(\gamma\) is likely to be found to stabilise a matrix \(M\).

\hypertarget{Feasibility}{\section{Feasibility of complex
systems}\label{Feasibility}}

For complex systems in which individual system components (\(S\))
represent the density of some tangible quantity, it is important to
consider the feasibility of the system. Feasibile equilibria assume that
the values of all system components are positive at
equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{7}--\protect\hyperlink{ref-Song2018}{9}}.
This is of particular interest for ecological communities because
population density cannot take negative values, meaning that ecological
systems need to be feasible for stability to be biologically
realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
Consequently, the use of random matrices and traditional stability
critiera for making inferences in theoretical analyses of species
networks has recently been
criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
While the key results in the main text are intended to be general to all
complex systems, and not restricted to species networks, I have also
performed a feasibility analysis on all matrices \(M\). This analysis
reveals that feasibility is not affected by \(Var(\gamma)\), meaning
that for pure interacting species networks, variation in component
response time (i.e., species generation time) does not affect stability
at biologically realistic species densities. Nevertheless, ecological
interactions do not exist in isolation in empirical systems, but instead
interact with
evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{10}},
abiotic, or social-economic systems. The relevance of \(\gamma\) for
complex system stability presented in the main text should therefore not
be ignored in the broader context of ecological communities.

Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}
define the following feasibility criteria for ecological systems
characterised by \(S\) interacting species with varying densities.

\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]

In the above, \(x^{*}\) is the vector of species abundances at
equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
off-diagonal elements), and the value \(\theta\) is strength of
intraspecific competition (diagonal elements). As I have done elsewhere,
diagonal values are set to \(-1\), so \(\theta = -1\). The variable
\(C\) is the inter-connectivity (i.e., `connectance') of the community,
which was set to \(C = 1\) throughout the manuscript and supplemental
information, except \protect\hyperlink{connectance}{where otherwise
noted}. The variable \(\delta\) is a normalisation parameter that
modulates the strength of interactions (\(\sigma\) in the main text),
which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
\(A\), which has a diagonal of all zeros and an off-diagonal affecting
species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} stability
criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}} by
\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
text and Supplementary Information). The above criteria is therefore
reduced to the below; note that the parenthetical in both equations
produces an \(M\) matrix as used throughout the main text and
supplemental information,

\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]

To check the feasibility criteria, I therefore inverted
\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
the resulting matrix by the vector of population growth rates \(r\).
Feasibility is satisfied if all of the elements of the resulting vector
are positive.

The population growth rate for an individual species \(i\) is sampled
from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
shown in the \texttt{rand\_gen\_var} function in
\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
is assumed to be a species with a growth rate of \(r_{i}\). Note that
negative intrinsic growth rates are not unrealistic, and will occur in
obligate mutualists in the absence of a partner.

When feasibility was evaluated with and without variation in \(\gamma\),
there was no increase in stability for \(M\) where \(\gamma\) varied as
compared to where \(\gamma = 1\). Results below illustrate this result,
which was general to all other simulations performed.

\begin{longtable}[]{@{}rrrrrrr@{}}
\toprule
S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
\midrule
\endhead
2 & 749978 & 250022 & 749942 & 250058 & 35552 & 35516\tabularnewline
3 & 874519 & 125481 & 874296 & 125704 & 36803 & 36580\tabularnewline
4 & 937192 & 62808 & 937215 & 62785 & 26440 & 26463\tabularnewline
5 & 968776 & 31224 & 968639 & 31361 & 16319 & 16182\tabularnewline
6 & 984313 & 15687 & 984463 & 15537 & 9006 & 9156\tabularnewline
7 & 992149 & 7851 & 992161 & 7839 & 4991 & 5003\tabularnewline
8 & 996124 & 3876 & 996103 & 3897 & 2644 & 2623\tabularnewline
9 & 998014 & 1986 & 998027 & 1973 & 1361 & 1374\tabularnewline
10 & 999031 & 969 & 999040 & 960 & 698 & 707\tabularnewline
11 & 999546 & 454 & 999514 & 486 & 377 & 345\tabularnewline
12 & 999764 & 236 & 999792 & 208 & 160 & 188\tabularnewline
13 & 999883 & 117 & 999865 & 135 & 105 & 87\tabularnewline
14 & 999938 & 62 & 999945 & 55 & 40 & 47\tabularnewline
15 & 999971 & 29 & 999964 & 36 & 31 & 24\tabularnewline
16 & 999988 & 12 & 999991 & 9 & 8 & 11\tabularnewline
17 & 999996 & 4 & 999991 & 9 & 8 & 3\tabularnewline
18 & 999997 & 3 & 999999 & 1 & 1 & 3\tabularnewline
19 & 999998 & 2 & 999997 & 3 & 3 & 2\tabularnewline
20 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
21 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
22 & 999999 & 1 & 1000000 & 0 & 0 & 1\tabularnewline
23 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
24 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
25 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
26 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
27 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
28 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
29 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
30 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
31 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
in pure species interaction networks.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-May1972}{}
1. May, R. M. Will a large complex system be stable? \emph{Nature}
\textbf{238,} 413--414 (1972).

\hypertarget{ref-Allesina2012}{}
2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
\emph{Nature} \textbf{483,} 205--208 (2012).

\hypertarget{ref-Allesina2015a}{}
3. Allesina, S. \& Tang, S. The stability--complexity relationship at
age 40: a random matrix perspective. \emph{Population Ecology} 63--75
(2015).
doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}

\hypertarget{ref-Tang2014b}{}
4. Tang, S. \& Allesina, S. Reactivity and stability of large
ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
(2014).

\hypertarget{ref-Allesina2011}{}
5. Allesina, S. \& Levine, J. M. A competitive network theory of species
diversity. \emph{Proceedings of the National Academy of Sciences of the
United States of America} \textbf{108,} 5638--5642 (2011).

\hypertarget{ref-Hamblin2013}{}
6. Hamblin, S. On the practical usage of genetic algorithms in ecology
and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
184--194 (2013).

\hypertarget{ref-Grilli2017}{}
7. Grilli, J. \emph{et al.} Feasibility and coexistence of large
ecological communities. \emph{Nature Communications} \textbf{8,} (2017).

\hypertarget{ref-Dougoud2018}{}
8. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
The feasibility of equilibria in large ecosystems: a primary but
neglected concept in the complexity-stability debate. \emph{PLOS
Computational Biology} \textbf{14,} e1005988 (2018).

\hypertarget{ref-Song2018}{}
9. Song, C. \& Saavedra, S. Will a small randomly assembled community be
feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).

\hypertarget{ref-Patel2018}{}
10. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
effects of eco-evolutionary feedbacks on community stability.
\emph{American Naturalist} \textbf{191,} 1--29 (2018).

\end{document}

### 2018-06-03 notebook/arxiv/ms.tex
New semi-major edits to the manuscript

@@
 perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of interacting components; \(S\)), its
-inter-connectivity (i.e., the probability that one component will affect
+system (i.e., the number of potentially interacting components; \(S\)), its 
+connectance (i.e., the probability that one component will interact with 
 another; \(C\)), and the variance of interaction strengths
@@
 banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
-element of such a matrix \(M_{ij}\) defines how component \(j\) affects
+element of such a matrix, \(M_{ij}\), defines how component \(j\) affects
 component \(i\) in the system at a point of
@@
 \(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
-centred at \(\gamma = -0.05\). In a, the maximum real eigenvalaue
+centred at \(\gamma = -0.05\). In a, the maximum real eigenvalue
 \(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
@@
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32 (see \hyperlink{SIstart}{Supplementary Information}
-for different values of \(\sigma\) and \(C\)). One million \(M\) were simulated 
+\(S\) ranging from 2-32. One million \(M\) were simulated 
 for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\) (note that under these
-conditions, \(\sigma\sqrt{SC} = 1\) given \(S = 25\) when
-\(\gamma = 1\)). I found that the number of stable random systems was
+versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of stable 
+random systems was
 consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
@@
 For the highest values of \(S\), nearly all systems that were stable
-given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\),
-and the maximum observed \(S\) for which a system was stable was \(31\)
-given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see
+given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\) (see
 \hyperlink{SIstart}{Supplementary Information} for full results). This 
@@
 
-Some caution is needed when interpreting these results. First, I
-emphasise that \(Var(\gamma)\) is not stabilising per se; that is,
+It is important to point out 
+that \(Var(\gamma)\) is not stabilising per se; that is,
 adding variation in \(\gamma\) to a particular system \(M\) does not
@@
 further investigate the potential of \(Var(\gamma)\) to be stabilising,
-I used a simple genetic algorithm (the space of possible \(\gamma\)
+I used a genetic algorithm (the space of possible \(\gamma\)
 values was too large to search
@@
 genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis of
-each set of \(\gamma\) values was performed, and the 20 sets with the
+\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
+performed on \(M\) using 
+each set of \(\gamma\) values, and the 20 sets with the
 lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
@@
 of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
-\(M\) was found, 20 generations occurred, or a converge criteria of
+\(M\) was found, 20 generations occurred, or a convergence criteria of
 minimum fitness increase between generations was satisfied. Across
@@
 resulted in stable systems with probabilities that were orders of
-magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary Information}).
+magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
+Information}), meaning that stability could often be achieved by 
+manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
 Hence, managing the response rates of system components in a targetted
-way can potentially facilitate the stabilisation of complex systems.
+way can potentially facilitate the stabilisation of complex systems 
+through a reduction in dimensionality.
 
@@
 networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (e.g.,
-competitive networks), positive (e.g., mutualist networks), or \(i\) and
-\(j\) pairs have opposing signs (e.g., predator-prey networks). In
+include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or \(i\) and
+\(j\) pairs have opposing signs (predator-prey networks). In
 general, competitive and mutualist networks tend to be destabilising,
@@
 
-\hypertarget{connectance}{\section{Different inter-connectivity (C)
+\hypertarget{connectance}{\section{Different connectance (C)
 values}\label{connectance}}
 
-In the main text, for simplicity, I assumed inter-connectivity values of
+In the main text, for simplicity, I assumed connectance values of
 \(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
@@
 diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the inter-connectivity (i.e., `connectance') of the community,
+\(C\) is the connectance of the community,
 which was set to \(C = 1\) throughout the manuscript and supplemental

### 2018-06-03 notebook/ms.Rmd
New semi-major edits to the manuscript

@@
 output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-  html_document: default
   word_document:
@@
     reference_docx: docx_template.docx
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+  html_document: default
 bibliography: references.bib
@@
 <!---
-Current word count: 1683
+Current word count: 1672
 --->
@@
 In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations.
-Complexity in this case is defined by the size of the system (i.e., the number of interacting components; $S$), its inter-connectivity (i.e., the probability that one component will affect another; $C$), and the variance of interaction strengths ($\sigma^{2}$)[@Allesina2012].
+Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$)[@Allesina2012].
 May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ has profound consequences across multiple disciplines, raising the question of how complex systems in, e.g., ecology[@Mougi2012; @Allesina2012; @Allesina2015; @Grilli2017] or banking [@May2008; @Haldane2011; @Bardoscia2017] are predicted to persist or change.
@@
 Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., species[@Allesina2012] or banks[@Haldane2011]).
-One element of such a matrix $M_{ij}$ defines how component $j$ affects component $i$ in the system at a point of equilibrium[@Allesina2012].
+One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium[@Allesina2012].
 Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
@@
 **b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
-In a, the maximum real eigenvalaue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
+In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
 In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` was stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
@@
 
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32 (see Supplementary Information for different values of $\sigma$ and $C$).
-One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$ (note that under these conditions, $\sigma\sqrt{SC} = 1$ given $S = 25$ when $\gamma = 1$).
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
+One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
 I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$, and the maximum observed $S$ for which a system was stable was $31$ given $Var(\gamma)$ versus $27$ given $\gamma = 1$ (see Supplementary Information for full results).
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
 This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
 
-Some caution is needed when interpreting these results.
-First, I emphasise that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
+It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
 Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a simple genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
 For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis of each set of $\gamma$ values was performed, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a converge criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supporting Information).
-Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems.
+Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supporting Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
 
 I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
-These include systems in which all interactions are negative (e.g., competitive networks), positive (e.g., mutualist networks), or $i$ and $j$ pairs have opposing signs (e.g., predator-prey networks).
+These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
 In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].

### 2018-06-03 notebook/ms.tex
Add more notebook files; final draft edits
New semi-major edits to the manuscript

@@
 perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of interacting components; \(S\)), its
-inter-connectivity (i.e., the probability that one component will affect
-another; \(C\)), and the variance of interaction strengths
+system (i.e., the number of potentially interacting components; \(S\)),
+its connectance (i.e., the probability that one component will interact
+with another; \(C\)), and the variance of interaction strengths
 (\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
-element of such a matrix \(M_{ij}\) defines how component \(j\) affects
-component \(i\) in the system at a point of
+element of such a matrix, \(M_{ij}\), defines how component \(j\)
+affects component \(i\) in the system at a point of
 equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.549 and 0.557 for a and b, respectively.
+of \(\Re(\lambda) < 0\) are 0.549 and 0.547 for a and b, respectively.
 
@@
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32 (see Supplementary Information for different
-values of \(\sigma\) and \(C\)). One million \(M\) were simulated for
-each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\) (note that under these
-conditions, \(\sigma\sqrt{SC} = 1\) given \(S = 25\) when
-\(\gamma = 1\)). I found that the number of stable random systems was
-consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
-3), and that the difference between the probabilities of observing a
-stable system increased with an increase in \(S\); i.e., the potential
-for \(Var(\gamma)\) to drive stability increased with system complexity.
-For the highest values of \(S\), nearly all systems that were stable
-given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\),
-and the maximum observed \(S\) for which a system was stable was \(31\)
-given \(Var(\gamma)\) versus \(27\) given \(\gamma = 1\) (see
-Supplementary Information for full results). This suggests that the
-stability of large systems might be dependent upon variation in the
-response rate of their individual components, meaning that factors such
-as generation time (in ecological networks), transaction speed (in
-economic networks), or communication speed (in social networks) needs to
-be considered when investigating the stability of complex systems.
+\(S\) ranging from 2-32. One million \(M\) were simulated for each
+\(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
+versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of
+stable random systems was consistently higher given \(Var(\gamma)\) than
+when \(\gamma = 1\) (Fig. 3), and that the difference between the
+probabilities of observing a stable system increased with an increase in
+\(S\); i.e., the potential for \(Var(\gamma)\) to drive stability
+increased with system complexity. For the highest values of \(S\),
+nearly all systems that were stable given \(Var(\gamma)\) would not have
+been stable given \(\gamma = 1\) (see Supplementary Information for full
+results). This suggests that the stability of large systems might be
+dependent upon variation in the response rate of their individual
+components, meaning that factors such as generation time (in ecological
+networks), transaction speed (in economic networks), or communication
+speed (in social networks) needs to be considered when investigating the
+stability of complex systems.
 
-Some caution is needed when interpreting these results. First, I
-emphasise that \(Var(\gamma)\) is not stabilising per se; that is,
-adding variation in \(\gamma\) to a particular system \(M\) does not
-necessarily increase the probability that the system will be stable (see
-Supplementary Information). Rather, systems that are observed to be
-stable are more likely to vary in \(\gamma\), and for this
-\(Var(\gamma)\) to be critical to their stability. This is caused by the
-shift in the distribution of eigenvalues that occurs by introducing
-\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
-\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
-further investigate the potential of \(Var(\gamma)\) to be stabilising,
-I used a simple genetic algorithm (the space of possible \(\gamma\)
-values was too large to search
+It is important to point out that \(Var(\gamma)\) is not stabilising per
+se; that is, adding variation in \(\gamma\) to a particular system \(M\)
+does not necessarily increase the probability that the system will be
+stable (see Supplementary Information). Rather, systems that are
+observed to be stable are more likely to vary in \(\gamma\), and for
+this \(Var(\gamma)\) to be critical to their stability. This is caused
+by the shift in the distribution of eigenvalues that occurs by
+introducing \(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in
+all \(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\)
+values. To further investigate the potential of \(Var(\gamma)\) to be
+stabilising, I used a genetic algorithm (the space of possible
+\(\gamma\) values was too large to search
 exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
@@
 genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis of
-each set of \(\gamma\) values was performed, and the 20 sets with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
-subsequent mutation and crossover between the resulting new population
-of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
-\(M\) was found, 20 generations occurred, or a converge criteria of
-minimum fitness increase between generations was satisfied. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were orders of
-magnitude higher than when \(\gamma = 1\) (see Supporting Information).
-Hence, managing the response rates of system components in a targetted
-way can potentially facilitate the stabilisation of complex systems.
+\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis
+was performed on \(M\) using each set of \(\gamma\) values, and the 20
+sets with the lowest \(\max\left(\Re(\lambda)\right)\) each produced 50
+offspring with subsequent mutation and crossover between the resulting
+new population of 1000 \(\gamma\) sets. The genetic algorithm terminated
+if a stable \(M\) was found, 20 generations occurred, or a convergence
+criteria of minimum fitness increase between generations was satisfied.
+Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were
+found that resulted in stable systems with probabilities that were
+orders of magnitude higher than when \(\gamma = 1\) (see Supporting
+Information), meaning that stability could often be achieved by
+manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\)
+elements. Hence, managing the response rates of system components in a
+targetted way can potentially facilitate the stabilisation of complex
+systems through a reduction in dimensionality.
 
@@
 networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (e.g.,
-competitive networks), positive (e.g., mutualist networks), or \(i\) and
-\(j\) pairs have opposing signs (e.g., predator-prey networks). In
-general, competitive and mutualist networks tend to be destabilising,
-and predator-prey network tend to be
+include systems in which all interactions are negative (competitive
+networks), positive (mutualist networks), or \(i\) and \(j\) pairs have
+opposing signs (predator-prey networks). In general, competitive and
+mutualist networks tend to be destabilising, and predator-prey network
+tend to be
 stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.

### 2018-06-03 notebook/SI.Rmd
Fix package name and edit links

@@
 **This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with all of the code required to recreate the analysis in the main text, and with additional analyses to support its conclusions.
-All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/random_matrix_stability) as part of the RandomMatrixStability package.**
+All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability package.**
 
-The [RandomMatrixStability package](https://github.com/bradduthie/random_matrix_stability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for functions as part of the package.
-The RandomMatrixStability package package is available on [GitHub](https://github.com/bradduthie/random_matrix_stability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
+The [RandomMatrixStability package](https://github.com/bradduthie/RandomMatrixStability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for functions as part of the package.
+The RandomMatrixStability package package is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
 
@@
 ```{r, eval = FALSE}
-install_github("bradduthie/random_matrix_stability")
+install_github("bradduthie/RandomMatrixStability")
 ```
@@
 The sample $M$ used for the eigenvalue distributions in Fig.
-1 of the text is available on [GitHub](https://github.com/bradduthie/random_matrix_stability/tree/master/sim_results/bi_gamma), and was produced by running the following function.
+1 of the text is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma), and was produced by running the following function.
 
@@
 The frequencies of stable $M$ were `r sum(pr_st[,1])` given $\gamma = 1$ and `r sum(pr_st[,2])` given $\gamma = \{1.95, 0.05\}$, as reported in the main text and legend of Fig.
-1 (raw data are [available on GitHub](https://github.com/bradduthie/random_matrix_stability/blob/master/sim_results/bi_gamma/bi_pr_st.csv)).
+1 (raw data are [available on GitHub](https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv)).
 

### 2018-06-03 notebook/ms.Rmd
Fix package name and edit links

@@
     fig_caption: yes
-    keep_tex: yes
+    keep_tex: no
   html_document: default

### 2018-06-03 notebook/ms.tex
Fix package name and edit links

@@
 
+\usepackage{lmodern}
+\usepackage{setspace}
+\setstretch{1}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \ifxetex
+    \usepackage{mathspec}
+  \else
+    \usepackage{fontspec}
+  \fi
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\usepackage[margin=1in]{geometry}
+\usepackage{hyperref}
+\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
+\hypersetup{unicode=true,
+            pdftitle={Component response rate variation drives stability in large complex systems},
+            pdfauthor={A. Bradley Duthie},
+            colorlinks=true,
+            linkcolor=blue,
+            citecolor=Blue,
+            urlcolor=Blue,
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\usepackage{longtable,booktabs}
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+%%% Use protect on footnotes to avoid problems with footnotes in titles
+\let\rmarkdownfootnote\footnote%
+\def\footnote{\protect\rmarkdownfootnote}
+
+%%% Change title format to be more compact
+\usepackage{titling}
+
 \setlength{\droptitle}{-2em}
@@
 \usepackage[utf8]{inputenc}
-\linenumbers
 \bibliographystyle{amnatnat}
@@
 \maketitle
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of potentially interacting components; \(S\)),
-its connectance (i.e., the probability that one component will interact
-with another; \(C\)), and the variance of interaction strengths
+system (i.e., the number of potentially interacting components; \(S\)), its 
+connectance (i.e., the probability that one component will interact with 
+another; \(C\)), and the variance of interaction strengths
 (\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
-element of such a matrix, \(M_{ij}\), defines how component \(j\)
-affects component \(i\) in the system at a point of
+element of such a matrix, \(M_{ij}\), defines how component \(j\) affects
+component \(i\) in the system at a point of
 equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{unnamed-chunk-4-1.pdf}
 
@@
 respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.549 and 0.547 for a and b, respectively.
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-7-1.pdf}
+of \(\Re(\lambda) < 0\) are 0.548 and 0.552 for a and b, respectively.
+
+\includegraphics{unnamed-chunk-7-1.pdf}
 
@@
 focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
-(see Supplementary Information for other distributions, which give
+(see \hyperlink{SIstart}{Supplementary Information} for other distributions, 
+which give
 similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{unnamed-chunk-9-1.pdf}
 
@@
 simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32. One million \(M\) were simulated for each
-\(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of
-stable random systems was consistently higher given \(Var(\gamma)\) than
-when \(\gamma = 1\) (Fig. 3), and that the difference between the
-probabilities of observing a stable system increased with an increase in
-\(S\); i.e., the potential for \(Var(\gamma)\) to drive stability
-increased with system complexity. For the highest values of \(S\),
-nearly all systems that were stable given \(Var(\gamma)\) would not have
-been stable given \(\gamma = 1\) (see Supplementary Information for full
-results). This suggests that the stability of large systems might be
-dependent upon variation in the response rate of their individual
-components, meaning that factors such as generation time (in ecological
-networks), transaction speed (in economic networks), or communication
-speed (in social networks) needs to be considered when investigating the
-stability of complex systems.
-
-It is important to point out that \(Var(\gamma)\) is not stabilising per
-se; that is, adding variation in \(\gamma\) to a particular system \(M\)
-does not necessarily increase the probability that the system will be
-stable (see Supplementary Information). Rather, systems that are
-observed to be stable are more likely to vary in \(\gamma\), and for
-this \(Var(\gamma)\) to be critical to their stability. This is caused
-by the shift in the distribution of eigenvalues that occurs by
-introducing \(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in
-all \(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\)
-values. To further investigate the potential of \(Var(\gamma)\) to be
-stabilising, I used a genetic algorithm (the space of possible
-\(\gamma\) values was too large to search
+\(S\) ranging from 2-32. One million \(M\) were simulated 
+for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
+versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of stable 
+random systems was
+consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
+3), and that the difference between the probabilities of observing a
+stable system increased with an increase in \(S\); i.e., the potential
+for \(Var(\gamma)\) to drive stability increased with system complexity.
+For the highest values of \(S\), nearly all systems that were stable
+given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\) (see
+\hyperlink{SIstart}{Supplementary Information} for full results). This 
+suggests that the
+stability of large systems might be dependent upon variation in the
+response rate of their individual components, meaning that factors such
+as generation time (in ecological networks), transaction speed (in
+economic networks), or communication speed (in social networks) needs to
+be considered when investigating the stability of complex systems.
+
+It is important to point out 
+that \(Var(\gamma)\) is not stabilising per se; that is,
+adding variation in \(\gamma\) to a particular system \(M\) does not
+necessarily increase the probability that the system will be stable (see
+\hyperlink{SIstart}{Supplementary Information}). Rather, systems that are 
+observed to be stable are more likely to vary in \(\gamma\), and for this
+\(Var(\gamma)\) to be critical to their stability. This is caused by the
+shift in the distribution of eigenvalues that occurs by introducing
+\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
+\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
+further investigate the potential of \(Var(\gamma)\) to be stabilising,
+I used a genetic algorithm (the space of possible \(\gamma\)
+values was too large to search
 exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
-see Supplementary Information). For each of 10000 random \(M\), the
+see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
+random \(M\), the
 genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis
-was performed on \(M\) using each set of \(\gamma\) values, and the 20
-sets with the lowest \(\max\left(\Re(\lambda)\right)\) each produced 50
-offspring with subsequent mutation and crossover between the resulting
-new population of 1000 \(\gamma\) sets. The genetic algorithm terminated
-if a stable \(M\) was found, 20 generations occurred, or a convergence
-criteria of minimum fitness increase between generations was satisfied.
-Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were
-found that resulted in stable systems with probabilities that were
-orders of magnitude higher than when \(\gamma = 1\) (see Supporting
-Information), meaning that stability could often be achieved by
-manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\)
-elements. Hence, managing the response rates of system components in a
-targetted way can potentially facilitate the stabilisation of complex
-systems through a reduction in dimensionality.
+\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
+performed on \(M\) using 
+each set of \(\gamma\) values, and the 20 sets with the
+lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
+subsequent mutation and crossover between the resulting new population
+of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
+\(M\) was found, 20 generations occurred, or a convergence criteria of
+minimum fitness increase between generations was satisfied. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were orders of
+magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
+Information}), meaning that stability could often be achieved by 
+manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
+Hence, managing the response rates of system components in a targetted
+way can potentially facilitate the stabilisation of complex systems 
+through a reduction in dimensionality.
 
@@
 networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (competitive
-networks), positive (mutualist networks), or \(i\) and \(j\) pairs have
-opposing signs (predator-prey networks). In general, competitive and
-mutualist networks tend to be destabilising, and predator-prey network
-tend to be
+include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or \(i\) and
+\(j\) pairs have opposing signs (predator-prey networks). In
+general, competitive and mutualist networks tend to be destabilising,
+and predator-prey network tend to be
 stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
@@
 competitive and predator-prey networks increases, but the proportion of
-stable mutualist networks does not (see Supplementary Information).
+stable mutualist networks does not (see 
+\hyperlink{SIstart}{Supplementary Information}).
 Additionally, when each component of \(M\) is interpreted as a unique
@@
 variation in species generation time might be unlikely to drive
-stability in purely multi-species networks (see Supplementary
-Information).
+stability in purely multi-species networks (see \hyperlink{SIstart}{Supplementary
+Information}).
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\textbf{Acknowledgements}
+
+I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
+Conversations with L. Bussi\`ere and N. Bunnefeld improved the quality of
+this work.
 
@@
 
+\clearpage
+
+\begin{center}
+\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
+\end{center}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{This supplementary information supports the manuscript
+``Component response rate variation drives stability in large complex
+systems'' with all of the code required to recreate the analysis in the
+main text, and with additional analyses to support its conclusions. All
+text, code, and data underlying this manuscript are publicly available
+on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}
+as part of the RandomMatrixStability package.}
+
+The
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
+package} includes all functions and tools for recreating the text, this
+supplemental information, and running all code; additional documentation
+is also provided for functions as part of the package. The
+RandomMatrixStability package package is available on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
+download it, the
+\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
+library} is needed.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
+\KeywordTok{library}\NormalTok{(devtools);}
+\end{Highlighting}
+\end{Shaded}
+
+The code below installs the RandomMatrixStability package using
+devtools.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+While downloading this package is recommended, all relevant code is also
+reproduced below with explanation, so it is possible to recreate all
+analyses using only this Supplementary Information.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\section{Supplementary Information table of
+contents}\label{supplemental-information-table-of-contents}
+
+\begin{itemize}
+\tightlist
+\item
+  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
+\item
+  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
+\item
+  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
+\item
+  \protect\hyperlink{ecological}{Stability of ecological networks}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{competition}{Competitor networks}
+  \item
+    \protect\hyperlink{mutualism}{Mutualist networks}
+  \item
+    \protect\hyperlink{pred-prey}{Predator-prey networks}
+  \end{itemize}
+\item
+  \protect\hyperlink{connectance}{Different connectance (C) values}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{connect3}{C = 0.3}
+  \item
+    \protect\hyperlink{connect5}{C = 0.5}
+  \item
+    \protect\hyperlink{connect7}{C = 0.7}
+  \item
+    \protect\hyperlink{connect9}{C = 0.9}
+  \end{itemize}
+\item
+  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
+\item
+  \protect\hyperlink{ga}{Genetic algorithm}
+\item
+  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
+\end{itemize}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
+1}\label{Fig1}}
+
+The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
+text is available on
+\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
+and was produced by running the following function.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
+    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
+        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
+        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
+        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
+        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
+        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
+        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
+        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
+        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
+        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
+        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
+        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
+        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
+        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
+        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
+        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
+        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
+            \NormalTok{break;}
+        \NormalTok{\}}
+        \KeywordTok{print}\NormalTok{(iters);}
+        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
+    \NormalTok{\}}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
+is found that is not stable when all component response rates are set to
+\(\gamma = 1\), but is stable when half of component response rates are
+\(1.95\) and half are \(0.05\). The function is used to illustrate the
+concept of how fast versus slow component responses can cause a system
+to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
+but terminated once an acceptable \texttt{A0} and \texttt{A1} were
+found. The code below plots the eigenvalue distributions of \texttt{A0}
+and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
+plot itself can be recreated with the function and code below.
+
+\includegraphics{unnamed-chunk-4-1.pdf}
+
+To find out how frequently \(M\) was stable given that all
+\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
+was created.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
+    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
+    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
+        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
+        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
+        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
+        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
+        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
+        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
+        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
+        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
+        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
+        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
+        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
+        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
+        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
+        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
+        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
+        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
+        \NormalTok{\}}
+        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
+        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(ress);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The above functions produced the \texttt{bi\_pr\_st} data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
+\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
+\end{Highlighting}
+\end{Shaded}
+
+The function \texttt{stab\_bgamma} was run for
+\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
+returned. Each row of \texttt{ress} represents a single \(M\) given
+\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
+Values of 0 indicate that \(M\) was found to be unstable (at least one
+real component of its eigenvalues greater than or equal to zero),
+whereas values of 1 indicate that \(M\) was found to be stable (all real
+components of eigenvalues are negative). The frequencies of stable \(M\)
+were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
+reported in the main text and legend of Fig. 1 (raw data are
+\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
+on GitHub}).
+
+\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
+2}\label{Fig2}}
+
+Figure 2 of the main text shows eigenvalue distributions in a system
+where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
+reproduced using the code below for when \(\gamma = 1\) (panel a) and
+\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
+reproduces the figure.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
+    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
+    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
+    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
+    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
+    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
+    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
+    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
+    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
+    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
+    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
+    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
+    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
+    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
+    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
+    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
+    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
+    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
+    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
+    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
+    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
+    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
+    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
+         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
+         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
+    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
+    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
+    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
+    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
+         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+\NormalTok{\}}
+\KeywordTok{plot_Fig_2}\NormalTok{();}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{unnamed-chunk-8-1.pdf}
+
+\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
+\(S\)}{Stability across increasing S}}\label{IncrS}}
+
+Figure 3 of the main text reports the number of stable random complex
+systems found over 1 million iterations. The data used to make this
+figure are read into R below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_1/random_all.csv"}\NormalTok{);}
+\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[,-}\DecValTok{1}\NormalTok{]; }\CommentTok{# Extra row-indicating column removed}
+\end{Highlighting}
+\end{Shaded}
+
+The table below shows the results for all simulations of random \(M\)
+matrices at \(\sigma = 0.4\) and \(C = 1\) given a range of
+\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
+matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
+after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
+Each row summarises data for a given \(S\) over 1 million randomly
+simulated \(M\) (\texttt{A0} and \texttt{A1}). The column
+\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
+unstable, and the column \texttt{A0\_stable} shows the number of
+\texttt{A0} matrices that are stable (these two columns sum to 1
+million). Similarly, the column \texttt{A1\_unstable} shows the number
+of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
+the number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 293 & 999707 & 293 & 999707 & 0 & 0\tabularnewline
+3 & 3602 & 996398 & 3609 & 996391 & 0 & 7\tabularnewline
+4 & 14937 & 985063 & 15008 & 984992 & 0 & 71\tabularnewline
+5 & 39289 & 960711 & 39783 & 960217 & 36 & 530\tabularnewline
+6 & 78845 & 921155 & 80207 & 919793 & 389 & 1751\tabularnewline
+7 & 133764 & 866236 & 136904 & 863096 & 1679 & 4819\tabularnewline
+8 & 204112 & 795888 & 208241 & 791759 & 5391 & 9520\tabularnewline
+9 & 288041 & 711959 & 291775 & 708225 & 12619 & 16353\tabularnewline
+10 & 384024 & 615976 & 384931 & 615069 & 23153 & 24060\tabularnewline
+11 & 485975 & 514025 & 481019 & 518981 & 35681 & 30725\tabularnewline
+12 & 590453 & 409547 & 577439 & 422561 & 48302 & 35288\tabularnewline
+13 & 689643 & 310357 & 669440 & 330560 & 57194 & 36991\tabularnewline
+14 & 777496 & 222504 & 751433 & 248567 & 60959 & 34896\tabularnewline
+15 & 850159 & 149841 & 821613 & 178387 & 58567 & 30021\tabularnewline
+16 & 905057 & 94943 & 877481 & 122519 & 51255 & 23679\tabularnewline
+17 & 943192 & 56808 & 919536 & 80464 & 40854 & 17198\tabularnewline
+18 & 969018 & 30982 & 949944 & 50056 & 30102 & 11028\tabularnewline
+19 & 984301 & 15699 & 970703 & 29297 & 20065 & 6467\tabularnewline
+20 & 992601 & 7399 & 983507 & 16493 & 12587 & 3493\tabularnewline
+21 & 996765 & 3235 & 991532 & 8468 & 7030 & 1797\tabularnewline
+22 & 998693 & 1307 & 995567 & 4433 & 3884 & 758\tabularnewline
+23 & 999503 & 497 & 997941 & 2059 & 1883 & 321\tabularnewline
+24 & 999861 & 139 & 999059 & 941 & 899 & 97\tabularnewline
+25 & 999964 & 36 & 999617 & 383 & 380 & 33\tabularnewline
+26 & 999993 & 7 & 999878 & 122 & 121 & 6\tabularnewline
+27 & 999995 & 5 & 999946 & 54 & 53 & 4\tabularnewline
+28 & 1000000 & 0 & 999975 & 25 & 25 & 0\tabularnewline
+29 & 1000000 & 0 & 999997 & 3 & 3 & 0\tabularnewline
+30 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+31 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
+matrices found is greater than 1 (compare column 5 to column 3), and
+this ratio increases with increasing \(S\) (column 1). Hence, more
+randomly created complex systems (\(M\)) are generated given variation
+in \(\gamma\) than when \(\gamma = 1\). The results underlying this
+table were produced with the \texttt{rand\_gen\_var} function below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
+    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
+    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
+    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
+        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
+        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
+        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
+        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
+            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
+            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
+            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
+            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
+            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
+            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
+            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
+            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
+            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
+            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
+            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
+        \NormalTok{\}}
+        \KeywordTok{print}\NormalTok{(i);}
+    \NormalTok{\}}
+    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
+    \KeywordTok{return}\NormalTok{(all_res);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The above function calls the two functions
+\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
+are provided below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
+        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
+        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
+        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
+            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
+                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
+                \NormalTok{\}}
+            \NormalTok{\}}
+        \NormalTok{\}}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(mat);}
+\NormalTok{\}}
+
+\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
+    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
+    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
+    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
+        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
+        \CommentTok{# Stable and unstable}
+        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \CommentTok{# Stabilised and destabilised}
+        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
+\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
+\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+        \CommentTok{# Feasible and infeasible}
+        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \CommentTok{# Feased and defeased}
+        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
+\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
+        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
+\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
+    \NormalTok{\}}
+    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
+                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
+                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
+                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
+    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
+    \KeywordTok{return}\NormalTok{(all_res);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+Note that feasibility results were ommited for the table above, but are
+\protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ecological}{\section{Stability of ecological
+networks}\label{ecological}}
+
+While the foundational work of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
+to complex networks, much attention has been given specifically to
+ecological networks of interacting species. In these networks, the
+matrix \(M\) is interpreted as a community matrix and each row and
+column is interpreted as a single species. The effect that the density
+of any species \(i\) has on the population dynamics of species \(j\) is
+found in \(M_{ij}\), meaning that \(M\) holds the effects of pair-wise
+interactions between \(S\)
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Tang2014b}{4}}.
+While May's original
+work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
+randomly assembled communities, recent work has specifically looked at
+more restricted ecological communities including competitive networks
+(all off-diagonal elements of \(M\) are negative), mutualist networks
+(all off-diagonal elements of \(M\) are positive), and predator-prey
+networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
+is negative and \(j\) on \(i\) is positive, or vice
+versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Allesina2011}{5}}.
+In general, competitor and mutualist networks tend to be unstable, while
+predator-prey networks tend to be highly stabilising.
+
+I investigated competitor, mutualist, and predator-prey networks
+following Allesina et
+al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
+these networks, I first generated a random matrix \(M\), then changed
+the elements of \(M\) accordingly. If \(M\) was a competitive network,
+then the sign of any positive off-diagonal elements was reversed to be
+negative. If \(M\) was a mutualist network, then the sign of any
+positive off-diagonal elements was reversed to be positive. And if \(M\)
+was a predator-prey network, then all \(i\) and \(j\) pairs of elements
+were checked; any pairs of the same sign were changed so that one was
+negative and the other was positive. The \texttt{species\_interaction}
+function used to do this is below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
+        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
+        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
+        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
+            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
+                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
+                \NormalTok{\}}
+            \NormalTok{\}}
+        \NormalTok{\}}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(mat);}
+\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
+\end{Highlighting}
+\end{Shaded}
+
+This function was applied to all created matrices \(M\), then the number
+of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
+as it was} in the main text for random matrices for values of \(S\) from
+2 to 50 (100 in the case of the relatively more stable predator-prey
+interactions), except that only 100000 random \(M\) were generated
+instead of 1 million. This produced the data set below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
+\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
+\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
+\end{Highlighting}
+\end{Shaded}
+
+The following tables for restricted ecological communities can therefore
+be compared with the random \(M\) \protect\hyperlink{IncrS}{results
+above} (but note that counts from systems with comparable probabilities
+of stability will be an order of magnitude lower in the tables below due
+to the smaller number of \(M\) matrices generated). As with the
+\protect\hyperlink{IncrS}{results above}, in the tables below,
+\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
+refers to matrices after \(Var(\gamma)\) is added. The column
+\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
+unstable, and the column \texttt{A0\_stable} shows the number of
+\texttt{A0} matrices that are stable (these two columns sum to 100000).
+Similarly, the column \texttt{A1\_unstable} shows the number of
+\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
+number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+\textbf{Competition}
+
+Results for competitor interaction networks are shown below
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised\tabularnewline
+\midrule
+\endhead
+1 & 2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
+2 & 3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
+3 & 4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
+4 & 5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
+5 & 6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
+6 & 7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
+7 & 8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
+8 & 9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
+9 & 10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
+10 & 11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
+11 & 12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
+12 & 13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
+13 & 14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
+14 & 15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
+15 & 16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
+16 & 17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
+17 & 18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
+18 & 19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
+19 & 20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
+20 & 21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
+21 & 22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
+22 & 23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
+23 & 24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
+24 & 25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
+25 & 26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
+26 & 27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
+27 & 28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
+28 & 29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
+29 & 30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
+30 & 31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
+31 & 32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
+32 & 33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
+33 & 34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Mutualism}
+
+Results for mutualist interaction networks are shown below
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised\tabularnewline
+\midrule
+\endhead
+1 & 2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
+2 & 3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
+3 & 4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
+4 & 5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
+5 & 6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
+6 & 7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
+7 & 8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+8 & 9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+9 & 10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+10 & 11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+11 & 12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Predator-prey}
+
+Results for predator-prey interaction networks are shown below
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised\tabularnewline
+\midrule
+\endhead
+1 & 2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+2 & 3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+3 & 4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+4 & 5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
+5 & 6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
+6 & 7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
+7 & 8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
+8 & 9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
+9 & 10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
+10 & 11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
+11 & 12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
+12 & 13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
+13 & 14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
+14 & 15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
+15 & 16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
+16 & 17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
+17 & 18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
+18 & 19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
+19 & 20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
+20 & 21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
+21 & 22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
+22 & 23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
+23 & 24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
+24 & 25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
+25 & 26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
+26 & 27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
+27 & 28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
+28 & 29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
+29 & 30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
+30 & 31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
+31 & 32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
+32 & 33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
+33 & 34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
+34 & 35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
+35 & 36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
+36 & 37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
+37 & 38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
+38 & 39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
+39 & 40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
+40 & 41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
+41 & 42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
+42 & 43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
+43 & 44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
+44 & 45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
+45 & 46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
+46 & 47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
+47 & 48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
+48 & 49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
+49 & 50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
+50 & 51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
+51 & 52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
+52 & 53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
+53 & 54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
+54 & 55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
+55 & 56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
+56 & 57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
+57 & 58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
+58 & 59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
+59 & 60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
+60 & 61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
+61 & 62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
+62 & 63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
+63 & 64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
+64 & 65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
+65 & 66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
+66 & 67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
+67 & 68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
+68 & 69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
+69 & 70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
+70 & 71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
+71 & 72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
+72 & 73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
+73 & 74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
+74 & 75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
+75 & 76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
+76 & 77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
+77 & 78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
+78 & 79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
+79 & 80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
+80 & 81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
+81 & 82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
+82 & 83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
+83 & 84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
+84 & 85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
+85 & 86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
+86 & 87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
+87 & 88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
+88 & 89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
+89 & 90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
+90 & 91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
+91 & 92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
+92 & 93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
+93 & 94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
+94 & 95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
+95 & 96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
+96 & 97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
+97 & 98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
+98 & 99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
+99 & 100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
+\bottomrule
+\end{longtable}
+
+Overall, as
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
+predator-prey communities are relatively stable while mutualist
+communties are highly unstable. But interestingly, while \(Var(\gamma)\)
+stabilises predator-prey and competitor communities, it does not
+stabilise mutualist communities. This is unsurprising because purely
+mutualist communities are characterised by a very
+positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
+alone will shift all real parts of eigenvalues to negative values.
+
+\hypertarget{connectance}{\section{Different connectance (C)
+values}\label{connectance}}
+
+In the main text, for simplicity, I assumed connectance values of
+\(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
+potentially nonzero and sampled from a normal distribution
+\(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here I present
+four tables showing the number of stable communities given
+\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
+component response time (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
+higher number of stable communities than when \(\gamma\) did not vary
+(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
+million \(M\) were simulated. As with the results on
+\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
+in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
+and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
+column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
+that are unstable, and the column \texttt{A0\_stable} shows the number
+of \texttt{A0} matrices that are stable (these two columns sum to
+100000). Similarly, the column \texttt{A1\_unstable} shows the number of
+\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
+number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+All data reported below for various values of \(C\) are accessible using
+the below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{C3dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt3.csv"}\NormalTok{);}
+\NormalTok{C5dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt5.csv"}\NormalTok{);}
+\NormalTok{C7dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt7.csv"}\NormalTok{);}
+\NormalTok{C9dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt9.csv"}\NormalTok{);}
+\end{Highlighting}
+\end{Shaded}
+
+These objects \texttt{C3dat}, \texttt{C5dat}, \texttt{C7dat}, and
+\texttt{C9dat} include the results for \(C = 0.3\), \(C = 0.5\),
+\(C = 0.7\), and \(C = 0.9\), respectively.
+
+\textbf{Connectance \(C = 0.3\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
+4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
+5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
+6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
+7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
+8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
+9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
+10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
+11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
+12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
+13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
+14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
+15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
+16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
+17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
+18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
+19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
+20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
+21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
+22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
+23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
+24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
+25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
+26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
+27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
+28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
+29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
+30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
+31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
+32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
+33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
+34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
+35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
+36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
+37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
+38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
+39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
+40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
+41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
+42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
+43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
+44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
+45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
+46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
+47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
+48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
+49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
+50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
+51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
+52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
+53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
+54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
+55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(C = 0.5\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
+3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
+4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
+5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
+6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
+7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
+8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
+9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
+10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
+11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
+12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
+13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
+14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
+15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
+16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
+17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
+18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
+19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
+20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
+21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
+22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
+23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
+24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
+25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
+26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
+27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
+28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
+29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
+30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
+31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
+32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
+33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
+34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
+35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
+36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
+37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
+38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
+39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
+40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(C = 0.7\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
+3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
+4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
+5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
+6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
+7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
+8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
+9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
+10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
+11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
+12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
+13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
+14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
+15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
+16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
+17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
+18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
+19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
+20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
+21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
+22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
+23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
+24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
+25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
+26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
+27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
+28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
+29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
+30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
+31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
+32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
+33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
+34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(C = 0.9\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
+3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
+4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
+5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
+6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
+7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
+8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
+9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
+10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
+11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
+12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
+13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
+14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
+15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
+16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
+17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
+18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
+19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
+20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
+21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
+22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
+23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
+24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
+25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
+26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
+27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
+28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
+29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
+30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{gam_dist}{\section{\texorpdfstring{Different distributions
+of
+\(\gamma\)}{Different distributions of \textbackslash{}gamma}}\label{gam_dist}}
+
+In the main text, I considered a uniform distribution of component
+response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
+and stable \(M\) matrices are reported in \protect\hyperlink{IncrS}{a
+table above} across different values of \(S\). Here I show complementary
+results for three different distributions including an exponential,
+beta, and gamma distribution of \(\gamma\) values. The shape of these
+distributions is shown in the figure below.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Distributions of component response rate
+(\(\boldsymbol{\gamma}\)) values in complex systems.} The stabilities of
+simulated complex systems with these \(\gamma\) distributions are
+compared to otherwise identical complex systems with a fixed component
+response rate of \(\gamma = 1\) across different system sizes (\(S\);
+i.e., component numbers) given a unit \(\gamma\) standard deviation
+(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
+uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
+\(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
+panel shows 1 million randomly generated \(\gamma\) values.
+
+\includegraphics{unnamed-chunk-23-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+The same 100000 \(M\) matrices were used to investigate stability when
+applying each of these different distributions of \(\gamma\) values. The
+table below shows the number of \(M\) that were unstable (\_unst) and
+stable (\_stbl) for the exponential (Exp), beta, and gamma
+distributions.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
+\KeywordTok{kable}\NormalTok{(fourdists);}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
+gamma\_stbl\tabularnewline
+\midrule
+\endhead
+2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
+3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
+4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
+5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
+6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
+7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
+8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
+9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
+10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
+11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
+12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
+13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
+14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
+15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
+16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
+17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
+18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
+19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
+20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
+21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
+22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
+23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
+24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
+25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
+26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
+27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
+28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
+29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
+30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
+32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+In comparison to the uniform distribution (a), proportionally fewer
+random systems are found with the exponential distribution (b), while
+more are found with the beta (c) and gamma (d) distributions.
+
+\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}
+
+Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
+the proportion of stable complex systems, the search space of all
+possible \(\gamma\) vectors would be evaluated for each unique \(M\).
+This is technically impossible because \(\gamma_{i}\) can take any real
+value between 0-2, but even rounding \(\gamma\) to reasonable values
+would result in a search space too large to practically explore. Under
+these conditions, genetic algorithms are highly useful tools for finding
+practical solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{6}}. In
+this case, the practical solution is finding vectors of \(\gamma\) that
+decrease the most positive real eigenvalue of \(M\). The genetic
+algorithm below achieves this by initialising a large population of 1000
+different potential \(\gamma\) vectors and allowing this population to
+evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
+values between vectors), selection, and reproduction until either a
+\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
+``giving up'' critiera is met (in the below, this ``giving up''"
+criteria is met when 20 generations pass, or if the fitness increase
+from one generation to the next is below a certain criteria). The
+genetic algorithm relies on five functions. The first outer function
+\texttt{Evo\_rand\_gen\_var} runs all of the simulations
+(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
+\texttt{iters} refers to the number of \(M\) to try for each \(S\)).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
+    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
+    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
+    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
+        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
+        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
+        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
+        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
+        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
+            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
+            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
+            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
+            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
+            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
+            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
+            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
+            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
+            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
+            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
+            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
+            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
+                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
+            \NormalTok{\}}
+            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
+        \NormalTok{\}}
+        \KeywordTok{print}\NormalTok{(i);}
+    \NormalTok{\}}
+    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
+    \KeywordTok{return}\NormalTok{(all_res);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
+\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
+\texttt{summarise\_randmat}. The first simply allows for non-random
+interactions between components (e.g., modelling
+\protect\hyperlink{ecological}{ecological interactions} of random,
+competition, mutualism, or predator-prey).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
+        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
+        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
+    \NormalTok{\}}
+    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
+        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
+            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
+                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
+                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
+                \NormalTok{\}}
+            \NormalTok{\}}
+        \NormalTok{\}}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(mat);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
+algorithm, searching for \(\gamma\) vectors that are stabilising.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
+    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
+    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
+    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
+    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
+    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
+    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
+    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
+        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
+        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
+        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
+        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
+            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
+            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
+            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
+        \NormalTok{\}}
+        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
+        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
+        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
+                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
+        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
+        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
+        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
+        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
+        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
+        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
+        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
+        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
+        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
+        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
+        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
+        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
+        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
+        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
+    \NormalTok{\}}
+    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
+        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
+        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
+        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
+        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(find_st);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+The while loop in \texttt{rand\_mat\_ga} continues until either
+\texttt{iter} generations have occured, a solution \(\gamma\) vector is
+found that results in all \(\Re(\lambda) < 0\), or some criteria of
+minimum fitness increase is observed (by default,
+\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
+values are mutated, crossover occurs between \(\gamma\) vectors, and
+selection occurs in each generation such that the 20 \(\gamma\) vectors
+that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
+offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
+zero are multiplied by \(-1\), and any values that mutate above 2 are
+set to 2. Note also that if a solution is found, then one such
+\(\gamma\) vector causing stability is printed to a file.
+
+Crossover occurs in the \texttt{crossover} function below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
+    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
+    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
+    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
+    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
+        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
+        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
+        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
+        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
+        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(inds);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
+\texttt{summarise\_randmat} formats the data into a table.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{summarise_randmat_ga <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
+    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
+    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{10}\NormalTok{);}
+    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
+        \NormalTok{unstables <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{stables   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{non_feas  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{feasibl   <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{unfeased  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{feased    <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{foundd    <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{3}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
+        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstables);}
+        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stables);}
+        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
+        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
+        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(non_feas);}
+        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feasibl);}
+        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unfeased);}
+        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feased);}
+        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(foundd);}
+    \NormalTok{\}}
+    \KeywordTok{return}\NormalTok{(all_res);}
+\NormalTok{\}}
+\end{Highlighting}
+\end{Shaded}
+
+Some stability results from this table are shown below. Each histogram
+shows a different distribution of \(\gamma\) that was found to be
+stabilising.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
+3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
+4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
+5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
+6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
+7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
+8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
+9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
+10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
+11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
+12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
+13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
+14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
+15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
+16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
+17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
+18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
+19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
+20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
+21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
+22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
+23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
+24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
+25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
+26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
+27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
+28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
+29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
+30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
+31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
+32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
+33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
+34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
+35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
+36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
+37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
+38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
+39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
+40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. Recall that 1 million random matrices were
+generated for the less computationally intense task of
+\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
+when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
+compare stability in column 5 above with column 3 above. \textbf{This
+comparison shows the high number of stable \(M\) that can be produced
+through a targetted search of \(\gamma\) values, and suggests that many
+otherwise unstable systems could potentially be stabilised by an
+informed manipulation of their component response times. Such a
+possibility might conceivably reduce the dimensionality of problems
+involving stability in social-ecological or economic systems.}
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
+\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{unnamed-chunk-31-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
+
+\hypertarget{Feasibility}{\section{Feasibility of complex
+systems}\label{Feasibility}}
+
+For complex systems in which individual system components (\(S\))
+represent the density of some tangible quantity, it is important to
+consider the feasibility of the system. Feasibile equilibria assume that
+the values of all system components are positive at
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{7}--\protect\hyperlink{ref-Song2018}{9}}.
+This is of particular interest for ecological communities because
+population density cannot take negative values, meaning that ecological
+systems need to be feasible for stability to be biologically
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
+Consequently, the use of random matrices and traditional stability
+critiera for making inferences in theoretical analyses of species
+networks has recently been
+criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
+While the key results in the main text are intended to be general to all
+complex systems, and not restricted to species networks, I have also
+performed a feasibility analysis on all matrices \(M\). This analysis
+reveals that feasibility is not affected by \(Var(\gamma)\), meaning
+that for pure interacting species networks, variation in component
+response time (i.e., species generation time) does not affect stability
+at biologically realistic species densities. Nevertheless, ecological
+interactions do not exist in isolation in empirical systems, but instead
+interact with
+evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{10}},
+abiotic, or social-economic systems. The relevance of \(\gamma\) for
+complex system stability presented in the main text should therefore not
+be ignored in the broader context of ecological communities.
+
+Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}
+define the following feasibility criteria for ecological systems
+characterised by \(S\) interacting species with varying densities.
+
+\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]
+
+In the above, \(x^{*}\) is the vector of species abundances at
+equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
+The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
+off-diagonal elements), and the value \(\theta\) is strength of
+intraspecific competition (diagonal elements). As I have done elsewhere,
+diagonal values are set to \(-1\), so \(\theta = -1\). The variable
+\(C\) is the connectance of the community,
+which was set to \(C = 1\) throughout the manuscript and supplemental
+information, except \protect\hyperlink{connectance}{where otherwise
+noted}. The variable \(\delta\) is a normalisation parameter that
+modulates the strength of interactions (\(\sigma\) in the main text),
+which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
+underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
+so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
+\(A\), which has a diagonal of all zeros and an off-diagonal affecting
+species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} stability
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}} by
+\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
+\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
+text and Supplementary Information). The above criteria is therefore
+reduced to the below; note that the parenthetical in both equations
+produces an \(M\) matrix as used throughout the main text and
+supplemental information,
+
+\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]
+
+To check the feasibility criteria, I therefore inverted
+\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
+the resulting matrix by the vector of population growth rates \(r\).
+Feasibility is satisfied if all of the elements of the resulting vector
+are positive.
+
+The population growth rate for an individual species \(i\) is sampled
+from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
+shown in the \texttt{rand\_gen\_var} function in
+\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
+\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
+is assumed to be a species with a growth rate of \(r_{i}\). Note that
+negative intrinsic growth rates are not unrealistic, and will occur in
+obligate mutualists in the absence of a partner.
+
+When feasibility was evaluated with and without variation in \(\gamma\),
+there was no increase in stability for \(M\) where \(\gamma\) varied as
+compared to where \(\gamma = 1\). Results below illustrate this result,
+which was general to all other simulations performed.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
+A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 749978 & 250022 & 749942 & 250058 & 35552 & 35516\tabularnewline
+3 & 874519 & 125481 & 874296 & 125704 & 36803 & 36580\tabularnewline
+4 & 937192 & 62808 & 937215 & 62785 & 26440 & 26463\tabularnewline
+5 & 968776 & 31224 & 968639 & 31361 & 16319 & 16182\tabularnewline
+6 & 984313 & 15687 & 984463 & 15537 & 9006 & 9156\tabularnewline
+7 & 992149 & 7851 & 992161 & 7839 & 4991 & 5003\tabularnewline
+8 & 996124 & 3876 & 996103 & 3897 & 2644 & 2623\tabularnewline
+9 & 998014 & 1986 & 998027 & 1973 & 1361 & 1374\tabularnewline
+10 & 999031 & 969 & 999040 & 960 & 698 & 707\tabularnewline
+11 & 999546 & 454 & 999514 & 486 & 377 & 345\tabularnewline
+12 & 999764 & 236 & 999792 & 208 & 160 & 188\tabularnewline
+13 & 999883 & 117 & 999865 & 135 & 105 & 87\tabularnewline
+14 & 999938 & 62 & 999945 & 55 & 40 & 47\tabularnewline
+15 & 999971 & 29 & 999964 & 36 & 31 & 24\tabularnewline
+16 & 999988 & 12 & 999991 & 9 & 8 & 11\tabularnewline
+17 & 999996 & 4 & 999991 & 9 & 8 & 3\tabularnewline
+18 & 999997 & 3 & 999999 & 1 & 1 & 3\tabularnewline
+19 & 999998 & 2 & 999997 & 3 & 3 & 2\tabularnewline
+20 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+21 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+22 & 999999 & 1 & 1000000 & 0 & 0 & 1\tabularnewline
+23 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+24 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+25 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+26 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+27 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+28 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+29 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+30 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+31 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
+in pure species interaction networks.
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-May1972}{}
+1. May, R. M. Will a large complex system be stable? \emph{Nature}
+\textbf{238,} 413--414 (1972).
+
+\hypertarget{ref-Allesina2012}{}
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+\emph{Nature} \textbf{483,} 205--208 (2012).
+
+\hypertarget{ref-Allesina2015a}{}
+3. Allesina, S. \& Tang, S. The stability--complexity relationship at
+age 40: a random matrix perspective. \emph{Population Ecology} 63--75
+(2015).
+doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
+
+\hypertarget{ref-Tang2014b}{}
+4. Tang, S. \& Allesina, S. Reactivity and stability of large
+ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
+(2014).
+
+\hypertarget{ref-Allesina2011}{}
+5. Allesina, S. \& Levine, J. M. A competitive network theory of species
+diversity. \emph{Proceedings of the National Academy of Sciences of the
+United States of America} \textbf{108,} 5638--5642 (2011).
+
+\hypertarget{ref-Hamblin2013}{}
+6. Hamblin, S. On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+184--194 (2013).
+
+\hypertarget{ref-Grilli2017}{}
+7. Grilli, J. \emph{et al.} Feasibility and coexistence of large
+ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
+
+\hypertarget{ref-Dougoud2018}{}
+8. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+The feasibility of equilibria in large ecosystems: a primary but
+neglected concept in the complexity-stability debate. \emph{PLOS
+Computational Biology} \textbf{14,} e1005988 (2018).
+
+\hypertarget{ref-Song2018}{}
+9. Song, C. \& Saavedra, S. Will a small randomly assembled community be
+feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
+
+\hypertarget{ref-Patel2018}{}
+10. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+effects of eco-evolutionary feedbacks on community stability.
+\emph{American Naturalist} \textbf{191,} 1--29 (2018).
+
 \end{document}

### 2018-06-04 notebook/arxiv/ms.tex
Minor grammar issues

@@
 stability of large systems might be dependent upon variation in the
-response rate of their individual components, meaning that factors such
+response rates of their individual components, meaning that factors such
 as generation time (in ecological networks), transaction speed (in
-economic networks), or communication speed (in social networks) needs to
+economic networks), or communication speed (in social networks) need to
 be considered when investigating the stability of complex systems.
@@
 \(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
-\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values. To
-further investigate the potential of \(Var(\gamma)\) to be stabilising,
+\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values.
+
+To further investigate the potential of \(Var(\gamma)\) to be stabilising,
 I used a genetic algorithm (the space of possible \(\gamma\)
@@
 each set of \(\gamma\) values, and the 20 sets with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 offspring with
+lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
 subsequent mutation and crossover between the resulting new population
@@
 text, code, and data underlying this manuscript are publicly available
-on \href{https://github.com/bradduthie/random_matrix_stability}{GitHub}
+on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}
 as part of the RandomMatrixStability package.}
@@
 The
-\href{https://github.com/bradduthie/random_matrix_stability}{RandomMatrixStability
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
 package} includes all functions and tools for recreating the text, this
@@
 RandomMatrixStability package package is available on
-\href{https://github.com/bradduthie/random_matrix_stability}{GitHub}; to
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
 download it, the
@@
 \begin{Highlighting}[]
-\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/random_matrix_stability"}\NormalTok{)}
+\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{);}
 \end{Highlighting}
@@
 text is available on
-\href{https://github.com/bradduthie/random_matrix_stability/tree/master/sim_results/bi_gamma}{GitHub},
+\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
 and was produced by running the following function.
@@
 reported in the main text and legend of Fig. 1 (raw data are
-\href{https://github.com/bradduthie/random_matrix_stability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
+\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
 on GitHub}).

### 2018-06-04 notebook/ms.Rmd
Minor grammar issues

@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
+
 To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
 For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.

### 2018-06-05 notebook/SI.Rmd
Semi-colon
Add in the text of Issue #1 to remind myself to write and include further detail on the matsh of gamma in future SI versoins

@@
 ```{r, eval = FALSE}
-install_github("bradduthie/RandomMatrixStability")
+install_github("bradduthie/RandomMatrixStability");
 ```
@@
 ********************************************************************************
+
+<!---
+
+Further explanation of $\gamma$ {#moregamma}
+
+It would be perhaps useful to include a more detailed explanation of what $\gamma$ does in the context of the $M$.
+That is, as in @patel2018, I'm looking at $\frac{d N_{i}}{dt}$ and $\frac{d N_{j}}{dt}$, with $N_{i}$ and $N_{j}$ potentially affecting one another -- the $\gamma$ parameter modulates the change in each with respect to time at local equilibrium; in the language of @patel2018, it "determines the timescale separation" between the dynamics of individual components in the system $M$.
+The best place to provide further detail for this is likely to be early on in the Supplemental Information.
+
+--->
 

### 2018-06-05 vignettes/R_old/evolve_gammas.R
Initialise Website files

@@ -1,0 +1,156 @@
+Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
+    tot_res <- NULL;
+    fea_res <- NULL;
+    if(from >= to){
+        stop("Argument 'from' must be less than argument 'to'");
+    }
+    for(i in from:to){
+        nn             <- i;
+        A1_stt         <- 0;
+        A2_stt         <- 0;
+        A1_fet         <- 0;
+        A2_fet         <- 0;
+        iter           <- iters;
+        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
+        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
+        while(iter > 0){
+            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            A0       <- species_interactions(mat = A0, type = int_type);
+            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            A0       <- A0 * C_mat;
+            diag(A0) <- -1;
+            gam1     <- runif(n = i, min = 0, max = 2);
+            A1       <- A0 * gam1;
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            A1_stb   <- rand_mat_ga(A1);
+            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
+            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
+            if(A0_stb == TRUE){
+                tot_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_stb == TRUE){
+                tot_res[[i-1]][iter, 2] <- 1;
+            }
+            if(A0_fea == TRUE){
+                fea_res[[i-1]][iter, 1] <- 1;
+            }
+            if(A1_fea == TRUE){
+                fea_res[[i-1]][iter, 2] <- 1;
+            }
+            iter    <- iter - 1;
+        }
+        print(i);
+    }
[…]

### 2018-06-05 vignettes/R_old/find_bi_gamma.R

@@ -1,0 +1,108 @@
+stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
+    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
+    A0_count <- 0;
+    A1_count <- 0;
+    while(iters > 0){
+        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
+        A_mat  <- matrix(data = A_dat, nrow = S);
+        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+        A_mat  <- A_mat * C_mat;
+        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
+        mu_gam <- mean(gammas);
+        diag(A_mat) <- -1;
+        A1     <- gammas * A_mat;
+        A0     <- mu_gam * A_mat;
+        A0_e   <- eigen(A0)$values;
+        A0_r   <- Re(A0_e);
+        A0_i   <- Im(A0_e);
+        A1_e   <- eigen(A1)$values;
+        A1_r   <- Re(A1_e);
+        A1_i   <- Im(A1_e);
+        if(max(A0_r) < 0){
+            ress[iters, 1] <- 1;
+            A0_count       <- A0_count + 1;
+            }
+        if(max(A1_r) < 0){
+            ress[iters, 2] <- 1;
+            A1_count       <- A1_count + 1;
+        }
+        print(c(iters, A0_count, A1_count));
+        iters <- iters - 1;
+    }
+    return(ress);
+}
+
+
+find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
+    while(iters > 0){
+        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
+        A_mat  <- matrix(data = A_dat, nrow = S);
+        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+        A_mat  <- A_mat * C_mat;
+        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
+        mu_gam <- mean(gammas);
+        diag(A_mat) <- -1;
[…]

### 2018-06-05 vignettes/R_old/plot_figs.R

@@ -1,0 +1,397 @@
+ make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- sdd * sqrt(12);
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
+
+show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
+    y1 <- make_gammas(nn = nn, distribution = 1, sdd = sdd, mn = mn);
+    y2 <- make_gammas(nn = nn, distribution = 2, sdd = sdd, mn = mn);
+    y3 <- make_gammas(nn = nn, distribution = 3, sdd = sdd, mn = mn);
+    y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
+    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
+    h1 <- hist(y1, breaks = 1000, plot = FALSE);
+    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+         ylim = c(0, max(h1$counts)*1.2));
+    mtext("a", adj = 0, line = -2.3, cex = 2, 
+          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
+    box();
[…]

### 2018-06-05 vignettes/R_old/scratch.R

@@ -1,0 +1,706 @@
+
+
+eout <- read.csv("Evo_out.txt");
+
+dat <- eout[eout[,1] == 40,];
+
+
+
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1, 
+                         gamma_mn = 1, C = 1){
+  tot_res  <- NULL;
+  fea_res  <- NULL;
+  vars_t   <- NULL;
+  off_cors <- NULL;
+  for(i in 2:max_sp){
+    iter           <- iters;
+    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+    fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+    while(iter > 0){
+      r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
+      A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
+      A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+      A0       <- species_interactions(mat = A0, type = int_type);
+      C_dat    <- rbinom(n = i * i, size = 1, prob = C);
+      C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+      A0       <- A0 * C_mat;
+      diag(A0) <- -1;
+      gam0     <- make_gammas(nn = i, distribution = 0, sdd = gamma_sd);
+      
+      gam1     <- runif(n = i, min = 0, max = 999) #make_gammas(nn = i, distribution = 1, sdd = gamma_sd);
+      gam0     <- rep(mean(gam1), i);
+      
+      gam2     <- make_gammas(nn = i, distribution = 2, sdd = gamma_sd);
+      gam3     <- make_gammas(nn = i, distribution = 3, sdd = gamma_sd);
+      gam4     <- make_gammas(nn = i, distribution = 4, sdd = gamma_sd);
+      
+      A1       <- A0 * gam1;
+      A2       <- A0 * gam1;
+      A3       <- A0 * gam0;
+      diag(A2) <- -1 * gam0;
+      diag(A3) <- -1 * gam1;
+      
+
[…]

### 2018-06-05 vignettes/R_old/sim_mat.R

@@ -1,0 +1,95 @@
+species_interactions <- function(mat, type = 0){
+    if(type == 1){
+        mat[mat > 0] <- -1*mat[mat > 0];
+    }
+    if(type == 2){
+        mat[mat < 0] <- -1*mat[mat < 0];
+    }
+    if(type == 3){
+        for(i in 1:dim(mat)[1]){
+            for(j in 1:dim(mat)[2]){
+                if(mat[i, j] * mat[j, i] > 0){
+                    mat[j, i] <- -1 * mat[j, i];
+                }
+            }
+        }
+    }
+    return(mat);
+}
+
+summarise_randmat <- function(tot_res, fea_res){
+    sims    <- length(tot_res);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
+    for(i in 1:sims){
+        all_res[i, 1]  <- i + 1;
+        # Stable and unstable
+        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
+        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
+        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
+        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
+        # Stabilised and destabilised
+        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
+                                  tot_res[[i]][,2] == TRUE);
+        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
+                                  tot_res[[i]][,2] == FALSE);
+        # Feasible and infeasible
+        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
+        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
+        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
+        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
+        # Feased and defeased
+        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
+                                  fea_res[[i]][,2] == TRUE);
+        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
+                                  fea_res[[i]][,2] == FALSE);
+    }
+    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
[…]

### 2018-06-05 vignettes/R_old/test_mat.R

@@ -1,0 +1,1555 @@
+
+################################################################################
+################################################################################
+################################################################################
+
+################################################################################
+################################################################################
+################################################################################
+
+################################################################################
+################################################################################
+################################################################################
+
+################################################################################
+################################################################################
+################################################################################
+
+################################################################################
+################################################################################
+
+image(-1*egi$hmat, col = heat.colors(2000));
+
+make_e_image <- function(ei_ret){
+     hmat   <- ei_ret$hmat;
+     xcoord <- ei_ret$xcoord;
+     ycoord <- ei_ret$ycoord;
+     adj_ei <- -1 * log(hmat + 0.001);
+     xaxis  <- which(xcoord == 0) / length(xcoord);
+     yaxis  <- which(ycoord == 0) / length(ycoord);
+     coomat <- matrix(data = 0, nrow = dim(hmat)[1], ncol = dim(hmat)[2]);
+     coomat[,xaxis] <- 1;
+     coomat[yaxis,] <- 1;
+     image(adj_ei, col = heat.colors(2000), axes = FALSE);
+     abline(h = xaxis, lwd = 2);
+     abline(v = yaxis, lwd = 2);
+     box();
+}
+
+
+eigen_image  <- function(eigen_list){
+    reigs <- round(eigen_list, digits = 1);
+    xcoord <- seq(from = -40, to = 40, by = 0.1);
+    ycoord <- seq(from = -40, to = 40, by = 0.1);
[…]

### 2018-06-07 notebook/arxiv/ms.tex
Remove a couple extraneous words

@@
 self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, values of off-diagonal elements are assigned non-zero
+Traditionally, off-diagonal elements are assigned non-zero
 values with a probability \(C\), which are sampled from a distribution

### 2018-06-07 notebook/ms.Rmd
Remove a couple extraneous words

@@
 Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, values of off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
+Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
 Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$)[@May1972; @Allesina2012].

### 2018-06-07 notebook/ms.tex
Semi-colon
Remove a couple extraneous words

@@
 \begin{Highlighting}[]
-\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{)}
+\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{);}
 \end{Highlighting}

### 2018-06-08 notebook/SI.Rmd
Fix Issue #2 in the SI; dummy index removed from all ecological tables

@@
 ```{r, echo = FALSE}
-cd_tab <- cdat[,1:7];
+cd_tab <- cdat[,2:7];
 kable(cd_tab);
@@
 ```{r, echo = FALSE}
-md_tab <- mdat[,1:7];
+md_tab <- mdat[,2:7];
 kable(md_tab);
@@
 ```{r, echo = FALSE}
-pd_tab <- pdat[,1:7];
+pd_tab <- pdat[,2:7];
 kable(pd_tab);

### 2018-06-08 notebook/SI.tex
Fix Issue #2 in the SI; dummy index removed from all ecological tables

--- 2018-06-07 notebook/arxiv/ms.tex
@@
 \documentclass[]{article}
-\usepackage{lmodern}
-\usepackage{setspace}
-\setstretch{1}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            pdftitle={Component response rate variation drives stability in large complex systems},
-            pdfauthor={A. Bradley Duthie},
-            colorlinks=true,
-            linkcolor=blue,
-            citecolor=Blue,
-            urlcolor=Blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
 \usepackage{lmodern}
@@
 
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
 \setlength{\droptitle}{-2em}
@@
   \posttitle{\par}
+\subtitle{Supplemental Information}
   \author{A. Bradley Duthie}
@@
 \usepackage[utf8]{inputenc}
+\linenumbers
 \bibliographystyle{amnatnat}
@@
 
-\textbf{The stability of a complex system generally decreases with
-increasing system size, as is demonstrated by random matrix
-theory\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-This counter-intuitive result, first shown by
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}, is broadly
-relevant for understanding the dynamics and persistence of systems such
-as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{3},\protect\hyperlink{ref-Gray2009}{4}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{5},\protect\hyperlink{ref-MacArthur2010}{6}}
-and
-socio-economic\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}--\protect\hyperlink{ref-Bardoscia2017}{9}}
-networks. Much attention has especially been given to the stability of
-ecological communities such as food webs or mutualist networks, with
-recent work investigating how different community structures affect
-stability\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10}--\protect\hyperlink{ref-Patel2018}{14}}.
-But more broadly, stabilising mechanisms in complex systems remain
-under-developed, and the effect of variation in the response rate of
-individual system components remains an open
-problem\textsuperscript{\protect\hyperlink{ref-Allesina2015}{15}}. Here
-I show that when components of a complex system respond to system
-dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
-system stability is markedly increased. Stability is caused by the
-clustering of some eigenvalues toward the centre of eigenvalue
-distributions despite the destabilising effect of higher interaction
-strength variation (\(\boldsymbol{\sigma^{2}}\)). This effect of
-variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
-system size increases, to the extent that the largest stable complex
-systems would otherwise be unstable if not for
-\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
-unconsidered driver of system stability that is likely to be pervasive
-across all complex systems. Future research in complex systems should
-therefore account for the varying response rates of individual system
-components when assessing whole system stability.}
-
-In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
-demonstrated that randomly assembled systems of sufficient complexity
-are almost inevitably unstable given infinitesimally small
-perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of potentially interacting components; \(S\)), its 
-connectance (i.e., the probability that one component will interact with 
-another; \(C\)), and the variance of interaction strengths
-(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-May's finding that the probability of local stability falls to near zero
-given a sufficiently high threshold of \(\sigma\sqrt{SC}\) has profound
-consequences across multiple disciplines, raising the question of how
-complex systems in, e.g.,
-ecology\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10},\protect\hyperlink{ref-Grilli2017}{13},\protect\hyperlink{ref-Allesina2015}{15}}
-or
-banking\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7},\protect\hyperlink{ref-Bardoscia2017}{9},\protect\hyperlink{ref-May2008}{16}}
-are predicted to persist or change.
-
-Randomly assembled complex systems can be represented as large square
-matrices (\(M\)) with \(S\) components (e.g.,
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
-element of such a matrix, \(M_{ij}\), defines how component \(j\) affects
-component \(i\) in the system at a point of
-equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-Off-diagonal elements (\(i \neq j\)) therefore define interactions
-between components, while diagonal elements (\(i = j\)) define component
-self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero
-values with a probability \(C\), which are sampled from a distribution
-with variance \(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}.
-Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of \(M\)
-are negative
-(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}} within a
-circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
-\(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}
-(Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma^{2}\) increase.
-
-The above stability criterion assumes that individual components respond
-to perturbations of the system at the same rate (\(\gamma\)), but this
-is highly unlikely in any complex system. In ecological communities, for
-example, the rate at which population density changes following
-perturbation will depend on the generation time of individuals, which
-might vary by orders of magnitude among species. Species with short
-generation times will respond quickly (high \(\gamma\)) to perturbations
-relative to species with long generation times (low \(\gamma\)).
-Similarly, the speed at which individual banks respond to perturbations
-in financial networks, or individuals or institutions respond to
-perturbations in complex social networks, is likely to vary. The effect
-of such variance has not been investigated in complex systems theory.
-Intuitively, variation in \(\gamma\) might be expected to decrease
-system stability by introducing a new source of variation into the
-system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\gamma\) varies are
-actually more likely to be stable, especially when \(S\) is high.
-
 \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the grey circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
-centred at \(\gamma = -0.05\). In a, the maximum real eigenvalue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 was stable when
-\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
-
-\includegraphics{unnamed-chunk-4-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-Rows in \(M\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(M\) where \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
-the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
-\(\gamma_{i} = 0.05\). This models one system in which \(\gamma\) is
-invariant and one in which \(\gamma\) varies, but systems are otherwise
-identical (note \(E[\gamma_{i}] = 1\) in both cases). I assume
-\(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
-set to \(-1\) and non-zero off-diagonal elements are drawn randomly from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
-\(\gamma_{i}\) to generate \(M\). When \(\gamma_{i} = 1\), eigenvalues
-of \(M\) are distributed uniformly within a circle centred at
-(\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real components
-of eigenvalues are highly unlikely to all be negative when all
-\(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated into
-two groups, eigenvalues are no longer uniformly distributed (Fig. 1b).
-Instead, two distinct clusters of eigenvalues appear (grey circles in
-Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
-(\(-0.05, 0\)). The former has a large radius, but the real components
-have shifted to the left (in comparison to when \(\gamma = 1\)) and all
-\(\Re({\lambda}) < 0\). The latter cluster has real components that have
-shifted to the right, but has a smaller radius. Overall, for 1 million
-randomly assembled \(M\), this division between slow and fast component
-response rates results in more stable systems: 1 stable given
-\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
-Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
-the distribution's centre (-1, 0). Black and red elipses in both panels
-show the circle centred on the distribution in panels a and b,
-respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.548 and 0.552 for a and b, respectively.
-
-\includegraphics{unnamed-chunk-7-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I now
-focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
-(see \hyperlink{SIstart}{Supplementary Information} for other distributions, 
-which give
-similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
-1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(M\) before and after variation in component response
-rate. Figure 2 shows a comparison of eigenvalue distributions given
-\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues, some of which are clustered tightly around the centre of
-the distribution, but others of which are spread outside the former
-radius of 12.649 (red circle Fig 2b). This larger radius occurs because
-the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
-\(M\). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold of \(\sigma\sqrt{SC} > 1\).
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate(\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\)) given \(C = 1\), and the proportion of systems in which
-variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Light and dark grey bars show the number of stable systems in the
-absence and presence of \(Var(\gamma)\), respectively. The black line
-shows the proportion of systems that are stable when
-\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-9-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-To investigate the effect of \(Var(\gamma)\) on system stability, I
-simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32. One million \(M\) were simulated 
-for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of stable 
-random systems was
-consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
-3), and that the difference between the probabilities of observing a
-stable system increased with an increase in \(S\); i.e., the potential
-for \(Var(\gamma)\) to drive stability increased with system complexity.
-For the highest values of \(S\), nearly all systems that were stable
-given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\) (see
-\hyperlink{SIstart}{Supplementary Information} for full results). This 
-suggests that the
-stability of large systems might be dependent upon variation in the
-response rates of their individual components, meaning that factors such
-as generation time (in ecological networks), transaction speed (in
-economic networks), or communication speed (in social networks) need to
-be considered when investigating the stability of complex systems.
-
-It is important to point out 
-that \(Var(\gamma)\) is not stabilising per se; that is,
-adding variation in \(\gamma\) to a particular system \(M\) does not
-necessarily increase the probability that the system will be stable (see
-\hyperlink{SIstart}{Supplementary Information}). Rather, systems that are 
-observed to be stable are more likely to vary in \(\gamma\), and for this
-\(Var(\gamma)\) to be critical to their stability. This is caused by the
-shift in the distribution of eigenvalues that occurs by introducing
-\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
-\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values.
-
-To further investigate the potential of \(Var(\gamma)\) to be stabilising,
-I used a genetic algorithm (the space of possible \(\gamma\)
-values was too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
-see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
-random \(M\), the
-genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
-performed on \(M\) using 
-each set of \(\gamma\) values, and the 20 sets with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
-subsequent mutation and crossover between the resulting new population
-of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
-\(M\) was found, 20 generations occurred, or a convergence criteria of
-minimum fitness increase between generations was satisfied. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were orders of
-magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
-Information}), meaning that stability could often be achieved by 
-manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
-Hence, managing the response rates of system components in a targetted
-way can potentially facilitate the stabilisation of complex systems 
-through a reduction in dimensionality.
-
-I have focused broadly on random complex systems, but it is also
-worthwhile to consider more restricted interactions such as those of
-specific ecological
-networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or \(i\) and
-\(j\) pairs have opposing signs (predator-prey networks). In
-general, competitive and mutualist networks tend to be destabilising,
-and predator-prey network tend to be
-stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
-When \(Var(\gamma)\) is applied to each, the proportion of stable
-competitive and predator-prey networks increases, but the proportion of
-stable mutualist networks does not (see 
-\hyperlink{SIstart}{Supplementary Information}).
-Additionally, when each component of \(M\) is interpreted as a unique
-species and given a random intrinsic growth
-rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
-feasibility is not increased by \(Var(\gamma)\), suggesting that
-variation in species generation time might be unlikely to drive
-stability in purely multi-species networks (see \hyperlink{SIstart}{Supplementary
-Information}).
-
-My results show that complex systems are more likely to be stable when
-the response rates of system components vary. These results are broadly
-applicable to complex biological and social networks.
-
-\textbf{Acknowledgements}
-
-I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L. Bussi\`ere and N. Bunnefeld improved the quality of
-this work.
-
-\textbf{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
-
-\hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
-\emph{Nature} \textbf{483,} 205--208 (2012).
-
-\hypertarget{ref-Gray2008}{}
-3. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
-random brain networks with a distribution of connection strengths.
-\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).
-
-\hypertarget{ref-Gray2009}{}
-4. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
-with excitatory and inhibitory connections. \emph{Neurocomputing}
-\textbf{72,} 1849--1858 (2009).
-
-\hypertarget{ref-Rosenfeld2009}{}
-5. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
-high-dimensional biochemical networks. \emph{Gene Regulation and Systems
-Biology} \textbf{3,} 1--10 (2009).
-
-\hypertarget{ref-MacArthur2010}{}
-6. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
-and criticality of adaptive regulatory networks. \emph{Physics Review
-Letters} \textbf{104,} 168701 (2010).
-
-\hypertarget{ref-Haldane2011}{}
-7. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
-\emph{Nature} \textbf{469,} 351--355 (2011).
-
-\hypertarget{ref-Suweis2014}{}
-8. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
-networks. \emph{PLoS ONE} \textbf{9,} (2014).
-
-\hypertarget{ref-Bardoscia2017}{}
-9. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G. Pathways
-towards instability in financial networks. \emph{Nature Communications}
-\textbf{8,} 1--7 (2017).
-
-\hypertarget{ref-Mougi2012}{}
-10. Mougi, A. \& Kondoh, M. Diversity of interaction types and
-ecological community stability. \emph{Science} \textbf{337,} 349--351
-(2012).
-
-\hypertarget{ref-Allesina2015a}{}
-11. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
-(2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Gao2016}{}
-12. Gao, J., Barzel, B. \& Barabási, A. L. Universal resilience patterns
-in complex networks. \emph{Nature} \textbf{530,} 307--312 (2016).
-
-\hypertarget{ref-Grilli2017}{}
-13. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
-
-\hypertarget{ref-Patel2018}{}
-14. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
-effects of eco-evolutionary feedbacks on community stability.
-\emph{American Naturalist} \textbf{191,} 1--29 (2018).
-
-\hypertarget{ref-Allesina2015}{}
-15. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
-(2015).
-
-\hypertarget{ref-May2008}{}
-16. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
-for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
-
-\hypertarget{ref-Tao2010}{}
-17. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
-(2010).
-
-\hypertarget{ref-Hamblin2013}{}
-18. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
-
-\hypertarget{ref-Allesina2011}{}
-19. Allesina, S. \& Levine, J. M. A competitive network theory of
-species diversity. \emph{Proceedings of the National Academy of Sciences
-of the United States of America} \textbf{108,} 5638--5642 (2011).
-
-\hypertarget{ref-Dougoud2018}{}
-20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
-
-\clearpage
-
-\begin{center}
-\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
-\end{center}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{This supplementary information supports the manuscript
+\textbf{This supplemental information supports the manuscript
 ``Component response rate variation drives stability in large complex
@@
 text, code, and data underlying this manuscript are publicly available
-on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}
-as part of the RandomMatrixStability package.}
+on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
+part of the RandomMatrixStability package.}
 
@@
 reproduced below with explanation, so it is possible to recreate all
-analyses using only this Supplementary Information.
+analyses using only this supplemental information.
 
@@
 
-\section{Supplementary Information table of
+\section{Supplemental Information table of
 contents}\label{supplemental-information-table-of-contents}
@@
 
-\includegraphics{unnamed-chunk-4-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-8-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\begin{longtable}[]{@{}lllllll@{}}
+\begin{longtable}[]{@{}llllll@{}}
 \toprule
-X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
 A1\_stabilised\tabularnewline
@@
 \endhead
-1 & 2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
-2 & 3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
-3 & 4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
-4 & 5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
-5 & 6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
-6 & 7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
-7 & 8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
-8 & 9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
-9 & 10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
-10 & 11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
-11 & 12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
-12 & 13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
-13 & 14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
-14 & 15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
-15 & 16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
-16 & 17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
-17 & 18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
-18 & 19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
-19 & 20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
-20 & 21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
-21 & 22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
-22 & 23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
-23 & 24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
-24 & 25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
-25 & 26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
-26 & 27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
-27 & 28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
-28 & 29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
-29 & 30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
-30 & 31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
-31 & 32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
-32 & 33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
-33 & 34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
+3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
+4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
+5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
+6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
+7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
+8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
+9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
+10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
+11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
+12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
+13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
+14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
+15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
+16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
+17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
+18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
+19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
+20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
+21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
+22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
+23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
+24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
+25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
+26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
+27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
+28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
+29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
+30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
+31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
+32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
+33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
+34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \bottomrule
@@
 \endhead
-1 & 2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
-2 & 3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
-3 & 4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
-4 & 5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
-5 & 6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
-6 & 7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
-7 & 8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-8 & 9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-9 & 10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-10 & 11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-11 & 12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
+3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
+4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
+5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
+6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
+7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
+8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}rrrrrrr@{}}
+\begin{longtable}[]{@{}rrrrrr@{}}
 \toprule
-X & N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
 A1\_stabilised\tabularnewline
@@
 \endhead
-1 & 2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-2 & 3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-3 & 4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-4 & 5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
-5 & 6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
-6 & 7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
-7 & 8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
-8 & 9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
-9 & 10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
-10 & 11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
-11 & 12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
-12 & 13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
-13 & 14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
-14 & 15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
-15 & 16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
-16 & 17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
-17 & 18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
-18 & 19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
-19 & 20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
-20 & 21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
-21 & 22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
-22 & 23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
-23 & 24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
-24 & 25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
-25 & 26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
-26 & 27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
-27 & 28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
-28 & 29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
-29 & 30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
-30 & 31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
-31 & 32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
-32 & 33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
-33 & 34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
-34 & 35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
-35 & 36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
-36 & 37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
-37 & 38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
-38 & 39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
-39 & 40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
-40 & 41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
-41 & 42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
-42 & 43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
-43 & 44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
-44 & 45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
-45 & 46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
-46 & 47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
-47 & 48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
-48 & 49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
-49 & 50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
-50 & 51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
-51 & 52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
-52 & 53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
-53 & 54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
-54 & 55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
-55 & 56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
-56 & 57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
-57 & 58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
-58 & 59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
-59 & 60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
-60 & 61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
-61 & 62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
-62 & 63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
-63 & 64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
-64 & 65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
-65 & 66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
-66 & 67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
-67 & 68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
-68 & 69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
-69 & 70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
-70 & 71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
-71 & 72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
-72 & 73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
-73 & 74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
-74 & 75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
-75 & 76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
-76 & 77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
-77 & 78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
-78 & 79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
-79 & 80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
-80 & 81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
-81 & 82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
-82 & 83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
-83 & 84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
-84 & 85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
-85 & 86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
-86 & 87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
-87 & 88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
-88 & 89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
-89 & 90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
-90 & 91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
-91 & 92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
-92 & 93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
-93 & 94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
-94 & 95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
-95 & 96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
-96 & 97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
-97 & 98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
-98 & 99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
-99 & 100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
+2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
+6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
+7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
+8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
+9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
+10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
+11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
+12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
+13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
+14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
+15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
+16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
+17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
+18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
+19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
+20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
+21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
+22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
+23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
+24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
+25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
+26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
+27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
+28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
+29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
+30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
+31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
+32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
+33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
+34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
+35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
+36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
+37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
+38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
+39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
+40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
+41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
+42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
+43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
+44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
+45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
+46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
+47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
+48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
+49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
+50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
+51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
+52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
+53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
+54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
+55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
+56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
+57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
+58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
+59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
+60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
+61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
+62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
+63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
+64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
+65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
+66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
+67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
+68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
+69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
+70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
+71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
+72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
+73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
+74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
+75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
+76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
+77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
+78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
+79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
+80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
+81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
+82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
+83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
+84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
+85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
+86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
+87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
+88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
+89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
+90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
+91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
+92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
+93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
+94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
+95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
+96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
+97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
+98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
+99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
+100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
 \bottomrule
@@
 
-\hypertarget{connectance}{\section{Different connectance (C)
+\hypertarget{connectance}{\section{Different inter-connectivity (C)
 values}\label{connectance}}
 
-In the main text, for simplicity, I assumed connectance values of
+In the main text, for simplicity, I assumed inter-connectivity values of
 \(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
@@
 
-\includegraphics{unnamed-chunk-23-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-31-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
 
@@
 diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the connectance of the community,
+\(C\) is the inter-connectivity (i.e., `connectance') of the community,
 which was set to \(C = 1\) throughout the manuscript and supplemental
@@
 \((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
-text and Supplementary Information). The above criteria is therefore
+text and supplemental information). The above criteria is therefore
 reduced to the below; note that the parenthetical in both equations

### 2018-06-08 notebook/SI.Rmd
Add further explanation of gamma based on May and Patel et al., thereby resolving Issue #1 in the Supplemental Information

--- 2018-06-03 notebook/SI.Rmd
@@
 
+- [Further explanation of $\gamma$](#moregamma)
 - [Code and simulations underlying Fig. 1](#Fig1) 
@@
 
-********************************************************************************
+Further explanation of $\gamma$ {#moregamma}
+================================================================================
+
+In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
+
+$$\frac{dN_{i}(t)}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+
+$$\frac{dx_{j}(t)}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+
+In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
+Patel et al.
+were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
+The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
+
+I use the same principle that Patel et al.
+use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
+Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+
+$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).$$
+
+In matrix notation[@May1973],
+
+$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).$$
+
+In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
+Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
 

### 2018-06-08 notebook/SI.tex
Add further explanation of gamma based on May and Patel et al., thereby resolving Issue #1 in the Supplemental Information

@@
 \item
+  \protect\hyperlink{moregamma}{Further explanation of \(\gamma\)}
+\item
   \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\hypertarget{moregamma}{\section{\texorpdfstring{Further explanation of
+\(\gamma\)}{Further explanation of \textbackslash{}gamma}}\label{moregamma}}
+
+In a synthesis of eco-evolutionary feedbacks on community stability,
+Patel et al. model a system that includes a vector of potentially
+changing species densities (\(\mathbf{N}\)) and a vector of potentially
+evolving traits
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
+For any species \(i\) or trait \(j\), change in species density
+(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+
+\[\frac{dN_{i}(t)}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
+
+\[\frac{dx_{j}(t)}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+
+In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
+effects of all species densities and trait values on the density of a
+species \(i\) and the value of trait \(j\), respectively. Patel et al.
+were interested in stability when the evolution of traits was relatively
+slow or fast in comparison with the change in species
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}, and
+this is modulated in the above by the scalar \(\epsilon\). The value of
+\(\epsilon\) thereby determines the timescale separation between ecology
+and evolution, with high \(\epsilon\) modelling relatively fast
+evolution and low \(\epsilon\) modelling relative slow
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
+
+I use the same principle that Patel et al. use to modulate the relative
+rate of evolution to modulate rates of component responses for \(S\)
+components. Following
+May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}},
+the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
+by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
+\(i\) (\(m_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
+
+\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).\]
+
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{3}},
+
+\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).\]
+
+In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
+correspond to individual component response rates. Therefore,
+\(\mathbf{\gamma} \mathbf{M}\) modulates the values of components and
+can be analysed using the techniques of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}}.
 
@@
 While the foundational work of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
+May\textsuperscript{\protect\hyperlink{ref-May1972}{2}} applies broadly
 to complex networks, much attention has been given specifically to
@@
 interactions between \(S\)
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Tang2014b}{4}}.
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}--\protect\hyperlink{ref-Tang2014b}{6}}.
 While May's original
-work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
+work\textsuperscript{\protect\hyperlink{ref-May1972}{2}} considered only
 randomly assembled communities, recent work has specifically looked at
@@
 is negative and \(j\) on \(i\) is positive, or vice
-versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Allesina2011}{5}}.
+versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}--\protect\hyperlink{ref-Allesina2011}{7}}.
 In general, competitor and mutualist networks tend to be unstable, while
@@
 following Allesina et
-al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
+al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}. To create
 these networks, I first generated a random matrix \(M\), then changed
@@
 Overall, as
-expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}},
 predator-prey communities are relatively stable while mutualist
@@
 mutualist communities are characterised by a very
-positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}
 leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
@@
 practical solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{6}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{8}}. In
 this case, the practical solution is finding vectors of \(\gamma\) that
@@
 the values of all system components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{7}--\protect\hyperlink{ref-Song2018}{9}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{9}--\protect\hyperlink{ref-Song2018}{11}}.
 This is of particular interest for ecological communities because
@@
 systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
 Consequently, the use of random matrices and traditional stability
@@
 networks has recently been
-criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
+criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
 While the key results in the main text are intended to be general to all
@@
 interact with
-evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{10}},
+evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}},
 abiotic, or social-economic systems. The relevance of \(\gamma\) for
@@
 
-Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}
+Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}
 define the following feasibility criteria for ecological systems
@@
 species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}} by
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{2}} stability
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Patel2018}{}
+1. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the effects
+of eco-evolutionary feedbacks on community stability. \emph{American
+Naturalist} \textbf{191,} 1--29 (2018).
+
 \hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
+2. May, R. M. Will a large complex system be stable? \emph{Nature}
 \textbf{238,} 413--414 (1972).
 
+\hypertarget{ref-May1973}{}
+3. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+\textbf{54,} 638--641 (1973).
+
 \hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+4. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
 \emph{Nature} \textbf{483,} 205--208 (2012).
@@
 \hypertarget{ref-Allesina2015a}{}
-3. Allesina, S. \& Tang, S. The stability--complexity relationship at
+5. Allesina, S. \& Tang, S. The stability--complexity relationship at
 age 40: a random matrix perspective. \emph{Population Ecology} 63--75
@@
 \hypertarget{ref-Tang2014b}{}
-4. Tang, S. \& Allesina, S. Reactivity and stability of large
+6. Tang, S. \& Allesina, S. Reactivity and stability of large
 ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
@@
 \hypertarget{ref-Allesina2011}{}
-5. Allesina, S. \& Levine, J. M. A competitive network theory of species
+7. Allesina, S. \& Levine, J. M. A competitive network theory of species
 diversity. \emph{Proceedings of the National Academy of Sciences of the
@@
 \hypertarget{ref-Hamblin2013}{}
-6. Hamblin, S. On the practical usage of genetic algorithms in ecology
+8. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
@@
 \hypertarget{ref-Grilli2017}{}
-7. Grilli, J. \emph{et al.} Feasibility and coexistence of large
+9. Grilli, J. \emph{et al.} Feasibility and coexistence of large
 ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
@@
 \hypertarget{ref-Dougoud2018}{}
-8. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+10. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-9. Song, C. \& Saavedra, S. Will a small randomly assembled community be
-feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
-
-\hypertarget{ref-Patel2018}{}
-10. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
-effects of eco-evolutionary feedbacks on community stability.
-\emph{American Naturalist} \textbf{191,} 1--29 (2018).
+11. Song, C. \& Saavedra, S. Will a small randomly assembled community
+be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
 

### 2018-06-08 notebook/arxiv/ms.tex
Add further explanation of gamma based on May and Patel et al., thereby resolving Issue #1 in the Supplemental Information

@@
 \endhead
-1 & 2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
-2 & 3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
-3 & 4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
-4 & 5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
-5 & 6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
-6 & 7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
-7 & 8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
-8 & 9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
-9 & 10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
-10 & 11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
-11 & 12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
-12 & 13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
-13 & 14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
-14 & 15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
-15 & 16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
-16 & 17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
-17 & 18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
-18 & 19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
-19 & 20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
-20 & 21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
-21 & 22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
-22 & 23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
-23 & 24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
-24 & 25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
-25 & 26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
-26 & 27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
-27 & 28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
-28 & 29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
-29 & 30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
-30 & 31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
-31 & 32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
-32 & 33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
-33 & 34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
+3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
+4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
+5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
+6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
+7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
+8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
+9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
+10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
+11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
+12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
+13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
+14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
+15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
+16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
+17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
+18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
+19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
+20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
+21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
+22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
+23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
+24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
+25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
+26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
+27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
+28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
+29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
+30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
+31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
+32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
+33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
+34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \bottomrule
@@
 \endhead
-1 & 2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
-2 & 3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
-3 & 4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
-4 & 5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
-5 & 6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
-6 & 7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
-7 & 8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-8 & 9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-9 & 10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-10 & 11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-11 & 12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
+3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
+4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
+5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
+6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
+7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
+8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-49 & 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
 \bottomrule
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Patel2018}{}
+1. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the effects
+of eco-evolutionary feedbacks on community stability. \emph{American
+Naturalist} \textbf{191,} 1--29 (2018).
+
+\hypertarget{refs}{}
 \hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
+2. May, R. M. Will a large complex system be stable? \emph{Nature}
 \textbf{238,} 413--414 (1972).
 
+\hypertarget{ref-May1973}{}
+3. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+\textbf{54,} 638--641 (1973).
+
 \hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+4. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
 \emph{Nature} \textbf{483,} 205--208 (2012).
@@
 \hypertarget{ref-Song2018}{}
-9. Song, C. \& Saavedra, S. Will a small randomly assembled community be
+11. Song, C. \& Saavedra, S. Will a small randomly assembled community be
 feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
 
-\hypertarget{ref-Patel2018}{}
-10. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
-effects of eco-evolutionary feedbacks on community stability.
-\emph{American Naturalist} \textbf{191,} 1--29 (2018).
-
 \end{document}

### 2018-06-11 notebook/arxiv/ms.tex
Fix notation
Fix a word in the abstract

@@
 I show that when components of a complex system respond to system
-dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
+perturbation at different rates (\(\boldsymbol{\gamma}\)), the potential for
 system stability is markedly increased. Stability is caused by the
@@
 
-\[\frac{dN_{i}(t)}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}(t)}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
+
+\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
 

### 2018-06-11 notebook/ms.Rmd
Fix a word in the abstract

@@
 output:
+  pdf_document:
+    fig_caption: yes
+    keep_tex: no
+  html_document: default
   word_document:
@@
     reference_docx: docx_template.docx
-  pdf_document:
-    fig_caption: yes
-    keep_tex: no
-  html_document: default
 bibliography: references.bib
@@
 But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem [@Allesina2015].
-Here I show that when components of a complex system respond to system dynamics at different rates ($\boldsymbol{\gamma}$), the potential for system stability is markedly increased.
+Here I show that when components of a complex system respond to system perturbation at different rates ($\boldsymbol{\gamma}$), the potential for system stability is markedly increased.
 Stability is caused by the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher interaction strength variation ($\boldsymbol{\sigma^{2}}$).

### 2018-06-12 notebook/ms.Rmd
Add acknowledgements

@@
 
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018].
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018] (see Supplementary Information for details).
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
@@
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supporting Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
 Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
@@
 
+**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
+Conversations with L.
+Bussi&#xe8;re and N.
+Bunnefeld, and helpful comments from J.
+J.
+Cusack and I.
+L.
+Jones, improved the quality of this work.
+
 ********************************************************************************

### 2018-06-12 notebook/ms_for_review.Rmd
Add file for review

--- 2018-06-12 notebook/ms.Rmd
@@
 title: "Component response rate variation drives stability in large complex systems"
-author: "A. Bradley Duthie"
+author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
 date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
-  FK9 4LA alexander.duthie@stir.ac.uk
+  FK9 4LA
 output:
@@
 - \bibliographystyle{amnatnat}
-linestretch: 1
+linestretch: 2
 link-citations: yes
@@
 ```
-
-********************************************************************************
 
@@
 
-********************************************************************************
+```{r, echo = FALSE}
+A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
+A1 <- as.matrix(A1[,-1]);
+
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(t(A0_vm));
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(t(A1_vm));
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+fhalf       <- 1:(0.5*length(A1vec));
+shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st       <- pr_st[,-1];
+```
+
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018] (see Supplementary Information for details).
+The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
+This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
+I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
+Rows are then multiplied by $\gamma_{i}$ to generate $M$.
+When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+1a).
+Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
+But when $\gamma_{i}$ values are separated into two groups, eigenvalues are no longer uniformly distributed (Fig.
+1b).
+Instead, two distinct clusters of eigenvalues appear (grey circles in Fig.
+1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
+The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$) and all $\Re({\lambda}) < 0$.
+The latter cluster has real components that have shifted to the right, but has a smaller radius.
+Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+
+```{r echo = FALSE}
+S     <- 1000;
+C     <- 1;
+sigma <- 0.4;
+```
+
+```{r, echo = FALSE}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+gammas <- runif(n = 1000, min = 0, max = 2);
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+```
+
+Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
+I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
+As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
+1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
+Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered tightly around the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (red circle Fig 2b).
+This larger radius occurs because the addition of $Var(\gamma)$ increases the realised $\sigma$ of $M$.
+The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
+But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold of $\sigma\sqrt{SC} > 1$.
+
+```{r, echo = FALSE}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+```
+
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
+One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
+I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
+This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
+
+It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
+Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
+This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
+1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
+
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
+For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
+Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+
+I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
+These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
+In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
+When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
+Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
+
+My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
+
+**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
+Conversations with L.
+Bussi&#xe8;re and N.
+Bunnefeld, and helpful comments from J.
+J.
+Cusack and I.
+L.
+Jones, improved the quality of this work.
+
+**References**
+
+<div id="refs"></div>
+
+\clearpage
+
+```{r echo = FALSE}
+S     <- 200;
+C     <- 0.05;
+sigma <- 0.4;
+pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st <- pr_st[,-1];
+```
 
@@
 
-********************************************************************************
-
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018] (see Supplementary Information for details).
-The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
-An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
-This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
-I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
-Rows are then multiplied by $\gamma_{i}$ to generate $M$.
-When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
-1a).
-Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
-But when $\gamma_{i}$ values are separated into two groups, eigenvalues are no longer uniformly distributed (Fig.
-1b).
-Instead, two distinct clusters of eigenvalues appear (grey circles in Fig.
-1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
-The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$) and all $\Re({\lambda}) < 0$.
-The latter cluster has real components that have shifted to the right, but has a smaller radius.
-Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+\clearpage
 
@@
 
-********************************************************************************
-
-Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
-As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
-1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
-Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered tightly around the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (red circle Fig 2b).
-This larger radius occurs because the addition of $Var(\gamma)$ increases the realised $\sigma$ of $M$.
-The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
-But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold of $\sigma\sqrt{SC} > 1$.
+\clearpage
 
@@
 ```
-
-********************************************************************************
-
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
-One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
-I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
-3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
-This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
-
-It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
-Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
-This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
-For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
-Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
-
-I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
-These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
-In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
-When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
-
-My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
-
-**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L.
-Bussi&#xe8;re and N.
-Bunnefeld, and helpful comments from J.
-J.
-Cusack and I.
-L.
-Jones, improved the quality of this work.
-
-********************************************************************************
-
-**References**
-
-<div id="refs"></div>

### 2018-06-14 notebook/arxiv/ms.tex
Add acknowledgements
Thorough edits

@@
 shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
+\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only 
+\(Var(\gamma)\) differs between panels). \textbf{a.}
+Eigenvalues plotted when all \(\gamma = 1\); points are
 uniformly distributed within the grey circle with a radius of
@@
 \(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
+large circle with a radius of \(\sigma\sqrt{SC} =\) 1.718 centred at
+\(\gamma = -1.95\) and one small circle with a radius of
+\(\sigma\sqrt{SC} =\) 0.044
 centred at \(\gamma = -0.05\). In a, the maximum real eigenvalue
@@
 value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}} (see 
+\hyperlink{SIstart}{Supplementary Information} for details).
 The distribution of \(\gamma\) over \(S\) components thereby models the
 distribution of component response rates. An instructive example
-compares one \(M\) where \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
+compares one \(M\) when \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
 the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
@@
 introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
+(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel shows the same
 system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
@@
 \textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate(\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\)) given \(C = 1\), and the proportion of systems in which
-variation in \(\gamma\) is critical for system stability. For each
+variation in component response rate(\(\boldsymbol{\gamma}\)).} Y-axes show
+the \(\ln\) number of systems that are stable across different system sizes
+(\(S\)) given \(C = 1\) (left axis), and the proportion of systems in which
+variation in \(\gamma\) is critical for system stability (right axis). For each
 \(S\), 1 million complex systems are randomly generated. Stability of
@@
 Light and dark grey bars show the number of stable systems in the
-absence and presence of \(Var(\gamma)\), respectively. The black line
+absence versus presence of \(Var(\gamma)\), respectively. The black line
 shows the proportion of systems that are stable when
@@
 \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were orders of
+resulted in stable systems with probabilities that were up to four orders of
 magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
@@
 general, competitive and mutualist networks tend to be destabilising,
-and predator-prey network tend to be
+and predator-prey networks tend to be
 stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
@@
 I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L. Bussi\`ere and N. Bunnefeld improved the quality of
-this work.
+Conversations with L. Bussi\`ere and N. Bunnefeld, and helpful comments from
+J. J. Cusack and I. L. Jones, improved the quality of this work.
 
@@
 34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} \tabularnewline
 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
@@
 12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} \tabularnewline
 50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline

### 2018-06-14 notebook/ms.Rmd
Thorough edits

@@
 
-**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
+**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only $Var(\gamma)$ differs between panels).
+**a.** Eigenvalues plotted when all $\gamma = 1$; points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
 In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
@@
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
-An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
+An instructive example compares one $M$ when $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
 This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
@@
 
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel shows the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
 **a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
@@
 
-**Figure 3: Stability of large complex systems with and without variation in component response rate($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 3: Stability of large complex systems with and without variation in component response rate($\boldsymbol{\gamma}$).** Y-axes show the $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$ (left axis), and the proportion of systems in which variation in $\gamma$ is critical for system stability (right axis).
 For each $S$, 1 million complex systems are randomly generated.
@@
 Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Light and dark grey bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
+Light and dark grey bars show the number of stable systems in the absence versus presence of $Var(\gamma)$, respectively.
 The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
@@
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
 Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
@@
 These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
-In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
+In general, competitive and mutualist networks tend to be destabilising, and predator-prey networks tend to be stabilising[@Allesina2011].
 When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).

### 2018-06-14 notebook/SI.Rmd
Fix notation
Additional files

@@
 The [RandomMatrixStability package](https://github.com/bradduthie/RandomMatrixStability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for functions as part of the package.
-The RandomMatrixStability package package is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
+The RandomMatrixStability package is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
 
@@
 
-$$\frac{dN_{i}(t)}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
 
-$$\frac{dx_{j}(t)}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
 
@@
 
-```{r, echo = FALSE, fig.height = 5, fig.width = 8}
+```{r, echo = FALSE}
 A0 <- read.csv("sim_results/bi_gamma/S200_A0.csv");
 A1 <- read.csv("sim_results/bi_gamma/S200_A1.csv");
+```
+
+```{r, echo = TRUE, fig.height = 5, fig.width = 8}
 A0 <- as.matrix(A0[,-1]);
@@
 
-Some stability results from this table are shown below. Each histogram shows a different distribution of $\gamma$ that was found to be stabilising.
+Some stability results from this table are shown below.
+Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
 

### 2018-06-14 notebook/SI.tex
Fix notation
Additional files

@@
 is also provided for functions as part of the package. The
-RandomMatrixStability package package is available on
+RandomMatrixStability package is available on
 \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{A0 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A0[,-}\DecValTok{1}\NormalTok{]);}
+\NormalTok{A1 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A1[,-}\DecValTok{1}\NormalTok{]);}
+\NormalTok{plot_Fig_1 <-}\StringTok{ }\NormalTok{function(A0, A1)\{}
+    \NormalTok{S_val       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A0)[}\DecValTok{1}\NormalTok{];}
+    \NormalTok{A0_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
+    \NormalTok{A0_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
+    \NormalTok{A0_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
+    \NormalTok{A1_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
+    \NormalTok{A1_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
+    \NormalTok{A1_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
+    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
+    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A0_vm));}
+    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
+    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A1_vm));}
+    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \NormalTok{fhalf       <-}\StringTok{ }\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec));}
+    \NormalTok{shalf       <-}\StringTok{ }\NormalTok{(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec)+}\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(A1vec);}
+    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
+    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{vl   <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
+    \NormalTok{A0x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
+    \NormalTok{A0y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x0, }\DataTypeTok{y =} \NormalTok{A0y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(A0_r, A0_i, }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);}
+    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }
+         \DataTypeTok{col =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
+    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
+    \NormalTok{A0x1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*S_val)]);}
+    \NormalTok{A0y1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1a, }\DataTypeTok{y =} \NormalTok{A0y1a, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \NormalTok{A0x1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }
+\StringTok{                  }\KeywordTok{mean}\NormalTok{( }\KeywordTok{diag}\NormalTok{(A1)[( (}\FloatTok{0.5}\NormalTok{*S_val) +}\StringTok{ }\DecValTok{1} \NormalTok{):S_val] );}
+    \NormalTok{A0y1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1b, }\DataTypeTok{y =} \NormalTok{A0y1b, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(A1_r[}\DecValTok{1}\NormalTok{:S_val], A1_i[}\DecValTok{1}\NormalTok{:S_val],}\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);   }
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+\NormalTok{\}}
+\KeywordTok{plot_Fig_1}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1);}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
 
@@
 \end{document}
+\grid

### 2018-06-14 notebook/arxiv/ms.tex
Additional files

@@
 text is available on
-\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
+\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/notebook/sim_results/bi_gamma}{GitHub},
 and was produced by running the following function.
@@
 plot itself can be recreated with the function and code below.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{A0 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A0[,-}\DecValTok{1}\NormalTok{]);}
+\NormalTok{A1 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A1[,-}\DecValTok{1}\NormalTok{]);}
+\NormalTok{plot_Fig_1 <-}\StringTok{ }\NormalTok{function(A0, A1)\{}
+    \NormalTok{S_val       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A0)[}\DecValTok{1}\NormalTok{];}
+    \NormalTok{A0_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
+    \NormalTok{A0_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
+    \NormalTok{A0_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
+    \NormalTok{A1_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
+    \NormalTok{A1_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
+    \NormalTok{A1_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
+    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
+    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A0_vm));}
+    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
+    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
+    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A1_vm));}
+    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
+    \NormalTok{fhalf       <-}\StringTok{ }\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec));}
+    \NormalTok{shalf       <-}\StringTok{ }\NormalTok{(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec)+}\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(A1vec);}
+    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
+    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{vl   <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
+    \NormalTok{A0x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
+    \NormalTok{A0y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x0, }\DataTypeTok{y =} \NormalTok{A0y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(A0_r, A0_i, }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);}
+    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }
+         \DataTypeTok{col =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
+    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
+    \NormalTok{A0x1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*S_val)]);}
+    \NormalTok{A0y1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1a, }\DataTypeTok{y =} \NormalTok{A0y1a, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \NormalTok{A0x1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }
+\StringTok{                  }\KeywordTok{mean}\NormalTok{( }\KeywordTok{diag}\NormalTok{(A1)[( (}\FloatTok{0.5}\NormalTok{*S_val) +}\StringTok{ }\DecValTok{1} \NormalTok{):S_val] );}
+    \NormalTok{A0y1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1b, }\DataTypeTok{y =} \NormalTok{A0y1b, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(A1_r[}\DecValTok{1}\NormalTok{:S_val], A1_i[}\DecValTok{1}\NormalTok{:S_val],}\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);   }
+    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
+\NormalTok{\}}
+\KeywordTok{plot_Fig_1}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1);}
+\end{Highlighting}
+\end{Shaded}
 
@@
 reported in the main text and legend of Fig. 1 (raw data are
-\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
+\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/notebook/sim_results/bi_gamma/bi_pr_st.csv}{available
 on GitHub}).
@@
 
-Some stability results from this table are shown below. Each histogram
-shows a different distribution of \(\gamma\) that was found to be
-stabilising.
+Some stability results from this table are shown below. Results for \texttt{A0}
+indicate systems in which \(\gamma = 1\), while \texttt{A1} refers to systems in 
+which the genetic algorithm searched for a set of \(\gamma\) values that 
+stabilised the system.
 

### 2018-06-18 R/ga.R
Change the output function

@@ -160,30 +160,26 @@ crossover <- function(inds, pr = 0.1){
 #'sim_ga      <- Evo_rand_gen_var(from = 2, to = 2, iters = 1);
 #'sum_rand_ga <- summarise_randmat_ga(sim_ga$tot_res, sim_ga$fea_res);
 #'@export
-summarise_randmat_ga <- function(tot_res, fea_res){
+summarise_randmat <- function(tot_res, fea_res){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 10);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 7);
     for(i in 1:sims){
-        unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
-        stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
-        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
+        A0_unst   <- tot_res[[i]][,1] == FALSE;
+        A0_stbl   <- tot_res[[i]][,1] == TRUE;
+        A1_unst   <- tot_res[[i]][,2] == FALSE;
+        A1_stbl   <- tot_res[[i]][,2] == TRUE;
         stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
-        non_feas  <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == FALSE;
-        feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
-        unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
-        feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
-        foundd    <- tot_res[[i]][,3] == TRUE;
+        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
         all_res[i, 1]  <- i + 1;
-        all_res[i, 2]  <- sum(unstables);
-        all_res[i, 3]  <- sum(stables);
-        all_res[i, 4]  <- sum(unstabled);
-        all_res[i, 5]  <- sum(stabled);
-        all_res[i, 6]  <- sum(non_feas);
-        all_res[i, 7]  <- sum(feasibl);
[…]

### 2018-06-18 notebook/SI.Rmd
Change the wording in the SI

@@
 ```{r}
-summarise_randmat_ga <- function(tot_res, fea_res){
+summarise_randmat <- function(tot_res, fea_res){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 10);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 7);
     for(i in 1:sims){
-        unstables <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == FALSE;
-        stables   <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == TRUE;
+        A0_unst   <- tot_res[[i]][,1] == FALSE;
+        A0_stbl   <- tot_res[[i]][,1] == TRUE;
+        A1_unst   <- tot_res[[i]][,2] == FALSE;
+        A1_stbl   <- tot_res[[i]][,2] == TRUE;
+        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
         unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
-        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
-        non_feas  <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == FALSE;
-        feasibl   <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == TRUE;
-        unfeased  <- fea_res[[i]][,1] == TRUE  & fea_res[[i]][,2] == FALSE;
-        feased    <- fea_res[[i]][,1] == FALSE & fea_res[[i]][,2] == TRUE;
-        foundd    <- tot_res[[i]][,3] == TRUE;
         all_res[i, 1]  <- i + 1;
-        all_res[i, 2]  <- sum(unstables);
-        all_res[i, 3]  <- sum(stables);
-        all_res[i, 4]  <- sum(unstabled);
-        all_res[i, 5]  <- sum(stabled);
-        all_res[i, 6]  <- sum(non_feas);
-        all_res[i, 7]  <- sum(feasibl);
-        all_res[i, 8]  <- sum(unfeased);
-        all_res[i, 9]  <- sum(feased);
-        all_res[i, 10] <- sum(foundd);
+        all_res[i, 2]  <- sum(A0_unst);
+        all_res[i, 3]  <- sum(A0_stbl);
+        all_res[i, 4]  <- sum(A1_unst);
+        all_res[i, 5]  <- sum(A1_stbl);
+        all_res[i, 6]  <- sum(stabled);
+        all_res[i, 7]  <- sum(unstabled);
     }
+    colnames(all_res) <- c("N", "A0_unstable", "A0_stable", "A1_unstable",
+                           "A1_stable", "A1_stabilised", "A1_destabilised");
     return(all_res);
@@
 
-Some stability results from this table are shown below.
+Stability results for 40000 $M$ for each $S$ from 2-40 are shown below.
 Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.

### 2018-06-18 notebook/SI.tex
Add updated (and corrected) table to SI

@@
 \begin{Highlighting}[]
-\NormalTok{summarise_randmat_ga <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
+\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
     \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{10}\NormalTok{);}
+    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
     \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{unstables <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{stables   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{A0_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{A0_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{A1_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
+        \NormalTok{A1_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
+        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
         \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{non_feas  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{feasibl   <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{unfeased  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{feased    <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{foundd    <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{3}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
         \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstables);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stables);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(non_feas);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feasibl);}
-        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unfeased);}
-        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feased);}
-        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(foundd);}
-    \NormalTok{\}}
+        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_unst);}
+        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_stbl);}
+        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_unst);}
+        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_stbl);}
+        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
+        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
+    \NormalTok{\}}
+    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{,}
+                           \StringTok{"A1_stable"}\NormalTok{, }\StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{);}
     \KeywordTok{return}\NormalTok{(all_res);}
@@
 
-Some stability results from this table are shown below. Each histogram
-shows a different distribution of \(\gamma\) that was found to be
-stabilising.
+Stability results for 40000 \(M\) for each \(S\) from 2-40 are shown
+below. Results for \texttt{A0} indicate systems in which \(\gamma = 1\),
+while \texttt{A1} refers to systems in which the genetic algorithm
+searched for a set of \(\gamma\) values that stabilised the system.
 
@@
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
 A1\_stabilised & A1\_destabilised\tabularnewline
@@
 \endhead
-2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
-3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
-4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
-5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
-6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
-7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
-8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
-9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
-10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
-11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
-12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
-13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
-14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
-15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
-16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
-17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
-18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
-19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
-20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
-21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
-22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
-23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
-24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
-25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
-26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
-27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
-28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
-29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
-30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
-31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
-32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
-33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
-34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
-35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
-36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
-37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
-38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
-39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
-40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
+2 & 13 & 39987 & 13 & 39987 & 0 & 0\tabularnewline
+3 & 157 & 39843 & 157 & 39843 & 0 & 0\tabularnewline
+4 & 563 & 39437 & 563 & 39437 & 0 & 0\tabularnewline
+5 & 1588 & 38412 & 1581 & 38419 & 7 & 0\tabularnewline
+6 & 3268 & 36732 & 3219 & 36781 & 49 & 0\tabularnewline
+7 & 5370 & 34630 & 5196 & 34804 & 175 & 1\tabularnewline
+8 & 8231 & 31769 & 7604 & 32396 & 627 & 0\tabularnewline
+9 & 11704 & 28296 & 10181 & 29819 & 1523 & 0\tabularnewline
+10 & 15215 & 24785 & 12303 & 27697 & 2912 & 0\tabularnewline
+11 & 19529 & 20471 & 14599 & 25401 & 4930 & 0\tabularnewline
+12 & 23480 & 16520 & 16210 & 23790 & 7270 & 0\tabularnewline
+13 & 27570 & 12430 & 17800 & 22200 & 9770 & 0\tabularnewline
+14 & 31096 & 8904 & 18899 & 21101 & 12198 & 1\tabularnewline
+15 & 34036 & 5964 & 19948 & 20052 & 14089 & 1\tabularnewline
+16 & 36231 & 3769 & 21000 & 19000 & 15231 & 0\tabularnewline
+17 & 37737 & 2263 & 21693 & 18307 & 16044 & 0\tabularnewline
+18 & 38808 & 1192 & 22291 & 17709 & 16518 & 1\tabularnewline
+19 & 39391 & 609 & 23221 & 16779 & 16170 & 0\tabularnewline
+20 & 39711 & 289 & 23943 & 16057 & 15768 & 0\tabularnewline
+21 & 39867 & 133 & 25361 & 14639 & 14506 & 0\tabularnewline
+22 & 39950 & 50 & 26584 & 13416 & 13366 & 0\tabularnewline
+23 & 39982 & 18 & 28162 & 11838 & 11820 & 0\tabularnewline
+24 & 39994 & 6 & 29660 & 10340 & 10334 & 0\tabularnewline
+25 & 39997 & 3 & 31405 & 8595 & 8592 & 0\tabularnewline
+26 & 39999 & 1 & 33057 & 6943 & 6942 & 0\tabularnewline
+27 & 40000 & 0 & 34602 & 5398 & 5398 & 0\tabularnewline
+28 & 40000 & 0 & 35957 & 4043 & 4043 & 0\tabularnewline
+29 & 40000 & 0 & 37154 & 2846 & 2846 & 0\tabularnewline
+30 & 40000 & 0 & 38104 & 1896 & 1896 & 0\tabularnewline
+31 & 40000 & 0 & 38722 & 1278 & 1278 & 0\tabularnewline
+32 & 40000 & 0 & 39259 & 741 & 741 & 0\tabularnewline
+33 & 40000 & 0 & 39551 & 449 & 449 & 0\tabularnewline
+34 & 40000 & 0 & 39742 & 258 & 258 & 0\tabularnewline
+35 & 40000 & 0 & 39879 & 121 & 121 & 0\tabularnewline
+36 & 40000 & 0 & 39929 & 71 & 71 & 0\tabularnewline
+37 & 40000 & 0 & 39967 & 33 & 33 & 0\tabularnewline
+38 & 40000 & 0 & 39989 & 11 & 11 & 0\tabularnewline
+39 & 40000 & 0 & 39994 & 6 & 6 & 0\tabularnewline
+40 & 40000 & 0 & 39996 & 4 & 4 & 0\tabularnewline
 \bottomrule
@@
 \end{document}
-\grid

### 2018-06-18 notebook/arxiv/ms.tex
Change the wording in the SI
Add updated (and corrected) table to SI
Update the simulation number total

@@
 exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
-see \hyperlink{SIstart}{Supplementary Information}). For each of 10000 
+see \hyperlink{SIstart}{Supplementary Information}). For each of 40000 
 random \(M\), the
@@
 
-Some stability results from this table are shown below. Results for \texttt{A0}
+Stability results for 40000 \(M\) for each \(S\) from 2-40 are shown below. Results for \texttt{A0}
 indicate systems in which \(\gamma = 1\), while \texttt{A1} refers to systems in 

### 2018-06-18 notebook/arxiv/ms.tex
Updated website and small edits to manuscript

@@
 
-The above stability criterion assumes that individual components respond
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}} stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual components respond
 to perturbations of the system at the same rate (\(\gamma\)), but this
@@
 see \hyperlink{SIstart}{Supplementary Information}). For each of 40000 
-random \(M\), the
+random \(M\) (\(\sigma = 0.4\), \(C = 1\)), the
 genetic algorithm initialised 1000 different sets of

### 2018-06-18 notebook/ms.Rmd
Update the simulation number total
Updated website and small edits to manuscript

@@
 
-The above stability criterion assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+May's[@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
@@
 To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
-For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
+For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
 Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.

### 2018-06-18 notebook/ms_for_review.Rmd
Add cover letter, minor edit

@@
 linestretch: 2
-link-citations: yes
+link-citations: no
 linkcolor: blue
@@
 <!---
-Current word count: 1672
+Abstract current word count: 225
+Text current word count: 1453
 --->
@@
 Bussi&#xe8;re and N.
-Bunnefeld, and helpful comments from J.
+Bunnefeld, and comments from J.
 J.

### 2018-06-18 notebook/arxiv/ms.tex
Minor changes

@@
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold of \(\sigma\sqrt{SC} > 1\).
+threshold \(\sigma\sqrt{SC} < 1\).
 

### 2018-06-18 notebook/ms_for_review.Rmd
Minor changes

@@
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
-But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold of $\sigma\sqrt{SC} > 1$.
+But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
 

### 2018-07-04 notebook/ms_for_review.Rmd
Add changes for a science report

@@
 output:
+  word_document:
+    fig_caption: yes
+    pandoc_args:
+    - --csl
+    - science.csl
+    reference_docx: docx_template.docx
   pdf_document:
@@
   html_document: default
-  word_document:
-    fig_caption: yes
-    pandoc_args:
-    - --csl
-    - evolution.csl
-    reference_docx: docx_template.docx
 bibliography: references.bib

### 2018-07-05 notebook/ms_for_review.Rmd
Adjustments for Science

@@
 linkcolor: blue
-csl: nature.csl
+csl: science.csl
 biblio-style: apalike
@@
 
+**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counter-intuitive result of widespread importance in the physical, life, and social sciences.
+Nevertheless, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem.
+Here I vary the component response rates ($\boldsymbol{\gamma}$) of randomly generated complex systems.
+I show that when component response rates vary, the potential for system stability is markedly increased.
+Variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, such that the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
+My results reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.**
+
+<!---
+
 **The stability of a complex system generally decreases with increasing system size, as is demonstrated by random matrix theory [@May1972; @Allesina2012].
@@
 
+--->
+
 In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations.
-Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$)[@Allesina2012].
-May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ has profound consequences across multiple disciplines, raising the question of how complex systems in, e.g., ecology[@Mougi2012; @Allesina2012; @Allesina2015; @Grilli2017] or banking [@May2008; @Haldane2011; @Bardoscia2017] are predicted to persist or change.
+Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$) [@Allesina2012].
+May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks.
+As such, identifying general principles that drive stability in complex systems is of wide-ranging importance.
 
 Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., species[@Allesina2012] or banks[@Haldane2011]).
-One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium[@Allesina2012].
+One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
 Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
 Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$)[@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly[@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$[@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
 Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma^{2}$ increase.
 
-May's[@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
@@
 A_mat  <- matrix(data = A_dat, nrow = 1000);
-C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
 C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
 A_mat     <- A_mat * C_mat;
-gammas <- runif(n = 1000, min = 0, max = 2);
+A_mat  <- species_interactions(A_mat, type = 0);
+gammas <- c(rep(0.05, 500), rep(1.95, 500));
+
 mu_gam <- mean(gammas);
@@
 diag(A1_vm) <- NA;
-A1vec       <- as.vector(A1_vm);
+A1vec       <- as.vector(A1_vm[1:500,]);
 A1vec       <- A1vec[is.na(A1vec) == FALSE];
@@
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
-But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's[@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
+But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's [@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
 
@@
 
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively [@Hamblin2013]; see Supplementary Information).
 For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
@@
 When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
+Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
 
-My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
+My results show that complex systems are more likely to be stable when the response rates of system components vary.
+These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
 
@@
 **a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
 In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
@@
 A_mat  <- matrix(data = A_dat, nrow = 1000);
-C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
 C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
 A_mat     <- A_mat * C_mat;
+A_mat  <- species_interactions(A_mat, type = 0);
 gammas <- runif(n = 1000, min = 0, max = 2);
+
 mu_gam <- mean(gammas);
@@
 A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(A1_vm);
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
 ```
 
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
-Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
+Black elipses in both panels show the circle centred on the distribution in panel a.
 Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
@@
 par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
      xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
@@
 y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
 text(x = -15.5, y = 19, labels = "a", cex = 2);
 points(x = x0, y = y0, type = "l", lwd = 3);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
      xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
@@
 text(x = -15.5, y = 19, labels = "b", cex = 2);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid");
 mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);

### 2018-07-05 notebook/ms_for_review.Rmd
Update manuscript for Science

@@
 
-**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counter-intuitive result of widespread importance in the physical, life, and social sciences.
-Nevertheless, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem.
+**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counter-intuitive result of widespread importance across the physical, life, and social sciences.
+Despite recent interest in the relationship between system properties and stability, the effect of variation in the response rate of individual system components remains unconsidered.
 Here I vary the component response rates ($\boldsymbol{\gamma}$) of randomly generated complex systems.
@@
 
-Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., species[@Allesina2012] or banks[@Haldane2011]).
+Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., networks of species [@Allesina2012] or banks [@Haldane2011]).
 One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
 Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
+Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1 [@May1972; @Allesina2012; @Allesina2015].
 Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
@@
 Here I show why, despite higher $\sigma$, complex systems in which $\gamma$ varies are actually more likely to be stable, especially when $S$ is high.
+
+<!---
+
+Introduction could end here.
+
+--->
 
@@
 
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$[@Patel2018] (see Supplementary Information for details).
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018] (see Supplementary Information for details).
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
@@
 S     <- 1000;
-C     <- 1;
+C     <- 0.05;
 sigma <- 0.4;
@@
 1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
-Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 1$, and $\sigma = 0.4$.
+Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
 As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered tightly around the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (red circle Fig 2b).
-This larger radius occurs because the addition of $Var(\gamma)$ increases the realised $\sigma$ of $M$.
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b) [see also ].
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
@@
 One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
-I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
 3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
@@
 
-I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
+I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks [@Allesina2012].
 These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
-In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
+In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising [@Allesina2011].
 When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
@@
 These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
+
+<!---
+
+If more discussion is needed, then talk about relation to Jirsa and Ding 2004.
+
+--->
 

### 2018-07-24 notebook/SI_science.Rmd
Last commit before shifting to natcom

--- 2018-06-18 notebook/SI.Rmd
@@
 In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
-model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$) (15).
 For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
@@
 Patel et al.
-were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
-The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
+were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities (15), and this is modulated in the above by the scalar $\epsilon$.
+The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution (15).
 
@@
 use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
-Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+Following May (1, 19), the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
 
@@
 
-In matrix notation[@May1973],
+In matrix notation (20),
 
@@
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
+Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May (1, 19).
 
@@
 
-```{r, echo = TRUE, fig.height = 5, fig.width = 8}
+```{r, echo = TRUE}
 A0 <- as.matrix(A0[,-1]);
@@
     A_mat  <- matrix(data = A_dat, nrow = 1000);
-    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
     C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
     A_mat     <- A_mat * C_mat;
+    A_mat  <- species_interactions(A_mat, type = 0);
     gammas <- runif(n = 1000, min = 0, max = 2);
@@
     A0vec       <- A0vec[is.na(A0vec) == FALSE];
-    A1_vm       <- A1;
-    diag(A1_vm) <- NA;
-    A1vec       <- as.vector(A1_vm);
-    A1vec       <- A1vec[is.na(A1vec) == FALSE];
     par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
-         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
-         asp = 1);
+    plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
     vl <- seq(from = 0, to = 2*pi, by = 0.001);
@@
     y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
     text(x = -15.5, y = 19, labels = "a", cex = 2);
     points(x = x0, y = y0, type = "l", lwd = 3);
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
+    plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
          xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
@@
     text(x = -15.5, y = 19, labels = "b", cex = 2);
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
-    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+    points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid");
     mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
@@
 
-While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
+While the foundational work of May (1) applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
 In these networks, the matrix $M$ is interpreted as a community matrix and each row and column is interpreted as a single species.
-The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $M$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015a; @Tang2014b].
-While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $M$ are negative), mutualist networks (all off-diagonal elements of $M$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2011; @Allesina2012; @Allesina2015a; @Tang2014b].
+The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $M$ holds the effects of pair-wise interactions between $S$ species (2, 4, 20).
+While May's original work (1) considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $M$ are negative), mutualist networks (all off-diagonal elements of $M$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa) (2, 4, 17, 20).
 In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.
 
-I investigated competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012].
+I investigated competitor, mutualist, and predator-prey networks following Allesina et al.
+(2).
 To create these networks, I first generated a random matrix $M$, then changed the elements of $M$ accordingly.
@@
 
-Overall, as expected[@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable.
+Overall, as expected (2), predator-prey communities are relatively stable while mutualist communties are highly unstable.
 But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
-This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
+This is unsurprising because purely mutualist communities are characterised by a very positive (2) leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
 
@@
 This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
-Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution (16).
 In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
@@
 For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
-Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
-Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
+Feasibile equilibria assume that the values of all system components are positive at equilibrium (5, 18, 21).
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic (18).
+Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised (18).
 While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
 This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
-Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary (15), abiotic, or social-economic systems.
 The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
 
-Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
+Dougoud et al. (18) define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
 
@@
 As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
-The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
+The variable $C$ is the connectance of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
 The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
 In the main text, implicitly, $\delta = 0$ underlying strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
+Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's (1) stability criterion (19) by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
 The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
@@
 Hence, each component $i$ of the complex system $M$ is assumed to be a species with a growth rate of $r_{i}$.
-Note that negative intrinsic growth rates are not unrealistic, and will occur in obligate mutualists in the absence of a partner.
 
@@
 Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks.
-
-References
-================================================================================

### 2018-07-24 notebook/SI_science.tex
Last commit before shifting to natcom

--- 2018-06-18 notebook/SI.tex
@@
 changing species densities (\(\mathbf{N}\)) and a vector of potentially
-evolving traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
-For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+evolving traits (\(\mathbf{x}\)) (15). For any species \(i\) or trait
+\(j\), change in species density (\(N_{i}\)) or trait value (\(x_{j}\))
+with time (\(t\)) is a function of the vectors \(\mathbf{N}\) and
+\(\mathbf{x}\),
 
@@
 were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}, and
-this is modulated in the above by the scalar \(\epsilon\). The value of
-\(\epsilon\) thereby determines the timescale separation between ecology
-and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
+slow or fast in comparison with the change in species densities (15),
+and this is modulated in the above by the scalar \(\epsilon\). The value
+of \(\epsilon\) thereby determines the timescale separation between
+ecology and evolution, with high \(\epsilon\) modelling relatively fast
+evolution and low \(\epsilon\) modelling relative slow evolution (15).
 
@@
 rate of evolution to modulate rates of component responses for \(S\)
-components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}},
-the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
-by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
-\(i\) (\(m_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
+components. Following May (1, 19), the value of a component \(i\) at
+time \(t\) (\(v_{i}(t)\)) is affected by the value of \(j\)
+(\(v_{j}(t)\)) and \(j\)'s marginal effect on \(i\) (\(m_{ij}\)), and by
+\(i\)'s response rate (\(\gamma_{i}\)),
 
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{3}},
+In matrix notation (20),
 
@@
 \(\mathbf{\gamma} \mathbf{M}\) modulates the values of components and
-can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}}.
+can be analysed using the techniques of May (1, 19).
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-6-1.pdf}
 
@@
     \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
+    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.05}\NormalTok{);}
     \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
     \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
+    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(A_mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{);}
     \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
@@
     \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
-    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
-    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
     \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
-         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
-         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
+    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{6}\NormalTok{, }\FloatTok{2.25}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{5}\NormalTok{,}\FloatTok{4.25}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
     \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
@@
     \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
-    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
     \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
     \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
+    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{6}\NormalTok{, }\FloatTok{2.25}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{5}\NormalTok{,}\FloatTok{4.25}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
          \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
@@
     \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
+    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"solid"}\NormalTok{);}
     \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-While the foundational work of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2}} applies broadly
-to complex networks, much attention has been given specifically to
-ecological networks of interacting species. In these networks, the
-matrix \(M\) is interpreted as a community matrix and each row and
-column is interpreted as a single species. The effect that the density
-of any species \(i\) has on the population dynamics of species \(j\) is
-found in \(M_{ij}\), meaning that \(M\) holds the effects of pair-wise
-interactions between \(S\)
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}--\protect\hyperlink{ref-Tang2014b}{6}}.
-While May's original
-work\textsuperscript{\protect\hyperlink{ref-May1972}{2}} considered only
+While the foundational work of May (1) applies broadly to complex
+networks, much attention has been given specifically to ecological
+networks of interacting species. In these networks, the matrix \(M\) is
+interpreted as a community matrix and each row and column is interpreted
+as a single species. The effect that the density of any species \(i\)
+has on the population dynamics of species \(j\) is found in \(M_{ij}\),
+meaning that \(M\) holds the effects of pair-wise interactions between
+\(S\) species (2, 4, 20). While May's original work (1) considered only
 randomly assembled communities, recent work has specifically looked at
@@
 networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
-is negative and \(j\) on \(i\) is positive, or vice
-versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}--\protect\hyperlink{ref-Allesina2011}{7}}.
-In general, competitor and mutualist networks tend to be unstable, while
-predator-prey networks tend to be highly stabilising.
+is negative and \(j\) on \(i\) is positive, or vice versa) (2, 4, 17,
+20). In general, competitor and mutualist networks tend to be unstable,
+while predator-prey networks tend to be highly stabilising.
 
 I investigated competitor, mutualist, and predator-prey networks
-following Allesina et
-al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}. To create
-these networks, I first generated a random matrix \(M\), then changed
-the elements of \(M\) accordingly. If \(M\) was a competitive network,
-then the sign of any positive off-diagonal elements was reversed to be
-negative. If \(M\) was a mutualist network, then the sign of any
-positive off-diagonal elements was reversed to be positive. And if \(M\)
-was a predator-prey network, then all \(i\) and \(j\) pairs of elements
-were checked; any pairs of the same sign were changed so that one was
-negative and the other was positive. The \texttt{species\_interaction}
-function used to do this is below.
+following Allesina et al. (2). To create these networks, I first
+generated a random matrix \(M\), then changed the elements of \(M\)
+accordingly. If \(M\) was a competitive network, then the sign of any
+positive off-diagonal elements was reversed to be negative. If \(M\) was
+a mutualist network, then the sign of any positive off-diagonal elements
+was reversed to be positive. And if \(M\) was a predator-prey network,
+then all \(i\) and \(j\) pairs of elements were checked; any pairs of
+the same sign were changed so that one was negative and the other was
+positive. The \texttt{species\_interaction} function used to do this is
+below.
 
@@
 
-Overall, as
-expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}},
-predator-prey communities are relatively stable while mutualist
-communties are highly unstable. But interestingly, while \(Var(\gamma)\)
-stabilises predator-prey and competitor communities, it does not
-stabilise mutualist communities. This is unsurprising because purely
-mutualist communities are characterised by a very
-positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}
-leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
-alone will shift all real parts of eigenvalues to negative values.
+Overall, as expected (2), predator-prey communities are relatively
+stable while mutualist communties are highly unstable. But
+interestingly, while \(Var(\gamma)\) stabilises predator-prey and
+competitor communities, it does not stabilise mutualist communities.
+This is unsurprising because purely mutualist communities are
+characterised by a very positive (2) leading \(\Re(\lambda)\), and it is
+highly unlikely that \(Var(\gamma)\) alone will shift all real parts of
+eigenvalues to negative values.
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-24-1.pdf}
 
@@
 these conditions, genetic algorithms are highly useful tools for finding
-practical solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{8}}. In
-this case, the practical solution is finding vectors of \(\gamma\) that
-decrease the most positive real eigenvalue of \(M\). The genetic
-algorithm below achieves this by initialising a large population of 1000
-different potential \(\gamma\) vectors and allowing this population to
-evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
-values between vectors), selection, and reproduction until either a
-\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
-``giving up'' critiera is met (in the below, this ``giving up''"
+practical solutions by mimicking the process of biological evolution
+(16). In this case, the practical solution is finding vectors of
+\(\gamma\) that decrease the most positive real eigenvalue of \(M\). The
+genetic algorithm below achieves this by initialising a large population
+of 1000 different potential \(\gamma\) vectors and allowing this
+population to evolve through a process of mutation, crossover (swaping
+\(\gamma_{i}\) values between vectors), selection, and reproduction
+until either a \(\gamma\) vector is found where all \(\Re(\lambda) < 0\)
+or some ``giving up'' critiera is met (in the below, this ``giving up''"
 criteria is met when 20 generations pass, or if the fitness increase
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-32-1.pdf}
 
@@
 consider the feasibility of the system. Feasibile equilibria assume that
-the values of all system components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{9}--\protect\hyperlink{ref-Song2018}{11}}.
-This is of particular interest for ecological communities because
+the values of all system components are positive at equilibrium (5, 18,
+21). This is of particular interest for ecological communities because
 population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
-Consequently, the use of random matrices and traditional stability
+systems need to be feasible for stability to be biologically realistic
+(18). Consequently, the use of random matrices and traditional stability
 critiera for making inferences in theoretical analyses of species
-networks has recently been
-criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
-While the key results in the main text are intended to be general to all
-complex systems, and not restricted to species networks, I have also
-performed a feasibility analysis on all matrices \(M\). This analysis
-reveals that feasibility is not affected by \(Var(\gamma)\), meaning
-that for pure interacting species networks, variation in component
-response time (i.e., species generation time) does not affect stability
-at biologically realistic species densities. Nevertheless, ecological
-interactions do not exist in isolation in empirical systems, but instead
-interact with
-evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}},
-abiotic, or social-economic systems. The relevance of \(\gamma\) for
-complex system stability presented in the main text should therefore not
-be ignored in the broader context of ecological communities.
-
-Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}
-define the following feasibility criteria for ecological systems
-characterised by \(S\) interacting species with varying densities.
+networks has recently been criticised (18). While the key results in the
+main text are intended to be general to all complex systems, and not
+restricted to species networks, I have also performed a feasibility
+analysis on all matrices \(M\). This analysis reveals that feasibility
+is not affected by \(Var(\gamma)\), meaning that for pure interacting
+species networks, variation in component response time (i.e., species
+generation time) does not affect stability at biologically realistic
+species densities. Nevertheless, ecological interactions do not exist in
+isolation in empirical systems, but instead interact with evolutionary
+(15), abiotic, or social-economic systems. The relevance of \(\gamma\)
+for complex system stability presented in the main text should therefore
+not be ignored in the broader context of ecological communities.
+
+Dougoud et al. (18) define the following feasibility criteria for
+ecological systems characterised by \(S\) interacting species with
+varying densities.
 
@@
 diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the inter-connectivity (i.e., `connectance') of the community,
-which was set to \(C = 1\) throughout the manuscript and supplemental
-information, except \protect\hyperlink{connectance}{where otherwise
-noted}. The variable \(\delta\) is a normalisation parameter that
-modulates the strength of interactions (\(\sigma\) in the main text),
-which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
-underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
-so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
-\(A\), which has a diagonal of all zeros and an off-diagonal affecting
-species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{2}} stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}} by
-\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
-\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
-text and supplemental information). The above criteria is therefore
+\(C\) is the connectance of the community, which was set to \(C = 1\)
+throughout the manuscript and supplemental information, except
+\protect\hyperlink{connectance}{where otherwise noted}. The variable
+\(\delta\) is a normalisation parameter that modulates the strength of
+interactions (\(\sigma\) in the main text), which are held in \(A\). In
+the main text, implicitly, \(\delta = 0\) underlying strong
+interactions. Hence, the whole \((CS)^{-\delta} = 1\), so in the above,
+a diagonal matrix of -1s (\(\theta I\)) is added to \(A\), which has a
+diagonal of all zeros and an off-diagonal affecting species interactions
+(i.e., the expression \((CS)^{-\delta}\) relates to May's (1) stability
+criterion (19) by \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and
+hence \((CS)^{-\delta} = 1\) for the randomly simulated systems in the
+main text and supplemental information). The above criteria is therefore
 reduced to the below; note that the parenthetical in both equations
@@
 \(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
-is assumed to be a species with a growth rate of \(r_{i}\). Note that
-negative intrinsic growth rates are not unrealistic, and will occur in
-obligate mutualists in the absence of a partner.
+is assumed to be a species with a growth rate of \(r_{i}\).
 
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-Patel2018}{}
-1. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the effects
-of eco-evolutionary feedbacks on community stability. \emph{American
-Naturalist} \textbf{191,} 1--29 (2018).
-
-\hypertarget{ref-May1972}{}
-2. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
-
-\hypertarget{ref-May1973}{}
-3. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
-\textbf{54,} 638--641 (1973).
-
-\hypertarget{ref-Allesina2012}{}
-4. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
-\emph{Nature} \textbf{483,} 205--208 (2012).
-
-\hypertarget{ref-Allesina2015a}{}
-5. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
-(2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Tang2014b}{}
-6. Tang, S. \& Allesina, S. Reactivity and stability of large
-ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
-(2014).
-
-\hypertarget{ref-Allesina2011}{}
-7. Allesina, S. \& Levine, J. M. A competitive network theory of species
-diversity. \emph{Proceedings of the National Academy of Sciences of the
-United States of America} \textbf{108,} 5638--5642 (2011).
-
-\hypertarget{ref-Hamblin2013}{}
-8. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
-
-\hypertarget{ref-Grilli2017}{}
-9. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
-
-\hypertarget{ref-Dougoud2018}{}
-10. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
-
-\hypertarget{ref-Song2018}{}
-11. Song, C. \& Saavedra, S. Will a small randomly assembled community
-be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
-
 \end{document}

### 2018-07-24 notebook/ms_for_review.Rmd
Last commit before shifting to natcom

@@
 output:
-  pdf_document:
-    fig_caption: yes
-    keep_tex: no
-  html_document: default
   word_document:
@@
     reference_docx: docx_template.docx
+  pdf_document:
+    fig_caption: yes
+    keep_tex: no
+  html_document: default
 bibliography: references.bib
@@
 As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b) [see also ].
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.

### 2018-07-24 R/random_mats.R

@@ -16,23 +16,27 @@
 #'@examples
 #'rand_gen_var(max_sp = 2, iters = 4);
 #'@export
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
+rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
+                         sigma = 0.4){
     tot_res <- NULL;
     fea_res <- NULL;
-    for(i in 2:max_sp){
+    sp_try  <- seq(from = by, to = max_sp, by = by);
+    for(i in 1:length(sp_try)){
         iter           <- iters;
-        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
-        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
+        tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
+        fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         while(iter > 0){
-            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
-            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
-            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
+            r_vec    <- rnorm(n = sp_try[i], mean = 0, sd = rmx);
+            A0_dat   <- rnorm(n = sp_try[i] * sp_try[i], mean = 0, sd = sigma);
+            A0       <- matrix(data = A0_dat, nrow = sp_try[i], 
+                               ncol = sp_try[i]);
             A0       <- species_interactions(mat = A0, type = int_type);
-            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
-            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
+            C_dat    <- rbinom(n = sp_try[i] * sp_try[i], size = 1, prob = C);
+            C_mat    <- matrix(data = C_dat, nrow = sp_try[i], 
+                               ncol = sp_try[i]);
             A0       <- A0 * C_mat;
             diag(A0) <- -1;
-            gam1     <- runif(n = i, min = 0, max = 2);
+            gam1     <- runif(n = sp_try[i], min = 0, max = 2);
             A1       <- A0 * gam1;
             A0       <- A0 * mean(gam1);
             A0_stb   <- max(Re(eigen(A0)$values)) < 0;
[…]

### 2018-07-26 R/Figure_1.R
Changes to R functions

@@ -81,20 +81,21 @@ plot_Fig_1 <- function(A0, A1){
     A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
     A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
     text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
+    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "dodgerblue4");
     points(A0_r, A0_i, pch = 4, cex = 0.7);
     plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
          xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-         col = "black", yaxt = "n");
+         col = "dodgerblue4", yaxt = "n");
     vl <- seq(from = 0, to = 2*pi, by = 0.001);
-    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:(0.5*S_val)]);
+    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + 
+                 mean(diag(A1)[1:(0.5*S_val)]);
     A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
-    points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+    #points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
     A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
         mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
     A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
-    points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
-    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7);   
+    #points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7, col = "firebrick");   
[…]

### 2018-07-26 R/Figure_2.R

@@ -38,20 +38,20 @@ plot_Fig_2 <- function(){
     par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
     plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
          cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
-         asp = 1);
+         asp = 1, col = "dodgerblue4");
     vl <- seq(from = 0, to = 2*pi, by = 0.001);
     x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
     y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
     x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
     y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
     text(x = -15.5, y = 19, labels = "a", cex = 2);
-    points(x = x0, y = y0, type = "l", lwd = 3);
+    points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
     points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
-         yaxt = "n");
+    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "firebrick", yaxt = "n");
     text(x = -15.5, y = 19, labels = "b", cex = 2);
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
+    points(x = x1, y = y1, type = "l", col = "firebrick", lwd = 3);
     points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
     mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
     mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);

### 2018-07-26 R/Figures_3_4.R

@@ -1,0 +1,95 @@
+#' Plot comparison of gamma = 1 versus var(gamma)
+#' 
+#' Produces a plot comparing random matrix results
+#'
+#'@return A plot as in Figures 3 of the manuscript, showing the stability of
+#'random matrices on a log scale in simulations where component response rate is
+#'constant versus simulations in which it varies
+#'@param dat Table of simulation results
+#'@param S_s Maximum size of S to use
+#'@export
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
[…]

### 2018-07-26 notebook/ms_for_review.Rmd
New files

@@
 linkcolor: blue
-csl: science.csl
+csl: nature.csl
 biblio-style: apalike
@@
 https://gist.github.com/JJ/ef9d3d8f1142df064bd5
+--->
+
+<!---
+
+Given the larger space of Nature Communications, one new figure will be used in the same style as that of Figure 3; the new Figure 4 will show the same results, but for sultions found with the genetic algorithm.
+
+Because Nature communcations is online, change all non-standard gamma results in figures to Red. The new genetic algorithm results will be in blue.
+
 --->
@@
 ```
+
+Results
+--------------------------------------------------------------------------------
 
@@
 
+Discussion
+--------------------------------------------------------------------------------
+
 My results show that complex systems are more likely to be stable when the response rates of system components vary.
 These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
+
+Methods
+--------------------------------------------------------------------------------
+
+**Component response rate variation ($\mathbf{\gamma}$)**.
+In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
+
+$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+
+$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+
+In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
+Patel et al.
+were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
+The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
+
+I use the same principle that Patel et al.
+use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
+Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+
+$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).$$
+
+In matrix notation[@May1973],
+
+$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).$$
+
+In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
+Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
+
+**Genetic algorithm**.
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $M$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
+In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
+The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
+
+**System feasibility**.
+For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
+Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
+Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
+While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
+This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
+
+Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
+
+$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+
+In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive).
+The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements).
+As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
+The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
+The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
+In the main text, implicitly, $\delta = 0$ underlying strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
+The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
+
+$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
+
+To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$.
+Feasibility is satisfied if all of the elements of the resulting vector are positive.
+
+The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$.
 
@@
 <div id="refs"></div>
+
+```{r, echo = FALSE}
+# Temporary place for revised functions
+plot_Fig_1 <- function(A0, A1){
+    S_val       <- dim(A0)[1];
+    A0_e        <- eigen(A0)$values;
+    A0_r        <- Re(A0_e);
+    A0_i        <- Im(A0_e);
+    A1_e        <- eigen(A1)$values;
+    A1_r        <- Re(A1_e);
+    A1_i        <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(t(A0_vm));
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(t(A1_vm));
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    fhalf       <- 1:(0.5*length(A1vec));
+    shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+    vl   <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
+    text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(A0_r, A0_i, pch = 4, cex = 0.7);
+    plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+         col = "dodgerblue4", yaxt = "n");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + 
+                 mean(diag(A1)[1:(0.5*S_val)]);
+    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
+    #points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
+        mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
+    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
+    #points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7, col = "firebrick");   
+    text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_Fig_2 <- function(){
+    A_comp <- NULL;
+    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+    A_mat  <- matrix(data = A_dat, nrow = 1000);
+    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+    A_mat     <- A_mat * C_mat;
+    gammas <- runif(n = 1000, min = 0, max = 2);
+    mu_gam <- mean(gammas);
+    diag(A_mat) <- -1;
+    A1     <- gammas * A_mat;
+    A0     <- mu_gam * A_mat;
+    A0_e   <- eigen(A0)$values;
+    A0_r   <- Re(A0_e);
+    A0_i   <- Im(A0_e);
+    A1_e   <- eigen(A1)$values;
+    A1_r   <- Re(A1_e);
+    A1_i   <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(A0_vm);
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(A1_vm);
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "dodgerblue4");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+    text(x = -15.5, y = 19, labels = "a", cex = 2);
+    points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "firebrick", yaxt = "n");
+    text(x = -15.5, y = 19, labels = "b", cex = 2);
+    points(x = x1, y = y1, type = "l", col = "firebrick", lwd = 3);
+    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
+
+plot_stable_4 <- function(dat, S_s = 39){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 14), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 116), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
+```
 
@@
 **Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the blue circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
 **b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
@@
 plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "dodgerblue4");
 vl <- seq(from = 0, to = 2*pi, by = 0.001);
@@
 text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
-points(A0_r, A0_i, pch = 4, cex = 0.7);
+points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "blue");
+points(A0_r, A0_i, pch = 4, cex = 0.7, col = "dodgerblue4");
 
@@
      xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "black", yaxt = "n");
+     col = "firebrick", yaxt = "n");
 
@@
 A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
-points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "red");
 A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
 A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
-points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "red");
 
-points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7);
+points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7, col = "firebrick");
 
@@
 **a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
-Black elipses in both panels show the circle centred on the distribution in panel a.
+Blue elipses in both panels show the circle centred on the distribution in panel a.
 Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
@@
 plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "dodgerblue4");
 vl <- seq(from = 0, to = 2*pi, by = 0.001);
@@
 text(x = -15.5, y = 19, labels = "a", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3);
+points(x = x0, y = y0, type = "l", lwd = 3, col = "blue");
+points(A0_r, A0_i, col = "dodgerblue4",pch = 4, cex = 0.7);
 plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
-     yaxt = "n");
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "firebrick", yaxt = "n");
 text(x = -15.5, y = 19, labels = "b", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid");
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid", col = "blue");
+points(A1_r, A1_i, col = "firebrick", pch = 4, cex = 0.7);
 mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
@@
 
-**Figure 3: Stability of large complex systems with and without variation in component response rate($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
@@
 Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Light and dark grey bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
+Blue and red bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
 The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
@@
 ```
+
+\clearpage
+
+**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targetted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targetted search of $\gamma$ values successfully resulted in system stability.
+For each $S$, 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
+Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
+The black line shows the proportion of systems that are stable when component response rate is targetted to increase stability, but would not be stable if $Var(\gamma) = 0$.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
+dat <- dat[,-1];
+plot_stable_4(dat);
+```

### 2018-07-26 notebook/ms_for_review.Rmd
Add more edits

@@
 
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018] (see Supplementary Information for details).
+**Component response rates of random complex systems**.
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018].
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
@@
 1b).
-Instead, two distinct clusters of eigenvalues appear (grey circles in Fig.
+Instead, two distinct clusters of eigenvalues appear (red circles in Fig.
 1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
@@
 Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I now focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
+I focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
@@
 
+**Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
 To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
@@
 For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
+
+**Targetted manipulation of $\mathbf{\gamma}$**.
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively [@Hamblin2013]).
+[Need a methods like sentence here](http://google.com).
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
+4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+
+Discussion
+--------------------------------------------------------------------------------
+
 This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
@@
 
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively [@Hamblin2013]; see Supplementary Information).
-For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
-Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
-
 I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks [@Allesina2012].
@@
 
-Discussion
---------------------------------------------------------------------------------
+Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
 
@@
 The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
+
+For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
+Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
 

### 2018-07-27 notebook/SI_science.tex
Reformat SI with new results

@@
 
-While downloading this package is recommended, all relevant code is also
-reproduced below with explanation, so it is possible to recreate all
-analyses using only this supplemental information.
-
 \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
@@
 \item
-  \protect\hyperlink{moregamma}{Further explanation of \(\gamma\)}
-\item
-  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
-\item
-  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
-\item
   \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
+\item
+  \protect\hyperlink{ga}{Stability given targetted manipulation of
+  \(\gamma\) (genetic algorithm)}
 \item
@@
 \item
-  \protect\hyperlink{connectance}{Different connectance (C) values}
+  \protect\hyperlink{connectance}{Sensitivity of connectance (C) values}
 
@@
 \item
-  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
-\item
-  \protect\hyperlink{ga}{Genetic algorithm}
+  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
+  and \(\sigma\)}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
+  \item
+    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
+  \item
+    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
+  \item
+    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
+  \end{itemize}
+\item
+  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
+  \(\gamma\)}
 \item
@@
 
-\hypertarget{moregamma}{\section{\texorpdfstring{Further explanation of
-\(\gamma\)}{Further explanation of \textbackslash{}gamma}}\label{moregamma}}
-
-In a synthesis of eco-evolutionary feedbacks on community stability,
-Patel et al. model a system that includes a vector of potentially
-changing species densities (\(\mathbf{N}\)) and a vector of potentially
-evolving traits (\(\mathbf{x}\)) (15). For any species \(i\) or trait
-\(j\), change in species density (\(N_{i}\)) or trait value (\(x_{j}\))
-with time (\(t\)) is a function of the vectors \(\mathbf{N}\) and
-\(\mathbf{x}\),
-
-\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
-
-In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
-effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
-were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species densities (15),
-and this is modulated in the above by the scalar \(\epsilon\). The value
-of \(\epsilon\) thereby determines the timescale separation between
-ecology and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow evolution (15).
-
-I use the same principle that Patel et al. use to modulate the relative
-rate of evolution to modulate rates of component responses for \(S\)
-components. Following May (1, 19), the value of a component \(i\) at
-time \(t\) (\(v_{i}(t)\)) is affected by the value of \(j\)
-(\(v_{j}(t)\)) and \(j\)'s marginal effect on \(i\) (\(m_{ij}\)), and by
-\(i\)'s response rate (\(\gamma_{i}\)),
-
-\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).\]
-
-In matrix notation (20),
-
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{\gamma} \mathbf{M}\) modulates the values of components and
-can be analysed using the techniques of May (1, 19).
-
-\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
-1}\label{Fig1}}
-
-The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
-text is available on
-\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
-and was produced by running the following function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
-            \NormalTok{break;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(iters);}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
-is found that is not stable when all component response rates are set to
-\(\gamma = 1\), but is stable when half of component response rates are
-\(1.95\) and half are \(0.05\). The function is used to illustrate the
-concept of how fast versus slow component responses can cause a system
-to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
-but terminated once an acceptable \texttt{A0} and \texttt{A1} were
-found. The code below plots the eigenvalue distributions of \texttt{A0}
-and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
-plot itself can be recreated with the function and code below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{A0 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A0[,-}\DecValTok{1}\NormalTok{]);}
-\NormalTok{A1 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A1[,-}\DecValTok{1}\NormalTok{]);}
-\NormalTok{plot_Fig_1 <-}\StringTok{ }\NormalTok{function(A0, A1)\{}
-    \NormalTok{S_val       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A0)[}\DecValTok{1}\NormalTok{];}
-    \NormalTok{A0_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-    \NormalTok{A0_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-    \NormalTok{A0_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-    \NormalTok{A1_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-    \NormalTok{A1_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-    \NormalTok{A1_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
-    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A0_vm));}
-    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
-    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A1_vm));}
-    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{fhalf       <-}\StringTok{ }\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec));}
-    \NormalTok{shalf       <-}\StringTok{ }\NormalTok{(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec)+}\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(A1vec);}
-    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{vl   <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{A0x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
-    \NormalTok{A0y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x0, }\DataTypeTok{y =} \NormalTok{A0y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(A0_r, A0_i, }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }
-         \DataTypeTok{col =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
-    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{A0x1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*S_val)]);}
-    \NormalTok{A0y1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1a, }\DataTypeTok{y =} \NormalTok{A0y1a, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \NormalTok{A0x1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }
-\StringTok{                  }\KeywordTok{mean}\NormalTok{( }\KeywordTok{diag}\NormalTok{(A1)[( (}\FloatTok{0.5}\NormalTok{*S_val) +}\StringTok{ }\DecValTok{1} \NormalTok{):S_val] );}
-    \NormalTok{A0y1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1b, }\DataTypeTok{y =} \NormalTok{A0y1b, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(A1_r[}\DecValTok{1}\NormalTok{:S_val], A1_i[}\DecValTok{1}\NormalTok{:S_val],}\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);   }
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-\NormalTok{\}}
-\KeywordTok{plot_Fig_1}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1);}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{SI_science_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-To find out how frequently \(M\) was stable given that all
-\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
-was created.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(ress);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above functions produced the \texttt{bi\_pr\_st} data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
-\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
-\end{Highlighting}
-\end{Shaded}
-
-The function \texttt{stab\_bgamma} was run for
-\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
-returned. Each row of \texttt{ress} represents a single \(M\) given
-\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
-Values of 0 indicate that \(M\) was found to be unstable (at least one
-real component of its eigenvalues greater than or equal to zero),
-whereas values of 1 indicate that \(M\) was found to be stable (all real
-components of eigenvalues are negative). The frequencies of stable \(M\)
-were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
-reported in the main text and legend of Fig. 1 (raw data are
-\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
-on GitHub}).
-
-\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
-2}\label{Fig2}}
-
-Figure 2 of the main text shows eigenvalue distributions in a system
-where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
-reproduced using the code below for when \(\gamma = 1\) (panel a) and
-\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
-reproduces the figure.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
-    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.05}\NormalTok{);}
-    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(A_mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{);}
-    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
-    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
-    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{6}\NormalTok{, }\FloatTok{2.25}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{5}\NormalTok{,}\FloatTok{4.25}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
-    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{6}\NormalTok{, }\FloatTok{2.25}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{5}\NormalTok{,}\FloatTok{4.25}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
-         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"solid"}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-\NormalTok{\}}
-\KeywordTok{plot_Fig_2}\NormalTok{();}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{SI_science_files/figure-latex/unnamed-chunk-9-1.pdf}
-
 \hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
@@
 Figure 3 of the main text reports the number of stable random complex
-systems found over 1 million iterations. The data used to make this
-figure are read into R below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_1/random_all.csv"}\NormalTok{);}
-\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[,-}\DecValTok{1}\NormalTok{]; }\CommentTok{# Extra row-indicating column removed}
-\end{Highlighting}
-\end{Shaded}
-
-The table below shows the results for all simulations of random \(M\)
-matrices at \(\sigma = 0.4\) and \(C = 1\) given a range of
-\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
-matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
-after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
-Each row summarises data for a given \(S\) over 1 million randomly
-simulated \(M\) (\texttt{A0} and \texttt{A1}). The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 1
-million). Similarly, the column \texttt{A1\_unstable} shows the number
-of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
-the number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+systems found over 1 million iterations. The table below shows the
+results for all simulations of random \(M\) matrices at \(\sigma = 0.4\)
+and \(C = 1\) given a range of \(S = \{2, 3, ..., 49, 50\}\). In this
+table, the \texttt{A0} refers to matrices where \(\gamma = 1\), while
+\texttt{A1} refers to matrices after \(Var(\gamma)\) is added and
+\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
+\(S\) over 1 million randomly simulated \(M\) (\texttt{A0} and
+\texttt{A1}). The column \texttt{A0\_unstable} shows the number of
+\texttt{A0} matrices that are unstable, and the column
+\texttt{A0\_stable} shows the number of \texttt{A0} matrices that are
+stable (these two columns sum to 1 million). Similarly, the column
+\texttt{A1\_unstable} shows the number of \texttt{A1} matrices that are
+unstable and \texttt{A1\_stable} shows the number that are stable. The
+columns \texttt{A1\_stabilised} and \texttt{A1\_destabilised} show how
+many \texttt{A0} matrices were stabilised or destabilised, respectively,
+by \(Var(\gamma)\).
 
@@
 randomly created complex systems (\(M\)) are generated given variation
-in \(\gamma\) than when \(\gamma = 1\). The results underlying this
-table were produced with the \texttt{rand\_gen\_var} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above function calls the two functions
-\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
-are provided below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-
-\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \CommentTok{# Stable and unstable}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Stabilised and destabilised}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \CommentTok{# Feasible and infeasible}
-        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Feased and defeased}
-        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-    \NormalTok{\}}
-    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
-                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
-                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
-                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
-    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that feasibility results were ommited for the table above, but are
+in \(\gamma\) than when \(\gamma = 1\). Note that feasibility results
+were ommited for the table above, but are
 \protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targetted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targetted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+Figure 4 of the main text reports the number of stable random complex
+systems found over 100000 using the genetic algorithm to maximise
+stability with a vector \(\mathbf{\gamma}\). Stability results for
+100000 \(M\) for each \(S\) from 2-40 are shown below. Results for
+\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
+refers to systems in which the genetic algorithm searched for a set of
+\(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(M\) that can be produced through a targetted search of \(\gamma\)
+values, and suggests that many otherwise unstable systems could
+potentially be stabilised by an informed manipulation of their component
+response times. Such a possibility might conceivably reduce the
+dimensionality of problems involving stability in social-ecological or
+economic systems.
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
 
@@
 the same sign were changed so that one was negative and the other was
-positive. The \texttt{species\_interaction} function used to do this is
-below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
-\end{Highlighting}
-\end{Shaded}
-
-This function was applied to all created matrices \(M\), then the number
-of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
-as it was} in the main text for random matrices for values of \(S\) from
-2 to 50 (100 in the case of the relatively more stable predator-prey
-interactions), except that only 100000 random \(M\) were generated
-instead of 1 million. This produced the data set below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
-\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
-\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
+positive.
+
+The number of stable \(M\) matrices was estimated
+\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
+matrices for values of \(S\) from 2 to 50 (100 in the case of the
+relatively more stable predator-prey interactions), except that only
+100000 random \(M\) were generated instead of 1 million.
 
@@
 
-Overall, as expected (2), predator-prey communities are relatively
-stable while mutualist communties are highly unstable. But
-interestingly, while \(Var(\gamma)\) stabilises predator-prey and
-competitor communities, it does not stabilise mutualist communities.
-This is unsurprising because purely mutualist communities are
-characterised by a very positive (2) leading \(\Re(\lambda)\), and it is
-highly unlikely that \(Var(\gamma)\) alone will shift all real parts of
-eigenvalues to negative values.
-
-\hypertarget{connectance}{\section{Different inter-connectivity (C)
+Overall, as
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{1}},
+predator-prey communities are relatively stable while mutualist
+communties are highly unstable. But interestingly, while \(Var(\gamma)\)
+stabilises predator-prey and competitor communities, it does not
+stabilise mutualist communities. This is unsurprising because purely
+mutualist communities are characterised by a very positive (2) leading
+\(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\) alone
+will shift all real parts of eigenvalues to negative values.
+
+\hypertarget{connectance}{\section{Sensitivity of inter-connectivity (C)
 values}\label{connectance}}
@@
 
-All data reported below for various values of \(C\) are accessible using
-the below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{C3dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt3.csv"}\NormalTok{);}
-\NormalTok{C5dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt5.csv"}\NormalTok{);}
-\NormalTok{C7dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt7.csv"}\NormalTok{);}
-\NormalTok{C9dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt9.csv"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
-
-These objects \texttt{C3dat}, \texttt{C5dat}, \texttt{C7dat}, and
-\texttt{C9dat} include the results for \(C = 0.3\), \(C = 0.5\),
-\(C = 0.7\), and \(C = 0.9\), respectively.
-
-\textbf{Connectance \(C = 0.3\)}
+\textbf{Connectance \(\mathbf{C = 0.3}\)}
 
@@
 
-\textbf{Connectance \(C = 0.5\)}
+\textbf{Connectance \(\mathbf{C = 0.5}\)}
 
@@
 
-\textbf{Connectance \(C = 0.7\)}
+\textbf{Connectance \(\mathbf{C = 0.7}\)}
 
@@
 
-\textbf{Connectance \(C = 0.9\)}
+\textbf{Connectance \(\mathbf{C = 0.9}\)}
 
@@
 
-\hypertarget{gam_dist}{\section{\texorpdfstring{Different distributions
-of
-\(\gamma\)}{Different distributions of \textbackslash{}gamma}}\label{gam_dist}}
+\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
+strength (\(\sigma\))
+values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
+
+\textbf{ADD TEXT ABOUT COMPARISON WITH ALLESINA AND VALUES ACROSS SIGMA}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+20 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+30 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+40 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+60 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+70 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+80 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
+90 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+100 & 11 & 99989 & 11 & 99989 & 0 & 0\tabularnewline
+110 & 12 & 99988 & 13 & 99987 & 0 & 1\tabularnewline
+120 & 23 & 99977 & 23 & 99977 & 0 & 0\tabularnewline
+130 & 40 & 99960 & 40 & 99960 & 0 & 0\tabularnewline
+140 & 62 & 99938 & 65 & 99935 & 0 & 3\tabularnewline
+150 & 162 & 99838 & 165 & 99835 & 0 & 3\tabularnewline
+160 & 325 & 99675 & 329 & 99671 & 2 & 6\tabularnewline
+170 & 829 & 99171 & 851 & 99149 & 6 & 28\tabularnewline
+180 & 1817 & 98183 & 1860 & 98140 & 31 & 74\tabularnewline
+190 & 3927 & 96073 & 3989 & 96011 & 143 & 205\tabularnewline
+200 & 8084 & 91916 & 8048 & 91952 & 557 & 521\tabularnewline
+210 & 15558 & 84442 & 15147 & 84853 & 1534 & 1123\tabularnewline
+220 & 26848 & 73152 & 25342 & 74658 & 3625 & 2119\tabularnewline
+230 & 43386 & 56614 & 39535 & 60465 & 6992 & 3141\tabularnewline
+240 & 62734 & 37266 & 56684 & 43316 & 9815 & 3765\tabularnewline
+250 & 80128 & 19872 & 73080 & 26920 & 10128 & 3080\tabularnewline
+260 & 92206 & 7794 & 86619 & 13381 & 7490 & 1903\tabularnewline
+270 & 97946 & 2054 & 94824 & 5176 & 3797 & 675\tabularnewline
+280 & 99659 & 341 & 98534 & 1466 & 1265 & 140\tabularnewline
+290 & 99962 & 38 & 99696 & 304 & 281 & 15\tabularnewline
+300 & 99994 & 6 & 99964 & 36 & 34 & 4\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 3 & 99997 & 3 & 99997 & 0 & 0\tabularnewline
+20 & 15 & 99985 & 15 & 99985 & 0 & 0\tabularnewline
+30 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+40 & 85 & 99915 & 85 & 99915 & 0 & 0\tabularnewline
+50 & 163 & 99837 & 163 & 99837 & 0 & 0\tabularnewline
+60 & 280 & 99720 & 282 & 99718 & 0 & 2\tabularnewline
+70 & 561 & 99439 & 566 & 99434 & 3 & 8\tabularnewline
+80 & 1009 & 98991 & 1029 & 98971 & 6 & 26\tabularnewline
+90 & 2126 & 97874 & 2175 & 97825 & 31 & 80\tabularnewline
+100 & 4580 & 95420 & 4653 & 95347 & 142 & 215\tabularnewline
+110 & 9540 & 90460 & 9632 & 90368 & 465 & 557\tabularnewline
+120 & 19090 & 80910 & 18668 & 81332 & 1676 & 1254\tabularnewline
+130 & 35047 & 64953 & 33220 & 66780 & 4172 & 2345\tabularnewline
+140 & 56411 & 43589 & 52439 & 47561 & 7297 & 3325\tabularnewline
+150 & 78003 & 21997 & 72574 & 27426 & 8477 & 3048\tabularnewline
+160 & 92678 & 7322 & 88438 & 11562 & 5901 & 1661\tabularnewline
+170 & 98614 & 1386 & 96670 & 3330 & 2397 & 453\tabularnewline
+180 & 99839 & 161 & 99418 & 582 & 499 & 78\tabularnewline
+190 & 99990 & 10 & 99945 & 55 & 52 & 7\tabularnewline
+200 & 100000 & 0 & 99995 & 5 & 5 & 0\tabularnewline
+210 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 36 & 99964 & 36 & 99964 & 0 & 0\tabularnewline
+20 & 195 & 99805 & 195 & 99805 & 0 & 0\tabularnewline
+30 & 519 & 99481 & 523 & 99477 & 0 & 4\tabularnewline
+40 & 1096 & 98904 & 1101 & 98899 & 2 & 7\tabularnewline
+50 & 2375 & 97625 & 2397 & 97603 & 9 & 31\tabularnewline
+60 & 4898 & 95102 & 4968 & 95032 & 83 & 153\tabularnewline
+70 & 10841 & 89159 & 10916 & 89084 & 432 & 507\tabularnewline
+80 & 22281 & 77719 & 21988 & 78012 & 1622 & 1329\tabularnewline
+90 & 42010 & 57990 & 39998 & 60002 & 4458 & 2446\tabularnewline
+100 & 67289 & 32711 & 63098 & 36902 & 7153 & 2962\tabularnewline
+110 & 88137 & 11863 & 84023 & 15977 & 6108 & 1994\tabularnewline
+120 & 97678 & 2322 & 95557 & 4443 & 2740 & 619\tabularnewline
+130 & 99795 & 205 & 99304 & 696 & 578 & 87\tabularnewline
+140 & 99989 & 11 & 99948 & 52 & 49 & 8\tabularnewline
+150 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 162 & 99838 & 162 & 99838 & 0 & 0\tabularnewline
+20 & 798 & 99202 & 799 & 99201 & 0 & 1\tabularnewline
+30 & 2273 & 97727 & 2289 & 97711 & 6 & 22\tabularnewline
+40 & 5259 & 94741 & 5298 & 94702 & 70 & 109\tabularnewline
+50 & 12084 & 87916 & 12054 & 87946 & 446 & 416\tabularnewline
+60 & 26072 & 73928 & 25511 & 74489 & 1810 & 1249\tabularnewline
+70 & 50121 & 49879 & 47747 & 52253 & 4748 & 2374\tabularnewline
+80 & 77806 & 22194 & 73810 & 26190 & 6421 & 2425\tabularnewline
+90 & 94862 & 5138 & 92069 & 7931 & 3842 & 1049\tabularnewline
+100 & 99527 & 473 & 98822 & 1178 & 870 & 165\tabularnewline
+110 & 99984 & 16 & 99912 & 88 & 80 & 8\tabularnewline
+120 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+130 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
+distribution of
+\(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
 
@@
 
-\includegraphics{SI_science_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{SI_science_files/figure-latex/unnamed-chunk-23-1.pdf}
 
@@
 
-\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}
-
-Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
-the proportion of stable complex systems, the search space of all
-possible \(\gamma\) vectors would be evaluated for each unique \(M\).
-This is technically impossible because \(\gamma_{i}\) can take any real
-value between 0-2, but even rounding \(\gamma\) to reasonable values
-would result in a search space too large to practically explore. Under
-these conditions, genetic algorithms are highly useful tools for finding
-practical solutions by mimicking the process of biological evolution
-(16). In this case, the practical solution is finding vectors of
-\(\gamma\) that decrease the most positive real eigenvalue of \(M\). The
-genetic algorithm below achieves this by initialising a large population
-of 1000 different potential \(\gamma\) vectors and allowing this
-population to evolve through a process of mutation, crossover (swaping
-\(\gamma_{i}\) values between vectors), selection, and reproduction
-until either a \(\gamma\) vector is found where all \(\Re(\lambda) < 0\)
-or some ``giving up'' critiera is met (in the below, this ``giving up''"
-criteria is met when 20 generations pass, or if the fitness increase
-from one generation to the next is below a certain criteria). The
-genetic algorithm relies on five functions. The first outer function
-\texttt{Evo\_rand\_gen\_var} runs all of the simulations
-(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
-\texttt{iters} refers to the number of \(M\) to try for each \(S\)).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
-        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
-\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
-\texttt{summarise\_randmat}. The first simply allows for non-random
-interactions between components (e.g., modelling
-\protect\hyperlink{ecological}{ecological interactions} of random,
-competition, mutualism, or predator-prey).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
-algorithm, searching for \(\gamma\) vectors that are stabilising.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
-    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
-    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
-    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
-    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
-    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
-        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
-            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
-            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{\}}
-        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
-        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
-                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
-        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
-        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
-        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
-        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
-        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
-        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
-        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
-        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
-        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
-        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
-        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
-        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(find_st);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The while loop in \texttt{rand\_mat\_ga} continues until either
-\texttt{iter} generations have occured, a solution \(\gamma\) vector is
-found that results in all \(\Re(\lambda) < 0\), or some criteria of
-minimum fitness increase is observed (by default,
-\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
-values are mutated, crossover occurs between \(\gamma\) vectors, and
-selection occurs in each generation such that the 20 \(\gamma\) vectors
-that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
-offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
-zero are multiplied by \(-1\), and any values that mutate above 2 are
-set to 2. Note also that if a solution is found, then one such
-\(\gamma\) vector causing stability is printed to a file.
-
-Crossover occurs in the \texttt{crossover} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
-    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
-    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
-        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
-        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
-        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(inds);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
-\texttt{summarise\_randmat} formats the data into a table.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{A0_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{A0_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{A1_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{A1_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_unst);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_stbl);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_unst);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_stbl);}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
-    \NormalTok{\}}
-    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{,}
-                           \StringTok{"A1_stable"}\NormalTok{, }\StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Stability results for 40000 \(M\) for each \(S\) from 2-40 are shown
-below. Results for \texttt{A0} indicate systems in which \(\gamma = 1\),
-while \texttt{A1} refers to systems in which the genetic algorithm
-searched for a set of \(\gamma\) values that stabilised the system.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 13 & 39987 & 13 & 39987 & 0 & 0\tabularnewline
-3 & 157 & 39843 & 157 & 39843 & 0 & 0\tabularnewline
-4 & 563 & 39437 & 563 & 39437 & 0 & 0\tabularnewline
-5 & 1588 & 38412 & 1581 & 38419 & 7 & 0\tabularnewline
-6 & 3268 & 36732 & 3219 & 36781 & 49 & 0\tabularnewline
-7 & 5370 & 34630 & 5196 & 34804 & 175 & 1\tabularnewline
-8 & 8231 & 31769 & 7604 & 32396 & 627 & 0\tabularnewline
-9 & 11704 & 28296 & 10181 & 29819 & 1523 & 0\tabularnewline
-10 & 15215 & 24785 & 12303 & 27697 & 2912 & 0\tabularnewline
-11 & 19529 & 20471 & 14599 & 25401 & 4930 & 0\tabularnewline
-12 & 23480 & 16520 & 16210 & 23790 & 7270 & 0\tabularnewline
-13 & 27570 & 12430 & 17800 & 22200 & 9770 & 0\tabularnewline
-14 & 31096 & 8904 & 18899 & 21101 & 12198 & 1\tabularnewline
-15 & 34036 & 5964 & 19948 & 20052 & 14089 & 1\tabularnewline
-16 & 36231 & 3769 & 21000 & 19000 & 15231 & 0\tabularnewline
-17 & 37737 & 2263 & 21693 & 18307 & 16044 & 0\tabularnewline
-18 & 38808 & 1192 & 22291 & 17709 & 16518 & 1\tabularnewline
-19 & 39391 & 609 & 23221 & 16779 & 16170 & 0\tabularnewline
-20 & 39711 & 289 & 23943 & 16057 & 15768 & 0\tabularnewline
-21 & 39867 & 133 & 25361 & 14639 & 14506 & 0\tabularnewline
-22 & 39950 & 50 & 26584 & 13416 & 13366 & 0\tabularnewline
-23 & 39982 & 18 & 28162 & 11838 & 11820 & 0\tabularnewline
-24 & 39994 & 6 & 29660 & 10340 & 10334 & 0\tabularnewline
-25 & 39997 & 3 & 31405 & 8595 & 8592 & 0\tabularnewline
-26 & 39999 & 1 & 33057 & 6943 & 6942 & 0\tabularnewline
-27 & 40000 & 0 & 34602 & 5398 & 5398 & 0\tabularnewline
-28 & 40000 & 0 & 35957 & 4043 & 4043 & 0\tabularnewline
-29 & 40000 & 0 & 37154 & 2846 & 2846 & 0\tabularnewline
-30 & 40000 & 0 & 38104 & 1896 & 1896 & 0\tabularnewline
-31 & 40000 & 0 & 38722 & 1278 & 1278 & 0\tabularnewline
-32 & 40000 & 0 & 39259 & 741 & 741 & 0\tabularnewline
-33 & 40000 & 0 & 39551 & 449 & 449 & 0\tabularnewline
-34 & 40000 & 0 & 39742 & 258 & 258 & 0\tabularnewline
-35 & 40000 & 0 & 39879 & 121 & 121 & 0\tabularnewline
-36 & 40000 & 0 & 39929 & 71 & 71 & 0\tabularnewline
-37 & 40000 & 0 & 39967 & 33 & 33 & 0\tabularnewline
-38 & 40000 & 0 & 39989 & 11 & 11 & 0\tabularnewline
-39 & 40000 & 0 & 39994 & 6 & 6 & 0\tabularnewline
-40 & 40000 & 0 & 39996 & 4 & 4 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. Recall that 1 million random matrices were
-generated for the less computationally intense task of
-\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
-when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
-compare stability in column 5 above with column 3 above. \textbf{This
-comparison shows the high number of stable \(M\) that can be produced
-through a targetted search of \(\gamma\) values, and suggests that many
-otherwise unstable systems could potentially be stabilised by an
-informed manipulation of their component response times. Such a
-possibility might conceivably reduce the dimensionality of problems
-involving stability in social-ecological or economic systems.}
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
-\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{SI_science_files/figure-latex/unnamed-chunk-32-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
-
 \hypertarget{Feasibility}{\section{Feasibility of complex
 systems}\label{Feasibility}}
-
-For complex systems in which individual system components (\(S\))
-represent the density of some tangible quantity, it is important to
-consider the feasibility of the system. Feasibile equilibria assume that
-the values of all system components are positive at equilibrium (5, 18,
-21). This is of particular interest for ecological communities because
-population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically realistic
-(18). Consequently, the use of random matrices and traditional stability
-critiera for making inferences in theoretical analyses of species
-networks has recently been criticised (18). While the key results in the
-main text are intended to be general to all complex systems, and not
-restricted to species networks, I have also performed a feasibility
-analysis on all matrices \(M\). This analysis reveals that feasibility
-is not affected by \(Var(\gamma)\), meaning that for pure interacting
-species networks, variation in component response time (i.e., species
-generation time) does not affect stability at biologically realistic
-species densities. Nevertheless, ecological interactions do not exist in
-isolation in empirical systems, but instead interact with evolutionary
-(15), abiotic, or social-economic systems. The relevance of \(\gamma\)
-for complex system stability presented in the main text should therefore
-not be ignored in the broader context of ecological communities.
-
-Dougoud et al. (18) define the following feasibility criteria for
-ecological systems characterised by \(S\) interacting species with
-varying densities.
-
-\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]
-
-In the above, \(x^{*}\) is the vector of species abundances at
-equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
-The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
-off-diagonal elements), and the value \(\theta\) is strength of
-intraspecific competition (diagonal elements). As I have done elsewhere,
-diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the connectance of the community, which was set to \(C = 1\)
-throughout the manuscript and supplemental information, except
-\protect\hyperlink{connectance}{where otherwise noted}. The variable
-\(\delta\) is a normalisation parameter that modulates the strength of
-interactions (\(\sigma\) in the main text), which are held in \(A\). In
-the main text, implicitly, \(\delta = 0\) underlying strong
-interactions. Hence, the whole \((CS)^{-\delta} = 1\), so in the above,
-a diagonal matrix of -1s (\(\theta I\)) is added to \(A\), which has a
-diagonal of all zeros and an off-diagonal affecting species interactions
-(i.e., the expression \((CS)^{-\delta}\) relates to May's (1) stability
-criterion (19) by \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and
-hence \((CS)^{-\delta} = 1\) for the randomly simulated systems in the
-main text and supplemental information). The above criteria is therefore
-reduced to the below; note that the parenthetical in both equations
-produces an \(M\) matrix as used throughout the main text and
-supplemental information,
-
-\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]
-
-To check the feasibility criteria, I therefore inverted
-\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
-the resulting matrix by the vector of population growth rates \(r\).
-Feasibility is satisfied if all of the elements of the resulting vector
-are positive.
-
-The population growth rate for an individual species \(i\) is sampled
-from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
-shown in the \texttt{rand\_gen\_var} function in
-\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
-\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
-is assumed to be a species with a growth rate of \(r_{i}\).
 
@@
 
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-Allesina2012}{}
+1. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+\emph{Nature} \textbf{483,} 205--208 (2012).
+
 \end{document}

### 2018-07-27 notebook/ms_for_review.Rmd
Small changes to text

@@
 The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
+
+The while loop in `rand_mat_ga` continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed (by default, `converg = 0.01`).
+Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
+In mutation, any $\gamma_{i}$ values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
+Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
 

### 2018-07-28 notebook/SI.Rmd
Update manuscript for Science
Clean the notebook respository

@@
 
-While downloading this package is recommended, all relevant code is also reproduced below with explanation, so it is possible to recreate all analyses using only this supplemental information.
-
 ********************************************************************************
@@
 
-- [Further explanation of $\gamma$](#moregamma)
-- [Code and simulations underlying Fig. 1](#Fig1) 
-- [Code and simulations underlying Fig. 2](#Fig2) 
 - [Stability across increasing $S$](#IncrS) 
+- [Stability given targetted manipulation of $\gamma$ (genetic algorithm)](#ga)
 - [Stability of ecological networks](#ecological) 
@@
     - [Predator-prey networks](#pred-prey) 
-- [Different connectance (C) values](#connectance)
+- [Sensitivity of connectance (C) values](#connectance)
     - [C = 0.3](#connect3)
@@
     - [C = 0.9](#connect9)
-- [Different distributions of $\gamma$](#gam_dist)
-- [Genetic algorithm](#ga) 
+- [Large networks of $C = 0.05$ across $S$ and $\sigma$](#sigma)  
+    - [$\sigma$ = 0.3](#sigma3)
+    - [$\sigma$ = 0.4](#sigma4)
+    - [$\sigma$ = 0.5](#sigma5)
+    - [$\sigma$ = 0.6](#sigma6)
+- [Sensitivity of distribution of $\gamma$](#gam_dist)
 - [Feasibility of complex systems](#Feasibility)
-
-Further explanation of $\gamma$ {#moregamma}
-================================================================================
-
-In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
-model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
-For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
-
-$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
-
-$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
-
-In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
-Patel et al.
-were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
-The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
-
-I use the same principle that Patel et al.
-use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
-Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
-
-$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).$$
-
-In matrix notation[@May1973],
-
-$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).$$
-
-In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
-
-Code and simulations underlying Fig. 1 {#Fig1}
-================================================================================
-
-The sample $M$ used for the eigenvalue distributions in Fig.
-1 of the text is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma), and was produced by running the following function.
-
-```{r}
-find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
-    while(iters > 0){
-        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
-        A_mat  <- matrix(data = A_dat, nrow = S);
-        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
-        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
-        A_mat  <- A_mat * C_mat;
-        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
-        mu_gam <- mean(gammas);
-        diag(A_mat) <- -1;
-        A1     <- gammas * A_mat;
-        A0     <- mu_gam * A_mat;
-        A0_e   <- eigen(A0)$values;
-        A0_r   <- Re(A0_e);
-        A0_i   <- Im(A0_e);
-        A1_e   <- eigen(A1)$values;
-        A1_r   <- Re(A1_e);
-        A1_i   <- Im(A1_e);
-        if(max(A0_r) >= 0 & max(A1_r) < 0){
-            return(list(A0 = A0, A1 = A1));
-            break;
-        }
-        print(iters);
-        iters <- iters - 1;
-    }
-}
-```
-
-The above `find_bgamma` function terminates when a matrix $M$ is found that is not stable when all component response rates are set to $\gamma = 1$, but is stable when half of component response rates are $1.95$ and half are $0.05$.
-The function is used to illustrate the concept of how fast versus slow component responses can cause a system to become stable.
-Simulations were run for `iter = 1000000`, but terminated once an acceptable `A0` and `A1` were found.
-The code below plots the eigenvalue distributions of `A0` and `A1` in panels **a** and **b**, respectively.
-The plot itself can be recreated with the function and code below.
-
-```{r, echo = FALSE}
-A0 <- read.csv("sim_results/bi_gamma/S200_A0.csv");
-A1 <- read.csv("sim_results/bi_gamma/S200_A1.csv");
-```
-
-```{r, echo = TRUE, fig.height = 5, fig.width = 8}
-A0 <- as.matrix(A0[,-1]);
-A1 <- as.matrix(A1[,-1]);
-plot_Fig_1 <- function(A0, A1){
-    S_val       <- dim(A0)[1];
-    A0_e        <- eigen(A0)$values;
-    A0_r        <- Re(A0_e);
-    A0_i        <- Im(A0_e);
-    A1_e        <- eigen(A1)$values;
-    A1_r        <- Re(A1_e);
-    A1_i        <- Im(A1_e);
-    A0_vm       <- A0;
-    diag(A0_vm) <- NA;
-    A0vec       <- as.vector(t(A0_vm));
-    A0vec       <- A0vec[is.na(A0vec) == FALSE];
-    A1_vm       <- A1;
-    diag(A1_vm) <- NA;
-    A1vec       <- as.vector(t(A1_vm));
-    A1vec       <- A1vec[is.na(A1vec) == FALSE];
-    fhalf       <- 1:(0.5*length(A1vec));
-    shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-    plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
-    vl   <- seq(from = 0, to = 2*pi, by = 0.001);
-    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
-    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
-    text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "grey");
-    points(A0_r, A0_i, pch = 4, cex = 0.7);
-    plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-         col = "black", yaxt = "n");
-    vl <- seq(from = 0, to = 2*pi, by = 0.001);
-    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:(0.5*S_val)]);
-    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
-    points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
-    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
-                  mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
-    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
-    points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
-    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7);   
-    text(x = -3.5, y = 2.25, labels = "b", cex = 2);
-    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-}
-plot_Fig_1(A0 = A0, A1 = A1);
-```
-
-To find out how frequently $M$ was stable given that all $\gamma = 1$ versus $\gamma = \{1.95, 0.05\}$, the function below was created.
-
-```{r}
-stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
-    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
-    A0_count <- 0;
-    A1_count <- 0;
-    while(iters > 0){
-        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
-        A_mat  <- matrix(data = A_dat, nrow = S);
-        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
-        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
-        A_mat  <- A_mat * C_mat;
-        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
-        mu_gam <- mean(gammas);
-        diag(A_mat) <- -1;
-        A1     <- gammas * A_mat;
-        A0     <- mu_gam * A_mat;
-        A0_e   <- eigen(A0)$values;
-        A0_r   <- Re(A0_e);
-        A0_i   <- Im(A0_e);
-        A1_e   <- eigen(A1)$values;
-        A1_r   <- Re(A1_e);
-        A1_i   <- Im(A1_e);
-        if(max(A0_r) < 0){
-            ress[iters, 1] <- 1;
-            A0_count       <- A0_count + 1;
-            }
-        if(max(A1_r) < 0){
-            ress[iters, 2] <- 1;
-            A1_count       <- A1_count + 1;
-        }
-        print(c(iters, A0_count, A1_count));
-        iters <- iters - 1;
-    }
-    return(ress);
-}
-```
-
-The above functions produced the `bi_pr_st` data.
-
-```{r, echo = TRUE}
-bi_pr_st <- read.csv("sim_results/bi_gamma/bi_pr_st.csv");
-pr_st    <- bi_pr_st[,-1];
-```
-
-The function `stab_bgamma` was run for `iters = 1000000`, and the resulting matrix `ress` was returned.
-Each row of `ress` represents a single $M$ given $\gamma = 1$ (column 1) versus $\gamma = \{1.95, 0.05\}$ (column 2).
-Values of 0 indicate that $M$ was found to be unstable (at least one real component of its eigenvalues greater than or equal to zero), whereas values of 1 indicate that $M$ was found to be stable (all real components of eigenvalues are negative).
-The frequencies of stable $M$ were `r sum(pr_st[,1])` given $\gamma = 1$ and `r sum(pr_st[,2])` given $\gamma = \{1.95, 0.05\}$, as reported in the main text and legend of Fig.
-1 (raw data are [available on GitHub](https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv)).
-
-Code and simulations underlying Fig. 2 {#Fig2}
-===============================================================================
-
-Figure 2 of the main text shows eigenvalue distributions in a system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-Eigenvalues can be reproduced using the code below for when $\gamma = 1$ (panel a) and $\gamma \sim \mathcal{U}(0, 2)$ (panel b).
-The function below reproduces the figure.
-
-```{r, fig.height = 6, fig.width = 9}
-plot_Fig_2 <- function(){
-    A_comp <- NULL;
-    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
-    A_mat  <- matrix(data = A_dat, nrow = 1000);
-    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
-    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
-    A_mat     <- A_mat * C_mat;
-    gammas <- runif(n = 1000, min = 0, max = 2);
-    mu_gam <- mean(gammas);
-    diag(A_mat) <- -1;
-    A1     <- gammas * A_mat;
-    A0     <- mu_gam * A_mat;
-    A0_e   <- eigen(A0)$values;
-    A0_r   <- Re(A0_e);
-    A0_i   <- Im(A0_e);
-    A1_e   <- eigen(A1)$values;
-    A1_r   <- Re(A1_e);
-    A1_i   <- Im(A1_e);
-    A0_vm       <- A0;
-    diag(A0_vm) <- NA;
-    A0vec       <- as.vector(A0_vm);
-    A0vec       <- A0vec[is.na(A0vec) == FALSE];
-    A1_vm       <- A1;
-    diag(A1_vm) <- NA;
-    A1vec       <- as.vector(A1_vm);
-    A1vec       <- A1vec[is.na(A1vec) == FALSE];
-    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
-         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
-         asp = 1);
-    vl <- seq(from = 0, to = 2*pi, by = 0.001);
-    x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
-    y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
-    text(x = -15.5, y = 19, labels = "a", cex = 2);
-    points(x = x0, y = y0, type = "l", lwd = 3);
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
-         yaxt = "n");
-    text(x = -15.5, y = 19, labels = "b", cex = 2);
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
-    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
-    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-}
-plot_Fig_2();
-```
 
@@
 
-Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations. The data used to make this figure are read into R below.
-
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 dat <- read.csv(file = "sim_results/C_1/random_all.csv");
@@
 
+Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations.
 The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
@@
 Hence, more randomly created complex systems ($M$) are generated given variation in $\gamma$ than when $\gamma = 1$.
-The results underlying this table were produced with the `rand_gen_var` function below.
+Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).
 
-```{r}
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
-    tot_res <- NULL;
-    fea_res <- NULL;
-    for(i in 2:max_sp){
-        iter           <- iters;
-        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
-        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
-        while(iter > 0){
-            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
-            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
-            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
-            A0       <- species_interactions(mat = A0, type = int_type);
-            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
-            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
-            A0       <- A0 * C_mat;
-            diag(A0) <- -1;
-            gam1     <- runif(n = i, min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0       <- A0 * mean(gam1);
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
-            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
-            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
-            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
-            if(A0_stb == TRUE){
-                tot_res[[i-1]][iter, 1] <- 1;
+Stability given targetted manipulation of $\gamma$ (genetic algorithm) {#ga}
+================================================================================
+
+Figure 4 of the main text reports the number of stable random complex systems found over 100000 using the genetic algorithm to maximise stability with a vector $\mathbf{\gamma}$.
+Stability results for 100000 $M$ for each $S$ from 2-40 are shown below.
+Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
+
+```{r, echo = FALSE}
+get_top_evo_out <- function(evo_out, size){
+    highest <- max(evo_out);
+    gammas  <- NULL;
+    while(size > 0){
+        pos <- which(evo_out == highest);
+        len <- length(pos);
+        nli <- NULL;
+        for(i in 1:len){
+            start          <- pos[i] + 1;
+            end            <- pos[i] + highest;
+            gammas[[size]] <- evo_out[start:end];
+            size           <- size - 1;
+            if(size == 0){
+                break;
             }
-            if(A1_stb == TRUE){
-                tot_res[[i-1]][iter, 2] <- 1;
         }
-            if(A0_fea == TRUE){
-                fea_res[[i-1]][iter, 1] <- 1;
+        highest <- highest - 1;
     }
-            if(A1_fea == TRUE){
-                fea_res[[i-1]][iter, 2] <- 1;
+    return(gammas);
 }
-            iter    <- iter - 1;
+plot_evo_out <- function(evo_out){
+    gammas <- get_top_evo_out(evo_out, size = 9);
+    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
+    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xlim = c(0, 1.1));
+    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1.1));
+    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 1.1));
+    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
+          text = expression(paste("Component response rate (",gamma,")")));
+    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
 }
-        print(i);
-    }
-    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
-    return(all_res);
-}
+evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
+evo_res <- evo_res[,-1];
+kable(evo_res);
 ```
 
-The above function calls the two functions `species_interactions` and `summarise_randmat`, which are provided below.
+The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
+This comparison shows the high number of stable $M$ that can be produced through a targetted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
+Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.
 
-```{r}
-species_interactions <- function(mat, type = 0){
-    if(type == 1){
-        mat[mat > 0] <- -1*mat[mat > 0];
-    }
-    if(type == 2){
-        mat[mat < 0] <- -1*mat[mat < 0];
-    }
-    if(type == 3){
-        for(i in 1:dim(mat)[1]){
-            for(j in 1:dim(mat)[2]){
-                if(mat[i, j] * mat[j, i] > 0){
-                    mat[j, i] <- -1 * mat[j, i];
-                }
-            }
-        }
-    }
-    return(mat);
-}
+Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
 
-summarise_randmat <- function(tot_res, fea_res){
-    sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
-    for(i in 1:sims){
-        all_res[i, 1]  <- i + 1;
-        # Stable and unstable
-        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
-        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
-        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
-        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
-        # Stabilised and destabilised
-        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,2] == TRUE);
-        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
-                                  tot_res[[i]][,2] == FALSE);
-        # Feasible and infeasible
-        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
-        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
-        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
-        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
-        # Feased and defeased
-        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
-                                  fea_res[[i]][,2] == TRUE);
-        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
-                                  fea_res[[i]][,2] == FALSE);
-    }
-    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
-                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
-                "A0_feasible", "A1_infeasible", "A1_feasible", 
-                "A1_made_feasible", "A1_made_infeasible");
-    colnames(all_res) <- cnames;
-    return(all_res);
-}
+```{r, fig.height = 4, fig.width = 6, echo = FALSE}
+evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
+plot_evo_out(evo_out);
 ```
 
-Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).
+The distribution of $\gamma$ values found by the genetic algorithm is uniform.
+A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $M$.
 
@@
 And if $M$ was a predator-prey network, then all $i$ and $j$ pairs of elements were checked; any pairs of the same sign were changed so that one was negative and the other was positive.
-The `species_interaction` function used to do this is below.
 
-```{r}
+```{r, echo = FALSE}
 species_interactions <- function(mat, type = 0){
@@
 
-This function was applied to all created matrices $M$, then the number of stable $M$ matrices was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
-This produced the data set below.
+The number of stable $M$ matrices was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
 
-```{r}
+```{r, echo = FALSE}
 cdat <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
@@
 
-Overall, as expected[@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable.
+Overall, as expected [@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable.
 But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
-This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
+This is unsurprising because purely mutualist communities are characterised by a very positive (2) leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
 
-Different inter-connectivity (C) values {#connectance}
+Sensitivity of inter-connectivity (C) values {#connectance}
 ================================================================================
@@
 
-All data reported below for various values of $C$ are accessible using the below.
-
-```{r}
+```{r, echo = FALSE}
 C3dat <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
@@
 
-These objects `C3dat`, `C5dat`, `C7dat`, and `C9dat` include the results for $C = 0.3$, $C = 0.5$, $C = 0.7$, and $C = 0.9$, respectively.
-
-<a name="connect3">**Connectance $C = 0.3$**</a>
+<a name="connect3">**Connectance $\mathbf{C = 0.3}$**</a>
 
@@
 
-<a name="connect5">**Connectance $C = 0.5$**</a>
+<a name="connect5">**Connectance $\mathbf{C = 0.5}$**</a>
 
@@
 
-<a name="connect7">**Connectance $C = 0.7$**</a>
+<a name="connect7">**Connectance $\mathbf{C = 0.7}$**</a>
 
@@
 
-<a name="connect9">**Connectance $C = 0.9$**</a>
+<a name="connect9">**Connectance $\mathbf{C = 0.9}$**</a>
 
@@
 
-Different distributions of $\gamma$ {#gam_dist}
+Sensitivity of interaction strength ($\sigma$) values {#sigma}  
+================================================================================
+
+**ADD TEXT ABOUT COMPARISON WITH ALLESINA AND VALUES ACROSS SIGMA**
+
+```{r, echo = FALSE}
+S3dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt3.csv");
+S4dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt4.csv");
+S5dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt5.csv");
+S6dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt6.csv");
+```
+
+<a name="sigma3">**Interaction strength $\mathbf{\sigma = 0.3}$**</a>
+
+```{r, echo = FALSE}
+S3dat   <- S3dat[,-1];
+S3d_tab <- S3dat[,1:7];
+kable(S3d_tab);
+```
+
+<a name="sigma4">**Interaction strength $\mathbf{\sigma = 0.4}$**</a>
+
+```{r, echo = FALSE}
+S4dat   <- S4dat[,-1];
+S4d_tab <- S4dat[,1:7];
+kable(S4d_tab);
+```
+
+<a name="sigma5">**Interaction strength $\mathbf{\sigma = 0.5}$**</a>
+
+```{r, echo = FALSE}
+S5dat   <- S5dat[,-1];
+S5d_tab <- S5dat[,1:7];
+kable(S5d_tab);
+```
+
+<a name="sigma6">**Interaction strength $\mathbf{\sigma = 0.6}$**</a>
+
+```{r, echo = FALSE}
+S6dat   <- S6dat[,-1];
+S6d_tab <- S6dat[,1:7];
+kable(S6d_tab);
+```
+
+Sensitivity of distribution of $\gamma$ {#gam_dist}
 ================================================================================
@@
 
-Genetic algorithm {#ga}
-================================================================================
-
-Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $M$.
-This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
-Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
-In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
-The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
-The genetic algorithm relies on five functions.
-The first outer function `Evo_rand_gen_var` runs all of the simulations (`max_sp` refers to the maximum $S$ value simulated, and `iters` refers to the number of $M$ to try for each $S$).
-
-```{r}
-Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
-    tot_res <- NULL;
-    fea_res <- NULL;
-    for(i in 2:max_sp){
-        nn             <- i;
-        A1_stt         <- 0;
-        A2_stt         <- 0;
-        A1_fet         <- 0;
-        A2_fet         <- 0;
-        iter           <- iters;
-        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
-        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
-        while(iter > 0){
-            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
-            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
-            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
-            A0       <- species_interactions(mat = A0, type = int_type);
-            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
-            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
-            A0       <- A0 * C_mat;
-            diag(A0) <- -1;
-            gam1     <- runif(n = i, min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
-            A1_stb   <- rand_mat_ga(A1);
-            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
-            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
-            if(A0_stb == TRUE){
-                tot_res[[i-1]][iter, 1] <- 1;
-            }
-            if(A1_stb == TRUE){
-                tot_res[[i-1]][iter, 2] <- 1;
-            }
-            if(A0_fea == TRUE){
-                fea_res[[i-1]][iter, 1] <- 1;
-            }
-            if(A1_fea == TRUE){
-                fea_res[[i-1]][iter, 2] <- 1;
-            }
-            iter    <- iter - 1;
-        }
-        print(i);
-    }
-    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
-    return(all_res);
-}
-```
-
-Note that `Evo_rand_gen_var` calls three custom sub-functions, `species_interactions`, `rand_mat_ga`, and `summarise_randmat`.
-The first simply allows for non-random interactions between components (e.g., modelling [ecological interactions](#ecological) of random, competition, mutualism, or predator-prey).
-
-```{r}
-species_interactions <- function(mat, type = 0){
-    if(type == 1){
-        mat[mat > 0] <- -1*mat[mat > 0];
-    }
-    if(type == 2){
-        mat[mat < 0] <- -1*mat[mat < 0];
-    }
-    if(type == 3){
-        for(i in 1:dim(mat)[1]){
-            for(j in 1:dim(mat)[2]){
-                if(mat[i, j] * mat[j, i] > 0){
-                    mat[j, i] <- -1 * mat[j, i];
-                }
-            }
-        }
-    }
-    return(mat);
-}
-```
-
-The sub-function `rand_mat_ga` does the work of the genetic algorithm, searching for $\gamma$ vectors that are stabilising.
-
-```{r}
-rand_mat_ga <- function(A1, max_it = 20, converg = 0.01){
-    nn       <- dim(A1)[1];
-    rind     <- runif(n = nn*1000, min = 0, max = 1);
-    inds     <- matrix(data = rind, nrow = 1000, ncol = nn);
-    lastf    <- -10;
-    ccrit    <- 10;
-    find_st  <- 0;
-    iter     <- max_it;
-    while(iter > 0 & find_st < 1 & ccrit > converg){
-        ivar  <- rep(x = 0, length = dim(inds)[1]);
-        ifit  <- rep(x = 0, length = dim(inds)[1]);
-        isst  <- rep(x = 0, length = dim(inds)[1]);
-        for(i in 1:dim(inds)[1]){
-            ifit[i] <- -1*max(Re(eigen(inds[i,]*A1)$values));
-            ivar[i] <- var(inds[i,]);
-            isst[i] <- max(Re(eigen(inds[i,]*A1)$values)) < 0;
-        }
-        most_fit <- order(ifit, decreasing = TRUE)[1:20];
-        parents  <- inds[most_fit,];
-        new_gen  <- matrix(data = t(parents), nrow = 1000, ncol = nn, 
-                           byrow = TRUE);
-        mu_dat   <- rbinom(n = nn*1000, size = 1, prob = 0.2);
-        mu_dat2  <- rnorm(n = nn*1000, mean = 0, sd = 0.02);
-        mu_dat2[mu_dat2 < 0] <- -mu_dat2[mu_dat2 < 0];
-        mu_dat2[mu_dat2 > 2] <- 2;
-        mu_dat3  <- mu_dat * mu_dat2;
-        mu_mat   <- matrix(data = mu_dat3, nrow = 1000, ncol = nn);
-        new_gen  <- new_gen + mu_mat;
-        new_gen  <- crossover(inds = new_gen, pr = 0.1);
-        inds     <- new_gen;
-        find_st  <- max(isst);
-        newf     <- mean(ifit);
-        ccrit    <- newf - lastf;
-        lastf    <- newf;
-        iter     <- iter - 1;
-    }
-    if(find_st == 1){
-        s_row <- which(isst == 1)[1];
-        writt <- c(nn, inds[s_row,]);
-        cat(writt, file = "evo_out.txt", append = TRUE);
-        cat("\n", file = "evo_out.txt", append = TRUE);
-    }
-    return(find_st);
-}
-```
-
-The while loop in `rand_mat_ga` continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed (by default, `converg = 0.01`).
-Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
-In mutation, any $\gamma_{i}$ values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
-Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
-
-Crossover occurs in the `crossover` function below.
-
-```{r} 
-crossover <- function(inds, pr = 0.1){
-    crossed <- floor(dim(inds)[1] * pr);
-    cross1  <- sample(x = 1:dim(inds)[1], size = crossed);
-    cross2  <- sample(x = 1:dim(inds)[1], size = crossed);
-    for(i in 1:length(cross1)){
-        fromv   <- sample(x = 1:dim(inds)[2], size = 1);
-        tov     <- sample(x = 1:dim(inds)[2], size = 1);
-        temp                   <- inds[cross1[i],fromv:tov];
-        inds[cross1[i],fromv:tov] <- inds[cross2[i],fromv:tov];
-        inds[cross2[i],fromv:tov] <- temp;
-    }
-    return(inds);
-}
-```
-
-After all $M$ are simulated in `Evo_rand_gen_var`, the `summarise_randmat` formats the data into a table.
-
-```{r}
-summarise_randmat <- function(tot_res, fea_res){
-    sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 7);
-    for(i in 1:sims){
-        A0_unst   <- tot_res[[i]][,1] == FALSE;
-        A0_stbl   <- tot_res[[i]][,1] == TRUE;
-        A1_unst   <- tot_res[[i]][,2] == FALSE;
-        A1_stbl   <- tot_res[[i]][,2] == TRUE;
-        stabled   <- tot_res[[i]][,1] == FALSE & tot_res[[i]][,2] == TRUE;
-        unstabled <- tot_res[[i]][,1] == TRUE  & tot_res[[i]][,2] == FALSE;
-        all_res[i, 1]  <- i + 1;
-        all_res[i, 2]  <- sum(A0_unst);
-        all_res[i, 3]  <- sum(A0_stbl);
-        all_res[i, 4]  <- sum(A1_unst);
-        all_res[i, 5]  <- sum(A1_stbl);
-        all_res[i, 6]  <- sum(stabled);
-        all_res[i, 7]  <- sum(unstabled);
-    }
-    colnames(all_res) <- c("N", "A0_unstable", "A0_stable", "A1_unstable",
-                           "A1_stable", "A1_stabilised", "A1_destabilised");
-    return(all_res);
-}
-```
-
-Stability results for 40000 $M$ for each $S$ from 2-40 are shown below.
-Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
-
-```{r, echo = FALSE}
-get_top_evo_out <- function(evo_out, size){
-    highest <- max(evo_out);
-    gammas  <- NULL;
-    while(size > 0){
-        pos <- which(evo_out == highest);
-        len <- length(pos);
-        nli <- NULL;
-        for(i in 1:len){
-            start          <- pos[i] + 1;
-            end            <- pos[i] + highest;
-            gammas[[size]] <- evo_out[start:end];
-            size           <- size - 1;
-            if(size == 0){
-                break;
-            }
-        }
-        highest <- highest - 1;
-    }
-    return(gammas);
-}
-plot_evo_out <- function(evo_out){
-    gammas <- get_top_evo_out(evo_out, size = 9);
-    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
-    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xlim = c(0, 1.1));
-    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1.1));
-    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1.1));
-    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
-          text = expression(paste("Component response rate (",gamma,")")));
-    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
-}
-evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
-kable(evo_res);
-```
-
-The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
-Recall that 1 million random matrices were generated for the less computationally intense task of [comparing](#IncrS) $M$ when $\gamma = 1$ versus when $\gamma \sim \mathcal{U}(0, 2)$, so it is more informative to compare stability in column 5 above with column 3 above.
-**This comparison shows the high number of stable $M$ that can be produced through a targetted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
-Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.**
-
-Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
-
-```{r, fig.height = 6, fig.width = 7}
-evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
-plot_evo_out(evo_out);
-```
-
-The distribution of $\gamma$ values found by the genetic algorithm is uniform.
-A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $M$.
-
 Feasibility of complex systems {#Feasibility}
 ================================================================================
-
-For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
-Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
-Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
-While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
-This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
-Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
-The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
-
-Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
-
-$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
-
-In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive).
-The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements).
-As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
-The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
-The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
-In the main text, implicitly, $\delta = 0$ underlying strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
-The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
-
-$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
-
-To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$.
-Feasibility is satisfied if all of the elements of the resulting vector are positive.
-
-The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$, as shown in the `rand_gen_var` function in [the section](#IncrS) on "Stability across increasing $S$" above.
-Hence, each component $i$ of the complex system $M$ is assumed to be a species with a growth rate of $r_{i}$.
-Note that negative intrinsic growth rates are not unrealistic, and will occur in obligate mutualists in the absence of a partner.
 

### 2018-07-28 notebook/ms.Rmd
Clean the notebook respository

--- 2018-06-12 notebook/ms.Rmd
@@
     - --csl
-    - evolution.csl
+    - science.csl
     reference_docx: docx_template.docx
@@
 - \bibliographystyle{amnatnat}
-linestretch: 1
-link-citations: yes
+linestretch: 2
+link-citations: no
 linkcolor: blue
@@
 
+<!---
+
+Given the larger space of Nature Communications, one new figure will be used in the same style as that of Figure 3; the new Figure 4 will show the same results, but for sultions found with the genetic algorithm.
+
+Because Nature communcations is online, change all non-standard gamma results in figures to Red. The new genetic algorithm results will be in blue.
+
+--->
+
 ```{r, echo = FALSE}
@@
 
-********************************************************************************
+**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counter-intuitive result of widespread importance across the physical, life, and social sciences.
+Despite recent interest in the relationship between system properties and stability, the effect of variation in the response rate of individual system components remains unconsidered.
+Here I vary the component response rates ($\boldsymbol{\gamma}$) of randomly generated complex systems.
+I show that when component response rates vary, the potential for system stability is markedly increased.
+Variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, such that the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
+My results reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.**
+
+<!---
 
@@
 
+--->
+
 In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations.
-Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$)[@Allesina2012].
-May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ has profound consequences across multiple disciplines, raising the question of how complex systems in, e.g., ecology[@Mougi2012; @Allesina2012; @Allesina2015; @Grilli2017] or banking [@May2008; @Haldane2011; @Bardoscia2017] are predicted to persist or change.
+Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$) [@Allesina2012].
+May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks.
+As such, identifying general principles that drive stability in complex systems is of wide-ranging importance.
 
-Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., species[@Allesina2012] or banks[@Haldane2011]).
-One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium[@Allesina2012].
+Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., networks of species [@Allesina2012] or banks [@Haldane2011]).
+One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
 Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1[@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$)[@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly[@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$[@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1 [@May1972; @Allesina2012; @Allesina2015].
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
 Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma^{2}$ increase.
 
-The above stability criterion assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
@@
 
-********************************************************************************
+```{r, echo = FALSE}
+A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
+A1 <- as.matrix(A1[,-1]);
+
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(t(A0_vm));
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(t(A1_vm));
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+fhalf       <- 1:(0.5*length(A1vec));
+shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st       <- pr_st[,-1];
+```
+
+Results
+--------------------------------------------------------------------------------
+
+**Component response rates of random complex systems**.
+Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018].
+The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
+This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
+I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
+Rows are then multiplied by $\gamma_{i}$ to generate $M$.
+When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+1a).
+Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
+But when $\gamma_{i}$ values are separated into two groups, eigenvalues are no longer uniformly distributed (Fig.
+1b).
+Instead, two distinct clusters of eigenvalues appear (red circles in Fig.
+1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
+The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$) and all $\Re({\lambda}) < 0$.
+The latter cluster has real components that have shifted to the right, but has a smaller radius.
+Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+
+```{r echo = FALSE}
+S     <- 1000;
+C     <- 0.05;
+sigma <- 0.4;
+```
+
+```{r, echo = FALSE}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat     <- A_mat * C_mat;
+A_mat  <- species_interactions(A_mat, type = 0);
+gammas <- c(rep(0.05, 500), rep(1.95, 500));
+
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(A1_vm[1:500,]);
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+```
+
+Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
+I focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
+As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
+1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
+Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
+As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
+The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
+But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's [@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
+
+```{r, echo = FALSE}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+```
+
+**Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
+One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
+For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
+
+**Targetted manipulation of $\mathbf{\gamma}$**.
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively [@Hamblin2013]).
+[Need a methods like sentence here](http://google.com).
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
+4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+
+Discussion
+--------------------------------------------------------------------------------
+
+This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
+*Cite @Ahmadian2015 here*?
+
+It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
+Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
+This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
+1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
+
+I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks [@Allesina2012].
+These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
+In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising [@Allesina2011].
+When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
+Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
+
+Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+
+My results show that complex systems are more likely to be stable when the response rates of system components vary.
+These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
+
+Methods
+--------------------------------------------------------------------------------
+
+**Component response rate variation ($\mathbf{\gamma}$)**.
+In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
+
+$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+
+$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+
+In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
+Patel et al.
+were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
+The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
+
+I use the same principle that Patel et al.
+use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
+Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+
+$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).$$
+
+In matrix notation[@May1973],
+
+$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).$$
+
+In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
+Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
+
+**Genetic algorithm**.
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $M$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
+In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
+The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
+
+For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
+Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
+
+The while loop in `rand_mat_ga` continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed (by default, `converg = 0.01`).
+Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
+In mutation, any $\gamma_{i}$ values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
+Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
+
+**System feasibility**.
+For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
+Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
+Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
+While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
+This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
+
+Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
+
+$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+
+In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive).
+The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements).
+As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
+The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
+The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
+In the main text, implicitly, $\delta = 0$ underlying strong interactions.
+Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
+The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
+
+$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
+
+To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$.
+Feasibility is satisfied if all of the elements of the resulting vector are positive.
+
+The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$.
+
+<!---
+
+If more discussion is needed, then talk about relation to Jirsa and Ding 2004.
+
+--->
+
+**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
+Conversations with L.
+Bussi&#xe8;re and N.
+Bunnefeld, and comments from J.
+J.
+Cusack and I.
+L.
+Jones, improved the quality of this work.
+
+**References**
+
+<div id="refs"></div>
+
+```{r, echo = FALSE}
+# Temporary place for revised functions
+plot_Fig_1 <- function(A0, A1){
+    S_val       <- dim(A0)[1];
+    A0_e        <- eigen(A0)$values;
+    A0_r        <- Re(A0_e);
+    A0_i        <- Im(A0_e);
+    A1_e        <- eigen(A1)$values;
+    A1_r        <- Re(A1_e);
+    A1_i        <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(t(A0_vm));
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(t(A1_vm));
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    fhalf       <- 1:(0.5*length(A1vec));
+    shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+    vl   <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
+    text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(A0_r, A0_i, pch = 4, cex = 0.7);
+    plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+         col = "dodgerblue4", yaxt = "n");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + 
+                 mean(diag(A1)[1:(0.5*S_val)]);
+    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
+    #points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
+        mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
+    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
+    #points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7, col = "firebrick");   
+    text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_Fig_2 <- function(){
+    A_comp <- NULL;
+    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+    A_mat  <- matrix(data = A_dat, nrow = 1000);
+    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+    A_mat     <- A_mat * C_mat;
+    gammas <- runif(n = 1000, min = 0, max = 2);
+    mu_gam <- mean(gammas);
+    diag(A_mat) <- -1;
+    A1     <- gammas * A_mat;
+    A0     <- mu_gam * A_mat;
+    A0_e   <- eigen(A0)$values;
+    A0_r   <- Re(A0_e);
+    A0_i   <- Im(A0_e);
+    A1_e   <- eigen(A1)$values;
+    A1_r   <- Re(A1_e);
+    A1_i   <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(A0_vm);
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(A1_vm);
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "dodgerblue4");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+    text(x = -15.5, y = 19, labels = "a", cex = 2);
+    points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "firebrick", yaxt = "n");
+    text(x = -15.5, y = 19, labels = "b", cex = 2);
+    points(x = x1, y = y1, type = "l", col = "firebrick", lwd = 3);
+    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
+
+plot_stable_4 <- function(dat, S_s = 39){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 14), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 116), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
+```
+
+\clearpage
+
+```{r echo = FALSE}
+S     <- 200;
+C     <- 0.05;
+sigma <- 0.4;
+pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st <- pr_st[,-1];
+```
 
@@
 **Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the grey circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[fhalf]), digits = 3)` centred at $\gamma = -1.95$ and one small circle of $\sigma\sqrt{SC} =$ `r round(sqrt(100) * sd(A1vec[shalf]), digits = 3)` centred at $\gamma = -0.05$.
+**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the blue circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
 In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
@@
 
-********************************************************************************
-
 ```{r, echo = FALSE}
@@
 A_mat  <- matrix(data = A_dat, nrow = 1000);
-C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
 C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
 A_mat  <- A_mat * C_mat;
+A_mat  <- species_interactions(A_mat, type = 0);
 gammas <- runif(n = 1000, min = 0, max = 2);
+
 mu_gam <- mean(gammas);
@@
 A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(A1_vm);
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
 ```
 
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 1$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but clusters eigenvalues toward the distribution's centre (-1, 0).
-Black and red elipses in both panels show the circle centred on the distribution in panels a and b, respectively, which have a radius of $\sigma \sqrt{SC}$.
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
+Blue elipses in both panels show the circle centred on the distribution in panel a.
 Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
@@
 par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "dodgerblue4");
 vl <- seq(from = 0, to = 2*pi, by = 0.001);
@@
 y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
 text(x = -15.5, y = 19, labels = "a", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, col = "red",
-     yaxt = "n");
+points(x = x0, y = y0, type = "l", lwd = 3, col = "blue");
+points(A0_r, A0_i, col = "dodgerblue4",pch = 4, cex = 0.7);
+plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "firebrick", yaxt = "n");
 text(x = -15.5, y = 19, labels = "b", cex = 2);
-points(x = x1, y = y1, type = "l", col = "red", lwd = 3)
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid", col = "blue");
+points(A1_r, A1_i, col = "firebrick", pch = 4, cex = 0.7);
 mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
@@
 
-********************************************************************************
-
-**Figure 3: Stability of large complex systems with and without variation in component response rate($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
@@
 
-********************************************************************************
+\clearpage
 
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
-One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
-I found that the number of stable random systems was consistently higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
-3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
-This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
+**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targetted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targetted search of $\gamma$ values successfully resulted in system stability.
+For each $S$, 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
+Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
+The black line shows the proportion of systems that are stable when component response rate is targetted to increase stability, but would not be stable if $Var(\gamma) = 0$.
 
-It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
-Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
-This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively[@Hamblin2013]; see Supplementary Information).
-For each of 10000 random $M$, the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
-Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
-
-I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks[@Allesina2012].
-These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
-In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising[@Allesina2011].
-When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate[@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
-
-My results show that complex systems are more likely to be stable when the response rates of system components vary. These results are broadly applicable to complex biological and social networks.
-
-**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L.
-Bussi&#xe8;re and N.
-Bunnefeld, and helpful comments from J.
-J.
-Cusack and I.
-L.
-Jones, improved the quality of this work.
-
-********************************************************************************
-
-**References**
-
-<div id="refs"></div>
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
+dat <- dat[,-1];
+plot_stable_4(dat);
+```

### 2018-07-28 notebook/ms.Rmd
Fix most of up stuff except Discussion

@@
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
-1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after variation in component response rate.
+1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after the addition of variation in component response rate.
 Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
 As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues, some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues[@Ahmadian2015; @Gibbs2017; @Stone2017], some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
@@
 
-**Targetted manipulation of $\mathbf{\gamma}$**.
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm (the space of possible $\gamma$ values was too large to search exhaustively [@Hamblin2013]).
-[Need a methods like sentence here](http://google.com).
+**targeted manipulation of $\mathbf{\gamma}$**.
+To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm.
+Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively[@Hamblin2013].
+Up to 20 generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
@@
 
-Hence, managing the response rates of system components in a targetted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+Hence, managing the response rates of system components in a targeted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
 
@@
 --------------------------------------------------------------------------------
+
+<!--- USE J INSTEAD TO CLARIFY M --->
 
@@
 use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
-Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($m_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($a_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
 
-$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).$$
+$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).$$
 
@@
 
-$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).$$
+$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).$$
 
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{\gamma} \mathbf{M}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973].
+Therefore, $M =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973; @Ahmadian2015].
 
 **Genetic algorithm**.
-Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $M$.
-This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma$ to reasonable values would result in a search space too large to practically explore.
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
 Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
-In this case, the practical solution is finding vectors of $\gamma$ that decrease the most positive real eigenvalue of $M$.
-The genetic algorithm below achieves this by initialising a large population of 1000 different potential $\gamma$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\gamma$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met (in the below, this "giving up"" criteria is met when 20 generations pass, or if the fitness increase from one generation to the next is below a certain criteria).
+In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $M$.
+The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
 
 For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $M$ using each set of $\gamma$ values, and the 20 sets with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\gamma$ sets.
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or a convergence criteria of minimum fitness increase between generations was satisfied.
-
-The while loop in `rand_mat_ga` continues until either `iter` generations have occured, a solution $\gamma$ vector is found that results in all $\Re(\lambda) < 0$, or some criteria of minimum fitness increase is observed (by default, `converg = 0.01`).
-Within the genetic algorithm, $\gamma$ values are mutated, crossover occurs between $\gamma$ vectors, and selection occurs in each generation such that the 20 $\gamma$ vectors that produce the lowest maximum $\Re(\lambda)$ are allowed to have 50 offspring each.
-In mutation, any $\gamma_{i}$ values that mutate below zero are multiplied by $-1$, and any values that mutate above 2 are set to 2.
-Note also that if a solution is found, then one such $\gamma$ vector causing stability is printed to a file.
+Eigenanalysis was performed on $\mathbf{M = \gamma A}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutate below zero were multiplied by $-1$, and any values that mutate above 2 were set to 2).
+The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $-\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
 
@@
 
-**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targetted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targetted search of $\gamma$ values successfully resulted in system stability.
+**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
 For each $S$, 100000 complex systems are randomly generated.
@@
 Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
-The black line shows the proportion of systems that are stable when component response rate is targetted to increase stability, but would not be stable if $Var(\gamma) = 0$.
+The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.
 

### 2018-07-28 notebook/SI.Rmd
Minor changes

@@
 - [Stability across increasing $S$](#IncrS) 
-- [Stability given targetted manipulation of $\gamma$ (genetic algorithm)](#ga)
+- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
 - [Stability of ecological networks](#ecological) 
@@
 
-Stability given targetted manipulation of $\gamma$ (genetic algorithm) {#ga}
+Stability given targeted manipulation of $\gamma$ (genetic algorithm) {#ga}
 ================================================================================
@@
 The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
-This comparison shows the high number of stable $M$ that can be produced through a targetted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
+This comparison shows the high number of stable $M$ that can be produced through a targeted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
 Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.

### 2018-07-28 notebook/SI.tex
Update manuscript for Science
Minor changes

--- 2018-06-08 notebook/SI.tex
@@
 
-\hypertarget{moregamma}{\section{\texorpdfstring{Further explanation of
-\(\gamma\)}{Further explanation of \textbackslash{}gamma}}\label{moregamma}}
-
-In a synthesis of eco-evolutionary feedbacks on community stability,
-Patel et al. model a system that includes a vector of potentially
-changing species densities (\(\mathbf{N}\)) and a vector of potentially
-evolving traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
-For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
-
-\[\frac{dN_{i}(t)}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}(t)}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
-
-In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
-effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
-were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}, and
-this is modulated in the above by the scalar \(\epsilon\). The value of
-\(\epsilon\) thereby determines the timescale separation between ecology
-and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
-
-I use the same principle that Patel et al. use to modulate the relative
-rate of evolution to modulate rates of component responses for \(S\)
-components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}},
-the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
-by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
-\(i\) (\(m_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
-
-\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).\]
-
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{3}},
-
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{\gamma} \mathbf{M}\) modulates the values of components and
-can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}}.
-
-\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
-1}\label{Fig1}}
-
-The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
-text is available on
-\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/sim_results/bi_gamma}{GitHub},
-and was produced by running the following function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
-            \NormalTok{break;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(iters);}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
-is found that is not stable when all component response rates are set to
-\(\gamma = 1\), but is stable when half of component response rates are
-\(1.95\) and half are \(0.05\). The function is used to illustrate the
-concept of how fast versus slow component responses can cause a system
-to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
-but terminated once an acceptable \texttt{A0} and \texttt{A1} were
-found. The code below plots the eigenvalue distributions of \texttt{A0}
-and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
-plot itself can be recreated with the function and code below.
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-5-1.pdf}
-
-To find out how frequently \(M\) was stable given that all
-\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
-was created.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(ress);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above functions produced the \texttt{bi\_pr\_st} data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
-\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
-\end{Highlighting}
-\end{Shaded}
-
-The function \texttt{stab\_bgamma} was run for
-\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
-returned. Each row of \texttt{ress} represents a single \(M\) given
-\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
-Values of 0 indicate that \(M\) was found to be unstable (at least one
-real component of its eigenvalues greater than or equal to zero),
-whereas values of 1 indicate that \(M\) was found to be stable (all real
-components of eigenvalues are negative). The frequencies of stable \(M\)
-were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
-reported in the main text and legend of Fig. 1 (raw data are
-\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/sim_results/bi_gamma/bi_pr_st.csv}{available
-on GitHub}).
-
-\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
-2}\label{Fig2}}
-
-Figure 2 of the main text shows eigenvalue distributions in a system
-where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
-reproduced using the code below for when \(\gamma = 1\) (panel a) and
-\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
-reproduces the figure.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
-    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
-    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
-    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
-    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
-    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
-         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
-         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
-    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
-    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
-         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-\NormalTok{\}}
-\KeywordTok{plot_Fig_2}\NormalTok{();}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
-
 \hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
@@
 randomly created complex systems (\(M\)) are generated given variation
-in \(\gamma\) than when \(\gamma = 1\). The results underlying this
-table were produced with the \texttt{rand\_gen\_var} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above function calls the two functions
-\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
-are provided below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-
-\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \CommentTok{# Stable and unstable}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Stabilised and destabilised}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \CommentTok{# Feasible and infeasible}
-        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Feased and defeased}
-        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-    \NormalTok{\}}
-    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
-                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
-                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
-                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
-    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that feasibility results were ommited for the table above, but are
+in \(\gamma\) than when \(\gamma = 1\). Note that feasibility results
+were ommited for the table above, but are
 \protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targetted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targetted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+Figure 4 of the main text reports the number of stable random complex
+systems found over 100000 using the genetic algorithm to maximise
+stability with a vector \(\mathbf{\gamma}\). Stability results for
+100000 \(M\) for each \(S\) from 2-40 are shown below. Results for
+\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
+refers to systems in which the genetic algorithm searched for a set of
+\(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(M\) that can be produced through a targetted search of \(\gamma\)
+values, and suggests that many otherwise unstable systems could
+potentially be stabilised by an informed manipulation of their component
+response times. Such a possibility might conceivably reduce the
+dimensionality of problems involving stability in social-ecological or
+economic systems.
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
 
@@
 networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
-is negative and \(j\) on \(i\) is positive, or vice
-versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}--\protect\hyperlink{ref-Allesina2011}{7}}.
-In general, competitor and mutualist networks tend to be unstable, while
-predator-prey networks tend to be highly stabilising.
+is negative and \(j\) on \(i\) is positive, or vice versa) (2, 4, 17,
+20). In general, competitor and mutualist networks tend to be unstable,
+while predator-prey networks tend to be highly stabilising.
 
 I investigated competitor, mutualist, and predator-prey networks
-following Allesina et
-al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}. To create
-these networks, I first generated a random matrix \(M\), then changed
-the elements of \(M\) accordingly. If \(M\) was a competitive network,
-then the sign of any positive off-diagonal elements was reversed to be
-negative. If \(M\) was a mutualist network, then the sign of any
-positive off-diagonal elements was reversed to be positive. And if \(M\)
-was a predator-prey network, then all \(i\) and \(j\) pairs of elements
-were checked; any pairs of the same sign were changed so that one was
-negative and the other was positive. The \texttt{species\_interaction}
-function used to do this is below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
-\end{Highlighting}
-\end{Shaded}
-
-This function was applied to all created matrices \(M\), then the number
-of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
-as it was} in the main text for random matrices for values of \(S\) from
-2 to 50 (100 in the case of the relatively more stable predator-prey
-interactions), except that only 100000 random \(M\) were generated
-instead of 1 million. This produced the data set below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
-\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
-\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
+following Allesina et al. (2). To create these networks, I first
+generated a random matrix \(M\), then changed the elements of \(M\)
+accordingly. If \(M\) was a competitive network, then the sign of any
+positive off-diagonal elements was reversed to be negative. If \(M\) was
+a mutualist network, then the sign of any positive off-diagonal elements
+was reversed to be positive. And if \(M\) was a predator-prey network,
+then all \(i\) and \(j\) pairs of elements were checked; any pairs of
+the same sign were changed so that one was negative and the other was
+positive.
+
+The number of stable \(M\) matrices was estimated
+\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
+matrices for values of \(S\) from 2 to 50 (100 in the case of the
+relatively more stable predator-prey interactions), except that only
+100000 random \(M\) were generated instead of 1 million.
 
@@
 Overall, as
-expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}},
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{1}},
 predator-prey communities are relatively stable while mutualist
@@
 stabilise mutualist communities. This is unsurprising because purely
-mutualist communities are characterised by a very
-positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{4}}
-leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
-alone will shift all real parts of eigenvalues to negative values.
-
-\hypertarget{connectance}{\section{Different inter-connectivity (C)
+mutualist communities are characterised by a very positive (2) leading
+\(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\) alone
+will shift all real parts of eigenvalues to negative values.
+
+\hypertarget{connectance}{\section{Sensitivity of inter-connectivity (C)
 values}\label{connectance}}
@@
 
-\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}
-
-Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
-the proportion of stable complex systems, the search space of all
-possible \(\gamma\) vectors would be evaluated for each unique \(M\).
-This is technically impossible because \(\gamma_{i}\) can take any real
-value between 0-2, but even rounding \(\gamma\) to reasonable values
-would result in a search space too large to practically explore. Under
-these conditions, genetic algorithms are highly useful tools for finding
-practical solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{8}}. In
-this case, the practical solution is finding vectors of \(\gamma\) that
-decrease the most positive real eigenvalue of \(M\). The genetic
-algorithm below achieves this by initialising a large population of 1000
-different potential \(\gamma\) vectors and allowing this population to
-evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
-values between vectors), selection, and reproduction until either a
-\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
-``giving up'' critiera is met (in the below, this ``giving up''"
-criteria is met when 20 generations pass, or if the fitness increase
-from one generation to the next is below a certain criteria). The
-genetic algorithm relies on five functions. The first outer function
-\texttt{Evo\_rand\_gen\_var} runs all of the simulations
-(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
-\texttt{iters} refers to the number of \(M\) to try for each \(S\)).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
-        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
-\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
-\texttt{summarise\_randmat}. The first simply allows for non-random
-interactions between components (e.g., modelling
-\protect\hyperlink{ecological}{ecological interactions} of random,
-competition, mutualism, or predator-prey).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
-algorithm, searching for \(\gamma\) vectors that are stabilising.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
-    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
-    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
-    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
-    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
-    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
-        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
-            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
-            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{\}}
-        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
-        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
-                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
-        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
-        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
-        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
-        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
-        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
-        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
-        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
-        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
-        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
-        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
-        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
-        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(find_st);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The while loop in \texttt{rand\_mat\_ga} continues until either
-\texttt{iter} generations have occured, a solution \(\gamma\) vector is
-found that results in all \(\Re(\lambda) < 0\), or some criteria of
-minimum fitness increase is observed (by default,
-\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
-values are mutated, crossover occurs between \(\gamma\) vectors, and
-selection occurs in each generation such that the 20 \(\gamma\) vectors
-that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
-offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
-zero are multiplied by \(-1\), and any values that mutate above 2 are
-set to 2. Note also that if a solution is found, then one such
-\(\gamma\) vector causing stability is printed to a file.
-
-Crossover occurs in the \texttt{crossover} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
-    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
-    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
-        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
-        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
-        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(inds);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
-\texttt{summarise\_randmat} formats the data into a table.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{summarise_randmat_ga <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{10}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{unstables <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{stables   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{non_feas  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{feasibl   <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{unfeased  <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{feased    <-}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{foundd    <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{3}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstables);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stables);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(non_feas);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feasibl);}
-        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unfeased);}
-        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(feased);}
-        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(foundd);}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Some stability results from this table are shown below. Each histogram
-shows a different distribution of \(\gamma\) that was found to be
-stabilising.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 4 & 9996 & 4 & 9996 & 0 & 0\tabularnewline
-3 & 42 & 9958 & 42 & 9958 & 0 & 0\tabularnewline
-4 & 133 & 9867 & 133 & 9867 & 0 & 0\tabularnewline
-5 & 414 & 9586 & 411 & 9589 & 3 & 0\tabularnewline
-6 & 809 & 9191 & 799 & 9201 & 10 & 0\tabularnewline
-7 & 1380 & 8620 & 1339 & 8661 & 41 & 0\tabularnewline
-8 & 2074 & 7926 & 1927 & 8073 & 147 & 0\tabularnewline
-9 & 2885 & 7115 & 2503 & 7497 & 382 & 0\tabularnewline
-10 & 3842 & 6158 & 3158 & 6842 & 684 & 0\tabularnewline
-11 & 4867 & 5133 & 3613 & 6387 & 1255 & 1\tabularnewline
-12 & 5932 & 4068 & 4148 & 5852 & 1784 & 0\tabularnewline
-13 & 6937 & 3063 & 4470 & 5530 & 2468 & 1\tabularnewline
-14 & 7784 & 2216 & 4724 & 5276 & 3060 & 0\tabularnewline
-15 & 8519 & 1481 & 5086 & 4914 & 3433 & 0\tabularnewline
-16 & 9081 & 919 & 5262 & 4738 & 3819 & 0\tabularnewline
-17 & 9431 & 569 & 5368 & 4632 & 4063 & 0\tabularnewline
-18 & 9671 & 329 & 5571 & 4429 & 4100 & 0\tabularnewline
-19 & 9844 & 156 & 5807 & 4193 & 4037 & 0\tabularnewline
-20 & 9934 & 66 & 6133 & 3867 & 3801 & 0\tabularnewline
-21 & 6387 & 34 & 6421 & 3579 & 3545 & 0\tabularnewline
-22 & 6634 & 11 & 6645 & 3355 & 3344 & 0\tabularnewline
-23 & 7037 & 8 & 7045 & 2955 & 2947 & 0\tabularnewline
-24 & 7468 & 3 & 7471 & 2529 & 2526 & 0\tabularnewline
-25 & 7816 & 0 & 7816 & 2184 & 2184 & 0\tabularnewline
-26 & 8192 & 0 & 8192 & 1808 & 1808 & 0\tabularnewline
-27 & 8680 & 0 & 8680 & 1320 & 1320 & 0\tabularnewline
-28 & 8936 & 0 & 8936 & 1064 & 1064 & 0\tabularnewline
-29 & 9296 & 0 & 9296 & 704 & 704 & 0\tabularnewline
-30 & 9523 & 0 & 9523 & 477 & 477 & 0\tabularnewline
-31 & 9705 & 0 & 9705 & 295 & 295 & 0\tabularnewline
-32 & 9816 & 0 & 9816 & 184 & 184 & 0\tabularnewline
-33 & 9894 & 0 & 9894 & 106 & 106 & 0\tabularnewline
-34 & 9941 & 0 & 9941 & 59 & 59 & 0\tabularnewline
-35 & 9968 & 0 & 9968 & 32 & 32 & 0\tabularnewline
-36 & 9991 & 0 & 9991 & 9 & 9 & 0\tabularnewline
-37 & 9993 & 0 & 9993 & 7 & 7 & 0\tabularnewline
-38 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
-39 & 9999 & 0 & 9999 & 1 & 1 & 0\tabularnewline
-40 & 10000 & 0 & 10000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. Recall that 1 million random matrices were
-generated for the less computationally intense task of
-\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
-when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
-compare stability in column 5 above with column 3 above. \textbf{This
-comparison shows the high number of stable \(M\) that can be produced
-through a targetted search of \(\gamma\) values, and suggests that many
-otherwise unstable systems could potentially be stabilised by an
-informed manipulation of their component response times. Such a
-possibility might conceivably reduce the dimensionality of problems
-involving stability in social-ecological or economic systems.}
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
-\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
-
 \hypertarget{Feasibility}{\section{Feasibility of complex
 systems}\label{Feasibility}}
-
-For complex systems in which individual system components (\(S\))
-represent the density of some tangible quantity, it is important to
-consider the feasibility of the system. Feasibile equilibria assume that
-the values of all system components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{9}--\protect\hyperlink{ref-Song2018}{11}}.
-This is of particular interest for ecological communities because
-population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
-Consequently, the use of random matrices and traditional stability
-critiera for making inferences in theoretical analyses of species
-networks has recently been
-criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}.
-While the key results in the main text are intended to be general to all
-complex systems, and not restricted to species networks, I have also
-performed a feasibility analysis on all matrices \(M\). This analysis
-reveals that feasibility is not affected by \(Var(\gamma)\), meaning
-that for pure interacting species networks, variation in component
-response time (i.e., species generation time) does not affect stability
-at biologically realistic species densities. Nevertheless, ecological
-interactions do not exist in isolation in empirical systems, but instead
-interact with
-evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}},
-abiotic, or social-economic systems. The relevance of \(\gamma\) for
-complex system stability presented in the main text should therefore not
-be ignored in the broader context of ecological communities.
-
-Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}}
-define the following feasibility criteria for ecological systems
-characterised by \(S\) interacting species with varying densities.
-
-\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]
-
-In the above, \(x^{*}\) is the vector of species abundances at
-equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
-The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
-off-diagonal elements), and the value \(\theta\) is strength of
-intraspecific competition (diagonal elements). As I have done elsewhere,
-diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the inter-connectivity (i.e., `connectance') of the community,
-which was set to \(C = 1\) throughout the manuscript and supplemental
-information, except \protect\hyperlink{connectance}{where otherwise
-noted}. The variable \(\delta\) is a normalisation parameter that
-modulates the strength of interactions (\(\sigma\) in the main text),
-which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
-underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
-so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
-\(A\), which has a diagonal of all zeros and an off-diagonal affecting
-species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{2}} stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{10}} by
-\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
-\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
-text and supplemental information). The above criteria is therefore
-reduced to the below; note that the parenthetical in both equations
-produces an \(M\) matrix as used throughout the main text and
-supplemental information,
-
-\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]
-
-To check the feasibility criteria, I therefore inverted
-\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
-the resulting matrix by the vector of population growth rates \(r\).
-Feasibility is satisfied if all of the elements of the resulting vector
-are positive.
-
-The population growth rate for an individual species \(i\) is sampled
-from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
-shown in the \texttt{rand\_gen\_var} function in
-\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
-\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
-is assumed to be a species with a growth rate of \(r_{i}\). Note that
-negative intrinsic growth rates are not unrealistic, and will occur in
-obligate mutualists in the absence of a partner.
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Patel2018}{}
-1. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the effects
-of eco-evolutionary feedbacks on community stability. \emph{American
-Naturalist} \textbf{191,} 1--29 (2018).
-
-\hypertarget{ref-May1972}{}
-2. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
-
-\hypertarget{ref-May1973}{}
-3. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
-\textbf{54,} 638--641 (1973).
-
 \hypertarget{ref-Allesina2012}{}
-4. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+1. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
 \emph{Nature} \textbf{483,} 205--208 (2012).
 
-\hypertarget{ref-Allesina2015a}{}
-5. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
-(2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Tang2014b}{}
-6. Tang, S. \& Allesina, S. Reactivity and stability of large
-ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
-(2014).
-
-\hypertarget{ref-Allesina2011}{}
-7. Allesina, S. \& Levine, J. M. A competitive network theory of species
-diversity. \emph{Proceedings of the National Academy of Sciences of the
-United States of America} \textbf{108,} 5638--5642 (2011).
-
-\hypertarget{ref-Hamblin2013}{}
-8. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
-
-\hypertarget{ref-Grilli2017}{}
-9. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
-
-\hypertarget{ref-Dougoud2018}{}
-10. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
-
-\hypertarget{ref-Song2018}{}
-11. Song, C. \& Saavedra, S. Will a small randomly assembled community
-be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
-
 \end{document}

### 2018-07-28 notebook/ms.Rmd
Feasibility stuff

@@
 
-Randomly assembled complex systems can be represented as large square matrices ($M$) with $S$ components (e.g., networks of species [@Allesina2012] or banks [@Haldane2011]).
+Randomly assembled complex systems can be represented as large square matrices ($\mathbf{M}$) with $S$ components (e.g., networks of species [@Allesina2012] or banks [@Haldane2011]).
 One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
@@
 Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1 [@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $M$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
 In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
@@
 **Component response rates of random complex systems**.
-Rows in $M$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018].
+Rows in $\mathbf{M}$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018].
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
-An instructive example compares one $M$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $M$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
+An instructive example compares one $\mathbf{M}$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $\mathbf{M}$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
 This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
 I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
-Rows are then multiplied by $\gamma_{i}$ to generate $M$.
-When $\gamma_{i} = 1$, eigenvalues of $M$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
+Rows are then multiplied by $\gamma_{i}$ to generate $\mathbf{M}$.
+When $\gamma_{i} = 1$, eigenvalues of $\mathbf{M}$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
 1a).
@@
 The latter cluster has real components that have shifted to the right, but has a smaller radius.
-Overall, for 1 million randomly assembled $M$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
+Overall, for 1 million randomly assembled $\mathbf{M}$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
 
@@
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
-1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $M$ before and after the addition of variation in component response rate.
+1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $\mathbf{M}$ before and after the addition of variation in component response rate.
 Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $M$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
-One million $M$ were simulated for each $S$, and the stability of $M$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
+One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
 For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
@@
 
-**targeted manipulation of $\mathbf{\gamma}$**.
+**Targeted manipulation of $\mathbf{\gamma}$**.
 To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm.
@@
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
-4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $M$ elements.
+4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements.
+
+**System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
+Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
+While my key results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities[@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
+Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
 
@@
 
-It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $M$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
+It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
 Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
@@
 When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of $M$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
+Additionally, when each component of $\mathbf{M}$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
+
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
 
@@
 Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
-In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $M$.
+In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
 The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
 
-For each of 40000 random $M$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $\mathbf{M = \gamma A}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutate below zero were multiplied by $-1$, and any values that mutate above 2 were set to 2).
-The genetic algorithm terminated if a stable $M$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $-\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
+For each of 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma_{i} \sim \mathcal{U}(0, 2)$ values of size $S$.
+Eigenanalysis was performed on $\mathbf{M = \gamma A}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M = \gamma A}$).
 
-**System feasibility**.
-For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is important to consider the feasibility of the system.
-Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
-Consequently, the use of random matrices and traditional stability critiera for making inferences in theoretical analyses of species networks has recently been criticised[@Dougoud2018].
-While the key results in the main text are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $M$.
-This analysis reveals that feasibility is not affected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
-Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
-The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
+**System feasibility**. Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
 
-Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
+$$\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.$$
 
-$$x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.$$
+In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium (for feasibility, all values in $\mathbf{x^{*}}$ must be positive).
+The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is strength of intraspecific competition (diagonal elements).
+Diagonal values are set to $-1$, so $\theta = -1$.
+The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$) for $\mathbf{J}$.
+Implicitly, here $\delta = 0$ underlying strong interactions.
+Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence for my purposes $(CS)^{-\delta} = 1$).
+Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below,
 
-In the above, $x^{*}$ is the vector of species abundances at equilibrium (for feasibility, all values in $x^{*}$ must be positive).
-The matrix $I$ is the identity matrix (1s on the diagonal, 0s on the off-diagonal elements), and the value $\theta$ is strength of intraspecific competition (diagonal elements).
-As I have done elsewhere, diagonal values are set to $-1$, so $\theta = -1$.
-The variable $C$ is the inter-connectivity (i.e., 'connectance') of the community, which was set to $C = 1$ throughout the manuscript and supplemental information, except [where otherwise noted](#connectance).
-The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$ in the main text), which are held in $A$.
-In the main text, implicitly, $\delta = 0$ underlying strong interactions.
-Hence, the whole $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta I$) is added to $A$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence $(CS)^{-\delta} = 1$ for the randomly simulated systems in the main text and supplemental information).
-The above criteria is therefore reduced to the below; note that the parenthetical in both equations produces an $M$ matrix as used throughout the main text and supplemental information,
+$$\mathbf{x^{*} = -A^{-1}r}.$$
 
-$$x^{*} = -\left(\theta I + A\right)^{-1}r.$$
-
-To check the feasibility criteria, I therefore inverted $M = (\theta I + A)$ and multiplied elements by -1, then multiplied the resulting matrix by the vector of population growth rates $r$.
+To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were randomly sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
 Feasibility is satisfied if all of the elements of the resulting vector are positive.
-
-The population growth rate for an individual species $i$ is sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$.
 

### 2018-07-28 notebook/ms.Rmd
All but Discussion

@@
 Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively[@Hamblin2013].
-Up to 20 generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
+Selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
@@
 
-**System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components ($S$) represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
+**System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
 Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
@@
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $M =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973; @Ahmadian2015].
+Therefore, $\mathbf{M} =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973; @Ahmadian2015].
 
@@
 
-For each of 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), the genetic algorithm initialised 1000 different sets of $\gamma_{i} \sim \mathcal{U}(0, 2)$ values of size $S$.
-Eigenanalysis was performed on $\mathbf{M = \gamma A}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
-The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M = \gamma A}$).
+For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), where $\mathbf{M = \gamma A}$.
+The genetic algorithm was initialised with a populatoin of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
 
-**System feasibility**. Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities.
+**System feasibility**.
+Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
 
@@
 
-In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium (for feasibility, all values in $\mathbf{x^{*}}$ must be positive).
+In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium (feasibility is satisfied if all values in $\mathbf{x^{*}}$ are positive).
 The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is strength of intraspecific competition (diagonal elements).
@@
 
-To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were randomly sampled from a normal distribution of $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
+To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were sampled i.i.d.
+from $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
 Feasibility is satisfied if all of the elements of the resulting vector are positive.
@@
 Jones, improved the quality of this work.
+
+**Supplementary Information:** Full tables of stability results for simulations across different system size ($S$) values, ecological community types, connectance ($C$) values, interaction strengths ($\sigma$), and $\gamma$ distributions are provided as supplementary material.
+An additional table also shows results for how feasibility changes across $S$.
+All code and simulation outputs are publicly available as part of the RandomMatrixStability package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
 

### 2018-07-28 notebook/SI.Rmd
Fix SI

@@
 
-**This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with all of the code required to recreate the analysis in the main text, and with additional analyses to support its conclusions.
+**This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with additional analyses to support its conclusions.
 All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability package.**
 
-The [RandomMatrixStability package](https://github.com/bradduthie/RandomMatrixStability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for functions as part of the package.
+The [RandomMatrixStability package](https://github.com/bradduthie/RandomMatrixStability) includes all functions and tools for recreating the text, this supplemental information, and running all code; additional documentation is also provided for package functions.
 The RandomMatrixStability package is available on [GitHub](https://github.com/bradduthie/RandomMatrixStability); to download it, the  [`devtools` library](https://cran.r-project.org/web/packages/devtools/index.html) is needed.
@@
 Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations.
-The table below shows the results for all simulations of random $M$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
+The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
 In this table, the `A0` refers to matrices where $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
-Each row summarises data for a given $S$ over 1 million randomly simulated $M$ (`A0` and `A1`).
+Each row summarises data for a given $S$ over 1 million randomly simulated $\mathbf{M}$ (`A0` and `A1`).
 The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 1 million).
@@
 Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
-Hence, more randomly created complex systems ($M$) are generated given variation in $\gamma$ than when $\gamma = 1$.
+Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$.
 Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).
@@
 Figure 4 of the main text reports the number of stable random complex systems found over 100000 using the genetic algorithm to maximise stability with a vector $\mathbf{\gamma}$.
-Stability results for 100000 $M$ for each $S$ from 2-40 are shown below.
+Stability results for 100000 $\mathbf{M}$ for each $S$ from 2-40 are shown below.
 Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
@@
 The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
-This comparison shows the high number of stable $M$ that can be produced through a targeted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
+This comparison shows the high number of stable $\mathbf{M}$ that can be produced through a targeted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
 Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.
@@
 The distribution of $\gamma$ values found by the genetic algorithm is uniform.
-A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $M$.
+A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $\mathbf{M}$.
 
@@
 
-While the foundational work of May (1) applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
-In these networks, the matrix $M$ is interpreted as a community matrix and each row and column is interpreted as a single species.
-The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $M$ holds the effects of pair-wise interactions between $S$ species (2, 4, 20).
-While May's original work (1) considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $M$ are negative), mutualist networks (all off-diagonal elements of $M$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa) (2, 4, 17, 20).
+While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
+In these networks, the matrix $\mathbf{M}$ is interpreted as a community matrix and each row and column is interpreted as a single species.
+The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $\mathbf{M}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015].
+While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{M}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{M}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015].
 In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.
 
-I investigated competitor, mutualist, and predator-prey networks following Allesina et al.
-(2).
-To create these networks, I first generated a random matrix $M$, then changed the elements of $M$ accordingly.
-If $M$ was a competitive network, then the sign of any positive off-diagonal elements was reversed to be negative.
-If $M$ was a mutualist network, then the sign of any positive off-diagonal elements was reversed to be positive.
-And if $M$ was a predator-prey network, then all $i$ and $j$ pairs of elements were checked; any pairs of the same sign were changed so that one was negative and the other was positive.
+I investigated competitor, mutualist, and predator-prey networks following Allesina et al.[@Allesina2012].
+To create these networks, I first generated a random matrix $\mathbf{A}$, then changed the elements of $\mathbf{A}$ accordingly.
+If $\mathbf{A}$ was a competitive network, then the sign of any positive off-diagonal elements was reversed to be negative.
+If $\mathbf{A}$ was a mutualist network, then the sign of any positive off-diagonal elements was reversed to be positive.
+And if $\mathbf{A}$ was a predator-prey network, then all $i$ and $j$ pairs of elements were checked; any pairs of the same sign were changed so that one was negative and the other was positive.
 
@@
 
-The number of stable $M$ matrices was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $M$ were generated instead of 1 million.
+The number of stable $\mathbf{M = \gamma A}$ systems was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $\mathbf{M}$ were generated instead of 1 million.
 
@@
 
-The following tables for restricted ecological communities can therefore be compared with the random $M$ [results above](#IncrS) (but note that counts from systems with comparable probabilities of stability will be an order of magnitude lower in the tables below due to the smaller number of $M$ matrices generated).
+The following tables for restricted ecological communities can therefore be compared with the random $\mathbf{M}$ [results above](#IncrS) (but note that counts from systems with comparable probabilities of stability will be an order of magnitude lower in the tables below due to the smaller number of $\mathbf{M}$ matrices generated).
 As with the [results above](#IncrS), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added.
@@
 But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
-This is unsurprising because purely mutualist communities are characterised by a very positive (2) leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
+This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
 
@@
 
-In the main text, for simplicity, I assumed inter-connectivity values of $C = 1$, meaning that all off-diagonal elements of a matrix $M$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
+In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $\mathbf{M}$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
 Here I present four tables showing the number of stable communities given $C = \{0.3, 0.
@@
 In all cases, uniform variation in component response time ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
-In contrast to the main text, 100000 rather than 1 million $M$ were simulated.
+In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.
 As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added.
@@
 
-**ADD TEXT ABOUT COMPARISON WITH ALLESINA AND VALUES ACROSS SIGMA**
+Results below show stability results given varying interaction strengths ($\sigma$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows).
+In the tables below (as [above](#IncrS)), `A0` and `A1` refers to matrices for $\gamma = 1$ and $Var(\gamma)$, respectively.
 
@@
 In the main text, I considered a uniform distribution of component response rates $\gamma \sim \mathcal{U}(0, 2)$.
-The number of unstable and stable $M$ matrices are reported in [a table above](#IncrS) across different values of $S$.
+The number of unstable and stable $\mathbf{M}$ matrices are reported in [a table above](#IncrS) across different values of $S$.
 Here I show complementary results for three different distributions including an exponential, beta, and gamma distribution of $\gamma$ values.
@@
 
-The same 100000 $M$ matrices were used to investigate stability when applying each of these different distributions of $\gamma$ values.
-The table below shows the number of $M$ that were unstable (_unst) and stable (_stbl) for the exponential (Exp), beta, and gamma distributions.
+The same 100000 $\mathbf{M}$ matrices were used to investigate stability when applying each of these different distributions of $\gamma$ values.
+The table below shows the number of $\mathbf{M}$ that were unstable (_unst) and stable (_stbl) for the exponential (Exp), beta, and gamma distributions.
 
@@
 
-When feasibility was evaluated with and without variation in $\gamma$, there was no increase in stability for $M$ where $\gamma$ varied as compared to where $\gamma = 1$.
+When feasibility was evaluated with and without variation in $\gamma$, there was no increase in stability for $\mathbf{M}$ where $\gamma$ varied as compared to where $\gamma = 1$.
 Results below illustrate this result, which was general to all other simulations performed.

### 2018-07-28 notebook/SI.tex
Fix SI

@@
 ``Component response rate variation drives stability in large complex
-systems'' with all of the code required to recreate the analysis in the
-main text, and with additional analyses to support its conclusions. All
-text, code, and data underlying this manuscript are publicly available
-on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
+systems'' with additional analyses to support its conclusions. All text,
+code, and data underlying this manuscript are publicly available on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
 part of the RandomMatrixStability package.}
@@
 supplemental information, and running all code; additional documentation
-is also provided for functions as part of the package. The
-RandomMatrixStability package is available on
+is also provided for package functions. The RandomMatrixStability
+package is available on
 \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
@@
 \item
-  \protect\hyperlink{ga}{Stability given targetted manipulation of
+  \protect\hyperlink{ga}{Stability given targeted manipulation of
   \(\gamma\) (genetic algorithm)}
@@
 systems found over 1 million iterations. The table below shows the
-results for all simulations of random \(M\) matrices at \(\sigma = 0.4\)
-and \(C = 1\) given a range of \(S = \{2, 3, ..., 49, 50\}\). In this
-table, the \texttt{A0} refers to matrices where \(\gamma = 1\), while
-\texttt{A1} refers to matrices after \(Var(\gamma)\) is added and
-\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
-\(S\) over 1 million randomly simulated \(M\) (\texttt{A0} and
-\texttt{A1}). The column \texttt{A0\_unstable} shows the number of
-\texttt{A0} matrices that are unstable, and the column
-\texttt{A0\_stable} shows the number of \texttt{A0} matrices that are
-stable (these two columns sum to 1 million). Similarly, the column
-\texttt{A1\_unstable} shows the number of \texttt{A1} matrices that are
-unstable and \texttt{A1\_stable} shows the number that are stable. The
-columns \texttt{A1\_stabilised} and \texttt{A1\_destabilised} show how
-many \texttt{A0} matrices were stabilised or destabilised, respectively,
-by \(Var(\gamma)\).
+results for all simulations of random \(\mathbf{M}\) matrices at
+\(\sigma = 0.4\) and \(C = 1\) given a range of
+\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
+matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
+after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
+Each row summarises data for a given \(S\) over 1 million randomly
+simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
+\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
+unstable, and the column \texttt{A0\_stable} shows the number of
+\texttt{A0} matrices that are stable (these two columns sum to 1
+million). Similarly, the column \texttt{A1\_unstable} shows the number
+of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
+the number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
 
@@
 this ratio increases with increasing \(S\) (column 1). Hence, more
-randomly created complex systems (\(M\)) are generated given variation
-in \(\gamma\) than when \(\gamma = 1\). Note that feasibility results
-were ommited for the table above, but are
+randomly created complex systems (\(\mathbf{M}\)) are stable given
+variation in \(\gamma\) than when \(\gamma = 1\). Note that feasibility
+results were ommited for the table above, but are
 \protect\hyperlink{Feasibility}{reported below}.
 
-\hypertarget{ga}{\section{\texorpdfstring{Stability given targetted
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
 manipulation of \(\gamma\) (genetic
-algorithm)}{Stability given targetted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
 
@@
 stability with a vector \(\mathbf{\gamma}\). Stability results for
-100000 \(M\) for each \(S\) from 2-40 are shown below. Results for
-\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
-refers to systems in which the genetic algorithm searched for a set of
-\(\gamma\) values that stabilised the system.
+100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
+for \texttt{A0} indicate systems in which \(\gamma = 1\), while
+\texttt{A1} refers to systems in which the genetic algorithm searched
+for a set of \(\gamma\) values that stabilised the system.
 
@@
 values are shown below. This comparison shows the high number of stable
-\(M\) that can be produced through a targetted search of \(\gamma\)
-values, and suggests that many otherwise unstable systems could
-potentially be stabilised by an informed manipulation of their component
-response times. Such a possibility might conceivably reduce the
-dimensionality of problems involving stability in social-ecological or
-economic systems.
+\(\mathbf{M}\) that can be produced through a targeted search of
+\(\gamma\) values, and suggests that many otherwise unstable systems
+could potentially be stabilised by an informed manipulation of their
+component response times. Such a possibility might conceivably reduce
+the dimensionality of problems involving stability in social-ecological
+or economic systems.
 
@@
 values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
+of \(\gamma\) is likely to be found to stabilise a matrix
+\(\mathbf{M}\).
 
@@
 
-While the foundational work of May (1) applies broadly to complex
-networks, much attention has been given specifically to ecological
-networks of interacting species. In these networks, the matrix \(M\) is
-interpreted as a community matrix and each row and column is interpreted
-as a single species. The effect that the density of any species \(i\)
-has on the population dynamics of species \(j\) is found in \(M_{ij}\),
-meaning that \(M\) holds the effects of pair-wise interactions between
-\(S\) species (2, 4, 20). While May's original work (1) considered only
+While the foundational work of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
+to complex networks, much attention has been given specifically to
+ecological networks of interacting species. In these networks, the
+matrix \(\mathbf{M}\) is interpreted as a community matrix and each row
+and column is interpreted as a single species. The effect that the
+density of any species \(i\) has on the population dynamics of species
+\(j\) is found in \(M_{ij}\), meaning that \(\mathbf{M}\) holds the
+effects of pair-wise interactions between \(S\)
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
+While May's original
+work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
 randomly assembled communities, recent work has specifically looked at
 more restricted ecological communities including competitive networks
-(all off-diagonal elements of \(M\) are negative), mutualist networks
-(all off-diagonal elements of \(M\) are positive), and predator-prey
-networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
-is negative and \(j\) on \(i\) is positive, or vice versa) (2, 4, 17,
-20). In general, competitor and mutualist networks tend to be unstable,
-while predator-prey networks tend to be highly stabilising.
+(all off-diagonal elements of \(\mathbf{M}\) are negative), mutualist
+networks (all off-diagonal elements of \(\mathbf{M}\) are positive), and
+predator-prey networks (for any pair of \(i\) and \(j\), the effect of
+\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
+versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
+In general, competitor and mutualist networks tend to be unstable, while
+predator-prey networks tend to be highly stabilising.
 
 I investigated competitor, mutualist, and predator-prey networks
-following Allesina et al. (2). To create these networks, I first
-generated a random matrix \(M\), then changed the elements of \(M\)
-accordingly. If \(M\) was a competitive network, then the sign of any
-positive off-diagonal elements was reversed to be negative. If \(M\) was
-a mutualist network, then the sign of any positive off-diagonal elements
-was reversed to be positive. And if \(M\) was a predator-prey network,
-then all \(i\) and \(j\) pairs of elements were checked; any pairs of
-the same sign were changed so that one was negative and the other was
-positive.
-
-The number of stable \(M\) matrices was estimated
+following Allesina et
+al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
+these networks, I first generated a random matrix \(\mathbf{A}\), then
+changed the elements of \(\mathbf{A}\) accordingly. If \(\mathbf{A}\)
+was a competitive network, then the sign of any positive off-diagonal
+elements was reversed to be negative. If \(\mathbf{A}\) was a mutualist
+network, then the sign of any positive off-diagonal elements was
+reversed to be positive. And if \(\mathbf{A}\) was a predator-prey
+network, then all \(i\) and \(j\) pairs of elements were checked; any
+pairs of the same sign were changed so that one was negative and the
+other was positive.
+
+The number of stable \(\mathbf{M = \gamma A}\) systems was estimated
 \protect\hyperlink{IncrS}{exactly as it was} in the main text for random
@@
 relatively more stable predator-prey interactions), except that only
-100000 random \(M\) were generated instead of 1 million.
+100000 random \(\mathbf{M}\) were generated instead of 1 million.
 
 The following tables for restricted ecological communities can therefore
-be compared with the random \(M\) \protect\hyperlink{IncrS}{results
-above} (but note that counts from systems with comparable probabilities
-of stability will be an order of magnitude lower in the tables below due
-to the smaller number of \(M\) matrices generated). As with the
+be compared with the random \(\mathbf{M}\)
+\protect\hyperlink{IncrS}{results above} (but note that counts from
+systems with comparable probabilities of stability will be an order of
+magnitude lower in the tables below due to the smaller number of
+\(\mathbf{M}\) matrices generated). As with the
 \protect\hyperlink{IncrS}{results above}, in the tables below,
@@
 Overall, as
-expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{1}},
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
 predator-prey communities are relatively stable while mutualist
@@
 stabilise mutualist communities. This is unsurprising because purely
-mutualist communities are characterised by a very positive (2) leading
-\(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\) alone
-will shift all real parts of eigenvalues to negative values.
+mutualist communities are characterised by a very
+positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
+alone will shift all real parts of eigenvalues to negative values.
 
@@
 
-In the main text, for simplicity, I assumed inter-connectivity values of
-\(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
-potentially nonzero and sampled from a normal distribution
-\(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here I present
-four tables showing the number of stable communities given
+In the main text, for simplicity, I assumed connectance values of
+\(C = 1\), meaning that all off-diagonal elements of a matrix
+\(\mathbf{M}\) were potentially nonzero and sampled from a normal
+distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
+I present four tables showing the number of stable communities given
 \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
@@
 (\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
-million \(M\) were simulated. As with the results on
+million \(\mathbf{M}\) were simulated. As with the results on
 \protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
@@
 
-\textbf{ADD TEXT ABOUT COMPARISON WITH ALLESINA AND VALUES ACROSS SIGMA}
+Results below show stability results given varying interaction strengths
+(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
+larger and increase by 10 with increasing rows). In the tables below (as
+\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
+matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
 
@@
 response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
-and stable \(M\) matrices are reported in \protect\hyperlink{IncrS}{a
-table above} across different values of \(S\). Here I show complementary
-results for three different distributions including an exponential,
-beta, and gamma distribution of \(\gamma\) values. The shape of these
-distributions is shown in the figure below.
+and stable \(\mathbf{M}\) matrices are reported in
+\protect\hyperlink{IncrS}{a table above} across different values of
+\(S\). Here I show complementary results for three different
+distributions including an exponential, beta, and gamma distribution of
+\(\gamma\) values. The shape of these distributions is shown in the
+figure below.
 
@@
 
-The same 100000 \(M\) matrices were used to investigate stability when
-applying each of these different distributions of \(\gamma\) values. The
-table below shows the number of \(M\) that were unstable (\_unst) and
-stable (\_stbl) for the exponential (Exp), beta, and gamma
-distributions.
+The same 100000 \(\mathbf{M}\) matrices were used to investigate
+stability when applying each of these different distributions of
+\(\gamma\) values. The table below shows the number of \(\mathbf{M}\)
+that were unstable (\_unst) and stable (\_stbl) for the exponential
+(Exp), beta, and gamma distributions.
 
@@
 When feasibility was evaluated with and without variation in \(\gamma\),
-there was no increase in stability for \(M\) where \(\gamma\) varied as
-compared to where \(\gamma = 1\). Results below illustrate this result,
-which was general to all other simulations performed.
+there was no increase in stability for \(\mathbf{M}\) where \(\gamma\)
+varied as compared to where \(\gamma = 1\). Results below illustrate
+this result, which was general to all other simulations performed.
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-May1972}{}
+1. May, R. M. Will a large complex system be stable? \emph{Nature}
+\textbf{238,} 413--414 (1972).
+
 \hypertarget{ref-Allesina2012}{}
-1. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
 \emph{Nature} \textbf{483,} 205--208 (2012).
 
+\hypertarget{ref-Allesina2015}{}
+3. Allesina, S. \emph{et al.} Predicting the stability of large
+structured food webs. \emph{Nature Communications} \textbf{6,} 7842
+(2015).
+
 \end{document}

### 2018-07-28 notebook/ms.Rmd
First Discussion sentence

@@
 
-This suggests that the stability of large systems might be dependent upon variation in the response rate of their individual components, meaning that factors such as generation time (in ecological networks), transaction speed (in economic networks), or communication speed (in social networks) needs to be considered when investigating the stability of complex systems.
+Here I have shown that the stability of large systems might be contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
 *Cite @Ahmadian2015 here*?

### 2018-07-28 notebook/ms.Rmd
More discussion

@@
 Here I have shown that the stability of large systems might be contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
-*Cite @Ahmadian2015 here*?
 
-It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable (see Supplementary Information).
+It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable.
 Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
@@
 1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
+The mathematics underlying this shift in eigenvalue distribution have been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but have not been interpreted as rates of response of individual componetents to system perturbation.
+
+Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
+The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
 
@@
 Additionally, when each component of $\mathbf{M}$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
-
-Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
-The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
 

### 2018-07-29 notebook/ms.Rmd
Some major edits to all but the Discussion

@@
 Similarly, the speed at which individual banks respond to perturbations in financial networks, or individuals or institutions respond to perturbations in complex social networks, is likely to vary.
-The effect of such variance has not been investigated in complex systems theory.
+The effect of such variance on stability has not been investigated in complex systems theory.
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-32.
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-50.
 One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
 For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
-3), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$ (see Supplementary Information for full results).
+3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
+For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$.
 
@@
 Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively[@Hamblin2013].
-Selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
+Generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
@@
 This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
-While my key results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities[@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
+While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities[@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
 Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
@@
 
-Here I have shown that the stability of large systems might be contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
+Here I have shown that the stability of large systems might be often contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
 
@@
 For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), where $\mathbf{M = \gamma A}$.
-The genetic algorithm was initialised with a populatoin of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new population of 1000 $\mathbf{\gamma}$ vectors (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
+Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
+Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
 The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
@@
 
-In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium (feasibility is satisfied if all values in $\mathbf{x^{*}}$ are positive).
+In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium.
+Feasibility is satisfied if all elements in $\mathbf{x^{*}}$ are positive.
 The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is strength of intraspecific competition (diagonal elements).
@@
 Implicitly, here $\delta = 0$ underlying strong interactions.
-Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1$, and hence for my purposes $(CS)^{-\delta} = 1$).
+Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
 Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below,
@@
 
-**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.

### 2018-07-29 notebook/SI.Rmd
Discussion draft written and SI edits

@@
 
-Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
+Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 whenever $S > 10$ (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
 Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$.
-Note that feasibility results were ommited for the table above, but are [reported below](#Feasibility).
+Note that feasibility results were omitted for the table above, but are [reported below](#Feasibility).
 
@@
     gammas <- get_top_evo_out(evo_out, size = 9);
+    for(i in 1:9){
+        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
+    }
     par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
     hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 1.1));
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xlim = c(0, 1.1));
+         xlim = c(0, 2.1));
     hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1.1));
+         yaxt = "n", xlim = c(0, 2.1));
     hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 1.1));
+         yaxt = "n", xlim = c(0, 2.1));
     mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
@@
 While the foundational work of May[@May1972] applies broadly to complex networks, much attention has been given specifically to ecological networks of interacting species.
-In these networks, the matrix $\mathbf{M}$ is interpreted as a community matrix and each row and column is interpreted as a single species.
-The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $M_{ij}$, meaning that $\mathbf{M}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015].
-While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{M}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{M}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015].
+In these networks, the matrix $\mathbf{A}$ is interpreted as a community matrix and each row and column is interpreted as a single species.
+The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $A_{ij}$, meaning that $\mathbf{A}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015].
+While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{A}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{A}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015].
 In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.
@@
 ```{r, echo = FALSE}
-cd_tab <- cdat[,2:7];
+cd_tab <- cdat[,2:8];
 kable(cd_tab);
@@
 ```{r, echo = FALSE}
-md_tab <- mdat[,2:7];
+md_tab <- mdat[,2:8];
 kable(md_tab);
@@
 ```{r, echo = FALSE}
-pd_tab <- pdat[,2:7];
+pd_tab <- pdat[,2:8];
 kable(pd_tab);
@@
 
-Sensitivity of inter-connectivity (C) values {#connectance}
+Sensitivity of connectance (C) values {#connectance}
 ================================================================================
@@
 C3dat   <- C3dat[,-1];
-C3d_tab <- C3dat[,1:7];
+C3d_tab <- C3dat[,1:8];
 kable(C3d_tab);
@@
 C5dat   <- C5dat[,-1];
-C5d_tab <- C5dat[,1:7];
+C5d_tab <- C5dat[,1:8];
 kable(C5d_tab);
@@
 C7dat   <- C7dat[,-1];
-C7d_tab <- C7dat[,1:7];
+C7d_tab <- C7dat[,1:8];
 kable(C7d_tab);
@@
 C9dat   <- C9dat[,-1];
-C9d_tab <- C9dat[,1:7];
+C9d_tab <- C9dat[,1:8];
 kable(C9d_tab);

### 2018-07-29 notebook/SI.tex
Discussion draft written and SI edits

@@
 Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
-matrices found is greater than 1 (compare column 5 to column 3), and
-this ratio increases with increasing \(S\) (column 1). Hence, more
-randomly created complex systems (\(\mathbf{M}\)) are stable given
-variation in \(\gamma\) than when \(\gamma = 1\). Note that feasibility
-results were ommited for the table above, but are
+matrices found is greater than 1 whenever \(S > 10\) (compare column 5
+to column 3), and this ratio increases with increasing \(S\) (column 1).
+Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
+given variation in \(\gamma\) than when \(\gamma = 1\). Note that
+feasibility results were omitted for the table above, but are
 \protect\hyperlink{Feasibility}{reported below}.
@@
 ecological networks of interacting species. In these networks, the
-matrix \(\mathbf{M}\) is interpreted as a community matrix and each row
+matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
 and column is interpreted as a single species. The effect that the
 density of any species \(i\) has on the population dynamics of species
-\(j\) is found in \(M_{ij}\), meaning that \(\mathbf{M}\) holds the
+\(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds the
 effects of pair-wise interactions between \(S\)
@@
 more restricted ecological communities including competitive networks
-(all off-diagonal elements of \(\mathbf{M}\) are negative), mutualist
-networks (all off-diagonal elements of \(\mathbf{M}\) are positive), and
+(all off-diagonal elements of \(\mathbf{A}\) are negative), mutualist
+networks (all off-diagonal elements of \(\mathbf{A}\) are positive), and
 predator-prey networks (for any pair of \(i\) and \(j\), the effect of
@@
 
-\begin{longtable}[]{@{}llllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
-\midrule
-\endhead
-2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
-3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
-4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
-5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
-6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
-7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
-8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
-9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
-10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
-11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
-12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
-13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
-14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
-15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
-16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
-17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
-18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
-19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
-20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
-21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
-22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
-23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
-24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
-25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
-26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
-27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
-28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
-29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
-30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
-31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
-32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
-33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
-34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+3 & 229 & 99771 & 231 & 99769 & 0 & 2\tabularnewline
+4 & 701 & 99299 & 704 & 99296 & 0 & 3\tabularnewline
+5 & 1579 & 98421 & 1587 & 98413 & 0 & 8\tabularnewline
+6 & 3218 & 96782 & 3253 & 96747 & 6 & 41\tabularnewline
+7 & 5519 & 94481 & 5619 & 94381 & 23 & 123\tabularnewline
+8 & 9062 & 90938 & 9237 & 90763 & 77 & 252\tabularnewline
+9 & 13436 & 86564 & 13729 & 86271 & 230 & 523\tabularnewline
+10 & 18911 & 81089 & 19303 & 80697 & 505 & 897\tabularnewline
+11 & 25594 & 74406 & 25961 & 74039 & 1011 & 1378\tabularnewline
+12 & 33207 & 66793 & 33382 & 66618 & 1724 & 1899\tabularnewline
+13 & 41160 & 58840 & 41089 & 58911 & 2655 & 2584\tabularnewline
+14 & 50575 & 49425 & 49894 & 50106 & 3777 & 3096\tabularnewline
+15 & 59250 & 40750 & 57892 & 42108 & 4824 & 3466\tabularnewline
+16 & 67811 & 32189 & 65740 & 34260 & 5634 & 3563\tabularnewline
+17 & 75483 & 24517 & 73056 & 26944 & 5943 & 3516\tabularnewline
+18 & 82551 & 17449 & 79878 & 20122 & 5780 & 3107\tabularnewline
+19 & 88030 & 11970 & 85204 & 14796 & 5417 & 2591\tabularnewline
+20 & 92254 & 7746 & 89766 & 10234 & 4544 & 2056\tabularnewline
+21 & 95233 & 4767 & 93002 & 6998 & 3695 & 1464\tabularnewline
+22 & 97317 & 2683 & 95451 & 4549 & 2803 & 937\tabularnewline
+23 & 98508 & 1492 & 97122 & 2878 & 1991 & 605\tabularnewline
+24 & 99240 & 760 & 98407 & 1593 & 1216 & 383\tabularnewline
+25 & 99669 & 331 & 99082 & 918 & 739 & 152\tabularnewline
+26 & 99871 & 129 & 99490 & 510 & 452 & 71\tabularnewline
+27 & 99938 & 62 & 99732 & 268 & 240 & 34\tabularnewline
+28 & 99985 & 15 & 99888 & 112 & 108 & 11\tabularnewline
+29 & 99990 & 10 & 99951 & 49 & 46 & 7\tabularnewline
+30 & 100000 & 0 & 99981 & 19 & 19 & 0\tabularnewline
+31 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
+32 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
+33 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+34 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
-\midrule
-\endhead
-2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
-3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
-4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
-5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
-6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
-7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
-8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 56 & 99944 & 56 & 99944 & 0 & 0\tabularnewline
+3 & 3301 & 96699 & 3301 & 96699 & 0 & 0\tabularnewline
+4 & 34446 & 65554 & 34446 & 65554 & 0 & 0\tabularnewline
+5 & 86520 & 13480 & 86520 & 13480 & 0 & 0\tabularnewline
+6 & 99683 & 317 & 99683 & 317 & 0 & 0\tabularnewline
+7 & 99998 & 2 & 99998 & 2 & 0 & 0\tabularnewline
+8 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+9 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+11 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+12 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}rrrrrr@{}}
+\begin{longtable}[]{@{}rrrrrrr@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
-\midrule
-\endhead
-2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
-6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
-7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
-8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
-9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
-10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
-11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
-12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
-13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
-14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
-15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
-16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
-17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
-18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
-19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
-20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
-21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
-22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
-23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
-24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
-25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
-26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
-27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
-28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
-29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
-30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
-31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
-32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
-33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
-34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
-35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
-36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
-37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
-38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
-39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
-40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
-41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
-42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
-43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
-44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
-45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
-46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
-47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
-48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
-49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
-50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
-51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
-52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
-53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
-54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
-55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
-56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
-57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
-58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
-59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
-60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
-61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
-62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
-63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
-64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
-65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
-66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
-67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
-68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
-69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
-70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
-71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
-72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
-73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
-74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
-75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
-76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
-77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
-78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
-79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
-80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
-81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
-82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
-83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
-84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
-85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
-86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
-87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
-88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
-89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
-90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
-91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
-92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
-93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
-94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
-95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
-96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
-97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
-98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
-99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
-100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+3 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+4 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+5 & 1 & 99999 & 1 & 99999 & 0 & 0\tabularnewline
+6 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+7 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+8 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+9 & 20 & 99980 & 21 & 99979 & 0 & 1\tabularnewline
+10 & 20 & 99980 & 22 & 99978 & 0 & 2\tabularnewline
+11 & 38 & 99962 & 39 & 99961 & 0 & 1\tabularnewline
+12 & 64 & 99936 & 66 & 99934 & 0 & 2\tabularnewline
+13 & 87 & 99913 & 91 & 99909 & 0 & 4\tabularnewline
+14 & 157 & 99843 & 159 & 99841 & 0 & 2\tabularnewline
+15 & 215 & 99785 & 227 & 99773 & 0 & 12\tabularnewline
+16 & 293 & 99707 & 310 & 99690 & 0 & 17\tabularnewline
+17 & 383 & 99617 & 408 & 99592 & 0 & 25\tabularnewline
+18 & 443 & 99557 & 473 & 99527 & 3 & 33\tabularnewline
+19 & 642 & 99358 & 675 & 99325 & 4 & 37\tabularnewline
+20 & 836 & 99164 & 887 & 99113 & 7 & 58\tabularnewline
+21 & 1006 & 98994 & 1058 & 98942 & 10 & 62\tabularnewline
+22 & 1153 & 98847 & 1228 & 98772 & 20 & 95\tabularnewline
+23 & 1501 & 98499 & 1593 & 98407 & 30 & 122\tabularnewline
+24 & 1841 & 98159 & 1996 & 98004 & 40 & 195\tabularnewline
+25 & 2146 & 97854 & 2316 & 97684 & 58 & 228\tabularnewline
+26 & 2643 & 97357 & 2809 & 97191 & 119 & 285\tabularnewline
+27 & 3034 & 96966 & 3258 & 96742 & 158 & 382\tabularnewline
+28 & 3690 & 96310 & 3928 & 96072 & 201 & 439\tabularnewline
+29 & 4257 & 95743 & 4532 & 95468 & 290 & 565\tabularnewline
+30 & 4964 & 95036 & 5221 & 94779 & 424 & 681\tabularnewline
+31 & 5627 & 94373 & 5978 & 94022 & 452 & 803\tabularnewline
+32 & 6543 & 93457 & 6891 & 93109 & 666 & 1014\tabularnewline
+33 & 7425 & 92575 & 7777 & 92223 & 818 & 1170\tabularnewline
+34 & 8540 & 91460 & 8841 & 91159 & 1071 & 1372\tabularnewline
+35 & 9526 & 90474 & 9842 & 90158 & 1337 & 1653\tabularnewline
+36 & 10617 & 89383 & 10891 & 89109 & 1624 & 1898\tabularnewline
+37 & 12344 & 87656 & 12508 & 87492 & 2021 & 2185\tabularnewline
+38 & 13675 & 86325 & 13877 & 86123 & 2442 & 2644\tabularnewline
+39 & 15264 & 84736 & 15349 & 84651 & 2870 & 2955\tabularnewline
+40 & 17026 & 82974 & 17053 & 82947 & 3363 & 3390\tabularnewline
+41 & 18768 & 81232 & 18614 & 81386 & 3905 & 3751\tabularnewline
+42 & 20791 & 79209 & 20470 & 79530 & 4579 & 4258\tabularnewline
+43 & 23150 & 76850 & 22754 & 77246 & 5217 & 4821\tabularnewline
+44 & 25449 & 74551 & 24184 & 75816 & 6285 & 5020\tabularnewline
+45 & 27702 & 72298 & 26464 & 73536 & 6754 & 5516\tabularnewline
+46 & 30525 & 69475 & 28966 & 71034 & 7646 & 6087\tabularnewline
+47 & 32832 & 67168 & 31125 & 68875 & 8487 & 6780\tabularnewline
+48 & 36152 & 63848 & 33865 & 66135 & 9479 & 7192\tabularnewline
+49 & 38714 & 61286 & 36242 & 63758 & 10125 & 7653\tabularnewline
+50 & 41628 & 58372 & 38508 & 61492 & 11036 & 7916\tabularnewline
+51 & 44483 & 55517 & 41023 & 58977 & 11704 & 8244\tabularnewline
+52 & 48134 & 51866 & 44287 & 55713 & 12573 & 8726\tabularnewline
+53 & 51138 & 48862 & 46721 & 53279 & 13223 & 8806\tabularnewline
+54 & 54261 & 45739 & 49559 & 50441 & 13757 & 9055\tabularnewline
+55 & 57647 & 42353 & 52403 & 47597 & 14324 & 9080\tabularnewline
+56 & 60630 & 39370 & 55293 & 44707 & 14669 & 9332\tabularnewline
+57 & 63647 & 36353 & 57787 & 42213 & 15103 & 9243\tabularnewline
+58 & 66961 & 33039 & 60439 & 39561 & 15450 & 8928\tabularnewline
+59 & 69968 & 30032 & 63708 & 36292 & 15246 & 8986\tabularnewline
+60 & 72838 & 27162 & 66270 & 33730 & 15177 & 8609\tabularnewline
+61 & 75609 & 24391 & 68873 & 31127 & 15006 & 8270\tabularnewline
+62 & 77999 & 22001 & 71318 & 28682 & 14538 & 7857\tabularnewline
+63 & 80616 & 19384 & 73517 & 26483 & 14510 & 7411\tabularnewline
+64 & 83089 & 16911 & 76209 & 23791 & 13784 & 6904\tabularnewline
+65 & 85150 & 14850 & 78086 & 21914 & 13412 & 6348\tabularnewline
+66 & 86908 & 13092 & 80437 & 19563 & 12477 & 6006\tabularnewline
+67 & 88671 & 11329 & 82379 & 17621 & 11718 & 5426\tabularnewline
+68 & 90537 & 9463 & 84483 & 15517 & 10878 & 4824\tabularnewline
+69 & 91969 & 8031 & 86233 & 13767 & 10033 & 4297\tabularnewline
+70 & 93181 & 6819 & 87914 & 12086 & 9070 & 3803\tabularnewline
+71 & 94330 & 5670 & 89200 & 10800 & 8401 & 3271\tabularnewline
+72 & 95324 & 4676 & 90833 & 9167 & 7359 & 2868\tabularnewline
+73 & 96143 & 3857 & 91805 & 8195 & 6726 & 2388\tabularnewline
+74 & 96959 & 3041 & 93065 & 6935 & 5900 & 2006\tabularnewline
+75 & 97543 & 2457 & 93987 & 6013 & 5222 & 1666\tabularnewline
+76 & 97969 & 2031 & 94900 & 5100 & 4481 & 1412\tabularnewline
+77 & 98497 & 1503 & 95756 & 4244 & 3809 & 1068\tabularnewline
+78 & 98744 & 1256 & 96442 & 3558 & 3269 & 967\tabularnewline
+79 & 99045 & 955 & 96942 & 3058 & 2837 & 734\tabularnewline
+80 & 99276 & 724 & 97528 & 2472 & 2329 & 581\tabularnewline
+81 & 99481 & 519 & 97996 & 2004 & 1894 & 409\tabularnewline
+82 & 99556 & 444 & 98321 & 1679 & 1597 & 362\tabularnewline
+83 & 99691 & 309 & 98722 & 1278 & 1227 & 258\tabularnewline
+84 & 99752 & 248 & 98943 & 1057 & 1015 & 206\tabularnewline
+85 & 99833 & 167 & 99144 & 856 & 837 & 148\tabularnewline
+86 & 99895 & 105 & 99346 & 654 & 642 & 93\tabularnewline
+87 & 99925 & 75 & 99461 & 539 & 530 & 66\tabularnewline
+88 & 99945 & 55 & 99566 & 434 & 428 & 49\tabularnewline
+89 & 99976 & 24 & 99675 & 325 & 324 & 23\tabularnewline
+90 & 99977 & 23 & 99756 & 244 & 243 & 22\tabularnewline
+91 & 99982 & 18 & 99839 & 161 & 155 & 12\tabularnewline
+92 & 99988 & 12 & 99865 & 135 & 135 & 12\tabularnewline
+93 & 99994 & 6 & 99885 & 115 & 115 & 6\tabularnewline
+94 & 99993 & 7 & 99911 & 89 & 88 & 6\tabularnewline
+95 & 99998 & 2 & 99953 & 47 & 47 & 2\tabularnewline
+96 & 99999 & 1 & 99965 & 35 & 35 & 1\tabularnewline
+97 & 99999 & 1 & 99979 & 21 & 21 & 1\tabularnewline
+98 & 100000 & 0 & 99973 & 27 & 27 & 0\tabularnewline
+99 & 100000 & 0 & 99984 & 16 & 16 & 0\tabularnewline
+100 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
 \bottomrule
@@
 
-\hypertarget{connectance}{\section{Sensitivity of inter-connectivity (C)
+\hypertarget{connectance}{\section{Sensitivity of connectance (C)
 values}\label{connectance}}
@@
 
-\begin{longtable}[]{@{}lllllll@{}}
+\begin{longtable}[]{@{}llllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
-3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
-4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
-5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
-6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
-7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
-8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
-9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
-10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
-11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
-12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
-13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
-14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
-15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
-16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
-17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
-18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
-19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
-20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
-21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
-22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
-23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
-24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
-25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
-26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
-27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
-28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
-29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
-30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
-31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
-32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
-33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
-34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
-35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
-36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
-37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
-38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
-39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
-40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
-41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
-42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
-43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
-44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
-45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
-46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
-47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
-48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
-49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
-50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
-51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
-52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
-53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
-54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
-55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 5 & 99995 & 5 & 99995 & 0 & 0 & 75110\tabularnewline
+3 & 6 & 99994 & 6 & 99994 & 0 & 0 & 87526\tabularnewline
+4 & 24 & 99976 & 24 & 99976 & 0 & 0 & 93713\tabularnewline
+5 & 59 & 99941 & 59 & 99941 & 0 & 0 & 96929\tabularnewline
+6 & 98 & 99902 & 98 & 99902 & 0 & 0 & 98492\tabularnewline
+7 & 160 & 99840 & 161 & 99839 & 0 & 1 & 99243\tabularnewline
+8 & 290 & 99710 & 293 & 99707 & 0 & 3 & 99582\tabularnewline
+9 & 430 & 99570 & 434 & 99566 & 0 & 4 & 99821\tabularnewline
+10 & 648 & 99352 & 653 & 99347 & 1 & 6 & 99895\tabularnewline
+11 & 946 & 99054 & 957 & 99043 & 0 & 11 & 99945\tabularnewline
+12 & 1392 & 98608 & 1415 & 98585 & 4 & 27 & 99978\tabularnewline
+13 & 2032 & 97968 & 2065 & 97935 & 5 & 38 & 99987\tabularnewline
+14 & 2627 & 97373 & 2688 & 97312 & 10 & 71 & 99988\tabularnewline
+15 & 3588 & 96412 & 3647 & 96353 & 35 & 94 & 99996\tabularnewline
+16 & 5019 & 94981 & 5124 & 94876 & 51 & 156 & 99998\tabularnewline
+17 & 6512 & 93488 & 6673 & 93327 & 79 & 240 & 99999\tabularnewline
+18 & 8444 & 91556 & 8600 & 91400 & 165 & 321 & 100000\tabularnewline
+19 & 10416 & 89584 & 10667 & 89333 & 244 & 495 & 100000\tabularnewline
+20 & 13254 & 86746 & 13477 & 86523 & 425 & 648 & 100000\tabularnewline
+21 & 16248 & 83752 & 16481 & 83519 & 642 & 875 & 100000\tabularnewline
+22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151 & 100000\tabularnewline
+23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490 & 100000\tabularnewline
+24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818 & 100000\tabularnewline
+25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137 & 100000\tabularnewline
+26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506 & 100000\tabularnewline
+27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811 & 100000\tabularnewline
+28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072 & 100000\tabularnewline
+29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321 & 100000\tabularnewline
+30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480 & 100000\tabularnewline
+31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568 & 100000\tabularnewline
+32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531 & 100000\tabularnewline
+33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360 & 100000\tabularnewline
+34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014 & 100000\tabularnewline
+35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773 & 100000\tabularnewline
+36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393 & 100000\tabularnewline
+37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913 & 100000\tabularnewline
+38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566 & 100000\tabularnewline
+39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187 & 100000\tabularnewline
+40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936 & 100000\tabularnewline
+41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683 & 100000\tabularnewline
+42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449 & 100000\tabularnewline
+43 & 99275 & 725 & 98312 & 1688 & 1251 & 288 & 100000\tabularnewline
+44 & 99583 & 417 & 98872 & 1128 & 903 & 192 & 100000\tabularnewline
+45 & 99776 & 224 & 99339 & 661 & 576 & 139 & 100000\tabularnewline
+46 & 99865 & 135 & 99518 & 482 & 413 & 66 & 100000\tabularnewline
+47 & 99938 & 62 & 99744 & 256 & 226 & 32 & 100000\tabularnewline
+48 & 99956 & 44 & 99824 & 176 & 151 & 19 & 100000\tabularnewline
+49 & 99980 & 20 & 99914 & 86 & 85 & 19 & 100000\tabularnewline
+50 & 99993 & 7 & 99950 & 50 & 46 & 3 & 100000\tabularnewline
+51 & 99998 & 2 & 99971 & 29 & 28 & 1 & 100000\tabularnewline
+52 & 99998 & 2 & 99986 & 14 & 14 & 2 & 100000\tabularnewline
+53 & 99999 & 1 & 99992 & 8 & 7 & 0 & 100000\tabularnewline
+54 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+55 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+56 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+57 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+58 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{}\tabularnewline
+100 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}lllllll@{}}
+\begin{longtable}[]{@{}llllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
-3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
-4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
-5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
-6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
-7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
-8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
-9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
-10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
-11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
-12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
-13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
-14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
-15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
-16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
-17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
-18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
-19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
-20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
-21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
-22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
-23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
-24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
-25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
-26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
-27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
-28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
-29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
-30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
-31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
-32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
-33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
-34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
-35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
-36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
-37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
-38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
-39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
-40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 74863\tabularnewline
+3 & 32 & 99968 & 32 & 99968 & 0 & 0 & 87434\tabularnewline
+4 & 122 & 99878 & 122 & 99878 & 0 & 0 & 93761\tabularnewline
+5 & 320 & 99680 & 321 & 99679 & 0 & 1 & 96830\tabularnewline
+6 & 667 & 99333 & 673 & 99327 & 0 & 6 & 98481\tabularnewline
+7 & 1233 & 98767 & 1252 & 98748 & 0 & 19 & 99187\tabularnewline
+8 & 2123 & 97877 & 2156 & 97844 & 3 & 36 & 99654\tabularnewline
+9 & 3415 & 96585 & 3471 & 96529 & 16 & 72 & 99816\tabularnewline
+10 & 5349 & 94651 & 5450 & 94550 & 30 & 131 & 99900\tabularnewline
+11 & 7990 & 92010 & 8185 & 91815 & 81 & 276 & 99958\tabularnewline
+12 & 11073 & 88927 & 11301 & 88699 & 219 & 447 & 99973\tabularnewline
+13 & 14971 & 85029 & 15204 & 84796 & 445 & 678 & 99986\tabularnewline
+14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002 & 99991\tabularnewline
+15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404 & 99996\tabularnewline
+16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980 & 99999\tabularnewline
+17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476 & 100000\tabularnewline
+18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937 & 99999\tabularnewline
+19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262 & 99999\tabularnewline
+20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527 & 100000\tabularnewline
+21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673 & 100000\tabularnewline
+22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491 & 100000\tabularnewline
+23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244 & 100000\tabularnewline
+24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876 & 100000\tabularnewline
+25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382 & 100000\tabularnewline
+26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901 & 100000\tabularnewline
+27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358 & 100000\tabularnewline
+28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974 & 100000\tabularnewline
+29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657 & 100000\tabularnewline
+30 & 99103 & 897 & 98033 & 1967 & 1466 & 396 & 100000\tabularnewline
+31 & 99502 & 498 & 98665 & 1335 & 1068 & 231 & 100000\tabularnewline
+32 & 99745 & 255 & 99185 & 815 & 696 & 136 & 100000\tabularnewline
+33 & 99881 & 119 & 99572 & 428 & 375 & 66 & 100000\tabularnewline
+34 & 99955 & 45 & 99788 & 212 & 191 & 24 & 100000\tabularnewline
+35 & 99979 & 21 & 99900 & 100 & 95 & 16 & 100000\tabularnewline
+36 & 99995 & 5 & 99950 & 50 & 50 & 5 & 100000\tabularnewline
+37 & 99997 & 3 & 99970 & 30 & 28 & 1 & 100000\tabularnewline
+38 & 99998 & 2 & 99986 & 14 & 13 & 1 & 100000\tabularnewline
+39 & 99999 & 1 & 99991 & 9 & 9 & 1 & 100000\tabularnewline
+40 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+41 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+42 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+43 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}lllllll@{}}
+\begin{longtable}[]{@{}llllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
-3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
-4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
-5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
-6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
-7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
-8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
-9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
-10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
-11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
-12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
-13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
-14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
-15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
-16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
-17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
-18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
-19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
-20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
-21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
-22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
-23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
-24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
-25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
-26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
-27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
-28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
-29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
-30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
-31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
-32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
-33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
-34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 75160\tabularnewline
+3 & 106 & 99894 & 106 & 99894 & 0 & 0 & 87447\tabularnewline
+4 & 395 & 99605 & 397 & 99603 & 0 & 2 & 93849\tabularnewline
+5 & 1117 & 98883 & 1123 & 98877 & 0 & 6 & 96827\tabularnewline
+6 & 2346 & 97654 & 2367 & 97633 & 6 & 27 & 98402\tabularnewline
+7 & 4314 & 95686 & 4388 & 95612 & 16 & 90 & 99218\tabularnewline
+8 & 7327 & 92673 & 7456 & 92544 & 61 & 190 & 99611\tabularnewline
+9 & 11514 & 88486 & 11792 & 88208 & 150 & 428 & 99808\tabularnewline
+10 & 16247 & 83753 & 16584 & 83416 & 415 & 752 & 99904\tabularnewline
+11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162 & 99952\tabularnewline
+12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818 & 99977\tabularnewline
+13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355 & 99984\tabularnewline
+14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927 & 99995\tabularnewline
+15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421 & 99994\tabularnewline
+16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527 & 99999\tabularnewline
+17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495 & 100000\tabularnewline
+18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351 & 100000\tabularnewline
+19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979 & 99999\tabularnewline
+20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339 & 100000\tabularnewline
+21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869 & 100000\tabularnewline
+22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267 & 100000\tabularnewline
+23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791 & 100000\tabularnewline
+24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504 & 100000\tabularnewline
+25 & 99433 & 567 & 98630 & 1370 & 1079 & 276 & 100000\tabularnewline
+26 & 99760 & 240 & 99259 & 741 & 655 & 154 & 100000\tabularnewline
+27 & 99895 & 105 & 99576 & 424 & 377 & 58 & 100000\tabularnewline
+28 & 99950 & 50 & 99790 & 210 & 194 & 34 & 100000\tabularnewline
+29 & 99981 & 19 & 99915 & 85 & 80 & 14 & 100000\tabularnewline
+30 & 99994 & 6 & 99952 & 48 & 47 & 5 & 100000\tabularnewline
+31 & 99998 & 2 & 99972 & 28 & 28 & 2 & 100000\tabularnewline
+32 & 99999 & 1 & 99992 & 8 & 8 & 1 & 100000\tabularnewline
+33 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+34 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+35 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}lllllll@{}}
+\begin{longtable}[]{@{}llllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
-3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
-4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
-5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
-6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
-7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
-8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
-9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
-10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
-11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
-12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
-13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
-14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
-15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
-16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
-17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
-18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
-19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
-20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
-21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
-22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
-23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
-24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
-25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
-26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
-27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
-28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
-29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
-30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 14 & 99986 & 14 & 99986 & 0 & 0 & 75187\tabularnewline
+3 & 240 & 99760 & 240 & 99760 & 0 & 0 & 87443\tabularnewline
+4 & 1008 & 98992 & 1016 & 98984 & 0 & 8 & 93795\tabularnewline
+5 & 2708 & 97292 & 2729 & 97271 & 2 & 23 & 96814\tabularnewline
+6 & 5669 & 94331 & 5755 & 94245 & 13 & 99 & 98439\tabularnewline
+7 & 9848 & 90152 & 10057 & 89943 & 91 & 300 & 99208\tabularnewline
+8 & 15903 & 84097 & 16201 & 83799 & 336 & 634 & 99603\tabularnewline
+9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168 & 99803\tabularnewline
+10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852 & 99909\tabularnewline
+11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507 & 99951\tabularnewline
+12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127 & 99977\tabularnewline
+13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649 & 99986\tabularnewline
+14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765 & 99995\tabularnewline
+15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550 & 100000\tabularnewline
+16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092 & 100000\tabularnewline
+17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424 & 100000\tabularnewline
+18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746 & 100000\tabularnewline
+19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167 & 99999\tabularnewline
+20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729 & 100000\tabularnewline
+21 & 99111 & 889 & 98133 & 1867 & 1389 & 411 & 100000\tabularnewline
+22 & 99588 & 412 & 98905 & 1095 & 903 & 220 & 100000\tabularnewline
+23 & 99837 & 163 & 99480 & 520 & 452 & 95 & 100000\tabularnewline
+24 & 99932 & 68 & 99744 & 256 & 228 & 40 & 100000\tabularnewline
+25 & 99976 & 24 & 99863 & 137 & 133 & 20 & 100000\tabularnewline
+26 & 99995 & 5 & 99950 & 50 & 49 & 4 & 100000\tabularnewline
+27 & 99996 & 4 & 99986 & 14 & 13 & 3 & 100000\tabularnewline
+28 & 100000 & 0 & 99993 & 7 & 7 & 0 & 100000\tabularnewline
+29 & 100000 & 0 & 99996 & 4 & 4 & 0 & 100000\tabularnewline
+30 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
 \bottomrule

### 2018-07-29 notebook/ms.Rmd
Discussion draft written and SI edits

@@
 Here I have shown that the stability of large systems might be often contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
+Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's[@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
+The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
 
-It is important to point out that $Var(\gamma)$ is not stabilising per se; that is, adding variation in $\gamma$ to a particular system $\mathbf{M}$ does not necessarily increase the probability that the system will be stable.
-Rather, systems that are observed to be stable are more likely to vary in $\gamma$, and for this $Var(\gamma)$ to be critical to their stability.
-This is caused by the shift in the distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes result in all $\Re(\lambda) < 0$ but might also increase $\Re(\lambda)$ values.
-The mathematics underlying this shift in eigenvalue distribution have been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but have not been interpreted as rates of response of individual componetents to system perturbation.
+It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
+Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
+3).
+This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
+1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
+The mathematics underlying this shift in eigenvalue distribution has been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual componetents to system perturbation.
 
-Nevertheless, ecological interactions do not exist in isolation in empirical systems, but instead interact with evolutionary[@Patel2018], abiotic, or social-economic systems.
-The relevance of $\gamma$ for complex system stability presented in the main text should therefore not be ignored in the broader context of ecological communities.
+The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
+The probability that some combination of component respose rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
+Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
+Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
 
-I have focused broadly on random complex systems, but it is also worthwhile to consider more restricted interactions such as those of specific ecological networks [@Allesina2012].
-These include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or $i$ and $j$ pairs have opposing signs (predator-prey networks).
-In general, competitive and mutualist networks tend to be destabilising, and predator-prey network tend to be stabilising [@Allesina2011].
-When $Var(\gamma)$ is applied to each, the proportion of stable competitive and predator-prey networks increases, but the proportion of stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of $\mathbf{M}$ is interpreted as a unique species and given a random intrinsic growth rate [@Dougoud2018], feasibility is not increased by $Var(\gamma)$, suggesting that variation in species generation time might be unlikely to drive stability in purely multi-species networks (see Supplementary Information).
-
-Hence, managing the response rates of system components in a targeted way can potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+Interestingly, while complex systems were more likely to be stable given variation in component respose rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
+Feasibility is important to consider, particularly for the study of ecological networks of species[@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative.
+My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in empirical systems[@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability should therefore not be ignored in the broader context of ecological communities.
 

### 2018-07-30 notebook/SI.Rmd
Final touch ups

@@
 **This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with additional analyses to support its conclusions.
-All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability package.**
+All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability R package.**
 
@@
 In these networks, the matrix $\mathbf{A}$ is interpreted as a community matrix and each row and column is interpreted as a single species.
-The effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $A_{ij}$, meaning that $\mathbf{A}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015].
+The per capita effect that the density of any species $i$ has on the population dynamics of species $j$ is found in $A_{ij}$, meaning that $\mathbf{A}$ holds the effects of pair-wise interactions between $S$ species[@Allesina2012; @Allesina2015].
 While May's original work[@May1972] considered only randomly assembled communities, recent work has specifically looked at more restricted ecological communities including competitive networks (all off-diagonal elements of $\mathbf{A}$ are negative), mutualist networks (all off-diagonal elements of $\mathbf{A}$ are positive), and predator-prey networks (for any pair of $i$ and $j$, the effect of $i$ on $j$ is negative and $j$ on $i$ is positive, or vice versa)[@Allesina2012; @Allesina2015].
-In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising.
+In general, competitor and mutualist networks tend to be unstable, while predator-prey networks tend to be highly stabilising[@Allesina2012].
 
@@
 5, 0.7, 0.9 \}$.
-In all cases, uniform variation in component response time ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
+In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
 In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.

### 2018-07-30 notebook/SI.tex
Final touch ups

@@
 \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
-part of the RandomMatrixStability package.}
+part of the RandomMatrixStability R package.}
 
@@
 matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
-and column is interpreted as a single species. The effect that the
-density of any species \(i\) has on the population dynamics of species
-\(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds the
-effects of pair-wise interactions between \(S\)
+and column is interpreted as a single species. The per capita effect
+that the density of any species \(i\) has on the population dynamics of
+species \(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds
+the effects of pair-wise interactions between \(S\)
 species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
@@
 In general, competitor and mutualist networks tend to be unstable, while
-predator-prey networks tend to be highly stabilising.
+predator-prey networks tend to be highly
+stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
 
@@
 \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
-component response time (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
+component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
 higher number of stable communities than when \(\gamma\) did not vary

### 2018-07-30 notebook/ms.Rmd
Final touch ups

@@
 
-**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counter-intuitive result of widespread importance across the physical, life, and social sciences.
+**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counterintuitive result of widespread importance across the physical, life, and social sciences.
 Despite recent interest in the relationship between system properties and stability, the effect of variation in the response rate of individual system components remains unconsidered.
@@
 May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
-In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of individuals, which might vary by orders of magnitude among species.
+In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
 Species with short generation times will respond quickly (high $\gamma$) to perturbations relative to species with long generation times (low $\gamma$).
@@
 1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
-The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$) and all $\Re({\lambda}) < 0$.
-The latter cluster has real components that have shifted to the right, but has a smaller radius.
+The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$), while the latter cluster has a smaller radius, but real components have shifted to the right; all $\Re({\lambda}) < 0$.
 Overall, for 1 million randomly assembled $\mathbf{M}$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S$ ranging from 2-50.
+To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
 One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
@@
 **System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
-Feasibile equilibria assume that the values of all system components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
+Feasibilility assumes that the system is stable and values of all components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
 This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
 While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities[@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
-Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response time (i.e., species generation time) does not affect stability at biologically realistic species densities.
+Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response rate (i.e., species generation time) does not affect stability at biologically realistic species densities.
 
@@
 
-Here I have shown that the stability of large systems might be often contigent upon variation in the response rates of their individual components, meaning that factors such as generation time and rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
+Here I have shown that the stability of large systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
 Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's[@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
@@
 1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
-The mathematics underlying this shift in eigenvalue distribution has been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual componetents to system perturbation.
+The mathematics underlying this shift in eigenvalue distribution has been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual system components to perturbation.
 
 The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
-The probability that some combination of component respose rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
+The probability that some combination of component response rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
 Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
@@
 
-Interestingly, while complex systems were more likely to be stable given variation in component respose rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
+Interestingly, while complex systems were more likely to be stable given variation in component response rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
 Feasibility is important to consider, particularly for the study of ecological networks of species[@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative.
@@
 
-In the above, $\mathbf{x^{*}}$ is the vector of species abundances at equilibrium.
+In the above, $\mathbf{x^{*}}$ is the vector of species densities at equilibrium.
 Feasibility is satisfied if all elements in $\mathbf{x^{*}}$ are positive.
-The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is strength of intraspecific competition (diagonal elements).
+The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is the strength of intraspecific competition (diagonal elements).
 Diagonal values are set to $-1$, so $\theta = -1$.

### 2018-07-30 notebook/ms.Rmd
Rework all files for arXiv

@@
     fig_caption: yes
-    keep_tex: no
+    keep_tex: yes
   html_document: default
@@
 - \usepackage{natbib}
-- \usepackage{lineno}
 - \usepackage[utf8]{inputenc}
-- \linenumbers
 - \bibliographystyle{amnatnat}
-linestretch: 2
-link-citations: no
+linestretch: 1
+link-citations: yes
 linkcolor: blue
@@
 source(file = "R_old/plot_figs.R");
+```
+
+```{r, echo = FALSE}
+# Temporary place for revised functions
+plot_Fig_1 <- function(A0, A1){
+    S_val       <- dim(A0)[1];
+    A0_e        <- eigen(A0)$values;
+    A0_r        <- Re(A0_e);
+    A0_i        <- Im(A0_e);
+    A1_e        <- eigen(A1)$values;
+    A1_r        <- Re(A1_e);
+    A1_i        <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(t(A0_vm));
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(t(A1_vm));
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    fhalf       <- 1:(0.5*length(A1vec));
+    shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
+    vl   <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
+    text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(A0_r, A0_i, pch = 4, cex = 0.7);
+    plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+         col = "dodgerblue4", yaxt = "n");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + 
+                 mean(diag(A1)[1:(0.5*S_val)]);
+    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
+    #points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
+    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
+        mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
+    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
+    #points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
+    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7, col = "firebrick");   
+    text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_Fig_2 <- function(){
+    A_comp <- NULL;
+    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+    A_mat  <- matrix(data = A_dat, nrow = 1000);
+    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
+    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+    A_mat     <- A_mat * C_mat;
+    gammas <- runif(n = 1000, min = 0, max = 2);
+    mu_gam <- mean(gammas);
+    diag(A_mat) <- -1;
+    A1     <- gammas * A_mat;
+    A0     <- mu_gam * A_mat;
+    A0_e   <- eigen(A0)$values;
+    A0_r   <- Re(A0_e);
+    A0_i   <- Im(A0_e);
+    A1_e   <- eigen(A1)$values;
+    A1_r   <- Re(A1_e);
+    A1_i   <- Im(A1_e);
+    A0_vm       <- A0;
+    diag(A0_vm) <- NA;
+    A0vec       <- as.vector(A0_vm);
+    A0vec       <- A0vec[is.na(A0vec) == FALSE];
+    A1_vm       <- A1;
+    diag(A1_vm) <- NA;
+    A1vec       <- as.vector(A1_vm);
+    A1vec       <- A1vec[is.na(A1vec) == FALSE];
+    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "dodgerblue4");
+    vl <- seq(from = 0, to = 2*pi, by = 0.001);
+    x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+    y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
+    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
+    text(x = -15.5, y = 19, labels = "a", cex = 2);
+    points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
+    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
+    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
+         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+         asp = 1, col = "firebrick", yaxt = "n");
+    text(x = -15.5, y = 19, labels = "b", cex = 2);
+    points(x = x1, y = y1, type = "l", col = "firebrick", lwd = 3);
+    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
+    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
+
+plot_stable_4 <- function(dat, S_s = 39){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 14), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 116), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(gamma," = 1")), 
+                        expression(paste("Var(",gamma,")"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
 ```
@@
 
+********************************************************************************
+
+```{r echo = FALSE}
+S     <- 200;
+C     <- 0.05;
+sigma <- 0.4;
+pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st <- pr_st[,-1];
+```
+
+```{r, echo = FALSE}
+A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
+A0 <- as.matrix(A0[,-1]);
+A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
+A1 <- as.matrix(A1[,-1]);
+
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(t(A0_vm));
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+A1_vm       <- A1;
+diag(A1_vm) <- NA;
+A1vec       <- as.vector(t(A1_vm));
+A1vec       <- A1vec[is.na(A1vec) == FALSE];
+fhalf       <- 1:(0.5*length(A1vec));
+shalf       <- (0.5*length(A1vec)+1):length(A1vec);
+pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
+pr_st       <- pr_st[,-1];
+```
+
+**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
+**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the blue circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
+**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
+In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
+In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` was stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
+Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "dodgerblue4");
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
+A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
+text(x = -3.5, y = 2.25, labels = "a", cex = 2);
+points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "blue");
+points(A0_r, A0_i, pch = 4, cex = 0.7, col = "dodgerblue4");
+
+plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "firebrick", yaxt = "n");
+
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
+A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "red");
+A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
+A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
+points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "red");
+
+points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7, col = "firebrick");
+
+text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+********************************************************************************
+
 Discussion
@@
 The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
+
+********************************************************************************
+
+```{r echo = FALSE}
+S     <- 1000;
+C     <- 1;
+sigma <- 0.4;
+```
+
+```{r, echo = FALSE}
+A_comp <- NULL;
+A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
+A_mat  <- matrix(data = A_dat, nrow = 1000);
+C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
+C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
+A_mat  <- A_mat * C_mat;
+A_mat  <- species_interactions(A_mat, type = 0);
+gammas <- runif(n = 1000, min = 0, max = 2);
+
+mu_gam <- mean(gammas);
+diag(A_mat) <- -1;
+A1     <- gammas * A_mat;
+A0     <- mu_gam * A_mat;
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+A0_vm       <- A0;
+diag(A0_vm) <- NA;
+A0vec       <- as.vector(A0_vm);
+A0vec       <- A0vec[is.na(A0vec) == FALSE];
+```
+
+**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
+**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
+Blue elipses in both panels show the circle centred on the distribution in panel a.
+Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
+
+```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "dodgerblue4");
+vl <- seq(from = 0, to = 2*pi, by = 0.001);
+x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
+y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
+text(x = -5.5, y = 4, labels = "a", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3, col = "blue");
+points(A0_r, A0_i, col = "dodgerblue4",pch = 4, cex = 0.7);
+plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
+     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+     col = "firebrick", yaxt = "n");
+text(x = -5.5, y = 4, labels = "b", cex = 2);
+points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid", col = "blue");
+points(A1_r, A1_i, col = "firebrick", pch = 4, cex = 0.7);
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+********************************************************************************
 
@@
 
+********************************************************************************
+
+```{r, echo = FALSE}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+```
+
+**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+For each $S$, 1 million complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
+Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
+Blue and red bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
+The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/C_1/random_all.csv");
+dat <- dat[,-1];
+plot_stables(dat);
+```
+
+********************************************************************************
+
 Interestingly, while complex systems were more likely to be stable given variation in component response rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
@@
 The relevance of component response rate for complex system stability should therefore not be ignored in the broader context of ecological communities.
+
+********************************************************************************
+
+**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
+For each $S$, 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
+Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
+The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
+dat <- dat[,-1];
+plot_stable_4(dat);
+```
+
+********************************************************************************
 
@@
 <div id="refs"></div>
-
-```{r, echo = FALSE}
-# Temporary place for revised functions
-plot_Fig_1 <- function(A0, A1){
-    S_val       <- dim(A0)[1];
-    A0_e        <- eigen(A0)$values;
-    A0_r        <- Re(A0_e);
-    A0_i        <- Im(A0_e);
-    A1_e        <- eigen(A1)$values;
-    A1_r        <- Re(A1_e);
-    A1_i        <- Im(A1_e);
-    A0_vm       <- A0;
-    diag(A0_vm) <- NA;
-    A0vec       <- as.vector(t(A0_vm));
-    A0vec       <- A0vec[is.na(A0vec) == FALSE];
-    A1_vm       <- A1;
-    diag(A1_vm) <- NA;
-    A1vec       <- as.vector(t(A1_vm));
-    A1vec       <- A1vec[is.na(A1vec) == FALSE];
-    fhalf       <- 1:(0.5*length(A1vec));
-    shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-    plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1);
-    vl   <- seq(from = 0, to = 2*pi, by = 0.001);
-    A0x0 <- sqrt(S_val) * sd(A0vec) * cos(vl) + mean(diag(A0));
-    A0y0 <- sqrt(S_val) * sd(A0vec) * sin(vl);
-    text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-    points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "dodgerblue4");
-    points(A0_r, A0_i, pch = 4, cex = 0.7);
-    plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-         xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-         col = "dodgerblue4", yaxt = "n");
-    vl <- seq(from = 0, to = 2*pi, by = 0.001);
-    A0x1a <- sqrt(0.5*S_val) * sd(A1vec[fhalf]) * cos(vl) + 
-                 mean(diag(A1)[1:(0.5*S_val)]);
-    A0y1a <- sqrt(S_val) * sd(A1vec[fhalf]) * sin(vl);
-    #points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "grey");
-    A0x1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * cos(vl) + 
-        mean( diag(A1)[( (0.5*S_val) + 1 ):S_val] );
-    A0y1b <- sqrt(0.5*S_val) * sd(A1vec[shalf]) * sin(vl);
-    #points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "grey");
-    points(A1_r[1:S_val], A1_i[1:S_val],pch = 4, cex = 0.7, col = "firebrick");   
-    text(x = -3.5, y = 2.25, labels = "b", cex = 2);
-    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-}
-
-plot_Fig_2 <- function(){
-    A_comp <- NULL;
-    A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
-    A_mat  <- matrix(data = A_dat, nrow = 1000);
-    C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 1);
-    C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
-    A_mat     <- A_mat * C_mat;
-    gammas <- runif(n = 1000, min = 0, max = 2);
-    mu_gam <- mean(gammas);
-    diag(A_mat) <- -1;
-    A1     <- gammas * A_mat;
-    A0     <- mu_gam * A_mat;
-    A0_e   <- eigen(A0)$values;
-    A0_r   <- Re(A0_e);
-    A0_i   <- Im(A0_e);
-    A1_e   <- eigen(A1)$values;
-    A1_r   <- Re(A1_e);
-    A1_i   <- Im(A1_e);
-    A0_vm       <- A0;
-    diag(A0_vm) <- NA;
-    A0vec       <- as.vector(A0_vm);
-    A0vec       <- A0vec[is.na(A0vec) == FALSE];
-    A1_vm       <- A1;
-    diag(A1_vm) <- NA;
-    A1vec       <- as.vector(A1_vm);
-    A1vec       <- A1vec[is.na(A1vec) == FALSE];
-    par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-    plot(A0_r, A0_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
-         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
-         asp = 1, col = "dodgerblue4");
-    vl <- seq(from = 0, to = 2*pi, by = 0.001);
-    x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
-    y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-    x1 <- sqrt(1000) * sd(A1vec) * cos(vl) + mean(diag(A1));
-    y1 <- sqrt(1000) * sd(A1vec) * sin(vl);
-    text(x = -15.5, y = 19, labels = "a", cex = 2);
-    points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
-    points(x = x1, y = y1, type = "l", col = "red", lwd = 3, lty = "dashed");
-    plot(A1_r, A1_i, xlim = c(-16.5, 15.5), ylim = c(-16.5,15.5), pch = 4, 
-         cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
-         asp = 1, col = "firebrick", yaxt = "n");
-    text(x = -15.5, y = 19, labels = "b", cex = 2);
-    points(x = x1, y = y1, type = "l", col = "firebrick", lwd = 3);
-    points(x = x0, y = y0, type = "l", lwd = 3, lty = "dashed");
-    mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-    mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-}
-
-plot_stables <- function(dat, S_s = 32){
-    Ns          <- 1:S_s;
-    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
-    #=================================
-    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
-    log_bar_dat                  <- log(bar_dat);
-    log_bar_dat[log_bar_dat < 0] <- 0; 
-    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
-            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
-            ylab = "Ln number of stable communities", cex.lab = 1, 
-            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
-    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
-    box(lwd = 2);
-    par(new = TRUE);
-    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
-    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
-    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
-         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
-         cex = 1, col = "black", yaxs="i");
-    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
-    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c(expression(paste(gamma," = 1")), 
-                        expression(paste("Var(",gamma,")"))), 
-           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
-    #=================================
-    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
-          line = 3.0);
-    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
-          outer = TRUE, line = 3.5);
-    mtext(side = 4, text = expression(
-        paste("Pr. of systems stable due to Var(",gamma,")")),
-        cex = 2, outer = TRUE, line = 3.5);
-}
-
-plot_stable_4 <- function(dat, S_s = 39){
-    Ns          <- 1:S_s;
-    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
-    #=================================
-    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
-    log_bar_dat                  <- log(bar_dat);
-    log_bar_dat[log_bar_dat < 0] <- 0; 
-    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
-            names.arg = dat[Ns,1], ylim = c(0, 14), xlab = "",
-            ylab = "Ln number of stable communities", cex.lab = 1, 
-            cex.axis = 1.25, xlim = c(1, 116), cex.names = 1, yaxt = "n");
-    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12), cex.axis = 1.5);
-    box(lwd = 2);
-    par(new = TRUE);
-    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
-    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
-    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
-         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
-         cex = 1, col = "black", yaxs="i");
-    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
-    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c(expression(paste(gamma," = 1")), 
-                        expression(paste("Var(",gamma,")"))), 
-           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
-    #=================================
-    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
-          line = 3.0);
-    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
-          outer = TRUE, line = 3.5);
-    mtext(side = 4, text = expression(
-        paste("Pr. of systems stable due to Var(",gamma,")")),
-        cex = 2, outer = TRUE, line = 3.5);
-}
-```
-
-\clearpage
-
-```{r echo = FALSE}
-S     <- 200;
-C     <- 0.05;
-sigma <- 0.4;
-pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st <- pr_st[,-1];
-```
-
-```{r, echo = FALSE}
-A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
-A0 <- as.matrix(A0[,-1]);
-A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
-A1 <- as.matrix(A1[,-1]);
-
-A0_e   <- eigen(A0)$values;
-A0_r   <- Re(A0_e);
-A0_i   <- Im(A0_e);
-A1_e   <- eigen(A1)$values;
-A1_r   <- Re(A1_e);
-A1_i   <- Im(A1_e);
-
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(t(A0_vm));
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(t(A1_vm));
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
-fhalf       <- 1:(0.5*length(A1vec));
-shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st       <- pr_st[,-1];
-```
-
-**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the blue circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
-In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
-In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` was stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
-Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "dodgerblue4");
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
-A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
-text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "blue");
-points(A0_r, A0_i, pch = 4, cex = 0.7, col = "dodgerblue4");
-
-plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "firebrick", yaxt = "n");
-
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
-A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
-points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "red");
-A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
-A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
-points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "red");
-
-points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7, col = "firebrick");
-
-text(x = -3.5, y = 2.25, labels = "b", cex = 2);
-mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-```
-
-\clearpage
-
-```{r echo = FALSE}
-S     <- 1000;
-C     <- 1;
-sigma <- 0.4;
-```
-
-```{r, echo = FALSE}
-A_comp <- NULL;
-A_dat  <- rnorm(n = 1000000, mean = 0, sd = 0.4);
-A_mat  <- matrix(data = A_dat, nrow = 1000);
-C_dat  <- rbinom(n = 1000 * 1000, size = 1, prob = 0.05);
-C_mat  <- matrix(data = C_dat, nrow = 1000, ncol = 1000);
-A_mat  <- A_mat * C_mat;
-A_mat  <- species_interactions(A_mat, type = 0);
-gammas <- runif(n = 1000, min = 0, max = 2);
-
-mu_gam <- mean(gammas);
-diag(A_mat) <- -1;
-A1     <- gammas * A_mat;
-A0     <- mu_gam * A_mat;
-A0_e   <- eigen(A0)$values;
-A0_r   <- Re(A0_e);
-A0_i   <- Im(A0_e);
-A1_e   <- eigen(A1)$values;
-A1_r   <- Re(A1_e);
-A1_i   <- Im(A1_e);
-
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(A0_vm);
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-```
-
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
-Blue elipses in both panels show the circle centred on the distribution in panel a.
-Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
-
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "dodgerblue4");
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
-y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-text(x = -15.5, y = 19, labels = "a", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3, col = "blue");
-points(A0_r, A0_i, col = "dodgerblue4",pch = 4, cex = 0.7);
-plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "firebrick", yaxt = "n");
-text(x = -15.5, y = 19, labels = "b", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid", col = "blue");
-points(A1_r, A1_i, col = "firebrick", pch = 4, cex = 0.7);
-mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
-```
-
-\clearpage
-
-```{r, echo = FALSE}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
-```
-
-**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
-For each $S$, 1 million complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
-Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Blue and red bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
-The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
-plot_stables(dat);
-```
-
-\clearpage
-
-**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
-For each $S$, 100000 complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
-Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
-The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
-dat <- dat[,-1];
-plot_stable_4(dat);
-```

### 2018-07-30 notebook/ms.tex
Rework all files for arXiv

--- 2018-06-03 notebook/ms.tex
@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\NormalTok}[1]{{#1}}
+\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
   \posttitle{\par}
-  \author{A. Bradley Duthie}
+  \author{A. Bradley Duthie (
+\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}
+)}
   \preauthor{\centering\large\emph}
@@
   \date{Biological and Environmental Sciences, University of Stirling, Stirling,
-UK, FK9 4LA
-\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}
+UK, FK9 4LA}
 
@@
 \usepackage{natbib}
-\usepackage{lineno}
 \usepackage[utf8]{inputenc}
-\linenumbers
 \bibliographystyle{amnatnat}
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
 \textbf{The stability of a complex system generally decreases with
-increasing system size, as is demonstrated by random matrix
-theory\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-This counter-intuitive result, first shown by
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}, is broadly
-relevant for understanding the dynamics and persistence of systems such
-as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{3},\protect\hyperlink{ref-Gray2009}{4}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{5},\protect\hyperlink{ref-MacArthur2010}{6}}
-and
-socio-economic\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}--\protect\hyperlink{ref-Bardoscia2017}{9}}
-networks. Much attention has especially been given to the stability of
-ecological communities such as food webs or mutualist networks, with
-recent work investigating how different community structures affect
-stability\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10}--\protect\hyperlink{ref-Patel2018}{14}}.
-But more broadly, stabilising mechanisms in complex systems remain
-under-developed, and the effect of variation in the response rate of
-individual system components remains an open
-problem\textsuperscript{\protect\hyperlink{ref-Allesina2015}{15}}. Here
-I show that when components of a complex system respond to system
-dynamics at different rates (\(\boldsymbol{\gamma}\)), the potential for
-system stability is markedly increased. Stability is caused by the
-clustering of some eigenvalues toward the centre of eigenvalue
-distributions despite the destabilising effect of higher interaction
-strength variation (\(\boldsymbol{\sigma^{2}}\)). This effect of
-variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
-system size increases, to the extent that the largest stable complex
-systems would otherwise be unstable if not for
-\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
-unconsidered driver of system stability that is likely to be pervasive
-across all complex systems. Future research in complex systems should
-therefore account for the varying response rates of individual system
-components when assessing whole system stability.}
+increasing system size and interconnectivity, a counterintuitive result
+of widespread importance across the physical, life, and social sciences.
+Despite recent interest in the relationship between system properties
+and stability, the effect of variation in the response rate of
+individual system components remains unconsidered. Here I vary the
+component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
+complex systems. I show that when component response rates vary, the
+potential for system stability is markedly increased. Variation in
+\(\boldsymbol{\gamma}\) becomes increasingly important as system size
+increases, such that the largest stable complex systems would be
+unstable if not for \(\boldsymbol{Var(\gamma)}\). My results reveal a
+previously unconsidered driver of system stability that is likely to be
+pervasive across all complex systems.}
 
@@
 May's finding that the probability of local stability falls to near zero
-given a sufficiently high threshold of \(\sigma\sqrt{SC}\) has profound
-consequences across multiple disciplines, raising the question of how
-complex systems in, e.g.,
-ecology\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10},\protect\hyperlink{ref-Grilli2017}{13},\protect\hyperlink{ref-Allesina2015}{15}}
-or
-banking\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7},\protect\hyperlink{ref-Bardoscia2017}{9},\protect\hyperlink{ref-May2008}{16}}
-are predicted to persist or change.
+given a sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly
+relevant for understanding the dynamics and persistence of systems such
+as
+ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
+neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
+biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
+and
+socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
+networks. As such, identifying general principles that drive stability
+in complex systems is of wide-ranging importance.
 
 Randomly assembled complex systems can be represented as large square
-matrices (\(M\)) with \(S\) components (e.g.,
+matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
 species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
+banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
 element of such a matrix, \(M_{ij}\), defines how component \(j\)
@@
 self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, values of off-diagonal elements are assigned non-zero
-values with a probability \(C\), which are sampled from a distribution
-with variance \(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}.
+Traditionally, off-diagonal elements are assigned non-zero values with a
+probability \(C\), which are sampled from a distribution with variance
+\(\sigma^{2}\); diagonal elements are set to
+-1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
 Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of \(M\)
-are negative
+being stable if the real parts of all eigenvalues (\(\lambda\)) of
+\(\mathbf{M}\) are negative
 (\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
 In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}} within a
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
 circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
 \(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
@@
 
-The above stability criterion assumes that individual components respond
-to perturbations of the system at the same rate (\(\gamma\)), but this
-is highly unlikely in any complex system. In ecological communities, for
-example, the rate at which population density changes following
-perturbation will depend on the generation time of individuals, which
-might vary by orders of magnitude among species. Species with short
-generation times will respond quickly (high \(\gamma\)) to perturbations
-relative to species with long generation times (low \(\gamma\)).
-Similarly, the speed at which individual banks respond to perturbations
-in financial networks, or individuals or institutions respond to
-perturbations in complex social networks, is likely to vary. The effect
-of such variance has not been investigated in complex systems theory.
-Intuitively, variation in \(\gamma\) might be expected to decrease
-system stability by introducing a new source of variation into the
-system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\gamma\) varies are
-actually more likely to be stable, especially when \(S\) is high.
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
+stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
+components respond to perturbations of the system at the same rate
+(\(\gamma\)), but this is highly unlikely in any complex system. In
+ecological communities, for example, the rate at which population
+density changes following perturbation will depend on the generation
+time of organisms, which might vary by orders of magnitude among
+species. Species with short generation times will respond quickly (high
+\(\gamma\)) to perturbations relative to species with long generation
+times (low \(\gamma\)). Similarly, the speed at which individual banks
+respond to perturbations in financial networks, or individuals or
+institutions respond to perturbations in complex social networks, is
+likely to vary. The effect of such variance on stability has not been
+investigated in complex systems theory. Intuitively, variation in
+\(\gamma\) might be expected to decrease system stability by introducing
+a new source of variation into the system and thereby increasing
+\(\sigma\). Here I show why, despite higher \(\sigma\), complex systems
+in which \(\gamma\) varies are actually more likely to be stable,
+especially when \(S\) is high.
+
+\subsection{Results}\label{results}
+
+\textbf{Component response rates of random complex systems}. Rows in
+\(\mathbf{M}\) define how a given component \(i\) is affected by other
+components of the system, meaning that the rate of component response
+time can be modelled by multiplying all row elements by a real scalar
+value
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+The distribution of \(\gamma\) over \(S\) components thereby models the
+distribution of component response rates. An instructive example
+compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
+\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
+half of \(\gamma_{i} = 0.05\). This models one system in which
+\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
+are otherwise identical (note \(E[\gamma_{i}] = 1\) in both cases). I
+assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal
+elements are set to \(-1\) and non-zero off-diagonal elements are drawn
+randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied
+by \(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
+eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
+centred at (\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real
+components of eigenvalues are highly unlikely to all be negative when
+all \(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated
+into two groups, eigenvalues are no longer uniformly distributed (Fig.
+1b). Instead, two distinct clusters of eigenvalues appear (red circles
+in Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
+(\(-0.05, 0\)). The former has a large radius, but the real components
+have shifted to the left (in comparison to when \(\gamma = 1\)), while
+the latter cluster has a smaller radius, but real components have
+shifted to the right; all \(\Re({\lambda}) < 0\). Overall, for 1 million
+randomly assembled \(\mathbf{M}\), this division between slow and fast
+component response rates results in more stable systems: 1 stable given
+\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+
+Higher stability in systems with variation in \(\gamma\) can be observed
+by sampling \(\gamma_{i}\) values from various distributions. I focus on
+a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
+Supplementary Information for other distributions, which give similar
+results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b),
+\(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
+comparison of \(\mathbf{M}\) before and after the addition of variation
+in component response rate. Figure 2 shows a comparison of eigenvalue
+distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
+expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
+\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
+centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
+Uniform variation in \(\gamma\) leads to a non-uniform distribution of
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
+some of which are clustered locally near the centre of the distribution,
+but others of which are spread outside the former radius of 2.828 (Fig
+2b). The clustering and spreading of eigenvalues introduced by
+\(Var(\gamma)\) can destabilise previously stable systems or stabilise
+systems that are otherwise unstable. But where systems are otherwise too
+complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
+can often lead to stability above
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
+threshold \(\sigma\sqrt{SC} < 1\).
+
+\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
+investigate the effect of \(Var(\gamma)\) on system stability, I
+simulated random \(\mathbf{M}\) matrices at \(\sigma = 0.4\) and
+\(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One million
+\(\mathbf{M}\) were simulated for each \(S\), and the stability of
+\(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
+\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
+number of stable random systems was higher given \(Var(\gamma)\) than
+when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
+table of results), and that the difference between the probabilities of
+observing a stable system increased with an increase in \(S\); i.e., the
+potential for \(Var(\gamma)\) to drive stability increased with system
+complexity. For the highest values of \(S\), nearly all systems that
+were stable given \(Var(\gamma)\) would not have been stable given
+\(\gamma = 1\).
+
+\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
+investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
+genetic algorithm. Genetic algorithms are heuristic tools that mimic
+evolution by natural selection, and are useful when the space of
+potential solutions (in this case, possible combinations of \(\gamma\)
+values leading to stability in a large complex system) is too large to
+search
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
+Generations of selection on \(\gamma\) value combinations to minimise
+\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
+\(Var(\gamma)\) to increase system stability. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were up to four
+orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
+that stability could often be achieved by manipulating \(S\) \(\gamma\)
+values rather than \(S \times S\) \(\mathbf{M}\) elements.
+
+\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
+systems in which individual system components represent the density of
+some tangible quantity, it is relevant to consider the feasibility of
+the system. Feasibilility assumes that the system is stable and values
+of all components are positive at
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
+This is of particular interest for ecological communities because
+population density cannot take negative values, meaning that ecological
+systems need to be feasible for stability to be biologically
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
+While my results are intended to be general to all complex systems, and
+not restricted to species networks, I have also performed a feasibility
+analysis on all matrices \(\mathbf{M}\) tested for stability, and
+additionally for specific types of ecological
+communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+(e.g., competitive, mutualist, predator-prey; see Supplementary
+Information). Feasibility was unaffected by \(Var(\gamma)\), meaning
+that for pure interacting species networks, variation in component
+response rate (i.e., species generation time) does not affect stability
+at biologically realistic species densities.
 
@@
 Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the grey circle with a radius of
+uniformly distributed within the blue circle with a radius of
 \(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
@@
 \(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\gamma = -1.95\) and one small circle of \(\sigma\sqrt{SC} =\) 0.044
-centred at \(\gamma = -0.05\). In a, the maximum real eigenvalue
+large circle centred at \(\gamma = -1.95\) and one small circle centred
+at \(\gamma = -0.05\). In a, the maximum real eigenvalue
 \(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-4-1.pdf}
+\includegraphics{unnamed-chunk-10-1.pdf}
 
@@
 
-Rows in \(M\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}}.
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(M\) where \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
-the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
-\(\gamma_{i} = 0.05\). This models one system in which \(\gamma\) is
-invariant and one in which \(\gamma\) varies, but systems are otherwise
-identical (note \(E[\gamma_{i}] = 1\) in both cases). I assume
-\(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
-set to \(-1\) and non-zero off-diagonal elements are drawn randomly from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
-\(\gamma_{i}\) to generate \(M\). When \(\gamma_{i} = 1\), eigenvalues
-of \(M\) are distributed uniformly within a circle centred at
-(\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real components
-of eigenvalues are highly unlikely to all be negative when all
-\(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated into
-two groups, eigenvalues are no longer uniformly distributed (Fig. 1b).
-Instead, two distinct clusters of eigenvalues appear (grey circles in
-Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
-(\(-0.05, 0\)). The former has a large radius, but the real components
-have shifted to the left (in comparison to when \(\gamma = 1\)) and all
-\(\Re({\lambda}) < 0\). The latter cluster has real components that have
-shifted to the right, but has a smaller radius. Overall, for 1 million
-randomly assembled \(M\), this division between slow and fast component
-response rates results in more stable systems: 1 stable given
-\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+\subsection{Discussion}\label{discussion}
+
+Here I have shown that the stability of large systems might often be
+contigent upon variation in the response rates of their individual
+components, meaning that factors such as rate of trait evolution (in
+biological networks), transaction speed (in economic networks), or
+communication speed (in social networks) need to be considered when
+investigating the stability of complex systems. Variation in component
+response rate becomes more likely to be critical for stability as system
+size increases, and can ultimately increase the overall probability that
+system stability is observed above that predicted by
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
+derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
+general, and potentially applies to any complex system in which
+individual system components can vary in their reaction rates to system
+perturbation.
 
@@
 (\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
-Eigenvalues plotted in the absence of \(Var(\gamma)\) where
+system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
+\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
 \(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
 \(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
-the distribution's centre (-1, 0). Black and red elipses in both panels
-show the circle centred on the distribution in panels a and b,
-respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.549 and 0.547 for a and b, respectively.
+interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
+eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
+both panels show the circle centred on the distribution in panel a.
+Proportions of \(\Re(\lambda) < 0\) are 0.719 and 0.735 for a and b,
+respectively.
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{unnamed-chunk-13-1.pdf}
 
@@
 
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I now
-focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
-(see Supplementary Information for other distributions, which give
-similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
-1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(M\) before and after variation in component response
-rate. Figure 2 shows a comparison of eigenvalue distributions given
-\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues, some of which are clustered tightly around the centre of
-the distribution, but others of which are spread outside the former
-radius of 12.649 (red circle Fig 2b). This larger radius occurs because
-the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
-\(M\). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold of \(\sigma\sqrt{SC} > 1\).
+It is important to emphasise that variation in component response rate
+is not stabilising per se; that is, adding variation in component
+response rate to a particular system does not necessarily increase the
+probability that the system will be stable. Rather, systems that are
+observed to be stable are more likely to have varying component response
+rates, and for this variation to be critical to their stability (Fig.
+3). This is caused by the shift to a non-uniform distribution of
+eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
+which can sometimes cause all of the real components of the eigenvalues
+of the community matrix \(\mathbf{M}\) to become negative, but might
+also increase the real components of eigenvalues. The mathematics
+underlying this shift in eigenvalue distribution has been
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+and recently applied to questions concerning species density and
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
+but has not been interpreted as rates of response of individual system
+components to perturbation.
+
+The potential importance of component response rate variation was most
+evident from the results of simulations in which the genetic algorithm
+was used in attempt to maximise the probability of system stability. The
+probability that some combination of component response rates could be
+found to stabilise the system was shown to be up to four orders of
+magnitude higher than the background probabilities of stability in the
+absence of any component response rate variation. Instead of
+manipulating the \(S \times S\) interactions between system components,
+it might therefore be possible to manipulate only the \(S\) response
+rates of individual system components to achieve stability. Hence,
+managing the response rates of system components in a targeted way could
+potentially facilitate the stabilisation of complex systems through a
+reduction in dimensionality.
 
@@
 \textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate(\(\boldsymbol{\gamma}\)).} The
+variation in component response rate (\(\boldsymbol{\gamma}\)).} The
 \(\ln\) number of systems that are stable across different system sizes
-(\(S\)) given \(C = 1\), and the proportion of systems in which
-variation in \(\gamma\) is critical for system stability. For each
+(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
+which variation in \(\gamma\) is critical for system stability. For each
 \(S\), 1 million complex systems are randomly generated. Stability of
@@
 system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Light and dark grey bars show the number of stable systems in the
-absence and presence of \(Var(\gamma)\), respectively. The black line
-shows the proportion of systems that are stable when
-\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
+Blue and red bars show the number of stable systems in the absence and
+presence of \(Var(\gamma)\), respectively. The black line shows the
+proportion of systems that are stable when \(Var(\gamma) > 0\), but
+would be unstable if \(Var(\gamma) = 0\).
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{unnamed-chunk-15-1.pdf}
 
@@
 
-To investigate the effect of \(Var(\gamma)\) on system stability, I
-simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32. One million \(M\) were simulated for each
-\(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of
-stable random systems was consistently higher given \(Var(\gamma)\) than
-when \(\gamma = 1\) (Fig. 3), and that the difference between the
-probabilities of observing a stable system increased with an increase in
-\(S\); i.e., the potential for \(Var(\gamma)\) to drive stability
-increased with system complexity. For the highest values of \(S\),
-nearly all systems that were stable given \(Var(\gamma)\) would not have
-been stable given \(\gamma = 1\) (see Supplementary Information for full
-results). This suggests that the stability of large systems might be
-dependent upon variation in the response rate of their individual
-components, meaning that factors such as generation time (in ecological
-networks), transaction speed (in economic networks), or communication
-speed (in social networks) needs to be considered when investigating the
-stability of complex systems.
+Interestingly, while complex systems were more likely to be stable given
+variation in component response rate, they were not more likely to be
+feasible, meaning that stability was not increased when component values
+were also restricted to being positive at equilibrium. Feasibility is
+important to consider, particularly for the study of ecological networks
+of
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20}}
+because population densities cannot realistically be negative. My
+results therefore suggest that variation in the rate of population
+responses to perturbation (e.g., due to differences in generation time
+among species) is unlikely to be critical to the stability of purely
+multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in
+empirical
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
+instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability
+should therefore not be ignored in the broader context of ecological
+communities.
 
-It is important to point out that \(Var(\gamma)\) is not stabilising per
-se; that is, adding variation in \(\gamma\) to a particular system \(M\)
-does not necessarily increase the probability that the system will be
-stable (see Supplementary Information). Rather, systems that are
-observed to be stable are more likely to vary in \(\gamma\), and for
-this \(Var(\gamma)\) to be critical to their stability. This is caused
-by the shift in the distribution of eigenvalues that occurs by
-introducing \(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in
-all \(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\)
-values. To further investigate the potential of \(Var(\gamma)\) to be
-stabilising, I used a genetic algorithm (the space of possible
-\(\gamma\) values was too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
-see Supplementary Information). For each of 10000 random \(M\), the
-genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis
-was performed on \(M\) using each set of \(\gamma\) values, and the 20
-sets with the lowest \(\max\left(\Re(\lambda)\right)\) each produced 50
-offspring with subsequent mutation and crossover between the resulting
-new population of 1000 \(\gamma\) sets. The genetic algorithm terminated
-if a stable \(M\) was found, 20 generations occurred, or a convergence
-criteria of minimum fitness increase between generations was satisfied.
-Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were
-found that resulted in stable systems with probabilities that were
-orders of magnitude higher than when \(\gamma = 1\) (see Supporting
-Information), meaning that stability could often be achieved by
-manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\)
-elements. Hence, managing the response rates of system components in a
-targetted way can potentially facilitate the stabilisation of complex
-systems through a reduction in dimensionality.
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
-I have focused broadly on random complex systems, but it is also
-worthwhile to consider more restricted interactions such as those of
-specific ecological
-networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (competitive
-networks), positive (mutualist networks), or \(i\) and \(j\) pairs have
-opposing signs (predator-prey networks). In general, competitive and
-mutualist networks tend to be destabilising, and predator-prey network
-tend to be
-stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
-When \(Var(\gamma)\) is applied to each, the proportion of stable
-competitive and predator-prey networks increases, but the proportion of
-stable mutualist networks does not (see Supplementary Information).
-Additionally, when each component of \(M\) is interpreted as a unique
-species and given a random intrinsic growth
-rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
-feasibility is not increased by \(Var(\gamma)\), suggesting that
-variation in species generation time might be unlikely to drive
-stability in purely multi-species networks (see Supplementary
-Information).
+\textbf{Figure 4: Stability of large complex systems given
+\(\boldsymbol{\gamma = 1}\) versus targeted
+\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
+stable across different system sizes (\(S\), max \(S=40\)) for
+\(C = 1\), and the proportion of systems wherein a targeted search of
+\(\gamma\) values successfully resulted in system stability. For each
+\(S\), 100000 complex systems are randomly generated. Stability of each
+complex system is tested given variation in \(\gamma\) using a genetic
+algorithm to maximise the effect of \(\gamma\) values on increasing
+stability, as compared to stability in an otherwise identical system in
+which \(\gamma\) is the same for all components. Blue bars show the
+number of stable systems in the absence of component response rate
+variation, while red bars show the number of stable systems that can be
+generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(Var(\gamma) = 0\).
+
+\includegraphics{unnamed-chunk-16-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 the response rates of system components vary. These results are broadly
-applicable to complex biological and social networks.
+applicable to understanding stability of complex networks in the
+physical, life, and social sciences.
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\subsection{Methods}\label{methods}
+
+\textbf{Component response rate variation (\(\mathbf{\gamma}\))}. In a
+synthesis of eco-evolutionary feedbacks on community stability, Patel et
+al. model a system that includes a vector of potentially changing
+species densities (\(\mathbf{N}\)) and a vector of potentially evolving
+traits
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+For any species \(i\) or trait \(j\), change in species density
+(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+
+\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
+
+\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+
+In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
+effects of all species densities and trait values on the density of a
+species \(i\) and the value of trait \(j\), respectively. Patel et al.
+were interested in stability when the evolution of traits was relatively
+slow or fast in comparison with the change in species
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
+this is modulated in the above by the scalar \(\epsilon\). The value of
+\(\epsilon\) thereby determines the timescale separation between ecology
+and evolution, with high \(\epsilon\) modelling relatively fast
+evolution and low \(\epsilon\) modelling relative slow
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+
+I use the same principle that Patel et al. use to modulate the relative
+rate of evolution to modulate rates of component responses for \(S\)
+components. Following
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{22}},
+the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
+by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
+\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
+
+\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).\]
+
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{22}},
+
+\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).\]
+
+In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
+correspond to individual component response rates. Therefore,
+\(\mathbf{M} =\mathbf{\gamma} \mathbf{A}\) modulates the values of
+components and can be analysed using the techniques of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{22}}.
+
+\textbf{Genetic algorithm}. Ideally, to investigate the potential of
+\(Var(\gamma)\) for increasing the proportion of stable complex systems,
+the search space of all possible \(\gamma\) vectors would be evaluated
+for each unique \(\mathbf{M = \gamma A}\). This is technically
+impossible because \(\gamma_{i}\) can take any real value between 0-2,
+but even rounding \(\gamma_{i}\) to reasonable values would result in a
+search space too large to practically explore. Under these conditions,
+genetic algorithms are highly useful tools for finding practical
+solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
+this case, the practical solution is finding vectors of
+\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
+\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
+a large population of 1000 different potential \(\mathbf{\gamma}\)
+vectors and allowing this population to evolve through a process of
+mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
+selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
+found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
+met.
+
+For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
+for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)), where
+\(\mathbf{M = \gamma A}\). The genetic algorithm was initialised with a
+population of 1000 different \(\mathbf{\gamma}\) vectors with elements
+sampled i.i.d from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis
+was performed on the \(\mathbf{M}\) resulting from each
+\(\mathbf{\gamma}\) vector, and the 20 \(\mathbf{\gamma}\) vectors
+resulting in \(\mathbf{M}\) with the lowest
+\(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
+subsequent random mutation and crossover between the resulting new
+generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
+\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+probability of 0.2, resulting in a mutation effect of size
+\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
+\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
+multiplied by \(-1\), and any values that mutated above 2 were set to
+2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
+vectors paired in each generation; vectors were randomly sampled with
+replacement among but not within sets. Vector pairs selected for
+crossover swapped all elements between and including two \(\gamma_{i}\)
+randomly selected with replacement (this allowed for reversal of vector
+element positions during crossover; e.g.,
+\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
+). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
+found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
+fitness increase between generations was less than 0.01 (where fitness
+was defined as \(W_{\gamma} = -\max\left(\Re(\lambda)\right)\) for
+\(\mathbf{M}\)).
+
+\textbf{System feasibility}. Dougoud et
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} define the
+following feasibility criteria for ecological systems characterised by
+\(S\) interacting species with varying densities in a classical
+Lotka-Volterra model,
+
+\[\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]
+
+In the above, \(\mathbf{x^{*}}\) is the vector of species densities at
+equilibrium. Feasibility is satisfied if all elements in
+\(\mathbf{x^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
+identity matrix, and the value \(\theta\) is the strength of
+intraspecific competition (diagonal elements). Diagonal values are set
+to \(-1\), so \(\theta = -1\). The variable \(\delta\) is a
+normalisation parameter that modulates the strength of interactions
+(\(\sigma\)) for \(\mathbf{J}\). Implicitly, here \(\delta = 0\)
+underlying strong interactions. Hence, \((CS)^{-\delta} = 1\), so in the
+above, a diagonal matrix of -1s (\(\theta \mathbf{I}\)) is added to
+\(\mathbf{J}\), which has a diagonal of all zeros and an off-diagonal
+affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
+relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+stability
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
+\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
+purposes \((CS)^{-\delta} = 1\)). Given
+\(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
+reduced to the below,
+
+\[\mathbf{x^{*} = -A^{-1}r}.\]
+
+To check the feasibility criteria for \(\mathbf{M = \gamma A}\), I
+therefore evaluated \(\mathbf{-M^{-1}r}\) (\(\mathbf{r}\) elements were
+sampled i.i.d. from \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility
+is satisfied if all of the elements of the resulting vector are
+positive.
+
+\textbf{Acknowledgements:} I am supported by a Leverhulme Trust Early
+Career Fellowship (ECF-2016-376). Conversations with L. Bussière and N.
+Bunnefeld, and comments from J. J. Cusack and I. L. Jones, improved the
+quality of this work.
+
+\textbf{Supplementary Information:} Full tables of stability results for
+simulations across different system size (\(S\)) values, ecological
+community types, connectance (\(C\)) values, interaction strengths
+(\(\sigma\)), and \(\gamma\) distributions are provided as supplementary
+material. An additional table also shows results for how feasibility
+changes across \(S\). All code and simulation outputs are publicly
+available as part of the RandomMatrixStability package on GitHub
+(\url{https://github.com/bradduthie/RandomMatrixStability}).
 
@@
 
+\hypertarget{ref-Mougi2012}{}
+3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
+community stability. \emph{Science} \textbf{337,} 349--351 (2012).
+
+\hypertarget{ref-Allesina2015}{}
+4. Allesina, S. \emph{et al.} Predicting the stability of large
+structured food webs. \emph{Nature Communications} \textbf{6,} 7842
+(2015).
+
+\hypertarget{ref-Grilli2017}{}
+5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
+ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
+
 \hypertarget{ref-Gray2008}{}
-3. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
+6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
 random brain networks with a distribution of connection strengths.
@@
 \hypertarget{ref-Gray2009}{}
-4. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
+7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
 with excitatory and inhibitory connections. \emph{Neurocomputing}
@@
 \hypertarget{ref-Rosenfeld2009}{}
-5. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
+8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
 high-dimensional biochemical networks. \emph{Gene Regulation and Systems
@@
 \hypertarget{ref-MacArthur2010}{}
-6. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
+9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
 and criticality of adaptive regulatory networks. \emph{Physics Review
@@
 
+\hypertarget{ref-May2008}{}
+10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
+for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
+
 \hypertarget{ref-Haldane2011}{}
-7. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
+11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
 \emph{Nature} \textbf{469,} 351--355 (2011).
@@
 \hypertarget{ref-Suweis2014}{}
-8. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
+12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
 networks. \emph{PLoS ONE} \textbf{9,} (2014).
@@
 \hypertarget{ref-Bardoscia2017}{}
-9. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G. Pathways
-towards instability in financial networks. \emph{Nature Communications}
-\textbf{8,} 1--7 (2017).
+13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
+Pathways towards instability in financial networks. \emph{Nature
+Communications} \textbf{8,} 1--7 (2017).
 
-\hypertarget{ref-Mougi2012}{}
-10. Mougi, A. \& Kondoh, M. Diversity of interaction types and
-ecological community stability. \emph{Science} \textbf{337,} 349--351
-(2012).
-
-\hypertarget{ref-Allesina2015a}{}
-11. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
-(2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Gao2016}{}
-12. Gao, J., Barzel, B. \& Barabási, A. L. Universal resilience patterns
-in complex networks. \emph{Nature} \textbf{530,} 307--312 (2016).
-
-\hypertarget{ref-Grilli2017}{}
-13. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
+\hypertarget{ref-Tao2010}{}
+14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
+circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
+(2010).
 
 \hypertarget{ref-Patel2018}{}
-14. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 
-\hypertarget{ref-Allesina2015}{}
-15. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
-(2015).
+\hypertarget{ref-Ahmadian2015}{}
+16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+with partially structured and partially random connectivity.
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{91,} 012820 (2015).
 
-\hypertarget{ref-May2008}{}
-16. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
-for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
+\hypertarget{ref-Gibbs2017}{}
+17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{arXiv} (2017).
 
-\hypertarget{ref-Tao2010}{}
-17. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
-(2010).
+\hypertarget{ref-Stone2017}{}
+18. Stone, L. The feasibility and stability of large complex biological
+networks: a random matrix approach. \emph{Scientific Reports}
+\textbf{8,} 8246 (2018).
 
 \hypertarget{ref-Hamblin2013}{}
-18. Hamblin, S. On the practical usage of genetic algorithms in ecology
+19. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
 184--194 (2013).
-
-\hypertarget{ref-Allesina2011}{}
-19. Allesina, S. \& Levine, J. M. A competitive network theory of
-species diversity. \emph{Proceedings of the National Academy of Sciences
-of the United States of America} \textbf{108,} 5638--5642 (2011).
 
@@
 
+\hypertarget{ref-Song2018}{}
+21. Song, C. \& Saavedra, S. Will a small randomly assembled community
+be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
+
+\hypertarget{ref-May1973}{}
+22. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+\textbf{54,} 638--641 (1973).
+
+\clearpage
+
+\begin{center}
+\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
+\end{center}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{This supplemental information supports the manuscript
+``Component response rate variation drives stability in large complex
+systems'' with additional analyses to support its conclusions. All text,
+code, and data underlying this manuscript are publicly available on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
+part of the RandomMatrixStability R package.}
+
+The
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
+package} includes all functions and tools for recreating the text, this
+supplemental information, and running all code; additional documentation
+is also provided for package functions. The RandomMatrixStability
+package is available on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
+download it, the
+\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
+library} is needed.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
+\KeywordTok{library}\NormalTok{(devtools);}
+\end{Highlighting}
+\end{Shaded}
+
+The code below installs the RandomMatrixStability package using
+devtools.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{);}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\section{Supplemental Information table of
+contents}\label{supplemental-information-table-of-contents}
+
+\begin{itemize}
+\tightlist
+\item
+  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
+\item
+  \protect\hyperlink{ga}{Stability given targeted manipulation of
+  \(\gamma\) (genetic algorithm)}
+\item
+  \protect\hyperlink{ecological}{Stability of ecological networks}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{competition}{Competitor networks}
+  \item
+    \protect\hyperlink{mutualism}{Mutualist networks}
+  \item
+    \protect\hyperlink{pred-prey}{Predator-prey networks}
+  \end{itemize}
+\item
+  \protect\hyperlink{connectance}{Sensitivity of connectance (C) values}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{connect3}{C = 0.3}
+  \item
+    \protect\hyperlink{connect5}{C = 0.5}
+  \item
+    \protect\hyperlink{connect7}{C = 0.7}
+  \item
+    \protect\hyperlink{connect9}{C = 0.9}
+  \end{itemize}
+\item
+  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
+  and \(\sigma\)}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
+  \item
+    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
+  \item
+    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
+  \item
+    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
+  \end{itemize}
+\item
+  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
+  \(\gamma\)}
+\item
+  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
+\end{itemize}
+
+\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
+\(S\)}{Stability across increasing S}}\label{IncrS}}
+
+Figure 3 of the main text reports the number of stable random complex
+systems found over 1 million iterations. The table below shows the
+results for all simulations of random \(\mathbf{M}\) matrices at
+\(\sigma = 0.4\) and \(C = 1\) given a range of
+\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
+matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
+after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
+Each row summarises data for a given \(S\) over 1 million randomly
+simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
+\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
+unstable, and the column \texttt{A0\_stable} shows the number of
+\texttt{A0} matrices that are stable (these two columns sum to 1
+million). Similarly, the column \texttt{A1\_unstable} shows the number
+of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
+the number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 293 & 999707 & 293 & 999707 & 0 & 0\tabularnewline
+3 & 3602 & 996398 & 3609 & 996391 & 0 & 7\tabularnewline
+4 & 14937 & 985063 & 15008 & 984992 & 0 & 71\tabularnewline
+5 & 39289 & 960711 & 39783 & 960217 & 36 & 530\tabularnewline
+6 & 78845 & 921155 & 80207 & 919793 & 389 & 1751\tabularnewline
+7 & 133764 & 866236 & 136904 & 863096 & 1679 & 4819\tabularnewline
+8 & 204112 & 795888 & 208241 & 791759 & 5391 & 9520\tabularnewline
+9 & 288041 & 711959 & 291775 & 708225 & 12619 & 16353\tabularnewline
+10 & 384024 & 615976 & 384931 & 615069 & 23153 & 24060\tabularnewline
+11 & 485975 & 514025 & 481019 & 518981 & 35681 & 30725\tabularnewline
+12 & 590453 & 409547 & 577439 & 422561 & 48302 & 35288\tabularnewline
+13 & 689643 & 310357 & 669440 & 330560 & 57194 & 36991\tabularnewline
+14 & 777496 & 222504 & 751433 & 248567 & 60959 & 34896\tabularnewline
+15 & 850159 & 149841 & 821613 & 178387 & 58567 & 30021\tabularnewline
+16 & 905057 & 94943 & 877481 & 122519 & 51255 & 23679\tabularnewline
+17 & 943192 & 56808 & 919536 & 80464 & 40854 & 17198\tabularnewline
+18 & 969018 & 30982 & 949944 & 50056 & 30102 & 11028\tabularnewline
+19 & 984301 & 15699 & 970703 & 29297 & 20065 & 6467\tabularnewline
+20 & 992601 & 7399 & 983507 & 16493 & 12587 & 3493\tabularnewline
+21 & 996765 & 3235 & 991532 & 8468 & 7030 & 1797\tabularnewline
+22 & 998693 & 1307 & 995567 & 4433 & 3884 & 758\tabularnewline
+23 & 999503 & 497 & 997941 & 2059 & 1883 & 321\tabularnewline
+24 & 999861 & 139 & 999059 & 941 & 899 & 97\tabularnewline
+25 & 999964 & 36 & 999617 & 383 & 380 & 33\tabularnewline
+26 & 999993 & 7 & 999878 & 122 & 121 & 6\tabularnewline
+27 & 999995 & 5 & 999946 & 54 & 53 & 4\tabularnewline
+28 & 1000000 & 0 & 999975 & 25 & 25 & 0\tabularnewline
+29 & 1000000 & 0 & 999997 & 3 & 3 & 0\tabularnewline
+30 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+31 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
+matrices found is greater than 1 whenever \(S > 10\) (compare column 5
+to column 3), and this ratio increases with increasing \(S\) (column 1).
+Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
+given variation in \(\gamma\) than when \(\gamma = 1\). Note that
+feasibility results were omitted for the table above, but are
+\protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+Figure 4 of the main text reports the number of stable random complex
+systems found over 100000 using the genetic algorithm to maximise
+stability with a vector \(\mathbf{\gamma}\). Stability results for
+100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
+for \texttt{A0} indicate systems in which \(\gamma = 1\), while
+\texttt{A1} refers to systems in which the genetic algorithm searched
+for a set of \(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(\mathbf{M}\) that can be produced through a targeted search of
+\(\gamma\) values, and suggests that many otherwise unstable systems
+could potentially be stabilised by an informed manipulation of their
+component response times. Such a possibility might conceivably reduce
+the dimensionality of problems involving stability in social-ecological
+or economic systems.
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\includegraphics{unnamed-chunk-7-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix
+\(\mathbf{M}\).
+
+\hypertarget{ecological}{\section{Stability of ecological
+networks}\label{ecological}}
+
+While the foundational work of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
+to complex networks, much attention has been given specifically to
+ecological networks of interacting species. In these networks, the
+matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
+and column is interpreted as a single species. The per capita effect
+that the density of any species \(i\) has on the population dynamics of
+species \(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds
+the effects of pair-wise interactions between \(S\)
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
+While May's original
+work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
+randomly assembled communities, recent work has specifically looked at
+more restricted ecological communities including competitive networks
+(all off-diagonal elements of \(\mathbf{A}\) are negative), mutualist
+networks (all off-diagonal elements of \(\mathbf{A}\) are positive), and
+predator-prey networks (for any pair of \(i\) and \(j\), the effect of
+\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
+versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
+In general, competitor and mutualist networks tend to be unstable, while
+predator-prey networks tend to be highly
+stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
+
+I investigated competitor, mutualist, and predator-prey networks
+following Allesina et
+al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
+these networks, I first generated a random matrix \(\mathbf{A}\), then
+changed the elements of \(\mathbf{A}\) accordingly. If \(\mathbf{A}\)
+was a competitive network, then the sign of any positive off-diagonal
+elements was reversed to be negative. If \(\mathbf{A}\) was a mutualist
+network, then the sign of any positive off-diagonal elements was
+reversed to be positive. And if \(\mathbf{A}\) was a predator-prey
+network, then all \(i\) and \(j\) pairs of elements were checked; any
+pairs of the same sign were changed so that one was negative and the
+other was positive.
+
+The number of stable \(\mathbf{M = \gamma A}\) systems was estimated
+\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
+matrices for values of \(S\) from 2 to 50 (100 in the case of the
+relatively more stable predator-prey interactions), except that only
+100000 random \(\mathbf{M}\) were generated instead of 1 million.
+
+The following tables for restricted ecological communities can therefore
+be compared with the random \(\mathbf{M}\)
+\protect\hyperlink{IncrS}{results above} (but note that counts from
+systems with comparable probabilities of stability will be an order of
+magnitude lower in the tables below due to the smaller number of
+\(\mathbf{M}\) matrices generated). As with the
+\protect\hyperlink{IncrS}{results above}, in the tables below,
+\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
+refers to matrices after \(Var(\gamma)\) is added. The column
+\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
+unstable, and the column \texttt{A0\_stable} shows the number of
+\texttt{A0} matrices that are stable (these two columns sum to 100000).
+Similarly, the column \texttt{A1\_unstable} shows the number of
+\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
+number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+\textbf{Competition}
+
+Results for competitor interaction networks are shown below
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+3 & 229 & 99771 & 231 & 99769 & 0 & 2\tabularnewline
+4 & 701 & 99299 & 704 & 99296 & 0 & 3\tabularnewline
+5 & 1579 & 98421 & 1587 & 98413 & 0 & 8\tabularnewline
+6 & 3218 & 96782 & 3253 & 96747 & 6 & 41\tabularnewline
+7 & 5519 & 94481 & 5619 & 94381 & 23 & 123\tabularnewline
+8 & 9062 & 90938 & 9237 & 90763 & 77 & 252\tabularnewline
+9 & 13436 & 86564 & 13729 & 86271 & 230 & 523\tabularnewline
+10 & 18911 & 81089 & 19303 & 80697 & 505 & 897\tabularnewline
+11 & 25594 & 74406 & 25961 & 74039 & 1011 & 1378\tabularnewline
+12 & 33207 & 66793 & 33382 & 66618 & 1724 & 1899\tabularnewline
+13 & 41160 & 58840 & 41089 & 58911 & 2655 & 2584\tabularnewline
+14 & 50575 & 49425 & 49894 & 50106 & 3777 & 3096\tabularnewline
+15 & 59250 & 40750 & 57892 & 42108 & 4824 & 3466\tabularnewline
+16 & 67811 & 32189 & 65740 & 34260 & 5634 & 3563\tabularnewline
+17 & 75483 & 24517 & 73056 & 26944 & 5943 & 3516\tabularnewline
+18 & 82551 & 17449 & 79878 & 20122 & 5780 & 3107\tabularnewline
+19 & 88030 & 11970 & 85204 & 14796 & 5417 & 2591\tabularnewline
+20 & 92254 & 7746 & 89766 & 10234 & 4544 & 2056\tabularnewline
+21 & 95233 & 4767 & 93002 & 6998 & 3695 & 1464\tabularnewline
+22 & 97317 & 2683 & 95451 & 4549 & 2803 & 937\tabularnewline
+23 & 98508 & 1492 & 97122 & 2878 & 1991 & 605\tabularnewline
+24 & 99240 & 760 & 98407 & 1593 & 1216 & 383\tabularnewline
+25 & 99669 & 331 & 99082 & 918 & 739 & 152\tabularnewline
+26 & 99871 & 129 & 99490 & 510 & 452 & 71\tabularnewline
+27 & 99938 & 62 & 99732 & 268 & 240 & 34\tabularnewline
+28 & 99985 & 15 & 99888 & 112 & 108 & 11\tabularnewline
+29 & 99990 & 10 & 99951 & 49 & 46 & 7\tabularnewline
+30 & 100000 & 0 & 99981 & 19 & 19 & 0\tabularnewline
+31 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
+32 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
+33 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+34 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Mutualism}
+
+Results for mutualist interaction networks are shown below
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 56 & 99944 & 56 & 99944 & 0 & 0\tabularnewline
+3 & 3301 & 96699 & 3301 & 96699 & 0 & 0\tabularnewline
+4 & 34446 & 65554 & 34446 & 65554 & 0 & 0\tabularnewline
+5 & 86520 & 13480 & 86520 & 13480 & 0 & 0\tabularnewline
+6 & 99683 & 317 & 99683 & 317 & 0 & 0\tabularnewline
+7 & 99998 & 2 & 99998 & 2 & 0 & 0\tabularnewline
+8 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+9 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+11 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+12 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Predator-prey}
+
+Results for predator-prey interaction networks are shown below
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+3 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+4 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+5 & 1 & 99999 & 1 & 99999 & 0 & 0\tabularnewline
+6 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+7 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+8 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+9 & 20 & 99980 & 21 & 99979 & 0 & 1\tabularnewline
+10 & 20 & 99980 & 22 & 99978 & 0 & 2\tabularnewline
+11 & 38 & 99962 & 39 & 99961 & 0 & 1\tabularnewline
+12 & 64 & 99936 & 66 & 99934 & 0 & 2\tabularnewline
+13 & 87 & 99913 & 91 & 99909 & 0 & 4\tabularnewline
+14 & 157 & 99843 & 159 & 99841 & 0 & 2\tabularnewline
+15 & 215 & 99785 & 227 & 99773 & 0 & 12\tabularnewline
+16 & 293 & 99707 & 310 & 99690 & 0 & 17\tabularnewline
+17 & 383 & 99617 & 408 & 99592 & 0 & 25\tabularnewline
+18 & 443 & 99557 & 473 & 99527 & 3 & 33\tabularnewline
+19 & 642 & 99358 & 675 & 99325 & 4 & 37\tabularnewline
+20 & 836 & 99164 & 887 & 99113 & 7 & 58\tabularnewline
+21 & 1006 & 98994 & 1058 & 98942 & 10 & 62\tabularnewline
+22 & 1153 & 98847 & 1228 & 98772 & 20 & 95\tabularnewline
+23 & 1501 & 98499 & 1593 & 98407 & 30 & 122\tabularnewline
+24 & 1841 & 98159 & 1996 & 98004 & 40 & 195\tabularnewline
+25 & 2146 & 97854 & 2316 & 97684 & 58 & 228\tabularnewline
+26 & 2643 & 97357 & 2809 & 97191 & 119 & 285\tabularnewline
+27 & 3034 & 96966 & 3258 & 96742 & 158 & 382\tabularnewline
+28 & 3690 & 96310 & 3928 & 96072 & 201 & 439\tabularnewline
+29 & 4257 & 95743 & 4532 & 95468 & 290 & 565\tabularnewline
+30 & 4964 & 95036 & 5221 & 94779 & 424 & 681\tabularnewline
+31 & 5627 & 94373 & 5978 & 94022 & 452 & 803\tabularnewline
+32 & 6543 & 93457 & 6891 & 93109 & 666 & 1014\tabularnewline
+33 & 7425 & 92575 & 7777 & 92223 & 818 & 1170\tabularnewline
+34 & 8540 & 91460 & 8841 & 91159 & 1071 & 1372\tabularnewline
+35 & 9526 & 90474 & 9842 & 90158 & 1337 & 1653\tabularnewline
+36 & 10617 & 89383 & 10891 & 89109 & 1624 & 1898\tabularnewline
+37 & 12344 & 87656 & 12508 & 87492 & 2021 & 2185\tabularnewline
+38 & 13675 & 86325 & 13877 & 86123 & 2442 & 2644\tabularnewline
+39 & 15264 & 84736 & 15349 & 84651 & 2870 & 2955\tabularnewline
+40 & 17026 & 82974 & 17053 & 82947 & 3363 & 3390\tabularnewline
+41 & 18768 & 81232 & 18614 & 81386 & 3905 & 3751\tabularnewline
+42 & 20791 & 79209 & 20470 & 79530 & 4579 & 4258\tabularnewline
+43 & 23150 & 76850 & 22754 & 77246 & 5217 & 4821\tabularnewline
+44 & 25449 & 74551 & 24184 & 75816 & 6285 & 5020\tabularnewline
+45 & 27702 & 72298 & 26464 & 73536 & 6754 & 5516\tabularnewline
+46 & 30525 & 69475 & 28966 & 71034 & 7646 & 6087\tabularnewline
+47 & 32832 & 67168 & 31125 & 68875 & 8487 & 6780\tabularnewline
+48 & 36152 & 63848 & 33865 & 66135 & 9479 & 7192\tabularnewline
+49 & 38714 & 61286 & 36242 & 63758 & 10125 & 7653\tabularnewline
+50 & 41628 & 58372 & 38508 & 61492 & 11036 & 7916\tabularnewline
+51 & 44483 & 55517 & 41023 & 58977 & 11704 & 8244\tabularnewline
+52 & 48134 & 51866 & 44287 & 55713 & 12573 & 8726\tabularnewline
+53 & 51138 & 48862 & 46721 & 53279 & 13223 & 8806\tabularnewline
+54 & 54261 & 45739 & 49559 & 50441 & 13757 & 9055\tabularnewline
+55 & 57647 & 42353 & 52403 & 47597 & 14324 & 9080\tabularnewline
+56 & 60630 & 39370 & 55293 & 44707 & 14669 & 9332\tabularnewline
+57 & 63647 & 36353 & 57787 & 42213 & 15103 & 9243\tabularnewline
+58 & 66961 & 33039 & 60439 & 39561 & 15450 & 8928\tabularnewline
+59 & 69968 & 30032 & 63708 & 36292 & 15246 & 8986\tabularnewline
+60 & 72838 & 27162 & 66270 & 33730 & 15177 & 8609\tabularnewline
+61 & 75609 & 24391 & 68873 & 31127 & 15006 & 8270\tabularnewline
+62 & 77999 & 22001 & 71318 & 28682 & 14538 & 7857\tabularnewline
+63 & 80616 & 19384 & 73517 & 26483 & 14510 & 7411\tabularnewline
+64 & 83089 & 16911 & 76209 & 23791 & 13784 & 6904\tabularnewline
+65 & 85150 & 14850 & 78086 & 21914 & 13412 & 6348\tabularnewline
+66 & 86908 & 13092 & 80437 & 19563 & 12477 & 6006\tabularnewline
+67 & 88671 & 11329 & 82379 & 17621 & 11718 & 5426\tabularnewline
+68 & 90537 & 9463 & 84483 & 15517 & 10878 & 4824\tabularnewline
+69 & 91969 & 8031 & 86233 & 13767 & 10033 & 4297\tabularnewline
+70 & 93181 & 6819 & 87914 & 12086 & 9070 & 3803\tabularnewline
+71 & 94330 & 5670 & 89200 & 10800 & 8401 & 3271\tabularnewline
+72 & 95324 & 4676 & 90833 & 9167 & 7359 & 2868\tabularnewline
+73 & 96143 & 3857 & 91805 & 8195 & 6726 & 2388\tabularnewline
+74 & 96959 & 3041 & 93065 & 6935 & 5900 & 2006\tabularnewline
+75 & 97543 & 2457 & 93987 & 6013 & 5222 & 1666\tabularnewline
+76 & 97969 & 2031 & 94900 & 5100 & 4481 & 1412\tabularnewline
+77 & 98497 & 1503 & 95756 & 4244 & 3809 & 1068\tabularnewline
+78 & 98744 & 1256 & 96442 & 3558 & 3269 & 967\tabularnewline
+79 & 99045 & 955 & 96942 & 3058 & 2837 & 734\tabularnewline
+80 & 99276 & 724 & 97528 & 2472 & 2329 & 581\tabularnewline
+81 & 99481 & 519 & 97996 & 2004 & 1894 & 409\tabularnewline
+82 & 99556 & 444 & 98321 & 1679 & 1597 & 362\tabularnewline
+83 & 99691 & 309 & 98722 & 1278 & 1227 & 258\tabularnewline
+84 & 99752 & 248 & 98943 & 1057 & 1015 & 206\tabularnewline
+85 & 99833 & 167 & 99144 & 856 & 837 & 148\tabularnewline
+86 & 99895 & 105 & 99346 & 654 & 642 & 93\tabularnewline
+87 & 99925 & 75 & 99461 & 539 & 530 & 66\tabularnewline
+88 & 99945 & 55 & 99566 & 434 & 428 & 49\tabularnewline
+89 & 99976 & 24 & 99675 & 325 & 324 & 23\tabularnewline
+90 & 99977 & 23 & 99756 & 244 & 243 & 22\tabularnewline
+91 & 99982 & 18 & 99839 & 161 & 155 & 12\tabularnewline
+92 & 99988 & 12 & 99865 & 135 & 135 & 12\tabularnewline
+93 & 99994 & 6 & 99885 & 115 & 115 & 6\tabularnewline
+94 & 99993 & 7 & 99911 & 89 & 88 & 6\tabularnewline
+95 & 99998 & 2 & 99953 & 47 & 47 & 2\tabularnewline
+96 & 99999 & 1 & 99965 & 35 & 35 & 1\tabularnewline
+97 & 99999 & 1 & 99979 & 21 & 21 & 1\tabularnewline
+98 & 100000 & 0 & 99973 & 27 & 27 & 0\tabularnewline
+99 & 100000 & 0 & 99984 & 16 & 16 & 0\tabularnewline
+100 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+Overall, as
+expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
+predator-prey communities are relatively stable while mutualist
+communties are highly unstable. But interestingly, while \(Var(\gamma)\)
+stabilises predator-prey and competitor communities, it does not
+stabilise mutualist communities. This is unsurprising because purely
+mutualist communities are characterised by a very
+positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
+alone will shift all real parts of eigenvalues to negative values.
+
+\hypertarget{connectance}{\section{Sensitivity of connectance (C)
+values}\label{connectance}}
+
+In the main text, for simplicity, I assumed connectance values of
+\(C = 1\), meaning that all off-diagonal elements of a matrix
+\(\mathbf{M}\) were potentially nonzero and sampled from a normal
+distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
+I present four tables showing the number of stable communities given
+\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
+component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
+higher number of stable communities than when \(\gamma\) did not vary
+(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
+million \(\mathbf{M}\) were simulated. As with the results on
+\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
+in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
+and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
+column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
+that are unstable, and the column \texttt{A0\_stable} shows the number
+of \texttt{A0} matrices that are stable (these two columns sum to
+100000). Similarly, the column \texttt{A1\_unstable} shows the number of
+\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
+number that are stable. The columns \texttt{A1\_stabilised} and
+\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
+stabilised or destabilised, respectively, by \(Var(\gamma)\).
+
+\textbf{Connectance \(\mathbf{C = 0.3}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 5 & 99995 & 5 & 99995 & 0 & 0 & 75110\tabularnewline
+3 & 6 & 99994 & 6 & 99994 & 0 & 0 & 87526\tabularnewline
+4 & 24 & 99976 & 24 & 99976 & 0 & 0 & 93713\tabularnewline
+5 & 59 & 99941 & 59 & 99941 & 0 & 0 & 96929\tabularnewline
+6 & 98 & 99902 & 98 & 99902 & 0 & 0 & 98492\tabularnewline
+7 & 160 & 99840 & 161 & 99839 & 0 & 1 & 99243\tabularnewline
+8 & 290 & 99710 & 293 & 99707 & 0 & 3 & 99582\tabularnewline
+9 & 430 & 99570 & 434 & 99566 & 0 & 4 & 99821\tabularnewline
+10 & 648 & 99352 & 653 & 99347 & 1 & 6 & 99895\tabularnewline
+11 & 946 & 99054 & 957 & 99043 & 0 & 11 & 99945\tabularnewline
+12 & 1392 & 98608 & 1415 & 98585 & 4 & 27 & 99978\tabularnewline
+13 & 2032 & 97968 & 2065 & 97935 & 5 & 38 & 99987\tabularnewline
+14 & 2627 & 97373 & 2688 & 97312 & 10 & 71 & 99988\tabularnewline
+15 & 3588 & 96412 & 3647 & 96353 & 35 & 94 & 99996\tabularnewline
+16 & 5019 & 94981 & 5124 & 94876 & 51 & 156 & 99998\tabularnewline
+17 & 6512 & 93488 & 6673 & 93327 & 79 & 240 & 99999\tabularnewline
+18 & 8444 & 91556 & 8600 & 91400 & 165 & 321 & 100000\tabularnewline
+19 & 10416 & 89584 & 10667 & 89333 & 244 & 495 & 100000\tabularnewline
+20 & 13254 & 86746 & 13477 & 86523 & 425 & 648 & 100000\tabularnewline
+21 & 16248 & 83752 & 16481 & 83519 & 642 & 875 & 100000\tabularnewline
+22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151 & 100000\tabularnewline
+23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490 & 100000\tabularnewline
+24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818 & 100000\tabularnewline
+25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137 & 100000\tabularnewline
+26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506 & 100000\tabularnewline
+27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811 & 100000\tabularnewline
+28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072 & 100000\tabularnewline
+29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321 & 100000\tabularnewline
+30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480 & 100000\tabularnewline
+31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568 & 100000\tabularnewline
+32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531 & 100000\tabularnewline
+33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360 & 100000\tabularnewline
+34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014 & 100000\tabularnewline
+35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773 & 100000\tabularnewline
+36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393 & 100000\tabularnewline
+37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913 & 100000\tabularnewline
+38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566 & 100000\tabularnewline
+39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187 & 100000\tabularnewline
+40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936 & 100000\tabularnewline
+41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683 & 100000\tabularnewline
+42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449 & 100000\tabularnewline
+43 & 99275 & 725 & 98312 & 1688 & 1251 & 288 & 100000\tabularnewline
+44 & 99583 & 417 & 98872 & 1128 & 903 & 192 & 100000\tabularnewline
+45 & 99776 & 224 & 99339 & 661 & 576 & 139 & 100000\tabularnewline
+46 & 99865 & 135 & 99518 & 482 & 413 & 66 & 100000\tabularnewline
+47 & 99938 & 62 & 99744 & 256 & 226 & 32 & 100000\tabularnewline
+48 & 99956 & 44 & 99824 & 176 & 151 & 19 & 100000\tabularnewline
+49 & 99980 & 20 & 99914 & 86 & 85 & 19 & 100000\tabularnewline
+50 & 99993 & 7 & 99950 & 50 & 46 & 3 & 100000\tabularnewline
+51 & 99998 & 2 & 99971 & 29 & 28 & 1 & 100000\tabularnewline
+52 & 99998 & 2 & 99986 & 14 & 14 & 2 & 100000\tabularnewline
+53 & 99999 & 1 & 99992 & 8 & 7 & 0 & 100000\tabularnewline
+54 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+55 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+56 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+57 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+58 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+100 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.5}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 74863\tabularnewline
+3 & 32 & 99968 & 32 & 99968 & 0 & 0 & 87434\tabularnewline
+4 & 122 & 99878 & 122 & 99878 & 0 & 0 & 93761\tabularnewline
+5 & 320 & 99680 & 321 & 99679 & 0 & 1 & 96830\tabularnewline
+6 & 667 & 99333 & 673 & 99327 & 0 & 6 & 98481\tabularnewline
+7 & 1233 & 98767 & 1252 & 98748 & 0 & 19 & 99187\tabularnewline
+8 & 2123 & 97877 & 2156 & 97844 & 3 & 36 & 99654\tabularnewline
+9 & 3415 & 96585 & 3471 & 96529 & 16 & 72 & 99816\tabularnewline
+10 & 5349 & 94651 & 5450 & 94550 & 30 & 131 & 99900\tabularnewline
+11 & 7990 & 92010 & 8185 & 91815 & 81 & 276 & 99958\tabularnewline
+12 & 11073 & 88927 & 11301 & 88699 & 219 & 447 & 99973\tabularnewline
+13 & 14971 & 85029 & 15204 & 84796 & 445 & 678 & 99986\tabularnewline
+14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002 & 99991\tabularnewline
+15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404 & 99996\tabularnewline
+16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980 & 99999\tabularnewline
+17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476 & 100000\tabularnewline
+18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937 & 99999\tabularnewline
+19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262 & 99999\tabularnewline
+20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527 & 100000\tabularnewline
+21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673 & 100000\tabularnewline
+22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491 & 100000\tabularnewline
+23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244 & 100000\tabularnewline
+24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876 & 100000\tabularnewline
+25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382 & 100000\tabularnewline
+26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901 & 100000\tabularnewline
+27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358 & 100000\tabularnewline
+28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974 & 100000\tabularnewline
+29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657 & 100000\tabularnewline
+30 & 99103 & 897 & 98033 & 1967 & 1466 & 396 & 100000\tabularnewline
+31 & 99502 & 498 & 98665 & 1335 & 1068 & 231 & 100000\tabularnewline
+32 & 99745 & 255 & 99185 & 815 & 696 & 136 & 100000\tabularnewline
+33 & 99881 & 119 & 99572 & 428 & 375 & 66 & 100000\tabularnewline
+34 & 99955 & 45 & 99788 & 212 & 191 & 24 & 100000\tabularnewline
+35 & 99979 & 21 & 99900 & 100 & 95 & 16 & 100000\tabularnewline
+36 & 99995 & 5 & 99950 & 50 & 50 & 5 & 100000\tabularnewline
+37 & 99997 & 3 & 99970 & 30 & 28 & 1 & 100000\tabularnewline
+38 & 99998 & 2 & 99986 & 14 & 13 & 1 & 100000\tabularnewline
+39 & 99999 & 1 & 99991 & 9 & 9 & 1 & 100000\tabularnewline
+40 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+41 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+42 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+43 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.7}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 75160\tabularnewline
+3 & 106 & 99894 & 106 & 99894 & 0 & 0 & 87447\tabularnewline
+4 & 395 & 99605 & 397 & 99603 & 0 & 2 & 93849\tabularnewline
+5 & 1117 & 98883 & 1123 & 98877 & 0 & 6 & 96827\tabularnewline
+6 & 2346 & 97654 & 2367 & 97633 & 6 & 27 & 98402\tabularnewline
+7 & 4314 & 95686 & 4388 & 95612 & 16 & 90 & 99218\tabularnewline
+8 & 7327 & 92673 & 7456 & 92544 & 61 & 190 & 99611\tabularnewline
+9 & 11514 & 88486 & 11792 & 88208 & 150 & 428 & 99808\tabularnewline
+10 & 16247 & 83753 & 16584 & 83416 & 415 & 752 & 99904\tabularnewline
+11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162 & 99952\tabularnewline
+12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818 & 99977\tabularnewline
+13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355 & 99984\tabularnewline
+14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927 & 99995\tabularnewline
+15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421 & 99994\tabularnewline
+16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527 & 99999\tabularnewline
+17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495 & 100000\tabularnewline
+18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351 & 100000\tabularnewline
+19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979 & 99999\tabularnewline
+20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339 & 100000\tabularnewline
+21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869 & 100000\tabularnewline
+22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267 & 100000\tabularnewline
+23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791 & 100000\tabularnewline
+24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504 & 100000\tabularnewline
+25 & 99433 & 567 & 98630 & 1370 & 1079 & 276 & 100000\tabularnewline
+26 & 99760 & 240 & 99259 & 741 & 655 & 154 & 100000\tabularnewline
+27 & 99895 & 105 & 99576 & 424 & 377 & 58 & 100000\tabularnewline
+28 & 99950 & 50 & 99790 & 210 & 194 & 34 & 100000\tabularnewline
+29 & 99981 & 19 & 99915 & 85 & 80 & 14 & 100000\tabularnewline
+30 & 99994 & 6 & 99952 & 48 & 47 & 5 & 100000\tabularnewline
+31 & 99998 & 2 & 99972 & 28 & 28 & 2 & 100000\tabularnewline
+32 & 99999 & 1 & 99992 & 8 & 8 & 1 & 100000\tabularnewline
+33 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+34 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+35 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.9}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 14 & 99986 & 14 & 99986 & 0 & 0 & 75187\tabularnewline
+3 & 240 & 99760 & 240 & 99760 & 0 & 0 & 87443\tabularnewline
+4 & 1008 & 98992 & 1016 & 98984 & 0 & 8 & 93795\tabularnewline
+5 & 2708 & 97292 & 2729 & 97271 & 2 & 23 & 96814\tabularnewline
+6 & 5669 & 94331 & 5755 & 94245 & 13 & 99 & 98439\tabularnewline
+7 & 9848 & 90152 & 10057 & 89943 & 91 & 300 & 99208\tabularnewline
+8 & 15903 & 84097 & 16201 & 83799 & 336 & 634 & 99603\tabularnewline
+9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168 & 99803\tabularnewline
+10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852 & 99909\tabularnewline
+11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507 & 99951\tabularnewline
+12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127 & 99977\tabularnewline
+13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649 & 99986\tabularnewline
+14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765 & 99995\tabularnewline
+15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550 & 100000\tabularnewline
+16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092 & 100000\tabularnewline
+17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424 & 100000\tabularnewline
+18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746 & 100000\tabularnewline
+19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167 & 99999\tabularnewline
+20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729 & 100000\tabularnewline
+21 & 99111 & 889 & 98133 & 1867 & 1389 & 411 & 100000\tabularnewline
+22 & 99588 & 412 & 98905 & 1095 & 903 & 220 & 100000\tabularnewline
+23 & 99837 & 163 & 99480 & 520 & 452 & 95 & 100000\tabularnewline
+24 & 99932 & 68 & 99744 & 256 & 228 & 40 & 100000\tabularnewline
+25 & 99976 & 24 & 99863 & 137 & 133 & 20 & 100000\tabularnewline
+26 & 99995 & 5 & 99950 & 50 & 49 & 4 & 100000\tabularnewline
+27 & 99996 & 4 & 99986 & 14 & 13 & 3 & 100000\tabularnewline
+28 & 100000 & 0 & 99993 & 7 & 7 & 0 & 100000\tabularnewline
+29 & 100000 & 0 & 99996 & 4 & 4 & 0 & 100000\tabularnewline
+30 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
+strength (\(\sigma\))
+values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
+
+Results below show stability results given varying interaction strengths
+(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
+larger and increase by 10 with increasing rows). In the tables below (as
+\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
+matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+20 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+30 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+40 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+60 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+70 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+80 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
+90 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+100 & 11 & 99989 & 11 & 99989 & 0 & 0\tabularnewline
+110 & 12 & 99988 & 13 & 99987 & 0 & 1\tabularnewline
+120 & 23 & 99977 & 23 & 99977 & 0 & 0\tabularnewline
+130 & 40 & 99960 & 40 & 99960 & 0 & 0\tabularnewline
+140 & 62 & 99938 & 65 & 99935 & 0 & 3\tabularnewline
+150 & 162 & 99838 & 165 & 99835 & 0 & 3\tabularnewline
+160 & 325 & 99675 & 329 & 99671 & 2 & 6\tabularnewline
+170 & 829 & 99171 & 851 & 99149 & 6 & 28\tabularnewline
+180 & 1817 & 98183 & 1860 & 98140 & 31 & 74\tabularnewline
+190 & 3927 & 96073 & 3989 & 96011 & 143 & 205\tabularnewline
+200 & 8084 & 91916 & 8048 & 91952 & 557 & 521\tabularnewline
+210 & 15558 & 84442 & 15147 & 84853 & 1534 & 1123\tabularnewline
+220 & 26848 & 73152 & 25342 & 74658 & 3625 & 2119\tabularnewline
+230 & 43386 & 56614 & 39535 & 60465 & 6992 & 3141\tabularnewline
+240 & 62734 & 37266 & 56684 & 43316 & 9815 & 3765\tabularnewline
+250 & 80128 & 19872 & 73080 & 26920 & 10128 & 3080\tabularnewline
+260 & 92206 & 7794 & 86619 & 13381 & 7490 & 1903\tabularnewline
+270 & 97946 & 2054 & 94824 & 5176 & 3797 & 675\tabularnewline
+280 & 99659 & 341 & 98534 & 1466 & 1265 & 140\tabularnewline
+290 & 99962 & 38 & 99696 & 304 & 281 & 15\tabularnewline
+300 & 99994 & 6 & 99964 & 36 & 34 & 4\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 3 & 99997 & 3 & 99997 & 0 & 0\tabularnewline
+20 & 15 & 99985 & 15 & 99985 & 0 & 0\tabularnewline
+30 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+40 & 85 & 99915 & 85 & 99915 & 0 & 0\tabularnewline
+50 & 163 & 99837 & 163 & 99837 & 0 & 0\tabularnewline
+60 & 280 & 99720 & 282 & 99718 & 0 & 2\tabularnewline
+70 & 561 & 99439 & 566 & 99434 & 3 & 8\tabularnewline
+80 & 1009 & 98991 & 1029 & 98971 & 6 & 26\tabularnewline
+90 & 2126 & 97874 & 2175 & 97825 & 31 & 80\tabularnewline
+100 & 4580 & 95420 & 4653 & 95347 & 142 & 215\tabularnewline
+110 & 9540 & 90460 & 9632 & 90368 & 465 & 557\tabularnewline
+120 & 19090 & 80910 & 18668 & 81332 & 1676 & 1254\tabularnewline
+130 & 35047 & 64953 & 33220 & 66780 & 4172 & 2345\tabularnewline
+140 & 56411 & 43589 & 52439 & 47561 & 7297 & 3325\tabularnewline
+150 & 78003 & 21997 & 72574 & 27426 & 8477 & 3048\tabularnewline
+160 & 92678 & 7322 & 88438 & 11562 & 5901 & 1661\tabularnewline
+170 & 98614 & 1386 & 96670 & 3330 & 2397 & 453\tabularnewline
+180 & 99839 & 161 & 99418 & 582 & 499 & 78\tabularnewline
+190 & 99990 & 10 & 99945 & 55 & 52 & 7\tabularnewline
+200 & 100000 & 0 & 99995 & 5 & 5 & 0\tabularnewline
+210 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 36 & 99964 & 36 & 99964 & 0 & 0\tabularnewline
+20 & 195 & 99805 & 195 & 99805 & 0 & 0\tabularnewline
+30 & 519 & 99481 & 523 & 99477 & 0 & 4\tabularnewline
+40 & 1096 & 98904 & 1101 & 98899 & 2 & 7\tabularnewline
+50 & 2375 & 97625 & 2397 & 97603 & 9 & 31\tabularnewline
+60 & 4898 & 95102 & 4968 & 95032 & 83 & 153\tabularnewline
+70 & 10841 & 89159 & 10916 & 89084 & 432 & 507\tabularnewline
+80 & 22281 & 77719 & 21988 & 78012 & 1622 & 1329\tabularnewline
+90 & 42010 & 57990 & 39998 & 60002 & 4458 & 2446\tabularnewline
+100 & 67289 & 32711 & 63098 & 36902 & 7153 & 2962\tabularnewline
+110 & 88137 & 11863 & 84023 & 15977 & 6108 & 1994\tabularnewline
+120 & 97678 & 2322 & 95557 & 4443 & 2740 & 619\tabularnewline
+130 & 99795 & 205 & 99304 & 696 & 578 & 87\tabularnewline
+140 & 99989 & 11 & 99948 & 52 & 49 & 8\tabularnewline
+150 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+10 & 162 & 99838 & 162 & 99838 & 0 & 0\tabularnewline
+20 & 798 & 99202 & 799 & 99201 & 0 & 1\tabularnewline
+30 & 2273 & 97727 & 2289 & 97711 & 6 & 22\tabularnewline
+40 & 5259 & 94741 & 5298 & 94702 & 70 & 109\tabularnewline
+50 & 12084 & 87916 & 12054 & 87946 & 446 & 416\tabularnewline
+60 & 26072 & 73928 & 25511 & 74489 & 1810 & 1249\tabularnewline
+70 & 50121 & 49879 & 47747 & 52253 & 4748 & 2374\tabularnewline
+80 & 77806 & 22194 & 73810 & 26190 & 6421 & 2425\tabularnewline
+90 & 94862 & 5138 & 92069 & 7931 & 3842 & 1049\tabularnewline
+100 & 99527 & 473 & 98822 & 1178 & 870 & 165\tabularnewline
+110 & 99984 & 16 & 99912 & 88 & 80 & 8\tabularnewline
+120 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+130 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
+distribution of
+\(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
+
+In the main text, I considered a uniform distribution of component
+response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
+and stable \(\mathbf{M}\) matrices are reported in
+\protect\hyperlink{IncrS}{a table above} across different values of
+\(S\). Here I show complementary results for three different
+distributions including an exponential, beta, and gamma distribution of
+\(\gamma\) values. The shape of these distributions is shown in the
+figure below.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Distributions of component response rate
+(\(\boldsymbol{\gamma}\)) values in complex systems.} The stabilities of
+simulated complex systems with these \(\gamma\) distributions are
+compared to otherwise identical complex systems with a fixed component
+response rate of \(\gamma = 1\) across different system sizes (\(S\);
+i.e., component numbers) given a unit \(\gamma\) standard deviation
+(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
+uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
+\(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
+panel shows 1 million randomly generated \(\gamma\) values.
+
+\includegraphics{unnamed-chunk-23-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+The same 100000 \(\mathbf{M}\) matrices were used to investigate
+stability when applying each of these different distributions of
+\(\gamma\) values. The table below shows the number of \(\mathbf{M}\)
+that were unstable (\_unst) and stable (\_stbl) for the exponential
+(Exp), beta, and gamma distributions.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
+\KeywordTok{kable}\NormalTok{(fourdists);}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lllllll@{}}
+\toprule
+S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
+gamma\_stbl\tabularnewline
+\midrule
+\endhead
+2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
+3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
+4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
+5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
+6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
+7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
+8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
+9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
+10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
+11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
+12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
+13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
+14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
+15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
+16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
+17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
+18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
+19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
+20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
+21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
+22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
+23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
+24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
+25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
+26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
+27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
+28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
+29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
+30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
+32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+In comparison to the uniform distribution (a), proportionally fewer
+random systems are found with the exponential distribution (b), while
+more are found with the beta (c) and gamma (d) distributions.
+
+\hypertarget{Feasibility}{\section{Feasibility of complex
+systems}\label{Feasibility}}
+
+When feasibility was evaluated with and without variation in \(\gamma\),
+there was no increase in stability for \(\mathbf{M}\) where \(\gamma\)
+varied as compared to where \(\gamma = 1\). Results below illustrate
+this result, which was general to all other simulations performed.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
+A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 749978 & 250022 & 749942 & 250058 & 35552 & 35516\tabularnewline
+3 & 874519 & 125481 & 874296 & 125704 & 36803 & 36580\tabularnewline
+4 & 937192 & 62808 & 937215 & 62785 & 26440 & 26463\tabularnewline
+5 & 968776 & 31224 & 968639 & 31361 & 16319 & 16182\tabularnewline
+6 & 984313 & 15687 & 984463 & 15537 & 9006 & 9156\tabularnewline
+7 & 992149 & 7851 & 992161 & 7839 & 4991 & 5003\tabularnewline
+8 & 996124 & 3876 & 996103 & 3897 & 2644 & 2623\tabularnewline
+9 & 998014 & 1986 & 998027 & 1973 & 1361 & 1374\tabularnewline
+10 & 999031 & 969 & 999040 & 960 & 698 & 707\tabularnewline
+11 & 999546 & 454 & 999514 & 486 & 377 & 345\tabularnewline
+12 & 999764 & 236 & 999792 & 208 & 160 & 188\tabularnewline
+13 & 999883 & 117 & 999865 & 135 & 105 & 87\tabularnewline
+14 & 999938 & 62 & 999945 & 55 & 40 & 47\tabularnewline
+15 & 999971 & 29 & 999964 & 36 & 31 & 24\tabularnewline
+16 & 999988 & 12 & 999991 & 9 & 8 & 11\tabularnewline
+17 & 999996 & 4 & 999991 & 9 & 8 & 3\tabularnewline
+18 & 999997 & 3 & 999999 & 1 & 1 & 3\tabularnewline
+19 & 999998 & 2 & 999997 & 3 & 3 & 2\tabularnewline
+20 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
+21 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+22 & 999999 & 1 & 1000000 & 0 & 0 & 1\tabularnewline
+23 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+24 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+25 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+26 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+27 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+28 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+29 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+30 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+31 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
+in pure species interaction networks.
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\hypertarget{ref-May1972}{}
+1. May, R. M. Will a large complex system be stable? \emph{Nature}
+\textbf{238,} 413--414 (1972).
+
+\hypertarget{ref-Allesina2012}{}
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+\emph{Nature} \textbf{483,} 205--208 (2012).
+
+\hypertarget{ref-Allesina2015}{}
+3. Allesina, S. \emph{et al.} Predicting the stability of large
+structured food webs. \emph{Nature Communications} \textbf{6,} 7842
+(2015).
+
 \end{document}

### 2018-08-02 notebook/ms.Rmd
Reformat for PNAS

@@
 - \usepackage{natbib}
+- \usepackage{lineno}
 - \usepackage[utf8]{inputenc}
+- \linenumbers
 - \bibliographystyle{amnatnat}
-linestretch: 1
-link-citations: yes
+linestretch: 2
+link-citations: no
 linkcolor: blue
-csl: nature.csl
+csl: science.csl
 biblio-style: apalike
@@
 As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues[@Ahmadian2015; @Gibbs2017; @Stone2017], some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
+Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
 The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
@@
 To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm.
-Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively[@Hamblin2013].
+Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively [@Hamblin2013].
 Generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
@@
 **System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
-Feasibilility assumes that the system is stable and values of all components are positive at equilibrium[@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic[@Dougoud2018].
-While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities[@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
+Feasibilility assumes that the system is stable and values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
+This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
+While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities [@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
 Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response rate (i.e., species generation time) does not affect stability at biologically realistic species densities.
 
-********************************************************************************
+Discussion
+--------------------------------------------------------------------------------
+
+Here I have shown that the stability of large systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
+Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's [@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
+The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
+
+It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
+Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
+3).
+This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
+1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
+The mathematics underlying this shift in eigenvalue distribution has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual system components to perturbation.
+
+The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
+The probability that some combination of component response rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
+Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
+Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+
+Interestingly, while complex systems were more likely to be stable given variation in component response rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
+Feasibility is important to consider, particularly for the study of ecological networks of species [@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative.
+My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in empirical systems [@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability should therefore not be ignored in the broader context of ecological communities.
+
+My results show that complex systems are more likely to be stable when the response rates of system components vary.
+These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
+
+Methods
+--------------------------------------------------------------------------------
+
+<!--- USE J INSTEAD TO CLARIFY M --->
+
+**Component response rate variation ($\mathbf{\gamma}$)**.
+In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
+
+$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+
+$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+
+In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
+Patel et al.
+were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities [@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
+The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution [@Patel2018].
+
+I use the same principle that Patel et al.
+use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
+Following May [@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($a_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
+
+$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).$$
+
+In matrix notation[@May1973],
+
+$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).$$
+
+In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
+Therefore, $\mathbf{M} =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+
+**Genetic algorithm**.
+Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution [@Hamblin2013].
+In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
+The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
+
+For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), where $\mathbf{M = \gamma A}$.
+The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
+Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
+Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
+
+**System feasibility**.
+Dougoud et al.
+[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
+
+$$\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.$$
+
+In the above, $\mathbf{x^{*}}$ is the vector of species densities at equilibrium.
+Feasibility is satisfied if all elements in $\mathbf{x^{*}}$ are positive.
+The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is the strength of intraspecific competition (diagonal elements).
+Diagonal values are set to $-1$, so $\theta = -1$.
+The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$) for $\mathbf{J}$.
+Implicitly, here $\delta = 0$ underlying strong interactions.
+Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion [@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
+Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below,
+
+$$\mathbf{x^{*} = -A^{-1}r}.$$
+
+To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were sampled i.i.d.
+from $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
+Feasibility is satisfied if all of the elements of the resulting vector are positive.
+
+<!---
+
+If more discussion is needed, then talk about relation to Jirsa and Ding 2004.
+
+--->
+
+**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
+Conversations with L.
+Bussi&#xe8;re and N.
+Bunnefeld, and comments from J.
+J.
+Cusack and I.
+L.
+Jones, improved the quality of this work.
+
+**Supplementary Information:** Full tables of stability results for simulations across different system size ($S$) values, ecological community types, connectance ($C$) values, interaction strengths ($\sigma$), and $\gamma$ distributions are provided as supplementary material.
+An additional table also shows results for how feasibility changes across $S$.
+All code and simulation outputs are publicly available as part of the RandomMatrixStability package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
+
+**References**
+
+<div id="refs"></div>
+
+\clearpage
 
@@
 
-********************************************************************************
-
-Discussion
---------------------------------------------------------------------------------
-
-Here I have shown that the stability of large systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
-Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's[@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
-The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
-
-********************************************************************************
+\clearpage
 
@@
 
-********************************************************************************
-
-It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
-Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
-3).
-This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
-The mathematics underlying this shift in eigenvalue distribution has been investigated[@Ahmadian2015] and recently applied to questions concerning species density and feasibility[@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual system components to perturbation.
-
-The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
-The probability that some combination of component response rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
-Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
-Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
-
-********************************************************************************
+\clearpage
 
@@
 
-********************************************************************************
-
-Interestingly, while complex systems were more likely to be stable given variation in component response rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
-Feasibility is important to consider, particularly for the study of ecological networks of species[@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative.
-My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information).
-Nevertheless, ecological interactions do not exist in isolation in empirical systems[@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems.
-The relevance of component response rate for complex system stability should therefore not be ignored in the broader context of ecological communities.
-
-********************************************************************************
+\clearpage
 
@@
 ```
-
-********************************************************************************
-
-My results show that complex systems are more likely to be stable when the response rates of system components vary.
-These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
-
-Methods
---------------------------------------------------------------------------------
-
-<!--- USE J INSTEAD TO CLARIFY M --->
-
-**Component response rate variation ($\mathbf{\gamma}$)**.
-In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
-model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
-For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
-
-$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
-
-$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
-
-In the above, $f_{i}$ and $g_{j}$ are functions that define the effects of all species densities and trait values on the density of a species $i$ and the value of trait $j$, respectively.
-Patel et al.
-were interested in stability when the evolution of traits was relatively slow or fast in comparison with the change in species densities[@Patel2018], and this is modulated in the above by the scalar $\epsilon$.
-The value of $\epsilon$ thereby determines the timescale separation between ecology and evolution, with high $\epsilon$ modelling relatively fast evolution and low $\epsilon$ modelling relative slow evolution[@Patel2018].
-
-I use the same principle that Patel et al.
-use to modulate the relative rate of evolution to modulate rates of component responses for $S$ components.
-Following May[@May1972; @May1973], the value of a component $i$ at time $t$ ($v_{i}(t)$) is affected by the value of $j$ ($v_{j}(t)$) and $j$'s marginal effect on $i$ ($a_{ij}$), and by $i$'s response rate ($\gamma_{i}$),
-
-$$\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).$$
-
-In matrix notation[@May1973],
-
-$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).$$
-
-In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{M} =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May[@May1972; @May1973; @Ahmadian2015].
-
-**Genetic algorithm**.
-Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
-This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
-Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution[@Hamblin2013].
-In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
-The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
-
-For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), where $\mathbf{M = \gamma A}$.
-The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
-Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
-Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
-Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
-The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
-
-**System feasibility**.
-Dougoud et al.[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
-
-$$\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.$$
-
-In the above, $\mathbf{x^{*}}$ is the vector of species densities at equilibrium.
-Feasibility is satisfied if all elements in $\mathbf{x^{*}}$ are positive.
-The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is the strength of intraspecific competition (diagonal elements).
-Diagonal values are set to $-1$, so $\theta = -1$.
-The variable $\delta$ is a normalisation parameter that modulates the strength of interactions ($\sigma$) for $\mathbf{J}$.
-Implicitly, here $\delta = 0$ underlying strong interactions.
-Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion[@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
-Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below,
-
-$$\mathbf{x^{*} = -A^{-1}r}.$$
-
-To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were sampled i.i.d.
-from $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
-Feasibility is satisfied if all of the elements of the resulting vector are positive.
-
-<!---
-
-If more discussion is needed, then talk about relation to Jirsa and Ding 2004.
-
---->
-
-**Acknowledgements:** I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L.
-Bussi&#xe8;re and N.
-Bunnefeld, and comments from J.
-J.
-Cusack and I.
-L.
-Jones, improved the quality of this work.
-
-**Supplementary Information:** Full tables of stability results for simulations across different system size ($S$) values, ecological community types, connectance ($C$) values, interaction strengths ($\sigma$), and $\gamma$ distributions are provided as supplementary material.
-An additional table also shows results for how feasibility changes across $S$.
-All code and simulation outputs are publicly available as part of the RandomMatrixStability package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
-
-**References**
-
-<div id="refs"></div>

### 2018-08-02 notebook/ms.tex
Reformat for PNAS

@@
 \usepackage{setspace}
-\setstretch{1}
+\setstretch{2}
 \usepackage{amssymb,amsmath}
@@
             pdftitle={Component response rate variation drives stability in large complex systems},
-            pdfauthor={A. Bradley Duthie},
+            pdfauthor={A. Bradley Duthie ( alexander.duthie@stir.ac.uk )},
             colorlinks=true,
@@
 \urlstyle{same}  % don't use monospace font for urls
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
-\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 \usepackage{natbib}
+\usepackage{lineno}
 \usepackage[utf8]{inputenc}
+\linenumbers
 \bibliographystyle{amnatnat}
@@
 
-In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
-demonstrated that randomly assembled systems of sufficient complexity
-are almost inevitably unstable given infinitesimally small
-perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of potentially interacting components; \(S\)),
-its connectance (i.e., the probability that one component will interact
-with another; \(C\)), and the variance of interaction strengths
-(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-May's finding that the probability of local stability falls to near zero
-given a sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly
-relevant for understanding the dynamics and persistence of systems such
-as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
-and
-socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
-networks. As such, identifying general principles that drive stability
-in complex systems is of wide-ranging importance.
+In 1972, May (\emph{1}) first demonstrated that randomly assembled
+systems of sufficient complexity are almost inevitably unstable given
+infinitesimally small perturbations. Complexity in this case is defined
+by the size of the system (i.e., the number of potentially interacting
+components; \(S\)), its connectance (i.e., the probability that one
+component will interact with another; \(C\)), and the variance of
+interaction strengths (\(\sigma^{2}\)) (\emph{2}). May's finding that
+the probability of local stability falls to near zero given a
+sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly relevant
+for understanding the dynamics and persistence of systems such as
+ecological (\emph{1}--\emph{5}), neurological (\emph{6}, \emph{7}),
+biochemical (\emph{8}, \emph{9}), and socio-economic
+(\emph{10}--\emph{13}) networks. As such, identifying general principles
+that drive stability in complex systems is of wide-ranging importance.
 
@@
 matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
-element of such a matrix, \(M_{ij}\), defines how component \(j\)
-affects component \(i\) in the system at a point of
-equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-Off-diagonal elements (\(i \neq j\)) therefore define interactions
-between components, while diagonal elements (\(i = j\)) define component
-self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a
-probability \(C\), which are sampled from a distribution with variance
-\(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
-Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of
-\(\mathbf{M}\) are negative
-(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
-circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
-\(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
-(Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma^{2}\) increase.
-
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
-components respond to perturbations of the system at the same rate
-(\(\gamma\)), but this is highly unlikely in any complex system. In
-ecological communities, for example, the rate at which population
-density changes following perturbation will depend on the generation
-time of organisms, which might vary by orders of magnitude among
-species. Species with short generation times will respond quickly (high
-\(\gamma\)) to perturbations relative to species with long generation
-times (low \(\gamma\)). Similarly, the speed at which individual banks
-respond to perturbations in financial networks, or individuals or
-institutions respond to perturbations in complex social networks, is
-likely to vary. The effect of such variance on stability has not been
-investigated in complex systems theory. Intuitively, variation in
-\(\gamma\) might be expected to decrease system stability by introducing
-a new source of variation into the system and thereby increasing
-\(\sigma\). Here I show why, despite higher \(\sigma\), complex systems
-in which \(\gamma\) varies are actually more likely to be stable,
-especially when \(S\) is high.
+species (\emph{2}) or banks (\emph{11})). One element of such a matrix,
+\(M_{ij}\), defines how component \(j\) affects component \(i\) in the
+system at a point of equilibrium (\emph{2}). Off-diagonal elements
+(\(i \neq j\)) therefore define interactions between components, while
+diagonal elements (\(i = j\)) define component self-regulation (e.g.,
+carrying capacity in ecological communities). Traditionally,
+off-diagonal elements are assigned non-zero values with a probability
+\(C\), which are sampled from a distribution with variance
+\(\sigma^{2}\); diagonal elements are set to -1 (\emph{1}, \emph{2},
+\emph{4}). Local system stability is assessed using eigenanalysis, with
+the system being stable if the real parts of all eigenvalues
+(\(\lambda\)) of \(\mathbf{M}\) are negative
+(\(\max\left(\Re(\lambda)\right) < 0\)) (\emph{1}, \emph{2}). In a large
+system (high \(S\)), eigenvalues are distributed uniformly (\emph{14})
+within a circle centred at \(\Re = -1\) (the mean value of diagonal
+elements) and \(\Im = 0\), with a radius of \(\sigma\sqrt{SC}\)
+(\emph{1}, \emph{2}, \emph{4}) (Figs 1a and 2a). Local stability of
+randomly assembled systems therefore becomes increasingly unlikely as
+\(S\), \(C\), and \(\sigma^{2}\) increase.
+
+May's (\emph{1}, \emph{2}) stability criterion \(\sigma\sqrt{SC} < 1\)
+assumes that individual components respond to perturbations of the
+system at the same rate (\(\gamma\)), but this is highly unlikely in any
+complex system. In ecological communities, for example, the rate at
+which population density changes following perturbation will depend on
+the generation time of organisms, which might vary by orders of
+magnitude among species. Species with short generation times will
+respond quickly (high \(\gamma\)) to perturbations relative to species
+with long generation times (low \(\gamma\)). Similarly, the speed at
+which individual banks respond to perturbations in financial networks,
+or individuals or institutions respond to perturbations in complex
+social networks, is likely to vary. The effect of such variance on
+stability has not been investigated in complex systems theory.
+Intuitively, variation in \(\gamma\) might be expected to decrease
+system stability by introducing a new source of variation into the
+system and thereby increasing \(\sigma\). Here I show why, despite
+higher \(\sigma\), complex systems in which \(\gamma\) varies are
+actually more likely to be stable, especially when \(S\) is high.
 
@@
 time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
-\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
-half of \(\gamma_{i} = 0.05\). This models one system in which
-\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
-are otherwise identical (note \(E[\gamma_{i}] = 1\) in both cases). I
-assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal
-elements are set to \(-1\) and non-zero off-diagonal elements are drawn
-randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied
-by \(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
+value \(\gamma_{i}\) (\emph{15}). The distribution of \(\gamma\) over
+\(S\) components thereby models the distribution of component response
+rates. An instructive example compares one \(\mathbf{M}\) where
+\(\gamma_{i} = 1\) for all \(i\) in \(S\) to the same \(\mathbf{M}\)
+when half of \(\gamma_{i} = 1.95\) and half of \(\gamma_{i} = 0.05\).
+This models one system in which \(\gamma\) is invariant and one in which
+\(\gamma\) varies, but systems are otherwise identical (note
+\(E[\gamma_{i}] = 1\) in both cases). I assume \(S = 200\),
+\(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are set to \(-1\)
+and non-zero off-diagonal elements are drawn randomly from
+\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
+\(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
 eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
@@
 distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
-some of which are clustered locally near the centre of the distribution,
-but others of which are spread outside the former radius of 2.828 (Fig
-2b). The clustering and spreading of eigenvalues introduced by
+expected (\emph{14}), when \(\gamma = 1\), eigenvalues are distributed
+uniformly in a circle centred at (\(-1, 0\)) with a radius of
+\(\sigma\sqrt{SC} =\) 2.828. Uniform variation in \(\gamma\) leads to a
+non-uniform distribution of eigenvalues (\emph{16}--\emph{18}), some of
+which are clustered locally near the centre of the distribution, but
+others of which are spread outside the former radius of 2.828 (Fig 2b).
+The clustering and spreading of eigenvalues introduced by
 \(Var(\gamma)\) can destabilise previously stable systems or stabilise
@@
 complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold \(\sigma\sqrt{SC} < 1\).
+can often lead to stability above May's (\emph{1}, \emph{2}) threshold
+\(\sigma\sqrt{SC} < 1\).
 
@@
 values leading to stability in a large complex system) is too large to
-search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
-Generations of selection on \(\gamma\) value combinations to minimise
-\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
-\(Var(\gamma)\) to increase system stability. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were up to four
-orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
-that stability could often be achieved by manipulating \(S\) \(\gamma\)
-values rather than \(S \times S\) \(\mathbf{M}\) elements.
+search exhaustively (\emph{19}). Generations of selection on \(\gamma\)
+value combinations to minimise \(\max\left(\Re(\lambda)\right)\)
+demonstrated the potential for \(Var(\gamma)\) to increase system
+stability. Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\)
+values were found that resulted in stable systems with probabilities
+that were up to four orders of magnitude higher than when \(\gamma = 1\)
+(Fig. 4), meaning that stability could often be achieved by manipulating
+\(S\) \(\gamma\) values rather than \(S \times S\) \(\mathbf{M}\)
+elements.
 
@@
 the system. Feasibilility assumes that the system is stable and values
-of all components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
-This is of particular interest for ecological communities because
-population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
-While my results are intended to be general to all complex systems, and
-not restricted to species networks, I have also performed a feasibility
-analysis on all matrices \(\mathbf{M}\) tested for stability, and
-additionally for specific types of ecological
-communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-(e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). Feasibility was unaffected by \(Var(\gamma)\), meaning
-that for pure interacting species networks, variation in component
-response rate (i.e., species generation time) does not affect stability
-at biologically realistic species densities.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the blue circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle centred at \(\gamma = -1.95\) and one small circle centred
-at \(\gamma = -0.05\). In a, the maximum real eigenvalue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 was stable when
-\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
-
-\includegraphics{unnamed-chunk-10-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+of all components are positive at equilibrium (\emph{5}, \emph{20},
+\emph{21}). This is of particular interest for ecological communities
+because population density cannot take negative values, meaning that
+ecological systems need to be feasible for stability to be biologically
+realistic (\emph{20}). While my results are intended to be general to
+all complex systems, and not restricted to species networks, I have also
+performed a feasibility analysis on all matrices \(\mathbf{M}\) tested
+for stability, and additionally for specific types of ecological
+communities (\emph{2}) (e.g., competitive, mutualist, predator-prey; see
+Supplementary Information). Feasibility was unaffected by
+\(Var(\gamma)\), meaning that for pure interacting species networks,
+variation in component response rate (i.e., species generation time)
+does not affect stability at biologically realistic species densities.
 
@@
 size increases, and can ultimately increase the overall probability that
-system stability is observed above that predicted by
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
-derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
-general, and potentially applies to any complex system in which
+system stability is observed above that predicted by May's (\emph{1})
+classically derived \(\sigma \sqrt{SC}\) criterion. The logic outlined
+here is general, and potentially applies to any complex system in which
 individual system components can vary in their reaction rates to system
 perturbation.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
-\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
-eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
-both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.719 and 0.735 for a and b,
-respectively.
-
-\includegraphics{unnamed-chunk-13-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
@@
 also increase the real components of eigenvalues. The mathematics
-underlying this shift in eigenvalue distribution has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
-and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
-but has not been interpreted as rates of response of individual system
-components to perturbation.
+underlying this shift in eigenvalue distribution has been investigated
+(\emph{16}) and recently applied to questions concerning species density
+and feasibility (\emph{17}, \emph{18}), but has not been interpreted as
+rates of response of individual system components to perturbation.
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate (\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
-which variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Blue and red bars show the number of stable systems in the absence and
-presence of \(Var(\gamma)\), respectively. The black line shows the
-proportion of systems that are stable when \(Var(\gamma) > 0\), but
-would be unstable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-15-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
 Interestingly, while complex systems were more likely to be stable given
@@
 important to consider, particularly for the study of ecological networks
-of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20}}
-because population densities cannot realistically be negative. My
-results therefore suggest that variation in the rate of population
-responses to perturbation (e.g., due to differences in generation time
-among species) is unlikely to be critical to the stability of purely
-multi-species interaction networks (see also Supplementary Information).
-Nevertheless, ecological interactions do not exist in isolation in
-empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
-instead interact with evolutionary, abiotic, or social-economic systems.
-The relevance of component response rate for complex system stability
-should therefore not be ignored in the broader context of ecological
-communities.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 4: Stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted
-\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
-stable across different system sizes (\(S\), max \(S=40\)) for
-\(C = 1\), and the proportion of systems wherein a targeted search of
-\(\gamma\) values successfully resulted in system stability. For each
-\(S\), 100000 complex systems are randomly generated. Stability of each
-complex system is tested given variation in \(\gamma\) using a genetic
-algorithm to maximise the effect of \(\gamma\) values on increasing
-stability, as compared to stability in an otherwise identical system in
-which \(\gamma\) is the same for all components. Blue bars show the
-number of stable systems in the absence of component response rate
-variation, while red bars show the number of stable systems that can be
-generated if component response rate is varied to maximise system
-stability. The black line shows the proportion of systems that are
-stable when component response rate is targeted to increase stability,
-but would not be stable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-16-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+of species (\emph{5}, \emph{18}, \emph{20}) because population densities
+cannot realistically be negative. My results therefore suggest that
+variation in the rate of population responses to perturbation (e.g., due
+to differences in generation time among species) is unlikely to be
+critical to the stability of purely multi-species interaction networks
+(see also Supplementary Information). Nevertheless, ecological
+interactions do not exist in isolation in empirical systems (\emph{15}),
+but instead interact with evolutionary, abiotic, or social-economic
+systems. The relevance of component response rate for complex system
+stability should therefore not be ignored in the broader context of
+ecological communities.
 
@@
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
-traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+traits (\(\mathbf{x}\))(\emph{15}). For any species \(i\) or trait
+\(j\), change in species density (\(N_{i}\)) or trait value (\(x_{j}\))
+with time (\(t\)) is a function of the vectors \(\mathbf{N}\) and
+\(\mathbf{x}\),
 
@@
 were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
-this is modulated in the above by the scalar \(\epsilon\). The value of
-\(\epsilon\) thereby determines the timescale separation between ecology
-and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+slow or fast in comparison with the change in species densities
+(\emph{15}), and this is modulated in the above by the scalar
+\(\epsilon\). The value of \(\epsilon\) thereby determines the timescale
+separation between ecology and evolution, with high \(\epsilon\)
+modelling relatively fast evolution and low \(\epsilon\) modelling
+relative slow evolution (\emph{15}).
 
@@
 rate of evolution to modulate rates of component responses for \(S\)
-components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{22}},
-the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
-by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
-\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
+components. Following May (\emph{1}, \emph{22}), the value of a
+component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected by the value of
+\(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on \(i\) (\(a_{ij}\)),
+and by \(i\)'s response rate (\(\gamma_{i}\)),
 
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{22}},
+In matrix notation(\emph{22}),
 
@@
 \(\mathbf{M} =\mathbf{\gamma} \mathbf{A}\) modulates the values of
-components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{22}}.
+components and can be analysed using the techniques of May (\emph{1},
+\emph{16}, \emph{22}).
 
@@
 genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
-this case, the practical solution is finding vectors of
+solutions by mimicking the process of biological evolution (\emph{19}).
+In this case, the practical solution is finding vectors of
 \(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
@@
 
-\textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} define the
+\textbf{System feasibility}. Dougoud et al. (\emph{20}) define the
 following feasibility criteria for ecological systems characterised by
@@
 affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
-relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
-stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
+relates to May's(\emph{1}) stability criterion (\emph{20}) by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 \hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
+1. R. M. May, Will a large complex system be stable? \emph{Nature}.
+\textbf{238}, 413--414 (1972).
 
 \hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
-\emph{Nature} \textbf{483,} 205--208 (2012).
+2. S. Allesina, S. Tang, Stability criteria for complex ecosystems.
+\emph{Nature}. \textbf{483}, 205--208 (2012).
 
 \hypertarget{ref-Mougi2012}{}
-3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
-community stability. \emph{Science} \textbf{337,} 349--351 (2012).
+3. A. Mougi, M. Kondoh, Diversity of interaction types and ecological
+community stability. \emph{Science}. \textbf{337}, 349--351 (2012).
 
 \hypertarget{ref-Allesina2015}{}
-4. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
+4. S. Allesina \emph{et al.}, Predicting the stability of large
+structured food webs. \emph{Nature Communications}. \textbf{6}, 7842
 (2015).
@@
 \hypertarget{ref-Grilli2017}{}
-5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
+5. J. Grilli \emph{et al.}, Feasibility and coexistence of large
+ecological communities. \emph{Nature Communications}. \textbf{8} (2017),
+doi:\href{https://doi.org/10.1038/ncomms14389}{10.1038/ncomms14389}.
 
 \hypertarget{ref-Gray2008}{}
-6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
-random brain networks with a distribution of connection strengths.
-\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).
+6. R. T. Gray, P. A. Robinson, Stability and synchronization of random
+brain networks with a distribution of connection strengths.
+\emph{Neurocomputing}. \textbf{71}, 1373--1387 (2008).
 
 \hypertarget{ref-Gray2009}{}
-7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
-with excitatory and inhibitory connections. \emph{Neurocomputing}
-\textbf{72,} 1849--1858 (2009).
+7. R. T. Gray, P. A. Robinson, Stability of random brain networks with
+excitatory and inhibitory connections. \emph{Neurocomputing}.
+\textbf{72}, 1849--1858 (2009).
 
 \hypertarget{ref-Rosenfeld2009}{}
-8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
+8. S. Rosenfeld, Patterns of stochastic behavior in dynamically unstable
 high-dimensional biochemical networks. \emph{Gene Regulation and Systems
-Biology} \textbf{3,} 1--10 (2009).
+Biology}. \textbf{3}, 1--10 (2009).
 
 \hypertarget{ref-MacArthur2010}{}
-9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
-and criticality of adaptive regulatory networks. \emph{Physics Review
-Letters} \textbf{104,} 168701 (2010).
+9. B. D. MacArthur, R. J. Sanchez-Garcia, A. Ma'ayan, Microdynamics and
+criticality of adaptive regulatory networks. \emph{Physics Review
+Letters}. \textbf{104}, 168701 (2010).
 
 \hypertarget{ref-May2008}{}
-10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
-for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
+10. R. M. May, S. A. Levin, G. Sugihara, Complex systems: Ecology for
+bankers. \emph{Nature}. \textbf{451}, 893--895 (2008).
 
 \hypertarget{ref-Haldane2011}{}
-11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
-\emph{Nature} \textbf{469,} 351--355 (2011).
+11. A. G. Haldane, R. M. May, Systemic risk in banking ecosystems.
+\emph{Nature}. \textbf{469}, 351--355 (2011).
 
 \hypertarget{ref-Suweis2014}{}
-12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
-networks. \emph{PLoS ONE} \textbf{9,} (2014).
+12. S. Suweis, P. D'Odorico, Early warning signs in social-ecological
+networks. \emph{PLoS ONE}. \textbf{9} (2014),
+doi:\href{https://doi.org/10.1371/journal.pone.0101851}{10.1371/journal.pone.0101851}.
 
 \hypertarget{ref-Bardoscia2017}{}
-13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
-Pathways towards instability in financial networks. \emph{Nature
-Communications} \textbf{8,} 1--7 (2017).
+13. M. Bardoscia, S. Battiston, F. Caccioli, G. Caldarelli, Pathways
+towards instability in financial networks. \emph{Nature Communications}.
+\textbf{8}, 1--7 (2017).
 
 \hypertarget{ref-Tao2010}{}
-14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
+14. T. Tao, V. Vu, Random matrices: Universality of ESDs and the
+circular law. \emph{Annals of Probability}. \textbf{38}, 2023--2065
 (2010).
@@
 \hypertarget{ref-Patel2018}{}
-15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
-effects of eco-evolutionary feedbacks on community stability.
-\emph{American Naturalist} \textbf{191,} 1--29 (2018).
+15. S. Patel, M. H. Cortez, S. J. Schreiber, Partitioning the effects of
+eco-evolutionary feedbacks on community stability. \emph{American
+Naturalist}. \textbf{191}, 1--29 (2018).
 
 \hypertarget{ref-Ahmadian2015}{}
-16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
-with partially structured and partially random connectivity.
-\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
-Physics} \textbf{91,} 012820 (2015).
+16. Y. Ahmadian, F. Fumarola, K. D. Miller, Properties of networks with
+partially structured and partially random connectivity. \emph{Physical
+Review E - Statistical, Nonlinear, and Soft Matter Physics}.
+\textbf{91}, 012820 (2015).
 
 \hypertarget{ref-Gibbs2017}{}
-17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+17. T. Gibbs, J. Grilli, T. Rogers, S. Allesina, The effect of
 population abundances on the stability of large random ecosystems.
-\emph{arXiv} (2017).
+\emph{arXiv} (2017) (available at
+\url{http://arxiv.org/abs/1708.08837}).
 
 \hypertarget{ref-Stone2017}{}
-18. Stone, L. The feasibility and stability of large complex biological
-networks: a random matrix approach. \emph{Scientific Reports}
-\textbf{8,} 8246 (2018).
+18. L. Stone, The feasibility and stability of large complex biological
+networks: a random matrix approach. \emph{Scientific Reports}.
+\textbf{8}, 8246 (2018).
 
 \hypertarget{ref-Hamblin2013}{}
-19. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+19. S. Hamblin, On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution}. \textbf{4},
 184--194 (2013).
@@
 \hypertarget{ref-Dougoud2018}{}
-20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
+20. M. Dougoud, L. Vinckenbosch, R. Rohr, L.-F. Bersier, C. Mazza, The
+feasibility of equilibria in large ecosystems: a primary but neglected
+concept in the complexity-stability debate. \emph{PLOS Computational
+Biology}. \textbf{14}, e1005988 (2018).
 
 \hypertarget{ref-Song2018}{}
-21. Song, C. \& Saavedra, S. Will a small randomly assembled community
-be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
+21. C. Song, S. Saavedra, Will a small randomly assembled community be
+feasible and stable? \emph{Ecology}. \textbf{99}, 743--751 (2018).
 
 \hypertarget{ref-May1973}{}
-22. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
-\textbf{54,} 638--641 (1973).
+22. R. M. May, Qualitative stability in model ecosystems.
+\emph{Ecology}. \textbf{54}, 638--641 (1973).
 
@@
 
-\begin{center}
-\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
-\end{center}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{This supplemental information supports the manuscript
-``Component response rate variation drives stability in large complex
-systems'' with additional analyses to support its conclusions. All text,
-code, and data underlying this manuscript are publicly available on
-\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
-part of the RandomMatrixStability R package.}
-
-The
-\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability
-package} includes all functions and tools for recreating the text, this
-supplemental information, and running all code; additional documentation
-is also provided for package functions. The RandomMatrixStability
-package is available on
-\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
-download it, the
-\href{https://cran.r-project.org/web/packages/devtools/index.html}{\texttt{devtools}
-library} is needed.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{);}
-\KeywordTok{library}\NormalTok{(devtools);}
-\end{Highlighting}
-\end{Shaded}
-
-The code below installs the RandomMatrixStability package using
-devtools.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{install_github}\NormalTok{(}\StringTok{"bradduthie/RandomMatrixStability"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\section{Supplemental Information table of
-contents}\label{supplemental-information-table-of-contents}
-
-\begin{itemize}
-\tightlist
-\item
-  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
-\item
-  \protect\hyperlink{ga}{Stability given targeted manipulation of
-  \(\gamma\) (genetic algorithm)}
-\item
-  \protect\hyperlink{ecological}{Stability of ecological networks}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{competition}{Competitor networks}
-  \item
-    \protect\hyperlink{mutualism}{Mutualist networks}
-  \item
-    \protect\hyperlink{pred-prey}{Predator-prey networks}
-  \end{itemize}
-\item
-  \protect\hyperlink{connectance}{Sensitivity of connectance (C) values}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{connect3}{C = 0.3}
-  \item
-    \protect\hyperlink{connect5}{C = 0.5}
-  \item
-    \protect\hyperlink{connect7}{C = 0.7}
-  \item
-    \protect\hyperlink{connect9}{C = 0.9}
-  \end{itemize}
-\item
-  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
-  and \(\sigma\)}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
-  \item
-    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
-  \item
-    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
-  \item
-    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
-  \end{itemize}
-\item
-  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
-  \(\gamma\)}
-\item
-  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
-\end{itemize}
-
-\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
-\(S\)}{Stability across increasing S}}\label{IncrS}}
-
-Figure 3 of the main text reports the number of stable random complex
-systems found over 1 million iterations. The table below shows the
-results for all simulations of random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) given a range of
-\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
-matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
-after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
-Each row summarises data for a given \(S\) over 1 million randomly
-simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 1
-million). Similarly, the column \texttt{A1\_unstable} shows the number
-of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
-the number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 293 & 999707 & 293 & 999707 & 0 & 0\tabularnewline
-3 & 3602 & 996398 & 3609 & 996391 & 0 & 7\tabularnewline
-4 & 14937 & 985063 & 15008 & 984992 & 0 & 71\tabularnewline
-5 & 39289 & 960711 & 39783 & 960217 & 36 & 530\tabularnewline
-6 & 78845 & 921155 & 80207 & 919793 & 389 & 1751\tabularnewline
-7 & 133764 & 866236 & 136904 & 863096 & 1679 & 4819\tabularnewline
-8 & 204112 & 795888 & 208241 & 791759 & 5391 & 9520\tabularnewline
-9 & 288041 & 711959 & 291775 & 708225 & 12619 & 16353\tabularnewline
-10 & 384024 & 615976 & 384931 & 615069 & 23153 & 24060\tabularnewline
-11 & 485975 & 514025 & 481019 & 518981 & 35681 & 30725\tabularnewline
-12 & 590453 & 409547 & 577439 & 422561 & 48302 & 35288\tabularnewline
-13 & 689643 & 310357 & 669440 & 330560 & 57194 & 36991\tabularnewline
-14 & 777496 & 222504 & 751433 & 248567 & 60959 & 34896\tabularnewline
-15 & 850159 & 149841 & 821613 & 178387 & 58567 & 30021\tabularnewline
-16 & 905057 & 94943 & 877481 & 122519 & 51255 & 23679\tabularnewline
-17 & 943192 & 56808 & 919536 & 80464 & 40854 & 17198\tabularnewline
-18 & 969018 & 30982 & 949944 & 50056 & 30102 & 11028\tabularnewline
-19 & 984301 & 15699 & 970703 & 29297 & 20065 & 6467\tabularnewline
-20 & 992601 & 7399 & 983507 & 16493 & 12587 & 3493\tabularnewline
-21 & 996765 & 3235 & 991532 & 8468 & 7030 & 1797\tabularnewline
-22 & 998693 & 1307 & 995567 & 4433 & 3884 & 758\tabularnewline
-23 & 999503 & 497 & 997941 & 2059 & 1883 & 321\tabularnewline
-24 & 999861 & 139 & 999059 & 941 & 899 & 97\tabularnewline
-25 & 999964 & 36 & 999617 & 383 & 380 & 33\tabularnewline
-26 & 999993 & 7 & 999878 & 122 & 121 & 6\tabularnewline
-27 & 999995 & 5 & 999946 & 54 & 53 & 4\tabularnewline
-28 & 1000000 & 0 & 999975 & 25 & 25 & 0\tabularnewline
-29 & 1000000 & 0 & 999997 & 3 & 3 & 0\tabularnewline
-30 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
-31 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
-32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
-matrices found is greater than 1 whenever \(S > 10\) (compare column 5
-to column 3), and this ratio increases with increasing \(S\) (column 1).
-Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
-given variation in \(\gamma\) than when \(\gamma = 1\). Note that
-feasibility results were omitted for the table above, but are
-\protect\hyperlink{Feasibility}{reported below}.
-
-\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
-manipulation of \(\gamma\) (genetic
-algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
-
-Figure 4 of the main text reports the number of stable random complex
-systems found over 100000 using the genetic algorithm to maximise
-stability with a vector \(\mathbf{\gamma}\). Stability results for
-100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
-for \texttt{A0} indicate systems in which \(\gamma = 1\), while
-\texttt{A1} refers to systems in which the genetic algorithm searched
-for a set of \(\gamma\) values that stabilised the system.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
-3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
-4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
-5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
-6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
-7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
-8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
-9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
-10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
-11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
-12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
-13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
-14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
-15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
-16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
-17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
-18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
-19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
-20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
-21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
-22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
-23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
-24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
-25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
-26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
-27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
-28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
-29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
-30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
-31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
-32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
-33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
-34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
-35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
-36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
-37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
-38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
-39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
-40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. This comparison shows the high number of stable
-\(\mathbf{M}\) that can be produced through a targeted search of
-\(\gamma\) values, and suggests that many otherwise unstable systems
-could potentially be stabilised by an informed manipulation of their
-component response times. Such a possibility might conceivably reduce
-the dimensionality of problems involving stability in social-ecological
-or economic systems.
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\includegraphics{unnamed-chunk-7-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix
-\(\mathbf{M}\).
-
-\hypertarget{ecological}{\section{Stability of ecological
-networks}\label{ecological}}
-
-While the foundational work of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} applies broadly
-to complex networks, much attention has been given specifically to
-ecological networks of interacting species. In these networks, the
-matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
-and column is interpreted as a single species. The per capita effect
-that the density of any species \(i\) has on the population dynamics of
-species \(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds
-the effects of pair-wise interactions between \(S\)
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
-While May's original
-work\textsuperscript{\protect\hyperlink{ref-May1972}{1}} considered only
-randomly assembled communities, recent work has specifically looked at
-more restricted ecological communities including competitive networks
-(all off-diagonal elements of \(\mathbf{A}\) are negative), mutualist
-networks (all off-diagonal elements of \(\mathbf{A}\) are positive), and
-predator-prey networks (for any pair of \(i\) and \(j\), the effect of
-\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
-versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
-In general, competitor and mutualist networks tend to be unstable, while
-predator-prey networks tend to be highly
-stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-
-I investigated competitor, mutualist, and predator-prey networks
-following Allesina et
-al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
-these networks, I first generated a random matrix \(\mathbf{A}\), then
-changed the elements of \(\mathbf{A}\) accordingly. If \(\mathbf{A}\)
-was a competitive network, then the sign of any positive off-diagonal
-elements was reversed to be negative. If \(\mathbf{A}\) was a mutualist
-network, then the sign of any positive off-diagonal elements was
-reversed to be positive. And if \(\mathbf{A}\) was a predator-prey
-network, then all \(i\) and \(j\) pairs of elements were checked; any
-pairs of the same sign were changed so that one was negative and the
-other was positive.
-
-The number of stable \(\mathbf{M = \gamma A}\) systems was estimated
-\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
-matrices for values of \(S\) from 2 to 50 (100 in the case of the
-relatively more stable predator-prey interactions), except that only
-100000 random \(\mathbf{M}\) were generated instead of 1 million.
-
-The following tables for restricted ecological communities can therefore
-be compared with the random \(\mathbf{M}\)
-\protect\hyperlink{IncrS}{results above} (but note that counts from
-systems with comparable probabilities of stability will be an order of
-magnitude lower in the tables below due to the smaller number of
-\(\mathbf{M}\) matrices generated). As with the
-\protect\hyperlink{IncrS}{results above}, in the tables below,
-\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
-refers to matrices after \(Var(\gamma)\) is added. The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 100000).
-Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
-
-\textbf{Competition}
-
-Results for competitor interaction networks are shown below
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
-3 & 229 & 99771 & 231 & 99769 & 0 & 2\tabularnewline
-4 & 701 & 99299 & 704 & 99296 & 0 & 3\tabularnewline
-5 & 1579 & 98421 & 1587 & 98413 & 0 & 8\tabularnewline
-6 & 3218 & 96782 & 3253 & 96747 & 6 & 41\tabularnewline
-7 & 5519 & 94481 & 5619 & 94381 & 23 & 123\tabularnewline
-8 & 9062 & 90938 & 9237 & 90763 & 77 & 252\tabularnewline
-9 & 13436 & 86564 & 13729 & 86271 & 230 & 523\tabularnewline
-10 & 18911 & 81089 & 19303 & 80697 & 505 & 897\tabularnewline
-11 & 25594 & 74406 & 25961 & 74039 & 1011 & 1378\tabularnewline
-12 & 33207 & 66793 & 33382 & 66618 & 1724 & 1899\tabularnewline
-13 & 41160 & 58840 & 41089 & 58911 & 2655 & 2584\tabularnewline
-14 & 50575 & 49425 & 49894 & 50106 & 3777 & 3096\tabularnewline
-15 & 59250 & 40750 & 57892 & 42108 & 4824 & 3466\tabularnewline
-16 & 67811 & 32189 & 65740 & 34260 & 5634 & 3563\tabularnewline
-17 & 75483 & 24517 & 73056 & 26944 & 5943 & 3516\tabularnewline
-18 & 82551 & 17449 & 79878 & 20122 & 5780 & 3107\tabularnewline
-19 & 88030 & 11970 & 85204 & 14796 & 5417 & 2591\tabularnewline
-20 & 92254 & 7746 & 89766 & 10234 & 4544 & 2056\tabularnewline
-21 & 95233 & 4767 & 93002 & 6998 & 3695 & 1464\tabularnewline
-22 & 97317 & 2683 & 95451 & 4549 & 2803 & 937\tabularnewline
-23 & 98508 & 1492 & 97122 & 2878 & 1991 & 605\tabularnewline
-24 & 99240 & 760 & 98407 & 1593 & 1216 & 383\tabularnewline
-25 & 99669 & 331 & 99082 & 918 & 739 & 152\tabularnewline
-26 & 99871 & 129 & 99490 & 510 & 452 & 71\tabularnewline
-27 & 99938 & 62 & 99732 & 268 & 240 & 34\tabularnewline
-28 & 99985 & 15 & 99888 & 112 & 108 & 11\tabularnewline
-29 & 99990 & 10 & 99951 & 49 & 46 & 7\tabularnewline
-30 & 100000 & 0 & 99981 & 19 & 19 & 0\tabularnewline
-31 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
-32 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
-33 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-34 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Mutualism}
-
-Results for mutualist interaction networks are shown below
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 56 & 99944 & 56 & 99944 & 0 & 0\tabularnewline
-3 & 3301 & 96699 & 3301 & 96699 & 0 & 0\tabularnewline
-4 & 34446 & 65554 & 34446 & 65554 & 0 & 0\tabularnewline
-5 & 86520 & 13480 & 86520 & 13480 & 0 & 0\tabularnewline
-6 & 99683 & 317 & 99683 & 317 & 0 & 0\tabularnewline
-7 & 99998 & 2 & 99998 & 2 & 0 & 0\tabularnewline
-8 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-9 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-10 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-11 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-12 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Predator-prey}
-
-Results for predator-prey interaction networks are shown below
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-3 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-4 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-5 & 1 & 99999 & 1 & 99999 & 0 & 0\tabularnewline
-6 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
-7 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
-8 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
-9 & 20 & 99980 & 21 & 99979 & 0 & 1\tabularnewline
-10 & 20 & 99980 & 22 & 99978 & 0 & 2\tabularnewline
-11 & 38 & 99962 & 39 & 99961 & 0 & 1\tabularnewline
-12 & 64 & 99936 & 66 & 99934 & 0 & 2\tabularnewline
-13 & 87 & 99913 & 91 & 99909 & 0 & 4\tabularnewline
-14 & 157 & 99843 & 159 & 99841 & 0 & 2\tabularnewline
-15 & 215 & 99785 & 227 & 99773 & 0 & 12\tabularnewline
-16 & 293 & 99707 & 310 & 99690 & 0 & 17\tabularnewline
-17 & 383 & 99617 & 408 & 99592 & 0 & 25\tabularnewline
-18 & 443 & 99557 & 473 & 99527 & 3 & 33\tabularnewline
-19 & 642 & 99358 & 675 & 99325 & 4 & 37\tabularnewline
-20 & 836 & 99164 & 887 & 99113 & 7 & 58\tabularnewline
-21 & 1006 & 98994 & 1058 & 98942 & 10 & 62\tabularnewline
-22 & 1153 & 98847 & 1228 & 98772 & 20 & 95\tabularnewline
-23 & 1501 & 98499 & 1593 & 98407 & 30 & 122\tabularnewline
-24 & 1841 & 98159 & 1996 & 98004 & 40 & 195\tabularnewline
-25 & 2146 & 97854 & 2316 & 97684 & 58 & 228\tabularnewline
-26 & 2643 & 97357 & 2809 & 97191 & 119 & 285\tabularnewline
-27 & 3034 & 96966 & 3258 & 96742 & 158 & 382\tabularnewline
-28 & 3690 & 96310 & 3928 & 96072 & 201 & 439\tabularnewline
-29 & 4257 & 95743 & 4532 & 95468 & 290 & 565\tabularnewline
-30 & 4964 & 95036 & 5221 & 94779 & 424 & 681\tabularnewline
-31 & 5627 & 94373 & 5978 & 94022 & 452 & 803\tabularnewline
-32 & 6543 & 93457 & 6891 & 93109 & 666 & 1014\tabularnewline
-33 & 7425 & 92575 & 7777 & 92223 & 818 & 1170\tabularnewline
-34 & 8540 & 91460 & 8841 & 91159 & 1071 & 1372\tabularnewline
-35 & 9526 & 90474 & 9842 & 90158 & 1337 & 1653\tabularnewline
-36 & 10617 & 89383 & 10891 & 89109 & 1624 & 1898\tabularnewline
-37 & 12344 & 87656 & 12508 & 87492 & 2021 & 2185\tabularnewline
-38 & 13675 & 86325 & 13877 & 86123 & 2442 & 2644\tabularnewline
-39 & 15264 & 84736 & 15349 & 84651 & 2870 & 2955\tabularnewline
-40 & 17026 & 82974 & 17053 & 82947 & 3363 & 3390\tabularnewline
-41 & 18768 & 81232 & 18614 & 81386 & 3905 & 3751\tabularnewline
-42 & 20791 & 79209 & 20470 & 79530 & 4579 & 4258\tabularnewline
-43 & 23150 & 76850 & 22754 & 77246 & 5217 & 4821\tabularnewline
-44 & 25449 & 74551 & 24184 & 75816 & 6285 & 5020\tabularnewline
-45 & 27702 & 72298 & 26464 & 73536 & 6754 & 5516\tabularnewline
-46 & 30525 & 69475 & 28966 & 71034 & 7646 & 6087\tabularnewline
-47 & 32832 & 67168 & 31125 & 68875 & 8487 & 6780\tabularnewline
-48 & 36152 & 63848 & 33865 & 66135 & 9479 & 7192\tabularnewline
-49 & 38714 & 61286 & 36242 & 63758 & 10125 & 7653\tabularnewline
-50 & 41628 & 58372 & 38508 & 61492 & 11036 & 7916\tabularnewline
-51 & 44483 & 55517 & 41023 & 58977 & 11704 & 8244\tabularnewline
-52 & 48134 & 51866 & 44287 & 55713 & 12573 & 8726\tabularnewline
-53 & 51138 & 48862 & 46721 & 53279 & 13223 & 8806\tabularnewline
-54 & 54261 & 45739 & 49559 & 50441 & 13757 & 9055\tabularnewline
-55 & 57647 & 42353 & 52403 & 47597 & 14324 & 9080\tabularnewline
-56 & 60630 & 39370 & 55293 & 44707 & 14669 & 9332\tabularnewline
-57 & 63647 & 36353 & 57787 & 42213 & 15103 & 9243\tabularnewline
-58 & 66961 & 33039 & 60439 & 39561 & 15450 & 8928\tabularnewline
-59 & 69968 & 30032 & 63708 & 36292 & 15246 & 8986\tabularnewline
-60 & 72838 & 27162 & 66270 & 33730 & 15177 & 8609\tabularnewline
-61 & 75609 & 24391 & 68873 & 31127 & 15006 & 8270\tabularnewline
-62 & 77999 & 22001 & 71318 & 28682 & 14538 & 7857\tabularnewline
-63 & 80616 & 19384 & 73517 & 26483 & 14510 & 7411\tabularnewline
-64 & 83089 & 16911 & 76209 & 23791 & 13784 & 6904\tabularnewline
-65 & 85150 & 14850 & 78086 & 21914 & 13412 & 6348\tabularnewline
-66 & 86908 & 13092 & 80437 & 19563 & 12477 & 6006\tabularnewline
-67 & 88671 & 11329 & 82379 & 17621 & 11718 & 5426\tabularnewline
-68 & 90537 & 9463 & 84483 & 15517 & 10878 & 4824\tabularnewline
-69 & 91969 & 8031 & 86233 & 13767 & 10033 & 4297\tabularnewline
-70 & 93181 & 6819 & 87914 & 12086 & 9070 & 3803\tabularnewline
-71 & 94330 & 5670 & 89200 & 10800 & 8401 & 3271\tabularnewline
-72 & 95324 & 4676 & 90833 & 9167 & 7359 & 2868\tabularnewline
-73 & 96143 & 3857 & 91805 & 8195 & 6726 & 2388\tabularnewline
-74 & 96959 & 3041 & 93065 & 6935 & 5900 & 2006\tabularnewline
-75 & 97543 & 2457 & 93987 & 6013 & 5222 & 1666\tabularnewline
-76 & 97969 & 2031 & 94900 & 5100 & 4481 & 1412\tabularnewline
-77 & 98497 & 1503 & 95756 & 4244 & 3809 & 1068\tabularnewline
-78 & 98744 & 1256 & 96442 & 3558 & 3269 & 967\tabularnewline
-79 & 99045 & 955 & 96942 & 3058 & 2837 & 734\tabularnewline
-80 & 99276 & 724 & 97528 & 2472 & 2329 & 581\tabularnewline
-81 & 99481 & 519 & 97996 & 2004 & 1894 & 409\tabularnewline
-82 & 99556 & 444 & 98321 & 1679 & 1597 & 362\tabularnewline
-83 & 99691 & 309 & 98722 & 1278 & 1227 & 258\tabularnewline
-84 & 99752 & 248 & 98943 & 1057 & 1015 & 206\tabularnewline
-85 & 99833 & 167 & 99144 & 856 & 837 & 148\tabularnewline
-86 & 99895 & 105 & 99346 & 654 & 642 & 93\tabularnewline
-87 & 99925 & 75 & 99461 & 539 & 530 & 66\tabularnewline
-88 & 99945 & 55 & 99566 & 434 & 428 & 49\tabularnewline
-89 & 99976 & 24 & 99675 & 325 & 324 & 23\tabularnewline
-90 & 99977 & 23 & 99756 & 244 & 243 & 22\tabularnewline
-91 & 99982 & 18 & 99839 & 161 & 155 & 12\tabularnewline
-92 & 99988 & 12 & 99865 & 135 & 135 & 12\tabularnewline
-93 & 99994 & 6 & 99885 & 115 & 115 & 6\tabularnewline
-94 & 99993 & 7 & 99911 & 89 & 88 & 6\tabularnewline
-95 & 99998 & 2 & 99953 & 47 & 47 & 2\tabularnewline
-96 & 99999 & 1 & 99965 & 35 & 35 & 1\tabularnewline
-97 & 99999 & 1 & 99979 & 21 & 21 & 1\tabularnewline
-98 & 100000 & 0 & 99973 & 27 & 27 & 0\tabularnewline
-99 & 100000 & 0 & 99984 & 16 & 16 & 0\tabularnewline
-100 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-Overall, as
-expected\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}},
-predator-prey communities are relatively stable while mutualist
-communties are highly unstable. But interestingly, while \(Var(\gamma)\)
-stabilises predator-prey and competitor communities, it does not
-stabilise mutualist communities. This is unsurprising because purely
-mutualist communities are characterised by a very
-positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
-alone will shift all real parts of eigenvalues to negative values.
-
-\hypertarget{connectance}{\section{Sensitivity of connectance (C)
-values}\label{connectance}}
-
-In the main text, for simplicity, I assumed connectance values of
-\(C = 1\), meaning that all off-diagonal elements of a matrix
-\(\mathbf{M}\) were potentially nonzero and sampled from a normal
-distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
-I present four tables showing the number of stable communities given
-\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
-component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
-higher number of stable communities than when \(\gamma\) did not vary
-(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
-million \(\mathbf{M}\) were simulated. As with the results on
-\protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
-in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
-and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
-column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
-that are unstable, and the column \texttt{A0\_stable} shows the number
-of \texttt{A0} matrices that are stable (these two columns sum to
-100000). Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
-
-\textbf{Connectance \(\mathbf{C = 0.3}\)}
-
-\begin{longtable}[]{@{}llllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
-\midrule
-\endhead
-2 & 5 & 99995 & 5 & 99995 & 0 & 0 & 75110\tabularnewline
-3 & 6 & 99994 & 6 & 99994 & 0 & 0 & 87526\tabularnewline
-4 & 24 & 99976 & 24 & 99976 & 0 & 0 & 93713\tabularnewline
-5 & 59 & 99941 & 59 & 99941 & 0 & 0 & 96929\tabularnewline
-6 & 98 & 99902 & 98 & 99902 & 0 & 0 & 98492\tabularnewline
-7 & 160 & 99840 & 161 & 99839 & 0 & 1 & 99243\tabularnewline
-8 & 290 & 99710 & 293 & 99707 & 0 & 3 & 99582\tabularnewline
-9 & 430 & 99570 & 434 & 99566 & 0 & 4 & 99821\tabularnewline
-10 & 648 & 99352 & 653 & 99347 & 1 & 6 & 99895\tabularnewline
-11 & 946 & 99054 & 957 & 99043 & 0 & 11 & 99945\tabularnewline
-12 & 1392 & 98608 & 1415 & 98585 & 4 & 27 & 99978\tabularnewline
-13 & 2032 & 97968 & 2065 & 97935 & 5 & 38 & 99987\tabularnewline
-14 & 2627 & 97373 & 2688 & 97312 & 10 & 71 & 99988\tabularnewline
-15 & 3588 & 96412 & 3647 & 96353 & 35 & 94 & 99996\tabularnewline
-16 & 5019 & 94981 & 5124 & 94876 & 51 & 156 & 99998\tabularnewline
-17 & 6512 & 93488 & 6673 & 93327 & 79 & 240 & 99999\tabularnewline
-18 & 8444 & 91556 & 8600 & 91400 & 165 & 321 & 100000\tabularnewline
-19 & 10416 & 89584 & 10667 & 89333 & 244 & 495 & 100000\tabularnewline
-20 & 13254 & 86746 & 13477 & 86523 & 425 & 648 & 100000\tabularnewline
-21 & 16248 & 83752 & 16481 & 83519 & 642 & 875 & 100000\tabularnewline
-22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151 & 100000\tabularnewline
-23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490 & 100000\tabularnewline
-24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818 & 100000\tabularnewline
-25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137 & 100000\tabularnewline
-26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506 & 100000\tabularnewline
-27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811 & 100000\tabularnewline
-28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072 & 100000\tabularnewline
-29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321 & 100000\tabularnewline
-30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480 & 100000\tabularnewline
-31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568 & 100000\tabularnewline
-32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531 & 100000\tabularnewline
-33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360 & 100000\tabularnewline
-34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014 & 100000\tabularnewline
-35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773 & 100000\tabularnewline
-36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393 & 100000\tabularnewline
-37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913 & 100000\tabularnewline
-38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566 & 100000\tabularnewline
-39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187 & 100000\tabularnewline
-40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936 & 100000\tabularnewline
-41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683 & 100000\tabularnewline
-42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449 & 100000\tabularnewline
-43 & 99275 & 725 & 98312 & 1688 & 1251 & 288 & 100000\tabularnewline
-44 & 99583 & 417 & 98872 & 1128 & 903 & 192 & 100000\tabularnewline
-45 & 99776 & 224 & 99339 & 661 & 576 & 139 & 100000\tabularnewline
-46 & 99865 & 135 & 99518 & 482 & 413 & 66 & 100000\tabularnewline
-47 & 99938 & 62 & 99744 & 256 & 226 & 32 & 100000\tabularnewline
-48 & 99956 & 44 & 99824 & 176 & 151 & 19 & 100000\tabularnewline
-49 & 99980 & 20 & 99914 & 86 & 85 & 19 & 100000\tabularnewline
-50 & 99993 & 7 & 99950 & 50 & 46 & 3 & 100000\tabularnewline
-51 & 99998 & 2 & 99971 & 29 & 28 & 1 & 100000\tabularnewline
-52 & 99998 & 2 & 99986 & 14 & 14 & 2 & 100000\tabularnewline
-53 & 99999 & 1 & 99992 & 8 & 7 & 0 & 100000\tabularnewline
-54 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
-55 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-56 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
-57 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-58 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-100 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Connectance \(\mathbf{C = 0.5}\)}
-
-\begin{longtable}[]{@{}llllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
-\midrule
-\endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 74863\tabularnewline
-3 & 32 & 99968 & 32 & 99968 & 0 & 0 & 87434\tabularnewline
-4 & 122 & 99878 & 122 & 99878 & 0 & 0 & 93761\tabularnewline
-5 & 320 & 99680 & 321 & 99679 & 0 & 1 & 96830\tabularnewline
-6 & 667 & 99333 & 673 & 99327 & 0 & 6 & 98481\tabularnewline
-7 & 1233 & 98767 & 1252 & 98748 & 0 & 19 & 99187\tabularnewline
-8 & 2123 & 97877 & 2156 & 97844 & 3 & 36 & 99654\tabularnewline
-9 & 3415 & 96585 & 3471 & 96529 & 16 & 72 & 99816\tabularnewline
-10 & 5349 & 94651 & 5450 & 94550 & 30 & 131 & 99900\tabularnewline
-11 & 7990 & 92010 & 8185 & 91815 & 81 & 276 & 99958\tabularnewline
-12 & 11073 & 88927 & 11301 & 88699 & 219 & 447 & 99973\tabularnewline
-13 & 14971 & 85029 & 15204 & 84796 & 445 & 678 & 99986\tabularnewline
-14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002 & 99991\tabularnewline
-15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404 & 99996\tabularnewline
-16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980 & 99999\tabularnewline
-17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476 & 100000\tabularnewline
-18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937 & 99999\tabularnewline
-19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262 & 99999\tabularnewline
-20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527 & 100000\tabularnewline
-21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673 & 100000\tabularnewline
-22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491 & 100000\tabularnewline
-23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244 & 100000\tabularnewline
-24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876 & 100000\tabularnewline
-25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382 & 100000\tabularnewline
-26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901 & 100000\tabularnewline
-27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358 & 100000\tabularnewline
-28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974 & 100000\tabularnewline
-29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657 & 100000\tabularnewline
-30 & 99103 & 897 & 98033 & 1967 & 1466 & 396 & 100000\tabularnewline
-31 & 99502 & 498 & 98665 & 1335 & 1068 & 231 & 100000\tabularnewline
-32 & 99745 & 255 & 99185 & 815 & 696 & 136 & 100000\tabularnewline
-33 & 99881 & 119 & 99572 & 428 & 375 & 66 & 100000\tabularnewline
-34 & 99955 & 45 & 99788 & 212 & 191 & 24 & 100000\tabularnewline
-35 & 99979 & 21 & 99900 & 100 & 95 & 16 & 100000\tabularnewline
-36 & 99995 & 5 & 99950 & 50 & 50 & 5 & 100000\tabularnewline
-37 & 99997 & 3 & 99970 & 30 & 28 & 1 & 100000\tabularnewline
-38 & 99998 & 2 & 99986 & 14 & 13 & 1 & 100000\tabularnewline
-39 & 99999 & 1 & 99991 & 9 & 9 & 1 & 100000\tabularnewline
-40 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-41 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-42 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-43 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Connectance \(\mathbf{C = 0.7}\)}
-
-\begin{longtable}[]{@{}llllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
-\midrule
-\endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 75160\tabularnewline
-3 & 106 & 99894 & 106 & 99894 & 0 & 0 & 87447\tabularnewline
-4 & 395 & 99605 & 397 & 99603 & 0 & 2 & 93849\tabularnewline
-5 & 1117 & 98883 & 1123 & 98877 & 0 & 6 & 96827\tabularnewline
-6 & 2346 & 97654 & 2367 & 97633 & 6 & 27 & 98402\tabularnewline
-7 & 4314 & 95686 & 4388 & 95612 & 16 & 90 & 99218\tabularnewline
-8 & 7327 & 92673 & 7456 & 92544 & 61 & 190 & 99611\tabularnewline
-9 & 11514 & 88486 & 11792 & 88208 & 150 & 428 & 99808\tabularnewline
-10 & 16247 & 83753 & 16584 & 83416 & 415 & 752 & 99904\tabularnewline
-11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162 & 99952\tabularnewline
-12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818 & 99977\tabularnewline
-13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355 & 99984\tabularnewline
-14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927 & 99995\tabularnewline
-15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421 & 99994\tabularnewline
-16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527 & 99999\tabularnewline
-17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495 & 100000\tabularnewline
-18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351 & 100000\tabularnewline
-19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979 & 99999\tabularnewline
-20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339 & 100000\tabularnewline
-21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869 & 100000\tabularnewline
-22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267 & 100000\tabularnewline
-23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791 & 100000\tabularnewline
-24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504 & 100000\tabularnewline
-25 & 99433 & 567 & 98630 & 1370 & 1079 & 276 & 100000\tabularnewline
-26 & 99760 & 240 & 99259 & 741 & 655 & 154 & 100000\tabularnewline
-27 & 99895 & 105 & 99576 & 424 & 377 & 58 & 100000\tabularnewline
-28 & 99950 & 50 & 99790 & 210 & 194 & 34 & 100000\tabularnewline
-29 & 99981 & 19 & 99915 & 85 & 80 & 14 & 100000\tabularnewline
-30 & 99994 & 6 & 99952 & 48 & 47 & 5 & 100000\tabularnewline
-31 & 99998 & 2 & 99972 & 28 & 28 & 2 & 100000\tabularnewline
-32 & 99999 & 1 & 99992 & 8 & 8 & 1 & 100000\tabularnewline
-33 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
-34 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-35 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Connectance \(\mathbf{C = 0.9}\)}
-
-\begin{longtable}[]{@{}llllllll@{}}
-\toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
-\midrule
-\endhead
-2 & 14 & 99986 & 14 & 99986 & 0 & 0 & 75187\tabularnewline
-3 & 240 & 99760 & 240 & 99760 & 0 & 0 & 87443\tabularnewline
-4 & 1008 & 98992 & 1016 & 98984 & 0 & 8 & 93795\tabularnewline
-5 & 2708 & 97292 & 2729 & 97271 & 2 & 23 & 96814\tabularnewline
-6 & 5669 & 94331 & 5755 & 94245 & 13 & 99 & 98439\tabularnewline
-7 & 9848 & 90152 & 10057 & 89943 & 91 & 300 & 99208\tabularnewline
-8 & 15903 & 84097 & 16201 & 83799 & 336 & 634 & 99603\tabularnewline
-9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168 & 99803\tabularnewline
-10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852 & 99909\tabularnewline
-11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507 & 99951\tabularnewline
-12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127 & 99977\tabularnewline
-13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649 & 99986\tabularnewline
-14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765 & 99995\tabularnewline
-15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550 & 100000\tabularnewline
-16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092 & 100000\tabularnewline
-17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424 & 100000\tabularnewline
-18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746 & 100000\tabularnewline
-19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167 & 99999\tabularnewline
-20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729 & 100000\tabularnewline
-21 & 99111 & 889 & 98133 & 1867 & 1389 & 411 & 100000\tabularnewline
-22 & 99588 & 412 & 98905 & 1095 & 903 & 220 & 100000\tabularnewline
-23 & 99837 & 163 & 99480 & 520 & 452 & 95 & 100000\tabularnewline
-24 & 99932 & 68 & 99744 & 256 & 228 & 40 & 100000\tabularnewline
-25 & 99976 & 24 & 99863 & 137 & 133 & 20 & 100000\tabularnewline
-26 & 99995 & 5 & 99950 & 50 & 49 & 4 & 100000\tabularnewline
-27 & 99996 & 4 & 99986 & 14 & 13 & 3 & 100000\tabularnewline
-28 & 100000 & 0 & 99993 & 7 & 7 & 0 & 100000\tabularnewline
-29 & 100000 & 0 & 99996 & 4 & 4 & 0 & 100000\tabularnewline
-30 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
-strength (\(\sigma\))
-values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
-
-Results below show stability results given varying interaction strengths
-(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
-larger and increase by 10 with increasing rows). In the tables below (as
-\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
-matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-10 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-20 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-30 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-40 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-50 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-60 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
-70 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
-80 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
-90 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
-100 & 11 & 99989 & 11 & 99989 & 0 & 0\tabularnewline
-110 & 12 & 99988 & 13 & 99987 & 0 & 1\tabularnewline
-120 & 23 & 99977 & 23 & 99977 & 0 & 0\tabularnewline
-130 & 40 & 99960 & 40 & 99960 & 0 & 0\tabularnewline
-140 & 62 & 99938 & 65 & 99935 & 0 & 3\tabularnewline
-150 & 162 & 99838 & 165 & 99835 & 0 & 3\tabularnewline
-160 & 325 & 99675 & 329 & 99671 & 2 & 6\tabularnewline
-170 & 829 & 99171 & 851 & 99149 & 6 & 28\tabularnewline
-180 & 1817 & 98183 & 1860 & 98140 & 31 & 74\tabularnewline
-190 & 3927 & 96073 & 3989 & 96011 & 143 & 205\tabularnewline
-200 & 8084 & 91916 & 8048 & 91952 & 557 & 521\tabularnewline
-210 & 15558 & 84442 & 15147 & 84853 & 1534 & 1123\tabularnewline
-220 & 26848 & 73152 & 25342 & 74658 & 3625 & 2119\tabularnewline
-230 & 43386 & 56614 & 39535 & 60465 & 6992 & 3141\tabularnewline
-240 & 62734 & 37266 & 56684 & 43316 & 9815 & 3765\tabularnewline
-250 & 80128 & 19872 & 73080 & 26920 & 10128 & 3080\tabularnewline
-260 & 92206 & 7794 & 86619 & 13381 & 7490 & 1903\tabularnewline
-270 & 97946 & 2054 & 94824 & 5176 & 3797 & 675\tabularnewline
-280 & 99659 & 341 & 98534 & 1466 & 1265 & 140\tabularnewline
-290 & 99962 & 38 & 99696 & 304 & 281 & 15\tabularnewline
-300 & 99994 & 6 & 99964 & 36 & 34 & 4\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-10 & 3 & 99997 & 3 & 99997 & 0 & 0\tabularnewline
-20 & 15 & 99985 & 15 & 99985 & 0 & 0\tabularnewline
-30 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
-40 & 85 & 99915 & 85 & 99915 & 0 & 0\tabularnewline
-50 & 163 & 99837 & 163 & 99837 & 0 & 0\tabularnewline
-60 & 280 & 99720 & 282 & 99718 & 0 & 2\tabularnewline
-70 & 561 & 99439 & 566 & 99434 & 3 & 8\tabularnewline
-80 & 1009 & 98991 & 1029 & 98971 & 6 & 26\tabularnewline
-90 & 2126 & 97874 & 2175 & 97825 & 31 & 80\tabularnewline
-100 & 4580 & 95420 & 4653 & 95347 & 142 & 215\tabularnewline
-110 & 9540 & 90460 & 9632 & 90368 & 465 & 557\tabularnewline
-120 & 19090 & 80910 & 18668 & 81332 & 1676 & 1254\tabularnewline
-130 & 35047 & 64953 & 33220 & 66780 & 4172 & 2345\tabularnewline
-140 & 56411 & 43589 & 52439 & 47561 & 7297 & 3325\tabularnewline
-150 & 78003 & 21997 & 72574 & 27426 & 8477 & 3048\tabularnewline
-160 & 92678 & 7322 & 88438 & 11562 & 5901 & 1661\tabularnewline
-170 & 98614 & 1386 & 96670 & 3330 & 2397 & 453\tabularnewline
-180 & 99839 & 161 & 99418 & 582 & 499 & 78\tabularnewline
-190 & 99990 & 10 & 99945 & 55 & 52 & 7\tabularnewline
-200 & 100000 & 0 & 99995 & 5 & 5 & 0\tabularnewline
-210 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-10 & 36 & 99964 & 36 & 99964 & 0 & 0\tabularnewline
-20 & 195 & 99805 & 195 & 99805 & 0 & 0\tabularnewline
-30 & 519 & 99481 & 523 & 99477 & 0 & 4\tabularnewline
-40 & 1096 & 98904 & 1101 & 98899 & 2 & 7\tabularnewline
-50 & 2375 & 97625 & 2397 & 97603 & 9 & 31\tabularnewline
-60 & 4898 & 95102 & 4968 & 95032 & 83 & 153\tabularnewline
-70 & 10841 & 89159 & 10916 & 89084 & 432 & 507\tabularnewline
-80 & 22281 & 77719 & 21988 & 78012 & 1622 & 1329\tabularnewline
-90 & 42010 & 57990 & 39998 & 60002 & 4458 & 2446\tabularnewline
-100 & 67289 & 32711 & 63098 & 36902 & 7153 & 2962\tabularnewline
-110 & 88137 & 11863 & 84023 & 15977 & 6108 & 1994\tabularnewline
-120 & 97678 & 2322 & 95557 & 4443 & 2740 & 619\tabularnewline
-130 & 99795 & 205 & 99304 & 696 & 578 & 87\tabularnewline
-140 & 99989 & 11 & 99948 & 52 & 49 & 8\tabularnewline
-150 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-10 & 162 & 99838 & 162 & 99838 & 0 & 0\tabularnewline
-20 & 798 & 99202 & 799 & 99201 & 0 & 1\tabularnewline
-30 & 2273 & 97727 & 2289 & 97711 & 6 & 22\tabularnewline
-40 & 5259 & 94741 & 5298 & 94702 & 70 & 109\tabularnewline
-50 & 12084 & 87916 & 12054 & 87946 & 446 & 416\tabularnewline
-60 & 26072 & 73928 & 25511 & 74489 & 1810 & 1249\tabularnewline
-70 & 50121 & 49879 & 47747 & 52253 & 4748 & 2374\tabularnewline
-80 & 77806 & 22194 & 73810 & 26190 & 6421 & 2425\tabularnewline
-90 & 94862 & 5138 & 92069 & 7931 & 3842 & 1049\tabularnewline
-100 & 99527 & 473 & 98822 & 1178 & 870 & 165\tabularnewline
-110 & 99984 & 16 & 99912 & 88 & 80 & 8\tabularnewline
-120 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-130 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
-distribution of
-\(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
-
-In the main text, I considered a uniform distribution of component
-response rates \(\gamma \sim \mathcal{U}(0, 2)\). The number of unstable
-and stable \(\mathbf{M}\) matrices are reported in
-\protect\hyperlink{IncrS}{a table above} across different values of
-\(S\). Here I show complementary results for three different
-distributions including an exponential, beta, and gamma distribution of
-\(\gamma\) values. The shape of these distributions is shown in the
-figure below.
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Distributions of component response rate
-(\(\boldsymbol{\gamma}\)) values in complex systems.} The stabilities of
-simulated complex systems with these \(\gamma\) distributions are
-compared to otherwise identical complex systems with a fixed component
-response rate of \(\gamma = 1\) across different system sizes (\(S\);
-i.e., component numbers) given a unit \(\gamma\) standard deviation
-(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
-uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
-\(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
-panel shows 1 million randomly generated \(\gamma\) values.
-
-\includegraphics{unnamed-chunk-23-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-The same 100000 \(\mathbf{M}\) matrices were used to investigate
-stability when applying each of these different distributions of
-\(\gamma\) values. The table below shows the number of \(\mathbf{M}\)
-that were unstable (\_unst) and stable (\_stbl) for the exponential
-(Exp), beta, and gamma distributions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
-\KeywordTok{kable}\NormalTok{(fourdists);}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{longtable}[]{@{}lllllll@{}}
-\toprule
-S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
-gamma\_stbl\tabularnewline
-\midrule
-\endhead
-2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
-3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
-4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
-5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
-6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
-7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
-8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
-9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
-10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
-11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
-12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
-13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
-14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
-15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
-16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
-17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
-18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
-19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
-20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
-21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
-22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
-23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
-24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
-25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
-26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
-27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
-28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
-29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
-30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
-32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-In comparison to the uniform distribution (a), proportionally fewer
-random systems are found with the exponential distribution (b), while
-more are found with the beta (c) and gamma (d) distributions.
-
-\hypertarget{Feasibility}{\section{Feasibility of complex
-systems}\label{Feasibility}}
-
-When feasibility was evaluated with and without variation in \(\gamma\),
-there was no increase in stability for \(\mathbf{M}\) where \(\gamma\)
-varied as compared to where \(\gamma = 1\). Results below illustrate
-this result, which was general to all other simulations performed.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
-A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
-\midrule
-\endhead
-2 & 749978 & 250022 & 749942 & 250058 & 35552 & 35516\tabularnewline
-3 & 874519 & 125481 & 874296 & 125704 & 36803 & 36580\tabularnewline
-4 & 937192 & 62808 & 937215 & 62785 & 26440 & 26463\tabularnewline
-5 & 968776 & 31224 & 968639 & 31361 & 16319 & 16182\tabularnewline
-6 & 984313 & 15687 & 984463 & 15537 & 9006 & 9156\tabularnewline
-7 & 992149 & 7851 & 992161 & 7839 & 4991 & 5003\tabularnewline
-8 & 996124 & 3876 & 996103 & 3897 & 2644 & 2623\tabularnewline
-9 & 998014 & 1986 & 998027 & 1973 & 1361 & 1374\tabularnewline
-10 & 999031 & 969 & 999040 & 960 & 698 & 707\tabularnewline
-11 & 999546 & 454 & 999514 & 486 & 377 & 345\tabularnewline
-12 & 999764 & 236 & 999792 & 208 & 160 & 188\tabularnewline
-13 & 999883 & 117 & 999865 & 135 & 105 & 87\tabularnewline
-14 & 999938 & 62 & 999945 & 55 & 40 & 47\tabularnewline
-15 & 999971 & 29 & 999964 & 36 & 31 & 24\tabularnewline
-16 & 999988 & 12 & 999991 & 9 & 8 & 11\tabularnewline
-17 & 999996 & 4 & 999991 & 9 & 8 & 3\tabularnewline
-18 & 999997 & 3 & 999999 & 1 & 1 & 3\tabularnewline
-19 & 999998 & 2 & 999997 & 3 & 3 & 2\tabularnewline
-20 & 1000000 & 0 & 999999 & 1 & 1 & 0\tabularnewline
-21 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-22 & 999999 & 1 & 1000000 & 0 & 0 & 1\tabularnewline
-23 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-24 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-25 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-26 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-27 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-28 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-29 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-30 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-31 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-32 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-33 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-34 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-35 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-36 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-37 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-38 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-39 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-40 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-41 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-42 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-43 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-44 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-45 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-46 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-47 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-48 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-49 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-50 & 1000000 & 0 & 1000000 & 0 & 0 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
-in pure species interaction networks.
-
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
-
-\hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
-\emph{Nature} \textbf{483,} 205--208 (2012).
-
-\hypertarget{ref-Allesina2015}{}
-3. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
-(2015).
+\textbf{Figure 1: Example distribution of eigenvalues before (a) and
+after (b) separating a randomly generated complex system into fast
+(\(\boldsymbol{\gamma} = 1.95\)) and slow
+(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
+shows the same system where \(S = 200\), \(C = 0.05\), and
+\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
+distribution of \(\gamma\) differs between panels). \textbf{a.}
+Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
+uniformly distributed within the blue circle with a radius of
+\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
+Eigenvalues plotted when half \(\gamma = 1.95\) and half
+\(\gamma = 0.05\); distributions of points can be partitioned into one
+large circle centred at \(\gamma = -1.95\) and one small circle centred
+at \(\gamma = -0.05\). In a, the maximum real eigenvalue
+\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
+\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
+complex system in b but not a is stable because in b
+\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
+complex systems under the same parameter values, 1 was stable when
+\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
+Overall, complex systems that are separated into fast versus slow
+components tend to be more stable than otherwise identical systems with
+identical component response rates.
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+\clearpage
+
+\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
+introducing variation in component response rate
+(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
+system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
+\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
+\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
+\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
+interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
+eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
+both panels show the circle centred on the distribution in panel a.
+Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.734 for a and b,
+respectively.
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+
+\clearpage
+
+\textbf{Figure 3: Stability of large complex systems with and without
+variation in component response rate (\(\boldsymbol{\gamma}\)).} The
+\(\ln\) number of systems that are stable across different system sizes
+(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
+which variation in \(\gamma\) is critical for system stability. For each
+\(S\), 1 million complex systems are randomly generated. Stability of
+each complex system is tested given variation in \(\gamma\) by randomly
+sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
+\(Var(\gamma)\) is then compared to stability in an otherwise identical
+system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
+Blue and red bars show the number of stable systems in the absence and
+presence of \(Var(\gamma)\), respectively. The black line shows the
+proportion of systems that are stable when \(Var(\gamma) > 0\), but
+would be unstable if \(Var(\gamma) = 0\).
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-15-1.pdf}
+
+\clearpage
+
+\textbf{Figure 4: Stability of large complex systems given
+\(\boldsymbol{\gamma = 1}\) versus targeted
+\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
+stable across different system sizes (\(S\), max \(S=40\)) for
+\(C = 1\), and the proportion of systems wherein a targeted search of
+\(\gamma\) values successfully resulted in system stability. For each
+\(S\), 100000 complex systems are randomly generated. Stability of each
+complex system is tested given variation in \(\gamma\) using a genetic
+algorithm to maximise the effect of \(\gamma\) values on increasing
+stability, as compared to stability in an otherwise identical system in
+which \(\gamma\) is the same for all components. Blue bars show the
+number of stable systems in the absence of component response rate
+variation, while red bars show the number of stable systems that can be
+generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(Var(\gamma) = 0\).
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
 

### 2018-08-10 notebook/ms.Rmd
Update files for PLoS Biol

@@
 linestretch: 2
-link-citations: no
+link-citations: yes
 linkcolor: blue
-csl: science.csl
+csl: plos.csl
 biblio-style: apalike
@@
 
-**The stability of a complex system generally decreases with increasing system size and interconnectivity, a counterintuitive result of widespread importance across the physical, life, and social sciences.
+**Key words:** Ecological networks, gene-regulatory networks, neural networks, financial networks, system stability, random matrix theory
+
+Abstract
+--------------------------------------------------------------------------------
+
+The stability of a complex system generally decreases with increasing system size and interconnectivity, a counterintuitive result of widespread importance across the physical, life, and social sciences.
 Despite recent interest in the relationship between system properties and stability, the effect of variation in the response rate of individual system components remains unconsidered.
@@
 Variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, such that the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
-My results reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.**
+My results reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
+
+Author Summary
+--------------------------------------------------------------------------------
+
+Any set of discrete components that potentially interact can define a complex system.
+Typically, properties of individual components will cause them to respond to system perturbation at different rates, but the consequences of such variation for system stability are unexplored.
+This work makes three novel theoretical contributions: First, I show that large complex systems with varying component response rates are more likely to be stable than when response rates are uniform.
+Second, I show that a targeted manipulation of component response rates can increase the probability of system stability by several orders of magnitude.
+Lastly, I show that varying component response rate does not affect system feasibility.
+These results are general to any physical, biological, social, or mixed system.
 
@@
 --->
+
+Introduction
+--------------------------------------------------------------------------------
 

### 2018-08-10 notebook/ms.tex
Update files for PLoS Biol

@@
 
-\textbf{The stability of a complex system generally decreases with
-increasing system size and interconnectivity, a counterintuitive result
-of widespread importance across the physical, life, and social sciences.
+\textbf{Key words:} Ecological networks, gene-regulatory networks,
+neural networks, financial networks, system stability, random matrix
+theory
+
+\subsection{Abstract}\label{abstract}
+
+The stability of a complex system generally decreases with increasing
+system size and interconnectivity, a counterintuitive result of
+widespread importance across the physical, life, and social sciences.
 Despite recent interest in the relationship between system properties
@@
 previously unconsidered driver of system stability that is likely to be
-pervasive across all complex systems.}
+pervasive across all complex systems.
 
-In 1972, May (\emph{1}) first demonstrated that randomly assembled
-systems of sufficient complexity are almost inevitably unstable given
-infinitesimally small perturbations. Complexity in this case is defined
-by the size of the system (i.e., the number of potentially interacting
-components; \(S\)), its connectance (i.e., the probability that one
-component will interact with another; \(C\)), and the variance of
-interaction strengths (\(\sigma^{2}\)) (\emph{2}). May's finding that
-the probability of local stability falls to near zero given a
-sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly relevant
-for understanding the dynamics and persistence of systems such as
-ecological (\emph{1}--\emph{5}), neurological (\emph{6}, \emph{7}),
-biochemical (\emph{8}, \emph{9}), and socio-economic
-(\emph{10}--\emph{13}) networks. As such, identifying general principles
-that drive stability in complex systems is of wide-ranging importance.
+\subsection{Author Summary}\label{author-summary}
+
+Any set of discrete components that potentially interact can define a
+complex system. Typically, properties of individual components will
+cause them to respond to system perturbation at different rates, but the
+consequences of such variation for system stability are unexplored. This
+work makes three novel theoretical contributions: First, I show that
+large complex systems with varying component response rates are more
+likely to be stable than when response rates are uniform. Second, I show
+that a targeted manipulation of component response rates can increase
+the probability of system stability by several orders of magnitude.
+Lastly, I show that varying component response rate does not affect
+system feasibility. These results are general to any physical,
+biological, social, or mixed system.
+
+\subsection{Introduction}\label{introduction}
+
+In 1972, May {[}\protect\hyperlink{ref-May1972}{1}{]} first demonstrated
+that randomly assembled systems of sufficient complexity are almost
+inevitably unstable given infinitesimally small perturbations.
+Complexity in this case is defined by the size of the system (i.e., the
+number of potentially interacting components; \(S\)), its connectance
+(i.e., the probability that one component will interact with another;
+\(C\)), and the variance of interaction strengths (\(\sigma^{2}\))
+{[}\protect\hyperlink{ref-Allesina2012}{2}{]}. May's finding that the
+probability of local stability falls to near zero given a sufficiently
+high threshold of \(\sigma\sqrt{SC}\) is broadly relevant for
+understanding the dynamics and persistence of systems such as ecological
+{[}\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}{]},
+neurological
+{[}\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}{]},
+biochemical
+{[}\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}{]},
+and socio-economic
+{[}\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}{]}
+networks. As such, identifying general principles that drive stability
+in complex systems is of wide-ranging importance.
 
@@
 matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
-species (\emph{2}) or banks (\emph{11})). One element of such a matrix,
-\(M_{ij}\), defines how component \(j\) affects component \(i\) in the
-system at a point of equilibrium (\emph{2}). Off-diagonal elements
+species {[}\protect\hyperlink{ref-Allesina2012}{2}{]} or banks
+{[}\protect\hyperlink{ref-Haldane2011}{11}{]}). One element of such a
+matrix, \(M_{ij}\), defines how component \(j\) affects component \(i\)
+in the system at a point of equilibrium
+{[}\protect\hyperlink{ref-Allesina2012}{2}{]}. Off-diagonal elements
 (\(i \neq j\)) therefore define interactions between components, while
@@
 \(C\), which are sampled from a distribution with variance
-\(\sigma^{2}\); diagonal elements are set to -1 (\emph{1}, \emph{2},
-\emph{4}). Local system stability is assessed using eigenanalysis, with
-the system being stable if the real parts of all eigenvalues
-(\(\lambda\)) of \(\mathbf{M}\) are negative
-(\(\max\left(\Re(\lambda)\right) < 0\)) (\emph{1}, \emph{2}). In a large
-system (high \(S\)), eigenvalues are distributed uniformly (\emph{14})
-within a circle centred at \(\Re = -1\) (the mean value of diagonal
-elements) and \(\Im = 0\), with a radius of \(\sigma\sqrt{SC}\)
-(\emph{1}, \emph{2}, \emph{4}) (Figs 1a and 2a). Local stability of
-randomly assembled systems therefore becomes increasingly unlikely as
-\(S\), \(C\), and \(\sigma^{2}\) increase.
+\(\sigma^{2}\); diagonal elements are set to -1
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}{]}.
+Local system stability is assessed using eigenanalysis, with the system
+being stable if the real parts of all eigenvalues (\(\lambda\)) of
+\(\mathbf{M}\) are negative (\(\max\left(\Re(\lambda)\right) < 0\))
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}.
+In a large system (high \(S\)), eigenvalues are distributed uniformly
+{[}\protect\hyperlink{ref-Tao2010}{14}{]} within a circle centred at
+\(\Re = -1\) (the mean value of diagonal elements) and \(\Im = 0\), with
+a radius of \(\sigma\sqrt{SC}\)
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}{]}
+(Figs 1a and 2a). Local stability of randomly assembled systems
+therefore becomes increasingly unlikely as \(S\), \(C\), and
+\(\sigma^{2}\) increase.
 
-May's (\emph{1}, \emph{2}) stability criterion \(\sigma\sqrt{SC} < 1\)
-assumes that individual components respond to perturbations of the
-system at the same rate (\(\gamma\)), but this is highly unlikely in any
-complex system. In ecological communities, for example, the rate at
-which population density changes following perturbation will depend on
-the generation time of organisms, which might vary by orders of
-magnitude among species. Species with short generation times will
-respond quickly (high \(\gamma\)) to perturbations relative to species
-with long generation times (low \(\gamma\)). Similarly, the speed at
-which individual banks respond to perturbations in financial networks,
-or individuals or institutions respond to perturbations in complex
-social networks, is likely to vary. The effect of such variance on
-stability has not been investigated in complex systems theory.
-Intuitively, variation in \(\gamma\) might be expected to decrease
-system stability by introducing a new source of variation into the
-system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\gamma\) varies are
-actually more likely to be stable, especially when \(S\) is high.
+May's
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}
+stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
+components respond to perturbations of the system at the same rate
+(\(\gamma\)), but this is highly unlikely in any complex system. In
+ecological communities, for example, the rate at which population
+density changes following perturbation will depend on the generation
+time of organisms, which might vary by orders of magnitude among
+species. Species with short generation times will respond quickly (high
+\(\gamma\)) to perturbations relative to species with long generation
+times (low \(\gamma\)). Similarly, the speed at which individual banks
+respond to perturbations in financial networks, or individuals or
+institutions respond to perturbations in complex social networks, is
+likely to vary. The effect of such variance on stability has not been
+investigated in complex systems theory. Intuitively, variation in
+\(\gamma\) might be expected to decrease system stability by introducing
+a new source of variation into the system and thereby increasing
+\(\sigma\). Here I show why, despite higher \(\sigma\), complex systems
+in which \(\gamma\) varies are actually more likely to be stable,
+especially when \(S\) is high.
 
@@
 time can be modelled by multiplying all row elements by a real scalar
-value \(\gamma_{i}\) (\emph{15}). The distribution of \(\gamma\) over
-\(S\) components thereby models the distribution of component response
-rates. An instructive example compares one \(\mathbf{M}\) where
-\(\gamma_{i} = 1\) for all \(i\) in \(S\) to the same \(\mathbf{M}\)
-when half of \(\gamma_{i} = 1.95\) and half of \(\gamma_{i} = 0.05\).
-This models one system in which \(\gamma\) is invariant and one in which
-\(\gamma\) varies, but systems are otherwise identical (note
-\(E[\gamma_{i}] = 1\) in both cases). I assume \(S = 200\),
-\(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are set to \(-1\)
-and non-zero off-diagonal elements are drawn randomly from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
-\(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
+value \(\gamma_{i}\) {[}\protect\hyperlink{ref-Patel2018}{15}{]}. The
+distribution of \(\gamma\) over \(S\) components thereby models the
+distribution of component response rates. An instructive example
+compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
+\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
+half of \(\gamma_{i} = 0.05\). This models one system in which
+\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
+are otherwise identical (note \(E[\gamma_{i}] = 1\) in both cases). I
+assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal
+elements are set to \(-1\) and non-zero off-diagonal elements are drawn
+randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied
+by \(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
 eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
@@
 distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected (\emph{14}), when \(\gamma = 1\), eigenvalues are distributed
-uniformly in a circle centred at (\(-1, 0\)) with a radius of
-\(\sigma\sqrt{SC} =\) 2.828. Uniform variation in \(\gamma\) leads to a
-non-uniform distribution of eigenvalues (\emph{16}--\emph{18}), some of
-which are clustered locally near the centre of the distribution, but
-others of which are spread outside the former radius of 2.828 (Fig 2b).
-The clustering and spreading of eigenvalues introduced by
+expected {[}\protect\hyperlink{ref-Tao2010}{14}{]}, when \(\gamma = 1\),
+eigenvalues are distributed uniformly in a circle centred at (\(-1, 0\))
+with a radius of \(\sigma\sqrt{SC} =\) 2.828. Uniform variation in
+\(\gamma\) leads to a non-uniform distribution of eigenvalues
+{[}\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}{]},
+some of which are clustered locally near the centre of the distribution,
+but others of which are spread outside the former radius of 2.828 (Fig
+2b). The clustering and spreading of eigenvalues introduced by
 \(Var(\gamma)\) can destabilise previously stable systems or stabilise
@@
 complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above May's (\emph{1}, \emph{2}) threshold
-\(\sigma\sqrt{SC} < 1\).
+can often lead to stability above May's
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}
+threshold \(\sigma\sqrt{SC} < 1\).
 
@@
 values leading to stability in a large complex system) is too large to
-search exhaustively (\emph{19}). Generations of selection on \(\gamma\)
-value combinations to minimise \(\max\left(\Re(\lambda)\right)\)
-demonstrated the potential for \(Var(\gamma)\) to increase system
-stability. Across \(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\)
-values were found that resulted in stable systems with probabilities
-that were up to four orders of magnitude higher than when \(\gamma = 1\)
-(Fig. 4), meaning that stability could often be achieved by manipulating
-\(S\) \(\gamma\) values rather than \(S \times S\) \(\mathbf{M}\)
-elements.
+search exhaustively {[}\protect\hyperlink{ref-Hamblin2013}{19}{]}.
+Generations of selection on \(\gamma\) value combinations to minimise
+\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
+\(Var(\gamma)\) to increase system stability. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were up to four
+orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
+that stability could often be achieved by manipulating \(S\) \(\gamma\)
+values rather than \(S \times S\) \(\mathbf{M}\) elements.
 
@@
 the system. Feasibilility assumes that the system is stable and values
-of all components are positive at equilibrium (\emph{5}, \emph{20},
-\emph{21}). This is of particular interest for ecological communities
-because population density cannot take negative values, meaning that
-ecological systems need to be feasible for stability to be biologically
-realistic (\emph{20}). While my results are intended to be general to
-all complex systems, and not restricted to species networks, I have also
-performed a feasibility analysis on all matrices \(\mathbf{M}\) tested
-for stability, and additionally for specific types of ecological
-communities (\emph{2}) (e.g., competitive, mutualist, predator-prey; see
-Supplementary Information). Feasibility was unaffected by
-\(Var(\gamma)\), meaning that for pure interacting species networks,
-variation in component response rate (i.e., species generation time)
-does not affect stability at biologically realistic species densities.
+of all components are positive at equilibrium
+{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}{]}.
+This is of particular interest for ecological communities because
+population density cannot take negative values, meaning that ecological
+systems need to be feasible for stability to be biologically realistic
+{[}\protect\hyperlink{ref-Dougoud2018}{20}{]}. While my results are
+intended to be general to all complex systems, and not restricted to
+species networks, I have also performed a feasibility analysis on all
+matrices \(\mathbf{M}\) tested for stability, and additionally for
+specific types of ecological communities
+{[}\protect\hyperlink{ref-Allesina2012}{2}{]} (e.g., competitive,
+mutualist, predator-prey; see Supplementary Information). Feasibility
+was unaffected by \(Var(\gamma)\), meaning that for pure interacting
+species networks, variation in component response rate (i.e., species
+generation time) does not affect stability at biologically realistic
+species densities.
 
@@
 size increases, and can ultimately increase the overall probability that
-system stability is observed above that predicted by May's (\emph{1})
-classically derived \(\sigma \sqrt{SC}\) criterion. The logic outlined
-here is general, and potentially applies to any complex system in which
-individual system components can vary in their reaction rates to system
-perturbation.
+system stability is observed above that predicted by May's
+{[}\protect\hyperlink{ref-May1972}{1}{]} classically derived
+\(\sigma \sqrt{SC}\) criterion. The logic outlined here is general, and
+potentially applies to any complex system in which individual system
+components can vary in their reaction rates to system perturbation.
 
@@
 underlying this shift in eigenvalue distribution has been investigated
-(\emph{16}) and recently applied to questions concerning species density
-and feasibility (\emph{17}, \emph{18}), but has not been interpreted as
-rates of response of individual system components to perturbation.
+{[}\protect\hyperlink{ref-Ahmadian2015}{16}{]} and recently applied to
+questions concerning species density and feasibility
+{[}\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}{]},
+but has not been interpreted as rates of response of individual system
+components to perturbation.
 
@@
 important to consider, particularly for the study of ecological networks
-of species (\emph{5}, \emph{18}, \emph{20}) because population densities
-cannot realistically be negative. My results therefore suggest that
-variation in the rate of population responses to perturbation (e.g., due
-to differences in generation time among species) is unlikely to be
-critical to the stability of purely multi-species interaction networks
-(see also Supplementary Information). Nevertheless, ecological
-interactions do not exist in isolation in empirical systems (\emph{15}),
-but instead interact with evolutionary, abiotic, or social-economic
-systems. The relevance of component response rate for complex system
-stability should therefore not be ignored in the broader context of
-ecological communities.
+of species
+{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20}{]}
+because population densities cannot realistically be negative. My
+results therefore suggest that variation in the rate of population
+responses to perturbation (e.g., due to differences in generation time
+among species) is unlikely to be critical to the stability of purely
+multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in
+empirical systems {[}\protect\hyperlink{ref-Patel2018}{15}{]}, but
+instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability
+should therefore not be ignored in the broader context of ecological
+communities.
 
@@
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
-traits (\(\mathbf{x}\))(\emph{15}). For any species \(i\) or trait
-\(j\), change in species density (\(N_{i}\)) or trait value (\(x_{j}\))
-with time (\(t\)) is a function of the vectors \(\mathbf{N}\) and
-\(\mathbf{x}\),
+traits (\(\mathbf{x}\)){[}\protect\hyperlink{ref-Patel2018}{15}{]}. For
+any species \(i\) or trait \(j\), change in species density (\(N_{i}\))
+or trait value (\(x_{j}\)) with time (\(t\)) is a function of the
+vectors \(\mathbf{N}\) and \(\mathbf{x}\),
 
@@
 slow or fast in comparison with the change in species densities
-(\emph{15}), and this is modulated in the above by the scalar
-\(\epsilon\). The value of \(\epsilon\) thereby determines the timescale
-separation between ecology and evolution, with high \(\epsilon\)
-modelling relatively fast evolution and low \(\epsilon\) modelling
-relative slow evolution (\emph{15}).
+{[}\protect\hyperlink{ref-Patel2018}{15}{]}, and this is modulated in
+the above by the scalar \(\epsilon\). The value of \(\epsilon\) thereby
+determines the timescale separation between ecology and evolution, with
+high \(\epsilon\) modelling relatively fast evolution and low
+\(\epsilon\) modelling relative slow evolution
+{[}\protect\hyperlink{ref-Patel2018}{15}{]}.
 
@@
 rate of evolution to modulate rates of component responses for \(S\)
-components. Following May (\emph{1}, \emph{22}), the value of a
-component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected by the value of
-\(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on \(i\) (\(a_{ij}\)),
-and by \(i\)'s response rate (\(\gamma_{i}\)),
+components. Following May
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{22}{]},
+the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
+by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
+\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
 
@@
 
-In matrix notation(\emph{22}),
+In matrix notation{[}\protect\hyperlink{ref-May1973}{22}{]},
 
@@
 \(\mathbf{M} =\mathbf{\gamma} \mathbf{A}\) modulates the values of
-components and can be analysed using the techniques of May (\emph{1},
-\emph{16}, \emph{22}).
+components and can be analysed using the techniques of May
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{22}{]}.
 
@@
 genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological evolution (\emph{19}).
-In this case, the practical solution is finding vectors of
-\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
-\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
-a large population of 1000 different potential \(\mathbf{\gamma}\)
-vectors and allowing this population to evolve through a process of
-mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
-selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
-found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
-met.
+solutions by mimicking the process of biological evolution
+{[}\protect\hyperlink{ref-Hamblin2013}{19}{]}. In this case, the
+practical solution is finding vectors of \(\mathbf{\gamma}\) that
+decrease the most positive real eigenvalue of \(\mathbf{M}\). The
+genetic algorithm used achieves this by initialising a large population
+of 1000 different potential \(\mathbf{\gamma}\) vectors and allowing
+this population to evolve through a process of mutation, crossover
+(swaping \(\gamma_{i}\) values between vectors), selection, and
+reproduction until either a \(\mathbf{\gamma}\) vector is found where
+all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is met.
 
@@
 
-\textbf{System feasibility}. Dougoud et al. (\emph{20}) define the
-following feasibility criteria for ecological systems characterised by
-\(S\) interacting species with varying densities in a classical
-Lotka-Volterra model,
+\textbf{System feasibility}. Dougoud et al.
+{[}\protect\hyperlink{ref-Dougoud2018}{20}{]} define the following
+feasibility criteria for ecological systems characterised by \(S\)
+interacting species with varying densities in a classical Lotka-Volterra
+model,
 
@@
 affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
-relates to May's(\emph{1}) stability criterion (\emph{20}) by
+relates to May's{[}\protect\hyperlink{ref-May1972}{1}{]} stability
+criterion {[}\protect\hyperlink{ref-Dougoud2018}{20}{]} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 \hypertarget{ref-May1972}{}
-1. R. M. May, Will a large complex system be stable? \emph{Nature}.
-\textbf{238}, 413--414 (1972).
+1. May RM. Will a large complex system be stable? Nature. 1972;238:
+413--414.
 
 \hypertarget{ref-Allesina2012}{}
-2. S. Allesina, S. Tang, Stability criteria for complex ecosystems.
-\emph{Nature}. \textbf{483}, 205--208 (2012).
+2. Allesina S, Tang S. Stability criteria for complex ecosystems.
+Nature. Nature Publishing Group; 2012;483: 205--208.
+doi:\href{https://doi.org/10.1038/nature10832}{10.1038/nature10832}
 
 \hypertarget{ref-Mougi2012}{}
-3. A. Mougi, M. Kondoh, Diversity of interaction types and ecological
-community stability. \emph{Science}. \textbf{337}, 349--351 (2012).
+3. Mougi A, Kondoh M. Diversity of interaction types and ecological
+community stability. Science. 2012;337: 349--351.
+doi:\href{https://doi.org/10.1126/science.1220529}{10.1126/science.1220529}
 
 \hypertarget{ref-Allesina2015}{}
-4. S. Allesina \emph{et al.}, Predicting the stability of large
-structured food webs. \emph{Nature Communications}. \textbf{6}, 7842
-(2015).
+4. Allesina S, Grilli J, Barabás G, Tang S, Aljadeff J, Maritan A.
+Predicting the stability of large structured food webs. Nature
+Communications. 2015;6: 7842.
+doi:\href{https://doi.org/10.1038/ncomms8842}{10.1038/ncomms8842}
 
 \hypertarget{ref-Grilli2017}{}
-5. J. Grilli \emph{et al.}, Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications}. \textbf{8} (2017),
-doi:\href{https://doi.org/10.1038/ncomms14389}{10.1038/ncomms14389}.
+5. Grilli J, Adorisio M, Suweis S, Barabás G, Banavar JR, Allesina S, et
+al. Feasibility and coexistence of large ecological communities. Nature
+Communications. 2017;8.
+doi:\href{https://doi.org/10.1038/ncomms14389}{10.1038/ncomms14389}
 
 \hypertarget{ref-Gray2008}{}
-6. R. T. Gray, P. A. Robinson, Stability and synchronization of random
-brain networks with a distribution of connection strengths.
-\emph{Neurocomputing}. \textbf{71}, 1373--1387 (2008).
+6. Gray RT, Robinson PA. Stability and synchronization of random brain
+networks with a distribution of connection strengths. Neurocomputing.
+2008;71: 1373--1387.
+doi:\href{https://doi.org/10.1016/j.neucom.2007.06.002}{10.1016/j.neucom.2007.06.002}
 
 \hypertarget{ref-Gray2009}{}
-7. R. T. Gray, P. A. Robinson, Stability of random brain networks with
-excitatory and inhibitory connections. \emph{Neurocomputing}.
-\textbf{72}, 1849--1858 (2009).
+7. Gray RT, Robinson PA. Stability of random brain networks with
+excitatory and inhibitory connections. Neurocomputing. 2009;72:
+1849--1858.
+doi:\href{https://doi.org/10.1016/j.neucom.2008.06.001}{10.1016/j.neucom.2008.06.001}
 
 \hypertarget{ref-Rosenfeld2009}{}
-8. S. Rosenfeld, Patterns of stochastic behavior in dynamically unstable
-high-dimensional biochemical networks. \emph{Gene Regulation and Systems
-Biology}. \textbf{3}, 1--10 (2009).
+8. Rosenfeld S. Patterns of stochastic behavior in dynamically unstable
+high-dimensional biochemical networks. Gene Regulation and Systems
+Biology. 2009;3: 1--10.
 
 \hypertarget{ref-MacArthur2010}{}
-9. B. D. MacArthur, R. J. Sanchez-Garcia, A. Ma'ayan, Microdynamics and
-criticality of adaptive regulatory networks. \emph{Physics Review
-Letters}. \textbf{104}, 168701 (2010).
+9. MacArthur BD, Sanchez-Garcia RJ, Ma'ayan A. Microdynamics and
+criticality of adaptive regulatory networks. Physics Review Letters.
+2010;104: 168701.
+doi:\href{https://doi.org/10.1002/ana.22528.Toll-like}{10.1002/ana.22528.Toll-like}
 
 \hypertarget{ref-May2008}{}
-10. R. M. May, S. A. Levin, G. Sugihara, Complex systems: Ecology for
-bankers. \emph{Nature}. \textbf{451}, 893--895 (2008).
+10. May RM, Levin SA, Sugihara G. Complex systems: Ecology for bankers.
+Nature. 2008;451: 893--895.
+doi:\href{https://doi.org/10.1038/451893a}{10.1038/451893a}
 
 \hypertarget{ref-Haldane2011}{}
-11. A. G. Haldane, R. M. May, Systemic risk in banking ecosystems.
-\emph{Nature}. \textbf{469}, 351--355 (2011).
+11. Haldane AG, May RM. Systemic risk in banking ecosystems. Nature.
+Nature Publishing Group; 2011;469: 351--355.
+doi:\href{https://doi.org/10.1038/nature09659}{10.1038/nature09659}
 
 \hypertarget{ref-Suweis2014}{}
-12. S. Suweis, P. D'Odorico, Early warning signs in social-ecological
-networks. \emph{PLoS ONE}. \textbf{9} (2014),
-doi:\href{https://doi.org/10.1371/journal.pone.0101851}{10.1371/journal.pone.0101851}.
+12. Suweis S, D'Odorico P. Early warning signs in social-ecological
+networks. PLoS ONE. 2014;9.
+doi:\href{https://doi.org/10.1371/journal.pone.0101851}{10.1371/journal.pone.0101851}
 
 \hypertarget{ref-Bardoscia2017}{}
-13. M. Bardoscia, S. Battiston, F. Caccioli, G. Caldarelli, Pathways
-towards instability in financial networks. \emph{Nature Communications}.
-\textbf{8}, 1--7 (2017).
+13. Bardoscia M, Battiston S, Caccioli F, Caldarelli G. Pathways towards
+instability in financial networks. Nature Communications. Nature
+Publishing Group; 2017;8: 1--7.
+doi:\href{https://doi.org/10.1038/ncomms14416}{10.1038/ncomms14416}
 
 \hypertarget{ref-Tao2010}{}
-14. T. Tao, V. Vu, Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability}. \textbf{38}, 2023--2065
-(2010).
+14. Tao T, Vu V. Random matrices: Universality of ESDs and the circular
+law. Annals of Probability. 2010;38: 2023--2065.
+doi:\href{https://doi.org/10.1214/10-AOP534}{10.1214/10-AOP534}
 
 \hypertarget{ref-Patel2018}{}
-15. S. Patel, M. H. Cortez, S. J. Schreiber, Partitioning the effects of
-eco-evolutionary feedbacks on community stability. \emph{American
-Naturalist}. \textbf{191}, 1--29 (2018).
+15. Patel S, Cortez MH, Schreiber SJ. Partitioning the effects of
+eco-evolutionary feedbacks on community stability. American Naturalist.
+2018;191: 1--29.
+doi:\href{https://doi.org/10.1101/104505}{10.1101/104505}
 
 \hypertarget{ref-Ahmadian2015}{}
-16. Y. Ahmadian, F. Fumarola, K. D. Miller, Properties of networks with
-partially structured and partially random connectivity. \emph{Physical
-Review E - Statistical, Nonlinear, and Soft Matter Physics}.
-\textbf{91}, 012820 (2015).
+16. Ahmadian Y, Fumarola F, Miller KD. Properties of networks with
+partially structured and partially random connectivity. Physical Review
+E - Statistical, Nonlinear, and Soft Matter Physics. 2015;91: 012820.
+doi:\href{https://doi.org/10.1103/PhysRevE.91.012820}{10.1103/PhysRevE.91.012820}
 
 \hypertarget{ref-Gibbs2017}{}
-17. T. Gibbs, J. Grilli, T. Rogers, S. Allesina, The effect of
-population abundances on the stability of large random ecosystems.
-\emph{arXiv} (2017) (available at
-\url{http://arxiv.org/abs/1708.08837}).
+17. Gibbs T, Grilli J, Rogers T, Allesina S. The effect of population
+abundances on the stability of large random ecosystems. arXiv. 2017;
+Available: \url{http://arxiv.org/abs/1708.08837}
 
 \hypertarget{ref-Stone2017}{}
-18. L. Stone, The feasibility and stability of large complex biological
-networks: a random matrix approach. \emph{Scientific Reports}.
-\textbf{8}, 8246 (2018).
+18. Stone L. The feasibility and stability of large complex biological
+networks: a random matrix approach. Scientific Reports. 2018;8: 8246.
+doi:\href{https://doi.org/10.1101/223651}{10.1101/223651}
 
 \hypertarget{ref-Hamblin2013}{}
-19. S. Hamblin, On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution}. \textbf{4},
-184--194 (2013).
+19. Hamblin S. On the practical usage of genetic algorithms in ecology
+and evolution. Methods in Ecology and Evolution. 2013;4: 184--194.
+doi:\href{https://doi.org/10.1111/2041-210X.12000}{10.1111/2041-210X.12000}
 
 \hypertarget{ref-Dougoud2018}{}
-20. M. Dougoud, L. Vinckenbosch, R. Rohr, L.-F. Bersier, C. Mazza, The
+20. Dougoud M, Vinckenbosch L, Rohr R, Bersier L-F, Mazza C. The
 feasibility of equilibria in large ecosystems: a primary but neglected
-concept in the complexity-stability debate. \emph{PLOS Computational
-Biology}. \textbf{14}, e1005988 (2018).
+concept in the complexity-stability debate. PLOS Computational Biology.
+2018;14: e1005988. Available: \url{http://arxiv.org/abs/1612.06735}
 
 \hypertarget{ref-Song2018}{}
-21. C. Song, S. Saavedra, Will a small randomly assembled community be
-feasible and stable? \emph{Ecology}. \textbf{99}, 743--751 (2018).
+21. Song C, Saavedra S. Will a small randomly assembled community be
+feasible and stable? Ecology. 2018;99: 743--751.
+doi:\href{https://doi.org/10.1002/ecy.2125}{10.1002/ecy.2125}
 
 \hypertarget{ref-May1973}{}
-22. R. M. May, Qualitative stability in model ecosystems.
-\emph{Ecology}. \textbf{54}, 638--641 (1973).
+22. May RM. Qualitative stability in model ecosystems. Ecology. 1973;54:
+638--641. doi:\href{https://doi.org/10.2307/1935352}{10.2307/1935352}
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.734 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.725 and 0.741 for a and b,
 respectively.

### 2018-08-15 notebook/ms.Rmd
Updates for PLoS Comp Biol -- new Servan et al reference add

@@
 
-<!---
-
-**The stability of a complex system generally decreases with increasing system size, as is demonstrated by random matrix theory [@May1972; @Allesina2012].
-This counter-intuitive result, first shown by May [@May1972], is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010] and socio-economic [@Haldane2011; @Suweis2014; @Bardoscia2017] networks.
-Much attention has especially been given to the stability of ecological communities such as food webs or mutualist networks, with recent work investigating how different community structures affect stability [@Allesina2012; @Mougi2012; @Allesina2015a; @Gao2016; @Grilli2017; @Patel2018].
-But more broadly, stabilising mechanisms in complex systems remain under-developed, and the effect of variation in the response rate of individual system components remains an open problem [@Allesina2015].
-Here I show that when components of a complex system respond to system perturbation at different rates ($\boldsymbol{\gamma}$), the potential for system stability is markedly increased.
-Stability is caused by the clustering of some eigenvalues toward the centre of eigenvalue distributions despite the destabilising effect of higher interaction strength variation ($\boldsymbol{\sigma^{2}}$).
-This effect of variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, to the extent that the largest stable complex systems would otherwise be unstable if not for $\boldsymbol{Var(\gamma)}$.
-My results therefore reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
-Future research in complex systems should therefore account for the varying response rates of individual system components when assessing whole system stability.**
-
---->
-
 Introduction
@@
 While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities [@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
-Feasibility was unaffected by $Var(\gamma)$, meaning that for pure interacting species networks, variation in component response rate (i.e., species generation time) does not affect stability at biologically realistic species densities.
+Feasibility was unaffected by $Var(\gamma)$ and instead occurred with a fixed probability of $1/2^{S}$, consistent with a recent proof by Serv&aacute;n et al.
+[@Servan2018] (see Supplementary Information).
+Hence, for pure interacting species networks, variation in component response rate (i.e., species generation time) does not affect stability at biologically realistic species densities.
 
@@
 Interestingly, while complex systems were more likely to be stable given variation in component response rate, they were not more likely to be feasible, meaning that stability was not increased when component values were also restricted to being positive at equilibrium.
-Feasibility is important to consider, particularly for the study of ecological networks of species [@Stone2017; @Grilli2017; @Dougoud2018] because population densities cannot realistically be negative.
-My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information).
+Feasibility is important to consider, particularly for the study of ecological networks of species [@Stone2017; @Grilli2017; @Dougoud2018; @Servan2018] because population densities cannot realistically be negative.
+My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the feasibility of purely multi-species interaction networks (see also Supplementary Information).
 Nevertheless, ecological interactions do not exist in isolation in empirical systems [@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems.
@@
 
-<!--- USE J INSTEAD TO CLARIFY M --->
-
 **Component response rate variation ($\mathbf{\gamma}$)**.
 In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
-model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$)[@Patel2018].
+model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$) [@Patel2018].
 For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
@@
 
-In matrix notation[@May1973],
+In matrix notation [@May1973],
 
@@
 Dougoud et al.
-[@Dougoud2018] define the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
+[@Dougoud2018] identify the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
 
@@
 Implicitly, here $\delta = 0$ underlying strong interactions.
-Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's[@May1972] stability criterion [@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
-Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below,
+Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's [@May1972] stability criterion [@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
+Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below (see also [@Servan2018]),
 
@@
 Feasibility is satisfied if all of the elements of the resulting vector are positive.
-
-<!---
-
-If more discussion is needed, then talk about relation to Jirsa and Ding 2004.
-
---->
 

### 2018-08-15 notebook/ms.tex
Updates for PLoS Comp Biol -- new Servan et al reference add

@@
 mutualist, predator-prey; see Supplementary Information). Feasibility
-was unaffected by \(Var(\gamma)\), meaning that for pure interacting
-species networks, variation in component response rate (i.e., species
-generation time) does not affect stability at biologically realistic
-species densities.
+was unaffected by \(Var(\gamma)\) and instead occurred with a fixed
+probability of \(1/2^{S}\), consistent with a recent proof by Serván et
+al. {[}\protect\hyperlink{ref-Servan2018}{22}{]} (see Supplementary
+Information). Hence, for pure interacting species networks, variation in
+component response rate (i.e., species generation time) does not affect
+stability at biologically realistic species densities.
 
@@
 of species
-{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20}{]}
+{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}{]}
 because population densities cannot realistically be negative. My
@@
 responses to perturbation (e.g., due to differences in generation time
-among species) is unlikely to be critical to the stability of purely
+among species) is unlikely to be critical to the feasibility of purely
 multi-species interaction networks (see also Supplementary Information).
@@
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
-traits (\(\mathbf{x}\)){[}\protect\hyperlink{ref-Patel2018}{15}{]}. For
+traits (\(\mathbf{x}\)) {[}\protect\hyperlink{ref-Patel2018}{15}{]}. For
 any species \(i\) or trait \(j\), change in species density (\(N_{i}\))
@@
 components. Following May
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{22}{]},
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}{]},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation{[}\protect\hyperlink{ref-May1973}{22}{]},
+In matrix notation {[}\protect\hyperlink{ref-May1973}{23}{]},
 
@@
 components and can be analysed using the techniques of May
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{22}{]}.
+{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}{]}.
 
@@
 \textbf{System feasibility}. Dougoud et al.
-{[}\protect\hyperlink{ref-Dougoud2018}{20}{]} define the following
+{[}\protect\hyperlink{ref-Dougoud2018}{20}{]} identify the following
 feasibility criteria for ecological systems characterised by \(S\)
@@
 affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
-relates to May's{[}\protect\hyperlink{ref-May1972}{1}{]} stability
+relates to May's {[}\protect\hyperlink{ref-May1972}{1}{]} stability
 criterion {[}\protect\hyperlink{ref-Dougoud2018}{20}{]} by
@@
 \(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below,
+reduced to the below (see also
+{[}\protect\hyperlink{ref-Servan2018}{22}{]}),
 
@@
 
+\hypertarget{ref-Servan2018}{}
+22. Serván CA, Capitán JA, Grilli J, Morrison KE, Allesina S.
+Coexistence of many species in random ecosystems. Nature Ecology and
+Evolution. 2018;2: 1237--1242.
+doi:\href{https://doi.org/10.1038/s41559-018-0603-6}{10.1038/s41559-018-0603-6}
+
 \hypertarget{ref-May1973}{}
-22. May RM. Qualitative stability in model ecosystems. Ecology. 1973;54:
+23. May RM. Qualitative stability in model ecosystems. Ecology. 1973;54:
 638--641. doi:\href{https://doi.org/10.2307/1935352}{10.2307/1935352}
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.725 and 0.741 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.721 and 0.739 for a and b,
 respectively.

### 2018-10-20 notebook/ms.Rmd
Add an idea about resilience

@@
 https://gist.github.com/JJ/ef9d3d8f1142df064bd5
+--->
+
+<!---
+
+NOTE: Consider looking at the resilience of the stable networks given $\gamma = 1$ and $Var(\gamma)$, which is defined by the inverse of the leading eigenvalue.
+The inverse of the real part of the leading eigenvalue indicates how quickly the system returns to equilibrium (resilience).
+See @Landi2018 for more.
+Intuitively, resilience should perhaps increase a bit if we're only looking at the subset of systems that are stable (many of which will have had their eigenvalues shifted left)?
+
+Landi, P., Minoarivelo, H.
+O., Brännström, Å., Hui, C., & Dieckmann, U.
+(2018).
+Complexity and stability of ecological networks: A review of the theory.
+Population Ecology, 0(0), In press.
+https://doi.org/10.1007/s10144-018-0628-3
+
 --->

### 2019-01-18 notebook/ms.Rmd
More setup before writing

@@
 
-Given the larger space of Nature Communications, one new figure will be used in the same style as that of Figure 3; the new Figure 4 will show the same results, but for sultions found with the genetic algorithm.
+1.
+As above, check out the change in resilience defined as the difference beween the inverse of the real part of the leading eigenvalue in $Var(\gamma)$ versus $\gamma = 1$.
+Initial simulaitons suggest that resilience is increased on average by $Var(\gamma)$, and increasingly so as S increases.
 
-Because Nature communcations is online, change all non-standard gamma results in figures to Red. The new genetic algorithm results will be in blue.
+2.
+Check for induced correlations causing stabilisation (and E1, E2, and EC, more generally), as per Reviewer 2 specific comment 3.
+If possible, give an analytical explanation for the increase in stability that has to do with the error inducing changes in correlations.
 

### 2019-01-18 notebook/PLoS_Compu_reviews_code.R

@@ -39,3 +39,78 @@ while(iter > 0){
 }
 unstable_A0 <- A0_mx[A0_mx > 0];
 unstable_A1 <- A1_mx[A0_mx > 0];
+
+
+
+
+# Reviewer 2 specific comment 3
+Ec_calc <- function(A0, E1, E2){
+    S   <- dim(A0)[1];
+    smm <- 0;
+    for(i in 1:S){
+        for(j in 1:S){
+            if(i != j){
+                smm <- smm + A0[i,j] * A0[j, i] - (E1*E1 / E2*E2);
+            }
+        }
+    }
+    EC <- (1 / S*(S - 1) * E2*E2) * smm;
+    return(EC);
+}
+#----------------------------
+S         <- 28;
+A0rE1     <- NULL;
+A0rE2     <- NULL;
+A0rEC     <- NULL;
+A1rE1     <- NULL;
+A1rE2     <- NULL;
+A1rEC     <- NULL;
+A0_eig    <- NULL;
+A1_eig    <- NULL;
+iter      <- 10000;
+while(iter > 0){
+    r_vec     <- rnorm(n =S, mean = 0, sd = 0.4);
+    A0_dat    <- rnorm(n =S *S, mean = 0, sd = sigma);
+    A0        <- matrix(data = A0_dat, nrow =S, 
+                        ncol =S);
+    C_dat     <- rbinom(n =S *S, size = 1, prob = C);
+    C_mat     <- matrix(data = C_dat, nrow =S, 
+                        ncol =S);
+    A0        <- A0 * C_mat;
+    diag(A0)  <- -1;
+    gam1      <- runif(n =S, min = 0, max = 2);
+    A1        <- A0 * gam1;
+    A0        <- A0 * mean(gam1);
+    A0_stb    <- max(Re(eigen(A0)$values));
+    A1_stb    <- max(Re(eigen(A1)$values));
+    Ad0       <- A0;
+    diag(Ad0) <- NA;
+    Ad1       <- A1;
+    diag(Ad1) <- NA;
+    A0E1      <- mean(Ad0, na.rm = TRUE);
+    A0E2      <- var(as.vector(Ad0), na.rm = TRUE);
+    A0EC      <- Ec_calc(A0, E1, E2);
+    A1E1      <- mean(Ad1, na.rm = TRUE);
[…]

### 2019-02-22 notebook/PLoS_Compu_reviews_code.R
Update review

@@ -14,16 +14,18 @@ eg_run <- SD_mn_row_vals(S = 50);
 
 
 # Reviewer 2 specific comment 4
-S       <- 30;
+S       <- 5;
 A0_mx   <- NULL;
 A1_mx   <- NULL;
 iter    <- 10000;
+sigma   <- 0.4;
+C       <- 1;
 while(iter > 0){
-    r_vec    <- rnorm(n =S, mean = 0, sd = 0.4);
-    A0_dat   <- rnorm(n =S *S, mean = 0, sd = sigma);
+    r_vec    <- rnorm(n = S, mean = 0, sd = sigma);
+    A0_dat   <- rnorm(n = S * S, mean = 0, sd = sigma);
     A0       <- matrix(data = A0_dat, nrow =S, 
-                       ncol =S);
-    C_dat    <- rbinom(n =S *S, size = 1, prob = C);
+                       ncol = S);
+    C_dat    <- rbinom(n = S * S, size = 1, prob = C);
     C_mat    <- matrix(data = C_dat, nrow =S, 
                        ncol =S);
     A0       <- A0 * C_mat;

### 2019-02-22 notebook/ms.Rmd
Change the wording

@@
 ---
-title: "Component response rate variation drives stability in large complex systems"
+title: "Component response rate variation underlies stability in large complex systems"
 author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
@@
 Variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, such that the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
-My results reveal a previously unconsidered driver of system stability that is likely to be pervasive across all complex systems.
+My results reveal a previously unconsidered aspect of system stability that is likely to be pervasive across all complex systems.
 

### 2019-02-22 notebook/ms.Rmd
Add some to the abstract to emphasise finite systems

@@
 output:
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
+  html_document: default
   word_document:
@@
     - --csl
-    - science.csl
+    - nature.csl
     reference_docx: docx_template.docx
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-  html_document: default
 bibliography: references.bib
@@
 linkcolor: blue
-csl: plos.csl
+csl: nature.csl
 biblio-style: apalike
@@
 I show that when component response rates vary, the potential for system stability is markedly increased.
-Variation in $\boldsymbol{\gamma}$ becomes increasingly important as system size increases, such that the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
-My results reveal a previously unconsidered aspect of system stability that is likely to be pervasive across all complex systems.
+Variation in $\boldsymbol{\gamma}$ becomes increasingly important for stability in large but finite systems, where the probability of stability would otherwise be negligible.
+At such extremes of simulated system size, the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
+My results therefore reveal a previously unconsidered aspect of system stability that is likely to be pervasive across all realistic complex systems.
 
@@
 May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks.
-As such, identifying general principles that drive stability in complex systems is of wide-ranging importance.
+As such, identifying general principles that affect stability in complex systems is of wide-ranging importance.
 
@@
 
-May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that individual components respond to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rate of individual components to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, complex systems in which $\gamma$ varies are actually more likely to be stable, especially when $S$ is high.
+Here I show why, despite higher $\sigma$, realistic complex systems (where $S$ is high but finite) in which $\gamma$ varies are actually more likely to be stable.
+<!--- Need to emphasise that this is for *real* systems at the thresholds of stability --->
 

### 2019-02-22 notebook/ms.tex
Add some to the abstract to emphasise finite systems

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation drives stability in large complex systems},
+            pdftitle={Component response rate variation underlies stability in large complex systems},
             pdfauthor={A. Bradley Duthie ( alexander.duthie@stir.ac.uk )},
@@
 \setlength{\droptitle}{-2em}
-  \title{Component response rate variation drives stability in large complex
+
+  \title{Component response rate variation underlies stability in large complex
 systems}
@@
 unstable if not for \(\boldsymbol{Var(\gamma)}\). My results reveal a
-previously unconsidered driver of system stability that is likely to be
+previously unconsidered aspect of system stability that is likely to be
 pervasive across all complex systems.
@@
 
-In 1972, May {[}\protect\hyperlink{ref-May1972}{1}{]} first demonstrated
-that randomly assembled systems of sufficient complexity are almost
-inevitably unstable given infinitesimally small perturbations.
-Complexity in this case is defined by the size of the system (i.e., the
-number of potentially interacting components; \(S\)), its connectance
-(i.e., the probability that one component will interact with another;
-\(C\)), and the variance of interaction strengths (\(\sigma^{2}\))
-{[}\protect\hyperlink{ref-Allesina2012}{2}{]}. May's finding that the
-probability of local stability falls to near zero given a sufficiently
-high threshold of \(\sigma\sqrt{SC}\) is broadly relevant for
-understanding the dynamics and persistence of systems such as ecological
-{[}\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}{]},
-neurological
-{[}\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}{]},
-biochemical
-{[}\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}{]},
-and socio-economic
-{[}\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}{]}
-networks. As such, identifying general principles that drive stability
+In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
+demonstrated that randomly assembled systems of sufficient complexity
+are almost inevitably unstable given infinitesimally small
+perturbations. Complexity in this case is defined by the size of the
+system (i.e., the number of potentially interacting components; \(S\)),
+its connectance (i.e., the probability that one component will interact
+with another; \(C\)), and the variance of interaction strengths
+(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
+May's finding that the probability of local stability falls to near zero
+given a sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly
+relevant for understanding the dynamics and persistence of systems such
+as
+ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
+neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
+biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
+and
+socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
+networks. As such, identifying general principles that affect stability
 in complex systems is of wide-ranging importance.
@@
 matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
-species {[}\protect\hyperlink{ref-Allesina2012}{2}{]} or banks
-{[}\protect\hyperlink{ref-Haldane2011}{11}{]}). One element of such a
-matrix, \(M_{ij}\), defines how component \(j\) affects component \(i\)
-in the system at a point of equilibrium
-{[}\protect\hyperlink{ref-Allesina2012}{2}{]}. Off-diagonal elements
-(\(i \neq j\)) therefore define interactions between components, while
-diagonal elements (\(i = j\)) define component self-regulation (e.g.,
-carrying capacity in ecological communities). Traditionally,
-off-diagonal elements are assigned non-zero values with a probability
-\(C\), which are sampled from a distribution with variance
-\(\sigma^{2}\); diagonal elements are set to -1
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}{]}.
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
+banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
+element of such a matrix, \(M_{ij}\), defines how component \(j\)
+affects component \(i\) in the system at a point of
+equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
+Off-diagonal elements (\(i \neq j\)) therefore define interactions
+between components, while diagonal elements (\(i = j\)) define component
+self-regulation (e.g., carrying capacity in ecological communities).
+Traditionally, off-diagonal elements are assigned non-zero values with a
+probability \(C\), which are sampled from a distribution with variance
+\(\sigma^{2}\); diagonal elements are set to
+-1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
 Local system stability is assessed using eigenanalysis, with the system
 being stable if the real parts of all eigenvalues (\(\lambda\)) of
-\(\mathbf{M}\) are negative (\(\max\left(\Re(\lambda)\right) < 0\))
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}.
-In a large system (high \(S\)), eigenvalues are distributed uniformly
-{[}\protect\hyperlink{ref-Tao2010}{14}{]} within a circle centred at
-\(\Re = -1\) (the mean value of diagonal elements) and \(\Im = 0\), with
-a radius of \(\sigma\sqrt{SC}\)
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}{]}
+\(\mathbf{M}\) are negative
+(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
+In a large system (high \(S\)), eigenvalues are distributed
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
+circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
+\(\Im = 0\), with a radius of
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
@@
 
-May's
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
 stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
@@
 time can be modelled by multiplying all row elements by a real scalar
-value \(\gamma_{i}\) {[}\protect\hyperlink{ref-Patel2018}{15}{]}. The
-distribution of \(\gamma\) over \(S\) components thereby models the
+value
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+The distribution of \(\gamma\) over \(S\) components thereby models the
 distribution of component response rates. An instructive example
@@
 distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected {[}\protect\hyperlink{ref-Tao2010}{14}{]}, when \(\gamma = 1\),
-eigenvalues are distributed uniformly in a circle centred at (\(-1, 0\))
-with a radius of \(\sigma\sqrt{SC} =\) 2.828. Uniform variation in
-\(\gamma\) leads to a non-uniform distribution of eigenvalues
-{[}\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}{]},
+expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
+\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
+centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
+Uniform variation in \(\gamma\) leads to a non-uniform distribution of
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
 some of which are clustered locally near the centre of the distribution,
@@
 complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above May's
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}{]}
+can often lead to stability above
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
 threshold \(\sigma\sqrt{SC} < 1\).
@@
 values leading to stability in a large complex system) is too large to
-search exhaustively {[}\protect\hyperlink{ref-Hamblin2013}{19}{]}.
+search
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
 Generations of selection on \(\gamma\) value combinations to minimise
@@
 the system. Feasibilility assumes that the system is stable and values
-of all components are positive at equilibrium
-{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}{]}.
+of all components are positive at
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
 This is of particular interest for ecological communities because
 population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically realistic
-{[}\protect\hyperlink{ref-Dougoud2018}{20}{]}. While my results are
-intended to be general to all complex systems, and not restricted to
-species networks, I have also performed a feasibility analysis on all
-matrices \(\mathbf{M}\) tested for stability, and additionally for
-specific types of ecological communities
-{[}\protect\hyperlink{ref-Allesina2012}{2}{]} (e.g., competitive,
-mutualist, predator-prey; see Supplementary Information). Feasibility
-was unaffected by \(Var(\gamma)\) and instead occurred with a fixed
-probability of \(1/2^{S}\), consistent with a recent proof by Serván et
-al. {[}\protect\hyperlink{ref-Servan2018}{22}{]} (see Supplementary
-Information). Hence, for pure interacting species networks, variation in
-component response rate (i.e., species generation time) does not affect
-stability at biologically realistic species densities.
+systems need to be feasible for stability to be biologically
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
+While my results are intended to be general to all complex systems, and
+not restricted to species networks, I have also performed a feasibility
+analysis on all matrices \(\mathbf{M}\) tested for stability, and
+additionally for specific types of ecological
+communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+(e.g., competitive, mutualist, predator-prey; see Supplementary
+Information). Feasibility was unaffected by \(Var(\gamma)\) and instead
+occurred with a fixed probability of \(1/2^{S}\), consistent with a
+recent proof by Serván et
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}} (see
+Supplementary Information). Hence, for pure interacting species
+networks, variation in component response rate (i.e., species generation
+time) does not affect stability at biologically realistic species
+densities.
 
@@
 size increases, and can ultimately increase the overall probability that
-system stability is observed above that predicted by May's
-{[}\protect\hyperlink{ref-May1972}{1}{]} classically derived
-\(\sigma \sqrt{SC}\) criterion. The logic outlined here is general, and
-potentially applies to any complex system in which individual system
-components can vary in their reaction rates to system perturbation.
+system stability is observed above that predicted by
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
+derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
+general, and potentially applies to any complex system in which
+individual system components can vary in their reaction rates to system
+perturbation.
 
@@
 also increase the real components of eigenvalues. The mathematics
-underlying this shift in eigenvalue distribution has been investigated
-{[}\protect\hyperlink{ref-Ahmadian2015}{16}{]} and recently applied to
-questions concerning species density and feasibility
-{[}\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}{]},
+underlying this shift in eigenvalue distribution has been
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+and recently applied to questions concerning species density and
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
 but has not been interpreted as rates of response of individual system
@@
 important to consider, particularly for the study of ecological networks
-of species
-{[}\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}{]}
+of
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}}
 because population densities cannot realistically be negative. My
@@
 Nevertheless, ecological interactions do not exist in isolation in
-empirical systems {[}\protect\hyperlink{ref-Patel2018}{15}{]}, but
+empirical
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
 instead interact with evolutionary, abiotic, or social-economic systems.
@@
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
-traits (\(\mathbf{x}\)) {[}\protect\hyperlink{ref-Patel2018}{15}{]}. For
-any species \(i\) or trait \(j\), change in species density (\(N_{i}\))
-or trait value (\(x_{j}\)) with time (\(t\)) is a function of the
-vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+traits
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+For any species \(i\) or trait \(j\), change in species density
+(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
 
@@
 were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species densities
-{[}\protect\hyperlink{ref-Patel2018}{15}{]}, and this is modulated in
-the above by the scalar \(\epsilon\). The value of \(\epsilon\) thereby
-determines the timescale separation between ecology and evolution, with
-high \(\epsilon\) modelling relatively fast evolution and low
-\(\epsilon\) modelling relative slow evolution
-{[}\protect\hyperlink{ref-Patel2018}{15}{]}.
+slow or fast in comparison with the change in species
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
+this is modulated in the above by the scalar \(\epsilon\). The value of
+\(\epsilon\) thereby determines the timescale separation between ecology
+and evolution, with high \(\epsilon\) modelling relatively fast
+evolution and low \(\epsilon\) modelling relative slow
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
 
@@
 rate of evolution to modulate rates of component responses for \(S\)
-components. Following May
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}{]},
+components. Following
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation {[}\protect\hyperlink{ref-May1973}{23}{]},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{23}},
 
@@
 \(\mathbf{M} =\mathbf{\gamma} \mathbf{A}\) modulates the values of
-components and can be analysed using the techniques of May
-{[}\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}{]}.
+components and can be analysed using the techniques of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
 
@@
 genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological evolution
-{[}\protect\hyperlink{ref-Hamblin2013}{19}{]}. In this case, the
-practical solution is finding vectors of \(\mathbf{\gamma}\) that
-decrease the most positive real eigenvalue of \(\mathbf{M}\). The
-genetic algorithm used achieves this by initialising a large population
-of 1000 different potential \(\mathbf{\gamma}\) vectors and allowing
-this population to evolve through a process of mutation, crossover
-(swaping \(\gamma_{i}\) values between vectors), selection, and
-reproduction until either a \(\mathbf{\gamma}\) vector is found where
-all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is met.
+solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
+this case, the practical solution is finding vectors of
+\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
+\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
+a large population of 1000 different potential \(\mathbf{\gamma}\)
+vectors and allowing this population to evolve through a process of
+mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
+selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
+found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
+met.
 
@@
 
-\textbf{System feasibility}. Dougoud et al.
-{[}\protect\hyperlink{ref-Dougoud2018}{20}{]} identify the following
-feasibility criteria for ecological systems characterised by \(S\)
-interacting species with varying densities in a classical Lotka-Volterra
-model,
+\textbf{System feasibility}. Dougoud et
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} identify
+the following feasibility criteria for ecological systems characterised
+by \(S\) interacting species with varying densities in a classical
+Lotka-Volterra model,
 
@@
 affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
-relates to May's {[}\protect\hyperlink{ref-May1972}{1}{]} stability
-criterion {[}\protect\hyperlink{ref-Dougoud2018}{20}{]} by
+relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+stability
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 \(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below (see also
-{[}\protect\hyperlink{ref-Servan2018}{22}{]}),
+reduced to the below (see
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),
 
@@
 \hypertarget{ref-May1972}{}
-1. May RM. Will a large complex system be stable? Nature. 1972;238:
-413--414.
+1. May, R. M. Will a large complex system be stable? \emph{Nature}
+\textbf{238,} 413--414 (1972).
 
 \hypertarget{ref-Allesina2012}{}
-2. Allesina S, Tang S. Stability criteria for complex ecosystems.
-Nature. Nature Publishing Group; 2012;483: 205--208.
-doi:\href{https://doi.org/10.1038/nature10832}{10.1038/nature10832}
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+\emph{Nature} \textbf{483,} 205--208 (2012).
 
 \hypertarget{ref-Mougi2012}{}
-3. Mougi A, Kondoh M. Diversity of interaction types and ecological
-community stability. Science. 2012;337: 349--351.
-doi:\href{https://doi.org/10.1126/science.1220529}{10.1126/science.1220529}
+3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
+community stability. \emph{Science} \textbf{337,} 349--351 (2012).
 
 \hypertarget{ref-Allesina2015}{}
-4. Allesina S, Grilli J, Barabás G, Tang S, Aljadeff J, Maritan A.
-Predicting the stability of large structured food webs. Nature
-Communications. 2015;6: 7842.
-doi:\href{https://doi.org/10.1038/ncomms8842}{10.1038/ncomms8842}
+4. Allesina, S. \emph{et al.} Predicting the stability of large
+structured food webs. \emph{Nature Communications} \textbf{6,} 7842
+(2015).
 
 \hypertarget{ref-Grilli2017}{}
-5. Grilli J, Adorisio M, Suweis S, Barabás G, Banavar JR, Allesina S, et
-al. Feasibility and coexistence of large ecological communities. Nature
-Communications. 2017;8.
-doi:\href{https://doi.org/10.1038/ncomms14389}{10.1038/ncomms14389}
+5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
+ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
 
 \hypertarget{ref-Gray2008}{}
-6. Gray RT, Robinson PA. Stability and synchronization of random brain
-networks with a distribution of connection strengths. Neurocomputing.
-2008;71: 1373--1387.
-doi:\href{https://doi.org/10.1016/j.neucom.2007.06.002}{10.1016/j.neucom.2007.06.002}
+6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
+random brain networks with a distribution of connection strengths.
+\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).
 
 \hypertarget{ref-Gray2009}{}
-7. Gray RT, Robinson PA. Stability of random brain networks with
-excitatory and inhibitory connections. Neurocomputing. 2009;72:
-1849--1858.
-doi:\href{https://doi.org/10.1016/j.neucom.2008.06.001}{10.1016/j.neucom.2008.06.001}
+7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
+with excitatory and inhibitory connections. \emph{Neurocomputing}
+\textbf{72,} 1849--1858 (2009).
 
 \hypertarget{ref-Rosenfeld2009}{}
-8. Rosenfeld S. Patterns of stochastic behavior in dynamically unstable
-high-dimensional biochemical networks. Gene Regulation and Systems
-Biology. 2009;3: 1--10.
+8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
+high-dimensional biochemical networks. \emph{Gene Regulation and Systems
+Biology} \textbf{3,} 1--10 (2009).
 
 \hypertarget{ref-MacArthur2010}{}
-9. MacArthur BD, Sanchez-Garcia RJ, Ma'ayan A. Microdynamics and
-criticality of adaptive regulatory networks. Physics Review Letters.
-2010;104: 168701.
-doi:\href{https://doi.org/10.1002/ana.22528.Toll-like}{10.1002/ana.22528.Toll-like}
+9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
+and criticality of adaptive regulatory networks. \emph{Physics Review
+Letters} \textbf{104,} 168701 (2010).
 
 \hypertarget{ref-May2008}{}
-10. May RM, Levin SA, Sugihara G. Complex systems: Ecology for bankers.
-Nature. 2008;451: 893--895.
-doi:\href{https://doi.org/10.1038/451893a}{10.1038/451893a}
+10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
+for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
 
 \hypertarget{ref-Haldane2011}{}
-11. Haldane AG, May RM. Systemic risk in banking ecosystems. Nature.
-Nature Publishing Group; 2011;469: 351--355.
-doi:\href{https://doi.org/10.1038/nature09659}{10.1038/nature09659}
+11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
+\emph{Nature} \textbf{469,} 351--355 (2011).
 
 \hypertarget{ref-Suweis2014}{}
-12. Suweis S, D'Odorico P. Early warning signs in social-ecological
-networks. PLoS ONE. 2014;9.
-doi:\href{https://doi.org/10.1371/journal.pone.0101851}{10.1371/journal.pone.0101851}
+12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
+networks. \emph{PLoS ONE} \textbf{9,} (2014).
 
 \hypertarget{ref-Bardoscia2017}{}
-13. Bardoscia M, Battiston S, Caccioli F, Caldarelli G. Pathways towards
-instability in financial networks. Nature Communications. Nature
-Publishing Group; 2017;8: 1--7.
-doi:\href{https://doi.org/10.1038/ncomms14416}{10.1038/ncomms14416}
+13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
+Pathways towards instability in financial networks. \emph{Nature
+Communications} \textbf{8,} 1--7 (2017).
 
 \hypertarget{ref-Tao2010}{}
-14. Tao T, Vu V. Random matrices: Universality of ESDs and the circular
-law. Annals of Probability. 2010;38: 2023--2065.
-doi:\href{https://doi.org/10.1214/10-AOP534}{10.1214/10-AOP534}
+14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
+circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
+(2010).
 
 \hypertarget{ref-Patel2018}{}
-15. Patel S, Cortez MH, Schreiber SJ. Partitioning the effects of
-eco-evolutionary feedbacks on community stability. American Naturalist.
-2018;191: 1--29.
-doi:\href{https://doi.org/10.1101/104505}{10.1101/104505}
+15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+effects of eco-evolutionary feedbacks on community stability.
+\emph{American Naturalist} \textbf{191,} 1--29 (2018).
 
 \hypertarget{ref-Ahmadian2015}{}
-16. Ahmadian Y, Fumarola F, Miller KD. Properties of networks with
-partially structured and partially random connectivity. Physical Review
-E - Statistical, Nonlinear, and Soft Matter Physics. 2015;91: 012820.
-doi:\href{https://doi.org/10.1103/PhysRevE.91.012820}{10.1103/PhysRevE.91.012820}
+16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+with partially structured and partially random connectivity.
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{91,} 012820 (2015).
 
 \hypertarget{ref-Gibbs2017}{}
-17. Gibbs T, Grilli J, Rogers T, Allesina S. The effect of population
-abundances on the stability of large random ecosystems. arXiv. 2017;
-Available: \url{http://arxiv.org/abs/1708.08837}
+17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{arXiv} (2017).
 
 \hypertarget{ref-Stone2017}{}
-18. Stone L. The feasibility and stability of large complex biological
-networks: a random matrix approach. Scientific Reports. 2018;8: 8246.
-doi:\href{https://doi.org/10.1101/223651}{10.1101/223651}
+18. Stone, L. The feasibility and stability of large complex biological
+networks: a random matrix approach. \emph{Scientific Reports}
+\textbf{8,} 8246 (2018).
 
 \hypertarget{ref-Hamblin2013}{}
-19. Hamblin S. On the practical usage of genetic algorithms in ecology
-and evolution. Methods in Ecology and Evolution. 2013;4: 184--194.
-doi:\href{https://doi.org/10.1111/2041-210X.12000}{10.1111/2041-210X.12000}
+19. Hamblin, S. On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+184--194 (2013).
 
 \hypertarget{ref-Dougoud2018}{}
-20. Dougoud M, Vinckenbosch L, Rohr R, Bersier L-F, Mazza C. The
-feasibility of equilibria in large ecosystems: a primary but neglected
-concept in the complexity-stability debate. PLOS Computational Biology.
-2018;14: e1005988. Available: \url{http://arxiv.org/abs/1612.06735}
+20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+The feasibility of equilibria in large ecosystems: a primary but
+neglected concept in the complexity-stability debate. \emph{PLOS
+Computational Biology} \textbf{14,} e1005988 (2018).
 
 \hypertarget{ref-Song2018}{}
-21. Song C, Saavedra S. Will a small randomly assembled community be
-feasible and stable? Ecology. 2018;99: 743--751.
-doi:\href{https://doi.org/10.1002/ecy.2125}{10.1002/ecy.2125}
+21. Song, C. \& Saavedra, S. Will a small randomly assembled community
+be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
 
 \hypertarget{ref-Servan2018}{}
-22. Serván CA, Capitán JA, Grilli J, Morrison KE, Allesina S.
-Coexistence of many species in random ecosystems. Nature Ecology and
-Evolution. 2018;2: 1237--1242.
-doi:\href{https://doi.org/10.1038/s41559-018-0603-6}{10.1038/s41559-018-0603-6}
+22. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+Allesina, S. Coexistence of many species in random ecosystems.
+\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).
 
 \hypertarget{ref-May1973}{}
-23. May RM. Qualitative stability in model ecosystems. Ecology. 1973;54:
-638--641. doi:\href{https://doi.org/10.2307/1935352}{10.2307/1935352}
+23. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+\textbf{54,} 638--641 (1973).
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.721 and 0.739 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.738 for a and b,
 respectively.

### 2019-02-22 notebook/ms.Rmd
First pass, fix some notation

@@
 
-May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rate of individual components to perturbations of the system at the same rate ($\gamma$), but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rates ($E[\gamma]$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, realistic complex systems (where $S$ is high but finite) in which $\gamma$ varies are actually more likely to be stable.
+Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) in which $\gamma$ varies are actually more likely to be stable.
 <!--- Need to emphasise that this is for *real* systems at the thresholds of stability --->
@@
 An instructive example compares one $\mathbf{M}$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $\mathbf{M}$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
-This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note $E[\gamma_{i}] = 1$ in both cases).
+This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note that mean $\gamma_{i} = 1$ in both cases).
 I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
@@
 As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
-1b), $E[\gamma] = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $\mathbf{M}$ before and after the addition of variation in component response rate.
+1b), mean $\gamma = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $\mathbf{M}$ before and after the addition of variation in component response rate.
 Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
@@
 For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
-3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$; i.e., the potential for $Var(\gamma)$ to drive stability increased with system complexity.
+3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
+In other words, the potential for $Var(\gamma)$ to drive stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
 For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$.
@@
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
-4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements.
+4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements (i.e., by manipulating component response rates rather than interactions between components).
 
@@
 Feasibilility assumes that the system is stable and values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
+This is of particular interest for ecological communities because population density ($N$) cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
 While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities [@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
+I want to emphasise that $\gamma$ is not interpreted as population density in this analysis, but instead as a fundamental property of species life history such as expected generation time.
 Feasibility was unaffected by $Var(\gamma)$ and instead occurred with a fixed probability of $1/2^{S}$, consistent with a recent proof by Serv&aacute;n et al.
@@
 Here I have shown that the stability of large systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
-Variation in component response rate becomes more likely to be critical for stability as system size increases, and can ultimately increase the overall probability that system stability is observed above that predicted by May's [@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
+Variation in component response rate is more likely to be critical for stability in systems that are especially large (but finite, see Gibbs et al.[@Gibbs2017]), and it can ultimately increase the probability that system stability is observed above that predicted by May's [@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
 The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
@@
 The mathematics underlying this shift in eigenvalue distribution has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual system components to perturbation.
+
+<!--- Also important to emphasise Gibbs result -- I'm doing this for finite systems, and I deliberate stressed the system complexity to the point of collapse. --->
 
@@
 
-$$\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.$$
+$$\mathbf{n^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.$$
 
-In the above, $\mathbf{x^{*}}$ is the vector of species densities at equilibrium.
-Feasibility is satisfied if all elements in $\mathbf{x^{*}}$ are positive.
+In the above, $\mathbf{n^{*}}$ is the vector of species densities at equilibrium.
+Feasibility is satisfied if all elements in $\mathbf{n^{*}}$ are positive.
 The matrix $\mathbf{I}$ is the identity matrix, and the value $\theta$ is the strength of intraspecific competition (diagonal elements).
@@
 
-$$\mathbf{x^{*} = -A^{-1}r}.$$
+$$\mathbf{n^{*} = -A^{-1}r}.$$
 

### 2019-02-22 notebook/ms.tex
First pass, fix some notation

@@
 potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) becomes increasingly important as system size
-increases, such that the largest stable complex systems would be
-unstable if not for \(\boldsymbol{Var(\gamma)}\). My results reveal a
-previously unconsidered aspect of system stability that is likely to be
-pervasive across all complex systems.
+\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
+large but finite systems, where the probability of stability would
+otherwise be negligible. At such extremes of simulated system size, the
+largest stable complex systems would be unstable if not for
+\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
+unconsidered aspect of system stability that is likely to be pervasive
+across all realistic complex systems.
 
@@
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual
-components respond to perturbations of the system at the same rate
-(\(\gamma\)), but this is highly unlikely in any complex system. In
-ecological communities, for example, the rate at which population
-density changes following perturbation will depend on the generation
-time of organisms, which might vary by orders of magnitude among
-species. Species with short generation times will respond quickly (high
-\(\gamma\)) to perturbations relative to species with long generation
-times (low \(\gamma\)). Similarly, the speed at which individual banks
-respond to perturbations in financial networks, or individuals or
-institutions respond to perturbations in complex social networks, is
-likely to vary. The effect of such variance on stability has not been
-investigated in complex systems theory. Intuitively, variation in
-\(\gamma\) might be expected to decrease system stability by introducing
-a new source of variation into the system and thereby increasing
-\(\sigma\). Here I show why, despite higher \(\sigma\), complex systems
-in which \(\gamma\) varies are actually more likely to be stable,
-especially when \(S\) is high.
+stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
+response rates (\(E[\gamma]\)) of individual components to perturbations
+of the system are identical, but this is highly unlikely in any complex
+system. In ecological communities, for example, the rate at which
+population density changes following perturbation will depend on the
+generation time of organisms, which might vary by orders of magnitude
+among species. Species with short generation times will respond quickly
+(high \(\gamma\)) to perturbations relative to species with long
+generation times (low \(\gamma\)). Similarly, the speed at which
+individual banks respond to perturbations in financial networks, or
+individuals or institutions respond to perturbations in complex social
+networks, is likely to vary. The effect of such variance on stability
+has not been investigated in complex systems theory. Intuitively,
+variation in \(\gamma\) might be expected to decrease system stability
+by introducing a new source of variation into the system and thereby
+increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
+realistic complex systems (such that \(S\) is high but finite) in which
+\(\gamma\) varies are actually more likely to be stable.
 
@@
 \(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
-are otherwise identical (note \(E[\gamma_{i}] = 1\) in both cases). I
-assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal
-elements are set to \(-1\) and non-zero off-diagonal elements are drawn
-randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied
-by \(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
-eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
-centred at (\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real
-components of eigenvalues are highly unlikely to all be negative when
-all \(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated
-into two groups, eigenvalues are no longer uniformly distributed (Fig.
-1b). Instead, two distinct clusters of eigenvalues appear (red circles
-in Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
-(\(-0.05, 0\)). The former has a large radius, but the real components
-have shifted to the left (in comparison to when \(\gamma = 1\)), while
-the latter cluster has a smaller radius, but real components have
-shifted to the right; all \(\Re({\lambda}) < 0\). Overall, for 1 million
-randomly assembled \(\mathbf{M}\), this division between slow and fast
-component response rates results in more stable systems: 1 stable given
-\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
+cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
+diagonal elements are set to \(-1\) and non-zero off-diagonal elements
+are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
+multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
+\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
+uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
+(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
+to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
+values are separated into two groups, eigenvalues are no longer
+uniformly distributed (Fig. 1b). Instead, two distinct clusters of
+eigenvalues appear (red circles in Fig. 1b), one centred at
+(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
+large radius, but the real components have shifted to the left (in
+comparison to when \(\gamma = 1\)), while the latter cluster has a
+smaller radius, but real components have shifted to the right; all
+\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
+\(\mathbf{M}\), this division between slow and fast component response
+rates results in more stable systems: 1 stable given \(\gamma = 1\)
+versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
 
@@
 Supplementary Information for other distributions, which give similar
-results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b),
-\(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
+results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
+\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
 comparison of \(\mathbf{M}\) before and after the addition of variation
@@
 table of results), and that the difference between the probabilities of
-observing a stable system increased with an increase in \(S\); i.e., the
-potential for \(Var(\gamma)\) to drive stability increased with system
-complexity. For the highest values of \(S\), nearly all systems that
-were stable given \(Var(\gamma)\) would not have been stable given
-\(\gamma = 1\).
+observing a stable system increased with an increase in \(S\). In other
+words, the potential for \(Var(\gamma)\) to drive stability increased
+with system complexity and was most relevant for systems on the cusp of
+being too complex to be realistically stable. For the highest values of
+\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
+not have been stable given \(\gamma = 1\).
 
@@
 that stability could often be achieved by manipulating \(S\) \(\gamma\)
-values rather than \(S \times S\) \(\mathbf{M}\) elements.
+values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
+manipulating component response rates rather than interactions between
+components).
 
@@
 This is of particular interest for ecological communities because
-population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically
+population density (\(N\)) cannot take negative values, meaning that
+ecological systems need to be feasible for stability to be biologically
 realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
@@
 (e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). Feasibility was unaffected by \(Var(\gamma)\) and instead
-occurred with a fixed probability of \(1/2^{S}\), consistent with a
-recent proof by Serván et
-al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}} (see
-Supplementary Information). Hence, for pure interacting species
+Information). I want to emphasise that \(\gamma\) is not interpreted as
+population density in this analysis, but instead as a fundamental
+property of species life history such as expected generation time.
+Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
+a fixed probability of \(1/2^{S}\), consistent with a recent proof by
+Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
+(see Supplementary Information). Hence, for pure interacting species
 networks, variation in component response rate (i.e., species generation
@@
 investigating the stability of complex systems. Variation in component
-response rate becomes more likely to be critical for stability as system
-size increases, and can ultimately increase the overall probability that
-system stability is observed above that predicted by
+response rate is more likely to be critical for stability in systems
+that are especially large (but finite, see Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}), and it can
+ultimately increase the probability that system stability is observed
+above that predicted by
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
@@
 
-\[\mathbf{x^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]
+\[\mathbf{n^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]
 
-In the above, \(\mathbf{x^{*}}\) is the vector of species densities at
+In the above, \(\mathbf{n^{*}}\) is the vector of species densities at
 equilibrium. Feasibility is satisfied if all elements in
-\(\mathbf{x^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
+\(\mathbf{n^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
 identity matrix, and the value \(\theta\) is the strength of
@@
 
-\[\mathbf{x^{*} = -A^{-1}r}.\]
+\[\mathbf{n^{*} = -A^{-1}r}.\]
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.738 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.732 for a and b,
 respectively.

### 2019-02-27 notebook/ms.Rmd
Edit the manuscript up to Reviewer 2 specific comment 3 -- these are next on the list, specifically the new Discussion paragraph

@@
 
-May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rates ($E[\gamma]$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rates ($\gamma$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
@@
 
-<!--- Also important to emphasise Gibbs result -- I'm doing this for finite systems, and I deliberate stressed the system complexity to the point of collapse. --->
+<!---
+
+Also important to emphasise Gibbs result -- I'm doing this for finite systems, and I deliberate stressed the system complexity to the point of collapse.
+
+But Gibbs was more interested in first assuming a stable matrix and then showing that the vector of abundances would not change it as S \to infinity -- Instead of assuming the system was stable prior to adding component response rate, I allow both the system and component response rate to vary to see the effects on stability.
+
+--->
 
@@
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{M} =\mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+In these analyses, row means of $\mathbf{A}$ are expected to be identical, but variation around this expectation will naturally arise due to random sampling of $\mathbf{A}$ off-diagonal elements and finite $S$.
+In simulations, the total variation in $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ row means that is attributable to $\mathbf{A}$ is small relative to that attributable to $\mathbf{\gamma}$, especially at high $S$.
+Variation in $\mathbf{\gamma}$ specifically isolates the effects of differing component response rates, hence causing differences in expected $\mathbf{M}$ row means.
 

### 2019-02-27 notebook/ms.tex
Edit the manuscript up to Reviewer 2 specific comment 3 -- these are next on the list, specifically the new Discussion paragraph

@@
 correspond to individual component response rates. Therefore,
-\(\mathbf{M} =\mathbf{\gamma} \mathbf{A}\) modulates the values of
+\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) modulates the values of
 components and can be analysed using the techniques of
 May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
+In these analyses, row means of \(\mathbf{A}\) are expected to be
+identical, but variation around this expectation will naturally arise
+due to random sampling of \(\mathbf{A}\) off-diagonal elements and
+finite \(S\). In simulations, the total variation in
+\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) row means that is
+attributable to \(\mathbf{A}\) is small relative to that attributable to
+\(\mathbf{\gamma}\), especially at high \(S\). Variation in
+\(\mathbf{\gamma}\) specifically isolates the effects of differing
+component response rates, hence causing differences in expected
+\(\mathbf{M}\) row means.
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.732 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.719 and 0.737 for a and b,
 respectively.

### 2019-03-01 notebook/ms.Rmd
Major restructure and revision of the Discussion to compare to Gibbs et al.

@@
 Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) in which $\gamma$ varies are actually more likely to be stable.
-<!--- Need to emphasise that this is for *real* systems at the thresholds of stability --->
-
-<!---
-
-Introduction could end here.
-
---->
 
@@
 It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
-Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
+Rather, highly complex systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
 3).
@@
 1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
-The mathematics underlying this shift in eigenvalue distribution has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but has not been interpreted as rates of response of individual system components to perturbation.
-
-<!---
-
-Also important to emphasise Gibbs result -- I'm doing this for finite systems, and I deliberate stressed the system complexity to the point of collapse.
-
-But Gibbs was more interested in first assuming a stable matrix and then showing that the vector of abundances would not change it as S \to infinity -- Instead of assuming the system was stable prior to adding component response rate, I allow both the system and component response rate to vary to see the effects on stability.
-
---->
+In this way, my focus is distinct from that of Gibbs et al.
+[@Gibbs2017], who applied the same mathematical framework to investigate how a diagonal matrix $\mathbf{X}$ (equivalent to $\gamma$ in my model) affects the stability of a community matrix $\mathbf{M}$ given an interaction matrix $\mathbf{A}$ within a generalised Lotka-Volterra model, where $\mathbf{M} = \mathbf{XA}$.
+Gibbs et al.
+[@Gibbs2017] analytically demonstrated that the effect of $\mathbf{X}$ on system stability decreases exponentially as system size becomes arbitrarily large ($S \to \infty$).
+My numerical results are entirely consistent with this prediction (see Supplementary Information).
+Given that a complex system is known to be stable (or unstable), $Var(\gamma)$ is not expected to increase or decrease the probability of stability.
+Rather, component response rate variation increases the upper bound at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
 
@@
 
-My results show that complex systems are more likely to be stable when the response rates of system components vary.
-These results are broadly applicable to understanding stability of complex networks in the physical, life, and social sciences.
+A general mathematical framework encompassing shifts in eigenvalue distributions caused by a vector $\gamma$ has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but $\gamma$ has not been interpreted as rates of response of individual system components to perturbation.
+My model focuses on component response rates for systems of a finite size, in which complexity is high but not yet high enough to make the probability of stability unrealistically low for actual empirical systems.
+For this upper range of system size, randomly assembled complex systems are more likely to be stable if their component response rates vary (e.g., $10 < S < 30$ for parameter values in Fig.
+3).
+Overall, I suggest that variation in component response rate might therefore be critical for maintaining stability in many highly complex empirical systems.
+These results are broadly applicable to understanding stability of complex networks across the physical, life, and social sciences.
 

### 2019-03-01 notebook/ms.tex
Major restructure and revision of the Discussion to compare to Gibbs et al.

@@
 stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
-response rates (\(E[\gamma]\)) of individual components to perturbations
-of the system are identical, but this is highly unlikely in any complex
+response rates (\(\gamma\)) of individual components to perturbations of
+the system are identical, but this is highly unlikely in any complex
 system. In ecological communities, for example, the rate at which
@@
 response rate to a particular system does not necessarily increase the
-probability that the system will be stable. Rather, systems that are
-observed to be stable are more likely to have varying component response
-rates, and for this variation to be critical to their stability (Fig.
-3). This is caused by the shift to a non-uniform distribution of
-eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
-which can sometimes cause all of the real components of the eigenvalues
-of the community matrix \(\mathbf{M}\) to become negative, but might
-also increase the real components of eigenvalues. The mathematics
-underlying this shift in eigenvalue distribution has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
-and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
-but has not been interpreted as rates of response of individual system
-components to perturbation.
+probability that the system will be stable. Rather, highly complex
+systems that are observed to be stable are more likely to have varying
+component response rates, and for this variation to be critical to their
+stability (Fig. 3). This is caused by the shift to a non-uniform
+distribution of eigenvalues that occurs by introducing \(Var(\gamma)\)
+(Fig. 1b, 2b), which can sometimes cause all of the real components of
+the eigenvalues of the community matrix \(\mathbf{M}\) to become
+negative, but might also increase the real components of eigenvalues. In
+this way, my focus is distinct from that of Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
+the same mathematical framework to investigate how a diagonal matrix
+\(\mathbf{X}\) (equivalent to \(\gamma\) in my model) affects the
+stability of a community matrix \(\mathbf{M}\) given an interaction
+matrix \(\mathbf{A}\) within a generalised Lotka-Volterra model, where
+\(\mathbf{M} = \mathbf{XA}\). Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}} analytically
+demonstrated that the effect of \(\mathbf{X}\) on system stability
+decreases exponentially as system size becomes arbitrarily large
+(\(S \to \infty\)). My numerical results are entirely consistent with
+this prediction (see Supplementary Information). Given that a complex
+system is known to be stable (or unstable), \(Var(\gamma)\) is not
+expected to increase or decrease the probability of stability. Rather,
+component response rate variation increases the upper bound at which
+stability can be realistically observed, meaning that highly complex
+systems are more likely than not to vary in their component response
+rates, and for this variation to be critical for system stability.
 
@@
 
-My results show that complex systems are more likely to be stable when
-the response rates of system components vary. These results are broadly
-applicable to understanding stability of complex networks in the
-physical, life, and social sciences.
+A general mathematical framework encompassing shifts in eigenvalue
+distributions caused by a vector \(\gamma\) has been
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+and recently applied to questions concerning species density and
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
+but \(\gamma\) has not been interpreted as rates of response of
+individual system components to perturbation. My model focuses on
+component response rates for systems of a finite size, in which
+complexity is high but not yet high enough to make the probability of
+stability unrealistically low for actual empirical systems. For this
+upper range of system size, randomly assembled complex systems are more
+likely to be stable if their component response rates vary (e.g.,
+\(10 < S < 30\) for parameter values in Fig. 3). Overall, I suggest that
+variation in component response rate might therefore be critical for
+maintaining stability in many highly complex empirical systems. These
+results are broadly applicable to understanding stability of complex
+networks across the physical, life, and social sciences.
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.719 and 0.737 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.738 for a and b,
 respectively.

### 2019-03-01 notebook/SI.Rmd
Some work on the SI

@@
 - [Feasibility of complex systems](#Feasibility)
+- [Consistency with Gibbs et al. (2018)](#Gibbs)
 
@@
 
+Consistency with Gibbs et al. (2018) {#Gibbs}
+===============================================================================
+
+The question that I address in the main text is distinct from that of Gibbs et al [@Gibbs2017], which focused instead on the effect of a diagonal matrix of biological species densities $\mathbf{X}$ on a community matrix $\mathbf{M}$ given a species interaction matrix $\mathbf{A}$.
+This is modelled as below,
+
+$$\mathbf{M} = \mathbf{XA}$$.
+
+Mathematically, the above is identical to my model in the main text where the system $\mathbf{M}$ is defined by component interaction strengths $\mathbf{A}$ and individual component response rates $\mathbf{\gamma}$,
+
+$$\mathbf{M} = \mathbf{\gamma A}$$.
+
+While I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ for finite by high system size ($S$), Gibbs et al.
+[@Gibbs2017] instead focused on the different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$.
+It is easy to show that my primary results in [the table above](#IncrS) are consistent with Gibbs et al.
+[@Gibbs2017] by calculating the probability that system stability changes as $S$ increases.
+Gibbs et al.
+[@Gibbs2017] predict that the probability that stability is altered should decrease with increasing $S$.
+
+```{r, echo = FALSE}
+d_tab    <- dat[,1:7];
+p_chnged <- (dat[,6] + dat[,7]) / (dat[,2] + dat[,3]);
+gibbtb   <- cbind(d_tab[,1], p_chnged);
+colnames(gibbtb) <- c("S", "Pr_changed");
+plot(x = gibbtb[,1], y = gibbtb[,2], type = "l");
+```
+
+Note, however, that within this finite range of $S$, the probability that an unstable system will be stabilised by $\gamma$ increases as $S$ increases.
+In other words, the proportion of unstable systems that are stabilised by $\gamma$ increases with $S$ in simulations.
+
+```{r, echo = FALSE}
+Pr_stbled <- (dat[,6]) / (dat[,6] + dat[,7]);
+gibbtb2   <- cbind(d_tab[,1], Pr_stbled);
+colnames(gibbtb2) <- c("S", "Pr_stbled");
+plot(x = gibbtb2[,1], y = gibbtb2[,2], type = "l");
+```
+
+```
+
 References

### 2019-03-01 notebook/SI.tex
Some work on the SI

@@
 \setlength{\droptitle}{-2em}
+
   \title{Component response rate variation drives stability in large complex
@@
   \protect\hyperlink{Feasibility}{Feasibility of complex systems}
+\item
+  \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
 \end{itemize}
@@
 
+\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
+(2018)}\label{Gibbs}}
+
+The question that I address in the main text is distinct from that of
+Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}}, which
+focused instead on the effect of a diagonal matrix of biological species
+densities \(\mathbf{X}\) on a community matrix \(\mathbf{M}\) given a
+species interaction matrix \(\mathbf{A}\). This is modelled as below,
+
+\[\mathbf{M} = \mathbf{XA}\].
+
+Mathematically, the above is identical to my model in the main text
+where the system \(\mathbf{M}\) is defined by component interaction
+strengths \(\mathbf{A}\) and individual component response rates
+\(\mathbf{\gamma}\),
+
+\[\mathbf{M} = \mathbf{\gamma A}\].
+
+While I focused on the probability of observing a stable versus unstable
+system given variation in \(\mathbf{\gamma}\) for finite by high system
+size (\(S\)), Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}} instead
+focused on the different question concerning the effect of
+\(\mathbf{\gamma}\) on the stability of \(\mathbf{M}\) given
+\(\mathbf{A}\) as \(S \to \infty\). It is easy to show that my primary
+results in \protect\hyperlink{IncrS}{the table above} are consistent
+with Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}}
+by calculating the probability that system stability changes as \(S\)
+increases. Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}} predict that
+the probability that stability is altered should decrease with
+increasing \(S\).
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+
+Note, however, that within this finite range of \(S\), the probability
+that an unstable system will be stabilised by \(\gamma\) increases as
+\(S\) increases. In other words, the proportion of unstable systems that
+are stabilised by \(\gamma\) increases with \(S\) in simulations.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-27-1.pdf}
+
+```
+
 \section*{References}\label{references}
@@
 
+\hypertarget{ref-Gibbs2017}{}
+4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{arXiv} (2017).
+
 \end{document}

### 2019-03-02 notebook/SI.Rmd
Revisions, need to add some still to the SI

@@
 
-While I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ for finite by high system size ($S$), Gibbs et al.
-[@Gibbs2017] instead focused on the different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$.
-It is easy to show that my primary results in [the table above](#IncrS) are consistent with Gibbs et al.
-[@Gibbs2017] by calculating the probability that system stability changes as $S$ increases.
-Gibbs et al.
-[@Gibbs2017] predict that the probability that stability is altered should decrease with increasing $S$.
+I focused on the probability of observing a stable versus unstable system given variation in $\mathbf{\gamma}$ as system complexity ($\sigma\sqrt{SC}$) increased.
+I increased system complexity by holding $C$ and $\sigma$ constant and incrementally increasing $S$ to obtain numerical results.
+In contrast, Gibbs et al.
+[@Gibbs2017] applied analytical techniques to instead focused on a different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$, with $\sigma$ scaled so that $\sigma = 1/\sqrt{S}$.
+Under such scaling, Gibbs et al.
+[@Gibbs2017] showed that the effect of $\gamma$ on stability should decrease exponentially as $S$ increases, which I can replicate below by running simulations in which $\sigma = 1/\sqrt{S}$.
 
-```{r, echo = FALSE}
-d_tab    <- dat[,1:7];
-p_chnged <- (dat[,6] + dat[,7]) / (dat[,2] + dat[,3]);
-gibbtb   <- cbind(d_tab[,1], p_chnged);
-colnames(gibbtb) <- c("S", "Pr_changed");
-plot(x = gibbtb[,1], y = gibbtb[,2], type = "l");
-```
-
-Note, however, that within this finite range of $S$, the probability that an unstable system will be stabilised by $\gamma$ increases as $S$ increases.
-In other words, the proportion of unstable systems that are stabilised by $\gamma$ increases with $S$ in simulations.
-
-```{r, echo = FALSE}
-Pr_stbled <- (dat[,6]) / (dat[,6] + dat[,7]);
-gibbtb2   <- cbind(d_tab[,1], Pr_stbled);
-colnames(gibbtb2) <- c("S", "Pr_stbled");
-plot(x = gibbtb2[,1], y = gibbtb2[,2], type = "l");
-```
-
-```
+The results in the table above can be compared to those of the [main results] (#IncrS).
+Note that only 100000, rather than 1 million, simulations are run to confirm consistency with Gibbs et al.
+[@Gibbs2017].
 

### 2019-03-02 notebook/ms.Rmd
Revisions, need to add some still to the SI

@@
 
+<!---
+
 Author Summary
@@
 These results are general to any physical, biological, social, or mixed system.
+
+--->
 
@@
 
-It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
+It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable (at high $S$, the opposite is true; see Supplementary Information).
 Rather, highly complex systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
@@
 [@Gibbs2017] analytically demonstrated that the effect of $\mathbf{X}$ on system stability decreases exponentially as system size becomes arbitrarily large ($S \to \infty$).
-My numerical results are entirely consistent with this prediction (see Supplementary Information).
-Given that a complex system is known to be stable (or unstable), $Var(\gamma)$ is not expected to increase or decrease the probability of stability.
-Rather, component response rate variation increases the upper bound at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
+My numerical results are not inconsistent with this prediction because I did not scale $\sigma = 1 / \sqrt{S}$, but instead fixed $\sigma$ and increased $S$ to thereby increase total system complexity, $\sigma\sqrt{SC}$ (see Supplemental Information for results simulated across $\sigma$ and $C$.
+Overall, my results show that component response rate variation increases the upper bound of complexity at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
 
@@
 The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
-dat <- dat[,-1];
-plot_stable_4(dat);
-```

### 2019-03-02 notebook/ms.tex
Revisions, need to add some still to the SI

@@
 across all realistic complex systems.
-
-\subsection{Author Summary}\label{author-summary}
-
-Any set of discrete components that potentially interact can define a
-complex system. Typically, properties of individual components will
-cause them to respond to system perturbation at different rates, but the
-consequences of such variation for system stability are unexplored. This
-work makes three novel theoretical contributions: First, I show that
-large complex systems with varying component response rates are more
-likely to be stable than when response rates are uniform. Second, I show
-that a targeted manipulation of component response rates can increase
-the probability of system stability by several orders of magnitude.
-Lastly, I show that varying component response rate does not affect
-system feasibility. These results are general to any physical,
-biological, social, or mixed system.
 
@@
 response rate to a particular system does not necessarily increase the
-probability that the system will be stable. Rather, highly complex
-systems that are observed to be stable are more likely to have varying
-component response rates, and for this variation to be critical to their
-stability (Fig. 3). This is caused by the shift to a non-uniform
-distribution of eigenvalues that occurs by introducing \(Var(\gamma)\)
-(Fig. 1b, 2b), which can sometimes cause all of the real components of
-the eigenvalues of the community matrix \(\mathbf{M}\) to become
-negative, but might also increase the real components of eigenvalues. In
-this way, my focus is distinct from that of Gibbs et
+probability that the system will be stable (at high \(S\), the opposite
+is true; see Supplementary Information). Rather, highly complex systems
+that are observed to be stable are more likely to have varying component
+response rates, and for this variation to be critical to their stability
+(Fig. 3). This is caused by the shift to a non-uniform distribution of
+eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
+which can sometimes cause all of the real components of the eigenvalues
+of the community matrix \(\mathbf{M}\) to become negative, but might
+also increase the real components of eigenvalues. In this way, my focus
+is distinct from that of Gibbs et
 al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
@@
 decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)). My numerical results are entirely consistent with
-this prediction (see Supplementary Information). Given that a complex
-system is known to be stable (or unstable), \(Var(\gamma)\) is not
-expected to increase or decrease the probability of stability. Rather,
-component response rate variation increases the upper bound at which
-stability can be realistically observed, meaning that highly complex
-systems are more likely than not to vary in their component response
-rates, and for this variation to be critical for system stability.
+(\(S \to \infty\)). My numerical results are not inconsistent with this
+prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
+instead fixed \(\sigma\) and increased \(S\) to thereby increase total
+system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
+results simulated across \(\sigma\) and \(C\). Overall, my results show
+that component response rate variation increases the upper bound of
+complexity at which stability can be realistically observed, meaning
+that highly complex systems are more likely than not to vary in their
+component response rates, and for this variation to be critical for
+system stability.
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.738 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.726 and 0.742 for a and b,
 respectively.
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
-
 \end{document}

### 2019-03-02 notebook/arxiv/ms.tex
Prep the manuscript for both Scientific Reports and arxiv (SI still needs to be fixed in the arxiv tex)

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation drives stability in large complex systems},
+            pdftitle={Component response rate variation underlies the stability of complex systems},
             pdfauthor={A. Bradley Duthie},
@@
 \setlength{\droptitle}{-2em}
-  \title{Component response rate variation drives stability in large complex
-systems}
+  \title{Component response rate variation underlies the stability of complex systems}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
-\textbf{The stability of a complex system generally decreases with
-increasing system size, as is demonstrated by random matrix
-theory\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-This counter-intuitive result, first shown by
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}, is broadly
-relevant for understanding the dynamics and persistence of systems such
-as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{3},\protect\hyperlink{ref-Gray2009}{4}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{5},\protect\hyperlink{ref-MacArthur2010}{6}}
-and
-socio-economic\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}--\protect\hyperlink{ref-Bardoscia2017}{9}}
-networks. Much attention has especially been given to the stability of
-ecological communities such as food webs or mutualist networks, with
-recent work investigating how different community structures affect
-stability\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10}--\protect\hyperlink{ref-Patel2018}{14}}.
-But more broadly, stabilising mechanisms in complex systems remain
-under-developed, and the effect of variation in the response rate of
-individual system components remains an open
-problem\textsuperscript{\protect\hyperlink{ref-Allesina2015}{15}}. Here
-I show that when components of a complex system respond to system
-perturbation at different rates (\(\boldsymbol{\gamma}\)), the potential for
-system stability is markedly increased. Stability is caused by the
-clustering of some eigenvalues toward the centre of eigenvalue
-distributions despite the destabilising effect of higher interaction
-strength variation (\(\boldsymbol{\sigma^{2}}\)). This effect of
-variation in \(\boldsymbol{\gamma}\) becomes increasingly important as
-system size increases, to the extent that the largest stable complex
-systems would otherwise be unstable if not for
-\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
-unconsidered driver of system stability that is likely to be pervasive
-across all complex systems. Future research in complex systems should
-therefore account for the varying response rates of individual system
-components when assessing whole system stability.}
+\textbf{The stability of a complex system generally decreases with increasing
+system size and interconnectivity, a counterintuitive result of
+widespread importance across the physical, life, and social sciences.
+Despite recent interest in the relationship between system properties
+and stability, the effect of variation in the response rate of
+individual system components remains unconsidered. Here I vary the
+component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
+complex systems. I show that when component response rates vary, the
+potential for system stability is markedly increased. Variation in
+\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
+large but finite systems, where the probability of stability would
+otherwise be negligible. At such extremes of simulated system
+complexity, the largest stable complex systems would be unstable if not
+for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
+previously unconsidered aspect of system stability that is likely to be
+pervasive across all realistic complex systems.}
 
@@
 perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of potentially interacting components; \(S\)), its 
-connectance (i.e., the probability that one component will interact with 
-another; \(C\)), and the variance of interaction strengths
+system (i.e., the number of potentially interacting components; \(S\)),
+its connectance (i.e., the probability that one component will interact
+with another; \(C\)), and the variance of interaction strengths
 (\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
 May's finding that the probability of local stability falls to near zero
-given a sufficiently high threshold of \(\sigma\sqrt{SC}\) has profound
-consequences across multiple disciplines, raising the question of how
-complex systems in, e.g.,
-ecology\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Mougi2012}{10},\protect\hyperlink{ref-Grilli2017}{13},\protect\hyperlink{ref-Allesina2015}{15}}
-or
-banking\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7},\protect\hyperlink{ref-Bardoscia2017}{9},\protect\hyperlink{ref-May2008}{16}}
-are predicted to persist or change.
+given a sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly
+relevant for understanding the dynamics and persistence of systems such
+as
+ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
+neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
+biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
+and
+socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
+networks. As such, identifying general principles that affect stability
+in complex systems is of wide-ranging importance.
 
 Randomly assembled complex systems can be represented as large square
-matrices (\(M\)) with \(S\) components (e.g.,
+matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
 species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{7}}). One
-element of such a matrix, \(M_{ij}\), defines how component \(j\) affects
-component \(i\) in the system at a point of
+banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
+element of such a matrix, \(M_{ij}\), defines how component \(j\)
+affects component \(i\) in the system at a point of
 equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero
-values with a probability \(C\), which are sampled from a distribution
-with variance \(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}.
+Traditionally, off-diagonal elements are assigned non-zero values with a
+probability \(C\), which are sampled from a distribution with variance
+\(\sigma^{2}\); diagonal elements are set to
+-1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
 Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of \(M\)
-are negative
+being stable if the real parts of all eigenvalues (\(\lambda\)) of
+\(\mathbf{M}\) are negative
 (\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
 In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}} within a
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
 circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
 \(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{15}}
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
@@
 
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}} stability criterion \(\sigma\sqrt{SC} < 1\) assumes that individual components respond
-to perturbations of the system at the same rate (\(\gamma\)), but this
-is highly unlikely in any complex system. In ecological communities, for
-example, the rate at which population density changes following
-perturbation will depend on the generation time of individuals, which
-might vary by orders of magnitude among species. Species with short
-generation times will respond quickly (high \(\gamma\)) to perturbations
-relative to species with long generation times (low \(\gamma\)).
-Similarly, the speed at which individual banks respond to perturbations
-in financial networks, or individuals or institutions respond to
-perturbations in complex social networks, is likely to vary. The effect
-of such variance has not been investigated in complex systems theory.
-Intuitively, variation in \(\gamma\) might be expected to decrease
-system stability by introducing a new source of variation into the
-system and thereby increasing \(\sigma\). Here I show why, despite
-higher \(\sigma\), complex systems in which \(\gamma\) varies are
-actually more likely to be stable, especially when \(S\) is high.
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
+stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
+response rates (\(\gamma\)) of individual components to perturbations of
+the system are identical, but this is highly unlikely in any complex
+system. In ecological communities, for example, the rate at which
+population density changes following perturbation will depend on the
+generation time of organisms, which might vary by orders of magnitude
+among species. Species with short generation times will respond quickly
+(high \(\gamma\)) to perturbations relative to species with long
+generation times (low \(\gamma\)). Similarly, the speed at which
+individual banks respond to perturbations in financial networks, or
+individuals or institutions respond to perturbations in complex social
+networks, is likely to vary. The effect of such variance on stability
+has not been investigated in complex systems theory. Intuitively,
+variation in \(\gamma\) might be expected to decrease system stability
+by introducing a new source of variation into the system and thereby
+increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
+realistic complex systems (such that \(S\) is high but finite) in which
+\(\gamma\) varies are actually more likely to be stable.
+
+\subsection{Results}\label{results}
+
+\textbf{Component response rates of random complex systems}. Rows in
+\(\mathbf{M}\) define how a given component \(i\) is affected by other
+components of the system, meaning that the rate of component response
+time can be modelled by multiplying all row elements by a real scalar
+value
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+The distribution of \(\gamma\) over \(S\) components thereby models the
+distribution of component response rates. An instructive example
+compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
+\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
+half of \(\gamma_{i} = 0.05\). This models one system in which
+\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
+are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
+cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
+diagonal elements are set to \(-1\) and non-zero off-diagonal elements
+are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
+multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
+\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
+uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
+(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
+to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
+values are separated into two groups, eigenvalues are no longer
+uniformly distributed (Fig. 1b). Instead, two distinct clusters of
+eigenvalues appear (red circles in Fig. 1b), one centred at
+(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
+large radius, but the real components have shifted to the left (in
+comparison to when \(\gamma = 1\)), while the latter cluster has a
+smaller radius, but real components have shifted to the right; all
+\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
+\(\mathbf{M}\), this division between slow and fast component response
+rates results in more stable systems: 1 stable given \(\gamma = 1\)
+versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+
+Higher stability in systems with variation in \(\gamma\) can be observed
+by sampling \(\gamma_{i}\) values from various distributions. I focus on
+a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
+Supplementary Information for other distributions, which give similar
+results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
+\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
+comparison of \(\mathbf{M}\) before and after the addition of variation
+in component response rate. Figure 2 shows a comparison of eigenvalue
+distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
+expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
+\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
+centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
+Uniform variation in \(\gamma\) leads to a non-uniform distribution of
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
+some of which are clustered locally near the centre of the distribution,
+but others of which are spread outside the former radius of 2.828 (Fig
+2b). The clustering and spreading of eigenvalues introduced by
+\(Var(\gamma)\) can destabilise previously stable systems or stabilise
+systems that are otherwise unstable. But where systems are otherwise too
+complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
+can often lead to stability above
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
+threshold \(\sigma\sqrt{SC} < 1\).
+
+\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
+investigate the effect of \(Var(\gamma)\) on stability across systems of
+varying complexity, I simulated random \(\mathbf{M}\) matrices at
+\(\sigma = 0.4\) and \(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One
+million \(\mathbf{M}\) were simulated for each \(S\), and the stability
+of \(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
+\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
+number of stable random systems was higher given \(Var(\gamma)\) than
+when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
+table of results), and that the difference between the probabilities of
+observing a stable system increased with an increase in \(S\). In other
+words, the potential for \(Var(\gamma)\) to affect stability increased
+with system complexity and was most relevant for systems on the cusp of
+being too complex to be realistically stable. For the highest values of
+\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
+not have been stable given \(\gamma = 1\).
+
+\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
+investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
+genetic algorithm. Genetic algorithms are heuristic tools that mimic
+evolution by natural selection, and are useful when the space of
+potential solutions (in this case, possible combinations of \(\gamma\)
+values leading to stability in a complex system) is too large to search
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
+Generations of selection on \(\gamma\) value combinations to minimise
+\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
+\(Var(\gamma)\) to increase system stability. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were up to four
+orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
+that stability could often be achieved by manipulating \(S\) \(\gamma\)
+values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
+manipulating component response rates rather than interactions between
+components).
+
+\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
+systems in which individual system components represent the density of
+some tangible quantity, it is relevant to consider the feasibility of
+the system. Feasibilility assumes that values of all components are
+positive at
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
+This is of particular interest for ecological communities because
+population density (\(N\)) cannot take negative values, meaning that
+ecological systems need to be feasible for stability to be biologically
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
+While my results are intended to be general to all complex systems, and
+not restricted to species networks, I have also performed a feasibility
+analysis on all matrices \(\mathbf{M}\) tested for stability, and
+additionally for specific types of ecological
+communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+(e.g., competitive, mutualist, predator-prey; see Supplementary
+Information). I emphasise that \(\gamma\) is not interpreted as
+population density in this analysis, but instead as a fundamental
+property of species life history such as expected generation time.
+Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
+a fixed probability of \(1/2^{S}\), consistent with a recent proof by
+Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
+(see Supplementary Information). Hence, for pure interacting species
+networks, variation in component response rate (i.e., species generation
+time) does not affect stability at biologically realistic species
+densities.
 
@@
 shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only 
-\(Var(\gamma)\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); points are
-uniformly distributed within the grey circle with a radius of
+\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
+distribution of \(\gamma\) differs between panels). \textbf{a.}
+Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
+uniformly distributed within the blue circle with a radius of
 \(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
@@
 \(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle with a radius of \(\sigma\sqrt{SC} =\) 1.718 centred at
-\(\gamma = -1.95\) and one small circle with a radius of
-\(\sigma\sqrt{SC} =\) 0.044
-centred at \(\gamma = -0.05\). In a, the maximum real eigenvalue
+large circle centred at \(\gamma = -1.95\) and one small circle centred
+at \(\gamma = -0.05\). In a, the maximum real eigenvalue
 \(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
@@
 
-\includegraphics{unnamed-chunk-4-1.pdf}
+\includegraphics{unnamed-chunk-10-1.pdf}
 
@@
 
-Rows in \(M\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{14}} (see 
-\hyperlink{SIstart}{Supplementary Information} for details).
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(M\) when \(\gamma_{i} = 1\) for all \(i\) in \(S\) to
-the same \(M\) when half of \(\gamma_{i} = 1.95\) and half of
-\(\gamma_{i} = 0.05\). This models one system in which \(\gamma\) is
-invariant and one in which \(\gamma\) varies, but systems are otherwise
-identical (note \(E[\gamma_{i}] = 1\) in both cases). I assume
-\(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\); diagonal elements are
-set to \(-1\) and non-zero off-diagonal elements are drawn randomly from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
-\(\gamma_{i}\) to generate \(M\). When \(\gamma_{i} = 1\), eigenvalues
-of \(M\) are distributed uniformly within a circle centred at
-(\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real components
-of eigenvalues are highly unlikely to all be negative when all
-\(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated into
-two groups, eigenvalues are no longer uniformly distributed (Fig. 1b).
-Instead, two distinct clusters of eigenvalues appear (grey circles in
-Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
-(\(-0.05, 0\)). The former has a large radius, but the real components
-have shifted to the left (in comparison to when \(\gamma = 1\)) and all
-\(\Re({\lambda}) < 0\). The latter cluster has real components that have
-shifted to the right, but has a smaller radius. Overall, for 1 million
-randomly assembled \(M\), this division between slow and fast component
-response rates results in more stable systems: 1 stable given
-\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+\subsection{Discussion}\label{discussion}
+
+I have shown that the stability of complex systems might often be
+contigent upon variation in the response rates of their individual
+components, meaning that factors such as rate of trait evolution (in
+biological networks), transaction speed (in economic networks), or
+communication speed (in social networks) need to be considered when
+investigating the stability of complex systems. Variation in component
+response rate is more likely to be critical for stability in systems
+that are especially complex, and it can ultimately increase the
+probability that system stability is observed above that predicted by
+May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
+derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
+general, and potentially applies to any complex system in which
+individual system components can vary in their reaction rates to system
+perturbation.
 
@@
 introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel shows the same
-system where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). \textbf{a.}
-Eigenvalues plotted in the absence of \(Var(\gamma)\) where
+(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
+system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
+\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
 \(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
 \(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but clusters eigenvalues toward
-the distribution's centre (-1, 0). Black and red elipses in both panels
-show the circle centred on the distribution in panels a and b,
-respectively, which have a radius of \(\sigma \sqrt{SC}\). Proportions
-of \(\Re(\lambda) < 0\) are 0.548 and 0.552 for a and b, respectively.
-
-\includegraphics{unnamed-chunk-7-1.pdf}
+interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
+eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
+both panels show the circle centred on the distribution in panel a.
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.742 for a and b,
+respectively.
+
+\includegraphics{unnamed-chunk-13-1.pdf}
 
@@
 
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I now
-focus on a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\)
-(see \hyperlink{SIstart}{Supplementary Information} for other distributions, 
-which give
-similar results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig.
-1b), \(E[\gamma] = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(M\) before and after variation in component response
-rate. Figure 2 shows a comparison of eigenvalue distributions given
-\(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{17}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 12.649.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues, some of which are clustered tightly around the centre of
-the distribution, but others of which are spread outside the former
-radius of 12.649 (red circle Fig 2b). This larger radius occurs because
-the addition of \(Var(\gamma)\) increases the realised \(\sigma\) of
-\(M\). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold \(\sigma\sqrt{SC} < 1\).
+It is important to recognise that variation in component response rate
+is not stabilising per se; that is, adding variation in component
+response rates to a particular system does not increase the probability
+that the system will be stable. Rather, highly complex systems that are
+observed to be stable are more likely to have varying component response
+rates, and for this variation to be critical to their stability (Fig.
+3). This is caused by the shift to a non-uniform distribution of
+eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
+which can sometimes cause all of the real components of the eigenvalues
+of the system matrix to become negative, but might also increase the
+real components of eigenvalues.
 
@@
 \textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate(\(\boldsymbol{\gamma}\)).} Y-axes show
-the \(\ln\) number of systems that are stable across different system sizes
-(\(S\)) given \(C = 1\) (left axis), and the proportion of systems in which
-variation in \(\gamma\) is critical for system stability (right axis). For each
+variation in component response rate (\(\boldsymbol{\gamma}\)).} The
+\(\ln\) number of systems that are stable across different system sizes
+(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
+which variation in \(\gamma\) is critical for system stability. For each
 \(S\), 1 million complex systems are randomly generated. Stability of
@@
 system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Light and dark grey bars show the number of stable systems in the
-absence versus presence of \(Var(\gamma)\), respectively. The black line
-shows the proportion of systems that are stable when
-\(Var(\gamma) > 0\), but would be unstable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-9-1.pdf}
+Blue and red bars show the number of stable systems in the absence and
+presence of \(Var(\gamma)\), respectively. The black line shows the
+proportion of systems that are stable when \(Var(\gamma) > 0\), but
+would be unstable if \(Var(\gamma) = 0\).
+
+\includegraphics{unnamed-chunk-15-1.pdf}
 
@@
 
-To investigate the effect of \(Var(\gamma)\) on system stability, I
-simulated random \(M\) matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S\) ranging from 2-32. One million \(M\) were simulated 
-for each \(S\), and the stability of \(M\) was assessed given \(\gamma = 1\)
-versus \(\gamma \sim \mathcal{U}(0, 2)\). I found that the number of stable 
-random systems was
-consistently higher given \(Var(\gamma)\) than when \(\gamma = 1\) (Fig.
-3), and that the difference between the probabilities of observing a
-stable system increased with an increase in \(S\); i.e., the potential
-for \(Var(\gamma)\) to drive stability increased with system complexity.
-For the highest values of \(S\), nearly all systems that were stable
-given \(Var(\gamma)\) would not have been stable given \(\gamma = 1\) (see
-\hyperlink{SIstart}{Supplementary Information} for full results). This 
-suggests that the
-stability of large systems might be dependent upon variation in the
-response rates of their individual components, meaning that factors such
-as generation time (in ecological networks), transaction speed (in
-economic networks), or communication speed (in social networks) need to
-be considered when investigating the stability of complex systems.
-
-It is important to point out 
-that \(Var(\gamma)\) is not stabilising per se; that is,
-adding variation in \(\gamma\) to a particular system \(M\) does not
-necessarily increase the probability that the system will be stable (see
-\hyperlink{SIstart}{Supplementary Information}). Rather, systems that are 
-observed to be stable are more likely to vary in \(\gamma\), and for this
-\(Var(\gamma)\) to be critical to their stability. This is caused by the
-shift in the distribution of eigenvalues that occurs by introducing
-\(Var(\gamma)\) (Fig. 1b, 2b), which can sometimes result in all
-\(\Re(\lambda) < 0\) but might also increase \(\Re(\lambda)\) values.
-
-To further investigate the potential of \(Var(\gamma)\) to be stabilising,
-I used a genetic algorithm (the space of possible \(\gamma\)
-values was too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{18}};
-see \hyperlink{SIstart}{Supplementary Information}). For each of 40000 
-random \(M\) (\(\sigma = 0.4\), \(C = 1\)), the
-genetic algorithm initialised 1000 different sets of
-\(\gamma \sim \mathcal{U}(0, 2)\) values of size \(S\). Eigenanalysis was
-performed on \(M\) using 
-each set of \(\gamma\) values, and the 20 sets with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
-subsequent mutation and crossover between the resulting new population
-of 1000 \(\gamma\) sets. The genetic algorithm terminated if a stable
-\(M\) was found, 20 generations occurred, or a convergence criteria of
-minimum fitness increase between generations was satisfied. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were up to four orders of
-magnitude higher than when \(\gamma = 1\) (see \hyperlink{SIstart}{Supplementary
-Information}), meaning that stability could often be achieved by 
-manipulating \(S\) \(\gamma\) values rather than \(S \times S\) \(M\) elements.
-Hence, managing the response rates of system components in a targetted
-way can potentially facilitate the stabilisation of complex systems 
-through a reduction in dimensionality.
-
-I have focused broadly on random complex systems, but it is also
-worthwhile to consider more restricted interactions such as those of
-specific ecological
-networks\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. These
-include systems in which all interactions are negative (competitive networks), positive (mutualist networks), or \(i\) and
-\(j\) pairs have opposing signs (predator-prey networks). In
-general, competitive and mutualist networks tend to be destabilising,
-and predator-prey networks tend to be
-stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2011}{19}}.
-When \(Var(\gamma)\) is applied to each, the proportion of stable
-competitive and predator-prey networks increases, but the proportion of
-stable mutualist networks does not (see 
-\hyperlink{SIstart}{Supplementary Information}).
-Additionally, when each component of \(M\) is interpreted as a unique
-species and given a random intrinsic growth
-rate\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}},
-feasibility is not increased by \(Var(\gamma)\), suggesting that
-variation in species generation time might be unlikely to drive
-stability in purely multi-species networks (see \hyperlink{SIstart}{Supplementary
-Information}).
-
-My results show that complex systems are more likely to be stable when
-the response rates of system components vary. These results are broadly
-applicable to complex biological and social networks.
-
-\textbf{Acknowledgements}
-
-I am supported by a Leverhulme Trust Early Career Fellowship (ECF-2016-376).
-Conversations with L. Bussi\`ere and N. Bunnefeld, and helpful comments from
-J. J. Cusack and I. L. Jones, improved the quality of this work.
+My focus is distinct from Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
+the same mathematical framework to investigate how a diagonal matrix
+\(\mathbf{X}\) (equivalent to \(\gamma\) in my model) affects the
+stability of a community matrix \(\mathbf{M}\) given an interaction
+matrix \(\mathbf{A}\) within a generalised Lotka-Volterra model, where
+\(\mathbf{M} = \mathbf{XA}\). Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}} analytically
+demonstrated that the effect of \(\mathbf{X}\) on system stability
+decreases exponentially as system size becomes arbitrarily large
+(\(S \to \infty\)). My numerical results do not contradict this
+prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
+instead fixed \(\sigma\) and increased \(S\) to thereby increase total
+system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
+results simulated across \(\sigma\) and \(C\)). Overall, I show that
+component response rate variation increases the upper bound of
+complexity at which stability can be realistically observed, meaning
+that highly complex systems are more likely than not to vary in their
+component response rates, and for this variation to be critical for
+system stability.
+
+The potential importance of component response rate variation was most
+evident from the results of simulations in which the genetic algorithm
+was used in attempt to maximise the probability of system stability. The
+probability that some combination of component response rates could be
+found to stabilise the system was shown to be up to four orders of
+magnitude higher than the background probabilities of stability in the
+absence of any component response rate variation. Instead of
+manipulating the \(S \times S\) interactions between system components,
+it might therefore be possible to manipulate only the \(S\) response
+rates of individual system components to achieve stability. Hence,
+managing the response rates of system components in a targeted way could
+potentially facilitate the stabilisation of complex systems through a
+reduction in dimensionality.
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+\textbf{Figure 4: Stability of large complex systems given
+\(\boldsymbol{\gamma = 1}\) versus targeted
+\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
+stable across different system sizes (\(S\), max \(S=40\)) for
+\(C = 1\), and the proportion of systems wherein a targeted search of
+\(\gamma\) values successfully resulted in system stability. For each
+\(S\), 100000 complex systems are randomly generated. Stability of each
+complex system is tested given variation in \(\gamma\) using a genetic
+algorithm to maximise the effect of \(\gamma\) values on increasing
+stability, as compared to stability in an otherwise identical system in
+which \(\gamma\) is the same for all components. Blue bars show the
+number of stable systems in the absence of component response rate
+variation, while red bars show the number of stable systems that can be
+generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(Var(\gamma) = 0\).
+
+\includegraphics{unnamed-chunk-16-1.pdf}
+
+\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+
+Interestingly, while complex systems were more likely to be stable given
+variation in component response rate, they were not more likely to be
+feasible, meaning that stability was not increased when component values
+were also restricted to being positive at equilibrium. Feasibility is
+important to consider, particularly for the study of ecological networks
+of
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}}
+because population densities cannot realistically be negative. My
+results therefore suggest that variation in the rate of population
+responses to perturbation (e.g., due to differences in generation time
+among species) is unlikely to be critical to the stability of purely
+multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in
+empirical
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
+instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability
+should therefore not be ignored in the broader context of ecological
+communities.
+
+A general mathematical framework encompassing shifts in eigenvalue
+distributions caused by a vector \(\gamma\) has been
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+and recently applied to questions concerning species density and
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
+but \(\gamma\) has not been interpreted as rates of response of
+individual system components to perturbation. My model focuses on
+component response rates for systems of a finite size, in which
+complexity is high but not yet high enough to make the probability of
+stability unrealistically low for actual empirical systems. For this
+upper range of system size, randomly assembled complex systems are more
+likely to be stable if their component response rates vary (e.g.,
+\(10 < S < 30\) for parameter values in Fig. 3). Overall, I suggest that
+variation in component response rate might therefore be critical for
+maintaining stability in many highly complex empirical systems. These
+results are broadly applicable for understanding the stability of
+complex networks across the physical, life, and social sciences.
+
+\subsection{Methods}\label{methods}
+
+\textbf{Component response rate variation (\(\mathbf{\gamma}\))}. In a
+synthesis of eco-evolutionary feedbacks on community stability, Patel et
+al. model a system that includes a vector of potentially changing
+species densities (\(\mathbf{N}\)) and a vector of potentially evolving
+traits
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+For any species \(i\) or trait \(j\), change in species density
+(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+
+\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
+
+\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+
+In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
+effects of all species densities and trait values on the density of a
+species \(i\) and the value of trait \(j\), respectively. Patel et al.
+were interested in stability when the evolution of traits was relatively
+slow or fast in comparison with the change in species
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
+this is modulated in the above by the scalar \(\epsilon\). The value of
+\(\epsilon\) thereby determines the timescale separation between ecology
+and evolution, with high \(\epsilon\) modelling relatively fast
+evolution and low \(\epsilon\) modelling relative slow
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+
+I use the same principle that Patel et al. use to modulate the relative
+rate of evolution to modulate rates of component responses for \(S\)
+components. Following
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}},
+the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
+by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
+\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
+
+\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).\]
+
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{23}},
+
+\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).\]
+
+In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
+correspond to individual component response rates. Therefore,
+\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the values of system
+components and can be analysed using the techniques of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
+In these analyses, row means of \(\mathbf{A}\) are expected to be
+identical, but variation around this expectation will naturally arise
+due to random sampling of \(\mathbf{A}\) off-diagonal elements and
+finite \(S\). In simulations, the total variation in \(\mathbf{M}\) row
+means that is attributable to \(\mathbf{A}\) is small relative to that
+attributable to \(\mathbf{\gamma}\), especially at high \(S\). Variation
+in \(\mathbf{\gamma}\) specifically isolates the effects of differing
+component response rates, hence causing differences in expected
+\(\mathbf{M}\) row means.
+
+\textbf{Genetic algorithm}. Ideally, to investigate the potential of
+\(Var(\gamma)\) for increasing the proportion of stable complex systems,
+the search space of all possible \(\gamma\) vectors would be evaluated
+for each unique \(\mathbf{M = \gamma A}\). This is technically
+impossible because \(\gamma_{i}\) can take any real value between 0-2,
+but even rounding \(\gamma_{i}\) to reasonable values would result in a
+search space too large to practically explore. Under these conditions,
+genetic algorithms are highly useful tools for finding practical
+solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
+this case, the practical solution is finding vectors of
+\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
+\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
+a large population of 1000 different potential \(\mathbf{\gamma}\)
+vectors and allowing this population to evolve through a process of
+mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
+selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
+found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
+met.
+
+For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
+for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)). The
+genetic algorithm was initialised with a population of 1000 different
+\(\mathbf{\gamma}\) vectors with elements sampled i.i.d from
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed on
+the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector, and
+the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
+lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal
+offspring with subsequent random mutation and crossover between the
+resulting new generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation
+of each \(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+probability of 0.2, resulting in a mutation effect of size
+\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
+\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
+multiplied by \(-1\), and any values that mutated above 2 were set to
+2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
+vectors paired in each generation; vectors were randomly sampled with
+replacement among but not within sets. Vector pairs selected for
+crossover swapped all elements between and including two \(\gamma_{i}\)
+randomly selected with replacement (this allowed for reversal of vector
+element positions during crossover; e.g.,
+\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
+). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
+found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
+fitness increase between generations was less than 0.01 (where fitness
+was defined as \(W_{\gamma} = -\max\left(\Re(\lambda)\right)\) for
+\(\mathbf{M}\)).
+
+\textbf{System feasibility}. Dougoud et
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} identify
+the following feasibility criteria for ecological systems characterised
+by \(S\) interacting species with varying densities in a generalised
+Lotka-Volterra model,
+
+\[\mathbf{n^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]
+
+In the above, \(\mathbf{n^{*}}\) is the vector of species densities at
+equilibrium. Feasibility is satisfied if all elements in
+\(\mathbf{n^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
+identity matrix, and the value \(\theta\) is the strength of
+intraspecific competition (diagonal elements). Diagonal values are set
+to \(-1\), so \(\theta = -1\). The variable \(\delta\) is a
+normalisation parameter that modulates the strength of interactions
+(\(\sigma\)) for \(\mathbf{J}\). Implicitly, here \(\delta = 0\)
+underlying strong interactions. Hence, \((CS)^{-\delta} = 1\), so in the
+above, a diagonal matrix of -1s (\(\theta \mathbf{I}\)) is added to
+\(\mathbf{J}\), which has a diagonal of all zeros and an off-diagonal
+affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
+relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+stability
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
+\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
+purposes \((CS)^{-\delta} = 1\)). Given
+\(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
+reduced to the below (see
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),
+
+\[\mathbf{n^{*} = -A^{-1}r}.\]
+
+To check the feasibility criteria for \(\mathbf{M = \gamma A}\), I
+therefore evaluated \(\mathbf{-M^{-1}r}\) (\(\mathbf{r}\) elements were
+sampled i.i.d. from \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility
+is satisfied if all of the elements of the resulting vector are
+positive.
+
+\textbf{Acknowledgements:} I am supported by a Leverhulme Trust Early
+Career Fellowship (ECF-2016-376). Conversations with L. Bussière and N.
+Bunnefeld, and comments from J. J. Cusack and I. L. Jones, improved the
+quality of this work.
 
@@
 \hypertarget{ref-Bardoscia2017}{}
-9. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G. Pathways
-towards instability in financial networks. \emph{Nature Communications}
-\textbf{8,} 1--7 (2017).
-
-\hypertarget{ref-Mougi2012}{}
-10. Mougi, A. \& Kondoh, M. Diversity of interaction types and
-ecological community stability. \emph{Science} \textbf{337,} 349--351
-(2012).
-
-\hypertarget{ref-Allesina2015a}{}
-11. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
-(2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Gao2016}{}
-12. Gao, J., Barzel, B. \& Barabási, A. L. Universal resilience patterns
-in complex networks. \emph{Nature} \textbf{530,} 307--312 (2016).
-
-\hypertarget{ref-Grilli2017}{}
-13. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
+13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
+Pathways towards instability in financial networks. \emph{Nature
+Communications} \textbf{8,} 1--7 (2017).
+
+\hypertarget{ref-Tao2010}{}
+14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
+circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
+(2010).
 
 \hypertarget{ref-Patel2018}{}
-14. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 
-\hypertarget{ref-Allesina2015}{}
-15. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
-(2015).
-
-\hypertarget{ref-May2008}{}
-16. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
-for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
-
-\hypertarget{ref-Tao2010}{}
-17. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
-(2010).
+\hypertarget{ref-Ahmadian2015}{}
+16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+with partially structured and partially random connectivity.
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{91,} 012820 (2015).
+
+\hypertarget{ref-Gibbs2017}{}
+17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{arXiv} (2017).
+
+\hypertarget{ref-Stone2017}{}
+18. Stone, L. The feasibility and stability of large complex biological
+networks: a random matrix approach. \emph{Scientific Reports}
+\textbf{8,} 8246 (2018).
 
 \hypertarget{ref-Hamblin2013}{}
-18. Hamblin, S. On the practical usage of genetic algorithms in ecology
+19. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
 184--194 (2013).
-
-\hypertarget{ref-Allesina2011}{}
-19. Allesina, S. \& Levine, J. M. A competitive network theory of
-species diversity. \emph{Proceedings of the National Academy of Sciences
-of the United States of America} \textbf{108,} 5638--5642 (2011).
 
@@
 Computational Biology} \textbf{14,} e1005988 (2018).
+
+\hypertarget{ref-Song2018}{}
+21. Song, C. \& Saavedra, S. Will a small randomly assembled community
+be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
+
+\hypertarget{ref-Servan2018}{}
+22. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+Allesina, S. Coexistence of many species in random ecosystems.
+\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).
+
+\hypertarget{ref-May1973}{}
+23. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+\textbf{54,} 638--641 (1973).
 

### 2019-03-02 notebook/ms.Rmd
Prep the manuscript for both Scientific Reports and arxiv (SI still needs to be fixed in the arxiv tex)

--- 2019-03-01 notebook/ms.Rmd
@@
 ---
-title: "Component response rate variation underlies stability in large complex systems"
+title: "Component response rate variation underlies the stability of complex systems"
 author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
@@
 Variation in $\boldsymbol{\gamma}$ becomes increasingly important for stability in large but finite systems, where the probability of stability would otherwise be negligible.
-At such extremes of simulated system size, the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
+At such extremes of simulated system complexity, the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
 My results therefore reveal a previously unconsidered aspect of system stability that is likely to be pervasive across all realistic complex systems.
+
+<!---
 
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $Var(\gamma)$ on system stability, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
+To investigate the effect of $Var(\gamma)$ on stability across systems of varying complexity, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
 One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
@@
 3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
-In other words, the potential for $Var(\gamma)$ to drive stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
+In other words, the potential for $Var(\gamma)$ to affect stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
 For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$.
@@
 To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm.
-Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a large complex system) is too large to search exhaustively [@Hamblin2013].
+Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013].
 Generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
@@
 **System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
-Feasibilility assumes that the system is stable and values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
+Feasibilility assumes that values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
 This is of particular interest for ecological communities because population density ($N$) cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
 While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities [@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
-I want to emphasise that $\gamma$ is not interpreted as population density in this analysis, but instead as a fundamental property of species life history such as expected generation time.
+I emphasise that $\gamma$ is not interpreted as population density in this analysis, but instead as a fundamental property of species life history such as expected generation time.
 Feasibility was unaffected by $Var(\gamma)$ and instead occurred with a fixed probability of $1/2^{S}$, consistent with a recent proof by Serv&aacute;n et al.
@@
 
-Here I have shown that the stability of large systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
-Variation in component response rate is more likely to be critical for stability in systems that are especially large (but finite, see Gibbs et al.[@Gibbs2017]), and it can ultimately increase the probability that system stability is observed above that predicted by May's [@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
+I have shown that the stability of complex systems might often be contigent upon variation in the response rates of their individual components, meaning that factors such as rate of trait evolution (in biological networks), transaction speed (in economic networks), or communication speed (in social networks) need to be considered when investigating the stability of complex systems.
+Variation in component response rate is more likely to be critical for stability in systems that are especially complex, and it can ultimately increase the probability that system stability is observed above that predicted by May's [@May1972] classically derived $\sigma \sqrt{SC}$ criterion.
 The logic outlined here is general, and potentially applies to any complex system in which individual system components can vary in their reaction rates to system perturbation.
 
-It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable.
+It is important to recognise that variation in component response rate is not stabilising per se; that is, adding variation in component response rates to a particular system does not increase the probability that the system will be stable.
 Rather, highly complex systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
@@
 This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the community matrix $\mathbf{M}$ to become negative, but might also increase the real components of eigenvalues.
-In this way, my focus is distinct from that of Gibbs et al.
+1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the system matrix to become negative, but might also increase the real components of eigenvalues.
+
+My focus is distinct from Gibbs et al.
 [@Gibbs2017], who applied the same mathematical framework to investigate how a diagonal matrix $\mathbf{X}$ (equivalent to $\gamma$ in my model) affects the stability of a community matrix $\mathbf{M}$ given an interaction matrix $\mathbf{A}$ within a generalised Lotka-Volterra model, where $\mathbf{M} = \mathbf{XA}$.
@@
 [@Gibbs2017] analytically demonstrated that the effect of $\mathbf{X}$ on system stability decreases exponentially as system size becomes arbitrarily large ($S \to \infty$).
-My numerical results are entirely consistent with this prediction (see Supplementary Information).
-Given that a complex system is known to be stable (or unstable), $Var(\gamma)$ is not expected to increase or decrease the probability of stability.
-Rather, component response rate variation increases the upper bound at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
+My numerical results do not contradict this prediction because I did not scale $\sigma = 1 / \sqrt{S}$, but instead fixed $\sigma$ and increased $S$ to thereby increase total system complexity, $\sigma\sqrt{SC}$ (see Supplemental Information for results simulated across $\sigma$ and $C$).
+Overall, I show that component response rate variation increases the upper bound of complexity at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
 
@@
 Feasibility is important to consider, particularly for the study of ecological networks of species [@Stone2017; @Grilli2017; @Dougoud2018; @Servan2018] because population densities cannot realistically be negative.
-My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the feasibility of purely multi-species interaction networks (see also Supplementary Information).
+My results therefore suggest that variation in the rate of population responses to perturbation (e.g., due to differences in generation time among species) is unlikely to be critical to the stability of purely multi-species interaction networks (see also Supplementary Information).
 Nevertheless, ecological interactions do not exist in isolation in empirical systems [@Patel2018], but instead interact with evolutionary, abiotic, or social-economic systems.
@@
 Overall, I suggest that variation in component response rate might therefore be critical for maintaining stability in many highly complex empirical systems.
-These results are broadly applicable to understanding stability of complex networks across the physical, life, and social sciences.
+These results are broadly applicable for understanding the stability of complex networks across the physical, life, and social sciences.
 
@@
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ modulates the values of components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ defines the values of system components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
 In these analyses, row means of $\mathbf{A}$ are expected to be identical, but variation around this expectation will naturally arise due to random sampling of $\mathbf{A}$ off-diagonal elements and finite $S$.
-In simulations, the total variation in $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ row means that is attributable to $\mathbf{A}$ is small relative to that attributable to $\mathbf{\gamma}$, especially at high $S$.
+In simulations, the total variation in $\mathbf{M}$ row means that is attributable to $\mathbf{A}$ is small relative to that attributable to $\mathbf{\gamma}$, especially at high $S$.
 Variation in $\mathbf{\gamma}$ specifically isolates the effects of differing component response rates, hence causing differences in expected $\mathbf{M}$ row means.
@@
 
-For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$), where $\mathbf{M = \gamma A}$.
+For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$).
 The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
@@
 Dougoud et al.
-[@Dougoud2018] identify the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a classical Lotka-Volterra model,
+[@Dougoud2018] identify the following feasibility criteria for ecological systems characterised by $S$ interacting species with varying densities in a generalised Lotka-Volterra model,
 

### 2019-03-02 notebook/ms.tex
Prep the manuscript for both Scientific Reports and arxiv (SI still needs to be fixed in the arxiv tex)

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies stability in large complex systems},
+            pdftitle={Component response rate variation underlies the stability of complex systems},
             pdfauthor={A. Bradley Duthie ( alexander.duthie@stir.ac.uk )},
@@
 
-  \title{Component response rate variation underlies stability in large complex
+  \title{Component response rate variation underlies the stability of complex
 systems}
@@
 large but finite systems, where the probability of stability would
-otherwise be negligible. At such extremes of simulated system size, the
-largest stable complex systems would be unstable if not for
-\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
-unconsidered aspect of system stability that is likely to be pervasive
-across all realistic complex systems.
+otherwise be negligible. At such extremes of simulated system
+complexity, the largest stable complex systems would be unstable if not
+for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
+previously unconsidered aspect of system stability that is likely to be
+pervasive across all realistic complex systems.
 
@@
 \textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
-investigate the effect of \(Var(\gamma)\) on system stability, I
-simulated random \(\mathbf{M}\) matrices at \(\sigma = 0.4\) and
-\(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One million
-\(\mathbf{M}\) were simulated for each \(S\), and the stability of
-\(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
+investigate the effect of \(Var(\gamma)\) on stability across systems of
+varying complexity, I simulated random \(\mathbf{M}\) matrices at
+\(\sigma = 0.4\) and \(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One
+million \(\mathbf{M}\) were simulated for each \(S\), and the stability
+of \(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
 \(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
@@
 observing a stable system increased with an increase in \(S\). In other
-words, the potential for \(Var(\gamma)\) to drive stability increased
+words, the potential for \(Var(\gamma)\) to affect stability increased
 with system complexity and was most relevant for systems on the cusp of
@@
 potential solutions (in this case, possible combinations of \(\gamma\)
-values leading to stability in a large complex system) is too large to
-search
+values leading to stability in a complex system) is too large to search
 exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
@@
 some tangible quantity, it is relevant to consider the feasibility of
-the system. Feasibilility assumes that the system is stable and values
-of all components are positive at
+the system. Feasibilility assumes that values of all components are
+positive at
 equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
@@
 (e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). I want to emphasise that \(\gamma\) is not interpreted as
+Information). I emphasise that \(\gamma\) is not interpreted as
 population density in this analysis, but instead as a fundamental
@@
 
-Here I have shown that the stability of large systems might often be
+I have shown that the stability of complex systems might often be
 contigent upon variation in the response rates of their individual
@@
 response rate is more likely to be critical for stability in systems
-that are especially large (but finite, see Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}), and it can
-ultimately increase the probability that system stability is observed
-above that predicted by
+that are especially complex, and it can ultimately increase the
+probability that system stability is observed above that predicted by
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
@@
 
-It is important to emphasise that variation in component response rate
+It is important to recognise that variation in component response rate
 is not stabilising per se; that is, adding variation in component
-response rate to a particular system does not necessarily increase the
-probability that the system will be stable (at high \(S\), the opposite
-is true; see Supplementary Information). Rather, highly complex systems
-that are observed to be stable are more likely to have varying component
-response rates, and for this variation to be critical to their stability
-(Fig. 3). This is caused by the shift to a non-uniform distribution of
+response rates to a particular system does not increase the probability
+that the system will be stable. Rather, highly complex systems that are
+observed to be stable are more likely to have varying component response
+rates, and for this variation to be critical to their stability (Fig.
+3). This is caused by the shift to a non-uniform distribution of
 eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
 which can sometimes cause all of the real components of the eigenvalues
-of the community matrix \(\mathbf{M}\) to become negative, but might
-also increase the real components of eigenvalues. In this way, my focus
-is distinct from that of Gibbs et
+of the system matrix to become negative, but might also increase the
+real components of eigenvalues.
+
+My focus is distinct from Gibbs et
 al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
@@
 decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)). My numerical results are not inconsistent with this
+(\(S \to \infty\)). My numerical results do not contradict this
 prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
@@
 system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
-results simulated across \(\sigma\) and \(C\). Overall, my results show
-that component response rate variation increases the upper bound of
+results simulated across \(\sigma\) and \(C\)). Overall, I show that
+component response rate variation increases the upper bound of
 complexity at which stability can be realistically observed, meaning
@@
 responses to perturbation (e.g., due to differences in generation time
-among species) is unlikely to be critical to the feasibility of purely
+among species) is unlikely to be critical to the stability of purely
 multi-species interaction networks (see also Supplementary Information).
@@
 maintaining stability in many highly complex empirical systems. These
-results are broadly applicable to understanding stability of complex
-networks across the physical, life, and social sciences.
+results are broadly applicable for understanding the stability of
+complex networks across the physical, life, and social sciences.
 
@@
 correspond to individual component response rates. Therefore,
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) modulates the values of
+\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the values of system
 components and can be analysed using the techniques of
@@
 due to random sampling of \(\mathbf{A}\) off-diagonal elements and
-finite \(S\). In simulations, the total variation in
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) row means that is
-attributable to \(\mathbf{A}\) is small relative to that attributable to
-\(\mathbf{\gamma}\), especially at high \(S\). Variation in
-\(\mathbf{\gamma}\) specifically isolates the effects of differing
+finite \(S\). In simulations, the total variation in \(\mathbf{M}\) row
+means that is attributable to \(\mathbf{A}\) is small relative to that
+attributable to \(\mathbf{\gamma}\), especially at high \(S\). Variation
+in \(\mathbf{\gamma}\) specifically isolates the effects of differing
 component response rates, hence causing differences in expected
@@
 For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
-for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)), where
-\(\mathbf{M = \gamma A}\). The genetic algorithm was initialised with a
-population of 1000 different \(\mathbf{\gamma}\) vectors with elements
-sampled i.i.d from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis
-was performed on the \(\mathbf{M}\) resulting from each
-\(\mathbf{\gamma}\) vector, and the 20 \(\mathbf{\gamma}\) vectors
-resulting in \(\mathbf{M}\) with the lowest
-\(\max\left(\Re(\lambda)\right)\) each produced 50 clonal offspring with
-subsequent random mutation and crossover between the resulting new
-generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
-\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)). The
+genetic algorithm was initialised with a population of 1000 different
+\(\mathbf{\gamma}\) vectors with elements sampled i.i.d from
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed on
+the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector, and
+the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
+lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal
+offspring with subsequent random mutation and crossover between the
+resulting new generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation
+of each \(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
 probability of 0.2, resulting in a mutation effect of size
@@
 the following feasibility criteria for ecological systems characterised
-by \(S\) interacting species with varying densities in a classical
+by \(S\) interacting species with varying densities in a generalised
 Lotka-Volterra model,
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.726 and 0.742 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.742 for a and b,
 respectively.
@@
 
+\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
+
 \end{document}

### 2019-03-03 notebook/SI.Rmd
Update the SI with new figures showing difference between my model and Gibbs et al

@@
 - \usepackage[utf8]{inputenc}
+- \usepackage{float}
 - \linenumbers
@@
 
-```{r, echo = TRUE}
+```{r, echo = FALSE}
 fourdists <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
@@
 
-Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks.
+Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks [@Servan2018].
 
@@
 
-$$\mathbf{M} = \mathbf{XA}$$.
+$$\mathbf{M} = \mathbf{XA}.$$
 
@@
 
-$$\mathbf{M} = \mathbf{\gamma A}$$.
+$$\mathbf{M} = \mathbf{\gamma A}.$$
 
@@
 In contrast, Gibbs et al.
-[@Gibbs2017] applied analytical techniques to instead focused on a different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$, with $\sigma$ scaled so that $\sigma = 1/\sqrt{S}$.
+[@Gibbs2017] applied analytical techniques to instead focus on a different question concerning the effect of $\mathbf{\gamma}$ on the stability of $\mathbf{M}$ given $\mathbf{A}$ as $S \to \infty$, with $\sigma$ scaled so that $\sigma = 1/\sqrt{S}$.
 Under such scaling, Gibbs et al.
-[@Gibbs2017] showed that the effect of $\gamma$ on stability should decrease exponentially as $S$ increases, which I can replicate below by running simulations in which $\sigma = 1/\sqrt{S}$.
+[@Gibbs2017] showed that the effect of $\gamma$ on stability should decrease exponentially as $S$ increases, which I demonstrate below by running simulations in which $\sigma = 1/\sqrt{S}$.
 
-The results in the table above can be compared to those of the [main results] (#IncrS).
-Note that only 100000, rather than 1 million, simulations are run to confirm consistency with Gibbs et al.
+```{r, echo = FALSE}
+load("../data/constant_complexity.rda");
+kable(constant_complexity);
+```
+
+Above table results can be compared to those of the [main results](#IncrS).
+Note that 100000 (not 1 million), simulations are run to confirm consistency with Gibbs et al.
 [@Gibbs2017].
+The difference between my model and Gibbs et al.
+[@Gibbs2017] is that in the latter, $\sigma\sqrt{SC} = 1$ remains constant with increasing $S$.
+In the former, $\sigma\sqrt{SC}$ increases with $S$, so the expected complexity of the system also increases accordingly.
+Consequently, for the scaled $\sigma$ in the table above, systems are not more likely to be stabilised by $\gamma$ as $S$ increases, consistent with Gibbs et al.
+[@Gibbs2017].
+Note that overall stability does decrease with increasing $S$ due to the increased density of eigenvalues (see below).
+
+```{r, echo = FALSE, fig.height = 4, fig.width = 4, fig.align = "center", fig.cap = "",  fig.pos="H"}
+S_vals     <- 2:32;
+complexity <- 0.4 * sqrt(S_vals);
+par(mfrow = c(1, 1), mar = c(5, 5, 1, 1), oma = c(0.2, 0.2, 0.2, 0.2));
+plot(x = S_vals, y = complexity, type = "l", lwd = 2, xlab = "System size (S)",
+     ylab = expression(paste("System complexity (",sigma,sqrt(SC),")")),
+     cex.lab = 1.5, cex.axis = 1.5);
+abline(h = 1, lwd = 2, lty = "dashed");
+```
+
+**Complexity as a function of $S$ in the main text (solid) versus in Gibbs et al. [@Gibbs2017] (dashed).**
+
+When the complexity is scaled to $\sigma\sqrt{SC} = 1$, an increase in $S$ increases the eigenvalue density within a circle with a unit radius centred at $(-1, 0)$ on the complex plane.
+As $S \to \infty$, this circle becomes increasingly saturated.
+Gibbs et al.
+[@Gibbs2017] showed that a diagonal matrix $\mathbf{\gamma}$ will have an exponentially decreasing effect on stability with increasing $S$.
+Increasing $S$ is visualised below, first with a system size $S = 100$.
+
+```{r, echo = FALSE}
+comp_scale <- function(S, C = 1, sigma = 0.4){
+  complx <- sigma * sqrt(S * C);
+  A_comp <- NULL;
+  A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
+  A_mat  <- matrix(data = A_dat, nrow = S);
+  C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+  C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+  A_mat     <- A_mat * C_mat;
+  gammas <- runif(n = S, min = 0, max = 2);
+  mu_gam <- mean(gammas);
+  diag(A_mat) <- -1;
+  A1     <- gammas * A_mat;
+  A0     <- mu_gam * A_mat;
+  A0_e   <- eigen(A0)$values;
+  A0_r   <- Re(A0_e);
+  A0_i   <- Im(A0_e);
+  A1_e   <- eigen(A1)$values;
+  A1_r   <- Re(A1_e);
+  A1_i   <- Im(A1_e);
+  A0_vm       <- A0;
+  diag(A0_vm) <- NA;
+  A0vec       <- as.vector(A0_vm);
+  A0vec       <- A0vec[is.na(A0vec) == FALSE];
+  A1_vm       <- A1;
+  diag(A1_vm) <- NA;
+  A1vec       <- as.vector(A1_vm);
+  A1vec       <- A1vec[is.na(A1vec) == FALSE];
+  par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+  plot(A0_r, A0_i, xlim = c(-2.75 * complx, 0.5 * complx), 
+       ylim = c(-1.5,1.5), pch = 4, 
+       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+       asp = 1, col = "dodgerblue4");
+  abline(v = 0, lty = "dotted");
+  vl <- seq(from = 0, to = 2*pi, by = 0.001);
+  x0 <- sqrt(S*C) * sigma * cos(vl) - 1;
+  y0 <- sqrt(S*C) * sigma * sin(vl);
+  x1 <- sqrt(S*C) * sigma * cos(vl) - 1;
+  y1 <- sqrt(S*C) * sigma * sin(vl);
+  text(x = -15.5, y = 19, labels = "a", cex = 2);
+  points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
+  plot(A1_r, A1_i, xlim = c(-2.75*complx, 0.5*complx), 
+       ylim = c(-1.5, 1.5), pch = 4, 
+       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+       asp = 1, col = "firebrick", yaxt = "n");
+  abline(v = 0, lty = "dotted");
+  text(x = -15.5, y = 19, labels = "b", cex = 2);
+  mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+  mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+```
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_scale(S = 100, C = 1, sigma = 1 / sqrt(100));
+```
+
+The left panel above shows the distribution of eigenvalues; the blue ellipse shows the unit radius within which eigenvalues are expected to be contained.
+The right panel shows how eigenvalue distributions change given $\gamma \sim \mathcal{U}(0,2)$.
+The vertical dotted line shows the threshold of stability, $\Re = 0$.
+Increasing to $S = 200$, the scaling $\sigma = 1 / \sqrt{S}$ maintains the expected distribution of eigenvalues but increases eigenvalue density.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_scale(S = 200, C = 1, sigma = 1 / sqrt(200));
+```
+
+We can increase the system size to $S = 500$ and see the corresponding increase in eigenvalue density.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_scale(S = 500, C = 1, sigma = 1 / sqrt(500));
+```
+
+Finally, below shows a increase in system size to $S = 1000$.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_scale(S = 1000, C = 1, sigma = 1 / sqrt(1000));
+```
+
+In contrast, in the model of the main text, the complexity of system is not scaled to $\sigma\sqrt{SC} = 1$.
+Rather, the density of eigenvalues within a circle centred at $(-1, 0)$ with a radius $\sigma\sqrt{SC}$ is held constant such that there are $S / \pi(\sigma\sqrt{SC})^2$ eigenvalues per unit area of the circle.
+As $S$ increases, so does the expected complexity of the system, but the density of eigenvalues remains finite causing error around this expectation.
+Below shows a system where $S = 100$, $C = 0.0625$, and $\sigma = 0.4$, where $\sigma \sqrt{SC} = 1$ (identical to the first example distribution above in which $S = 100$ and $\sigma = 1/\sqrt{S}$).
+
+```{r, echo = FALSE}
+comp_dens <- function(S, C = 1, sigma = 0.4){
+  complx <- sigma * sqrt(S * C);
+  A_comp <- NULL;
+  A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
+  A_mat  <- matrix(data = A_dat, nrow = S);
+  C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+  C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+  A_mat     <- A_mat * C_mat;
+  gammas <- runif(n = S, min = 0, max = 2);
+  mu_gam <- mean(gammas);
+  diag(A_mat) <- -1;
+  A1     <- gammas * A_mat;
+  A0     <- mu_gam * A_mat;
+  A0_e   <- eigen(A0)$values;
+  A0_r   <- Re(A0_e);
+  A0_i   <- Im(A0_e);
+  A1_e   <- eigen(A1)$values;
+  A1_r   <- Re(A1_e);
+  A1_i   <- Im(A1_e);
+  A0_vm       <- A0;
+  diag(A0_vm) <- NA;
+  A0vec       <- as.vector(A0_vm);
+  A0vec       <- A0vec[is.na(A0vec) == FALSE];
+  A1_vm       <- A1;
+  diag(A1_vm) <- NA;
+  A1vec       <- as.vector(A1_vm);
+  A1vec       <- A1vec[is.na(A1vec) == FALSE];
+  par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
+  plot(A0_r, A0_i, xlim = c(-5, 3), 
+       ylim = c(-4,4), pch = 4, 
+       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+       asp = 1, col = "dodgerblue4");
+  abline(v = 0, lty = "dotted");
+  vl <- seq(from = 0, to = 2*pi, by = 0.001);
+  x0 <- sqrt(S*C) * sigma * cos(vl) - 1;
+  y0 <- sqrt(S*C) * sigma * sin(vl);
+  x1 <- sqrt(S*C) * sigma * cos(vl) - 1;
+  y1 <- sqrt(S*C) * sigma * sin(vl);
+  text(x = -15.5, y = 19, labels = "a", cex = 2);
+  points(x = x0, y = y0, type = "l", lwd = 3, col = "dodgerblue4");
+  plot(A1_r, A1_i, xlim = c(-4, 4), 
+       ylim = c(-1.5, 1.5), pch = 4, 
+       cex = 0.7, xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+       asp = 1, col = "firebrick", yaxt = "n");
+  abline(v = 0, lty = "dotted");
+  text(x = -15.5, y = 19, labels = "b", cex = 2);
+  mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+  mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+}
+```
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_dens(S = 100, C = 0.0625, sigma = 0.4);
+```
+
+Now when $S$ is increased to $200$ while keeping $C = 0.0625$ and $\sigma = 0.4$, the area of the circle within which eigenvalues are contained increases to keep the density of eigenvalues constant.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_dens(S = 200, C = 0.0625, sigma = 0.4);
+```
+
+Note that the expected distribution of eigenvalues increases so that the threshold $\Re = 0$ is exceeded. Below, system size is increased to $S = 500$.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_dens(S = 500, C = 0.0625, sigma = 0.4);
+```
+
+Finally, $S = 1000$ is shown below.
+Again, the density of eigenvalues per unit remains constant at ca 2, but the system has increased in complexity such that some real components of eigenvalues are almost assured to be greater than zero.
+
+```{r, echo = FALSE, fig.height = 3.9}
+comp_dens(S = 1000, C = 0.0625, sigma = 0.4);
+```
 

### 2019-03-03 notebook/SI.tex
Update the SI with new figures showing difference between my model and Gibbs et al

--- 2018-07-30 notebook/SI.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{float}
 \linenumbers
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
-\KeywordTok{kable}\NormalTok{(fourdists);}
-\end{Highlighting}
-\end{Shaded}
-
 \begin{longtable}[]{@{}lllllll@{}}
@@
 Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
-in pure species interaction networks.
+in pure species interaction
+networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.
+
+\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
+(2018)}\label{Gibbs}}
+
+The question that I address in the main text is distinct from that of
+Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
+focused instead on the effect of a diagonal matrix of biological species
+densities \(\mathbf{X}\) on a community matrix \(\mathbf{M}\) given a
+species interaction matrix \(\mathbf{A}\). This is modelled as below,
+
+\[\mathbf{M} = \mathbf{XA}.\]
+
+Mathematically, the above is identical to my model in the main text
+where the system \(\mathbf{M}\) is defined by component interaction
+strengths \(\mathbf{A}\) and individual component response rates
+\(\mathbf{\gamma}\),
+
+\[\mathbf{M} = \mathbf{\gamma A}.\]
+
+I focused on the probability of observing a stable versus unstable
+system given variation in \(\mathbf{\gamma}\) as system complexity
+(\(\sigma\sqrt{SC}\)) increased. I increased system complexity by
+holding \(C\) and \(\sigma\) constant and incrementally increasing \(S\)
+to obtain numerical results. In contrast, Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} applied
+analytical techniques to instead focus on a different question
+concerning the effect of \(\mathbf{\gamma}\) on the stability of
+\(\mathbf{M}\) given \(\mathbf{A}\) as \(S \to \infty\), with \(\sigma\)
+scaled so that \(\sigma = 1/\sqrt{S}\). Under such scaling, Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that
+the effect of \(\gamma\) on stability should decrease exponentially as
+\(S\) increases, which I demonstrate below by running simulations in
+which \(\sigma = 1/\sqrt{S}\).
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 3111 & 96889 & 3111 & 96889 & 0 & 0\tabularnewline
+3 & 5203 & 94797 & 5237 & 94763 & 1 & 35\tabularnewline
+4 & 6743 & 93257 & 6818 & 93182 & 6 & 81\tabularnewline
+5 & 7889 & 92111 & 8005 & 91995 & 20 & 136\tabularnewline
+6 & 8834 & 91166 & 8991 & 91009 & 55 & 212\tabularnewline
+7 & 9885 & 90115 & 10072 & 89928 & 81 & 268\tabularnewline
+8 & 10516 & 89484 & 10764 & 89236 & 108 & 356\tabularnewline
+9 & 11135 & 88865 & 11383 & 88617 & 145 & 393\tabularnewline
+10 & 11819 & 88181 & 12095 & 87905 & 181 & 457\tabularnewline
+11 & 12414 & 87586 & 12700 & 87300 & 213 & 499\tabularnewline
+12 & 12865 & 87135 & 13136 & 86864 & 283 & 554\tabularnewline
+13 & 13530 & 86470 & 13836 & 86164 & 324 & 630\tabularnewline
+14 & 13745 & 86255 & 14042 & 85958 & 362 & 659\tabularnewline
+15 & 14401 & 85599 & 14720 & 85280 & 387 & 706\tabularnewline
+16 & 14793 & 85207 & 15123 & 84877 & 428 & 758\tabularnewline
+17 & 15004 & 84996 & 15356 & 84644 & 444 & 796\tabularnewline
+18 & 15361 & 84639 & 15735 & 84265 & 472 & 846\tabularnewline
+19 & 16062 & 83938 & 16303 & 83697 & 592 & 833\tabularnewline
+20 & 15814 & 84186 & 16184 & 83816 & 566 & 936\tabularnewline
+21 & 16171 & 83829 & 16492 & 83508 & 640 & 961\tabularnewline
+22 & 16671 & 83329 & 17049 & 82951 & 641 & 1019\tabularnewline
+23 & 17000 & 83000 & 17291 & 82709 & 718 & 1009\tabularnewline
+24 & 17411 & 82589 & 17666 & 82334 & 765 & 1020\tabularnewline
+25 & 17414 & 82586 & 17742 & 82258 & 783 & 1111\tabularnewline
+26 & 17697 & 82303 & 18027 & 81973 & 806 & 1136\tabularnewline
+27 & 18010 & 81990 & 18316 & 81684 & 880 & 1186\tabularnewline
+28 & 18584 & 81416 & 18735 & 81265 & 1008 & 1159\tabularnewline
+29 & 18401 & 81599 & 18572 & 81428 & 942 & 1113\tabularnewline
+30 & 18497 & 81503 & 18754 & 81246 & 952 & 1209\tabularnewline
+31 & 18744 & 81256 & 18942 & 81058 & 991 & 1189\tabularnewline
+32 & 18936 & 81064 & 19194 & 80806 & 1022 & 1280\tabularnewline
+33 & 19174 & 80826 & 19346 & 80654 & 1113 & 1285\tabularnewline
+34 & 19477 & 80523 & 19632 & 80368 & 1120 & 1275\tabularnewline
+35 & 19659 & 80341 & 19777 & 80223 & 1206 & 1324\tabularnewline
+36 & 19883 & 80117 & 19929 & 80071 & 1275 & 1321\tabularnewline
+37 & 20275 & 79725 & 20348 & 79652 & 1308 & 1381\tabularnewline
+38 & 20067 & 79933 & 20190 & 79810 & 1275 & 1398\tabularnewline
+39 & 20416 & 79584 & 20516 & 79484 & 1340 & 1440\tabularnewline
+40 & 20370 & 79630 & 20489 & 79511 & 1359 & 1478\tabularnewline
+41 & 20295 & 79705 & 20430 & 79570 & 1382 & 1517\tabularnewline
+42 & 20767 & 79233 & 20839 & 79161 & 1418 & 1490\tabularnewline
+43 & 20688 & 79312 & 20705 & 79295 & 1471 & 1488\tabularnewline
+44 & 21049 & 78951 & 21028 & 78972 & 1555 & 1534\tabularnewline
+45 & 21114 & 78886 & 21034 & 78966 & 1572 & 1492\tabularnewline
+46 & 21163 & 78837 & 21195 & 78805 & 1463 & 1495\tabularnewline
+47 & 21373 & 78627 & 21353 & 78647 & 1535 & 1515\tabularnewline
+48 & 21338 & 78662 & 21285 & 78715 & 1632 & 1579\tabularnewline
+49 & 21547 & 78453 & 21566 & 78434 & 1575 & 1594\tabularnewline
+50 & 21738 & 78262 & 21633 & 78367 & 1636 & 1531\tabularnewline
+51 & 21967 & 78033 & 21892 & 78108 & 1698 & 1623\tabularnewline
+\bottomrule
+\end{longtable}
+
+Above table results can be compared to those of the
+\protect\hyperlink{IncrS}{main results}. Note that 100000 (not 1
+million), simulations are run to confirm consistency with Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. The
+difference between my model and Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} is that in the
+latter, \(\sigma\sqrt{SC} = 1\) remains constant with increasing \(S\).
+In the former, \(\sigma\sqrt{SC}\) increases with \(S\), so the expected
+complexity of the system also increases accordingly. Consequently, for
+the scaled \(\sigma\) in the table above, systems are not more likely to
+be stabilised by \(\gamma\) as \(S\) increases, consistent with Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. Note that
+overall stability does decrease with increasing \(S\) due to the
+increased density of eigenvalues (see below).
+
+\begin{figure}[H]
+
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-27-1}
+
+}
+
+\end{figure}
+
+\textbf{Complexity as a function of \(S\) in the main text (solid)
+versus in Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} (dashed).}
+
+When the complexity is scaled to \(\sigma\sqrt{SC} = 1\), an increase in
+\(S\) increases the eigenvalue density within a circle with a unit
+radius centred at \((-1, 0)\) on the complex plane. As \(S \to \infty\),
+this circle becomes increasingly saturated. Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that a
+diagonal matrix \(\mathbf{\gamma}\) will have an exponentially
+decreasing effect on stability with increasing \(S\). Increasing \(S\)
+is visualised below, first with a system size \(S = 100\).
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-29-1.pdf}
+
+The left panel above shows the distribution of eigenvalues; the blue
+ellipse shows the unit radius within which eigenvalues are expected to
+be contained. The right panel shows how eigenvalue distributions change
+given \(\gamma \sim \mathcal{U}(0,2)\). The vertical dotted line shows
+the threshold of stability, \(\Re = 0\). Increasing to \(S = 200\), the
+scaling \(\sigma = 1 / \sqrt{S}\) maintains the expected distribution of
+eigenvalues but increases eigenvalue density.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
+
+We can increase the system size to \(S = 500\) and see the corresponding
+increase in eigenvalue density.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
+
+Finally, below shows a increase in system size to \(S = 1000\).
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
+
+In contrast, in the model of the main text, the complexity of system is
+not scaled to \(\sigma\sqrt{SC} = 1\). Rather, the density of
+eigenvalues within a circle centred at \((-1, 0)\) with a radius
+\(\sigma\sqrt{SC}\) is held constant such that there are
+\(S / \pi(\sigma\sqrt{SC})^2\) eigenvalues per unit area of the circle.
+As \(S\) increases, so does the expected complexity of the system, but
+the density of eigenvalues remains finite causing error around this
+expectation. Below shows a system where \(S = 100\), \(C = 0.0625\), and
+\(\sigma = 0.4\), where \(\sigma \sqrt{SC} = 1\) (identical to the first
+example distribution above in which \(S = 100\) and
+\(\sigma = 1/\sqrt{S}\)).
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-34-1.pdf}
+
+Now when \(S\) is increased to \(200\) while keeping \(C = 0.0625\) and
+\(\sigma = 0.4\), the area of the circle within which eigenvalues are
+contained increases to keep the density of eigenvalues constant.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
+
+Note that the expected distribution of eigenvalues increases so that the
+threshold \(\Re = 0\) is exceeded. Below, system size is increased to
+\(S = 500\).
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
+
+Finally, \(S = 1000\) is shown below. Again, the density of eigenvalues
+per unit remains constant at ca 2, but the system has increased in
+complexity such that some real components of eigenvalues are almost
+assured to be greater than zero.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
 
@@
 
+\hypertarget{ref-Servan2018}{}
+4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+Allesina, S. Coexistence of many species in random ecosystems.
+\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).
+
+\hypertarget{ref-Gibbs2017}{}
+5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{arXiv} (2017).
+
 \end{document}

### 2019-03-03 notebook/arxiv/ms.tex
Tweaks

@@
 increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
-realistic complex systems (such that \(S\) is high but finite) in which
-\(\gamma\) varies are actually more likely to be stable.
+realistic complex systems (such that \(S\) is high but finite) are actually 
+more likely to be stable if their individual component response rates vary.
 
@@
 instead fixed \(\sigma\) and increased \(S\) to thereby increase total
-system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
+system complexity, \(\sigma\sqrt{SC}\) (see Supplementary Information for
 results simulated across \(\sigma\) and \(C\)). Overall, I show that
@@
 \(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),
+reduced to the below\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}},
 
@@
 population abundances on the stability of large random ecosystems.
-\emph{arXiv} (2017).
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{98,} 022410 (2018).
 
@@
 
-\begin{center}
-\hypertarget{SIstart}{\section{Supplementary Information}\label{SIstart}}
-\end{center}
-
 \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
 
-\textbf{This supplementary information supports the manuscript
+\textbf{This Supplementary information supports the manuscript
 ``Component response rate variation drives stability in large complex
-systems'' with all of the code required to recreate the analysis in the
-main text, and with additional analyses to support its conclusions. All
-text, code, and data underlying this manuscript are publicly available
-on \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}
-as part of the RandomMatrixStability package.}
+systems'' with additional analyses to support its conclusions. All text,
+code, and data underlying this manuscript are publicly available on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
+part of the RandomMatrixStability R package.}
 
@@
 package} includes all functions and tools for recreating the text, this
-supplemental information, and running all code; additional documentation
-is also provided for functions as part of the package. The
-RandomMatrixStability package is available on
+Supplementary information, and running all code; additional documentation
+is also provided for package functions. The RandomMatrixStability
+package is available on
 \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}; to
@@
 
-While downloading this package is recommended, all relevant code is also
-reproduced below with explanation, so it is possible to recreate all
-analyses using only this Supplementary Information.
-
 \begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
@@
 \section{Supplementary Information table of
-contents}\label{supplemental-information-table-of-contents}
+contents}\label{Supplementary-information-table-of-contents}
 
@@
 \item
-  \protect\hyperlink{moregamma}{Further explanation of \(\gamma\)}
+  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
 \item
-  \protect\hyperlink{Fig1}{Code and simulations underlying Fig. 1}
-\item
-  \protect\hyperlink{Fig2}{Code and simulations underlying Fig. 2}
-\item
-  \protect\hyperlink{IncrS}{Stability across increasing \(S\)}
+  \protect\hyperlink{ga}{Stability given targeted manipulation of
+  \(\gamma\) (genetic algorithm)}
 \item
@@
 \item
-  \protect\hyperlink{gam_dist}{Different distributions of \(\gamma\)}
+  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
+  and \(\sigma\)}
+
+  \begin{itemize}
+  \tightlist
   \item
-  \protect\hyperlink{ga}{Genetic algorithm}
+    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
+  \item
+    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
+  \item
+    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
+  \item
+    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
+  \end{itemize}
+\item
+  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
+  \(\gamma\)}
 \item
   \protect\hyperlink{Feasibility}{Feasibility of complex systems}
+\item
+  \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
 \end{itemize}
-
-\hypertarget{moregamma}{\section{\texorpdfstring{Further explanation of
-\(\gamma\)}{Further explanation of \textbackslash{}gamma}}\label{moregamma}}
-
-In a synthesis of eco-evolutionary feedbacks on community stability,
-Patel et al. model a system that includes a vector of potentially
-changing species densities (\(\mathbf{N}\)) and a vector of potentially
-evolving traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
-For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
-
-\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
-
-In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
-effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
-were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}, and
-this is modulated in the above by the scalar \(\epsilon\). The value of
-\(\epsilon\) thereby determines the timescale separation between ecology
-and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{1}}.
-
-I use the same principle that Patel et al. use to modulate the relative
-rate of evolution to modulate rates of component responses for \(S\)
-components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}},
-the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
-by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
-\(i\) (\(m_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
-
-\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}m_{ij}v_{j}(t).\]
-
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{3}},
-
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{M}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{\gamma} \mathbf{M}\) modulates the values of components and
-can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{2},\protect\hyperlink{ref-May1973}{3}}.
-
-\hypertarget{Fig1}{\section{Code and simulations underlying Fig.
-1}\label{Fig1}}
-
-The sample \(M\) used for the eigenvalue distributions in Fig. 1 of the
-text is available on
-\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/notebook/sim_results/bi_gamma}{GitHub},
-and was produced by running the following function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{find_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) >=}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \KeywordTok{return}\NormalTok{(}\KeywordTok{list}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1));}
-            \NormalTok{break;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(iters);}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above \texttt{find\_bgamma} function terminates when a matrix \(M\)
-is found that is not stable when all component response rates are set to
-\(\gamma = 1\), but is stable when half of component response rates are
-\(1.95\) and half are \(0.05\). The function is used to illustrate the
-concept of how fast versus slow component responses can cause a system
-to become stable. Simulations were run for \texttt{iter\ =\ 1000000},
-but terminated once an acceptable \texttt{A0} and \texttt{A1} were
-found. The code below plots the eigenvalue distributions of \texttt{A0}
-and \texttt{A1} in panels \textbf{a} and \textbf{b}, respectively. The
-plot itself can be recreated with the function and code below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{A0 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A0[,-}\DecValTok{1}\NormalTok{]);}
-\NormalTok{A1 <-}\StringTok{ }\KeywordTok{as.matrix}\NormalTok{(A1[,-}\DecValTok{1}\NormalTok{]);}
-\NormalTok{plot_Fig_1 <-}\StringTok{ }\NormalTok{function(A0, A1)\{}
-    \NormalTok{S_val       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A0)[}\DecValTok{1}\NormalTok{];}
-    \NormalTok{A0_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-    \NormalTok{A0_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-    \NormalTok{A0_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-    \NormalTok{A1_e        <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-    \NormalTok{A1_r        <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-    \NormalTok{A1_i        <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
-    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A0_vm));}
-    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
-    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(}\KeywordTok{t}\NormalTok{(A1_vm));}
-    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{fhalf       <-}\StringTok{ }\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec));}
-    \NormalTok{shalf       <-}\StringTok{ }\NormalTok{(}\FloatTok{0.5}\NormalTok{*}\KeywordTok{length}\NormalTok{(A1vec)+}\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(A1vec);}
-    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{vl   <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{A0x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
-    \NormalTok{A0y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x0, }\DataTypeTok{y =} \NormalTok{A0y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(A0_r, A0_i, }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{3.7}\NormalTok{, }\FloatTok{0.3}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }
-         \DataTypeTok{col =} \StringTok{"black"}\NormalTok{, }\DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
-    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{A0x1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{:(}\FloatTok{0.5}\NormalTok{*S_val)]);}
-    \NormalTok{A0y1a <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[fhalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1a, }\DataTypeTok{y =} \NormalTok{A0y1a, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \NormalTok{A0x1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }
-\StringTok{                  }\KeywordTok{mean}\NormalTok{( }\KeywordTok{diag}\NormalTok{(A1)[( (}\FloatTok{0.5}\NormalTok{*S_val) +}\StringTok{ }\DecValTok{1} \NormalTok{):S_val] );}
-    \NormalTok{A0y1b <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\FloatTok{0.5}\NormalTok{*S_val) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec[shalf]) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{A0x1b, }\DataTypeTok{y =} \NormalTok{A0y1b, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{col =} \StringTok{"grey"}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(A1_r[}\DecValTok{1}\NormalTok{:S_val], A1_i[}\DecValTok{1}\NormalTok{:S_val],}\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{);   }
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{3.5}\NormalTok{, }\DataTypeTok{y =} \FloatTok{2.25}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-\NormalTok{\}}
-\KeywordTok{plot_Fig_1}\NormalTok{(}\DataTypeTok{A0 =} \NormalTok{A0, }\DataTypeTok{A1 =} \NormalTok{A1);}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{unnamed-chunk-4-1.pdf}
-
-To find out how frequently \(M\) was stable given that all
-\(\gamma = 1\) versus \(\gamma = \{1.95, 0.05\}\), the function below
-was created.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{stab_bgamma <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{200}\NormalTok{, }\DataTypeTok{C =} \FloatTok{0.05}\NormalTok{, }\DataTypeTok{Osd =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{10000}\NormalTok{)\{}
-    \NormalTok{ress     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iters, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{A0_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{A1_count <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{while(iters >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-        \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{Osd);}
-        \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \NormalTok{S);}
-        \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-        \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{S, }\DataTypeTok{ncol =} \NormalTok{S);}
-        \NormalTok{A_mat  <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-        \NormalTok{gammas <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\FloatTok{1.95}\NormalTok{, S/}\DecValTok{2}\NormalTok{), }\KeywordTok{rep}\NormalTok{(}\FloatTok{0.05}\NormalTok{, S/}\DecValTok{2}\NormalTok{))}
-        \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-        \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-        \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-        \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-        \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-        \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-        \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-        \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-        \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A0_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A0_count       <-}\StringTok{ }\NormalTok{A0_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-        \NormalTok{if(}\KeywordTok{max}\NormalTok{(A1_r) <}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{ress[iters, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{A1_count       <-}\StringTok{ }\NormalTok{A1_count +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(}\KeywordTok{c}\NormalTok{(iters, A0_count, A1_count));}
-        \NormalTok{iters <-}\StringTok{ }\NormalTok{iters -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(ress);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above functions produced the \texttt{bi\_pr\_st} data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{bi_pr_st <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\StringTok{"sim_results/bi_gamma/bi_pr_st.csv"}\NormalTok{);}
-\NormalTok{pr_st    <-}\StringTok{ }\NormalTok{bi_pr_st[,-}\DecValTok{1}\NormalTok{];}
-\end{Highlighting}
-\end{Shaded}
-
-The function \texttt{stab\_bgamma} was run for
-\texttt{iters\ =\ 1000000}, and the resulting matrix \texttt{ress} was
-returned. Each row of \texttt{ress} represents a single \(M\) given
-\(\gamma = 1\) (column 1) versus \(\gamma = \{1.95, 0.05\}\) (column 2).
-Values of 0 indicate that \(M\) was found to be unstable (at least one
-real component of its eigenvalues greater than or equal to zero),
-whereas values of 1 indicate that \(M\) was found to be stable (all real
-components of eigenvalues are negative). The frequencies of stable \(M\)
-were 1 given \(\gamma = 1\) and 32 given \(\gamma = \{1.95, 0.05\}\), as
-reported in the main text and legend of Fig. 1 (raw data are
-\href{https://github.com/bradduthie/RandomMatrixStability/blob/master/notebook/sim_results/bi_gamma/bi_pr_st.csv}{available
-on GitHub}).
-
-\hypertarget{Fig2}{\section{Code and simulations underlying Fig.
-2}\label{Fig2}}
-
-Figure 2 of the main text shows eigenvalue distributions in a system
-where \(S = 1000\), \(C = 1\), and \(\sigma = 0.4\). Eigenvalues can be
-reproduced using the code below for when \(\gamma = 1\) (panel a) and
-\(\gamma \sim \mathcal{U}(0, 2)\) (panel b). The function below
-reproduces the figure.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{plot_Fig_2 <-}\StringTok{ }\NormalTok{function()\{}
-    \NormalTok{A_comp <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{A_dat  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-    \NormalTok{A_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{C_dat  <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000} \NormalTok{*}\StringTok{ }\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{C_mat  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \DecValTok{1000}\NormalTok{);}
-    \NormalTok{A_mat     <-}\StringTok{ }\NormalTok{A_mat *}\StringTok{ }\NormalTok{C_mat;}
-    \NormalTok{gammas <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-    \NormalTok{mu_gam <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(gammas);}
-    \KeywordTok{diag}\NormalTok{(A_mat) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-    \NormalTok{A1     <-}\StringTok{ }\NormalTok{gammas *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0     <-}\StringTok{ }\NormalTok{mu_gam *}\StringTok{ }\NormalTok{A_mat;}
-    \NormalTok{A0_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A0)$values;}
-    \NormalTok{A0_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A0_e);}
-    \NormalTok{A0_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A0_e);}
-    \NormalTok{A1_e   <-}\StringTok{ }\KeywordTok{eigen}\NormalTok{(A1)$values;}
-    \NormalTok{A1_r   <-}\StringTok{ }\KeywordTok{Re}\NormalTok{(A1_e);}
-    \NormalTok{A1_i   <-}\StringTok{ }\KeywordTok{Im}\NormalTok{(A1_e);}
-    \NormalTok{A0_vm       <-}\StringTok{ }\NormalTok{A0;}
-    \KeywordTok{diag}\NormalTok{(A0_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A0vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A0_vm);}
-    \NormalTok{A0vec       <-}\StringTok{ }\NormalTok{A0vec[}\KeywordTok{is.na}\NormalTok{(A0vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \NormalTok{A1_vm       <-}\StringTok{ }\NormalTok{A1;}
-    \KeywordTok{diag}\NormalTok{(A1_vm) <-}\StringTok{ }\OtherTok{NA}\NormalTok{;}
-    \NormalTok{A1vec       <-}\StringTok{ }\KeywordTok{as.vector}\NormalTok{(A1_vm);}
-    \NormalTok{A1vec       <-}\StringTok{ }\NormalTok{A1vec[}\KeywordTok{is.na}\NormalTok{(A1vec) ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{];}
-    \KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{), }\DataTypeTok{mar =} \KeywordTok{c}\NormalTok{(}\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\DecValTok{5}\NormalTok{, }\DecValTok{5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{));}
-    \KeywordTok{plot}\NormalTok{(A0_r, A0_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }
-         \DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }
-         \DataTypeTok{asp =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{vl <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \DecValTok{0}\NormalTok{, }\DataTypeTok{to =} \DecValTok{2}\NormalTok{*pi, }\DataTypeTok{by =} \FloatTok{0.001}\NormalTok{);}
-    \NormalTok{x0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A0));}
-    \NormalTok{y0 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A0vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \NormalTok{x1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{cos}\NormalTok{(vl) +}\StringTok{ }\KeywordTok{mean}\NormalTok{(}\KeywordTok{diag}\NormalTok{(A1));}
-    \NormalTok{y1 <-}\StringTok{ }\KeywordTok{sqrt}\NormalTok{(}\DecValTok{1000}\NormalTok{) *}\StringTok{ }\KeywordTok{sd}\NormalTok{(A1vec) *}\StringTok{ }\KeywordTok{sin}\NormalTok{(vl);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"a"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
-    \KeywordTok{plot}\NormalTok{(A1_r, A1_i, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{, }\FloatTok{15.5}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(-}\FloatTok{16.5}\NormalTok{,}\FloatTok{15.5}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{4}\NormalTok{, }\DataTypeTok{cex =} \FloatTok{0.7}\NormalTok{,}
-         \DataTypeTok{xlab =} \StringTok{""}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{""}\NormalTok{, }\DataTypeTok{cex.lab =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex.axis =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{asp =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{,}
-         \DataTypeTok{yaxt =} \StringTok{"n"}\NormalTok{);}
-    \KeywordTok{text}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\FloatTok{15.5}\NormalTok{, }\DataTypeTok{y =} \DecValTok{19}\NormalTok{, }\DataTypeTok{labels =} \StringTok{"b"}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x1, }\DataTypeTok{y =} \NormalTok{y1, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{col =} \StringTok{"red"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{)}
-    \KeywordTok{points}\NormalTok{(}\DataTypeTok{x =} \NormalTok{x0, }\DataTypeTok{y =} \NormalTok{y0, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{, }\DataTypeTok{lwd =} \DecValTok{3}\NormalTok{, }\DataTypeTok{lty =} \StringTok{"dashed"}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{1}\NormalTok{, }\StringTok{"Real"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \DecValTok{3}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-    \KeywordTok{mtext}\NormalTok{(}\DataTypeTok{side =} \DecValTok{2}\NormalTok{, }\StringTok{"Imaginary"}\NormalTok{, }\DataTypeTok{outer =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{line =} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{cex =} \DecValTok{2}\NormalTok{);}
-\NormalTok{\}}
-\KeywordTok{plot_Fig_2}\NormalTok{();}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{unnamed-chunk-8-1.pdf}
 
@@
 Figure 3 of the main text reports the number of stable random complex
-systems found over 1 million iterations. The data used to make this
-figure are read into R below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_1/random_all.csv"}\NormalTok{);}
-\NormalTok{dat <-}\StringTok{ }\NormalTok{dat[,-}\DecValTok{1}\NormalTok{]; }\CommentTok{# Extra row-indicating column removed}
-\end{Highlighting}
-\end{Shaded}
-
-The table below shows the results for all simulations of random \(M\)
-matrices at \(\sigma = 0.4\) and \(C = 1\) given a range of
+systems found over 1 million iterations. The table below shows the
+results for all simulations of random \(\mathbf{M}\) matrices at
+\(\sigma = 0.4\) and \(C = 1\) given a range of
 \(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
@@
 Each row summarises data for a given \(S\) over 1 million randomly
-simulated \(M\) (\texttt{A0} and \texttt{A1}). The column
+simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
 \texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
@@
 Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
-matrices found is greater than 1 (compare column 5 to column 3), and
-this ratio increases with increasing \(S\) (column 1). Hence, more
-randomly created complex systems (\(M\)) are generated given variation
-in \(\gamma\) than when \(\gamma = 1\). The results underlying this
-table were produced with the \texttt{rand\_gen\_var} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The above function calls the two functions
-\texttt{species\_interactions} and \texttt{summarise\_randmat}, which
-are provided below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-
-\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{13}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \CommentTok{# Stable and unstable}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Stabilised and destabilised}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \CommentTok{# Feasible and infeasible}
-        \NormalTok{all_res[i, }\DecValTok{8}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{9}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{10}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{11}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \CommentTok{# Feased and defeased}
-        \NormalTok{all_res[i, }\DecValTok{12}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{all_res[i, }\DecValTok{13}\NormalTok{] <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(fea_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE} \NormalTok{&}\StringTok{ }
-\StringTok{                                  }\NormalTok{fea_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{);}
-    \NormalTok{\}}
-    \NormalTok{cnames <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{, }\StringTok{"A1_stable"}\NormalTok{, }
-                \StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{, }\StringTok{"A0_infeasible"}\NormalTok{, }
-                \StringTok{"A0_feasible"}\NormalTok{, }\StringTok{"A1_infeasible"}\NormalTok{, }\StringTok{"A1_feasible"}\NormalTok{, }
-                \StringTok{"A1_made_feasible"}\NormalTok{, }\StringTok{"A1_made_infeasible"}\NormalTok{);}
-    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\NormalTok{cnames;}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that feasibility results were ommited for the table above, but are
+matrices found is greater than 1 whenever \(S > 10\) (compare column 5
+to column 3), and this ratio increases with increasing \(S\) (column 1).
+Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
+given variation in \(\gamma\) than when \(\gamma = 1\). Note that
+feasibility results were omitted for the table above, but are
 \protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+Figure 4 of the main text reports the number of stable random complex
+systems found over 100000 using the genetic algorithm to maximise
+stability with a vector \(\mathbf{\gamma}\). Stability results for
+100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
+for \texttt{A0} indicate systems in which \(\gamma = 1\), while
+\texttt{A1} refers to systems in which the genetic algorithm searched
+for a set of \(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(\mathbf{M}\) that can be produced through a targeted search of
+\(\gamma\) values, and suggests that many otherwise unstable systems
+could potentially be stabilised by an informed manipulation of their
+component response times. Such a possibility might conceivably reduce
+the dimensionality of problems involving stability in social-ecological
+or economic systems.
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\includegraphics{unnamed-chunk-7-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix
+\(\mathbf{M}\).
 
@@
 ecological networks of interacting species. In these networks, the
-matrix \(M\) is interpreted as a community matrix and each row and
-column is interpreted as a single species. The effect that the density
-of any species \(i\) has on the population dynamics of species \(j\) is
-found in \(M_{ij}\), meaning that \(M\) holds the effects of pair-wise
-interactions between \(S\)
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Tang2014b}{4}}.
+matrix \(\mathbf{A}\) is interpreted as a community matrix and each row
+and column is interpreted as a single species. The per capita effect
+that the density of any species \(i\) has on the population dynamics of
+species \(j\) is found in \(A_{ij}\), meaning that \(\mathbf{A}\) holds
+the effects of pair-wise interactions between \(S\)
+species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
 While May's original
@@
 more restricted ecological communities including competitive networks
-(all off-diagonal elements of \(M\) are negative), mutualist networks
-(all off-diagonal elements of \(M\) are positive), and predator-prey
-networks (for any pair of \(i\) and \(j\), the effect of \(i\) on \(j\)
-is negative and \(j\) on \(i\) is positive, or vice
-versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}--\protect\hyperlink{ref-Allesina2011}{5}}.
+(all off-diagonal elements of \(\mathbf{A}\) are negative), mutualist
+networks (all off-diagonal elements of \(\mathbf{A}\) are positive), and
+predator-prey networks (for any pair of \(i\) and \(j\), the effect of
+\(i\) on \(j\) is negative and \(j\) on \(i\) is positive, or vice
+versa)\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{3}}.
 In general, competitor and mutualist networks tend to be unstable, while
-predator-prey networks tend to be highly stabilising.
+predator-prey networks tend to be highly
+stabilising\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
 
@@
 al.\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. To create
-these networks, I first generated a random matrix \(M\), then changed
-the elements of \(M\) accordingly. If \(M\) was a competitive network,
-then the sign of any positive off-diagonal elements was reversed to be
-negative. If \(M\) was a mutualist network, then the sign of any
-positive off-diagonal elements was reversed to be positive. And if \(M\)
-was a predator-prey network, then all \(i\) and \(j\) pairs of elements
-were checked; any pairs of the same sign were changed so that one was
-negative and the other was positive. The \texttt{species\_interaction}
-function used to do this is below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\} }\CommentTok{# Note: -1 values are added in the diagonal later}
-\end{Highlighting}
-\end{Shaded}
-
-This function was applied to all created matrices \(M\), then the number
-of stable \(M\) matrices was estimated \protect\hyperlink{IncrS}{exactly
-as it was} in the main text for random matrices for values of \(S\) from
-2 to 50 (100 in the case of the relatively more stable predator-prey
-interactions), except that only 100000 random \(M\) were generated
-instead of 1 million. This produced the data set below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{cdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/competition_C_1.csv"}\NormalTok{);}
-\NormalTok{mdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/mutualism_C_1.csv"}\NormalTok{);}
-\NormalTok{pdat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/ecology/pred-prey_C_1.csv"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
+these networks, I first generated a random matrix \(\mathbf{A}\), then
+changed the elements of \(\mathbf{A}\) accordingly. If \(\mathbf{A}\)
+was a competitive network, then the sign of any positive off-diagonal
+elements was reversed to be negative. If \(\mathbf{A}\) was a mutualist
+network, then the sign of any positive off-diagonal elements was
+reversed to be positive. And if \(\mathbf{A}\) was a predator-prey
+network, then all \(i\) and \(j\) pairs of elements were checked; any
+pairs of the same sign were changed so that one was negative and the
+other was positive.
+
+The number of stable \(\mathbf{M = \gamma A}\) systems was estimated
+\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
+matrices for values of \(S\) from 2 to 50 (100 in the case of the
+relatively more stable predator-prey interactions), except that only
+100000 random \(\mathbf{M}\) were generated instead of 1 million.
 
 The following tables for restricted ecological communities can therefore
-be compared with the random \(M\) \protect\hyperlink{IncrS}{results
-above} (but note that counts from systems with comparable probabilities
-of stability will be an order of magnitude lower in the tables below due
-to the smaller number of \(M\) matrices generated). As with the
+be compared with the random \(\mathbf{M}\)
+\protect\hyperlink{IncrS}{results above} (but note that counts from
+systems with comparable probabilities of stability will be an order of
+magnitude lower in the tables below due to the smaller number of
+\(\mathbf{M}\) matrices generated). As with the
 \protect\hyperlink{IncrS}{results above}, in the tables below,
@@
 
-\begin{longtable}[]{@{}llllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
+A1\_stabilised & A1\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 48 & 99952 & 48 & 99952 & 0\tabularnewline
-3 & 229 & 99771 & 231 & 99769 & 0\tabularnewline
-4 & 701 & 99299 & 704 & 99296 & 0\tabularnewline
-5 & 1579 & 98421 & 1587 & 98413 & 0\tabularnewline
-6 & 3218 & 96782 & 3253 & 96747 & 6\tabularnewline
-7 & 5519 & 94481 & 5619 & 94381 & 23\tabularnewline
-8 & 9062 & 90938 & 9237 & 90763 & 77\tabularnewline
-9 & 13436 & 86564 & 13729 & 86271 & 230\tabularnewline
-10 & 18911 & 81089 & 19303 & 80697 & 505\tabularnewline
-11 & 25594 & 74406 & 25961 & 74039 & 1011\tabularnewline
-12 & 33207 & 66793 & 33382 & 66618 & 1724\tabularnewline
-13 & 41160 & 58840 & 41089 & 58911 & 2655\tabularnewline
-14 & 50575 & 49425 & 49894 & 50106 & 3777\tabularnewline
-15 & 59250 & 40750 & 57892 & 42108 & 4824\tabularnewline
-16 & 67811 & 32189 & 65740 & 34260 & 5634\tabularnewline
-17 & 75483 & 24517 & 73056 & 26944 & 5943\tabularnewline
-18 & 82551 & 17449 & 79878 & 20122 & 5780\tabularnewline
-19 & 88030 & 11970 & 85204 & 14796 & 5417\tabularnewline
-20 & 92254 & 7746 & 89766 & 10234 & 4544\tabularnewline
-21 & 95233 & 4767 & 93002 & 6998 & 3695\tabularnewline
-22 & 97317 & 2683 & 95451 & 4549 & 2803\tabularnewline
-23 & 98508 & 1492 & 97122 & 2878 & 1991\tabularnewline
-24 & 99240 & 760 & 98407 & 1593 & 1216\tabularnewline
-25 & 99669 & 331 & 99082 & 918 & 739\tabularnewline
-26 & 99871 & 129 & 99490 & 510 & 452\tabularnewline
-27 & 99938 & 62 & 99732 & 268 & 240\tabularnewline
-28 & 99985 & 15 & 99888 & 112 & 108\tabularnewline
-29 & 99990 & 10 & 99951 & 49 & 46\tabularnewline
-30 & 100000 & 0 & 99981 & 19 & 19\tabularnewline
-31 & 100000 & 0 & 99993 & 7 & 7\tabularnewline
-32 & 100000 & 0 & 99996 & 4 & 4\tabularnewline
-33 & 100000 & 0 & 99998 & 2 & 2\tabularnewline
-34 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} \tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+2 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+3 & 229 & 99771 & 231 & 99769 & 0 & 2\tabularnewline
+4 & 701 & 99299 & 704 & 99296 & 0 & 3\tabularnewline
+5 & 1579 & 98421 & 1587 & 98413 & 0 & 8\tabularnewline
+6 & 3218 & 96782 & 3253 & 96747 & 6 & 41\tabularnewline
+7 & 5519 & 94481 & 5619 & 94381 & 23 & 123\tabularnewline
+8 & 9062 & 90938 & 9237 & 90763 & 77 & 252\tabularnewline
+9 & 13436 & 86564 & 13729 & 86271 & 230 & 523\tabularnewline
+10 & 18911 & 81089 & 19303 & 80697 & 505 & 897\tabularnewline
+11 & 25594 & 74406 & 25961 & 74039 & 1011 & 1378\tabularnewline
+12 & 33207 & 66793 & 33382 & 66618 & 1724 & 1899\tabularnewline
+13 & 41160 & 58840 & 41089 & 58911 & 2655 & 2584\tabularnewline
+14 & 50575 & 49425 & 49894 & 50106 & 3777 & 3096\tabularnewline
+15 & 59250 & 40750 & 57892 & 42108 & 4824 & 3466\tabularnewline
+16 & 67811 & 32189 & 65740 & 34260 & 5634 & 3563\tabularnewline
+17 & 75483 & 24517 & 73056 & 26944 & 5943 & 3516\tabularnewline
+18 & 82551 & 17449 & 79878 & 20122 & 5780 & 3107\tabularnewline
+19 & 88030 & 11970 & 85204 & 14796 & 5417 & 2591\tabularnewline
+20 & 92254 & 7746 & 89766 & 10234 & 4544 & 2056\tabularnewline
+21 & 95233 & 4767 & 93002 & 6998 & 3695 & 1464\tabularnewline
+22 & 97317 & 2683 & 95451 & 4549 & 2803 & 937\tabularnewline
+23 & 98508 & 1492 & 97122 & 2878 & 1991 & 605\tabularnewline
+24 & 99240 & 760 & 98407 & 1593 & 1216 & 383\tabularnewline
+25 & 99669 & 331 & 99082 & 918 & 739 & 152\tabularnewline
+26 & 99871 & 129 & 99490 & 510 & 452 & 71\tabularnewline
+27 & 99938 & 62 & 99732 & 268 & 240 & 34\tabularnewline
+28 & 99985 & 15 & 99888 & 112 & 108 & 11\tabularnewline
+29 & 99990 & 10 & 99951 & 49 & 46 & 7\tabularnewline
+30 & 100000 & 0 & 99981 & 19 & 19 & 0\tabularnewline
+31 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
+32 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
+33 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+34 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
+A1\_stabilised & A1\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 56 & 99944 & 56 & 99944 & 0\tabularnewline
-3 & 3301 & 96699 & 3301 & 96699 & 0\tabularnewline
-4 & 34446 & 65554 & 34446 & 65554 & 0\tabularnewline
-5 & 86520 & 13480 & 86520 & 13480 & 0\tabularnewline
-6 & 99683 & 317 & 99683 & 317 & 0\tabularnewline
-7 & 99998 & 2 & 99998 & 2 & 0\tabularnewline
-8 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-9 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-10 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-11 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-12 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} \tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+2 & 56 & 99944 & 56 & 99944 & 0 & 0\tabularnewline
+3 & 3301 & 96699 & 3301 & 96699 & 0 & 0\tabularnewline
+4 & 34446 & 65554 & 34446 & 65554 & 0 & 0\tabularnewline
+5 & 86520 & 13480 & 86520 & 13480 & 0 & 0\tabularnewline
+6 & 99683 & 317 & 99683 & 317 & 0 & 0\tabularnewline
+7 & 99998 & 2 & 99998 & 2 & 0 & 0\tabularnewline
+8 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+9 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+11 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+12 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}rrrrrr@{}}
+\begin{longtable}[]{@{}rrrrrrr@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised\tabularnewline
+A1\_stabilised & A1\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-3 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-4 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-5 & 1 & 99999 & 1 & 99999 & 0\tabularnewline
-6 & 4 & 99996 & 4 & 99996 & 0\tabularnewline
-7 & 2 & 99998 & 2 & 99998 & 0\tabularnewline
-8 & 5 & 99995 & 5 & 99995 & 0\tabularnewline
-9 & 20 & 99980 & 21 & 99979 & 0\tabularnewline
-10 & 20 & 99980 & 22 & 99978 & 0\tabularnewline
-11 & 38 & 99962 & 39 & 99961 & 0\tabularnewline
-12 & 64 & 99936 & 66 & 99934 & 0\tabularnewline
-13 & 87 & 99913 & 91 & 99909 & 0\tabularnewline
-14 & 157 & 99843 & 159 & 99841 & 0\tabularnewline
-15 & 215 & 99785 & 227 & 99773 & 0\tabularnewline
-16 & 293 & 99707 & 310 & 99690 & 0\tabularnewline
-17 & 383 & 99617 & 408 & 99592 & 0\tabularnewline
-18 & 443 & 99557 & 473 & 99527 & 3\tabularnewline
-19 & 642 & 99358 & 675 & 99325 & 4\tabularnewline
-20 & 836 & 99164 & 887 & 99113 & 7\tabularnewline
-21 & 1006 & 98994 & 1058 & 98942 & 10\tabularnewline
-22 & 1153 & 98847 & 1228 & 98772 & 20\tabularnewline
-23 & 1501 & 98499 & 1593 & 98407 & 30\tabularnewline
-24 & 1841 & 98159 & 1996 & 98004 & 40\tabularnewline
-25 & 2146 & 97854 & 2316 & 97684 & 58\tabularnewline
-26 & 2643 & 97357 & 2809 & 97191 & 119\tabularnewline
-27 & 3034 & 96966 & 3258 & 96742 & 158\tabularnewline
-28 & 3690 & 96310 & 3928 & 96072 & 201\tabularnewline
-29 & 4257 & 95743 & 4532 & 95468 & 290\tabularnewline
-30 & 4964 & 95036 & 5221 & 94779 & 424\tabularnewline
-31 & 5627 & 94373 & 5978 & 94022 & 452\tabularnewline
-32 & 6543 & 93457 & 6891 & 93109 & 666\tabularnewline
-33 & 7425 & 92575 & 7777 & 92223 & 818\tabularnewline
-34 & 8540 & 91460 & 8841 & 91159 & 1071\tabularnewline
-35 & 9526 & 90474 & 9842 & 90158 & 1337\tabularnewline
-36 & 10617 & 89383 & 10891 & 89109 & 1624\tabularnewline
-37 & 12344 & 87656 & 12508 & 87492 & 2021\tabularnewline
-38 & 13675 & 86325 & 13877 & 86123 & 2442\tabularnewline
-39 & 15264 & 84736 & 15349 & 84651 & 2870\tabularnewline
-40 & 17026 & 82974 & 17053 & 82947 & 3363\tabularnewline
-41 & 18768 & 81232 & 18614 & 81386 & 3905\tabularnewline
-42 & 20791 & 79209 & 20470 & 79530 & 4579\tabularnewline
-43 & 23150 & 76850 & 22754 & 77246 & 5217\tabularnewline
-44 & 25449 & 74551 & 24184 & 75816 & 6285\tabularnewline
-45 & 27702 & 72298 & 26464 & 73536 & 6754\tabularnewline
-46 & 30525 & 69475 & 28966 & 71034 & 7646\tabularnewline
-47 & 32832 & 67168 & 31125 & 68875 & 8487\tabularnewline
-48 & 36152 & 63848 & 33865 & 66135 & 9479\tabularnewline
-49 & 38714 & 61286 & 36242 & 63758 & 10125\tabularnewline
-50 & 41628 & 58372 & 38508 & 61492 & 11036\tabularnewline
-51 & 44483 & 55517 & 41023 & 58977 & 11704\tabularnewline
-52 & 48134 & 51866 & 44287 & 55713 & 12573\tabularnewline
-53 & 51138 & 48862 & 46721 & 53279 & 13223\tabularnewline
-54 & 54261 & 45739 & 49559 & 50441 & 13757\tabularnewline
-55 & 57647 & 42353 & 52403 & 47597 & 14324\tabularnewline
-56 & 60630 & 39370 & 55293 & 44707 & 14669\tabularnewline
-57 & 63647 & 36353 & 57787 & 42213 & 15103\tabularnewline
-58 & 66961 & 33039 & 60439 & 39561 & 15450\tabularnewline
-59 & 69968 & 30032 & 63708 & 36292 & 15246\tabularnewline
-60 & 72838 & 27162 & 66270 & 33730 & 15177\tabularnewline
-61 & 75609 & 24391 & 68873 & 31127 & 15006\tabularnewline
-62 & 77999 & 22001 & 71318 & 28682 & 14538\tabularnewline
-63 & 80616 & 19384 & 73517 & 26483 & 14510\tabularnewline
-64 & 83089 & 16911 & 76209 & 23791 & 13784\tabularnewline
-65 & 85150 & 14850 & 78086 & 21914 & 13412\tabularnewline
-66 & 86908 & 13092 & 80437 & 19563 & 12477\tabularnewline
-67 & 88671 & 11329 & 82379 & 17621 & 11718\tabularnewline
-68 & 90537 & 9463 & 84483 & 15517 & 10878\tabularnewline
-69 & 91969 & 8031 & 86233 & 13767 & 10033\tabularnewline
-70 & 93181 & 6819 & 87914 & 12086 & 9070\tabularnewline
-71 & 94330 & 5670 & 89200 & 10800 & 8401\tabularnewline
-72 & 95324 & 4676 & 90833 & 9167 & 7359\tabularnewline
-73 & 96143 & 3857 & 91805 & 8195 & 6726\tabularnewline
-74 & 96959 & 3041 & 93065 & 6935 & 5900\tabularnewline
-75 & 97543 & 2457 & 93987 & 6013 & 5222\tabularnewline
-76 & 97969 & 2031 & 94900 & 5100 & 4481\tabularnewline
-77 & 98497 & 1503 & 95756 & 4244 & 3809\tabularnewline
-78 & 98744 & 1256 & 96442 & 3558 & 3269\tabularnewline
-79 & 99045 & 955 & 96942 & 3058 & 2837\tabularnewline
-80 & 99276 & 724 & 97528 & 2472 & 2329\tabularnewline
-81 & 99481 & 519 & 97996 & 2004 & 1894\tabularnewline
-82 & 99556 & 444 & 98321 & 1679 & 1597\tabularnewline
-83 & 99691 & 309 & 98722 & 1278 & 1227\tabularnewline
-84 & 99752 & 248 & 98943 & 1057 & 1015\tabularnewline
-85 & 99833 & 167 & 99144 & 856 & 837\tabularnewline
-86 & 99895 & 105 & 99346 & 654 & 642\tabularnewline
-87 & 99925 & 75 & 99461 & 539 & 530\tabularnewline
-88 & 99945 & 55 & 99566 & 434 & 428\tabularnewline
-89 & 99976 & 24 & 99675 & 325 & 324\tabularnewline
-90 & 99977 & 23 & 99756 & 244 & 243\tabularnewline
-91 & 99982 & 18 & 99839 & 161 & 155\tabularnewline
-92 & 99988 & 12 & 99865 & 135 & 135\tabularnewline
-93 & 99994 & 6 & 99885 & 115 & 115\tabularnewline
-94 & 99993 & 7 & 99911 & 89 & 88\tabularnewline
-95 & 99998 & 2 & 99953 & 47 & 47\tabularnewline
-96 & 99999 & 1 & 99965 & 35 & 35\tabularnewline
-97 & 99999 & 1 & 99979 & 21 & 21\tabularnewline
-98 & 100000 & 0 & 99973 & 27 & 27\tabularnewline
-99 & 100000 & 0 & 99984 & 16 & 16\tabularnewline
-100 & 100000 & 0 & 99989 & 11 & 11\tabularnewline
+2 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+3 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+4 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+5 & 1 & 99999 & 1 & 99999 & 0 & 0\tabularnewline
+6 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+7 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+8 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+9 & 20 & 99980 & 21 & 99979 & 0 & 1\tabularnewline
+10 & 20 & 99980 & 22 & 99978 & 0 & 2\tabularnewline
+11 & 38 & 99962 & 39 & 99961 & 0 & 1\tabularnewline
+12 & 64 & 99936 & 66 & 99934 & 0 & 2\tabularnewline
+13 & 87 & 99913 & 91 & 99909 & 0 & 4\tabularnewline
+14 & 157 & 99843 & 159 & 99841 & 0 & 2\tabularnewline
+15 & 215 & 99785 & 227 & 99773 & 0 & 12\tabularnewline
+16 & 293 & 99707 & 310 & 99690 & 0 & 17\tabularnewline
+17 & 383 & 99617 & 408 & 99592 & 0 & 25\tabularnewline
+18 & 443 & 99557 & 473 & 99527 & 3 & 33\tabularnewline
+19 & 642 & 99358 & 675 & 99325 & 4 & 37\tabularnewline
+20 & 836 & 99164 & 887 & 99113 & 7 & 58\tabularnewline
+21 & 1006 & 98994 & 1058 & 98942 & 10 & 62\tabularnewline
+22 & 1153 & 98847 & 1228 & 98772 & 20 & 95\tabularnewline
+23 & 1501 & 98499 & 1593 & 98407 & 30 & 122\tabularnewline
+24 & 1841 & 98159 & 1996 & 98004 & 40 & 195\tabularnewline
+25 & 2146 & 97854 & 2316 & 97684 & 58 & 228\tabularnewline
+26 & 2643 & 97357 & 2809 & 97191 & 119 & 285\tabularnewline
+27 & 3034 & 96966 & 3258 & 96742 & 158 & 382\tabularnewline
+28 & 3690 & 96310 & 3928 & 96072 & 201 & 439\tabularnewline
+29 & 4257 & 95743 & 4532 & 95468 & 290 & 565\tabularnewline
+30 & 4964 & 95036 & 5221 & 94779 & 424 & 681\tabularnewline
+31 & 5627 & 94373 & 5978 & 94022 & 452 & 803\tabularnewline
+32 & 6543 & 93457 & 6891 & 93109 & 666 & 1014\tabularnewline
+33 & 7425 & 92575 & 7777 & 92223 & 818 & 1170\tabularnewline
+34 & 8540 & 91460 & 8841 & 91159 & 1071 & 1372\tabularnewline
+35 & 9526 & 90474 & 9842 & 90158 & 1337 & 1653\tabularnewline
+36 & 10617 & 89383 & 10891 & 89109 & 1624 & 1898\tabularnewline
+37 & 12344 & 87656 & 12508 & 87492 & 2021 & 2185\tabularnewline
+38 & 13675 & 86325 & 13877 & 86123 & 2442 & 2644\tabularnewline
+39 & 15264 & 84736 & 15349 & 84651 & 2870 & 2955\tabularnewline
+40 & 17026 & 82974 & 17053 & 82947 & 3363 & 3390\tabularnewline
+41 & 18768 & 81232 & 18614 & 81386 & 3905 & 3751\tabularnewline
+42 & 20791 & 79209 & 20470 & 79530 & 4579 & 4258\tabularnewline
+43 & 23150 & 76850 & 22754 & 77246 & 5217 & 4821\tabularnewline
+44 & 25449 & 74551 & 24184 & 75816 & 6285 & 5020\tabularnewline
+45 & 27702 & 72298 & 26464 & 73536 & 6754 & 5516\tabularnewline
+46 & 30525 & 69475 & 28966 & 71034 & 7646 & 6087\tabularnewline
+47 & 32832 & 67168 & 31125 & 68875 & 8487 & 6780\tabularnewline
+48 & 36152 & 63848 & 33865 & 66135 & 9479 & 7192\tabularnewline
+49 & 38714 & 61286 & 36242 & 63758 & 10125 & 7653\tabularnewline
+50 & 41628 & 58372 & 38508 & 61492 & 11036 & 7916\tabularnewline
+51 & 44483 & 55517 & 41023 & 58977 & 11704 & 8244\tabularnewline
+52 & 48134 & 51866 & 44287 & 55713 & 12573 & 8726\tabularnewline
+53 & 51138 & 48862 & 46721 & 53279 & 13223 & 8806\tabularnewline
+54 & 54261 & 45739 & 49559 & 50441 & 13757 & 9055\tabularnewline
+55 & 57647 & 42353 & 52403 & 47597 & 14324 & 9080\tabularnewline
+56 & 60630 & 39370 & 55293 & 44707 & 14669 & 9332\tabularnewline
+57 & 63647 & 36353 & 57787 & 42213 & 15103 & 9243\tabularnewline
+58 & 66961 & 33039 & 60439 & 39561 & 15450 & 8928\tabularnewline
+59 & 69968 & 30032 & 63708 & 36292 & 15246 & 8986\tabularnewline
+60 & 72838 & 27162 & 66270 & 33730 & 15177 & 8609\tabularnewline
+61 & 75609 & 24391 & 68873 & 31127 & 15006 & 8270\tabularnewline
+62 & 77999 & 22001 & 71318 & 28682 & 14538 & 7857\tabularnewline
+63 & 80616 & 19384 & 73517 & 26483 & 14510 & 7411\tabularnewline
+64 & 83089 & 16911 & 76209 & 23791 & 13784 & 6904\tabularnewline
+65 & 85150 & 14850 & 78086 & 21914 & 13412 & 6348\tabularnewline
+66 & 86908 & 13092 & 80437 & 19563 & 12477 & 6006\tabularnewline
+67 & 88671 & 11329 & 82379 & 17621 & 11718 & 5426\tabularnewline
+68 & 90537 & 9463 & 84483 & 15517 & 10878 & 4824\tabularnewline
+69 & 91969 & 8031 & 86233 & 13767 & 10033 & 4297\tabularnewline
+70 & 93181 & 6819 & 87914 & 12086 & 9070 & 3803\tabularnewline
+71 & 94330 & 5670 & 89200 & 10800 & 8401 & 3271\tabularnewline
+72 & 95324 & 4676 & 90833 & 9167 & 7359 & 2868\tabularnewline
+73 & 96143 & 3857 & 91805 & 8195 & 6726 & 2388\tabularnewline
+74 & 96959 & 3041 & 93065 & 6935 & 5900 & 2006\tabularnewline
+75 & 97543 & 2457 & 93987 & 6013 & 5222 & 1666\tabularnewline
+76 & 97969 & 2031 & 94900 & 5100 & 4481 & 1412\tabularnewline
+77 & 98497 & 1503 & 95756 & 4244 & 3809 & 1068\tabularnewline
+78 & 98744 & 1256 & 96442 & 3558 & 3269 & 967\tabularnewline
+79 & 99045 & 955 & 96942 & 3058 & 2837 & 734\tabularnewline
+80 & 99276 & 724 & 97528 & 2472 & 2329 & 581\tabularnewline
+81 & 99481 & 519 & 97996 & 2004 & 1894 & 409\tabularnewline
+82 & 99556 & 444 & 98321 & 1679 & 1597 & 362\tabularnewline
+83 & 99691 & 309 & 98722 & 1278 & 1227 & 258\tabularnewline
+84 & 99752 & 248 & 98943 & 1057 & 1015 & 206\tabularnewline
+85 & 99833 & 167 & 99144 & 856 & 837 & 148\tabularnewline
+86 & 99895 & 105 & 99346 & 654 & 642 & 93\tabularnewline
+87 & 99925 & 75 & 99461 & 539 & 530 & 66\tabularnewline
+88 & 99945 & 55 & 99566 & 434 & 428 & 49\tabularnewline
+89 & 99976 & 24 & 99675 & 325 & 324 & 23\tabularnewline
+90 & 99977 & 23 & 99756 & 244 & 243 & 22\tabularnewline
+91 & 99982 & 18 & 99839 & 161 & 155 & 12\tabularnewline
+92 & 99988 & 12 & 99865 & 135 & 135 & 12\tabularnewline
+93 & 99994 & 6 & 99885 & 115 & 115 & 6\tabularnewline
+94 & 99993 & 7 & 99911 & 89 & 88 & 6\tabularnewline
+95 & 99998 & 2 & 99953 & 47 & 47 & 2\tabularnewline
+96 & 99999 & 1 & 99965 & 35 & 35 & 1\tabularnewline
+97 & 99999 & 1 & 99979 & 21 & 21 & 1\tabularnewline
+98 & 100000 & 0 & 99973 & 27 & 27 & 0\tabularnewline
+99 & 100000 & 0 & 99984 & 16 & 16 & 0\tabularnewline
+100 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
 \bottomrule
@@
 
-\hypertarget{connectance}{\section{Different connectance (C)
+\hypertarget{connectance}{\section{Sensitivity of connectance (C)
 values}\label{connectance}}
@@
 In the main text, for simplicity, I assumed connectance values of
-\(C = 1\), meaning that all off-diagonal elements of a matrix \(M\) were
-potentially nonzero and sampled from a normal distribution
-\(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here I present
-four tables showing the number of stable communities given
+\(C = 1\), meaning that all off-diagonal elements of a matrix
+\(\mathbf{M}\) were potentially nonzero and sampled from a normal
+distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
+I present four tables showing the number of stable communities given
 \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
-component response time (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
+component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
 higher number of stable communities than when \(\gamma\) did not vary
 (\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
-million \(M\) were simulated. As with the results on
+million \(\mathbf{M}\) were simulated. As with the results on
 \protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
@@
 
-All data reported below for various values of \(C\) are accessible using
-the below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{C3dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt3.csv"}\NormalTok{);}
-\NormalTok{C5dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt5.csv"}\NormalTok{);}
-\NormalTok{C7dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt7.csv"}\NormalTok{);}
-\NormalTok{C9dat <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/C_other/rand_c-0pt9.csv"}\NormalTok{);}
-\end{Highlighting}
-\end{Shaded}
-
-These objects \texttt{C3dat}, \texttt{C5dat}, \texttt{C7dat}, and
-\texttt{C9dat} include the results for \(C = 0.3\), \(C = 0.5\),
-\(C = 0.7\), and \(C = 0.9\), respectively.
-
-\textbf{Connectance \(C = 0.3\)}
-
-\begin{longtable}[]{@{}lllllll@{}}
+\textbf{Connectance \(\mathbf{C = 0.3}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
 \toprule
 N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 5 & 99995 & 5 & 99995 & 0 & 0 & 75110\tabularnewline
+3 & 6 & 99994 & 6 & 99994 & 0 & 0 & 87526\tabularnewline
+4 & 24 & 99976 & 24 & 99976 & 0 & 0 & 93713\tabularnewline
+5 & 59 & 99941 & 59 & 99941 & 0 & 0 & 96929\tabularnewline
+6 & 98 & 99902 & 98 & 99902 & 0 & 0 & 98492\tabularnewline
+7 & 160 & 99840 & 161 & 99839 & 0 & 1 & 99243\tabularnewline
+8 & 290 & 99710 & 293 & 99707 & 0 & 3 & 99582\tabularnewline
+9 & 430 & 99570 & 434 & 99566 & 0 & 4 & 99821\tabularnewline
+10 & 648 & 99352 & 653 & 99347 & 1 & 6 & 99895\tabularnewline
+11 & 946 & 99054 & 957 & 99043 & 0 & 11 & 99945\tabularnewline
+12 & 1392 & 98608 & 1415 & 98585 & 4 & 27 & 99978\tabularnewline
+13 & 2032 & 97968 & 2065 & 97935 & 5 & 38 & 99987\tabularnewline
+14 & 2627 & 97373 & 2688 & 97312 & 10 & 71 & 99988\tabularnewline
+15 & 3588 & 96412 & 3647 & 96353 & 35 & 94 & 99996\tabularnewline
+16 & 5019 & 94981 & 5124 & 94876 & 51 & 156 & 99998\tabularnewline
+17 & 6512 & 93488 & 6673 & 93327 & 79 & 240 & 99999\tabularnewline
+18 & 8444 & 91556 & 8600 & 91400 & 165 & 321 & 100000\tabularnewline
+19 & 10416 & 89584 & 10667 & 89333 & 244 & 495 & 100000\tabularnewline
+20 & 13254 & 86746 & 13477 & 86523 & 425 & 648 & 100000\tabularnewline
+21 & 16248 & 83752 & 16481 & 83519 & 642 & 875 & 100000\tabularnewline
+22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151 & 100000\tabularnewline
+23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490 & 100000\tabularnewline
+24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818 & 100000\tabularnewline
+25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137 & 100000\tabularnewline
+26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506 & 100000\tabularnewline
+27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811 & 100000\tabularnewline
+28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072 & 100000\tabularnewline
+29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321 & 100000\tabularnewline
+30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480 & 100000\tabularnewline
+31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568 & 100000\tabularnewline
+32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531 & 100000\tabularnewline
+33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360 & 100000\tabularnewline
+34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014 & 100000\tabularnewline
+35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773 & 100000\tabularnewline
+36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393 & 100000\tabularnewline
+37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913 & 100000\tabularnewline
+38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566 & 100000\tabularnewline
+39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187 & 100000\tabularnewline
+40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936 & 100000\tabularnewline
+41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683 & 100000\tabularnewline
+42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449 & 100000\tabularnewline
+43 & 99275 & 725 & 98312 & 1688 & 1251 & 288 & 100000\tabularnewline
+44 & 99583 & 417 & 98872 & 1128 & 903 & 192 & 100000\tabularnewline
+45 & 99776 & 224 & 99339 & 661 & 576 & 139 & 100000\tabularnewline
+46 & 99865 & 135 & 99518 & 482 & 413 & 66 & 100000\tabularnewline
+47 & 99938 & 62 & 99744 & 256 & 226 & 32 & 100000\tabularnewline
+48 & 99956 & 44 & 99824 & 176 & 151 & 19 & 100000\tabularnewline
+49 & 99980 & 20 & 99914 & 86 & 85 & 19 & 100000\tabularnewline
+50 & 99993 & 7 & 99950 & 50 & 46 & 3 & 100000\tabularnewline
+51 & 99998 & 2 & 99971 & 29 & 28 & 1 & 100000\tabularnewline
+52 & 99998 & 2 & 99986 & 14 & 14 & 2 & 100000\tabularnewline
+53 & 99999 & 1 & 99992 & 8 & 7 & 0 & 100000\tabularnewline
+54 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+55 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+56 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+57 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+58 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+100 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.5}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 74863\tabularnewline
+3 & 32 & 99968 & 32 & 99968 & 0 & 0 & 87434\tabularnewline
+4 & 122 & 99878 & 122 & 99878 & 0 & 0 & 93761\tabularnewline
+5 & 320 & 99680 & 321 & 99679 & 0 & 1 & 96830\tabularnewline
+6 & 667 & 99333 & 673 & 99327 & 0 & 6 & 98481\tabularnewline
+7 & 1233 & 98767 & 1252 & 98748 & 0 & 19 & 99187\tabularnewline
+8 & 2123 & 97877 & 2156 & 97844 & 3 & 36 & 99654\tabularnewline
+9 & 3415 & 96585 & 3471 & 96529 & 16 & 72 & 99816\tabularnewline
+10 & 5349 & 94651 & 5450 & 94550 & 30 & 131 & 99900\tabularnewline
+11 & 7990 & 92010 & 8185 & 91815 & 81 & 276 & 99958\tabularnewline
+12 & 11073 & 88927 & 11301 & 88699 & 219 & 447 & 99973\tabularnewline
+13 & 14971 & 85029 & 15204 & 84796 & 445 & 678 & 99986\tabularnewline
+14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002 & 99991\tabularnewline
+15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404 & 99996\tabularnewline
+16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980 & 99999\tabularnewline
+17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476 & 100000\tabularnewline
+18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937 & 99999\tabularnewline
+19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262 & 99999\tabularnewline
+20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527 & 100000\tabularnewline
+21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673 & 100000\tabularnewline
+22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491 & 100000\tabularnewline
+23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244 & 100000\tabularnewline
+24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876 & 100000\tabularnewline
+25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382 & 100000\tabularnewline
+26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901 & 100000\tabularnewline
+27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358 & 100000\tabularnewline
+28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974 & 100000\tabularnewline
+29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657 & 100000\tabularnewline
+30 & 99103 & 897 & 98033 & 1967 & 1466 & 396 & 100000\tabularnewline
+31 & 99502 & 498 & 98665 & 1335 & 1068 & 231 & 100000\tabularnewline
+32 & 99745 & 255 & 99185 & 815 & 696 & 136 & 100000\tabularnewline
+33 & 99881 & 119 & 99572 & 428 & 375 & 66 & 100000\tabularnewline
+34 & 99955 & 45 & 99788 & 212 & 191 & 24 & 100000\tabularnewline
+35 & 99979 & 21 & 99900 & 100 & 95 & 16 & 100000\tabularnewline
+36 & 99995 & 5 & 99950 & 50 & 50 & 5 & 100000\tabularnewline
+37 & 99997 & 3 & 99970 & 30 & 28 & 1 & 100000\tabularnewline
+38 & 99998 & 2 & 99986 & 14 & 13 & 1 & 100000\tabularnewline
+39 & 99999 & 1 & 99991 & 9 & 9 & 1 & 100000\tabularnewline
+40 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+41 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+42 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+43 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.7}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 75160\tabularnewline
+3 & 106 & 99894 & 106 & 99894 & 0 & 0 & 87447\tabularnewline
+4 & 395 & 99605 & 397 & 99603 & 0 & 2 & 93849\tabularnewline
+5 & 1117 & 98883 & 1123 & 98877 & 0 & 6 & 96827\tabularnewline
+6 & 2346 & 97654 & 2367 & 97633 & 6 & 27 & 98402\tabularnewline
+7 & 4314 & 95686 & 4388 & 95612 & 16 & 90 & 99218\tabularnewline
+8 & 7327 & 92673 & 7456 & 92544 & 61 & 190 & 99611\tabularnewline
+9 & 11514 & 88486 & 11792 & 88208 & 150 & 428 & 99808\tabularnewline
+10 & 16247 & 83753 & 16584 & 83416 & 415 & 752 & 99904\tabularnewline
+11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162 & 99952\tabularnewline
+12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818 & 99977\tabularnewline
+13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355 & 99984\tabularnewline
+14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927 & 99995\tabularnewline
+15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421 & 99994\tabularnewline
+16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527 & 99999\tabularnewline
+17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495 & 100000\tabularnewline
+18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351 & 100000\tabularnewline
+19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979 & 99999\tabularnewline
+20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339 & 100000\tabularnewline
+21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869 & 100000\tabularnewline
+22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267 & 100000\tabularnewline
+23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791 & 100000\tabularnewline
+24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504 & 100000\tabularnewline
+25 & 99433 & 567 & 98630 & 1370 & 1079 & 276 & 100000\tabularnewline
+26 & 99760 & 240 & 99259 & 741 & 655 & 154 & 100000\tabularnewline
+27 & 99895 & 105 & 99576 & 424 & 377 & 58 & 100000\tabularnewline
+28 & 99950 & 50 & 99790 & 210 & 194 & 34 & 100000\tabularnewline
+29 & 99981 & 19 & 99915 & 85 & 80 & 14 & 100000\tabularnewline
+30 & 99994 & 6 & 99952 & 48 & 47 & 5 & 100000\tabularnewline
+31 & 99998 & 2 & 99972 & 28 & 28 & 2 & 100000\tabularnewline
+32 & 99999 & 1 & 99992 & 8 & 8 & 1 & 100000\tabularnewline
+33 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
+34 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
+35 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\textbf{Connectance \(\mathbf{C = 0.9}\)}
+
+\begin{longtable}[]{@{}llllllll@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+\midrule
+\endhead
+2 & 14 & 99986 & 14 & 99986 & 0 & 0 & 75187\tabularnewline
+3 & 240 & 99760 & 240 & 99760 & 0 & 0 & 87443\tabularnewline
+4 & 1008 & 98992 & 1016 & 98984 & 0 & 8 & 93795\tabularnewline
+5 & 2708 & 97292 & 2729 & 97271 & 2 & 23 & 96814\tabularnewline
+6 & 5669 & 94331 & 5755 & 94245 & 13 & 99 & 98439\tabularnewline
+7 & 9848 & 90152 & 10057 & 89943 & 91 & 300 & 99208\tabularnewline
+8 & 15903 & 84097 & 16201 & 83799 & 336 & 634 & 99603\tabularnewline
+9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168 & 99803\tabularnewline
+10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852 & 99909\tabularnewline
+11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507 & 99951\tabularnewline
+12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127 & 99977\tabularnewline
+13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649 & 99986\tabularnewline
+14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765 & 99995\tabularnewline
+15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550 & 100000\tabularnewline
+16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092 & 100000\tabularnewline
+17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424 & 100000\tabularnewline
+18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746 & 100000\tabularnewline
+19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167 & 99999\tabularnewline
+20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729 & 100000\tabularnewline
+21 & 99111 & 889 & 98133 & 1867 & 1389 & 411 & 100000\tabularnewline
+22 & 99588 & 412 & 98905 & 1095 & 903 & 220 & 100000\tabularnewline
+23 & 99837 & 163 & 99480 & 520 & 452 & 95 & 100000\tabularnewline
+24 & 99932 & 68 & 99744 & 256 & 228 & 40 & 100000\tabularnewline
+25 & 99976 & 24 & 99863 & 137 & 133 & 20 & 100000\tabularnewline
+26 & 99995 & 5 & 99950 & 50 & 49 & 4 & 100000\tabularnewline
+27 & 99996 & 4 & 99986 & 14 & 13 & 3 & 100000\tabularnewline
+28 & 100000 & 0 & 99993 & 7 & 7 & 0 & 100000\tabularnewline
+29 & 100000 & 0 & 99996 & 4 & 4 & 0 & 100000\tabularnewline
+30 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
+\ldots{} & \ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\bottomrule
+\end{longtable}
+
+\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
+strength (\(\sigma\))
+values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
+
+Results below show stability results given varying interaction strengths
+(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
+larger and increase by 10 with increasing rows). In the tables below (as
+\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
+matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
+
+\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
 A1\_stabilised & A1\_destabilised\tabularnewline
@@
 \endhead
-2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
-3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
-4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
-5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
-6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
-7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
-8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
-9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
-10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
-11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
-12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
-13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
-14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
-15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
-16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
-17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
-18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
-19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
-20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
-21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
-22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
-23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
-24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
-25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
-26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
-27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
-28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
-29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
-30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
-31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
-32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
-33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
-34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
-35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
-36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
-37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
-38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
-39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
-40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
-41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
-42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
-43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
-44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
-45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
-46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
-47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
-48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
-49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
-50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
-51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
-52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
-53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
-54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
-55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-\ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{}\tabularnewline
-100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+20 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+30 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+40 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+50 & 0 & 100000 & 0 & 100000 & 0 & 0\tabularnewline
+60 & 2 & 99998 & 2 & 99998 & 0 & 0\tabularnewline
+70 & 4 & 99996 & 4 & 99996 & 0 & 0\tabularnewline
+80 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
+90 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+100 & 11 & 99989 & 11 & 99989 & 0 & 0\tabularnewline
+110 & 12 & 99988 & 13 & 99987 & 0 & 1\tabularnewline
+120 & 23 & 99977 & 23 & 99977 & 0 & 0\tabularnewline
+130 & 40 & 99960 & 40 & 99960 & 0 & 0\tabularnewline
+140 & 62 & 99938 & 65 & 99935 & 0 & 3\tabularnewline
+150 & 162 & 99838 & 165 & 99835 & 0 & 3\tabularnewline
+160 & 325 & 99675 & 329 & 99671 & 2 & 6\tabularnewline
+170 & 829 & 99171 & 851 & 99149 & 6 & 28\tabularnewline
+180 & 1817 & 98183 & 1860 & 98140 & 31 & 74\tabularnewline
+190 & 3927 & 96073 & 3989 & 96011 & 143 & 205\tabularnewline
+200 & 8084 & 91916 & 8048 & 91952 & 557 & 521\tabularnewline
+210 & 15558 & 84442 & 15147 & 84853 & 1534 & 1123\tabularnewline
+220 & 26848 & 73152 & 25342 & 74658 & 3625 & 2119\tabularnewline
+230 & 43386 & 56614 & 39535 & 60465 & 6992 & 3141\tabularnewline
+240 & 62734 & 37266 & 56684 & 43316 & 9815 & 3765\tabularnewline
+250 & 80128 & 19872 & 73080 & 26920 & 10128 & 3080\tabularnewline
+260 & 92206 & 7794 & 86619 & 13381 & 7490 & 1903\tabularnewline
+270 & 97946 & 2054 & 94824 & 5176 & 3797 & 675\tabularnewline
+280 & 99659 & 341 & 98534 & 1466 & 1265 & 140\tabularnewline
+290 & 99962 & 38 & 99696 & 304 & 281 & 15\tabularnewline
+300 & 99994 & 6 & 99964 & 36 & 34 & 4\tabularnewline
 \bottomrule
@@
 
-\textbf{Connectance \(C = 0.5\)}
+\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}
 
@@
 \endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
-3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
-4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
-5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
-6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
-7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
-8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
-9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
-10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
-11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
-12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
-13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
-14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
-15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
-16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
-17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
-18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
-19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
-20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
-21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
-22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
-23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
-24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
-25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
-26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
-27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
-28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
-29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
-30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
-31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
-32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
-33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
-34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
-35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
-36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
-37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
-38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
-39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
-40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
-41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 3 & 99997 & 3 & 99997 & 0 & 0\tabularnewline
+20 & 15 & 99985 & 15 & 99985 & 0 & 0\tabularnewline
+30 & 48 & 99952 & 48 & 99952 & 0 & 0\tabularnewline
+40 & 85 & 99915 & 85 & 99915 & 0 & 0\tabularnewline
+50 & 163 & 99837 & 163 & 99837 & 0 & 0\tabularnewline
+60 & 280 & 99720 & 282 & 99718 & 0 & 2\tabularnewline
+70 & 561 & 99439 & 566 & 99434 & 3 & 8\tabularnewline
+80 & 1009 & 98991 & 1029 & 98971 & 6 & 26\tabularnewline
+90 & 2126 & 97874 & 2175 & 97825 & 31 & 80\tabularnewline
+100 & 4580 & 95420 & 4653 & 95347 & 142 & 215\tabularnewline
+110 & 9540 & 90460 & 9632 & 90368 & 465 & 557\tabularnewline
+120 & 19090 & 80910 & 18668 & 81332 & 1676 & 1254\tabularnewline
+130 & 35047 & 64953 & 33220 & 66780 & 4172 & 2345\tabularnewline
+140 & 56411 & 43589 & 52439 & 47561 & 7297 & 3325\tabularnewline
+150 & 78003 & 21997 & 72574 & 27426 & 8477 & 3048\tabularnewline
+160 & 92678 & 7322 & 88438 & 11562 & 5901 & 1661\tabularnewline
+170 & 98614 & 1386 & 96670 & 3330 & 2397 & 453\tabularnewline
+180 & 99839 & 161 & 99418 & 582 & 499 & 78\tabularnewline
+190 & 99990 & 10 & 99945 & 55 & 52 & 7\tabularnewline
+200 & 100000 & 0 & 99995 & 5 & 5 & 0\tabularnewline
+210 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\textbf{Connectance \(C = 0.7\)}
+\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}
 
@@
 \endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
-3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
-4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
-5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
-6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
-7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
-8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
-9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
-10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
-11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
-12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
-13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
-14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
-15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
-16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
-17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
-18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
-19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
-20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
-21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
-22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
-23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
-24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
-25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
-26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
-27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
-28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
-29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
-30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
-31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
-32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
-33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
-34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
-35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 36 & 99964 & 36 & 99964 & 0 & 0\tabularnewline
+20 & 195 & 99805 & 195 & 99805 & 0 & 0\tabularnewline
+30 & 519 & 99481 & 523 & 99477 & 0 & 4\tabularnewline
+40 & 1096 & 98904 & 1101 & 98899 & 2 & 7\tabularnewline
+50 & 2375 & 97625 & 2397 & 97603 & 9 & 31\tabularnewline
+60 & 4898 & 95102 & 4968 & 95032 & 83 & 153\tabularnewline
+70 & 10841 & 89159 & 10916 & 89084 & 432 & 507\tabularnewline
+80 & 22281 & 77719 & 21988 & 78012 & 1622 & 1329\tabularnewline
+90 & 42010 & 57990 & 39998 & 60002 & 4458 & 2446\tabularnewline
+100 & 67289 & 32711 & 63098 & 36902 & 7153 & 2962\tabularnewline
+110 & 88137 & 11863 & 84023 & 15977 & 6108 & 1994\tabularnewline
+120 & 97678 & 2322 & 95557 & 4443 & 2740 & 619\tabularnewline
+130 & 99795 & 205 & 99304 & 696 & 578 & 87\tabularnewline
+140 & 99989 & 11 & 99948 & 52 & 49 & 8\tabularnewline
+150 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\textbf{Connectance \(C = 0.9\)}
+\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}
 
@@
 \endhead
-2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
-3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
-4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
-5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
-6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
-7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
-8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
-9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
-10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
-11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
-12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
-13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
-14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
-15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
-16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
-17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
-18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
-19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
-20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
-21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
-22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
-23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
-24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
-25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
-26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
-27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
-28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
-29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
-30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+10 & 162 & 99838 & 162 & 99838 & 0 & 0\tabularnewline
+20 & 798 & 99202 & 799 & 99201 & 0 & 1\tabularnewline
+30 & 2273 & 97727 & 2289 & 97711 & 6 & 22\tabularnewline
+40 & 5259 & 94741 & 5298 & 94702 & 70 & 109\tabularnewline
+50 & 12084 & 87916 & 12054 & 87946 & 446 & 416\tabularnewline
+60 & 26072 & 73928 & 25511 & 74489 & 1810 & 1249\tabularnewline
+70 & 50121 & 49879 & 47747 & 52253 & 4748 & 2374\tabularnewline
+80 & 77806 & 22194 & 73810 & 26190 & 6421 & 2425\tabularnewline
+90 & 94862 & 5138 & 92069 & 7931 & 3842 & 1049\tabularnewline
+100 & 99527 & 473 & 98822 & 1178 & 870 & 165\tabularnewline
+110 & 99984 & 16 & 99912 & 88 & 80 & 8\tabularnewline
+120 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+130 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+300 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\hypertarget{gam_dist}{\section{\texorpdfstring{Different distributions
-of
-\(\gamma\)}{Different distributions of \textbackslash{}gamma}}\label{gam_dist}}
+\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
+distribution of
+\(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
 
@@
 
-The same 100000 \(M\) matrices were used to investigate stability when
-applying each of these different distributions of \(\gamma\) values. The
-table below shows the number of \(M\) that were unstable (\_unst) and
-stable (\_stbl) for the exponential (Exp), beta, and gamma
-distributions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{fourdists <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/different_distr/four_distr_rand.csv"}\NormalTok{);}
-\KeywordTok{kable}\NormalTok{(fourdists);}
-\end{Highlighting}
-\end{Shaded}
+The same 100000 \(\mathbf{M}\) matrices were used to investigate
+stability when applying each of these different distributions of
+\(\gamma\) values. The table below shows the number of \(\mathbf{M}\)
+that were unstable (\_unst) and stable (\_stbl) for the exponential
+(Exp), beta, and gamma distributions.
 
@@
 
-\hypertarget{ga}{\section{Genetic algorithm}\label{ga}}
-
-Ideally, to investigate the potential of \(Var(\gamma)\) for increasing
-the proportion of stable complex systems, the search space of all
-possible \(\gamma\) vectors would be evaluated for each unique \(M\).
-This is technically impossible because \(\gamma_{i}\) can take any real
-value between 0-2, but even rounding \(\gamma\) to reasonable values
-would result in a search space too large to practically explore. Under
-these conditions, genetic algorithms are highly useful tools for finding
-practical solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{6}}. In
-this case, the practical solution is finding vectors of \(\gamma\) that
-decrease the most positive real eigenvalue of \(M\). The genetic
-algorithm below achieves this by initialising a large population of 1000
-different potential \(\gamma\) vectors and allowing this population to
-evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
-values between vectors), selection, and reproduction until either a
-\(\gamma\) vector is found where all \(\Re(\lambda) < 0\) or some
-``giving up'' critiera is met (in the below, this ``giving up''"
-criteria is met when 20 generations pass, or if the fitness increase
-from one generation to the next is below a certain criteria). The
-genetic algorithm relies on five functions. The first outer function
-\texttt{Evo\_rand\_gen\_var} runs all of the simulations
-(\texttt{max\_sp} refers to the maximum \(S\) value simulated, and
-\texttt{iters} refers to the number of \(M\) to try for each \(S\)).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{Evo_rand_gen_var <-}\StringTok{ }\NormalTok{function(max_sp, iters, }\DataTypeTok{int_type =} \DecValTok{0}\NormalTok{, }\DataTypeTok{rmx =} \FloatTok{0.4}\NormalTok{, }\DataTypeTok{C =} \DecValTok{1}\NormalTok{)\{}
-    \NormalTok{tot_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{fea_res <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
-    \NormalTok{for(i in }\DecValTok{2}\NormalTok{:max_sp)\{}
-        \NormalTok{nn             <-}\StringTok{ }\NormalTok{i;}
-        \NormalTok{A1_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_stt         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A1_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{A2_fet         <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{iter           <-}\StringTok{ }\NormalTok{iters;}
-        \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{3}\NormalTok{);}
-        \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]] <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{iter, }\DataTypeTok{ncol =} \DecValTok{2}\NormalTok{);}
-        \NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-            \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{rmx);}
-            \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\KeywordTok{species_interactions}\NormalTok{(}\DataTypeTok{mat =} \NormalTok{A0, }\DataTypeTok{type =} \NormalTok{int_type);}
-            \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i *}\StringTok{ }\NormalTok{i, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
-            \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =} \NormalTok{i, }\DataTypeTok{ncol =} \NormalTok{i);}
-            \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
-            \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
-            \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{i, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
-            \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
-            \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{rand_mat_ga}\NormalTok{(A1);}
-            \NormalTok{A0_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A0) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{A1_fea   <-}\StringTok{ }\KeywordTok{min}\NormalTok{(-}\DecValTok{1}\NormalTok{*}\KeywordTok{solve}\NormalTok{(A1) %*%}\StringTok{ }\NormalTok{r_vec) >}\StringTok{ }\DecValTok{0}\NormalTok{;}
-            \NormalTok{if(A0_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_stb ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{tot_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A0_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{1}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{if(A1_fea ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{)\{}
-                \NormalTok{fea_res[[i}\DecValTok{-1}\NormalTok{]][iter, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{1}\NormalTok{;}
-            \NormalTok{\}}
-            \NormalTok{iter    <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{\}}
-        \KeywordTok{print}\NormalTok{(i);}
-    \NormalTok{\}}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{summarise_randmat}\NormalTok{(}\DataTypeTok{tot_res =} \NormalTok{tot_res, }\DataTypeTok{fea_res =} \NormalTok{fea_res);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Note that \texttt{Evo\_rand\_gen\_var} calls three custom sub-functions,
-\texttt{species\_interactions}, \texttt{rand\_mat\_ga}, and
-\texttt{summarise\_randmat}. The first simply allows for non-random
-interactions between components (e.g., modelling
-\protect\hyperlink{ecological}{ecological interactions} of random,
-competition, mutualism, or predator-prey).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{species_interactions <-}\StringTok{ }\NormalTok{function(mat, }\DataTypeTok{type =} \DecValTok{0}\NormalTok{)\{}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat >}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{2}\NormalTok{)\{}
-        \NormalTok{mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*mat[mat <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-    \NormalTok{\}}
-    \NormalTok{if(type ==}\StringTok{ }\DecValTok{3}\NormalTok{)\{}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{for(j in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(mat)[}\DecValTok{2}\NormalTok{])\{}
-                \NormalTok{if(mat[i, j] *}\StringTok{ }\NormalTok{mat[j, i] >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
-                    \NormalTok{mat[j, i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1} \NormalTok{*}\StringTok{ }\NormalTok{mat[j, i];}
-                \NormalTok{\}}
-            \NormalTok{\}}
-        \NormalTok{\}}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(mat);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The sub-function \texttt{rand\_mat\_ga} does the work of the genetic
-algorithm, searching for \(\gamma\) vectors that are stabilising.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{rand_mat_ga <-}\StringTok{ }\NormalTok{function(A1, }\DataTypeTok{max_it =} \DecValTok{20}\NormalTok{, }\DataTypeTok{converg =} \FloatTok{0.01}\NormalTok{)\{}
-    \NormalTok{nn       <-}\StringTok{ }\KeywordTok{dim}\NormalTok{(A1)[}\DecValTok{1}\NormalTok{];}
-    \NormalTok{rind     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{1}\NormalTok{);}
-    \NormalTok{inds     <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{rind, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-    \NormalTok{lastf    <-}\StringTok{ }\NormalTok{-}\DecValTok{10}\NormalTok{;}
-    \NormalTok{ccrit    <-}\StringTok{ }\DecValTok{10}\NormalTok{;}
-    \NormalTok{find_st  <-}\StringTok{ }\DecValTok{0}\NormalTok{;}
-    \NormalTok{iter     <-}\StringTok{ }\NormalTok{max_it;}
-    \NormalTok{while(iter >}\StringTok{ }\DecValTok{0} \NormalTok{&}\StringTok{ }\NormalTok{find_st <}\StringTok{ }\DecValTok{1} \NormalTok{&}\StringTok{ }\NormalTok{ccrit >}\StringTok{ }\NormalTok{converg)\{}
-        \NormalTok{ivar  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{ifit  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{isst  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{length =} \KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{]);}
-        \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{])\{}
-            \NormalTok{ifit[i] <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{*}\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values));}
-            \NormalTok{ivar[i] <-}\StringTok{ }\KeywordTok{var}\NormalTok{(inds[i,]);}
-            \NormalTok{isst[i] <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(inds[i,]*A1)$values)) <}\StringTok{ }\DecValTok{0}\NormalTok{;}
-        \NormalTok{\}}
-        \NormalTok{most_fit <-}\StringTok{ }\KeywordTok{order}\NormalTok{(ifit, }\DataTypeTok{decreasing =} \OtherTok{TRUE}\NormalTok{)[}\DecValTok{1}\NormalTok{:}\DecValTok{20}\NormalTok{];}
-        \NormalTok{parents  <-}\StringTok{ }\NormalTok{inds[most_fit,];}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \KeywordTok{t}\NormalTok{(parents), }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn, }
-                           \DataTypeTok{byrow =} \OtherTok{TRUE}\NormalTok{);}
-        \NormalTok{mu_dat   <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \FloatTok{0.2}\NormalTok{);}
-        \NormalTok{mu_dat2  <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{nn*}\DecValTok{1000}\NormalTok{, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.02}\NormalTok{);}
-        \NormalTok{mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{] <-}\StringTok{ }\NormalTok{-mu_dat2[mu_dat2 <}\StringTok{ }\DecValTok{0}\NormalTok{];}
-        \NormalTok{mu_dat2[mu_dat2 >}\StringTok{ }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\DecValTok{2}\NormalTok{;}
-        \NormalTok{mu_dat3  <-}\StringTok{ }\NormalTok{mu_dat *}\StringTok{ }\NormalTok{mu_dat2;}
-        \NormalTok{mu_mat   <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{mu_dat3, }\DataTypeTok{nrow =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{ncol =} \NormalTok{nn);}
-        \NormalTok{new_gen  <-}\StringTok{ }\NormalTok{new_gen +}\StringTok{ }\NormalTok{mu_mat;}
-        \NormalTok{new_gen  <-}\StringTok{ }\KeywordTok{crossover}\NormalTok{(}\DataTypeTok{inds =} \NormalTok{new_gen, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{);}
-        \NormalTok{inds     <-}\StringTok{ }\NormalTok{new_gen;}
-        \NormalTok{find_st  <-}\StringTok{ }\KeywordTok{max}\NormalTok{(isst);}
-        \NormalTok{newf     <-}\StringTok{ }\KeywordTok{mean}\NormalTok{(ifit);}
-        \NormalTok{ccrit    <-}\StringTok{ }\NormalTok{newf -}\StringTok{ }\NormalTok{lastf;}
-        \NormalTok{lastf    <-}\StringTok{ }\NormalTok{newf;}
-        \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
-    \NormalTok{\}}
-    \NormalTok{if(find_st ==}\StringTok{ }\DecValTok{1}\NormalTok{)\{}
-        \NormalTok{s_row <-}\StringTok{ }\KeywordTok{which}\NormalTok{(isst ==}\StringTok{ }\DecValTok{1}\NormalTok{)[}\DecValTok{1}\NormalTok{];}
-        \NormalTok{writt <-}\StringTok{ }\KeywordTok{c}\NormalTok{(nn, inds[s_row,]);}
-        \KeywordTok{cat}\NormalTok{(writt, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-        \KeywordTok{cat}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"evo_out.txt"}\NormalTok{, }\DataTypeTok{append =} \OtherTok{TRUE}\NormalTok{);}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(find_st);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-The while loop in \texttt{rand\_mat\_ga} continues until either
-\texttt{iter} generations have occured, a solution \(\gamma\) vector is
-found that results in all \(\Re(\lambda) < 0\), or some criteria of
-minimum fitness increase is observed (by default,
-\texttt{converg\ =\ 0.01}). Within the genetic algorithm, \(\gamma\)
-values are mutated, crossover occurs between \(\gamma\) vectors, and
-selection occurs in each generation such that the 20 \(\gamma\) vectors
-that produce the lowest maximum \(\Re(\lambda)\) are allowed to have 50
-offspring each. In mutation, any \(\gamma_{i}\) values that mutate below
-zero are multiplied by \(-1\), and any values that mutate above 2 are
-set to 2. Note also that if a solution is found, then one such
-\(\gamma\) vector causing stability is printed to a file.
-
-Crossover occurs in the \texttt{crossover} function below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{crossover <-}\StringTok{ }\NormalTok{function(inds, }\DataTypeTok{pr =} \FloatTok{0.1}\NormalTok{)\{}
-    \NormalTok{crossed <-}\StringTok{ }\KeywordTok{floor}\NormalTok{(}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{] *}\StringTok{ }\NormalTok{pr);}
-    \NormalTok{cross1  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{cross2  <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{1}\NormalTok{], }\DataTypeTok{size =} \NormalTok{crossed);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(cross1))\{}
-        \NormalTok{fromv   <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{tov     <-}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\DataTypeTok{x =} \DecValTok{1}\NormalTok{:}\KeywordTok{dim}\NormalTok{(inds)[}\DecValTok{2}\NormalTok{], }\DataTypeTok{size =} \DecValTok{1}\NormalTok{);}
-        \NormalTok{temp                   <-}\StringTok{ }\NormalTok{inds[cross1[i],fromv:tov];}
-        \NormalTok{inds[cross1[i],fromv:tov] <-}\StringTok{ }\NormalTok{inds[cross2[i],fromv:tov];}
-        \NormalTok{inds[cross2[i],fromv:tov] <-}\StringTok{ }\NormalTok{temp;}
-    \NormalTok{\}}
-    \KeywordTok{return}\NormalTok{(inds);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-After all \(M\) are simulated in \texttt{Evo\_rand\_gen\_var}, the
-\texttt{summarise\_randmat} formats the data into a table.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{summarise_randmat <-}\StringTok{ }\NormalTok{function(tot_res, fea_res)\{}
-    \NormalTok{sims    <-}\StringTok{ }\KeywordTok{length}\NormalTok{(tot_res);}
-    \NormalTok{all_res <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \DecValTok{0}\NormalTok{, }\DataTypeTok{nrow =} \NormalTok{sims, }\DataTypeTok{ncol =} \DecValTok{7}\NormalTok{);}
-    \NormalTok{for(i in }\DecValTok{1}\NormalTok{:sims)\{}
-        \NormalTok{A0_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{A0_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{A1_unst   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{A1_stbl   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{stabled   <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE} \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}\NormalTok{;}
-        \NormalTok{unstabled <-}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{1}\NormalTok{] ==}\StringTok{ }\OtherTok{TRUE}  \NormalTok{&}\StringTok{ }\NormalTok{tot_res[[i]][,}\DecValTok{2}\NormalTok{] ==}\StringTok{ }\OtherTok{FALSE}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{1}\NormalTok{]  <-}\StringTok{ }\NormalTok{i +}\StringTok{ }\DecValTok{1}\NormalTok{;}
-        \NormalTok{all_res[i, }\DecValTok{2}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_unst);}
-        \NormalTok{all_res[i, }\DecValTok{3}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A0_stbl);}
-        \NormalTok{all_res[i, }\DecValTok{4}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_unst);}
-        \NormalTok{all_res[i, }\DecValTok{5}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(A1_stbl);}
-        \NormalTok{all_res[i, }\DecValTok{6}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(stabled);}
-        \NormalTok{all_res[i, }\DecValTok{7}\NormalTok{]  <-}\StringTok{ }\KeywordTok{sum}\NormalTok{(unstabled);}
-    \NormalTok{\}}
-    \KeywordTok{colnames}\NormalTok{(all_res) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"N"}\NormalTok{, }\StringTok{"A0_unstable"}\NormalTok{, }\StringTok{"A0_stable"}\NormalTok{, }\StringTok{"A1_unstable"}\NormalTok{,}
-                           \StringTok{"A1_stable"}\NormalTok{, }\StringTok{"A1_stabilised"}\NormalTok{, }\StringTok{"A1_destabilised"}\NormalTok{);}
-    \KeywordTok{return}\NormalTok{(all_res);}
-\NormalTok{\}}
-\end{Highlighting}
-\end{Shaded}
-
-Stability results for 40000 \(M\) for each \(S\) from 2-40 are shown below. Results for \texttt{A0}
-indicate systems in which \(\gamma = 1\), while \texttt{A1} refers to systems in 
-which the genetic algorithm searched for a set of \(\gamma\) values that 
-stabilised the system.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 13 & 39987 & 13 & 39987 & 0 & 0\tabularnewline
-3 & 157 & 39843 & 157 & 39843 & 0 & 0\tabularnewline
-4 & 563 & 39437 & 563 & 39437 & 0 & 0\tabularnewline
-5 & 1588 & 38412 & 1581 & 38419 & 7 & 0\tabularnewline
-6 & 3268 & 36732 & 3219 & 36781 & 49 & 0\tabularnewline
-7 & 5370 & 34630 & 5196 & 34804 & 175 & 1\tabularnewline
-8 & 8231 & 31769 & 7604 & 32396 & 627 & 0\tabularnewline
-9 & 11704 & 28296 & 10181 & 29819 & 1523 & 0\tabularnewline
-10 & 15215 & 24785 & 12303 & 27697 & 2912 & 0\tabularnewline
-11 & 19529 & 20471 & 14599 & 25401 & 4930 & 0\tabularnewline
-12 & 23480 & 16520 & 16210 & 23790 & 7270 & 0\tabularnewline
-13 & 27570 & 12430 & 17800 & 22200 & 9770 & 0\tabularnewline
-14 & 31096 & 8904 & 18899 & 21101 & 12198 & 1\tabularnewline
-15 & 34036 & 5964 & 19948 & 20052 & 14089 & 1\tabularnewline
-16 & 36231 & 3769 & 21000 & 19000 & 15231 & 0\tabularnewline
-17 & 37737 & 2263 & 21693 & 18307 & 16044 & 0\tabularnewline
-18 & 38808 & 1192 & 22291 & 17709 & 16518 & 1\tabularnewline
-19 & 39391 & 609 & 23221 & 16779 & 16170 & 0\tabularnewline
-20 & 39711 & 289 & 23943 & 16057 & 15768 & 0\tabularnewline
-21 & 39867 & 133 & 25361 & 14639 & 14506 & 0\tabularnewline
-22 & 39950 & 50 & 26584 & 13416 & 13366 & 0\tabularnewline
-23 & 39982 & 18 & 28162 & 11838 & 11820 & 0\tabularnewline
-24 & 39994 & 6 & 29660 & 10340 & 10334 & 0\tabularnewline
-25 & 39997 & 3 & 31405 & 8595 & 8592 & 0\tabularnewline
-26 & 39999 & 1 & 33057 & 6943 & 6942 & 0\tabularnewline
-27 & 40000 & 0 & 34602 & 5398 & 5398 & 0\tabularnewline
-28 & 40000 & 0 & 35957 & 4043 & 4043 & 0\tabularnewline
-29 & 40000 & 0 & 37154 & 2846 & 2846 & 0\tabularnewline
-30 & 40000 & 0 & 38104 & 1896 & 1896 & 0\tabularnewline
-31 & 40000 & 0 & 38722 & 1278 & 1278 & 0\tabularnewline
-32 & 40000 & 0 & 39259 & 741 & 741 & 0\tabularnewline
-33 & 40000 & 0 & 39551 & 449 & 449 & 0\tabularnewline
-34 & 40000 & 0 & 39742 & 258 & 258 & 0\tabularnewline
-35 & 40000 & 0 & 39879 & 121 & 121 & 0\tabularnewline
-36 & 40000 & 0 & 39929 & 71 & 71 & 0\tabularnewline
-37 & 40000 & 0 & 39967 & 33 & 33 & 0\tabularnewline
-38 & 40000 & 0 & 39989 & 11 & 11 & 0\tabularnewline
-39 & 40000 & 0 & 39994 & 6 & 6 & 0\tabularnewline
-40 & 40000 & 0 & 39996 & 4 & 4 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. Recall that 1 million random matrices were
-generated for the less computationally intense task of
-\protect\hyperlink{IncrS}{comparing} \(M\) when \(\gamma = 1\) versus
-when \(\gamma \sim \mathcal{U}(0, 2)\), so it is more informative to
-compare stability in column 5 above with column 3 above. \textbf{This
-comparison shows the high number of stable \(M\) that can be produced
-through a targetted search of \(\gamma\) values, and suggests that many
-otherwise unstable systems could potentially be stabilised by an
-informed manipulation of their component response times. Such a
-possibility might conceivably reduce the dimensionality of problems
-involving stability in social-ecological or economic systems.}
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{evo_out <-}\StringTok{ }\KeywordTok{scan}\NormalTok{(}\DataTypeTok{file =} \StringTok{"sim_results/evolved/evo_out.txt"}\NormalTok{);}
-\KeywordTok{plot_evo_out}\NormalTok{(evo_out);}
-\end{Highlighting}
-\end{Shaded}
-
-\includegraphics{unnamed-chunk-31-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix \(M\).
-
 \hypertarget{Feasibility}{\section{Feasibility of complex
@@
 
-For complex systems in which individual system components (\(S\))
-represent the density of some tangible quantity, it is important to
-consider the feasibility of the system. Feasibile equilibria assume that
-the values of all system components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{7}--\protect\hyperlink{ref-Song2018}{9}}.
-This is of particular interest for ecological communities because
-population density cannot take negative values, meaning that ecological
-systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
-Consequently, the use of random matrices and traditional stability
-critiera for making inferences in theoretical analyses of species
-networks has recently been
-criticised\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}.
-While the key results in the main text are intended to be general to all
-complex systems, and not restricted to species networks, I have also
-performed a feasibility analysis on all matrices \(M\). This analysis
-reveals that feasibility is not affected by \(Var(\gamma)\), meaning
-that for pure interacting species networks, variation in component
-response time (i.e., species generation time) does not affect stability
-at biologically realistic species densities. Nevertheless, ecological
-interactions do not exist in isolation in empirical systems, but instead
-interact with
-evolutionary\textsuperscript{\protect\hyperlink{ref-Patel2018}{10}},
-abiotic, or social-economic systems. The relevance of \(\gamma\) for
-complex system stability presented in the main text should therefore not
-be ignored in the broader context of ecological communities.
-
-Dougoud et al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}}
-define the following feasibility criteria for ecological systems
-characterised by \(S\) interacting species with varying densities.
-
-\[x^{*} = -\left(\theta I + (CS)^{-\delta}A\right)^{-1}r.\]
-
-In the above, \(x^{*}\) is the vector of species abundances at
-equilibrium (for feasibility, all values in \(x^{*}\) must be positive).
-The matrix \(I\) is the identity matrix (1s on the diagonal, 0s on the
-off-diagonal elements), and the value \(\theta\) is strength of
-intraspecific competition (diagonal elements). As I have done elsewhere,
-diagonal values are set to \(-1\), so \(\theta = -1\). The variable
-\(C\) is the connectance of the community,
-which was set to \(C = 1\) throughout the manuscript and supplemental
-information, except \protect\hyperlink{connectance}{where otherwise
-noted}. The variable \(\delta\) is a normalisation parameter that
-modulates the strength of interactions (\(\sigma\) in the main text),
-which are held in \(A\). In the main text, implicitly, \(\delta = 0\)
-underlying strong interactions. Hence, the whole \((CS)^{-\delta} = 1\),
-so in the above, a diagonal matrix of -1s (\(\theta I\)) is added to
-\(A\), which has a diagonal of all zeros and an off-diagonal affecting
-species interactions (i.e., the expression \((CS)^{-\delta}\) relates to
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{8}} by
-\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < -1\), and hence
-\((CS)^{-\delta} = 1\) for the randomly simulated systems in the main
-text and Supplementary Information). The above criteria is therefore
-reduced to the below; note that the parenthetical in both equations
-produces an \(M\) matrix as used throughout the main text and
-supplemental information,
-
-\[x^{*} = -\left(\theta I + A\right)^{-1}r.\]
-
-To check the feasibility criteria, I therefore inverted
-\(M = (\theta I + A)\) and multiplied elements by -1, then multiplied
-the resulting matrix by the vector of population growth rates \(r\).
-Feasibility is satisfied if all of the elements of the resulting vector
-are positive.
-
-The population growth rate for an individual species \(i\) is sampled
-from a normal distribution of \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\), as
-shown in the \texttt{rand\_gen\_var} function in
-\protect\hyperlink{IncrS}{the section} on ``Stability across increasing
-\(S\)'' above. Hence, each component \(i\) of the complex system \(M\)
-is assumed to be a species with a growth rate of \(r_{i}\). Note that
-negative intrinsic growth rates are not unrealistic, and will occur in
-obligate mutualists in the absence of a partner.
-
 When feasibility was evaluated with and without variation in \(\gamma\),
-there was no increase in stability for \(M\) where \(\gamma\) varied as
-compared to where \(\gamma = 1\). Results below illustrate this result,
-which was general to all other simulations performed.
+there was no increase in stability for \(\mathbf{M}\) where \(\gamma\)
+varied as compared to where \(\gamma = 1\). Results below illustrate
+this result, which was general to all other simulations performed.
 
@@
 Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
-in pure species interaction networks.
+in pure species interaction
+networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.
+
+\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
+(2018)}\label{Gibbs}}
+
+The question that I address in the main text is distinct from that of
+Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
+focused instead on the effect of a diagonal matrix of biological species
+densities \(\mathbf{X}\) on a community matrix \(\mathbf{M}\) given a
+species interaction matrix \(\mathbf{A}\). This is modelled as below,
+
+\[\mathbf{M} = \mathbf{XA}.\]
+
+Mathematically, the above is identical to my model in the main text
+where the system \(\mathbf{M}\) is defined by component interaction
+strengths \(\mathbf{A}\) and individual component response rates
+\(\mathbf{\gamma}\),
+
+\[\mathbf{M} = \mathbf{\gamma A}.\]
+
+I focused on the probability of observing a stable versus unstable
+system given variation in \(\mathbf{\gamma}\) as system complexity
+(\(\sigma\sqrt{SC}\)) increased. I increased system complexity by
+holding \(C\) and \(\sigma\) constant and incrementally increasing \(S\)
+to obtain numerical results. In contrast, Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} applied
+analytical techniques to instead focus on a different question
+concerning the effect of \(\mathbf{\gamma}\) on the stability of
+\(\mathbf{M}\) given \(\mathbf{A}\) as \(S \to \infty\), with \(\sigma\)
+scaled so that \(\sigma = 1/\sqrt{S}\). Under such scaling, Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that
+the effect of \(\gamma\) on stability should decrease exponentially as
+\(S\) increases, which I demonstrate below by running simulations in
+which \(\sigma = 1/\sqrt{S}\).
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 3111 & 96889 & 3111 & 96889 & 0 & 0\tabularnewline
+3 & 5203 & 94797 & 5237 & 94763 & 1 & 35\tabularnewline
+4 & 6743 & 93257 & 6818 & 93182 & 6 & 81\tabularnewline
+5 & 7889 & 92111 & 8005 & 91995 & 20 & 136\tabularnewline
+6 & 8834 & 91166 & 8991 & 91009 & 55 & 212\tabularnewline
+7 & 9885 & 90115 & 10072 & 89928 & 81 & 268\tabularnewline
+8 & 10516 & 89484 & 10764 & 89236 & 108 & 356\tabularnewline
+9 & 11135 & 88865 & 11383 & 88617 & 145 & 393\tabularnewline
+10 & 11819 & 88181 & 12095 & 87905 & 181 & 457\tabularnewline
+11 & 12414 & 87586 & 12700 & 87300 & 213 & 499\tabularnewline
+12 & 12865 & 87135 & 13136 & 86864 & 283 & 554\tabularnewline
+13 & 13530 & 86470 & 13836 & 86164 & 324 & 630\tabularnewline
+14 & 13745 & 86255 & 14042 & 85958 & 362 & 659\tabularnewline
+15 & 14401 & 85599 & 14720 & 85280 & 387 & 706\tabularnewline
+16 & 14793 & 85207 & 15123 & 84877 & 428 & 758\tabularnewline
+17 & 15004 & 84996 & 15356 & 84644 & 444 & 796\tabularnewline
+18 & 15361 & 84639 & 15735 & 84265 & 472 & 846\tabularnewline
+19 & 16062 & 83938 & 16303 & 83697 & 592 & 833\tabularnewline
+20 & 15814 & 84186 & 16184 & 83816 & 566 & 936\tabularnewline
+21 & 16171 & 83829 & 16492 & 83508 & 640 & 961\tabularnewline
+22 & 16671 & 83329 & 17049 & 82951 & 641 & 1019\tabularnewline
+23 & 17000 & 83000 & 17291 & 82709 & 718 & 1009\tabularnewline
+24 & 17411 & 82589 & 17666 & 82334 & 765 & 1020\tabularnewline
+25 & 17414 & 82586 & 17742 & 82258 & 783 & 1111\tabularnewline
+26 & 17697 & 82303 & 18027 & 81973 & 806 & 1136\tabularnewline
+27 & 18010 & 81990 & 18316 & 81684 & 880 & 1186\tabularnewline
+28 & 18584 & 81416 & 18735 & 81265 & 1008 & 1159\tabularnewline
+29 & 18401 & 81599 & 18572 & 81428 & 942 & 1113\tabularnewline
+30 & 18497 & 81503 & 18754 & 81246 & 952 & 1209\tabularnewline
+31 & 18744 & 81256 & 18942 & 81058 & 991 & 1189\tabularnewline
+32 & 18936 & 81064 & 19194 & 80806 & 1022 & 1280\tabularnewline
+33 & 19174 & 80826 & 19346 & 80654 & 1113 & 1285\tabularnewline
+34 & 19477 & 80523 & 19632 & 80368 & 1120 & 1275\tabularnewline
+35 & 19659 & 80341 & 19777 & 80223 & 1206 & 1324\tabularnewline
+36 & 19883 & 80117 & 19929 & 80071 & 1275 & 1321\tabularnewline
+37 & 20275 & 79725 & 20348 & 79652 & 1308 & 1381\tabularnewline
+38 & 20067 & 79933 & 20190 & 79810 & 1275 & 1398\tabularnewline
+39 & 20416 & 79584 & 20516 & 79484 & 1340 & 1440\tabularnewline
+40 & 20370 & 79630 & 20489 & 79511 & 1359 & 1478\tabularnewline
+41 & 20295 & 79705 & 20430 & 79570 & 1382 & 1517\tabularnewline
+42 & 20767 & 79233 & 20839 & 79161 & 1418 & 1490\tabularnewline
+43 & 20688 & 79312 & 20705 & 79295 & 1471 & 1488\tabularnewline
+44 & 21049 & 78951 & 21028 & 78972 & 1555 & 1534\tabularnewline
+45 & 21114 & 78886 & 21034 & 78966 & 1572 & 1492\tabularnewline
+46 & 21163 & 78837 & 21195 & 78805 & 1463 & 1495\tabularnewline
+47 & 21373 & 78627 & 21353 & 78647 & 1535 & 1515\tabularnewline
+48 & 21338 & 78662 & 21285 & 78715 & 1632 & 1579\tabularnewline
+49 & 21547 & 78453 & 21566 & 78434 & 1575 & 1594\tabularnewline
+50 & 21738 & 78262 & 21633 & 78367 & 1636 & 1531\tabularnewline
+51 & 21967 & 78033 & 21892 & 78108 & 1698 & 1623\tabularnewline
+\bottomrule
+\end{longtable}
+
+Above table results can be compared to those of the
+\protect\hyperlink{IncrS}{main results}. Note that 100000 (not 1
+million), simulations are run to confirm consistency with Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. The
+difference between my model and Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} is that in the
+latter, \(\sigma\sqrt{SC} = 1\) remains constant with increasing \(S\).
+In the former, \(\sigma\sqrt{SC}\) increases with \(S\), so the expected
+complexity of the system also increases accordingly. Consequently, for
+the scaled \(\sigma\) in the table above, systems are not more likely to
+be stabilised by \(\gamma\) as \(S\) increases, consistent with Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. Note that
+overall stability does decrease with increasing \(S\) due to the
+increased density of eigenvalues (see below).
+
+\begin{figure}[h]
+\includegraphics{unnamed-chunk-27-1}
+\end{figure}
+
+\textbf{Complexity as a function of \(S\) in the main text (solid)
+versus in Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} (dashed).}
+
+When the complexity is scaled to \(\sigma\sqrt{SC} = 1\), an increase in
+\(S\) increases the eigenvalue density within a circle with a unit
+radius centred at \((-1, 0)\) on the complex plane. As \(S \to \infty\),
+this circle becomes increasingly saturated. Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that a
+diagonal matrix \(\mathbf{\gamma}\) will have an exponentially
+decreasing effect on stability with increasing \(S\). Increasing \(S\)
+is visualised below, first with a system size \(S = 100\).
+
+\includegraphics{unnamed-chunk-29-1.pdf}
+
+The left panel above shows the distribution of eigenvalues; the blue
+ellipse shows the unit radius within which eigenvalues are expected to
+be contained. The right panel shows how eigenvalue distributions change
+given \(\gamma \sim \mathcal{U}(0,2)\). The vertical dotted line shows
+the threshold of stability, \(\Re = 0\). Increasing to \(S = 200\), the
+scaling \(\sigma = 1 / \sqrt{S}\) maintains the expected distribution of
+eigenvalues but increases eigenvalue density.
+
+\includegraphics{unnamed-chunk-30-1.pdf}
+
+We can increase the system size to \(S = 500\) and see the corresponding
+increase in eigenvalue density.
+
+\includegraphics{unnamed-chunk-31-1.pdf}
+
+Finally, below shows a increase in system size to \(S = 1000\).
+
+\includegraphics{unnamed-chunk-32-1.pdf}
+
+In contrast, in the model of the main text, the complexity of system is
+not scaled to \(\sigma\sqrt{SC} = 1\). Rather, the density of
+eigenvalues within a circle centred at \((-1, 0)\) with a radius
+\(\sigma\sqrt{SC}\) is held constant such that there are
+\(S / \pi(\sigma\sqrt{SC})^2\) eigenvalues per unit area of the circle.
+As \(S\) increases, so does the expected complexity of the system, but
+the density of eigenvalues remains finite causing error around this
+expectation. Below shows a system where \(S = 100\), \(C = 0.0625\), and
+\(\sigma = 0.4\), where \(\sigma \sqrt{SC} = 1\) (identical to the first
+example distribution above in which \(S = 100\) and
+\(\sigma = 1/\sqrt{S}\)).
+
+\includegraphics{unnamed-chunk-34-1.pdf}
+
+Now when \(S\) is increased to \(200\) while keeping \(C = 0.0625\) and
+\(\sigma = 0.4\), the area of the circle within which eigenvalues are
+contained increases to keep the density of eigenvalues constant.
+
+\includegraphics{unnamed-chunk-35-1.pdf}
+
+Note that the expected distribution of eigenvalues increases so that the
+threshold \(\Re = 0\) is exceeded. Below, system size is increased to
+\(S = 500\).
+
+\includegraphics{unnamed-chunk-36-1.pdf}
+
+Finally, \(S = 1000\) is shown below. Again, the density of eigenvalues
+per unit remains constant at ca 2, but the system has increased in
+complexity such that some real components of eigenvalues are almost
+assured to be greater than zero.
+
+\includegraphics{unnamed-chunk-37-1.pdf}
 
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Patel2018}{}
-1. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the effects
-of eco-evolutionary feedbacks on community stability. \emph{American
-Naturalist} \textbf{191,} 1--29 (2018).
-
-\hypertarget{refs}{}
 \hypertarget{ref-May1972}{}
-2. May, R. M. Will a large complex system be stable? \emph{Nature}
+1. May, R. M. Will a large complex system be stable? \emph{Nature}
 \textbf{238,} 413--414 (1972).
 
-\hypertarget{ref-May1973}{}
-3. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
-\textbf{54,} 638--641 (1973).
-
 \hypertarget{ref-Allesina2012}{}
-4. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
 \emph{Nature} \textbf{483,} 205--208 (2012).
 
-\hypertarget{ref-Allesina2015a}{}
-5. Allesina, S. \& Tang, S. The stability--complexity relationship at
-age 40: a random matrix perspective. \emph{Population Ecology} 63--75
+\hypertarget{ref-Allesina2015}{}
+3. Allesina, S. \emph{et al.} Predicting the stability of large
+structured food webs. \emph{Nature Communications} \textbf{6,} 7842
 (2015).
-doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
-
-\hypertarget{ref-Tang2014b}{}
-6. Tang, S. \& Allesina, S. Reactivity and stability of large
-ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
-(2014).
-
-\hypertarget{ref-Allesina2011}{}
-7. Allesina, S. \& Levine, J. M. A competitive network theory of species
-diversity. \emph{Proceedings of the National Academy of Sciences of the
-United States of America} \textbf{108,} 5638--5642 (2011).
-
-\hypertarget{ref-Hamblin2013}{}
-8. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
-
-\hypertarget{ref-Grilli2017}{}
-9. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
-
-\hypertarget{ref-Dougoud2018}{}
-10. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
-
-\hypertarget{ref-Song2018}{}
-11. Song, C. \& Saavedra, S. Will a small randomly assembled community be
-feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
+
+\hypertarget{ref-Servan2018}{}
+4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+Allesina, S. Coexistence of many species in random ecosystems.
+\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).
+
+\hypertarget{ref-Gibbs2017}{}
+5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{98,} 022410 (2018).
 

### 2019-03-03 notebook/ms.Rmd
Tweaks

@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) in which $\gamma$ varies are actually more likely to be stable.
+Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary
 

### 2019-03-06 notebook/SI.Rmd
Minor edits; arxiv ready

@@
 
-The question that I address in the main text is distinct from that of Gibbs et al [@Gibbs2017], which focused instead on the effect of a diagonal matrix of biological species densities $\mathbf{X}$ on a community matrix $\mathbf{M}$ given a species interaction matrix $\mathbf{A}$.
+The question that I address in the main text is distinct from that of Gibbs et al.
+[@Gibbs2017], who focused instead on the effect of a diagonal matrix of biological species densities $\mathbf{X}$ on a community matrix $\mathbf{M}$ given a species interaction matrix $\mathbf{A}$.
 This is modelled as below,

### 2019-03-06 notebook/arxiv/ms.tex
Minor edits; arxiv ready

@@
 potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
-large but finite systems, where the probability of stability would
+\(\boldsymbol{\gamma}\) is especially important for stability in
+highly complex systems, in which the probability of stability would
 otherwise be negligible. At such extremes of simulated system
@@
 decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)). My numerical results do not contradict this
+(\(S \to \infty\)) for a given magnitude of complexity \(\sigma\sqrt{SC}\). 
+My numerical results do not contradict this
 prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
 instead fixed \(\sigma\) and increased \(S\) to thereby increase total
-system complexity, \(\sigma\sqrt{SC}\) (see Supplementary Information for
+system complexity (see Supplementary Information for
 results simulated across \(\sigma\) and \(C\)). Overall, I show that
@@
 The question that I address in the main text is distinct from that of
-Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
 focused instead on the effect of a diagonal matrix of biological species

### 2019-03-06 notebook/ms.Rmd
Minor edits; arxiv ready

@@
 I show that when component response rates vary, the potential for system stability is markedly increased.
-Variation in $\boldsymbol{\gamma}$ becomes increasingly important for stability in large but finite systems, where the probability of stability would otherwise be negligible.
+Variation in $\boldsymbol{\gamma}$ is especially important for stability in highly complex systems, in which the probability of stability would otherwise be negligible.
 At such extremes of simulated system complexity, the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
@@
 Gibbs et al.
-[@Gibbs2017] analytically demonstrated that the effect of $\mathbf{X}$ on system stability decreases exponentially as system size becomes arbitrarily large ($S \to \infty$).
-My numerical results do not contradict this prediction because I did not scale $\sigma = 1 / \sqrt{S}$, but instead fixed $\sigma$ and increased $S$ to thereby increase total system complexity, $\sigma\sqrt{SC}$ (see Supplemental Information for results simulated across $\sigma$ and $C$).
+[@Gibbs2017] analytically demonstrated that the effect of $\mathbf{X}$ on system stability decreases exponentially as system size becomes arbitrarily large ($S \to \infty$) for a given magnitude of complexity $\sigma\sqrt{SC}$.
+My numerical results do not contradict this prediction because I did not scale $\sigma = 1 / \sqrt{S}$, but instead fixed $\sigma$ and increased $S$ to thereby increase total system complexity (see Supplemental Information for results simulated across $\sigma$ and $C$).
 Overall, I show that component response rate variation increases the upper bound of complexity at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.

### 2019-03-07 notebook/reviewer_comments_and_responses.Rmd
Update for submission to Scientific Reports

---
title: "Reviewer comments and repsonses"
author: "Brad Duthie"
date: "7 March 2019"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - nature.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

**The following notes acknowledge and respond to reviewer comments from a submission to *PLoS Computational Biology***.

Reviewer 1
================================================================================

Reviewer 1 General comments
--------------------------------------------------------------------------------

This paper takes on a timely question: given a random matrix A and diagonal non-negative matrix X, when will M= XA be stable.  It's important b/c this more accurately reflects the dynamics near equilibrium in a system where species abundances are variable (and then are equal to the diagonal elements of X).  The current paper also claims a different interpretation of the stability of a matrix of this form, in connection with variable response rates across species.

Even though this question probably did lie dormant for many years, it has been tackled more recently, and some of the work cited here seems to tackle very similar questions (e.g. Gibbs et al). But those connections/differences aren't really explained here. In short, I'd like to understand if what the author has identified goes beyond what is known from other work about the effect of a diagonal matrix X on the stability of M.

I found the presentation and framing of these results overall very confusing. Here are some comments:

Response to Reviewer 1 General comments
--------------------------------------------------------------------------------

I am grateful for the helpful comments of Reviewer 1. It is important to first emphasise that my focal questions on the stability of complex systems are not restricted to ecological communities (in which all rows and columns of matrices A and M are species densities), but instead address more general questions regarding how the stability of complex systems is affected by varying response rate of system components (with the exception of the specific question that I address concerning ecological community feasibility, of course). System components might be species densities, but they could additionally include a mix of evolving species traits (e.g., Patel et al. 2018) or human social decisions (e.g., management, harvesting, rewilding, land use, etc.; Hui and Richardson 2018). Components might also not include any species, as in the case of purely social systems (e.g., banks; Haldane & May 2011), or physiological or physical systems (e.g., brain networks; Gray & Robinson 2008; 2009). Reviewer 1 is correct that my paper offers a different interpretation of system stability given the diagonal non-negative matrix $X$, one in which elements $X_{i}$ ($\gamma_{i}$ in my manuscript) are interpreted as the rates at which components respond to perturbation.

The question that I address is distinct from that of Gibbs et al (2018; see also Reviewer 3 comments and responses). I have now revised the manuscript to explain these differences and connections more clearly, both mathematically and conceptually. Mathematically distinct from Gibbs et al. (2018), I am not focused on the effect of a diagonal non-negative matrix $X$ on a stable matrix A as $S \to \infty$ (i.e., as the number of system components becomes arbitrarily large given a stable system), and I specifically focus on a range of system complexities by not forcing the variance of interaction strengths to scale with increasing system size ($\sigma = 1/\sqrt{S}$). Conceptually distinct, I am not focused specifically on the effect of species abundances on ecosystem stability in a generalised Lotka-Volterra system as species number increases. Rather, I am interested in how the stability of M is affected by X and A at increasingly high levels of system complexity, in which the probability of a random complex system being stable becomes rare but non-negligible (under conditions originally identified by May 1972), and for any type of complex system. Clarifying revisions have been now made accordingly to the Introduction and Discussion (most notably, a new paragraph has been added to the Discussion).

Reviewer 1 specific comment 1
--------------------------------------------------------------------------------

(a) The most obvious interpretation (to me) for gamma_i is as the set of equilibrium abundances for the set of populations, N_i. But  this intepretation is not explained clearly, and indeed would presumably not lead one to restrict the support of gamma to (0,2), or (given what we know about realistic distributions of abundances) to consider cases with half the species abundant, and half very rare.

Response to Reviewer 1 specific comment 1
--------------------------------------------------------------------------------

The use of random matrix theory to investigate the stability of complex systems has perhaps most often been applied to ecological communities, in which the complex system is restricted entirely to interacting species where interactions are described by a square matrix A of dimensions $S \times S$. This matrix A can then be multiplied by a vector describing the density of each of S species ($N$) to find the change in species density ($N'$) over some time $t$: $N' = AV$ (May 1972; 1973). It is therefore understandable why the most obvious interpretation of a vector $\gamma$ of length S might be equilibrium species abundances, particularly for those with a background in theoretical ecology; nevertheless, this is an incorrect interpretation of $\gamma$.

In my original manuscript, I was careful to define the vector $\gamma$ as the rate at which the components of a complex system (potentially, but not necessarily, biological species) respond to system perturbation. For the specific question of ecological community feasibility, I also explicitly defined equilibrium abundances as the vector n. I now clarify the use and interpretation of $\gamma$ earlier in the manuscript, and explicitly state that $\gamma$ is not interpreted as a vector of species abundances.

Reviewer 1 specific comment 2
--------------------------------------------------------------------------------

(b) If there is another interpretation of gamma_i in terms of evolution of traits, then this is not adequately explained.  Patel et al is cited in support of this, where stability of an eco-evolutionary system is explored using a block matrix of S species and a set of traits that can evolve. But it isn't clear to me in what limit the stability of such a system will reduce to the stability of a matrix of the form M=XA for community matrix A.  Whereas the effect of having a distribution of equilibrium abundances is very clearly of this form.

In summary I would recommend making the connection with species equilibrium abundances clearer, and in parallel including the derivation of gamma_i by considering (presumably extremely fast) trait evolution.

Response to Reviewer 1 specific comment 2
--------------------------------------------------------------------------------

There is another interpretation of $\gamma$ that was explicitly defined in the original manuscript. This interpretation is the rate at which the components of a complex system (note, not just biological species or the values of their traits) respond to system perturbation. I have now revised my manuscript to explain this interpretation more thoroughly, and especially more explicitly with respect to the mathematics; $\gamma$ modifies the matrix A to account for different rates of response inherent to different system components.

Because $\gamma$ is not a vector of equilibrium species abundances, I do not make this connection as requested by Reviewer 1; instead, I have revised my manuscript to make the interpretation of $\gamma$ clearer, and further emphasised that the systems of interest in my manuscript are not restricted to species densities or biological traits.

Reviewer 2
================================================================================

Reviewer 2 General comments
--------------------------------------------------------------------------------

This work it is proposes that accounting for rate variation of nodes in responding to perturbation increases (it is incorrectly stated "drives") the stability in large complex systems.

The paper is well written and presented, and the results are reproducible. However, as I will try to explain, I do not find it meets the criteria for publication in PLOS COMPUTATIONAL BIOLOGY (i.e., Originality, Innovation, High importance to researchers in the field, Significant biological and/or methodological insight, Substantial evidence for its conclusions)

Response to Reviewer 2 General comments
--------------------------------------------------------------------------------

I have changed the title from, "Component response rate variation drives stability in large complex systems" to "Component response rate variation underlies stability in large complex systems". I agree with Reviewer 2 that the previous title was technically incorrect.

I am grateful for the comment on the quality of the writing and reproducibility of my results. And while I believe that this manuscript reports original and important work, this helpful review has helped convinced me that it would ultimately fit better in a more multi-disciplinary journal.

Reviewer 2 specific comment 1
--------------------------------------------------------------------------------

1) The Jacobian matrix A (in the notation of the author -  see pag. 8) already incorporates heterogeneity of the response rate, as $s_i=\sum_j A_ij$ is not a priori constant, i.e. different nodes do have different response rates based on their connections. Adding an external heterogeneity therefore, should be justified, having in mind a non-linear model that when linearized, gives a Jacobian that is $M_j=\gamma_i A_ij$.

Response to Reviewer 2 specific comment 1
--------------------------------------------------------------------------------

This is an interesting and helpful comment, as it is of course true that variation in mean row values will arise when constructing random matrices of finite size. And this random variation could quite reasonably be interpreted as variation in component response rates. However, under such conditions, it is important to also emphasise that the *expected* difference between mean row values (and hence component response rates) is still zero; i.e., differences only arise due to error in i.i.d. random sampling for matrix element values. The effect of this error on mean row value differences decreases with increasing matrix size S.

In fact, we can calculate the expected standard deviation of mean row values -- a metric of the natural variation in component response rates -- given a randomly generated matrix, A. This value is just the standard error of the row means. Hence for default values of $\sigma = 0.4$ and $C = 1$ for, e.g., $S = 10$, standard devation of mean row values should be as follows:

$$SD_{mean.row.vals} = \sigma / \sqrt{S} = 0.4 / \sqrt{10} = 0.1265$$

We can do the same for small values of $S = 2$ ($SD_{mean.row.vals} = 0.2828$) and large values of $S = 50$ ($SD_{mean.row.vals} = 0.0566$). The R code below recreates this for any $S$; note that the -1 values on the diagonal will cause slight deviations in practice.

```{r, eval = FALSE}
# Reviewer 2 specific comment 1
SD_mn_row_vals <- function(S = 10, iters = 1000, sigma = 0.4){
    sims  <- rep(x = 0, times = iters);
    for (i in 1:iters){
        A0_dat   <- rnorm(n = S * S, mean = 0, sd = sigma);
        A0       <- matrix(data = A0_dat, nrow = S); 
        diag(A0) <- -1;
        rowmns   <- apply(X = A0, MAR = 1, FUN = mean);
        sims[i]  <- sd(rowmns);
    }
    return(mean(sims));
}
eg_run <- SD_mn_row_vals(S = 50);
```

Importantly, these values are small compared to the equivalent values derived from imposing variation in component response rates a priori, and the inclusion of $Var(\gamma)$ adds to the variation that already is present in $A$. This added variation is ca 0.9523 for my illustrative example in which half $\gamma$ values equal 1.95 and the other half equal 0.05. For simulations in which $\gamma$ is randomly sampled from 0 to 2, the value is ca 0.577.

I have edited the text of the manuscript now to make it clear that such variation in row means of A exist prior to the inclusion of $\gamma$, and that such variation could reasonably be interpreted as variation in component response rate (see Methods). I now introduce $\gamma$ as an a priori increase in the *expected* difference between component response rates (see Introduction) and more carefully interpret the difference between results before and after $\gamma$ is included.

Reviewer 2 specific comment 2
--------------------------------------------------------------------------------

2) The latter case can be realised for a GLV, with $\gamma_i=x^*_i$. This case corresponds to the case when $\gamma_i$ and $A$ are correlated, and it has been recently fully analysed [ref].

Response to Reviewer 2 specific comment 2
--------------------------------------------------------------------------------

Reviewer 2 here points out the approach in which a vector affecting the community matrix is interpreted as equilibrium species abundances ($x^*_{i}$) in a generalised Lotka-Volterra ecological model, as, e.g., in Gibbs et al (2018). In my Response to Reviewer 3 specific comments 1 and 2, I explain how my manuscript differs from analyses of species equilibrium abundances.

Reviewer 2 specific comment 3
--------------------------------------------------------------------------------

3) There is no analytical analysis of the results, and yet it is known that stability of the linearized system dynamics only depends on the mean, the variance and the correlation of the matrix M (C, or S or sigma are not the right control parameters) [see Grilli et al, Nature Comm 2017]. As the variance of M increases with the heterogeneity in $\gamma$, it means that when $\gamma$ is different from zero, the induced correlations $M_ij$ $\gamma_i$ $M_ki$ $\gamma_i$ have sometimes a stabilize effect. Nevertheless, as it is now, this result is only shown with numerical simulations and there is not an analytical understanding of the result, preventing precise quantitative claims on the effect of the heterogeneity of response rates to the system stability. No mention to diagonal stability theory.

Response to Reviewer 2 specific comment 3
--------------------------------------------------------------------------------

Reviewer 2's statement that the stability of the linearised system dynamics depends only on the mean, variance, and correlation of the matrix M is correct specifically for the large S limit (Grilli et al. 2017; see also Gibbs et al. 2018). In other words, stability only depends on these three values as $S \to \infty$. Reviewer 3 also raised this issue in their specific comments 3 and 4. In my revised manuscript, I now clarify that my focus is specifically on the upper range of S (not $S \to \infty$) at which large random systems can reasonably be expected to be stable (see Introduction and Discussion). For such values of S, as Reviewer 2 acknowledges (see Reviewer 2 specific comment 4 below), systems are indeed more likely to be stable where there is heterogeneity in $\gamma$.

Given that my focus is specifically on finite values of $S$, I believe that the numerical simulation approach that I have chosen is appropriate.

Reviewer 2 specific comment 4
--------------------------------------------------------------------------------

4) The numerical results do no lead any substantial evidence for the conclusion of the manuscript ("rate variation in nodes drive the stability of large complex systems"). In fact, by replicating the analysis (see attached Figures) and by the tables provided by the authors, it is evident that the increase of stability is at best around 2-3% on the overall analysed matrix. Nevertheless, the unstable matrices are typically more unstable for the $\gamma \neq 1$ case. Too me these results shows that the variation of the stability between $\gamma_i=1$ and heterogeneous gamma_i is weak.

Response to Reviewer 2 specific comment 4
--------------------------------------------------------------------------------

Reviewer 2 states that the numerical results both do not produce substantial evidence for the manuscript's conclusions (sentence 1), but also acknowledges an increase in stability ("at best around 2-3%"; sentence 2 -- the attached figures replicating my analysis were much appreciated). Technically this is a contradiction, so I interpret the reviewer to mean that, while stability does in fact increase, the actual magnitude of this increase is not large enough to be interesting. While I agree that the effect here is modest, I would argue that it is an important effect to recognise when considering whether or not complex systems are predicted to be stable, and one that has very broad implications that might affect any complex system (i.e., not just ecological communities).

I disagree that the unstable matrices are typically more unstable when $\gamma \neq 1$. I presume Reviewer 2 is interpreting "more unstable" to mean that the real part of the leading eigenvalue is higher given $\gamma \neq 1$ for unstable systems (i.e., systems where the real part of the leading eigenvalue is positive). While some such values might become higher in the $\gamma \neq 1$ condition, there is no evidence that I can find to suggest that this is typical (i.e., occurs more often than not). Such a statement would seem to contradict both the work of Gibbs (2018) and my own results (see, e.g., the change in distribution of eigenvalues in Figure 2). Simulation also supports the conclusion that real parts of eigenvalues in unstable matrices do not tend to increase.

```{r, eval = FALSE}
# Reviewer 2 specific comment 4
S       <- 30;
A0_mx   <- NULL;
A1_mx   <- NULL;
iter    <- 10000;
while(iter > 0){
    r_vec    <- rnorm(n =S, mean = 0, sd = 0.4);
    A0_dat   <- rnorm(n =S *S, mean = 0, sd = sigma);
    A0       <- matrix(data = A0_dat, nrow =S, 
                       ncol =S);
    C_dat    <- rbinom(n =S *S, size = 1, prob = C);
    C_mat    <- matrix(data = C_dat, nrow =S, 
                       ncol =S);
    A0       <- A0 * C_mat;
    diag(A0) <- -1;
    gam1     <- runif(n =S, min = 0, max = 2);
    A1       <- A0 * gam1;
    A0       <- A0 * mean(gam1);
    A0_stb   <- max(Re(eigen(A0)$values));
    A1_stb   <- max(Re(eigen(A1)$values));
    iter     <- iter - 1;
    A0_mx[iter] <- A0_stb;
    A1_mx[iter] <- A1_stb;
}
unstable_A0 <- A0_mx[A0_mx > 0];
unstable_A1 <- A1_mx[A0_mx > 0];
```

Hence, the increase in stability I observe given $Var(\gamma)$ is consistent, and $Var(\gamma)$ does not appear to make unstable systems more unstable. Nevertheless, I do acknowledge that the evident increase in stability is modest, and I now emphasise this better in my revised manuscript.

Reviewer 3
================================================================================

Reviewer 3 General comments
--------------------------------------------------------------------------------

Reviewer #3: In this contribution, the author studies the variability of component response rates (different time scales) in a general model of population dynamics. He finds that the (asymptotic) stability of the system is largely increased if variability in rates is introduced, compared to the stability of a system where the vector of rates is constant. Although the paper is well written, I have serious concerns about its merits for publication in PLoS Computational Biology.

Response to Reviewer 3 General comments
--------------------------------------------------------------------------------

I am grateful for Reviewer 3's comments. In my revised manuscript, I further emphasise that my model is not restricted to population dynamics, as Reviewer 3 incorrectly interpreted, but encompasses any complex system with interacting components. Most of Reviewer 3's additional concerns are based on a comparison with Gibbs et al. (2018; Phys Rev E. 98:022410. doi: 10.1103/PhysRevE.98.022410. pre-print: https://arxiv.org/pdf/1708.08837.pdf). I have revised my manuscript to further emphasise the difference between my work and Gibbs et al. (2018), and I address more specific concerns below.

Reviewer 3 specific comment 1
--------------------------------------------------------------------------------

First, although the motivation of the author was (in principle) different from that of Gibbs et al., PRE (2018), at the end there is substantial overlap with the latter paper. Gibbs et al. conducted a thorough study of the effect of random population abundances on the stability of community matrices for generalized Lotka-Volterra systems, which reduces to analyzing the stability of the matrix M=XA, where X is a diagonal matrix with random (positive) abundances and A is a (random) interaction matrix. This turns out to be the same problem discussed in the present manuscript, so novelty is substantially reduced (in part because the analysis conducted by Gibbs et al. is deeper and much more extensive).

Response to Reviewer 3 specific comment 1
--------------------------------------------------------------------------------

Reviewer 3 acknowledges that my motivations were different in principle from Gibbs et al. (2018), and I have revised my manuscript to further emphasise these differences while also acknowledging where overlap occurs. My manuscript does not focus on primarily random population abundances (though considers them in sections on feasibility), but rather the reaction rates of system components to system perturbation. While the mathematics of underlying my approach is based on analyses of the stability of the form M = XA, the diagonal matrices are not interpreted as abundances, and my simulations do not address the same theoretical question as in Gibbs et al. (2018). The novelty of my manuscript is conceptual rather than mathematical, as is also the case for Gibbs et al. (2018). Indeed, Ahmadian et al. (2015; Phys. Rev. E Stat. Nonlin. Soft. Matter Phys. 91:012820. doi:10.1103/PhysRevE. 91.012820) provided a framework for investigating eigenvalue densities of random matrices of the even more general form $M = B + XAY$, which would include Gibbs et al.'s (2018) M = XA. What was unique about Gibbs et al. (2018) was its application of the properties of random matrices of this form to a long-standing question about the relationship between community stability and community feasibility. They showed that given an interaction matrix ($A$) is stable, the vector of population abundances ($X$) becomes decreasingly likely to destabilise the resulting community matrix ($M$) as the community size increases.

My work applies the same mathematical framework to a different theoretical question, and I have revised my manuscript to make this distinction clear.

Reviewer 3 specific comment 2
--------------------------------------------------------------------------------

Second, the main point of the present manuscript is that stability is largely increased by introducing variability in response rates. Although the author cites the work by Gibbs et al., is surprising that he didn't acknowledge that his results are contradictory with those provided by Gibbs et al. The key point of the latter paper is that population abundances have no effect on stability, as May assumed in his seminal paper. In fact, Gibbs et al. state that 'the community matrix is stable if an only if the interaction matrix is stable. In other words, the abundance of species seems to not affect the sign of eigenvalues'. This statement has to be interpreted in the limit of large systems (S going to infinity) and in probabilistic terms, i.e., it is true almost surely. In fact, Gibbs et al. show that the probability of matrix M=XA being unstable given that A is stable decays exponentially with S (see Fig. 5 of Gibbs el al.) Indeed, they showed this analytically for uniformly distributed abundances X and constant self-regulation terms, which is precisely the case studied in the present manuscript. Conversely, the probability of M becoming stable given that A is unstable also decays exponentially with S (Fig. 9 of Gibbs et al.). This means that multiplying a random matrix by a diagonal random matrix has no effect on stability in the limit of large S.

Response to Reviewer 3 specific comment 2
--------------------------------------------------------------------------------

Reviewer 3 has slightly mis-stated the main point of my manuscript, in part due to some imprecision in the wording of my previous text that Reviewer 1 noticed (particularly regarding the title, which has now been revised). The main point of the manuscript is not that "stability is largely increased by introducing variability in response rates", but that *systems in which there is variability in component response rates are more likely to be stable*. I made this point in the second paragraph of my Discussion:

"It is important to emphasise that variation in component response rate is not stabilising per se; that is, adding variation in component response rate to a particular system does not necessarily increase the probability that the system will be stable. Rather, systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability."

I now also emphasise this point in the abstract and clarify the finite range of S in which I am interested. Three points are particularly important:

(1) Randomly generated matrices ($M = XA$) that are stable are more likely than not to have variation in X across the range of system sizes (S) that I simulated. This is (to my knowledge) a novel result and fully reprodible using the code underlying my simulations (https://github.com/bradduthie/RandomMatrixStability).

(2) At high S, the probability that a system is stable becomes negligible; no such systems were found given S > 32. I now emphasise that my results apply specifically for the upper range of S > 10 for which stability is also non-negligible (i.e., observed at least once in one million simulations). In this range, systems in which there is variation in X are indeed more stable than in systems in which no such variation exists. Further, this general pattern occurs for all simulated systems at the upper range of S, regardless of system connectance (C), interaction stengths ($\sigma$), or X and A element distributions (see Supplementary Information). In this manuscript, I am not interested in the effect of varying component response rates in systems that are too complex to be stable; my focus is entirely restricted to effects in the range where some stability can be reasonably expected. I explain this more carefully now in my Introduction and revise to state that my general conclusions may not apply as $S \to \infty$.

(3) Related to the above, I have, deliberately, not scaled interaction strengths ($\sigma$) by system size ($S$). My reason for doing this is because my manuscript is focused on the effect of component response rates as system complexity ($\sigma \times \sqrt{SC}$) increases, not system size per se. Hence, S is increased while keeping all other parameter values constant so that I can evaluate the effect of $Var(\gamma)$ on the stability of increasingly complex systems. This is fundamentally different from  Gibbs et al. (2018), who scaled $\sigma$ by $S$ such that $\sigma = 1/ \sqrt{S}$. I now state this explicitly in a new paragraph in the Discussion. I also show that my results are consistent with Gibbs et al. (2018) given the assumption that $\sigma = 1/ \sqrt{S}$ in Supplementary Information.

I hope these points adequately address Reviewer 3's concerns. There is no contradiction between my results and those of Gibbs et al. (2018), and I have revised my manuscript to more carefully explain why the two works differ.

Reviewer 3 specific comment 3
--------------------------------------------------------------------------------

This result is seemingly contradictory with the simulations reported in the manuscript under review (Figs. 3 and 4). This is because the author didn't scale the variance of $A$ ($\sigma^2$) with $1/S$, indeed he maintained constant variance while increasing S. Increasing S without scaling variances is misleading because eigenvalue distributions are not comparable for different values of S. My feeling is that if he repeats the figures 3 and 4 using this scaling for sigma^2, then the artifactual difference will disappear in the limit of large S, as Gibbs et al found.

Response to Reviewer 3 specific comment 3
--------------------------------------------------------------------------------

Reviewer 3 is of course correct that scaling the variance of A (sigma^2) with 1/S makes the difference between $\gamma_{i} = 1$ and $Var(\gamma_{i})$ disappear. This was confirmed by simulation during model development, but was not discussed in great detail in the manuscript because it was not a focus of mine (this further highlights the difference between my work and that of Gibbs et al. 2018). I now explain this difference more, and I confirm Reviewer 3's feeling in the Supplementary Information. I also explain the rational for note scaling variances more clearly in the manuscript.

The reason for increasing values of S is to increase the total complexity of the system  (defined as $\sigma \times \sqrt{SC}$), and to isolate and ultimately understand the effect of $Var(\gamma_{i})$ when S approaches a size at which a random system becomes too complex to have a realistic probability of being stable (less than 1 in 1 million). This is the reason for showing increasing system size (S) on the x-axis of Figures 3 and 4. If I were to have scaled the variance of off-diagonal elements of A ($\sigma^{2}$) with $1/S$ as Reviewer 3 suggested (and as done in Gibbs et al. 2018), then the total complexity of the system would not change, defeating the purpose of increasing S. To my knowledge, there is also no a priori reason to expect systems to have variances of interaction strengths ($\sigma$) that scale to their size S, so forcing this to be the case was not justified.

Finally, while eigenvalue distributions are not *identical* for different values of S without scaling variances, it is not really accurate to say that eigenvalue distributions are "not comparable"; eigenvalue distributions change with S in a predictable way that can be compared when S is changed. It is well-established that the eigenvalues for a random matrix M such as that used in my manuscript, and originally by May (1972; Nature 238:413-414), will be uniformly distributed within a circle on the complex plane within a radius of $\sigma\sqrt{SC} < d$, where $-d$ is the mean of diagonal elements of M (Tao and Tu 2010; Annals of Probability 38:2023-2065). Hence, increasing S specifically increases the radius of this uniform distribution, and does so predictably such that the probability that real eigenvalues are negative (the criteria for system stability) becomes increasingly low with increasing S. Multiplying the random matrix A (which has the uniform distribution with a circle of radius $\sigma\sqrt{SC}$) by the vector $\gamma_{i}$ in my manuscript resulted in an eigenvalue distribution that was more likely to be stable at the upper range of S. I now clarify the difference between my result and that of Gibbs et al. (2018) in this context in my revised Discussion, and in a new section of the Supplementary Information.

Reviewer 3 specific comment 4
--------------------------------------------------------------------------------

The author also states in this manuscript that stability can be largely increased by 'selecting' response rates so that linear stability is forced to increase. This is an obvious technical result but (in the light of the results by Gibbs et al.) it is also obvious that these cases will form a set of null measure in the limit of large systems.

Response to Reviewer 3 specific comment 4
--------------------------------------------------------------------------------

I showed that manipulating component response rates can, sometimes greatly, increase the probability that a random system will be stable. As with Reviewer 3 specific comment 3 (see above), I now clarify that my focus is on the upper range of S at which large random systems can reasonably be expected to be stable, not at the limit of $S \to \infty$ (as in Gibbs et al. 2018). Clarifying revisions have been now made accordingly to the Introduction.

Author References Cited
================================================================================

Gibbs, T., Grilli, J., Rogers, T., & Allesina, S. (2018). Effect of population abundances on the stability of large random ecosystems. Physical Review E, 98(2), 022410.

Gray, R. T., & Robinson, P. A. (2009). Stability of random brain networks with excitatory and inhibitory connections. Neurocomputing, 72(7--9), 1849--1858. https://doi.org/10.1016/j.neucom.2008.06.001

Gray, R. T., & Robinson, P. A. (2008). Stability and synchronization of random brain networks with a distribution of connection strengths. Neurocomputing, 71(7--9), 1373--1387. https://doi.org/10.1016/j.neucom.2007.06.002

Grilli, J., Adorisio, M., Suweis, S., Barabás, G., Banavar, J. R., Allesina, S., & Maritan, A. (2017). Feasibility and coexistence of large ecological communities. Nature Communications, 8. https://doi.org/10.1038/ncomms14389

Hui, C., & Richardson, D. M. (2018). How to invade an ecological network. Trends in Ecology and Ecolution, xx, 1--11. https://doi.org/10.1016/j.tree.2018.11.003

May, R. M. (1972). Will a large complex system be stable? Nature, 238, 413--414.

May, R. M. (1973). Qualitative stability in model ecosystems. Ecology, 54(3), 638–641. https://doi.org/10.2307/1935352

Patel, S., Cortez, M. H., & Schreiber, S. J. (2018). Partitioning the effects of eco-evolutionary feedbacks on community stability. American Naturalist, 191(3), 1--29. https://doi.org/10.1101/104505

Haldane, A. G., & May, R. M. (2011). Systemic risk in banking ecosystems. Nature, 469(7330), 351--355. https://doi.org/10.1038/nature09659

### 2019-03-07 notebook/reviewer_comments_and_responses.tex
Update for submission to Scientific Reports

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{setspace}
\setstretch{1}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={Reviewer comments and repsonses},
            pdfauthor={Brad Duthie},
            colorlinks=true,
            linkcolor=blue,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Reviewer comments and repsonses}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Brad Duthie}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{7 March 2019}

\usepackage{amsmath}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[utf8]{inputenc}
\linenumbers
\bibliographystyle{amnatnat}

\begin{document}
\maketitle

\textbf{The following notes acknowledge and respond to reviewer comments
from a submission to \emph{PLoS Computational Biology}}.

\section{Reviewer 1}\label{reviewer-1}

\subsection{Reviewer 1 General
comments}\label{reviewer-1-general-comments}

This paper takes on a timely question: given a random matrix A and
diagonal non-negative matrix X, when will M= XA be stable. It's
important b/c this more accurately reflects the dynamics near
equilibrium in a system where species abundances are variable (and then
are equal to the diagonal elements of X). The current paper also claims
a different interpretation of the stability of a matrix of this form, in
connection with variable response rates across species.

Even though this question probably did lie dormant for many years, it
has been tackled more recently, and some of the work cited here seems to
tackle very similar questions (e.g.~Gibbs et al). But those
connections/differences aren't really explained here. In short, I'd like
to understand if what the author has identified goes beyond what is
known from other work about the effect of a diagonal matrix X on the
stability of M.

I found the presentation and framing of these results overall very
confusing. Here are some comments:

\subsection{Response to Reviewer 1 General
comments}\label{response-to-reviewer-1-general-comments}

I am grateful for the helpful comments of Reviewer 1. It is important to
first emphasise that my focal questions on the stability of complex
systems are not restricted to ecological communities (in which all rows
and columns of matrices A and M are species densities), but instead
address more general questions regarding how the stability of complex
systems is affected by varying response rate of system components (with
the exception of the specific question that I address concerning
ecological community feasibility, of course). System components might be
species densities, but they could additionally include a mix of evolving
species traits (e.g., Patel et al. 2018) or human social decisions
(e.g., management, harvesting, rewilding, land use, etc.; Hui and
Richardson 2018). Components might also not include any species, as in
the case of purely social systems (e.g., banks; Haldane \& May 2011), or
physiological or physical systems (e.g., brain networks; Gray \&
Robinson 2008; 2009). Reviewer 1 is correct that my paper offers a
different interpretation of system stability given the diagonal
non-negative matrix \(X\), one in which elements \(X_{i}\)
(\(\gamma_{i}\) in my manuscript) are interpreted as the rates at which
components respond to perturbation.

The question that I address is distinct from that of Gibbs et al (2018;
see also Reviewer 3 comments and responses). I have now revised the
manuscript to explain these differences and connections more clearly,
both mathematically and conceptually. Mathematically distinct from Gibbs
et al. (2018), I am not focused on the effect of a diagonal non-negative
matrix \(X\) on a stable matrix A as \(S \to \infty\) (i.e., as the
number of system components becomes arbitrarily large given a stable
system), and I specifically focus on a range of system complexities by
not forcing the variance of interaction strengths to scale with
increasing system size (\(\sigma = 1/\sqrt{S}\)). Conceptually distinct,
I am not focused specifically on the effect of species abundances on
ecosystem stability in a generalised Lotka-Volterra system as species
number increases. Rather, I am interested in how the stability of M is
affected by X and A at increasingly high levels of system complexity, in
which the probability of a random complex system being stable becomes
rare but non-negligible (under conditions originally identified by May
1972), and for any type of complex system. Clarifying revisions have
been now made accordingly to the Introduction and Discussion (most
notably, a new paragraph has been added to the Discussion).

\subsection{Reviewer 1 specific comment
1}\label{reviewer-1-specific-comment-1}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\tightlist
\item
  The most obvious interpretation (to me) for gamma\_i is as the set of
  equilibrium abundances for the set of populations, N\_i. But this
  intepretation is not explained clearly, and indeed would presumably
  not lead one to restrict the support of gamma to (0,2), or (given what
  we know about realistic distributions of abundances) to consider cases
  with half the species abundant, and half very rare.
\end{enumerate}

\subsection{Response to Reviewer 1 specific comment
1}\label{response-to-reviewer-1-specific-comment-1}

The use of random matrix theory to investigate the stability of complex
systems has perhaps most often been applied to ecological communities,
in which the complex system is restricted entirely to interacting
species where interactions are described by a square matrix A of
dimensions \(S \times S\). This matrix A can then be multiplied by a
vector describing the density of each of S species (\(N\)) to find the
change in species density (\(N'\)) over some time \(t\): \(N' = AV\)
(May 1972; 1973). It is therefore understandable why the most obvious
interpretation of a vector \(\gamma\) of length S might be equilibrium
species abundances, particularly for those with a background in
theoretical ecology; nevertheless, this is an incorrect interpretation
of \(\gamma\).

In my original manuscript, I was careful to define the vector \(\gamma\)
as the rate at which the components of a complex system (potentially,
but not necessarily, biological species) respond to system perturbation.
For the specific question of ecological community feasibility, I also
explicitly defined equilibrium abundances as the vector n. I now clarify
the use and interpretation of \(\gamma\) earlier in the manuscript, and
explicitly state that \(\gamma\) is not interpreted as a vector of
species abundances.

\subsection{Reviewer 1 specific comment
2}\label{reviewer-1-specific-comment-2}

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  If there is another interpretation of gamma\_i in terms of evolution
  of traits, then this is not adequately explained. Patel et al is cited
  in support of this, where stability of an eco-evolutionary system is
  explored using a block matrix of S species and a set of traits that
  can evolve. But it isn't clear to me in what limit the stability of
  such a system will reduce to the stability of a matrix of the form
  M=XA for community matrix A. Whereas the effect of having a
  distribution of equilibrium abundances is very clearly of this form.
\end{enumerate}

In summary I would recommend making the connection with species
equilibrium abundances clearer, and in parallel including the derivation
of gamma\_i by considering (presumably extremely fast) trait evolution.

\subsection{Response to Reviewer 1 specific comment
2}\label{response-to-reviewer-1-specific-comment-2}

There is another interpretation of \(\gamma\) that was explicitly
defined in the original manuscript. This interpretation is the rate at
which the components of a complex system (note, not just biological
species or the values of their traits) respond to system perturbation. I
have now revised my manuscript to explain this interpretation more
thoroughly, and especially more explicitly with respect to the
mathematics; \(\gamma\) modifies the matrix A to account for different
rates of response inherent to different system components.

Because \(\gamma\) is not a vector of equilibrium species abundances, I
do not make this connection as requested by Reviewer 1; instead, I have
revised my manuscript to make the interpretation of \(\gamma\) clearer,
and further emphasised that the systems of interest in my manuscript are
not restricted to species densities or biological traits.

\section{Reviewer 2}\label{reviewer-2}

\subsection{Reviewer 2 General
comments}\label{reviewer-2-general-comments}

This work it is proposes that accounting for rate variation of nodes in
responding to perturbation increases (it is incorrectly stated
``drives'') the stability in large complex systems.

The paper is well written and presented, and the results are
reproducible. However, as I will try to explain, I do not find it meets
the criteria for publication in PLOS COMPUTATIONAL BIOLOGY (i.e.,
Originality, Innovation, High importance to researchers in the field,
Significant biological and/or methodological insight, Substantial
evidence for its conclusions)

\subsection{Response to Reviewer 2 General
comments}\label{response-to-reviewer-2-general-comments}

I have changed the title from, ``Component response rate variation
drives stability in large complex systems'' to ``Component response rate
variation underlies stability in large complex systems''. I agree with
Reviewer 2 that the previous title was technically incorrect.

I am grateful for the comment on the quality of the writing and
reproducibility of my results. And while I believe that this manuscript
reports original and important work, this helpful review has helped
convinced me that it would ultimately fit better in a more
multi-disciplinary journal.

\subsection{Reviewer 2 specific comment
1}\label{reviewer-2-specific-comment-1}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  The Jacobian matrix A (in the notation of the author - see pag. 8)
  already incorporates heterogeneity of the response rate, as
  \(s_i=\sum_j A_ij\) is not a priori constant, i.e.~different nodes do
  have different response rates based on their connections. Adding an
  external heterogeneity therefore, should be justified, having in mind
  a non-linear model that when linearized, gives a Jacobian that is
  \(M_j=\gamma_i A_ij\).
\end{enumerate}

\subsection{Response to Reviewer 2 specific comment
1}\label{response-to-reviewer-2-specific-comment-1}

This is an interesting and helpful comment, as it is of course true that
variation in mean row values will arise when constructing random
matrices of finite size. And this random variation could quite
reasonably be interpreted as variation in component response rates.
However, under such conditions, it is important to also emphasise that
the \emph{expected} difference between mean row values (and hence
component response rates) is still zero; i.e., differences only arise
due to error in i.i.d. random sampling for matrix element values. The
effect of this error on mean row value differences decreases with
increasing matrix size S.

In fact, we can calculate the expected standard deviation of mean row
values -- a metric of the natural variation in component response rates
-- given a randomly generated matrix, A. This value is just the standard
error of the row means. Hence for default values of \(\sigma = 0.4\) and
\(C = 1\) for, e.g., \(S = 10\), standard devation of mean row values
should be as follows:

\[SD_{mean.row.vals} = \sigma / \sqrt{S} = 0.4 / \sqrt{10} = 0.1265\]

We can do the same for small values of \(S = 2\)
(\(SD_{mean.row.vals} = 0.2828\)) and large values of \(S = 50\)
(\(SD_{mean.row.vals} = 0.0566\)). The R code below recreates this for
any \(S\); note that the -1 values on the diagonal will cause slight
deviations in practice.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Reviewer 2 specific comment 1}
\NormalTok{SD_mn_row_vals <-}\StringTok{ }\NormalTok{function(}\DataTypeTok{S =} \DecValTok{10}\NormalTok{, }\DataTypeTok{iters =} \DecValTok{1000}\NormalTok{, }\DataTypeTok{sigma =} \FloatTok{0.4}\NormalTok{)\{}
    \NormalTok{sims  <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DataTypeTok{x =} \DecValTok{0}\NormalTok{, }\DataTypeTok{times =} \NormalTok{iters);}
    \NormalTok{for (i in }\DecValTok{1}\NormalTok{:iters)\{}
        \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =} \NormalTok{S *}\StringTok{ }\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{sigma);}
        \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =} \NormalTok{S); }
        \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
        \NormalTok{rowmns   <-}\StringTok{ }\KeywordTok{apply}\NormalTok{(}\DataTypeTok{X =} \NormalTok{A0, }\DataTypeTok{MAR =} \DecValTok{1}\NormalTok{, }\DataTypeTok{FUN =} \NormalTok{mean);}
        \NormalTok{sims[i]  <-}\StringTok{ }\KeywordTok{sd}\NormalTok{(rowmns);}
    \NormalTok{\}}
    \KeywordTok{return}\NormalTok{(}\KeywordTok{mean}\NormalTok{(sims));}
\NormalTok{\}}
\NormalTok{eg_run <-}\StringTok{ }\KeywordTok{SD_mn_row_vals}\NormalTok{(}\DataTypeTok{S =} \DecValTok{50}\NormalTok{);}
\end{Highlighting}
\end{Shaded}

Importantly, these values are small compared to the equivalent values
derived from imposing variation in component response rates a priori,
and the inclusion of \(Var(\gamma)\) adds to the variation that already
is present in \(A\). This added variation is ca 0.9523 for my
illustrative example in which half \(\gamma\) values equal 1.95 and the
other half equal 0.05. For simulations in which \(\gamma\) is randomly
sampled from 0 to 2, the value is ca 0.577.

I have edited the text of the manuscript now to make it clear that such
variation in row means of A exist prior to the inclusion of \(\gamma\),
and that such variation could reasonably be interpreted as variation in
component response rate (see Methods). I now introduce \(\gamma\) as an
a priori increase in the \emph{expected} difference between component
response rates (see Introduction) and more carefully interpret the
difference between results before and after \(\gamma\) is included.

\subsection{Reviewer 2 specific comment
2}\label{reviewer-2-specific-comment-2}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1}
\tightlist
\item
  The latter case can be realised for a GLV, with \(\gamma_i=x^*_i\).
  This case corresponds to the case when \(\gamma_i\) and \(A\) are
  correlated, and it has been recently fully analysed {[}ref{]}.
\end{enumerate}

\subsection{Response to Reviewer 2 specific comment
2}\label{response-to-reviewer-2-specific-comment-2}

Reviewer 2 here points out the approach in which a vector affecting the
community matrix is interpreted as equilibrium species abundances
(\(x^*_{i}\)) in a generalised Lotka-Volterra ecological model, as,
e.g., in Gibbs et al (2018). In my Response to Reviewer 3 specific
comments 1 and 2, I explain how my manuscript differs from analyses of
species equilibrium abundances.

\subsection{Reviewer 2 specific comment
3}\label{reviewer-2-specific-comment-3}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{2}
\tightlist
\item
  There is no analytical analysis of the results, and yet it is known
  that stability of the linearized system dynamics only depends on the
  mean, the variance and the correlation of the matrix M (C, or S or
  sigma are not the right control parameters) {[}see Grilli et al,
  Nature Comm 2017{]}. As the variance of M increases with the
  heterogeneity in \(\gamma\), it means that when \(\gamma\) is
  different from zero, the induced correlations \(M_ij\) \(\gamma_i\)
  \(M_ki\) \(\gamma_i\) have sometimes a stabilize effect. Nevertheless,
  as it is now, this result is only shown with numerical simulations and
  there is not an analytical understanding of the result, preventing
  precise quantitative claims on the effect of the heterogeneity of
  response rates to the system stability. No mention to diagonal
  stability theory.
\end{enumerate}

\subsection{Response to Reviewer 2 specific comment
3}\label{response-to-reviewer-2-specific-comment-3}

Reviewer 2's statement that the stability of the linearised system
dynamics depends only on the mean, variance, and correlation of the
matrix M is correct specifically for the large S limit (Grilli et al.
2017; see also Gibbs et al. 2018). In other words, stability only
depends on these three values as \(S \to \infty\). Reviewer 3 also
raised this issue in their specific comments 3 and 4. In my revised
manuscript, I now clarify that my focus is specifically on the upper
range of S (not \(S \to \infty\)) at which large random systems can
reasonably be expected to be stable (see Introduction and Discussion).
For such values of S, as Reviewer 2 acknowledges (see Reviewer 2
specific comment 4 below), systems are indeed more likely to be stable
where there is heterogeneity in \(\gamma\).

Given that my focus is specifically on finite values of \(S\), I believe
that the numerical simulation approach that I have chosen is
appropriate.

\subsection{Reviewer 2 specific comment
4}\label{reviewer-2-specific-comment-4}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{3}
\tightlist
\item
  The numerical results do no lead any substantial evidence for the
  conclusion of the manuscript (``rate variation in nodes drive the
  stability of large complex systems''). In fact, by replicating the
  analysis (see attached Figures) and by the tables provided by the
  authors, it is evident that the increase of stability is at best
  around 2-3\% on the overall analysed matrix. Nevertheless, the
  unstable matrices are typically more unstable for the
  \(\gamma \neq 1\) case. Too me these results shows that the variation
  of the stability between \(\gamma_i=1\) and heterogeneous gamma\_i is
  weak.
\end{enumerate}

\subsection{Response to Reviewer 2 specific comment
4}\label{response-to-reviewer-2-specific-comment-4}

Reviewer 2 states that the numerical results both do not produce
substantial evidence for the manuscript's conclusions (sentence 1), but
also acknowledges an increase in stability (``at best around 2-3\%'';
sentence 2 -- the attached figures replicating my analysis were much
appreciated). Technically this is a contradiction, so I interpret the
reviewer to mean that, while stability does in fact increase, the actual
magnitude of this increase is not large enough to be interesting. While
I agree that the effect here is modest, I would argue that it is an
important effect to recognise when considering whether or not complex
systems are predicted to be stable, and one that has very broad
implications that might affect any complex system (i.e., not just
ecological communities).

I disagree that the unstable matrices are typically more unstable when
\(\gamma \neq 1\). I presume Reviewer 2 is interpreting ``more
unstable'' to mean that the real part of the leading eigenvalue is
higher given \(\gamma \neq 1\) for unstable systems (i.e., systems where
the real part of the leading eigenvalue is positive). While some such
values might become higher in the \(\gamma \neq 1\) condition, there is
no evidence that I can find to suggest that this is typical (i.e.,
occurs more often than not). Such a statement would seem to contradict
both the work of Gibbs (2018) and my own results (see, e.g., the change
in distribution of eigenvalues in Figure 2). Simulation also supports
the conclusion that real parts of eigenvalues in unstable matrices do
not tend to increase.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Reviewer 2 specific comment 4}
\NormalTok{S       <-}\StringTok{ }\DecValTok{30}\NormalTok{;}
\NormalTok{A0_mx   <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
\NormalTok{A1_mx   <-}\StringTok{ }\OtherTok{NULL}\NormalTok{;}
\NormalTok{iter    <-}\StringTok{ }\DecValTok{10000}\NormalTok{;}
\NormalTok{while(iter >}\StringTok{ }\DecValTok{0}\NormalTok{)\{}
    \NormalTok{r_vec    <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =}\NormalTok{S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \FloatTok{0.4}\NormalTok{);}
    \NormalTok{A0_dat   <-}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DataTypeTok{n =}\NormalTok{S *S, }\DataTypeTok{mean =} \DecValTok{0}\NormalTok{, }\DataTypeTok{sd =} \NormalTok{sigma);}
    \NormalTok{A0       <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{A0_dat, }\DataTypeTok{nrow =}\NormalTok{S, }
                       \DataTypeTok{ncol =}\NormalTok{S);}
    \NormalTok{C_dat    <-}\StringTok{ }\KeywordTok{rbinom}\NormalTok{(}\DataTypeTok{n =}\NormalTok{S *S, }\DataTypeTok{size =} \DecValTok{1}\NormalTok{, }\DataTypeTok{prob =} \NormalTok{C);}
    \NormalTok{C_mat    <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DataTypeTok{data =} \NormalTok{C_dat, }\DataTypeTok{nrow =}\NormalTok{S, }
                       \DataTypeTok{ncol =}\NormalTok{S);}
    \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{C_mat;}
    \KeywordTok{diag}\NormalTok{(A0) <-}\StringTok{ }\NormalTok{-}\DecValTok{1}\NormalTok{;}
    \NormalTok{gam1     <-}\StringTok{ }\KeywordTok{runif}\NormalTok{(}\DataTypeTok{n =}\NormalTok{S, }\DataTypeTok{min =} \DecValTok{0}\NormalTok{, }\DataTypeTok{max =} \DecValTok{2}\NormalTok{);}
    \NormalTok{A1       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\NormalTok{gam1;}
    \NormalTok{A0       <-}\StringTok{ }\NormalTok{A0 *}\StringTok{ }\KeywordTok{mean}\NormalTok{(gam1);}
    \NormalTok{A0_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A0)$values));}
    \NormalTok{A1_stb   <-}\StringTok{ }\KeywordTok{max}\NormalTok{(}\KeywordTok{Re}\NormalTok{(}\KeywordTok{eigen}\NormalTok{(A1)$values));}
    \NormalTok{iter     <-}\StringTok{ }\NormalTok{iter -}\StringTok{ }\DecValTok{1}\NormalTok{;}
    \NormalTok{A0_mx[iter] <-}\StringTok{ }\NormalTok{A0_stb;}
    \NormalTok{A1_mx[iter] <-}\StringTok{ }\NormalTok{A1_stb;}
\NormalTok{\}}
\NormalTok{unstable_A0 <-}\StringTok{ }\NormalTok{A0_mx[A0_mx >}\StringTok{ }\DecValTok{0}\NormalTok{];}
\NormalTok{unstable_A1 <-}\StringTok{ }\NormalTok{A1_mx[A0_mx >}\StringTok{ }\DecValTok{0}\NormalTok{];}
\end{Highlighting}
\end{Shaded}

Hence, the increase in stability I observe given \(Var(\gamma)\) is
consistent, and \(Var(\gamma)\) does not appear to make unstable systems
more unstable. Nevertheless, I do acknowledge that the evident increase
in stability is modest, and I now emphasise this better in my revised
manuscript.

\section{Reviewer 3}\label{reviewer-3}

\subsection{Reviewer 3 General
comments}\label{reviewer-3-general-comments}

Reviewer \#3: In this contribution, the author studies the variability
of component response rates (different time scales) in a general model
of population dynamics. He finds that the (asymptotic) stability of the
system is largely increased if variability in rates is introduced,
compared to the stability of a system where the vector of rates is
constant. Although the paper is well written, I have serious concerns
about its merits for publication in PLoS Computational Biology.

\subsection{Response to Reviewer 3 General
comments}\label{response-to-reviewer-3-general-comments}

I am grateful for Reviewer 3's comments. In my revised manuscript, I
further emphasise that my model is not restricted to population
dynamics, as Reviewer 3 incorrectly interpreted, but encompasses any
complex system with interacting components. Most of Reviewer 3's
additional concerns are based on a comparison with Gibbs et al. (2018;
Phys Rev E. 98:022410. doi: 10.1103/PhysRevE.98.022410. pre-print:
\url{https://arxiv.org/pdf/1708.08837.pdf}). I have revised my
manuscript to further emphasise the difference between my work and Gibbs
et al. (2018), and I address more specific concerns below.

\subsection{Reviewer 3 specific comment
1}\label{reviewer-3-specific-comment-1}

First, although the motivation of the author was (in principle)
different from that of Gibbs et al., PRE (2018), at the end there is
substantial overlap with the latter paper. Gibbs et al. conducted a
thorough study of the effect of random population abundances on the
stability of community matrices for generalized Lotka-Volterra systems,
which reduces to analyzing the stability of the matrix M=XA, where X is
a diagonal matrix with random (positive) abundances and A is a (random)
interaction matrix. This turns out to be the same problem discussed in
the present manuscript, so novelty is substantially reduced (in part
because the analysis conducted by Gibbs et al. is deeper and much more
extensive).

\subsection{Response to Reviewer 3 specific comment
1}\label{response-to-reviewer-3-specific-comment-1}

Reviewer 3 acknowledges that my motivations were different in principle
from Gibbs et al. (2018), and I have revised my manuscript to further
emphasise these differences while also acknowledging where overlap
occurs. My manuscript does not focus on primarily random population
abundances (though considers them in sections on feasibility), but
rather the reaction rates of system components to system perturbation.
While the mathematics of underlying my approach is based on analyses of
the stability of the form M = XA, the diagonal matrices are not
interpreted as abundances, and my simulations do not address the same
theoretical question as in Gibbs et al. (2018). The novelty of my
manuscript is conceptual rather than mathematical, as is also the case
for Gibbs et al. (2018). Indeed, Ahmadian et al. (2015; Phys. Rev.~E
Stat. Nonlin. Soft. Matter Phys. 91:012820. \url{doi:10.1103/PhysRevE}.
91.012820) provided a framework for investigating eigenvalue densities
of random matrices of the even more general form \(M = B + XAY\), which
would include Gibbs et al.'s (2018) M = XA. What was unique about Gibbs
et al. (2018) was its application of the properties of random matrices
of this form to a long-standing question about the relationship between
community stability and community feasibility. They showed that given an
interaction matrix (\(A\)) is stable, the vector of population
abundances (\(X\)) becomes decreasingly likely to destabilise the
resulting community matrix (\(M\)) as the community size increases.

My work applies the same mathematical framework to a different
theoretical question, and I have revised my manuscript to make this
distinction clear.

\subsection{Reviewer 3 specific comment
2}\label{reviewer-3-specific-comment-2}

Second, the main point of the present manuscript is that stability is
largely increased by introducing variability in response rates. Although
the author cites the work by Gibbs et al., is surprising that he didn't
acknowledge that his results are contradictory with those provided by
Gibbs et al. The key point of the latter paper is that population
abundances have no effect on stability, as May assumed in his seminal
paper. In fact, Gibbs et al. state that `the community matrix is stable
if an only if the interaction matrix is stable. In other words, the
abundance of species seems to not affect the sign of eigenvalues'. This
statement has to be interpreted in the limit of large systems (S going
to infinity) and in probabilistic terms, i.e., it is true almost surely.
In fact, Gibbs et al. show that the probability of matrix M=XA being
unstable given that A is stable decays exponentially with S (see Fig. 5
of Gibbs el al.) Indeed, they showed this analytically for uniformly
distributed abundances X and constant self-regulation terms, which is
precisely the case studied in the present manuscript. Conversely, the
probability of M becoming stable given that A is unstable also decays
exponentially with S (Fig. 9 of Gibbs et al.). This means that
multiplying a random matrix by a diagonal random matrix has no effect on
stability in the limit of large S.

\subsection{Response to Reviewer 3 specific comment
2}\label{response-to-reviewer-3-specific-comment-2}

Reviewer 3 has slightly mis-stated the main point of my manuscript, in
part due to some imprecision in the wording of my previous text that
Reviewer 1 noticed (particularly regarding the title, which has now been
revised). The main point of the manuscript is not that ``stability is
largely increased by introducing variability in response rates'', but
that \emph{systems in which there is variability in component response
rates are more likely to be stable}. I made this point in the second
paragraph of my Discussion:

``It is important to emphasise that variation in component response rate
is not stabilising per se; that is, adding variation in component
response rate to a particular system does not necessarily increase the
probability that the system will be stable. Rather, systems that are
observed to be stable are more likely to have varying component response
rates, and for this variation to be critical to their stability.''

I now also emphasise this point in the abstract and clarify the finite
range of S in which I am interested. Three points are particularly
important:

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\item
  Randomly generated matrices (\(M = XA\)) that are stable are more
  likely than not to have variation in X across the range of system
  sizes (S) that I simulated. This is (to my knowledge) a novel result
  and fully reprodible using the code underlying my simulations
  (\url{https://github.com/bradduthie/RandomMatrixStability}).
\item
  At high S, the probability that a system is stable becomes negligible;
  no such systems were found given S \textgreater{} 32. I now emphasise
  that my results apply specifically for the upper range of S
  \textgreater{} 10 for which stability is also non-negligible (i.e.,
  observed at least once in one million simulations). In this range,
  systems in which there is variation in X are indeed more stable than
  in systems in which no such variation exists. Further, this general
  pattern occurs for all simulated systems at the upper range of S,
  regardless of system connectance (C), interaction stengths
  (\(\sigma\)), or X and A element distributions (see Supplementary
  Information). In this manuscript, I am not interested in the effect of
  varying component response rates in systems that are too complex to be
  stable; my focus is entirely restricted to effects in the range where
  some stability can be reasonably expected. I explain this more
  carefully now in my Introduction and revise to state that my general
  conclusions may not apply as \(S \to \infty\).
\item
  Related to the above, I have, deliberately, not scaled interaction
  strengths (\(\sigma\)) by system size (\(S\)). My reason for doing
  this is because my manuscript is focused on the effect of component
  response rates as system complexity (\(\sigma \times \sqrt{SC}\))
  increases, not system size per se. Hence, S is increased while keeping
  all other parameter values constant so that I can evaluate the effect
  of \(Var(\gamma)\) on the stability of increasingly complex systems.
  This is fundamentally different from Gibbs et al. (2018), who scaled
  \(\sigma\) by \(S\) such that \(\sigma = 1/ \sqrt{S}\). I now state
  this explicitly in a new paragraph in the Discussion. I also show that
  my results are consistent with Gibbs et al. (2018) given the
  assumption that \(\sigma = 1/ \sqrt{S}\) in Supplementary Information.
\end{enumerate}

I hope these points adequately address Reviewer 3's concerns. There is
no contradiction between my results and those of Gibbs et al. (2018),
and I have revised my manuscript to more carefully explain why the two
works differ.

\subsection{Reviewer 3 specific comment
3}\label{reviewer-3-specific-comment-3}

This result is seemingly contradictory with the simulations reported in
the manuscript under review (Figs. 3 and 4). This is because the author
didn't scale the variance of \(A\) (\(\sigma^2\)) with \(1/S\), indeed
he maintained constant variance while increasing S. Increasing S without
scaling variances is misleading because eigenvalue distributions are not
comparable for different values of S. My feeling is that if he repeats
the figures 3 and 4 using this scaling for sigma\^{}2, then the
artifactual difference will disappear in the limit of large S, as Gibbs
et al found.

\subsection{Response to Reviewer 3 specific comment
3}\label{response-to-reviewer-3-specific-comment-3}

Reviewer 3 is of course correct that scaling the variance of A
(sigma\^{}2) with 1/S makes the difference between \(\gamma_{i} = 1\)
and \(Var(\gamma_{i})\) disappear. This was confirmed by simulation
during model development, but was not discussed in great detail in the
manuscript because it was not a focus of mine (this further highlights
the difference between my work and that of Gibbs et al. 2018). I now
explain this difference more, and I confirm Reviewer 3's feeling in the
Supplementary Information. I also explain the rational for note scaling
variances more clearly in the manuscript.

The reason for increasing values of S is to increase the total
complexity of the system (defined as \(\sigma \times \sqrt{SC}\)), and
to isolate and ultimately understand the effect of \(Var(\gamma_{i})\)
when S approaches a size at which a random system becomes too complex to
have a realistic probability of being stable (less than 1 in 1 million).
This is the reason for showing increasing system size (S) on the x-axis
of Figures 3 and 4. If I were to have scaled the variance of
off-diagonal elements of A (\(\sigma^{2}\)) with \(1/S\) as Reviewer 3
suggested (and as done in Gibbs et al. 2018), then the total complexity
of the system would not change, defeating the purpose of increasing S.
To my knowledge, there is also no a priori reason to expect systems to
have variances of interaction strengths (\(\sigma\)) that scale to their
size S, so forcing this to be the case was not justified.

Finally, while eigenvalue distributions are not \emph{identical} for
different values of S without scaling variances, it is not really
accurate to say that eigenvalue distributions are ``not comparable'';
eigenvalue distributions change with S in a predictable way that can be
compared when S is changed. It is well-established that the eigenvalues
for a random matrix M such as that used in my manuscript, and originally
by May (1972; Nature 238:413-414), will be uniformly distributed within
a circle on the complex plane within a radius of
\(\sigma\sqrt{SC} < d\), where \(-d\) is the mean of diagonal elements
of M (Tao and Tu 2010; Annals of Probability 38:2023-2065). Hence,
increasing S specifically increases the radius of this uniform
distribution, and does so predictably such that the probability that
real eigenvalues are negative (the criteria for system stability)
becomes increasingly low with increasing S. Multiplying the random
matrix A (which has the uniform distribution with a circle of radius
\(\sigma\sqrt{SC}\)) by the vector \(\gamma_{i}\) in my manuscript
resulted in an eigenvalue distribution that was more likely to be stable
at the upper range of S. I now clarify the difference between my result
and that of Gibbs et al. (2018) in this context in my revised
Discussion, and in a new section of the Supplementary Information.

\subsection{Reviewer 3 specific comment
4}\label{reviewer-3-specific-comment-4}

The author also states in this manuscript that stability can be largely
increased by `selecting' response rates so that linear stability is
forced to increase. This is an obvious technical result but (in the
light of the results by Gibbs et al.) it is also obvious that these
cases will form a set of null measure in the limit of large systems.

\subsection{Response to Reviewer 3 specific comment
4}\label{response-to-reviewer-3-specific-comment-4}

I showed that manipulating component response rates can, sometimes
greatly, increase the probability that a random system will be stable.
As with Reviewer 3 specific comment 3 (see above), I now clarify that my
focus is on the upper range of S at which large random systems can
reasonably be expected to be stable, not at the limit of
\(S \to \infty\) (as in Gibbs et al. 2018). Clarifying revisions have
been now made accordingly to the Introduction.

\section{Author References Cited}\label{author-references-cited}

Gibbs, T., Grilli, J., Rogers, T., \& Allesina, S. (2018). Effect of
population abundances on the stability of large random ecosystems.
Physical Review E, 98(2), 022410.

Gray, R. T., \& Robinson, P. A. (2009). Stability of random brain
networks with excitatory and inhibitory connections. Neurocomputing,
72(7--9), 1849--1858. \url{https://doi.org/10.1016/j.neucom.2008.06.001}

Gray, R. T., \& Robinson, P. A. (2008). Stability and synchronization of
random brain networks with a distribution of connection strengths.
Neurocomputing, 71(7--9), 1373--1387.
\url{https://doi.org/10.1016/j.neucom.2007.06.002}

Grilli, J., Adorisio, M., Suweis, S., Barabás, G., Banavar, J. R.,
Allesina, S., \& Maritan, A. (2017). Feasibility and coexistence of
large ecological communities. Nature Communications, 8.
\url{https://doi.org/10.1038/ncomms14389}

Hui, C., \& Richardson, D. M. (2018). How to invade an ecological
network. Trends in Ecology and Ecolution, xx, 1--11.
\url{https://doi.org/10.1016/j.tree.2018.11.003}

May, R. M. (1972). Will a large complex system be stable? Nature, 238,
413--414.

May, R. M. (1973). Qualitative stability in model ecosystems. Ecology,
54(3), 638--641. \url{https://doi.org/10.2307/1935352}

Patel, S., Cortez, M. H., \& Schreiber, S. J. (2018). Partitioning the
effects of eco-evolutionary feedbacks on community stability. American
Naturalist, 191(3), 1--29. \url{https://doi.org/10.1101/104505}

Haldane, A. G., \& May, R. M. (2011). Systemic risk in banking
ecosystems. Nature, 469(7330), 351--355.
\url{https://doi.org/10.1038/nature09659}

\end{document}

### 2019-03-07 notebook/SI.tex
Update for submission to Scientific Reports

@@
 \setlength{\droptitle}{-2em}
-  \title{Component response rate variation drives stability in large complex
+
+  \title{Component response rate variation underlies the stability of complex
 systems}
@@
 The question that I address in the main text is distinct from that of
-Gibbs et al\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, who
 focused instead on the effect of a diagonal matrix of biological species

### 2019-03-07 notebook/ms.tex
Update for submission to Scientific Reports

@@
 potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
-large but finite systems, where the probability of stability would
-otherwise be negligible. At such extremes of simulated system
-complexity, the largest stable complex systems would be unstable if not
-for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
-previously unconsidered aspect of system stability that is likely to be
-pervasive across all realistic complex systems.
+\(\boldsymbol{\gamma}\) is especially important for stability in highly
+complex systems, in which the probability of stability would otherwise
+be negligible. At such extremes of simulated system complexity, the
+largest stable complex systems would be unstable if not for
+\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
+unconsidered aspect of system stability that is likely to be pervasive
+across all realistic complex systems.
 
@@
 increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
-realistic complex systems (such that \(S\) is high but finite) in which
-\(\gamma\) varies are actually more likely to be stable.
+realistic complex systems (such that \(S\) is high but finite) are
+actually more likely to be stable if their individual component response
+rates vary
 
@@
 decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)). My numerical results do not contradict this
+(\(S \to \infty\)) for a given magnitude of complexity
+\(\sigma\sqrt{SC}\). My numerical results do not contradict this
 prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
 instead fixed \(\sigma\) and increased \(S\) to thereby increase total
-system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
-results simulated across \(\sigma\) and \(C\)). Overall, I show that
-component response rate variation increases the upper bound of
-complexity at which stability can be realistically observed, meaning
-that highly complex systems are more likely than not to vary in their
-component response rates, and for this variation to be critical for
-system stability.
+system complexity (see Supplemental Information for results simulated
+across \(\sigma\) and \(C\)). Overall, I show that component response
+rate variation increases the upper bound of complexity at which
+stability can be realistically observed, meaning that highly complex
+systems are more likely than not to vary in their component response
+rates, and for this variation to be critical for system stability.
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.742 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.74 for a and b,
 respectively.

### 2019-12-30 notebook/sci_rep_reviews.Rmd
Sci Rep reviews

---
title: "Reviewer Comments (Scientific Reports)"
author: "Brad Duthie"
date: "30 December 2019"
output:
  pdf_document: default
  html_document: default
---

Reviewer 1 comments
================================================================================

This paper considers timely questions concerning the stability of complex systems, mainly using random matrix theory. The central problem consists in studying the stability of random matrices which were originally seen as random community matrices from ecology. In a seminal paper, R May challenged the consensus that complexity begets stability in complex ecosystems. These issues are of current concern in ecology. Most of the works dealing with such problems (either for stability or feasibility) consider networks of high species richness S, and develop or use techniques and results from random matrix theory in the large S limit, like e.g. Girko circular law that describes the asymptotic behavior of the related spectra measures (a system is stable when all of its eigenvalues have negative real parts).

The paper of Duthie is interesting since it deals with finite size systems where S is not too large. Mathematically, the author focus on random matrices of the form D A-id where D is a random diagonal matrix and A is a random matrix having i.i.d. centered entries. The author suggests numerically that stability is enhanced by multiplying A by D. This idea seems to be new.

Recent works dealing with similar models in the large S limit have been published recently (Gibbs et al (2018) and Ahmadian (2018). Such works are not dealing with finite S and focus on limiting spectral measures. On the other hand, real empirical ecological networks have finite sizes S, so that the paper might be of practical interest for ecologists dealing with empirical data. In this setting, most interaction weights that define Lotka-Volterra dynamics are unknown at present time, and a natural idea consists in randomizing such weights to arrive at random community matrices of finite size, see, e.g. Ensemble ecosystem modeling for predicting ecosystem response to predator reintroduction, Baker et al. (2016) Conservation Biology, Volume 31, No. 2, 376–384.

In summary I think that the paper is interesting and timely, but I strongly recommend that the author
considers also empirical webs (like for example those given in Dougoud et al. (2018)), the feasibility
of equilibria in large ecosystems: A primary but neglected concept in the complexity-stability debate,
PLoS Comput. Biol, and check numerically if the observation that variations of component responses
rates enhance stability and feasibility is valid in experimentally observed webs. The author should
also test this hypothesis on standard webs models for structured food web models like the niche, or
the nested-hierarchy random web models.

Comment:

The manuscript is well written. I have just a remark concerning Figures 3 and 4. The black line is
intended to show the proportion of systems that are stable when the variance is positive but would
be unstable for zero variance. Is the author considering the conditional probability that A-id is
unstable given that DA-id is stable ? I think that this point should be explained in more details.

Reviewer 2 comments
================================================================================

The manuscript explores the stability of matrices of the form

$A = diag(\gamma)  M$,

where gamma is an S-component vector, diag(gamma) is the diagonal matrix with gamma along its diagonal, \* denotes matrix multiplication, and M is a circular random matrix (independently and randomly sampled offdiagonal entries with mean 0 and variance V=C*sigma^2 where C is the connectance and sigma^2 the variance of nonzero entries; and with all diagonal entries equal to -1). The main claim is that whenever the random matrix M has parameters putting it near the boundary of stability and instability, the likelihood of A being stable is increased by having a larger variance in gamma.

The manuscript is very clear and its results were easily reproducible. My concerns are threefold: first, there is no attempt at an analytical understanding, making the breadth of the results difficult to judge; second, the numerical exploration is restricted to a very narrow class of random matrices, further hampering a good appreciation of the results' breadth; and third, it feels as if the effect of response rate variation was a second- or even third-order effect, and thus only of minor importance in explaining the stability of large complex systems. Below I elaborate on each of these concerns in turn.

The results crucially build on the matrices being dangerously close to the stability boundary. At that point, minor fluctuations in the position of the leading eigenvalue can make or break a system. In fact, there are known results on the uncertainty in the leading eigenvalue's position - but, to my knowledge, those are only available for circular matrices (here: constant gamma). I am wondering whether combining those with the methods of Gibbs et al. (2018) could perhaps be used to gain further understanding. In either case, the manuscript would of course be much stronger with analytical results, but I do realize this may be uncharted territory.

Since numerical explorations are used in lieu of analytical ones, I would look at a broader class of matrices M. For example, the average pairwise correlation rho between the (i,j)th and (j,i) th matrix entries is known to strongly affect the local stability of a random matrix - in the large S limit, the leading eigenvalue converges to

$\lambda = \sqrt{SV}*(+\rho)$.

I did some preliminary exploration of having different rho values, and this did not change the Author's results. So adding this new aspect to the manuscript would strengthen its results. I would also maybe consider matrices with even more structure - e.g., varying diagonal entries in M, or having different means/variances in the top and bottom triangles (corresponding to food webs with cascade structures; Allesina et al. 2015 Nature Communications).

Finally, the general relevance of the results is hampered by being confined to a narrow parameter space, and even there the fraction of occurrences where var(gamma) makes a difference is minuscule. Looking at Figure 3, the only cases in which there is an appreciable effect of a variance in the gammas is when only a small handful of the matrices end up being stable in the first place. As such, var(gamma) only makes a difference when there is not much difference to be made anyway, because the vast majority of matrices end up unstable regardless of the magnitude of var(gamma). In light of this, I would be much more careful with the manuscript's framing - in particular, I would change the title to express better that we are dealing with second-order effects, and would not say in the Abstract that system stability is "markedly increased" by variation in the gammas.

Reviewer 3 commments
================================================================================

This manuscript describe the effect of variation in the response rate of individual systems to the stability of a complex system by A. Bradley Duthie. The manuscript is well-written and the calculations, codes and so on sound technically correct, but I have problem to see the importance of the response rates to understand the stability of the systems.

In my opinion 3 main questions are more important in all these discussions:

1. The diagonal of the matrix, thinking in species interactions, represents the intraspecific interactions in the same species, and in the study of the stability is the main point because it gives us the center of the cloud of eigenvalues. In these study, as many all of them, is taken as -1. All infraspecific interactions in every species are negatives and equal? It is difficult to agree with this idea.

2. In these discussions the author use random matrix to implement complex systems. Again it is difficult to understand how a random matrix can represent a complex system. If we think in species again, different ecological interactions have different distributions, has can be observed in the paper of Tang and I Sundstrand that the random matrices can be a perfect null model, but nothing more.

3. I can understand the importance of the study of the responds rates to understand the stability of the matrices. Maybe I am wrong but sound like the author is comparing apples and oranges.

Remarks to the author:

This manuscript describe the effect of variation in the response rate of individual systems to the stability of a complex system by A. Bradley Duthie. The manuscript is well-written and the calculations, codes and so on sound technically correct, but I have problem to see the importance of the response rates to understand the stability of the systems.

In my opinion 3 main questions are more important in all these discussions:

1. The diagonal of the matrix, thinking in species interactions, represents the intraspecific interactions in the same species, and in the study of the stability is the main point because it gives us the center of the cloud of eigenvalues. In these study, as many all of them, is taken as -1. All infraspecific interactions in every species are negatives and equal? It is difficult to agree with this idea.

2. In these discussions the author use random matrix to implement complex systems. Again it is difficult to understand how a random matrix can represent a complex system. If we think in species again, different ecological interactions have different distributions, has can be observed in the paper of Tang and I Sundstrand that the random matrices can be a perfect null model, but nothing more.

3. I can understand the importance of the study of the responds rates to understand the stability of the matrices. Maybe I am wrong but sound like the author is comparing apples and oranges.

Other minor questions.

I've missed some classic articles from random-matrix literature as:

H.J. Sommers, A. Crisanti, H. Sompolinsky, and Y. Stein, Spectrum of Large Random Asymmetric Variables, Phys. Rev. Lett. 60, 1895 (1988).

S.E. Townsend, D.T. Haydon, and L. Matthews, On the generality of stability-complexity relationships in Lotka-Volterra ecosystems, J. Theor. Biol. 267, 243 (2010).

A. Mougi, and M. Kondoh, Diversity of Interaction Types and Ecolog- ical Community Stability, Science 337, 349 (2012).

K.Z. Coyte, J. Schluter, and K.R. Foster, The ecology of the micro- biome: Networks, competition, and stability, Science 350, 663 (2015).

### 2019-12-30 R/random_mats.R
Remove the analytical stuff and add a cascade model

@@ -1,199 +1,3 @@
-# NOTE FOR RESEARCH 27 OCT 2019
-# SEE EQN 2.4-2.9 OF AHMADIAN ET AL 2015
-# PARTICULARLY 2.8, IN WHICH THE SPECTRAL DENSITY IS DEPENDENT UP ON THE MATRIX
-# L. THE DISTRIBUTION APPEARS TO DEPEND ON L^-1 IN A WAY (TRACE) THAT MIGHT 
-# CAUSE THE STABILITY. NEED TO FIGURE OUT HOW THIS WORKS FOR N < 10 AND N > 10?
-N  <- 4000;
-z  <- 10;
-M  <- matrix(data = rnorm(N * N, sd = 1/N), nrow = N, ncol = N);
-Mz <- z - M;
-ml <- Mz %*% t(Mz);
-ms <- solve(ml);
-tr <- (1/N) * sum(diag(ms));
-tr;
-
-
-Rg <- Re(eigen(M)$values);
-Ig <- Im(eigen(M)$values);
-plot(x = Rg, y = Ig);
-
-# Maybe instead, define L as {N, N, N, N}, so the SD is preserved? Then
-# use the tools above?
-
-# Another check: Show unusually high values can be shunted into the very
-# negative group.
-N <- 800;
-L <- matrix(data = 0, nrow = N, ncol = N);
-J <- matrix(data = rnorm(N*N, sd = sqrt(1/N)), nrow = N, ncol = N);
-diag(J)  <- 1;
-diag(L)  <- -1 * c(rep(x = 0.01, times = N/2), rep(x = 1.98, times = N/2));
-M1       <- L %*% J
-eM1      <- eigen(M1)$values;
-rM1      <- Re(eM1);
-iM1      <- Im(eM1);
-plot(x = rM1, y = iM1, asp = 1, cex = 0.8, pch = "+", xlim = c(-0.015, 0.015),
-     ylim = c(-0.015, 0.015));
-abline(v = 0, col = "red");
-plot(x = rM1, y = iM1, asp = 1, cex = 0.8, pch = "+", 
-     xlim = c(-1.5, -1.5), ylim = c(-1.5, 1.5));
-rm1 <- tail(sort(rM1));
-
-L        <- matrix(data = 0, nrow = N, ncol = N);
-diag(L)  <- -0.995;
-M2       <- L %*% J;
-eM2      <- eigen(M2)$values;
[…]

### 2019-12-30 R/analytical.R
Paste the analytical stuff

@@ -1,0 +1,193 @@
+# NOTE FOR RESEARCH 27 OCT 2019
+# SEE EQN 2.4-2.9 OF AHMADIAN ET AL 2015
+# PARTICULARLY 2.8, IN WHICH THE SPECTRAL DENSITY IS DEPENDENT UP ON THE MATRIX
+# L. THE DISTRIBUTION APPEARS TO DEPEND ON L^-1 IN A WAY (TRACE) THAT MIGHT 
+# CAUSE THE STABILITY. NEED TO FIGURE OUT HOW THIS WORKS FOR N < 10 AND N > 10?
+N  <- 4000;
+z  <- 10;
+M  <- matrix(data = rnorm(N * N, sd = 1/N), nrow = N, ncol = N);
+Mz <- z - M;
+ml <- Mz %*% t(Mz);
+ms <- solve(ml);
+tr <- (1/N) * sum(diag(ms));
+tr;
+
+
+Rg <- Re(eigen(M)$values);
+Ig <- Im(eigen(M)$values);
+plot(x = Rg, y = Ig);
+
+# Maybe instead, define L as {N, N, N, N}, so the SD is preserved? Then
+# use the tools above?
+
+# Another check: Show unusually high values can be shunted into the very
+# negative group.
+N <- 800;
+L <- matrix(data = 0, nrow = N, ncol = N);
+J <- matrix(data = rnorm(N*N, sd = sqrt(1/N)), nrow = N, ncol = N);
+diag(J)  <- 1;
+diag(L)  <- -1 * c(rep(x = 0.01, times = N/2), rep(x = 1.98, times = N/2));
+M1       <- L %*% J
+eM1      <- eigen(M1)$values;
+rM1      <- Re(eM1);
+iM1      <- Im(eM1);
+plot(x = rM1, y = iM1, asp = 1, cex = 0.8, pch = "+", xlim = c(-0.015, 0.015),
+     ylim = c(-0.015, 0.015));
+abline(v = 0, col = "red");
+plot(x = rM1, y = iM1, asp = 1, cex = 0.8, pch = "+", 
+     xlim = c(-1.5, -1.5), ylim = c(-1.5, 1.5));
+rm1 <- tail(sort(rM1));
+
+L        <- matrix(data = 0, nrow = N, ncol = N);
+diag(L)  <- -0.995;
+M2       <- L %*% J;
+eM2      <- eigen(M2)$values;
[…]

### 2019-12-30 notebook/ms.tex
Some side additions

@@
 
+\textbf{Additional Information:} The author declares no competing
+interests. All work was carried out by A. Bradley Duthie, and all code
+and data are accessible on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}.
+
 \textbf{References}
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.722 and 0.74 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.723 and 0.737 for a and b,
 respectively.

### 2020-01-08 R/random_mats.R
Address the rho issue with a new function

@@ -326,3 +326,30 @@ rand_A1_test <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
     ret_mat <- ret_mat[ret_mat[,1] > 2,];
     return(ret_mat);
 }
+
+# Build a matrix with fixed correlation between A[i, j] and A[j, i]
+build_rho_mat <- function(S, sigma, rho){
+    mat  <- matrix(data = 0, nrow = S, ncol = S);
+    ia   <- 0.5 * ((S * S) - S);
+    ut   <- rnorm(n = ia, mean = 0, sd = sigma);
+    lt   <- rnorm(n = ia, mean = 0, sd = sigma);
+    od   <- cbind(ut, lt);
+    co   <- var(od);
+    ch   <- solve(chol(co));
+    nx   <- od %*% ch;
+    ms   <- matrix(data = c(1, rho, rho, 1), ncol = 2);
+    c2   <- chol(ms);
+    el   <- nx %*% c2 * sd(ut) + mean(ut);
+    fc   <- 1;
+    for(i in 1:S){
+        for(j in 1:S){
+            if(i < j){
+                mat[i, j] <- el[fc, 1];
+                mat[j, i] <- el[fc, 2];
+                fc        <- fc + 1;
+            }
+        }
+    }
+    diag(mat)           <- -1;
+    return(mat);
+}

### 2020-01-08 notebook/sci_rep_reviews.Rmd
Start a new set of responses

@@
 ---
+
+I have taken the following actions:
+
+1.
+Re-run simulations while calculating the correlation between off-diagonal elements of A[i,j] and A[j,i] to see if the correlation changes with $\gamma$ across $S$.
+It does not.
+I also generated systems of known correlation coefficients to assess how this modulated $\gamma$'s affect on stability.
+It did not.
+
+2. Run simulations across multiple structured network types to confirm the robustness of results. Network types include (a) cascade food webs, (b) small world networks, and (c) scale-free network.
 

### 2020-01-08 R/random_mats.R
Added absolute change in rho

@@ -62,7 +62,7 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 3);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         while(iter > 0){
             r_vec    <- rnorm(n = sp_try[i], mean = 0, sd = rmx);
@@ -85,6 +85,7 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
             A0_rho   <- mat_rho(A0);
             A1_rho   <- mat_rho(A1);
             rho_diff <- A1_rho - A0_rho;
+            rho_abs  <- abs(A1_rho) - abs(A0_rho);
             if(A0_stb == TRUE){
                 tot_res[[i]][iter, 1] <- 1;
             }
@@ -100,6 +101,7 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
             rho_res[[i]][iter, 1] <- A0_rho;
             rho_res[[i]][iter, 2] <- A1_rho;
             rho_res[[i]][iter, 3] <- rho_diff;
+            rho_res[[i]][iter, 4] <- rho_abs;
             cmplxty[[i]][iter, 1] <- get_complexity(A0);
             cmplxty[[i]][iter, 2] <- get_complexity(A1);
             iter                  <- iter - 1;
@@ -170,7 +172,7 @@ species_interactions <- function(mat, type = 0){
 #'@export
 summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 18);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 19);
     for(i in 1:sims){
         all_res[i, 1]  <- i + 1;
         # Stable and unstable
@@ -197,14 +199,15 @@ summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
[…]

### 2020-01-08 R/small_world.R
New small world with abs rho

@@ -10,7 +10,7 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 3);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         real_Cs[[i]]   <- matrix(data = 0, nrow = iter, ncol = 3);
         while(iter > 0){
@@ -40,6 +40,7 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
             A0_rho   <- mat_rho(A0);
             A1_rho   <- mat_rho(A1);
             rho_diff <- A1_rho - A0_rho;
+            rho_abs  <- abs(A1_rho) - abs(A0_rho);
             if(A0_stb == TRUE){
                 tot_res[[i]][iter, 1] <- 1;
             }
@@ -55,6 +56,7 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
             rho_res[[i]][iter, 1] <- A0_rho;
             rho_res[[i]][iter, 2] <- A1_rho;
             rho_res[[i]][iter, 3] <- rho_diff;
+            rho_res[[i]][iter, 4] <- rho_abs;
             cmplxty[[i]][iter, 1] <- get_complexity(A0);
             cmplxty[[i]][iter, 2] <- get_complexity(A1);
             iter                  <- iter - 1;

### 2020-01-10 R/analytical.R
Some more playing around

@@ -430,7 +430,7 @@ mtext(text = "E leading real part eigenvalue", side = 2,
 
 
 rand_fishing <- function(S, iters, count_stop, int_type = 0, C = 1, 
-                         by = 1, sigma = 0.4){
+                         sigma = 0.4){
     tot_res <- NULL;
     iter    <- iters;
     count   <- 0;
@@ -438,6 +438,7 @@ rand_fishing <- function(S, iters, count_stop, int_type = 0, C = 1,
         A0_dat   <- rnorm(n = S * S, mean = 0, sd = sigma);
         A0       <- matrix(data = A0_dat, nrow = S, ncol = S);
         gam1     <- runif(n = S, min = 0, max = 2);
+        diag(A0) <- -1;
         A1       <- A0 * gam1;
         A0       <- A0 * mean(gam1);
         A0_eig   <- max(Re(eigen(A0)$values));
@@ -463,4 +464,203 @@ rand_fishing <- function(S, iters, count_stop, int_type = 0, C = 1,
     }
     return(tot_res);
 }
+
+
+
+fish <- rand_fishing(S = 30, iters = 10000000, count_stop = 5)
+
+
+
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+
+A0       <- build_rho_mat(S = 1000, sigma = 0.4, rho = 0);
+gam1     <- runif(n = 1000, min = 0, max = 2);
+A1       <- A0 * gam1;
+A0       <- A0 * mean(gam1);
+A0_r     <- Re(eigen(A0)$values);
+A0_i     <- Im(eigen(A0)$values);
+A1_r     <- Re(eigen(A1)$values);
+A1_i     <- Im(eigen(A1)$values);
+
+par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));
+plot(x = A1_r, y = A1_i, pch = 4, cex = 0.7, col = "firebrick", 
+     ylim = c(-24, 24), xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, 
+     asp = 1);
+points(x = A0_r, y = A0_i, pch = 4, cex = 0.7,  col = "dodgerblue4");
+
+
[…]

### 2020-01-10 notebook/revision_notes.Rmd
Starting to write up the general idea

---
title: "Component response rate variation underlies the stability of complex systems (revision notes)"
author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - nature.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

Role of correlated matrices in stabilisation
================================================================================

```{r, echo = FALSE}
source("../R/random_mats.R");
source("../R/small_world.R");
rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1, 
                         by = 1, sigma = 0.4){
    tot_res <- NULL;
    fea_res <- NULL;
    rho_res <- NULL;
    cmplxty <- NULL;
    for(i in 1:length(rhos)){
        iter           <- iters;
        tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 8);
        fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
        cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
        while(iter > 0){
            r_vec    <- rnorm(n = S, mean = 0, sd = rmx);
            A0       <- build_rho_mat(S = S, sigma = sigma, rho = rhos[i]);
            gam1     <- runif(n = S, min = 0, max = 2);
            A1       <- A0 * gam1;
            A0       <- A0 * mean(gam1);
            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
            A0_rho   <- mat_rho(A0);
            A1_rho   <- mat_rho(A1);
            rho_diff <- A1_rho - A0_rho;
            rho_abs  <- abs(A1_rho) - abs(A0_rho);
            if(A0_stb == TRUE){
                tot_res[[i]][iter, 1] <- 1;
            }
            if(A1_stb == TRUE){
                tot_res[[i]][iter, 2] <- 1;
            }
            if(A0_fea == TRUE){
                fea_res[[i]][iter, 1] <- 1;
            }
            if(A1_fea == TRUE){
                fea_res[[i]][iter, 2] <- 1;
            }
            rho_res[[i]][iter, 1] <- A0_rho;
            rho_res[[i]][iter, 2] <- A1_rho;
            rho_res[[i]][iter, 3] <- rho_diff;
            rho_res[[i]][iter, 4] <- rho_abs;
            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
            cmplxty[[i]][iter, 1] <- get_complexity(A0);
            cmplxty[[i]][iter, 2] <- get_complexity(A1);
            iter                  <- iter - 1;
        }
    }
    all_res     <- summarise_randmat(tot_res = tot_res, fea_res = fea_res,
                                     rho_res = rho_res, cmplxty = cmplxty);
    all_res[,1]          <- rhos;
    colnames(all_res)[1] <- "rho_set";
    return(all_res);
}

simpleboot <- function(freqs,repli=1000,alpha=0.05){    
    vals  <- NULL;                                                          
    i     <- 0;                                         
    while(i < repli){                                                   
        boot  <- sample(x=freqs,size=length(freqs),replace=TRUE);           
        strap <- mean(boot);                                                
        vals  <- c(vals,strap);                                             
        i     <- i + 1;                                                     
    }                                                                       
    vals   <- sort(x=vals,decreasing=FALSE);                            
    lowCI  <- vals[round((alpha*0.5)*repli)];                               
    highCI <- vals[round((1-(alpha*0.5))*repli)];                           
    CIs    <- c(lowCI,highCI);                                          
    return(CIs);                                                            
}

summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
    sims    <- length(tot_res);
    all_res <- matrix(data = 0, nrow = sims, ncol = 25);
    for(i in 1:sims){
        all_res[i, 1]  <- i + 1;
        # Stable and unstable
        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
        # Stabilised and destabilised
        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
                                 tot_res[[i]][,2] == TRUE);
        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
                                 tot_res[[i]][,2] == FALSE);
        # Feasible and infeasible
        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
        # Feased and defeased
        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
                                  fea_res[[i]][,2] == TRUE);
        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
                                  fea_res[[i]][,2] == FALSE);
        # Mean correlations between A[i,j] and A[j,i] off-diag elements
        all_res[i, 14] <- mean(rho_res[[i]][,1]);
        all_res[i, 15] <- mean(rho_res[[i]][,2]);
        all_res[i, 16] <- mean(rho_res[[i]][,3]);
        all_res[i, 17] <- mean(rho_res[[i]][,4]);
        all_res[i, 18] <- mean(cmplxty[[i]][,1]);
        all_res[i, 19] <- mean(cmplxty[[i]][,2]);
        all_res[i, 20] <- mean(rho_res[[i]][,5]);
        all_res[i, 21] <- mean(rho_res[[i]][,6]);
        all_res[i, 22:23] <- range(rho_res[[i]][,5]);
        all_res[i, 24:25] <- range(rho_res[[i]][,6]);
    }
    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
                "A0_feasible", "A1_infeasible", "A1_feasible", 
                "A1_made_feasible", "A1_made_infeasible", "A0_rho", "A1_rho",
                "rho_diff", "rho_abs", "complex_A0", "complex_A1", "A0_eig",
                "A1_eig", "LCI_A0", "UCI_A0", "LCI_A1", "UCI_A1");
    colnames(all_res) <- cnames;
    return(all_res);
}
```

In complex systems represented by large random matrices, correlation between matrix elements $A_{ij}$ and $A_{ji}$ affects the distribution of eigenvalues and therefore local stability. As the correlation between matrix elements ($\rho$) decreases, the eigenvalue spectra changes such that more variation falls along the imaginary axis. The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which ($\rho$ = -0.5). In both cases, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $0$ and $\sigma = 0.4$.

```{r, echo = FALSE}
mat_rho0  <- build_rho_mat(S = 1000, sigma = 0.4, rho = 0);
mat_rho5  <- build_rho_mat(S = 1000, sigma = 0.4, rho = -0.5);
mr0_eig   <- eigen(mat_rho0)$values;
mr5_eig   <- eigen(mat_rho5)$values;
mr0_r     <- Re(mr0_eig);
mr0_i     <- Im(mr0_eig);
mr5_r     <- Re(mr5_eig);
mr5_i     <- Im(mr5_eig);
par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));
plot(x = mr0_r, y = mr0_i, pch = 4, cex = 0.7, col = "firebrick",
     xlab = "", ylab = "", cex.lab = 1.25, cex.axis = 1.25, asp = 1, 
     ylim = c(min(mr5_i), max(mr5_i)));
plot(x = mr5_r, y = mr5_i, pch = 4, cex = 0.7, col = "firebrick", 
     xlab = "", ylab = "", cex.lab = 1.25, cex.axis = 1.25, asp = 1, 
     yaxt = "n");
mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 1.5);
mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 1.5);
```

Because of this affect of $\rho$ on the eigenvalue spectra, decreasing values of $\rho$ will also decrease the rightmost eigenvalue of the matrix $A$. This makes it more likely that the complex system represented by $A$ is locally stable, as stability occurs when all real parts of eigenvalues are negative. Note that this elongation along the imaginary axis is also characteristic of predator-prey communities (in which, by definition $A_{ij}$ and $A_{ji}$ have opposing signs). Also note that as $\rho$ increases such that $\rho > 0$, the same elongation happens along the real axis, making random complex systems less likely to be stable.

```{r, echo = FALSE}
sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
                         iters = 1000);
```

A simple numerical analysis illustrates the linear relationship between $\rho$ and the expected value of the real part of the leading eigenvalue $\max(\Re(\lambda))$. Below, I have run 1000 simulations across values of $\rho$ from `r min(sim_rhos[,1])` to `r max(sim_rhos[,1])` for complex systems with $S = 25$ components. Error bars show 95\% bootstrapped confidence intervals.

```{r, echo = FALSE}
par(mar = c(5, 5, 1, 1));
plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
     xlab = expression(paste("Correlation between A"[ij]," and A"[ji])),
     ylab = "Leading real part eigenvalue", 
     ylim = c(min(sim[,22]), max(sim[,23])));
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 2);
arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 2);
```

```{r, echo = FALSE}
dat <- read.csv(file = "sim_results/C_1/random_all.csv");
dat <- dat[,-1];
```

In the main text, I demonstrated that when $S$ is finite but system complexity $\sigma\sqrt{SC}$ is high ($C$ defines the connectance of $A$, or the proportion of non-zero off-diagonal elements), variation in component response rate $\gamma$ often underlies system stability. Mathematically, this means that by multiplying $A$ by a diagonal matrix $\gamma$ with variable elements, the sign of $\max(\Re(\lambda))$ is sometimes flipped from positive to negative in these finite systems of high complexity. Interestingly, this increase in stability given $Var(\gamma) > 0$ is not caused by $\gamma$ decrease $\rho$. In fact, for $S = 25$, $\sigma = 0.4$, and $C = 1$, random complex systems that are stabilised by $\gamma$ typically have increased $\rho$ values. Note that for these parameter values, 1 million simulations found $M = \gamma A$ to be stable for `r dat[24, 3]` systems when all $\gamma_{i} = 1$, but `r dat[24, 5]` systems when $\gamma_{i} \sim \mathcal{U}(0, 2)$. Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.

### 2020-01-10 notebook/revision_notes.tex
Starting to write up the general idea

--- 2019-03-02 notebook/ms.tex
@@
 \usepackage{setspace}
-\setstretch{2}
+\setstretch{1}
 \usepackage{amssymb,amsmath}
@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
+            pdftitle={Component response rate variation underlies the stability of complex systems (revision notes)},
             pdfauthor={A. Bradley Duthie ( alexander.duthie@stir.ac.uk )},
@@
 % Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
+\providecommand{\subtitle}[1]{
   \posttitle{
@@
   \title{Component response rate variation underlies the stability of complex
-systems}
+systems (revision notes)}
     \pretitle{\vspace{\droptitle}\centering\huge}
@@
 
-\textbf{Key words:} Ecological networks, gene-regulatory networks,
-neural networks, financial networks, system stability, random matrix
-theory
+\hypertarget{role-of-correlated-matrices-in-stabilisation}{%
+\section{Role of correlated matrices in
+stabilisation}\label{role-of-correlated-matrices-in-stabilisation}}
 
-\subsection{Abstract}\label{abstract}
+In complex systems represented by large random matrices, correlation
+between matrix elements \(A_{ij}\) and \(A_{ji}\) affects the
+distribution of eigenvalues and therefore local stability. As the
+correlation between matrix elements (\(\rho\)) decreases, the eigenvalue
+spectra changes such that more variation falls along the imaginary axis.
+The figure panels below compare a random matrix in which \(\rho = 0\)
+(left) to one in which (\(\rho\) = -0.5). In both cases, complex systems
+include \(S = 1000\) components, with diagonal elements of \(-1\) and
+off-diagonal elements drawn from a normal distribution with a mean of
+\(0\) and \(\sigma = 0.4\).
 
-The stability of a complex system generally decreases with increasing
-system size and interconnectivity, a counterintuitive result of
-widespread importance across the physical, life, and social sciences.
-Despite recent interest in the relationship between system properties
-and stability, the effect of variation in the response rate of
-individual system components remains unconsidered. Here I vary the
-component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
-complex systems. I show that when component response rates vary, the
-potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) becomes increasingly important for stability in
-large but finite systems, where the probability of stability would
-otherwise be negligible. At such extremes of simulated system
-complexity, the largest stable complex systems would be unstable if not
-for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
-previously unconsidered aspect of system stability that is likely to be
-pervasive across all realistic complex systems.
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-2-1.pdf}
 
-\subsection{Introduction}\label{introduction}
+Because of this affect of \(\rho\) on the eigenvalue spectra, decreasing
+values of \(\rho\) will also decrease the rightmost eigenvalue of the
+matrix \(A\). This makes it more likely that the complex system
+represented by \(A\) is locally stable, as stability occurs when all
+real parts of eigenvalues are negative. Note that this elongation along
+the imaginary axis is also characteristic of predator-prey communities
+(in which, by definition \(A_{ij}\) and \(A_{ji}\) have opposing signs).
+Also note that as \(\rho\) increases such that \(\rho > 0\), the same
+elongation happens along the real axis, making random complex systems
+less likely to be stable.
 
-In 1972, May\textsuperscript{\protect\hyperlink{ref-May1972}{1}} first
-demonstrated that randomly assembled systems of sufficient complexity
-are almost inevitably unstable given infinitesimally small
-perturbations. Complexity in this case is defined by the size of the
-system (i.e., the number of potentially interacting components; \(S\)),
-its connectance (i.e., the probability that one component will interact
-with another; \(C\)), and the variance of interaction strengths
-(\(\sigma^{2}\))\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-May's finding that the probability of local stability falls to near zero
-given a sufficiently high threshold of \(\sigma\sqrt{SC}\) is broadly
-relevant for understanding the dynamics and persistence of systems such
-as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
-and
-socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
-networks. As such, identifying general principles that affect stability
-in complex systems is of wide-ranging importance.
+A simple numerical analysis illustrates the linear relationship between
+\(\rho\) and the expected value of the real part of the leading
+eigenvalue \(\max(\Re(\lambda))\). Below, I have run 1000 simulations
+across values of \(\rho\) from -0.9 to 0.9 for complex systems with
+\(S = 25\) components.
 
-Randomly assembled complex systems can be represented as large square
-matrices (\(\mathbf{M}\)) with \(S\) components (e.g., networks of
-species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
-element of such a matrix, \(M_{ij}\), defines how component \(j\)
-affects component \(i\) in the system at a point of
-equilibrium\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}.
-Off-diagonal elements (\(i \neq j\)) therefore define interactions
-between components, while diagonal elements (\(i = j\)) define component
-self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a
-probability \(C\), which are sampled from a distribution with variance
-\(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
-Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of
-\(\mathbf{M}\) are negative
-(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
-In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
-circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
-\(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
-(Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma^{2}\) increase.
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-4-1.pdf}
 
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
-response rates (\(\gamma\)) of individual components to perturbations of
-the system are identical, but this is highly unlikely in any complex
-system. In ecological communities, for example, the rate at which
-population density changes following perturbation will depend on the
-generation time of organisms, which might vary by orders of magnitude
-among species. Species with short generation times will respond quickly
-(high \(\gamma\)) to perturbations relative to species with long
-generation times (low \(\gamma\)). Similarly, the speed at which
-individual banks respond to perturbations in financial networks, or
-individuals or institutions respond to perturbations in complex social
-networks, is likely to vary. The effect of such variance on stability
-has not been investigated in complex systems theory. Intuitively,
-variation in \(\gamma\) might be expected to decrease system stability
-by introducing a new source of variation into the system and thereby
-increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
-realistic complex systems (such that \(S\) is high but finite) in which
-\(\gamma\) varies are actually more likely to be stable.
-
-\subsection{Results}\label{results}
-
-\textbf{Component response rates of random complex systems}. Rows in
-\(\mathbf{M}\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
-\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
-half of \(\gamma_{i} = 0.05\). This models one system in which
-\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
-are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
-cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
-diagonal elements are set to \(-1\) and non-zero off-diagonal elements
-are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
-multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
-\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
-uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
-(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
-to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
-values are separated into two groups, eigenvalues are no longer
-uniformly distributed (Fig. 1b). Instead, two distinct clusters of
-eigenvalues appear (red circles in Fig. 1b), one centred at
-(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
-large radius, but the real components have shifted to the left (in
-comparison to when \(\gamma = 1\)), while the latter cluster has a
-smaller radius, but real components have shifted to the right; all
-\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
-\(\mathbf{M}\), this division between slow and fast component response
-rates results in more stable systems: 1 stable given \(\gamma = 1\)
-versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
-
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I focus on
-a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
-Supplementary Information for other distributions, which give similar
-results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
-\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(\mathbf{M}\) before and after the addition of variation
-in component response rate. Figure 2 shows a comparison of eigenvalue
-distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
-some of which are clustered locally near the centre of the distribution,
-but others of which are spread outside the former radius of 2.828 (Fig
-2b). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold \(\sigma\sqrt{SC} < 1\).
-
-\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
-investigate the effect of \(Var(\gamma)\) on stability across systems of
-varying complexity, I simulated random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One
-million \(\mathbf{M}\) were simulated for each \(S\), and the stability
-of \(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
-\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
-number of stable random systems was higher given \(Var(\gamma)\) than
-when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
-table of results), and that the difference between the probabilities of
-observing a stable system increased with an increase in \(S\). In other
-words, the potential for \(Var(\gamma)\) to affect stability increased
-with system complexity and was most relevant for systems on the cusp of
-being too complex to be realistically stable. For the highest values of
-\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
-not have been stable given \(\gamma = 1\).
-
-\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
-investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
-genetic algorithm. Genetic algorithms are heuristic tools that mimic
-evolution by natural selection, and are useful when the space of
-potential solutions (in this case, possible combinations of \(\gamma\)
-values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
-Generations of selection on \(\gamma\) value combinations to minimise
-\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
-\(Var(\gamma)\) to increase system stability. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were up to four
-orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
-that stability could often be achieved by manipulating \(S\) \(\gamma\)
-values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
-manipulating component response rates rather than interactions between
-components).
-
-\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
-systems in which individual system components represent the density of
-some tangible quantity, it is relevant to consider the feasibility of
-the system. Feasibilility assumes that values of all components are
-positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
-This is of particular interest for ecological communities because
-population density (\(N\)) cannot take negative values, meaning that
-ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
-While my results are intended to be general to all complex systems, and
-not restricted to species networks, I have also performed a feasibility
-analysis on all matrices \(\mathbf{M}\) tested for stability, and
-additionally for specific types of ecological
-communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-(e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). I emphasise that \(\gamma\) is not interpreted as
-population density in this analysis, but instead as a fundamental
-property of species life history such as expected generation time.
-Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
-a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
-(see Supplementary Information). Hence, for pure interacting species
-networks, variation in component response rate (i.e., species generation
-time) does not affect stability at biologically realistic species
-densities.
-
-\subsection{Discussion}\label{discussion}
-
-I have shown that the stability of complex systems might often be
-contigent upon variation in the response rates of their individual
-components, meaning that factors such as rate of trait evolution (in
-biological networks), transaction speed (in economic networks), or
-communication speed (in social networks) need to be considered when
-investigating the stability of complex systems. Variation in component
-response rate is more likely to be critical for stability in systems
-that are especially complex, and it can ultimately increase the
-probability that system stability is observed above that predicted by
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}} classically
-derived \(\sigma \sqrt{SC}\) criterion. The logic outlined here is
-general, and potentially applies to any complex system in which
-individual system components can vary in their reaction rates to system
-perturbation.
-
-It is important to recognise that variation in component response rate
-is not stabilising per se; that is, adding variation in component
-response rates to a particular system does not increase the probability
-that the system will be stable. Rather, highly complex systems that are
-observed to be stable are more likely to have varying component response
-rates, and for this variation to be critical to their stability (Fig.
-3). This is caused by the shift to a non-uniform distribution of
-eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
-which can sometimes cause all of the real components of the eigenvalues
-of the system matrix to become negative, but might also increase the
-real components of eigenvalues.
-
-My focus is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
-the same mathematical framework to investigate how a diagonal matrix
-\(\mathbf{X}\) (equivalent to \(\gamma\) in my model) affects the
-stability of a community matrix \(\mathbf{M}\) given an interaction
-matrix \(\mathbf{A}\) within a generalised Lotka-Volterra model, where
-\(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}} analytically
-demonstrated that the effect of \(\mathbf{X}\) on system stability
-decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)). My numerical results do not contradict this
-prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
-instead fixed \(\sigma\) and increased \(S\) to thereby increase total
-system complexity, \(\sigma\sqrt{SC}\) (see Supplemental Information for
-results simulated across \(\sigma\) and \(C\)). Overall, I show that
-component response rate variation increases the upper bound of
-complexity at which stability can be realistically observed, meaning
-that highly complex systems are more likely than not to vary in their
-component response rates, and for this variation to be critical for
-system stability.
-
-The potential importance of component response rate variation was most
-evident from the results of simulations in which the genetic algorithm
-was used in attempt to maximise the probability of system stability. The
-probability that some combination of component response rates could be
-found to stabilise the system was shown to be up to four orders of
-magnitude higher than the background probabilities of stability in the
-absence of any component response rate variation. Instead of
-manipulating the \(S \times S\) interactions between system components,
-it might therefore be possible to manipulate only the \(S\) response
-rates of individual system components to achieve stability. Hence,
-managing the response rates of system components in a targeted way could
-potentially facilitate the stabilisation of complex systems through a
-reduction in dimensionality.
-
-Interestingly, while complex systems were more likely to be stable given
-variation in component response rate, they were not more likely to be
-feasible, meaning that stability was not increased when component values
-were also restricted to being positive at equilibrium. Feasibility is
-important to consider, particularly for the study of ecological networks
-of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}}
-because population densities cannot realistically be negative. My
-results therefore suggest that variation in the rate of population
-responses to perturbation (e.g., due to differences in generation time
-among species) is unlikely to be critical to the stability of purely
-multi-species interaction networks (see also Supplementary Information).
-Nevertheless, ecological interactions do not exist in isolation in
-empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
-instead interact with evolutionary, abiotic, or social-economic systems.
-The relevance of component response rate for complex system stability
-should therefore not be ignored in the broader context of ecological
-communities.
-
-A general mathematical framework encompassing shifts in eigenvalue
-distributions caused by a vector \(\gamma\) has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
-and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
-but \(\gamma\) has not been interpreted as rates of response of
-individual system components to perturbation. My model focuses on
-component response rates for systems of a finite size, in which
-complexity is high but not yet high enough to make the probability of
-stability unrealistically low for actual empirical systems. For this
-upper range of system size, randomly assembled complex systems are more
-likely to be stable if their component response rates vary (e.g.,
-\(10 < S < 30\) for parameter values in Fig. 3). Overall, I suggest that
-variation in component response rate might therefore be critical for
-maintaining stability in many highly complex empirical systems. These
-results are broadly applicable for understanding the stability of
-complex networks across the physical, life, and social sciences.
-
-\subsection{Methods}\label{methods}
-
-\textbf{Component response rate variation (\(\mathbf{\gamma}\))}. In a
-synthesis of eco-evolutionary feedbacks on community stability, Patel et
-al. model a system that includes a vector of potentially changing
-species densities (\(\mathbf{N}\)) and a vector of potentially evolving
-traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
-
-\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
-
-In the above, \(f_{i}\) and \(g_{j}\) are functions that define the
-effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
-were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
-this is modulated in the above by the scalar \(\epsilon\). The value of
-\(\epsilon\) thereby determines the timescale separation between ecology
-and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-
-I use the same principle that Patel et al. use to modulate the relative
-rate of evolution to modulate rates of component responses for \(S\)
-components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}},
-the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
-by the value of \(j\) (\(v_{j}(t)\)) and \(j\)'s marginal effect on
-\(i\) (\(a_{ij}\)), and by \(i\)'s response rate (\(\gamma_{i}\)),
-
-\[\frac{dv_{i}(t)}{dt} = \gamma_{i} \sum_{j=1}^{S}a_{ij}v_{j}(t).\]
-
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{23}},
-
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the values of system
-components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
-In these analyses, row means of \(\mathbf{A}\) are expected to be
-identical, but variation around this expectation will naturally arise
-due to random sampling of \(\mathbf{A}\) off-diagonal elements and
-finite \(S\). In simulations, the total variation in \(\mathbf{M}\) row
-means that is attributable to \(\mathbf{A}\) is small relative to that
-attributable to \(\mathbf{\gamma}\), especially at high \(S\). Variation
-in \(\mathbf{\gamma}\) specifically isolates the effects of differing
-component response rates, hence causing differences in expected
-\(\mathbf{M}\) row means.
-
-\textbf{Genetic algorithm}. Ideally, to investigate the potential of
-\(Var(\gamma)\) for increasing the proportion of stable complex systems,
-the search space of all possible \(\gamma\) vectors would be evaluated
-for each unique \(\mathbf{M = \gamma A}\). This is technically
-impossible because \(\gamma_{i}\) can take any real value between 0-2,
-but even rounding \(\gamma_{i}\) to reasonable values would result in a
-search space too large to practically explore. Under these conditions,
-genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
-this case, the practical solution is finding vectors of
-\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
-\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
-a large population of 1000 different potential \(\mathbf{\gamma}\)
-vectors and allowing this population to evolve through a process of
-mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
-selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
-found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
-met.
-
-For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
-for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)). The
-genetic algorithm was initialised with a population of 1000 different
-\(\mathbf{\gamma}\) vectors with elements sampled i.i.d from
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed on
-the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector, and
-the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal
-offspring with subsequent random mutation and crossover between the
-resulting new generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation
-of each \(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
-probability of 0.2, resulting in a mutation effect of size
-\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
-\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
-multiplied by \(-1\), and any values that mutated above 2 were set to
-2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
-vectors paired in each generation; vectors were randomly sampled with
-replacement among but not within sets. Vector pairs selected for
-crossover swapped all elements between and including two \(\gamma_{i}\)
-randomly selected with replacement (this allowed for reversal of vector
-element positions during crossover; e.g.,
-\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
-). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
-found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
-fitness increase between generations was less than 0.01 (where fitness
-was defined as \(W_{\gamma} = -\max\left(\Re(\lambda)\right)\) for
-\(\mathbf{M}\)).
-
-\textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} identify
-the following feasibility criteria for ecological systems characterised
-by \(S\) interacting species with varying densities in a generalised
-Lotka-Volterra model,
-
-\[\mathbf{n^{*}} = -\left(\theta \mathbf{I} + (CS)^{-\delta}\mathbf{J} \right)^{-1}\mathbf{r}.\]
-
-In the above, \(\mathbf{n^{*}}\) is the vector of species densities at
-equilibrium. Feasibility is satisfied if all elements in
-\(\mathbf{n^{*}}\) are positive. The matrix \(\mathbf{I}\) is the
-identity matrix, and the value \(\theta\) is the strength of
-intraspecific competition (diagonal elements). Diagonal values are set
-to \(-1\), so \(\theta = -1\). The variable \(\delta\) is a
-normalisation parameter that modulates the strength of interactions
-(\(\sigma\)) for \(\mathbf{J}\). Implicitly, here \(\delta = 0\)
-underlying strong interactions. Hence, \((CS)^{-\delta} = 1\), so in the
-above, a diagonal matrix of -1s (\(\theta \mathbf{I}\)) is added to
-\(\mathbf{J}\), which has a diagonal of all zeros and an off-diagonal
-affecting species interactions (i.e., the expression \((CS)^{-\delta}\)
-relates to May's\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
-stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
-\(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
-purposes \((CS)^{-\delta} = 1\)). Given
-\(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),
-
-\[\mathbf{n^{*} = -A^{-1}r}.\]
-
-To check the feasibility criteria for \(\mathbf{M = \gamma A}\), I
-therefore evaluated \(\mathbf{-M^{-1}r}\) (\(\mathbf{r}\) elements were
-sampled i.i.d. from \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility
-is satisfied if all of the elements of the resulting vector are
-positive.
-
-\textbf{Acknowledgements:} I am supported by a Leverhulme Trust Early
-Career Fellowship (ECF-2016-376). Conversations with L. Bussière and N.
-Bunnefeld, and comments from J. J. Cusack and I. L. Jones, improved the
-quality of this work.
-
-\textbf{Supplementary Information:} Full tables of stability results for
-simulations across different system size (\(S\)) values, ecological
-community types, connectance (\(C\)) values, interaction strengths
-(\(\sigma\)), and \(\gamma\) distributions are provided as supplementary
-material. An additional table also shows results for how feasibility
-changes across \(S\). All code and simulation outputs are publicly
-available as part of the RandomMatrixStability package on GitHub
-(\url{https://github.com/bradduthie/RandomMatrixStability}).
-
-\textbf{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-May1972}{}
-1. May, R. M. Will a large complex system be stable? \emph{Nature}
-\textbf{238,} 413--414 (1972).
-
-\hypertarget{ref-Allesina2012}{}
-2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
-\emph{Nature} \textbf{483,} 205--208 (2012).
-
-\hypertarget{ref-Mougi2012}{}
-3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
-community stability. \emph{Science} \textbf{337,} 349--351 (2012).
-
-\hypertarget{ref-Allesina2015}{}
-4. Allesina, S. \emph{et al.} Predicting the stability of large
-structured food webs. \emph{Nature Communications} \textbf{6,} 7842
-(2015).
-
-\hypertarget{ref-Grilli2017}{}
-5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
-ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
-
-\hypertarget{ref-Gray2008}{}
-6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
-random brain networks with a distribution of connection strengths.
-\emph{Neurocomputing} \textbf{71,} 1373--1387 (2008).
-
-\hypertarget{ref-Gray2009}{}
-7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
-with excitatory and inhibitory connections. \emph{Neurocomputing}
-\textbf{72,} 1849--1858 (2009).
-
-\hypertarget{ref-Rosenfeld2009}{}
-8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
-high-dimensional biochemical networks. \emph{Gene Regulation and Systems
-Biology} \textbf{3,} 1--10 (2009).
-
-\hypertarget{ref-MacArthur2010}{}
-9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
-and criticality of adaptive regulatory networks. \emph{Physics Review
-Letters} \textbf{104,} 168701 (2010).
-
-\hypertarget{ref-May2008}{}
-10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
-for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
-
-\hypertarget{ref-Haldane2011}{}
-11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
-\emph{Nature} \textbf{469,} 351--355 (2011).
-
-\hypertarget{ref-Suweis2014}{}
-12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
-networks. \emph{PLoS ONE} \textbf{9,} (2014).
-
-\hypertarget{ref-Bardoscia2017}{}
-13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
-Pathways towards instability in financial networks. \emph{Nature
-Communications} \textbf{8,} 1--7 (2017).
-
-\hypertarget{ref-Tao2010}{}
-14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
-circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
-(2010).
-
-\hypertarget{ref-Patel2018}{}
-15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
-effects of eco-evolutionary feedbacks on community stability.
-\emph{American Naturalist} \textbf{191,} 1--29 (2018).
-
-\hypertarget{ref-Ahmadian2015}{}
-16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
-with partially structured and partially random connectivity.
-\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
-Physics} \textbf{91,} 012820 (2015).
-
-\hypertarget{ref-Gibbs2017}{}
-17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
-population abundances on the stability of large random ecosystems.
-\emph{arXiv} (2017).
-
-\hypertarget{ref-Stone2017}{}
-18. Stone, L. The feasibility and stability of large complex biological
-networks: a random matrix approach. \emph{Scientific Reports}
-\textbf{8,} 8246 (2018).
-
-\hypertarget{ref-Hamblin2013}{}
-19. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
-
-\hypertarget{ref-Dougoud2018}{}
-20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
-The feasibility of equilibria in large ecosystems: a primary but
-neglected concept in the complexity-stability debate. \emph{PLOS
-Computational Biology} \textbf{14,} e1005988 (2018).
-
-\hypertarget{ref-Song2018}{}
-21. Song, C. \& Saavedra, S. Will a small randomly assembled community
-be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
-
-\hypertarget{ref-Servan2018}{}
-22. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
-Allesina, S. Coexistence of many species in random ecosystems.
-\emph{Nature Ecology and Evolution} \textbf{2,} 1237--1242 (2018).
-
-\hypertarget{ref-May1973}{}
-23. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
-\textbf{54,} 638--641 (1973).
-
-\clearpage
-
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the blue circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle centred at \(\gamma = -1.95\) and one small circle centred
-at \(\gamma = -0.05\). In a, the maximum real eigenvalue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 was stable when
-\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-\clearpage
-
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
-\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
-eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
-both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.742 for a and b,
-respectively.
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
-
-\clearpage
-
-\textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate (\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
-which variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Blue and red bars show the number of stable systems in the absence and
-presence of \(Var(\gamma)\), respectively. The black line shows the
-proportion of systems that are stable when \(Var(\gamma) > 0\), but
-would be unstable if \(Var(\gamma) = 0\).
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-15-1.pdf}
-
-\clearpage
-
-\textbf{Figure 4: Stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted
-\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
-stable across different system sizes (\(S\), max \(S=40\)) for
-\(C = 1\), and the proportion of systems wherein a targeted search of
-\(\gamma\) values successfully resulted in system stability. For each
-\(S\), 100000 complex systems are randomly generated. Stability of each
-complex system is tested given variation in \(\gamma\) using a genetic
-algorithm to maximise the effect of \(\gamma\) values on increasing
-stability, as compared to stability in an otherwise identical system in
-which \(\gamma\) is the same for all components. Blue bars show the
-number of stable systems in the absence of component response rate
-variation, while red bars show the number of stable systems that can be
-generated if component response rate is varied to maximise system
-stability. The black line shows the proportion of systems that are
-stable when component response rate is targeted to increase stability,
-but would not be stable if \(Var(\gamma) = 0\).
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
+Error bars show 95\% bootstrapped confidence intervals.
 

### 2020-01-11 notebook/revision_notes.Rmd
Some more notes

@@
 As the correlation between matrix elements ($\rho$) decreases, the eigenvalue spectra changes such that more variation falls along the imaginary axis.
-The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which ($\rho$ = -0.5).
-In both cases, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $0$ and $\sigma = 0.4$.
+The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which $\rho$ = -0.5.
+In both panels, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $0$ and $\sigma = 0.4$.
 
@@
 
-Because of this affect of $\rho$ on the eigenvalue spectra, decreasing values of $\rho$ will also decrease the rightmost eigenvalue of the matrix $A$.
-This makes it more likely that the complex system represented by $A$ is locally stable, as stability occurs when all real parts of eigenvalues are negative.
+Because of this effect of $\rho$ on the eigenvalue spectra, decreasing values of $\rho$ will also tend to decrease the rightmost eigenvalue of the matrix $\textbf{A}$.
+This makes it more likely that the complex system represented by $\textbf{A}$ is locally stable, as stability occurs when all real parts of eigenvalues are negative.
 Note that this elongation along the imaginary axis is also characteristic of predator-prey communities (in which, by definition $A_{ij}$ and $A_{ji}$ have opposing signs).
@@
 ```{r, echo = FALSE}
+rho_its  <- 1000;
 sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
-                         iters = 1000);
+                         iters = rho_its);
 ```
 
-A simple numerical analysis illustrates the linear relationship between $\rho$ and the expected value of the real part of the leading eigenvalue $\max(\Re(\lambda))$.
-Below, I have run 1000 simulations across values of $\rho$ from `r min(sim_rhos[,1])` to `r max(sim_rhos[,1])` for complex systems with $S = 25$ components.
+A simple numerical analysis illustrates the linear relationship between $\rho$ and the expected value of the real part of the leading eigenvalue, $\max(\Re(\lambda))$.
+Below, I have run `r rho_its` simulations across values of $\rho$ from `r min(sim_rhos[,1])` to `r max(sim_rhos[,1])` for complex systems with $S = 25$ components.
 Error bars show 95\% bootstrapped confidence intervals.
@@
 plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
-     xlab = expression(paste("Correlation between A"[ij]," and A"[ji])),
+     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
+                             "(",rho,")")),
      ylab = "Leading real part eigenvalue", 
-     ylim = c(min(sim[,22]), max(sim[,23])));
+     ylim = c(min(sim_rhos[,22]), max(sim_rhos[,23])));
 arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
@@
 
-In the main text, I demonstrated that when $S$ is finite but system complexity $\sigma\sqrt{SC}$ is high ($C$ defines the connectance of $A$, or the proportion of non-zero off-diagonal elements), variation in component response rate $\gamma$ often underlies system stability.
-Mathematically, this means that by multiplying $A$ by a diagonal matrix $\gamma$ with variable elements, the sign of $\max(\Re(\lambda))$ is sometimes flipped from positive to negative in these finite systems of high complexity.
-Interestingly, this increase in stability given $Var(\gamma) > 0$ is not caused by $\gamma$ decrease $\rho$.
+```{r, echo = FALSE}
+sim25 <- read.csv(file = "sim_results/rhos/sim25.csv")
+```
+
+In the main text, I demonstrated that when $S$ is finite but system complexity $\sigma\sqrt{SC}$ is high ($C$ defines the connectance of $\textbf{A}$, or the proportion of non-zero off-diagonal elements), variation in component response rate $\gamma$ often underlies system stability.
+In other words, highly complex systems that are observed to be stable would not be if we removed the variation in their component response rates.
+Mathematically, this means that by multiplying $\textbf{A}$ by a diagonal matrix $\gamma$ with variable elements, the sign of $\max(\Re(\lambda))$ is sometimes flipped from positive to negative in these finite systems of high complexity.
+Interestingly, this increase in stability given $Var(\gamma) > 0$ is not necessarily caused by $\gamma$ decreasing $\rho$.
 In fact, for $S = 25$, $\sigma = 0.4$, and $C = 1$, random complex systems that are stabilised by $\gamma$ typically have increased $\rho$ values.
-Note that for these parameter values, 1 million simulations found $M = \gamma A$ to be stable for `r dat[24, 3]` systems when all $\gamma_{i} = 1$, but `r dat[24, 5]` systems when $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.
+Note that for these parameter values, 1 million simulations found $\textbf{M} = \gamma \textbf{A}$ to be stable for `r dat[24, 3]` systems when all $\gamma_{i} = 1$, but `r dat[24, 5]` systems when $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for `r dim(sim25)[1]` stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.
+
+```{r, echo = FALSE}
+rho_change <- sim25[,4] - sim25[,3];
+par(mar = c(5, 5, 1, 1));
+hist(x = rho_change, main = "", ylab = "Frequency", cex.lab = 1.25,
+     xlab = expression(paste(rho[change])), col = "grey", cex.axis = 1.25)
+```
+
+When $\textbf{A}$ was stabilised by $Var(\gamma)$, the change in $\rho$ was normally distributed around zero.
+Hence, decreasing the correlation between $\textbf{A}_{ij}$ and $\textbf{A}_{ji}$ was not by itself the cause of stability.
+For a clearer picture of the effect of $\gamma$, it is useful to show the relationship between $\rho$ and $\max(\Re(\lambda))$ again as above with, but this time with how the relationship changes given $Var(\gamma)$.

### 2020-01-11 notebook/revision_notes.tex
Some more notes

@@
 % Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
+\newcommand{\subtitle}[1]{
   \posttitle{
@@
 
-\hypertarget{role-of-correlated-matrices-in-stabilisation}{%
 \section{Role of correlated matrices in
-stabilisation}\label{role-of-correlated-matrices-in-stabilisation}}
+stabilisation}\label{role-of-correlated-matrices-in-stabilisation}
 
@@
 The figure panels below compare a random matrix in which \(\rho = 0\)
-(left) to one in which (\(\rho\) = -0.5). In both cases, complex systems
+(left) to one in which \(\rho\) = -0.5. In both panels, complex systems
 include \(S = 1000\) components, with diagonal elements of \(-1\) and
@@
 
-Because of this affect of \(\rho\) on the eigenvalue spectra, decreasing
-values of \(\rho\) will also decrease the rightmost eigenvalue of the
-matrix \(A\). This makes it more likely that the complex system
-represented by \(A\) is locally stable, as stability occurs when all
-real parts of eigenvalues are negative. Note that this elongation along
-the imaginary axis is also characteristic of predator-prey communities
-(in which, by definition \(A_{ij}\) and \(A_{ji}\) have opposing signs).
-Also note that as \(\rho\) increases such that \(\rho > 0\), the same
-elongation happens along the real axis, making random complex systems
-less likely to be stable.
+Because of this effect of \(\rho\) on the eigenvalue spectra, decreasing
+values of \(\rho\) will also tend to decrease the rightmost eigenvalue
+of the matrix \(\textbf{A}\). This makes it more likely that the complex
+system represented by \(\textbf{A}\) is locally stable, as stability
+occurs when all real parts of eigenvalues are negative. Note that this
+elongation along the imaginary axis is also characteristic of
+predator-prey communities (in which, by definition \(A_{ij}\) and
+\(A_{ji}\) have opposing signs). Also note that as \(\rho\) increases
+such that \(\rho > 0\), the same elongation happens along the real axis,
+making random complex systems less likely to be stable.
 
@@
 \(\rho\) and the expected value of the real part of the leading
-eigenvalue \(\max(\Re(\lambda))\). Below, I have run 1000 simulations
+eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 1000 simulations
 across values of \(\rho\) from -0.9 to 0.9 for complex systems with
-\(S = 25\) components.
+\(S = 25\) components. Error bars show 95\% bootstrapped confidence
+intervals.
 
@@
 
-Error bars show 95\% bootstrapped confidence intervals.
+In the main text, I demonstrated that when \(S\) is finite but system
+complexity \(\sigma\sqrt{SC}\) is high (\(C\) defines the connectance of
+\(\textbf{A}\), or the proportion of non-zero off-diagonal elements),
+variation in component response rate \(\gamma\) often underlies system
+stability. In other words, highly complex systems that are observed to
+be stable would not be if we removed the variation in their component
+response rates. Mathematically, this means that by multiplying
+\(\textbf{A}\) by a diagonal matrix \(\gamma\) with variable elements,
+the sign of \(\max(\Re(\lambda))\) is sometimes flipped from positive to
+negative in these finite systems of high complexity. Interestingly, this
+increase in stability given \(Var(\gamma) > 0\) is not necessarily
+caused by \(\gamma\) decreasing \(\rho\). In fact, for \(S = 25\),
+\(\sigma = 0.4\), and \(C = 1\), random complex systems that are
+stabilised by \(\gamma\) typically have increased \(\rho\) values. Note
+that for these parameter values, 1 million simulations found
+\(\textbf{M} = \gamma \textbf{A}\) to be stable for 36 systems when all
+\(\gamma_{i} = 1\), but 383 systems when
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Below shows the distribution of
+the difference in \(\rho\) between systems with versus without
+\(var(\gamma)\) for 100 stabilised systems; that is,
+\(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+For all cases in which \(\textbf{A}\) was stabilised by \(Var(\gamma)\),
+the resulting \(\textbf{M} = \gamma\textbf{A}\) had increased rather
+than decreased \(\rho\) (\(\rho_{var(\gamma)} > \rho_{\gamma = 1}\)).
+Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
+\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
+clearer picture of the effectof \(\gamma\), it is useful to show the
+relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above
+with, but this time with how the relationship changes given
+\(Var(\gamma)\).
 

### 2020-01-11 notebook/revision_notes.Rmd
More notes

@@
                              "(",rho,")")),
-     ylab = "Leading real part eigenvalue", 
-     ylim = c(min(sim_rhos[,22]), max(sim_rhos[,23])));
+     ylab = "Real part of leading eigenvalue", 
+     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])));
 arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
@@
        y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 2);
+abline(h = 0, lwd = 0.8, lty = "dotted");
 ```
@@
 par(mar = c(5, 5, 1, 1));
-hist(x = rho_change, main = "", ylab = "Frequency", cex.lab = 1.25,
-     xlab = expression(paste(rho[change])), col = "grey", cex.axis = 1.25)
+hist(x = rho_change, main = "", ylab = "Frequency", cex.lab = 1.25, breaks = 15,
+     xlab = expression(paste(rho[change])), col = "grey", cex.axis = 1.25);
+rho_change_CI <- simpleboot(rho_change);
 ```
 
-When $\textbf{A}$ was stabilised by $Var(\gamma)$, the change in $\rho$ was normally distributed around zero.
+When $\textbf{A}$ was stabilised by $Var(\gamma)$, the change in $\rho$ was normally distributed around `r mean(rho_change)` (95\% CIs: `r rho_change_CI[1]`, `r rho_change_CI[2]`).
 Hence, decreasing the correlation between $\textbf{A}_{ij}$ and $\textbf{A}_{ji}$ was not by itself the cause of stability.
 For a clearer picture of the effect of $\gamma$, it is useful to show the relationship between $\rho$ and $\max(\Re(\lambda))$ again as above with, but this time with how the relationship changes given $Var(\gamma)$.
+
+```{r, echo = FALSE}
+par(mar = c(5, 5, 1, 1));
+plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
+     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
+                             " (",rho,")")),
+     ylab = "Real part of leading eigenvalue", col = "grey",
+     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])));
+for(i in 1:dim(sim_rhos)[1]){
+    lines(x = c(sim_rhos[i,1], sim_rhos[i,15]), col = "grey60",
+          y = c(sim_rhos[i,20], sim_rhos[i,21]), lwd = 0.8);
+}
+points(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
+       col = "grey40");
+points(x = sim_rhos[,15], y = sim_rhos[,21], type = "b", pch = 20, lwd = 2,
+       col = "red");
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 2, col = "grey40");
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 2, col = "grey40");
+arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
+       y1 = sim_rhos[,24], angle = 90, length = 0.05, lwd = 2, col = "red");
+arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
+       y1 = sim_rhos[,25], angle = 90, length = 0.05, lwd = 2, col = "red");
+abline(h = 0, lwd = 0.8, lty = "dotted");
+legend("topleft", fill = c("grey", "red"), cex = 1.5,
+       legend = c(expression(gamma == 1), expression(paste("Var(",gamma,")"))));
+```
+
+Including $Var(\gamma)$ introduces a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$.
+Points along the x-axis are spaced more closely together given $Var(\gamma)$ because $Var(\gamma)$ tends to decrease the absolute magnitude of $\rho$.
+Grey and red points centred on $\rho = 0$ represent the same simulations, before (grey) and after (red) including $Var(\gamma)$.
+Grey and red points to the left and right show decreasing and increasing simulated $\rho$ values, respectively.
+Faint grey lines connect points for the same set of simulations, and where the red point is lower than the black point, the expected $\max(\Re(\lambda))$ was lower given $Var(\gamma)$.
+
+The region of $\rho$ values for $\textbf{A}$ that would result in increased stability given $Var(\gamma)$ is highlighted with red shading below.
+
+```{r, echo = FALSE}
+mbox <- function(x0, x1, y0, y1){
+    xx <- seq(from=x0, to=x1, length.out = 100);
+    yy <- seq(from=y0, to=y1, length.out = 100);
+    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
+    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
+    return(list(x=xd, y=yd));
+}
+more_stable_1 <- sim_rhos[,21] - sim_rhos[,20]
+more_stable_2 <- which(more_stable_1 < 0);
+min_x         <- min(more_stable_2);
+max_x         <- max(more_stable_2);
+shade_box     <- mbox(x0 = sim_rhos[min_x, 1], x1 = sim_rhos[max_x, 1], 
+                      y0 = -12, y1 = 15);
+par(mar = c(5, 5, 1, 1));
+plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
+     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
+                             " (",rho,")")),
+     ylab = "Real part of leading eigenvalue", col = "black",
+     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])));
+polygon(x = shade_box$x, y = shade_box$y, col = "red");
+points(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2);
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 2, col = "black");
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 2, col = "black");
+abline(h = 0, lwd = 0.8, lty = "dotted");
+```
+
+In this red shaded region above, $\max(\Re(\lambda))$ is decreased by $Var(\gamma)$.
+The shaded region encompasses values of $\rho$ between `r sim_rhos[min_x,1]` and `r sim_rhos[max_x,1]`.
+But `r dim(sim25)[1]` simulated $\textbf{A}$ had values that ranged between `r min(sim25[,3])` and `r max(sim25[,3])`, meaning that $\max(\Re(\lambda))$ was always expected to decrease given $Var(\gamma)$ even if $Var(\gamma)$ caused $\rho$ to increase.
+
+The curvature of the relationship between $\rho$ and $\max(\Re(\lambda))$ is consistent across different values of $S$, as shown below.

### 2020-01-11 notebook/revision_notes.tex
More notes

@@
 the difference in \(\rho\) between systems with versus without
-\(var(\gamma)\) for 100 stabilised systems; that is,
+\(var(\gamma)\) for 1000 stabilised systems; that is,
 \(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
@@
 
-For all cases in which \(\textbf{A}\) was stabilised by \(Var(\gamma)\),
-the resulting \(\textbf{M} = \gamma\textbf{A}\) had increased rather
-than decreased \(\rho\) (\(\rho_{var(\gamma)} > \rho_{\gamma = 1}\)).
-Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
-\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
-clearer picture of the effectof \(\gamma\), it is useful to show the
-relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above
-with, but this time with how the relationship changes given
-\(Var(\gamma)\).
+When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
+\(\rho\) was normally distributed around 0.001414 (95\% CIs: -0.0010192,
+0.0040992). Hence, decreasing the correlation between
+\(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself the cause
+of stability. For a clearer picture of the effect of \(\gamma\), it is
+useful to show the relationship between \(\rho\) and
+\(\max(\Re(\lambda))\) again as above with, but this time with how the
+relationship changes given \(Var(\gamma)\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+Including \(Var(\gamma)\) introduces a nonlinear relationship between
+\(\rho\) and \(\max(\Re(\lambda))\). Points along the x-axis are spaced
+more closely together given \(Var(\gamma)\) because \(Var(\gamma)\)
+tends to decrease the absolute magnitude of \(\rho\). Grey and red
+points centred on \(\rho = 0\) represent the same simulations, before
+(grey) and after (red) including \(Var(\gamma)\). Grey and red points to
+the left and right show decreasing and increasing simulated \(\rho\)
+values, respectively. Faint grey lines connect points for the same set
+of simulations, and where the red point is lower than the black point,
+the expected \(\max(\Re(\lambda))\) was lower given \(Var(\gamma)\).
+
+The region of \(\rho\) values for \(\textbf{A}\) that would result in
+increased stability given \(Var(\gamma)\) is highlighted with red
+shading below.
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-9-1.pdf}
+
+In this red shaded region above, \(\max(\Re(\lambda))\) is decreased by
+\(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
+between -0.4 and 0.8. But 1000 simulated \(\textbf{A}\) had values that
+ranged between -0.240986 and 0.0704592, meaning that
+\(\max(\Re(\lambda))\) was always expected to decrease given
+\(Var(\gamma)\) even if \(Var(\gamma)\) caused \(\rho\) to increase.
 

### 2020-01-13 notebook/revision_notes.Rmd
Work through notes

@@
 The curvature of the relationship between $\rho$ and $\max(\Re(\lambda))$ is consistent across different values of $S$, as shown below.
+Including $\gamma$ always results in a concave upward relationship between $\rho$ and $\max(\Re(\lambda))$.
+
+<!---
+
+```{r, echo = FALSE}
+sim10 <- read.csv(file = "sim_results/rhos/sim10.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sim15.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sim20.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sim30.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sim35.csv");
+par(mar = c(1, 1, 1, 1), mfrow = c(3, 2), oma = c(5, 5, 1, 1));
+plot(x = sim10[,1], y = sim10[,20], type = "b", pch = 20, lwd = 2, xaxt = "n");
+points(x = sim10[,15], y = sim10[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+plot(x = sim15[,1], y = sim15[,20], type = "b", pch = 20, lwd = 2, xaxt = "n");
+points(x = sim15[,15], y = sim15[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+plot(x = sim20[,1], y = sim20[,20], type = "b", pch = 20, lwd = 2, xaxt = "n");
+points(x = sim20[,15], y = sim20[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+plot(x = sim25[,1], y = sim25[,20], type = "b", pch = 20, lwd = 2, xaxt = "n");
+points(x = sim25[,15], y = sim25[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+plot(x = sim30[,1], y = sim30[,20], type = "b", pch = 20, lwd = 2);
+points(x = sim30[,15], y = sim30[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+plot(x = sim35[,1], y = sim35[,20], type = "b", pch = 20, lwd = 2);
+points(x = sim35[,15], y = sim35[,21], type = "b", lwd = 2);
+abline(h = 0, col = "red");
+mtext(text = "E correlation between A_ij and A_ji", side = 1,
+      line = 0, outer = TRUE, cex = 1.5);
+mtext(text = "E leading real part eigenvalue", side = 2,
+      line = 1, outer = TRUE, cex = 1.5);
+```
+
+--->

### 2020-01-13 notebook/revision_notes.tex
Work through notes

@@
 
+\hypertarget{role-of-correlated-matrices-in-stabilisation}{%
 \section{Role of correlated matrices in
-stabilisation}\label{role-of-correlated-matrices-in-stabilisation}
+stabilisation}\label{role-of-correlated-matrices-in-stabilisation}}
 
@@
 When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around 0.001414 (95\% CIs: -0.0010192,
-0.0040992). Hence, decreasing the correlation between
+\(\rho\) was normally distributed around 0.001414 (95\% CIs: -0.0011599,
+0.0040885). Hence, decreasing the correlation between
 \(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself the cause
@@
 \(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
-between -0.4 and 0.8. But 1000 simulated \(\textbf{A}\) had values that
+between -0.4 and 0.7. But 1000 simulated \(\textbf{A}\) had values that
 ranged between -0.240986 and 0.0704592, meaning that
@@
 
+The curvature of the relationship between \(\rho\) and
+\(\max(\Re(\lambda))\) is consistent across different values of \(S\),
+as shown below. Including \(\gamma\) always results in a concave upward
+relationship between \(\rho\) and \(\max(\Re(\lambda))\).
+
 \end{document}

### 2020-01-13 notebook/revision_notes.Rmd
New simulations and notes to show the curvature for the leading eigenvalue with rho

--- 2020-01-11 notebook/revision_notes.Rmd
@@
 rho_its  <- 1000;
-sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
-                         iters = rho_its);
+# sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
+#                         iters = rho_its);
+sim_rhos <- read.csv(file = "sim_results/rhos/sim_rhos.csv");
 ```
@@
 The curvature of the relationship between $\rho$ and $\max(\Re(\lambda))$ is consistent across different values of $S$, as shown below.
+Including $\gamma$ always results in a concave upward relationship between $\rho$ and $\max(\Re(\lambda))$.
+
+```{r, echo = FALSE, fig.height = 8, fig.width = 6}
+sim10 <- read.csv(file = "sim_results/rhos/sim10_rho.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sim15_rho.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sim20_rho.csv");
+sim25 <- read.csv(file = "sim_results/rhos/sim25_rho.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sim30_rho.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sim35_rho.csv");
+par(mar = c(0.25, 0.25, 0.25, 0.25), mfrow = c(3, 2), oma = c(6, 6, 1, 1));
+plot(x = sim10[,1], y = sim10[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim10[,15], y = sim10[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 10", cex = 2.5);
+plot(x = sim15[,1], y = sim15[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+points(x = sim15[,15], y = sim15[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 15", cex = 2.5);
+plot(x = sim20[,1], y = sim20[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim20[,15], y = sim20[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 20", cex = 2.5);
+plot(x = sim25[,1], y = sim25[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.25);
+points(x = sim25[,15], y = sim25[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 25", cex = 2.5);
+plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim30[,15], y = sim30[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 30", cex = 2.5);
+plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+points(x = sim35[,15], y = sim35[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 35", cex = 2.5);
+mtext(text = "E correlation between A_ij and A_ji", side = 1,
+      line = 4, outer = TRUE, cex = 1.5);
+mtext(text = "E leading real part eigenvalue", side = 2,
+      line = 3.5, outer = TRUE, cex = 1.5);
+```

### 2020-01-17 R/analytical.R
tweak R code

@@ -316,23 +316,82 @@ summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
     return(all_res);
 }
 
-sim10 <- rand_rho_var(S = 10, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
+sim10 <- rand_rho_var(S = 10, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
                       iters = 10000);
 
-sim15 <- rand_rho_var(S = 15, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
+sim15 <- rand_rho_var(S = 15, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
                       iters = 10000);
 
-sim20 <- rand_rho_var(S = 20, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
+sim20 <- rand_rho_var(S = 20, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
                       iters = 10000);
 
 sim25 <- rand_rho_var(S = 25, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
                       iters = 10000);
 
-sim30 <- rand_rho_var(S = 30, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
+sim30 <- rand_rho_var(S = 30, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
                       iters = 10000);
 
-sim35 <- rand_rho_var(S = 35, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
+sim35 <- rand_rho_var(S = 35, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
                       iters = 10000);
+
+write.csv(sim10, "sim_results/rhos/sim10.csv");
+write.csv(sim15, "sim_results/rhos/sim15.csv");
+write.csv(sim20, "sim_results/rhos/sim20.csv");
+write.csv(sim30, "sim_results/rhos/sim30.csv");
+write.csv(sim35, "sim_results/rhos/sim35.csv");
+
+
+
+sim10 <- read.csv(file = "sim_results/rhos/sim10.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sim15.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sim20.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sim30.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sim35.csv");
[…]

### 2020-01-17 R/random_mats.R

@@ -416,3 +416,15 @@ rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1,
 
 
 
+get_complexity <- function(mat){
+    S         <- length(diag(mat));
+    mat_gz    <- sum(mat != 0);
+    trace_gz  <- sum(diag(mat) != 0);
+    offdiagC  <- mat_gz - trace_gz;
+    offdiags  <- (dim(mat)[1] * dim(mat)[2]) - S;
+    calc_C    <- offdiagC / offdiags;
+    diag(mat) <- NA;
+    sigma     <- sd(mat, na.rm = TRUE);
+    complx    <- sigma * sqrt(S * calc_C);
+    return(complx);
+}

### 2020-01-17 notebook/revision_notes.Rmd
More work on the notes, getting into the analytical reasoning for rho gamma relationship

--- 2020-01-11 notebook/revision_notes.Rmd
@@
 As the correlation between matrix elements ($\rho$) decreases, the eigenvalue spectra changes such that more variation falls along the imaginary axis.
-The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which $\rho$ = -0.5.
-In both panels, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $0$ and $\sigma = 0.4$.
+The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which $\rho$ = -0.5 (right).
+In both panels, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $\mu = 0$ and standard deviation of $\sigma = 0.4$.
 
@@
 ```{r, echo = FALSE}
-rho_its  <- 1000;
-sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
-                         iters = rho_its);
+# rho_its  <- 1000;
+# sim_rhos <- rand_rho_var(S = 25, rhos = seq(from = -0.9, to = 0.9, by = 0.1), 
+#                         iters = rho_its);
+sim_rhos <- read.csv(file = "sim_results/rhos/sim_rhos.csv");
+rho_its  <- sim_rhos[1, 2] + sim_rhos[1, 3];
 ```
@@
 ```{r, echo = FALSE}
-sim25 <- read.csv(file = "sim_results/rhos/sim25.csv")
+sim_rho_distr <- read.csv(file = "sim_results/rhos/sim25.csv")
 ```
@@
 In the main text, I demonstrated that when $S$ is finite but system complexity $\sigma\sqrt{SC}$ is high ($C$ defines the connectance of $\textbf{A}$, or the proportion of non-zero off-diagonal elements), variation in component response rate $\gamma$ often underlies system stability.
-In other words, highly complex systems that are observed to be stable would not be if we removed the variation in their component response rates.
+In other words, highly complex systems that are observed to be stable would typically not be if we removed the variation in their component response rates.
 Mathematically, this means that by multiplying $\textbf{A}$ by a diagonal matrix $\gamma$ with variable elements, the sign of $\max(\Re(\lambda))$ is sometimes flipped from positive to negative in these finite systems of high complexity.
@@
 Note that for these parameter values, 1 million simulations found $\textbf{M} = \gamma \textbf{A}$ to be stable for `r dat[24, 3]` systems when all $\gamma_{i} = 1$, but `r dat[24, 5]` systems when $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for `r dim(sim25)[1]` stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.
+Below shows the distribution of the difference in $\rho$ between systems with versus without $var(\gamma)$ for `r dim(sim_rho_distr)[1]` stabilised systems; that is, $\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}$.
 
 ```{r, echo = FALSE}
-rho_change <- sim25[,4] - sim25[,3];
+rho_change <- sim_rho_distr[,8] - sim_rho_distr[,7];
 par(mar = c(5, 5, 1, 1));
-hist(x = rho_change, main = "", ylab = "Frequency", cex.lab = 1.25, breaks = 15,
+hist(x = rho_change, main = "", ylab = "Frequency", cex.lab = 1.25, breaks = 25,
      xlab = expression(paste(rho[change])), col = "grey", cex.axis = 1.25);
@@
 Hence, decreasing the correlation between $\textbf{A}_{ij}$ and $\textbf{A}_{ji}$ was not by itself the cause of stability.
-For a clearer picture of the effect of $\gamma$, it is useful to show the relationship between $\rho$ and $\max(\Re(\lambda))$ again as above with, but this time with how the relationship changes given $Var(\gamma)$.
+For a clearer picture of the effect of $\gamma$, it is useful to show the relationship between $\rho$ and $\max(\Re(\lambda))$ again as above, but this time also for how the relationship changes given $Var(\gamma)$.
+That is, given $\textbf{M} = \textbf{A}$ in addition to $\textbf{M} = \gamma\textbf{A}$.
 
@@
 The shaded region encompasses values of $\rho$ between `r sim_rhos[min_x,1]` and `r sim_rhos[max_x,1]`.
-But `r dim(sim25)[1]` simulated $\textbf{A}$ had values that ranged between `r min(sim25[,3])` and `r max(sim25[,3])`, meaning that $\max(\Re(\lambda))$ was always expected to decrease given $Var(\gamma)$ even if $Var(\gamma)$ caused $\rho$ to increase.
+But `r dim(sim_rho_distr)[1]` simulated $\textbf{A}$ had values that ranged between `r min(sim_rho_distr[,7])` and `r max(sim_rho_distr[7])`, meaning that $\max(\Re(\lambda))$ was always expected to decrease given $Var(\gamma)$ even if $Var(\gamma)$ caused $\rho$ to increase.
 
 The curvature of the relationship between $\rho$ and $\max(\Re(\lambda))$ is consistent across different values of $S$, as shown below.
+Including $\gamma$ always results in a concave upward relationship between $\rho$ and $\max(\Re(\lambda))$.
+
+```{r, echo = FALSE, fig.height = 8, fig.width = 6}
+sim10 <- read.csv(file = "sim_results/rhos/sim10_rho.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sim15_rho.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sim20_rho.csv");
+sim25 <- read.csv(file = "sim_results/rhos/sim25_rho.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sim30_rho.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sim35_rho.csv");
+par(mar = c(0.25, 0.25, 0.25, 0.25), mfrow = c(3, 2), oma = c(6, 6, 1, 1));
+plot(x = sim10[,1], y = sim10[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim10[,15], y = sim10[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 10", cex = 2.5);
+plot(x = sim15[,1], y = sim15[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+points(x = sim15[,15], y = sim15[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 15", cex = 2.5);
+plot(x = sim20[,1], y = sim20[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim20[,15], y = sim20[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 20", cex = 2.5);
+plot(x = sim25[,1], y = sim25[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.25);
+points(x = sim25[,15], y = sim25[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 25", cex = 2.5);
+plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim30[,15], y = sim30[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 30", cex = 2.5);
+plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+points(x = sim35[,15], y = sim35[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 35", cex = 2.5);
+mtext(text =  expression(paste("E Correlation between A"[ij]," and A"[ji], 
+                             " (",rho,")")), side = 1,
+      line = 4, outer = TRUE, cex = 1.5);
+mtext(text = "E leading real part eigenvalue", side = 2,
+      line = 3.5, outer = TRUE, cex = 1.5);
+```
+
+In all panels above, the solid line shows the relationship between the expected $\rho$ between $\textbf{A}$ elements $A_{ij}$ and $A_{ji}$ given no variation in component response rates (i.e., the diagonal matrix equals the identity matrix, $\gamma = \textbf{I}$, so $\textbf{M} = \textbf{A}$).
+The dotted line shows the same relationship given variation in component response rates (i.e., the diagonal matrix contains elements drawn from a random uniform distribution between 0 and 2, so $\textbf{M} = \gamma\textbf{A}$).
+
+Trying to think analytically
+================================================================================
+
+Trying to think analytically, note that the correlation $\rho$ adjusts the criteria for stability as follows [@Allesina2015a],
+
+$$\sigma\sqrt{SC}\left(1 + \rho\right) < 1.$$
+
+Given that diagonal values of $\textbf{M}$ are set to $-1$, We can know the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$.
+Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$.
+Hence, the expected value of $\max(\Re(\lambda)) = 0$
+
+```{r, echo = FALSE}
+mat_S16   <- build_rho_mat(S = 1600, sigma = (1/40), rho = 0);
+mat_S16e  <- eigen(mat_S16)$values;
+mat_S16r  <- Re(mat_S16e);
+mat_S16i  <- Im(mat_S16e);
+plot(x = mat_S16e, y = mat_S16i, pch = 4, cex = 0.7, col = "firebrick",
+     xlab = "Real", ylab = "Imaginary", cex.lab = 1.25, cex.axis = 1.25, 
+     asp = 1);
+```
+
+Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$.
+
+Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],
+
+$$\rho = \frac{E[M_{ij}M_{ji}] - E[M_{ij}]E[M_{ji}]}{Var[M_{ij}]}.$$
+
+We can rewrite as the more typical way of writing covariance (top),
+
+References
+================================================================================

### 2020-01-17 notebook/revision_notes.tex
New simulations and notes to show the curvature for the leading eigenvalue with rho
More work on the notes, getting into the analytical reasoning for rho gamma relationship

@@
 The figure panels below compare a random matrix in which \(\rho = 0\)
-(left) to one in which \(\rho\) = -0.5. In both panels, complex systems
-include \(S = 1000\) components, with diagonal elements of \(-1\) and
-off-diagonal elements drawn from a normal distribution with a mean of
-\(0\) and \(\sigma = 0.4\).
+(left) to one in which \(\rho\) = -0.5 (right). In both panels, complex
+systems include \(S = 1000\) components, with diagonal elements of
+\(-1\) and off-diagonal elements drawn from a normal distribution with a
+mean of \(\mu = 0\) and standard deviation of \(\sigma = 0.4\).
 
@@
 \(\rho\) and the expected value of the real part of the leading
-eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 1000 simulations
+eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 10000 simulations
 across values of \(\rho\) from -0.9 to 0.9 for complex systems with
@@
 stability. In other words, highly complex systems that are observed to
-be stable would not be if we removed the variation in their component
-response rates. Mathematically, this means that by multiplying
+be stable would typically not be if we removed the variation in their
+component response rates. Mathematically, this means that by multiplying
 \(\textbf{A}\) by a diagonal matrix \(\gamma\) with variable elements,
@@
 the difference in \(\rho\) between systems with versus without
-\(var(\gamma)\) for 1000 stabilised systems; that is,
+\(var(\gamma)\) for 1000000 stabilised systems; that is,
 \(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
@@
 When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around 0.001414 (95\% CIs: -0.0011599,
-0.0040885). Hence, decreasing the correlation between
-\(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself the cause
-of stability. For a clearer picture of the effect of \(\gamma\), it is
-useful to show the relationship between \(\rho\) and
-\(\max(\Re(\lambda))\) again as above with, but this time with how the
-relationship changes given \(Var(\gamma)\).
+\(\rho\) was normally distributed around
+\ensuremath{4.8394472\times 10^{-6}} (95\% CIs:
+\ensuremath{-7.3124139\times 10^{-5}},
+\ensuremath{8.7487968\times 10^{-5}}). Hence, decreasing the correlation
+between \(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself
+the cause of stability. For a clearer picture of the effect of
+\(\gamma\), it is useful to show the relationship between \(\rho\) and
+\(\max(\Re(\lambda))\) again as above, but this time also for how the
+relationship changes given \(Var(\gamma)\). That is, given
+\(\textbf{M} = \textbf{A}\) in addition to
+\(\textbf{M} = \gamma\textbf{A}\).
 
@@
 \(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
-between -0.4 and 0.7. But 1000 simulated \(\textbf{A}\) had values that
-ranged between -0.240986 and 0.0704592, meaning that
+between -0.4 and 0.7. But 1000000 simulated \(\textbf{A}\) had values
+that ranged between -0.2642846 and 0.2668648, meaning that
 \(\max(\Re(\lambda))\) was always expected to decrease given
@@
 
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+In all panels above, the solid line shows the relationship between the
+expected \(\rho\) between \(\textbf{A}\) elements \(A_{ij}\) and
+\(A_{ji}\) given no variation in component response rates (i.e., the
+diagonal matrix equals the identity matrix, \(\gamma = \textbf{I}\), so
+\(\textbf{M} = \textbf{A}\)). The dotted line shows the same
+relationship given variation in component response rates (i.e., the
+diagonal matrix contains elements drawn from a random uniform
+distribution between 0 and 2, so \(\textbf{M} = \gamma\textbf{A}\)).
+
+\hypertarget{trying-to-think-analytically}{%
+\section{Trying to think
+analytically}\label{trying-to-think-analytically}}
+
+Trying to think analytically, note that the correlation \(\rho\) adjusts
+the criteria for stability as
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
+
+\[\sigma\sqrt{SC}\left(1 + \rho\right) < 1.\]
+
+Given that diagonal values of \(\textbf{M}\) are set to \(-1\), We can
+know the relationship \(S = 1600\), \(C = 1\), and \(\sigma = 1/40\).
+Under these parameter values, given \(\rho = 0\), we know that
+\(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the expected value
+of \(\max(\Re(\lambda)) = 0\)
+
+\begin{verbatim}
+## Warning in xy.coords(x, y, xlabel, ylabel, log): imaginary parts discarded
+## in coercion
+\end{verbatim}
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-11-1.pdf}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Allesina2015a}{}%
+1. Allesina, S. \& Tang, S. The stability--complexity relationship at
+age 40: a random matrix perspective. \emph{Population Ecology} 63--75
+(2015).
+doi:\href{https://doi.org/10.1007/s10144-014-0471-0}{10.1007/s10144-014-0471-0}
+
 \end{document}

### 2020-01-18 R/analytical.R
Some more scratch work

@@ -825,3 +825,99 @@ points(x = fxeig5[,15], y = fxeig5[,21], type = "l", lwd = 2, lty = "dashed");
 
 
 
+fxeig6 <- rand_rho_var(S = 1600, iters = 10, int_type = 0, sigma = (1/40), C = 1,
+                       rhos = seq(from = -0.95, to = 0.95, by = 0.05));
+
+plot(x = fxeig6[,1], y = fxeig6[,20], type = "l", pch = 20, lwd = 2, 
+     cex.axis = 1.5, asp = 1);
+points(x = fxeig6[,15], y = fxeig6[,21], type = "l", lwd = 2, lty = "dashed");
+
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+
+mat_cov <- function(mat){
+    if(dim(mat)[1] != dim(mat)[2]){
+        stop("Error: Not a square matrix");
+    }
+    S  <- dim(mat)[1];
+    od <- 0.5 * (S * S - S);
+    ut <- rep(x = NA, times = od);
+    lt <- rep(x = NA, times = od);
+    ct <- 1;
+    for(i in 1:dim(mat)[1]){
+        for(j in 1:dim(mat)[2]){
+            if(i < j){
+                ut[ct] <- mat[i, j];
+                lt[ct] <- mat[j, i]; 
+                ct     <- ct + 1;
+            }
+        }
+    }
+    rho <- cov(ut, lt);
+    return(rho);
+}
+
+mat_cor <- function(mat){
+    if(dim(mat)[1] != dim(mat)[2]){
+        stop("Error: Not a square matrix");
+    }
+    S  <- dim(mat)[1];
+    od <- 0.5 * (S * S - S);
+    ut <- rep(x = NA, times = od);
+    lt <- rep(x = NA, times = od);
+    ct <- 1;
+    for(i in 1:dim(mat)[1]){
+        for(j in 1:dim(mat)[2]){
+            if(i < j){
+                ut[ct] <- mat[i, j];
+                lt[ct] <- mat[j, i]; 
+                ct     <- ct + 1;
+            }
+        }
+    }
+    rho <- cor(ut, lt);
+    return(rho);
+}
+
[…]

### 2020-01-18 R/scale_free.R
Build scale free networks

@@ -1,0 +1,45 @@
+# Build a scale free network
+find_sfree <- function(S){
+    min_vert <- 0;
+    while(min_vert == 0){
+        mat      <- sfree(S = S);
+        verts    <- apply(X = mat, MARGIN = 1, FUN = sum);
+        min_vert <- min(verts);
+    }
+    return(mat);
+}
+
+sfree <- function(S){
+    mat <- matrix(data = 0, nrow = S, ncol = S);
+    mat <- seed_mat(mat);
+    for(i in 3:S){
+        new_p <- get_p(mat = mat[1:i, 1:i]);
+        mat   <- add_p(mat = mat, new_p = new_p);
+    }
+    return(mat);
+}
+
+seed_mat <- function(mat){
+    mat[1, 2] <- 1;
+    mat[2, 1] <- 1;
+    return(mat);
+}
+
+# This function gets links for the new node
+get_p <- function(mat){
+    msize <- dim(mat)[1];
+    edges <- apply(X = mat, MARGIN = 1, FUN = sum);
+    prs   <- edges / sum(edges);
+    newe  <- rbinom(n = msize, size = 1, prob = prs);
+    return(newe);
+}
+
+add_p <- function(mat, new_p){
+    row             <- length(new_p);
+    mat[row, 1:row] <- new_p;
+    mat[1:row, row] <- new_p;
+    return(mat);
+}
+
+
+

### 2020-01-18 R/scale_free.R
Build scale free networks faster

@@ -1,12 +1,17 @@
 # Build a scale free network
 find_sfree <- function(S){
-    min_vert <- 0;
-    while(min_vert == 0){
-        mat      <- sfree(S = S);
+    tverts   <- 0;
+    while(tverts < S){
+        mat      <- sfree(S = 2*S);
         verts    <- apply(X = mat, MARGIN = 1, FUN = sum);
-        min_vert <- min(verts);
+        overts   <- which(verts > 0);
+        cmat     <- mat[overts, overts];
+        tverts   <- dim(cmat)[1];
+    }    
+    if(dim(cmat)[1] > S){
+        nmat <- cmat[1:S, 1:S];
     }
-    return(mat);
+    return(nmat);
 }
 
 sfree <- function(S){

### 2020-01-18 notebook/revision_notes.Rmd
Minor changes

@@
 
-Given that diagonal values of $\textbf{M}$ are set to $-1$, We can know the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$.
+Given that diagonal values of $\textbf{M}$ are set to $-1$, we can consider the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$.
 Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$.
@@
 Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$.
-
 Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],
@@
 
-We can rewrite as the more typical way of writing covariance (top),
+Since we define $\sigma = \sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity.
+Further, since we set $diag(\textbf{M}) = -1$, we know that $E[\max(\Re(\lambda))] = 0$.
+
+Gibbs et al.
+[@Gibbs2017] showed analytically that as $S \to \infty$, the effect of a diagonal matrix $\textbf{X}$ on the stability of $\textbf{M} = \textbf{XA}$ becomes negligible.
+Hence, assuming high $S$ and $\rho = 0$, we should not see an effect of $\gamma$ on $E[\max(\Re(\lambda))]$.
+But there is clearly an effect of $\gamma$ for nonzero $\rho$ such that $\partial E[\max(\Re(\lambda))] / \partial\rho > 0$ and $\partial^{2} E[\max(\Re(\lambda))] / \partial\rho > 0$.
+
+Note that, by definition, we have not changed $E[M_{ij}]$ with the addition of $\gamma$, since $E[\gamma_{ii}] = 1$.
+Elements of $\gamma$ are pulled from a random uniform distribution between 0 and 2, meaning that no covariance should be induced betwen $M_{ij}$ and $M_{ji}$.
 

### 2020-01-18 notebook/revision_notes.tex
Minor changes

@@
 When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around
-\ensuremath{4.8394472\times 10^{-6}} (95\% CIs:
-\ensuremath{-7.3124139\times 10^{-5}},
-\ensuremath{8.7487968\times 10^{-5}}). Hence, decreasing the correlation
-between \(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself
-the cause of stability. For a clearer picture of the effect of
-\(\gamma\), it is useful to show the relationship between \(\rho\) and
-\(\max(\Re(\lambda))\) again as above, but this time also for how the
-relationship changes given \(Var(\gamma)\). That is, given
-\(\textbf{M} = \textbf{A}\) in addition to
-\(\textbf{M} = \gamma\textbf{A}\).
+\(\rho\) was normally distributed around 4.8394472\times 10\^{}\{-6\}
+(95\% CIs: -7.0762643\times 10\^{}\{-5\}, 8.0991595\times 10\^{}\{-5\}).
+Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
+\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
+clearer picture of the effect of \(\gamma\), it is useful to show the
+relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
+but this time also for how the relationship changes given
+\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
+to \(\textbf{M} = \gamma\textbf{A}\).
 
@@
 
-\hypertarget{trying-to-think-analytically}{%
 \section{Trying to think
-analytically}\label{trying-to-think-analytically}}
+analytically}\label{trying-to-think-analytically}
 
@@
 
-Given that diagonal values of \(\textbf{M}\) are set to \(-1\), We can
-know the relationship \(S = 1600\), \(C = 1\), and \(\sigma = 1/40\).
-Under these parameter values, given \(\rho = 0\), we know that
-\(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the expected value
-of \(\max(\Re(\lambda)) = 0\)
+Given that diagonal values of \(\textbf{M}\) are set to \(-1\), we can
+consider the relationship \(S = 1600\), \(C = 1\), and
+\(\sigma = 1/40\). Under these parameter values, given \(\rho = 0\), we
+know that \(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the
+expected value of \(\max(\Re(\lambda)) = 0\)
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+Hence, under the same set of parameters, we predict a linear effect of
+\(\rho\) on \(\max(\Re(\lambda)) = 0\), such that
+\(\max(\Re(\lambda)) = \rho\).
+
+Allesina and Tang note that \(\rho\) is defined as
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
+
+\[\rho = \frac{E[M_{ij}M_{ji}] - E[M_{ij}]E[M_{ji}]}{Var[M_{ij}]}.\]
+
+We can rewrite as the more typical way of writing covariance (top),
+
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Allesina2015a}{}%
+\hypertarget{ref-Allesina2015a}{}
 1. Allesina, S. \& Tang, S. The stability--complexity relationship at

### 2020-01-18 notebook/revision_notes.Rmd
A bit more after re-reading Grilli et al.

@@
 
-Since we define $\sigma = \sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity.
+Since we define $\sigma = 1/\sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity.
 Further, since we set $diag(\textbf{M}) = -1$, we know that $E[\max(\Re(\lambda))] = 0$.
@@
 [@Gibbs2017] showed analytically that as $S \to \infty$, the effect of a diagonal matrix $\textbf{X}$ on the stability of $\textbf{M} = \textbf{XA}$ becomes negligible.
+They also note that local stability of a random matrix can be predicted from simple moments affecting three parameters,
+
+$$E_{1} = \frac{1}{S(1-S)}\sum_{i \neq j}A_{ij},$$
+
+$$E_{2} = \frac{1}{S(1-S)}\sum_{i \neq j}A^{2}_{ij} - E^{2}_{1},$$
+
+and,
+
+$$E_{c} = \frac{1}{S(1-S)E^{2}_{2}}\sum_{i \neq j} A_{ij} A_{ji} - \frac{E^{2}_{1}}{E^{2}_{2}}.$$
+
+Note that given $\mu = 0$ and $C = 1$, $E_{1} = 0$, while $E^{2}_{2} = Var(A_{ij})$, and $E_{c} = \rho$.
+This suggests that my parameterisations should only be affected by $Var(A_{ij})$ and $\rho$.
+This does indeed appear to be the case if we restrict $\sigma = 1/\sqrt{S}$, where if $\rho = 0$, random matrices with $\gamma$ have similar $E[\max(\Re(\lambda))]$ to those without.
+But the relationship between $\rho$ and $E[\max(\Re(\lambda))]$ clearly differs when $\gamma$ is included, and as $S$ increases such that $\sigma > 1/\sqrt{S}$.
+
 Hence, assuming high $S$ and $\rho = 0$, we should not see an effect of $\gamma$ on $E[\max(\Re(\lambda))]$.

### 2020-01-18 R/scale_free.R
Fix scale-free network

@@ -14,28 +14,31 @@ find_sfree <- function(S){
     return(nmat);
 }
 
-sfree <- function(S){
+# Need to seed with some m such that m can be a fixed node number
+sfree <- function(S, m){
     mat <- matrix(data = 0, nrow = S, ncol = S);
-    mat <- seed_mat(mat);
-    for(i in 3:S){
-        new_p <- get_p(mat = mat[1:i, 1:i]);
+    mat <- seed_mat(mat, m);
+    for(i in m:S){
+        new_p <- get_p(mat = mat[1:i, 1:i], m = m);
         mat   <- add_p(mat = mat, new_p = new_p);
     }
     return(mat);
 }
 
-seed_mat <- function(mat){
-    mat[1, 2] <- 1;
-    mat[2, 1] <- 1;
+seed_mat <- function(mat, m){
+    mat[1:m, 1:m]  <- 1;
+    diag(mat[1:m]) <- 0;
     return(mat);
 }
 
 # This function gets links for the new node
-get_p <- function(mat){
-    msize <- dim(mat)[1];
-    edges <- apply(X = mat, MARGIN = 1, FUN = sum);
-    prs   <- edges / sum(edges);
-    newe  <- rbinom(n = msize, size = 1, prob = prs);
+get_p <- function(mat, m){
+    msize    <- dim(mat)[1];
+    edges    <- apply(X = mat, MARGIN = 1, FUN = sum);
+    prs      <- edges / sum(edges);
+    ch       <- sample(x = 1:msize, size = m, prob = prs);
+    newe     <- rep(x = 0, times = msize);
+    newe[ch] <- 1;
     return(newe);
 }
 

### 2020-01-18 R/scale_free.R
No need for find function anymore

@@ -1,19 +1,3 @@
-# Build a scale free network
-find_sfree <- function(S){
-    tverts   <- 0;
-    while(tverts < S){
-        mat      <- sfree(S = 2*S);
-        verts    <- apply(X = mat, MARGIN = 1, FUN = sum);
-        overts   <- which(verts > 0);
-        cmat     <- mat[overts, overts];
-        tverts   <- dim(cmat)[1];
-    }    
-    if(dim(cmat)[1] > S){
-        nmat <- cmat[1:S, 1:S];
-    }
-    return(nmat);
-}
-
 # Need to seed with some m such that m can be a fixed node number
 sfree <- function(S, m){
     mat <- matrix(data = 0, nrow = S, ncol = S);
@@ -27,7 +11,7 @@ sfree <- function(S, m){
 
 seed_mat <- function(mat, m){
     mat[1:m, 1:m]  <- 1;
-    diag(mat[1:m]) <- 0;
+    diag(mat)[1:m] <- 0;
     return(mat);
 }
 

### 2020-01-19 notebook/revision_notes.Rmd
Edit to fix compile error

@@
 As the correlation between matrix elements ($\rho$) decreases, the eigenvalue spectra changes such that more variation falls along the imaginary axis.
-The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which $\rho$ = -0.5 (right).
+The figure panels below compare a random matrix in which $\rho = 0$ (left) to one in which $\rho = -0.5$ (right).
 In both panels, complex systems include $S = 1000$ components, with diagonal elements of $-1$ and off-diagonal elements drawn from a normal distribution with a mean of $\mu = 0$ and standard deviation of $\sigma = 0.4$.
@@
 
-When $\textbf{A}$ was stabilised by $Var(\gamma)$, the change in $\rho$ was normally distributed around `r mean(rho_change)` (95\% CIs: `r rho_change_CI[1]`, `r rho_change_CI[2]`).
+When $\textbf{A}$ was stabilised by $Var(\gamma)$, the change in $\rho$ was normally distributed around `r mean(rho_change) * 10^6`$\times 10^{-6}$ (95\% CIs: `r rho_change_CI[1] * 10^6`$\times 10^{-6}$, `r rho_change_CI[2] * 10^6`$\times 10^{-6}$).
 Hence, decreasing the correlation between $\textbf{A}_{ij}$ and $\textbf{A}_{ji}$ was not by itself the cause of stability.
@@
 mat_S16i  <- Im(mat_S16e);
-plot(x = mat_S16e, y = mat_S16i, pch = 4, cex = 0.7, col = "firebrick",
+plot(x = mat_S16r, y = mat_S16i, pch = 4, cex = 0.7, col = "firebrick",
      xlab = "Real", ylab = "Imaginary", cex.lab = 1.25, cex.axis = 1.25, 

### 2020-01-19 notebook/revision_notes.tex
Edit to fix compile error

@@
 The figure panels below compare a random matrix in which \(\rho = 0\)
-(left) to one in which \(\rho\) = -0.5 (right). In both panels, complex
+(left) to one in which \(\rho = -0.5\) (right). In both panels, complex
 systems include \(S = 1000\) components, with diagonal elements of
@@
 When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around 4.8394472\times 10\^{}\{-6\}
-(95\% CIs: -7.0762643\times 10\^{}\{-5\}, 8.0991595\times 10\^{}\{-5\}).
+\(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
+(95\% CIs: -73.6979659\(\times 10^{-6}\), 83.9237307\(\times 10^{-6}\)).
 Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
@@
 
-\begin{verbatim}
-## Warning in xy.coords(x, y, xlabel, ylabel, log): imaginary parts discarded
-## in coercion
-\end{verbatim}
-
 \includegraphics{revision_notes_files/figure-latex/unnamed-chunk-11-1.pdf}
@@
 \(\rho\) on \(\max(\Re(\lambda)) = 0\), such that
-\(\max(\Re(\lambda)) = \rho\).
-
-Allesina and Tang note that \(\rho\) is defined as
+\(\max(\Re(\lambda)) = \rho\). Allesina and Tang note that \(\rho\) is
+defined as
 follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
@@
 
-We can rewrite as the more typical way of writing covariance (top),
+Since we define \(\sigma = 1/\sqrt{S}\) and \(C = 1\),
+\(\sigma\sqrt{SC} = 1\), so the distribution of eigenvalues is circular
+with a radius of unity. Further, since we set \(diag(\textbf{M}) = -1\),
+we know that \(E[\max(\Re(\lambda))] = 0\).
+
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{2}}
+showed analytically that as \(S \to \infty\), the effect of a diagonal
+matrix \(\textbf{X}\) on the stability of \(\textbf{M} = \textbf{XA}\)
+becomes negligible. They also note that local stability of a random
+matrix can be predicted from simple moments affecting three parameters,
+
+\[E_{1} = \frac{1}{S(1-S)}\sum_{i \neq j}A_{ij},\]
+
+\[E_{2} = \frac{1}{S(1-S)}\sum_{i \neq j}A^{2}_{ij} - E^{2}_{1},\]
+
+and,
+
+\[E_{c} = \frac{1}{S(1-S)E^{2}_{2}}\sum_{i \neq j} A_{ij} A_{ji} - \frac{E^{2}_{1}}{E^{2}_{2}}.\]
+
+Note that given \(\mu = 0\) and \(C = 1\), \(E_{1} = 0\), while
+\(E^{2}_{2} = Var(A_{ij})\), and \(E_{c} = \rho\). This suggests that my
+parameterisations should only be affected by \(Var(A_{ij})\) and
+\(\rho\). This does indeed appear to be the case if we restrict
+\(\sigma = 1/\sqrt{S}\), where if \(\rho = 0\), random matrices with
+\(\gamma\) have similar \(E[\max(\Re(\lambda))]\) to those without. But
+the relationship between \(\rho\) and \(E[\max(\Re(\lambda))]\) clearly
+differs when \(\gamma\) is included, and as \(S\) increases such that
+\(\sigma > 1/\sqrt{S}\).
+
+Hence, assuming high \(S\) and \(\rho = 0\), we should not see an effect
+of \(\gamma\) on \(E[\max(\Re(\lambda))]\). But there is clearly an
+effect of \(\gamma\) for nonzero \(\rho\) such that
+\(\partial E[\max(\Re(\lambda))] / \partial\rho > 0\) and
+\(\partial^{2} E[\max(\Re(\lambda))] / \partial\rho > 0\).
+
+Note that, by definition, we have not changed \(E[M_{ij}]\) with the
+addition of \(\gamma\), since \(E[\gamma_{ii}] = 1\). Elements of
+\(\gamma\) are pulled from a random uniform distribution between 0 and
+2, meaning that no covariance should be induced betwen \(M_{ij}\) and
+\(M_{ji}\).
 
@@
 
+\hypertarget{ref-Gibbs2017}{}
+2. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+population abundances on the stability of large random ecosystems.
+\emph{Physical Review E - Statistical, Nonlinear, and Soft Matter
+Physics} \textbf{98,} 022410 (2018).
+
 \end{document}

### 2020-01-19 R/analytical.R
Small changes to better show rho effect

@@ -317,22 +317,28 @@ summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
 }
 
 sim10 <- rand_rho_var(S = 10, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
 
 sim15 <- rand_rho_var(S = 15, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
 
 sim20 <- rand_rho_var(S = 20, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
 
 sim25 <- rand_rho_var(S = 25, rhos = seq(from = -0.5, to = 0.5, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
 
 sim30 <- rand_rho_var(S = 30, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
 
 sim35 <- rand_rho_var(S = 35, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
-                      iters = 10000);
+                      iters = 10000, sigma = 0.2);
+
+sim40 <- rand_rho_var(S = 40, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
+
+sim45 <- rand_rho_var(S = 45, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
 
 write.csv(sim10, "sim_results/rhos/sim10.csv");
 write.csv(sim15, "sim_results/rhos/sim15.csv");
@@ -831,6 +837,14 @@ fxeig6 <- rand_rho_var(S = 1600, iters = 10, int_type = 0, sigma = (1/40), C = 1
 plot(x = fxeig6[,1], y = fxeig6[,20], type = "l", pch = 20, lwd = 2, 
      cex.axis = 1.5, asp = 1);
 points(x = fxeig6[,15], y = fxeig6[,21], type = "l", lwd = 2, lty = "dashed");
+
+
+
+
[…]

### 2020-01-20 notebook/revision_notes.Rmd
Figuring out variances

@@
 
+Effects of correlations interactions
+================================================================================
+
+If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\max(\Re(\lambda))$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
+These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+When investigating the effect of varying component response rate $Var(\gamma)$ on stability by defining $\textbf{M} = \gamma\textbf{A}$, $S$ remains unchanged.
+Further, values of $\gamma_{i}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
+What $Var(\gamma)$ does change is the variation in component interaction strengths, and $\rho$.
+
+Variation in $\gamma$ increases the total variation such that for $\textbf{M} = \gamma \textbf{A}$, $\sigma^{2}_{M_{ij}} = \sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}} \right)$.
+Eigenvalues of $\textbf{A}$ are therefore enclosed within a circle centred at $d$ with a radius of $\sigma_{A_{ij}}\sqrt{SC} = \sqrt{SC\sigma^{2}_{A_{ij}}}$, while eigenvalues of $\textbf{M}$ have a larger radius of $\sqrt{SC \sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}}\right)}$ (see blue versus red points below).
+
+Note that the correlation $\rho$ adjusts the criteria for stability as follows [@Allesina2015a],
+
+$$\sigma\sqrt{SC}\left(1 + \rho\right) < 1.$$
+
+Given that diagonal values of $\textbf{M}$ are set to $-1$, we can consider the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$.
+Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$.
+Hence, the expected value of $\max(\Re(\lambda)) = 0$
+
+```{r, echo = FALSE}
+mat_S16   <- build_rho_mat(S = 1600, sigma = (1/40), rho = 0);
+mat_S16e  <- eigen(mat_S16)$values;
+mat_S16r  <- Re(mat_S16e);
+mat_S16i  <- Im(mat_S16e);
+plot(x = mat_S16r, y = mat_S16i, pch = 4, cex = 0.7, col = "firebrick",
+     xlab = "Real", ylab = "Imaginary", cex.lab = 1.25, cex.axis = 1.25, 
+     asp = 1);
+```
+
+Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$.
+Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],
+
+$$\rho = \frac{E[M_{ij}M_{ji}] - E[M_{ij}]E[M_{ji}]}{Var[M_{ij}]}.$$
+
+Since we define $\sigma = 1/\sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity.
+Further, since we set $diag(\textbf{M}) = -1$, we know that $E[\max(\Re(\lambda))] = 0$.
+
+Gibbs et al.
+[@Gibbs2017] showed analytically that as $S \to \infty$, the effect of a diagonal matrix $\textbf{X}$ on the stability of $\textbf{M} = \textbf{XA}$ becomes negligible.
+They also note that local stability of a random matrix can be predicted from simple moments affecting three parameters,
+
+$$E_{1} = \frac{1}{S(1-S)}\sum_{i \neq j}A_{ij},$$
+
+$$E_{2} = \frac{1}{S(1-S)}\sum_{i \neq j}A^{2}_{ij} - E^{2}_{1},$$
+
+and,
+
+$$E_{c} = \frac{1}{S(1-S)E^{2}_{2}}\sum_{i \neq j} A_{ij} A_{ji} - \frac{E^{2}_{1}}{E^{2}_{2}}.$$
+
+Note that given $\mu = 0$ and $C = 1$, $E_{1} = 0$, while $E^{2}_{2} = Var(A_{ij})$, and $E_{c} = \rho$.
+This suggests that my parameterisations should only be affected by $Var(A_{ij})$ and $\rho$.
+This does indeed appear to be the case if we restrict $\sigma = 1/\sqrt{S}$, where if $\rho = 0$, random matrices with $\gamma$ have similar $E[\max(\Re(\lambda))]$ to those without.
+But the relationship between $\rho$ and $E[\max(\Re(\lambda))]$ clearly differs when $\gamma$ is included, and as $S$ increases such that $\sigma > 1/\sqrt{S}$.
+
+Hence, assuming high $S$ and $\rho = 0$, we should not see an effect of $\gamma$ on $E[\max(\Re(\lambda))]$.
+But there is clearly an effect of $\gamma$ for nonzero $\rho$ such that $\partial E[\max(\Re(\lambda))] / \partial\rho > 0$ and $\partial^{2} E[\max(\Re(\lambda))] / \partial\rho > 0$.
+
+Note that, by definition, we have not changed $E[M_{ij}]$ with the addition of $\gamma$, since $E[\gamma_{ii}] = 1$.
+Elements of $\gamma$ are pulled from a random uniform distribution between 0 and 2, meaning that no covariance should be induced betwen $M_{ij}$ and $M_{ji}$.
+
 Role of correlated matrices in stabilisation
@@
 source("../R/small_world.R");
-rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1, 
-                         by = 1, sigma = 0.4){
-    tot_res <- NULL;
-    fea_res <- NULL;
-    rho_res <- NULL;
-    cmplxty <- NULL;
-    for(i in 1:length(rhos)){
-        iter           <- iters;
-        tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 8);
-        fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
-        cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
-        while(iter > 0){
-            r_vec    <- rnorm(n = S, mean = 0, sd = rmx);
-            A0       <- build_rho_mat(S = S, sigma = sigma, rho = rhos[i]);
-            gam1     <- runif(n = S, min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0       <- A0 * mean(gam1);
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
-            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
-            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
-            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
-            A0_rho   <- mat_rho(A0);
-            A1_rho   <- mat_rho(A1);
-            rho_diff <- A1_rho - A0_rho;
-            rho_abs  <- abs(A1_rho) - abs(A0_rho);
-            if(A0_stb == TRUE){
-                tot_res[[i]][iter, 1] <- 1;
-            }
-            if(A1_stb == TRUE){
-                tot_res[[i]][iter, 2] <- 1;
-            }
-            if(A0_fea == TRUE){
-                fea_res[[i]][iter, 1] <- 1;
-            }
-            if(A1_fea == TRUE){
-                fea_res[[i]][iter, 2] <- 1;
-            }
-            rho_res[[i]][iter, 1] <- A0_rho;
-            rho_res[[i]][iter, 2] <- A1_rho;
-            rho_res[[i]][iter, 3] <- rho_diff;
-            rho_res[[i]][iter, 4] <- rho_abs;
-            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
-            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
-            cmplxty[[i]][iter, 1] <- get_complexity(A0);
-            cmplxty[[i]][iter, 2] <- get_complexity(A1);
-            iter                  <- iter - 1;
-        }
-    }
-    all_res     <- summarise_randmat(tot_res = tot_res, fea_res = fea_res,
-                                     rho_res = rho_res, cmplxty = cmplxty);
-    all_res[,1]          <- rhos;
-    colnames(all_res)[1] <- "rho_set";
-    return(all_res);
-}
-
-simpleboot <- function(freqs,repli=1000,alpha=0.05){    
-    vals  <- NULL;                                                          
-    i     <- 0;                                         
-    while(i < repli){                                                   
-        boot  <- sample(x=freqs,size=length(freqs),replace=TRUE);           
-        strap <- mean(boot);                                                
-        vals  <- c(vals,strap);                                             
-        i     <- i + 1;                                                     
-    }                                                                       
-    vals   <- sort(x=vals,decreasing=FALSE);                            
-    lowCI  <- vals[round((alpha*0.5)*repli)];                               
-    highCI <- vals[round((1-(alpha*0.5))*repli)];                           
-    CIs    <- c(lowCI,highCI);                                          
-    return(CIs);                                                            
-}
-
-summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
-    sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 25);
-    for(i in 1:sims){
-        all_res[i, 1]  <- i + 1;
-        # Stable and unstable
-        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
-        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
-        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
-        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
-        # Stabilised and destabilised
-        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
-                                 tot_res[[i]][,2] == TRUE);
-        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
-                                 tot_res[[i]][,2] == FALSE);
-        # Feasible and infeasible
-        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
-        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
-        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
-        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
-        # Feased and defeased
-        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
-                                  fea_res[[i]][,2] == TRUE);
-        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
-                                  fea_res[[i]][,2] == FALSE);
-        # Mean correlations between A[i,j] and A[j,i] off-diag elements
-        all_res[i, 14] <- mean(rho_res[[i]][,1]);
-        all_res[i, 15] <- mean(rho_res[[i]][,2]);
-        all_res[i, 16] <- mean(rho_res[[i]][,3]);
-        all_res[i, 17] <- mean(rho_res[[i]][,4]);
-        all_res[i, 18] <- mean(cmplxty[[i]][,1]);
-        all_res[i, 19] <- mean(cmplxty[[i]][,2]);
-        all_res[i, 20] <- mean(rho_res[[i]][,5]);
-        all_res[i, 21] <- mean(rho_res[[i]][,6]);
-        all_res[i, 22:23] <- range(rho_res[[i]][,5]);
-        all_res[i, 24:25] <- range(rho_res[[i]][,6]);
-    }
-    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
-                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
-                "A0_feasible", "A1_infeasible", "A1_feasible", 
-                "A1_made_feasible", "A1_made_infeasible", "A0_rho", "A1_rho",
-                "rho_diff", "rho_abs", "complex_A0", "complex_A1", "A0_eig",
-                "A1_eig", "LCI_A0", "UCI_A0", "LCI_A1", "UCI_A1");
-    colnames(all_res) <- cnames;
-    return(all_res);
-}
+source("../R/scale_free.R");
+source("../R/utilities.R");
 ```
@@
 
-Trying to think analytically
-================================================================================
-
-Trying to think analytically, note that the correlation $\rho$ adjusts the criteria for stability as follows [@Allesina2015a],
-
-$$\sigma\sqrt{SC}\left(1 + \rho\right) < 1.$$
-
-Given that diagonal values of $\textbf{M}$ are set to $-1$, we can consider the relationship $S = 1600$, $C = 1$, and $\sigma = 1/40$.
-Under these parameter values, given $\rho = 0$, we know that $\sigma\sqrt{SC}\left(1 + \rho\right) = 1$.
-Hence, the expected value of $\max(\Re(\lambda)) = 0$
-
-```{r, echo = FALSE}
-mat_S16   <- build_rho_mat(S = 1600, sigma = (1/40), rho = 0);
-mat_S16e  <- eigen(mat_S16)$values;
-mat_S16r  <- Re(mat_S16e);
-mat_S16i  <- Im(mat_S16e);
-plot(x = mat_S16r, y = mat_S16i, pch = 4, cex = 0.7, col = "firebrick",
-     xlab = "Real", ylab = "Imaginary", cex.lab = 1.25, cex.axis = 1.25, 
-     asp = 1);
-```
-
-Hence, under the same set of parameters, we predict a linear effect of $\rho$ on $\max(\Re(\lambda)) = 0$, such that $\max(\Re(\lambda)) = \rho$.
-Allesina and Tang note that $\rho$ is defined as follows [@Allesina2015a],
-
-$$\rho = \frac{E[M_{ij}M_{ji}] - E[M_{ij}]E[M_{ji}]}{Var[M_{ij}]}.$$
-
-Since we define $\sigma = 1/\sqrt{S}$ and $C = 1$, $\sigma\sqrt{SC} = 1$, so the distribution of eigenvalues is circular with a radius of unity.
-Further, since we set $diag(\textbf{M}) = -1$, we know that $E[\max(\Re(\lambda))] = 0$.
-
-Gibbs et al.
-[@Gibbs2017] showed analytically that as $S \to \infty$, the effect of a diagonal matrix $\textbf{X}$ on the stability of $\textbf{M} = \textbf{XA}$ becomes negligible.
-They also note that local stability of a random matrix can be predicted from simple moments affecting three parameters,
-
-$$E_{1} = \frac{1}{S(1-S)}\sum_{i \neq j}A_{ij},$$
-
-$$E_{2} = \frac{1}{S(1-S)}\sum_{i \neq j}A^{2}_{ij} - E^{2}_{1},$$
-
-and,
-
-$$E_{c} = \frac{1}{S(1-S)E^{2}_{2}}\sum_{i \neq j} A_{ij} A_{ji} - \frac{E^{2}_{1}}{E^{2}_{2}}.$$
-
-Note that given $\mu = 0$ and $C = 1$, $E_{1} = 0$, while $E^{2}_{2} = Var(A_{ij})$, and $E_{c} = \rho$.
-This suggests that my parameterisations should only be affected by $Var(A_{ij})$ and $\rho$.
-This does indeed appear to be the case if we restrict $\sigma = 1/\sqrt{S}$, where if $\rho = 0$, random matrices with $\gamma$ have similar $E[\max(\Re(\lambda))]$ to those without.
-But the relationship between $\rho$ and $E[\max(\Re(\lambda))]$ clearly differs when $\gamma$ is included, and as $S$ increases such that $\sigma > 1/\sqrt{S}$.
-
-Hence, assuming high $S$ and $\rho = 0$, we should not see an effect of $\gamma$ on $E[\max(\Re(\lambda))]$.
-But there is clearly an effect of $\gamma$ for nonzero $\rho$ such that $\partial E[\max(\Re(\lambda))] / \partial\rho > 0$ and $\partial^{2} E[\max(\Re(\lambda))] / \partial\rho > 0$.
-
-Note that, by definition, we have not changed $E[M_{ij}]$ with the addition of $\gamma$, since $E[\gamma_{ii}] = 1$.
-Elements of $\gamma$ are pulled from a random uniform distribution between 0 and 2, meaning that no covariance should be induced betwen $M_{ij}$ and $M_{ji}$.
-
 References

### 2020-01-20 notebook/revision_notes.tex
Figuring out variances

--- 2020-01-17 notebook/revision_notes.tex
@@
 
-\hypertarget{role-of-correlated-matrices-in-stabilisation}{%
-\section{Role of correlated matrices in
-stabilisation}\label{role-of-correlated-matrices-in-stabilisation}}
+\hypertarget{effects-of-correlations-interactions}{%
+\section{Effects of correlations
+interactions}\label{effects-of-correlations-interactions}}
 
-In complex systems represented by large random matrices, correlation
-between matrix elements \(A_{ij}\) and \(A_{ji}\) affects the
-distribution of eigenvalues and therefore local stability. As the
-correlation between matrix elements (\(\rho\)) decreases, the eigenvalue
-spectra changes such that more variation falls along the imaginary axis.
-The figure panels below compare a random matrix in which \(\rho = 0\)
-(left) to one in which \(\rho\) = -0.5 (right). In both panels, complex
-systems include \(S = 1000\) components, with diagonal elements of
-\(-1\) and off-diagonal elements drawn from a normal distribution with a
-mean of \(\mu = 0\) and standard deviation of \(\sigma = 0.4\).
+If the off-diagonal elements of \(\textbf{A}\) are sampled independently
+from an identical distribution, then \(\max(\Re(\lambda))\) for
+\(\textbf{M} = \textbf{A}\) can be estimated from five
+values\textsuperscript{\protect\hyperlink{ref-Tang2014b}{1}}. These
+values include (1) system size (\(S\)), (2) mean self-regulation of
+components (\(d\)), (3) mean interaction strength between components
+(\(\mu\)), (4) the standard deviation between component interaction
+strengths (\(\sigma\)), and (5) the correlation of interaction strengths
+between components, \(M_{ij}\) and \(M_{ji}\) (\(\rho\)). When
+investigating the effect of varying component response rate
+\(Var(\gamma)\) on stability by defining
+\(\textbf{M} = \gamma\textbf{A}\), \(S\) remains unchanged. Further,
+values of \(\gamma_{i}\) were sampled such that \(E[d]\) and \(E[\mu]\)
+also remained unchanged (in practice, diagonal elements of
+\(\textbf{M}\) were standardised so that mean values were identical
+before and after adding \(\gamma\)). What \(Var(\gamma)\) does change is
+the variation in component interaction strengths, and \(\rho\).
 
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-2-1.pdf}
+Variation in \(\gamma\) increases the total variation such that for
+\(\textbf{M} = \gamma \textbf{A}\),
+\(\sigma^{2}_{M_{ij}} = \sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}} \right)\).
+Eigenvalues of \(\textbf{S}\) are therefore enclosed within a circle
+centred at \(d\) with a radius of
+\(\sigma_{A_{ij}}\sqrt{SC} = \sqrt{SC\sigma^{2}_{A_{ij}}}\), while
+eigenvalues of \(\textbf{A}\) have a larger radius of
+\(\sqrt{SC\sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}\right)}}\)
 
-Because of this effect of \(\rho\) on the eigenvalue spectra, decreasing
-values of \(\rho\) will also tend to decrease the rightmost eigenvalue
-of the matrix \(\textbf{A}\). This makes it more likely that the complex
-system represented by \(\textbf{A}\) is locally stable, as stability
-occurs when all real parts of eigenvalues are negative. Note that this
-elongation along the imaginary axis is also characteristic of
-predator-prey communities (in which, by definition \(A_{ij}\) and
-\(A_{ji}\) have opposing signs). Also note that as \(\rho\) increases
-such that \(\rho > 0\), the same elongation happens along the real axis,
-making random complex systems less likely to be stable.
-
-A simple numerical analysis illustrates the linear relationship between
-\(\rho\) and the expected value of the real part of the leading
-eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 10000 simulations
-across values of \(\rho\) from -0.9 to 0.9 for complex systems with
-\(S = 25\) components. Error bars show 95\% bootstrapped confidence
-intervals.
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-4-1.pdf}
-
-In the main text, I demonstrated that when \(S\) is finite but system
-complexity \(\sigma\sqrt{SC}\) is high (\(C\) defines the connectance of
-\(\textbf{A}\), or the proportion of non-zero off-diagonal elements),
-variation in component response rate \(\gamma\) often underlies system
-stability. In other words, highly complex systems that are observed to
-be stable would typically not be if we removed the variation in their
-component response rates. Mathematically, this means that by multiplying
-\(\textbf{A}\) by a diagonal matrix \(\gamma\) with variable elements,
-the sign of \(\max(\Re(\lambda))\) is sometimes flipped from positive to
-negative in these finite systems of high complexity. Interestingly, this
-increase in stability given \(Var(\gamma) > 0\) is not necessarily
-caused by \(\gamma\) decreasing \(\rho\). In fact, for \(S = 25\),
-\(\sigma = 0.4\), and \(C = 1\), random complex systems that are
-stabilised by \(\gamma\) typically have increased \(\rho\) values. Note
-that for these parameter values, 1 million simulations found
-\(\textbf{M} = \gamma \textbf{A}\) to be stable for 36 systems when all
-\(\gamma_{i} = 1\), but 383 systems when
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Below shows the distribution of
-the difference in \(\rho\) between systems with versus without
-\(var(\gamma)\) for 1000000 stabilised systems; that is,
-\(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-7-1.pdf}
-
-When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around
-\ensuremath{4.8394472\times 10^{-6}} (95\% CIs:
-\ensuremath{-7.3124139\times 10^{-5}},
-\ensuremath{8.7487968\times 10^{-5}}). Hence, decreasing the correlation
-between \(\textbf{A}_{ij}\) and \(\textbf{A}_{ji}\) was not by itself
-the cause of stability. For a clearer picture of the effect of
-\(\gamma\), it is useful to show the relationship between \(\rho\) and
-\(\max(\Re(\lambda))\) again as above, but this time also for how the
-relationship changes given \(Var(\gamma)\). That is, given
-\(\textbf{M} = \textbf{A}\) in addition to
-\(\textbf{M} = \gamma\textbf{A}\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-Including \(Var(\gamma)\) introduces a nonlinear relationship between
-\(\rho\) and \(\max(\Re(\lambda))\). Points along the x-axis are spaced
-more closely together given \(Var(\gamma)\) because \(Var(\gamma)\)
-tends to decrease the absolute magnitude of \(\rho\). Grey and red
-points centred on \(\rho = 0\) represent the same simulations, before
-(grey) and after (red) including \(Var(\gamma)\). Grey and red points to
-the left and right show decreasing and increasing simulated \(\rho\)
-values, respectively. Faint grey lines connect points for the same set
-of simulations, and where the red point is lower than the black point,
-the expected \(\max(\Re(\lambda))\) was lower given \(Var(\gamma)\).
-
-The region of \(\rho\) values for \(\textbf{A}\) that would result in
-increased stability given \(Var(\gamma)\) is highlighted with red
-shading below.
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-In this red shaded region above, \(\max(\Re(\lambda))\) is decreased by
-\(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
-between -0.4 and 0.7. But 1000000 simulated \(\textbf{A}\) had values
-that ranged between -0.2642846 and 0.2668648, meaning that
-\(\max(\Re(\lambda))\) was always expected to decrease given
-\(Var(\gamma)\) even if \(Var(\gamma)\) caused \(\rho\) to increase.
-
-The curvature of the relationship between \(\rho\) and
-\(\max(\Re(\lambda))\) is consistent across different values of \(S\),
-as shown below. Including \(\gamma\) always results in a concave upward
-relationship between \(\rho\) and \(\max(\Re(\lambda))\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-In all panels above, the solid line shows the relationship between the
-expected \(\rho\) between \(\textbf{A}\) elements \(A_{ij}\) and
-\(A_{ji}\) given no variation in component response rates (i.e., the
-diagonal matrix equals the identity matrix, \(\gamma = \textbf{I}\), so
-\(\textbf{M} = \textbf{A}\)). The dotted line shows the same
-relationship given variation in component response rates (i.e., the
-diagonal matrix contains elements drawn from a random uniform
-distribution between 0 and 2, so \(\textbf{M} = \gamma\textbf{A}\)).
-
-\hypertarget{trying-to-think-analytically}{%
-\section{Trying to think
-analytically}\label{trying-to-think-analytically}}
-
-Trying to think analytically, note that the correlation \(\rho\) adjusts
-the criteria for stability as
-follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
-
-\[\sigma\sqrt{SC}\left(1 + \rho\right) < 1.\]
-
-Given that diagonal values of \(\textbf{M}\) are set to \(-1\), We can
-know the relationship \(S = 1600\), \(C = 1\), and \(\sigma = 1/40\).
-Under these parameter values, given \(\rho = 0\), we know that
-\(\sigma\sqrt{SC}\left(1 + \rho\right) = 1\). Hence, the expected value
-of \(\max(\Re(\lambda)) = 0\)
-
-\begin{verbatim}
-## Warning in xy.coords(x, y, xlabel, ylabel, log): imaginary parts discarded
-## in coercion
-\end{verbatim}
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-11-1.pdf}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+Note that the correlation \(\rho\) adjusts the criteria for stability as
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{2}},
 
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-Tang2014b}{}%
+1. Tang, S. \& Allesina, S. Reactivity and stability of large
+ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
+(2014).
+
 \leavevmode\hypertarget{ref-Allesina2015a}{}%
-1. Allesina, S. \& Tang, S. The stability--complexity relationship at
+2. Allesina, S. \& Tang, S. The stability--complexity relationship at
 age 40: a random matrix perspective. \emph{Population Ecology} 63--75

### 2020-01-20 notebook/analytical.R

@@ -343,11 +343,14 @@ sim45 <- rand_rho_var(S = 45, rhos = seq(from = -0.9, to = 0.9, by = 0.05),
 write.csv(sim10, "../notebook/sim_results/rhos/sigma0pt2/sim10.csv");
 write.csv(sim15, "../notebook/sim_results/rhos/sigma0pt2/sim15.csv");
 write.csv(sim20, "../notebook/sim_results/rhos/sigma0pt2/sim20.csv");
-write.csv(sim30, "../notebook/sim_results/rhos/sigma0pt2/sim25.csv");
-write.csv(sim35, "../notebook/sim_results/rhos/sigma0pt2/sim30.csv");
-write.csv(sim30, "../notebook/sim_results/rhos/sigma0pt2/sim35.csv");
-write.csv(sim35, "../notebook/sim_results/rhos/sigma0pt2/sim40.csv");
-write.csv(sim30, "../notebook/sim_results/rhos/sigma0pt2/sim45.csv");
+write.csv(sim25, "../notebook/sim_results/rhos/sigma0pt2/sim25.csv");
+
+
+write.csv(sim30, "notebook/sim_results/rhos/sigma0pt2/sim30.csv");
+write.csv(sim35, "notebook/sim_results/rhos/sigma0pt2/sim35.csv");
+write.csv(sim40, "notebook/sim_results/rhos/sigma0pt2/sim40.csv");
+write.csv(sim45, "notebook/sim_results/rhos/sigma0pt2/sim45.csv");
+
 
 
 sim10 <- read.csv(file = "sim_results/rhos/sim10.csv");

### 2020-01-20 R/utilities.R
Add a small function to find variances

@@ -154,3 +154,23 @@ add_C_stats <- function(sim){
     colnames(new_all_res)[dim(new_all_res)[2]] <- "C";
     return(new_all_res);
 }
+
+
+
+#' Get variance of interaction strengths
+#' 
+#' Returns the variance of interaction stengths (sigma^2).
+#'
+#'@return The variance of off-diagonal elements of 'mat'
+#'@param mat The matrix to be valuated
+#'@examples
+#'eg_mat  <- matrix(dat = rnorm(n = 16), nrow = 4);
+#'mat_cmp <- get_sigma_sqd(mat = eg_mat);
+#'@export
+get_sigma_sqd <- function(mat){
+    diag(mat) <- 0;
+    mat_offs  <- mat[mat != 0];
+    var_offs  <- var(mat_offs);
+    return(var_offs);
+}
+

### 2020-01-22 notebook/revision_notes.Rmd
Working through stuff again

@@
 
-Variation in $\gamma$ increases the total variation such that for $\textbf{M} = \gamma \textbf{A}$, $\sigma^{2}_{M_{ij}} = \sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}} \right)$.
-Eigenvalues of $\textbf{A}$ are therefore enclosed within a circle centred at $d$ with a radius of $\sigma_{A_{ij}}\sqrt{SC} = \sqrt{SC\sigma^{2}_{A_{ij}}}$, while eigenvalues of $\textbf{M}$ have a larger radius of $\sqrt{SC \sigma^{2}_{A_{ij}} \left(1 + \sigma^{2}_{\gamma_{i}}\right)}$ (see blue versus red points below).
+Variation in $\gamma$ ($\sigma^{2}_{\gamma}$) **can increase or decrease the radius of the circle**.
+
+<!--- still wrong below -- no simple analytical solution seems possible --->
+
+$$2 \sigma_{\gamma} \sigma_{\textbf{A}} \sqrt{SC}.$$
+
+The figure below shows the relationship between $\sigma_{\gamma}$ and the radius of the circle in which eigenvalue spectra are contained given $S = 1024$, $C = 1$, and $\sigma_{\textbf{A}} = 1/\sqrt{S}$.
+Values of $\gamma$ were drawn from a uniform distribution $\gamma \sim \mathcal{U}(0, b)$ at increasing values of $b$ so that $\sigma_{\gamma} = (1/12)b^2$ gradually increased.
+
+For poisson, it's $$\sigma^{2}_{\gamma} \sigma_{\textbf{A}} \sqrt{SC}.$$
+
+The mean of a uniform is $(1/2)b$, while the mean of the Poisson is $\lambda = \sigma^{2}_{\gamma}$
+
+Note that the error around the expected radius of eigenvalues is not normally distributed. It is log distributed (check).
 

### 2020-01-22 notebook/analytical.R

@@ -995,26 +995,95 @@ cor.test(xx, yy);
 
 
 
+get_varchange <- function(S, bs){
+  sigma   <- 1/(1*sqrt(S));
+  a       <- 0;
+  res_mat <- NULL;
+  for(b in bs){
+      A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
+      A_mat  <- matrix(data = A_dat, nrow = S);
+      C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+      C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+      A_mat  <- A_mat * C_mat;
+      gammas <- rpois(n = S, lambda = b); # runif(n = S, min = a, max = b);
+      vargam <- b; # (1/12) * (b-a)^2;
+      diag(A_mat) <- 0;
+      A1     <- gammas * A_mat;
+      A0     <- A_mat;
+      A0_e   <- eigen(A0)$values;
+      A0_r   <- Re(A0_e);
+      A0_i   <- Im(A0_e);
+      A1_e   <- eigen(A1)$values;
+      A1_r   <- Re(A1_e);
+      A1_i   <- Im(A1_e);
+      A_test <- A0;
+      diag(A_test) <- NA;
+      A0_var <-var(as.vector(A_test), na.rm = TRUE);
+      A_test <- A1;
+      diag(A_test) <- NA;
+      A1_var <-var(as.vector(A_test), na.rm = TRUE);
+      hackradii <- (range(A1_r)[2] - range(A1_r)[1]) / 2
+      keepit <- c(sigma, sigma^2, b, vargam, A0_var, A1_var, hackradii);
+      res_mat <- rbind(res_mat, keepit);
+      print(b);
+  }
+  colnames(res_mat) <- c("sigma", "sigma^2", "b", "vargam", "A0_var", "A1_var",
+                         "hackradii");
+  return(res_mat);
+}
+
+
+dat <- get_varchange(S = 1024, bs = seq(from = 0.1, to = 48, by = 1));
+
+
+plot(x = dat[,4], dat[,7], xlab = "Variation of gamma", 
+     ylab = "Radius of eigenvalue circle");
+var_form <- 2 * SCsigma * sqrt(dat[,4]);
[…]

### 2020-01-23 notebook/revision_notes.Rmd
Some progress, finally get the eigenvalue spectra variance correct

@@
 
+```{r, echo = FALSE}
+source("../R/utilities.R");
+source("../R/random_mats.R");
+```
+
 If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\max(\Re(\lambda))$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
@@
 
-Variation in $\gamma$ ($\sigma^{2}_{\gamma}$) **can increase or decrease the radius of the circle**.
+Introducing variation in $\gamma$ increases the total variation in the system, making it more likely that $\textbf{M} = \gamma\textbf{A}$ is unstable.
+Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables (to simplify the notation, $\sigma^{2}_{M}$ and $\sigma^{2}_{A}$ refer to variances of off-diagonal elements only),
 
-<!--- still wrong below -- no simple analytical solution seems possible --->
+$$\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.$$
 
-$$2 \sigma_{\gamma} \sigma_{\textbf{A}} \sqrt{SC}.$$
+Note that in my simulations $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, so the above can be simplified,
+
+$$\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).$$
+
+The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$.
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{M}$ and $\textbf{A}$ lie within a circle of a radius $\sigma^{2}_{M}\sqrt{S}$ and $\sigma^{2}_{A}\sqrt{S}$, respectively.
+
+```{r, fig.width = 7, fig.height = 7, echo = FALSE}
+S     <- 400;
+C     <- 1;
+dval  <- 0;
+sigma <- 1/(1*sqrt(S));
+a     <- 0;
+b     <- 2;
+mn    <- 0;
+
+A_dat  <- rnorm(n = S * S, mean = mn, sd = sigma);
+A_mat  <- matrix(data = A_dat, nrow = S);
+C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+A_mat  <- A_mat * C_mat;
+gammas <- runif(n = S, min = a, max = b);
+vargam <- (1/12) * (b-a)^2;
+
+diag(A_mat) <- dval;
+A1     <- gammas * A_mat;
+A0     <- A_mat;
+A0     <- A0 * mean(gammas);
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+joint_var <- ((sigma^2) * vargam) + ((sigma^2) * mean(gammas)^2) + 
+             (vargam * (dval*mn)^2);
+
+par(mfrow = c(2, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));
+plot(A0_r, A0_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "dodgerblue4", xlim = c(-3, 1.5),
+     xaxt = "n", ylim = c(-1.5, 1.5));
+vl        <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x1a     <- mean(gammas) * sigma * sqrt(S) * cos(vl) + mean(diag(A0));
+A0y1a     <- mean(gammas) * sigma * sqrt(S) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "dodgerblue4");
+text(x = -2.75, y = 1.75, labels = "a", cex = 2);
+
+plot(A1_r, A1_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "firebrick", xlim = c(-3, 1.5),
+     xaxt = "n", yaxt = "n", ylim = c(-1.5, 1.5));
+crad      <- sqrt(joint_var) * sqrt(S);
+A1x1a     <- crad * cos(vl) + mean(diag(A1));
+A1y1a     <- crad * sin(vl);
+points(x = A1x1a, y = A1y1a, type = "l", lwd = 3, col = "firebrick");
+text(x = -2.75, y = 1.75, labels = "b", cex = 2);
+
+calc_var <- A1;
+diag(calc_var) <- NA;
+calc_var  <- as.numeric(calc_var);
+cvar1 <- var(calc_var[!is.na(calc_var)]);
+
+################################################################################
+################################################################################
+
+S     <- 400;
+C     <- 1;
+dval  <- -1;
+sigma <- 1/(1*sqrt(S));
+a     <- 0;
+b     <- 2;
+mn    <- 0;
+
+A_dat  <- rnorm(n = S * S, mean = mn, sd = sigma);
+A_mat  <- matrix(data = A_dat, nrow = S);
+C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+A_mat  <- A_mat * C_mat;
+gammas <- runif(n = S, min = a, max = b);
+vargam <- (1/12) * (b-a)^2;
+
+diag(A_mat) <- dval;
+A1     <- gammas * A_mat;
+A0     <- A_mat;
+A0     <- A0 * mean(gammas);
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+joint_var <- ((sigma^2) * vargam) + ((sigma^2) * mean(gammas)^2) + 
+             (vargam * (dval*mn)^2);
+
+plot(A0_r, A0_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "dodgerblue4", ylim = c(-1.5, 1.5),
+     xlim = c(-3, 1.5));
+A0x1a     <- mean(gammas) * sigma * sqrt(S) * cos(vl) + mean(diag(A0));
+A0y1a     <- mean(gammas) * sigma * sqrt(S) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "dodgerblue4");
+text(x = -2.75, y = 1.75, labels = "c", cex = 2);
+
+plot(A1_r, A1_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "firebrick", yaxt = "n", 
+     xlim = c(-3, 1.5), ylim = c(-1.5, 1.5));
+text(x = -2.75, y = 1.75, labels = "d", cex = 2);
+
+calc_var <- A1;
+diag(calc_var) <- NA;
+calc_var  <- as.numeric(calc_var);
+cvar2 <- var(calc_var[!is.na(calc_var)]);
+
+mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
+mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+```
+
+The effect of $\gamma$ does not necessarily result in a circular eigenvalue spectra of $\textbf{M}$.
+In both panels of Figure 1, $S =$ `r S` and $\sigma_{A} = 1/\sqrt{S}$.
+In Figures 1a,b and 1c,d, $d = 0$ and $d = 1$, respectively, causing the eigenvalue spectra of $\textbf{A}$ to differ on the real axis (Figure a,c).
+Figure panels b and c show eigenvalue spectra after $\gamma$ is included such that $M = \gamma A$, where elements of $\gamma$ are drawn from a uniform distribution $\gamma \sim \mathcal{U}(0, 2)$.
+In both cases, $\sigma^{2}_{A} =$ `r round(sigma, digits = 4)` and $\sigma^{2}_{\gamma} =$ `r round(joint_var, digits = 4)`.
+Nevertheless, the change in eigenvalue spectra caused by $\gamma$ can affect $\max(\Re(\lambda))$ and therefore system stability.
+
+Stability also decreases with increasing $\rho$ such that $\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)$.
+Hence, when $\rho$ values are negative, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability is expected to increase [@Allesina2012].
+When elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the covariance between $M_{ij}$ and $M_{ji}$,
+
+$$\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.$$ 
+Because $\gamma$ does not affect $cov(M_{ij}, M_{ji})$, the sign of $\rho$ remains unchanged.
+
+More down
+======================
 
@@
 Values of $\gamma$ were drawn from a uniform distribution $\gamma \sim \mathcal{U}(0, b)$ at increasing values of $b$ so that $\sigma_{\gamma} = (1/12)b^2$ gradually increased.
-
-For poisson, it's $$\sigma^{2}_{\gamma} \sigma_{\textbf{A}} \sqrt{SC}.$$
-
-The mean of a uniform is $(1/2)b$, while the mean of the Poisson is $\lambda = \sigma^{2}_{\gamma}$
-
-Note that the error around the expected radius of eigenvalues is not normally distributed. It is log distributed (check).
 

### 2020-01-23 notebook/revision_notes.tex
Some progress, finally get the eigenvalue spectra variance correct

--- 2020-01-19 notebook/revision_notes.tex
@@
 
-\section{Role of correlated matrices in
-stabilisation}\label{role-of-correlated-matrices-in-stabilisation}
+\section{Effects of correlations
+interactions}\label{effects-of-correlations-interactions}
 
-In complex systems represented by large random matrices, correlation
-between matrix elements \(A_{ij}\) and \(A_{ji}\) affects the
-distribution of eigenvalues and therefore local stability. As the
-correlation between matrix elements (\(\rho\)) decreases, the eigenvalue
-spectra changes such that more variation falls along the imaginary axis.
-The figure panels below compare a random matrix in which \(\rho = 0\)
-(left) to one in which \(\rho = -0.5\) (right). In both panels, complex
-systems include \(S = 1000\) components, with diagonal elements of
-\(-1\) and off-diagonal elements drawn from a normal distribution with a
-mean of \(\mu = 0\) and standard deviation of \(\sigma = 0.4\).
+If the off-diagonal elements of \(\textbf{A}\) are sampled independently
+from an identical distribution, then \(\max(\Re(\lambda))\) for
+\(\textbf{M} = \textbf{A}\) can be estimated from five
+values\textsuperscript{\protect\hyperlink{ref-Tang2014b}{1}}. These
+values include (1) system size (\(S\)), (2) mean self-regulation of
+components (\(d\)), (3) mean interaction strength between components
+(\(\mu\)), (4) the standard deviation between component interaction
+strengths (\(\sigma\)), and (5) the correlation of interaction strengths
+between components, \(M_{ij}\) and \(M_{ji}\) (\(\rho\)). When
+investigating the effect of varying component response rate
+\(Var(\gamma)\) on stability by defining
+\(\textbf{M} = \gamma\textbf{A}\), \(S\) remains unchanged. Further,
+values of \(\gamma_{i}\) were sampled such that \(E[d]\) and \(E[\mu]\)
+also remained unchanged (in practice, diagonal elements of
+\(\textbf{M}\) were standardised so that mean values were identical
+before and after adding \(\gamma\)). What \(Var(\gamma)\) does change is
+the variation in component interaction strengths, and \(\rho\).
+
+Introducing variation in \(\gamma\) increases the total variation in the
+system, making it more likely that \(\textbf{M} = \gamma\textbf{A}\) is
+unstable. Variation of the off-diagonal elements in \(\textbf{M}\) is
+described by the joint variation of two random variables (to simplify
+the notation, \(\sigma^{2}_{M}\) and \(\sigma^{2}_{A}\) refer to
+variances of off-diagonal elements only),
+
+\[\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.\]
+
+Note that in my simulations \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\),
+so the above can be simplified,
+
+\[\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).\]
+
+The increase caused by \(\sigma^{2}_\gamma\) can be visualised from the
+eigenvalue spectra of \(\textbf{A}\) versus
+\(\textbf{M} = \gamma\textbf{A}\). Given \(d = 0\) and \(C = 1\), the
+distribution of eigenvalues of \(\textbf{M}\) and \(\textbf{A}\) lie
+within a circle of a radius \(\sigma^{2}_{M}\sqrt{S}\) and
+\(\sigma^{2}_{A}\sqrt{S}\), respectively.
 
@@
 
-Because of this effect of \(\rho\) on the eigenvalue spectra, decreasing
-values of \(\rho\) will also tend to decrease the rightmost eigenvalue
-of the matrix \(\textbf{A}\). This makes it more likely that the complex
-system represented by \(\textbf{A}\) is locally stable, as stability
-occurs when all real parts of eigenvalues are negative. Note that this
-elongation along the imaginary axis is also characteristic of
-predator-prey communities (in which, by definition \(A_{ij}\) and
-\(A_{ji}\) have opposing signs). Also note that as \(\rho\) increases
-such that \(\rho > 0\), the same elongation happens along the real axis,
-making random complex systems less likely to be stable.
+The effect of \(\gamma\) does not necessarily result in a circular
+eigenvalue spectra of \(\textbf{M}\). In both panels of Figure 1,
+\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a and 1b,
+\(d = 0\) and \(d = 1\), respectively, causing the eigen spectra of
+\(\textbf{A}\) to differ on the real axis (blue points). Red points she
+eigen spectra after \(\gamma\) is include such that \(M = \gamma A\),
+where elements of \(\gamma\) are drawn from a uniform distribution
+\(\gamma \sim \mathcal{U}(0, 2)\). In both cases, \(\sigma^{2}_{A} =\)
+0.05 and \(\sigma^{2}_{\gamma} =\) 0.0033.
 
-A simple numerical analysis illustrates the linear relationship between
-\(\rho\) and the expected value of the real part of the leading
-eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 10000 simulations
-across values of \(\rho\) from -0.9 to 0.9 for complex systems with
-\(S = 25\) components. Error bars show 95\% bootstrapped confidence
-intervals.
+\section{More down}\label{more-down}
 
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-4-1.pdf}
+The figure below shows the relationship between \(\sigma_{\gamma}\) and
+the radius of the circle in which eigenvalue spectra are contained given
+\(S = 1024\), \(C = 1\), and \(\sigma_{\textbf{A}} = 1/\sqrt{S}\).
+Values of \(\gamma\) were drawn from a uniform distribution
+\(\gamma \sim \mathcal{U}(0, b)\) at increasing values of \(b\) so that
+\(\sigma_{\gamma} = (1/12)b^2\) gradually increased.
 
-In the main text, I demonstrated that when \(S\) is finite but system
-complexity \(\sigma\sqrt{SC}\) is high (\(C\) defines the connectance of
-\(\textbf{A}\), or the proportion of non-zero off-diagonal elements),
-variation in component response rate \(\gamma\) often underlies system
-stability. In other words, highly complex systems that are observed to
-be stable would typically not be if we removed the variation in their
-component response rates. Mathematically, this means that by multiplying
-\(\textbf{A}\) by a diagonal matrix \(\gamma\) with variable elements,
-the sign of \(\max(\Re(\lambda))\) is sometimes flipped from positive to
-negative in these finite systems of high complexity. Interestingly, this
-increase in stability given \(Var(\gamma) > 0\) is not necessarily
-caused by \(\gamma\) decreasing \(\rho\). In fact, for \(S = 25\),
-\(\sigma = 0.4\), and \(C = 1\), random complex systems that are
-stabilised by \(\gamma\) typically have increased \(\rho\) values. Note
-that for these parameter values, 1 million simulations found
-\(\textbf{M} = \gamma \textbf{A}\) to be stable for 36 systems when all
-\(\gamma_{i} = 1\), but 383 systems when
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Below shows the distribution of
-the difference in \(\rho\) between systems with versus without
-\(var(\gamma)\) for 1000000 stabilised systems; that is,
-\(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-7-1.pdf}
-
-When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
-\(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
-(95\% CIs: -73.6979659\(\times 10^{-6}\), 83.9237307\(\times 10^{-6}\)).
-Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
-\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
-clearer picture of the effect of \(\gamma\), it is useful to show the
-relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
-but this time also for how the relationship changes given
-\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
-to \(\textbf{M} = \gamma\textbf{A}\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-Including \(Var(\gamma)\) introduces a nonlinear relationship between
-\(\rho\) and \(\max(\Re(\lambda))\). Points along the x-axis are spaced
-more closely together given \(Var(\gamma)\) because \(Var(\gamma)\)
-tends to decrease the absolute magnitude of \(\rho\). Grey and red
-points centred on \(\rho = 0\) represent the same simulations, before
-(grey) and after (red) including \(Var(\gamma)\). Grey and red points to
-the left and right show decreasing and increasing simulated \(\rho\)
-values, respectively. Faint grey lines connect points for the same set
-of simulations, and where the red point is lower than the black point,
-the expected \(\max(\Re(\lambda))\) was lower given \(Var(\gamma)\).
-
-The region of \(\rho\) values for \(\textbf{A}\) that would result in
-increased stability given \(Var(\gamma)\) is highlighted with red
-shading below.
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-9-1.pdf}
-
-In this red shaded region above, \(\max(\Re(\lambda))\) is decreased by
-\(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
-between -0.4 and 0.7. But 1000000 simulated \(\textbf{A}\) had values
-that ranged between -0.2642846 and 0.2668648, meaning that
-\(\max(\Re(\lambda))\) was always expected to decrease given
-\(Var(\gamma)\) even if \(Var(\gamma)\) caused \(\rho\) to increase.
-
-The curvature of the relationship between \(\rho\) and
-\(\max(\Re(\lambda))\) is consistent across different values of \(S\),
-as shown below. Including \(\gamma\) always results in a concave upward
-relationship between \(\rho\) and \(\max(\Re(\lambda))\).
-
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-10-1.pdf}
-
-In all panels above, the solid line shows the relationship between the
-expected \(\rho\) between \(\textbf{A}\) elements \(A_{ij}\) and
-\(A_{ji}\) given no variation in component response rates (i.e., the
-diagonal matrix equals the identity matrix, \(\gamma = \textbf{I}\), so
-\(\textbf{M} = \textbf{A}\)). The dotted line shows the same
-relationship given variation in component response rates (i.e., the
-diagonal matrix contains elements drawn from a random uniform
-distribution between 0 and 2, so \(\textbf{M} = \gamma\textbf{A}\)).
-
-\section{Trying to think
-analytically}\label{trying-to-think-analytically}
-
-Trying to think analytically, note that the correlation \(\rho\) adjusts
-the criteria for stability as
-follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
+Note that the correlation \(\rho\) adjusts the criteria for stability as
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{2}},
 
@@
 
-\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-3-1.pdf}
 
@@
 defined as
-follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{1}},
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{2}},
 
@@
 
-Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{2}}
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{3}}
 showed analytically that as \(S \to \infty\), the effect of a diagonal
@@
 
+\section{Role of correlated matrices in
+stabilisation}\label{role-of-correlated-matrices-in-stabilisation}
+
+In complex systems represented by large random matrices, correlation
+between matrix elements \(A_{ij}\) and \(A_{ji}\) affects the
+distribution of eigenvalues and therefore local stability. As the
+correlation between matrix elements (\(\rho\)) decreases, the eigenvalue
+spectra changes such that more variation falls along the imaginary axis.
+The figure panels below compare a random matrix in which \(\rho = 0\)
+(left) to one in which \(\rho = -0.5\) (right). In both panels, complex
+systems include \(S = 1000\) components, with diagonal elements of
+\(-1\) and off-diagonal elements drawn from a normal distribution with a
+mean of \(\mu = 0\) and standard deviation of \(\sigma = 0.4\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-5-1.pdf}
+
+Because of this effect of \(\rho\) on the eigenvalue spectra, decreasing
+values of \(\rho\) will also tend to decrease the rightmost eigenvalue
+of the matrix \(\textbf{A}\). This makes it more likely that the complex
+system represented by \(\textbf{A}\) is locally stable, as stability
+occurs when all real parts of eigenvalues are negative. Note that this
+elongation along the imaginary axis is also characteristic of
+predator-prey communities (in which, by definition \(A_{ij}\) and
+\(A_{ji}\) have opposing signs). Also note that as \(\rho\) increases
+such that \(\rho > 0\), the same elongation happens along the real axis,
+making random complex systems less likely to be stable.
+
+A simple numerical analysis illustrates the linear relationship between
+\(\rho\) and the expected value of the real part of the leading
+eigenvalue, \(\max(\Re(\lambda))\). Below, I have run 10000 simulations
+across values of \(\rho\) from -0.9 to 0.9 for complex systems with
+\(S = 25\) components. Error bars show 95\% bootstrapped confidence
+intervals.
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+In the main text, I demonstrated that when \(S\) is finite but system
+complexity \(\sigma\sqrt{SC}\) is high (\(C\) defines the connectance of
+\(\textbf{A}\), or the proportion of non-zero off-diagonal elements),
+variation in component response rate \(\gamma\) often underlies system
+stability. In other words, highly complex systems that are observed to
+be stable would typically not be if we removed the variation in their
+component response rates. Mathematically, this means that by multiplying
+\(\textbf{A}\) by a diagonal matrix \(\gamma\) with variable elements,
+the sign of \(\max(\Re(\lambda))\) is sometimes flipped from positive to
+negative in these finite systems of high complexity. Interestingly, this
+increase in stability given \(Var(\gamma) > 0\) is not necessarily
+caused by \(\gamma\) decreasing \(\rho\). In fact, for \(S = 25\),
+\(\sigma = 0.4\), and \(C = 1\), random complex systems that are
+stabilised by \(\gamma\) typically have increased \(\rho\) values. Note
+that for these parameter values, 1 million simulations found
+\(\textbf{M} = \gamma \textbf{A}\) to be stable for 36 systems when all
+\(\gamma_{i} = 1\), but 383 systems when
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Below shows the distribution of
+the difference in \(\rho\) between systems with versus without
+\(var(\gamma)\) for 1000000 stabilised systems; that is,
+\(\rho_{change} = \rho_{var(\gamma)} - \rho_{\gamma = 1}\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+When \(\textbf{A}\) was stabilised by \(Var(\gamma)\), the change in
+\(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
+(95\% CIs: -74.3396367\(\times 10^{-6}\), 77.6796914\(\times 10^{-6}\)).
+Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
+\(\textbf{A}_{ji}\) was not by itself the cause of stability. For a
+clearer picture of the effect of \(\gamma\), it is useful to show the
+relationship between \(\rho\) and \(\max(\Re(\lambda))\) again as above,
+but this time also for how the relationship changes given
+\(Var(\gamma)\). That is, given \(\textbf{M} = \textbf{A}\) in addition
+to \(\textbf{M} = \gamma\textbf{A}\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-11-1.pdf}
+
+Including \(Var(\gamma)\) introduces a nonlinear relationship between
+\(\rho\) and \(\max(\Re(\lambda))\). Points along the x-axis are spaced
+more closely together given \(Var(\gamma)\) because \(Var(\gamma)\)
+tends to decrease the absolute magnitude of \(\rho\). Grey and red
+points centred on \(\rho = 0\) represent the same simulations, before
+(grey) and after (red) including \(Var(\gamma)\). Grey and red points to
+the left and right show decreasing and increasing simulated \(\rho\)
+values, respectively. Faint grey lines connect points for the same set
+of simulations, and where the red point is lower than the black point,
+the expected \(\max(\Re(\lambda))\) was lower given \(Var(\gamma)\).
+
+The region of \(\rho\) values for \(\textbf{A}\) that would result in
+increased stability given \(Var(\gamma)\) is highlighted with red
+shading below.
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+In this red shaded region above, \(\max(\Re(\lambda))\) is decreased by
+\(Var(\gamma)\). The shaded region encompasses values of \(\rho\)
+between -0.4 and 0.7. But 1000000 simulated \(\textbf{A}\) had values
+that ranged between -0.2642846 and 0.2668648, meaning that
+\(\max(\Re(\lambda))\) was always expected to decrease given
+\(Var(\gamma)\) even if \(Var(\gamma)\) caused \(\rho\) to increase.
+
+The curvature of the relationship between \(\rho\) and
+\(\max(\Re(\lambda))\) is consistent across different values of \(S\),
+as shown below. Including \(\gamma\) always results in a concave upward
+relationship between \(\rho\) and \(\max(\Re(\lambda))\).
+
+\includegraphics{revision_notes_files/figure-latex/unnamed-chunk-13-1.pdf}
+
+In all panels above, the solid line shows the relationship between the
+expected \(\rho\) between \(\textbf{A}\) elements \(A_{ij}\) and
+\(A_{ji}\) given no variation in component response rates (i.e., the
+diagonal matrix equals the identity matrix, \(\gamma = \textbf{I}\), so
+\(\textbf{M} = \textbf{A}\)). The dotted line shows the same
+relationship given variation in component response rates (i.e., the
+diagonal matrix contains elements drawn from a random uniform
+distribution between 0 and 2, so \(\textbf{M} = \gamma\textbf{A}\)).
+
 \section*{References}\label{references}
@@
 \hypertarget{refs}{}
+\hypertarget{ref-Tang2014b}{}
+1. Tang, S. \& Allesina, S. Reactivity and stability of large
+ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
+(2014).
+
 \hypertarget{ref-Allesina2015a}{}
-1. Allesina, S. \& Tang, S. The stability--complexity relationship at
+2. Allesina, S. \& Tang, S. The stability--complexity relationship at
 age 40: a random matrix perspective. \emph{Population Ecology} 63--75
@@
 \hypertarget{ref-Gibbs2017}{}
-2. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+3. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.

### 2020-01-23 R/utilities.R
Small change

@@ -174,3 +174,19 @@ get_sigma_sqd <- function(mat){
     return(var_offs);
 }
 
+
+simpleboot <- function(freqs , repli = 1000, alpha = 0.05){    
+    vals  <- NULL;                                                          
+    i     <- 0;                                         
+    while(i < repli){                                                   
+        boot  <- sample(x = freqs, size = length(freqs), replace = TRUE);           
+        strap <- mean(boot);                                                
+        vals  <- c(vals,strap);                                             
+        i     <- i + 1;                                                     
+    }                                                                       
+    vals   <- sort(x = vals, decreasing = FALSE);                            
+    lowCI  <- vals[round((alpha*0.5)*repli)];                               
+    highCI <- vals[round((1-(alpha*0.5))*repli)];                           
+    CIs    <- c(lowCI,highCI);                                          
+    return(CIs);                                                            
+}  

### 2020-01-23 notebook/analytical.R
Analytical now more scratch paper

@@ -839,11 +839,14 @@ points(x = fxeig5[,15], y = fxeig5[,21], type = "l", lwd = 2, lty = "dashed");
 fxeig6 <- rand_rho_var(S = 1600, iters = 10, int_type = 0, sigma = (1/40), C = 1,
                        rhos = seq(from = -0.95, to = 0.95, by = 0.05));
 
+
+fxeig6 <- read.csv("../fxeig6.csv");
+
 plot(x = fxeig6[,1], y = fxeig6[,20], type = "l", pch = 20, lwd = 2, 
      cex.axis = 1.5, asp = 1);
 points(x = fxeig6[,15], y = fxeig6[,21], type = "l", lwd = 2, lty = "dashed");
 
-
+points(x = 0, y = 0, cex = 2, col = "red");
 
 
 points(x = fxeig6[,15], y = fxeig6[,24], type = "l", col = "blue", 
@@ -1062,18 +1065,35 @@ plot(x = dat[,4], tt, xlab = "Variation of gamma",
 
 
 
-
-
-
-# Is it really just a radius of the minimum 'd'?
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
 S     <- 2500;
 C     <- 1;
 dval  <- 0;
 sigma <- 1/(1*sqrt(S));
 a     <- 0;
-b     <- 8;
+b     <- 2;
+mn    <- 0;
 
-A_dat  <- rnorm(n = S * S, mean = 0, sd = sigma);
+A_dat  <- rnorm(n = S * S, mean = mn, sd = sigma);
[…]

### 2020-01-23 temp_rho_sim.R
Add some temporary files

@@ -1,0 +1,17 @@
+
+
+sim30 <- rand_rho_var(S = 30, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
+write.csv(sim30, "notebook/sim_results/rhos/sigma0pt2/sim30.csv");
+
+sim35 <- rand_rho_var(S = 35, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
+write.csv(sim35, "notebook/sim_results/rhos/sigma0pt2/sim35.csv");
+
+sim40 <- rand_rho_var(S = 40, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
+write.csv(sim40, "notebook/sim_results/rhos/sigma0pt2/sim40.csv");
+
+sim45 <- rand_rho_var(S = 45, rhos = seq(from = -0.9, to = 0.9, by = 0.05), 
+                      iters = 10000, sigma = 0.2);
+write.csv(sim45, "notebook/sim_results/rhos/sigma0pt2/sim45.csv");

### 2020-01-24 R/random_mats.R
Get an updated summarise_randmat function

@@ -147,7 +147,7 @@ species_interactions <- function(mat, type = 0){
 #'@export
 summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
     sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 19);
+    all_res <- matrix(data = 0, nrow = sims, ncol = 25);
     for(i in 1:sims){
         all_res[i, 1]  <- i + 1;
         # Stable and unstable
@@ -177,12 +177,17 @@ summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
         all_res[i, 17] <- mean(rho_res[[i]][,4]);
         all_res[i, 18] <- mean(cmplxty[[i]][,1]);
         all_res[i, 19] <- mean(cmplxty[[i]][,2]);
+        all_res[i, 20] <- mean(rho_res[[i]][,5]);
+        all_res[i, 21] <- mean(rho_res[[i]][,6]);
+        all_res[i, 22:23] <- range(rho_res[[i]][,5]);
+        all_res[i, 24:25] <- range(rho_res[[i]][,6]);
     }
     cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
                 "A1_stabilised", "A1_destabilised", "A0_infeasible", 
                 "A0_feasible", "A1_infeasible", "A1_feasible", 
                 "A1_made_feasible", "A1_made_infeasible", "A0_rho", "A1_rho",
-                "rho_diff", "rho_abs", "complex_A0", "complex_A1");
+                "rho_diff", "rho_abs", "complex_A0", "complex_A1", "A0_eig",
+                "A1_eig", "LCI_A0", "UCI_A0", "LCI_A1", "UCI_A1");
     colnames(all_res) <- cnames;
     return(all_res);
 }

### 2020-01-24 notebook/sci_rep_reviews.Rmd
Respond to reviewer 1's comments

@@
 
-Reviewer 1 comments
+Reviewer 1 comment 1
 ================================================================================
@@
 
+Author response to Reviewer 1 comment 1
+================================================================================
+
+I thank the reviewer for these comments.
+I share their belief that the issues in this paper are of current concern in ecology, and that the ideas presented in this manuscript are, to the best of my knowledge, novel.
+
+Reviewer 1 comment 2
+================================================================================
+
 Recent works dealing with similar models in the large S limit have been published recently (Gibbs et al (2018) and Ahmadian (2018).
@@
 2, 376–384.
+
+Author response to Reviewer 1 comment 2
+================================================================================
+
+I again thank the reviewer for their comments.
+I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (line XXX).
+
+Reviewer 1 comment 3
+================================================================================
 
@@
 
+Author response to Reviewer 1 comment 3
+================================================================================
+
+I agree with the relevance of Dougoud et al.
+(2018), and have made appropriate links to their results on feasibility.
+Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility.
+Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic systems in which all population densities are positive).
+Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application.
+
+I welcome Reviewer 1's call for expanding the scope of models considered for simulations, which was also recommended by Reviewer 2.
+Nevertheless, my manuscript focuses broadly on complex systems rather than pure ecological systems, so I have revised my manuscript to test more general models of common network structures.
+In terms of ecological models, in addition to random competition, mutualism, and predator-prey networks, I have now also included a cascade food web model (which is a foundation for the extended niche and nested-heirarchy models that Reviewer 1 mentions; Williams and Martinez 2000).
+Other network types that I have incorporated into my revision include small-world networks (Watts and Strogatz 1998) and scale-free networks (Albert and Barab\'{a}si 2002).
+Overall, the structured networks that I now include in the main text and supporting information now encompass a wide range of physical, biological, and social systems.
+
+More emphasis on feasibility and how it relates to ecological systems has now been included (line XXX).
+I have also added new sections in the Results (line XXX) and methods (line XXX) for the new simulations of structured networks.
+Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
+
+Reviewer 1 comment 4
+================================================================================
+
 Comment:
@@
 unstable given that DA-id is stable ? I think that this point should be explained in more details.
+
+Author response to Reviewer 1 comment 3
+================================================================================
+
+Reviewer 1 is correct that the line illustrates the conditional probability that $A$ is unstable given that $\gamma A$ is stable.
+In other words, the proportion of complex systems found to be stable, but which would have been unstable if it were not for the variation in component response rates.
+
+I have explained the result shown by the black line of Figures 3 and 4 now more clearly in my revision (line XXX).
 

### 2020-01-25 notebook/revision_notes.Rmd
More to the argument for a numerical approach, why an analytical one probably won't work

@@
 In both panels of Figure 1, $S =$ `r S` and $\sigma_{A} = 1/\sqrt{S}$.
-In Figures 1a,b and 1c,d, $d = 0$ and $d = 1$, respectively, causing the eigenvalue spectra of $\textbf{A}$ to differ on the real axis (Figure a,c).
-Figure panels b and c show eigenvalue spectra after $\gamma$ is included such that $M = \gamma A$, where elements of $\gamma$ are drawn from a uniform distribution $\gamma \sim \mathcal{U}(0, 2)$.
+In Figures 1a, b and 1c, d, $d = 0$ and $d = 1$, respectively, causing the eigenvalue spectra of $\textbf{A}$ to differ on the real axis (Figure a,c).
+Figure panels b and c show eigenvalue spectra after $\gamma$ is included such that $\textbf{M} = \gamma \textbf{A}$, where elements of $\gamma$ are drawn from a uniform distribution $\gamma \sim \mathcal{U}(0, 2)$.
 In both cases, $\sigma^{2}_{A} =$ `r round(sigma, digits = 4)` and $\sigma^{2}_{\gamma} =$ `r round(joint_var, digits = 4)`.
@@
 Stability also decreases with increasing $\rho$ such that $\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)$.
-Hence, when $\rho$ values are negative, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability is expected to increase [@Allesina2012].
+When $\rho$ values are negative, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability is expected to increase [@Allesina2012].
 When elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the covariance between $M_{ij}$ and $M_{ji}$,
@@
 ```{r, echo = FALSE, fig.height = 8, fig.width = 6}
-sim10 <- read.csv(file = "sim_results/rhos/sim10_rho.csv");
-sim15 <- read.csv(file = "sim_results/rhos/sim15_rho.csv");
-sim20 <- read.csv(file = "sim_results/rhos/sim20_rho.csv");
-sim25 <- read.csv(file = "sim_results/rhos/sim25_rho.csv");
-sim30 <- read.csv(file = "sim_results/rhos/sim30_rho.csv");
-sim35 <- read.csv(file = "sim_results/rhos/sim35_rho.csv");
+sim10 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim10.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim15.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim20.csv");
+sim25 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim25.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim30.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim35.csv");
+sim40 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim40.csv");
+sim45 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim45.csv");
 par(mar = c(0.25, 0.25, 0.25, 0.25), mfrow = c(3, 2), oma = c(6, 6, 1, 1));
-plot(x = sim10[,1], y = sim10[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), cex.axis = 1.5);
-points(x = sim10[,15], y = sim10[,21], type = "l", lwd = 2, lty = "dashed");
-abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 10", cex = 2.5);
-plot(x = sim15[,1], y = sim15[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
-points(x = sim15[,15], y = sim15[,21], type = "l", lwd = 2, lty = "dashed");
-abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 15", cex = 2.5);
 plot(x = sim20[,1], y = sim20[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
@@
 plot(x = sim25[,1], y = sim25[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.25);
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
 points(x = sim25[,15], y = sim25[,21], type = "l", lwd = 2, lty = "dashed");
@@
 text(x = -0.6, y = 3.2, labels = "S = 25", cex = 2.5);
-plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, 
+plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
      ylim = c(-1, 3.5), cex.axis = 1.5);
@@
 text(x = -0.6, y = 3.2, labels = "S = 30", cex = 2.5);
-plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, 
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.25);
 points(x = sim35[,15], y = sim35[,21], type = "l", lwd = 2, lty = "dashed");
@@
 text(x = -0.6, y = 3.2, labels = "S = 35", cex = 2.5);
+plot(x = sim40[,1], y = sim40[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), cex.axis = 1.5);
+points(x = sim40[,15], y = sim40[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 40", cex = 2.5);
+plot(x = sim45[,1], y = sim45[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+points(x = sim45[,15], y = sim45[,21], type = "l", lwd = 2, lty = "dashed");
+abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 3.2, labels = "S = 45", cex = 2.5);
 mtext(text =  expression(paste("E Correlation between A"[ij]," and A"[ji], 

### 2020-01-25 notebook/revision_notes.tex
More to the argument for a numerical approach, why an analytical one probably won't work

@@
 
+\hypertarget{effects-of-correlations-interactions}{%
 \section{Effects of correlations
-interactions}\label{effects-of-correlations-interactions}
+interactions}\label{effects-of-correlations-interactions}}
 
@@
 eigenvalue spectra of \(\textbf{M}\). In both panels of Figure 1,
-\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a and 1b,
-\(d = 0\) and \(d = 1\), respectively, causing the eigen spectra of
-\(\textbf{A}\) to differ on the real axis (blue points). Red points she
-eigen spectra after \(\gamma\) is include such that \(M = \gamma A\),
-where elements of \(\gamma\) are drawn from a uniform distribution
-\(\gamma \sim \mathcal{U}(0, 2)\). In both cases, \(\sigma^{2}_{A} =\)
-0.05 and \(\sigma^{2}_{\gamma} =\) 0.0033.
+\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a,b and 1c,d,
+\(d = 0\) and \(d = 1\), respectively, causing the eigenvalue spectra of
+\(\textbf{A}\) to differ on the real axis (Figure a,c). Figure panels b
+and c show eigenvalue spectra after \(\gamma\) is included such that
+\(M = \gamma A\), where elements of \(\gamma\) are drawn from a uniform
+distribution \(\gamma \sim \mathcal{U}(0, 2)\). In both cases,
+\(\sigma^{2}_{A} =\) 0.05 and \(\sigma^{2}_{\gamma} =\) 0.0031.
+Nevertheless, the change in eigenvalue spectra caused by \(\gamma\) can
+affect \(\max(\Re(\lambda))\) and therefore system stability.
 
-\section{More down}\label{more-down}
+Stability also decreases with increasing \(\rho\) such that
+\(\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)\).
+Hence, when \(\rho\) values are negative, such as when \(\textbf{M}\)
+models a predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have
+opposing signs, stability is expected to
+increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
+elements of \(\gamma\) vary independently, the magnitude of \(\rho\) is
+decreased because \(\sigma^{2}_{\gamma}\) increases the variance of
+\(M_{ij}\) without affecting the covariance between \(M_{ij}\) and
+\(M_{ji}\),
+
+\[\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.\]
+Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
+\(\rho\) remains unchanged.
+
+\hypertarget{more-down}{%
+\section{More down}\label{more-down}}
 
@@
 Note that the correlation \(\rho\) adjusts the criteria for stability as
-follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{2}},
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{3}},
 
@@
 defined as
-follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{2}},
+follows\textsuperscript{\protect\hyperlink{ref-Allesina2015a}{3}},
 
@@
 
-Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{3}}
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{4}}
 showed analytically that as \(S \to \infty\), the effect of a diagonal
@@
 \(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
-(95\% CIs: -74.3396367\(\times 10^{-6}\), 77.6796914\(\times 10^{-6}\)).
+(95\% CIs: -75.9050127\(\times 10^{-6}\), 84.5804827\(\times 10^{-6}\)).
 Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Tang2014b}{}
+\leavevmode\hypertarget{ref-Tang2014b}{}%
 1. Tang, S. \& Allesina, S. Reactivity and stability of large
@@
 
-\hypertarget{ref-Allesina2015a}{}
-2. Allesina, S. \& Tang, S. The stability--complexity relationship at
+\leavevmode\hypertarget{ref-Allesina2012}{}%
+2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
+\emph{Nature} \textbf{483,} 205--208 (2012).
+
+\leavevmode\hypertarget{ref-Allesina2015a}{}%
+3. Allesina, S. \& Tang, S. The stability--complexity relationship at
 age 40: a random matrix perspective. \emph{Population Ecology} 63--75
@@
 
-\hypertarget{ref-Gibbs2017}{}
-3. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+\leavevmode\hypertarget{ref-Gibbs2017}{}%
+4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.

### 2020-01-25 notebook/sci_rep_reviews.Rmd
Some more analytical playing, and rho results

@@
 
-Reviewer 2 comments
+Reviewer 2 comment 1
 ================================================================================
@@
 
+Author response to Reviewer 1 comment 1
+================================================================================
+
+I thank Reviewer 2 for their helpful comments.
+In my revision, I have provided more analytical context, expanded the types of networks explored (see also Author response to Reviewer 1 comment 3), and further clarified the role response rate variation.
+I explain these changes in more detail with the reviewers elaborated concerns below.
+
+Reviewer 2 comment 2
+================================================================================
+
 The results crucially build on the matrices being dangerously close to the stability boundary.
@@
 In either case, the manuscript would of course be much stronger with analytical results, but I do realize this may be uncharted territory.
+
+Author response to Reviewer 1 comment 1
+================================================================================
+
+I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (line XXX).
+In making these changes, I have slightly adjusted the notation so that the text is more readable.
+
+As Reviewer 2 correctly points out, my focus is on finite values of S and \sigma, in which complexity is gradually increased until it is dangerously close to the stability boundary.
+Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases stability) would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$.
+For circular matrices in which elements are i.i.d ($A$), it has been proved that eigenvalues are distributed uniformly within a known boundary of support of the spectral density (Tao and Vu 2010).
+While recent papers have found the support for more general types of random matrices that include the form $M = AX$ (e.g, Gibbs et al.
+2018; Neri and Metz 2012; Ahmadian et al.
+2015), to the best of my understanding, finding the actual density of eigenvalues within the support has not been attempted.
+This likely matters because for finite S, even though the real boundary of the support might be positive for both $M$ and $A$ (implying instability as $S \to \infty$; Grilli et al.
+2017), the right most eigenvalue for a system still might be less than zero.
+
+My approach follows others with similar goals (Sinha 2005).
+
+Others have found the support
+
+Reviewer 2 is correct that analytical results here are uncharted territory, despite
+
+In recent papers, analytical solutions have been provided to find the support of the spectral densities of matrices of the form $M = XA$,
+
+This is unique, as most studies make the simplifying assumption of \sigma = \sqrt{S} such that the circular matrix has a radius of unity.
+
+Ideally, an analytical proof could demonstrate the uncertainty in the position of the leading eigenvalue at finite S, which would likely require knowledge of the distribution of M = \gamma A for arbitrary values of \sigma and S.
+
+For matrices of the form $M = XA$, Gibbs et al. (2018) do find the boundaries of support for spectral densities.
+
+My focus is uniquely on a finite S.
+This is important, as Ahmadian et al.
+(2015) showed that any realisation of the eigenvalue distribution for a matrix of the form Z = M + LJR approaches the expected distribution as S \to infty.
+
+Elements of M = XA are no longer independent.
 

### 2020-01-25 notebook/analytical.R

@@ -1235,6 +1235,159 @@ for(i in 1:100000){
 }
 
 
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
+################################################################################
[…]

### 2020-01-25 R/random_mats.R
Small tweaks corrections

@@ -34,7 +34,7 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         while(iter > 0){
             r_vec    <- rnorm(n = sp_try[i], mean = mn, sd = rmx);
@@ -74,6 +74,8 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
             rho_res[[i]][iter, 2] <- A1_rho;
             rho_res[[i]][iter, 3] <- rho_diff;
             rho_res[[i]][iter, 4] <- rho_abs;
+            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
+            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
             cmplxty[[i]][iter, 1] <- get_complexity(A0);
             cmplxty[[i]][iter, 2] <- get_complexity(A1);
             iter                  <- iter - 1;
@@ -226,7 +228,7 @@ rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         while(iter > 0){
             r_vec    <- rnorm(n = S, mean = 0, sd = rmx);
@@ -259,6 +261,8 @@ rand_rho_var <- function(S, rhos, iters, int_type = 0, rmx = 0.4, C = 1,
             rho_res[[i]][iter, 2] <- A1_rho;
[…]

### 2020-01-25 R/scale_free.R

@@ -41,7 +41,7 @@ rand_gen_sfn <- function(S, m_vals, iters, int_type = 0, rmx = 0.4, sigma = 0.4,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         real_Cs[[i]]   <- matrix(data = 0, nrow = iter, ncol = 3);
         while(iter > 0){
@@ -82,6 +82,8 @@ rand_gen_sfn <- function(S, m_vals, iters, int_type = 0, rmx = 0.4, sigma = 0.4,
             rho_res[[i]][iter, 2] <- A1_rho;
             rho_res[[i]][iter, 3] <- rho_diff;
             rho_res[[i]][iter, 4] <- rho_abs;
+            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
+            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
             cmplxty[[i]][iter, 1] <- get_complexity(A0);
             cmplxty[[i]][iter, 2] <- get_complexity(A1);
             iter                  <- iter - 1;

### 2020-01-25 R/small_world.R

@@ -43,7 +43,7 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
         fea_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
-        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 4);
+        rho_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 6);
         cmplxty[[i]]   <- matrix(data = 0, nrow = iter, ncol = 2);
         real_Cs[[i]]   <- matrix(data = 0, nrow = iter, ncol = 3);
         while(iter > 0){
@@ -90,6 +90,8 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
             rho_res[[i]][iter, 2] <- A1_rho;
             rho_res[[i]][iter, 3] <- rho_diff;
             rho_res[[i]][iter, 4] <- rho_abs;
+            rho_res[[i]][iter, 5] <- max(Re(eigen(A0)$values));
+            rho_res[[i]][iter, 6] <- max(Re(eigen(A1)$values));
             cmplxty[[i]][iter, 1] <- get_complexity(A0);
             cmplxty[[i]][iter, 2] <- get_complexity(A1);
             iter                  <- iter - 1;

### 2020-02-03 notebook/sci_rep_reviews.Rmd
Address Reviewer 2 comments

@@
 
-Author response to Reviewer 1 comment 1
+Author response to Reviewer 2 comment 1
 ================================================================================
@@
 
-Author response to Reviewer 1 comment 1
+Author response to Reviewer 2 comment 2
 ================================================================================
@@
 As Reviewer 2 correctly points out, my focus is on finite values of S and \sigma, in which complexity is gradually increased until it is dangerously close to the stability boundary.
-Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases stability) would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$.
+Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases the stability of $M$), would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$.
 For circular matrices in which elements are i.i.d ($A$), it has been proved that eigenvalues are distributed uniformly within a known boundary of support of the spectral density (Tao and Vu 2010).
@@
 
-My approach follows others with similar goals (Sinha 2005).
+Reviewer 2 is correct that this is uncharted territory, and after carefully considering the possibility of analytical demonstration of my manuscript's results, I am convinced that such a demonstration is beyond the scope of this manuscript.
+My approach follows others with similar objectives.
+For example, Sinha (2005) also used a numerical approach to understand the stability of small-world networks (in contrast to the classic random networks of May).
 
-Others have found the support
-
-Reviewer 2 is correct that analytical results here are uncharted territory, despite
-
-In recent papers, analytical solutions have been provided to find the support of the spectral densities of matrices of the form $M = XA$,
-
-This is unique, as most studies make the simplifying assumption of \sigma = \sqrt{S} such that the circular matrix has a radius of unity.
-
-Ideally, an analytical proof could demonstrate the uncertainty in the position of the leading eigenvalue at finite S, which would likely require knowledge of the distribution of M = \gamma A for arbitrary values of \sigma and S.
-
-For matrices of the form $M = XA$, Gibbs et al. (2018) do find the boundaries of support for spectral densities.
-
-My focus is uniquely on a finite S.
-This is important, as Ahmadian et al.
-(2015) showed that any realisation of the eigenvalue distribution for a matrix of the form Z = M + LJR approaches the expected distribution as S \to infty.
-
-Elements of M = XA are no longer independent.
+Reviewer 2 comment 3
+================================================================================
 
@@
 
+Author response to Reviewer 2 comment 3
+================================================================================
+
+These were very helpful comments, and I am especially grateful to Reviewer 2 for doing a bit of preliminary exploration themselves to suggest improvements.
+I have now run simulations for different values of $\rho$, as suggested.
+Reviewer 2 is correct that the nonzero $\rho$ values do not change my results, but an analysis of stability across $\rho$ values also revealed an interesting pattern.
+While the relationship between $\rho$ and the leading eigenvalue is linear for circular matrices ($A$), it is convex when $\gamma$ is included such that $M = \gamma A$.
+I have now included these results into the main text of the manuscript, as I believe that they are important.
+
+As suggested, I also considered matrices with even more structure.
+Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems.
+These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002).
+
+Reviewer 2 comment 4
+================================================================================
+
 Finally, the general relevance of the results is hampered by being confined to a narrow parameter space, and even there the fraction of occurrences where var(gamma) makes a difference is minuscule.
@@
 
+Author response to Reviewer 2 comment 4
+================================================================================
+
+While I agree that the effect of $var(\gamma)$ is restricted to a small proportion systems at the margins of stability, I would argue that this effect is actually of quite general relevance because of the breadth of complex systems to which my results apply.
+Addressing Reviewer 2's previous comment helped emphasise this; the effect of variation in $\gamma$ on stability, while small in overall magnitude, is predicted to be general across any physical, social, or biological network.
+Since nearly all such networks will have components that vary in the rate at which they respond to perturbation, my results contribute to a general understanding of system stability.
+
+I have revised the writing to more carefully frame the manuscript's key results (line XXX).
+This includes a change of the title, which now makes it more clear that the effects on stability are small while retaining an emphasis on generality across systems.
+I have also removed the word "markedly" from the abstract (line XXX).
+
 Reviewer 3 commments
 ================================================================================
-
-This manuscript describe the effect of variation in the response rate of individual systems to the stability of a complex system by A.
-Bradley Duthie.
-The manuscript is well-written and the calculations, codes and so on sound technically correct, but I have problem to see the importance of the response rates to understand the stability of the systems.
-
-In my opinion 3 main questions are more important in all these discussions:
-
-1.
-The diagonal of the matrix, thinking in species interactions, represents the intraspecific interactions in the same species, and in the study of the stability is the main point because it gives us the center of the cloud of eigenvalues.
-In these study, as many all of them, is taken as -1.
-All infraspecific interactions in every species are negatives and equal?
-It is difficult to agree with this idea.
-
-2.
-In these discussions the author use random matrix to implement complex systems.
-Again it is difficult to understand how a random matrix can represent a complex system.
-If we think in species again, different ecological interactions have different distributions, has can be observed in the paper of Tang and I Sundstrand that the random matrices can be a perfect null model, but nothing more.
-
-3. I can understand the importance of the study of the responds rates to understand the stability of the matrices. Maybe I am wrong but sound like the author is comparing apples and oranges.
 

### 2020-02-03 notebook/sci_rep_reviews.Rmd
Update some response to Reviewer 3

@@
 
-Reviewer 3 commments
+Reviewer 3 commment 1
 ================================================================================
@@
 
+Author response to Reviewer 3 commment 1
+================================================================================
+
+I thank Reviewer 3 for their review, including their comment on the quality of the writing and verification that the calculations and and code are technically correct.
+In my revision, in line with the comments of the other two reviewers, I have improved the way in which the importance of my results are communicated.
+I address Reviewer 3's main three questions below.
+
+Reviewer 3 commment 2
+================================================================================
+
 1.
@@
 It is difficult to agree with this idea.
+
+Author response to Reviewer 3 commment 2
+================================================================================
+
+The diagonal of the matrix $M$ models component self-regulation, such as species carrying capacity in an ecological community.
+The relationship between diagonal elements of $M$ and stability is well-known for large circular matrices (e.g., Tang et al.
+2014; Allesina and Tang 2015), in which the mean value of diagonal elements determines the center of the eigenvalue distribution on the real axis as $S \to \infty$.
+
+Theoretical studies of community stability assign negative diagonal values (typically -1) to make the simplifying assumption that components of the system are self-regulating (May 1972; Sinha 2005; Allesina and Tang 2012; Grilli et al.
+2017; Song and Saavedra 2018; Stone 2018).
+If components are not self-regulating, then the system will almost never be stable.
+I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
+The effect of $\gamma$ on increasing stability is retained, but stability of all systems drops to zero out of one million given size $S > 14$.
+And in the random systems that are found to be stable, all diagonal elements are nearly always negative (or, rarely, with one or two very low positive values).
 

### 2020-02-03 notebook/sci_rep_reviews.Rmd
Draft of all responses to reviewers, aside from the initial response to the editor

@@
 
-Theoretical studies of community stability assign negative diagonal values (typically -1) to make the simplifying assumption that components of the system are self-regulating (May 1972; Sinha 2005; Allesina and Tang 2012; Grilli et al.
+Theoretical studies of community stability assign negative diagonal values (typically -1) to make the simplifying assumption that components of the system are self-regulating (e.g., May 1972, 1973; Sinha 2005; Allesina and Tang 2012; Grilli et al.
 2017; Song and Saavedra 2018; Stone 2018).
-If components are not self-regulating, then the system will almost never be stable.
+If individual components are not self-regulating, then the system will almost never be stable.
 I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
@@
 And in the random systems that are found to be stable, all diagonal elements are nearly always negative (or, rarely, with one or two very low positive values).
+In other words, if diagonal elements were randomly sampled from a distribution with positive support, then the only systems that would be stable would be those that by chance had all (or nearly all) diagonal elements.
+
+Hence, in this paper and nearly all related papers, when assessing stability, it makes sense to assume that individual system components are self-regulating.
+But this is not necessarily just a simplifying assumption to avoid unnecessary computations.
+In the case of ecological species, a positive intraspecific interaction value does not make sense.
+This would imply that the species being modelled did not compete with itself and had no carrying capacity (because intraspecific competition is directly related to carrying capacity).
+Since infinite growth is clearly unrealistic, ecological systems represented by random matrices really should have negative diagonal elements.
+
+Unequal diagonal elements could be used, but doing so would only be expected to add some variation around the expected centre of eigenvalue distributions, and would not change my primary result concerning the effect of $\gamma$ (note that this would lead to the same results as mentioned above where diagonal elements were sampled in the same way as off diagonals, but with more total stable systems since the mean diagonal value would be less lower).
+
+I have tried to clarify and more clearly state the assumption of self-regulation in my revision (line XXX).
+
+Reviewer 3 commment 3
+================================================================================
 
@@
 
+Author response to Reviewer 3 commment 3
+================================================================================
+
+Reviewer 3 is correct that random matrices are an ideal null model for understanding stability in complex systems such as communities of species.
+In this manuscript, I use the random matrix approach as a null model to investigate the role played by a property of nearly all complex systems (component response variation) in affecting stability.
+This approach to investigating stability in complex systems has been used widely since May (1972), and has been refined to more specifically to reflect the properties of systems with a particular structure (e.g., specific ecological interactions; Allesina and Tang 2012).
+My revision now more extensively simulates the structures of real-world networks, including the cascade foode web model  (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002; see also Author response to Reviewer 2 comment 3).
+
+I have tried to clarify the role of random matrices as a null model of complex system stability (line XXX).
+
+Reviewer 3 commment 4
+================================================================================
+
 3. I can understand the importance of the study of the responds rates to understand the stability of the matrices. Maybe I am wrong but sound like the author is comparing apples and oranges.
+
+Author response to Reviewer 3 commment 4
+================================================================================
+
+Component response rates can be modelled using a diagonal matrix in which elements correspond to the rate at which individual system components respond to perturbations in the system (Patel et al.
+2018).
+Hence, such a diagonal matrix $\gamma$ can be multiplied by a random matrix representing a complex system $A$ to produce a complex system in which component response rates vary, $M = \gamma A$.
+By doing an eigenanalysis on $M$ versus $A$, it is possible to understand how response rates affect stability.
+Hence, I am not comparing apples and oranges; response rates of components have a clear and important role in complex system stability that can be investigated with the mathematical techniques used in this manuscript.
+I have tried to explain this role more clearly in my revision (line XXX).
+
+Reviewer 3 commment 5
+================================================================================
 
@@
 
-A. Mougi, and M. Kondoh, Diversity of Interaction Types and Ecolog- ical Community Stability, Science 337, 349 (2012).
+A. Mougi, and M. Kondoh, Diversity of Interaction Types and Ecological Community Stability, Science 337, 349 (2012).
 
-K.Z. Coyte, J. Schluter, and K.R. Foster, The ecology of the micro- biome: Networks, competition, and stability, Science 350, 663 (2015).
+K.Z. Coyte, J. Schluter, and K.R. Foster, The ecology of the microbiome: Networks, competition, and stability, Science 350, 663 (2015).
+
+Author response to Reviewer 3 commment 3
+================================================================================
+
+I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (lines XXX).
+
+Author References within Review
+================================================================================
+
+Ahmadian, Y., Fumarola, F., & Miller, K.
+D.
+(2015).
+Properties of networks with partially structured and partially random connectivity.
+Physical Review E - Statistical, Nonlinear, and Soft Matter Physics, 91(1), 012820.
+https://doi.org/10.1103/PhysRevE.91.012820
+
+Albert, R., & Barab\'{a}si, A. L. (2002). Statistical mechanics of complex networks. Reviews of Modern Physics, 74(1), 47–97. https://doi.org/10.1103/RevModPhys.74.47
+
+Allesina, S., & Tang, S. (2012). Stability criteria for complex ecosystems. Nature, 483(7388), 205–208. https://doi.org/10.1038/nature10832
+
+Allesina, S., & Tang, S. (2015). The stability–complexity relationship at age 40: a random matrix perspective. Population Ecology, 63–75. https://doi.org/10.1007/s10144-014-0471-0
+
+Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F., & Mazza, C.
+(2018).
+The feasibility of equilibria in large ecosystems: a primary but neglected concept in the complexity-stability debate.
+PLOS Computational Biology, 14(2), e1005988.
+Retrieved from http://arxiv.org/abs/1612.06735
+
+Gibbs, T., Grilli, J., Rogers, T., & Allesina, S. (2017). The effect of population abundances on the stability of large random ecosystems. Arxiv. Retrieved from http://arxiv.org/abs/1708.08837
+
+Grilli, J., Adorisio, M., Suweis, S., Barabás, G., Banavar, J.
+R., Allesina, S., & Maritan, A.
+(2017).
+Feasibility and coexistence of large ecological communities.
+Nature Communications, 8.
+https://doi.org/10.1038/ncomms14389
+
+May, R. M. (1972). Will a large complex system be stable? Nature, 238, 413–414.
+
+May, R. M. (1973). Qualitative stability in model ecosystems. Ecology, 54(3), 638–641. https://doi.org/10.2307/1935352
+
+Neri, I., & Metz, F. L. (2012). Spectra of sparse non-Hermitian random matrices: An analytical solution. Physical Review Letters, 109(3). https://doi.org/10.1103/PhysRevLett.109.030602
+
+Patel, S., Cortez, M. H., & Schreiber, S. J. (2018). Partitioning the effects of eco-evolutionary feedbacks on community stability. American Naturalist, 191(3), 1–29. https://doi.org/10.1101/104505
+
+Sinha, S. (2005). Complexity vs. stability in small-world networks. Physica A: Statistical Mechanics and Its Applications, 346(1-2 SPEC. ISS.), 147–153. https://doi.org/10.1016/j.physa.2004.08.062
+
+Song, C., & Saavedra, S. (2018). Will a small randomly assembled community be feasible and stable? Ecology, 99(3), 743–751. https://doi.org/10.1002/ecy.2125
+
+Stone, L. (2018). The feasibility and stability of large complex biological networks: a random matrix approach. Scientific Reports, 8, 8246. https://doi.org/10.1101/223651
+
+Tang, S., & Allesina, S. (2014). Reactivity and stability of large ecosystems. Frontiers in Ecology and Evolution, 2(June), 1–8. https://doi.org/10.3389/fevo.2014.00021
+
+Tao, T., & Vu, V. (2010). Random matrices: Universality of ESDs and the circular law. Annals of Probability, 38(5), 2023–2065. https://doi.org/10.1214/10-AOP534
+
+Watts, D. J., & Strogatz, S. H. (1998). Collective dynamics of “small world” networks. Nature, 393(June), 440–442.
+
+Williams, R. J., & Martinez, N. D. (2000). Simple rules yield complex food webs. Nature, 404(6774), 180–183. https://doi.org/10.1038/35004572

### 2020-02-23 notebook/sci_rep_reviews.Rmd
Need to write now
Small edits to response to reviewers

@@
 2. Run simulations across multiple structured network types to confirm the robustness of results. Network types include (a) cascade food webs, (b) small world networks, and (c) scale-free network.
+
+3. Rewrite the manuscript, or the relevant parts of it.
 
@@
 Other network types that I have incorporated into my revision include small-world networks (Watts and Strogatz 1998) and scale-free networks (Albert and Barab\'{a}si 2002).
-Overall, the structured networks that I now include in the main text and supporting information now encompass a wide range of physical, biological, and social systems.
+Overall, the structured networks that I now include in the main text and supporting information encompass a wide range of physical, biological, and social systems.
 
@@
 I thank Reviewer 2 for their helpful comments.
-In my revision, I have provided more analytical context, expanded the types of networks explored (see also Author response to Reviewer 1 comment 3), and further clarified the role response rate variation.
-I explain these changes in more detail with the reviewers elaborated concerns below.
+In my revision, I have provided more analytical context, expanded the types of networks explored (see also Author response to Reviewer 1 comment 3), and further clarified the role of response rate variation.
+I explain these changes in more detail with the reviewer's elaborated concerns below.
 
@@
 2018; Neri and Metz 2012; Ahmadian et al.
-2015), to the best of my understanding, finding the actual density of eigenvalues within the support has not been attempted.
+2015), to the best of my knowledge, finding the actual density of eigenvalues within the support has not been attempted.
 This likely matters because for finite S, even though the real boundary of the support might be positive for both $M$ and $A$ (implying instability as $S \to \infty$; Grilli et al.
@@
 If individual components are not self-regulating, then the system will almost never be stable.
-I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
+I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI, which was subsequently improved prior to submission: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
 The effect of $\gamma$ on increasing stability is retained, but stability of all systems drops to zero out of one million given size $S > 14$.
 And in the random systems that are found to be stable, all diagonal elements are nearly always negative (or, rarely, with one or two very low positive values).
-In other words, if diagonal elements were randomly sampled from a distribution with positive support, then the only systems that would be stable would be those that by chance had all (or nearly all) diagonal elements.
+In other words, if diagonal elements were randomly sampled from a distribution with positive support, then the only systems that would be stable would be those that by chance had all (or nearly all) negative diagonal elements.
 
@@
 In the case of ecological species, a positive intraspecific interaction value does not make sense.
-This would imply that the species being modelled did not compete with itself and had no carrying capacity (because intraspecific competition is directly related to carrying capacity).
+This would imply that the species being modelled did not compete with itself and had no carrying capacity (because intraspecific competition is directly and inversely related to carrying capacity).
 Since infinite growth is clearly unrealistic, ecological systems represented by random matrices really should have negative diagonal elements.

### 2020-02-23 notebook/ms.Rmd
Some side additions
New sentence in the abstract

--- 2019-03-06 notebook/ms.Rmd
@@
 Here I vary the component response rates ($\boldsymbol{\gamma}$) of randomly generated complex systems.
-I show that when component response rates vary, the potential for system stability is markedly increased.
+I use numerical simulations to show that when component response rates vary, the potential for system stability is markedly increased.
+These results are robust to common network structures, including small-world and scale-free networks, and cascade food webs.
 Variation in $\boldsymbol{\gamma}$ is especially important for stability in highly complex systems, in which the probability of stability would otherwise be negligible.
@@
 
+**Additional Information:** The author declares no competing interests.
+All work was carried out by A.
+Bradley Duthie, and all code and data are accessible on [GitHub](https://github.com/bradduthie/RandomMatrixStability).
+
 **References**

### 2020-02-23 notebook/revision_notes.tex
Run the revision notes again

@@
 
-\hypertarget{effects-of-correlations-interactions}{%
 \section{Effects of correlations
-interactions}\label{effects-of-correlations-interactions}}
+interactions}\label{effects-of-correlations-interactions}
 
@@
 eigenvalue spectra of \(\textbf{M}\). In both panels of Figure 1,
-\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a,b and 1c,d,
+\(S =\) 400 and \(\sigma_{A} = 1/\sqrt{S}\). In Figures 1a, b and 1c, d,
 \(d = 0\) and \(d = 1\), respectively, causing the eigenvalue spectra of
@@
 and c show eigenvalue spectra after \(\gamma\) is included such that
-\(M = \gamma A\), where elements of \(\gamma\) are drawn from a uniform
-distribution \(\gamma \sim \mathcal{U}(0, 2)\). In both cases,
-\(\sigma^{2}_{A} =\) 0.05 and \(\sigma^{2}_{\gamma} =\) 0.0031.
-Nevertheless, the change in eigenvalue spectra caused by \(\gamma\) can
-affect \(\max(\Re(\lambda))\) and therefore system stability.
+\(\textbf{M} = \gamma \textbf{A}\), where elements of \(\gamma\) are
+drawn from a uniform distribution \(\gamma \sim \mathcal{U}(0, 2)\). In
+both cases, \(\sigma^{2}_{A} =\) 0.05 and \(\sigma^{2}_{\gamma} =\)
+0.0032. Nevertheless, the change in eigenvalue spectra caused by
+\(\gamma\) can affect \(\max(\Re(\lambda))\) and therefore system
+stability.
 
@@
 \(\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right)\).
-Hence, when \(\rho\) values are negative, such as when \(\textbf{M}\)
-models a predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have
-opposing signs, stability is expected to
+When \(\rho\) values are negative, such as when \(\textbf{M}\) models a
+predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have opposing
+signs, stability is expected to
 increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
@@
 \[\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.\]
+
 Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
@@
 
-\hypertarget{more-down}{%
-\section{More down}\label{more-down}}
+\section{More down}\label{more-down}
 
@@
 \(\rho\) was normally distributed around 4.8394472\(\times 10^{-6}\)
-(95\% CIs: -75.9050127\(\times 10^{-6}\), 84.5804827\(\times 10^{-6}\)).
+(95\% CIs: -80.284156\(\times 10^{-6}\), 86.1295891\(\times 10^{-6}\)).
 Hence, decreasing the correlation between \(\textbf{A}_{ij}\) and
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Tang2014b}{}%
+\hypertarget{ref-Tang2014b}{}
 1. Tang, S. \& Allesina, S. Reactivity and stability of large
@@
 
-\leavevmode\hypertarget{ref-Allesina2012}{}%
+\hypertarget{ref-Allesina2012}{}
 2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
@@
 
-\leavevmode\hypertarget{ref-Allesina2015a}{}%
+\hypertarget{ref-Allesina2015a}{}
 3. Allesina, S. \& Tang, S. The stability--complexity relationship at
@@
 
-\leavevmode\hypertarget{ref-Gibbs2017}{}%
+\hypertarget{ref-Gibbs2017}{}
 4. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of

### 2020-02-23 notebook/ms.Rmd
Add in some of the places where the writing needs to change, including figures and tables that need to be inserted or removed

@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary
+Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
+<!--- Something about including realistic structured networks (scale free, small world, and cascade food webs) --->
 
@@
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+<!--- Replace the below with some more informative text concerning the variation that $\var(gamma})$ adds to the system $M$ overall.
+This might require starting with just $M = A$ as an explanation.
+Borrowing the text from revision_notes.Rmd below --->
+
+<!--- The current Figure 1 should be replaced by the new figure first shown in revision_notes.Rmd --->
+
+<!---
+
+If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\max(\Re(\lambda))$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
+These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+When investigating the effect of varying component response rate $Var(\gamma)$ on stability by defining $\textbf{M} = \gamma\textbf{A}$, $S$ remains unchanged.
+Further, values of $\gamma_{i}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
+What $Var(\gamma)$ does change is the variation in component interaction strengths, and $\rho$.
+
+Introducing variation in $\gamma$ increases the total variation in the system, making it more likely that $\textbf{M} = \gamma\textbf{A}$ is unstable.
+Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables (to simplify the notation, $\sigma^{2}_{M}$ and $\sigma^{2}_{A}$ refer to variances of off-diagonal elements only),
+
+$$\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.$$
+
+Note that in my simulations $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, so the above can be simplified,
+
+$$\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).$$
+
+The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$.
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{M}$ and $\textbf{A}$ lie within a circle of a radius $\sigma^{2}_{M}\sqrt{S}$ and $\sigma^{2}_{A}\sqrt{S}$, respectively.
+
+--->
+
+<!--- Remove this instructive example; replace it with something better. This entire paragraph and the code underneath it can be deleted. --->
 An instructive example compares one $\mathbf{M}$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $\mathbf{M}$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
@@
 
+<!--- Some of this paragraph can be removed.
+Some of the middle can focus on the new Figure 1 panels C and D, which illustrate the effect of gamma on the eigenvalue spectra.
+The rest, focusing on the results, can be retained.--->
 Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
 I focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
-As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
+<!--- Cut from here ---> As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
 1b), mean $\gamma = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $\mathbf{M}$ before and after the addition of variation in component response rate.
 Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
-As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
+<!--- end cut here, but now refer to Figure 1 panels C and D ---> As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
 Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
@@
 But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's [@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
+<!--- Okay, so most can be retained --->
 
@@
 
+<!--- Here a new section is needed --->
+<!---
+
+**Effects of $\gamma$ and correlated interactions on stability**
+
+This is going to require that a new figure be inserted showing correlation versus stability.
+Some general points about how $rho$ is known to affect stability is also needed, but this should not be too difficult to explain in a sentence or two.
+
+--->
+
+<!--- Another new section is needed --->
+<!---
+
+**Effects of $\gamma$ on structured networks**
+
+This will report results for small-world, scale-free, and cascade food-webs.
+It will also briefly mention random predator-prey, mutualist, and competitor networks.
+These resutls will require a small table, at least.
+
+--->
+
+<!--- Tempting to remove this section entirely, but if there is enough space for it, then maybe retain it; reviewers did not mention a problem with it, and it's a short and relevant point to make.
+In any case, it should be swapped with the section on feasibility below.
+--->
 **Targeted manipulation of $\mathbf{\gamma}$**.
@@
 
+<!--- I don't think much here needs to change --->
 **System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
@@
 Variation in $\mathbf{\gamma}$ specifically isolates the effects of differing component response rates, hence causing differences in expected $\mathbf{M}$ row means.
+
+<!--- Need a section here on how the structured networks were built --->
 

### 2020-02-23 notebook/ms.Rmd
Add new sentence to the introduction listing robustness of network types, and add associated bibtex entries.

@@
 
-<!---
-
-Author Summary
---------------------------------------------------------------------------------
-
-Any set of discrete components that potentially interact can define a complex system.
-Typically, properties of individual components will cause them to respond to system perturbation at different rates, but the consequences of such variation for system stability are unexplored.
-This work makes three novel theoretical contributions: First, I show that large complex systems with varying component response rates are more likely to be stable than when response rates are uniform.
-Second, I show that a targeted manipulation of component response rates can increase the probability of system stability by several orders of magnitude.
-Lastly, I show that varying component response rate does not affect system feasibility.
-These results are general to any physical, biological, social, or mixed system.
-
---->
-
 Introduction
@@
 Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show why, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
-<!--- Something about including realistic structured networks (scale free, small world, and cascade food webs) --->
+Here I show that, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
+My results are robust across commonly observed network structures, including random [@May1972], small-world [@Watts1998], scale-free [@Albert2002], cascade food web [@Williams2000] networks.
 

### 2020-02-23 notebook/ms.tex
LaTeX files after knitting the new version to check that all looks good

@@
 component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
-complex systems. I show that when component response rates vary, the
-potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) is especially important for stability in highly
-complex systems, in which the probability of stability would otherwise
-be negligible. At such extremes of simulated system complexity, the
-largest stable complex systems would be unstable if not for
-\(\boldsymbol{Var(\gamma)}\). My results therefore reveal a previously
-unconsidered aspect of system stability that is likely to be pervasive
-across all realistic complex systems.
+complex systems. I use numerical simulations to show that when component
+response rates vary, the potential for system stability is markedly
+increased. These results are robust to common network structures,
+including small-world and scale-free networks, and cascade food webs.
+Variation in \(\boldsymbol{\gamma}\) is especially important for
+stability in highly complex systems, in which the probability of
+stability would otherwise be negligible. At such extremes of simulated
+system complexity, the largest stable complex systems would be unstable
+if not for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
+previously unconsidered aspect of system stability that is likely to be
+pervasive across all realistic complex systems.
 
@@
 by introducing a new source of variation into the system and thereby
-increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
+increasing \(\sigma\). Here I show that, despite higher \(\sigma\),
 realistic complex systems (such that \(S\) is high but finite) are
 actually more likely to be stable if their individual component response
-rates vary
+rates vary. My results are robust across commonly observed network
+structures, including
+random\textsuperscript{\protect\hyperlink{ref-May1972}{1}},
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}},
+cascade food
+web\textsuperscript{\protect\hyperlink{ref-Williams2000}{17}} networks.
 
@@
 value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
 The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
-\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
-half of \(\gamma_{i} = 0.05\). This models one system in which
-\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
-are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
-cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
-diagonal elements are set to \(-1\) and non-zero off-diagonal elements
-are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
-multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
-\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
-uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
-(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
-to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
-values are separated into two groups, eigenvalues are no longer
-uniformly distributed (Fig. 1b). Instead, two distinct clusters of
-eigenvalues appear (red circles in Fig. 1b), one centred at
-(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
-large radius, but the real components have shifted to the left (in
-comparison to when \(\gamma = 1\)), while the latter cluster has a
-smaller radius, but real components have shifted to the right; all
-\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
-\(\mathbf{M}\), this division between slow and fast component response
-rates results in more stable systems: 1 stable given \(\gamma = 1\)
-versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+distribution of component response rates.
+
+An instructive example compares one \(\mathbf{M}\) where
+\(\gamma_{i} = 1\) for all \(i\) in \(S\) to the same \(\mathbf{M}\)
+when half of \(\gamma_{i} = 1.95\) and half of \(\gamma_{i} = 0.05\).
+This models one system in which \(\gamma\) is invariant and one in which
+\(\gamma\) varies, but systems are otherwise identical (note that mean
+\(\gamma_{i} = 1\) in both cases). I assume \(S = 200\), \(C = 0.05\),
+and \(\sigma = 0.4\); diagonal elements are set to \(-1\) and non-zero
+off-diagonal elements are drawn randomly from
+\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
+\(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
+eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
+centred at (\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real
+components of eigenvalues are highly unlikely to all be negative when
+all \(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated
+into two groups, eigenvalues are no longer uniformly distributed (Fig.
+1b). Instead, two distinct clusters of eigenvalues appear (red circles
+in Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
+(\(-0.05, 0\)). The former has a large radius, but the real components
+have shifted to the left (in comparison to when \(\gamma = 1\)), while
+the latter cluster has a smaller radius, but real components have
+shifted to the right; all \(\Re({\lambda}) < 0\). Overall, for 1 million
+randomly assembled \(\mathbf{M}\), this division between slow and fast
+component response rates results in more stable systems: 1 stable given
+\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
 
@@
 Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{19}--\protect\hyperlink{ref-Stone2017}{21}},
 some of which are clustered locally near the centre of the distribution,
@@
 values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{22}}.
 Generations of selection on \(\gamma\) value combinations to minimise
@@
 positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{23},\protect\hyperlink{ref-Song2018}{24}}.
 This is of particular interest for ecological communities because
@@
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}}.
 While my results are intended to be general to all complex systems, and
@@
 a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
+Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{25}}
 (see Supplementary Information). Hence, for pure interacting species
@@
 My focus is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20}}, who applied
 the same mathematical framework to investigate how a diagonal matrix
@@
 \(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}} analytically
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20}} analytically
 demonstrated that the effect of \(\mathbf{X}\) on system stability
@@
 of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}}
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{21},\protect\hyperlink{ref-Dougoud2018}{23},\protect\hyperlink{ref-Servan2018}{25}}
 because population densities cannot realistically be negative. My
@@
 empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, but
 instead interact with evolutionary, abiotic, or social-economic systems.
@@
 distributions caused by a vector \(\gamma\) has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{19}}
 and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20},\protect\hyperlink{ref-Stone2017}{21}},
 but \(\gamma\) has not been interpreted as rates of response of
@@
 traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
 For any species \(i\) or trait \(j\), change in species density
@@
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
 
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{26}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{23}},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{26}},
 
@@
 components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{19},\protect\hyperlink{ref-May1973}{26}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{22}}. In
 this case, the practical solution is finding vectors of
@@
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}),
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{25}}),
 
@@
 
+\hypertarget{ref-Watts1998}{}
+15. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
+networks. \emph{Nature} \textbf{393,} 440--442 (1998).
+
+\hypertarget{ref-Albert2002}{}
+16. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
+networks. \emph{Reviews of Modern Physics} \textbf{74,} 47--97 (2002).
+
+\hypertarget{ref-Williams2000}{}
+17. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
+webs. \emph{Nature} \textbf{404,} 180--183 (2000).
+
 \hypertarget{ref-Patel2018}{}
-15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+18. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 \hypertarget{ref-Ahmadian2015}{}
-16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+19. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
 with partially structured and partially random connectivity.
@@
 \hypertarget{ref-Gibbs2017}{}
-17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+20. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.
@@
 \hypertarget{ref-Stone2017}{}
-18. Stone, L. The feasibility and stability of large complex biological
+21. Stone, L. The feasibility and stability of large complex biological
 networks: a random matrix approach. \emph{Scientific Reports}
@@
 \hypertarget{ref-Hamblin2013}{}
-19. Hamblin, S. On the practical usage of genetic algorithms in ecology
+22. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
@@
 \hypertarget{ref-Dougoud2018}{}
-20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+23. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-21. Song, C. \& Saavedra, S. Will a small randomly assembled community
+24. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-22. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+25. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 \hypertarget{ref-May1973}{}
-23. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+26. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.723 and 0.737 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.736 for a and b,
 respectively.

### 2020-02-23 notebook/ms.Rmd
Large reworking of the first paragraph of the results, more focus on analytical equations

@@
 **Component response rates of random complex systems**.
-Rows in $\mathbf{M}$ define how a given component $i$ is affected by other components of the system, meaning that the rate of component response time can be modelled by multiplying all row elements by a real scalar value $\gamma_{i}$ [@Patel2018].
+Complex systems $\mathbf{M}$ are built from two matrices, one modelling component interactions ($\mathbf{A}$), and second modelling component response rates ($\gamma$).
+Both $\mathbf{A}$ and $\gamma$ are square $S \times S$ matrices.
+Rows in $\mathbf{A}$ define how a given component $i$ is affected by each component $j$ in the system, including itself (where $i = j$).
+Off-diagonal elments of $\mathbf{A}$ are independent and identically distributed (i.i.d), and diagonal elements are set to $A_{ii} = -1$ as in May [@May1972].
+Diagonal elements of $\gamma$ are positive, and off-diagonal elements are set to zero (i.e.
+$\gamma$ is a diagonal matrix with positive support).
 The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+The dynamics of the entire system $\mathbf{M}$ can be defined as follows [@Patel2018],
+
+\begin{equation} \label{defM}
+M = \gamma A.
+\end{equation}
+
+Equation \ref{defM} thereby serves as a null model to investigate how varying component response rates affect complex systems.
+
  <!--- Replace the below with some more informative text concerning the variation that $\var(gamma})$ adds to the system $M$ overall.
@@
 --->
-
-<!--- Remove this instructive example; replace it with something better. This entire paragraph and the code underneath it can be deleted. --->
-An instructive example compares one $\mathbf{M}$ where $\gamma_{i} = 1$ for all $i$ in $S$ to the same $\mathbf{M}$ when half of $\gamma_{i} = 1.95$ and half of $\gamma_{i} = 0.05$.
-This models one system in which $\gamma$ is invariant and one in which $\gamma$ varies, but systems are otherwise identical (note that mean $\gamma_{i} = 1$ in both cases).
-I assume $S = 200$, $C = 0.05$, and $\sigma = 0.4$; diagonal elements are set to $-1$ and non-zero off-diagonal elements are drawn randomly from $\mathcal{N}(0, \sigma^{2})$.
-Rows are then multiplied by $\gamma_{i}$ to generate $\mathbf{M}$.
-When $\gamma_{i} = 1$, eigenvalues of $\mathbf{M}$ are distributed uniformly within a circle centred at ($-1, 0$) with a radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig.
-1a).
-Hence, the real components of eigenvalues are highly unlikely to all be negative when all $\gamma_{i} = 1$.
-But when $\gamma_{i}$ values are separated into two groups, eigenvalues are no longer uniformly distributed (Fig.
-1b).
-Instead, two distinct clusters of eigenvalues appear (red circles in Fig.
-1b), one centred at ($-1.95, 0$) and the other centred at ($-0.05, 0$).
-The former has a large radius, but the real components have shifted to the left (in comparison to when $\gamma = 1$), while the latter cluster has a smaller radius, but real components have shifted to the right; all $\Re({\lambda}) < 0$.
-Overall, for 1 million randomly assembled $\mathbf{M}$, this division between slow and fast component response rates results in more stable systems: `r sum(pr_st[,1])` stable given $\gamma = 1$ versus `r sum(pr_st[,2])` stable given $\gamma = \{1.95, 0.5\}$.
 

### 2020-02-23 notebook/ms.tex
Large reworking of the first paragraph of the results, more focus on analytical equations

@@
 
-\textbf{Component response rates of random complex systems}. Rows in
-\(\mathbf{M}\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
+\textbf{Component response rates of random complex systems}. Complex
+systems \(\mathbf{M}\) are built from two matrices, one modelling
+component interactions (\(\mathbf{A}\)), and second modelling component
+response rates (\(\gamma\)). Both \(\mathbf{A}\) and \(\gamma\) are
+square \(S \times S\) matrices. Rows in \(\mathbf{A}\) define how a
+given component \(i\) is affected by each component \(j\) in the system,
+including itself (where \(i = j\)). Off-diagonal elments of
+\(\mathbf{A}\) are independent and identically distributed (i.i.d), and
+diagonal elements are set to \(A_{ii} = -1\) as in
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}. Diagonal
+elements of \(\gamma\) are positive, and off-diagonal elements are set
+to zero (i.e. \(\gamma\) is a diagonal matrix with positive support).
 The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates.
+distribution of component response rates. The dynamics of the entire
+system \(\mathbf{M}\) can be defined
+as\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}},
 
-An instructive example compares one \(\mathbf{M}\) where
-\(\gamma_{i} = 1\) for all \(i\) in \(S\) to the same \(\mathbf{M}\)
-when half of \(\gamma_{i} = 1.95\) and half of \(\gamma_{i} = 0.05\).
-This models one system in which \(\gamma\) is invariant and one in which
-\(\gamma\) varies, but systems are otherwise identical (note that mean
-\(\gamma_{i} = 1\) in both cases). I assume \(S = 200\), \(C = 0.05\),
-and \(\sigma = 0.4\); diagonal elements are set to \(-1\) and non-zero
-off-diagonal elements are drawn randomly from
-\(\mathcal{N}(0, \sigma^{2})\). Rows are then multiplied by
-\(\gamma_{i}\) to generate \(\mathbf{M}\). When \(\gamma_{i} = 1\),
-eigenvalues of \(\mathbf{M}\) are distributed uniformly within a circle
-centred at (\(-1, 0\)) with a radius of 1.265 (Fig. 1a). Hence, the real
-components of eigenvalues are highly unlikely to all be negative when
-all \(\gamma_{i} = 1\). But when \(\gamma_{i}\) values are separated
-into two groups, eigenvalues are no longer uniformly distributed (Fig.
-1b). Instead, two distinct clusters of eigenvalues appear (red circles
-in Fig. 1b), one centred at (\(-1.95, 0\)) and the other centred at
-(\(-0.05, 0\)). The former has a large radius, but the real components
-have shifted to the left (in comparison to when \(\gamma = 1\)), while
-the latter cluster has a smaller radius, but real components have
-shifted to the right; all \(\Re({\lambda}) < 0\). Overall, for 1 million
-randomly assembled \(\mathbf{M}\), this division between slow and fast
-component response rates results in more stable systems: 1 stable given
-\(\gamma = 1\) versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
+\begin{equation} \label{defM}
+M = \gamma A.
+\end{equation}
+
+Equation \ref{defM} thereby serves as a null model to investigate how
+varying component response rates affect complex systems.
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.736 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.738 for a and b,
 respectively.

### 2020-03-02 notebook/ms.Rmd
Edit the notation a bit in the Introduction to make the results easier to read. Also, add some more analytical explanation to the start of the results

@@
 One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
-Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to -1 [@May1972; @Allesina2012; @Allesina2015].
+Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation ($d$, e.g., carrying capacity in ecological communities).
+Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma_{M}^{2}$; diagonal elements are set to $d = -1$ [@May1972; @Allesina2012; @Allesina2015].
 Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
-Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma^{2}$ increase.
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S_{M}$, $C_{M}$, and $\sigma_{M}^{2}$ increase.
 
-May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < 1$ assumes that the expected response rates ($\gamma$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma_{M}\sqrt{SC} < 1$ assumes that the expected response rates ($\gamma$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
@@
 The effect of such variance on stability has not been investigated in complex systems theory.
-Intuitively, variation in $\gamma$ might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
-Here I show that, despite higher $\sigma$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
+Intuitively, variation in $\gamma$ ($\sigma^{2}_{\gamma}$) might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma_{M}$.
+Here I show that, despite higher $\sigma_{M}$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
 My results are robust across commonly observed network structures, including random [@May1972], small-world [@Watts1998], scale-free [@Albert2002], cascade food web [@Williams2000] networks.
@@
 
-Equation \ref{defM} thereby serves as a null model to investigate how varying component response rates affect complex systems.
+Equation \ref{defM} thereby serves as a null model to investigate how variation in component response rate ($\sigma_{\gamma}$) affects complex systems.
+In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\textbf{M} = \textbf{A}$.
+For this baseline condition, $\max(\Re(\lambda))$ can be estimated from five values [@Tang2014b].
+These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+
+When investigating the effect of varying component response rate $\sigma_{\gamma}$ on stability by defining $\textbf{M} = \gamma\textbf{A}$, $S$ remains unchanged.
+Further, values of $\gamma_{i}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
+What $\sigma_{\gamma}$ does change is the variation in component interaction strengths, and $\rho$.
+Introducing variation in $\gamma$ increases the total variation in the system, making it more likely that $\textbf{M} = \gamma\textbf{A}$ is unstable.
+Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables (to simplify the notation, $\sigma^{2}_{M}$ and $\sigma^{2}_{A}$ refer to variances of off-diagonal elements only),
+
+\begin{equation} \label{var_full}
+\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.
+\end{equation}
+
+Note that in my simulations $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, so equation \ref{var_full} can be simplified,
+
+\begin{equation} \label{var_reduced}
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\end{equation}
+
+The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$.
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{M}$ and $\textbf{A}$ lie within a circle of a radius $\sigma^{2}_{M}\sqrt{S}$ and $\sigma^{2}_{A}\sqrt{S}$, respectively.
 

### 2020-03-02 notebook/ms.tex
Edit the notation a bit in the Introduction to make the results easier to read. Also, add some more analytical explanation to the start of the results

@@
 distribution of component response rates. The dynamics of the entire
-system \(\mathbf{M}\) can be defined
-as\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}},
+system \(\mathbf{M}\) can be defined as
+follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}},
 
@@
 Equation \ref{defM} thereby serves as a null model to investigate how
-varying component response rates affect complex systems.
+variation in component response rate (\(\sigma_{\gamma}\)) affect
+complex systems. In the absence of such variation
+(\(\sigma_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
+(diagonal elements all equal 1) and \(\textbf{M} = \textbf{A}\). Given
+this baseline condition, \(\max(\Re(\lambda))\) for
+\(\textbf{M} = \textbf{A}\) can be estimated from five
+values\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}. These
+values include (1) system size (\(S\)), (2) mean self-regulation of
+components (\(d\); i.e., mean diagonal element of \(\textbf{M}\)), (3)
+mean interaction strength between components (\(\mu\)), (4) the standard
+deviation between component interaction strengths (\(\sigma\)), and (5)
+the correlation of interaction strengths between components, \(M_{ij}\)
+and \(M_{ji}\) (\(\rho\)).
+
+When investigating the effect of varying component response rate
+\(\sigma_{\gamma}\) on stability by defining
+\(\textbf{M} = \gamma\textbf{A}\), \(S\) remains unchanged. Further,
+values of \(\gamma_{i}\) were sampled such that \(E[d]\) and \(E[\mu]\)
+also remained unchanged (in practice, diagonal elements of
+\(\textbf{M}\) were standardised so that mean values were identical
+before and after adding \(\gamma\)). What \(\sigma_{\gamma}\) does
+change is the variation in component interaction strengths, and
+\(\rho\). Introducing variation in \(\gamma\) increases the total
+variation in the system, making it more likely that
+\(\textbf{M} = \gamma\textbf{A}\) is unstable. Variation of the
+off-diagonal elements in \(\textbf{M}\) is described by the joint
+variation of two random variables (to simplify the notation,
+\(\sigma^{2}_{M}\) and \(\sigma^{2}_{A}\) refer to variances of
+off-diagonal elements only),
+
+\begin{equation} \label{var_full}
+\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.
+\end{equation}
+
+Note that in my simulations \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\),
+so equation \ref{var_full} can be simplified,
+
+\begin{equation} \label{var_reduced}
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\end{equation}
+
+The increase caused by \(\sigma^{2}_\gamma\) can be visualised from the
+eigenvalue spectra of \(\textbf{A}\) versus
+\(\textbf{M} = \gamma\textbf{A}\). Given \(d = 0\) and \(C = 1\), the
+distribution of eigenvalues of \(\textbf{M}\) and \(\textbf{A}\) lie
+within a circle of a radius \(\sigma^{2}_{M}\sqrt{S}\) and
+\(\sigma^{2}_{A}\sqrt{S}\), respectively.
 
@@
 Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{19}--\protect\hyperlink{ref-Stone2017}{21}},
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}--\protect\hyperlink{ref-Stone2017}{22}},
 some of which are clustered locally near the centre of the distribution,
@@
 values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{22}}.
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{23}}.
 Generations of selection on \(\gamma\) value combinations to minimise
@@
 positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{23},\protect\hyperlink{ref-Song2018}{24}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{24},\protect\hyperlink{ref-Song2018}{25}}.
 This is of particular interest for ecological communities because
@@
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}}.
 While my results are intended to be general to all complex systems, and
@@
 a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{25}}
+Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{26}}
 (see Supplementary Information). Hence, for pure interacting species
@@
 My focus is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20}}, who applied
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21}}, who applied
 the same mathematical framework to investigate how a diagonal matrix
@@
 \(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20}} analytically
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21}} analytically
 demonstrated that the effect of \(\mathbf{X}\) on system stability
@@
 of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{21},\protect\hyperlink{ref-Dougoud2018}{23},\protect\hyperlink{ref-Servan2018}{25}}
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{22},\protect\hyperlink{ref-Dougoud2018}{24},\protect\hyperlink{ref-Servan2018}{26}}
 because population densities cannot realistically be negative. My
@@
 distributions caused by a vector \(\gamma\) has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{19}}
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}}
 and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{20},\protect\hyperlink{ref-Stone2017}{21}},
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21},\protect\hyperlink{ref-Stone2017}{22}},
 but \(\gamma\) has not been interpreted as rates of response of
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{26}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{27}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{26}},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{27}},
 
@@
 components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{19},\protect\hyperlink{ref-May1973}{26}}.
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{20},\protect\hyperlink{ref-May1973}{27}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{22}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{23}}. In
 this case, the practical solution is finding vectors of
@@
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{23}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{25}}),
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{26}}),
 
@@
 
+\hypertarget{ref-Tang2014b}{}
+19. Tang, S. \& Allesina, S. Reactivity and stability of large
+ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
+(2014).
+
 \hypertarget{ref-Ahmadian2015}{}
-19. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+20. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
 with partially structured and partially random connectivity.
@@
 \hypertarget{ref-Gibbs2017}{}
-20. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+21. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.
@@
 \hypertarget{ref-Stone2017}{}
-21. Stone, L. The feasibility and stability of large complex biological
+22. Stone, L. The feasibility and stability of large complex biological
 networks: a random matrix approach. \emph{Scientific Reports}
@@
 \hypertarget{ref-Hamblin2013}{}
-22. Hamblin, S. On the practical usage of genetic algorithms in ecology
+23. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
@@
 \hypertarget{ref-Dougoud2018}{}
-23. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+24. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-24. Song, C. \& Saavedra, S. Will a small randomly assembled community
+25. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-25. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+26. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 \hypertarget{ref-May1973}{}
-26. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+27. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.738 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.718 and 0.742 for a and b,
 respectively.

### 2020-03-02 notebook/ms.Rmd
Rearrange some of the text to improve the flow between analytical and numerical simulation sections of the Results

@@
 Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
 Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S_{M}$, $C_{M}$, and $\sigma_{M}^{2}$ increase.
@@
 In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\textbf{M} = \textbf{A}$.
-For this baseline condition, $\max(\Re(\lambda))$ can be estimated from five values [@Tang2014b].
-These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
 
-When investigating the effect of varying component response rate $\sigma_{\gamma}$ on stability by defining $\textbf{M} = \gamma\textbf{A}$, $S$ remains unchanged.
-Further, values of $\gamma_{i}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
-What $\sigma_{\gamma}$ does change is the variation in component interaction strengths, and $\rho$.
-Introducing variation in $\gamma$ increases the total variation in the system, making it more likely that $\textbf{M} = \gamma\textbf{A}$ is unstable.
-Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables (to simplify the notation, $\sigma^{2}_{M}$ and $\sigma^{2}_{A}$ refer to variances of off-diagonal elements only),
+**Analytical considerations**.
+The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\textbf{M}$ [@Tang2014b].
+These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+Positive $\sigma_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$.
+What $\sigma_{\gamma}$ does change is the total variation in component interaction strengths ($\sigma_{M}$), and $\rho$.
+Introducing variation in $\gamma$ increases the total variation in the system.
+Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables,
 
@@
 
-Note that in my simulations $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, so equation \ref{var_full} can be simplified,
+Given $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, equation \ref{var_full} can be simplified,
 
@@
 
-The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$.
-Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{M}$ and $\textbf{A}$ lie within a circle of a radius $\sigma^{2}_{M}\sqrt{S}$ and $\sigma^{2}_{A}\sqrt{S}$, respectively.
+The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Figure 1).
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{A}$ and $\textbf{M}$ lie within a circle of a radius $\sigma^{2}_{A}\sqrt{S}$ and $\sigma^{2}_{M}\sqrt{S}$, respectively (Figure 1a vs.
+1b).
+When components are self-regulating (e.g., $d = -1$), positive $\sigma^{2}_\gamma$ changes the distribution of eigenvalues, potentially affecting stability (Figure 1c vs.
+1d).
+
+Given $\sigma^{2}_\gamma = 0$, $\max(\Re(\lambda))$ decrease linearly with $\rho$ such that,
+
+\begin{equation} \label{rho_stab}
+\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\end{equation}
+
+When $\rho$ values are negative, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability is expected to increase [@Allesina2012].
+When elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the covariance between $M_{ij}$ and $M_{ji}$,
+
+\begin{equation} \label{rho_def}
+\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.
+\end{equation}
+
+Because $\gamma$ does not affect $cov(M_{ij}, M_{ji})$, the sign of $\rho$ remains unchanged.
+
+**Numerical simulations**.
+I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on the stability of $\textbf{M} = \gamma \textbf{A}$ by comparing $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$.
+Values of $\gamma_{ii}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
 

### 2020-03-02 notebook/ms.tex
Rearrange some of the text to improve the flow between analytical and numerical simulation sections of the Results

@@
 between components, while diagonal elements (\(i = j\)) define component
-self-regulation (e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a
-probability \(C\), which are sampled from a distribution with variance
-\(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
+self-regulation (\(d\), e.g., carrying capacity in ecological
+communities). Traditionally, off-diagonal elements are assigned non-zero
+values with a probability \(C\), which are sampled from a distribution
+with variance \(\sigma_{M}^{2}\); diagonal elements are set to
+\(d = -1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
 Local system stability is assessed using eigenanalysis, with the system
@@
 \(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
+\(\sigma_{M}\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma^{2}\) increase.
+therefore becomes increasingly unlikely as \(S_{M}\), \(C_{M}\), and
+\(\sigma_{M}^{2}\) increase.
 
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
-response rates (\(\gamma\)) of individual components to perturbations of
-the system are identical, but this is highly unlikely in any complex
-system. In ecological communities, for example, the rate at which
-population density changes following perturbation will depend on the
-generation time of organisms, which might vary by orders of magnitude
-among species. Species with short generation times will respond quickly
-(high \(\gamma\)) to perturbations relative to species with long
-generation times (low \(\gamma\)). Similarly, the speed at which
-individual banks respond to perturbations in financial networks, or
-individuals or institutions respond to perturbations in complex social
-networks, is likely to vary. The effect of such variance on stability
-has not been investigated in complex systems theory. Intuitively,
-variation in \(\gamma\) might be expected to decrease system stability
-by introducing a new source of variation into the system and thereby
-increasing \(\sigma\). Here I show that, despite higher \(\sigma\),
-realistic complex systems (such that \(S\) is high but finite) are
-actually more likely to be stable if their individual component response
-rates vary. My results are robust across commonly observed network
-structures, including
+stability criterion \(\sigma_{M}\sqrt{SC} < 1\) assumes that the
+expected response rates (\(\gamma\)) of individual components to
+perturbations of the system are identical, but this is highly unlikely
+in any complex system. In ecological communities, for example, the rate
+at which population density changes following perturbation will depend
+on the generation time of organisms, which might vary by orders of
+magnitude among species. Species with short generation times will
+respond quickly (high \(\gamma\)) to perturbations relative to species
+with long generation times (low \(\gamma\)). Similarly, the speed at
+which individual banks respond to perturbations in financial networks,
+or individuals or institutions respond to perturbations in complex
+social networks, is likely to vary. The effect of such variance on
+stability has not been investigated in complex systems theory.
+Intuitively, variation in \(\gamma\) (\(\sigma^{2}_{\gamma}\)) might be
+expected to decrease system stability by introducing a new source of
+variation into the system and thereby increasing \(\sigma_{M}\). Here I
+show that, despite higher \(\sigma_{M}\), realistic complex systems
+(such that \(S\) is high but finite) are actually more likely to be
+stable if their individual component response rates vary. My results are
+robust across commonly observed network structures, including
 random\textsuperscript{\protect\hyperlink{ref-May1972}{1}},
@@
 Equation \ref{defM} thereby serves as a null model to investigate how
-variation in component response rate (\(\sigma_{\gamma}\)) affect
+variation in component response rate (\(\sigma_{\gamma}\)) affects
 complex systems. In the absence of such variation
 (\(\sigma_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
-(diagonal elements all equal 1) and \(\textbf{M} = \textbf{A}\). Given
-this baseline condition, \(\max(\Re(\lambda))\) for
-\(\textbf{M} = \textbf{A}\) can be estimated from five
-values\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}. These
-values include (1) system size (\(S\)), (2) mean self-regulation of
-components (\(d\); i.e., mean diagonal element of \(\textbf{M}\)), (3)
-mean interaction strength between components (\(\mu\)), (4) the standard
-deviation between component interaction strengths (\(\sigma\)), and (5)
-the correlation of interaction strengths between components, \(M_{ij}\)
-and \(M_{ji}\) (\(\rho\)).
+(diagonal elements all equal 1) and \(\textbf{M} = \textbf{A}\).
 
-When investigating the effect of varying component response rate
-\(\sigma_{\gamma}\) on stability by defining
-\(\textbf{M} = \gamma\textbf{A}\), \(S\) remains unchanged. Further,
-values of \(\gamma_{i}\) were sampled such that \(E[d]\) and \(E[\mu]\)
-also remained unchanged (in practice, diagonal elements of
-\(\textbf{M}\) were standardised so that mean values were identical
-before and after adding \(\gamma\)). What \(\sigma_{\gamma}\) does
-change is the variation in component interaction strengths, and
-\(\rho\). Introducing variation in \(\gamma\) increases the total
-variation in the system, making it more likely that
-\(\textbf{M} = \gamma\textbf{A}\) is unstable. Variation of the
-off-diagonal elements in \(\textbf{M}\) is described by the joint
-variation of two random variables (to simplify the notation,
-\(\sigma^{2}_{M}\) and \(\sigma^{2}_{A}\) refer to variances of
-off-diagonal elements only),
+\textbf{Analytical considerations}. The value of \(\max(\Re(\lambda))\),
+and therefore system stability, can be estimated from five properties of
+\(\textbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}.
+These properties include (1) system size (\(S\)), (2) mean
+self-regulation of components (\(d\)), (3) mean interaction strength
+between components (\(\mu\)), (4) the standard deviation between
+component interaction strengths (\(\sigma_{M}\)), and (5) the
+correlation of interaction strengths between components, \(M_{ij}\) and
+\(M_{ji}\) (\(\rho\)). Positive \(\sigma_{\gamma}\) does not change
+\(S\), nor does it necessarily change \(E[d]\) or \(E[\mu]\). What
+\(\sigma_{\gamma}\) does change is the total variation in component
+interaction strengths (\(\sigma_{M}\)), and \(\rho\). Introducing
+variation in \(\gamma\) increases the total variation in the system.
+Variation of the off-diagonal elements in \(\textbf{M}\) is described by
+the joint variation of two random variables,
 
@@
 
-Note that in my simulations \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\),
-so equation \ref{var_full} can be simplified,
+Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), equation
+\ref{var_full} can be simplified,
 
@@
 eigenvalue spectra of \(\textbf{A}\) versus
-\(\textbf{M} = \gamma\textbf{A}\). Given \(d = 0\) and \(C = 1\), the
-distribution of eigenvalues of \(\textbf{M}\) and \(\textbf{A}\) lie
-within a circle of a radius \(\sigma^{2}_{M}\sqrt{S}\) and
-\(\sigma^{2}_{A}\sqrt{S}\), respectively.
+\(\textbf{M} = \gamma\textbf{A}\) (Figure 1). Given \(d = 0\) and
+\(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
+\(\textbf{M}\) lie within a circle of a radius
+\(\sigma^{2}_{A}\sqrt{S}\) and \(\sigma^{2}_{M}\sqrt{S}\), respectively
+(Figure 1a vs.~1b). When components are self-regulating (e.g.,
+\(d = -1\)), positive \(\sigma^{2}_\gamma\) changes the distribution of
+eigenvalues, potentially affecting stability (Figure 1c vs.~1d).
+
+Given \(\sigma^{2}_\gamma = 0\), \(\max(\Re(\lambda))\) decrease
+linearly with \(\rho\) such that,
+
+\begin{equation} \label{rho_stab}
+\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\end{equation}
+
+When \(\rho\) values are negative, such as when \(\textbf{M}\) models a
+predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have opposing
+signs, stability is expected to
+increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
+elements of \(\gamma\) vary independently, the magnitude of \(\rho\) is
+decreased because \(\sigma^{2}_{\gamma}\) increases the variance of
+\(M_{ij}\) without affecting the covariance between \(M_{ij}\) and
+\(M_{ji}\),
+
+\begin{equation} \label{rho_def}
+\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.
+\end{equation}
+
+Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
+\(\rho\) remains unchanged.
+
+\textbf{Numerical simulations}. I used numerical simulations to isolate
+the effect of \(\sigma_{\gamma}\) on the stability of
+\(\textbf{M} = \gamma \textbf{A}\) by comparing \(\sigma_{\gamma} > 0\)
+versus \(\sigma_{\gamma} = 0\). Values of \(\gamma_{ii}\) were sampled
+such that \(E[d]\) and \(E[\mu]\) also remained unchanged (in practice,
+diagonal elements of \(\textbf{M}\) were standardised so that mean
+values were identical before and after adding \(\gamma\)).
 
@@
 both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.718 and 0.742 for a and b,
+Proportions of \(\Re(\lambda) < 0\) are 0.727 and 0.733 for a and b,
 respectively.

### 2020-03-02 notebook/ms.Rmd
Reworking the model so that simulations are introduced in a more seemless way

@@
 In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\textbf{M} = \textbf{A}$.
+Under these conditions, eigenvalues of $\textbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Figure 1a).
 
@@
 1b).
-When components are self-regulating (e.g., $d = -1$), positive $\sigma^{2}_\gamma$ changes the distribution of eigenvalues, potentially affecting stability (Figure 1c vs.
+If $d \neq 0$, positive $\sigma^{2}_\gamma$ changes the distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], potentially affecting stability (Figure 1c vs.
 1d).
 
-Given $\sigma^{2}_\gamma = 0$, $\max(\Re(\lambda))$ decrease linearly with $\rho$ such that,
+Given $\sigma^{2}_\gamma = 0$, $\max(\Re(\lambda))$ decrease linearly with $\rho$ such that [@Tang2014c],
 
@@
 
-When $\rho$ values are negative, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability is expected to increase [@Allesina2012].
-When elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the covariance between $M_{ij}$ and $M_{ji}$,
+If $\rho < 0$, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability increases [@Allesina2012].
+If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$,
 
@@
 
-Because $\gamma$ does not affect $cov(M_{ij}, M_{ji})$, the sign of $\rho$ remains unchanged.
+Neverthless, numerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
 
 **Numerical simulations**.
-I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on the stability of $\textbf{M} = \gamma \textbf{A}$ by comparing $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$.
-Values of $\gamma_{ii}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
+I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$.
+Values of $\gamma_{ii}$ were sampled from a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
+For $\sigma_{\gamma} = 0$, diagonal elements were standardised to the mean value of sampled $\gamma_{ii}$ so that $E[d]$ of $\textbf{M}$ were identical.
 
@@
 --------------------------------------------------------------------------------
+
+<!---  (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$) --->
 

### 2020-03-02 notebook/ms.Rmd
Remove a paragraph that has now been made redundant

@@
 **Numerical simulations**.
-I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$.
+I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$ across a rane of parameter combinations.
 Values of $\gamma_{ii}$ were sampled from a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
@@
 ```
-
-<!--- Some of this paragraph can be removed.
-Some of the middle can focus on the new Figure 1 panels C and D, which illustrate the effect of gamma on the eigenvalue spectra.
-The rest, focusing on the results, can be retained.--->
-Higher stability in systems with variation in $\gamma$ can be observed by sampling $\gamma_{i}$ values from various distributions.
-I focus on a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
-<!--- Cut from here ---> As with the case of $\gamma = \{1.95, 0.5\}$ (Fig.
-1b), mean $\gamma = 1$ when $\gamma \sim \mathcal{U}(0, 2)$, allowing comparison of $\mathbf{M}$ before and after the addition of variation in component response rate.
-Figure 2 shows a comparison of eigenvalue distributions given $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
-<!--- end cut here, but now refer to Figure 1 panels C and D ---> As expected [@Tao2010], when $\gamma = 1$, eigenvalues are distributed uniformly in a circle centred at ($-1, 0$) with a radius of $\sigma\sqrt{SC} =$ `r round(sigma*sqrt(S*C), digits = 3)`.
-Uniform variation in $\gamma$ leads to a non-uniform distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], some of which are clustered locally near the centre of the distribution, but others of which are spread outside the former radius of `r round(sigma*sqrt(S*C), digits = 3)` (Fig 2b).
-The clustering and spreading of eigenvalues introduced by $Var(\gamma)$ can destabilise previously stable systems or stabilise systems that are otherwise unstable.
-But where systems are otherwise too complex to be stable given $\gamma = 1$, the effect of $Var(\gamma)$ can often lead to stability above May's [@May1972; @Allesina2012] threshold $\sigma\sqrt{SC} < 1$.
-<!--- Okay, so most can be retained --->
 

### 2020-03-02 notebook/revision_notes.Rmd
Small update, and fix column in very old data set mentioned in response to Reviewer 3
Update the revision notes with a better figure

@@
 $$\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.$$
+
 Because $\gamma$ does not affect $cov(M_{ij}, M_{ji})$, the sign of $\rho$ remains unchanged.
@@
 plot(x = sim20[,1], y = sim20[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), cex.axis = 1.5);
+     ylim = c(-1, 1.4), cex.axis = 1.5);
 points(x = sim20[,15], y = sim20[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 20", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 20", cex = 2.5);
 plot(x = sim25[,1], y = sim25[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.5);
 points(x = sim25[,15], y = sim25[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 25", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 25", cex = 2.5);
 plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), cex.axis = 1.5);
+     ylim = c(-1, 1.4), cex.axis = 1.5);
 points(x = sim30[,15], y = sim30[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 30", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 30", cex = 2.5);
 plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.25);
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.25);
 points(x = sim35[,15], y = sim35[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 35", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 35", cex = 2.5);
 plot(x = sim40[,1], y = sim40[,20], type = "l", pch = 20, lwd = 2, 
-     ylim = c(-1, 3.5), cex.axis = 1.5);
+     ylim = c(-1, 1.4), cex.axis = 1.5);
 points(x = sim40[,15], y = sim40[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 40", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 40", cex = 2.5);
 plot(x = sim45[,1], y = sim45[,20], type = "l", pch = 20, lwd = 2, 
-     ylim = c(-1, 3.5), yaxt = "n", cex.axis = 1.5);
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.5);
 points(x = sim45[,15], y = sim45[,21], type = "l", lwd = 2, lty = "dashed");
 abline(h = 0, col = "red", lty = "dotted", lwd = 0.8);
-text(x = -0.6, y = 3.2, labels = "S = 45", cex = 2.5);
+text(x = -0.6, y = 1.2, labels = "S = 45", cex = 2.5);
 mtext(text =  expression(paste("E Correlation between A"[ij]," and A"[ji], 

### 2020-03-02 notebook/ms.Rmd
Replace Figure 1 with the four panel figure from the notebook

@@
 
-**Analytical considerations**.
+**Analytical considerations of $\textbf{\gamma}$ effects on \textbf{M}$**.
 The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\textbf{M}$ [@Tang2014b].
@@
 
-**Numerical simulations**.
+**Numerical simulations of random $\textbf{M}$ with and without $\textbf{\sigma_{\gamma}}$**.
 I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$ across a rane of parameter combinations.
@@
 For $\sigma_{\gamma} = 0$, diagonal elements were standardised to the mean value of sampled $\gamma_{ii}$ so that $E[d]$ of $\textbf{M}$ were identical.
+
+**Simulation of random $\mathbf{M}$ across $\mathbf{\rho}$**
 
@@
 
-```{r echo = FALSE}
-S     <- 200;
-C     <- 0.05;
-sigma <- 0.4;
-pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st <- pr_st[,-1];
-```
+**Figure 1: Insert a new legend here**
 
-```{r, echo = FALSE}
-A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
-A0 <- as.matrix(A0[,-1]);
-A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
-A1 <- as.matrix(A1[,-1]);
+```{r, fig.width = 7, fig.height = 7, echo = FALSE}
+S     <- 400;
+C     <- 1;
+dval  <- 0;
+sigma <- 1/(1*sqrt(S));
+a     <- 0;
+b     <- 2;
+mn    <- 0;
 
+A_dat  <- rnorm(n = S * S, mean = mn, sd = sigma);
+A_mat  <- matrix(data = A_dat, nrow = S);
+C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+A_mat  <- A_mat * C_mat;
+gammas <- runif(n = S, min = a, max = b);
+vargam <- (1/12) * (b-a)^2;
+
+diag(A_mat) <- dval;
+A1     <- gammas * A_mat;
+A0     <- A_mat;
+A0     <- A0 * mean(gammas);
 A0_e   <- eigen(A0)$values;
@@
 
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(t(A0_vm));
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(t(A1_vm));
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
-fhalf       <- 1:(0.5*length(A1vec));
-shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st       <- pr_st[,-1];
-```
+joint_var <- ((sigma^2) * vargam) + ((sigma^2) * mean(gammas)^2) + 
+             (vargam * (dval*mn)^2);
 
-**Figure 1: Example distribution of eigenvalues before (a) and after (b) separating a randomly generated complex system into fast ($\boldsymbol{\gamma} = 1.95$) and slow ($\boldsymbol{\gamma} = 0.05$) component response rates.** Each panel shows the same system where $S = 200$, $C = 0.05$, and $\sigma = 0.4$, and in each case $E[\gamma] = 1$ (i.e., only the distribution of $\gamma$ differs between panels).
-**a.** Eigenvalues plotted when all $\gamma = 1$; distributions of points are uniformly distributed within the blue circle with a radius of $\sigma\sqrt{SC} =$ `r round(sqrt(200) * sd(A0vec), digits =3)` centred at -1 on the real axis.
-**b.** Eigenvalues plotted when half $\gamma = 1.95$ and half $\gamma = 0.05$; distributions of points can be partitioned into one large circle centred at $\gamma = -1.95$ and one small circle centred at $\gamma = -0.05$.
-In a, the maximum real eigenvalue $\max\left(\Re(\lambda)\right) =$ `r format(max(A0_r), scientific = FALSE)`, while in b $\max\left(\Re(\lambda)\right) =$ `r format(max(A1_r), scientific = FALSE)`, meaning that the complex system in b but not a is stable because in b $\max\left(\Re(\lambda)\right) < 0$.
-In 1 million randomly generated complex systems under the same parameter values, `r sum(pr_st[,1])` was stable when $\gamma = 1$ while `r sum(pr_st[,2])` were stable when $\gamma = \{1.95, 0.05\}$.
-Overall, complex systems that are separated into fast versus slow components tend to be more stable than otherwise identical systems with identical component response rates.
+par(mfrow = c(2, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));
+plot(A0_r, A0_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "dodgerblue4", xlim = c(-3, 1.5),
+     xaxt = "n", ylim = c(-1.5, 1.5));
+vl        <- seq(from = 0, to = 2*pi, by = 0.001);
+A0x1a     <- mean(gammas) * sigma * sqrt(S) * cos(vl) + mean(diag(A0));
+A0y1a     <- mean(gammas) * sigma * sqrt(S) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "dodgerblue4");
+text(x = -2.75, y = 1.75, labels = "a", cex = 2);
 
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "dodgerblue4");
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x0 <- sqrt(200) * sd(A0vec) * cos(vl) + mean(diag(A0));
-A0y0 <- sqrt(200) * sd(A0vec) * sin(vl);
-text(x = -3.5, y = 2.25, labels = "a", cex = 2);
-points(x = A0x0, y = A0y0, type = "l", lwd = 3, col = "blue");
-points(A0_r, A0_i, pch = 4, cex = 0.7, col = "dodgerblue4");
+plot(A1_r, A1_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "firebrick", xlim = c(-3, 1.5),
+     xaxt = "n", yaxt = "n", ylim = c(-1.5, 1.5));
+crad      <- sqrt(joint_var) * sqrt(S);
+A1x1a     <- crad * cos(vl) + mean(diag(A1));
+A1y1a     <- crad * sin(vl);
+points(x = A1x1a, y = A1y1a, type = "l", lwd = 3, col = "firebrick");
+text(x = -2.75, y = 1.75, labels = "b", cex = 2);
 
-plot(A1_r, A1_i, xlim = c(-3.7, 0.3), ylim = c(-2, 2), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "firebrick", yaxt = "n");
+calc_var <- A1;
+diag(calc_var) <- NA;
+calc_var  <- as.numeric(calc_var);
+cvar1 <- var(calc_var[!is.na(calc_var)]);
 
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-A0x1a <- sqrt(100) * sd(A1vec[fhalf]) * cos(vl) + mean(diag(A1)[1:100]);
-A0y1a <- sqrt(100) * sd(A1vec[fhalf]) * sin(vl);
-points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "red");
-A0x1b <- sqrt(100) * sd(A1vec[shalf]) * cos(vl) + mean(diag(A1)[101:200]);
-A0y1b <- sqrt(100) * sd(A1vec[shalf]) * sin(vl);
-points(x = A0x1b, y = A0y1b, type = "l", lwd = 3, col = "red");
+################################################################################
+################################################################################
 
-points(A1_r[1:100], A1_i[1:100],pch = 4, cex = 0.7, col = "firebrick");
+S     <- 400;
+C     <- 1;
+dval  <- -1;
+sigma <- 1/(1*sqrt(S));
+a     <- 0;
+b     <- 2;
+mn    <- 0;
 
-text(x = -3.5, y = 2.25, labels = "b", cex = 2);
+A_dat  <- rnorm(n = S * S, mean = mn, sd = sigma);
+A_mat  <- matrix(data = A_dat, nrow = S);
+C_dat  <- rbinom(n = S * S, size = 1, prob = C);
+C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
+A_mat  <- A_mat * C_mat;
+gammas <- runif(n = S, min = a, max = b);
+vargam <- (1/12) * (b-a)^2;
+
+diag(A_mat) <- dval;
+A1     <- gammas * A_mat;
+A0     <- A_mat;
+A0     <- A0 * mean(gammas);
+A0_e   <- eigen(A0)$values;
+A0_r   <- Re(A0_e);
+A0_i   <- Im(A0_e);
+A1_e   <- eigen(A1)$values;
+A1_r   <- Re(A1_e);
+A1_i   <- Im(A1_e);
+
+joint_var <- ((sigma^2) * vargam) + ((sigma^2) * mean(gammas)^2) + 
+             (vargam * (dval*mn)^2);
+
+plot(A0_r, A0_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "dodgerblue4", ylim = c(-1.5, 1.5),
+     xlim = c(-3, 1.5));
+A0x1a     <- mean(gammas) * sigma * sqrt(S) * cos(vl) + mean(diag(A0));
+A0y1a     <- mean(gammas) * sigma * sqrt(S) * sin(vl);
+points(x = A0x1a, y = A0y1a, type = "l", lwd = 3, col = "dodgerblue4");
+text(x = -2.75, y = 1.75, labels = "c", cex = 2);
+
+plot(A1_r, A1_i, pch = 4, cex = 0.7, xlab = "", ylab = "", cex.lab = 1.3, 
+     cex.axis = 1.25, asp = 1, col = "firebrick", yaxt = "n", 
+     xlim = c(-3, 1.5), ylim = c(-1.5, 1.5));
+text(x = -2.75, y = 1.75, labels = "d", cex = 2);
+
+calc_var <- A1;
+diag(calc_var) <- NA;
+calc_var  <- as.numeric(calc_var);
+cvar2 <- var(calc_var[!is.na(calc_var)]);
+
 mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);

### 2020-03-02 notebook/ms.Rmd
Rework Figure 2 to get it the way that I want it

@@
 Equation \ref{defM} thereby serves as a null model to investigate how variation in component response rate ($\sigma_{\gamma}$) affects complex systems.
-In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\textbf{M} = \textbf{A}$.
-Under these conditions, eigenvalues of $\textbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Figure 1a).
+In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
+Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Figure 1a).
 
-**Analytical considerations of $\textbf{\gamma}$ effects on \textbf{M}$**.
-The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\textbf{M}$ [@Tang2014b].
+**Analytical considerations of $\mathbf{\gamma}$ effects on $\mathbf{M}$**.
+The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
 These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
@@
 Introducing variation in $\gamma$ increases the total variation in the system.
-Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables,
+Variation of the off-diagonal elements in $\mathbf{M}$ is described by the joint variation of two random variables,
 
@@
 
-**Numerical simulations of random $\textbf{M}$ with and without $\textbf{\sigma_{\gamma}}$**.
+**Numerical simulations of random $\mathbf{M}$ with and without $\mathbf{\sigma_{\gamma}}$**.
 I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$ across a rane of parameter combinations.
@@
 
-**Figure 2: Distributions of eigenvalues before (a) and after (b) introducing variation in component response rate ($\boldsymbol{\gamma}$) in complex systems.** Each panel show the same system where $S = 1000$, $C = 0.05$, and $\sigma = 0.4$.
-**a.** Eigenvalues plotted in the absence of $Var(\gamma)$ where $E[\gamma] = 1$, versus **b.** eigenvalues plotted given $\gamma \sim \mathcal{U}(0, 2)$, which increases the variance of interaction strengths ($\sigma^{2}$) but also creates a cluster of eigenvalues toward the distribution's centre (-1, 0).
-Blue elipses in both panels show the circle centred on the distribution in panel a.
-Proportions of $\Re(\lambda) < 0$ are `r sum(A0_r < 0) / length(A0_r)` and `r sum(A1_r < 0) / length(A1_r)` for a and b, respectively.
+**Figure 2: Insert a new caption here**.
 
-```{r, eval = TRUE, echo = FALSE, fig.height = 6, fig.width = 9}
-par(mfrow = c(1, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0, 0));
-plot(A0_r, A0_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
+```{r, echo = FALSE, fig.height = 6, fig.width = 6}
+sim_rhos <- read.csv(file = "sim_results/rhos/sim_rhos.csv");
+par(mar = c(5, 5, 1, 1));
+plot(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 2,
+     xlab = expression(paste("Correlation between A"[ij]," and A"[ji], 
+                             " (",rho,")")),
+     ylab = "Real part of leading eigenvalue", col = "dodgerblue4",
+     ylim = c(min(sim_rhos[,22:25]), max(sim_rhos[,22:25])),
+     cex.lab = 1.25, cex.axis = 1.25);
+mbox <- function(x0, x1, y0, y1){
+    xx <- seq(from=x0, to=x1, length.out = 100);
+    yy <- seq(from=y0, to=y1, length.out = 100);
+    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
+    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
+    return(list(x=xd, y=yd));
+}
+more_stable_1 <- sim_rhos[,21] - sim_rhos[,20]
+more_stable_2 <- which(more_stable_1 < 0);
+min_x         <- min(more_stable_2);
+max_x         <- max(more_stable_2);
+shade_box     <- mbox(x0 = sim_rhos[min_x, 1], x1 = sim_rhos[max_x, 1], 
+                      y0 = -12, y1 = 15);
+polygon(x = shade_box$x, y = shade_box$y, col = "grey80", border = NA);
+abline(h = 0, lwd = 0.8, lty = "dotted", col = "black");
+for(i in 1:dim(sim_rhos)[1]){
+    lines(x = c(sim_rhos[i,1], sim_rhos[i,15]), col = "purple",
+          y = c(sim_rhos[i,20], sim_rhos[i,21]), lwd = 1.5);
+}
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,22], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
+arrows(x0 = sim_rhos[,1], x1 = sim_rhos[,1], y0 = sim_rhos[,20], 
+       y1 = sim_rhos[,23], angle = 90, length = 0.05, lwd = 1, col = "dodgerblue4");
+arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
+       y1 = sim_rhos[,24], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
+arrows(x0 = sim_rhos[,15], x1 = sim_rhos[,15], y0 = sim_rhos[,21], 
+       y1 = sim_rhos[,25], angle = 90, length = 0.05, lwd = 1, col = "firebrick");
+points(x = sim_rhos[,1], y = sim_rhos[,20], type = "b", pch = 20, lwd = 3,
        col = "dodgerblue4");
-vl <- seq(from = 0, to = 2*pi, by = 0.001);
-x0 <- sqrt(1000) * sd(A0vec) * cos(vl) + mean(diag(A0));
-y0 <- sqrt(1000) * sd(A0vec) * sin(vl);
-text(x = -5.5, y = 4, labels = "a", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3, col = "blue");
-points(A0_r, A0_i, col = "dodgerblue4",pch = 4, cex = 0.7);
-plot(A1_r, A1_i, xlim = c(-6, 2.25), ylim = c(-5,4.25), pch = 4, cex = 0.7,
-     xlab = "", ylab = "", cex.lab = 1.5, cex.axis = 1.5, asp = 1, 
-     col = "firebrick", yaxt = "n");
-text(x = -5.5, y = 4, labels = "b", cex = 2);
-points(x = x0, y = y0, type = "l", lwd = 3, lty = "solid", col = "blue");
-points(A1_r, A1_i, col = "firebrick", pch = 4, cex = 0.7);
-mtext(side = 1, "Real", outer = TRUE, line = 3, cex = 2);
-mtext(side = 2, "Imaginary", outer = TRUE, line = 2.5, cex = 2);
+points(x = sim_rhos[,15], y = sim_rhos[,21], type = "b", pch = 15, lwd = 3,
+       col = "firebrick");
+points(x = sim_rhos[,1], y = sim_rhos[,20], type = "l", pch = 20, lwd = 3,
+       col = "dodgerblue4");
+points(x = sim_rhos[,15], y = sim_rhos[,21], type = "l", pch = 15, lwd = 3,
+       col = "firebrick");
+legend("topleft", fill = c("dodgerblue4", "firebrick"), cex = 1.5,
+       legend = c(expression(paste("Var(",gamma,") = 0")), 
+                  expression(paste("Var(",gamma,") = 1/3"))),
+       bg = "white");
+box();
+# Note that the 1/3 comes from the definition of unif distr. from 0 to 2.
 ```

### 2020-03-02 notebook/ms.tex
Rework Figure 2 to get it the way that I want it

@@
 (\(\sigma_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
-(diagonal elements all equal 1) and \(\textbf{M} = \textbf{A}\).
+(diagonal elements all equal 1) and \(\mathbf{M} = \mathbf{A}\). Under
+these conditions, eigenvalues of \(\mathbf{M}\) are distributed
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} in a
+circle centred at \((-d, 0)\) with a radius of
+\(\sigma_{M} \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+(Figure 1a).
 
-\textbf{Analytical considerations}. The value of \(\max(\Re(\lambda))\),
-and therefore system stability, can be estimated from five properties of
-\(\textbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}.
+\textbf{Analytical considerations of \(\mathbf{\gamma}\) effects on
+\(\mathbf{M}\)}. The value of \(\max(\Re(\lambda))\), and therefore
+system stability, can be estimated from five properties of
+\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}.
 These properties include (1) system size (\(S\)), (2) mean
@@
 variation in \(\gamma\) increases the total variation in the system.
-Variation of the off-diagonal elements in \(\textbf{M}\) is described by
+Variation of the off-diagonal elements in \(\mathbf{M}\) is described by
 the joint variation of two random variables,
@@
 \(\sigma^{2}_{A}\sqrt{S}\) and \(\sigma^{2}_{M}\sqrt{S}\), respectively
-(Figure 1a vs.~1b). When components are self-regulating (e.g.,
-\(d = -1\)), positive \(\sigma^{2}_\gamma\) changes the distribution of
-eigenvalues, potentially affecting stability (Figure 1c vs.~1d).
+(Figure 1a vs.~1b). If \(d \neq 0\), positive \(\sigma^{2}_\gamma\)
+changes the distribution of
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}--\protect\hyperlink{ref-Stone2017}{22}},
+potentially affecting stability (Figure 1c vs.~1d).
 
 Given \(\sigma^{2}_\gamma = 0\), \(\max(\Re(\lambda))\) decrease
-linearly with \(\rho\) such that,
+linearly with \(\rho\) such
+that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{23}},
 
@@
 
-When \(\rho\) values are negative, such as when \(\textbf{M}\) models a
-predator-prey system in which \(M_{ij}\) and \(M_{ji}\) have opposing
-signs, stability is expected to
-increase\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. When
-elements of \(\gamma\) vary independently, the magnitude of \(\rho\) is
-decreased because \(\sigma^{2}_{\gamma}\) increases the variance of
-\(M_{ij}\) without affecting the covariance between \(M_{ij}\) and
-\(M_{ji}\),
+If \(\rho < 0\), such as when \(\textbf{M}\) models a predator-prey
+system in which \(M_{ij}\) and \(M_{ji}\) have opposing signs, stability
+increases\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. If
+diagonal elements of \(\gamma\) vary independently, the magnitude of
+\(\rho\) is decreased because \(\sigma^{2}_{\gamma}\) increases the
+variance of \(M_{ij}\) without affecting the expected covariance between
+\(M_{ij}\) and \(M_{ji}\),
 
@@
 
-Because \(\gamma\) does not affect \(cov(M_{ij}, M_{ji})\), the sign of
-\(\rho\) remains unchanged.
+Neverthless, numerical simulations reveal that
+\(\sigma^{2}_{\gamma} > 0\) results in a nonlinear relationship between
+\(\rho\) and \(\max(\Re(\lambda))\), which can sometimes increase the
+stability of the system.
 
-\textbf{Numerical simulations}. I used numerical simulations to isolate
-the effect of \(\sigma_{\gamma}\) on the stability of
-\(\textbf{M} = \gamma \textbf{A}\) by comparing \(\sigma_{\gamma} > 0\)
-versus \(\sigma_{\gamma} = 0\). Values of \(\gamma_{ii}\) were sampled
-such that \(E[d]\) and \(E[\mu]\) also remained unchanged (in practice,
-diagonal elements of \(\textbf{M}\) were standardised so that mean
-values were identical before and after adding \(\gamma\)).
+\textbf{Numerical simulations of random \(\mathbf{M}\) with and without
+\(\mathbf{\sigma_{\gamma}}\)}. I used numerical simulations to isolate
+the effect of \(\sigma_{\gamma}\) on system stability by comparing the
+same \(\textbf{M} = \gamma \textbf{A}\) given \(\sigma_{\gamma} > 0\)
+versus \(\sigma_{\gamma} = 0\) across a rane of parameter combinations.
+Values of \(\gamma_{ii}\) were sampled from a uniform distribution where
+\(\gamma \sim \mathcal{U}(0, 2)\) (see Supplementary Information for
+other distributions, which give similar results). For
+\(\sigma_{\gamma} = 0\), diagonal elements were standardised to the mean
+value of sampled \(\gamma_{ii}\) so that \(E[d]\) of \(\textbf{M}\) were
+identical.
 
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I focus on
-a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
-Supplementary Information for other distributions, which give similar
-results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
-\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(\mathbf{M}\) before and after the addition of variation
-in component response rate. Figure 2 shows a comparison of eigenvalue
-distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}--\protect\hyperlink{ref-Stone2017}{22}},
-some of which are clustered locally near the centre of the distribution,
-but others of which are spread outside the former radius of 2.828 (Fig
-2b). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold \(\sigma\sqrt{SC} < 1\).
+\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{\rho}\)}
 
@@
 values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{23}}.
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{24}}.
 Generations of selection on \(\gamma\) value combinations to minimise
@@
 positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{24},\protect\hyperlink{ref-Song2018}{25}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{25},\protect\hyperlink{ref-Song2018}{26}}.
 This is of particular interest for ecological communities because
@@
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}}.
 While my results are intended to be general to all complex systems, and
@@
 a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{26}}
+Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{27}}
 (see Supplementary Information). Hence, for pure interacting species
@@
 of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{22},\protect\hyperlink{ref-Dougoud2018}{24},\protect\hyperlink{ref-Servan2018}{26}}
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{22},\protect\hyperlink{ref-Dougoud2018}{25},\protect\hyperlink{ref-Servan2018}{27}}
 because population densities cannot realistically be negative. My
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{27}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{28}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{27}},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{28}},
 
@@
 components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{20},\protect\hyperlink{ref-May1973}{27}}.
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{20},\protect\hyperlink{ref-May1973}{28}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{23}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{24}}. In
 this case, the practical solution is finding vectors of
@@
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{24}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{26}}),
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{27}}),
 
@@
 
+\hypertarget{ref-Tang2014c}{}
+23. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
+strengths drives stability in large ecological networks. \textbf{17,}
+1094--1100 (2014).
+
 \hypertarget{ref-Hamblin2013}{}
-23. Hamblin, S. On the practical usage of genetic algorithms in ecology
+24. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
@@
 \hypertarget{ref-Dougoud2018}{}
-24. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+25. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-25. Song, C. \& Saavedra, S. Will a small randomly assembled community
+26. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-26. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+27. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 \hypertarget{ref-May1973}{}
-27. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+28. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
@@
 
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the blue circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle centred at \(\gamma = -1.95\) and one small circle centred
-at \(\gamma = -0.05\). In a, the maximum real eigenvalue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 was stable when
-\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
+\textbf{Figure 1: Insert a new legend here}
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-10-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
-\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
-eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
-both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.727 and 0.733 for a and b,
-respectively.
+\textbf{Figure 2: Insert a new caption here}.
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-15-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-14-1.pdf}
 

### 2020-03-02 notebook/ms.Rmd
Fix some small errors that had crept up in the notation

@@
 In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
-Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S_{M}$, $C_{M}$, and $\sigma_{M}^{2}$ increase.
+Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma_{M}$ increase.
 
@@
 
-**Analytical considerations of $\mathbf{\gamma}$ effects on $\mathbf{M}$**.
+**Effect of $\mathbf{\sigma^{2}_{\gamma}}$ on $\mathbf{M}$ (co)variation**.
 The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
@@
 The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Figure 1).
-Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{A}$ and $\textbf{M}$ lie within a circle of a radius $\sigma^{2}_{A}\sqrt{S}$ and $\sigma^{2}_{M}\sqrt{S}$, respectively (Figure 1a vs.
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{A}$ and $\textbf{M}$ lie within a circle of a radius $\sigma_{A}\sqrt{S}$ and $\sigma_{M}\sqrt{S}$, respectively (Figure 1a vs.
 1b).
@@
 
+**Figure 3: Insert a new caption here**.
+
+```{r, echo = FALSE, fig.height = 8, fig.width = 6}
+sim10 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim10.csv");
+sim15 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim15.csv");
+sim20 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim20.csv");
+sim25 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim25.csv");
+sim30 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim30.csv");
+sim35 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim35.csv");
+sim40 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim40.csv");
+sim45 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim45.csv");
+fstlt <- c(1, dim(sim20)[1]);
+par(mar = c(0.25, 0.25, 0.25, 0.25), mfrow = c(3, 2), oma = c(6, 6, 1, 1));
+plot(x = sim20[,1], y = sim20[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 1.4), yaxt = "n", col = "dodgerblue4");
+axis(side = 2, at = c(-0.75, 0, 0.75), cex.axis = 1.75);
+points(x = sim20[fstlt, 1], y = sim20[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+points(x = sim20[,15], y = sim20[,21], type = "l", lwd = 2, col = "firebrick");
+points(x = sim20[fstlt,15], y = sim20[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 20", cex = 2.5);
+plot(x = sim25[,1], y = sim25[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.75, col = "dodgerblue4");
+points(x = sim25[fstlt, 1], y = sim25[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+points(x = sim25[,15], y = sim25[,21], type = "l", lwd = 2, col = "firebrick");
+points(x = sim25[fstlt,15], y = sim25[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 25", cex = 2.5);
+plot(x = sim30[,1], y = sim30[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 1.4), cex.axis = 1.75, col = "dodgerblue4", yaxt = "n");
+points(x = sim30[fstlt, 1], y = sim30[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+points(x = sim30[fstlt,15], y = sim30[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+axis(side = 2, at = c(-0.75, 0, 0.75), cex.axis = 1.75);
+points(x = sim30[,15], y = sim30[,21], type = "l", lwd = 2, col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 30", cex = 2.5);
+plot(x = sim35[,1], y = sim35[,20], type = "l", pch = 20, lwd = 2, xaxt = "n",
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.75, col = "dodgerblue4");
+points(x = sim35[fstlt, 1], y = sim35[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+points(x = sim35[,15], y = sim35[,21], type = "l", lwd = 2, col = "firebrick");
+points(x = sim35[fstlt,15], y = sim35[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 35", cex = 2.5);
+plot(x = sim40[,1], y = sim40[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 1.4), cex.axis = 1.75, col = "dodgerblue4", yaxt = "n");
+points(x = sim40[fstlt, 1], y = sim40[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+axis(side = 2, at = c(-0.75, 0, 0.75), cex.axis = 1.75);
+points(x = sim40[,15], y = sim40[,21], type = "l", lwd = 2,col = "firebrick");
+points(x = sim40[fstlt,15], y = sim40[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 40", cex = 2.5);
+plot(x = sim45[,1], y = sim45[,20], type = "l", pch = 20, lwd = 2, 
+     ylim = c(-1, 1.4), yaxt = "n", cex.axis = 1.75, col = "dodgerblue4");
+points(x = sim45[fstlt, 1], y = sim45[fstlt, 20], pch = 20, cex = 2, 
+       col = "dodgerblue4");
+points(x = sim45[,15], y = sim45[,21], type = "l", lwd = 2, col = "firebrick");
+points(x = sim45[fstlt,15], y = sim45[fstlt,21], pch = 15, cex = 2, 
+       col = "firebrick");
+abline(h = 0, col = "black", lty = "dotted", lwd = 0.8);
+text(x = -0.6, y = 1.2, labels = "S = 45", cex = 2.5);
+mtext(text =  expression(paste("E Correlation between A"[ij]," and A"[ji], 
+                               " (",rho,")")), side = 1,
+      line = 4, outer = TRUE, cex = 1.5);
+mtext(text = "E leading real part eigenvalue", side = 2,
+      line = 3.5, outer = TRUE, cex = 1.5);
+```
+
+\clearpage
+
 ```{r, echo = FALSE}
@@
 
-**Figure 3: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
@@
 
-**Figure 4: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
+**Figure 5: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
 For each $S$, 100000 complex systems are randomly generated.

### 2020-03-02 notebook/ms.tex
Fix some small errors that had crept up in the notation

@@
 (Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S_{M}\), \(C_{M}\), and
-\(\sigma_{M}^{2}\) increase.
+therefore becomes increasingly unlikely as \(S\), \(C\), and
+\(\sigma_{M}\) increase.
 
@@
 
-\textbf{Analytical considerations of \(\mathbf{\gamma}\) effects on
-\(\mathbf{M}\)}. The value of \(\max(\Re(\lambda))\), and therefore
+\textbf{Effect of \(\mathbf{\sigma^{2}_{\gamma}}\) on \(\mathbf{M}\)
+(co)variation}. The value of \(\max(\Re(\lambda))\), and therefore
 system stability, can be estimated from five properties of
@@
 \(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
-\(\textbf{M}\) lie within a circle of a radius
-\(\sigma^{2}_{A}\sqrt{S}\) and \(\sigma^{2}_{M}\sqrt{S}\), respectively
-(Figure 1a vs.~1b). If \(d \neq 0\), positive \(\sigma^{2}_\gamma\)
-changes the distribution of
+\(\textbf{M}\) lie within a circle of a radius \(\sigma_{A}\sqrt{S}\)
+and \(\sigma_{M}\sqrt{S}\), respectively (Figure 1a vs.~1b). If
+\(d \neq 0\), positive \(\sigma^{2}_\gamma\) changes the distribution of
 eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}--\protect\hyperlink{ref-Stone2017}{22}},
@@
 
-\textbf{Figure 3: Stability of large complex systems with and without
+\textbf{Figure 3: Insert a new caption here}.
+
+\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
+
+\clearpage
+
+\textbf{Figure 4: Stability of large complex systems with and without
 variation in component response rate (\(\boldsymbol{\gamma}\)).} The
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-14-1.pdf}
 
@@
 
-\textbf{Figure 4: Stability of large complex systems given
+\textbf{Figure 5: Stability of large complex systems given
 \(\boldsymbol{\gamma = 1}\) versus targeted
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-14-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-15-1.pdf}
 

### 2020-03-03 notebook/ms.Rmd
Draft a legend for the new Figure 1

@@
 
-**Figure 1: Insert a new legend here**
-
 ```{r, fig.width = 7, fig.height = 7, echo = FALSE}
@@
              (vargam * (dval*mn)^2);
+```
 
+**Figure 1: Eigenvalue distributions of random complex systems.** Each panel shows the real (x-axis) and imaginary (y-axis) parts of $S =$ `r S` eigenvalues from random $S \times S$ matrices.
+($\textbf{a}$) A system represented by a matrix $\mathbf{A}$, in which all elements are sampled from a normal distribution with $\mu = 0$ and $\sigma_{A} = 1/\sqrt{S}$.
+Points are uniformly distributed within a circle centred at the origin with a radius of $\sigma_{A} \sqrt{S} = 1$.
+($\textbf{b}$) The same system after including variation in the response rates of $S$ components, represented by the diagonal matrix $\gamma$, such that $\mathbf{M} = \gamma\mathbf{A}$.
+Elements of $\gamma$ are randomly sampled from a uniform distribution of $\min = 0$ and $\max = 2$.
+Eigenvalues of $\mathbf{M}$ are distributed non-uniformly within a circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
+($\textbf{c}$) A different random system $\mathbf{A}$ constructed from the same parameters as in $\textbf{a}$, except with diagonal values of $-1$.
+($\textbf{d}$) The same system $\textbf{c}$ after including variation in component response rates, sampled from $\mathcal{U}(0, 2)$ as in $\textbf{b}$.
+
+```{r, fig.width = 7, fig.height = 7, echo = FALSE}
 par(mfrow = c(2, 2), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 0.2, 0.2));

### 2020-03-03 notebook/ms.tex
Draft a legend for the new Figure 1

@@
 
-\textbf{Figure 1: Insert a new legend here}
+\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
+Each panel shows the real (x-axis) and imaginary (y-axis) parts of
+\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
+(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
+which all elements are sampled from a normal distribution with
+\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
+distributed within a circle centred at the origin with a radius of
+\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system after
+including variation in the response rates of \(S\) components,
+represented by the diagonal matrix \(\gamma\), such that
+\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
+sampled from a uniform distribution of \(\min = 0\) and \(\max = 2\).
+Eigenvalues of \(\mathbf{M}\) are distributed non-uniformly within a
+circle centred at the origin with a radius of
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.19.
+(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
+from the same parameters as in \(\textbf{a}\), except with diagonal
+values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
+including variation in component response rates, sampled from
+\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-15-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
 

### 2020-03-05 notebook/ms.Rmd
Small edits to Figure 1 to improve readability

@@
 ($\textbf{b}$) The same system after including variation in the response rates of $S$ components, represented by the diagonal matrix $\gamma$, such that $\mathbf{M} = \gamma\mathbf{A}$.
-Elements of $\gamma$ are randomly sampled from a uniform distribution of $\min = 0$ and $\max = 2$.
-Eigenvalues of $\mathbf{M}$ are distributed non-uniformly within a circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
+Elements of $\gamma$ are randomly sampled from a uniform distribution from $\min = 0$ to $\max = 2$.
+Eigenvalues of $\mathbf{M}$ are then distributed non-uniformly within a circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
 ($\textbf{c}$) A different random system $\mathbf{A}$ constructed from the same parameters as in $\textbf{a}$, except with diagonal values of $-1$.

### 2020-03-05 notebook/ms.tex
Small edits to Figure 1 to improve readability

@@
 
-\subsection{Abstract}\label{abstract}
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
@@
 
-\subsection{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
@@
 
-\subsection{Results}\label{results}
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
@@
 
-\subsection{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
@@
 
-\subsection{Methods}\label{methods}
+\hypertarget{methods}{%
+\subsection{Methods}\label{methods}}
 
@@
 synthesis of eco-evolutionary feedbacks on community stability, Patel et
-al. model a system that includes a vector of potentially changing
+al.~model a system that includes a vector of potentially changing
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
@@
 effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
-were interested in stability when the evolution of traits was relatively
-slow or fast in comparison with the change in species
+species \(i\) and the value of trait \(j\), respectively. Patel et
+al.~were interested in stability when the evolution of traits was
+relatively slow or fast in comparison with the change in species
 densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, and
@@
 
-I use the same principle that Patel et al. use to modulate the relative
+I use the same principle that Patel et al.~use to modulate the relative
 rate of evolution to modulate rates of component responses for \(S\)
@@
 \hypertarget{refs}{}
-\hypertarget{ref-May1972}{}
+\leavevmode\hypertarget{ref-May1972}{}%
 1. May, R. M. Will a large complex system be stable? \emph{Nature}
@@
 
-\hypertarget{ref-Allesina2012}{}
+\leavevmode\hypertarget{ref-Allesina2012}{}%
 2. Allesina, S. \& Tang, S. Stability criteria for complex ecosystems.
@@
 
-\hypertarget{ref-Mougi2012}{}
+\leavevmode\hypertarget{ref-Mougi2012}{}%
 3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
@@
 
-\hypertarget{ref-Allesina2015}{}
+\leavevmode\hypertarget{ref-Allesina2015}{}%
 4. Allesina, S. \emph{et al.} Predicting the stability of large
@@
 
-\hypertarget{ref-Grilli2017}{}
+\leavevmode\hypertarget{ref-Grilli2017}{}%
 5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
@@
 
-\hypertarget{ref-Gray2008}{}
+\leavevmode\hypertarget{ref-Gray2008}{}%
 6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
@@
 
-\hypertarget{ref-Gray2009}{}
+\leavevmode\hypertarget{ref-Gray2009}{}%
 7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
@@
 
-\hypertarget{ref-Rosenfeld2009}{}
+\leavevmode\hypertarget{ref-Rosenfeld2009}{}%
 8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
@@
 
-\hypertarget{ref-MacArthur2010}{}
+\leavevmode\hypertarget{ref-MacArthur2010}{}%
 9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
@@
 
-\hypertarget{ref-May2008}{}
+\leavevmode\hypertarget{ref-May2008}{}%
 10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
@@
 
-\hypertarget{ref-Haldane2011}{}
+\leavevmode\hypertarget{ref-Haldane2011}{}%
 11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
@@
 
-\hypertarget{ref-Suweis2014}{}
+\leavevmode\hypertarget{ref-Suweis2014}{}%
 12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
@@
 
-\hypertarget{ref-Bardoscia2017}{}
+\leavevmode\hypertarget{ref-Bardoscia2017}{}%
 13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
@@
 
-\hypertarget{ref-Tao2010}{}
+\leavevmode\hypertarget{ref-Tao2010}{}%
 14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
@@
 
-\hypertarget{ref-Watts1998}{}
+\leavevmode\hypertarget{ref-Watts1998}{}%
 15. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
@@
 
-\hypertarget{ref-Albert2002}{}
+\leavevmode\hypertarget{ref-Albert2002}{}%
 16. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
@@
 
-\hypertarget{ref-Williams2000}{}
+\leavevmode\hypertarget{ref-Williams2000}{}%
 17. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
@@
 
-\hypertarget{ref-Patel2018}{}
+\leavevmode\hypertarget{ref-Patel2018}{}%
 18. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
@@
 
-\hypertarget{ref-Tang2014b}{}
+\leavevmode\hypertarget{ref-Tang2014b}{}%
 19. Tang, S. \& Allesina, S. Reactivity and stability of large
@@
 
-\hypertarget{ref-Ahmadian2015}{}
+\leavevmode\hypertarget{ref-Ahmadian2015}{}%
 20. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
@@
 
-\hypertarget{ref-Gibbs2017}{}
+\leavevmode\hypertarget{ref-Gibbs2017}{}%
 21. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
@@
 
-\hypertarget{ref-Stone2017}{}
+\leavevmode\hypertarget{ref-Stone2017}{}%
 22. Stone, L. The feasibility and stability of large complex biological
@@
 
-\hypertarget{ref-Tang2014c}{}
+\leavevmode\hypertarget{ref-Tang2014c}{}%
 23. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
@@
 
-\hypertarget{ref-Hamblin2013}{}
+\leavevmode\hypertarget{ref-Hamblin2013}{}%
 24. Hamblin, S. On the practical usage of genetic algorithms in ecology
@@
 
-\hypertarget{ref-Dougoud2018}{}
+\leavevmode\hypertarget{ref-Dougoud2018}{}%
 25. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
@@
 
-\hypertarget{ref-Song2018}{}
+\leavevmode\hypertarget{ref-Song2018}{}%
 26. Song, C. \& Saavedra, S. Will a small randomly assembled community
@@
 
-\hypertarget{ref-Servan2018}{}
+\leavevmode\hypertarget{ref-Servan2018}{}%
 27. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
@@
 
-\hypertarget{ref-May1973}{}
+\leavevmode\hypertarget{ref-May1973}{}%
 28. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
@@
 circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.19.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-05 notebook/ms.Rmd
Add a caption for the new Figure 2

@@
 
-**Figure 2: Insert a new caption here**.
+**Figure 2: Complex system correlation versus stability with and without variation in component response rates**.
+Each point represents 10000 replicate numerical simulations of a random complex system $\mathbf{M} = \gamma \mathbf{A}$ with a fixed correlation between off-diagonal elements $A_{ij}$ and $A_{ji}$ ($\rho$, x-axis).
+Where real parts of eigenvalues of $\mathbf{M}$ are negative (y-axis), $\mathbf{M}$ is stable (black dotted line).
+Blue circles show systems in the absence of variation in component response rates ($\sigma_{\gamma} = 0$).
+Red squares show systems in which $\sigma_{\gamma} = 1/3$.
+Arrows show the range of real parts of leading eigenvalues observed.
+Because $\gamma$ decreases the magnitude of $\rho$, purple lines are included to link replicate simulations before (blue circles) and after (red squares) including $\gamma$.
+The range of $\rho$ values in which $\gamma$ decreases the mean real part of the leading eigenvalue is indicated with grey shading.
+In all simulations, system size and connectence were set to $S = 25$ and $C = 1$, respectively.
+Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.4)$, and diagonal elements were set to $-1$.
+Elements of $\gamma \sim \mathcal{U}(0, 2)$.
 

### 2020-03-05 notebook/ms.tex
Add a caption for the new Figure 2

@@
 \(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
-sampled from a uniform distribution of \(\min = 0\) and \(\max = 2\).
-Eigenvalues of \(\mathbf{M}\) are distributed non-uniformly within a
-circle centred at the origin with a radius of
+sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
+Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
+a circle centred at the origin with a radius of
 \(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
@@
 
-\textbf{Figure 2: Insert a new caption here}.
+\textbf{Figure 2: Complex system correlation versus stability with and
+without variation in component response rates}. Each point represents
+10000 replicate numerical simulations of a random complex system
+\(\mathbf{M} = \gamma \mathbf{A}\) with a fixed correlation between
+off-diagonal elements \(A_{ij}\) and \(A_{ji}\) (\(\rho\), x-axis).
+Where real parts of eigenvalues of \(\mathbf{M}\) are negative (y-axis),
+\(\mathbf{M}\) is stable (black dotted line). Blue circles show systems
+in the absence of variation in component response rates
+(\(\sigma_{\gamma} = 0\)). Red squares show systems in which
+\(\sigma_{\gamma} = 1/3\). Arrows show the range of real parts of
+leading eigenvalues observed. Because \(\gamma\) decreases the magnitude
+of \(\rho\), purple lines are included to link replicate simulations
+before (blue circles) and after (red squares) including \(\gamma\). The
+range of \(\rho\) values in which \(\gamma\) decreases the mean real
+part of the leading eigenvalue is indicated with grey shading. In all
+simulations, system size and connectence were set to \(S = 25\) and
+\(C = 1\), respectively. Off-diagonal elements of \(\textbf{A}\) were
+randomly sampled from \(A_{ij} \sim \mathcal{N}(0, 0.4)\), and diagonal
+elements were set to \(-1\). Elements of
+\(\gamma \sim \mathcal{U}(0, 2)\).
 

### 2020-03-05 notebook/ms.Rmd
Fix just a bit of the text again on Figure 2 to better explain gamma sampling
Change the figure 2 legend so that the notation matches the text

@@
 Where real parts of eigenvalues of $\mathbf{M}$ are negative (y-axis), $\mathbf{M}$ is stable (black dotted line).
-Blue circles show systems in the absence of variation in component response rates ($\sigma_{\gamma} = 0$).
-Red squares show systems in which $\sigma_{\gamma} = 1/3$.
+Blue circles show systems in the absence of variation in component response rates ($\sigma^{2}_{\gamma} = 0$).
+Red squares show systems in which $\sigma^{2}_{\gamma} = 1/3$.
 Arrows show the range of real parts of leading eigenvalues observed.
@@
 Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.4)$, and diagonal elements were set to $-1$.
-Elements of $\gamma \sim \mathcal{U}(0, 2)$.
+Elements of $\gamma$ were sampled, $\gamma_{i} \sim \mathcal{U}(0, 2)$.
 
@@
        col = "firebrick");
-legend("topleft", fill = c("dodgerblue4", "firebrick"), cex = 1.5,
-       legend = c(expression(paste("Var(",gamma,") = 0")), 
-                  expression(paste("Var(",gamma,") = 1/3"))),
-       bg = "white");
+legend("topleft", fill = c("dodgerblue4", "firebrick"), cex = 1.55,
+       legend = c(expression(paste(sigma[gamma]^2," = 0")), 
+                  expression(paste(sigma[gamma]^2," = 1/3"))),
+       bg = "white", y.intersp = 1.5);
 box();
@@
 
-**Figure 3: Insert a new caption here**.
+**Figure 3: System correlation versus stability across different system sizes**.
 

### 2020-03-05 notebook/ms.tex
Fix just a bit of the text again on Figure 2 to better explain gamma sampling
Change the figure 2 legend so that the notation matches the text

@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.21.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 in the absence of variation in component response rates
-(\(\sigma_{\gamma} = 0\)). Red squares show systems in which
-\(\sigma_{\gamma} = 1/3\). Arrows show the range of real parts of
+(\(\sigma^{2}_{\gamma} = 0\)). Red squares show systems in which
+\(\sigma^{2}_{\gamma} = 1/3\). Arrows show the range of real parts of
 leading eigenvalues observed. Because \(\gamma\) decreases the magnitude
@@
 randomly sampled from \(A_{ij} \sim \mathcal{N}(0, 0.4)\), and diagonal
-elements were set to \(-1\). Elements of
-\(\gamma \sim \mathcal{U}(0, 2)\).
+elements were set to \(-1\). Elements of \(\gamma\) were sampled,
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\).
 
@@
 
-\textbf{Figure 3: Insert a new caption here}.
+\textbf{Figure 3: System correlation versus stability across different
+system sizes}.
 

### 2020-03-05 notebook/ms.Rmd
Remove the last figure, which will now be placed in the supporting information

@@
 **Figure 3: System correlation versus stability across different system sizes**.
+In each panel, 10000 random complex systems $\mathbf{M} = \gamma \mathbf{A}$ are simulated for each correlation $\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}$ between off-diagonal elements $A_{ij}$ and $A_{ji}$.
+Lines show the expected real part of the leading eigenvalues of $\mathbf{M}$ (red squares) versus $\mathbf{A}$ (blue circles) across $\rho$, where negative values (below the dotted black line) indicate system stability.
+Differences between lines thereby show the effect of component response rate variation ($\gamma$) on system stability across system correlations and sizes ($S$).
+For all simulations, system connectance was $C = 1$.
+Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.2)$, and diagonal elements were set to $-1$.
+Elements of $\gamma$ were sampled such that $\gamma_{i} \sim \mathcal{U}(0, 2)$, so $\sigma^{2}_{\gamma} = 1/3$.
 
@@
       line = 4, outer = TRUE, cex = 1.5);
-mtext(text = "E leading real part eigenvalue", side = 2,
+mtext(text = "E real part of leading eigenvalue", side = 2,
       line = 3.5, outer = TRUE, cex = 1.5);
@@
 ```
-
-\clearpage
-
-**Figure 5: Stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted $\boldsymbol{Var(\gamma)}$.** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=40$) for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
-For each $S$, 100000 complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
-Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
-The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $Var(\gamma) = 0$.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
-dat <- dat[,-1];
-plot_stable_4(dat);
-```

### 2020-03-05 notebook/ms.tex
Remove the last figure, which will now be placed in the supporting information

@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.21.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 \textbf{Figure 3: System correlation versus stability across different
-system sizes}.
+system sizes}. In each panel, 10000 random complex systems
+\(\mathbf{M} = \gamma \mathbf{A}\) are simulated for each correlation
+\(\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}\) between off-diagonal
+elements \(A_{ij}\) and \(A_{ji}\). Lines show the expected real part of
+the leading eigenvalues of \(\mathbf{M}\) (red squares) versus
+\(\mathbf{A}\) (blue circles) across \(\rho\), where negative values
+(below the dotted black line) indicate system stability. Differences
+between lines thereby show the effect of component response rate
+variation (\(\gamma\)) on system stability across system correlations
+and sizes (\(S\)). For all simulations, system connectance was
+\(C = 1\). Off-diagonal elements of \(\textbf{A}\) were randomly sampled
+from \(A_{ij} \sim \mathcal{N}(0, 0.2)\), and diagonal elements were set
+to \(-1\). Elements of \(\gamma\) were sampled such that
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
 
@@
 
-\clearpage
-
-\textbf{Figure 5: Stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted
-\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
-stable across different system sizes (\(S\), max \(S=40\)) for
-\(C = 1\), and the proportion of systems wherein a targeted search of
-\(\gamma\) values successfully resulted in system stability. For each
-\(S\), 100000 complex systems are randomly generated. Stability of each
-complex system is tested given variation in \(\gamma\) using a genetic
-algorithm to maximise the effect of \(\gamma\) values on increasing
-stability, as compared to stability in an otherwise identical system in
-which \(\gamma\) is the same for all components. Blue bars show the
-number of stable systems in the absence of component response rate
-variation, while red bars show the number of stable systems that can be
-generated if component response rate is varied to maximise system
-stability. The black line shows the proportion of systems that are
-stable when component response rate is targeted to increase stability,
-but would not be stable if \(Var(\gamma) = 0\).
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-16-1.pdf}
-
 \end{document}

### 2020-03-05 notebook/SI.Rmd
Update for submission to Scientific Reports
Some more analytical playing, and rho results
Add the genetic algorithm figure to the SI and change the text to incorporate it

@@
 ---
-title: "Component response rate variation drives stability in large complex systems"
+title: "Component response rate variation underlies the stability of complex systems"
 author: "A. Bradley Duthie"
@@
 ```{r, echo = FALSE}
+source(file = "../R/Figures_3_4.R");
 source(file = "R_old/sim_mat.R");
@@
 - [Consistency with Gibbs et al. (2018)](#Gibbs)
+- [Effects of matrix correlations](#rhos)
 
@@
 
-Figure 3 of the main text reports the number of stable random complex systems found over 1 million iterations.
+Figure 4 of the main text reports the number of stable random complex systems found over 1 million iterations.
 The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
@@
 
-Figure 4 of the main text reports the number of stable random complex systems found over 100000 using the genetic algorithm to maximise stability with a vector $\mathbf{\gamma}$.
-Stability results for 100000 $\mathbf{M}$ for each $S$ from 2-40 are shown below.
+The figure below compares the stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted manipulation of $\gamma$ elements.
+For each $S$, 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
+Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
+The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $\sigma^{2}_{\gamma} = 0$.
+The y-axis shows the $\ln$ number of systems that are stable across $S = \{1, 2, ..., 39, 40\}$ for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
+dat <- dat[,-1];
+plot_stable_4(dat);
+```
+
+Stability results are also shown in the table below.
 Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
@@
 
+Effects of matrix correlations
+================================================================================
+
 References

### 2020-03-05 notebook/SI.tex
Add the genetic algorithm figure to the SI and change the text to incorporate it

@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs}
@@
 
+\hypertarget{supplemental-information-table-of-contents}{%
 \section{Supplemental Information table of
-contents}\label{supplemental-information-table-of-contents}
+contents}\label{supplemental-information-table-of-contents}}
 
@@
   \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
+\item
+  \protect\hyperlink{rhos}{Effects of matrix correlations}
 \end{itemize}
 
-\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
+\hypertarget{IncrS}{%
+\section{\texorpdfstring{Stability across increasing
 \(S\)}{Stability across increasing S}}\label{IncrS}}
 
-Figure 3 of the main text reports the number of stable random complex
+Figure 4 of the main text reports the number of stable random complex
 systems found over 1 million iterations. The table below shows the
@@
 
-\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
-manipulation of \(\gamma\) (genetic
+\hypertarget{ga}{%
+\section{\texorpdfstring{Stability given targeted manipulation of
+\(\gamma\) (genetic
 algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
 
-Figure 4 of the main text reports the number of stable random complex
-systems found over 100000 using the genetic algorithm to maximise
-stability with a vector \(\mathbf{\gamma}\). Stability results for
-100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
-for \texttt{A0} indicate systems in which \(\gamma = 1\), while
-\texttt{A1} refers to systems in which the genetic algorithm searched
-for a set of \(\gamma\) values that stabilised the system.
+The figure below compares the stability of large complex systems given
+\(\boldsymbol{\gamma = 1}\) versus targeted manipulation of \(\gamma\)
+elements. For each \(S\), 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in \(\gamma\)
+using a genetic algorithm to maximise the effect of \(\gamma\) values on
+increasing stability, as compared to stability in an otherwise identical
+system in which \(\gamma\) is the same for all components. Blue bars
+show the number of stable systems in the absence of component response
+rate variation, while red bars show the number of stable systems that
+can be generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(\sigma^{2}_{\gamma} = 0\). The y-axis shows
+the \(\ln\) number of systems that are stable across
+\(S = \{1, 2, ..., 39, 40\}\) for \(C = 1\), and the proportion of
+systems wherein a targeted search of \(\gamma\) values successfully
+resulted in system stability.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
+
+Stability results are also shown in the table below. Results for
+\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
+refers to systems in which the genetic algorithm searched for a set of
+\(\gamma\) values that stabilised the system.
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\hypertarget{ecological}{\section{Stability of ecological
-networks}\label{ecological}}
+\hypertarget{ecological}{%
+\section{Stability of ecological networks}\label{ecological}}
 
@@
 
-\hypertarget{connectance}{\section{Sensitivity of connectance (C)
-values}\label{connectance}}
+\hypertarget{connectance}{%
+\section{Sensitivity of connectance (C) values}\label{connectance}}
 
@@
 
-\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
-strength (\(\sigma\))
+\hypertarget{sigma}{%
+\section{\texorpdfstring{Sensitivity of interaction strength
+(\(\sigma\))
 values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
@@
 
-\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
-distribution of
+\hypertarget{gam_dist}{%
+\section{\texorpdfstring{Sensitivity of distribution of
 \(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
@@
 
-\hypertarget{Feasibility}{\section{Feasibility of complex
-systems}\label{Feasibility}}
+\hypertarget{Feasibility}{%
+\section{Feasibility of complex systems}\label{Feasibility}}
 
@@
 
-\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
-(2018)}\label{Gibbs}}
+\hypertarget{Gibbs}{%
+\section{Consistency with Gibbs et al. (2018)}\label{Gibbs}}
 
@@
 
-{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-27-1}
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-28-1}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-29-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-33-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-34-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
-
-\section*{References}\label{references}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
+
+\hypertarget{effects-of-matrix-correlations}{%
+\section{Effects of matrix
+correlations}\label{effects-of-matrix-correlations}}
+
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 
-\hypertarget{ref-Allesina2015}{}
+\leavevmode\hypertarget{ref-Allesina2015}{}%
 3. Allesina, S. \emph{et al.} Predicting the stability of large
@@
 
-\hypertarget{ref-Servan2018}{}
+\leavevmode\hypertarget{ref-Servan2018}{}%
 4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
@@
 
-\hypertarget{ref-Gibbs2017}{}
+\leavevmode\hypertarget{ref-Gibbs2017}{}%
 5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of

### 2020-03-05 notebook/ms.Rmd
Edit the legend of Figure 4 so that the notation is consistent

@@
 
-**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S$, max $S=50$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
 Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
-Stability given $Var(\gamma)$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
-Blue and red bars show the number of stable systems in the absence and presence of $Var(\gamma)$, respectively.
-The black line shows the proportion of systems that are stable when $Var(\gamma) > 0$, but would be unstable if $Var(\gamma) = 0$.
+Stability given $\sigma^{2}_{\gamma}>0$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
+Blue and red bars show the number of stable systems in the absence and presence of $\sigma^{2}_{\gamma}$, respectively.
+The black line shows the proportion of systems that are stable when $\sigma^{2}_{\gamma}>0$, but would be unstable if $\sigma^{2}_{\gamma}=0$.
 

### 2020-03-05 notebook/ms.tex
Edit the legend of Figure 4 so that the notation is consistent

@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.16.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 \(\ln\) number of systems that are stable across different system sizes
-(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
-which variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Blue and red bars show the number of stable systems in the absence and
-presence of \(Var(\gamma)\), respectively. The black line shows the
-proportion of systems that are stable when \(Var(\gamma) > 0\), but
-would be unstable if \(Var(\gamma) = 0\).
+(\(S = \{2, 3, ..., 49, 50 \}\)) given \(C = 1\), and the proportion of
+systems in which variation in \(\gamma\) is critical for system
+stability. For each \(S\), 1 million complex systems are randomly
+generated. Stability of each complex system is tested given variation in
+\(\gamma\) by randomly sampling \(\gamma \sim \mathcal{U}(0, 2)\).
+Stability given \(\sigma^{2}_{\gamma}>0\) is then compared to stability
+in an otherwise identical system in which
+\(\gamma = E[\mathcal{U}(0, 2)]\) for all components. Blue and red bars
+show the number of stable systems in the absence and presence of
+\(\sigma^{2}_{\gamma}\), respectively. The black line shows the
+proportion of systems that are stable when \(\sigma^{2}_{\gamma}>0\),
+but would be unstable if \(\sigma^{2}_{\gamma}=0\).
 

### 2020-03-05 notebook/ms.Rmd
Add a paragraph introducing the numerical simulations and generally what is going on

@@
 If $\rho < 0$, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability increases [@Allesina2012].
-If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$,
+If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$.
 
-\begin{equation} \label{rho_def}
-\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.
-\end{equation}
-
-Neverthless, numerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
-
-**Numerical simulations of random $\mathbf{M}$ with and without $\mathbf{\sigma_{\gamma}}$**.
-I used numerical simulations to isolate the effect of $\sigma_{\gamma}$ on system stability by comparing the same $\textbf{M} = \gamma \textbf{A}$ given $\sigma_{\gamma} > 0$ versus $\sigma_{\gamma} = 0$ across a rane of parameter combinations.
-Values of $\gamma_{ii}$ were sampled from a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ (see Supplementary Information for other distributions, which give similar results).
-For $\sigma_{\gamma} = 0$, diagonal elements were standardised to the mean value of sampled $\gamma_{ii}$ so that $E[d]$ of $\textbf{M}$ were identical.
+**Numerical simulations of random $\mathbf{M}$ with and without $\mathbf{\sigma^{2}_{\gamma}}$**.
+I used numerical simulations and eigenanalysis to test how variation in $\gamma$ affects stability in random matrices with known properties, comparing the stability of $\textbf{A}$ versus $\mathbf{M} = \gamma\mathbf{A}$.
+Values of $\gamma$ were sampled from a uniform distribution where $\gamma_{i} \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other distributions, which give similar results).
+Diagonal elements of simulated $\mathbf{A}$ were standardised to the mean value of sampled $\gamma$ so that $E[d]$ of $\textbf{M}$ and $\textbf{M}$ were guaranteed to be identical.
+Here I focus on the effect of $\gamma$ across values of $\rho$, and for increasing system sizes ($S$) in random and structured networks.
+By increasing $S$, the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system.
 
 **Simulation of random $\mathbf{M}$ across $\mathbf{\rho}$**
+
+umerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
 

### 2020-03-08 notebook/ms.Rmd
Remove a bit of the notes to myself, as the sections are now more or less clear

@@
 
-umerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
+Numerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
 
@@
 Borrowing the text from revision_notes.Rmd below --->
-
-<!--- The current Figure 1 should be replaced by the new figure first shown in revision_notes.Rmd --->
 
@@
 
-<!--- Here a new section is needed --->
-<!---
-
-**Effects of $\gamma$ and correlated interactions on stability**
-
-This is going to require that a new figure be inserted showing correlation versus stability.
-Some general points about how $rho$ is known to affect stability is also needed, but this should not be too difficult to explain in a sentence or two.
-
---->
-
 <!--- Another new section is needed --->
@@
 It will also briefly mention random predator-prey, mutualist, and competitor networks.
-These resutls will require a small table, at least.
+These results will require a small table, at least.
 

### 2020-03-11 notebook/ms.Rmd
Add another paragraph on the relationship between rho and the stability

@@
 
-Numerical simulations reveal that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
+Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
+Figure 2 shows a comparison of $\max(\Re(\lambda))$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$.
+For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\max(\Re(\lambda))$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
+For $\rho \geq -0.1$, the range of $\max(\Re(\lambda))$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$.
+Hence, across a wide range of system correlations, variation in the response rate of system components had a stabilising effect.
+
+The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increases with increasing $S$.
+Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$.
 

### 2020-03-11 notebook/ms.tex
Add another paragraph on the relationship between rho and the stability

@@
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+\subsection{Abstract}\label{abstract}
 
@@
 
-\hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\subsection{Introduction}\label{introduction}
 
@@
 
-\hypertarget{results}{%
-\subsection{Results}\label{results}}
+\subsection{Results}\label{results}
 
@@
 variance of \(M_{ij}\) without affecting the expected covariance between
-\(M_{ij}\) and \(M_{ji}\),
-
-\begin{equation} \label{rho_def}
-\rho = \frac{cov(M_{ij}, M_{ji})}{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})}.
-\end{equation}
-
-Neverthless, numerical simulations reveal that
-\(\sigma^{2}_{\gamma} > 0\) results in a nonlinear relationship between
-\(\rho\) and \(\max(\Re(\lambda))\), which can sometimes increase the
-stability of the system.
+\(M_{ij}\) and \(M_{ji}\).
 
 \textbf{Numerical simulations of random \(\mathbf{M}\) with and without
-\(\mathbf{\sigma_{\gamma}}\)}. I used numerical simulations to isolate
-the effect of \(\sigma_{\gamma}\) on system stability by comparing the
-same \(\textbf{M} = \gamma \textbf{A}\) given \(\sigma_{\gamma} > 0\)
-versus \(\sigma_{\gamma} = 0\) across a rane of parameter combinations.
-Values of \(\gamma_{ii}\) were sampled from a uniform distribution where
-\(\gamma \sim \mathcal{U}(0, 2)\) (see Supplementary Information for
-other distributions, which give similar results). For
-\(\sigma_{\gamma} = 0\), diagonal elements were standardised to the mean
-value of sampled \(\gamma_{ii}\) so that \(E[d]\) of \(\textbf{M}\) were
-identical.
+\(\mathbf{\sigma^{2}_{\gamma}}\)}. I used numerical simulations and
+eigenanalysis to test how variation in \(\gamma\) affects stability in
+random matrices with known properties, comparing the stability of
+\(\textbf{A}\) versus \(\mathbf{M} = \gamma\mathbf{A}\). Values of
+\(\gamma\) were sampled from a uniform distribution where
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\) and \(\sigma^{2}_{\gamma} = 1/3\)
+(see Supplementary Information for other distributions, which give
+similar results). Diagonal elements of simulated \(\mathbf{A}\) were
+standardised to the mean value of sampled \(\gamma\) so that \(E[d]\) of
+\(\textbf{M}\) and \(\textbf{M}\) were guaranteed to be identical. Here
+I focus on the effect of \(\gamma\) across values of \(\rho\), and for
+increasing system sizes (\(S\)) in random and structured networks. By
+increasing \(S\), the objective is to determine the effect of \(\gamma\)
+as system complexity increases toward the boundary at which stability is
+realistic for a finite system.
 
 \textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{\rho}\)}
+
+Numerical simulations reveal that \(\sigma^{2}_{\gamma} > 0\) results in
+a nonlinear relationship between \(\rho\) and \(\max(\Re(\lambda))\),
+which can sometimes increase the stability of the system.
 
@@
 
-\hypertarget{discussion}{%
-\subsection{Discussion}\label{discussion}}
+\subsection{Discussion}\label{discussion}
 
@@
 
-\hypertarget{methods}{%
-\subsection{Methods}\label{methods}}
+\subsection{Methods}\label{methods}
 
@@
 synthesis of eco-evolutionary feedbacks on community stability, Patel et
-al.~model a system that includes a vector of potentially changing
+al. model a system that includes a vector of potentially changing
 species densities (\(\mathbf{N}\)) and a vector of potentially evolving
@@
 effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et
-al.~were interested in stability when the evolution of traits was
-relatively slow or fast in comparison with the change in species
+species \(i\) and the value of trait \(j\), respectively. Patel et al.
+were interested in stability when the evolution of traits was relatively
+slow or fast in comparison with the change in species
 densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, and
@@
 
-I use the same principle that Patel et al.~use to modulate the relative
+I use the same principle that Patel et al. use to modulate the relative
 rate of evolution to modulate rates of component responses for \(S\)
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-May1972}{}%
+\hypertarget{ref-May1972}{}
 1. May, R. M. Will a large complex system be stable? \emph{Nature}
@@
 
-\leavevmode\hypertarget{ref-Mougi2012}{}%
+\hypertarget{ref-Mougi2012}{}
 3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
@@
 
-\leavevmode\hypertarget{ref-Allesina2015}{}%
+\hypertarget{ref-Allesina2015}{}
 4. Allesina, S. \emph{et al.} Predicting the stability of large
@@
 
-\leavevmode\hypertarget{ref-Grilli2017}{}%
+\hypertarget{ref-Grilli2017}{}
 5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
@@
 
-\leavevmode\hypertarget{ref-Gray2008}{}%
+\hypertarget{ref-Gray2008}{}
 6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
@@
 
-\leavevmode\hypertarget{ref-Gray2009}{}%
+\hypertarget{ref-Gray2009}{}
 7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
@@
 
-\leavevmode\hypertarget{ref-Rosenfeld2009}{}%
+\hypertarget{ref-Rosenfeld2009}{}
 8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
@@
 
-\leavevmode\hypertarget{ref-MacArthur2010}{}%
+\hypertarget{ref-MacArthur2010}{}
 9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
@@
 
-\leavevmode\hypertarget{ref-May2008}{}%
+\hypertarget{ref-May2008}{}
 10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
@@
 
-\leavevmode\hypertarget{ref-Haldane2011}{}%
+\hypertarget{ref-Haldane2011}{}
 11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
@@
 
-\leavevmode\hypertarget{ref-Suweis2014}{}%
+\hypertarget{ref-Suweis2014}{}
 12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
@@
 
-\leavevmode\hypertarget{ref-Bardoscia2017}{}%
+\hypertarget{ref-Bardoscia2017}{}
 13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
@@
 
-\leavevmode\hypertarget{ref-Tao2010}{}%
+\hypertarget{ref-Tao2010}{}
 14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
@@
 
-\leavevmode\hypertarget{ref-Watts1998}{}%
+\hypertarget{ref-Watts1998}{}
 15. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
@@
 
-\leavevmode\hypertarget{ref-Albert2002}{}%
+\hypertarget{ref-Albert2002}{}
 16. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
@@
 
-\leavevmode\hypertarget{ref-Williams2000}{}%
+\hypertarget{ref-Williams2000}{}
 17. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
@@
 
-\leavevmode\hypertarget{ref-Patel2018}{}%
+\hypertarget{ref-Patel2018}{}
 18. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
@@
 
-\leavevmode\hypertarget{ref-Tang2014b}{}%
+\hypertarget{ref-Tang2014b}{}
 19. Tang, S. \& Allesina, S. Reactivity and stability of large
@@
 
-\leavevmode\hypertarget{ref-Ahmadian2015}{}%
+\hypertarget{ref-Ahmadian2015}{}
 20. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
@@
 
-\leavevmode\hypertarget{ref-Gibbs2017}{}%
+\hypertarget{ref-Gibbs2017}{}
 21. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
@@
 
-\leavevmode\hypertarget{ref-Stone2017}{}%
+\hypertarget{ref-Stone2017}{}
 22. Stone, L. The feasibility and stability of large complex biological
@@
 
-\leavevmode\hypertarget{ref-Tang2014c}{}%
+\hypertarget{ref-Tang2014c}{}
 23. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
@@
 
-\leavevmode\hypertarget{ref-Hamblin2013}{}%
+\hypertarget{ref-Hamblin2013}{}
 24. Hamblin, S. On the practical usage of genetic algorithms in ecology
@@
 
-\leavevmode\hypertarget{ref-Dougoud2018}{}%
+\hypertarget{ref-Dougoud2018}{}
 25. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
@@
 
-\leavevmode\hypertarget{ref-Song2018}{}%
+\hypertarget{ref-Song2018}{}
 26. Song, C. \& Saavedra, S. Will a small randomly assembled community
@@
 
-\leavevmode\hypertarget{ref-Servan2018}{}%
+\hypertarget{ref-Servan2018}{}
 27. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
@@
 
-\leavevmode\hypertarget{ref-May1973}{}%
+\hypertarget{ref-May1973}{}
 28. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.16.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-11 notebook/ms.Rmd
Another paragraph on the relationship between rho and max eigenvalues, and a bit of tweaking to improve the Figure 3 legend

@@
 Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
-Figure 2 shows a comparison of $\max(\Re(\lambda))$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$.
+Figure 2 shows a comparison of $\max(\Re(\lambda))$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$ and $\sigma = 0.4$.
 For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\max(\Re(\lambda))$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
@@
 
-The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increases with increasing $S$.
-Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$.
+The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$.
+Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
+For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\max(\Re(\lambda))$.
+But as $S$ increased, the curvilinear relationship between $\rho$ and $\max(\Re(\lambda))$ resulted in $\mathbf{M}$ with lower expected $\max(\Re(\lambda))$ for low magnitudes of $\rho$.
+In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
 
@@
 In each panel, 10000 random complex systems $\mathbf{M} = \gamma \mathbf{A}$ are simulated for each correlation $\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}$ between off-diagonal elements $A_{ij}$ and $A_{ji}$.
-Lines show the expected real part of the leading eigenvalues of $\mathbf{M}$ (red squares) versus $\mathbf{A}$ (blue circles) across $\rho$, where negative values (below the dotted black line) indicate system stability.
+Lines show the expected real part of the leading eigenvalues of $\mathbf{M}$ (red squares; $\sigma^{2}_{\gamma} = 1/3$) versus $\mathbf{A}$ (blue circles; $\sigma^{2}_{\gamma} = 0$) across $\rho$, where negative values (below the dotted black line) indicate system stability.
 Differences between lines thereby show the effect of component response rate variation ($\gamma$) on system stability across system correlations and sizes ($S$).

### 2020-03-11 notebook/ms.tex
Another paragraph on the relationship between rho and max eigenvalues, and a bit of tweaking to improve the Figure 3 legend

@@
 
-Numerical simulations reveal that \(\sigma^{2}_{\gamma} > 0\) results in
-a nonlinear relationship between \(\rho\) and \(\max(\Re(\lambda))\),
-which can sometimes increase the stability of the system.
+Ten thousand numerical simulations of \(\mathbf{M}\) revealed that
+\(\sigma^{2}_{\gamma} > 0\) results in a nonlinear relationship between
+\(\rho\) and \(\max(\Re(\lambda))\), which can sometimes increase the
+stability of the system. Figure 2 shows a comparison of
+\(\max(\Re(\lambda))\) across \(\rho\) values when
+\(\sigma^{2}_{\gamma} = 0\) versus \(\sigma^{2}_{\gamma} = 1/3\) given
+\(S = 25\) and \(\sigma = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded
+region of Figure 2), expected \(\max(\Re(\lambda))\) was lower given
+\(\sigma^{2}_{\gamma} = 1/3\). For \(\rho \geq -0.1\), the range of
+\(\max(\Re(\lambda))\) values observed given
+\(\sigma^{2}_{\gamma} = 1/3\) also had a lower bound, which included
+negative values for \(\rho = -0.1\) and \(\rho = 0\). Hence, across a
+wide range of system correlations, variation in the response rate of
+system components had a stabilising effect.
+
+The stabilising effect of \(\sigma^{2}_{\gamma}\) across \(\rho\)
+increased with increasing \(S\). Figure 3 shows 10000 numerical
+simulations of \(\mathbf{M}\) for \(\sigma^{2}_{\gamma} = 0\) versus
+\(\sigma^{2}_{\gamma} = 1/3\) across increasing \(S\) given
+\(\sigma = 0.2\) (\(\sigma\) has been lowered here to better illustrate
+the effect of \(S\); note that now given \(S = 25\),
+\(1 = \sigma\sqrt{SC}\)). For relatively small systems (\(S \leq 25\)),
+positive \(\sigma^{2}_{\gamma}\) never decreased the expected
+\(\max(\Re(\lambda))\). But as \(S\) increased, the curvilinear
+relationship between \(\rho\) and \(\max(\Re(\lambda))\) resulted in
+\(\mathbf{M}\) with lower expected \(\max(\Re(\lambda))\) for low
+magnitudes of \(\rho\). In turn, as \(S\) increased, and systems became
+more complex, positive \(\sigma^{2}_{\gamma}\) increased the proportion
+of numerical simulations that were observed to be stable (see below).
 
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 elements \(A_{ij}\) and \(A_{ji}\). Lines show the expected real part of
-the leading eigenvalues of \(\mathbf{M}\) (red squares) versus
-\(\mathbf{A}\) (blue circles) across \(\rho\), where negative values
+the leading eigenvalues of \(\mathbf{M}\) (red squares;
+\(\sigma^{2}_{\gamma} = 1/3\)) versus \(\mathbf{A}\) (blue circles;
+\(\sigma^{2}_{\gamma} = 0\)) across \(\rho\), where negative values
 (below the dotted black line) indicate system stability. Differences

### 2020-03-11 notebook/ms.Rmd
Change the notation of minimum real part of eigenvalues

@@
 Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma_{M}^{2}$; diagonal elements are set to $d = -1$ [@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\max\left(\Re(\lambda)\right) < 0$) [@May1972; @Allesina2012].
+Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\Re(\lambda_{max}) < 0$) [@May1972; @Allesina2012].
 In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
@@
 **Effect of $\mathbf{\sigma^{2}_{\gamma}}$ on $\mathbf{M}$ (co)variation**.
-The value of $\max(\Re(\lambda))$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
+The value of $\Re(\lambda_{max})$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
 These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
@@
 
-Given $\sigma^{2}_\gamma = 0$, $\max(\Re(\lambda))$ decrease linearly with $\rho$ such that [@Tang2014c],
+Given $\sigma^{2}_\gamma = 0$, $\Re(\lambda_{max})$ decrease linearly with $\rho$ such that [@Tang2014c],
 
@@
 
-Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\max(\Re(\lambda))$, which can sometimes increase the stability of the system.
-Figure 2 shows a comparison of $\max(\Re(\lambda))$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$ and $\sigma = 0.4$.
-For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\max(\Re(\lambda))$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
-For $\rho \geq -0.1$, the range of $\max(\Re(\lambda))$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$.
+Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
+Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$ and $\sigma = 0.4$.
+For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
+For $\rho \geq -0.1$, the range of $\Re(\lambda_{max})$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$.
 Hence, across a wide range of system correlations, variation in the response rate of system components had a stabilising effect.
@@
 Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
-For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\max(\Re(\lambda))$.
-But as $S$ increased, the curvilinear relationship between $\rho$ and $\max(\Re(\lambda))$ resulted in $\mathbf{M}$ with lower expected $\max(\Re(\lambda))$ for low magnitudes of $\rho$.
+For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$.
+But as $S$ increased, the curvilinear relationship between $\rho$ and $\Re(\lambda_{max})$ resulted in $\mathbf{M}$ with lower expected $\Re(\lambda_{max})$ for low magnitudes of $\rho$.
 In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
@@
 
-If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\max(\Re(\lambda))$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
+If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\Re(\lambda_{max})$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
 These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
@@
 Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013].
-Generations of selection on $\gamma$ value combinations to minimise $\max\left(\Re(\lambda)\right)$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
+Generations of selection on $\gamma$ value combinations to minimise $\Re(\lambda_{max})$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
 Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
@@
 The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\max\left(\Re(\lambda)\right)$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
 Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
@@
 Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
-The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\max\left(\Re(\lambda)\right)$ for $\mathbf{M}$).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\Re(\lambda_{max})$ for $\mathbf{M}$).
 

### 2020-03-11 notebook/ms.tex
Change the notation of minimum real part of eigenvalues

@@
 \(\mathbf{M}\) are negative
-(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
+(\(\Re(\lambda_{max}) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
 In a large system (high \(S\)), eigenvalues are distributed
@@
 \textbf{Effect of \(\mathbf{\sigma^{2}_{\gamma}}\) on \(\mathbf{M}\)
-(co)variation}. The value of \(\max(\Re(\lambda))\), and therefore
+(co)variation}. The value of \(\Re(\lambda_{max})\), and therefore
 system stability, can be estimated from five properties of
@@
 
-Given \(\sigma^{2}_\gamma = 0\), \(\max(\Re(\lambda))\) decrease
+Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) decrease
 linearly with \(\rho\) such
@@
 \(\sigma^{2}_{\gamma} > 0\) results in a nonlinear relationship between
-\(\rho\) and \(\max(\Re(\lambda))\), which can sometimes increase the
+\(\rho\) and \(\Re(\lambda_{max})\), which can sometimes increase the
 stability of the system. Figure 2 shows a comparison of
-\(\max(\Re(\lambda))\) across \(\rho\) values when
+\(\Re(\lambda_{max})\) across \(\rho\) values when
 \(\sigma^{2}_{\gamma} = 0\) versus \(\sigma^{2}_{\gamma} = 1/3\) given
 \(S = 25\) and \(\sigma = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded
-region of Figure 2), expected \(\max(\Re(\lambda))\) was lower given
+region of Figure 2), expected \(\Re(\lambda_{max})\) was lower given
 \(\sigma^{2}_{\gamma} = 1/3\). For \(\rho \geq -0.1\), the range of
-\(\max(\Re(\lambda))\) values observed given
+\(\Re(\lambda_{max})\) values observed given
 \(\sigma^{2}_{\gamma} = 1/3\) also had a lower bound, which included
@@
 positive \(\sigma^{2}_{\gamma}\) never decreased the expected
-\(\max(\Re(\lambda))\). But as \(S\) increased, the curvilinear
-relationship between \(\rho\) and \(\max(\Re(\lambda))\) resulted in
-\(\mathbf{M}\) with lower expected \(\max(\Re(\lambda))\) for low
+\(\Re(\lambda_{max})\). But as \(S\) increased, the curvilinear
+relationship between \(\rho\) and \(\Re(\lambda_{max})\) resulted in
+\(\mathbf{M}\) with lower expected \(\Re(\lambda_{max})\) for low
 magnitudes of \(\rho\). In turn, as \(S\) increased, and systems became
@@
 Generations of selection on \(\gamma\) value combinations to minimise
-\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
-\(Var(\gamma)\) to increase system stability. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were up to four
-orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
-that stability could often be achieved by manipulating \(S\) \(\gamma\)
-values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
-manipulating component response rates rather than interactions between
-components).
+\(\Re(\lambda_{max})\) demonstrated the potential for \(Var(\gamma)\) to
+increase system stability. Across \(S = \{2, 3, ..., 39, 40\}\), sets of
+\(\gamma\) values were found that resulted in stable systems with
+probabilities that were up to four orders of magnitude higher than when
+\(\gamma = 1\) (Fig. 4), meaning that stability could often be achieved
+by manipulating \(S\) \(\gamma\) values rather than \(S \times S\)
+\(\mathbf{M}\) elements (i.e., by manipulating component response rates
+rather than interactions between components).
 
@@
 the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal
-offspring with subsequent random mutation and crossover between the
-resulting new generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation
-of each \(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+lowest \(\Re(\lambda_{max})\) each produced 50 clonal offspring with
+subsequent random mutation and crossover between the resulting new
+generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
+\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
 probability of 0.2, resulting in a mutation effect of size
@@
 fitness increase between generations was less than 0.01 (where fitness
-was defined as \(W_{\gamma} = -\max\left(\Re(\lambda)\right)\) for
-\(\mathbf{M}\)).
+was defined as \(W_{\gamma} = -\Re(\lambda_{max})\) for \(\mathbf{M}\)).
 
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.18.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-11 notebook/ms.Rmd
Add the mention of C values into the results for rho versus eigenvalues

@@
 Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
-Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$ and $\sigma = 0.4$.
+Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$, $C = 1$, and $\sigma = 0.4$.
 For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
@@
 The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$.
-Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
+Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $C = 1$ and $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
 For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$.

### 2020-03-11 notebook/ms.Rmd
Fix some notation and rearrange the Results sections so that feasibility comes before the genetic algorithm

@@
 In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
-Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Figure 1a).
+Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Fig.
+1a).
 
@@
 
-The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Figure 1).
-Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{A}$ and $\textbf{M}$ lie within a circle of a radius $\sigma_{A}\sqrt{S}$ and $\sigma_{M}\sqrt{S}$, respectively (Figure 1a vs.
+The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Fig.
+1).
+Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{A}$ and $\textbf{M}$ lie within a circle of a radius $\sigma_{A}\sqrt{S}$ and $\sigma_{M}\sqrt{S}$, respectively (Fig.
+1a vs.
 1b).
-If $d \neq 0$, positive $\sigma^{2}_\gamma$ changes the distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], potentially affecting stability (Figure 1c vs.
+If $d \neq 0$, positive $\sigma^{2}_\gamma$ changes the distribution of eigenvalues [@Ahmadian2015; @Gibbs2017; @Stone2017], potentially affecting stability (Fig.
+1c vs.
 1d).
@@
 Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$, $C = 1$, and $\sigma = 0.4$.
-For $-0.4 \leq \rho \leq 0.7$ (shaded region of Figure 2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
+For $-0.4 \leq \rho \leq 0.7$ (shaded region of Fig.
+2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
 For $\rho \geq -0.1$, the range of $\Re(\lambda_{max})$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$.
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $Var(\gamma)$ on stability across systems of varying complexity, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
-One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ versus $\gamma \sim \mathcal{U}(0, 2)$.
-For all $S > 10$, I found that the number of stable random systems was higher given $Var(\gamma)$ than when $\gamma = 1$ (Fig.
+To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of varying complexity, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
+One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ ($\sigma^{2}_{\gamma} = 0$) versus $\gamma \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
+For all $S > 10$, I found that the number of stable random systems was higher given $\sigma^{2}_{\gamma} > 0$ than when $\sigma^{2}_{\gamma} = 0$ (Fig.
 3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
-In other words, the potential for $Var(\gamma)$ to affect stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
-For the highest values of $S$, nearly all systems that were stable given $Var(\gamma)$ would not have been stable given $\gamma = 1$.
+In other words, the potential for $\sigma^{2}_{\gamma}$ to affect stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
+For the highest values of $S$, nearly all systems that were stable given varying $\gamma$ would not have been stable given $\gamma = 1$.
 
@@
 
-<!--- Tempting to remove this section entirely, but if there is enough space for it, then maybe retain it; reviewers did not mention a problem with it, and it's a short and relevant point to make.
-In any case, it should be swapped with the section on feasibility below.
---->
-**Targeted manipulation of $\mathbf{\gamma}$**.
-To further investigate the potential of $Var(\gamma)$ to be stabilising, I used a genetic algorithm.
-Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013].
-Generations of selection on $\gamma$ value combinations to minimise $\Re(\lambda_{max})$ demonstrated the potential for $Var(\gamma)$ to increase system stability.
-Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (Fig.
-4), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements (i.e., by manipulating component response rates rather than interactions between components).
-
 <!--- I don't think much here needs to change --->
@@
 I emphasise that $\gamma$ is not interpreted as population density in this analysis, but instead as a fundamental property of species life history such as expected generation time.
-Feasibility was unaffected by $Var(\gamma)$ and instead occurred with a fixed probability of $1/2^{S}$, consistent with a recent proof by Serv&aacute;n et al.
+Feasibility was unaffected by $\sigma^{2}_{\gamma}$ and instead occurred with a fixed probability of $1/2^{S}$, consistent with a recent proof by Serv&aacute;n et al.
 [@Servan2018] (see Supplementary Information).
 Hence, for pure interacting species networks, variation in component response rate (i.e., species generation time) does not affect stability at biologically realistic species densities.
+
+<!--- Tempting to remove this section entirely, but if there is enough space for it, then maybe retain it; reviewers did not mention a problem with it, and it's a short and relevant point to make.
+In any case, it should be swapped with the section on feasibility below.
+--->
+**Targeted manipulation of $\mathbf{\gamma}$**.
+To further investigate the potential of $\sigma^{2}_{\gamma}$ to be stabilising, I used a genetic algorithm.
+Genetic algorithms are heuristic tools that mimic evolution by natural selection, and are useful when the space of potential solutions (in this case, possible combinations of $\gamma$ values leading to stability in a complex system) is too large to search exhaustively [@Hamblin2013].
+Generations of selection on $\gamma$ value combinations to minimise $\Re(\lambda_{max})$ demonstrated the potential for $\sigma^{2}_{\gamma}$ to increase system stability.
+Across $S = \{2, 3, ..., 39, 40\}$, sets of $\gamma$ values were found that resulted in stable systems with probabilities that were up to four orders of magnitude higher than when $\gamma = 1$ (see Supplementary Information), meaning that stability could often be achieved by manipulating $S$ $\gamma$ values rather than $S \times S$ $\mathbf{M}$ elements (i.e., by manipulating component response rates rather than interactions between components).
 

### 2020-03-11 notebook/ms.Rmd
Write a paragraph on structured networks in the Results

@@
 For the highest values of $S$, nearly all systems that were stable given varying $\gamma$ would not have been stable given $\gamma = 1$.
+
+**Simulation of structured $\mathbf{M}$ across $\mathbf{S}$**.
+
+To investigate how $\sigma^{2}_{\gamma}$ affects stability for different network structures, I simulated $\mathbf{M}$ with fixed properties across $S$.
+This included three types of random networks that are typically interpreted in terms of differing interspecific ecological interaction types [@Allesina2011; @Allesina2012].
+These interactions types are competitive, mutualist, or predator-prey, as modelled by off-diagonal elements of $\mathbf{M}$ that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012].
+I also simulated three commonly observed system structures, including cascade food webs [@Williams2000], small-world networks [@Watts1998], and scale-free networks [@Albert2002].
+For all except mutualist networks, as $S$ increased, a higher number of stable systems were observed given $\sigma^{2}_{\gamma}>0$ versus $\sigma^{2}_{\gamma}=0$.
 

### 2020-03-11 notebook/ms.Rmd
Add more text for a draft of the results

--- 2020-03-11 notebook/ms.Rmd
@@
 Here I show that, despite higher $\sigma_{M}$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
-My results are robust across commonly observed network structures, including random [@May1972], small-world [@Watts1998], scale-free [@Albert2002], cascade food web [@Williams2000] networks.
+My results are robust across commonly observed network structures, including random [@May1972], small-world [@Watts1998], scale-free [@Albert2002], cascade food web [@Solow1998; @Williams2000] networks.
 
@@
 \begin{equation} \label{rho_stab}
-\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\Re(\lambda_{max}) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
 \end{equation}
@@
 In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
-
- <!--- Replace the below with some more informative text concerning the variation that $\var(gamma})$ adds to the system $M$ overall.
-This might require starting with just $M = A$ as an explanation.
-Borrowing the text from revision_notes.Rmd below --->
-
-<!---
-
-If the off-diagonal elements of $\textbf{A}$ are sampled independently from an identical distribution, then  $\Re(\lambda_{max})$ for $\textbf{M} = \textbf{A}$ can be estimated from five values [@Tang2014b].
-These values include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
-When investigating the effect of varying component response rate $Var(\gamma)$ on stability by defining $\textbf{M} = \gamma\textbf{A}$, $S$ remains unchanged.
-Further, values of $\gamma_{i}$ were sampled such that $E[d]$ and $E[\mu]$ also remained unchanged (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$).
-What $Var(\gamma)$ does change is the variation in component interaction strengths, and $\rho$.
-
-Introducing variation in $\gamma$ increases the total variation in the system, making it more likely that $\textbf{M} = \gamma\textbf{A}$ is unstable.
-Variation of the off-diagonal elements in $\textbf{M}$ is described by the joint variation of two random variables (to simplify the notation, $\sigma^{2}_{M}$ and $\sigma^{2}_{A}$ refer to variances of off-diagonal elements only),
-
-$$\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.$$
-
-Note that in my simulations $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, so the above can be simplified,
-
-$$\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).$$
-
-The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$.
-Given $d = 0$ and $C = 1$, the distribution of eigenvalues of $\textbf{M}$ and $\textbf{A}$ lie within a circle of a radius $\sigma^{2}_{M}\sqrt{S}$ and $\sigma^{2}_{A}\sqrt{S}$, respectively.
-
---->
 
@@
 
-<!--- Another new section is needed --->
+I also simulated 100000 $\mathbf{M}$ for three types of random networks that are typically interpreted as differing interspecific ecological interaction types [@Allesina2011; @Allesina2012].
+These interactions types are competitive, mutualist, or predator-prey, as modelled by off-diagonal elements of $\mathbf{M}$ that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{M}$).
+As $S$ increased, a higher number of stable systems were observed given $\sigma^{2}_{\gamma}>0$ versus $\sigma^{2}_{\gamma}=0$ for competitor and predator-prey, but not mutualist, systems.
+A higher number of stable systems were observed whenever $S > 12$ and $S > 40$ for competitive and predator-prey systems, respectively (note that $\rho < 0$ by definition for predator-prey systems, making stability more likely overall).
+The stability of mutualist systems was never affected by $\sigma^{2}_{\gamma}$.
+
+The effect of $\sigma^{2}_{\gamma}$ on stability did not change qualitatively across values of $C$, $\sigma$, or different distributions of $\gamma$ (see Supporting Information).
+
+**Simulation of structured $\mathbf{M}$ across $\mathbf{S}$**.
+To investigate how $\sigma^{2}_{\gamma}$ affects the stability of commonly observed network structures, I simulated 1 million $\textbf{M}$ for small-world [@Watts1998], scale-free [@Albert2002], and cascade food web [@Solow1998; @Williams2000] networks.
+In all of these networks, rules determining the presence or absence of an interaction between components $i$ and $j$ constrained the overall structure of the network.
+In small-world networks, interactions between components are constrained so that the expected degree of separation between any two components increases in proportion to $\log(S)$ [@Watts1998].
+In scale-free networks, the number of components with which a focal component interacts follows a power law; typically a few components have many interactions while most components have few interactions [@Albert2002].
+In cascade food webs, species are ranked and interactions are constrained such that a species $i$ can only feed on $j$ if the rank of $i > j$.
+
+Network structure did not strongly modulate the effect that $\sigma^{2}_{\gamma}$ had on stability.
+For comparable magnitudes of complexity, structured networks still had a higher number of stable $\mathbf{M}$ for $\sigma^{2}_{\gamma}>0$ compared to $\sigma^{2}_{\gamma}=0$.
+For random networks, positive $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma \sqrt{SC} \gtrapprox 1.26$.
+This threshold of complexity was comparable for small-world networks, and slightly lower for scale-free networks (note that algorithms for generating small-world and scale-free networks necessarily led to varying $C$, see methods).
+Varying $\gamma$ increased stability in cascade food webs for $S > 27$, and therefore at a relatively low complexity magnitudes compared to random predator-prey networks.
+Overall, network structure does not greatly change the effect that $\sigma^{2}_{\gamma}$ has on increasing the upper bound of complexity within which stability might reasonably be observed.
+
 <!---
 
-**Effects of $\gamma$ on structured networks**
-
-This will report results for small-world, scale-free, and cascade food-webs.
-It will also briefly mention random predator-prey, mutualist, and competitor networks.
-These results will require a small table, at least.
+For all simulated network structures, interaction strengths between components were sampled as in random networks ($\mathcal{N}(0, 0.4^{2})$), but algorithms for building networks necessarily resulted in variable $C$.
 
@@
 
-<!--- I don't think much here needs to change --->
 **System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.

### 2020-03-11 notebook/ms.tex
Add more text for a draft of the results

@@
 cascade food
-web\textsuperscript{\protect\hyperlink{ref-Williams2000}{17}} networks.
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{17},\protect\hyperlink{ref-Williams2000}{18}}
+networks.
 
@@
 system \(\mathbf{M}\) can be defined as
-follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}},
+follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}},
 
@@
 \(\sigma_{M} \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
-(Figure 1a).
+(Fig. 1a).
 
@@
 system stability, can be estimated from five properties of
-\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{19}}.
+\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{20}}.
 These properties include (1) system size (\(S\)), (2) mean
@@
 eigenvalue spectra of \(\textbf{A}\) versus
-\(\textbf{M} = \gamma\textbf{A}\) (Figure 1). Given \(d = 0\) and
+\(\textbf{M} = \gamma\textbf{A}\) (Fig. 1). Given \(d = 0\) and
 \(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
 \(\textbf{M}\) lie within a circle of a radius \(\sigma_{A}\sqrt{S}\)
-and \(\sigma_{M}\sqrt{S}\), respectively (Figure 1a vs.~1b). If
+and \(\sigma_{M}\sqrt{S}\), respectively (Fig. 1a vs.~1b). If
 \(d \neq 0\), positive \(\sigma^{2}_\gamma\) changes the distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}--\protect\hyperlink{ref-Stone2017}{22}},
-potentially affecting stability (Figure 1c vs.~1d).
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}--\protect\hyperlink{ref-Stone2017}{23}},
+potentially affecting stability (Fig. 1c vs.~1d).
 
@@
 linearly with \(\rho\) such
-that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{23}},
+that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{24}},
 
 \begin{equation} \label{rho_stab}
-\max(\Re(\lambda)) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\Re(\lambda_{max}) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
 \end{equation}
@@
 \(\sigma^{2}_{\gamma} = 0\) versus \(\sigma^{2}_{\gamma} = 1/3\) given
-\(S = 25\) and \(\sigma = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded
-region of Figure 2), expected \(\Re(\lambda_{max})\) was lower given
-\(\sigma^{2}_{\gamma} = 1/3\). For \(\rho \geq -0.1\), the range of
-\(\Re(\lambda_{max})\) values observed given
-\(\sigma^{2}_{\gamma} = 1/3\) also had a lower bound, which included
-negative values for \(\rho = -0.1\) and \(\rho = 0\). Hence, across a
-wide range of system correlations, variation in the response rate of
-system components had a stabilising effect.
+\(S = 25\), \(C = 1\), and \(\sigma = 0.4\). For
+\(-0.4 \leq \rho \leq 0.7\) (shaded region of Fig. 2), expected
+\(\Re(\lambda_{max})\) was lower given \(\sigma^{2}_{\gamma} = 1/3\).
+For \(\rho \geq -0.1\), the range of \(\Re(\lambda_{max})\) values
+observed given \(\sigma^{2}_{\gamma} = 1/3\) also had a lower bound,
+which included negative values for \(\rho = -0.1\) and \(\rho = 0\).
+Hence, across a wide range of system correlations, variation in the
+response rate of system components had a stabilising effect.
 
@@
 simulations of \(\mathbf{M}\) for \(\sigma^{2}_{\gamma} = 0\) versus
-\(\sigma^{2}_{\gamma} = 1/3\) across increasing \(S\) given
-\(\sigma = 0.2\) (\(\sigma\) has been lowered here to better illustrate
-the effect of \(S\); note that now given \(S = 25\),
+\(\sigma^{2}_{\gamma} = 1/3\) across increasing \(S\) given \(C = 1\)
+and \(\sigma = 0.2\) (\(\sigma\) has been lowered here to better
+illustrate the effect of \(S\); note that now given \(S = 25\),
 \(1 = \sigma\sqrt{SC}\)). For relatively small systems (\(S \leq 25\)),
@@
 \textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
-investigate the effect of \(Var(\gamma)\) on stability across systems of
-varying complexity, I simulated random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One
-million \(\mathbf{M}\) were simulated for each \(S\), and the stability
-of \(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
-\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
-number of stable random systems was higher given \(Var(\gamma)\) than
-when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
-table of results), and that the difference between the probabilities of
-observing a stable system increased with an increase in \(S\). In other
-words, the potential for \(Var(\gamma)\) to affect stability increased
-with system complexity and was most relevant for systems on the cusp of
-being too complex to be realistically stable. For the highest values of
-\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
-not have been stable given \(\gamma = 1\).
+investigate the effect of \(\sigma^{2}_{\gamma}\) on stability across
+systems of varying complexity, I simulated random \(\mathbf{M}\)
+matrices at \(\sigma = 0.4\) and \(C = 1\) across
+\(S = \{2, 3, ..., 49, 50\}\). One million \(\mathbf{M}\) were simulated
+for each \(S\), and the stability of \(\mathbf{M}\) was assessed given
+\(\gamma = 1\) (\(\sigma^{2}_{\gamma} = 0\)) versus
+\(\gamma \sim \mathcal{U}(0, 2)\) (\(\sigma^{2}_{\gamma} = 1/3\)). For
+all \(S > 10\), I found that the number of stable random systems was
+higher given \(\sigma^{2}_{\gamma} > 0\) than when
+\(\sigma^{2}_{\gamma} = 0\) (Fig. 3; see Supplementary Information for
+full table of results), and that the difference between the
+probabilities of observing a stable system increased with an increase in
+\(S\). In other words, the potential for \(\sigma^{2}_{\gamma}\) to
+affect stability increased with system complexity and was most relevant
+for systems on the cusp of being too complex to be realistically stable.
+For the highest values of \(S\), nearly all systems that were stable
+given varying \(\gamma\) would not have been stable given
+\(\gamma = 1\).
 
-\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
-investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
-genetic algorithm. Genetic algorithms are heuristic tools that mimic
-evolution by natural selection, and are useful when the space of
-potential solutions (in this case, possible combinations of \(\gamma\)
-values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{24}}.
-Generations of selection on \(\gamma\) value combinations to minimise
-\(\Re(\lambda_{max})\) demonstrated the potential for \(Var(\gamma)\) to
-increase system stability. Across \(S = \{2, 3, ..., 39, 40\}\), sets of
-\(\gamma\) values were found that resulted in stable systems with
-probabilities that were up to four orders of magnitude higher than when
-\(\gamma = 1\) (Fig. 4), meaning that stability could often be achieved
-by manipulating \(S\) \(\gamma\) values rather than \(S \times S\)
-\(\mathbf{M}\) elements (i.e., by manipulating component response rates
-rather than interactions between components).
+I also simulated 100000 \(\mathbf{M}\) for three types of random
+networks that are typically interpreted as differing interspecific
+ecological interaction
+types\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{25}}.
+These interactions types are competitive, mutualist, or predator-prey,
+as modelled by off-diagonal elements of \(\mathbf{M}\) that are
+constrained to be negative, positive, or paired such that if
+\(A_{ij} > 0\) then \(A_{ji} < 0\),
+respectively\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+(but are otherwise identical to the purely random \(\mathbf{M}\)). As
+\(S\) increased, a higher number of stable systems were observed given
+\(\sigma^{2}_{\gamma}>0\) versus \(\sigma^{2}_{\gamma}=0\) for
+competitor and predator-prey, but not mutualist, systems. A higher
+number of stable systems were observed whenever \(S > 12\) and
+\(S > 40\) for competitive and predator-prey systems, respectively (note
+that \(\rho < 0\) by definition for predator-prey systems, making
+stability more likely overall). The stability of mutualist systems was
+never affected by \(\sigma^{2}_{\gamma}\).
+
+The effect of \(\sigma^{2}_{\gamma}\) on stability did not change
+qualitatively across values of \(C\), \(\sigma\), or different
+distributions of \(\gamma\) (see Supporting Information).
+
+\textbf{Simulation of structured \(\mathbf{M}\) across \(\mathbf{S}\)}.
+To investigate how \(\sigma^{2}_{\gamma}\) affects the stability of
+commonly observed network structures, I simulated 1 million
+\(\textbf{M}\) for
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}, and
+cascade food
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{17},\protect\hyperlink{ref-Williams2000}{18}}
+networks. In all of these networks, rules determining the presence or
+absence of an interaction between components \(i\) and \(j\) constrained
+the overall structure of the network. In small-world networks,
+interactions between components are constrained so that the expected
+degree of separation between any two components increases in proportion
+to \(\log(S)\)\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.
+In scale-free networks, the number of components with which a focal
+component interacts follows a power law; typically a few components have
+many interactions while most components have few
+interactions\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. In
+cascade food webs, species are ranked and interactions are constrained
+such that a species \(i\) can only feed on \(j\) if the rank of
+\(i > j\).
+
+Network structure did not strongly modulate the effect that
+\(\sigma^{2}_{\gamma}\) had on stability. For comparable magnitudes of
+complexity, structured networks still had a higher number of stable
+\(\mathbf{M}\) for \(\sigma^{2}_{\gamma}>0\) compared to
+\(\sigma^{2}_{\gamma}=0\). For random networks, positive
+\(\sigma^{2}_{\gamma}\) increased stability given \(S > 10\)
+(\(\sigma = 0.4\) and \(C = 1\)), and therefore complexity
+\(\sigma \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity was
+comparable for small-world networks, and slightly lower for scale-free
+networks (note that algorithms for generating small-world and scale-free
+networks necessarily led to varying \(C\), see methods). Varying
+\(\gamma\) increased stability in cascade food webs for \(S > 27\), and
+therefore at a relatively low complexity magnitudes compared to random
+predator-prey networks. Overall, network structure does not greatly
+change the effect that \(\sigma^{2}_{\gamma}\) has on increasing the
+upper bound of complexity within which stability might reasonably be
+observed.
 
@@
 positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{25},\protect\hyperlink{ref-Song2018}{26}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{26},\protect\hyperlink{ref-Song2018}{27}}.
 This is of particular interest for ecological communities because
@@
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}}.
 While my results are intended to be general to all complex systems, and
@@
 property of species life history such as expected generation time.
-Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
-a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{27}}
-(see Supplementary Information). Hence, for pure interacting species
+Feasibility was unaffected by \(\sigma^{2}_{\gamma}\) and instead
+occurred with a fixed probability of \(1/2^{S}\), consistent with a
+recent proof by Serván et
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}} (see
+Supplementary Information). Hence, for pure interacting species
 networks, variation in component response rate (i.e., species generation
@@
 densities.
+
+\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
+investigate the potential of \(\sigma^{2}_{\gamma}\) to be stabilising,
+I used a genetic algorithm. Genetic algorithms are heuristic tools that
+mimic evolution by natural selection, and are useful when the space of
+potential solutions (in this case, possible combinations of \(\gamma\)
+values leading to stability in a complex system) is too large to search
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}.
+Generations of selection on \(\gamma\) value combinations to minimise
+\(\Re(\lambda_{max})\) demonstrated the potential for
+\(\sigma^{2}_{\gamma}\) to increase system stability. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were up to four
+orders of magnitude higher than when \(\gamma = 1\) (see Supplementary
+Information), meaning that stability could often be achieved by
+manipulating \(S\) \(\gamma\) values rather than \(S \times S\)
+\(\mathbf{M}\) elements (i.e., by manipulating component response rates
+rather than interactions between components).
 
@@
 My focus is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21}}, who applied
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}}, who applied
 the same mathematical framework to investigate how a diagonal matrix
@@
 \(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21}} analytically
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}} analytically
 demonstrated that the effect of \(\mathbf{X}\) on system stability
@@
 of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{22},\protect\hyperlink{ref-Dougoud2018}{25},\protect\hyperlink{ref-Servan2018}{27}}
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{23},\protect\hyperlink{ref-Dougoud2018}{26},\protect\hyperlink{ref-Servan2018}{28}}
 because population densities cannot realistically be negative. My
@@
 empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, but
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}, but
 instead interact with evolutionary, abiotic, or social-economic systems.
@@
 distributions caused by a vector \(\gamma\) has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{20}}
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}}
 and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{21},\protect\hyperlink{ref-Stone2017}{22}},
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22},\protect\hyperlink{ref-Stone2017}{23}},
 but \(\gamma\) has not been interpreted as rates of response of
@@
 traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}.
 For any species \(i\) or trait \(j\), change in species density
@@
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}, and
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{18}}.
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}.
 
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{28}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{30}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{28}},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{30}},
 
@@
 components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{20},\protect\hyperlink{ref-May1973}{28}}.
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{21},\protect\hyperlink{ref-May1973}{30}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{24}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}. In
 this case, the practical solution is finding vectors of
@@
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{25}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{27}}),
+also\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}}),
 
@@
 
+\hypertarget{ref-Solow1998}{}
+17. Solow, A. R. \& Beet, A. R. On lumping species in food webs.
+\emph{Ecology} \textbf{79,} 2013--2018 (1998).
+
 \hypertarget{ref-Williams2000}{}
-17. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
+18. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
 webs. \emph{Nature} \textbf{404,} 180--183 (2000).
@@
 \hypertarget{ref-Patel2018}{}
-18. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+19. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 \hypertarget{ref-Tang2014b}{}
-19. Tang, S. \& Allesina, S. Reactivity and stability of large
+20. Tang, S. \& Allesina, S. Reactivity and stability of large
 ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
@@
 \hypertarget{ref-Ahmadian2015}{}
-20. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+21. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
 with partially structured and partially random connectivity.
@@
 \hypertarget{ref-Gibbs2017}{}
-21. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+22. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.
@@
 \hypertarget{ref-Stone2017}{}
-22. Stone, L. The feasibility and stability of large complex biological
+23. Stone, L. The feasibility and stability of large complex biological
 networks: a random matrix approach. \emph{Scientific Reports}
@@
 \hypertarget{ref-Tang2014c}{}
-23. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
+24. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
 strengths drives stability in large ecological networks. \textbf{17,}
@@
 
-\hypertarget{ref-Hamblin2013}{}
-24. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
+\hypertarget{ref-Allesina2011}{}
+25. Allesina, S. \& Levine, J. M. A competitive network theory of
+species diversity. \emph{Proceedings of the National Academy of Sciences
+of the United States of America} \textbf{108,} 5638--5642 (2011).
 
 \hypertarget{ref-Dougoud2018}{}
-25. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+26. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-26. Song, C. \& Saavedra, S. Will a small randomly assembled community
+27. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-27. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+28. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 
+\hypertarget{ref-Hamblin2013}{}
+29. Hamblin, S. On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+184--194 (2013).
+
 \hypertarget{ref-May1973}{}
-28. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+30. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.18.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.12.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-11 notebook/SI.Rmd
Rearrange the SI to better match the main text

@@
 - [Stability across increasing $S$](#IncrS) 
-- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
 - [Stability of ecological networks](#ecological) 
@@
 - [Feasibility of complex systems](#Feasibility)
+- [Stability given targeted manipulation of $\gamma$ (genetic algorithm)](#ga)
 - [Consistency with Gibbs et al. (2018)](#Gibbs)
@@
 Note that feasibility results were omitted for the table above, but are [reported below](#Feasibility).
-
-Stability given targeted manipulation of $\gamma$ (genetic algorithm) {#ga}
-================================================================================
-
-The figure below compares the stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted manipulation of $\gamma$ elements.
-For each $S$, 100000 complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
-Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
-The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $\sigma^{2}_{\gamma} = 0$.
-The y-axis shows the $\ln$ number of systems that are stable across $S = \{1, 2, ..., 39, 40\}$ for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
-
-```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
-dat <- dat[,-1];
-plot_stable_4(dat);
-```
-
-Stability results are also shown in the table below.
-Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
-
-```{r, echo = FALSE}
-get_top_evo_out <- function(evo_out, size){
-    highest <- max(evo_out);
-    gammas  <- NULL;
-    while(size > 0){
-        pos <- which(evo_out == highest);
-        len <- length(pos);
-        nli <- NULL;
-        for(i in 1:len){
-            start          <- pos[i] + 1;
-            end            <- pos[i] + highest;
-            gammas[[size]] <- evo_out[start:end];
-            size           <- size - 1;
-            if(size == 0){
-                break;
-            }
-        }
-        highest <- highest - 1;
-    }
-    return(gammas);
-}
-plot_evo_out <- function(evo_out){
-    gammas <- get_top_evo_out(evo_out, size = 9);
-    for(i in 1:9){
-        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
-    }
-    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
-    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         xlim = c(0, 2.1));
-    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 2.1));
-    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
-         yaxt = "n", xlim = c(0, 2.1));
-    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
-          text = expression(paste("Component response rate (",gamma,")")));
-    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
-}
-evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
-evo_res <- evo_res[,-1];
-kable(evo_res);
-```
-
-The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
-This comparison shows the high number of stable $\mathbf{M}$ that can be produced through a targeted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
-Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.
-
-Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
-
-```{r, fig.height = 4, fig.width = 6, echo = FALSE}
-evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
-plot_evo_out(evo_out);
-```
-
-The distribution of $\gamma$ values found by the genetic algorithm is uniform.
-A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $\mathbf{M}$.
 
@@
 
+Stability given targeted manipulation of $\gamma$ (genetic algorithm) {#ga}
+================================================================================
+
+The figure below compares the stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted manipulation of $\gamma$ elements.
+For each $S$, 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in $\gamma$ using a genetic algorithm to maximise the effect of $\gamma$ values on increasing stability, as compared to stability in an otherwise identical system in which $\gamma$ is the same for all components.
+Blue bars show the number of stable systems in the absence of component response rate variation, while red bars show the number of stable systems that can be generated if component response rate is varied to maximise system stability.
+The black line shows the proportion of systems that are stable when component response rate is targeted to increase stability, but would not be stable if $\sigma^{2}_{\gamma} = 0$.
+The y-axis shows the $\ln$ number of systems that are stable across $S = \{1, 2, ..., 39, 40\}$ for $C = 1$, and the proportion of systems wherein a targeted search of $\gamma$ values successfully resulted in system stability.
+
+```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
+dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
+dat <- dat[,-1];
+plot_stable_4(dat);
+```
+
+Stability results are also shown in the table below.
+Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
+
+```{r, echo = FALSE}
+get_top_evo_out <- function(evo_out, size){
+    highest <- max(evo_out);
+    gammas  <- NULL;
+    while(size > 0){
+        pos <- which(evo_out == highest);
+        len <- length(pos);
+        nli <- NULL;
+        for(i in 1:len){
+            start          <- pos[i] + 1;
+            end            <- pos[i] + highest;
+            gammas[[size]] <- evo_out[start:end];
+            size           <- size - 1;
+            if(size == 0){
+                break;
+            }
+        }
+        highest <- highest - 1;
+    }
+    return(gammas);
+}
+plot_evo_out <- function(evo_out){
+    gammas <- get_top_evo_out(evo_out, size = 9);
+    for(i in 1:9){
+        gammas[[i]] <- 2 * gammas[[i]]; # Account for mistake in return
+    }
+    par(mfrow = c(3, 3), mar = c(0.5, 0.5, 0.5, 0.5), oma = c(5, 5, 1, 1));
+    hist(gammas[[1]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[2]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[3]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[4]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[5]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[6]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xaxt = "n", yaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[7]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         xlim = c(0, 2.1));
+    hist(gammas[[8]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 2.1));
+    hist(gammas[[9]], main = "", breaks = 20, col = "grey", ylim = c(0, 6),
+         yaxt = "n", xlim = c(0, 2.1));
+    mtext(side = 1, outer = TRUE, line = 3, cex = 1.5,
+          text = expression(paste("Component response rate (",gamma,")")));
+    mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
+}
+evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
+evo_res <- evo_res[,-1];
+kable(evo_res);
+```
+
+The distributions of nine $\gamma$ vectors from the highest $S$ values are shown below.
+This comparison shows the high number of stable $\mathbf{M}$ that can be produced through a targeted search of $\gamma$ values, and suggests that many otherwise unstable systems could potentially be stabilised by an informed manipulation of their component response times.
+Such a possibility might conceivably reduce the dimensionality of problems involving stability in social-ecological or economic systems.
+
+Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
+
+```{r, fig.height = 4, fig.width = 6, echo = FALSE}
+evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
+plot_evo_out(evo_out);
+```
+
+The distribution of $\gamma$ values found by the genetic algorithm is uniform.
+A uniform distribution was used to initialise $\gamma$ values, so there is therefore no evidence that a particular distribution of $\gamma$ is likely to be found to stabilise a matrix $\mathbf{M}$.
+
 Consistency with Gibbs et al. (2018) {#Gibbs}

### 2020-03-11 notebook/SI.tex
Rearrange the SI to better match the main text

@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
 \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
 \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
 \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
 \newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
 \newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{longtable,booktabs}
@@
 
-\hypertarget{supplemental-information-table-of-contents}{%
 \section{Supplemental Information table of
-contents}\label{supplemental-information-table-of-contents}}
+contents}\label{supplemental-information-table-of-contents}
 
@@
 \item
+  \protect\hyperlink{ecological}{Stability of ecological networks}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{competition}{Competitor networks}
+  \item
+    \protect\hyperlink{mutualism}{Mutualist networks}
+  \item
+    \protect\hyperlink{pred-prey}{Predator-prey networks}
+  \end{itemize}
+\item
+  \protect\hyperlink{connectance}{Sensitivity of connectance (C) values}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{connect3}{C = 0.3}
+  \item
+    \protect\hyperlink{connect5}{C = 0.5}
+  \item
+    \protect\hyperlink{connect7}{C = 0.7}
+  \item
+    \protect\hyperlink{connect9}{C = 0.9}
+  \end{itemize}
+\item
+  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
+  and \(\sigma\)}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
+  \item
+    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
+  \item
+    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
+  \item
+    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
+  \end{itemize}
+\item
+  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
+  \(\gamma\)}
+\item
+  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
+\item
   \protect\hyperlink{ga}{Stability given targeted manipulation of
@@
 \item
-  \protect\hyperlink{ecological}{Stability of ecological networks}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{competition}{Competitor networks}
-  \item
-    \protect\hyperlink{mutualism}{Mutualist networks}
-  \item
-    \protect\hyperlink{pred-prey}{Predator-prey networks}
-  \end{itemize}
-\item
-  \protect\hyperlink{connectance}{Sensitivity of connectance (C) values}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{connect3}{C = 0.3}
-  \item
-    \protect\hyperlink{connect5}{C = 0.5}
-  \item
-    \protect\hyperlink{connect7}{C = 0.7}
-  \item
-    \protect\hyperlink{connect9}{C = 0.9}
-  \end{itemize}
-\item
-  \protect\hyperlink{sigma}{Large networks of \(C = 0.05\) across \(S\)
-  and \(\sigma\)}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
-  \item
-    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
-  \item
-    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
-  \item
-    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
-  \end{itemize}
-\item
-  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
-  \(\gamma\)}
-\item
-  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
-\item
   \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
@@
 
-\hypertarget{IncrS}{%
-\section{\texorpdfstring{Stability across increasing
+\hypertarget{IncrS}{\section{\texorpdfstring{Stability across increasing
 \(S\)}{Stability across increasing S}}\label{IncrS}}
@@
 
-\hypertarget{ga}{%
-\section{\texorpdfstring{Stability given targeted manipulation of
-\(\gamma\) (genetic
-algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
-
-The figure below compares the stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted manipulation of \(\gamma\)
-elements. For each \(S\), 100000 complex systems are randomly generated.
-Stability of each complex system is tested given variation in \(\gamma\)
-using a genetic algorithm to maximise the effect of \(\gamma\) values on
-increasing stability, as compared to stability in an otherwise identical
-system in which \(\gamma\) is the same for all components. Blue bars
-show the number of stable systems in the absence of component response
-rate variation, while red bars show the number of stable systems that
-can be generated if component response rate is varied to maximise system
-stability. The black line shows the proportion of systems that are
-stable when component response rate is targeted to increase stability,
-but would not be stable if \(\sigma^{2}_{\gamma} = 0\). The y-axis shows
-the \(\ln\) number of systems that are stable across
-\(S = \{1, 2, ..., 39, 40\}\) for \(C = 1\), and the proportion of
-systems wherein a targeted search of \(\gamma\) values successfully
-resulted in system stability.
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-6-1.pdf}
-
-Stability results are also shown in the table below. Results for
-\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
-refers to systems in which the genetic algorithm searched for a set of
-\(\gamma\) values that stabilised the system.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
-3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
-4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
-5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
-6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
-7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
-8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
-9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
-10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
-11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
-12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
-13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
-14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
-15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
-16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
-17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
-18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
-19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
-20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
-21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
-22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
-23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
-24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
-25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
-26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
-27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
-28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
-29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
-30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
-31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
-32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
-33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
-34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
-35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
-36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
-37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
-38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
-39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
-40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. This comparison shows the high number of stable
-\(\mathbf{M}\) that can be produced through a targeted search of
-\(\gamma\) values, and suggests that many otherwise unstable systems
-could potentially be stabilised by an informed manipulation of their
-component response times. Such a possibility might conceivably reduce
-the dimensionality of problems involving stability in social-ecological
-or economic systems.
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix
-\(\mathbf{M}\).
-
-\hypertarget{ecological}{%
-\section{Stability of ecological networks}\label{ecological}}
+\hypertarget{ecological}{\section{Stability of ecological
+networks}\label{ecological}}
 
@@
 
-\hypertarget{connectance}{%
-\section{Sensitivity of connectance (C) values}\label{connectance}}
+\hypertarget{connectance}{\section{Sensitivity of connectance (C)
+values}\label{connectance}}
 
@@
 
-\hypertarget{sigma}{%
-\section{\texorpdfstring{Sensitivity of interaction strength
-(\(\sigma\))
+\hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
+strength (\(\sigma\))
 values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
@@
 
-\hypertarget{gam_dist}{%
-\section{\texorpdfstring{Sensitivity of distribution of
+\hypertarget{gam_dist}{\section{\texorpdfstring{Sensitivity of
+distribution of
 \(\gamma\)}{Sensitivity of distribution of \textbackslash{}gamma}}\label{gam_dist}}
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-21-1.pdf}
 
@@
 
-\hypertarget{Feasibility}{%
-\section{Feasibility of complex systems}\label{Feasibility}}
+\hypertarget{Feasibility}{\section{Feasibility of complex
+systems}\label{Feasibility}}
 
@@
 
-\hypertarget{Gibbs}{%
-\section{Consistency with Gibbs et al. (2018)}\label{Gibbs}}
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+The figure below compares the stability of large complex systems given
+\(\boldsymbol{\gamma = 1}\) versus targeted manipulation of \(\gamma\)
+elements. For each \(S\), 100000 complex systems are randomly generated.
+Stability of each complex system is tested given variation in \(\gamma\)
+using a genetic algorithm to maximise the effect of \(\gamma\) values on
+increasing stability, as compared to stability in an otherwise identical
+system in which \(\gamma\) is the same for all components. Blue bars
+show the number of stable systems in the absence of component response
+rate variation, while red bars show the number of stable systems that
+can be generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(\sigma^{2}_{\gamma} = 0\). The y-axis shows
+the \(\ln\) number of systems that are stable across
+\(S = \{1, 2, ..., 39, 40\}\) for \(C = 1\), and the proportion of
+systems wherein a targeted search of \(\gamma\) values successfully
+resulted in system stability.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+
+Stability results are also shown in the table below. Results for
+\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
+refers to systems in which the genetic algorithm searched for a set of
+\(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+A1\_stabilised & A1\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(\mathbf{M}\) that can be produced through a targeted search of
+\(\gamma\) values, and suggests that many otherwise unstable systems
+could potentially be stabilised by an informed manipulation of their
+component response times. Such a possibility might conceivably reduce
+the dimensionality of problems involving stability in social-ecological
+or economic systems.
+
+Distributions of \(\gamma\) values in vectors for the highest values of
+\(S\) are shown below.
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix
+\(\mathbf{M}\).
+
+\hypertarget{Gibbs}{\section{Consistency with Gibbs et al.
+(2018)}\label{Gibbs}}
 
@@
 
-\hypertarget{effects-of-matrix-correlations}{%
 \section{Effects of matrix
-correlations}\label{effects-of-matrix-correlations}}
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
+correlations}\label{effects-of-matrix-correlations}
+
+\section*{References}\label{references}
 \addcontentsline{toc}{section}{References}
@@
 
-\leavevmode\hypertarget{ref-Allesina2015}{}%
+\hypertarget{ref-Allesina2015}{}
 3. Allesina, S. \emph{et al.} Predicting the stability of large
@@
 
-\leavevmode\hypertarget{ref-Servan2018}{}%
+\hypertarget{ref-Servan2018}{}
 4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
@@
 
-\leavevmode\hypertarget{ref-Gibbs2017}{}%
+\hypertarget{ref-Gibbs2017}{}
 5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of

### 2020-03-11 notebook/ms.Rmd
Smooth the text to compare matrices rather than variances in component response rate (which was a bit more difficult to read

@@
 
-**Numerical simulations of random $\mathbf{M}$ with and without $\mathbf{\sigma^{2}_{\gamma}}$**.
+**Numerical simulations of random systems with and without $\mathbf{\sigma^{2}_{\gamma}}$**.
 I used numerical simulations and eigenanalysis to test how variation in $\gamma$ affects stability in random matrices with known properties, comparing the stability of $\textbf{A}$ versus $\mathbf{M} = \gamma\mathbf{A}$.
-Values of $\gamma$ were sampled from a uniform distribution where $\gamma_{i} \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other distributions, which give similar results).
+Values of $\gamma$ were sampled from a uniform distribution where $\gamma_{i} \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other $\gamma$ distributions, which gave similar results).
 Diagonal elements of simulated $\mathbf{A}$ were standardised to the mean value of sampled $\gamma$ so that $E[d]$ of $\textbf{M}$ and $\textbf{M}$ were guaranteed to be identical.
@@
 
-Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma} > 0$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
-Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values when $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ given $S = 25$, $C = 1$, and $\sigma = 0.4$.
+Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma}$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
+Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values for $\mathbf{A}$ ($\sigma^{2}_{\gamma} = 0$) versus $\mathbf{M}$ ($\sigma^{2}_{\gamma} = 1/3$) given $S = 25$, $C = 1$, and $\sigma = 0.4$.
 For $-0.4 \leq \rho \leq 0.7$ (shaded region of Fig.
-2), expected $\Re(\lambda_{max})$ was lower given $\sigma^{2}_{\gamma} = 1/3$.
-For $\rho \geq -0.1$, the range of $\Re(\lambda_{max})$ values observed given $\sigma^{2}_{\gamma} = 1/3$ also had a lower bound, which included negative values for $\rho = -0.1$ and $\rho = 0$.
+2), expected $\Re(\lambda_{max})$ was lower in $\mathbf{M}$ than $\mathbf{A}$.
+For $\rho \geq -0.1$, the lower bound of the range of $\Re(\lambda_{max})$ values also decreased given $\sigma^{2}_{\gamma}$, resulting in negative $\Re(\lambda_{max})$ in $\mathbf{M}$ for $\rho = -0.1$ and $\rho = 0$.
 Hence, across a wide range of system correlations, variation in the response rate of system components had a stabilising effect.
@@
 The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$.
-Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ for $\sigma^{2}_{\gamma} = 0$ versus $\sigma^{2}_{\gamma} = 1/3$ across increasing $S$ given $C = 1$ and $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
-For relatively small systems ($S \leq 25$), positive $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$.
-But as $S$ increased, the curvilinear relationship between $\rho$ and $\Re(\lambda_{max})$ resulted in $\mathbf{M}$ with lower expected $\Re(\lambda_{max})$ for low magnitudes of $\rho$.
-In turn, as $S$ increased, and systems became more complex, positive $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
+Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ across increasing $S$ given $C = 1$ and $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
+For relatively small systems ($S \leq 25$), $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$.
+But as $S$ increased, the curvilinear relationship between $\rho$ and $\Re(\lambda_{max})$ decreased expected $\Re(\lambda_{max})$ for $\mathbf{M}$ given low magnitudes of $\rho$.
+In turn, as $S$ increased, and systems became more complex, $\sigma^{2}_{\gamma}$ increased the proportion of numerical simulations that were observed to be stable (see below).
 
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of varying complexity, I simulated random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
-One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{M}$ was assessed given $\gamma = 1$ ($\sigma^{2}_{\gamma} = 0$) versus $\gamma \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
-For all $S > 10$, I found that the number of stable random systems was higher given $\sigma^{2}_{\gamma} > 0$ than when $\sigma^{2}_{\gamma} = 0$ (Fig.
+To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of varying complexity, I simulated random $\mathbf{M = \gamma A}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
+One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{A}$ vesus $\mathbf{M}$ was assessed given $\gamma_{i} \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
+For all $S > 10$, I found that the number of stable random systems was higher in $\mathbf{M}$ than $\mathbf{A}$ (Fig.
 3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
@@
 
-I also simulated 100000 $\mathbf{M}$ for three types of random networks that are typically interpreted as differing interspecific ecological interaction types [@Allesina2011; @Allesina2012].
-These interactions types are competitive, mutualist, or predator-prey, as modelled by off-diagonal elements of $\mathbf{M}$ that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{M}$).
-As $S$ increased, a higher number of stable systems were observed given $\sigma^{2}_{\gamma}>0$ versus $\sigma^{2}_{\gamma}=0$ for competitor and predator-prey, but not mutualist, systems.
-A higher number of stable systems were observed whenever $S > 12$ and $S > 40$ for competitive and predator-prey systems, respectively (note that $\rho < 0$ by definition for predator-prey systems, making stability more likely overall).
+I also simulated 100000 $\mathbf{M = \gamma A}$ for three types of random networks that are typically interpreted as differing interspecific ecological interaction types [@Allesina2011; @Allesina2012].
+These interactions types are competitive, mutualist, or predator-prey, as modelled by off-diagonal elements that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{A}$).
+As $S$ increased, a higher number of stable $\mathbf{M}$ relative to $\mathbf{A}$ were observed for competitor and predator-prey, but not mutualist, systems.
+A higher number of stable systems were observed whenever $S > 12$ and $S > 40$ for competitive and predator-prey systems, respectively (note that $\rho < 0$ for predator-prey systems, making stability more likely overall).
 The stability of mutualist systems was never affected by $\sigma^{2}_{\gamma}$.
@@
 **Simulation of structured $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate how $\sigma^{2}_{\gamma}$ affects the stability of commonly observed network structures, I simulated 1 million $\textbf{M}$ for small-world [@Watts1998], scale-free [@Albert2002], and cascade food web [@Solow1998; @Williams2000] networks.
+To investigate how $\sigma^{2}_{\gamma}$ affects the stability of commonly observed network structures, I simulated 1 million $\mathbf{M = \gamma A}$ for small-world [@Watts1998], scale-free [@Albert2002], and cascade food web [@Solow1998; @Williams2000] networks.
 In all of these networks, rules determining the presence or absence of an interaction between components $i$ and $j$ constrained the overall structure of the network.
@@
 Network structure did not strongly modulate the effect that $\sigma^{2}_{\gamma}$ had on stability.
-For comparable magnitudes of complexity, structured networks still had a higher number of stable $\mathbf{M}$ for $\sigma^{2}_{\gamma}>0$ compared to $\sigma^{2}_{\gamma}=0$.
-For random networks, positive $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma \sqrt{SC} \gtrapprox 1.26$.
-This threshold of complexity was comparable for small-world networks, and slightly lower for scale-free networks (note that algorithms for generating small-world and scale-free networks necessarily led to varying $C$, see methods).
-Varying $\gamma$ increased stability in cascade food webs for $S > 27$, and therefore at a relatively low complexity magnitudes compared to random predator-prey networks.
-Overall, network structure does not greatly change the effect that $\sigma^{2}_{\gamma}$ has on increasing the upper bound of complexity within which stability might reasonably be observed.
+For comparable magnitudes of complexity, structured networks still had a higher number of stable $\mathbf{M}$ than $\mathbf{A}$.
+For random networks, $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma \sqrt{SC} \gtrapprox 1.26$.
+This threshold of complexity, above which more $\mathbf{M}$ than $\mathbf{A}$ were stable, was comparable for small-world networks, and slightly lower for scale-free networks (note that algorithms for generating small-world and scale-free networks necessarily led to varying $C$; see methods).
+Varying $\gamma$ increased stability in cascade food webs for $S > 27$, and therefore at a relatively low complexity magnitudes compared to random predator-prey networks ($S > 40$).
+Overall, network structure did not greatly change the effect that $\sigma^{2}_{\gamma}$ had on increasing the upper bound of complexity within which stability might reasonably be observed.
 
@@
 This is of particular interest for ecological communities because population density ($N$) cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
-While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices $\mathbf{M}$ tested for stability, and additionally for specific types of ecological communities [@Allesina2012] (e.g., competitive, mutualist, predator-prey; see Supplementary Information).
+While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices tested for stability.
 I emphasise that $\gamma$ is not interpreted as population density in this analysis, but instead as a fundamental property of species life history such as expected generation time.

### 2020-03-11 notebook/ms.tex
Smooth the text to compare matrices rather than variances in component response rate (which was a bit more difficult to read

@@
 
-\textbf{Numerical simulations of random \(\mathbf{M}\) with and without
+\textbf{Numerical simulations of random systems with and without
 \(\mathbf{\sigma^{2}_{\gamma}}\)}. I used numerical simulations and
@@
 \(\gamma_{i} \sim \mathcal{U}(0, 2)\) and \(\sigma^{2}_{\gamma} = 1/3\)
-(see Supplementary Information for other distributions, which give
-similar results). Diagonal elements of simulated \(\mathbf{A}\) were
-standardised to the mean value of sampled \(\gamma\) so that \(E[d]\) of
-\(\textbf{M}\) and \(\textbf{M}\) were guaranteed to be identical. Here
-I focus on the effect of \(\gamma\) across values of \(\rho\), and for
-increasing system sizes (\(S\)) in random and structured networks. By
-increasing \(S\), the objective is to determine the effect of \(\gamma\)
-as system complexity increases toward the boundary at which stability is
-realistic for a finite system.
+(see Supplementary Information for other \(\gamma\) distributions, which
+gave similar results). Diagonal elements of simulated \(\mathbf{A}\)
+were standardised to the mean value of sampled \(\gamma\) so that
+\(E[d]\) of \(\textbf{M}\) and \(\textbf{M}\) were guaranteed to be
+identical. Here I focus on the effect of \(\gamma\) across values of
+\(\rho\), and for increasing system sizes (\(S\)) in random and
+structured networks. By increasing \(S\), the objective is to determine
+the effect of \(\gamma\) as system complexity increases toward the
+boundary at which stability is realistic for a finite system.
 
@@
 Ten thousand numerical simulations of \(\mathbf{M}\) revealed that
-\(\sigma^{2}_{\gamma} > 0\) results in a nonlinear relationship between
+\(\sigma^{2}_{\gamma}\) results in a nonlinear relationship between
 \(\rho\) and \(\Re(\lambda_{max})\), which can sometimes increase the
 stability of the system. Figure 2 shows a comparison of
-\(\Re(\lambda_{max})\) across \(\rho\) values when
-\(\sigma^{2}_{\gamma} = 0\) versus \(\sigma^{2}_{\gamma} = 1/3\) given
-\(S = 25\), \(C = 1\), and \(\sigma = 0.4\). For
-\(-0.4 \leq \rho \leq 0.7\) (shaded region of Fig. 2), expected
-\(\Re(\lambda_{max})\) was lower given \(\sigma^{2}_{\gamma} = 1/3\).
-For \(\rho \geq -0.1\), the range of \(\Re(\lambda_{max})\) values
-observed given \(\sigma^{2}_{\gamma} = 1/3\) also had a lower bound,
-which included negative values for \(\rho = -0.1\) and \(\rho = 0\).
-Hence, across a wide range of system correlations, variation in the
-response rate of system components had a stabilising effect.
+\(\Re(\lambda_{max})\) across \(\rho\) values for \(\mathbf{A}\)
+(\(\sigma^{2}_{\gamma} = 0\)) versus \(\mathbf{M}\)
+(\(\sigma^{2}_{\gamma} = 1/3\)) given \(S = 25\), \(C = 1\), and
+\(\sigma = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded region of Fig.
+2), expected \(\Re(\lambda_{max})\) was lower in \(\mathbf{M}\) than
+\(\mathbf{A}\). For \(\rho \geq -0.1\), the lower bound of the range of
+\(\Re(\lambda_{max})\) values also decreased given
+\(\sigma^{2}_{\gamma}\), resulting in negative \(\Re(\lambda_{max})\) in
+\(\mathbf{M}\) for \(\rho = -0.1\) and \(\rho = 0\). Hence, across a
+wide range of system correlations, variation in the response rate of
+system components had a stabilising effect.
 
@@
 increased with increasing \(S\). Figure 3 shows 10000 numerical
-simulations of \(\mathbf{M}\) for \(\sigma^{2}_{\gamma} = 0\) versus
-\(\sigma^{2}_{\gamma} = 1/3\) across increasing \(S\) given \(C = 1\)
+simulations of \(\mathbf{M}\) across increasing \(S\) given \(C = 1\)
 and \(\sigma = 0.2\) (\(\sigma\) has been lowered here to better
@@
 \(1 = \sigma\sqrt{SC}\)). For relatively small systems (\(S \leq 25\)),
-positive \(\sigma^{2}_{\gamma}\) never decreased the expected
+\(\sigma^{2}_{\gamma}\) never decreased the expected
 \(\Re(\lambda_{max})\). But as \(S\) increased, the curvilinear
-relationship between \(\rho\) and \(\Re(\lambda_{max})\) resulted in
-\(\mathbf{M}\) with lower expected \(\Re(\lambda_{max})\) for low
-magnitudes of \(\rho\). In turn, as \(S\) increased, and systems became
-more complex, positive \(\sigma^{2}_{\gamma}\) increased the proportion
-of numerical simulations that were observed to be stable (see below).
+relationship between \(\rho\) and \(\Re(\lambda_{max})\) decreased
+expected \(\Re(\lambda_{max})\) for \(\mathbf{M}\) given low magnitudes
+of \(\rho\). In turn, as \(S\) increased, and systems became more
+complex, \(\sigma^{2}_{\gamma}\) increased the proportion of numerical
+simulations that were observed to be stable (see below).
 
@@
 investigate the effect of \(\sigma^{2}_{\gamma}\) on stability across
-systems of varying complexity, I simulated random \(\mathbf{M}\)
-matrices at \(\sigma = 0.4\) and \(C = 1\) across
-\(S = \{2, 3, ..., 49, 50\}\). One million \(\mathbf{M}\) were simulated
-for each \(S\), and the stability of \(\mathbf{M}\) was assessed given
-\(\gamma = 1\) (\(\sigma^{2}_{\gamma} = 0\)) versus
-\(\gamma \sim \mathcal{U}(0, 2)\) (\(\sigma^{2}_{\gamma} = 1/3\)). For
-all \(S > 10\), I found that the number of stable random systems was
-higher given \(\sigma^{2}_{\gamma} > 0\) than when
-\(\sigma^{2}_{\gamma} = 0\) (Fig. 3; see Supplementary Information for
-full table of results), and that the difference between the
-probabilities of observing a stable system increased with an increase in
-\(S\). In other words, the potential for \(\sigma^{2}_{\gamma}\) to
-affect stability increased with system complexity and was most relevant
-for systems on the cusp of being too complex to be realistically stable.
-For the highest values of \(S\), nearly all systems that were stable
-given varying \(\gamma\) would not have been stable given
-\(\gamma = 1\).
+systems of varying complexity, I simulated random
+\(\mathbf{M = \gamma A}\) matrices at \(\sigma = 0.4\) and \(C = 1\)
+across \(S = \{2, 3, ..., 49, 50\}\). One million \(\mathbf{M}\) were
+simulated for each \(S\), and the stability of \(\mathbf{A}\) vesus
+\(\mathbf{M}\) was assessed given \(\gamma_{i} \sim \mathcal{U}(0, 2)\)
+(\(\sigma^{2}_{\gamma} = 1/3\)). For all \(S > 10\), I found that the
+number of stable random systems was higher in \(\mathbf{M}\) than
+\(\mathbf{A}\) (Fig. 3; see Supplementary Information for full table of
+results), and that the difference between the probabilities of observing
+a stable system increased with an increase in \(S\). In other words, the
+potential for \(\sigma^{2}_{\gamma}\) to affect stability increased with
+system complexity and was most relevant for systems on the cusp of being
+too complex to be realistically stable. For the highest values of \(S\),
+nearly all systems that were stable given varying \(\gamma\) would not
+have been stable given \(\gamma = 1\).
 
-I also simulated 100000 \(\mathbf{M}\) for three types of random
-networks that are typically interpreted as differing interspecific
-ecological interaction
+I also simulated 100000 \(\mathbf{M = \gamma A}\) for three types of
+random networks that are typically interpreted as differing
+interspecific ecological interaction
 types\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{25}}.
 These interactions types are competitive, mutualist, or predator-prey,
-as modelled by off-diagonal elements of \(\mathbf{M}\) that are
-constrained to be negative, positive, or paired such that if
-\(A_{ij} > 0\) then \(A_{ji} < 0\),
+as modelled by off-diagonal elements that are constrained to be
+negative, positive, or paired such that if \(A_{ij} > 0\) then
+\(A_{ji} < 0\),
 respectively\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-(but are otherwise identical to the purely random \(\mathbf{M}\)). As
-\(S\) increased, a higher number of stable systems were observed given
-\(\sigma^{2}_{\gamma}>0\) versus \(\sigma^{2}_{\gamma}=0\) for
-competitor and predator-prey, but not mutualist, systems. A higher
-number of stable systems were observed whenever \(S > 12\) and
-\(S > 40\) for competitive and predator-prey systems, respectively (note
-that \(\rho < 0\) by definition for predator-prey systems, making
-stability more likely overall). The stability of mutualist systems was
-never affected by \(\sigma^{2}_{\gamma}\).
+(but are otherwise identical to the purely random \(\mathbf{A}\)). As
+\(S\) increased, a higher number of stable \(\mathbf{M}\) relative to
+\(\mathbf{A}\) were observed for competitor and predator-prey, but not
+mutualist, systems. A higher number of stable systems were observed
+whenever \(S > 12\) and \(S > 40\) for competitive and predator-prey
+systems, respectively (note that \(\rho < 0\) for predator-prey systems,
+making stability more likely overall). The stability of mutualist
+systems was never affected by \(\sigma^{2}_{\gamma}\).
 
@@
 commonly observed network structures, I simulated 1 million
-\(\textbf{M}\) for
+\(\mathbf{M = \gamma A}\) for
 small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}},
@@
 complexity, structured networks still had a higher number of stable
-\(\mathbf{M}\) for \(\sigma^{2}_{\gamma}>0\) compared to
-\(\sigma^{2}_{\gamma}=0\). For random networks, positive
+\(\mathbf{M}\) than \(\mathbf{A}\). For random networks,
 \(\sigma^{2}_{\gamma}\) increased stability given \(S > 10\)
 (\(\sigma = 0.4\) and \(C = 1\)), and therefore complexity
-\(\sigma \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity was
+\(\sigma \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity,
+above which more \(\mathbf{M}\) than \(\mathbf{A}\) were stable, was
 comparable for small-world networks, and slightly lower for scale-free
 networks (note that algorithms for generating small-world and scale-free
-networks necessarily led to varying \(C\), see methods). Varying
+networks necessarily led to varying \(C\); see methods). Varying
 \(\gamma\) increased stability in cascade food webs for \(S > 27\), and
 therefore at a relatively low complexity magnitudes compared to random
-predator-prey networks. Overall, network structure does not greatly
-change the effect that \(\sigma^{2}_{\gamma}\) has on increasing the
-upper bound of complexity within which stability might reasonably be
+predator-prey networks (\(S > 40\)). Overall, network structure did not
+greatly change the effect that \(\sigma^{2}_{\gamma}\) had on increasing
+the upper bound of complexity within which stability might reasonably be
 observed.
@@
 not restricted to species networks, I have also performed a feasibility
-analysis on all matrices \(\mathbf{M}\) tested for stability, and
-additionally for specific types of ecological
-communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-(e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). I emphasise that \(\gamma\) is not interpreted as
-population density in this analysis, but instead as a fundamental
-property of species life history such as expected generation time.
-Feasibility was unaffected by \(\sigma^{2}_{\gamma}\) and instead
-occurred with a fixed probability of \(1/2^{S}\), consistent with a
-recent proof by Serván et
+analysis on all matrices tested for stability. I emphasise that
+\(\gamma\) is not interpreted as population density in this analysis,
+but instead as a fundamental property of species life history such as
+expected generation time. Feasibility was unaffected by
+\(\sigma^{2}_{\gamma}\) and instead occurred with a fixed probability of
+\(1/2^{S}\), consistent with a recent proof by Serván et
 al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}} (see
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.12.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.18.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-13 notebook/ms.Rmd
Edit the introduction to smooth the writing and fix the notation again, and add a bit to Figure 1 legend

@@
 - \bibliographystyle{amnatnat}
-linestretch: 2
+linestretch: 1
 link-citations: yes
@@
 The stability of a complex system generally decreases with increasing system size and interconnectivity, a counterintuitive result of widespread importance across the physical, life, and social sciences.
-Despite recent interest in the relationship between system properties and stability, the effect of variation in the response rate of individual system components remains unconsidered.
+Despite recent interest in the relationship between system properties and stability, the effect of variation in response rate across system components remains unconsidered.
 Here I vary the component response rates ($\boldsymbol{\gamma}$) of randomly generated complex systems.
-I use numerical simulations to show that when component response rates vary, the potential for system stability is markedly increased.
+I use numerical simulations to show that when component response rates vary, the potential for system stability increases.
 These results are robust to common network structures, including small-world and scale-free networks, and cascade food webs.
 Variation in $\boldsymbol{\gamma}$ is especially important for stability in highly complex systems, in which the probability of stability would otherwise be negligible.
-At such extremes of simulated system complexity, the largest stable complex systems would be unstable if not for $\boldsymbol{Var(\gamma)}$.
+At such extremes of simulated system complexity, the largest stable complex systems would be unstable if not for variation in $\boldsymbol{\gamma}$.
 My results therefore reveal a previously unconsidered aspect of system stability that is likely to be pervasive across all realistic complex systems.
@@
 One element of such a matrix, $M_{ij}$, defines how component $j$ affects component $i$ in the system at a point of equilibrium [@Allesina2012].
-Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation ($d$, e.g., carrying capacity in ecological communities).
-Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma_{M}^{2}$; diagonal elements are set to $d = -1$ [@May1972; @Allesina2012; @Allesina2015].
-Local system stability is assessed using eigenanalysis, with the system being stable if the real parts of all eigenvalues ($\lambda$) of $\mathbf{M}$ are negative ($\Re(\lambda_{max}) < 0$) [@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -1$ (the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma_{M}\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
-Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma_{M}$ increase.
+Off-diagonal elements ($i \neq j$) therefore define interactions between components, while diagonal elements ($i = j$) define component self-regulation (e.g., carrying capacity in ecological communities).
+Traditionally, off-diagonal elements are assigned non-zero values with a probability $C$, which are sampled from a distribution with variance $\sigma^{2}$; diagonal elements are set to $-1$ [@May1972; @Allesina2012; @Allesina2015].
+Local system stability is assessed using eigenanalysis on $\mathbf{M}$, with the system being stable if the real parts of all eigenvalues ($\lambda$), and therefore the leading eigenvalue ($\lambda_{max}$), are negative ($\Re(\lambda_{max}) < 0$) [@May1972; @Allesina2012].
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -d$ ($-d$ is the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma$ increase.
 
-May's [@May1972; @Allesina2012] stability criterion $\sigma_{M}\sqrt{SC} < 1$ assumes that the expected response rates ($\gamma$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
+May's [@May1972; @Allesina2012] stability criterion $\sigma\sqrt{SC} < d$ assumes that the expected response rates ($\gamma$) of individual components to perturbations of the system are identical, but this is highly unlikely in any complex system.
 In ecological communities, for example, the rate at which population density changes following perturbation will depend on the generation time of organisms, which might vary by orders of magnitude among species.
@@
 The effect of such variance on stability has not been investigated in complex systems theory.
-Intuitively, variation in $\gamma$ ($\sigma^{2}_{\gamma}$) might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma_{M}$.
-Here I show that, despite higher $\sigma_{M}$, realistic complex systems (such that $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
+Intuitively, variation in $\gamma$ ($\sigma^{2}_{\gamma}$) might be expected to decrease system stability by introducing a new source of variation into the system and thereby increasing $\sigma$.
+Here I show that, despite higher $\sigma$, realistic complex systems (in which $S$ is high but finite) are actually more likely to be stable if their individual component response rates vary.
 My results are robust across commonly observed network structures, including random [@May1972], small-world [@Watts1998], scale-free [@Albert2002], cascade food web [@Solow1998; @Williams2000] networks.
@@
 Points are uniformly distributed within a circle centred at the origin with a radius of $\sigma_{A} \sqrt{S} = 1$.
-($\textbf{b}$) The same system after including variation in the response rates of $S$ components, represented by the diagonal matrix $\gamma$, such that $\mathbf{M} = \gamma\mathbf{A}$.
+($\textbf{b}$) The same system as $\textbf{a}$ after including variation in the response rates of $S$ components, represented by the diagonal matrix $\gamma$, such that $\mathbf{M} = \gamma\mathbf{A}$.
 Elements of $\gamma$ are randomly sampled from a uniform distribution from $\min = 0$ to $\max = 2$.

### 2020-03-13 notebook/ms.tex
Edit the introduction to smooth the writing and fix the notation again, and add a bit to Figure 1 legend

@@
 Despite recent interest in the relationship between system properties
-and stability, the effect of variation in the response rate of
-individual system components remains unconsidered. Here I vary the
-component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
-complex systems. I use numerical simulations to show that when component
-response rates vary, the potential for system stability is markedly
-increased. These results are robust to common network structures,
-including small-world and scale-free networks, and cascade food webs.
-Variation in \(\boldsymbol{\gamma}\) is especially important for
-stability in highly complex systems, in which the probability of
-stability would otherwise be negligible. At such extremes of simulated
-system complexity, the largest stable complex systems would be unstable
-if not for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
-previously unconsidered aspect of system stability that is likely to be
-pervasive across all realistic complex systems.
+and stability, the effect of variation in response rate across system
+components remains unconsidered. Here I vary the component response
+rates (\(\boldsymbol{\gamma}\)) of randomly generated complex systems. I
+use numerical simulations to show that when component response rates
+vary, the potential for system stability increases. These results are
+robust to common network structures, including small-world and
+scale-free networks, and cascade food webs. Variation in
+\(\boldsymbol{\gamma}\) is especially important for stability in highly
+complex systems, in which the probability of stability would otherwise
+be negligible. At such extremes of simulated system complexity, the
+largest stable complex systems would be unstable if not for variation in
+\(\boldsymbol{\gamma}\). My results therefore reveal a previously
+unconsidered aspect of system stability that is likely to be pervasive
+across all realistic complex systems.
 
@@
 between components, while diagonal elements (\(i = j\)) define component
-self-regulation (\(d\), e.g., carrying capacity in ecological
-communities). Traditionally, off-diagonal elements are assigned non-zero
-values with a probability \(C\), which are sampled from a distribution
-with variance \(\sigma_{M}^{2}\); diagonal elements are set to
-\(d = -1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
-Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of
-\(\mathbf{M}\) are negative
+self-regulation (e.g., carrying capacity in ecological communities).
+Traditionally, off-diagonal elements are assigned non-zero values with a
+probability \(C\), which are sampled from a distribution with variance
+\(\sigma^{2}\); diagonal elements are set to
+\(-1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
+Local system stability is assessed using eigenanalysis on
+\(\mathbf{M}\), with the system being stable if the real parts of all
+eigenvalues (\(\lambda\)), and therefore the leading eigenvalue
+(\(\lambda_{max}\)), are negative
 (\(\Re(\lambda_{max}) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
@@
 uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
-circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
-\(\Im = 0\), with a radius of
-\(\sigma_{M}\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
+circle centred at \(\Re = -d\) (\(-d\) is the mean value of diagonal
+elements) and \(\Im = 0\), with a radius of
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma_{M}\) increase.
+therefore becomes increasingly unlikely as \(S\), \(C\), and \(\sigma\)
+increase.
 
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma_{M}\sqrt{SC} < 1\) assumes that the
-expected response rates (\(\gamma\)) of individual components to
-perturbations of the system are identical, but this is highly unlikely
-in any complex system. In ecological communities, for example, the rate
-at which population density changes following perturbation will depend
-on the generation time of organisms, which might vary by orders of
-magnitude among species. Species with short generation times will
-respond quickly (high \(\gamma\)) to perturbations relative to species
-with long generation times (low \(\gamma\)). Similarly, the speed at
-which individual banks respond to perturbations in financial networks,
-or individuals or institutions respond to perturbations in complex
-social networks, is likely to vary. The effect of such variance on
-stability has not been investigated in complex systems theory.
-Intuitively, variation in \(\gamma\) (\(\sigma^{2}_{\gamma}\)) might be
-expected to decrease system stability by introducing a new source of
-variation into the system and thereby increasing \(\sigma_{M}\). Here I
-show that, despite higher \(\sigma_{M}\), realistic complex systems
-(such that \(S\) is high but finite) are actually more likely to be
-stable if their individual component response rates vary. My results are
-robust across commonly observed network structures, including
+stability criterion \(\sigma\sqrt{SC} < d\) assumes that the expected
+response rates (\(\gamma\)) of individual components to perturbations of
+the system are identical, but this is highly unlikely in any complex
+system. In ecological communities, for example, the rate at which
+population density changes following perturbation will depend on the
+generation time of organisms, which might vary by orders of magnitude
+among species. Species with short generation times will respond quickly
+(high \(\gamma\)) to perturbations relative to species with long
+generation times (low \(\gamma\)). Similarly, the speed at which
+individual banks respond to perturbations in financial networks, or
+individuals or institutions respond to perturbations in complex social
+networks, is likely to vary. The effect of such variance on stability
+has not been investigated in complex systems theory. Intuitively,
+variation in \(\gamma\) (\(\sigma^{2}_{\gamma}\)) might be expected to
+decrease system stability by introducing a new source of variation into
+the system and thereby increasing \(\sigma\). Here I show that, despite
+higher \(\sigma\), realistic complex systems (in which \(S\) is high but
+finite) are actually more likely to be stable if their individual
+component response rates vary. My results are robust across commonly
+observed network structures, including
 random\textsuperscript{\protect\hyperlink{ref-May1972}{1}},
@@
 distributed within a circle centred at the origin with a radius of
-\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system after
-including variation in the response rates of \(S\) components,
-represented by the diagonal matrix \(\gamma\), such that
+\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
+\(\textbf{a}\) after including variation in the response rates of \(S\)
+components, represented by the diagonal matrix \(\gamma\), such that
 \(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
@@
 a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.18.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.21.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-13 notebook/ms.Rmd
Minor edits to improve the readability of the results up to reporting the simulation results

@@
 **Component response rates of random complex systems**.
-Complex systems $\mathbf{M}$ are built from two matrices, one modelling component interactions ($\mathbf{A}$), and second modelling component response rates ($\gamma$).
+Complex systems ($\mathbf{M}$) are built from two matrices, one modelling component interactions ($\mathbf{A}$), and second modelling component response rates ($\gamma$).
 Both $\mathbf{A}$ and $\gamma$ are square $S \times S$ matrices.
 Rows in $\mathbf{A}$ define how a given component $i$ is affected by each component $j$ in the system, including itself (where $i = j$).
-Off-diagonal elments of $\mathbf{A}$ are independent and identically distributed (i.i.d), and diagonal elements are set to $A_{ii} = -1$ as in May [@May1972].
+Off-diagonal elements of $\mathbf{A}$ are independent and identically distributed (i.i.d), and diagonal elements are set to $A_{ii} = -1$ as in May [@May1972].
 Diagonal elements of $\gamma$ are positive, and off-diagonal elements are set to zero (i.e.
 $\gamma$ is a diagonal matrix with positive support).
-The distribution of $\gamma$ over $S$ components thereby models the distribution of component response rates.
+The distribution of $diag(\gamma)$ over $S$ components thereby models the distribution of component response rates.
 The dynamics of the entire system $\mathbf{M}$ can be defined as follows [@Patel2018],
@@
 
-Equation \ref{defM} thereby serves as a null model to investigate how variation in component response rate ($\sigma_{\gamma}$) affects complex systems.
-In the absence of such variation ($\sigma_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
-Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma_{M} \sqrt{SC}$ [@May1972] (Fig.
+Equation \ref{defM} thereby serves as a null model to investigate how variation in component response rate ($\sigma^{2}_{\gamma}$) affects complex systems.
+In the absence of such variation ($\sigma^{2}_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
+Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma \sqrt{SC}$ [@May1972] (Fig.
 1a).
@@
 The value of $\Re(\lambda_{max})$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
-These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the standard deviation between component interaction strengths ($\sigma_{M}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
-Positive $\sigma_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$.
-What $\sigma_{\gamma}$ does change is the total variation in component interaction strengths ($\sigma_{M}$), and $\rho$.
+These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the variance of between component interaction strengths (hereafter $\sigma^{2}_{M}$, to distinguish from $\sigma^{2}_{A}$ and $\sigma^{2}_{\gamma}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+Positive $\sigma^{2}_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$.
+What $\sigma^{2}_{\gamma}$ does change is the total variation in component interaction strengths ($\sigma^{2}_{M}$), and $\rho$.
 Introducing variation in $\gamma$ increases the total variation in the system.
-Variation of the off-diagonal elements in $\mathbf{M}$ is described by the joint variation of two random variables,
+Variation in the off-diagonal elements of $\mathbf{M}$ is described by the joint variation of two random variables,
 
@@
 
-Given $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, equation \ref{var_full} can be simplified,
+Given $E[\gamma_{i}] = 1$ and $E[A_{ij}] = 0$, Eq. \ref{var_full} can be simplified,
 
@@
 
-The increase caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Fig.
+The increase in $\sigma^{2}_{M}$ caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Fig.
 1).
@@
 
-Given $\sigma^{2}_\gamma = 0$, $\Re(\lambda_{max})$ decrease linearly with $\rho$ such that [@Tang2014c],
+Given $\sigma^{2}_\gamma = 0$, $\Re(\lambda_{max})$ decreases linearly with $\rho$ such that [@Tang2014c],
 
@@
 If $\rho < 0$, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability increases [@Allesina2012].
-If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$.
+If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$ (Figure 2).
 
@@
 Values of $\gamma$ were sampled from a uniform distribution where $\gamma_{i} \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other $\gamma$ distributions, which gave similar results).
-Diagonal elements of simulated $\mathbf{A}$ were standardised to the mean value of sampled $\gamma$ so that $E[d]$ of $\textbf{M}$ and $\textbf{M}$ were guaranteed to be identical.
-Here I focus on the effect of $\gamma$ across values of $\rho$, and for increasing system sizes ($S$) in random and structured networks.
+In all simulations, diagonal elements were standardised to ensure that $-d$ between individual $\textbf{A}$ and $\textbf{M}$ pairs were identical (also note that $E[\gamma_{i}] = 1$).
+First I focus on the effect of $\gamma$ across values of $\rho$, then for increasing system sizes ($S$) in random and structured networks.
 By increasing $S$, the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system.

### 2020-03-13 notebook/ms.tex
Minor edits to improve the readability of the results up to reporting the simulation results

@@
 \textbf{Component response rates of random complex systems}. Complex
-systems \(\mathbf{M}\) are built from two matrices, one modelling
+systems (\(\mathbf{M}\)) are built from two matrices, one modelling
 component interactions (\(\mathbf{A}\)), and second modelling component
@@
 given component \(i\) is affected by each component \(j\) in the system,
-including itself (where \(i = j\)). Off-diagonal elments of
+including itself (where \(i = j\)). Off-diagonal elements of
 \(\mathbf{A}\) are independent and identically distributed (i.i.d), and
@@
 to zero (i.e. \(\gamma\) is a diagonal matrix with positive support).
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. The dynamics of the entire
-system \(\mathbf{M}\) can be defined as
+The distribution of \(diag(\gamma)\) over \(S\) components thereby
+models the distribution of component response rates. The dynamics of the
+entire system \(\mathbf{M}\) can be defined as
 follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}},
@@
 Equation \ref{defM} thereby serves as a null model to investigate how
-variation in component response rate (\(\sigma_{\gamma}\)) affects
+variation in component response rate (\(\sigma^{2}_{\gamma}\)) affects
 complex systems. In the absence of such variation
-(\(\sigma_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
+(\(\sigma^{2}_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
 (diagonal elements all equal 1) and \(\mathbf{M} = \mathbf{A}\). Under
@@
 circle centred at \((-d, 0)\) with a radius of
-\(\sigma_{M} \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+\(\sigma \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
 (Fig. 1a).
@@
 self-regulation of components (\(d\)), (3) mean interaction strength
-between components (\(\mu\)), (4) the standard deviation between
-component interaction strengths (\(\sigma_{M}\)), and (5) the
-correlation of interaction strengths between components, \(M_{ij}\) and
-\(M_{ji}\) (\(\rho\)). Positive \(\sigma_{\gamma}\) does not change
-\(S\), nor does it necessarily change \(E[d]\) or \(E[\mu]\). What
-\(\sigma_{\gamma}\) does change is the total variation in component
-interaction strengths (\(\sigma_{M}\)), and \(\rho\). Introducing
+between components (\(\mu\)), (4) the variance of between component
+interaction strengths (hereafter \(\sigma^{2}_{M}\), to distinguish from
+\(\sigma^{2}_{A}\) and \(\sigma^{2}_{\gamma}\)), and (5) the correlation
+of interaction strengths between components, \(M_{ij}\) and \(M_{ji}\)
+(\(\rho\)). Positive \(\sigma^{2}_{\gamma}\) does not change \(S\), nor
+does it necessarily change \(E[d]\) or \(E[\mu]\). What
+\(\sigma^{2}_{\gamma}\) does change is the total variation in component
+interaction strengths (\(\sigma^{2}_{M}\)), and \(\rho\). Introducing
 variation in \(\gamma\) increases the total variation in the system.
-Variation of the off-diagonal elements in \(\mathbf{M}\) is described by
+Variation in the off-diagonal elements of \(\mathbf{M}\) is described by
 the joint variation of two random variables,
@@
 
-Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), equation
-\ref{var_full} can be simplified,
+Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), Eq. \ref{var_full}
+can be simplified,
 
@@
 
-The increase caused by \(\sigma^{2}_\gamma\) can be visualised from the
-eigenvalue spectra of \(\textbf{A}\) versus
+The increase in \(\sigma^{2}_{M}\) caused by \(\sigma^{2}_\gamma\) can
+be visualised from the eigenvalue spectra of \(\textbf{A}\) versus
 \(\textbf{M} = \gamma\textbf{A}\) (Fig. 1). Given \(d = 0\) and
@@
 
-Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) decrease
+Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) decreases
 linearly with \(\rho\) such
@@
 variance of \(M_{ij}\) without affecting the expected covariance between
-\(M_{ij}\) and \(M_{ji}\).
+\(M_{ij}\) and \(M_{ji}\) (Figure 2).
 
@@
 (see Supplementary Information for other \(\gamma\) distributions, which
-gave similar results). Diagonal elements of simulated \(\mathbf{A}\)
-were standardised to the mean value of sampled \(\gamma\) so that
-\(E[d]\) of \(\textbf{M}\) and \(\textbf{M}\) were guaranteed to be
-identical. Here I focus on the effect of \(\gamma\) across values of
-\(\rho\), and for increasing system sizes (\(S\)) in random and
-structured networks. By increasing \(S\), the objective is to determine
-the effect of \(\gamma\) as system complexity increases toward the
-boundary at which stability is realistic for a finite system.
+gave similar results). In all simulations, diagonal elements were
+standardised to ensure that \(-d\) between individual \(\textbf{A}\) and
+\(\textbf{M}\) pairs were identical (also note that
+\(E[\gamma_{i}] = 1\)). First I focus on the effect of \(\gamma\) across
+values of \(\rho\), then for increasing system sizes (\(S\)) in random
+and structured networks. By increasing \(S\), the objective is to
+determine the effect of \(\gamma\) as system complexity increases toward
+the boundary at which stability is realistic for a finite system.
 

### 2020-03-13 notebook/ms.Rmd
Edits to the results and Discussion, improving readability and soothing the notation

@@
 
-Ten thousand numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma}$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
-Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values for $\mathbf{A}$ ($\sigma^{2}_{\gamma} = 0$) versus $\mathbf{M}$ ($\sigma^{2}_{\gamma} = 1/3$) given $S = 25$, $C = 1$, and $\sigma = 0.4$.
+Numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma}$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
+Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values for $\mathbf{A}$ ($\sigma^{2}_{\gamma} = 0$) versus $\mathbf{M}$ ($\sigma^{2}_{\gamma} = 1/3$) given $S = 25$, $C = 1$, and $\sigma_{A} = 0.4$.
 For $-0.4 \leq \rho \leq 0.7$ (shaded region of Fig.
@@
 The stabilising effect of $\sigma^{2}_{\gamma}$ across $\rho$ increased with increasing $S$.
-Figure 3 shows 10000 numerical simulations of $\mathbf{M}$ across increasing $S$ given $C = 1$ and $\sigma = 0.2$ ($\sigma$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma\sqrt{SC}$).
+Figure 3 shows numerical simulations of $\mathbf{M}$ across increasing $S$ given $C = 1$ and $\sigma_{A} = 0.2$ ($\sigma_{A}$ has been lowered here to better illustrate the effect of $S$; note that now given $S = 25$, $1 = \sigma_{A}\sqrt{SC}$).
 For relatively small systems ($S \leq 25$), $\sigma^{2}_{\gamma}$ never decreased the expected $\Re(\lambda_{max})$.
@@
 **Simulation of random $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of varying complexity, I simulated random $\mathbf{M = \gamma A}$ matrices at $\sigma = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
+To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of increasing complexity, I simulated random $\mathbf{M = \gamma A}$ matrices at $\sigma_{A} = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
 One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{A}$ vesus $\mathbf{M}$ was assessed given $\gamma_{i} \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
 For all $S > 10$, I found that the number of stable random systems was higher in $\mathbf{M}$ than $\mathbf{A}$ (Fig.
-3; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
+4; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
 In other words, the potential for $\sigma^{2}_{\gamma}$ to affect stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
@@
 
-I also simulated 100000 $\mathbf{M = \gamma A}$ for three types of random networks that are typically interpreted as differing interspecific ecological interaction types [@Allesina2011; @Allesina2012].
-These interactions types are competitive, mutualist, or predator-prey, as modelled by off-diagonal elements that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{A}$).
-As $S$ increased, a higher number of stable $\mathbf{M}$ relative to $\mathbf{A}$ were observed for competitor and predator-prey, but not mutualist, systems.
-A higher number of stable systems were observed whenever $S > 12$ and $S > 40$ for competitive and predator-prey systems, respectively (note that $\rho < 0$ for predator-prey systems, making stability more likely overall).
+I also simulated 100000 $\mathbf{M = \gamma A}$ for three types of random networks that are typically interpreted as modelling three types of interspecific ecological interactions [@Allesina2011; @Allesina2012].
+These interaction types are competitive, mutualist, and predator-prey, as modelled by off-diagonal elements that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{A}$).
+As $S$ increased, a higher number of stable $\mathbf{M}$ relative to $\mathbf{A}$ was observed for competitor and predator-prey, but not mutualist, systems.
+A higher number of stable systems was observed whenever $S > 12$ and $S > 40$ for competitive and predator-prey systems, respectively (note that $\rho < 0$ for predator-prey systems, making stability more likely overall).
 The stability of mutualist systems was never affected by $\sigma^{2}_{\gamma}$.
 
-The effect of $\sigma^{2}_{\gamma}$ on stability did not change qualitatively across values of $C$, $\sigma$, or different distributions of $\gamma$ (see Supporting Information).
+The effect of $\sigma^{2}_{\gamma}$ on stability did not change qualitatively across values of $C$, $\sigma_{A}$, or for different distributions of $\gamma$ (see Supporting Information).
 
 **Simulation of structured $\mathbf{M}$ across $\mathbf{S}$**.
-To investigate how $\sigma^{2}_{\gamma}$ affects the stability of commonly observed network structures, I simulated 1 million $\mathbf{M = \gamma A}$ for small-world [@Watts1998], scale-free [@Albert2002], and cascade food web [@Solow1998; @Williams2000] networks.
-In all of these networks, rules determining the presence or absence of an interaction between components $i$ and $j$ constrained the overall structure of the network.
+To investigate how $\sigma^{2}_{\gamma}$ affects the stability of commonly observed network structures, I simulated one million $\mathbf{M = \gamma A}$ for small-world [@Watts1998], scale-free [@Albert2002], and cascade food web [@Solow1998; @Williams2000] networks.
+In all of these networks, rules determining the presence or absence of an interaction between components $i$ and $j$ constrain the overall structure of the network.
 In small-world networks, interactions between components are constrained so that the expected degree of separation between any two components increases in proportion to $\log(S)$ [@Watts1998].
-In scale-free networks, the number of components with which a focal component interacts follows a power law; typically a few components have many interactions while most components have few interactions [@Albert2002].
+In scale-free networks, the distribution of the number of components with which a focal component interacts follows a power law; a few components have many interactions while most components have few interactions [@Albert2002].
 In cascade food webs, species are ranked and interactions are constrained such that a species $i$ can only feed on $j$ if the rank of $i > j$.
@@
 For comparable magnitudes of complexity, structured networks still had a higher number of stable $\mathbf{M}$ than $\mathbf{A}$.
-For random networks, $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma \sqrt{SC} \gtrapprox 1.26$.
+For random networks, $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma_{A} \sqrt{SC} \gtrapprox 1.26$.
 This threshold of complexity, above which more $\mathbf{M}$ than $\mathbf{A}$ were stable, was comparable for small-world networks, and slightly lower for scale-free networks (note that algorithms for generating small-world and scale-free networks necessarily led to varying $C$; see methods).
@@
 
-**System feasibility given $\mathbf{Var(\gamma)}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
+**System feasibility given $\mathbf{\sigma^{2}_{\gamma}}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
 Feasibilility assumes that values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
@@
 Rather, highly complex systems that are observed to be stable are more likely to have varying component response rates, and for this variation to be critical to their stability (Fig.
-3).
-This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing $Var(\gamma)$ (Fig.
-1b, 2b), which can sometimes cause all of the real components of the eigenvalues of the system matrix to become negative, but might also increase the real components of eigenvalues.
+4).
+This is caused by the shift to a non-uniform distribution of eigenvalues that occurs by introducing variation in $\gamma$ (Fig.
+1), which can sometimes cause all of the real components of the eigenvalues of the system matrix to become negative, but might also increase the real components of eigenvalues.
 
-My focus is distinct from Gibbs et al.
+My focus here is distinct from Gibbs et al.
 [@Gibbs2017], who applied the same mathematical framework to investigate how a diagonal matrix $\mathbf{X}$ (equivalent to $\gamma$ in my model) affects the stability of a community matrix $\mathbf{M}$ given an interaction matrix $\mathbf{A}$ within a generalised Lotka-Volterra model, where $\mathbf{M} = \mathbf{XA}$.
@@
 Overall, I show that component response rate variation increases the upper bound of complexity at which stability can be realistically observed, meaning that highly complex systems are more likely than not to vary in their component response rates, and for this variation to be critical for system stability.
-
-The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
-The probability that some combination of component response rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
-Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
-Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
 
@@
 
-A general mathematical framework encompassing shifts in eigenvalue distributions caused by a vector $\gamma$ has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but $\gamma$ has not been interpreted as rates of response of individual system components to perturbation.
+The potential importance of component response rate variation was most evident from the results of simulations in which the genetic algorithm was used in attempt to maximise the probability of system stability.
+The probability that some combination of component response rates could be found to stabilise the system was shown to be up to four orders of magnitude higher than the background probabilities of stability in the absence of any component response rate variation.
+Instead of manipulating the $S \times S$ interactions between system components, it might therefore be possible to manipulate only the $S$ response rates of individual system components to achieve stability.
+Hence, managing the response rates of system components in a targeted way could potentially facilitate the stabilisation of complex systems through a reduction in dimensionality.
+
+A general mathematical framework encompassing shifts in eigenvalue distributions caused by a diagona matrix $\gamma$ has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but $\gamma$ has not been interpreted as rates of response of individual system components to perturbation.
 My model focuses on component response rates for systems of a finite size, in which complexity is high but not yet high enough to make the probability of stability unrealistically low for actual empirical systems.
 For this upper range of system size, randomly assembled complex systems are more likely to be stable if their component response rates vary (e.g., $10 < S < 30$ for parameter values in Fig.
-3).
-Overall, I suggest that variation in component response rate might therefore be critical for maintaining stability in many highly complex empirical systems.
+4).
+Variation in component response rate might therefore be critical for maintaining stability in many highly complex empirical systems.
 These results are broadly applicable for understanding the stability of complex networks across the physical, life, and social sciences.

### 2020-03-13 notebook/ms.tex
Edits to the results and Discussion, improving readability and soothing the notation

@@
 
-Ten thousand numerical simulations of \(\mathbf{M}\) revealed that
+Numerical simulations of \(\mathbf{M}\) revealed that
 \(\sigma^{2}_{\gamma}\) results in a nonlinear relationship between
@@
 (\(\sigma^{2}_{\gamma} = 1/3\)) given \(S = 25\), \(C = 1\), and
-\(\sigma = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded region of Fig.
-2), expected \(\Re(\lambda_{max})\) was lower in \(\mathbf{M}\) than
-\(\mathbf{A}\). For \(\rho \geq -0.1\), the lower bound of the range of
-\(\Re(\lambda_{max})\) values also decreased given
+\(\sigma_{A} = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded region of
+Fig. 2), expected \(\Re(\lambda_{max})\) was lower in \(\mathbf{M}\)
+than \(\mathbf{A}\). For \(\rho \geq -0.1\), the lower bound of the
+range of \(\Re(\lambda_{max})\) values also decreased given
 \(\sigma^{2}_{\gamma}\), resulting in negative \(\Re(\lambda_{max})\) in
@@
 The stabilising effect of \(\sigma^{2}_{\gamma}\) across \(\rho\)
-increased with increasing \(S\). Figure 3 shows 10000 numerical
-simulations of \(\mathbf{M}\) across increasing \(S\) given \(C = 1\)
-and \(\sigma = 0.2\) (\(\sigma\) has been lowered here to better
+increased with increasing \(S\). Figure 3 shows numerical simulations of
+\(\mathbf{M}\) across increasing \(S\) given \(C = 1\) and
+\(\sigma_{A} = 0.2\) (\(\sigma_{A}\) has been lowered here to better
 illustrate the effect of \(S\); note that now given \(S = 25\),
-\(1 = \sigma\sqrt{SC}\)). For relatively small systems (\(S \leq 25\)),
-\(\sigma^{2}_{\gamma}\) never decreased the expected
+\(1 = \sigma_{A}\sqrt{SC}\)). For relatively small systems
+(\(S \leq 25\)), \(\sigma^{2}_{\gamma}\) never decreased the expected
 \(\Re(\lambda_{max})\). But as \(S\) increased, the curvilinear
@@
 investigate the effect of \(\sigma^{2}_{\gamma}\) on stability across
-systems of varying complexity, I simulated random
-\(\mathbf{M = \gamma A}\) matrices at \(\sigma = 0.4\) and \(C = 1\)
+systems of increasing complexity, I simulated random
+\(\mathbf{M = \gamma A}\) matrices at \(\sigma_{A} = 0.4\) and \(C = 1\)
 across \(S = \{2, 3, ..., 49, 50\}\). One million \(\mathbf{M}\) were
@@
 number of stable random systems was higher in \(\mathbf{M}\) than
-\(\mathbf{A}\) (Fig. 3; see Supplementary Information for full table of
+\(\mathbf{A}\) (Fig. 4; see Supplementary Information for full table of
 results), and that the difference between the probabilities of observing
@@
 I also simulated 100000 \(\mathbf{M = \gamma A}\) for three types of
-random networks that are typically interpreted as differing
-interspecific ecological interaction
-types\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{25}}.
-These interactions types are competitive, mutualist, or predator-prey,
+random networks that are typically interpreted as modelling three types
+of interspecific ecological
+interactions\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{25}}.
+These interaction types are competitive, mutualist, and predator-prey,
 as modelled by off-diagonal elements that are constrained to be
@@
 \(S\) increased, a higher number of stable \(\mathbf{M}\) relative to
-\(\mathbf{A}\) were observed for competitor and predator-prey, but not
-mutualist, systems. A higher number of stable systems were observed
+\(\mathbf{A}\) was observed for competitor and predator-prey, but not
+mutualist, systems. A higher number of stable systems was observed
 whenever \(S > 12\) and \(S > 40\) for competitive and predator-prey
@@
 The effect of \(\sigma^{2}_{\gamma}\) on stability did not change
-qualitatively across values of \(C\), \(\sigma\), or different
+qualitatively across values of \(C\), \(\sigma_{A}\), or for different
 distributions of \(\gamma\) (see Supporting Information).
@@
 To investigate how \(\sigma^{2}_{\gamma}\) affects the stability of
-commonly observed network structures, I simulated 1 million
+commonly observed network structures, I simulated one million
 \(\mathbf{M = \gamma A}\) for
@@
 networks. In all of these networks, rules determining the presence or
-absence of an interaction between components \(i\) and \(j\) constrained
+absence of an interaction between components \(i\) and \(j\) constrain
 the overall structure of the network. In small-world networks,
@@
 to \(\log(S)\)\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.
-In scale-free networks, the number of components with which a focal
-component interacts follows a power law; typically a few components have
-many interactions while most components have few
+In scale-free networks, the distribution of the number of components
+with which a focal component interacts follows a power law; a few
+components have many interactions while most components have few
 interactions\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. In
@@
 (\(\sigma = 0.4\) and \(C = 1\)), and therefore complexity
-\(\sigma \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity,
+\(\sigma_{A} \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity,
 above which more \(\mathbf{M}\) than \(\mathbf{A}\) were stable, was
@@
 
-\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
-systems in which individual system components represent the density of
-some tangible quantity, it is relevant to consider the feasibility of
-the system. Feasibilility assumes that values of all components are
-positive at
+\textbf{System feasibility given \(\mathbf{\sigma^{2}_{\gamma}}\)} For
+complex systems in which individual system components represent the
+density of some tangible quantity, it is relevant to consider the
+feasibility of the system. Feasibilility assumes that values of all
+components are positive at
 equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{26},\protect\hyperlink{ref-Song2018}{27}}.
@@
 rates, and for this variation to be critical to their stability (Fig.
-3). This is caused by the shift to a non-uniform distribution of
-eigenvalues that occurs by introducing \(Var(\gamma)\) (Fig. 1b, 2b),
+4). This is caused by the shift to a non-uniform distribution of
+eigenvalues that occurs by introducing variation in \(\gamma\) (Fig. 1),
 which can sometimes cause all of the real components of the eigenvalues
@@
 
-My focus is distinct from Gibbs et
+My focus here is distinct from Gibbs et
 al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}}, who applied
@@
 rates, and for this variation to be critical for system stability.
-
-The potential importance of component response rate variation was most
-evident from the results of simulations in which the genetic algorithm
-was used in attempt to maximise the probability of system stability. The
-probability that some combination of component response rates could be
-found to stabilise the system was shown to be up to four orders of
-magnitude higher than the background probabilities of stability in the
-absence of any component response rate variation. Instead of
-manipulating the \(S \times S\) interactions between system components,
-it might therefore be possible to manipulate only the \(S\) response
-rates of individual system components to achieve stability. Hence,
-managing the response rates of system components in a targeted way could
-potentially facilitate the stabilisation of complex systems through a
-reduction in dimensionality.
 
@@
 
+The potential importance of component response rate variation was most
+evident from the results of simulations in which the genetic algorithm
+was used in attempt to maximise the probability of system stability. The
+probability that some combination of component response rates could be
+found to stabilise the system was shown to be up to four orders of
+magnitude higher than the background probabilities of stability in the
+absence of any component response rate variation. Instead of
+manipulating the \(S \times S\) interactions between system components,
+it might therefore be possible to manipulate only the \(S\) response
+rates of individual system components to achieve stability. Hence,
+managing the response rates of system components in a targeted way could
+potentially facilitate the stabilisation of complex systems through a
+reduction in dimensionality.
+
 A general mathematical framework encompassing shifts in eigenvalue
-distributions caused by a vector \(\gamma\) has been
+distributions caused by a diagona matrix \(\gamma\) has been
 investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}}
@@
 likely to be stable if their component response rates vary (e.g.,
-\(10 < S < 30\) for parameter values in Fig. 3). Overall, I suggest that
-variation in component response rate might therefore be critical for
-maintaining stability in many highly complex empirical systems. These
-results are broadly applicable for understanding the stability of
-complex networks across the physical, life, and social sciences.
+\(10 < S < 30\) for parameter values in Fig. 4). Variation in component
+response rate might therefore be critical for maintaining stability in
+many highly complex empirical systems. These results are broadly
+applicable for understanding the stability of complex networks across
+the physical, life, and social sciences.
 

### 2020-03-13 notebook/ms.Rmd
Update Discussion, so all that is left is a Methods section now

@@
 ```{r, echo = FALSE}
-source(file = "R_old/sim_mat.R");
-source(file = "R_old/plot_figs.R");
+source(file = "../R/utilities.R");
+source(file = "../R/random_mats.R");
 ```
@@
 
-**Component response rate variation ($\mathbf{\gamma}$)**.
+**Component response rate ($\mathbf{\gamma}$) variation**.
 In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
-model a system that includes a vector of potentially changing species densities ($\mathbf{N}$) and a vector of potentially evolving traits ($\mathbf{x}$) [@Patel2018].
-For any species $i$ or trait $j$, change in species density ($N_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{N}$ and $\mathbf{x}$,
+model a system that includes a vector of potentially changing species densities ($\mathbf{n}$) and a vector of potentially evolving traits ($\mathbf{x}$) [@Patel2018].
+For any species $i$ or trait $j$, change in species density ($n_{i}$) or trait value ($x_{j}$) with time ($t$) is a function of the vectors $\mathbf{n}$ and $\mathbf{x}$,
 
-$$\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),$$
+$$\frac{dn_{i}}{dt} = n_{i}f_{i}(\mathbf{n}, \mathbf{x}),$$
 
-$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).$$
+$$\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{n}, \mathbf{x}).$$
 
@@
 In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ defines the values of system components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ defines the change in values of system components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
 In these analyses, row means of $\mathbf{A}$ are expected to be identical, but variation around this expectation will naturally arise due to random sampling of $\mathbf{A}$ off-diagonal elements and finite $S$.
@@
 <!--- Need a section here on how the structured networks were built --->
-
-**Genetic algorithm**.
-Ideally, to investigate the potential of $Var(\gamma)$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
-This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
-Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution [@Hamblin2013].
-In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
-The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
-
-For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M}$ ($\sigma = 0.4$, $C = 1$).
-The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
-Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
-Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
-Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
-The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\Re(\lambda_{max})$ for $\mathbf{M}$).
 
@@
 Hence, $(CS)^{-\delta} = 1$, so in the above, a diagonal matrix of -1s ($\theta \mathbf{I}$) is added to $\mathbf{J}$, which has a diagonal of all zeros and an off-diagonal affecting species interactions (i.e., the expression $(CS)^{-\delta}$ relates to May's [@May1972] stability criterion [@Dougoud2018] by $\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1$, and hence for my purposes $(CS)^{-\delta} = 1$).
-Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below (see also [@Servan2018]),
+Given $\mathbf{A} = \theta\mathbf{I + J}$, the above criteria is therefore reduced to the below (see also Serv&aacute;n et al.
+[@Servan2018]),
 
@@
 Feasibility is satisfied if all of the elements of the resulting vector are positive.
+
+**Genetic algorithm**.
+Ideally, to investigate the potential of $\sigma^{2}_{\gamma}$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
+This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
+Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution [@Hamblin2013].
+In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
+The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
+
+For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M = \gamma A}$ ($\sigma_{A} = 0.4$, $C = 1$).
+The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
+Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
+Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\Re(\lambda_{max})$ for $\mathbf{M}$).
 
@@
 ($\textbf{a}$) A system represented by a matrix $\mathbf{A}$, in which all elements are sampled from a normal distribution with $\mu = 0$ and $\sigma_{A} = 1/\sqrt{S}$.
-Points are uniformly distributed within a circle centred at the origin with a radius of $\sigma_{A} \sqrt{S} = 1$.
+Points are uniformly distributed within the blue circle centred at the origin with a radius of $\sigma_{A} \sqrt{S} = 1$.
 ($\textbf{b}$) The same system as $\textbf{a}$ after including variation in the response rates of $S$ components, represented by the diagonal matrix $\gamma$, such that $\mathbf{M} = \gamma\mathbf{A}$.
 Elements of $\gamma$ are randomly sampled from a uniform distribution from $\min = 0$ to $\max = 2$.
-Eigenvalues of $\mathbf{M}$ are then distributed non-uniformly within a circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
+Eigenvalues of $\mathbf{M}$ are then distributed non-uniformly within the red circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
 ($\textbf{c}$) A different random system $\mathbf{A}$ constructed from the same parameters as in $\textbf{a}$, except with diagonal values of $-1$.
@@
 
-**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems in which variation in $\gamma$ is critical for system stability.
+**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems for which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
@@
 dat <- dat[,-1];
+plot_stables <- function(dat, S_s = 32){
+    Ns          <- 1:S_s;
+    par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
+    #=================================
+    bar_dat                      <- t(cbind(dat[Ns,3], dat[Ns,5]));
+    log_bar_dat                  <- log(bar_dat);
+    log_bar_dat[log_bar_dat < 0] <- 0; 
+    barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
+            names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
+            ylab = "Ln number of stable communities", cex.lab = 1, 
+            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+    axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
+    box(lwd = 2);
+    par(new = TRUE);
+    y1     <- dat[1:S_s,6] / (dat[1:S_s,5]);
+    x1     <- seq(from = 2.132, to = 15.1112, length = S_s);
+    plot(x = x1, y = y1, xaxt = "n", yaxt = "n", lwd = 2, ylim = c(0, 1.1),
+         xlab = "", ylab = "", type = "b", xlim = c(2, 15), pch = 20, 
+         cex = 1, col = "black", yaxs="i");
+    points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
+    axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
+    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
+                        expression(paste(sigma[gamma]^2," = 1/3"))), 
+           pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
+    #=================================
+    mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
+          line = 3.0);
+    mtext(side = 2, text = "Ln number of stable systems", cex = 2, 
+          outer = TRUE, line = 3.5);
+    mtext(side = 4, text = expression(
+        paste("Pr. of systems stable due to Var(",gamma,")")),
+        cex = 2, outer = TRUE, line = 3.5);
+}
 plot_stables(dat);

### 2020-03-13 notebook/ms.tex
Update Discussion, so all that is left is a Methods section now

@@
 
-\textbf{Component response rate variation (\(\mathbf{\gamma}\))}. In a
+\textbf{Component response rate (\(\mathbf{\gamma}\)) variation}. In a
 synthesis of eco-evolutionary feedbacks on community stability, Patel et
 al. model a system that includes a vector of potentially changing
-species densities (\(\mathbf{N}\)) and a vector of potentially evolving
+species densities (\(\mathbf{n}\)) and a vector of potentially evolving
 traits
@@
 For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
+(\(n_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{n}\) and \(\mathbf{x}\),
 
-\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
+\[\frac{dn_{i}}{dt} = n_{i}f_{i}(\mathbf{n}, \mathbf{x}),\]
 
-\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{n}, \mathbf{x}).\]
 
@@
 correspond to individual component response rates. Therefore,
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the values of system
-components and can be analysed using the techniques of
+\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the change in values
+of system components and can be analysed using the techniques of
 May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{21},\protect\hyperlink{ref-May1973}{30}}.
@@
 \(\mathbf{M}\) row means.
-
-\textbf{Genetic algorithm}. Ideally, to investigate the potential of
-\(Var(\gamma)\) for increasing the proportion of stable complex systems,
-the search space of all possible \(\gamma\) vectors would be evaluated
-for each unique \(\mathbf{M = \gamma A}\). This is technically
-impossible because \(\gamma_{i}\) can take any real value between 0-2,
-but even rounding \(\gamma_{i}\) to reasonable values would result in a
-search space too large to practically explore. Under these conditions,
-genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}. In
-this case, the practical solution is finding vectors of
-\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
-\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
-a large population of 1000 different potential \(\mathbf{\gamma}\)
-vectors and allowing this population to evolve through a process of
-mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
-selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
-found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
-met.
-
-For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
-for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)). The
-genetic algorithm was initialised with a population of 1000 different
-\(\mathbf{\gamma}\) vectors with elements sampled i.i.d from
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed on
-the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector, and
-the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
-lowest \(\Re(\lambda_{max})\) each produced 50 clonal offspring with
-subsequent random mutation and crossover between the resulting new
-generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
-\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
-probability of 0.2, resulting in a mutation effect of size
-\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
-\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
-multiplied by \(-1\), and any values that mutated above 2 were set to
-2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
-vectors paired in each generation; vectors were randomly sampled with
-replacement among but not within sets. Vector pairs selected for
-crossover swapped all elements between and including two \(\gamma_{i}\)
-randomly selected with replacement (this allowed for reversal of vector
-element positions during crossover; e.g.,
-\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
-). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
-found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
-fitness increase between generations was less than 0.01 (where fitness
-was defined as \(W_{\gamma} = -\Re(\lambda_{max})\) for \(\mathbf{M}\)).
 
@@
 \(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below (see
-also\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}}),
+reduced to the below (see also Serván et
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}}),
 
@@
 positive.
+
+\textbf{Genetic algorithm}. Ideally, to investigate the potential of
+\(\sigma^{2}_{\gamma}\) for increasing the proportion of stable complex
+systems, the search space of all possible \(\gamma\) vectors would be
+evaluated for each unique \(\mathbf{M = \gamma A}\). This is technically
+impossible because \(\gamma_{i}\) can take any real value between 0-2,
+but even rounding \(\gamma_{i}\) to reasonable values would result in a
+search space too large to practically explore. Under these conditions,
+genetic algorithms are highly useful tools for finding practical
+solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}. In
+this case, the practical solution is finding vectors of
+\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
+\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
+a large population of 1000 different potential \(\mathbf{\gamma}\)
+vectors and allowing this population to evolve through a process of
+mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
+selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
+found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
+met.
+
+For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
+for 100000 random \(\mathbf{M = \gamma A}\) (\(\sigma_{A} = 0.4\),
+\(C = 1\)). The genetic algorithm was initialised with a population of
+1000 different \(\mathbf{\gamma}\) vectors with elements sampled i.i.d
+from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed
+on the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector,
+and the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with
+the lowest \(\Re(\lambda_{max})\) each produced 50 clonal offspring with
+subsequent random mutation and crossover between the resulting new
+generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
+\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+probability of 0.2, resulting in a mutation effect of size
+\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
+\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
+multiplied by \(-1\), and any values that mutated above 2 were set to
+2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
+vectors paired in each generation; vectors were randomly sampled with
+replacement among but not within sets. Vector pairs selected for
+crossover swapped all elements between and including two \(\gamma_{i}\)
+randomly selected with replacement (this allowed for reversal of vector
+element positions during crossover; e.g.,
+\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
+). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
+found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
+fitness increase between generations was less than 0.01 (where fitness
+was defined as \(W_{\gamma} = -\Re(\lambda_{max})\) for \(\mathbf{M}\)).
 
@@
 \(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
-distributed within a circle centred at the origin with a radius of
-\(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
+distributed within the blue circle centred at the origin with a radius
+of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
 \(\textbf{a}\) after including variation in the response rates of \(S\)
@@
 Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
-a circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
+the red circle centred at the origin with a radius of
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 (\(S = \{2, 3, ..., 49, 50 \}\)) given \(C = 1\), and the proportion of
-systems in which variation in \(\gamma\) is critical for system
+systems for which variation in \(\gamma\) is critical for system
 stability. For each \(S\), 1 million complex systems are randomly

### 2020-03-13 notebook/R_old/plot_figs.R

@@ -381,8 +381,8 @@ plot_stables <- function(dat, S_s = 32){
          cex = 1, col = "black", yaxs="i");
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c(expression(paste(gamma," = 1")), 
-                        expression(paste("Var(",gamma,")"))), 
+    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
+                        expression(paste(sigma[gamma]^2," = 1/3"))), 
            pch=15, col=c("grey80","grey40"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 

### 2020-03-13 R/Figures_3_4.R
Fix the legend on the big bar plot figure

@@ -29,8 +29,8 @@ plot_stables <- function(dat, S_s = 32){
          cex = 1, col = "black", yaxs="i");
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c(expression(paste(gamma," = 1")), 
-                        expression(paste("Var(",gamma,")"))), 
+    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
+                        expression(paste(sigma[gamma]^2," = 1/3"))), 
            pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
@@ -75,8 +75,8 @@ plot_stable_4 <- function(dat, S_s = 40){
          cex = 1, col = "black", yaxs="i");
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
-    legend("topleft", c(expression(paste(gamma," = 1")), 
-                        expression(paste("Var(",gamma,")"))), 
+    legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
+                        expression(paste(sigma[gamma]^2," = 0"))), 
            pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 
@@ -87,9 +87,3 @@ plot_stable_4 <- function(dat, S_s = 40){
         paste("Pr. of systems stable due to Var(",gamma,")")),
         cex = 2, outer = TRUE, line = 3.5);
 }
-
-
-
-
-
-

### 2020-03-13 notebook/ms.Rmd
Add first paragraph of methods for constructing random and structured networks, and start the second

@@
 
-<!--- Need a section here on how the structured networks were built --->
+**Construction of random and structured networks**.
+I used the R programming language for all numerical simulations and analyses [@Rproject].
+Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, \sigma_{A})$ with a probability $C$ (unsampled elements were set to zero).
+Diagonal elements $A_{ii}$ were set to $-1$.
+Elements of $\gamma$ were simulated i.i.d.
+from a distribution with positive support (typically $\gamma_{i} \sim \mathcal{U}(0, 2)$).
+Random $\mathbf{A}$ matrices with correlated elements $A_{ij}$ and $A_{ji}$ were built using Cholesky decomposition.
+Competitor networks in which off-diagonal elements $A_{ij} \geq 0$ were constructed by first building a random $\mathbf{A}$, then flipping the sign of any elements in which $A_{ij} < 0$.
+Similarly, mutualist networks were constructed by building a random $\mathbf{A}$, then flipping the sign of elements where $A_{ij} > 0$.
+Predator-prey networks were constructed by first buing a random $\mathbf{A}$, then flipping the sign of either $A_{ij}$ or $A_{ji}$ if $A_{ij} \times A_{ji} > 0$.
+
+Small-world networks were constructed using the method of Watts and Strogatz [@Watts1998].
+A random matrix $\mathbf{J}$ was generated, then a network $\mathbf{S}$ to get $\mathbf{A} = \mathbf{J} \odot \mathbf{S}$
 

### 2020-03-13 notebook/ms.tex
Add first paragraph of methods for constructing random and structured networks, and start the second

@@
 
+\textbf{Construction of random and structured networks}. I used the R
+programming language for all numerical simulations and
+analyses\textsuperscript{\protect\hyperlink{ref-Rproject}{31}}. Purely
+random networks were generated by sampling off-diagonal elements from an
+i.i.d \(A_{ij} \sim \mathcal{N}(0, \sigma_{A})\) with a probability
+\(C\) (unsampled elements were set to zero). Diagonal elements
+\(A_{ii}\) were set to \(-1\). Elements of \(\gamma\) were simulated
+i.i.d. from a distribution with positive support (typically
+\(\gamma_{i} \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices
+with correlated elements \(A_{ij}\) and \(A_{ji}\) were built using
+Cholesky decomposition. Competitor networks in which off-diagonal
+elements \(A_{ij} \geq 0\) were constructed by first building a random
+\(\mathbf{A}\), then flipping the sign of any elements in which
+\(A_{ij} < 0\). Similarly, mutualist networks were constructed by
+building a random \(\mathbf{A}\), then flipping the sign of elements
+where \(A_{ij} > 0\). Predator-prey networks were constructed by first
+buing a random \(\mathbf{A}\), then flipping the sign of either
+\(A_{ij}\) or \(A_{ji}\) if \(A_{ij} \times A_{ji} > 0\).
+
+Small-world networks were constructed using the method of Watts amd
+Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.
+
 \textbf{System feasibility}. Dougoud et
@@
 
+\hypertarget{ref-Rproject}{}
+31. R Core Team. \emph{R: A language and environment for statistical
+computing}. (R Foundation for Statistical Computing, 2018).
+
 \clearpage
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.15.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-14 notebook/ms.Rmd
Add a paragraph on small world networks, now edited a bit more clearly.

@@
 Small-world networks were constructed using the method of Watts and Strogatz [@Watts1998].
-A random matrix $\mathbf{J}$ was generated, then a network $\mathbf{S}$ to get $\mathbf{A} = \mathbf{J} \odot \mathbf{S}$
+First, a regular network [@Watts1998] was created such that components were arranged in a circle.
+Each component was initially set to interact with its $k/2$ closest neighbouring components on each side, where $k$ was an even natural number (e.g., for $k = 2$ the regular network forms a ring in which each component interacts with its two adjacent neighbours; see Supplemental Material for examples).
+Each interaction between a focal component and its neighbour was then removed and replaced with with a probability of $\beta$.
+In replacement, a new component was randomly selected to interact with the focal component; selection was done with equal probability among all but the focal component.
+The resulting small-world network was represented by a square $S \times S$ binary matrix $\mathbf{B}$ in which 1s represented interactions between components and 0s represented the absence of an interaction.
+A new random matrix $\mathbf{J}$ was then generated with elements $J_{ij}$ sampled i.i.d.
+from $\mathcal{N}(0, \sigma_{A})$.
+To build the interaction matrix $\mathbf{A}$, I used element-wise multiplication $\mathbf{A} = \mathbf{J} \odot \mathbf{B}$, then set $diag(\mathbf{A}) = -1$.
+I set $k = S/2$ and simulated across all combinations of $S = \{24, 48, 72, 96, 120, 144, 168\}$ and $\beta = \{0, 0.01, 0.1, 0.25, 1\}$.
+
+Scale-free networks were constructed using the method of Albert and Barab&aacute;si [@Albert2002].
 

### 2020-03-14 notebook/ms.tex
Add a paragraph on small world networks, now edited a bit more clearly.

@@
 
-Small-world networks were constructed using the method of Watts amd
-Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.
+Small-world networks were constructed using the method of Watts and
+Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}. First,
+a regular network\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}
+was created such that components were arranged in a circle. Each
+component was initially set to interact with its \(k/2\) closest
+neighbouring components on each side, where \(k\) was an even natural
+number (e.g., for \(k = 2\) the regular network forms a ring in which
+each component interacts with its two adjacent neighbours; see
+Supplemental Material for examples). Each interaction between a focal
+component and its neighbour was then removed and replaced with with a
+probability of \(\beta\). In replacement, a new component was randomly
+selected to interact with the focal component; selection was done with
+equal probability among all but the focal component. The resulting
+small-world network was represented by a square \(S \times S\) binary
+matrix \(\mathbf{B}\) in which 1s represented interactions between
+components and 0s represented the absence of an interaction. A new
+random matrix \(\mathbf{J}\) was then generated with elements \(J_{ij}\)
+sampled i.i.d. from \(\mathcal{N}(0, \sigma_{A})\). To build the
+interaction matrix \(\mathbf{A}\), I used element-wise multiplication
+\(\mathbf{A} = \mathbf{J} \odot \mathbf{B}\), then set
+\(diag(\mathbf{A}) = -1\). I set \(k = S/2\) and simulated across all
+combinations of \(S = \{24, 48, 72, 96, 120, 144, 168\}\) and
+\(\beta = \{0, 0.01, 0.1, 0.25, 1\}\).
+
+Scale-free networks were constructed using the method of Albert and
+Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}.
 
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.15.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-14 R/utilities.R
Fix the tool for visualising the networks, as it was omitting one node (or, rather, causing the last node to overlap with the first

@@ -125,7 +125,7 @@ get_C <- function(mat){
 #'@export
 visualise_network <- function(mat){
     N  <- dim(mat)[1];
-    rd <- seq(from = 0, to = 2*pi, length = N);
+    rd <- seq(from = 0, to = 2*pi, length = N + 1);
     yy <- sin(rd);
     xx <- cos(rd);
     par(bty = "n");

### 2020-03-14 notebook/ms.Rmd
Finish the explanation of scale free networks in the Methods, move onto cascade food webs

@@
 Scale-free networks were constructed using the method of Albert and Barab&aacute;si [@Albert2002].
+First, a saturated network (all components interact with each other) of size $m \leq S$ was created.
+New components were then added sequentially to the network; each newly added component was set to  interact with $m$ randomly selected existing components.
+When the system size reached $S$, the distribution of the number of total interactions that components had followed a power-law tail [@Albert2002].
+The resulting network was represented by an $S \times S$ binary matrix $\mathbf{G}$, where 1s and 0s represent the presence and absence of an interaction, respectively.
+As with small-world networks, a random matrix $\mathbf{J}$ was generated, and $\mathbf{A} = \mathbf{J} \odot \mathbf{G}$.
+Diagonal were set to $-1$.
+I simulated across all combinations of $S = \{24, 48, 72, 96, 120\}$ and $m = \{2, 3, ..., 11, 12\}$.
+
+Cascade food-webs were constructed following Solow and Beet [@Solow1998].
 

### 2020-03-14 notebook/ms.tex
Finish the explanation of scale free networks in the Methods, move onto cascade food webs

@@
 Scale-free networks were constructed using the method of Albert and
-Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}.
+Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. First,
+a saturated network (all components interact with each other) of size
+\(m \leq S\) was created. New components were then added sequentially to
+the network; each newly added component was set to interact with \(m\)
+randomly selected existing components. When the system size reached
+\(S\), the distribution of the number of total interactions that
+components had followed a power-law
+tail\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. The
+resulting network was represented by an \(S \times S\) binary matrix
+\(\mathbf{G}\), where 1s and 0s represent the presence and absence of an
+interaction, respectively. As with small-world networks, a random matrix
+\(\mathbf{J}\) was generated, and
+\(\mathbf{A} = \mathbf{J} \odot \mathbf{G}\). Diagonal were set to
+\(-1\). I simulated across all combinations of
+\(S = \{24, 48, 72, 96, 120\}\) and \(m = \{2, 3, ..., 11, 12\}\).
 
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.15.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-14 notebook/ms.Rmd
Finished a draft now, need to reread and smooth the text, and work on the SI

@@
 I used the R programming language for all numerical simulations and analyses [@Rproject].
-Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, \sigma_{A})$ with a probability $C$ (unsampled elements were set to zero).
+Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, 0.4)$ with a probability $C$ (unsampled elements were set to zero).
 Diagonal elements $A_{ii}$ were set to $-1$.
@@
 A new random matrix $\mathbf{J}$ was then generated with elements $J_{ij}$ sampled i.i.d.
-from $\mathcal{N}(0, \sigma_{A})$.
+from $\mathcal{N}(0, 0.4)$.
 To build the interaction matrix $\mathbf{A}$, I used element-wise multiplication $\mathbf{A} = \mathbf{J} \odot \mathbf{B}$, then set $diag(\mathbf{A}) = -1$.
-I set $k = S/2$ and simulated across all combinations of $S = \{24, 48, 72, 96, 120, 144, 168\}$ and $\beta = \{0, 0.01, 0.1, 0.25, 1\}$.
+I set $k = S/2$ and simulated small-world networks across all combinations of $S = \{24, 48, 72, 96, 120, 144, 168\}$ and $\beta = \{0, 0.01, 0.1, 0.25, 1\}$.
 
@@
 Diagonal were set to $-1$.
-I simulated across all combinations of $S = \{24, 48, 72, 96, 120\}$ and $m = \{2, 3, ..., 11, 12\}$.
+I simulated scale-free networks across all combinations of $S = \{24, 48, 72, 96, 120\}$ and $m = \{2, 3, ..., 11, 12\}$.
 
 Cascade food-webs were constructed following Solow and Beet [@Solow1998].
+First, a random matrix $\mathbf{A}$ was generated with off-diagonal elements sampled i.i.d so that $A_{ij} \sim \mathcal{N}(0, 0.4)$.
+Each component in the system was ranked from $1$ to $S$.
+If component $i$ had a higher rank than component $j$ and $A_{ij} < 0$, then $A_{ij}$ was multiplied by $-1$.
+If $i$ had a lower rank than $j$ and $A_{ji} < 0$, then $A_{ji}$ was multiplied by $-1$.
+In practice, this resulted in a matrix $\mathbf{A}$ with negative and positive values in the lower and upper triangles, respectively.
+Diagonal elements of $\mathbf{A}$ were set to $-1$ and $C = 1$.
+I simulated cascade food webs across all combinations of $S = \{2, 3, ..., 59, 60\}$.
 

### 2020-03-14 notebook/ms.tex
Finished a draft now, need to reread and smooth the text, and work on the SI

@@
 random networks were generated by sampling off-diagonal elements from an
-i.i.d \(A_{ij} \sim \mathcal{N}(0, \sigma_{A})\) with a probability
-\(C\) (unsampled elements were set to zero). Diagonal elements
-\(A_{ii}\) were set to \(-1\). Elements of \(\gamma\) were simulated
-i.i.d. from a distribution with positive support (typically
+i.i.d \(A_{ij} \sim \mathcal{N}(0, 0.4)\) with a probability \(C\)
+(unsampled elements were set to zero). Diagonal elements \(A_{ii}\) were
+set to \(-1\). Elements of \(\gamma\) were simulated i.i.d. from a
+distribution with positive support (typically
 \(\gamma_{i} \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices
@@
 random matrix \(\mathbf{J}\) was then generated with elements \(J_{ij}\)
-sampled i.i.d. from \(\mathcal{N}(0, \sigma_{A})\). To build the
-interaction matrix \(\mathbf{A}\), I used element-wise multiplication
+sampled i.i.d. from \(\mathcal{N}(0, 0.4)\). To build the interaction
+matrix \(\mathbf{A}\), I used element-wise multiplication
 \(\mathbf{A} = \mathbf{J} \odot \mathbf{B}\), then set
-\(diag(\mathbf{A}) = -1\). I set \(k = S/2\) and simulated across all
-combinations of \(S = \{24, 48, 72, 96, 120, 144, 168\}\) and
+\(diag(\mathbf{A}) = -1\). I set \(k = S/2\) and simulated small-world
+networks across all combinations of
+\(S = \{24, 48, 72, 96, 120, 144, 168\}\) and
 \(\beta = \{0, 0.01, 0.1, 0.25, 1\}\).
@@
 \(\mathbf{A} = \mathbf{J} \odot \mathbf{G}\). Diagonal were set to
-\(-1\). I simulated across all combinations of
+\(-1\). I simulated scale-free networks across all combinations of
 \(S = \{24, 48, 72, 96, 120\}\) and \(m = \{2, 3, ..., 11, 12\}\).
+
+Cascade food-webs were constructed following Solow and
+Beet\textsuperscript{\protect\hyperlink{ref-Solow1998}{17}}. First, a
+random matrix \(\mathbf{A}\) was generated with off-diagonal elements
+sampled i.i.d so that \(A_{ij} \sim \mathcal{N}(0, 0.4)\). Each
+component in the system was ranked from \(1\) to \(S\). If component
+\(i\) had a higher rank than component \(j\) and \(A_{ij} < 0\), then
+\(A_{ij}\) was multiplied by \(-1\). If \(i\) had a lower rank than
+\(j\) and \(A_{ji} < 0\), then \(A_{ji}\) was multiplied by \(-1\). In
+practice, this resulted in a matrix \(\mathbf{A}\) with negative and
+positive values in the lower and upper triangles, respectively. Diagonal
+elements of \(\mathbf{A}\) were set to \(-1\) and \(C = 1\). I simulated
+cascade food webs across all combinations of
+\(S = \{2, 3, ..., 59, 60\}\).
 
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.15.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.11.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-14 notebook/SI.Rmd
Update SI to put some examples of structured networks

@@
 source(file = "R_old/plot_figs.R");
+source(file = "../R/utilities.R")
 ```
@@
 - [Sensitivity of distribution of $\gamma$](#gam_dist)
+- [Structured network construction](#structured)
 - [Feasibility of complex systems](#Feasibility)
@@
 In comparison to the uniform distribution (a), proportionally fewer random systems are found with the exponential distribution (b), while more are found with the beta (c) and gamma (d) distributions.
+
+Structured network construction {#structured}
+================================================================================
+
+Example of a small-world network
+
+```{r, echo = FALSE}
+source("../R/small_world.R");
+par(mfrow = c(1, 2), mar = c(5, 5, 1, 1));
+swn <- create_swn(S = 72, K = 12, beta = 0.01);
+visualise_network(swn);
+swn <- create_swn(S = 72, K = 12, beta = 0.1);
+visualise_network(swn);
+```
+
+Example of a scale-free network
+
+```{r, echo = FALSE}
+source("../R/scale_free.R");
+sfn <- create_sfn(S = 72, m = 3);
+par(mfrow = c(1, 2), mar = c(5, 5, 1, 1));
+visualise_network(sfn);
+dd_dat       <- sfn;
+diag(dd_dat) <- 0;
+deg_dist     <- apply(X = dd_dat, MARGIN = 1, FUN = sum);
+hist(deg_dist, main = "", xlab = "Component interactions",
+     ylab = "Frequency", cex.lab = 1.25, cex.axis = 1.25,
+     col = "grey", breaks = 20);
+```
 

### 2020-03-14 notebook/SI.tex
Update SI to put some examples of structured networks

@@
 \item
+  \protect\hyperlink{structured}{Structured network construction}
+\item
   \protect\hyperlink{Feasibility}{Feasibility of complex systems}
@@
 
+\hypertarget{structured}{\section{Structured network
+construction}\label{structured}}
+
+Example of a small-world network
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1.pdf}
+
+Example of a scale-free network
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+
 \hypertarget{Feasibility}{\section{Feasibility of complex
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-28-1.pdf}
 
@@
 
-{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-28-1}
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-30-1}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-33-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-34-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-33-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-39-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-40-1.pdf}
 

### 2020-03-15 notebook/SI.Rmd
Remove the correlation effects on stability from the SI, as these are now in the main text

@@
 - [Sensitivity of distribution of $\gamma$](#gam_dist)
-- [Structured network construction](#structured)
+- [Stability of structured networks](#structured)
 - [Feasibility of complex systems](#Feasibility)
@@
 - [Consistency with Gibbs et al. (2018)](#Gibbs)
-- [Effects of matrix correlations](#rhos)
 
@@
 
-Effects of matrix correlations
-================================================================================
-
 References

### 2020-03-15 notebook/SI.Rmd
Report the SWN results and add the network figures to the SI

@@
 
-Structured network construction {#structured}
+Stability of structured networks {#structured}
 ================================================================================
 
-Example of a small-world network
+I tested the stability of one million random, small-world, scale-free, and cascade food web networks.
+Each of these networks is structured differently.
+In the main text, the random networks and cascade food webs that I built were saturated ($C = 1$), meaning that every component was connected to, and interacted with, every other component.
 
-```{r, echo = FALSE}
+\vspace{2mm}
+
+```{r, echo = FALSE, fig.height = 6, fig.width = 6, fig.align = "center"}
+A0_dat <- rnorm(n = 20 * 20, mean = 2, sd = 0);
+A0     <- matrix(data = A0_dat, nrow = 20, ncol = 20);
+visualise_network(A0);
+```
+
+\vspace{2mm}
+
+Small-world networks, in contrast, are not saturated.
+They are instead defined by components that interact mostly with other closely neighbouring components, but have a proportion of interactions ($\beta$) that are instead between non-neighbours.
+Two small-world networks are shown below.
+
+\vspace{2mm}
+
+```{r, echo = FALSE, fig.align = "center"}
 source("../R/small_world.R");
@@
 ```
+
+\vspace{2mm}
+
+The small-world network on the left shows a system in which $\beta = 0.01$, while the small-world network on the right shows one in which $\beta = 0.1$.
+At the extremes of $\beta = 0$ and $\beta = 1$, networks are regular and random, respectively.
+The table below shows how $\sigma^{2}_\gamma$ affects stability in small world networks across different values of $S$ and $\beta$.
+
+```{r, echo = FALSE, fig.align = "center"}
+swn <- read.csv(file = "sim_results/swn/old/swn.csv"); # Change when new results are in
+use <- c(1:6, 19:21);
+kable(swn[,use]);
+```
+
+In the above, the mean $C$, complexity of $\mathbf{A}$, and complexity of $\mathbf{M}$ are also shown.
 

### 2020-03-15 notebook/SI.tex
Report the SWN results and add the network figures to the SI

--- 2020-03-11 notebook/SI.tex
@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
+\newcommand{\ImportTok}[1]{{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
+\newcommand{\BuiltInTok}[1]{{#1}}
+\newcommand{\ExtensionTok}[1]{{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
+\newcommand{\RegionMarkerTok}[1]{{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
+\newcommand{\NormalTok}[1]{{#1}}
 \usepackage{longtable,booktabs}
@@
 \setlength{\droptitle}{-2em}
-
   \title{Component response rate variation underlies the stability of complex
@@
 \item
+  \protect\hyperlink{structured}{Stability of structured networks}
+\item
   \protect\hyperlink{Feasibility}{Feasibility of complex systems}
@@
   \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
-\item
-  \protect\hyperlink{rhos}{Effects of matrix correlations}
 \end{itemize}
@@
 
+\hypertarget{structured}{\section{Stability of structured
+networks}\label{structured}}
+
+I tested the stability of one million random, small-world, scale-free,
+and cascade food web networks. Each of these networks is structured
+differently. In the main text, the random networks and cascade food webs
+that I built were saturated (\(C = 1\)), meaning that every component
+was connected to, and interacted with, every other component.
+
+\vspace{2mm}
+
+\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1} \end{center}
+
+\vspace{2mm}
+
+Small-world networks, in contrast, are not saturated. They are instead
+defined by components that interact mostly with other closely
+neighbouring components, but have a proportion of interactions
+(\(\beta\)) that are instead between non-neighbours. Two small-world
+networks are shown below.
+
+\vspace{2mm}
+
+\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1} \end{center}
+
+\vspace{2mm}
+
+The small-world network on the left shows a system in which
+\(\beta = 0.01\), while the small-world network on the right shows one
+in which \(\beta = 0.1\). At the extremes of \(\beta = 0\) and
+\(\beta = 1\), networks are regular and random, respectively. The table
+below shows how \(\sigma^{2}_\gamma\) affects stability in small world
+networks across different values of \(S\) and \(\beta\).
+
+\begin{longtable}[]{@{}rrrrrrrrr@{}}
+\toprule
+beta & S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+complex\_A0 & complex\_A1 & C\tabularnewline
+\midrule
+\endhead
+0.00 & 24 & 17442 & 982558 & 17489 & 982511 & 0.1695021 & 0.1941223 &
+0.1304348\tabularnewline
+0.00 & 48 & 257013 & 742987 & 259517 & 740483 & 0.2355489 & 0.2711380 &
+0.1063830\tabularnewline
+0.00 & 72 & 715587 & 284413 & 723082 & 276918 & 0.2866561 & 0.3304009 &
+0.0985915\tabularnewline
+0.00 & 96 & 961817 & 38183 & 963276 & 36724 & 0.3299115 & 0.3804889 &
+0.0947368\tabularnewline
+0.00 & 120 & 998976 & 1024 & 998812 & 1188 & 0.3681518 & 0.4246688 &
+0.0924370\tabularnewline
+0.00 & 144 & 999998 & 2 & 999991 & 9 & 0.4027471 & 0.4646754 &
+0.0909091\tabularnewline
+0.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345714 & 0.5014674 &
+0.0898204\tabularnewline
+0.01 & 24 & 17386 & 982614 & 17433 & 982567 & 0.1695343 & 0.1941452 &
+0.1304321\tabularnewline
+0.01 & 48 & 255369 & 744631 & 257956 & 742044 & 0.2355569 & 0.2711250 &
+0.1063798\tabularnewline
+0.01 & 72 & 708407 & 291593 & 716019 & 283981 & 0.2866423 & 0.3303843 &
+0.0985884\tabularnewline
+0.01 & 96 & 960591 & 39409 & 961685 & 38315 & 0.3299322 & 0.3804641 &
+0.0947337\tabularnewline
+0.01 & 120 & 998991 & 1009 & 998813 & 1187 & 0.3681428 & 0.4246683 &
+0.0924339\tabularnewline
+0.01 & 144 & 999997 & 3 & 999993 & 7 & 0.4027696 & 0.4646907 &
+0.0909060\tabularnewline
+0.01 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345881 & 0.5014852 &
+0.0898172\tabularnewline
+0.10 & 24 & 20408 & 979592 & 20478 & 979522 & 0.1692658 & 0.1937917 &
+0.1302962\tabularnewline
+0.10 & 48 & 238101 & 761899 & 240787 & 759213 & 0.2351005 & 0.2705892 &
+0.1062307\tabularnewline
+0.10 & 72 & 680138 & 319862 & 685524 & 314476 & 0.2861156 & 0.3297663 &
+0.0984350\tabularnewline
+0.10 & 96 & 961691 & 38309 & 959740 & 40260 & 0.3293218 & 0.3797666 &
+0.0945786\tabularnewline
+0.10 & 120 & 999544 & 456 & 999282 & 718 & 0.3674529 & 0.4238479 &
+0.0922777\tabularnewline
+0.10 & 144 & 999999 & 1 & 999999 & 1 & 0.4019505 & 0.4637655 &
+0.0907492\tabularnewline
+0.10 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4337704 & 0.5005276 &
+0.0896598\tabularnewline
+0.25 & 24 & 23923 & 976077 & 24031 & 975969 & 0.1680052 & 0.1922420 &
+0.1296696\tabularnewline
+0.25 & 48 & 228389 & 771611 & 231312 & 768688 & 0.2331850 & 0.2683721 &
+0.1055242\tabularnewline
+0.25 & 72 & 666374 & 333626 & 668472 & 331528 & 0.2836431 & 0.3268807 &
+0.0977062\tabularnewline
+0.25 & 96 & 966462 & 33538 & 961616 & 38384 & 0.3263952 & 0.3763783 &
+0.0938386\tabularnewline
+0.25 & 120 & 999759 & 241 & 999510 & 490 & 0.3642255 & 0.4201168 &
+0.0915315\tabularnewline
+0.25 & 144 & 1000000 & 0 & 999999 & 1 & 0.3983171 & 0.4595509 &
+0.0899989\tabularnewline
+0.25 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4298720 & 0.4960198 &
+0.0889060\tabularnewline
+1.00 & 24 & 26159 & 973841 & 26309 & 973691 & 0.1588024 & 0.1815081 &
+0.1249673\tabularnewline
+1.00 & 48 & 211700 & 788300 & 214607 & 785393 & 0.2154544 & 0.2478696 &
+0.0991369\tabularnewline
+1.00 & 72 & 613966 & 386034 & 615862 & 384138 & 0.2602685 & 0.2998802 &
+0.0908174\tabularnewline
+1.00 & 96 & 943949 & 56051 & 936840 & 63160 & 0.2984622 & 0.3441159 &
+0.0867075\tabularnewline
+1.00 & 120 & 999134 & 866 & 998351 & 1649 & 0.3323169 & 0.3833061 &
+0.0842580\tabularnewline
+1.00 & 144 & 1000000 & 0 & 999997 & 3 & 0.3630114 & 0.4187917 &
+0.0826318\tabularnewline
+1.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.3913277 & 0.4515183 &
+0.0814745\tabularnewline
+\bottomrule
+\end{longtable}
+
+In the above, the mean \(C\), complexity of \(\mathbf{A}\), and
+complexity of \(\mathbf{M}\) are also shown.
+
+Example of a scale-free network
+
+\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+
 \hypertarget{Feasibility}{\section{Feasibility of complex
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-28-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
 
@@
 
-{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-28-1}
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-32-1}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-30-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-34-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-31-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-33-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-35-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-39-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-36-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-40-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-41-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
-
-\section{Effects of matrix
-correlations}\label{effects-of-matrix-correlations}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-42-1.pdf}
 

### 2020-03-15 notebook/SI.Rmd
Add section on scale-free networks to the SI, and fix one of the figure sizes

@@
 
-**Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$) for b-d.
+**Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to identical systems in which $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$) for b-d.
 Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$).
@@
 
-```{r, eval = TRUE, echo = FALSE, fig.height = 7, fig.width = 6}
+```{r, eval = TRUE, echo = FALSE, fig.height = 4, fig.width = 4, fig.align="center"}
 make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
@@
     y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
-    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
+    par(mfrow = c(2, 2), oma = c(4, 4, 1, 1), mar = c(2, 0.5, 0.5, 0.5));
     h1 <- hist(y1, breaks = 1000, plot = FALSE);
-    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.15,
          ylim = c(0, max(h1$counts)*1.2));
-    mtext("a", adj = 0, line = -2.3, cex = 2, 
+    mtext("a", adj = 0, line = -1.3, cex = 1.2, 
           at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
@@
     h2 <- hist(y2, breaks = 1000, plot = FALSE);
-    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.15,
          ylim = c(0, max(h2$counts)*1.2));
-    mtext("b", adj = 0, line = -2.3, cex = 2, 
+    mtext("b", adj = 0, line = -1.3, cex = 1.2, 
           at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
@@
     h3 <- hist(y3, breaks = 1000, plot = FALSE);
-    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.15,
          ylim = c(0, max(h3$counts)*1.2));
-    mtext("c", adj = 0, line = -2.3, cex = 2, 
+    mtext("c", adj = 0, line = -1.3, cex = 1.2, 
           at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
@@
     h4 <- hist(y4, breaks = 1000, plot = FALSE);
-    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
+    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.15,
          ylim = c(0, max(h4$counts)*1.2));
-    mtext("d", adj = 0, line = -2.3, cex = 2, 
+    mtext("d", adj = 0, line = -1.3, cex = 1.2, 
           at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
@@
     mtext(expression(paste("Component ",gamma," value")),
-          outer=TRUE,side=1,line=3.0,cex=2);
+          outer=TRUE,side=1,line=1.0,cex=1.25);
     mtext(expression(paste("Relative frequency")),
-          outer=TRUE,side=2,line=2.5,cex=2);
+          outer=TRUE,side=2,line=1.5,cex=1.25);
 }
@@
 
-```{r, echo = FALSE, fig.height = 6, fig.width = 6, fig.align = "center"}
+```{r, echo = FALSE, fig.height = 5, fig.width = 5, fig.align = "center"}
 A0_dat <- rnorm(n = 20 * 20, mean = 2, sd = 0);
@@
 In the above, the mean $C$, complexity of $\mathbf{A}$, and complexity of $\mathbf{M}$ are also shown.
+For similar magnitudes of complexity as in random networks of $\sigma\sqrt{SC} \gtrapprox 1.26$, variation in $\gamma$ typically results in more stable than unstable systems.
 
-Example of a scale-free network
+Scale-free networks are also not saturated, but are defined by an interaction number distribution that follows a power law.
+In other words, a small number of components interact with many other components, while most components interact with only a small number of other components.
+Scale-free networks can be built by adding new components, one by one, to an existing system, with each newly added component interacting with a randomly selected subset of $m$ existing components.
+The network on the left below shows an example of a scale-free network in which $m = 3$.
+The histogram on the right shows the number of other components with which each component interacts.
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.align = "center"}
 source("../R/scale_free.R");
@@
 ```
+
+The table below shows how $\sigma^{2}_\gamma$ affects stability across different scale-free networks with different $S$ and $m$ values.
+
+```{r, echo = FALSE, fig.align = "center"}
+sfn <- read.csv(file = "sim_results/sfn/sfn.csv"); # Change when new results are in
+use <- c(1:6, 19:21);
+kable(sfn[,use]);
+```
+
+As in small-world networks, the mean $C$ is shown, along with the mean complexities of $\mathbf{A}$ and $\mathbf{M}$.
+Like all other networks, $\sigma^{2}_\gamma$ increases the stability of scale-free networks given sufficiently complexity, but the complexity requires seems to be relatively low.
 

### 2020-03-15 notebook/SI.tex
Add section on scale-free networks to the SI, and fix one of the figure sizes

@@
 \newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
-\newcommand{\ImportTok}[1]{{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
-\newcommand{\BuiltInTok}[1]{{#1}}
-\newcommand{\ExtensionTok}[1]{{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
-\newcommand{\RegionMarkerTok}[1]{{#1}}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
-\newcommand{\NormalTok}[1]{{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
 \usepackage{longtable,booktabs}
@@
 \setlength{\droptitle}{-2em}
+
   \title{Component response rate variation underlies the stability of complex
@@
 simulated complex systems with these \(\gamma\) distributions are
-compared to otherwise identical complex systems with a fixed component
-response rate of \(\gamma = 1\) across different system sizes (\(S\);
-i.e., component numbers) given a unit \(\gamma\) standard deviation
-(\(\sigma_{\gamma} = 1\)) for b-d. Distributions are as follows: (a)
-uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
+compared to identical systems in which \(\gamma = 1\) across different
+system sizes (\(S\); i.e., component numbers) given a unit \(\gamma\)
+standard deviation (\(\sigma_{\gamma} = 1\)) for b-d. Distributions are
+as follows: (a) uniform, (b) exponential, (c) beta (\(\alpha = 0.5\) and
 \(\beta = 0.5\)), and (d) gamma (\(k = 2\) and \(\theta = 2\)). Each
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-21-1.pdf}
+\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-21-1} \end{center}
 
@@
 In the above, the mean \(C\), complexity of \(\mathbf{A}\), and
-complexity of \(\mathbf{M}\) are also shown.
-
-Example of a scale-free network
-
-\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1.pdf}
+complexity of \(\mathbf{M}\) are also shown. For similar magnitudes of
+complexity as in random networks of \(\sigma\sqrt{SC} \gtrapprox 1.26\),
+variation in \(\gamma\) typically results in more stable than unstable
+systems.
+
+Scale-free networks are also not saturated, but are defined by an
+interaction number distribution that follows a power law. In other
+words, a small number of components interact with many other components,
+while most components interact with only a small number of other
+components. Scale-free networks can be built by adding new components,
+one by one, to an existing system, with each newly added component
+interacting with a randomly selected subset of \(m\) existing
+components. The network on the left below shows an example of a
+scale-free network in which \(m = 3\). The histogram on the right shows
+the number of other components with which each component interacts.
+
+\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-26-1} \end{center}
+
+The table below shows how \(\sigma^{2}_\gamma\) affects stability across
+different scale-free networks with different \(S\) and \(m\) values.
+
+\begin{longtable}[]{@{}rrrrrrrrr@{}}
+\toprule
+S & M & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
+complex\_A0 & complex\_A1 & C\tabularnewline
+\midrule
+\endhead
+24 & 2 & 152791 & 847209 & 156034 & 843966 & 0.7891257 & 0.9034663 &
+0.1648551\tabularnewline
+24 & 3 & 320481 & 679519 & 326351 & 673649 & 0.9566487 & 1.0967499 &
+0.2409420\tabularnewline
+24 & 4 & 504433 & 495567 & 504826 & 495174 & 1.0922870 & 1.2532761 &
+0.3134058\tabularnewline
+24 & 5 & 670676 & 329324 & 660426 & 339574 & 1.2073054 & 1.3857169 &
+0.3822464\tabularnewline
+24 & 6 & 798637 & 201363 & 779345 & 220655 & 1.3067095 & 1.5004508 &
+0.4474638\tabularnewline
+24 & 7 & 884082 & 115918 & 862215 & 137785 & 1.3942577 & 1.6013368 &
+0.5090580\tabularnewline
+24 & 8 & 936190 & 63810 & 915630 & 84370 & 1.4722315 & 1.6908563 &
+0.5670290\tabularnewline
+24 & 9 & 964868 & 35132 & 948297 & 51703 & 1.5414455 & 1.7707292 &
+0.6213768\tabularnewline
+24 & 10 & 981460 & 18540 & 967911 & 32089 & 1.6030044 & 1.8417459 &
+0.6721014\tabularnewline
+24 & 11 & 989838 & 10162 & 980232 & 19768 & 1.6586511 & 1.9059313 &
+0.7192029\tabularnewline
+24 & 12 & 994393 & 5607 & 987436 & 12564 & 1.7081503 & 1.9628898 &
+0.7626812\tabularnewline
+48 & 2 & 303963 & 696037 & 310053 & 689947 & 0.7946875 & 0.9132519 &
+0.0828901\tabularnewline
+48 & 3 & 577855 & 422145 & 579996 & 420004 & 0.9685494 & 1.1141445 &
+0.1227837\tabularnewline
+48 & 4 & 810001 & 189999 & 799132 & 200868 & 1.1122992 & 1.2799335 &
+0.1617908\tabularnewline
+48 & 5 & 938004 & 61996 & 924613 & 75387 & 1.2369960 & 1.4236817 &
+0.1999113\tabularnewline
+48 & 6 & 984975 & 15025 & 976433 & 23567 & 1.3478291 & 1.5514420 &
+0.2371454\tabularnewline
+48 & 7 & 997160 & 2840 & 994005 & 5995 & 1.4473792 & 1.6663763 &
+0.2734929\tabularnewline
+48 & 8 & 999584 & 416 & 998590 & 1410 & 1.5385445 & 1.7716359 &
+0.3089539\tabularnewline
+48 & 9 & 999955 & 45 & 999707 & 293 & 1.6227742 & 1.8687074 &
+0.3435284\tabularnewline
+48 & 10 & 999992 & 8 & 999939 & 61 & 1.7006157 & 1.9583879 &
+0.3772163\tabularnewline
+48 & 11 & 999999 & 1 & 999990 & 10 & 1.7731759 & 2.0420990 &
+0.4100177\tabularnewline
+48 & 12 & 1000000 & 0 & 999999 & 1 & 1.8410402 & 2.1203112 &
+0.4419326\tabularnewline
+72 & 2 & 427243 & 572757 & 434600 & 565400 & 0.7964226 & 0.9166566 &
+0.0553599\tabularnewline
+72 & 3 & 741345 & 258655 & 739020 & 260980 & 0.9723446 & 1.1195788 &
+0.0823552\tabularnewline
+72 & 4 & 931043 & 68957 & 921145 & 78855 & 1.1188220 & 1.2888100 &
+0.1089593\tabularnewline
+72 & 5 & 989644 & 10356 & 984372 & 15628 & 1.2466268 & 1.4361875 &
+0.1351721\tabularnewline
+72 & 6 & 999131 & 869 & 997914 & 2086 & 1.3604666 & 1.5674966 &
+0.1609937\tabularnewline
+72 & 7 & 999946 & 54 & 999804 & 196 & 1.4642496 & 1.6872501 &
+0.1864241\tabularnewline
+72 & 8 & 999999 & 1 & 999988 & 12 & 1.5596340 & 1.7974044 &
+0.2114632\tabularnewline
+72 & 9 & 1000000 & 0 & 999999 & 1 & 1.6482181 & 1.8994441 &
+0.2361111\tabularnewline
+72 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7307859 & 1.9947150 &
+0.2603678\tabularnewline
+72 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8086766 & 2.0847262 &
+0.2842332\tabularnewline
+72 & 12 & 1000000 & 0 & 1000000 & 0 & 1.8817533 & 2.1689764 &
+0.3077074\tabularnewline
+96 & 2 & 527633 & 472367 & 535188 & 464812 & 0.7974024 & 0.9183557 &
+0.0415570\tabularnewline
+96 & 3 & 842274 & 157726 & 837756 & 162244 & 0.9741293 & 1.1224709 &
+0.0619518\tabularnewline
+96 & 4 & 975834 & 24166 & 969478 & 30522 & 1.1220115 & 1.2931371 &
+0.0821272\tabularnewline
+96 & 5 & 998391 & 1609 & 996991 & 3009 & 1.2511287 & 1.4422331 &
+0.1020833\tabularnewline
+96 & 6 & 999955 & 45 & 999838 & 162 & 1.3669903 & 1.5757699 &
+0.1218202\tabularnewline
+96 & 7 & 999999 & 1 & 999996 & 4 & 1.4725862 & 1.6977057 &
+0.1413377\tabularnewline
+96 & 8 & 1000000 & 0 & 1000000 & 0 & 1.5699145 & 1.8099762 &
+0.1606360\tabularnewline
+96 & 9 & 1000000 & 0 & 1000000 & 0 & 1.6606162 & 1.9146804 &
+0.1797149\tabularnewline
+96 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7457971 & 2.0129344 &
+0.1985746\tabularnewline
+96 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8260368 & 2.1055559 &
+0.2172149\tabularnewline
+96 & 12 & 1000000 & 0 & 1000000 & 0 & 1.9018608 & 2.1929362 &
+0.2356360\tabularnewline
+120 & 2 & 609563 & 390437 & 616036 & 383964 & 0.7979355 & 0.9194404 &
+0.0332633\tabularnewline
+120 & 3 & 904064 & 95936 & 899040 & 100960 & 0.9753815 & 1.1243251 &
+0.0496499\tabularnewline
+120 & 4 & 991710 & 8290 & 988410 & 11590 & 1.1239922 & 1.2957520 &
+0.0658964\tabularnewline
+120 & 5 & 999781 & 219 & 999477 & 523 & 1.2539362 & 1.4458518 &
+0.0820028\tabularnewline
+120 & 6 & 999999 & 1 & 999981 & 19 & 1.3707937 & 1.5806987 &
+0.0979692\tabularnewline
+120 & 7 & 1000000 & 0 & 999999 & 1 & 1.4775366 & 1.7038860 &
+0.1137955\tabularnewline
+120 & 8 & 1000000 & 0 & 1000000 & 0 & 1.5762636 & 1.8177236 &
+0.1294818\tabularnewline
+120 & 9 & 1000000 & 0 & 1000000 & 0 & 1.6680647 & 1.9238257 &
+0.1450280\tabularnewline
+120 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7545110 & 2.0233838 &
+0.1604342\tabularnewline
+120 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8363882 & 2.1178385 &
+0.1757003\tabularnewline
+120 & 12 & 1000000 & 0 & 1000000 & 0 & 1.9135798 & 2.2069806 &
+0.1908263\tabularnewline
+\bottomrule
+\end{longtable}
+
+As in small-world networks, the mean \(C\) is shown, along with the mean
+complexities of \(\mathbf{A}\) and \(\mathbf{M}\). Like all other
+networks, \(\sigma^{2}_\gamma\) increases the stability of scale-free
+networks given sufficiently complexity, but the complexity requires
+seems to be relatively low.
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-28-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-29-1.pdf}
 
@@
 
-{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-32-1}
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-33-1}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-37-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-39-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-40-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-40-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-41-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-41-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-42-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-42-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-43-1.pdf}
 

### 2020-03-15 notebook/SI.Rmd
Adust the SI so that the matrix references are closer to those used in the main text

@@
 The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
-In this table, the `A0` refers to matrices where $\gamma = 1$, while `A1` refers to matrices after $Var(\gamma)$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
-Each row summarises data for a given $S$ over 1 million randomly simulated $\mathbf{M}$ (`A0` and `A1`).
-The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 1 million).
-Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
-The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
+In this table, the `A` refers to $\mathbf{A}$ matrices where $\gamma = 1$, while `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
+Each row summarises data for a given $S$ over 1 million randomly simulated $\mathbf{M}$.
+The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 1 million).
+Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable.
+The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{M}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.
 
@@
 
-Overall, the ratio of stable `A1` matrices to stable `A0` matrices found is greater than 1 whenever $S > 10$ (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
+Overall, the ratio of stable $\mathbf{A}$ matrices to stable $\mathbf{M}$ matrices found is greater than 1 whenever $S > 10$ (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
 Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$.
@@
 The following tables for restricted ecological communities can therefore be compared with the random $\mathbf{M}$ [results above](#IncrS) (but note that counts from systems with comparable probabilities of stability will be an order of magnitude lower in the tables below due to the smaller number of $\mathbf{M}$ matrices generated).
-As with the [results above](#IncrS), in the tables below, `A0` refers to matrices when $\gamma = 1$ and `A1` refers to matrices after $Var(\gamma)$ is added.
-The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
-Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
-The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
+As with the [results above](#IncrS), in the tables below, `A` refers to matrices $\mathbf{A}$ when $\gamma = 1$ and `M` refers to matrices after $\sigma^{2}_{\gamma}$ is added.
+The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 100000).
+Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `A1_stable` shows the number that are stable.
+The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{A}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.
 
@@
 Overall, as expected [@Allesina2012], predator-prey communities are relatively stable while mutualist communties are highly unstable.
-But interestingly, while $Var(\gamma)$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
-This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $Var(\gamma)$ alone will shift all real parts of eigenvalues to negative values.
+But interestingly, while $\sigma^{2}_{\gamma}$ stabilises predator-prey and competitor communities, it does not stabilise mutualist communities.
+This is unsurprising because purely mutualist communities are characterised by a very positive[@Allesina2012] leading $\Re(\lambda)$, and it is highly unlikely that $\sigma^{2}_{\gamma}$ alone will shift all real parts of eigenvalues to negative values.
 
@@
 In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.
-As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A0` refers to matrices when $\gamma = 1$, and `A1` refers to matrices after $Var(\gamma)$ is added.
-The column `A0_unstable` shows the number of `A0` matrices that are unstable, and the column `A0_stable` shows the number of `A0` matrices that are stable (these two columns sum to 100000).
-Similarly, the column `A1_unstable` shows the number of `A1` matrices that are unstable and `A1_stable` shows the number that are stable.
-The columns `A1_stabilised` and `A1_destabilised` show how many `A0` matrices were stabilised or destabilised, respectively, by $Var(\gamma)$.
+As with the results on [stability with increasing $S$](#IncrS) shown above, in the tables below `A` refers to $\mathbf{A}$ matrices when $\gamma = 1$, and `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added.
+The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 100000).
+Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable.
+The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{A}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.
 
@@
 C3dat   <- C3dat[,-1];
-C3d_tab <- C3dat[,1:8];
+C3d_tab <- C3dat[,1:7];
 kable(C3d_tab);
@@
 C5dat   <- C5dat[,-1];
-C5d_tab <- C5dat[,1:8];
+C5d_tab <- C5dat[,1:7];
 kable(C5d_tab);
@@
 C7dat   <- C7dat[,-1];
-C7d_tab <- C7dat[,1:8];
+C7d_tab <- C7dat[,1:7];
 kable(C7d_tab);
@@
 C9dat   <- C9dat[,-1];
-C9d_tab <- C9dat[,1:8];
+C9d_tab <- C9dat[,1:7];
 kable(C9d_tab);
@@
 Results below show stability results given varying interaction strengths ($\sigma$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows).
-In the tables below (as [above](#IncrS)), `A0` and `A1` refers to matrices for $\gamma = 1$ and $Var(\gamma)$, respectively.
+In the tables below (as [above](#IncrS)), `A` and `M` refers to matrices for $\gamma = 1$ and $\sigma^{2}_{\gamma}$, respectively.
 
@@
 
-Hence, in general, $Var(\gamma)$ does not appear to affect feasibility in pure species interaction networks [@Servan2018].
+Hence, in general, $\sigma^{2}_{\gamma}$ does not appear to affect feasibility in pure species interaction networks [@Servan2018].
 
@@
 Stability results are also shown in the table below.
-Results for `A0` indicate systems in which $\gamma = 1$, while `A1` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
+Results for `A` indicate systems in which $\gamma = 1$, while `M` refers to systems in which the genetic algorithm searched for a set of $\gamma$ values that stabilised the system.
 
@@
 ```{r, echo = FALSE}
-load("../data/constant_complexity.rda");
+constant_complexity <- read.csv("sim_results/constant_complexity.csv")
 kable(constant_complexity);

### 2020-03-15 notebook/SI.tex
Adust the SI so that the matrix references are closer to those used in the main text

@@
 \(\sigma = 0.4\) and \(C = 1\) given a range of
-\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
-matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
-after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
-Each row summarises data for a given \(S\) over 1 million randomly
-simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 1
-million). Similarly, the column \texttt{A1\_unstable} shows the number
-of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
-the number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A} refers to
+\(\mathbf{A}\) matrices where \(\gamma = 1\), while \texttt{M} refers to
+\(\mathbf{M}\) matrices after \(\sigma^{2}_{\gamma}\) is added and
+\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
+\(S\) over 1 million randomly simulated \(\mathbf{M}\). The column
+\texttt{A\_unstable} shows the number of \(\mathbf{A}\) matrices that
+are unstable, and the column \texttt{A\_stable} shows the number of
+\(\mathbf{A}\) matrices that are stable (these two columns sum to 1
+million). Similarly, the column \texttt{M\_unstable} shows the number of
+\(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable} shows
+the number that are stable. The columns \texttt{A\_stabilised} and
+\texttt{A\_destabilised} show how many \(\mathbf{M}\) matrices were
+stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 \toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
@@
 
-Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
-matrices found is greater than 1 whenever \(S > 10\) (compare column 5
-to column 3), and this ratio increases with increasing \(S\) (column 1).
-Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
-given variation in \(\gamma\) than when \(\gamma = 1\). Note that
-feasibility results were omitted for the table above, but are
-\protect\hyperlink{Feasibility}{reported below}.
+Overall, the ratio of stable \(\mathbf{A}\) matrices to stable
+\(\mathbf{M}\) matrices found is greater than 1 whenever \(S > 10\)
+(compare column 5 to column 3), and this ratio increases with increasing
+\(S\) (column 1). Hence, more randomly created complex systems
+(\(\mathbf{M}\)) are stable given variation in \(\gamma\) than when
+\(\gamma = 1\). Note that feasibility results were omitted for the table
+above, but are \protect\hyperlink{Feasibility}{reported below}.
 
@@
 \protect\hyperlink{IncrS}{results above}, in the tables below,
-\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
-refers to matrices after \(Var(\gamma)\) is added. The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 100000).
-Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+\texttt{A} refers to matrices \(\mathbf{A}\) when \(\gamma = 1\) and
+\texttt{M} refers to matrices after \(\sigma^{2}_{\gamma}\) is added.
+The column \texttt{A\_unstable} shows the number of \(\mathbf{A}\)
+matrices that are unstable, and the column \texttt{A\_stable} shows the
+number of \(\mathbf{A}\) matrices that are stable (these two columns sum
+to 100000). Similarly, the column \texttt{M\_unstable} shows the number
+of \(\mathbf{M}\) matrices that are unstable and \texttt{A1\_stable}
+shows the number that are stable. The columns \texttt{A\_stabilised} and
+\texttt{A\_destabilised} show how many \(\mathbf{A}\) matrices were
+stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
@@
 predator-prey communities are relatively stable while mutualist
-communties are highly unstable. But interestingly, while \(Var(\gamma)\)
-stabilises predator-prey and competitor communities, it does not
-stabilise mutualist communities. This is unsurprising because purely
-mutualist communities are characterised by a very
-positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-leading \(\Re(\lambda)\), and it is highly unlikely that \(Var(\gamma)\)
-alone will shift all real parts of eigenvalues to negative values.
+communties are highly unstable. But interestingly, while
+\(\sigma^{2}_{\gamma}\) stabilises predator-prey and competitor
+communities, it does not stabilise mutualist communities. This is
+unsurprising because purely mutualist communities are characterised by a
+very positive\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+leading \(\Re(\lambda)\), and it is highly unlikely that
+\(\sigma^{2}_{\gamma}\) alone will shift all real parts of eigenvalues
+to negative values.
 
@@
 \protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
-in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
-and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
-column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
-that are unstable, and the column \texttt{A0\_stable} shows the number
-of \texttt{A0} matrices that are stable (these two columns sum to
-100000). Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+in the tables below \texttt{A} refers to \(\mathbf{A}\) matrices when
+\(\gamma = 1\), and \texttt{M} refers to \(\mathbf{M}\) matrices after
+\(\sigma^{2}_{\gamma}\) is added. The column \texttt{A\_unstable} shows
+the number of \(\mathbf{A}\) matrices that are unstable, and the column
+\texttt{A\_stable} shows the number of \(\mathbf{A}\) matrices that are
+stable (these two columns sum to 100000). Similarly, the column
+\texttt{M\_unstable} shows the number of \(\mathbf{M}\) matrices that
+are unstable and \texttt{M\_stable} shows the number that are stable.
+The columns \texttt{A\_stabilised} and \texttt{A\_destabilised} show how
+many \(\mathbf{A}\) matrices were stabilised or destabilised,
+respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 
-\begin{longtable}[]{@{}llllllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 5 & 99995 & 5 & 99995 & 0 & 0 & 75110\tabularnewline
-3 & 6 & 99994 & 6 & 99994 & 0 & 0 & 87526\tabularnewline
-4 & 24 & 99976 & 24 & 99976 & 0 & 0 & 93713\tabularnewline
-5 & 59 & 99941 & 59 & 99941 & 0 & 0 & 96929\tabularnewline
-6 & 98 & 99902 & 98 & 99902 & 0 & 0 & 98492\tabularnewline
-7 & 160 & 99840 & 161 & 99839 & 0 & 1 & 99243\tabularnewline
-8 & 290 & 99710 & 293 & 99707 & 0 & 3 & 99582\tabularnewline
-9 & 430 & 99570 & 434 & 99566 & 0 & 4 & 99821\tabularnewline
-10 & 648 & 99352 & 653 & 99347 & 1 & 6 & 99895\tabularnewline
-11 & 946 & 99054 & 957 & 99043 & 0 & 11 & 99945\tabularnewline
-12 & 1392 & 98608 & 1415 & 98585 & 4 & 27 & 99978\tabularnewline
-13 & 2032 & 97968 & 2065 & 97935 & 5 & 38 & 99987\tabularnewline
-14 & 2627 & 97373 & 2688 & 97312 & 10 & 71 & 99988\tabularnewline
-15 & 3588 & 96412 & 3647 & 96353 & 35 & 94 & 99996\tabularnewline
-16 & 5019 & 94981 & 5124 & 94876 & 51 & 156 & 99998\tabularnewline
-17 & 6512 & 93488 & 6673 & 93327 & 79 & 240 & 99999\tabularnewline
-18 & 8444 & 91556 & 8600 & 91400 & 165 & 321 & 100000\tabularnewline
-19 & 10416 & 89584 & 10667 & 89333 & 244 & 495 & 100000\tabularnewline
-20 & 13254 & 86746 & 13477 & 86523 & 425 & 648 & 100000\tabularnewline
-21 & 16248 & 83752 & 16481 & 83519 & 642 & 875 & 100000\tabularnewline
-22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151 & 100000\tabularnewline
-23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490 & 100000\tabularnewline
-24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818 & 100000\tabularnewline
-25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137 & 100000\tabularnewline
-26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506 & 100000\tabularnewline
-27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811 & 100000\tabularnewline
-28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072 & 100000\tabularnewline
-29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321 & 100000\tabularnewline
-30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480 & 100000\tabularnewline
-31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568 & 100000\tabularnewline
-32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531 & 100000\tabularnewline
-33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360 & 100000\tabularnewline
-34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014 & 100000\tabularnewline
-35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773 & 100000\tabularnewline
-36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393 & 100000\tabularnewline
-37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913 & 100000\tabularnewline
-38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566 & 100000\tabularnewline
-39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187 & 100000\tabularnewline
-40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936 & 100000\tabularnewline
-41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683 & 100000\tabularnewline
-42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449 & 100000\tabularnewline
-43 & 99275 & 725 & 98312 & 1688 & 1251 & 288 & 100000\tabularnewline
-44 & 99583 & 417 & 98872 & 1128 & 903 & 192 & 100000\tabularnewline
-45 & 99776 & 224 & 99339 & 661 & 576 & 139 & 100000\tabularnewline
-46 & 99865 & 135 & 99518 & 482 & 413 & 66 & 100000\tabularnewline
-47 & 99938 & 62 & 99744 & 256 & 226 & 32 & 100000\tabularnewline
-48 & 99956 & 44 & 99824 & 176 & 151 & 19 & 100000\tabularnewline
-49 & 99980 & 20 & 99914 & 86 & 85 & 19 & 100000\tabularnewline
-50 & 99993 & 7 & 99950 & 50 & 46 & 3 & 100000\tabularnewline
-51 & 99998 & 2 & 99971 & 29 & 28 & 1 & 100000\tabularnewline
-52 & 99998 & 2 & 99986 & 14 & 14 & 2 & 100000\tabularnewline
-53 & 99999 & 1 & 99992 & 8 & 7 & 0 & 100000\tabularnewline
-54 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
-55 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-56 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
-57 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-58 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+2 & 5 & 99995 & 5 & 99995 & 0 & 0\tabularnewline
+3 & 6 & 99994 & 6 & 99994 & 0 & 0\tabularnewline
+4 & 24 & 99976 & 24 & 99976 & 0 & 0\tabularnewline
+5 & 59 & 99941 & 59 & 99941 & 0 & 0\tabularnewline
+6 & 98 & 99902 & 98 & 99902 & 0 & 0\tabularnewline
+7 & 160 & 99840 & 161 & 99839 & 0 & 1\tabularnewline
+8 & 290 & 99710 & 293 & 99707 & 0 & 3\tabularnewline
+9 & 430 & 99570 & 434 & 99566 & 0 & 4\tabularnewline
+10 & 648 & 99352 & 653 & 99347 & 1 & 6\tabularnewline
+11 & 946 & 99054 & 957 & 99043 & 0 & 11\tabularnewline
+12 & 1392 & 98608 & 1415 & 98585 & 4 & 27\tabularnewline
+13 & 2032 & 97968 & 2065 & 97935 & 5 & 38\tabularnewline
+14 & 2627 & 97373 & 2688 & 97312 & 10 & 71\tabularnewline
+15 & 3588 & 96412 & 3647 & 96353 & 35 & 94\tabularnewline
+16 & 5019 & 94981 & 5124 & 94876 & 51 & 156\tabularnewline
+17 & 6512 & 93488 & 6673 & 93327 & 79 & 240\tabularnewline
+18 & 8444 & 91556 & 8600 & 91400 & 165 & 321\tabularnewline
+19 & 10416 & 89584 & 10667 & 89333 & 244 & 495\tabularnewline
+20 & 13254 & 86746 & 13477 & 86523 & 425 & 648\tabularnewline
+21 & 16248 & 83752 & 16481 & 83519 & 642 & 875\tabularnewline
+22 & 19497 & 80503 & 19719 & 80281 & 929 & 1151\tabularnewline
+23 & 23654 & 76346 & 23776 & 76224 & 1368 & 1490\tabularnewline
+24 & 28485 & 71515 & 28389 & 71611 & 1914 & 1818\tabularnewline
+25 & 32774 & 67226 & 32483 & 67517 & 2428 & 2137\tabularnewline
+26 & 38126 & 61874 & 37411 & 62589 & 3221 & 2506\tabularnewline
+27 & 43435 & 56565 & 42418 & 57582 & 3828 & 2811\tabularnewline
+28 & 49333 & 50667 & 47840 & 52160 & 4565 & 3072\tabularnewline
+29 & 55389 & 44611 & 53381 & 46619 & 5329 & 3321\tabularnewline
+30 & 60826 & 39174 & 58388 & 41612 & 5918 & 3480\tabularnewline
+31 & 66820 & 33180 & 64043 & 35957 & 6345 & 3568\tabularnewline
+32 & 72190 & 27810 & 69036 & 30964 & 6685 & 3531\tabularnewline
+33 & 77053 & 22947 & 73587 & 26413 & 6826 & 3360\tabularnewline
+34 & 81816 & 18184 & 78157 & 21843 & 6673 & 3014\tabularnewline
+35 & 85651 & 14349 & 82041 & 17959 & 6383 & 2773\tabularnewline
+36 & 88985 & 11015 & 85657 & 14343 & 5721 & 2393\tabularnewline
+37 & 92072 & 7928 & 88805 & 11195 & 5180 & 1913\tabularnewline
+38 & 94329 & 5671 & 91444 & 8556 & 4451 & 1566\tabularnewline
+39 & 95912 & 4088 & 93295 & 6705 & 3804 & 1187\tabularnewline
+40 & 97232 & 2768 & 95201 & 4799 & 2967 & 936\tabularnewline
+41 & 98179 & 1821 & 96506 & 3494 & 2356 & 683\tabularnewline
+42 & 98826 & 1174 & 97489 & 2511 & 1786 & 449\tabularnewline
+43 & 99275 & 725 & 98312 & 1688 & 1251 & 288\tabularnewline
+44 & 99583 & 417 & 98872 & 1128 & 903 & 192\tabularnewline
+45 & 99776 & 224 & 99339 & 661 & 576 & 139\tabularnewline
+46 & 99865 & 135 & 99518 & 482 & 413 & 66\tabularnewline
+47 & 99938 & 62 & 99744 & 256 & 226 & 32\tabularnewline
+48 & 99956 & 44 & 99824 & 176 & 151 & 19\tabularnewline
+49 & 99980 & 20 & 99914 & 86 & 85 & 19\tabularnewline
+50 & 99993 & 7 & 99950 & 50 & 46 & 3\tabularnewline
+51 & 99998 & 2 & 99971 & 29 & 28 & 1\tabularnewline
+52 & 99998 & 2 & 99986 & 14 & 14 & 2\tabularnewline
+53 & 99999 & 1 & 99992 & 8 & 7 & 0\tabularnewline
+54 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
+55 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+56 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+57 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+58 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-100 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{}\tabularnewline
+100 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llllllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 74863\tabularnewline
-3 & 32 & 99968 & 32 & 99968 & 0 & 0 & 87434\tabularnewline
-4 & 122 & 99878 & 122 & 99878 & 0 & 0 & 93761\tabularnewline
-5 & 320 & 99680 & 321 & 99679 & 0 & 1 & 96830\tabularnewline
-6 & 667 & 99333 & 673 & 99327 & 0 & 6 & 98481\tabularnewline
-7 & 1233 & 98767 & 1252 & 98748 & 0 & 19 & 99187\tabularnewline
-8 & 2123 & 97877 & 2156 & 97844 & 3 & 36 & 99654\tabularnewline
-9 & 3415 & 96585 & 3471 & 96529 & 16 & 72 & 99816\tabularnewline
-10 & 5349 & 94651 & 5450 & 94550 & 30 & 131 & 99900\tabularnewline
-11 & 7990 & 92010 & 8185 & 91815 & 81 & 276 & 99958\tabularnewline
-12 & 11073 & 88927 & 11301 & 88699 & 219 & 447 & 99973\tabularnewline
-13 & 14971 & 85029 & 15204 & 84796 & 445 & 678 & 99986\tabularnewline
-14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002 & 99991\tabularnewline
-15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404 & 99996\tabularnewline
-16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980 & 99999\tabularnewline
-17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476 & 100000\tabularnewline
-18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937 & 99999\tabularnewline
-19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262 & 99999\tabularnewline
-20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527 & 100000\tabularnewline
-21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673 & 100000\tabularnewline
-22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491 & 100000\tabularnewline
-23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244 & 100000\tabularnewline
-24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876 & 100000\tabularnewline
-25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382 & 100000\tabularnewline
-26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901 & 100000\tabularnewline
-27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358 & 100000\tabularnewline
-28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974 & 100000\tabularnewline
-29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657 & 100000\tabularnewline
-30 & 99103 & 897 & 98033 & 1967 & 1466 & 396 & 100000\tabularnewline
-31 & 99502 & 498 & 98665 & 1335 & 1068 & 231 & 100000\tabularnewline
-32 & 99745 & 255 & 99185 & 815 & 696 & 136 & 100000\tabularnewline
-33 & 99881 & 119 & 99572 & 428 & 375 & 66 & 100000\tabularnewline
-34 & 99955 & 45 & 99788 & 212 & 191 & 24 & 100000\tabularnewline
-35 & 99979 & 21 & 99900 & 100 & 95 & 16 & 100000\tabularnewline
-36 & 99995 & 5 & 99950 & 50 & 50 & 5 & 100000\tabularnewline
-37 & 99997 & 3 & 99970 & 30 & 28 & 1 & 100000\tabularnewline
-38 & 99998 & 2 & 99986 & 14 & 13 & 1 & 100000\tabularnewline
-39 & 99999 & 1 & 99991 & 9 & 9 & 1 & 100000\tabularnewline
-40 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
-41 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-42 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-43 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
+3 & 32 & 99968 & 32 & 99968 & 0 & 0\tabularnewline
+4 & 122 & 99878 & 122 & 99878 & 0 & 0\tabularnewline
+5 & 320 & 99680 & 321 & 99679 & 0 & 1\tabularnewline
+6 & 667 & 99333 & 673 & 99327 & 0 & 6\tabularnewline
+7 & 1233 & 98767 & 1252 & 98748 & 0 & 19\tabularnewline
+8 & 2123 & 97877 & 2156 & 97844 & 3 & 36\tabularnewline
+9 & 3415 & 96585 & 3471 & 96529 & 16 & 72\tabularnewline
+10 & 5349 & 94651 & 5450 & 94550 & 30 & 131\tabularnewline
+11 & 7990 & 92010 & 8185 & 91815 & 81 & 276\tabularnewline
+12 & 11073 & 88927 & 11301 & 88699 & 219 & 447\tabularnewline
+13 & 14971 & 85029 & 15204 & 84796 & 445 & 678\tabularnewline
+14 & 19754 & 80246 & 19992 & 80008 & 764 & 1002\tabularnewline
+15 & 25020 & 74980 & 25239 & 74761 & 1185 & 1404\tabularnewline
+16 & 30860 & 69140 & 30938 & 69062 & 1902 & 1980\tabularnewline
+17 & 37844 & 62156 & 37562 & 62438 & 2758 & 2476\tabularnewline
+18 & 44909 & 55091 & 44251 & 55749 & 3595 & 2937\tabularnewline
+19 & 52322 & 47678 & 51011 & 48989 & 4573 & 3262\tabularnewline
+20 & 60150 & 39850 & 58295 & 41705 & 5382 & 3527\tabularnewline
+21 & 67147 & 32853 & 64895 & 35105 & 5925 & 3673\tabularnewline
+22 & 74177 & 25823 & 71358 & 28642 & 6310 & 3491\tabularnewline
+23 & 80297 & 19703 & 77034 & 22966 & 6507 & 3244\tabularnewline
+24 & 85372 & 14628 & 82039 & 17961 & 6209 & 2876\tabularnewline
+25 & 89719 & 10281 & 86539 & 13461 & 5562 & 2382\tabularnewline
+26 & 92947 & 7053 & 90141 & 9859 & 4707 & 1901\tabularnewline
+27 & 95436 & 4564 & 92950 & 7050 & 3844 & 1358\tabularnewline
+28 & 97196 & 2804 & 95171 & 4829 & 2999 & 974\tabularnewline
+29 & 98300 & 1700 & 96842 & 3158 & 2115 & 657\tabularnewline
+30 & 99103 & 897 & 98033 & 1967 & 1466 & 396\tabularnewline
+31 & 99502 & 498 & 98665 & 1335 & 1068 & 231\tabularnewline
+32 & 99745 & 255 & 99185 & 815 & 696 & 136\tabularnewline
+33 & 99881 & 119 & 99572 & 428 & 375 & 66\tabularnewline
+34 & 99955 & 45 & 99788 & 212 & 191 & 24\tabularnewline
+35 & 99979 & 21 & 99900 & 100 & 95 & 16\tabularnewline
+36 & 99995 & 5 & 99950 & 50 & 50 & 5\tabularnewline
+37 & 99997 & 3 & 99970 & 30 & 28 & 1\tabularnewline
+38 & 99998 & 2 & 99986 & 14 & 13 & 1\tabularnewline
+39 & 99999 & 1 & 99991 & 9 & 9 & 1\tabularnewline
+40 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
+41 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+42 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+43 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llllllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 7 & 99993 & 7 & 99993 & 0 & 0 & 75160\tabularnewline
-3 & 106 & 99894 & 106 & 99894 & 0 & 0 & 87447\tabularnewline
-4 & 395 & 99605 & 397 & 99603 & 0 & 2 & 93849\tabularnewline
-5 & 1117 & 98883 & 1123 & 98877 & 0 & 6 & 96827\tabularnewline
-6 & 2346 & 97654 & 2367 & 97633 & 6 & 27 & 98402\tabularnewline
-7 & 4314 & 95686 & 4388 & 95612 & 16 & 90 & 99218\tabularnewline
-8 & 7327 & 92673 & 7456 & 92544 & 61 & 190 & 99611\tabularnewline
-9 & 11514 & 88486 & 11792 & 88208 & 150 & 428 & 99808\tabularnewline
-10 & 16247 & 83753 & 16584 & 83416 & 415 & 752 & 99904\tabularnewline
-11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162 & 99952\tabularnewline
-12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818 & 99977\tabularnewline
-13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355 & 99984\tabularnewline
-14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927 & 99995\tabularnewline
-15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421 & 99994\tabularnewline
-16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527 & 99999\tabularnewline
-17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495 & 100000\tabularnewline
-18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351 & 100000\tabularnewline
-19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979 & 99999\tabularnewline
-20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339 & 100000\tabularnewline
-21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869 & 100000\tabularnewline
-22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267 & 100000\tabularnewline
-23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791 & 100000\tabularnewline
-24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504 & 100000\tabularnewline
-25 & 99433 & 567 & 98630 & 1370 & 1079 & 276 & 100000\tabularnewline
-26 & 99760 & 240 & 99259 & 741 & 655 & 154 & 100000\tabularnewline
-27 & 99895 & 105 & 99576 & 424 & 377 & 58 & 100000\tabularnewline
-28 & 99950 & 50 & 99790 & 210 & 194 & 34 & 100000\tabularnewline
-29 & 99981 & 19 & 99915 & 85 & 80 & 14 & 100000\tabularnewline
-30 & 99994 & 6 & 99952 & 48 & 47 & 5 & 100000\tabularnewline
-31 & 99998 & 2 & 99972 & 28 & 28 & 2 & 100000\tabularnewline
-32 & 99999 & 1 & 99992 & 8 & 8 & 1 & 100000\tabularnewline
-33 & 100000 & 0 & 99997 & 3 & 3 & 0 & 100000\tabularnewline
-34 & 100000 & 0 & 99999 & 1 & 1 & 0 & 100000\tabularnewline
-35 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+2 & 7 & 99993 & 7 & 99993 & 0 & 0\tabularnewline
+3 & 106 & 99894 & 106 & 99894 & 0 & 0\tabularnewline
+4 & 395 & 99605 & 397 & 99603 & 0 & 2\tabularnewline
+5 & 1117 & 98883 & 1123 & 98877 & 0 & 6\tabularnewline
+6 & 2346 & 97654 & 2367 & 97633 & 6 & 27\tabularnewline
+7 & 4314 & 95686 & 4388 & 95612 & 16 & 90\tabularnewline
+8 & 7327 & 92673 & 7456 & 92544 & 61 & 190\tabularnewline
+9 & 11514 & 88486 & 11792 & 88208 & 150 & 428\tabularnewline
+10 & 16247 & 83753 & 16584 & 83416 & 415 & 752\tabularnewline
+11 & 22481 & 77519 & 22759 & 77241 & 884 & 1162\tabularnewline
+12 & 29459 & 70541 & 29729 & 70271 & 1548 & 1818\tabularnewline
+13 & 37631 & 62369 & 37567 & 62433 & 2419 & 2355\tabularnewline
+14 & 46317 & 53683 & 45696 & 54304 & 3548 & 2927\tabularnewline
+15 & 54945 & 45055 & 53695 & 46305 & 4671 & 3421\tabularnewline
+16 & 63683 & 36317 & 61643 & 38357 & 5567 & 3527\tabularnewline
+17 & 72004 & 27996 & 69375 & 30625 & 6124 & 3495\tabularnewline
+18 & 79220 & 20780 & 76158 & 23842 & 6413 & 3351\tabularnewline
+19 & 85286 & 14714 & 82283 & 17717 & 5982 & 2979\tabularnewline
+20 & 90240 & 9760 & 87181 & 12819 & 5398 & 2339\tabularnewline
+21 & 93676 & 6324 & 91077 & 8923 & 4468 & 1869\tabularnewline
+22 & 96203 & 3797 & 94045 & 5955 & 3425 & 1267\tabularnewline
+23 & 97866 & 2134 & 96161 & 3839 & 2496 & 791\tabularnewline
+24 & 98842 & 1158 & 97633 & 2367 & 1713 & 504\tabularnewline
+25 & 99433 & 567 & 98630 & 1370 & 1079 & 276\tabularnewline
+26 & 99760 & 240 & 99259 & 741 & 655 & 154\tabularnewline
+27 & 99895 & 105 & 99576 & 424 & 377 & 58\tabularnewline
+28 & 99950 & 50 & 99790 & 210 & 194 & 34\tabularnewline
+29 & 99981 & 19 & 99915 & 85 & 80 & 14\tabularnewline
+30 & 99994 & 6 & 99952 & 48 & 47 & 5\tabularnewline
+31 & 99998 & 2 & 99972 & 28 & 28 & 2\tabularnewline
+32 & 99999 & 1 & 99992 & 8 & 8 & 1\tabularnewline
+33 & 100000 & 0 & 99997 & 3 & 3 & 0\tabularnewline
+34 & 100000 & 0 & 99999 & 1 & 1 & 0\tabularnewline
+35 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 
-\begin{longtable}[]{@{}llllllll@{}}
+\begin{longtable}[]{@{}lllllll@{}}
 \toprule
-N & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised & A0\_infeasible\tabularnewline
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
 \midrule
 \endhead
-2 & 14 & 99986 & 14 & 99986 & 0 & 0 & 75187\tabularnewline
-3 & 240 & 99760 & 240 & 99760 & 0 & 0 & 87443\tabularnewline
-4 & 1008 & 98992 & 1016 & 98984 & 0 & 8 & 93795\tabularnewline
-5 & 2708 & 97292 & 2729 & 97271 & 2 & 23 & 96814\tabularnewline
-6 & 5669 & 94331 & 5755 & 94245 & 13 & 99 & 98439\tabularnewline
-7 & 9848 & 90152 & 10057 & 89943 & 91 & 300 & 99208\tabularnewline
-8 & 15903 & 84097 & 16201 & 83799 & 336 & 634 & 99603\tabularnewline
-9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168 & 99803\tabularnewline
-10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852 & 99909\tabularnewline
-11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507 & 99951\tabularnewline
-12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127 & 99977\tabularnewline
-13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649 & 99986\tabularnewline
-14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765 & 99995\tabularnewline
-15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550 & 100000\tabularnewline
-16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092 & 100000\tabularnewline
-17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424 & 100000\tabularnewline
-18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746 & 100000\tabularnewline
-19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167 & 99999\tabularnewline
-20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729 & 100000\tabularnewline
-21 & 99111 & 889 & 98133 & 1867 & 1389 & 411 & 100000\tabularnewline
-22 & 99588 & 412 & 98905 & 1095 & 903 & 220 & 100000\tabularnewline
-23 & 99837 & 163 & 99480 & 520 & 452 & 95 & 100000\tabularnewline
-24 & 99932 & 68 & 99744 & 256 & 228 & 40 & 100000\tabularnewline
-25 & 99976 & 24 & 99863 & 137 & 133 & 20 & 100000\tabularnewline
-26 & 99995 & 5 & 99950 & 50 & 49 & 4 & 100000\tabularnewline
-27 & 99996 & 4 & 99986 & 14 & 13 & 3 & 100000\tabularnewline
-28 & 100000 & 0 & 99993 & 7 & 7 & 0 & 100000\tabularnewline
-29 & 100000 & 0 & 99996 & 4 & 4 & 0 & 100000\tabularnewline
-30 & 100000 & 0 & 99998 & 2 & 2 & 0 & 100000\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+2 & 14 & 99986 & 14 & 99986 & 0 & 0\tabularnewline
+3 & 240 & 99760 & 240 & 99760 & 0 & 0\tabularnewline
+4 & 1008 & 98992 & 1016 & 98984 & 0 & 8\tabularnewline
+5 & 2708 & 97292 & 2729 & 97271 & 2 & 23\tabularnewline
+6 & 5669 & 94331 & 5755 & 94245 & 13 & 99\tabularnewline
+7 & 9848 & 90152 & 10057 & 89943 & 91 & 300\tabularnewline
+8 & 15903 & 84097 & 16201 & 83799 & 336 & 634\tabularnewline
+9 & 22707 & 77293 & 23110 & 76890 & 765 & 1168\tabularnewline
+10 & 30796 & 69204 & 31122 & 68878 & 1526 & 1852\tabularnewline
+11 & 40224 & 59776 & 40082 & 59918 & 2649 & 2507\tabularnewline
+12 & 49934 & 50066 & 49288 & 50712 & 3773 & 3127\tabularnewline
+13 & 60138 & 39862 & 58803 & 41197 & 4984 & 3649\tabularnewline
+14 & 69100 & 30900 & 67110 & 32890 & 5755 & 3765\tabularnewline
+15 & 77607 & 22393 & 74884 & 25116 & 6273 & 3550\tabularnewline
+16 & 84663 & 15337 & 81780 & 18220 & 5975 & 3092\tabularnewline
+17 & 90075 & 9925 & 87290 & 12710 & 5209 & 2424\tabularnewline
+18 & 93944 & 6056 & 91419 & 8581 & 4271 & 1746\tabularnewline
+19 & 96650 & 3350 & 94530 & 5470 & 3287 & 1167\tabularnewline
+20 & 98160 & 1840 & 96698 & 3302 & 2191 & 729\tabularnewline
+21 & 99111 & 889 & 98133 & 1867 & 1389 & 411\tabularnewline
+22 & 99588 & 412 & 98905 & 1095 & 903 & 220\tabularnewline
+23 & 99837 & 163 & 99480 & 520 & 452 & 95\tabularnewline
+24 & 99932 & 68 & 99744 & 256 & 228 & 40\tabularnewline
+25 & 99976 & 24 & 99863 & 137 & 133 & 20\tabularnewline
+26 & 99995 & 5 & 99950 & 50 & 49 & 4\tabularnewline
+27 & 99996 & 4 & 99986 & 14 & 13 & 3\tabularnewline
+28 & 100000 & 0 & 99993 & 7 & 7 & 0\tabularnewline
+29 & 100000 & 0 & 99996 & 4 & 4 & 0\tabularnewline
+30 & 100000 & 0 & 99998 & 2 & 2 & 0\tabularnewline
+31 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
-\ldots{} & \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 0 & 0 & 100000\tabularnewline
+\ldots{}\tabularnewline
+50 & 100000 & 0 & 100000 & 0 & 0 & 0\tabularnewline
 \bottomrule
@@
 larger and increase by 10 with increasing rows). In the tables below (as
-\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
-matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
+\protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers to
+matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\), respectively.
 
@@
 \toprule
-beta & S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-complex\_A0 & complex\_A1 & C\tabularnewline
+beta & S & A\_unstable & A\_stable & M\_unstable & M\_stable &
+complex\_A & complex\_M & C\tabularnewline
 \midrule
@@
 \toprule
-S & M & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-complex\_A0 & complex\_A1 & C\tabularnewline
+S & M & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
+complex\_M & C\tabularnewline
 \midrule
@@
 \toprule
-S & A0\_infeasible & A0\_feasible & A1\_infeasible & A1\_feasible &
-A1\_made\_feasible & A1\_made\_infeasible\tabularnewline
+S & A\_infeasible & A\_feasible & M\_infeasible & M\_feasible &
+A\_made\_feasible & A\_made\_infeasible\tabularnewline
 \midrule
@@
 
-Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
-in pure species interaction
+Hence, in general, \(\sigma^{2}_{\gamma}\) does not appear to affect
+feasibility in pure species interaction
 networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.
@@
 Stability results are also shown in the table below. Results for
-\texttt{A0} indicate systems in which \(\gamma = 1\), while \texttt{A1}
+\texttt{A} indicate systems in which \(\gamma = 1\), while \texttt{M}
 refers to systems in which the genetic algorithm searched for a set of

### 2020-03-15 R/Figures_3_4.R
Fix the legend of the genetic algorithm plot

@@ -76,7 +76,7 @@ plot_stable_4 <- function(dat, S_s = 40){
     points(x = x1, y = y1, lwd = 2, type = "l", col = "black");
     axis(side = 4, at = c(0, 0.2, 0.4, 0.6, 0.8, 1.0), cex.axis = 1.5);
     legend("topleft", c(expression(paste(sigma[gamma]^2," = 0")), 
-                        expression(paste(sigma[gamma]^2," = 0"))), 
+                        expression(paste(sigma[gamma]^2," > 0"))), 
            pch=15, col=c("dodgerblue4","firebrick"), cex = 1.5, horiz = TRUE);
     #=================================
     mtext(side = 1, text = "System size (S)", cex = 2, outer = TRUE, 

### 2020-03-20 notebook/SI.Rmd
Insert cascade food web model results into the SI

@@
 ```{r, echo = FALSE, fig.align = "center"}
-sfn <- read.csv(file = "sim_results/sfn/sfn.csv"); # Change when new results are in
+sfn <- read.csv(file = "sim_results/sfn/sfn.csv"); 
 use <- c(1:6, 19:21);
@@
 As in small-world networks, the mean $C$ is shown, along with the mean complexities of $\mathbf{A}$ and $\mathbf{M}$.
-Like all other networks, $\sigma^{2}_\gamma$ increases the stability of scale-free networks given sufficiently complexity, but the complexity requires seems to be relatively low.
+Like all other networks, $\sigma^{2}_\gamma$ increases the stability of scale-free networks given sufficiently high complexity.
+
+Cascade food webs are saturated, and similar to predator-prey random networks.
+What distinguishes them from predator-prey networks is that cascade food webs are also defined by intactions in which components are ranked such that if the rank of $i > j$, then $A_{ij} < 0$ and $A_{ji} > 0$.
+In other words, if interpreting components as ecological species, species can only feed off of a species of lower rank.
+The table below shows how $\sigma^{2}_\gamma$ affects stability across system sizes in cascade food webs.
+
+```{r, echo = FALSE, fig.align = "center"}
+csc <- read.csv(file = "sim_results/ecology/cascade_C_1.csv"); 
+use <- c(1:5, 18:19);
+kable(csc[,use]);
+```
+
+Cascade food webs are more likely to be stable than small-world or scale-free networks at equivalent magnitudes of complexity (note that $C = 1$ for all above rows).
+A higher number of stable $\mathbf{M}$ than $\mathbf{A}$ was found given $S \geq 27$.
 

### 2020-03-20 notebook/SI.tex
Insert cascade food web model results into the SI

@@
 networks, \(\sigma^{2}_\gamma\) increases the stability of scale-free
-networks given sufficiently complexity, but the complexity requires
-seems to be relatively low.
+networks given sufficiently high complexity.
+
+Cascade food webs are saturated, and similar to predator-prey random
+networks. What distinguishes them from predator-prey networks is that
+cascade food webs are also defined by intactions in which components are
+ranked such that if the rank of \(i > j\), then \(A_{ij} < 0\) and
+\(A_{ji} > 0\). In other words, if interpreting components as ecological
+species, species can only feed off of a species of lower rank. The table
+below shows how \(\sigma^{2}_\gamma\) affects stability across system
+sizes in cascade food webs.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
+complex\_M\tabularnewline
+\midrule
+\endhead
+2 & 0 & 1000000 & 0 & 1000000 & 0.6378839 & 0.6381485\tabularnewline
+3 & 1 & 999999 & 1 & 999999 & 0.7055449 & 0.7525143\tabularnewline
+4 & 2 & 999998 & 2 & 999998 & 0.8060500 & 0.8826100\tabularnewline
+5 & 17 & 999983 & 17 & 999983 & 0.8974749 & 0.9967594\tabularnewline
+6 & 42 & 999958 & 43 & 999957 & 0.9821323 & 1.0999762\tabularnewline
+7 & 124 & 999876 & 124 & 999876 & 1.0600906 & 1.1938910\tabularnewline
+8 & 303 & 999697 & 309 & 999691 & 1.1329713 & 1.2807302\tabularnewline
+9 & 653 & 999347 & 661 & 999339 & 1.2009135 & 1.3616372\tabularnewline
+10 & 1401 & 998599 & 1413 & 998587 & 1.2661142 &
+1.4387567\tabularnewline
+11 & 2534 & 997466 & 2566 & 997434 & 1.3276636 &
+1.5113096\tabularnewline
+12 & 4514 & 995486 & 4597 & 995403 & 1.3865754 &
+1.5804005\tabularnewline
+13 & 7570 & 992430 & 7722 & 992278 & 1.4424479 &
+1.6462780\tabularnewline
+14 & 12223 & 987777 & 12502 & 987498 & 1.4970134 &
+1.7102322\tabularnewline
+15 & 18433 & 981567 & 18879 & 981121 & 1.5498812 &
+1.7719564\tabularnewline
+16 & 26973 & 973027 & 27712 & 972288 & 1.6002970 &
+1.8310447\tabularnewline
+17 & 38272 & 961728 & 39499 & 960501 & 1.6494195 &
+1.8884211\tabularnewline
+18 & 52397 & 947603 & 54099 & 945901 & 1.6975099 &
+1.9443860\tabularnewline
+19 & 69986 & 930014 & 72342 & 927658 & 1.7439233 &
+1.9987398\tabularnewline
+20 & 92851 & 907149 & 95776 & 904224 & 1.7893524 &
+2.0514394\tabularnewline
+21 & 117487 & 882513 & 121095 & 878905 & 1.8335974 &
+2.1030121\tabularnewline
+22 & 147852 & 852148 & 151989 & 848011 & 1.8761874 &
+2.1527108\tabularnewline
+23 & 183501 & 816499 & 187888 & 812112 & 1.9186092 &
+2.2019827\tabularnewline
+24 & 222592 & 777408 & 226021 & 773979 & 1.9591518 &
+2.2491948\tabularnewline
+25 & 267691 & 732309 & 269822 & 730178 & 1.9999089 &
+2.2963949\tabularnewline
+26 & 316090 & 683910 & 316371 & 683629 & 2.0396325 &
+2.3427211\tabularnewline
+27 & 369830 & 630170 & 366550 & 633450 & 2.0785319 &
+2.3879356\tabularnewline
+28 & 426407 & 573593 & 419136 & 580864 & 2.1169703 &
+2.4324407\tabularnewline
+29 & 485068 & 514932 & 473666 & 526334 & 2.1545265 &
+2.4759539\tabularnewline
+30 & 544300 & 455700 & 527568 & 472432 & 2.1912376 &
+2.5187795\tabularnewline
+31 & 605803 & 394197 & 584385 & 415615 & 2.2271037 &
+2.5603818\tabularnewline
+32 & 664689 & 335311 & 638047 & 361953 & 2.2626270 &
+2.6016360\tabularnewline
+33 & 718848 & 281152 & 689172 & 310828 & 2.2979241 &
+2.6424881\tabularnewline
+34 & 770790 & 229210 & 737639 & 262361 & 2.3327303 &
+2.6828460\tabularnewline
+35 & 817531 & 182469 & 783112 & 216888 & 2.3666720 &
+2.7221952\tabularnewline
+36 & 858750 & 141250 & 823548 & 176452 & 2.3998286 &
+2.7608037\tabularnewline
+37 & 893017 & 106983 & 859194 & 140806 & 2.4332806 &
+2.7994470\tabularnewline
+38 & 921268 & 78732 & 890177 & 109823 & 2.4658414 &
+2.8372307\tabularnewline
+39 & 943551 & 56449 & 915655 & 84345 & 2.4974678 &
+2.8741350\tabularnewline
+40 & 961088 & 38912 & 936883 & 63117 & 2.5301278 &
+2.9116114\tabularnewline
+41 & 973664 & 26336 & 953645 & 46355 & 2.5616210 &
+2.9481298\tabularnewline
+42 & 982829 & 17171 & 967044 & 32956 & 2.5925309 &
+2.9841081\tabularnewline
+43 & 989464 & 10536 & 977033 & 22967 & 2.6228949 &
+3.0191690\tabularnewline
+44 & 993622 & 6378 & 984470 & 15530 & 2.6534626 &
+3.0548439\tabularnewline
+45 & 996221 & 3779 & 989678 & 10322 & 2.6832092 &
+3.0890543\tabularnewline
+46 & 997963 & 2037 & 993318 & 6682 & 2.7130588 &
+3.1236201\tabularnewline
+47 & 998818 & 1182 & 995957 & 4043 & 2.7423480 &
+3.1575904\tabularnewline
+48 & 999422 & 578 & 997446 & 2554 & 2.7714223 & 3.1912463\tabularnewline
+49 & 999746 & 254 & 998532 & 1468 & 2.7999596 & 3.2244020\tabularnewline
+50 & 999864 & 136 & 999132 & 868 & 2.8285547 & 3.2574510\tabularnewline
+51 & 999934 & 66 & 999561 & 439 & 2.8566907 & 3.2900943\tabularnewline
+52 & 999970 & 30 & 999761 & 239 & 2.8844703 & 3.3222721\tabularnewline
+53 & 999985 & 15 & 999873 & 127 & 2.9122645 & 3.3544290\tabularnewline
+54 & 999999 & 1 & 999935 & 65 & 2.9395400 & 3.3859103\tabularnewline
+55 & 1000000 & 0 & 999971 & 29 & 2.9665996 & 3.4173273\tabularnewline
+56 & 999999 & 1 & 999988 & 12 & 2.9936263 & 3.4486027\tabularnewline
+57 & 1000000 & 0 & 999989 & 11 & 3.0199283 & 3.4789408\tabularnewline
+58 & 1000000 & 0 & 999998 & 2 & 3.0460952 & 3.5094530\tabularnewline
+59 & 1000000 & 0 & 999999 & 1 & 3.0728115 & 3.5401634\tabularnewline
+60 & 1000000 & 0 & 1000000 & 0 & 3.0983367 & 3.5698067\tabularnewline
+\bottomrule
+\end{longtable}
+
+Cascade food webs are more likely to be stable than small-world or
+scale-free networks at equivalent magnitudes of complexity (note that
+\(C = 1\) for all above rows). A higher number of stable \(\mathbf{M}\)
+than \(\mathbf{A}\) was found given \(S \geq 27\).
 
@@
 
-{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-33-1}
+{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-34-1}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-39-1.pdf}
 
@@
 
-\includegraphics{SI_files/figure-latex/unnamed-chunk-43-1.pdf}
+\includegraphics{SI_files/figure-latex/unnamed-chunk-44-1.pdf}
 

### 2020-03-20 notebook/ms.Rmd
Fix the hyphenations of food webs to make things consistent
Add a sentence to the MS to address Reviewer 1's comment 3

@@
 
-Cascade food-webs were constructed following Solow and Beet [@Solow1998].
+Cascade food webs were constructed following Solow and Beet [@Solow1998].
 First, a random matrix $\mathbf{A}$ was generated with off-diagonal elements sampled i.i.d so that $A_{ij} \sim \mathcal{N}(0, 0.4)$.
@@
 Blue and red bars show the number of stable systems in the absence and presence of $\sigma^{2}_{\gamma}$, respectively.
-The black line shows the proportion of systems that are stable when $\sigma^{2}_{\gamma}>0$, but would be unstable if $\sigma^{2}_{\gamma}=0$.
+The black line shows the proportion of systems that are stable when $\sigma^{2}_{\gamma}>0$, but would be unstable if $\sigma^{2}_{\gamma}=0$ (i.e., the conditional probability that $\mathbf{A}$ is unstable given that $\mathbf{M}$ is stable).
 

### 2020-03-20 notebook/sci_rep_reviews.Rmd
Add a sentence to the MS to address Reviewer 1's comment 3

@@
 
-Reviewer 1 is correct that the line illustrates the conditional probability that $A$ is unstable given that $\gamma A$ is stable.
+Reviewer 1 is correct that the line illustrates the conditional probability that $A$ is unstable given that $M = \gamma A$ is stable.
 In other words, the proportion of complex systems found to be stable, but which would have been unstable if it were not for the variation in component response rates.
 
-I have explained the result shown by the black line of Figures 3 and 4 now more clearly in my revision (line XXX).
+I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (line XXX).
 

### 2020-03-20 notebook/SI.Rmd
Adjustments to the notation of the SI to match the main text

@@
 
-**This supplemental information supports the manuscript "Component response rate variation drives stability in large complex systems" with additional analyses to support its conclusions.
+**This supplemental information supports the manuscript "Component response rate variation underlies the stability of complex systems" with additional analyses to support its conclusions.
 All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability R package.**
@@
 - [Stability across increasing $S$](#IncrS) 
-- [Stability of ecological networks](#ecological) 
+- [Stability of random ecological networks](#ecological) 
     - [Competitor networks](#competition) 
@@
 Figure 4 of the main text reports the number of stable random complex systems found over 1 million iterations.
-The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
+The table below shows the results for all simulations of random $\mathbf{M}$ matrices at $\sigma_{A} = 0.4$ and $C = 1$ given a range of $S = \{2, 3, ..., 49, 50\}$.
 In this table, the `A` refers to $\mathbf{A}$ matrices where $\gamma = 1$, while `M` refers to $\mathbf{M}$ matrices after $\sigma^{2}_{\gamma}$ is added and $\gamma \sim \mathcal{U}(0, 2)$.
@@
 
-Stability of ecological networks {#ecological}
+Stability of random ecological networks {#ecological}
 ================================================================================
@@
 
-In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $\mathbf{M}$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2})$ where $\sigma = 0.4$.
+In the main text, for simplicity, I assumed connectance values of $C = 1$, meaning that all off-diagonal elements of a matrix $\mathbf{M}$ were potentially nonzero and sampled from a normal distribution $\mathcal{N}(0, \sigma^{2}_{A})$ where $\sigma_{A} = 0.4$.
 Here I present four tables showing the number of stable communities given $C = \{0.3, 0.
 5, 0.7, 0.9 \}$.
-In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
+In all cases, uniform variation in component response rate ($\gamma_{i} \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma_{i} = 1$).
 In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.
@@
 
-Sensitivity of interaction strength ($\sigma$) values {#sigma}  
+Sensitivity of interaction strength ($\sigma_{A}$) values {#sigma}  
 ================================================================================
 
-Results below show stability results given varying interaction strengths ($\sigma$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows).
+Results below show stability results given varying interaction strengths ($\sigma_{A}$) for $C = 0.05$ (note that system size $S$ values are larger and increase by 10 with increasing rows).
 In the tables below (as [above](#IncrS)), `A` and `M` refers to matrices for $\gamma = 1$ and $\sigma^{2}_{\gamma}$, respectively.
@@
 
-<a name="sigma3">**Interaction strength $\mathbf{\sigma = 0.3}$**</a>
+<a name="sigma3">**Interaction strength $\mathbf{\sigma_{A} = 0.3}$**</a>
 
@@
 
-<a name="sigma4">**Interaction strength $\mathbf{\sigma = 0.4}$**</a>
+<a name="sigma4">**Interaction strength $\mathbf{\sigma_{A} = 0.4}$**</a>
 
@@
 
-<a name="sigma5">**Interaction strength $\mathbf{\sigma = 0.5}$**</a>
+<a name="sigma5">**Interaction strength $\mathbf{\sigma_{A} = 0.5}$**</a>
 
@@
 
-<a name="sigma6">**Interaction strength $\mathbf{\sigma = 0.6}$**</a>
+<a name="sigma6">**Interaction strength $\mathbf{\sigma_{A} = 0.6}$**</a>
 

### 2020-03-20 notebook/SI.tex
Adjustments to the notation of the SI to match the main text

@@
 \textbf{This supplemental information supports the manuscript
-``Component response rate variation drives stability in large complex
+``Component response rate variation underlies the stability of complex
 systems'' with additional analyses to support its conclusions. All text,
@@
 \item
-  \protect\hyperlink{ecological}{Stability of ecological networks}
+  \protect\hyperlink{ecological}{Stability of random ecological
+  networks}
 
@@
 results for all simulations of random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) given a range of
+\(\sigma_{A} = 0.4\) and \(C = 1\) given a range of
 \(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A} refers to
@@
 
-\hypertarget{ecological}{\section{Stability of ecological
+\hypertarget{ecological}{\section{Stability of random ecological
 networks}\label{ecological}}
@@
 \(\mathbf{M}\) were potentially nonzero and sampled from a normal
-distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
-I present four tables showing the number of stable communities given
-\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
-component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
-higher number of stable communities than when \(\gamma\) did not vary
-(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
-million \(\mathbf{M}\) were simulated. As with the results on
+distribution \(\mathcal{N}(0, \sigma^{2}_{A})\) where
+\(\sigma_{A} = 0.4\). Here I present four tables showing the number of
+stable communities given \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases,
+uniform variation in component response rate
+(\(\gamma_{i} \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
+communities than when \(\gamma\) did not vary (\(\gamma_{i} = 1\)). In
+contrast to the main text, 100000 rather than 1 million \(\mathbf{M}\)
+were simulated. As with the results on
 \protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
@@
 \hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
-strength (\(\sigma\))
-values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
+strength (\(\sigma_{A}\))
+values}{Sensitivity of interaction strength (\textbackslash{}sigma\_\{A\}) values}}\label{sigma}}
 
 Results below show stability results given varying interaction strengths
-(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
-larger and increase by 10 with increasing rows). In the tables below (as
-\protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers to
-matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\), respectively.
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
+(\(\sigma_{A}\)) for \(C = 0.05\) (note that system size \(S\) values
+are larger and increase by 10 with increasing rows). In the tables below
+(as \protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers
+to matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\),
+respectively.
+
+\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.3}\)}
 
@@
 
-\textbf{Interaction strength \(\mathbf{\sigma = 0.4}\)}
+\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.4}\)}
 
@@
 
-\textbf{Interaction strength \(\mathbf{\sigma = 0.5}\)}
+\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.5}\)}
 
@@
 
-\textbf{Interaction strength \(\mathbf{\sigma = 0.6}\)}
+\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.6}\)}
 

### 2020-03-20 R/utilities.R
Add the make_gammas function to the utilities so that it is in the R package. Plan to redo this part of the SI to show more

@@ -155,7 +155,48 @@ add_C_stats <- function(sim){
     return(new_all_res);
 }
 
-
+#' Produce a vector (gamma) from one of four distributions
+#' 
+#' Returns  a vector (gamma) from one of four distributions.
+#'
+#'@return A vector sampled from one of four distributions
+#'@param nn The number of elements in the vector
+#'@param distribution The distribution from which elements will be sampled, 
+#'including (0) a repeated value with no variance, (1) a uniform distribution
+#'of min 0 and max 1, (2) an exponential distribution, (3) a beta distribution
+#'with parameters alpha = 0.5 and beta = 0.5, or (4) a gamma distribution with
+#'shape and scale parameters equal to k = 2 and theta = 2.
+#'@param mn The mean value of the distribution (only for distribution 0)
+#'@param sdd The standard deviatoin of the distribution (only for distributions
+#'2-4).
+#'@examples
+#'eg_gams  <- make_gammas(n = 20, distribution = 3, sdd = 1);
+#'@export
+make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
+    if(distribution == 0){
+        dat          <- rep(x = mn, times = nn);
+    }
+    if(distribution == 1){
+        mval         <- 2;
+        dat          <- runif(n = nn, min = 0, max = mval);
+    }
+    if(distribution == 2){
+        dat          <- rexp(n = nn);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 3){
+        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    if(distribution == 4){
+        dat          <- rgamma(n = nn, shape = 2, scale = 2);
+        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
+        dat          <- dat - min(dat) + min(abs(dat));
+    }
+    return(dat);
+}
 
 #' Get variance of interaction strengths
 #' 

### 2020-03-20 notebook/ms.Rmd
Make the gamma diagonal matrix bold in the text

@@
 **Component response rates of random complex systems**.
-Complex systems ($\mathbf{M}$) are built from two matrices, one modelling component interactions ($\mathbf{A}$), and second modelling component response rates ($\gamma$).
-Both $\mathbf{A}$ and $\gamma$ are square $S \times S$ matrices.
+Complex systems ($\mathbf{M}$) are built from two matrices, one modelling component interactions ($\mathbf{A}$), and second modelling component response rates ($\boldsymbol{\gamma}$).
+Both $\mathbf{A}$ and $\boldsymbol{\gamma}$ are square $S \times S$ matrices.
 Rows in $\mathbf{A}$ define how a given component $i$ is affected by each component $j$ in the system, including itself (where $i = j$).
 Off-diagonal elements of $\mathbf{A}$ are independent and identically distributed (i.i.d), and diagonal elements are set to $A_{ii} = -1$ as in May [@May1972].
-Diagonal elements of $\gamma$ are positive, and off-diagonal elements are set to zero (i.e.
-$\gamma$ is a diagonal matrix with positive support).
-The distribution of $diag(\gamma)$ over $S$ components thereby models the distribution of component response rates.
+Diagonal elements of $\boldsymbol{\gamma}$ are positive, and off-diagonal elements are set to zero (i.e.
+$\boldsymbol{\gamma}$ is a diagonal matrix with positive support).
+The distribution of $diag(\boldsymbol{\gamma})$ over $S$ components thereby models the distribution of component response rates.
 The dynamics of the entire system $\mathbf{M}$ can be defined as follows [@Patel2018],
@@
 \begin{equation} \label{defM}
-M = \gamma A.
+\mathbf{M} = \boldsymbol{\gamma} \mathbf{A}.
 \end{equation}
@@
 Equation \ref{defM} thereby serves as a null model to investigate how variation in component response rate ($\sigma^{2}_{\gamma}$) affects complex systems.
-In the absence of such variation ($\sigma^{2}_{\gamma} = 0$), $\gamma$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
+In the absence of such variation ($\sigma^{2}_{\gamma} = 0$), $\boldsymbol{\gamma}$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
 Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma \sqrt{SC}$ [@May1972] (Fig.
@@
 
-The increase in $\sigma^{2}_{M}$ caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \gamma\textbf{A}$ (Fig.
+The increase in $\sigma^{2}_{M}$ caused by $\sigma^{2}_\gamma$ can be visualised from the eigenvalue spectra of $\textbf{A}$ versus $\textbf{M} = \boldsymbol{\gamma}\textbf{A}$ (Fig.
 1).
@@
 If $\rho < 0$, such as when $\textbf{M}$ models a predator-prey system in which $M_{ij}$ and $M_{ji}$ have opposing signs, stability increases [@Allesina2012].
-If diagonal elements of $\gamma$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$ (Figure 2).
+If diagonal elements of $\boldsymbol{\gamma}$ vary independently, the magnitude of $\rho$ is decreased because $\sigma^{2}_{\gamma}$ increases the variance of $M_{ij}$ without affecting the expected covariance between $M_{ij}$ and $M_{ji}$ (Figure 2).
 
@@
 My focus here is distinct from Gibbs et al.
-[@Gibbs2017], who applied the same mathematical framework to investigate how a diagonal matrix $\mathbf{X}$ (equivalent to $\gamma$ in my model) affects the stability of a community matrix $\mathbf{M}$ given an interaction matrix $\mathbf{A}$ within a generalised Lotka-Volterra model, where $\mathbf{M} = \mathbf{XA}$.
+[@Gibbs2017], who applied the same mathematical framework to investigate how a diagonal matrix $\mathbf{X}$ (equivalent to $\boldsymbol{\gamma}$ in my model) affects the stability of a community matrix $\mathbf{M}$ given an interaction matrix $\mathbf{A}$ within a generalised Lotka-Volterra model, where $\mathbf{M} = \mathbf{XA}$.
 Gibbs et al.
@@
 
-A general mathematical framework encompassing shifts in eigenvalue distributions caused by a diagona matrix $\gamma$ has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but $\gamma$ has not been interpreted as rates of response of individual system components to perturbation.
+A general mathematical framework encompassing shifts in eigenvalue distributions caused by a diagonal matrix $\boldsymbol{\gamma}$ has been investigated [@Ahmadian2015] and recently applied to questions concerning species density and feasibility [@Gibbs2017; @Stone2017], but $\boldsymbol{\gamma}$ has not been interpreted as rates of response of individual system components to perturbation.
 My model focuses on component response rates for systems of a finite size, in which complexity is high but not yet high enough to make the probability of stability unrealistically low for actual empirical systems.
@@
 Diagonal elements $A_{ii}$ were set to $-1$.
-Elements of $\gamma$ were simulated i.i.d.
+Elements of $\boldsymbol{\gamma}$ were simulated i.i.d.
 from a distribution with positive support (typically $\gamma_{i} \sim \mathcal{U}(0, 2)$).

### 2020-03-20 notebook/ms.Rmd
Fix the notation in several places, particularly around the diagonal matrix gamma versus the vector diag(gamma)

@@
 --->
-**Targeted manipulation of $\mathbf{\gamma}$**.
+**Targeted manipulation of $\boldsymbol{\gamma}$**.
 To further investigate the potential of $\sigma^{2}_{\gamma}$ to be stabilising, I used a genetic algorithm.
@@
 
-**Component response rate ($\mathbf{\gamma}$) variation**.
+**Component response rate ($\boldsymbol{\gamma}$) variation**.
 In a synthesis of eco-evolutionary feedbacks on community stability, Patel et al.
@@
 
-$$\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).$$
-
-In the above, $\mathbf{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
-Therefore, $\mathbf{M} = \mathbf{\gamma} \mathbf{A}$ defines the change in values of system components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
+$$\frac{d\mathbf{v}(t)}{dt} = \boldsymbol{\gamma} \mathbf{A}\mathbf{v}(t).$$
+
+In the above, $\boldsymbol{\gamma}$ is a diagonal matrix in which elements correspond to individual component response rates.
+Therefore, $\mathbf{M} = \boldsymbol{\gamma} \mathbf{A}$ defines the change in values of system components and can be analysed using the techniques of May [@May1972; @May1973; @Ahmadian2015].
 In these analyses, row means of $\mathbf{A}$ are expected to be identical, but variation around this expectation will naturally arise due to random sampling of $\mathbf{A}$ off-diagonal elements and finite $S$.
-In simulations, the total variation in $\mathbf{M}$ row means that is attributable to $\mathbf{A}$ is small relative to that attributable to $\mathbf{\gamma}$, especially at high $S$.
-Variation in $\mathbf{\gamma}$ specifically isolates the effects of differing component response rates, hence causing differences in expected $\mathbf{M}$ row means.
+In simulations, the total variation in $\mathbf{M}$ row means that is attributable to $\mathbf{A}$ is small relative to that attributable to $\boldsymbol{\gamma}$, especially at high $S$.
+Variation in $\boldsymbol{\gamma}$ specifically isolates the effects of differing component response rates, hence causing differences in expected $\mathbf{M}$ row means.
 
@@
 **Genetic algorithm**.
-Ideally, to investigate the potential of $\sigma^{2}_{\gamma}$ for increasing the proportion of stable complex systems, the search space of all possible $\gamma$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
+Ideally, to investigate the potential of $\sigma^{2}_{\gamma}$ for increasing the proportion of stable complex systems, the search space of all possible $diag(\boldsymbol{\gamma})$ vectors would be evaluated for each unique $\mathbf{M = \gamma A}$.
 This is technically impossible because $\gamma_{i}$ can take any real value between 0-2, but even rounding $\gamma_{i}$ to reasonable values would result in a search space too large to practically explore.
 Under these conditions, genetic algorithms are highly useful tools for finding practical solutions by mimicking the process of biological evolution [@Hamblin2013].
-In this case, the practical solution is finding vectors of $\mathbf{\gamma}$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
-The genetic algorithm used achieves this by initialising a large population of 1000 different potential $\mathbf{\gamma}$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $\mathbf{\gamma}$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
+In this case, the practical solution is finding vectors of $diag(\boldsymbol{\gamma})$ that decrease the most positive real eigenvalue of $\mathbf{M}$.
+The genetic algorithm used achieves this by initialising a large population of 1000 different potential $diag(\boldsymbol{\gamma})$ vectors and allowing this population to evolve through a process of mutation, crossover (swaping $\gamma_{i}$ values between vectors), selection, and reproduction until either a $diag(\boldsymbol{\gamma})$ vector is found where all $\Re(\lambda) < 0$ or some "giving up" critiera is met.
 
 For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M = \gamma A}$ ($\sigma_{A} = 0.4$, $C = 1$).
-The genetic algorithm was initialised with a population of 1000 different $\mathbf{\gamma}$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
-Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\mathbf{\gamma}$ vector, and the 20 $\mathbf{\gamma}$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $\mathbf{\gamma}$ vectors.
-Mutation of each $\gamma_{i}$ in a $\mathbf{\gamma}$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
-Crossover occurred between two sets of 100 $\mathbf{\gamma}$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
+The genetic algorithm was initialised with a population of 1000 different $diag(\boldsymbol{\gamma})$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\boldsymbol{\gamma}$, and the 20 $diag(\boldsymbol{\gamma})$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $diag(\boldsymbol{\gamma})$ vectors.
+Mutation of each $\gamma_{i}$ in a $diag(\boldsymbol{\gamma})$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+Crossover occurred between two sets of 100 $diag(\boldsymbol{\gamma})$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
 Vector pairs selected for crossover swapped all elements between and including two $\gamma_{i}$ randomly selected with replacement (this allowed for reversal of vector element positions during crossover; e.g., $\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}$ ).
-The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\mathbf{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\Re(\lambda_{max})$ for $\mathbf{M}$).
+The genetic algorithm terminated if a stable $\mathbf{M}$ was found, 20 generations occurred, or if the mean $\boldsymbol{\gamma}$ fitness increase between generations was less than 0.01 (where fitness was defined as $W_{\gamma} = -\Re(\lambda_{max})$ for $\mathbf{M}$).
 

### 2020-03-20 notebook/ms.tex
Fix the notation in several places, particularly around the diagonal matrix gamma versus the vector diag(gamma)

@@
 component interactions (\(\mathbf{A}\)), and second modelling component
-response rates (\(\gamma\)). Both \(\mathbf{A}\) and \(\gamma\) are
-square \(S \times S\) matrices. Rows in \(\mathbf{A}\) define how a
-given component \(i\) is affected by each component \(j\) in the system,
-including itself (where \(i = j\)). Off-diagonal elements of
-\(\mathbf{A}\) are independent and identically distributed (i.i.d), and
-diagonal elements are set to \(A_{ii} = -1\) as in
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}. Diagonal
-elements of \(\gamma\) are positive, and off-diagonal elements are set
-to zero (i.e. \(\gamma\) is a diagonal matrix with positive support).
-The distribution of \(diag(\gamma)\) over \(S\) components thereby
-models the distribution of component response rates. The dynamics of the
-entire system \(\mathbf{M}\) can be defined as
+response rates (\(\boldsymbol{\gamma}\)). Both \(\mathbf{A}\) and
+\(\boldsymbol{\gamma}\) are square \(S \times S\) matrices. Rows in
+\(\mathbf{A}\) define how a given component \(i\) is affected by each
+component \(j\) in the system, including itself (where \(i = j\)).
+Off-diagonal elements of \(\mathbf{A}\) are independent and identically
+distributed (i.i.d), and diagonal elements are set to \(A_{ii} = -1\) as
+in May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}. Diagonal
+elements of \(\boldsymbol{\gamma}\) are positive, and off-diagonal
+elements are set to zero (i.e. \(\boldsymbol{\gamma}\) is a diagonal
+matrix with positive support). The distribution of
+\(diag(\boldsymbol{\gamma})\) over \(S\) components thereby models the
+distribution of component response rates. The dynamics of the entire
+system \(\mathbf{M}\) can be defined as
 follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}},
@@
 complex systems. In the absence of such variation
-(\(\sigma^{2}_{\gamma} = 0\)), \(\gamma\) is set to the identity matrix
-(diagonal elements all equal 1) and \(\mathbf{M} = \mathbf{A}\). Under
-these conditions, eigenvalues of \(\mathbf{M}\) are distributed
+(\(\sigma^{2}_{\gamma} = 0\)), \(\boldsymbol{\gamma}\) is set to the
+identity matrix (diagonal elements all equal 1) and
+\(\mathbf{M} = \mathbf{A}\). Under these conditions, eigenvalues of
+\(\mathbf{M}\) are distributed
 uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} in a
@@
 be visualised from the eigenvalue spectra of \(\textbf{A}\) versus
-\(\textbf{M} = \gamma\textbf{A}\) (Fig. 1). Given \(d = 0\) and
-\(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
+\(\textbf{M} = \boldsymbol{\gamma}\textbf{A}\) (Fig. 1). Given \(d = 0\)
+and \(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
 \(\textbf{M}\) lie within a circle of a radius \(\sigma_{A}\sqrt{S}\)
@@
 increases\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. If
-diagonal elements of \(\gamma\) vary independently, the magnitude of
-\(\rho\) is decreased because \(\sigma^{2}_{\gamma}\) increases the
-variance of \(M_{ij}\) without affecting the expected covariance between
-\(M_{ij}\) and \(M_{ji}\) (Figure 2).
+diagonal elements of \(\boldsymbol{\gamma}\) vary independently, the
+magnitude of \(\rho\) is decreased because \(\sigma^{2}_{\gamma}\)
+increases the variance of \(M_{ij}\) without affecting the expected
+covariance between \(M_{ij}\) and \(M_{ji}\) (Figure 2).
 
@@
 
-\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
+\textbf{Targeted manipulation of \(\boldsymbol{\gamma}\)}. To further
 investigate the potential of \(\sigma^{2}_{\gamma}\) to be stabilising,
@@
 the same mathematical framework to investigate how a diagonal matrix
-\(\mathbf{X}\) (equivalent to \(\gamma\) in my model) affects the
-stability of a community matrix \(\mathbf{M}\) given an interaction
-matrix \(\mathbf{A}\) within a generalised Lotka-Volterra model, where
-\(\mathbf{M} = \mathbf{XA}\). Gibbs et
+\(\mathbf{X}\) (equivalent to \(\boldsymbol{\gamma}\) in my model)
+affects the stability of a community matrix \(\mathbf{M}\) given an
+interaction matrix \(\mathbf{A}\) within a generalised Lotka-Volterra
+model, where \(\mathbf{M} = \mathbf{XA}\). Gibbs et
 al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}} analytically
@@
 A general mathematical framework encompassing shifts in eigenvalue
-distributions caused by a diagona matrix \(\gamma\) has been
+distributions caused by a diagonal matrix \(\boldsymbol{\gamma}\) has
+been
 investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}}
@@
 feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22},\protect\hyperlink{ref-Stone2017}{23}},
-but \(\gamma\) has not been interpreted as rates of response of
-individual system components to perturbation. My model focuses on
-component response rates for systems of a finite size, in which
-complexity is high but not yet high enough to make the probability of
-stability unrealistically low for actual empirical systems. For this
+but \(\boldsymbol{\gamma}\) has not been interpreted as rates of
+response of individual system components to perturbation. My model
+focuses on component response rates for systems of a finite size, in
+which complexity is high but not yet high enough to make the probability
+of stability unrealistically low for actual empirical systems. For this
 upper range of system size, randomly assembled complex systems are more
@@
 
-\textbf{Component response rate (\(\mathbf{\gamma}\)) variation}. In a
-synthesis of eco-evolutionary feedbacks on community stability, Patel et
-al. model a system that includes a vector of potentially changing
+\textbf{Component response rate (\(\boldsymbol{\gamma}\)) variation}. In
+a synthesis of eco-evolutionary feedbacks on community stability, Patel
+et al. model a system that includes a vector of potentially changing
 species densities (\(\mathbf{n}\)) and a vector of potentially evolving
@@
 
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the change in values
-of system components and can be analysed using the techniques of
+\[\frac{d\mathbf{v}(t)}{dt} = \boldsymbol{\gamma} \mathbf{A}\mathbf{v}(t).\]
+
+In the above, \(\boldsymbol{\gamma}\) is a diagonal matrix in which
+elements correspond to individual component response rates. Therefore,
+\(\mathbf{M} = \boldsymbol{\gamma} \mathbf{A}\) defines the change in
+values of system components and can be analysed using the techniques of
 May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{21},\protect\hyperlink{ref-May1973}{30}}.
@@
 means that is attributable to \(\mathbf{A}\) is small relative to that
-attributable to \(\mathbf{\gamma}\), especially at high \(S\). Variation
-in \(\mathbf{\gamma}\) specifically isolates the effects of differing
-component response rates, hence causing differences in expected
-\(\mathbf{M}\) row means.
+attributable to \(\boldsymbol{\gamma}\), especially at high \(S\).
+Variation in \(\boldsymbol{\gamma}\) specifically isolates the effects
+of differing component response rates, hence causing differences in
+expected \(\mathbf{M}\) row means.
 
@@
 (unsampled elements were set to zero). Diagonal elements \(A_{ii}\) were
-set to \(-1\). Elements of \(\gamma\) were simulated i.i.d. from a
-distribution with positive support (typically
+set to \(-1\). Elements of \(\boldsymbol{\gamma}\) were simulated i.i.d.
+from a distribution with positive support (typically
 \(\gamma_{i} \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices
@@
 
-Cascade food-webs were constructed following Solow and
+Cascade food webs were constructed following Solow and
 Beet\textsuperscript{\protect\hyperlink{ref-Solow1998}{17}}. First, a
@@
 \(\sigma^{2}_{\gamma}\) for increasing the proportion of stable complex
-systems, the search space of all possible \(\gamma\) vectors would be
-evaluated for each unique \(\mathbf{M = \gamma A}\). This is technically
-impossible because \(\gamma_{i}\) can take any real value between 0-2,
-but even rounding \(\gamma_{i}\) to reasonable values would result in a
-search space too large to practically explore. Under these conditions,
-genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological
+systems, the search space of all possible \(diag(\boldsymbol{\gamma})\)
+vectors would be evaluated for each unique \(\mathbf{M = \gamma A}\).
+This is technically impossible because \(\gamma_{i}\) can take any real
+value between 0-2, but even rounding \(\gamma_{i}\) to reasonable values
+would result in a search space too large to practically explore. Under
+these conditions, genetic algorithms are highly useful tools for finding
+practical solutions by mimicking the process of biological
 evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}. In
 this case, the practical solution is finding vectors of
-\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
-\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
-a large population of 1000 different potential \(\mathbf{\gamma}\)
-vectors and allowing this population to evolve through a process of
-mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
-selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
-found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
-met.
+\(diag(\boldsymbol{\gamma})\) that decrease the most positive real
+eigenvalue of \(\mathbf{M}\). The genetic algorithm used achieves this
+by initialising a large population of 1000 different potential
+\(diag(\boldsymbol{\gamma})\) vectors and allowing this population to
+evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
+values between vectors), selection, and reproduction until either a
+\(diag(\boldsymbol{\gamma})\) vector is found where all
+\(\Re(\lambda) < 0\) or some ``giving up'' critiera is met.
 
@@
 \(C = 1\)). The genetic algorithm was initialised with a population of
-1000 different \(\mathbf{\gamma}\) vectors with elements sampled i.i.d
-from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed
-on the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector,
-and the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with
-the lowest \(\Re(\lambda_{max})\) each produced 50 clonal offspring with
-subsequent random mutation and crossover between the resulting new
-generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation of each
-\(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
+1000 different \(diag(\boldsymbol{\gamma})\) vectors with elements
+sampled i.i.d from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis
+was performed on the \(\mathbf{M}\) resulting from each
+\(\boldsymbol{\gamma}\), and the 20 \(diag(\boldsymbol{\gamma})\)
+vectors resulting in \(\mathbf{M}\) with the lowest
+\(\Re(\lambda_{max})\) each produced 50 clonal offspring with subsequent
+random mutation and crossover between the resulting new generation of
+1000 \(diag(\boldsymbol{\gamma})\) vectors. Mutation of each
+\(\gamma_{i}\) in a \(diag(\boldsymbol{\gamma})\) vector occurred with a
 probability of 0.2, resulting in a mutation effect of size
@@
 multiplied by \(-1\), and any values that mutated above 2 were set to
-2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
-vectors paired in each generation; vectors were randomly sampled with
-replacement among but not within sets. Vector pairs selected for
-crossover swapped all elements between and including two \(\gamma_{i}\)
-randomly selected with replacement (this allowed for reversal of vector
-element positions during crossover; e.g.,
+2). Crossover occurred between two sets of 100
+\(diag(\boldsymbol{\gamma})\) vectors paired in each generation; vectors
+were randomly sampled with replacement among but not within sets. Vector
+pairs selected for crossover swapped all elements between and including
+two \(\gamma_{i}\) randomly selected with replacement (this allowed for
+reversal of vector element positions during crossover; e.g.,
 \(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
 ). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
-found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
+found, 20 generations occurred, or if the mean \(\boldsymbol{\gamma}\)
 fitness increase between generations was less than 0.01 (where fitness
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.11.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.13.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 proportion of systems that are stable when \(\sigma^{2}_{\gamma}>0\),
-but would be unstable if \(\sigma^{2}_{\gamma}=0\).
+but would be unstable if \(\sigma^{2}_{\gamma}=0\) (i.e., the
+conditional probability that \(\mathbf{A}\) is unstable given that
+\(\mathbf{M}\) is stable).
 

### 2020-03-20 notebook/SI.Rmd
Edits to the notation and a few small corrections in the SI

@@
 5, 0.7, 0.9 \}$.
-In all cases, uniform variation in component response rate ($\gamma_{i} \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma_{i} = 1$).
+In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma_{i} = 1$).
 In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.
@@
 
-I tested the stability of one million random, small-world, scale-free, and cascade food web networks.
+I tested the stability of one million random, small-world, scale-free, and cascade food web networks for different network parameters.
 Each of these networks is structured differently.
-In the main text, the random networks and cascade food webs that I built were saturated ($C = 1$), meaning that every component was connected to, and interacted with, every other component.
+In the main text, the random networks and cascade food webs that I built were saturated ($C = 1$), meaning that every component was connected to, and interacted with, every other component (see immediately below).
 
@@
 Small-world networks, in contrast, are not saturated.
-They are instead defined by components that interact mostly with other closely neighbouring components, but have a proportion of interactions ($\beta$) that are instead between non-neighbours.
+They are instead defined by components that interact mostly with other closely neighbouring components, but have a proportion of interactions ($\beta$) that are instead between non-neighbours [@Watts1998].
 Two small-world networks are shown below.
@@
 
-Scale-free networks are also not saturated, but are defined by an interaction number distribution that follows a power law.
+Scale-free networks are also not saturated, but are defined by an interaction frequency distribution that follows a power law.
 In other words, a small number of components interact with many other components, while most components interact with only a small number of other components.
-Scale-free networks can be built by adding new components, one by one, to an existing system, with each newly added component interacting with a randomly selected subset of $m$ existing components.
+Scale-free networks can be built by adding new components, one by one, to an existing system, with each newly added component interacting with a randomly selected subset of $m$ existing components [@Albert2002].
 The network on the left below shows an example of a scale-free network in which $m = 3$.
@@
 Cascade food webs are saturated, and similar to predator-prey random networks.
-What distinguishes them from predator-prey networks is that cascade food webs are also defined by intactions in which components are ranked such that if the rank of $i > j$, then $A_{ij} < 0$ and $A_{ji} > 0$.
+What distinguishes them from predator-prey networks is that cascade food webs are also defined by intactions in which components are ranked such that if the rank of $i > j$, then $A_{ij} < 0$ and $A_{ji} > 0$ [@Solow1998; @Williams2000].
 In other words, if interpreting components as ecological species, species can only feed off of a species of lower rank.
@@
 
-The figure below compares the stability of large complex systems given $\boldsymbol{\gamma = 1}$ versus targeted manipulation of $\gamma$ elements.
+The figure below compares the stability of large complex systems given $\gamma = 1$ versus targeted manipulation of $\gamma$ elements.
 For each $S$, 100000 complex systems are randomly generated.
@@
 
-Distributions of $\gamma$ values in vectors for the highest values of $S$ are shown below.
-
 ```{r, fig.height = 4, fig.width = 6, echo = FALSE}
@@
 
-Mathematically, the above is identical to my model in the main text where the system $\mathbf{M}$ is defined by component interaction strengths $\mathbf{A}$ and individual component response rates $\mathbf{\gamma}$,
+Mathematically, the above is identical to my model in the main text where the system $\mathbf{M}$ is defined by component interaction strengths $\mathbf{A}$ and individual component response rates $\boldsymbol{\gamma}$,
 

### 2020-03-20 notebook/SI.tex
Edits to the notation and a few small corrections in the SI

@@
 uniform variation in component response rate
-(\(\gamma_{i} \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
+(\(\gamma \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
 communities than when \(\gamma\) did not vary (\(\gamma_{i} = 1\)). In
@@
 I tested the stability of one million random, small-world, scale-free,
-and cascade food web networks. Each of these networks is structured
-differently. In the main text, the random networks and cascade food webs
-that I built were saturated (\(C = 1\)), meaning that every component
-was connected to, and interacted with, every other component.
+and cascade food web networks for different network parameters. Each of
+these networks is structured differently. In the main text, the random
+networks and cascade food webs that I built were saturated (\(C = 1\)),
+meaning that every component was connected to, and interacted with,
+every other component (see immediately below).
 
@@
 neighbouring components, but have a proportion of interactions
-(\(\beta\)) that are instead between non-neighbours. Two small-world
-networks are shown below.
+(\(\beta\)) that are instead between
+non-neighbours\textsuperscript{\protect\hyperlink{ref-Watts1998}{4}}.
+Two small-world networks are shown below.
 
@@
 Scale-free networks are also not saturated, but are defined by an
-interaction number distribution that follows a power law. In other
+interaction frequency distribution that follows a power law. In other
 words, a small number of components interact with many other components,
@@
 interacting with a randomly selected subset of \(m\) existing
-components. The network on the left below shows an example of a
-scale-free network in which \(m = 3\). The histogram on the right shows
-the number of other components with which each component interacts.
+components\textsuperscript{\protect\hyperlink{ref-Albert2002}{5}}. The
+network on the left below shows an example of a scale-free network in
+which \(m = 3\). The histogram on the right shows the number of other
+components with which each component interacts.
 
@@
 ranked such that if the rank of \(i > j\), then \(A_{ij} < 0\) and
-\(A_{ji} > 0\). In other words, if interpreting components as ecological
-species, species can only feed off of a species of lower rank. The table
-below shows how \(\sigma^{2}_\gamma\) affects stability across system
-sizes in cascade food webs.
+\(A_{ji} > 0\)\textsuperscript{\protect\hyperlink{ref-Solow1998}{6},\protect\hyperlink{ref-Williams2000}{7}}.
+In other words, if interpreting components as ecological species,
+species can only feed off of a species of lower rank. The table below
+shows how \(\sigma^{2}_\gamma\) affects stability across system sizes in
+cascade food webs.
 
@@
 feasibility in pure species interaction
-networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.
+networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{8}}.
 
@@
 The figure below compares the stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted manipulation of \(\gamma\)
-elements. For each \(S\), 100000 complex systems are randomly generated.
-Stability of each complex system is tested given variation in \(\gamma\)
-using a genetic algorithm to maximise the effect of \(\gamma\) values on
+\(\gamma = 1\) versus targeted manipulation of \(\gamma\) elements. For
+each \(S\), 100000 complex systems are randomly generated. Stability of
+each complex system is tested given variation in \(\gamma\) using a
+genetic algorithm to maximise the effect of \(\gamma\) values on
 increasing stability, as compared to stability in an otherwise identical
@@
 
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
 \includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}
@@
 The question that I address in the main text is distinct from that of
-Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, who
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}}, who
 focused instead on the effect of a diagonal matrix of biological species
@@
 strengths \(\mathbf{A}\) and individual component response rates
-\(\mathbf{\gamma}\),
+\(\boldsymbol{\gamma}\),
 
@@
 to obtain numerical results. In contrast, Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} applied
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}} applied
 analytical techniques to instead focus on a different question
@@
 scaled so that \(\sigma = 1/\sqrt{S}\). Under such scaling, Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}} showed that
 the effect of \(\gamma\) on stability should decrease exponentially as
@@
 million), simulations are run to confirm consistency with Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. The
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}}. The
 difference between my model and Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} is that in the
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}} is that in the
 latter, \(\sigma\sqrt{SC} = 1\) remains constant with increasing \(S\).
@@
 be stabilised by \(\gamma\) as \(S\) increases, consistent with Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}. Note that
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}}. Note that
 overall stability does decrease with increasing \(S\) due to the
@@
 versus in Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} (dashed).}
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}} (dashed).}
 
@@
 this circle becomes increasingly saturated. Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}} showed that a
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}} showed that a
 diagonal matrix \(\mathbf{\gamma}\) will have an exponentially
@@
 
+\hypertarget{ref-Watts1998}{}
+4. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
+networks. \emph{Nature} \textbf{393,} 440--442 (1998).
+
+\hypertarget{ref-Albert2002}{}
+5. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
+networks. \emph{Reviews of Modern Physics} \textbf{74,} 47--97 (2002).
+
+\hypertarget{ref-Solow1998}{}
+6. Solow, A. R. \& Beet, A. R. On lumping species in food webs.
+\emph{Ecology} \textbf{79,} 2013--2018 (1998).
+
+\hypertarget{ref-Williams2000}{}
+7. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
+webs. \emph{Nature} \textbf{404,} 180--183 (2000).
+
 \hypertarget{ref-Servan2018}{}
-4. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+8. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 \hypertarget{ref-Gibbs2017}{}
-5. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+9. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.

### 2020-03-21 R/utilities.R
Remove the duplicate gammas function. This is already in its own file in R

@@ -155,49 +155,6 @@ add_C_stats <- function(sim){
     return(new_all_res);
 }
 
-#' Produce a vector (gamma) from one of four distributions
-#' 
-#' Returns  a vector (gamma) from one of four distributions.
-#'
-#'@return A vector sampled from one of four distributions
-#'@param nn The number of elements in the vector
-#'@param distribution The distribution from which elements will be sampled, 
-#'including (0) a repeated value with no variance, (1) a uniform distribution
-#'of min 0 and max 1, (2) an exponential distribution, (3) a beta distribution
-#'with parameters alpha = 0.5 and beta = 0.5, or (4) a gamma distribution with
-#'shape and scale parameters equal to k = 2 and theta = 2.
-#'@param mn The mean value of the distribution (only for distribution 0)
-#'@param sdd The standard deviatoin of the distribution (only for distributions
-#'2-4).
-#'@examples
-#'eg_gams  <- make_gammas(n = 20, distribution = 3, sdd = 1);
-#'@export
-make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
-    if(distribution == 0){
-        dat          <- rep(x = mn, times = nn);
-    }
-    if(distribution == 1){
-        mval         <- 2;
-        dat          <- runif(n = nn, min = 0, max = mval);
-    }
-    if(distribution == 2){
-        dat          <- rexp(n = nn);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 3){
-        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 4){
-        dat          <- rgamma(n = nn, shape = 2, scale = 2);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    return(dat);
-}
-
 #' Get variance of interaction strengths
 #' 
 #' Returns the variance of interaction stengths (sigma^2).

### 2020-03-21 R/random_mats.R
Set the initial S in sims so that 1 is not allowed

@@ -29,7 +29,11 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
     fea_res <- NULL;
     rho_res <- NULL;
     cmplxty <- NULL;
-    sp_try  <- seq(from = by, to = max_sp, by = by);
+    ini_S   <- 2;
+    if(by > ini_S){
+        ini_S <- by;
+    }
+    sp_try  <- seq(from = ini_S, to = max_sp, by = by);
     for(i in 1:length(sp_try)){
         iter           <- iters;
         tot_res[[i]]   <- matrix(data = 0, nrow = iter, ncol = 7);
@@ -184,12 +188,12 @@ summarise_randmat <- function(tot_res, fea_res, rho_res, cmplxty){
         all_res[i, 22:23] <- range(rho_res[[i]][,5]);
         all_res[i, 24:25] <- range(rho_res[[i]][,6]);
     }
-    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
-                "A1_stabilised", "A1_destabilised", "A0_infeasible", 
-                "A0_feasible", "A1_infeasible", "A1_feasible", 
-                "A1_made_feasible", "A1_made_infeasible", "A0_rho", "A1_rho",
-                "rho_diff", "rho_abs", "complex_A0", "complex_A1", "A0_eig",
-                "A1_eig", "LCI_A0", "UCI_A0", "LCI_A1", "UCI_A1");
+    cnames <- c("S", "A_unstable", "A_stable", "M_unstable", "M_stable", 
+                "A_stabilised", "A_destabilised", "A_infeasible", 
+                "A_feasible", "M_infeasible", "M_feasible", 
+                "A_made_feasible", "A_made_infeasible", "A_rho", "M_rho",
+                "rho_diff", "rho_abs", "complex_A", "complex_M", "A_eig",
+                "M_eig", "LR_A", "UR_A", "LR_M", "UR_M");
     colnames(all_res) <- cnames;
     return(all_res);
 }

### 2020-03-21 R/ga.R
Change the genetic algorithm summary function so that it is specific to GA simulations

@@ -62,7 +62,7 @@ Evo_rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1){
         }
         print(i);
     }
-    all_res <- summarise_randmat(tot_res = tot_res, fea_res = fea_res);
+    all_res <- summarise_randmat_ga(tot_res = tot_res, fea_res = fea_res);
     return(all_res);
 }
 
@@ -160,7 +160,7 @@ crossover <- function(inds, pr = 0.1){
 #'sim_ga      <- Evo_rand_gen_var(from = 2, to = 2, iters = 1);
 #'sum_rand_ga <- summarise_randmat_ga(sim_ga$tot_res, sim_ga$fea_res);
 #'@export
-summarise_randmat <- function(tot_res, fea_res){
+summarise_randmat_ga <- function(tot_res, fea_res){
     sims    <- length(tot_res);
     all_res <- matrix(data = 0, nrow = sims, ncol = 7);
     for(i in 1:sims){
@@ -178,8 +178,8 @@ summarise_randmat <- function(tot_res, fea_res){
         all_res[i, 6]  <- sum(stabled);
         all_res[i, 7]  <- sum(unstabled);
     }
-    colnames(all_res) <- c("N", "A0_unstable", "A0_stable", "A1_unstable",
-                           "A1_stable", "A1_stabilised", "A1_destabilised");
+    colnames(all_res) <- c("S", "A_unstable", "A_stable", "M_unstable",
+                           "M_stable", "A_stabilised", "A_destabilised");
     return(all_res);
 }
 
@@ -249,10 +249,4 @@ plot_evo_out <- function(evo_out){
           text = expression(paste("Component response rate (",gamma,")")));
     mtext(side = 2, text = "Frequency", outer = TRUE, line = 3, cex = 1.5);
 }
-
-
-
-
-
-
 

### 2020-03-21 notebook/ms.Rmd
Rearrange the results files so that the manuscript calls them from the inst Rpackage folder

@@
 <!---
-Abstract current word count: 225
-Text current word count: 1453
+Abstract current word count: 
+Text current word count: 
 --->
@@
 https://gist.github.com/JJ/ef9d3d8f1142df064bd5
---->
-
-<!---
-
-NOTE: Consider looking at the resilience of the stable networks given $\gamma = 1$ and $Var(\gamma)$, which is defined by the inverse of the leading eigenvalue.
-The inverse of the real part of the leading eigenvalue indicates how quickly the system returns to equilibrium (resilience).
-See @Landi2018 for more.
-Intuitively, resilience should perhaps increase a bit if we're only looking at the subset of systems that are stable (many of which will have had their eigenvalues shifted left)?
-
-Landi, P., Minoarivelo, H.
-O., Brännström, Å., Hui, C., & Dieckmann, U.
-(2018).
-Complexity and stability of ecological networks: A review of the theory.
-Population Ecology, 0(0), In press.
-https://doi.org/10.1007/s10144-018-0628-3
-
---->
-
-<!---
-
-1.
-As above, check out the change in resilience defined as the difference beween the inverse of the real part of the leading eigenvalue in $Var(\gamma)$ versus $\gamma = 1$.
-Initial simulaitons suggest that resilience is increased on average by $Var(\gamma)$, and increasingly so as S increases.
-
-2.
-Check for induced correlations causing stabilisation (and E1, E2, and EC, more generally), as per Reviewer 2 specific comment 3.
-If possible, give an analytical explanation for the increase in stability that has to do with the error inducing changes in correlations.
-
 --->
@@
 sigma <- 0.4;
-pr_st <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st <- pr_st[,-1];
-```
-
-```{r, echo = FALSE}
-A0 <- read.csv(file = "sim_results/bi_gamma/S200_A0.csv");
-A0 <- as.matrix(A0[,-1]);
-A1 <- read.csv(file = "sim_results/bi_gamma/S200_A1.csv");
-A1 <- as.matrix(A1[,-1]);
-
-A0_e   <- eigen(A0)$values;
-A0_r   <- Re(A0_e);
-A0_i   <- Im(A0_e);
-A1_e   <- eigen(A1)$values;
-A1_r   <- Re(A1_e);
-A1_i   <- Im(A1_e);
-
-A0_vm       <- A0;
-diag(A0_vm) <- NA;
-A0vec       <- as.vector(t(A0_vm));
-A0vec       <- A0vec[is.na(A0vec) == FALSE];
-A1_vm       <- A1;
-diag(A1_vm) <- NA;
-A1vec       <- as.vector(t(A1_vm));
-A1vec       <- A1vec[is.na(A1vec) == FALSE];
-fhalf       <- 1:(0.5*length(A1vec));
-shalf       <- (0.5*length(A1vec)+1):length(A1vec);
-pr_st       <- read.csv(file = "sim_results/bi_gamma/bi_pr_st.csv");
-pr_st       <- pr_st[,-1];
 ```
@@
 ```{r, echo = FALSE}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
+dat <- read.csv(file = "../inst/extdata/random_all.csv");
 ```
@@
 
-<!---
-
-For all simulated network structures, interaction strengths between components were sampled as in random networks ($\mathcal{N}(0, 0.4^{2})$), but algorithms for building networks necessarily resulted in variable $C$.
-
---->
-
 **System feasibility given $\mathbf{\sigma^{2}_{\gamma}}$** For complex systems in which individual system components represent the density of some tangible quantity, it is relevant to consider the feasibility of the system.
@@
 
-<!--- Tempting to remove this section entirely, but if there is enough space for it, then maybe retain it; reviewers did not mention a problem with it, and it's a short and relevant point to make.
-In any case, it should be swapped with the section on feasibility below.
---->
 **Targeted manipulation of $\boldsymbol{\gamma}$**.
@@
 --------------------------------------------------------------------------------
-
-<!---  (in practice, diagonal elements of $\textbf{M}$ were standardised so that mean values were identical before and after adding $\gamma$) --->
 
@@
 ```{r, echo = FALSE, fig.height = 6, fig.width = 6}
-sim_rhos <- read.csv(file = "sim_results/rhos/sim_rhos.csv");
+sim_rhos <- read.csv(file = "../inst/extdata/rhos/sim_rhos.csv");
 par(mar = c(5, 5, 1, 1));
@@
 
-```{r, echo = FALSE, fig.height = 8, fig.width = 6}
-sim10 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim10.csv");
-sim15 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim15.csv");
-sim20 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim20.csv");
-sim25 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim25.csv");
-sim30 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim30.csv");
-sim35 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim35.csv");
-sim40 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim40.csv");
-sim45 <- read.csv(file = "sim_results/rhos/sigma0pt2/sim45.csv");
+```{r, echo = FALSE, fig.height = 7.5, fig.width = 6}
+sim10 <- read.csv(file = "../inst/extdata/rhos/sim10.csv");
+sim15 <- read.csv(file = "../inst/extdata/rhos/sim15.csv");
+sim20 <- read.csv(file = "../inst/extdata/rhos/sim20.csv");
+sim25 <- read.csv(file = "../inst/extdata/rhos/sim25.csv");
+sim30 <- read.csv(file = "../inst/extdata/rhos/sim30.csv");
+sim35 <- read.csv(file = "../inst/extdata/rhos/sim35.csv");
+sim40 <- read.csv(file = "../inst/extdata/rhos/sim40.csv");
+sim45 <- read.csv(file = "../inst/extdata/rhos/sim45.csv");
 fstlt <- c(1, dim(sim20)[1]);
@@
 
-```{r, echo = FALSE}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
-```
-
 **Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems for which variation in $\gamma$ is critical for system stability.
@@
 ```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1];
 plot_stables <- function(dat, S_s = 32){

### 2020-03-21 R/Figures_3_4.R
Very small change to the plotting figure to improve the x-axis

@@ -18,7 +18,7 @@ plot_stables <- function(dat, S_s = 32){
     barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
             names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
             ylab = "Ln number of stable communities", cex.lab = 1, 
-            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+            cex.axis = 1.25, xlim = c(1, 3*S_s), cex.names = 1, yaxt = "n");
     axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
     box(lwd = 2);
     par(new = TRUE);
@@ -64,7 +64,7 @@ plot_stable_4 <- function(dat, S_s = 40){
     barplot(log_bar_dat, beside = TRUE, col = c("dodgerblue4", "firebrick"),
             names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
             ylab = "Ln number of stable communities", cex.lab = 1, 
-            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+            cex.axis = 1.25, xlim = c(1, S_s*3), cex.names = 1, yaxt = "n");
     axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
     box(lwd = 2);
     par(new = TRUE);

### 2020-03-21 notebook/SI.Rmd
Edit the SI so that data are now called from inst instead of the notebook, this will payoff later when making the SI into a vignette

@@
 source(file = "../R/Figures_3_4.R");
-source(file = "R_old/sim_mat.R");
-source(file = "R_old/plot_figs.R");
 source(file = "../R/utilities.R")
 ```
-
-********************************************************************************
 
@@
 - [Consistency with Gibbs et al. (2018)](#Gibbs)
+- [Reproducing simulation results](#repr)
+- [Literature cited](#ref)
 
@@
 ```{r, echo = FALSE}
-dat <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat <- dat[,-1]; # Extra row-indicating column removed
+dat <- read.csv(file = "../inst/extdata/random_all.csv");
 ```
@@
 ```{r, echo = FALSE}
-cdat <- read.csv(file = "sim_results/ecology/competition_C_1.csv");
-mdat <- read.csv(file = "sim_results/ecology/mutualism_C_1.csv");
-pdat <- read.csv(file = "sim_results/ecology/pred-prey_C_1.csv");
+cdat <- read.csv(file = "../inst/extdata/competition_C_1.csv");
+mdat <- read.csv(file = "../inst/extdata/mutualism_C_1.csv");
+pdat <- read.csv(file = "../inst/extdata/pred-prey_C_1.csv");
 ```
@@
 ```{r, echo = FALSE}
-C3dat <- read.csv(file = "sim_results/C_other/rand_c-0pt3.csv");
-C5dat <- read.csv(file = "sim_results/C_other/rand_c-0pt5.csv");
-C7dat <- read.csv(file = "sim_results/C_other/rand_c-0pt7.csv");
-C9dat <- read.csv(file = "sim_results/C_other/rand_c-0pt9.csv");
+C3dat <- read.csv(file = "../inst/extdata/rand_c-0pt3.csv");
+C5dat <- read.csv(file = "../inst/extdata/rand_c-0pt5.csv");
+C7dat <- read.csv(file = "../inst/extdata/rand_c-0pt7.csv");
+C9dat <- read.csv(file = "../inst/extdata/rand_c-0pt9.csv");
 ```
@@
 ```{r, echo = FALSE}
-C3dat   <- C3dat[,-1];
 C3d_tab <- C3dat[,1:7];
@@
 ```{r, echo = FALSE}
-C5dat   <- C5dat[,-1];
 C5d_tab <- C5dat[,1:7];
@@
 ```{r, echo = FALSE}
-C7dat   <- C7dat[,-1];
 C7d_tab <- C7dat[,1:7];
@@
 ```{r, echo = FALSE}
-C9dat   <- C9dat[,-1];
 C9d_tab <- C9dat[,1:7];
@@
 ```{r, echo = FALSE}
-S3dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt3.csv");
-S4dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt4.csv");
-S5dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt5.csv");
-S6dat <- read.csv(file = "sim_results/C_0pt05/sim_SD_0pt6.csv");
+S3dat <- read.csv(file = "../inst/extdata/sim_SD_0pt3.csv");
+S4dat <- read.csv(file = "../inst/extdata/sim_SD_0pt4.csv");
+S5dat <- read.csv(file = "../inst/extdata/sim_SD_0pt5.csv");
+S6dat <- read.csv(file = "../inst/extdata/sim_SD_0pt6.csv");
 ```
@@
 ```{r, echo = FALSE}
-S3dat   <- S3dat[,-1];
 S3d_tab <- S3dat[,1:7];
@@
 ```{r, echo = FALSE}
-S4dat   <- S4dat[,-1];
 S4d_tab <- S4dat[,1:7];
@@
 ```{r, echo = FALSE}
-S5dat   <- S5dat[,-1];
 S5d_tab <- S5dat[,1:7];
@@
 ```{r, echo = FALSE}
-S6dat   <- S6dat[,-1];
 S6d_tab <- S6dat[,1:7];
@@
 ```{r, echo = FALSE, fig.align = "center"}
-swn <- read.csv(file = "sim_results/swn/old/swn.csv"); # Change when new results are in
-use <- c(1:6, 19:21);
+swn <- read.csv(file = "../inst/extdata/swn.csv"); 
+use <- c(1:6, 19:20, 27);
 kable(swn[,use]);
@@
 
-In the above, the mean $C$, complexity of $\mathbf{A}$, and complexity of $\mathbf{M}$ are also shown.
+In the above, the complexity of $\mathbf{A}$ and $\mathbf{M}$, and the mean $C$, are also shown.
 For similar magnitudes of complexity as in random networks of $\sigma\sqrt{SC} \gtrapprox 1.26$, variation in $\gamma$ typically results in more stable than unstable systems.
@@
 ```{r, echo = FALSE, fig.align = "center"}
-sfn <- read.csv(file = "sim_results/sfn/sfn.csv"); 
+sfn <- read.csv(file = "../inst/extdata/sfn.csv"); 
 use <- c(1:6, 19:21);
@@
 ```{r, echo = FALSE, fig.align = "center"}
-csc <- read.csv(file = "sim_results/ecology/cascade_C_1.csv"); 
+csc <- read.csv(file = "../inst/extdata/cascade_C_1.csv"); 
 use <- c(1:5, 18:19);
@@
 ```{r, echo = FALSE}
-dat      <- read.csv(file = "sim_results/C_1/random_all.csv");
-dat      <- dat[,-1]; # Extra row-indicating column removed
-feas_tab <- cbind(dat[,1], dat[,8:13]);
-colnames(feas_tab)[1] <- "S";
-kable(feas_tab);
+dat      <- read.csv(file = "../inst/extdata/random_all.csv");
+feas_tab <- c(1, 8:13);
+kable(dat[,feas_tab]);
 ```
@@
 ```{r, eval = TRUE, echo = FALSE, fig.width = 9, fig.height = 7}
-dat <- read.csv(file = "sim_results/evolved/evo_results.csv");
-dat <- dat[,-1];
-plot_stable_4(dat);
+evo_dat <- read.csv(file = "../inst/extdata/evo_results.csv");
+plot_stables(evo_dat, S_s = 40);
 ```
@@
 }
-evo_res <- read.csv(file = "sim_results/evolved/evo_results.csv");
-evo_res <- evo_res[,-1];
-kable(evo_res);
+kable(evo_dat);
 ```
@@
 ```{r, fig.height = 4, fig.width = 6, echo = FALSE}
-evo_out <- scan(file = "sim_results/evolved/evo_out.txt");
+evo_out <- scan(file = "../inst/extdata/evo_out.txt");
 plot_evo_out(evo_out);
@@
 ```{r, echo = FALSE}
-constant_complexity <- read.csv("sim_results/constant_complexity.csv")
+constant_complexity <- read.csv("../inst/extdata/constant_complexity.csv")
 kable(constant_complexity);
@@
 
-References
+\clearpage
+
+Reproducing simulation results {#repr}
 ================================================================================
+
+All results in the main text and the literature cited can be reproduced using the [RandomMatrixStability](https://github.com/bradduthie/RandomMatrixStability) R package, which can be downloaded as instructed at the beginning of this Supplemental Information document.
+The most relevant R functions for reproducing simulations include the following:
+
+1. `rand_gen_var`: Simulates random complex systems and cascade food webs
+2. `rand_rho_var`: Simulates random complex systems across a fixed correlation of $\rho = cor(A_{ij}, A_{ji})$
+3. `rand_gen_swn`: Simulates randomly generated small-world networks
+4. `rand_gen_sfn`: Simulates randomly generated scale-free networks
+5. `Evo_rand_gen_var`: Use a genetic algorithm to find stable random complex systems
+
+For the functions 1-4 above, R output will be a table of results. Below describes the headers of this table to more clearly explain what is being reported.
+
+\footnotesize
+
+```{r, echo = FALSE}
+header_descriptions <- read.csv("../inst/extdata/header_descriptions.csv")
+kable(header_descriptions);
+```
+
+\normalsize
+
+Note that output from `Evo_rand_gen_var` only includes the first seven rows of the table above, and `rand_gen_var` does not include $C$ (which can be defind as an argument).
+All results presented here and in the main text are available in the [inst/extdata](https://github.com/bradduthie/RandomMatrixStability/tree/master/inst/extdata) folder of the [RandomMatrixStability](https://github.com/bradduthie/RandomMatrixStability) R package.
+
+Literature cited {#ref}
+================================================================================

### 2020-03-21 notebook/SI.tex
Edit the SI so that data are now called from inst instead of the notebook, this will payoff later when making the SI into a vignette

@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
 \textbf{This supplemental information supports the manuscript
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
 \section{Supplemental Information table of
@@
   \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
+\item
+  \protect\hyperlink{repr}{Reproducing simulation results}
+\item
+  \protect\hyperlink{ref}{Literature cited}
 \end{itemize}
@@
 \endhead
-0.00 & 24 & 17442 & 982558 & 17489 & 982511 & 0.1695021 & 0.1941223 &
+0.00 & 24 & 17388 & 982612 & 17446 & 982554 & 0.5748066 & 0.6582632 &
 0.1304348\tabularnewline
-0.00 & 48 & 257013 & 742987 & 259517 & 740483 & 0.2355489 & 0.2711380 &
+0.00 & 48 & 258024 & 741976 & 260579 & 739421 & 0.8073918 & 0.9294192 &
 0.1063830\tabularnewline
-0.00 & 72 & 715587 & 284413 & 723082 & 276918 & 0.2866561 & 0.3304009 &
+0.00 & 72 & 715036 & 284964 & 722639 & 277361 & 0.9860840 & 1.1364805 &
 0.0985915\tabularnewline
-0.00 & 96 & 961817 & 38183 & 963276 & 36724 & 0.3299115 & 0.3804889 &
+0.00 & 96 & 961434 & 38566 & 962788 & 37212 & 1.1369395 & 1.3110263 &
 0.0947368\tabularnewline
-0.00 & 120 & 998976 & 1024 & 998812 & 1188 & 0.3681518 & 0.4246688 &
+0.00 & 120 & 999008 & 992 & 998857 & 1143 & 1.2700387 & 1.4649832 &
 0.0924370\tabularnewline
-0.00 & 144 & 999998 & 2 & 999991 & 9 & 0.4027471 & 0.4646754 &
+0.00 & 144 & 999997 & 3 & 999994 & 6 & 1.3903192 & 1.6041216 &
 0.0909091\tabularnewline
-0.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345714 & 0.5014674 &
+0.00 & 168 & 1000000 & 0 & 1000000 & 0 & 1.5010334 & 1.7320676 &
 0.0898204\tabularnewline
-0.01 & 24 & 17386 & 982614 & 17433 & 982567 & 0.1695343 & 0.1941452 &
-0.1304321\tabularnewline
-0.01 & 48 & 255369 & 744631 & 257956 & 742044 & 0.2355569 & 0.2711250 &
-0.1063798\tabularnewline
-0.01 & 72 & 708407 & 291593 & 716019 & 283981 & 0.2866423 & 0.3303843 &
+0.01 & 24 & 17673 & 982327 & 17720 & 982280 & 0.5747156 & 0.6581503 &
+0.1304319\tabularnewline
+0.01 & 48 & 255038 & 744962 & 257647 & 742353 & 0.8073388 & 0.9292952 &
+0.1063800\tabularnewline
+0.01 & 72 & 708892 & 291108 & 716829 & 283171 & 0.9859457 & 1.1363940 &
 0.0985884\tabularnewline
-0.01 & 96 & 960591 & 39409 & 961685 & 38315 & 0.3299322 & 0.3804641 &
+0.01 & 96 & 960635 & 39365 & 961876 & 38124 & 1.1370640 & 1.3112193 &
 0.0947337\tabularnewline
-0.01 & 120 & 998991 & 1009 & 998813 & 1187 & 0.3681428 & 0.4246683 &
-0.0924339\tabularnewline
-0.01 & 144 & 999997 & 3 & 999993 & 7 & 0.4027696 & 0.4646907 &
+0.01 & 120 & 999040 & 960 & 998794 & 1206 & 1.2698715 & 1.4648280 &
+0.0924338\tabularnewline
+0.01 & 144 & 999997 & 3 & 999994 & 6 & 1.3901601 & 1.6039285 &
 0.0909060\tabularnewline
-0.01 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345881 & 0.5014852 &
-0.0898172\tabularnewline
-0.10 & 24 & 20408 & 979592 & 20478 & 979522 & 0.1692658 & 0.1937917 &
-0.1302962\tabularnewline
-0.10 & 48 & 238101 & 761899 & 240787 & 759213 & 0.2351005 & 0.2705892 &
-0.1062307\tabularnewline
-0.10 & 72 & 680138 & 319862 & 685524 & 314476 & 0.2861156 & 0.3297663 &
-0.0984350\tabularnewline
-0.10 & 96 & 961691 & 38309 & 959740 & 40260 & 0.3293218 & 0.3797666 &
-0.0945786\tabularnewline
-0.10 & 120 & 999544 & 456 & 999282 & 718 & 0.3674529 & 0.4238479 &
-0.0922777\tabularnewline
-0.10 & 144 & 999999 & 1 & 999999 & 1 & 0.4019505 & 0.4637655 &
-0.0907492\tabularnewline
-0.10 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4337704 & 0.5005276 &
+0.01 & 168 & 1000000 & 0 & 1000000 & 0 & 1.5009490 & 1.7319739 &
+0.0898173\tabularnewline
+0.10 & 24 & 20382 & 979618 & 20455 & 979545 & 0.5742520 & 0.6573563 &
+0.1302974\tabularnewline
+0.10 & 48 & 237747 & 762253 & 240370 & 759630 & 0.8066604 & 0.9284434 &
+0.1062311\tabularnewline
+0.10 & 72 & 679874 & 320126 & 685575 & 314425 & 0.9849695 & 1.1352553 &
+0.0984349\tabularnewline
+0.10 & 96 & 961984 & 38016 & 960128 & 39872 & 1.1358912 & 1.3097957 &
+0.0945788\tabularnewline
+0.10 & 120 & 999546 & 454 & 999275 & 725 & 1.2687142 & 1.4634587 &
+0.0922779\tabularnewline
+0.10 & 144 & 1000000 & 0 & 1000000 & 0 & 1.3890356 & 1.6025900 &
+0.0907489\tabularnewline
+0.10 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4994818 & 1.7302649 &
 0.0896598\tabularnewline
-0.25 & 24 & 23923 & 976077 & 24031 & 975969 & 0.1680052 & 0.1922420 &
-0.1296696\tabularnewline
-0.25 & 48 & 228389 & 771611 & 231312 & 768688 & 0.2331850 & 0.2683721 &
-0.1055242\tabularnewline
-0.25 & 72 & 666374 & 333626 & 668472 & 331528 & 0.2836431 & 0.3268807 &
-0.0977062\tabularnewline
-0.25 & 96 & 966462 & 33538 & 961616 & 38384 & 0.3263952 & 0.3763783 &
-0.0938386\tabularnewline
-0.25 & 120 & 999759 & 241 & 999510 & 490 & 0.3642255 & 0.4201168 &
-0.0915315\tabularnewline
-0.25 & 144 & 1000000 & 0 & 999999 & 1 & 0.3983171 & 0.4595509 &
-0.0899989\tabularnewline
-0.25 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4298720 & 0.4960198 &
-0.0889060\tabularnewline
-1.00 & 24 & 26159 & 973841 & 26309 & 973691 & 0.1588024 & 0.1815081 &
-0.1249673\tabularnewline
-1.00 & 48 & 211700 & 788300 & 214607 & 785393 & 0.2154544 & 0.2478696 &
-0.0991369\tabularnewline
-1.00 & 72 & 613966 & 386034 & 615862 & 384138 & 0.2602685 & 0.2998802 &
-0.0908174\tabularnewline
-1.00 & 96 & 943949 & 56051 & 936840 & 63160 & 0.2984622 & 0.3441159 &
-0.0867075\tabularnewline
-1.00 & 120 & 999134 & 866 & 998351 & 1649 & 0.3323169 & 0.3833061 &
-0.0842580\tabularnewline
-1.00 & 144 & 1000000 & 0 & 999997 & 3 & 0.3630114 & 0.4187917 &
-0.0826318\tabularnewline
-1.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.3913277 & 0.4515183 &
-0.0814745\tabularnewline
+0.25 & 24 & 23654 & 976346 & 23775 & 976225 & 0.5722185 & 0.6546853 &
+0.1296712\tabularnewline
+0.25 & 48 & 228318 & 771682 & 231208 & 768792 & 0.8033257 & 0.9244966 &
+0.1055259\tabularnewline
+0.25 & 72 & 666982 & 333018 & 669104 & 330896 & 0.9808676 & 1.1304109 &
+0.0977066\tabularnewline
+0.25 & 96 & 966456 & 33544 & 961545 & 38455 & 1.1307841 & 1.3039452 &
+0.0938392\tabularnewline
+0.25 & 120 & 999749 & 251 & 999507 & 493 & 1.2632327 & 1.4571506 &
+0.0915316\tabularnewline
+0.25 & 144 & 1000000 & 0 & 1000000 & 0 & 1.3827642 & 1.5953248 &
+0.0899987\tabularnewline
+0.25 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4926700 & 1.7224506 &
+0.0889064\tabularnewline
+1.00 & 24 & 26331 & 973669 & 26478 & 973522 & 0.5561013 & 0.6356655 &
+0.1249651\tabularnewline
+1.00 & 48 & 211199 & 788801 & 214154 & 785846 & 0.7720342 & 0.8881302 &
+0.0991370\tabularnewline
+1.00 & 72 & 613621 & 386379 & 615771 & 384229 & 0.9394912 & 1.0825566 &
+0.0908153\tabularnewline
+1.00 & 96 & 943191 & 56809 & 936396 & 63604 & 1.0812364 & 1.2466510 &
+0.0867047\tabularnewline
+1.00 & 120 & 999157 & 843 & 998396 & 1604 & 1.2065026 & 1.3916458 &
+0.0842561\tabularnewline
+1.00 & 144 & 1000000 & 0 & 999997 & 3 & 1.3199179 & 1.5227509 &
+0.0826325\tabularnewline
+1.00 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4243560 & 1.6434386 &
+0.0814738\tabularnewline
 \bottomrule
@@
 
-In the above, the mean \(C\), complexity of \(\mathbf{A}\), and
-complexity of \(\mathbf{M}\) are also shown. For similar magnitudes of
-complexity as in random networks of \(\sigma\sqrt{SC} \gtrapprox 1.26\),
-variation in \(\gamma\) typically results in more stable than unstable
-systems.
+In the above, the complexity of \(\mathbf{A}\) and \(\mathbf{M}\), and
+the mean \(C\), are also shown. For similar magnitudes of complexity as
+in random networks of \(\sigma\sqrt{SC} \gtrapprox 1.26\), variation in
+\(\gamma\) typically results in more stable than unstable systems.
 
@@
 \toprule
-S & M & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
+m & S & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
 complex\_M & C\tabularnewline
@@
 \endhead
-24 & 2 & 152791 & 847209 & 156034 & 843966 & 0.7891257 & 0.9034663 &
+2 & 24 & 152791 & 847209 & 156034 & 843966 & 0.7891257 & 0.9034663 &
 0.1648551\tabularnewline
-24 & 3 & 320481 & 679519 & 326351 & 673649 & 0.9566487 & 1.0967499 &
+3 & 24 & 320481 & 679519 & 326351 & 673649 & 0.9566487 & 1.0967499 &
 0.2409420\tabularnewline
-24 & 4 & 504433 & 495567 & 504826 & 495174 & 1.0922870 & 1.2532761 &
+4 & 24 & 504433 & 495567 & 504826 & 495174 & 1.0922870 & 1.2532761 &
 0.3134058\tabularnewline
-24 & 5 & 670676 & 329324 & 660426 & 339574 & 1.2073054 & 1.3857169 &
+5 & 24 & 670676 & 329324 & 660426 & 339574 & 1.2073054 & 1.3857169 &
 0.3822464\tabularnewline
-24 & 6 & 798637 & 201363 & 779345 & 220655 & 1.3067095 & 1.5004508 &
+6 & 24 & 798637 & 201363 & 779345 & 220655 & 1.3067095 & 1.5004508 &
 0.4474638\tabularnewline
-24 & 7 & 884082 & 115918 & 862215 & 137785 & 1.3942577 & 1.6013368 &
+7 & 24 & 884082 & 115918 & 862215 & 137785 & 1.3942577 & 1.6013368 &
 0.5090580\tabularnewline
-24 & 8 & 936190 & 63810 & 915630 & 84370 & 1.4722315 & 1.6908563 &
+8 & 24 & 936190 & 63810 & 915630 & 84370 & 1.4722315 & 1.6908563 &
 0.5670290\tabularnewline
-24 & 9 & 964868 & 35132 & 948297 & 51703 & 1.5414455 & 1.7707292 &
+9 & 24 & 964868 & 35132 & 948297 & 51703 & 1.5414455 & 1.7707292 &
 0.6213768\tabularnewline
-24 & 10 & 981460 & 18540 & 967911 & 32089 & 1.6030044 & 1.8417459 &
+10 & 24 & 981460 & 18540 & 967911 & 32089 & 1.6030044 & 1.8417459 &
 0.6721014\tabularnewline
-24 & 11 & 989838 & 10162 & 980232 & 19768 & 1.6586511 & 1.9059313 &
+11 & 24 & 989838 & 10162 & 980232 & 19768 & 1.6586511 & 1.9059313 &
 0.7192029\tabularnewline
-24 & 12 & 994393 & 5607 & 987436 & 12564 & 1.7081503 & 1.9628898 &
+12 & 24 & 994393 & 5607 & 987436 & 12564 & 1.7081503 & 1.9628898 &
 0.7626812\tabularnewline
-48 & 2 & 303963 & 696037 & 310053 & 689947 & 0.7946875 & 0.9132519 &
+2 & 48 & 303963 & 696037 & 310053 & 689947 & 0.7946875 & 0.9132519 &
 0.0828901\tabularnewline
-48 & 3 & 577855 & 422145 & 579996 & 420004 & 0.9685494 & 1.1141445 &
+3 & 48 & 577855 & 422145 & 579996 & 420004 & 0.9685494 & 1.1141445 &
 0.1227837\tabularnewline
-48 & 4 & 810001 & 189999 & 799132 & 200868 & 1.1122992 & 1.2799335 &
+4 & 48 & 810001 & 189999 & 799132 & 200868 & 1.1122992 & 1.2799335 &
 0.1617908\tabularnewline
-48 & 5 & 938004 & 61996 & 924613 & 75387 & 1.2369960 & 1.4236817 &
+5 & 48 & 938004 & 61996 & 924613 & 75387 & 1.2369960 & 1.4236817 &
 0.1999113\tabularnewline
-48 & 6 & 984975 & 15025 & 976433 & 23567 & 1.3478291 & 1.5514420 &
+6 & 48 & 984975 & 15025 & 976433 & 23567 & 1.3478291 & 1.5514420 &
 0.2371454\tabularnewline
-48 & 7 & 997160 & 2840 & 994005 & 5995 & 1.4473792 & 1.6663763 &
+7 & 48 & 997160 & 2840 & 994005 & 5995 & 1.4473792 & 1.6663763 &
 0.2734929\tabularnewline
-48 & 8 & 999584 & 416 & 998590 & 1410 & 1.5385445 & 1.7716359 &
+8 & 48 & 999584 & 416 & 998590 & 1410 & 1.5385445 & 1.7716359 &
 0.3089539\tabularnewline
-48 & 9 & 999955 & 45 & 999707 & 293 & 1.6227742 & 1.8687074 &
+9 & 48 & 999955 & 45 & 999707 & 293 & 1.6227742 & 1.8687074 &
 0.3435284\tabularnewline
-48 & 10 & 999992 & 8 & 999939 & 61 & 1.7006157 & 1.9583879 &
+10 & 48 & 999992 & 8 & 999939 & 61 & 1.7006157 & 1.9583879 &
 0.3772163\tabularnewline
-48 & 11 & 999999 & 1 & 999990 & 10 & 1.7731759 & 2.0420990 &
+11 & 48 & 999999 & 1 & 999990 & 10 & 1.7731759 & 2.0420990 &
 0.4100177\tabularnewline
-48 & 12 & 1000000 & 0 & 999999 & 1 & 1.8410402 & 2.1203112 &
+12 & 48 & 1000000 & 0 & 999999 & 1 & 1.8410402 & 2.1203112 &
 0.4419326\tabularnewline
-72 & 2 & 427243 & 572757 & 434600 & 565400 & 0.7964226 & 0.9166566 &
+2 & 72 & 427243 & 572757 & 434600 & 565400 & 0.7964226 & 0.9166566 &
 0.0553599\tabularnewline
-72 & 3 & 741345 & 258655 & 739020 & 260980 & 0.9723446 & 1.1195788 &
+3 & 72 & 741345 & 258655 & 739020 & 260980 & 0.9723446 & 1.1195788 &
 0.0823552\tabularnewline
-72 & 4 & 931043 & 68957 & 921145 & 78855 & 1.1188220 & 1.2888100 &
+4 & 72 & 931043 & 68957 & 921145 & 78855 & 1.1188220 & 1.2888100 &
 0.1089593\tabularnewline
-72 & 5 & 989644 & 10356 & 984372 & 15628 & 1.2466268 & 1.4361875 &
+5 & 72 & 989644 & 10356 & 984372 & 15628 & 1.2466268 & 1.4361875 &
 0.1351721\tabularnewline
-72 & 6 & 999131 & 869 & 997914 & 2086 & 1.3604666 & 1.5674966 &
+6 & 72 & 999131 & 869 & 997914 & 2086 & 1.3604666 & 1.5674966 &
 0.1609937\tabularnewline
-72 & 7 & 999946 & 54 & 999804 & 196 & 1.4642496 & 1.6872501 &
+7 & 72 & 999946 & 54 & 999804 & 196 & 1.4642496 & 1.6872501 &
 0.1864241\tabularnewline
-72 & 8 & 999999 & 1 & 999988 & 12 & 1.5596340 & 1.7974044 &
+8 & 72 & 999999 & 1 & 999988 & 12 & 1.5596340 & 1.7974044 &
 0.2114632\tabularnewline
-72 & 9 & 1000000 & 0 & 999999 & 1 & 1.6482181 & 1.8994441 &
+9 & 72 & 1000000 & 0 & 999999 & 1 & 1.6482181 & 1.8994441 &
 0.2361111\tabularnewline
-72 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7307859 & 1.9947150 &
+10 & 72 & 1000000 & 0 & 1000000 & 0 & 1.7307859 & 1.9947150 &
 0.2603678\tabularnewline
-72 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8086766 & 2.0847262 &
+11 & 72 & 1000000 & 0 & 1000000 & 0 & 1.8086766 & 2.0847262 &
 0.2842332\tabularnewline
-72 & 12 & 1000000 & 0 & 1000000 & 0 & 1.8817533 & 2.1689764 &
+12 & 72 & 1000000 & 0 & 1000000 & 0 & 1.8817533 & 2.1689764 &
 0.3077074\tabularnewline
-96 & 2 & 527633 & 472367 & 535188 & 464812 & 0.7974024 & 0.9183557 &
+2 & 96 & 527633 & 472367 & 535188 & 464812 & 0.7974024 & 0.9183557 &
 0.0415570\tabularnewline
-96 & 3 & 842274 & 157726 & 837756 & 162244 & 0.9741293 & 1.1224709 &
+3 & 96 & 842274 & 157726 & 837756 & 162244 & 0.9741293 & 1.1224709 &
 0.0619518\tabularnewline
-96 & 4 & 975834 & 24166 & 969478 & 30522 & 1.1220115 & 1.2931371 &
+4 & 96 & 975834 & 24166 & 969478 & 30522 & 1.1220115 & 1.2931371 &
 0.0821272\tabularnewline
-96 & 5 & 998391 & 1609 & 996991 & 3009 & 1.2511287 & 1.4422331 &
+5 & 96 & 998391 & 1609 & 996991 & 3009 & 1.2511287 & 1.4422331 &
 0.1020833\tabularnewline
-96 & 6 & 999955 & 45 & 999838 & 162 & 1.3669903 & 1.5757699 &
+6 & 96 & 999955 & 45 & 999838 & 162 & 1.3669903 & 1.5757699 &
 0.1218202\tabularnewline
-96 & 7 & 999999 & 1 & 999996 & 4 & 1.4725862 & 1.6977057 &
+7 & 96 & 999999 & 1 & 999996 & 4 & 1.4725862 & 1.6977057 &
 0.1413377\tabularnewline
-96 & 8 & 1000000 & 0 & 1000000 & 0 & 1.5699145 & 1.8099762 &
+8 & 96 & 1000000 & 0 & 1000000 & 0 & 1.5699145 & 1.8099762 &
 0.1606360\tabularnewline
-96 & 9 & 1000000 & 0 & 1000000 & 0 & 1.6606162 & 1.9146804 &
+9 & 96 & 1000000 & 0 & 1000000 & 0 & 1.6606162 & 1.9146804 &
 0.1797149\tabularnewline
-96 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7457971 & 2.0129344 &
+10 & 96 & 1000000 & 0 & 1000000 & 0 & 1.7457971 & 2.0129344 &
 0.1985746\tabularnewline
-96 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8260368 & 2.1055559 &
+11 & 96 & 1000000 & 0 & 1000000 & 0 & 1.8260368 & 2.1055559 &
 0.2172149\tabularnewline
-96 & 12 & 1000000 & 0 & 1000000 & 0 & 1.9018608 & 2.1929362 &
+12 & 96 & 1000000 & 0 & 1000000 & 0 & 1.9018608 & 2.1929362 &
 0.2356360\tabularnewline
-120 & 2 & 609563 & 390437 & 616036 & 383964 & 0.7979355 & 0.9194404 &
+2 & 120 & 609563 & 390437 & 616036 & 383964 & 0.7979355 & 0.9194404 &
 0.0332633\tabularnewline
-120 & 3 & 904064 & 95936 & 899040 & 100960 & 0.9753815 & 1.1243251 &
+3 & 120 & 904064 & 95936 & 899040 & 100960 & 0.9753815 & 1.1243251 &
 0.0496499\tabularnewline
-120 & 4 & 991710 & 8290 & 988410 & 11590 & 1.1239922 & 1.2957520 &
+4 & 120 & 991710 & 8290 & 988410 & 11590 & 1.1239922 & 1.2957520 &
 0.0658964\tabularnewline
-120 & 5 & 999781 & 219 & 999477 & 523 & 1.2539362 & 1.4458518 &
+5 & 120 & 999781 & 219 & 999477 & 523 & 1.2539362 & 1.4458518 &
 0.0820028\tabularnewline
-120 & 6 & 999999 & 1 & 999981 & 19 & 1.3707937 & 1.5806987 &
+6 & 120 & 999999 & 1 & 999981 & 19 & 1.3707937 & 1.5806987 &
 0.0979692\tabularnewline
-120 & 7 & 1000000 & 0 & 999999 & 1 & 1.4775366 & 1.7038860 &
+7 & 120 & 1000000 & 0 & 999999 & 1 & 1.4775366 & 1.7038860 &
 0.1137955\tabularnewline
-120 & 8 & 1000000 & 0 & 1000000 & 0 & 1.5762636 & 1.8177236 &
+8 & 120 & 1000000 & 0 & 1000000 & 0 & 1.5762636 & 1.8177236 &
 0.1294818\tabularnewline
-120 & 9 & 1000000 & 0 & 1000000 & 0 & 1.6680647 & 1.9238257 &
+9 & 120 & 1000000 & 0 & 1000000 & 0 & 1.6680647 & 1.9238257 &
 0.1450280\tabularnewline
-120 & 10 & 1000000 & 0 & 1000000 & 0 & 1.7545110 & 2.0233838 &
+10 & 120 & 1000000 & 0 & 1000000 & 0 & 1.7545110 & 2.0233838 &
 0.1604342\tabularnewline
-120 & 11 & 1000000 & 0 & 1000000 & 0 & 1.8363882 & 2.1178385 &
+11 & 120 & 1000000 & 0 & 1000000 & 0 & 1.8363882 & 2.1178385 &
 0.1757003\tabularnewline
-120 & 12 & 1000000 & 0 & 1000000 & 0 & 1.9135798 & 2.2069806 &
+12 & 120 & 1000000 & 0 & 1000000 & 0 & 1.9135798 & 2.2069806 &
 0.1908263\tabularnewline
@@
 
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+\clearpage
+
+\hypertarget{repr}{\section{Reproducing simulation results}\label{repr}}
+
+All results in the main text and the literature cited can be reproduced
+using the
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability}
+R package, which can be downloaded as instructed at the beginning of
+this Supplemental Information document. The most relevant R functions
+for reproducing simulations include the following:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \texttt{rand\_gen\_var}: Simulates random complex systems and cascade
+  food webs
+\item
+  \texttt{rand\_rho\_var}: Simulates random complex systems across a
+  fixed correlation of \(\rho = cor(A_{ij}, A_{ji})\)
+\item
+  \texttt{rand\_gen\_swn}: Simulates randomly generated small-world
+  networks
+\item
+  \texttt{rand\_gen\_sfn}: Simulates randomly generated scale-free
+  networks
+\item
+  \texttt{Evo\_rand\_gen\_var}: Use a genetic algorithm to find stable
+  random complex systems
+\end{enumerate}
+
+For the functions 1-4 above, R output will be a table of results. Below
+describes the headers of this table to more clearly explain what is
+being reported.
+
+\footnotesize
+
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+Header & Description & Header\_cont. & Description\_cont.\tabularnewline
+\midrule
+\endhead
+S & The system size & A\_rho & Corr. between elements A{[}ij{]} and
+A{[}ji{]}\tabularnewline
+A\_unstable & No. of A that were unstable & M\_rho & Corr. between
+elements M{[}ij{]} and M{[}ji{]}\tabularnewline
+A\_stable & No. of A that were stable & rho\_diff & Diff. between A and
+M rho values\tabularnewline
+M\_unstable & No. of M that were unstable & rho\_abs & Diff. between A
+and M rho magnitudes\tabularnewline
+M\_stable & No. of M that were stable & complex\_A & Complexity of
+A\tabularnewline
+A\_stabilised & No. of A stabilised by gamma & complex\_M & Complexity
+of M\tabularnewline
+A\_destabilised & No. of A destabilised by gamma & A\_eig & Expected
+real part of leading A eigenvalue\tabularnewline
+A\_infeasible & No. of A that were infeasible & M\_eig & Expected real
+part of leading M eigenvalue\tabularnewline
+A\_feasible & No. of A that were feasible & LR\_A & Lowest obs. real
+part of leading A eigenvalue\tabularnewline
+M\_infeasible & No. of M that were infeasible & UR\_A & Highest obs.
+real part of leading A eigenvalue\tabularnewline
+M\_feasible & No. of M that were feasible & LR\_M & Lowest obs. real
+part of leading M eigenvalue\tabularnewline
+A\_made\_feasible & No. of A made feasible by gamma & UR\_M & Highest
+obs. real part of leading M eigenvalue\tabularnewline
+A\_made\_infeasible & No. of A made infeasible by gamma & C & Obs.
+network connectance\tabularnewline
+\bottomrule
+\end{longtable}
+
+\normalsize
+
+Note that output from \texttt{Evo\_rand\_gen\_var} only includes the
+first seven rows of the table above, and \texttt{rand\_gen\_var} does
+not include \(C\) (which can be defind as an argument). All results
+presented here and in the main text are available in the
+\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/inst/extdata}{inst/extdata}
+folder of the
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability}
+R package.
+
+\hypertarget{ref}{\section*{Literature cited}\label{ref}}
+\addcontentsline{toc}{section}{Literature cited}
 

### 2020-03-21 notebook/R_old/plot_figs.R
Small change to the old R files that won't affect the final presentation of the manuscript or SI

@@ -359,7 +359,16 @@ stable_N_plot <- function(dat, S_s = 32){
 }
 
 
-
+#' Plot comparison of gamma = 1 versus var(gamma)
+#' 
+#' Produces a plot comparing random matrix results
+#'
+#'@return A plot as in Figures 3 of the manuscript, showing the stability of
+#'random matrices on a log scale in simulations where component response rate is
+#'constant versus simulations in which it varies
+#'@param dat Table of simulation results
+#'@param S_s Maximum size of S to use
+#'@export
 plot_stables <- function(dat, S_s = 32){
     Ns          <- 1:S_s;
     par(oma = c(6, 6, 1, 6), mar = c(0.5, 0.5, 0.5, 0.5));
@@ -370,7 +379,7 @@ plot_stables <- function(dat, S_s = 32){
     barplot(log_bar_dat, beside = TRUE, col = c("grey80", "grey40"),
             names.arg = dat[Ns,1], ylim = c(0, 16), xlab = "",
             ylab = "Ln number of stable communities", cex.lab = 1, 
-            cex.axis = 1.25, xlim = c(1, 94), cex.names = 1, yaxt = "n");
+            cex.axis = 1.25, xlim = c(1, 3*S_s), cex.names = 1, yaxt = "n");
     axis(side = 2, at = c(0, 2, 4, 6, 8, 10, 12, 14), cex.axis = 1.5);
     box(lwd = 2);
     par(new = TRUE);

### 2020-03-21 notebook/ms.Rmd
Fix an error about k in small world networks, which was actually parameterised as S/12, not S/2

@@
 To build the interaction matrix $\mathbf{A}$, I used element-wise multiplication $\mathbf{A} = \mathbf{J} \odot \mathbf{B}$, then set $diag(\mathbf{A}) = -1$.
-I set $k = S/2$ and simulated small-world networks across all combinations of $S = \{24, 48, 72, 96, 120, 144, 168\}$ and $\beta = \{0, 0.01, 0.1, 0.25, 1\}$.
+I set $k = S/12$ and simulated small-world networks across all combinations of $S = \{24, 48, 72, 96, 120, 144, 168\}$ and $\beta = \{0, 0.01, 0.1, 0.25, 1\}$.
 

### 2020-03-21 notebook/ms.tex
Rearrange the results files so that the manuscript calls them from the inst Rpackage folder
Minor tweak to the files being read into the main text
Fix an error about k in small world networks, which was actually parameterised as S/12, not S/2

@@
 \(\mathbf{A} = \mathbf{J} \odot \mathbf{B}\), then set
-\(diag(\mathbf{A}) = -1\). I set \(k = S/2\) and simulated small-world
+\(diag(\mathbf{A}) = -1\). I set \(k = S/12\) and simulated small-world
 networks across all combinations of
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-9-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
 

### 2020-03-21 R/random_mats.R
Make the maths notation look more like that in the manuscript text -- confirm that it doesn't change any of the calculations, but just puts things in matrix algebra notation which I am now using in the Results and Methods

@@ -20,11 +20,15 @@
 #'@param sigma Standard deviation of interaction strength among network elements
 #'@param mn Mean interaction strength among network elements
 #'@param dval Self-regulation of network elements (1 by default)
+#'@param g_dist Gamma distribution to be used (1 by default)
+#'@param g_mn Mean value of gamma distribution (only for g_dist 0)
+#'@param g_sd Standard deviation of gamma distribution (only for g_dist 2-4)
 #'@examples
 #'rand_gen_var(max_sp = 2, iters = 4);
 #'@export
 rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
-                         sigma = 0.4, mn = 0, dval = 1){
+                         sigma = 0.4, mn = 0, dval = 1, g_dist = 1, g_mn = 1, 
+                         g_sd = 1){
     tot_res <- NULL;
     fea_res <- NULL;
     rho_res <- NULL;
@@ -51,12 +55,16 @@ rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 1,
                                ncol = sp_try[i]);
             A0       <- A0 * C_mat;
             diag(A0) <- -1 * dval;
-            gam1     <- runif(n = sp_try[i], min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0       <- A0 * mean(gam1);
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
-            A1_stb   <- max(Re(eigen(A1)$values)) < 0;
-            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
+            gam1     <- make_gammas(sp_try[i], g_dist, g_mn, g_sd);
+            gm       <- matrix(data = 0, nrow = sp_try[i], ncol = sp_try[i]);
+            diag(gm) <- gam1;
+            A1       <- gm %*% A0;
+            g0       <- matrix(data = 0, nrow = sp_try[i], ncol = sp_try[i]);
+            diag(g0) <- -1 * mean(diag(A1)); # Note: This standardisation will
+            A0       <- g0 %*% A0; # not affect stability, but I think helps
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0; # readers' interpretation
+            A1_stb   <- max(Re(eigen(A1)$values)) < 0; 
[…]

### 2020-03-21 R/scale_free.R

@@ -6,6 +6,7 @@
 #' are constructed using the method of Albert and Barab\'{a}si (Reviews of 
 #' Modern Physics, Vo. 74, 2002). Off-diagonal elements have a mean of 'mn' and 
 #' a standard deviation of 'sigma'.
+#' 
 #'@return A table of stability results, where rows summarise for each component
 #'number (S) the number of stable or unstable (also, feasible and infeasible)
 #'random matrices produced.
@@ -25,10 +26,10 @@
 #'@param mn Mean interaction strength among network elements
 #'@param dval Self-regulation of network elements (1 by default)
 #'@examples
-#'rand_gen_sfn(S = 24, mvals = c(2, 4), iters = 4);
+#'rand_gen_sfn(S = 24, m_vals = c(2, 4), iters = 4);
 #'@export
 rand_gen_sfn <- function(S, m_vals, iters, int_type = 0, rmx = 0.4, sigma = 0.4, 
-                         mn = 0, dval = 1){
+                         mn = 0, dval = 1, g_dist = 1, g_mn = 1, g_sd = 1){
     tot_res <- NULL;
     fea_res <- NULL;
     real_Cs <- NULL;
@@ -55,10 +56,14 @@ rand_gen_sfn <- function(S, m_vals, iters, int_type = 0, rmx = 0.4, sigma = 0.4,
             real_Cs[[i]][iter, 3] <- get_C(sfn);
             A0       <- A0 * sfn;
             diag(A0) <- -1 * dval;
-            gam1     <- runif(n = S, min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0       <- A0 * mean(gam1);
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            gam1     <- make_gammas(S, g_dist, g_mn, g_sd);
+            gm       <- matrix(data = 0, nrow = S, ncol = S);
+            diag(gm) <- gam1;
+            A1       <- gm %*% A0;
+            g0       <- matrix(data = 0, nrow = S, ncol = S);
+            diag(g0) <- -1 * mean(diag(A1)); # Note: This standardisation will
+            A0       <- g0 %*% A0; # not affect stability, but I think helps
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0; # readers' interpretation
             A1_stb   <- max(Re(eigen(A1)$values)) < 0;
[…]

### 2020-03-21 R/small_world.R

@@ -32,7 +32,8 @@
 #'rand_gen_swn(max_sp = 16, iters = 4);
 #'@export
 rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
-                         sigma = 0.4, Kdiv = 2, beta = 0.2, mn = 0, dval = 1){
+                         sigma = 0.4, Kdiv = 2, beta = 0.2, mn = 0, dval = 1,
+                         g_dist = 1, g_mn = 1, g_sd = 1){
     tot_res <- NULL;
     fea_res <- NULL;
     real_Cs <- NULL;
@@ -63,10 +64,14 @@ rand_gen_swn <- function(max_sp, iters, int_type = 0, rmx = 0.4, C = 1, by = 4,
             A0       <- A0 * C_mat;
             A0       <- A0 * swn;
             diag(A0) <- -1 * dval;
-            gam1     <- runif(n = sp_try[i], min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0       <- A0 * mean(gam1);
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
+            gam1     <- make_gammas(sp_try[i], g_dist, g_mn, g_sd);
+            gm       <- matrix(data = 0, nrow = sp_try[i], ncol = sp_try[i]);
+            diag(gm) <- gam1;
+            A1       <- gm %*% A0;
+            g0       <- matrix(data = 0, nrow = sp_try[i], ncol = sp_try[i]);
+            diag(g0) <- -1 * mean(diag(A1)); # Note: This standardisation will
+            A0       <- g0 %*% A0; # not affect stability, but I think helps
+            A0_stb   <- max(Re(eigen(A0)$values)) < 0; # readers' interpretation
             A1_stb   <- max(Re(eigen(A1)$values)) < 0;
             A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
             A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;

### 2020-03-22 notebook/SI.Rmd
Add the different gamma distribution stability results to the SI

@@
 
-The same 100000 $\mathbf{M}$ matrices were used to investigate stability when applying each of these different distributions of $\gamma$ values.
-The table below shows the number of $\mathbf{M}$ that were unstable (_unst) and stable (_stbl) for the exponential (Exp), beta, and gamma distributions.
+The stability of $\mathbf{A}$ versus $\mathbf{M}$ was investigated for each of the distributions of $\gamma$ shown in panels b-d above.
+The table below shows the number of $\mathbf{A}$ versus $\mathbf{M}$ that were stable for the exponential (exp), beta, and gamma distributions.
 
 ```{r, echo = FALSE}
-fourdists <- read.csv(file = "sim_results/different_distr/four_distr_rand.csv");
-kable(fourdists);
+g_dsts <- read.csv("../inst/extdata/g_dsts.csv");
+kable(g_dsts);
 ```

### 2020-03-22 notebook/SI.tex
Add the different gamma distribution stability results to the SI

@@
 
-The same 100000 \(\mathbf{M}\) matrices were used to investigate
-stability when applying each of these different distributions of
-\(\gamma\) values. The table below shows the number of \(\mathbf{M}\)
-that were unstable (\_unst) and stable (\_stbl) for the exponential
-(Exp), beta, and gamma distributions.
+The stability of \(\mathbf{A}\) versus \(\mathbf{M}\) was investigated
+for each of the distributions of \(\gamma\) shown in panels b-d above.
+The table below shows the number of \(\mathbf{A}\) versus \(\mathbf{M}\)
+that were stable for the exponential (exp), beta, and gamma
+distributions.
 
@@
 \toprule
-S & Exp\_unst & Exp\_stbl & beta\_unst & beta\_stbl & gamma\_unst &
-gamma\_stbl\tabularnewline
+S & exp\_A & exp\_M & beta\_A & beta\_M & gamma\_A &
+gamma\_M\tabularnewline
 \midrule
 \endhead
-2 & 30 & 99970 & 30 & 99970 & 30 & 99970\tabularnewline
-3 & 355 & 99645 & 355 & 99645 & 355 & 99645\tabularnewline
-4 & 1506 & 98494 & 1512 & 98488 & 1516 & 98484\tabularnewline
-5 & 3930 & 96070 & 3971 & 96029 & 4006 & 95994\tabularnewline
-6 & 7738 & 92262 & 7844 & 92156 & 7918 & 92082\tabularnewline
-7 & 13606 & 86394 & 13889 & 86111 & 13990 & 86010\tabularnewline
-8 & 20535 & 79465 & 21002 & 78998 & 21114 & 78886\tabularnewline
-9 & 28614 & 71386 & 29060 & 70940 & 29110 & 70890\tabularnewline
-10 & 38375 & 61625 & 38388 & 61612 & 38441 & 61559\tabularnewline
-11 & 48616 & 51384 & 48211 & 51789 & 47957 & 52043\tabularnewline
-12 & 59254 & 40746 & 58025 & 41975 & 57473 & 42527\tabularnewline
-13 & 68816 & 31184 & 66753 & 33247 & 66127 & 33873\tabularnewline
-14 & 77721 & 22279 & 75149 & 24851 & 74222 & 25778\tabularnewline
-15 & 84842 & 15158 & 82030 & 17970 & 81040 & 18960\tabularnewline
-16 & 90365 & 9635 & 87809 & 12191 & 86600 & 13400\tabularnewline
-17 & 94171 & 5829 & 91756 & 8244 & 90668 & 9332\tabularnewline
-18 & 96978 & 3022 & 94977 & 5023 & 94176 & 5824\tabularnewline
-19 & 98376 & 1624 & 97018 & 2982 & 96268 & 3732\tabularnewline
-20 & 99218 & 782 & 98357 & 1643 & 97765 & 2235\tabularnewline
-21 & 99678 & 322 & 99124 & 876 & 98746 & 1254\tabularnewline
-22 & 99864 & 136 & 99599 & 401 & 99323 & 677\tabularnewline
-23 & 99954 & 46 & 99783 & 217 & 99668 & 332\tabularnewline
-24 & 99978 & 22 & 99920 & 80 & 99821 & 179\tabularnewline
-25 & 99996 & 4 & 99967 & 33 & 99911 & 89\tabularnewline
-26 & 99999 & 1 & 99979 & 21 & 99960 & 40\tabularnewline
-27 & 99999 & 1 & 99990 & 10 & 99983 & 17\tabularnewline
-28 & 100000 & 0 & 99999 & 1 & 99991 & 9\tabularnewline
-29 & 100000 & 0 & 99999 & 1 & 99999 & 1\tabularnewline
-30 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
-31 & 100000 & 0 & 100000 & 0 & 99999 & 1\tabularnewline
-32 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+2 & 99965 & 99965 & 99974 & 99974 & 99977 & 99977\tabularnewline
+3 & 99636 & 99635 & 99650 & 99648 & 99628 & 99628\tabularnewline
+4 & 98576 & 98564 & 98482 & 98470 & 98508 & 98492\tabularnewline
+5 & 96053 & 95971 & 96156 & 96096 & 96068 & 96004\tabularnewline
+6 & 92036 & 91867 & 92104 & 91927 & 92233 & 92029\tabularnewline
+7 & 86667 & 86333 & 86456 & 86070 & 86604 & 86161\tabularnewline
+8 & 79670 & 79153 & 79392 & 78822 & 79393 & 78771\tabularnewline
+9 & 71389 & 70911 & 70998 & 70529 & 71070 & 70548\tabularnewline
+10 & 61674 & 61609 & 61794 & 61586 & 61265 & 61093\tabularnewline
+11 & 51150 & 51935 & 51352 & 51924 & 51313 & 51951\tabularnewline
+12 & 41209 & 42925 & 40954 & 42670 & 40708 & 42183\tabularnewline
+13 & 30827 & 33462 & 30969 & 33770 & 31046 & 33522\tabularnewline
+14 & 22203 & 25767 & 22208 & 25629 & 22342 & 25435\tabularnewline
+15 & 15003 & 18877 & 15206 & 18913 & 15025 & 18464\tabularnewline
+16 & 9613 & 13372 & 9504 & 13357 & 9418 & 12737\tabularnewline
+17 & 5579 & 8967 & 5570 & 8976 & 5719 & 8487\tabularnewline
+18 & 3104 & 5833 & 3048 & 5853 & 3060 & 5447\tabularnewline
+19 & 1516 & 3578 & 1553 & 3633 & 1600 & 3185\tabularnewline
+20 & 717 & 2067 & 799 & 2179 & 769 & 1862\tabularnewline
+21 & 312 & 1196 & 310 & 1200 & 331 & 1039\tabularnewline
+22 & 129 & 643 & 128 & 654 & 135 & 510\tabularnewline
+23 & 48 & 321 & 48 & 359 & 57 & 242\tabularnewline
+24 & 11 & 161 & 19 & 159 & 20 & 120\tabularnewline
+25 & 1 & 59 & 5 & 81 & 7 & 45\tabularnewline
+26 & 0 & 30 & 0 & 48 & 0 & 22\tabularnewline
+27 & 0 & 10 & 0 & 16 & 0 & 6\tabularnewline
+28 & 1 & 3 & 2 & 2 & 0 & 3\tabularnewline
+29 & 0 & 2 & 0 & 0 & 0 & 0\tabularnewline
+30 & 0 & 0 & 0 & 1 & 0 & 0\tabularnewline
+31 & 0 & 0 & 0 & 1 & 0 & 0\tabularnewline
+32 & 0 & 0 & 0 & 0 & 0 & 0\tabularnewline
 \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} & \ldots{} &
 \ldots{}\tabularnewline
-50 & 100000 & 0 & 100000 & 0 & 100000 & 0\tabularnewline
+50 & 0 & 0 & 0 & 0 & 0 & 0\tabularnewline
 \bottomrule

### 2020-03-23 notebook/sci_rep_reviews.Rmd
Revise response to reviewers, with line numbers, up to Reviewer 3

@@
 ---
-
-I have taken the following actions:
-
-1.
-Re-run simulations while calculating the correlation between off-diagonal elements of A[i,j] and A[j,i] to see if the correlation changes with $\gamma$ across $S$.
-It does not.
-I also generated systems of known correlation coefficients to assess how this modulated $\gamma$'s affect on stability.
-It did not.
-
-2. Run simulations across multiple structured network types to confirm the robustness of results. Network types include (a) cascade food webs, (b) small world networks, and (c) scale-free network.
-
-3. Rewrite the manuscript, or the relevant parts of it.
 
@@
 I again thank the reviewer for their comments.
-I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (line XXX).
+I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (L48, L88, L209, L234).
+In particular, in addition to the statement on L48, I now state more explicitly in describing the methods that "the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system." (L88)
 
@@
 I agree with the relevance of Dougoud et al.
-(2018), and have made appropriate links to their results on feasibility.
-Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility.
-Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic systems in which all population densities are positive).
+(2018), and have made appropriate links to their results on feasibility (L143, L274).
+Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility (L148).
+Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic ecological communities in which all population densities are positive).
 Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application.
@@
 I welcome Reviewer 1's call for expanding the scope of models considered for simulations, which was also recommended by Reviewer 2.
-Nevertheless, my manuscript focuses broadly on complex systems rather than pure ecological systems, so I have revised my manuscript to test more general models of common network structures.
+Nevertheless, my manuscript focuses broadly on complex systems rather than only on pure ecological systems, so I have revised my manuscript to test more general models of common network structures.
 In terms of ecological models, in addition to random competition, mutualism, and predator-prey networks, I have now also included a cascade food web model (which is a foundation for the extended niche and nested-heirarchy models that Reviewer 1 mentions; Williams and Martinez 2000).
@@
 Overall, the structured networks that I now include in the main text and supporting information encompass a wide range of physical, biological, and social systems.
+Results for thse simulations are presented in a new section entitled, "Simulations of structured M across S" (L123).
+Methods for constructing small-world (L247), scale-free (L260), and cascade food web (L268) networks are now explained, with additional details and figures now in the Supplemental Information.
+The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity.
 
-More emphasis on feasibility and how it relates to ecological systems has now been included (line XXX).
-I have also added new sections in the Results (line XXX) and methods (line XXX) for the new simulations of structured networks.
+Feasibility and how it relates to ecological systems is explained on L141, with discussion starting on L188 for how this relates to ecological systems.
 Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
@@
 
-I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (line XXX).
+I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L417).
 
@@
 
-I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (line XXX).
+I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the covariation of M (L65-L81).
+I agree that these analytical results improve the manuscript, and I think that they are particularly useful in setting the stage for the numerical simulations that now follow.
 In making these changes, I have slightly adjusted the notation so that the text is more readable.
 
-As Reviewer 2 correctly points out, my focus is on finite values of S and \sigma, in which complexity is gradually increased until it is dangerously close to the stability boundary.
-Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases the stability of $M$), would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$.
+As Reviewer 2 correctly points out, my focus is on finite values of S and $\sigma$, in which complexity is gradually increased until it is dangerously close to the stability boundary.
+Given this focus, analytically demonstrating my central result (that variation in $X$ when $M = XA$ increases the probability of finding a stable $M$) would require some sort of comparison between the probability distributions of eigenvalues for $M$ and $A$.
 For circular matrices in which elements are i.i.d ($A$), it has been proved that eigenvalues are distributed uniformly within a known boundary of support of the spectral density (Tao and Vu 2010).
@@
 This likely matters because for finite S, even though the real boundary of the support might be positive for both $M$ and $A$ (implying instability as $S \to \infty$; Grilli et al.
-2017), the right most eigenvalue for a system still might be less than zero.
+2017), the right-most eigenvalue for a system still might be less than zero.
 
-Reviewer 2 is correct that this is uncharted territory, and after carefully considering the possibility of analytical demonstration of my manuscript's results, I am convinced that such a demonstration is beyond the scope of this manuscript.
+Reviewer 2 is correct that this is uncharted territory, and after spending a lot of time considering the possibility of an analytical demonstration of my main results, I am convinced that such a demonstration is beyond the scope of this manuscript.
 My approach follows others with similar objectives.
@@
 
-These were very helpful comments, and I am especially grateful to Reviewer 2 for doing a bit of preliminary exploration themselves to suggest improvements.
-I have now run simulations for different values of $\rho$, as suggested.
+These were very helpful comments, and I have embraced them wholeheartedly in my revision.
+I am especially grateful to Reviewer 2 for doing a bit of preliminary exploration themselves to suggest improvements.
+I have now included a full section devoted to the role of $\rho$ on stability, and run simulations for different values of $\rho$, as suggested.
 Reviewer 2 is correct that the nonzero $\rho$ values do not change my results, but an analysis of stability across $\rho$ values also revealed an interesting pattern.
@@
 
-As suggested, I also considered matrices with even more structure.
+As suggested, I also considered matrices with even more structure, performing a new simulations on each.
 Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems.
 These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002).
+Results for these networks are presented in L123-L140, and methods are explained in L247-L273, and in the Supplemental Information.
+Overall, these structured networks were affected by variation in $\gamma$ in the same way as random networks (increasing stability given sufficiently high complexity).
 
@@
 
-I have revised the writing to more carefully frame the manuscript's key results (line XXX).
-This includes a change of the title, which now makes it more clear that the effects on stability are small while retaining an emphasis on generality across systems.
-I have also removed the word "markedly" from the abstract (line XXX).
+As recommended, I have changed the title, which now makes it more clear that the effects on stability are restricted to highly complex finite systems, while still retaining an emphasis on generality across systems.
+I have also removed the word "markedly" from the abstract, as recommended.
 

### 2020-03-23 notebook/ms.Rmd
Change the title, as requested by Reviewer 2
Insert two additional references that were recommended by reviewer 3

@@
 ---
-title: "Component response rate variation underlies the stability of complex systems"
+title: "Component response rate variation underlies the stability of highly complex finite systems"
 author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
@@
 Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$) [@Allesina2012].
-May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks.
+May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Townsend2010a; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks.
 As such, identifying general principles that affect stability in complex systems is of wide-ranging importance.
@@
 The value of $\Re(\lambda_{max})$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b].
-These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the variance of between component interaction strengths (hereafter $\sigma^{2}_{M}$, to distinguish from $\sigma^{2}_{A}$ and $\sigma^{2}_{\gamma}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$).
+These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the variance of between component interaction strengths (hereafter $\sigma^{2}_{M}$, to distinguish from $\sigma^{2}_{A}$ and $\sigma^{2}_{\gamma}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$) [@Sommers1988].
 Positive $\sigma^{2}_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$.

### 2020-03-23 notebook/ms.tex
Insert two additional references that were recommended by reviewer 3

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
+            pdftitle={Component response rate variation underlies the stability of highly complex finite systems},
             pdfauthor={A. Bradley Duthie ( alexander.duthie@stir.ac.uk )},
@@
 \setlength{\droptitle}{-2em}
-  \title{Component response rate variation underlies the stability of complex
-systems}
+  \title{Component response rate variation underlies the stability of highly
+complex finite systems}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
 as
-ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{5}},
-neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{6},\protect\hyperlink{ref-Gray2009}{7}},
-biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{8},\protect\hyperlink{ref-MacArthur2010}{9}},
+ecological\textsuperscript{\protect\hyperlink{ref-May1972}{1}--\protect\hyperlink{ref-Grilli2017}{6}},
+neurological\textsuperscript{\protect\hyperlink{ref-Gray2008}{7},\protect\hyperlink{ref-Gray2009}{8}},
+biochemical\textsuperscript{\protect\hyperlink{ref-Rosenfeld2009}{9},\protect\hyperlink{ref-MacArthur2010}{10}},
 and
-socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{10}--\protect\hyperlink{ref-Bardoscia2017}{13}}
+socio-economic\textsuperscript{\protect\hyperlink{ref-May2008}{11}--\protect\hyperlink{ref-Bardoscia2017}{14}}
 networks. As such, identifying general principles that affect stability
@@
 species\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}} or
-banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{11}}). One
+banks\textsuperscript{\protect\hyperlink{ref-Haldane2011}{12}}). One
 element of such a matrix, \(M_{ij}\), defines how component \(j\)
@@
 \(\sigma^{2}\); diagonal elements are set to
-\(-1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
+\(-1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{5}}.
 Local system stability is assessed using eigenanalysis on
@@
 In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{15}} within a
 circle centred at \(\Re = -d\) (\(-d\) is the mean value of diagonal
 elements) and \(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{5}}
 (Figs 1a and 2a). Local stability of randomly assembled systems
@@
 random\textsuperscript{\protect\hyperlink{ref-May1972}{1}},
-small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}},
-scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}},
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}},
 cascade food
-web\textsuperscript{\protect\hyperlink{ref-Solow1998}{17},\protect\hyperlink{ref-Williams2000}{18}}
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{18},\protect\hyperlink{ref-Williams2000}{19}}
 networks.
@@
 system \(\mathbf{M}\) can be defined as
-follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}},
+follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}},
 
@@
 \(\mathbf{M}\) are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} in a
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{15}} in a
 circle centred at \((-d, 0)\) with a radius of
@@
 system stability, can be estimated from five properties of
-\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{20}}.
+\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{21}}.
 These properties include (1) system size (\(S\)), (2) mean
@@
 of interaction strengths between components, \(M_{ij}\) and \(M_{ji}\)
-(\(\rho\)). Positive \(\sigma^{2}_{\gamma}\) does not change \(S\), nor
-does it necessarily change \(E[d]\) or \(E[\mu]\). What
-\(\sigma^{2}_{\gamma}\) does change is the total variation in component
-interaction strengths (\(\sigma^{2}_{M}\)), and \(\rho\). Introducing
-variation in \(\gamma\) increases the total variation in the system.
-Variation in the off-diagonal elements of \(\mathbf{M}\) is described by
-the joint variation of two random variables,
+(\(\rho\))\textsuperscript{\protect\hyperlink{ref-Sommers1988}{22}}.
+Positive \(\sigma^{2}_{\gamma}\) does not change \(S\), nor does it
+necessarily change \(E[d]\) or \(E[\mu]\). What \(\sigma^{2}_{\gamma}\)
+does change is the total variation in component interaction strengths
+(\(\sigma^{2}_{M}\)), and \(\rho\). Introducing variation in \(\gamma\)
+increases the total variation in the system. Variation in the
+off-diagonal elements of \(\mathbf{M}\) is described by the joint
+variation of two random variables,
 
@@
 \(d \neq 0\), positive \(\sigma^{2}_\gamma\) changes the distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}--\protect\hyperlink{ref-Stone2017}{23}},
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{23}--\protect\hyperlink{ref-Stone2017}{25}},
 potentially affecting stability (Fig. 1c vs.~1d).
@@
 linearly with \(\rho\) such
-that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{24}},
+that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{26}},
 
@@
 of interspecific ecological
-interactions\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{25}}.
+interactions\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{27}}.
 These interaction types are competitive, mutualist, and predator-prey,
@@
 \(\mathbf{M = \gamma A}\) for
-small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}},
-scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}, and
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}, and
 cascade food
-web\textsuperscript{\protect\hyperlink{ref-Solow1998}{17},\protect\hyperlink{ref-Williams2000}{18}}
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{18},\protect\hyperlink{ref-Williams2000}{19}}
 networks. In all of these networks, rules determining the presence or
@@
 degree of separation between any two components increases in proportion
-to \(\log(S)\)\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}.
+to \(\log(S)\)\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}.
 In scale-free networks, the distribution of the number of components
@@
 components have many interactions while most components have few
-interactions\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. In
+interactions\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. In
 cascade food webs, species are ranked and interactions are constrained
@@
 components are positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{26},\protect\hyperlink{ref-Song2018}{27}}.
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{6},\protect\hyperlink{ref-Dougoud2018}{28},\protect\hyperlink{ref-Song2018}{29}}.
 This is of particular interest for ecological communities because
@@
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}}.
 While my results are intended to be general to all complex systems, and
@@
 \(1/2^{S}\), consistent with a recent proof by Serván et
-al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}} (see
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{30}} (see
 Supplementary Information). Hence, for pure interacting species
@@
 values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}.
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{31}}.
 Generations of selection on \(\gamma\) value combinations to minimise
@@
 My focus here is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}}, who applied
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24}}, who applied
 the same mathematical framework to investigate how a diagonal matrix
@@
 model, where \(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22}} analytically
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24}} analytically
 demonstrated that the effect of \(\mathbf{X}\) on system stability
@@
 of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{23},\protect\hyperlink{ref-Dougoud2018}{26},\protect\hyperlink{ref-Servan2018}{28}}
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{6},\protect\hyperlink{ref-Stone2017}{25},\protect\hyperlink{ref-Dougoud2018}{28},\protect\hyperlink{ref-Servan2018}{30}}
 because population densities cannot realistically be negative. My
@@
 empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}, but
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, but
 instead interact with evolutionary, abiotic, or social-economic systems.
@@
 been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{21}}
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{23}}
 and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{22},\protect\hyperlink{ref-Stone2017}{23}},
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24},\protect\hyperlink{ref-Stone2017}{25}},
 but \(\boldsymbol{\gamma}\) has not been interpreted as rates of
@@
 traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}.
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
 For any species \(i\) or trait \(j\), change in species density
@@
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}, and
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{19}}.
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
 
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{30}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{32}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{30}},
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{32}},
 
@@
 values of system components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{21},\protect\hyperlink{ref-May1973}{30}}.
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{23},\protect\hyperlink{ref-May1973}{32}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 programming language for all numerical simulations and
-analyses\textsuperscript{\protect\hyperlink{ref-Rproject}{31}}. Purely
+analyses\textsuperscript{\protect\hyperlink{ref-Rproject}{33}}. Purely
 random networks were generated by sampling off-diagonal elements from an
@@
 Small-world networks were constructed using the method of Watts and
-Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}. First,
-a regular network\textsuperscript{\protect\hyperlink{ref-Watts1998}{15}}
+Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}. First,
+a regular network\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}
 was created such that components were arranged in a circle. Each
@@
 Scale-free networks were constructed using the method of Albert and
-Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. First,
+Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. First,
 a saturated network (all components interact with each other) of size
@@
 components had followed a power-law
-tail\textsuperscript{\protect\hyperlink{ref-Albert2002}{16}}. The
+tail\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. The
 resulting network was represented by an \(S \times S\) binary matrix
@@
 Cascade food webs were constructed following Solow and
-Beet\textsuperscript{\protect\hyperlink{ref-Solow1998}{17}}. First, a
+Beet\textsuperscript{\protect\hyperlink{ref-Solow1998}{18}}. First, a
 random matrix \(\mathbf{A}\) was generated with off-diagonal elements
@@
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{26}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 reduced to the below (see also Serván et
-al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{28}}),
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{30}}),
 
@@
 practical solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{29}}. In
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{31}}. In
 this case, the practical solution is finding vectors of
@@
 
+\hypertarget{ref-Townsend2010a}{}
+3. Townsend, S. E., Haydon, D. T. \& Matthews, L. On the generality of
+stability-complexity relationships in Lotka-Volterra ecosystems.
+\emph{Journal of Theoretical Biology} \textbf{267,} 243--251 (2010).
+
 \hypertarget{ref-Mougi2012}{}
-3. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
+4. Mougi, A. \& Kondoh, M. Diversity of interaction types and ecological
 community stability. \emph{Science} \textbf{337,} 349--351 (2012).
@@
 \hypertarget{ref-Allesina2015}{}
-4. Allesina, S. \emph{et al.} Predicting the stability of large
+5. Allesina, S. \emph{et al.} Predicting the stability of large
 structured food webs. \emph{Nature Communications} \textbf{6,} 7842
@@
 \hypertarget{ref-Grilli2017}{}
-5. Grilli, J. \emph{et al.} Feasibility and coexistence of large
+6. Grilli, J. \emph{et al.} Feasibility and coexistence of large
 ecological communities. \emph{Nature Communications} \textbf{8,} (2017).
@@
 \hypertarget{ref-Gray2008}{}
-6. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
+7. Gray, R. T. \& Robinson, P. A. Stability and synchronization of
 random brain networks with a distribution of connection strengths.
@@
 \hypertarget{ref-Gray2009}{}
-7. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
+8. Gray, R. T. \& Robinson, P. A. Stability of random brain networks
 with excitatory and inhibitory connections. \emph{Neurocomputing}
@@
 \hypertarget{ref-Rosenfeld2009}{}
-8. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
+9. Rosenfeld, S. Patterns of stochastic behavior in dynamically unstable
 high-dimensional biochemical networks. \emph{Gene Regulation and Systems
@@
 \hypertarget{ref-MacArthur2010}{}
-9. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
+10. MacArthur, B. D., Sanchez-Garcia, R. J. \& Ma'ayan, A. Microdynamics
 and criticality of adaptive regulatory networks. \emph{Physics Review
@@
 \hypertarget{ref-May2008}{}
-10. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
+11. May, R. M., Levin, S. A. \& Sugihara, G. Complex systems: Ecology
 for bankers. \emph{Nature} \textbf{451,} 893--895 (2008).
@@
 \hypertarget{ref-Haldane2011}{}
-11. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
+12. Haldane, A. G. \& May, R. M. Systemic risk in banking ecosystems.
 \emph{Nature} \textbf{469,} 351--355 (2011).
@@
 \hypertarget{ref-Suweis2014}{}
-12. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
+13. Suweis, S. \& D'Odorico, P. Early warning signs in social-ecological
 networks. \emph{PLoS ONE} \textbf{9,} (2014).
@@
 \hypertarget{ref-Bardoscia2017}{}
-13. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
+14. Bardoscia, M., Battiston, S., Caccioli, F. \& Caldarelli, G.
 Pathways towards instability in financial networks. \emph{Nature
@@
 \hypertarget{ref-Tao2010}{}
-14. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
+15. Tao, T. \& Vu, V. Random matrices: Universality of ESDs and the
 circular law. \emph{Annals of Probability} \textbf{38,} 2023--2065
@@
 \hypertarget{ref-Watts1998}{}
-15. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
+16. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
 networks. \emph{Nature} \textbf{393,} 440--442 (1998).
@@
 \hypertarget{ref-Albert2002}{}
-16. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
+17. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
 networks. \emph{Reviews of Modern Physics} \textbf{74,} 47--97 (2002).
@@
 \hypertarget{ref-Solow1998}{}
-17. Solow, A. R. \& Beet, A. R. On lumping species in food webs.
+18. Solow, A. R. \& Beet, A. R. On lumping species in food webs.
 \emph{Ecology} \textbf{79,} 2013--2018 (1998).
@@
 \hypertarget{ref-Williams2000}{}
-18. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
+19. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
 webs. \emph{Nature} \textbf{404,} 180--183 (2000).
@@
 \hypertarget{ref-Patel2018}{}
-19. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+20. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 \hypertarget{ref-Tang2014b}{}
-20. Tang, S. \& Allesina, S. Reactivity and stability of large
+21. Tang, S. \& Allesina, S. Reactivity and stability of large
 ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
@@
 
+\hypertarget{ref-Sommers1988}{}
+22. Sommers, H. J., Crisanti, A., Sompolinsky, H. \& Stein, Y. Spectrum
+of large random asymmetric matrices. \emph{Physical Review Letters}
+\textbf{60,} 1895--1898 (1988).
+
 \hypertarget{ref-Ahmadian2015}{}
-21. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+23. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
 with partially structured and partially random connectivity.
@@
 \hypertarget{ref-Gibbs2017}{}
-22. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+24. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.
@@
 \hypertarget{ref-Stone2017}{}
-23. Stone, L. The feasibility and stability of large complex biological
+25. Stone, L. The feasibility and stability of large complex biological
 networks: a random matrix approach. \emph{Scientific Reports}
@@
 \hypertarget{ref-Tang2014c}{}
-24. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
+26. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
 strengths drives stability in large ecological networks. \textbf{17,}
@@
 \hypertarget{ref-Allesina2011}{}
-25. Allesina, S. \& Levine, J. M. A competitive network theory of
+27. Allesina, S. \& Levine, J. M. A competitive network theory of
 species diversity. \emph{Proceedings of the National Academy of Sciences
@@
 \hypertarget{ref-Dougoud2018}{}
-26. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+28. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-27. Song, C. \& Saavedra, S. Will a small randomly assembled community
+29. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-28. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+30. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 \hypertarget{ref-Hamblin2013}{}
-29. Hamblin, S. On the practical usage of genetic algorithms in ecology
+31. Hamblin, S. On the practical usage of genetic algorithms in ecology
 and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
@@
 \hypertarget{ref-May1973}{}
-30. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+32. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
@@
 \hypertarget{ref-Rproject}{}
-31. R Core Team. \emph{R: A language and environment for statistical
+33. R Core Team. \emph{R: A language and environment for statistical
 computing}. (R Foundation for Statistical Computing, 2018).
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.13.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed

### 2020-03-23 notebook/sci_rep_reviews.Rmd
Adjust the line numbers for the new revision and finish addressing the comments from Reviewer 3

@@
 
-I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L417).
+I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L420).
 
@@
 If individual components are not self-regulating, then the system will almost never be stable.
-I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI, which was subsequently improved prior to submission: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
+I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI, which was subsequently improved prior to my original submission: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
 The effect of $\gamma$ on increasing stability is retained, but stability of all systems drops to zero out of one million given size $S > 14$.
@@
 
-I have tried to clarify and more clearly state the assumption of self-regulation in my revision (line XXX).
+I have tried to clarify and more clearly state the assumption of self-regulation in my revision (paragraphs starting on L28 and L65).
 
@@
 
-I have tried to clarify the role of random matrices as a null model of complex system stability (line XXX).
+I have tried to clarify the role of random matrices as a null model of complex system stability (L53-L64).
 
@@
 Hence, I am not comparing apples and oranges; response rates of components have a clear and important role in complex system stability that can be investigated with the mathematical techniques used in this manuscript.
-I have tried to explain this role more clearly in my revision (line XXX).
+I have tried to explain this role more clearly in my revision by more clearly presenting results in terms of differences between the matrices $A$ and $M$ (this affected many areas throughout the text and SI, with clarification of the comparison presented in L82-L89).
 
@@
 
-I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (lines XXX).
+I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (L25, L69, L323, L325, and L367).
 

### 2020-03-23 vignettes/R_old/evolve_gammas.R
Redo the vignettes to make room for all of the revised text

@@ -1,156 +1,0 @@
-Evo_rand_gen_var <- function(from, to, iters, int_type = 0, rmx = 0.4, C = 1){
-    tot_res <- NULL;
-    fea_res <- NULL;
-    if(from >= to){
-        stop("Argument 'from' must be less than argument 'to'");
-    }
-    for(i in from:to){
-        nn             <- i;
-        A1_stt         <- 0;
-        A2_stt         <- 0;
-        A1_fet         <- 0;
-        A2_fet         <- 0;
-        iter           <- iters;
-        tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 3);
-        fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 2);
-        while(iter > 0){
-            r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
-            A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
-            A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
-            A0       <- species_interactions(mat = A0, type = int_type);
-            C_dat    <- rbinom(n = i * i, size = 1, prob = C);
-            C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
-            A0       <- A0 * C_mat;
-            diag(A0) <- -1;
-            gam1     <- runif(n = i, min = 0, max = 2);
-            A1       <- A0 * gam1;
-            A0_stb   <- max(Re(eigen(A0)$values)) < 0;
-            A1_stb   <- rand_mat_ga(A1);
-            A0_fea   <- min(-1*solve(A0) %*% r_vec) > 0;
-            A1_fea   <- min(-1*solve(A1) %*% r_vec) > 0;
-            if(A0_stb == TRUE){
-                tot_res[[i-1]][iter, 1] <- 1;
-            }
-            if(A1_stb == TRUE){
-                tot_res[[i-1]][iter, 2] <- 1;
-            }
-            if(A0_fea == TRUE){
-                fea_res[[i-1]][iter, 1] <- 1;
-            }
-            if(A1_fea == TRUE){
-                fea_res[[i-1]][iter, 2] <- 1;
-            }
-            iter    <- iter - 1;
-        }
-        print(i);
-    }
[…]

### 2020-03-23 vignettes/R_old/find_bi_gamma.R

@@ -1,108 +1,0 @@
-stab_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
-    ress     <- matrix(data = 0, nrow = iters, ncol = 2);
-    A0_count <- 0;
-    A1_count <- 0;
-    while(iters > 0){
-        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
-        A_mat  <- matrix(data = A_dat, nrow = S);
-        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
-        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
-        A_mat  <- A_mat * C_mat;
-        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
-        mu_gam <- mean(gammas);
-        diag(A_mat) <- -1;
-        A1     <- gammas * A_mat;
-        A0     <- mu_gam * A_mat;
-        A0_e   <- eigen(A0)$values;
-        A0_r   <- Re(A0_e);
-        A0_i   <- Im(A0_e);
-        A1_e   <- eigen(A1)$values;
-        A1_r   <- Re(A1_e);
-        A1_i   <- Im(A1_e);
-        if(max(A0_r) < 0){
-            ress[iters, 1] <- 1;
-            A0_count       <- A0_count + 1;
-            }
-        if(max(A1_r) < 0){
-            ress[iters, 2] <- 1;
-            A1_count       <- A1_count + 1;
-        }
-        print(c(iters, A0_count, A1_count));
-        iters <- iters - 1;
-    }
-    return(ress);
-}
-
-
-find_bgamma <- function(S = 200, C = 0.05, Osd = 0.4, iters = 10000){
-    while(iters > 0){
-        A_dat  <- rnorm(n = S * S, mean = 0, sd = Osd);
-        A_mat  <- matrix(data = A_dat, nrow = S);
-        C_dat  <- rbinom(n = S * S, size = 1, prob = C);
-        C_mat  <- matrix(data = C_dat, nrow = S, ncol = S);
-        A_mat  <- A_mat * C_mat;
-        gammas <- c(rep(1.95, S/2), rep(0.05, S/2))
-        mu_gam <- mean(gammas);
-        diag(A_mat) <- -1;
[…]

### 2020-03-23 vignettes/R_old/plot_figs.R

@@ -1,397 +1,0 @@
- make_gammas <- function(nn = 10, distribution = 1, mn = 1, sdd = 1){
-    if(distribution == 0){
-        dat          <- rep(x = mn, times = nn);
-    }
-    if(distribution == 1){
-        mval         <- sdd * sqrt(12);
-        dat          <- runif(n = nn, min = 0, max = mval);
-    }
-    if(distribution == 2){
-        dat          <- rexp(n = nn);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 3){
-        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    if(distribution == 4){
-        dat          <- rgamma(n = nn, shape = 2, scale = 2);
-        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
-        dat          <- dat - min(dat) + min(abs(dat));
-    }
-    return(dat);
-}
-
-show_gammas <- function(nn = 1000000, sdd = 1, mn = 10){
-    y1 <- make_gammas(nn = nn, distribution = 1, sdd = sdd, mn = mn);
-    y2 <- make_gammas(nn = nn, distribution = 2, sdd = sdd, mn = mn);
-    y3 <- make_gammas(nn = nn, distribution = 3, sdd = sdd, mn = mn);
-    y4 <- make_gammas(nn = nn, distribution = 4, sdd = sdd, mn = mn);
-    par(mfrow = c(2, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
-    h1 <- hist(y1, breaks = 1000, plot = FALSE);
-    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5,
-         ylim = c(0, max(h1$counts)*1.2));
-    mtext("a", adj = 0, line = -2.3, cex = 2, 
-          at = par("usr")[1]+0.90*diff(par("usr")[1:2]));
-    box();
[…]

### 2020-03-23 vignettes/R_old/scratch.R

@@ -1,706 +1,0 @@
-
-
-eout <- read.csv("Evo_out.txt");
-
-dat <- eout[eout[,1] == 40,];
-
-
-
-rand_gen_var <- function(max_sp, iters, int_type = 0, rmx = 0.4, gamma_sd = 1, 
-                         gamma_mn = 1, C = 1){
-  tot_res  <- NULL;
-  fea_res  <- NULL;
-  vars_t   <- NULL;
-  off_cors <- NULL;
-  for(i in 2:max_sp){
-    iter           <- iters;
-    tot_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
-    fea_res[[i-1]] <- matrix(data = 0, nrow = iter, ncol = 7);
-    while(iter > 0){
-      r_vec    <- rnorm(n = i, mean = 0, sd = rmx);
-      A0_dat   <- rnorm(n = i * i, mean = 0, sd = 0.4);
-      A0       <- matrix(data = A0_dat, nrow = i, ncol = i);
-      A0       <- species_interactions(mat = A0, type = int_type);
-      C_dat    <- rbinom(n = i * i, size = 1, prob = C);
-      C_mat    <- matrix(data = C_dat, nrow = i, ncol = i);
-      A0       <- A0 * C_mat;
-      diag(A0) <- -1;
-      gam0     <- make_gammas(nn = i, distribution = 0, sdd = gamma_sd);
-      
-      gam1     <- runif(n = i, min = 0, max = 999) #make_gammas(nn = i, distribution = 1, sdd = gamma_sd);
-      gam0     <- rep(mean(gam1), i);
-      
-      gam2     <- make_gammas(nn = i, distribution = 2, sdd = gamma_sd);
-      gam3     <- make_gammas(nn = i, distribution = 3, sdd = gamma_sd);
-      gam4     <- make_gammas(nn = i, distribution = 4, sdd = gamma_sd);
-      
-      A1       <- A0 * gam1;
-      A2       <- A0 * gam1;
-      A3       <- A0 * gam0;
-      diag(A2) <- -1 * gam0;
-      diag(A3) <- -1 * gam1;
-      
-
[…]

### 2020-03-23 vignettes/R_old/sim_mat.R

@@ -1,95 +1,0 @@
-species_interactions <- function(mat, type = 0){
-    if(type == 1){
-        mat[mat > 0] <- -1*mat[mat > 0];
-    }
-    if(type == 2){
-        mat[mat < 0] <- -1*mat[mat < 0];
-    }
-    if(type == 3){
-        for(i in 1:dim(mat)[1]){
-            for(j in 1:dim(mat)[2]){
-                if(mat[i, j] * mat[j, i] > 0){
-                    mat[j, i] <- -1 * mat[j, i];
-                }
-            }
-        }
-    }
-    return(mat);
-}
-
-summarise_randmat <- function(tot_res, fea_res){
-    sims    <- length(tot_res);
-    all_res <- matrix(data = 0, nrow = sims, ncol = 13);
-    for(i in 1:sims){
-        all_res[i, 1]  <- i + 1;
-        # Stable and unstable
-        all_res[i, 2]  <- sum(tot_res[[i]][,1] == FALSE);
-        all_res[i, 3]  <- sum(tot_res[[i]][,1] == TRUE);
-        all_res[i, 4]  <- sum(tot_res[[i]][,2] == FALSE);
-        all_res[i, 5]  <- sum(tot_res[[i]][,2] == TRUE);
-        # Stabilised and destabilised
-        all_res[i, 6] <- sum(tot_res[[i]][,1] == FALSE & 
-                                  tot_res[[i]][,2] == TRUE);
-        all_res[i, 7] <- sum(tot_res[[i]][,1] == TRUE & 
-                                  tot_res[[i]][,2] == FALSE);
-        # Feasible and infeasible
-        all_res[i, 8]  <- sum(fea_res[[i]][,1] == FALSE);
-        all_res[i, 9]  <- sum(fea_res[[i]][,1] == TRUE);
-        all_res[i, 10]  <- sum(fea_res[[i]][,2] == FALSE);
-        all_res[i, 11]  <- sum(fea_res[[i]][,2] == TRUE);
-        # Feased and defeased
-        all_res[i, 12] <- sum(fea_res[[i]][,1] == FALSE & 
-                                  fea_res[[i]][,2] == TRUE);
-        all_res[i, 13] <- sum(fea_res[[i]][,1] == TRUE & 
-                                  fea_res[[i]][,2] == FALSE);
-    }
-    cnames <- c("N", "A0_unstable", "A0_stable", "A1_unstable", "A1_stable", 
[…]

### 2020-03-23 vignettes/R_old/test_mat.R

@@ -1,1555 +1,0 @@
-
-################################################################################
-################################################################################
-################################################################################
-
-################################################################################
-################################################################################
-################################################################################
-
-################################################################################
-################################################################################
-################################################################################
-
-################################################################################
-################################################################################
-################################################################################
-
-################################################################################
-################################################################################
-
-image(-1*egi$hmat, col = heat.colors(2000));
-
-make_e_image <- function(ei_ret){
-     hmat   <- ei_ret$hmat;
-     xcoord <- ei_ret$xcoord;
-     ycoord <- ei_ret$ycoord;
-     adj_ei <- -1 * log(hmat + 0.001);
-     xaxis  <- which(xcoord == 0) / length(xcoord);
-     yaxis  <- which(ycoord == 0) / length(ycoord);
-     coomat <- matrix(data = 0, nrow = dim(hmat)[1], ncol = dim(hmat)[2]);
-     coomat[,xaxis] <- 1;
-     coomat[yaxis,] <- 1;
-     image(adj_ei, col = heat.colors(2000), axes = FALSE);
-     abline(h = xaxis, lwd = 2);
-     abline(v = yaxis, lwd = 2);
-     box();
-}
-
-
-eigen_image  <- function(eigen_list){
-    reigs <- round(eigen_list, digits = 1);
-    xcoord <- seq(from = -40, to = 40, by = 0.1);
-    ycoord <- seq(from = -40, to = 40, by = 0.1);
[…]

### 2020-03-23 notebook/ms.Rmd
Correct the reference to figures in the introduction

@@
 Local system stability is assessed using eigenanalysis on $\mathbf{M}$, with the system being stable if the real parts of all eigenvalues ($\lambda$), and therefore the leading eigenvalue ($\lambda_{max}$), are negative ($\Re(\lambda_{max}) < 0$) [@May1972; @Allesina2012].
-In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -d$ ($-d$ is the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Figs 1a and 2a).
+In a large system (high $S$), eigenvalues are distributed uniformly [@Tao2010] within a circle centred at $\Re = -d$ ($-d$ is the mean value of diagonal elements) and $\Im = 0$, with a radius of $\sigma\sqrt{SC}$ [@May1972; @Allesina2012; @Allesina2015] (Fig.
+1a).
 Local stability of randomly assembled systems therefore becomes increasingly unlikely as $S$, $C$, and $\sigma$ increase.

### 2020-03-24 notebook/sci_rep_reviews.Rmd
Some more tweaks to the response to Reviewer 1's comments

@@
 I thank the reviewer for these comments.
-I share their belief that the issues in this paper are of current concern in ecology, and that the ideas presented in this manuscript are, to the best of my knowledge, novel.
+I share their belief that the issues in this paper are of current concern, and that the ideas presented in this manuscript are, to the best of my knowledge, novel.
 
@@
 I again thank the reviewer for their comments.
-I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (L48, L88, L209, L234).
+I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (e.g., L48, L88, L209, L234).
 In particular, in addition to the statement on L48, I now state more explicitly in describing the methods that "the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system." (L88)
@@
 
-In summary I think that the paper is interesting and timely, but I strongly recommend that the author
-considers also empirical webs (like for example those given in Dougoud et al. (2018)), the feasibility
-of equilibria in large ecosystems: A primary but neglected concept in the complexity-stability debate,
-PLoS Comput. Biol, and check numerically if the observation that variations of component responses
-rates enhance stability and feasibility is valid in experimentally observed webs. The author should
-also test this hypothesis on standard webs models for structured food web models like the niche, or
-the nested-hierarchy random web models.
+In summary I think that the paper is interesting and timely, but I strongly recommend that the author considers also empirical webs (like for example those given in Dougoud et al.
+(2018)), the feasibility of equilibria in large ecosystems: A primary but neglected concept in the complexity-stability debate, PLoS Comput.
+Biol, and check numerically if the observation that variations of component responses rates enhance stability and feasibility is valid in experimentally observed webs.
+The author should also test this hypothesis on standard webs models for structured food web models like the niche, or the nested-hierarchy random web models.
 
@@
 Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic ecological communities in which all population densities are positive).
-Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application.
+Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application (but, as I note in the Discussion L194, empirical food webs do not exist in a vacuum, and instead are parts of larger systems within which component response rate variation might indeed be relevant).
 
@@
 Methods for constructing small-world (L247), scale-free (L260), and cascade food web (L268) networks are now explained, with additional details and figures now in the Supplemental Information.
-The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity.
+The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity, thereby further showing the generality of my results across complex systems.
 
 Feasibility and how it relates to ecological systems is explained on L141, with discussion starting on L188 for how this relates to ecological systems.
-Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability).
+Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability) and accompanying website (https://bradduthie.github.io/RandomMatrixStability).
 

### 2020-03-24 notebook/sci_rep_reviews.Rmd
Make some edits to the response to Reviewer 2's comments

@@
 
-The manuscript is well written. I have just a remark concerning Figures 3 and 4. The black line is
-intended to show the proportion of systems that are stable when the variance is positive but would
-be unstable for zero variance. Is the author considering the conditional probability that A-id is
-unstable given that DA-id is stable ? I think that this point should be explained in more details.
+The manuscript is well written.
+I have just a remark concerning Figures 3 and 4.
+The black line is intended to show the proportion of systems that are stable when the variance is positive but would be unstable for zero variance.
+Is the author considering the conditional probability that A-id is unstable given that DA-id is stable ?
+I think that this point should be explained in more details.
 
@@
 
-I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the covariation of M (L65-L81).
+I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the (co)variation of M (L65-L81).
 I agree that these analytical results improve the manuscript, and I think that they are particularly useful in setting the stage for the numerical simulations that now follow.
@@
 
-As suggested, I also considered matrices with even more structure, performing a new simulations on each.
+As suggested, I also considered matrices with even more structure, performing new simulations on each.
 Specifically, I have simulated three broad classes of structured networks that are widespread in physical, social, and biological systems.

### 2020-03-24 notebook/sci_rep_reviews.Rmd
Make small edits to the response to Reviewer 3's comments

@@
 I tried this very early on in the process of investigating the effects of $\gamma$ by allowing diagonal elements to be sampled in the same way as off-diagonals: from a normal distribution centred at zero (data are here, note that the tables show the information in a bit more of a confusing format that in the SI, which was subsequently improved prior to my original submission: https://github.com/bradduthie/RandomMatrixStability/blob/dfe6f8e0225733c94348bc9e0dad6218c59cc109/notebook/sim_results_old/oldest/mat_dat.csv).
-The effect of $\gamma$ on increasing stability is retained, but stability of all systems drops to zero out of one million given size $S > 14$.
+The effect of $\gamma$ on increasing stability is retained, but the stability of all systems drops to zero out of one million given size $S > 14$.
 And in the random systems that are found to be stable, all diagonal elements are nearly always negative (or, rarely, with one or two very low positive values).
@@
 
-Unequal diagonal elements could be used, but doing so would only be expected to add some variation around the expected centre of eigenvalue distributions, and would not change my primary result concerning the effect of $\gamma$ (note that this would lead to the same results as mentioned above where diagonal elements were sampled in the same way as off diagonals, but with more total stable systems since the mean diagonal value would be less lower).
+Unequal diagonal elements could be used, but doing so would only be expected to add some variation around the expected centre of eigenvalue distributions, and would not change my primary result concerning the effect of $\gamma$ (note that this would lead to the same results as mentioned above where diagonal elements were sampled in the same way as off diagonals, but would result in more variation in the mean diagonal value, d).
 

### 2020-03-26 notebook/ms.Rmd
Make minor edits to the notation to make everything consistent

@@
 In the absence of such variation ($\sigma^{2}_{\gamma} = 0$), $\boldsymbol{\gamma}$ is set to the identity matrix (diagonal elements all equal 1) and $\mathbf{M} = \mathbf{A}$.
-Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-d, 0)$ with a radius of $\sigma \sqrt{SC}$ [@May1972] (Fig.
+Under these conditions, eigenvalues of $\mathbf{M}$ are distributed uniformly [@Tao2010] in a circle centred at $(-1, 0)$ with a radius of $\sigma \sqrt{SC}$ [@May1972] (Fig.
 1a).
@@
 
-Given $\sigma^{2}_\gamma = 0$, $\Re(\lambda_{max})$ decreases linearly with $\rho$ such that [@Tang2014c],
+Given $\sigma^{2}_\gamma = 0$, $\Re(\lambda_{max})$ increases linearly with $\rho$ such that [@Tang2014c],
 
@@
 
-**Simulation of random $\mathbf{M}$ across $\mathbf{\rho}$**
-
-Numerical simulations of $\mathbf{M}$ revealed that $\sigma^{2}_{\gamma}$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
+**Simulation of random $\mathbf{M}$ across $\mathbf{\rho}$**.
+Numerical simulations revealed that $\sigma^{2}_{\gamma}$ results in a nonlinear relationship between $\rho$ and $\Re(\lambda_{max})$, which can sometimes increase the stability of the system.
 Figure 2 shows a comparison of $\Re(\lambda_{max})$ across $\rho$ values for $\mathbf{A}$ ($\sigma^{2}_{\gamma} = 0$) versus $\mathbf{M}$ ($\sigma^{2}_{\gamma} = 1/3$) given $S = 25$, $C = 1$, and $\sigma_{A} = 0.4$.
@@
 4; see Supplementary Information for full table of results), and that the difference between the probabilities of observing a stable system increased with an increase in $S$.
-In other words, the potential for $\sigma^{2}_{\gamma}$ to affect stability increased with system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
+In other words, the potential for $\sigma^{2}_{\gamma}$ to affect stability increased with increasing system complexity and was most relevant for systems on the cusp of being too complex to be realistically stable.
 For the highest values of $S$, nearly all systems that were stable given varying $\gamma$ would not have been stable given $\gamma = 1$.
 
-I also simulated 100000 $\mathbf{M = \gamma A}$ for three types of random networks that are typically interpreted as modelling three types of interspecific ecological interactions [@Allesina2011; @Allesina2012].
+I also simulated 100000 $\mathbf{M}$ for three types of random networks that are typically interpreted as modelling three types of interspecific ecological interactions [@Allesina2011; @Allesina2012].
 These interaction types are competitive, mutualist, and predator-prey, as modelled by off-diagonal elements that are constrained to be negative, positive, or paired such that if $A_{ij} > 0$ then $A_{ji} < 0$, respectively [@Allesina2012] (but are otherwise identical to the purely random $\mathbf{A}$).
@@
 For comparable magnitudes of complexity, structured networks still had a higher number of stable $\mathbf{M}$ than $\mathbf{A}$.
-For random networks, $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma = 0.4$ and $C = 1$), and therefore complexity $\sigma_{A} \sqrt{SC} \gtrapprox 1.26$.
+For random networks, $\sigma^{2}_{\gamma}$ increased stability given $S > 10$ ($\sigma_{A} = 0.4$ and $C = 1$), and therefore complexity $\sigma_{A} \sqrt{SC} \gtrapprox 1.26$.
 This threshold of complexity, above which more $\mathbf{M}$ than $\mathbf{A}$ were stable, was comparable for small-world networks, and slightly lower for scale-free networks (note that algorithms for generating small-world and scale-free networks necessarily led to varying $C$; see methods).
@@
 Feasibilility assumes that values of all components are positive at equilibrium [@Grilli2017; @Dougoud2018; @Song2018].
-This is of particular interest for ecological communities because population density ($N$) cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
+This is of particular interest for ecological communities because population density ($n$) cannot take negative values, meaning that ecological systems need to be feasible for stability to be biologically realistic [@Dougoud2018].
 While my results are intended to be general to all complex systems, and not restricted to species networks, I have also performed a feasibility analysis on all matrices tested for stability.
@@
 I used the R programming language for all numerical simulations and analyses [@Rproject].
-Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, 0.4)$ with a probability $C$ (unsampled elements were set to zero).
+Purely random networks were generated by sampling off-diagonal elements from an i.i.d $A_{ij} \sim \mathcal{N}(0, 0.4^{2})$ with a probability $C$ (unsampled elements were set to zero).
 Diagonal elements $A_{ii}$ were set to $-1$.
@@
 Random $\mathbf{A}$ matrices with correlated elements $A_{ij}$ and $A_{ji}$ were built using Cholesky decomposition.
-Competitor networks in which off-diagonal elements $A_{ij} \geq 0$ were constructed by first building a random $\mathbf{A}$, then flipping the sign of any elements in which $A_{ij} < 0$.
-Similarly, mutualist networks were constructed by building a random $\mathbf{A}$, then flipping the sign of elements where $A_{ij} > 0$.
-Predator-prey networks were constructed by first buing a random $\mathbf{A}$, then flipping the sign of either $A_{ij}$ or $A_{ji}$ if $A_{ij} \times A_{ji} > 0$.
+Competitor networks in which off-diagonal elements $A_{ij} \leq 0$ were constructed by first building a random $\mathbf{A}$, then flipping the sign of any elements in which $A_{ij} > 0$.
+Similarly, mutualist networks were constructed by building a random $\mathbf{A}$, then flipping the sign of elements where $A_{ij} < 0$.
+Predator-prey networks were constructed by first building a random $\mathbf{A}$, then flipping the sign of either $A_{ij}$ or $A_{ji}$ if $A_{ij} \times A_{ji} > 0$.
 
@@
 A new random matrix $\mathbf{J}$ was then generated with elements $J_{ij}$ sampled i.i.d.
-from $\mathcal{N}(0, 0.4)$.
+from $\mathcal{N}(0, 0.4^{2})$.
 To build the interaction matrix $\mathbf{A}$, I used element-wise multiplication $\mathbf{A} = \mathbf{J} \odot \mathbf{B}$, then set $diag(\mathbf{A}) = -1$.
@@
 As with small-world networks, a random matrix $\mathbf{J}$ was generated, and $\mathbf{A} = \mathbf{J} \odot \mathbf{G}$.
-Diagonal were set to $-1$.
+Diagonal elements were set to $-1$.
 I simulated scale-free networks across all combinations of $S = \{24, 48, 72, 96, 120\}$ and $m = \{2, 3, ..., 11, 12\}$.
@@
 Cascade food webs were constructed following Solow and Beet [@Solow1998].
-First, a random matrix $\mathbf{A}$ was generated with off-diagonal elements sampled i.i.d so that $A_{ij} \sim \mathcal{N}(0, 0.4)$.
+First, a random matrix $\mathbf{A}$ was generated with off-diagonal elements sampled i.i.d so that $A_{ij} \sim \mathcal{N}(0, 0.4^2)$.
 Each component in the system was ranked from $1$ to $S$.
@@
 Diagonal elements of $\mathbf{A}$ were set to $-1$ and $C = 1$.
-I simulated cascade food webs across all combinations of $S = \{2, 3, ..., 59, 60\}$.
+I simulated cascade food webs for $S = \{2, 3, ..., 59, 60\}$.
 
@@
 Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\boldsymbol{\gamma}$, and the 20 $diag(\boldsymbol{\gamma})$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $diag(\boldsymbol{\gamma})$ vectors.
-Mutation of each $\gamma_{i}$ in a $diag(\boldsymbol{\gamma})$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02)$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
+Mutation of each $\gamma_{i}$ in a $diag(\boldsymbol{\gamma})$ vector occurred with a probability of 0.2, resulting in a mutation effect of size $\mathcal{N}(0, 0.02^{2})$ being added to generate the newly mutated $\gamma_{i}$ (any $\gamma_{i}$ values that mutated below zero were multiplied by $-1$, and any values that mutated above 2 were set to 2).
 Crossover occurred between two sets of 100 $diag(\boldsymbol{\gamma})$ vectors paired in each generation; vectors were randomly sampled with replacement among but not within sets.
@@
 Eigenvalues of $\mathbf{M}$ are then distributed non-uniformly within the red circle centred at the origin with a radius of $\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx$ `r round(sqrt(joint_var) * sqrt(S), digits = 2);`.
-($\textbf{c}$) A different random system $\mathbf{A}$ constructed from the same parameters as in $\textbf{a}$, except with diagonal values of $-1$.
+($\textbf{c}$) A different random system $\mathbf{A}$ constructed from the same parameters as in $\textbf{a}$, except with diagonal element values of $-1$.
 ($\textbf{d}$) The same system $\textbf{c}$ after including variation in component response rates, sampled from $\mathcal{U}(0, 2)$ as in $\textbf{b}$.
@@
 In all simulations, system size and connectence were set to $S = 25$ and $C = 1$, respectively.
-Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.4)$, and diagonal elements were set to $-1$.
+Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.4^{2})$, and diagonal elements were set to $-1$.
 Elements of $\gamma$ were sampled, $\gamma_{i} \sim \mathcal{U}(0, 2)$.
@@
 For all simulations, system connectance was $C = 1$.
-Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.2)$, and diagonal elements were set to $-1$.
+Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.2^{2})$, and diagonal elements were set to $-1$.
 Elements of $\gamma$ were sampled such that $\gamma_{i} \sim \mathcal{U}(0, 2)$, so $\sigma^{2}_{\gamma} = 1/3$.
@@
 
-**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\ln$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems for which variation in $\gamma$ is critical for system stability.
+**Figure 4: Stability of large complex systems with and without variation in component response rate ($\boldsymbol{\gamma}$).** The $\log$ number of systems that are stable across different system sizes ($S = \{2, 3, ..., 49, 50 \}$) given $C = 1$, and the proportion of systems for which variation in $\gamma$ is critical for system stability.
 For each $S$, 1 million complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
-Stability given $\sigma^{2}_{\gamma}>0$ is then compared to stability in an otherwise identical system in which $\gamma = E[\mathcal{U}(0, 2)]$ for all components.
+Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Stability given $\sigma^{2}_{\gamma}>0$ is then compared to stability in an otherwise identical system in which $\gamma_{i} = E[\mathcal{U}(0, 2)]$ for all components.
 Blue and red bars show the number of stable systems in the absence and presence of $\sigma^{2}_{\gamma}$, respectively.

### 2020-03-26 notebook/ms.tex
Make minor edits to the notation to make everything consistent

@@
 \(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{5}}
-(Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and \(\sigma\)
-increase.
+(Fig. 1a). Local stability of randomly assembled systems therefore
+becomes increasingly unlikely as \(S\), \(C\), and \(\sigma\) increase.
 
@@
 uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{15}} in a
-circle centred at \((-d, 0)\) with a radius of
+circle centred at \((-1, 0)\) with a radius of
 \(\sigma \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
@@
 
-Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) decreases
+Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) increases
 linearly with \(\rho\) such
@@
 
-\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{\rho}\)}
-
-Numerical simulations of \(\mathbf{M}\) revealed that
-\(\sigma^{2}_{\gamma}\) results in a nonlinear relationship between
-\(\rho\) and \(\Re(\lambda_{max})\), which can sometimes increase the
-stability of the system. Figure 2 shows a comparison of
-\(\Re(\lambda_{max})\) across \(\rho\) values for \(\mathbf{A}\)
-(\(\sigma^{2}_{\gamma} = 0\)) versus \(\mathbf{M}\)
+\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{\rho}\)}.
+Numerical simulations revealed that \(\sigma^{2}_{\gamma}\) results in a
+nonlinear relationship between \(\rho\) and \(\Re(\lambda_{max})\),
+which can sometimes increase the stability of the system. Figure 2 shows
+a comparison of \(\Re(\lambda_{max})\) across \(\rho\) values for
+\(\mathbf{A}\) (\(\sigma^{2}_{\gamma} = 0\)) versus \(\mathbf{M}\)
 (\(\sigma^{2}_{\gamma} = 1/3\)) given \(S = 25\), \(C = 1\), and
@@
 potential for \(\sigma^{2}_{\gamma}\) to affect stability increased with
-system complexity and was most relevant for systems on the cusp of being
-too complex to be realistically stable. For the highest values of \(S\),
-nearly all systems that were stable given varying \(\gamma\) would not
-have been stable given \(\gamma = 1\).
-
-I also simulated 100000 \(\mathbf{M = \gamma A}\) for three types of
-random networks that are typically interpreted as modelling three types
-of interspecific ecological
+increasing system complexity and was most relevant for systems on the
+cusp of being too complex to be realistically stable. For the highest
+values of \(S\), nearly all systems that were stable given varying
+\(\gamma\) would not have been stable given \(\gamma = 1\).
+
+I also simulated 100000 \(\mathbf{M}\) for three types of random
+networks that are typically interpreted as modelling three types of
+interspecific ecological
 interactions\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{27}}.
@@
 \(\sigma^{2}_{\gamma}\) increased stability given \(S > 10\)
-(\(\sigma = 0.4\) and \(C = 1\)), and therefore complexity
+(\(\sigma_{A} = 0.4\) and \(C = 1\)), and therefore complexity
 \(\sigma_{A} \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity,
@@
 This is of particular interest for ecological communities because
-population density (\(N\)) cannot take negative values, meaning that
+population density (\(n\)) cannot take negative values, meaning that
 ecological systems need to be feasible for stability to be biologically
@@
 random networks were generated by sampling off-diagonal elements from an
-i.i.d \(A_{ij} \sim \mathcal{N}(0, 0.4)\) with a probability \(C\)
+i.i.d \(A_{ij} \sim \mathcal{N}(0, 0.4^{2})\) with a probability \(C\)
 (unsampled elements were set to zero). Diagonal elements \(A_{ii}\) were
@@
 Cholesky decomposition. Competitor networks in which off-diagonal
-elements \(A_{ij} \geq 0\) were constructed by first building a random
+elements \(A_{ij} \leq 0\) were constructed by first building a random
 \(\mathbf{A}\), then flipping the sign of any elements in which
-\(A_{ij} < 0\). Similarly, mutualist networks were constructed by
+\(A_{ij} > 0\). Similarly, mutualist networks were constructed by
 building a random \(\mathbf{A}\), then flipping the sign of elements
-where \(A_{ij} > 0\). Predator-prey networks were constructed by first
-buing a random \(\mathbf{A}\), then flipping the sign of either
+where \(A_{ij} < 0\). Predator-prey networks were constructed by first
+building a random \(\mathbf{A}\), then flipping the sign of either
 \(A_{ij}\) or \(A_{ji}\) if \(A_{ij} \times A_{ji} > 0\).
@@
 random matrix \(\mathbf{J}\) was then generated with elements \(J_{ij}\)
-sampled i.i.d. from \(\mathcal{N}(0, 0.4)\). To build the interaction
-matrix \(\mathbf{A}\), I used element-wise multiplication
+sampled i.i.d. from \(\mathcal{N}(0, 0.4^{2})\). To build the
+interaction matrix \(\mathbf{A}\), I used element-wise multiplication
 \(\mathbf{A} = \mathbf{J} \odot \mathbf{B}\), then set
@@
 \(\mathbf{J}\) was generated, and
-\(\mathbf{A} = \mathbf{J} \odot \mathbf{G}\). Diagonal were set to
-\(-1\). I simulated scale-free networks across all combinations of
+\(\mathbf{A} = \mathbf{J} \odot \mathbf{G}\). Diagonal elements were set
+to \(-1\). I simulated scale-free networks across all combinations of
 \(S = \{24, 48, 72, 96, 120\}\) and \(m = \{2, 3, ..., 11, 12\}\).
@@
 random matrix \(\mathbf{A}\) was generated with off-diagonal elements
-sampled i.i.d so that \(A_{ij} \sim \mathcal{N}(0, 0.4)\). Each
+sampled i.i.d so that \(A_{ij} \sim \mathcal{N}(0, 0.4^2)\). Each
 component in the system was ranked from \(1\) to \(S\). If component
@@
 elements of \(\mathbf{A}\) were set to \(-1\) and \(C = 1\). I simulated
-cascade food webs across all combinations of
-\(S = \{2, 3, ..., 59, 60\}\).
+cascade food webs for \(S = \{2, 3, ..., 59, 60\}\).
 
@@
 probability of 0.2, resulting in a mutation effect of size
-\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
+\(\mathcal{N}(0, 0.02^{2})\) being added to generate the newly mutated
 \(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.14.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.19.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
 from the same parameters as in \(\textbf{a}\), except with diagonal
-values of \(-1\). (\(\textbf{d}\)) The same system \(\textbf{c}\) after
-including variation in component response rates, sampled from
-\(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).
+element values of \(-1\). (\(\textbf{d}\)) The same system
+\(\textbf{c}\) after including variation in component response rates,
+sampled from \(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).
 
@@
 \(C = 1\), respectively. Off-diagonal elements of \(\textbf{A}\) were
-randomly sampled from \(A_{ij} \sim \mathcal{N}(0, 0.4)\), and diagonal
-elements were set to \(-1\). Elements of \(\gamma\) were sampled,
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\).
+randomly sampled from \(A_{ij} \sim \mathcal{N}(0, 0.4^{2})\), and
+diagonal elements were set to \(-1\). Elements of \(\gamma\) were
+sampled, \(\gamma_{i} \sim \mathcal{U}(0, 2)\).
 
@@
 \(C = 1\). Off-diagonal elements of \(\textbf{A}\) were randomly sampled
-from \(A_{ij} \sim \mathcal{N}(0, 0.2)\), and diagonal elements were set
-to \(-1\). Elements of \(\gamma\) were sampled such that
+from \(A_{ij} \sim \mathcal{N}(0, 0.2^{2})\), and diagonal elements were
+set to \(-1\). Elements of \(\gamma\) were sampled such that
 \(\gamma_{i} \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
@@
 variation in component response rate (\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
+\(\log\) number of systems that are stable across different system sizes
 (\(S = \{2, 3, ..., 49, 50 \}\)) given \(C = 1\), and the proportion of
@@
 generated. Stability of each complex system is tested given variation in
-\(\gamma\) by randomly sampling \(\gamma \sim \mathcal{U}(0, 2)\).
+\(\gamma\) by randomly sampling \(\gamma_{i} \sim \mathcal{U}(0, 2)\).
 Stability given \(\sigma^{2}_{\gamma}>0\) is then compared to stability
 in an otherwise identical system in which
-\(\gamma = E[\mathcal{U}(0, 2)]\) for all components. Blue and red bars
-show the number of stable systems in the absence and presence of
+\(\gamma_{i} = E[\mathcal{U}(0, 2)]\) for all components. Blue and red
+bars show the number of stable systems in the absence and presence of
 \(\sigma^{2}_{\gamma}\), respectively. The black line shows the

### 2020-03-26 notebook/ms.Rmd
Fix notation issue again

@@
 I used numerical simulations and eigenanalysis to test how variation in $\gamma$ affects stability in random matrices with known properties, comparing the stability of $\textbf{A}$ versus $\mathbf{M} = \gamma\mathbf{A}$.
-Values of $\gamma$ were sampled from a uniform distribution where $\gamma_{i} \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other $\gamma$ distributions, which gave similar results).
+Values of $\gamma$ were sampled from a uniform distribution where $\gamma \sim \mathcal{U}(0, 2)$ and $\sigma^{2}_{\gamma} = 1/3$ (see Supplementary Information for other $\gamma$ distributions, which gave similar results).
 In all simulations, diagonal elements were standardised to ensure that $-d$ between individual $\textbf{A}$ and $\textbf{M}$ pairs were identical (also note that $E[\gamma_{i}] = 1$).
@@
 To investigate the effect of $\sigma^{2}_{\gamma}$ on stability across systems of increasing complexity, I simulated random $\mathbf{M = \gamma A}$ matrices at $\sigma_{A} = 0.4$ and $C = 1$ across $S = \{2, 3, ..., 49, 50\}$.
-One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{A}$ vesus $\mathbf{M}$ was assessed given $\gamma_{i} \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
+One million $\mathbf{M}$ were simulated for each $S$, and the stability of $\mathbf{A}$ vesus $\mathbf{M}$ was assessed given $\gamma \sim \mathcal{U}(0, 2)$ ($\sigma^{2}_{\gamma} = 1/3$).
 For all $S > 10$, I found that the number of stable random systems was higher in $\mathbf{M}$ than $\mathbf{A}$ (Fig.
@@
 Elements of $\boldsymbol{\gamma}$ were simulated i.i.d.
-from a distribution with positive support (typically $\gamma_{i} \sim \mathcal{U}(0, 2)$).
+from a distribution with positive support (typically $\gamma \sim \mathcal{U}(0, 2)$).
 Random $\mathbf{A}$ matrices with correlated elements $A_{ij}$ and $A_{ji}$ were built using Cholesky decomposition.
@@
 To check the feasibility criteria for $\mathbf{M = \gamma A}$, I therefore evaluated $\mathbf{-M^{-1}r}$ ($\mathbf{r}$ elements were sampled i.i.d.
-from $r_{i} \sim \mathcal{N}(0, 0.4^{2})$).
+from $r \sim \mathcal{N}(0, 0.4^{2})$).
 Feasibility is satisfied if all of the elements of the resulting vector are positive.
@@
 For each $S = \{2, 3, ..., 39, 40\}$, the genetic algorithm was run for 100000 random $\mathbf{M = \gamma A}$ ($\sigma_{A} = 0.4$, $C = 1$).
-The genetic algorithm was initialised with a population of 1000 different $diag(\boldsymbol{\gamma})$ vectors with elements sampled i.i.d from $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+The genetic algorithm was initialised with a population of 1000 different $diag(\boldsymbol{\gamma})$ vectors with elements sampled i.i.d from $\gamma \sim \mathcal{U}(0, 2)$.
 Eigenanalysis was performed on the $\mathbf{M}$ resulting from each $\boldsymbol{\gamma}$, and the 20 $diag(\boldsymbol{\gamma})$ vectors resulting in $\mathbf{M}$ with the lowest $\Re(\lambda_{max})$ each produced 50 clonal offspring with subsequent random mutation and crossover between the resulting new generation of 1000 $diag(\boldsymbol{\gamma})$ vectors.
@@
 Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.4^{2})$, and diagonal elements were set to $-1$.
-Elements of $\gamma$ were sampled, $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Elements of $\gamma$ were sampled, $\gamma \sim \mathcal{U}(0, 2)$.
 
@@
 Off-diagonal elements of $\textbf{A}$ were randomly sampled from $A_{ij} \sim \mathcal{N}(0, 0.2^{2})$, and diagonal elements were set to $-1$.
-Elements of $\gamma$ were sampled such that $\gamma_{i} \sim \mathcal{U}(0, 2)$, so $\sigma^{2}_{\gamma} = 1/3$.
+Elements of $\gamma$ were sampled such that $\gamma \sim \mathcal{U}(0, 2)$, so $\sigma^{2}_{\gamma} = 1/3$.
 
@@
 For each $S$, 1 million complex systems are randomly generated.
-Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma_{i} \sim \mathcal{U}(0, 2)$.
+Stability of each complex system is tested given variation in $\gamma$ by randomly sampling $\gamma \sim \mathcal{U}(0, 2)$.
 Stability given $\sigma^{2}_{\gamma}>0$ is then compared to stability in an otherwise identical system in which $\gamma_{i} = E[\mathcal{U}(0, 2)]$ for all components.

### 2020-03-26 notebook/SI.Rmd
Change the title, as requested by Reviewer 2
Make minor edits to the SI to correct small issues

@@
 ---
-title: "Component response rate variation underlies the stability of complex systems"
+title: "Component response rate variation underlies the stability of highly complex finite systems"
 author: "A. Bradley Duthie"
@@
 
-Overall, the ratio of stable $\mathbf{A}$ matrices to stable $\mathbf{M}$ matrices found is greater than 1 whenever $S > 10$ (compare column 5 to column 3), and this ratio increases with increasing $S$ (column 1).
+Overall, the ratio of stable $\mathbf{A}$ matrices to stable $\mathbf{M}$ matrices found is greater than 1 whenever $S > 10$ (compare column 3 to column 5), and this ratio increases with increasing $S$ (column 1).
 Hence, more randomly created complex systems ($\mathbf{M}$) are stable given variation in $\gamma$ than when $\gamma = 1$.
@@
 
-The number of stable $\mathbf{M = \gamma A}$ systems was estimated [exactly as it was](#IncrS) in the main text for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $\mathbf{M}$ were generated instead of 1 million.
+The number of stable $\mathbf{M = \gamma A}$ systems was calculated [exactly as it was](#IncrS) for random matrices for values of $S$ from 2 to 50 (100 in the case of the relatively more stable predator-prey interactions), except that only 100000 random $\mathbf{M}$ were generated instead of 1 million.
 
@@
 The column `A_unstable` shows the number of $\mathbf{A}$ matrices that are unstable, and the column `A_stable` shows the number of $\mathbf{A}$ matrices that are stable (these two columns sum to 100000).
-Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `A1_stable` shows the number that are stable.
+Similarly, the column `M_unstable` shows the number of $\mathbf{M}$ matrices that are unstable and `M_stable` shows the number that are stable.
 The columns `A_stabilised` and `A_destabilised` show how many $\mathbf{A}$ matrices were stabilised or destabilised, respectively, by $\sigma^{2}_{\gamma}$.
@@
 5, 0.7, 0.9 \}$.
-In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma_{i} = 1$).
+In all cases, uniform variation in component response rate ($\gamma \sim \mathcal{U}(0, 2)$) led to a higher number of stable communities than when $\gamma$ did not vary ($\gamma = 1$).
 In contrast to the main text, 100000 rather than 1 million $\mathbf{M}$ were simulated.
@@
 
-Note that output from `Evo_rand_gen_var` only includes the first seven rows of the table above, and `rand_gen_var` does not include $C$ (which can be defind as an argument).
+Note that output from `Evo_rand_gen_var` only includes the first seven rows of the table above, and `rand_gen_var` does not include $C$ (which can be defined as an argument).
 All results presented here and in the main text are available in the [inst/extdata](https://github.com/bradduthie/RandomMatrixStability/tree/master/inst/extdata) folder of the [RandomMatrixStability](https://github.com/bradduthie/RandomMatrixStability) R package.

### 2020-03-26 notebook/SI.tex
Make minor edits to the SI to correct small issues

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
+            pdftitle={Component response rate variation underlies the stability of highly complex finite systems},
             pdfauthor={A. Bradley Duthie},
@@
 \setlength{\droptitle}{-2em}
-
-  \title{Component response rate variation underlies the stability of complex
-systems}
+  \title{Component response rate variation underlies the stability of highly
+complex finite systems}
   \pretitle{\vspace{\droptitle}\centering\huge}
@@
 \(\mathbf{M}\) matrices found is greater than 1 whenever \(S > 10\)
-(compare column 5 to column 3), and this ratio increases with increasing
+(compare column 3 to column 5), and this ratio increases with increasing
 \(S\) (column 1). Hence, more randomly created complex systems
@@
 
-The number of stable \(\mathbf{M = \gamma A}\) systems was estimated
-\protect\hyperlink{IncrS}{exactly as it was} in the main text for random
-matrices for values of \(S\) from 2 to 50 (100 in the case of the
-relatively more stable predator-prey interactions), except that only
-100000 random \(\mathbf{M}\) were generated instead of 1 million.
+The number of stable \(\mathbf{M = \gamma A}\) systems was calculated
+\protect\hyperlink{IncrS}{exactly as it was} for random matrices for
+values of \(S\) from 2 to 50 (100 in the case of the relatively more
+stable predator-prey interactions), except that only 100000 random
+\(\mathbf{M}\) were generated instead of 1 million.
 
@@
 to 100000). Similarly, the column \texttt{M\_unstable} shows the number
-of \(\mathbf{M}\) matrices that are unstable and \texttt{A1\_stable}
+of \(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable}
 shows the number that are stable. The columns \texttt{A\_stabilised} and
@@
 (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
-communities than when \(\gamma\) did not vary (\(\gamma_{i} = 1\)). In
+communities than when \(\gamma\) did not vary (\(\gamma = 1\)). In
 contrast to the main text, 100000 rather than 1 million \(\mathbf{M}\)
@@
 first seven rows of the table above, and \texttt{rand\_gen\_var} does
-not include \(C\) (which can be defind as an argument). All results
+not include \(C\) (which can be defined as an argument). All results
 presented here and in the main text are available in the

### 2020-03-26 notebook/arxiv/ms.tex
Organise TEX files and prepare for sending to arXiv

@@
 \hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
-            pdfauthor={A. Bradley Duthie},
-            colorlinks=true,
-            linkcolor=blue,
-            citecolor=Blue,
-            urlcolor=Blue,
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{graphicx,grffile}
-\makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\setlength{\emergencystretch}{3em}  % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\usepackage{lmodern}
-\usepackage{setspace}
-\setstretch{1}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
-  \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
-\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
+            pdftitle={Component response rate variation underlies the stability of highly complex finite systems},
             pdfauthor={A. Bradley Duthie},
@@
 
+% Create subtitle command for use in maketitle
+\newcommand{\subtitle}[1]{
+  \posttitle{
+    \begin{center}\large#1\end{center}
+    }
+}
+
 \setlength{\droptitle}{-2em}
-  \title{Component response rate variation underlies the stability of complex systems}
+  \title{Component response rate variation underlies the stability of highly
+complex finite systems}
   \pretitle{\vspace{\droptitle}\centering\huge}
   \posttitle{\par}
-  \author{A. Bradley Duthie}
+  \author{A. Bradley Duthie (
+\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}
+)}
   \preauthor{\centering\large\emph}
@@
 \usepackage{natbib}
-\usepackage{lineno}
 \usepackage[utf8]{inputenc}
+\linenumbers
 \bibliographystyle{amnatnat}
@@
 \maketitle
+
+\textbf{Key words:} Ecological networks, gene-regulatory networks,
+neural networks, financial networks, system stability, random matrix
+theory
+
+\vspace{10mm}
+\hrule
+\vspace{2mm}
 
@@
 Despite recent interest in the relationship between system properties
-and stability, the effect of variation in the response rate of
-individual system components remains unconsidered. Here I vary the
-component response rates (\(\boldsymbol{\gamma}\)) of randomly generated
-complex systems. I show that when component response rates vary, the
-potential for system stability is markedly increased. Variation in
-\(\boldsymbol{\gamma}\) is especially important for stability in
-highly complex systems, in which the probability of stability would
-otherwise be negligible. At such extremes of simulated system
-complexity, the largest stable complex systems would be unstable if not
-for \(\boldsymbol{Var(\gamma)}\). My results therefore reveal a
-previously unconsidered aspect of system stability that is likely to be
-pervasive across all realistic complex systems.}
+and stability, the effect of variation in response rate across system
+components remains unconsidered. Here I vary the component response
+rates (\(\boldsymbol{\gamma}\)) of randomly generated complex systems. I
+use numerical simulations to show that when component response rates
+vary, the potential for system stability increases. These results are
+robust to common network structures, including small-world and
+scale-free networks, and cascade food webs. Variation in
+\(\boldsymbol{\gamma}\) is especially important for stability in highly
+complex systems, in which the probability of stability would otherwise
+be negligible. At such extremes of simulated system complexity, the
+largest stable complex systems would be unstable if not for variation in
+\(\boldsymbol{\gamma}\). My results therefore reveal a previously
+unconsidered aspect of system stability that is likely to be pervasive
+across all realistic complex systems.}
+
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\subsection{Introduction}\label{introduction}
 
@@
 \(\sigma^{2}\); diagonal elements are set to
--1\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}.
-Local system stability is assessed using eigenanalysis, with the system
-being stable if the real parts of all eigenvalues (\(\lambda\)) of
-\(\mathbf{M}\) are negative
-(\(\max\left(\Re(\lambda)\right) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
+\(-1\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{5}}.
+Local system stability is assessed using eigenanalysis on
+\(\mathbf{M}\), with the system being stable if the real parts of all
+eigenvalues (\(\lambda\)), and therefore the leading eigenvalue
+(\(\lambda_{max}\)), are negative
+(\(\Re(\lambda_{max}) < 0\))\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}.
 In a large system (high \(S\)), eigenvalues are distributed
-uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}} within a
-circle centred at \(\Re = -1\) (the mean value of diagonal elements) and
-\(\Im = 0\), with a radius of
-\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{4}}
-(Figs 1a and 2a). Local stability of randomly assembled systems
-therefore becomes increasingly unlikely as \(S\), \(C\), and
-\(\sigma^{2}\) increase.
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{15}} within a
+circle centred at \(\Re = -d\) (\(-d\) is the mean value of diagonal
+elements) and \(\Im = 0\), with a radius of
+\(\sigma\sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2015}{5}}
+(Fig. 1a). Local stability of randomly assembled systems therefore
+becomes increasingly unlikely as \(S\), \(C\), and \(\sigma\) increase.
+
+\clearpage
+\hrule
+\vspace{2mm}
+\textbf{Figure 1: Eigenvalue distributions of random complex systems.}
+Each panel shows the real (x-axis) and imaginary (y-axis) parts of
+\(S =\) 400 eigenvalues from random \(S \times S\) matrices.
+(\(\textbf{a}\)) A system represented by a matrix \(\mathbf{A}\), in
+which all elements are sampled from a normal distribution with
+\(\mu = 0\) and \(\sigma_{A} = 1/\sqrt{S}\). Points are uniformly
+distributed within the blue circle centred at the origin with a radius
+of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
+\(\textbf{a}\) after including variation in the response rates of \(S\)
+components, represented by the diagonal matrix \(\gamma\), such that
+\(\mathbf{M} = \gamma\mathbf{A}\). Elements of \(\gamma\) are randomly
+sampled from a uniform distribution from \(\min = 0\) to \(\max = 2\).
+Eigenvalues of \(\mathbf{M}\) are then distributed non-uniformly within
+the red circle centred at the origin with a radius of
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
+(\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
+from the same parameters as in \(\textbf{a}\), except with diagonal
+element values of \(-1\). (\(\textbf{d}\)) The same system
+\(\textbf{c}\) after including variation in component response rates,
+sampled from \(\mathcal{U}(0, 2)\) as in \(\textbf{b}\).
+
+\includegraphics{fig1.pdf}
+\vspace{2mm}
+\hrule
 
 May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-stability criterion \(\sigma\sqrt{SC} < 1\) assumes that the expected
+stability criterion \(\sigma\sqrt{SC} < d\) assumes that the expected
 response rates (\(\gamma\)) of individual components to perturbations of
@@
 has not been investigated in complex systems theory. Intuitively,
-variation in \(\gamma\) might be expected to decrease system stability
-by introducing a new source of variation into the system and thereby
-increasing \(\sigma\). Here I show why, despite higher \(\sigma\),
-realistic complex systems (such that \(S\) is high but finite) are actually 
-more likely to be stable if their individual component response rates vary.
+variation in \(\gamma\) (\(\sigma^{2}_{\gamma}\)) might be expected to
+decrease system stability by introducing a new source of variation into
+the system and thereby increasing \(\sigma\). Here I show that, despite
+higher \(\sigma\), realistic complex systems (in which \(S\) is high but
+finite) are actually more likely to be stable if their individual
+component response rates vary. My results are robust across commonly
+observed network structures, including
+random\textsuperscript{\protect\hyperlink{ref-May1972}{1}},
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}},
+cascade food
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{18},\protect\hyperlink{ref-Williams2000}{19}}
+networks.
 
@@
 
-\textbf{Component response rates of random complex systems}. Rows in
-\(\mathbf{M}\) define how a given component \(i\) is affected by other
-components of the system, meaning that the rate of component response
-time can be modelled by multiplying all row elements by a real scalar
-value
-\(\gamma_{i}\)\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
-The distribution of \(\gamma\) over \(S\) components thereby models the
-distribution of component response rates. An instructive example
-compares one \(\mathbf{M}\) where \(\gamma_{i} = 1\) for all \(i\) in
-\(S\) to the same \(\mathbf{M}\) when half of \(\gamma_{i} = 1.95\) and
-half of \(\gamma_{i} = 0.05\). This models one system in which
-\(\gamma\) is invariant and one in which \(\gamma\) varies, but systems
-are otherwise identical (note that mean \(\gamma_{i} = 1\) in both
-cases). I assume \(S = 200\), \(C = 0.05\), and \(\sigma = 0.4\);
-diagonal elements are set to \(-1\) and non-zero off-diagonal elements
-are drawn randomly from \(\mathcal{N}(0, \sigma^{2})\). Rows are then
-multiplied by \(\gamma_{i}\) to generate \(\mathbf{M}\). When
-\(\gamma_{i} = 1\), eigenvalues of \(\mathbf{M}\) are distributed
-uniformly within a circle centred at (\(-1, 0\)) with a radius of 1.265
-(Fig. 1a). Hence, the real components of eigenvalues are highly unlikely
-to all be negative when all \(\gamma_{i} = 1\). But when \(\gamma_{i}\)
-values are separated into two groups, eigenvalues are no longer
-uniformly distributed (Fig. 1b). Instead, two distinct clusters of
-eigenvalues appear (red circles in Fig. 1b), one centred at
-(\(-1.95, 0\)) and the other centred at (\(-0.05, 0\)). The former has a
-large radius, but the real components have shifted to the left (in
-comparison to when \(\gamma = 1\)), while the latter cluster has a
-smaller radius, but real components have shifted to the right; all
-\(\Re({\lambda}) < 0\). Overall, for 1 million randomly assembled
-\(\mathbf{M}\), this division between slow and fast component response
-rates results in more stable systems: 1 stable given \(\gamma = 1\)
-versus 32 stable given \(\gamma = \{1.95, 0.5\}\).
-
-Higher stability in systems with variation in \(\gamma\) can be observed
-by sampling \(\gamma_{i}\) values from various distributions. I focus on
-a uniform distribution where \(\gamma \sim \mathcal{U}(0, 2)\) (see
-Supplementary Information for other distributions, which give similar
-results). As with the case of \(\gamma = \{1.95, 0.5\}\) (Fig. 1b), mean
-\(\gamma = 1\) when \(\gamma \sim \mathcal{U}(0, 2)\), allowing
-comparison of \(\mathbf{M}\) before and after the addition of variation
-in component response rate. Figure 2 shows a comparison of eigenvalue
-distributions given \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\). As
-expected\textsuperscript{\protect\hyperlink{ref-Tao2010}{14}}, when
-\(\gamma = 1\), eigenvalues are distributed uniformly in a circle
-centred at (\(-1, 0\)) with a radius of \(\sigma\sqrt{SC} =\) 2.828.
-Uniform variation in \(\gamma\) leads to a non-uniform distribution of
-eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}--\protect\hyperlink{ref-Stone2017}{18}},
-some of which are clustered locally near the centre of the distribution,
-but others of which are spread outside the former radius of 2.828 (Fig
-2b). The clustering and spreading of eigenvalues introduced by
-\(Var(\gamma)\) can destabilise previously stable systems or stabilise
-systems that are otherwise unstable. But where systems are otherwise too
-complex to be stable given \(\gamma = 1\), the effect of \(Var(\gamma)\)
-can often lead to stability above
-May's\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Allesina2012}{2}}
-threshold \(\sigma\sqrt{SC} < 1\).
+\textbf{Component response rates of random complex systems}. Complex
+systems (\(\mathbf{M}\)) are built from two matrices, one modelling
+component interactions (\(\mathbf{A}\)), and second modelling component
+response rates (\(\boldsymbol{\gamma}\)). Both \(\mathbf{A}\) and
+\(\boldsymbol{\gamma}\) are square \(S \times S\) matrices. Rows in
+\(\mathbf{A}\) define how a given component \(i\) is affected by each
+component \(j\) in the system, including itself (where \(i = j\)).
+Off-diagonal elements of \(\mathbf{A}\) are independent and identically
+distributed (i.i.d), and diagonal elements are set to \(A_{ii} = -1\) as
+in May\textsuperscript{\protect\hyperlink{ref-May1972}{1}}. Diagonal
+elements of \(\boldsymbol{\gamma}\) are positive, and off-diagonal
+elements are set to zero (i.e. \(\boldsymbol{\gamma}\) is a diagonal
+matrix with positive support). The distribution of
+\(diag(\boldsymbol{\gamma})\) over \(S\) components thereby models the
+distribution of component response rates. The dynamics of the entire
+system \(\mathbf{M}\) can be defined as
+follows\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}},
+
+\begin{equation} \label{defM}
+\mathbf{M} = \boldsymbol{\gamma} \mathbf{A}.
+\end{equation}
+
+Equation 1 thereby serves as a null model to investigate how
+variation in component response rate (\(\sigma^{2}_{\gamma}\)) affects
+complex systems. In the absence of such variation
+(\(\sigma^{2}_{\gamma} = 0\)), \(\boldsymbol{\gamma}\) is set to the
+identity matrix (diagonal elements all equal 1) and
+\(\mathbf{M} = \mathbf{A}\). Under these conditions, eigenvalues of
+\(\mathbf{M}\) are distributed
+uniformly\textsuperscript{\protect\hyperlink{ref-Tao2010}{15}} in a
+circle centred at \((-1, 0)\) with a radius of
+\(\sigma \sqrt{SC}\)\textsuperscript{\protect\hyperlink{ref-May1972}{1}}
+(Fig. 1a).
+
+\textbf{Effect of \(\mathbf{\sigma^{2}_{\gamma}}\) on \(\mathbf{M}\)
+(co)variation}. The value of \(\Re(\lambda_{max})\), and therefore
+system stability, can be estimated from five properties of
+\(\mathbf{M}\)\textsuperscript{\protect\hyperlink{ref-Tang2014b}{21}}.
+These properties include (1) system size (\(S\)), (2) mean
+self-regulation of components (\(d\)), (3) mean interaction strength
+between components (\(\mu\)), (4) the variance of between component
+interaction strengths (hereafter \(\sigma^{2}_{M}\), to distinguish from
+\(\sigma^{2}_{A}\) and \(\sigma^{2}_{\gamma}\)), and (5) the correlation
+of interaction strengths between components, \(M_{ij}\) and \(M_{ji}\)
+(\(\rho\))\textsuperscript{\protect\hyperlink{ref-Sommers1988}{22}}.
+Positive \(\sigma^{2}_{\gamma}\) does not change \(S\), nor does it
+necessarily change \(E[d]\) or \(E[\mu]\). What \(\sigma^{2}_{\gamma}\)
+does change is the total variation in component interaction strengths
+(\(\sigma^{2}_{M}\)), and \(\rho\). Introducing variation in \(\gamma\)
+increases the total variation in the system. Variation in the
+off-diagonal elements of \(\mathbf{M}\) is described by the joint
+variation of two random variables,
+
+\begin{equation} \labe{var_ref}
+\sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.
+\end{equation}
+
+Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), Eq. 2
+can be simplified,
+
+\begin{equation} \label{var_reduced}
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\end{equation}
+
+The increase in \(\sigma^{2}_{M}\) caused by \(\sigma^{2}_\gamma\) can
+be visualised from the eigenvalue spectra of \(\textbf{A}\) versus
+\(\textbf{M} = \boldsymbol{\gamma}\textbf{A}\) (Fig. 1). Given \(d = 0\)
+and \(C = 1\), the distribution of eigenvalues of \(\textbf{A}\) and
+\(\textbf{M}\) lie within a circle of a radius \(\sigma_{A}\sqrt{S}\)
+and \(\sigma_{M}\sqrt{S}\), respectively (Fig. 1a vs.~1b). If
+\(d \neq 0\), positive \(\sigma^{2}_\gamma\) changes the distribution of
+eigenvalues\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{23}--\protect\hyperlink{ref-Stone2017}{25}},
+potentially affecting stability (Fig. 1c vs.~1d).
+
+Given \(\sigma^{2}_\gamma = 0\), \(\Re(\lambda_{max})\) increases
+linearly with \(\rho\) such
+that\textsuperscript{\protect\hyperlink{ref-Tang2014c}{26}},
+
+\begin{equation} \label{rho_stab}
+\Re(\lambda_{max}) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\end{equation}
+
+If \(\rho < 0\), such as when \(\textbf{M}\) models a predator-prey
+system in which \(M_{ij}\) and \(M_{ji}\) have opposing signs, stability
+increases\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}. If
+diagonal elements of \(\boldsymbol{\gamma}\) vary independently, the
+magnitude of \(\rho\) is decreased because \(\sigma^{2}_{\gamma}\)
+increases the variance of \(M_{ij}\) without affecting the expected
+covariance between \(M_{ij}\) and \(M_{ji}\) (Figure 2).
+
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\textbf{Figure 2: Complex system correlation versus stability with and
+without variation in component response rates}. Each point represents
+10000 replicate numerical simulations of a random complex system
+\(\mathbf{M} = \gamma \mathbf{A}\) with a fixed correlation between
+off-diagonal elements \(A_{ij}\) and \(A_{ji}\) (\(\rho\), x-axis).
+Where real parts of eigenvalues of \(\mathbf{M}\) are negative (y-axis),
+\(\mathbf{M}\) is stable (black dotted line). Blue circles show systems
+in the absence of variation in component response rates
+(\(\sigma^{2}_{\gamma} = 0\)). Red squares show systems in which
+\(\sigma^{2}_{\gamma} = 1/3\). Arrows show the range of real parts of
+leading eigenvalues observed. Because \(\gamma\) decreases the magnitude
+of \(\rho\), purple lines are included to link replicate simulations
+before (blue circles) and after (red squares) including \(\gamma\). The
+range of \(\rho\) values in which \(\gamma\) decreases the mean real
+part of the leading eigenvalue is indicated with grey shading. In all
+simulations, system size and connectence were set to \(S = 25\) and
+\(C = 1\), respectively. Off-diagonal elements of \(\textbf{A}\) were
+randomly sampled from \(A_{ij} \sim \mathcal{N}(0, 0.4^{2})\), and
+diagonal elements were set to \(-1\). Elements of \(\gamma\) were
+sampled, \(\gamma \sim \mathcal{U}(0, 2)\).
+
+\includegraphics[height=14cm,keepaspectratio]{fig2.pdf}
+
+\vspace{2mm}
+\hrule
+
+\textbf{Numerical simulations of random systems with and without
+\(\mathbf{\sigma^{2}_{\gamma}}\)}. I used numerical simulations and
+eigenanalysis to test how variation in \(\gamma\) affects stability in
+random matrices with known properties, comparing the stability of
+\(\textbf{A}\) versus \(\mathbf{M} = \gamma\mathbf{A}\). Values of
+\(\gamma\) were sampled from a uniform distribution where
+\(\gamma \sim \mathcal{U}(0, 2)\) and \(\sigma^{2}_{\gamma} = 1/3\) (see
+Supplementary Information for other \(\gamma\) distributions, which gave
+similar results). In all simulations, diagonal elements were
+standardised to ensure that \(-d\) between individual \(\textbf{A}\) and
+\(\textbf{M}\) pairs were identical (also note that
+\(E[\gamma_{i}] = 1\)). First I focus on the effect of \(\gamma\) across
+values of \(\rho\), then for increasing system sizes (\(S\)) in random
+and structured networks. By increasing \(S\), the objective is to
+determine the effect of \(\gamma\) as system complexity increases toward
+the boundary at which stability is realistic for a finite system.
+
+\textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{\rho}\)}.
+Numerical simulations revealed that \(\sigma^{2}_{\gamma}\) results in a
+nonlinear relationship between \(\rho\) and \(\Re(\lambda_{max})\),
+which can sometimes increase the stability of the system. Figure 2 shows
+a comparison of \(\Re(\lambda_{max})\) across \(\rho\) values for
+\(\mathbf{A}\) (\(\sigma^{2}_{\gamma} = 0\)) versus \(\mathbf{M}\)
+(\(\sigma^{2}_{\gamma} = 1/3\)) given \(S = 25\), \(C = 1\), and
+\(\sigma_{A} = 0.4\). For \(-0.4 \leq \rho \leq 0.7\) (shaded region of
+Fig. 2), expected \(\Re(\lambda_{max})\) was lower in \(\mathbf{M}\)
+than \(\mathbf{A}\). For \(\rho \geq -0.1\), the lower bound of the
+range of \(\Re(\lambda_{max})\) values also decreased given
+\(\sigma^{2}_{\gamma}\), resulting in negative \(\Re(\lambda_{max})\) in
+\(\mathbf{M}\) for \(\rho = -0.1\) and \(\rho = 0\). Hence, across a
+wide range of system correlations, variation in the response rate of
+system components had a stabilising effect.
+
+The stabilising effect of \(\sigma^{2}_{\gamma}\) across \(\rho\)
+increased with increasing \(S\). Figure 3 shows numerical simulations of
+\(\mathbf{M}\) across increasing \(S\) given \(C = 1\) and
+\(\sigma_{A} = 0.2\) (\(\sigma_{A}\) has been lowered here to better
+illustrate the effect of \(S\); note that now given \(S = 25\),
+\(1 = \sigma_{A}\sqrt{SC}\)). For relatively small systems
+(\(S \leq 25\)), \(\sigma^{2}_{\gamma}\) never decreased the expected
+\(\Re(\lambda_{max})\). But as \(S\) increased, the curvilinear
+relationship between \(\rho\) and \(\Re(\lambda_{max})\) decreased
+expected \(\Re(\lambda_{max})\) for \(\mathbf{M}\) given low magnitudes
+of \(\rho\). In turn, as \(S\) increased, and systems became more
+complex, \(\sigma^{2}_{\gamma}\) increased the proportion of numerical
+simulations that were observed to be stable (see below).
 
 \textbf{Simulation of random \(\mathbf{M}\) across \(\mathbf{S}\)}. To
-investigate the effect of \(Var(\gamma)\) on stability across systems of
-varying complexity, I simulated random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) across \(S = \{2, 3, ..., 49, 50\}\). One
-million \(\mathbf{M}\) were simulated for each \(S\), and the stability
-of \(\mathbf{M}\) was assessed given \(\gamma = 1\) versus
-\(\gamma \sim \mathcal{U}(0, 2)\). For all \(S > 10\), I found that the
-number of stable random systems was higher given \(Var(\gamma)\) than
-when \(\gamma = 1\) (Fig. 3; see Supplementary Information for full
-table of results), and that the difference between the probabilities of
-observing a stable system increased with an increase in \(S\). In other
-words, the potential for \(Var(\gamma)\) to affect stability increased
-with system complexity and was most relevant for systems on the cusp of
-being too complex to be realistically stable. For the highest values of
-\(S\), nearly all systems that were stable given \(Var(\gamma)\) would
-not have been stable given \(\gamma = 1\).
-
-\textbf{Targeted manipulation of \(\mathbf{\gamma}\)}. To further
-investigate the potential of \(Var(\gamma)\) to be stabilising, I used a
-genetic algorithm. Genetic algorithms are heuristic tools that mimic
-evolution by natural selection, and are useful when the space of
-potential solutions (in this case, possible combinations of \(\gamma\)
-values leading to stability in a complex system) is too large to search
-exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}.
-Generations of selection on \(\gamma\) value combinations to minimise
-\(\max\left(\Re(\lambda)\right)\) demonstrated the potential for
-\(Var(\gamma)\) to increase system stability. Across
-\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
-resulted in stable systems with probabilities that were up to four
-orders of magnitude higher than when \(\gamma = 1\) (Fig. 4), meaning
-that stability could often be achieved by manipulating \(S\) \(\gamma\)
-values rather than \(S \times S\) \(\mathbf{M}\) elements (i.e., by
-manipulating component response rates rather than interactions between
-components).
-
-\textbf{System feasibility given \(\mathbf{Var(\gamma)}\)} For complex
-systems in which individual system components represent the density of
-some tangible quantity, it is relevant to consider the feasibility of
-the system. Feasibilility assumes that values of all components are
-positive at
-equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Song2018}{21}}.
+investigate the effect of \(\sigma^{2}_{\gamma}\) on stability across
+systems of increasing complexity, I simulated random
+\(\mathbf{M = \gamma A}\) matrices at \(\sigma_{A} = 0.4\) and \(C = 1\)
+across \(S = \{2, 3, ..., 49, 50\}\). One million \(\mathbf{M}\) were
+simulated for each \(S\), and the stability of \(\mathbf{A}\) vesus
+\(\mathbf{M}\) was assessed given \(\gamma \sim \mathcal{U}(0, 2)\)
+(\(\sigma^{2}_{\gamma} = 1/3\)). For all \(S > 10\), I found that the
+number of stable random systems was higher in \(\mathbf{M}\) than
+\(\mathbf{A}\) (Fig. 4; see Supplementary Information for full table of
+results), and that the difference between the probabilities of observing
+a stable system increased with an increase in \(S\). In other words, the
+potential for \(\sigma^{2}_{\gamma}\) to affect stability increased with
+increasing system complexity and was most relevant for systems on the
+cusp of being too complex to be realistically stable. For the highest
+values of \(S\), nearly all systems that were stable given varying
+\(\gamma\) would not have been stable given \(\gamma = 1\).
+
+I also simulated 100000 \(\mathbf{M}\) for three types of random
+networks that are typically interpreted as modelling three types of
+interspecific ecological
+interactions\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2},\protect\hyperlink{ref-Allesina2011}{27}}.
+These interaction types are competitive, mutualist, and predator-prey,
+as modelled by off-diagonal elements that are constrained to be
+negative, positive, or paired such that if \(A_{ij} > 0\) then
+\(A_{ji} < 0\),
+respectively\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
+(but are otherwise identical to the purely random \(\mathbf{A}\)). As
+\(S\) increased, a higher number of stable \(\mathbf{M}\) relative to
+\(\mathbf{A}\) was observed for competitor and predator-prey, but not
+mutualist, systems. A higher number of stable systems was observed
+whenever \(S > 12\) and \(S > 40\) for competitive and predator-prey
+systems, respectively (note that \(\rho < 0\) for predator-prey systems,
+making stability more likely overall). The stability of mutualist
+systems was never affected by \(\sigma^{2}_{\gamma}\).
+
+The effect of \(\sigma^{2}_{\gamma}\) on stability did not change
+qualitatively across values of \(C\), \(\sigma_{A}\), or for different
+distributions of \(\gamma\) (see Supporting Information).
+
+\textbf{Simulation of structured \(\mathbf{M}\) across \(\mathbf{S}\)}.
+To investigate how \(\sigma^{2}_{\gamma}\) affects the stability of
+commonly observed network structures, I simulated one million
+\(\mathbf{M = \gamma A}\) for
+small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}},
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}, and
+cascade food
+web\textsuperscript{\protect\hyperlink{ref-Solow1998}{18},\protect\hyperlink{ref-Williams2000}{19}}
+networks. In all of these networks, rules determining the presence or
+absence of an interaction between components \(i\) and \(j\) constrain
+the overall structure of the network. In small-world networks,
+interactions between components are constrained so that the expected
+degree of separation between any two components increases in proportion
+to \(\log(S)\)\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}.
+In scale-free networks, the distribution of the number of components
+with which a focal component interacts follows a power law; a few
+components have many interactions while most components have few
+interactions\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. In
+cascade food webs, species are ranked and interactions are constrained
+such that a species \(i\) can only feed on \(j\) if the rank of
+\(i > j\).
+
+\clearpage
+
+\hrule
+\vspace{2mm}
+
+\textbf{Figure 3: System correlation versus stability across different
+system sizes}. In each panel, 10000 random complex systems
+\(\mathbf{M} = \gamma \mathbf{A}\) are simulated for each correlation
+\(\rho = \{-0.90, -0.85, ..., 0.85, 0.90 \}\) between off-diagonal
+elements \(A_{ij}\) and \(A_{ji}\). Lines show the expected real part of
+the leading eigenvalues of \(\mathbf{M}\) (red squares;
+\(\sigma^{2}_{\gamma} = 1/3\)) versus \(\mathbf{A}\) (blue circles;
+\(\sigma^{2}_{\gamma} = 0\)) across \(\rho\), where negative values
+(below the dotted black line) indicate system stability. Differences
+between lines thereby show the effect of component response rate
+variation (\(\gamma\)) on system stability across system correlations
+and sizes (\(S\)). For all simulations, system connectance was
+\(C = 1\). Off-diagonal elements of \(\textbf{A}\) were randomly sampled
+from \(A_{ij} \sim \mathcal{N}(0, 0.2^{2})\), and diagonal elements were
+set to \(-1\). Elements of \(\gamma\) were sampled such that
+\(\gamma \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
+
+\includegraphics{fig3.pdf}
+
+\vspace{2mm}
+\hrule
+
+Network structure did not strongly modulate the effect that
+\(\sigma^{2}_{\gamma}\) had on stability. For comparable magnitudes of
+complexity, structured networks still had a higher number of stable
+\(\mathbf{M}\) than \(\mathbf{A}\). For random networks,
+\(\sigma^{2}_{\gamma}\) increased stability given \(S > 10\)
+(\(\sigma_{A} = 0.4\) and \(C = 1\)), and therefore complexity
+\(\sigma_{A} \sqrt{SC} \gtrapprox 1.26\). This threshold of complexity,
+above which more \(\mathbf{M}\) than \(\mathbf{A}\) were stable, was
+comparable for small-world networks, and slightly lower for scale-free
+networks (note that algorithms for generating small-world and scale-free
+networks necessarily led to varying \(C\); see methods). Varying
+\(\gamma\) increased stability in cascade food webs for \(S > 27\), and
+therefore at a relatively low complexity magnitudes compared to random
+predator-prey networks (\(S > 40\)). Overall, network structure did not
+greatly change the effect that \(\sigma^{2}_{\gamma}\) had on increasing
+the upper bound of complexity within which stability might reasonably be
+observed.
+
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\textbf{Figure 4: Stability of large complex systems with and without
+variation in component response rate (\(\boldsymbol{\gamma}\)).} The
+\(\log\) number of systems that are stable across different system sizes
+(\(S = \{2, 3, ..., 49, 50 \}\)) given \(C = 1\), and the proportion of
+systems for which variation in \(\gamma\) is critical for system
+stability. For each \(S\), 1 million complex systems are randomly
+generated. Stability of each complex system is tested given variation in
+\(\gamma\) by randomly sampling \(\gamma \sim \mathcal{U}(0, 2)\).
+Stability given \(\sigma^{2}_{\gamma}>0\) is then compared to stability
+in an otherwise identical system in which
+\(\gamma_{i} = E[\mathcal{U}(0, 2)]\) for all components. Blue and red
+bars show the number of stable systems in the absence and presence of
+\(\sigma^{2}_{\gamma}\), respectively. The black line shows the
+proportion of systems that are stable when \(\sigma^{2}_{\gamma}>0\),
+but would be unstable if \(\sigma^{2}_{\gamma}=0\) (i.e., the
+conditional probability that \(\mathbf{A}\) is unstable given that
+\(\mathbf{M}\) is stable).
+
+\includegraphics{fig4.pdf}
+
+\vspace{2mm}
+\hrule
+
+\textbf{System feasibility given \(\mathbf{\sigma^{2}_{\gamma}}\)} For
+complex systems in which individual system components represent the
+density of some tangible quantity, it is relevant to consider the
+feasibility of the system. Feasibilility assumes that values of all
+components are positive at
+equilibrium\textsuperscript{\protect\hyperlink{ref-Grilli2017}{6},\protect\hyperlink{ref-Dougoud2018}{28},\protect\hyperlink{ref-Song2018}{29}}.
 This is of particular interest for ecological communities because
-population density (\(N\)) cannot take negative values, meaning that
+population density (\(n\)) cannot take negative values, meaning that
 ecological systems need to be feasible for stability to be biologically
-realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}}.
+realistic\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}}.
 While my results are intended to be general to all complex systems, and
 not restricted to species networks, I have also performed a feasibility
-analysis on all matrices \(\mathbf{M}\) tested for stability, and
-additionally for specific types of ecological
-communities\textsuperscript{\protect\hyperlink{ref-Allesina2012}{2}}
-(e.g., competitive, mutualist, predator-prey; see Supplementary
-Information). I emphasise that \(\gamma\) is not interpreted as
-population density in this analysis, but instead as a fundamental
-property of species life history such as expected generation time.
-Feasibility was unaffected by \(Var(\gamma)\) and instead occurred with
-a fixed probability of \(1/2^{S}\), consistent with a recent proof by
-Serván et al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}}
-(see Supplementary Information). Hence, for pure interacting species
+analysis on all matrices tested for stability. I emphasise that
+\(\gamma\) is not interpreted as population density in this analysis,
+but instead as a fundamental property of species life history such as
+expected generation time. Feasibility was unaffected by
+\(\sigma^{2}_{\gamma}\) and instead occurred with a fixed probability of
+\(1/2^{S}\), consistent with a recent proof by Serván et
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{30}} (see
+Supplementary Information). Hence, for pure interacting species
 networks, variation in component response rate (i.e., species generation
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 1: Example distribution of eigenvalues before (a) and
-after (b) separating a randomly generated complex system into fast
-(\(\boldsymbol{\gamma} = 1.95\)) and slow
-(\(\boldsymbol{\gamma} = 0.05\)) component response rates.} Each panel
-shows the same system where \(S = 200\), \(C = 0.05\), and
-\(\sigma = 0.4\), and in each case \(E[\gamma] = 1\) (i.e., only the
-distribution of \(\gamma\) differs between panels). \textbf{a.}
-Eigenvalues plotted when all \(\gamma = 1\); distributions of points are
-uniformly distributed within the blue circle with a radius of
-\(\sigma\sqrt{SC} =\) 1.238 centred at -1 on the real axis. \textbf{b.}
-Eigenvalues plotted when half \(\gamma = 1.95\) and half
-\(\gamma = 0.05\); distributions of points can be partitioned into one
-large circle centred at \(\gamma = -1.95\) and one small circle centred
-at \(\gamma = -0.05\). In a, the maximum real eigenvalue
-\(\max\left(\Re(\lambda)\right) =\) 0.2344871, while in b
-\(\max\left(\Re(\lambda)\right) =\) -0.0002273135, meaning that the
-complex system in b but not a is stable because in b
-\(\max\left(\Re(\lambda)\right) < 0\). In 1 million randomly generated
-complex systems under the same parameter values, 1 was stable when
-\(\gamma = 1\) while 32 were stable when \(\gamma = \{1.95, 0.05\}\).
-Overall, complex systems that are separated into fast versus slow
-components tend to be more stable than otherwise identical systems with
-identical component response rates.
-
-\includegraphics{unnamed-chunk-10-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\textbf{Targeted manipulation of \(\boldsymbol{\gamma}\)}. To further
+investigate the potential of \(\sigma^{2}_{\gamma}\) to be stabilising,
+I used a genetic algorithm. Genetic algorithms are heuristic tools that
+mimic evolution by natural selection, and are useful when the space of
+potential solutions (in this case, possible combinations of \(\gamma\)
+values leading to stability in a complex system) is too large to search
+exhaustively\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{31}}.
+Generations of selection on \(\gamma\) value combinations to minimise
+\(\Re(\lambda_{max})\) demonstrated the potential for
+\(\sigma^{2}_{\gamma}\) to increase system stability. Across
+\(S = \{2, 3, ..., 39, 40\}\), sets of \(\gamma\) values were found that
+resulted in stable systems with probabilities that were up to four
+orders of magnitude higher than when \(\gamma = 1\) (see Supplementary
+Information), meaning that stability could often be achieved by
+manipulating \(S\) \(\gamma\) values rather than \(S \times S\)
+\(\mathbf{M}\) elements (i.e., by manipulating component response rates
+rather than interactions between components).
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 2: Distributions of eigenvalues before (a) and after (b)
-introducing variation in component response rate
-(\(\boldsymbol{\gamma}\)) in complex systems.} Each panel show the same
-system where \(S = 1000\), \(C = 0.05\), and \(\sigma = 0.4\).
-\textbf{a.} Eigenvalues plotted in the absence of \(Var(\gamma)\) where
-\(E[\gamma] = 1\), versus \textbf{b.} eigenvalues plotted given
-\(\gamma \sim \mathcal{U}(0, 2)\), which increases the variance of
-interaction strengths (\(\sigma^{2}\)) but also creates a cluster of
-eigenvalues toward the distribution's centre (-1, 0). Blue elipses in
-both panels show the circle centred on the distribution in panel a.
-Proportions of \(\Re(\lambda) < 0\) are 0.724 and 0.742 for a and b,
-respectively.
-
-\includegraphics{unnamed-chunk-13-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
 It is important to recognise that variation in component response rate
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 3: Stability of large complex systems with and without
-variation in component response rate (\(\boldsymbol{\gamma}\)).} The
-\(\ln\) number of systems that are stable across different system sizes
-(\(S\), max \(S=50\)) given \(C = 1\), and the proportion of systems in
-which variation in \(\gamma\) is critical for system stability. For each
-\(S\), 1 million complex systems are randomly generated. Stability of
-each complex system is tested given variation in \(\gamma\) by randomly
-sampling \(\gamma \sim \mathcal{U}(0, 2)\). Stability given
-\(Var(\gamma)\) is then compared to stability in an otherwise identical
-system in which \(\gamma = E[\mathcal{U}(0, 2)]\) for all components.
-Blue and red bars show the number of stable systems in the absence and
-presence of \(Var(\gamma)\), respectively. The black line shows the
-proportion of systems that are stable when \(Var(\gamma) > 0\), but
-would be unstable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-15-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-My focus is distinct from Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}}, who applied
+My focus here is distinct from Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24}}, who applied
 the same mathematical framework to investigate how a diagonal matrix
-\(\mathbf{X}\) (equivalent to \(\gamma\) in my model) affects the
-stability of a community matrix \(\mathbf{M}\) given an interaction
-matrix \(\mathbf{A}\) within a generalised Lotka-Volterra model, where
-\(\mathbf{M} = \mathbf{XA}\). Gibbs et
-al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17}} analytically
+\(\mathbf{X}\) (equivalent to \(\boldsymbol{\gamma}\) in my model)
+affects the stability of a community matrix \(\mathbf{M}\) given an
+interaction matrix \(\mathbf{A}\) within a generalised Lotka-Volterra
+model, where \(\mathbf{M} = \mathbf{XA}\). Gibbs et
+al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24}} analytically
 demonstrated that the effect of \(\mathbf{X}\) on system stability
 decreases exponentially as system size becomes arbitrarily large
-(\(S \to \infty\)) for a given magnitude of complexity \(\sigma\sqrt{SC}\). 
-My numerical results do not contradict this
+(\(S \to \infty\)) for a given magnitude of complexity
+\(\sigma\sqrt{SC}\). My numerical results do not contradict this
 prediction because I did not scale \(\sigma = 1 / \sqrt{S}\), but
 instead fixed \(\sigma\) and increased \(S\) to thereby increase total
-system complexity (see Supplementary Information for
-results simulated across \(\sigma\) and \(C\)). Overall, I show that
-component response rate variation increases the upper bound of
-complexity at which stability can be realistically observed, meaning
-that highly complex systems are more likely than not to vary in their
-component response rates, and for this variation to be critical for
-system stability.
+system complexity (see Supplemental Information for results simulated
+across \(\sigma\) and \(C\)). Overall, I show that component response
+rate variation increases the upper bound of complexity at which
+stability can be realistically observed, meaning that highly complex
+systems are more likely than not to vary in their component response
+rates, and for this variation to be critical for system stability.
+
+Interestingly, while complex systems were more likely to be stable given
+variation in component response rate, they were not more likely to be
+feasible, meaning that stability was not increased when component values
+were also restricted to being positive at equilibrium. Feasibility is
+important to consider, particularly for the study of ecological networks
+of
+species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{6},\protect\hyperlink{ref-Stone2017}{25},\protect\hyperlink{ref-Dougoud2018}{28},\protect\hyperlink{ref-Servan2018}{30}}
+because population densities cannot realistically be negative. My
+results therefore suggest that variation in the rate of population
+responses to perturbation (e.g., due to differences in generation time
+among species) is unlikely to be critical to the stability of purely
+multi-species interaction networks (see also Supplementary Information).
+Nevertheless, ecological interactions do not exist in isolation in
+empirical
+systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, but
+instead interact with evolutionary, abiotic, or social-economic systems.
+The relevance of component response rate for complex system stability
+should therefore not be ignored in the broader context of ecological
+communities.
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{Figure 4: Stability of large complex systems given
-\(\boldsymbol{\gamma = 1}\) versus targeted
-\(\boldsymbol{Var(\gamma)}\).} The \(\ln\) number of systems that are
-stable across different system sizes (\(S\), max \(S=40\)) for
-\(C = 1\), and the proportion of systems wherein a targeted search of
-\(\gamma\) values successfully resulted in system stability. For each
-\(S\), 100000 complex systems are randomly generated. Stability of each
-complex system is tested given variation in \(\gamma\) using a genetic
-algorithm to maximise the effect of \(\gamma\) values on increasing
-stability, as compared to stability in an otherwise identical system in
-which \(\gamma\) is the same for all components. Blue bars show the
-number of stable systems in the absence of component response rate
-variation, while red bars show the number of stable systems that can be
-generated if component response rate is varied to maximise system
-stability. The black line shows the proportion of systems that are
-stable when component response rate is targeted to increase stability,
-but would not be stable if \(Var(\gamma) = 0\).
-
-\includegraphics{unnamed-chunk-16-1.pdf}
-
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-Interestingly, while complex systems were more likely to be stable given
-variation in component response rate, they were not more likely to be
-feasible, meaning that stability was not increased when component values
-were also restricted to being positive at equilibrium. Feasibility is
-important to consider, particularly for the study of ecological networks
-of
-species\textsuperscript{\protect\hyperlink{ref-Grilli2017}{5},\protect\hyperlink{ref-Stone2017}{18},\protect\hyperlink{ref-Dougoud2018}{20},\protect\hyperlink{ref-Servan2018}{22}}
-because population densities cannot realistically be negative. My
-results therefore suggest that variation in the rate of population
-responses to perturbation (e.g., due to differences in generation time
-among species) is unlikely to be critical to the stability of purely
-multi-species interaction networks (see also Supplementary Information).
-Nevertheless, ecological interactions do not exist in isolation in
-empirical
-systems\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, but
-instead interact with evolutionary, abiotic, or social-economic systems.
-The relevance of component response rate for complex system stability
-should therefore not be ignored in the broader context of ecological
-communities.
-
 A general mathematical framework encompassing shifts in eigenvalue
-distributions caused by a vector \(\gamma\) has been
-investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{16}}
+distributions caused by a diagonal matrix \(\boldsymbol{\gamma}\) has
+been
+investigated\textsuperscript{\protect\hyperlink{ref-Ahmadian2015}{23}}
 and recently applied to questions concerning species density and
-feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{17},\protect\hyperlink{ref-Stone2017}{18}},
-but \(\gamma\) has not been interpreted as rates of response of
-individual system components to perturbation. My model focuses on
-component response rates for systems of a finite size, in which
-complexity is high but not yet high enough to make the probability of
-stability unrealistically low for actual empirical systems. For this
+feasibility\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{24},\protect\hyperlink{ref-Stone2017}{25}},
+but \(\boldsymbol{\gamma}\) has not been interpreted as rates of
+response of individual system components to perturbation. My model
+focuses on component response rates for systems of a finite size, in
+which complexity is high but not yet high enough to make the probability
+of stability unrealistically low for actual empirical systems. For this
 upper range of system size, randomly assembled complex systems are more
 likely to be stable if their component response rates vary (e.g.,
-\(10 < S < 30\) for parameter values in Fig. 3). Overall, I suggest that
-variation in component response rate might therefore be critical for
-maintaining stability in many highly complex empirical systems. These
-results are broadly applicable for understanding the stability of
-complex networks across the physical, life, and social sciences.
+\(10 < S < 30\) for parameter values in Fig. 4). Variation in component
+response rate might therefore be critical for maintaining stability in
+many highly complex empirical systems. These results are broadly
+applicable for understanding the stability of complex networks across
+the physical, life, and social sciences.
 
@@
 
-\textbf{Component response rate variation (\(\mathbf{\gamma}\))}. In a
-synthesis of eco-evolutionary feedbacks on community stability, Patel et
-al. model a system that includes a vector of potentially changing
-species densities (\(\mathbf{N}\)) and a vector of potentially evolving
+\textbf{Component response rate (\(\boldsymbol{\gamma}\)) variation}. In
+a synthesis of eco-evolutionary feedbacks on community stability, Patel
+et al. model a system that includes a vector of potentially changing
+species densities (\(\mathbf{n}\)) and a vector of potentially evolving
 traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
 For any species \(i\) or trait \(j\), change in species density
-(\(N_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
-of the vectors \(\mathbf{N}\) and \(\mathbf{x}\),
-
-\[\frac{dN_{i}}{dt} = N_{i}f_{i}(\mathbf{N}, \mathbf{x}),\]
-
-\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{N}, \mathbf{x}).\]
+(\(n_{i}\)) or trait value (\(x_{j}\)) with time (\(t\)) is a function
+of the vectors \(\mathbf{n}\) and \(\mathbf{x}\),
+
+\[\frac{dn_{i}}{dt} = n_{i}f_{i}(\mathbf{n}, \mathbf{x}),\]
+
+\[\frac{dx_{j}}{dt} = \epsilon g_{j}(\mathbf{n}, \mathbf{x}).\]
 
@@
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}, and
+densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 evolution and low \(\epsilon\) modelling relative slow
-evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{15}}.
+evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
 
@@
 components. Following
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{23}},
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-May1973}{32}},
 the value of a component \(i\) at time \(t\) (\(v_{i}(t)\)) is affected
@@
 
-In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{23}},
-
-\[\frac{d\mathbf{v}(t)}{dt} = \mathbf{\gamma} \mathbf{A}\mathbf{v}(t).\]
-
-In the above, \(\mathbf{\gamma}\) is a diagonal matrix in which elements
-correspond to individual component response rates. Therefore,
-\(\mathbf{M} = \mathbf{\gamma} \mathbf{A}\) defines the values of system
-components and can be analysed using the techniques of
-May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{16},\protect\hyperlink{ref-May1973}{23}}.
+In matrix notation\textsuperscript{\protect\hyperlink{ref-May1973}{32}},
+
+\[\frac{d\mathbf{v}(t)}{dt} = \boldsymbol{\gamma} \mathbf{A}\mathbf{v}(t).\]
+
+In the above, \(\boldsymbol{\gamma}\) is a diagonal matrix in which
+elements correspond to individual component response rates. Therefore,
+\(\mathbf{M} = \boldsymbol{\gamma} \mathbf{A}\) defines the change in
+values of system components and can be analysed using the techniques of
+May\textsuperscript{\protect\hyperlink{ref-May1972}{1},\protect\hyperlink{ref-Ahmadian2015}{23},\protect\hyperlink{ref-May1973}{32}}.
 In these analyses, row means of \(\mathbf{A}\) are expected to be
@@
 means that is attributable to \(\mathbf{A}\) is small relative to that
-attributable to \(\mathbf{\gamma}\), especially at high \(S\). Variation
-in \(\mathbf{\gamma}\) specifically isolates the effects of differing
-component response rates, hence causing differences in expected
-\(\mathbf{M}\) row means.
-
-\textbf{Genetic algorithm}. Ideally, to investigate the potential of
-\(Var(\gamma)\) for increasing the proportion of stable complex systems,
-the search space of all possible \(\gamma\) vectors would be evaluated
-for each unique \(\mathbf{M = \gamma A}\). This is technically
-impossible because \(\gamma_{i}\) can take any real value between 0-2,
-but even rounding \(\gamma_{i}\) to reasonable values would result in a
-search space too large to practically explore. Under these conditions,
-genetic algorithms are highly useful tools for finding practical
-solutions by mimicking the process of biological
-evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{19}}. In
-this case, the practical solution is finding vectors of
-\(\mathbf{\gamma}\) that decrease the most positive real eigenvalue of
-\(\mathbf{M}\). The genetic algorithm used achieves this by initialising
-a large population of 1000 different potential \(\mathbf{\gamma}\)
-vectors and allowing this population to evolve through a process of
-mutation, crossover (swaping \(\gamma_{i}\) values between vectors),
-selection, and reproduction until either a \(\mathbf{\gamma}\) vector is
-found where all \(\Re(\lambda) < 0\) or some ``giving up'' critiera is
-met.
-
-For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
-for 100000 random \(\mathbf{M}\) (\(\sigma = 0.4\), \(C = 1\)). The
-genetic algorithm was initialised with a population of 1000 different
-\(\mathbf{\gamma}\) vectors with elements sampled i.i.d from
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis was performed on
-the \(\mathbf{M}\) resulting from each \(\mathbf{\gamma}\) vector, and
-the 20 \(\mathbf{\gamma}\) vectors resulting in \(\mathbf{M}\) with the
-lowest \(\max\left(\Re(\lambda)\right)\) each produced 50 clonal
-offspring with subsequent random mutation and crossover between the
-resulting new generation of 1000 \(\mathbf{\gamma}\) vectors. Mutation
-of each \(\gamma_{i}\) in a \(\mathbf{\gamma}\) vector occurred with a
-probability of 0.2, resulting in a mutation effect of size
-\(\mathcal{N}(0, 0.02)\) being added to generate the newly mutated
-\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
-multiplied by \(-1\), and any values that mutated above 2 were set to
-2). Crossover occurred between two sets of 100 \(\mathbf{\gamma}\)
-vectors paired in each generation; vectors were randomly sampled with
-replacement among but not within sets. Vector pairs selected for
-crossover swapped all elements between and including two \(\gamma_{i}\)
-randomly selected with replacement (this allowed for reversal of vector
-element positions during crossover; e.g.,
-\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
-). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
-found, 20 generations occurred, or if the mean \(\mathbf{\gamma}\)
-fitness increase between generations was less than 0.01 (where fitness
-was defined as \(W_{\gamma} = -\max\left(\Re(\lambda)\right)\) for
-\(\mathbf{M}\)).
+attributable to \(\boldsymbol{\gamma}\), especially at high \(S\).
+Variation in \(\boldsymbol{\gamma}\) specifically isolates the effects
+of differing component response rates, hence causing differences in
+expected \(\mathbf{M}\) row means.
+
+\textbf{Construction of random and structured networks}. I used the R
+programming language for all numerical simulations and
+analyses\textsuperscript{\protect\hyperlink{ref-Rproject}{33}}. Purely
+random networks were generated by sampling off-diagonal elements from an
+i.i.d \(A_{ij} \sim \mathcal{N}(0, 0.4^{2})\) with a probability \(C\)
+(unsampled elements were set to zero). Diagonal elements \(A_{ii}\) were
+set to \(-1\). Elements of \(\boldsymbol{\gamma}\) were simulated i.i.d.
+from a distribution with positive support (typically
+\(\gamma \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices with
+correlated elements \(A_{ij}\) and \(A_{ji}\) were built using Cholesky
+decomposition. Competitor networks in which off-diagonal elements
+\(A_{ij} \leq 0\) were constructed by first building a random
+\(\mathbf{A}\), then flipping the sign of any elements in which
+\(A_{ij} > 0\). Similarly, mutualist networks were constructed by
+building a random \(\mathbf{A}\), then flipping the sign of elements
+where \(A_{ij} < 0\). Predator-prey networks were constructed by first
+building a random \(\mathbf{A}\), then flipping the sign of either
+\(A_{ij}\) or \(A_{ji}\) if \(A_{ij} \times A_{ji} > 0\).
+
+Small-world networks were constructed using the method of Watts and
+Strogatz\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}. First,
+a regular network\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}}
+was created such that components were arranged in a circle. Each
+component was initially set to interact with its \(k/2\) closest
+neighbouring components on each side, where \(k\) was an even natural
+number (e.g., for \(k = 2\) the regular network forms a ring in which
+each component interacts with its two adjacent neighbours; see
+Supplemental Material for examples). Each interaction between a focal
+component and its neighbour was then removed and replaced with with a
+probability of \(\beta\). In replacement, a new component was randomly
+selected to interact with the focal component; selection was done with
+equal probability among all but the focal component. The resulting
+small-world network was represented by a square \(S \times S\) binary
+matrix \(\mathbf{B}\) in which 1s represented interactions between
+components and 0s represented the absence of an interaction. A new
+random matrix \(\mathbf{J}\) was then generated with elements \(J_{ij}\)
+sampled i.i.d. from \(\mathcal{N}(0, 0.4^{2})\). To build the
+interaction matrix \(\mathbf{A}\), I used element-wise multiplication
+\(\mathbf{A} = \mathbf{J} \odot \mathbf{B}\), then set
+\(diag(\mathbf{A}) = -1\). I set \(k = S/12\) and simulated small-world
+networks across all combinations of
+\(S = \{24, 48, 72, 96, 120, 144, 168\}\) and
+\(\beta = \{0, 0.01, 0.1, 0.25, 1\}\).
+
+Scale-free networks were constructed using the method of Albert and
+Barabási\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. First,
+a saturated network (all components interact with each other) of size
+\(m \leq S\) was created. New components were then added sequentially to
+the network; each newly added component was set to interact with \(m\)
+randomly selected existing components. When the system size reached
+\(S\), the distribution of the number of total interactions that
+components had followed a power-law
+tail\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}. The
+resulting network was represented by an \(S \times S\) binary matrix
+\(\mathbf{G}\), where 1s and 0s represent the presence and absence of an
+interaction, respectively. As with small-world networks, a random matrix
+\(\mathbf{J}\) was generated, and
+\(\mathbf{A} = \mathbf{J} \odot \mathbf{G}\). Diagonal elements were set
+to \(-1\). I simulated scale-free networks across all combinations of
+\(S = \{24, 48, 72, 96, 120\}\) and \(m = \{2, 3, ..., 11, 12\}\).
+
+Cascade food webs were constructed following Solow and
+Beet\textsuperscript{\protect\hyperlink{ref-Solow1998}{18}}. First, a
+random matrix \(\mathbf{A}\) was generated with off-diagonal elements
+sampled i.i.d so that \(A_{ij} \sim \mathcal{N}(0, 0.4^2)\). Each
+component in the system was ranked from \(1\) to \(S\). If component
+\(i\) had a higher rank than component \(j\) and \(A_{ij} < 0\), then
+\(A_{ij}\) was multiplied by \(-1\). If \(i\) had a lower rank than
+\(j\) and \(A_{ji} < 0\), then \(A_{ji}\) was multiplied by \(-1\). In
+practice, this resulted in a matrix \(\mathbf{A}\) with negative and
+positive values in the lower and upper triangles, respectively. Diagonal
+elements of \(\mathbf{A}\) were set to \(-1\) and \(C = 1\). I simulated
+cascade food webs for \(S = \{2, 3, ..., 59, 60\}\).
 
 \textbf{System feasibility}. Dougoud et
-al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} identify
+al.\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}} identify
 the following feasibility criteria for ecological systems characterised
@@
 stability
-criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{20}} by
+criterion\textsuperscript{\protect\hyperlink{ref-Dougoud2018}{28}} by
 \(\frac{\sigma}{(CS)^{-\delta}}\sqrt{SC} < 1\), and hence for my
@@
 \(\mathbf{A} = \theta\mathbf{I + J}\), the above criteria is therefore
-reduced to the below\textsuperscript{\protect\hyperlink{ref-Servan2018}{22}},
+reduced to the below (see also Serván et
+al.\textsuperscript{\protect\hyperlink{ref-Servan2018}{30}}),
 
@@
 therefore evaluated \(\mathbf{-M^{-1}r}\) (\(\mathbf{r}\) elements were
-sampled i.i.d. from \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility
-is satisfied if all of the elements of the resulting vector are
-positive.
+sampled i.i.d. from \(r \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility is
+satisfied if all of the elements of the resulting vector are positive.
+
+\textbf{Genetic algorithm}. Ideally, to investigate the potential of
+\(\sigma^{2}_{\gamma}\) for increasing the proportion of stable complex
+systems, the search space of all possible \(diag(\boldsymbol{\gamma})\)
+vectors would be evaluated for each unique \(\mathbf{M = \gamma A}\).
+This is technically impossible because \(\gamma_{i}\) can take any real
+value between 0-2, but even rounding \(\gamma_{i}\) to reasonable values
+would result in a search space too large to practically explore. Under
+these conditions, genetic algorithms are highly useful tools for finding
+practical solutions by mimicking the process of biological
+evolution\textsuperscript{\protect\hyperlink{ref-Hamblin2013}{31}}. In
+this case, the practical solution is finding vectors of
+\(diag(\boldsymbol{\gamma})\) that decrease the most positive real
+eigenvalue of \(\mathbf{M}\). The genetic algorithm used achieves this
+by initialising a large population of 1000 different potential
+\(diag(\boldsymbol{\gamma})\) vectors and allowing this population to
+evolve through a process of mutation, crossover (swaping \(\gamma_{i}\)
+values between vectors), selection, and reproduction until either a
+\(diag(\boldsymbol{\gamma})\) vector is found where all
+\(\Re(\lambda) < 0\) or some ``giving up'' critiera is met.
+
+For each \(S = \{2, 3, ..., 39, 40\}\), the genetic algorithm was run
+for 100000 random \(\mathbf{M = \gamma A}\) (\(\sigma_{A} = 0.4\),
+\(C = 1\)). The genetic algorithm was initialised with a population of
+1000 different \(diag(\boldsymbol{\gamma})\) vectors with elements
+sampled i.i.d from \(\gamma \sim \mathcal{U}(0, 2)\). Eigenanalysis was
+performed on the \(\mathbf{M}\) resulting from each
+\(\boldsymbol{\gamma}\), and the 20 \(diag(\boldsymbol{\gamma})\)
+vectors resulting in \(\mathbf{M}\) with the lowest
+\(\Re(\lambda_{max})\) each produced 50 clonal offspring with subsequent
+random mutation and crossover between the resulting new generation of
+1000 \(diag(\boldsymbol{\gamma})\) vectors. Mutation of each
+\(\gamma_{i}\) in a \(diag(\boldsymbol{\gamma})\) vector occurred with a
+probability of 0.2, resulting in a mutation effect of size
+\(\mathcal{N}(0, 0.02^{2})\) being added to generate the newly mutated
+\(\gamma_{i}\) (any \(\gamma_{i}\) values that mutated below zero were
+multiplied by \(-1\), and any values that mutated above 2 were set to
+2). Crossover occurred between two sets of 100
+\(diag(\boldsymbol{\gamma})\) vectors paired in each generation; vectors
+were randomly sampled with replacement among but not within sets. Vector
+pairs selected for crossover swapped all elements between and including
+two \(\gamma_{i}\) randomly selected with replacement (this allowed for
+reversal of vector element positions during crossover; e.g.,
+\(\{\gamma_{4}, \gamma_{5}, \gamma_{6}, \gamma_{7}\} \to \{\gamma_{7}, \gamma_{6}, \gamma_{5}, \gamma_{4}\}\)
+). The genetic algorithm terminated if a stable \(\mathbf{M}\) was
+found, 20 generations occurred, or if the mean \(\boldsymbol{\gamma}\)
+fitness increase between generations was less than 0.01 (where fitness
+was defined as \(W_{\gamma} = -\Re(\lambda_{max})\) for \(\mathbf{M}\)).
 
@@
 
+\textbf{Supplementary Information:} Full tables of stability results for
+simulations across different system size (\(S\)) values, ecological
+community types, connectance (\(C\)) values, interaction strengths
+(\(\sigma\)), and \(\gamma\) distributions are provided as supplementary
+material. An additional table also shows results for how feasibility
+changes across \(S\). All code and simulation outputs are publicly
+available as part of the RandomMatrixStability package on GitHub
+(\url{https://github.com/bradduthie/RandomMatrixStability}).
+
+\textbf{Additional Information:} The author declares no competing
+interests. All work was carried out by A. Bradley Duthie, and all code
+and data are accessible on
+\href{https://github.com/bradduthie/RandomMatrixStability}{GitHub}.
+
 \textbf{References}
@@
 
+\hypertarget{ref-Watts1998}{}
+16. Watts, D. J. \& Strogatz, S. H. Collective dynamics of 'small world'
+networks. \emph{Nature} \textbf{393,} 440--442 (1998).
+
+\hypertarget{ref-Albert2002}{}
+17. Albert, R. \& Barabási, A. L. Statistical mechanics of complex
+networks. \emph{Reviews of Modern Physics} \textbf{74,} 47--97 (2002).
+
+\hypertarget{ref-Solow1998}{}
+18. Solow, A. R. \& Beet, A. R. On lumping species in food webs.
+\emph{Ecology} \textbf{79,} 2013--2018 (1998).
+
+\hypertarget{ref-Williams2000}{}
+19. Williams, R. J. \& Martinez, N. D. Simple rules yield complex food
+webs. \emph{Nature} \textbf{404,} 180--183 (2000).
+
 \hypertarget{ref-Patel2018}{}
-15. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
+20. Patel, S., Cortez, M. H. \& Schreiber, S. J. Partitioning the
 effects of eco-evolutionary feedbacks on community stability.
@@
 
+\hypertarget{ref-Tang2014b}{}
+21. Tang, S. \& Allesina, S. Reactivity and stability of large
+ecosystems. \emph{Frontiers in Ecology and Evolution} \textbf{2,} 1--8
+(2014).
+
+\hypertarget{ref-Sommers1988}{}
+22. Sommers, H. J., Crisanti, A., Sompolinsky, H. \& Stein, Y. Spectrum
+of large random asymmetric matrices. \emph{Physical Review Letters}
+\textbf{60,} 1895--1898 (1988).
+
 \hypertarget{ref-Ahmadian2015}{}
-16. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
+23. Ahmadian, Y., Fumarola, F. \& Miller, K. D. Properties of networks
 with partially structured and partially random connectivity.
@@
 \hypertarget{ref-Gibbs2017}{}
-17. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
+24. Gibbs, T., Grilli, J., Rogers, T. \& Allesina, S. The effect of
 population abundances on the stability of large random ecosystems.
@@
 \hypertarget{ref-Stone2017}{}
-18. Stone, L. The feasibility and stability of large complex biological
+25. Stone, L. The feasibility and stability of large complex biological
 networks: a random matrix approach. \emph{Scientific Reports}
@@
 
-\hypertarget{ref-Hamblin2013}{}
-19. Hamblin, S. On the practical usage of genetic algorithms in ecology
-and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
-184--194 (2013).
+\hypertarget{ref-Tang2014c}{}
+26. Tang, S., Pawar, S. \& Allesina, S. Correlation between interaction
+strengths drives stability in large ecological networks. \textbf{17,}
+1094--1100 (2014).
+
+\hypertarget{ref-Allesina2011}{}
+27. Allesina, S. \& Levine, J. M. A competitive network theory of
+species diversity. \emph{Proceedings of the National Academy of Sciences
+of the United States of America} \textbf{108,} 5638--5642 (2011).
 
 \hypertarget{ref-Dougoud2018}{}
-20. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
+28. Dougoud, M., Vinckenbosch, L., Rohr, R., Bersier, L.-F. \& Mazza, C.
 The feasibility of equilibria in large ecosystems: a primary but
@@
 \hypertarget{ref-Song2018}{}
-21. Song, C. \& Saavedra, S. Will a small randomly assembled community
+29. Song, C. \& Saavedra, S. Will a small randomly assembled community
 be feasible and stable? \emph{Ecology} \textbf{99,} 743--751 (2018).
@@
 \hypertarget{ref-Servan2018}{}
-22. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
+30. Serván, C. A., Capitán, J. A., Grilli, J., Morrison, K. E. \&
 Allesina, S. Coexistence of many species in random ecosystems.
@@
 
+\hypertarget{ref-Hamblin2013}{}
+31. Hamblin, S. On the practical usage of genetic algorithms in ecology
+and evolution. \emph{Methods in Ecology and Evolution} \textbf{4,}
+184--194 (2013).
+
 \hypertarget{ref-May1973}{}
-23. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
+32. May, R. M. Qualitative stability in model ecosystems. \emph{Ecology}
 \textbf{54,} 638--641 (1973).
 
+\hypertarget{ref-Rproject}{}
+33. R Core Team. \emph{R: A language and environment for statistical
+computing}. (R Foundation for Statistical Computing, 2018).
+
 \clearpage
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\textbf{This Supplementary information supports the manuscript
-``Component response rate variation drives stability in large complex
+\section{Supplemental Information}
+
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\textbf{This supplemental information supports the manuscript
+``Component response rate variation underlies the stability of complex
 systems'' with additional analyses to support its conclusions. All text,
@@
 package} includes all functions and tools for recreating the text, this
-Supplementary information, and running all code; additional documentation
+supplemental information, and running all code; additional documentation
 is also provided for package functions. The RandomMatrixStability
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
-
-\section{Supplementary Information table of
-contents}\label{Supplementary-information-table-of-contents}
+\vspace{2mm}
+\hrule
+\vspace{2mm}
+
+\section{Supplemental Information table of
+contents}\label{supplemental-information-table-of-contents}
 
@@
 \item
-  \protect\hyperlink{ga}{Stability given targeted manipulation of
-  \(\gamma\) (genetic algorithm)}
-\item
-  \protect\hyperlink{ecological}{Stability of ecological networks}
+  \protect\hyperlink{ecological}{Stability of random ecological
+  networks}
 
@@
 \item
+  \protect\hyperlink{structured}{Stability of structured networks}
+\item
   \protect\hyperlink{Feasibility}{Feasibility of complex systems}
 \item
+  \protect\hyperlink{ga}{Stability given targeted manipulation of
+  \(\gamma\) (genetic algorithm)}
+\item
   \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
+\item
+  \protect\hyperlink{repr}{Reproducing simulation results}
+\item
+  \protect\hyperlink{ref}{Literature cited}
 \end{itemize}
+
+\clearpage
 
@@
 
-Figure 3 of the main text reports the number of stable random complex
+Figure 4 of the main text reports the number of stable random complex
 systems found over 1 million iterations. The table below shows the
 results for all simulations of random \(\mathbf{M}\) matrices at
-\(\sigma = 0.4\) and \(C = 1\) given a range of
-\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A0} refers to
-matrices where \(\gamma = 1\), while \texttt{A1} refers to matrices
-after \(Var(\gamma)\) is added and \(\gamma \sim \mathcal{U}(0, 2)\).
-Each row summarises data for a given \(S\) over 1 million randomly
-simulated \(\mathbf{M}\) (\texttt{A0} and \texttt{A1}). The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 1
-million). Similarly, the column \texttt{A1\_unstable} shows the number
-of \texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows
-the number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+\(\sigma_{A} = 0.4\) and \(C = 1\) given a range of
+\(S = \{2, 3, ..., 49, 50\}\). In this table, the \texttt{A} refers to
+\(\mathbf{A}\) matrices where \(\gamma = 1\), while \texttt{M} refers to
+\(\mathbf{M}\) matrices after \(\sigma^{2}_{\gamma}\) is added and
+\(\gamma \sim \mathcal{U}(0, 2)\). Each row summarises data for a given
+\(S\) over 1 million randomly simulated \(\mathbf{M}\). The column
+\texttt{A\_unstable} shows the number of \(\mathbf{A}\) matrices that
+are unstable, and the column \texttt{A\_stable} shows the number of
+\(\mathbf{A}\) matrices that are stable (these two columns sum to 1
+million). Similarly, the column \texttt{M\_unstable} shows the number of
+\(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable} shows
+the number that are stable. The columns \texttt{A\_stabilised} and
+\texttt{A\_destabilised} show how many \(\mathbf{M}\) matrices were
+stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 
-Overall, the ratio of stable \texttt{A1} matrices to stable \texttt{A0}
-matrices found is greater than 1 whenever \(S > 10\) (compare column 5
-to column 3), and this ratio increases with increasing \(S\) (column 1).
-Hence, more randomly created complex systems (\(\mathbf{M}\)) are stable
-given variation in \(\gamma\) than when \(\gamma = 1\). Note that
-feasibility results were omitted for the table above, but are
-\protect\hyperlink{Feasibility}{reported below}.
-
-\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
-manipulation of \(\gamma\) (genetic
-algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
-
-Figure 4 of the main text reports the number of stable random complex
-systems found over 100000 using the genetic algorithm to maximise
-stability with a vector \(\mathbf{\gamma}\). Stability results for
-100000 \(\mathbf{M}\) for each \(S\) from 2-40 are shown below. Results
-for \texttt{A0} indicate systems in which \(\gamma = 1\), while
-\texttt{A1} refers to systems in which the genetic algorithm searched
-for a set of \(\gamma\) values that stabilised the system.
-
-\begin{longtable}[]{@{}rrrrrrr@{}}
-\toprule
-S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
-A1\_stabilised & A1\_destabilised\tabularnewline
-\midrule
-\endhead
-2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
-3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
-4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
-5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
-6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
-7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
-8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
-9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
-10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
-11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
-12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
-13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
-14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
-15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
-16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
-17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
-18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
-19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
-20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
-21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
-22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
-23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
-24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
-25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
-26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
-27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
-28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
-29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
-30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
-31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
-32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
-33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
-34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
-35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
-36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
-37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
-38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
-39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
-40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
-\bottomrule
-\end{longtable}
-
-The distributions of nine \(\gamma\) vectors from the highest \(S\)
-values are shown below. This comparison shows the high number of stable
-\(\mathbf{M}\) that can be produced through a targeted search of
-\(\gamma\) values, and suggests that many otherwise unstable systems
-could potentially be stabilised by an informed manipulation of their
-component response times. Such a possibility might conceivably reduce
-the dimensionality of problems involving stability in social-ecological
-or economic systems.
-
-Distributions of \(\gamma\) values in vectors for the highest values of
-\(S\) are shown below.
-
-\includegraphics{unnamed-chunk-7-1.pdf}
-
-The distribution of \(\gamma\) values found by the genetic algorithm is
-uniform. A uniform distribution was used to initialise \(\gamma\)
-values, so there is therefore no evidence that a particular distribution
-of \(\gamma\) is likely to be found to stabilise a matrix
-\(\mathbf{M}\).
-
-\hypertarget{ecological}{\section{Stability of ecological
+Overall, the ratio of stable \(\mathbf{A}\) matrices to stable
+\(\mathbf{M}\) matrices found is greater than 1 whenever \(S > 10\)
+(compare column 3 to column 5), and this ratio increases with increasing
+\(S\) (column 1). Hence, more randomly created complex systems
+(\(\mathbf{M}\)) are stable given variation in \(\gamma\) than when
+\(\gamma = 1\). Note that feasibility results were omitted for the table
+above, but are \protect\hyperlink{Feasibility}{reported below}.
+
+\hypertarget{ecological}{\section{Stability of random ecological
 networks}\label{ecological}}
@@
 \protect\hyperlink{IncrS}{results above}, in the tables below,
-\texttt{A0} refers to matrices when \(\gamma = 1\) and \texttt{A1}
-refers to matrices after \(Var(\gamma)\) is added. The column
-\texttt{A0\_unstable} shows the number of \texttt{A0} matrices that are
-unstable, and the column \texttt{A0\_stable} shows the number of
-\texttt{A0} matrices that are stable (these two columns sum to 100000).
-Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+\texttt{A} refers to matrices \(\mathbf{A}\) when \(\gamma = 1\) and
+\texttt{M} refers to matrices after \(\sigma^{2}_{\gamma}\) is added.
+The column \texttt{A\_unstable} shows the number of \(\mathbf{A}\)
+matrices that are unstable, and the column \texttt{A\_stable} shows the
+number of \(\mathbf{A}\) matrices that are stable (these two columns sum
+to 100000). Similarly, the column \texttt{M\_unstable} shows the number
+of \(\mathbf{M}\) matrices that are unstable and \texttt{M\_stable}
+shows the number that are stable. The columns \texttt{A\_stabilised} and
+\texttt{A\_destabilised} show how many \(\mathbf{A}\) matrices were
+stabilised or destabilised, respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 \(\mathbf{M}\) were potentially nonzero and sampled from a normal
-distribution \(\mathcal{N}(0, \sigma^{2})\) where \(\sigma = 0.4\). Here
-I present four tables showing the number of stable communities given
-\(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases, uniform variation in
-component response rate (\(\gamma \sim \mathcal{U}(0, 2)\)) led to a
-higher number of stable communities than when \(\gamma\) did not vary
-(\(\gamma = 1\)). In contrast to the main text, 100000 rather than 1
-million \(\mathbf{M}\) were simulated. As with the results on
+distribution \(\mathcal{N}(0, \sigma^{2}_{A})\) where
+\(\sigma_{A} = 0.4\). Here I present four tables showing the number of
+stable communities given \(C = \{0.3, 0. 5, 0.7, 0.9 \}\). In all cases,
+uniform variation in component response rate
+(\(\gamma \sim \mathcal{U}(0, 2)\)) led to a higher number of stable
+communities than when \(\gamma\) did not vary (\(\gamma = 1\)). In
+contrast to the main text, 100000 rather than 1 million \(\mathbf{M}\)
+were simulated. As with the results on
 \protect\hyperlink{IncrS}{stability with increasing \(S\)} shown above,
-in the tables below \texttt{A0} refers to matrices when \(\gamma = 1\),
-and \texttt{A1} refers to matrices after \(Var(\gamma)\) is added. The
-column \texttt{A0\_unstable} shows the number of \texttt{A0} matrices
-that are unstable, and the column \texttt{A0\_stable} shows the number
-of \texttt{A0} matrices that are stable (these two columns sum to
-100000). Similarly, the column \texttt{A1\_unstable} shows the number of
-\texttt{A1} matrices that are unstable and \texttt{A1\_stable} shows the
-number that are stable. The columns \texttt{A1\_stabilised} and
-\texttt{A1\_destabilised} show how many \texttt{A0} matrices were
-stabilised or destabilised, respectively, by \(Var(\gamma)\).
+in the tables below \texttt{A} refers to \(\mathbf{A}\) matrices when
+\(\gamma = 1\), and \texttt{M} refers to \(\mathbf{M}\) matrices after
+\(\sigma^{2}_{\gamma}\) is added. The column \texttt{A\_unstable} shows
+the number of \(\mathbf{A}\) matrices that are unstable, and the column
+\texttt{A\_stable} shows the number of \(\mathbf{A}\) matrices that are
+stable (these two columns sum to 100000). Similarly, the column
+\texttt{M\_unstable} shows the number of \(\mathbf{M}\) matrices that
+are unstable and \texttt{M\_stable} shows the number that are stable.
+The columns \texttt{A\_stabilised} and \texttt{A\_destabilised} show how
+many \(\mathbf{A}\) matrices were stabilised or destabilised,
+respectively, by \(\sigma^{2}_{\gamma}\).
 
@@
 \hypertarget{sigma}{\section{\texorpdfstring{Sensitivity of interaction
-strength (\(\sigma\))
-values}{Sensitivity of interaction strength (\textbackslash{}sigma) values}}\label{sigma}}
+strength (\(\sigma_{A}\))
+values}{Sensitivity of interaction strength (\textbackslash{}sigma\_\{A\}) values}}\label{sigma}}
 
 Results below show stability results given varying interaction strengths
-(\(\sigma\)) for \(C = 0.05\) (note that system size \(S\) values are
-larger and increase by 10 with increasing rows). In the tables below (as
-\protect\hyperlink{IncrS}{above}), \texttt{A0} and \texttt{A1} refers to
-matrices for \(\gamma = 1\) and \(Var(\gamma)\), respectively.
-
-\textbf{Interaction strength \(\mathbf{\sigma = 0.3}\)}
+(\(\sigma_{A}\)) for \(C = 0.05\) (note that system size \(S\) values
+are larger and increase by 10 with increasing rows). In the tables below
+(as \protect\hyperlink{IncrS}{above}), \texttt{A} and \texttt{M} refers
+to matrices for \(\gamma = 1\) and \(\sigma^{2}_{\gamma}\),
+respectively.
+
+\textbf{Interaction strength \(\mathbf{\sigma_{A} = 0.3}\)}
 
@@
 
-\includegraphics{unnamed-chunk-23-1.pdf}
+\begin{center}\includegraphics{unnamed-chunk-21-1} \end{center}
 
@@
 more are found with the beta (c) and gamma (d) distributions.
+
+\hypertarget{structured}{\section{Stability of structured
+networks}\label{structured}}
+
+I tested the stability of one million random, small-world, scale-free,
+and cascade food web networks for different network parameters. Each of
+these networks is structured differently. In the main text, the random
+networks and cascade food webs that I built were saturated (\(C = 1\)),
+meaning that every component was connected to, and interacted with,
+every other component (see immediately below).
+
+\vspace{2mm}
+
+\begin{center}\includegraphics{unnamed-chunk-23-1} \end{center}
+
+\vspace{2mm}
+
+Small-world networks, in contrast, are not saturated. They are instead
+defined by components that interact mostly with other closely
+neighbouring components, but have a proportion of interactions
+(\(\beta\)) that are instead between
+non-neighbours\textsuperscript{\protect\hyperlink{ref-Watts1998}{4}}.
+Two small-world networks are shown below.
+
+\vspace{2mm}
+
+\begin{center}\includegraphics{unnamed-chunk-24-1} \end{center}
+
+\vspace{2mm}
+
+The small-world network on the left shows a system in which
+\(\beta = 0.01\), while the small-world network on the right shows one
+in which \(\beta = 0.1\). At the extremes of \(\beta = 0\) and
+\(\beta = 1\), networks are regular and random, respectively. The table
+below shows how \(\sigma^{2}_\gamma\) affects stability in small world
+networks across different values of \(S\) and \(\beta\).
+
+\begin{longtable}[]{@{}rrrrrrrrr@{}}
+\toprule
+beta & S & A\_unstable & A\_stable & M\_unstable & M\_stable &
+complex\_A & complex\_M & C\tabularnewline
+\midrule
+\endhead
+0.00 & 24 & 17388 & 982612 & 17446 & 982554 & 0.5748066 & 0.6582632 &
+0.1304348\tabularnewline
+0.00 & 48 & 258024 & 741976 & 260579 & 739421 & 0.8073918 & 0.9294192 &
+0.1063830\tabularnewline
+0.00 & 72 & 715036 & 284964 & 722639 & 277361 & 0.9860840 & 1.1364805 &
+0.0985915\tabularnewline
+0.00 & 96 & 961434 & 38566 & 962788 & 37212 & 1.1369395 & 1.3110263 &
+0.0947368\tabularnewline
+0.00 & 120 & 999008 & 992 & 998857 & 1143 & 1.2700387 & 1.4649832 &
+0.0924370\tabularnewline
+0.00 & 144 & 999997 & 3 & 999994 & 6 & 1.3903192 & 1.6041216 &
+0.0909091\tabularnewline
+0.00 & 168 & 1000000 & 0 & 1000000 & 0 & 1.5010334 & 1.7320676 &
+0.0898204\tabularnewline
+0.01 & 24 & 17673 & 982327 & 17720 & 982280 & 0.5747156 & 0.6581503 &
+0.1304319\tabularnewline
+0.01 & 48 & 255038 & 744962 & 257647 & 742353 & 0.8073388 & 0.9292952 &
+0.1063800\tabularnewline
+0.01 & 72 & 708892 & 291108 & 716829 & 283171 & 0.9859457 & 1.1363940 &
+0.0985884\tabularnewline
+0.01 & 96 & 960635 & 39365 & 961876 & 38124 & 1.1370640 & 1.3112193 &
+0.0947337\tabularnewline
+0.01 & 120 & 999040 & 960 & 998794 & 1206 & 1.2698715 & 1.4648280 &
+0.0924338\tabularnewline
+0.01 & 144 & 999997 & 3 & 999994 & 6 & 1.3901601 & 1.6039285 &
+0.0909060\tabularnewline
+0.01 & 168 & 1000000 & 0 & 1000000 & 0 & 1.5009490 & 1.7319739 &
+0.0898173\tabularnewline
+0.10 & 24 & 20382 & 979618 & 20455 & 979545 & 0.5742520 & 0.6573563 &
+0.1302974\tabularnewline
+0.10 & 48 & 237747 & 762253 & 240370 & 759630 & 0.8066604 & 0.9284434 &
+0.1062311\tabularnewline
+0.10 & 72 & 679874 & 320126 & 685575 & 314425 & 0.9849695 & 1.1352553 &
+0.0984349\tabularnewline
+0.10 & 96 & 961984 & 38016 & 960128 & 39872 & 1.1358912 & 1.3097957 &
+0.0945788\tabularnewline
+0.10 & 120 & 999546 & 454 & 999275 & 725 & 1.2687142 & 1.4634587 &
+0.0922779\tabularnewline
+0.10 & 144 & 1000000 & 0 & 1000000 & 0 & 1.3890356 & 1.6025900 &
+0.0907489\tabularnewline
+0.10 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4994818 & 1.7302649 &
+0.0896598\tabularnewline
+0.25 & 24 & 23654 & 976346 & 23775 & 976225 & 0.5722185 & 0.6546853 &
+0.1296712\tabularnewline
+0.25 & 48 & 228318 & 771682 & 231208 & 768792 & 0.8033257 & 0.9244966 &
+0.1055259\tabularnewline
+0.25 & 72 & 666982 & 333018 & 669104 & 330896 & 0.9808676 & 1.1304109 &
+0.0977066\tabularnewline
+0.25 & 96 & 966456 & 33544 & 961545 & 38455 & 1.1307841 & 1.3039452 &
+0.0938392\tabularnewline
+0.25 & 120 & 999749 & 251 & 999507 & 493 & 1.2632327 & 1.4571506 &
+0.0915316\tabularnewline
+0.25 & 144 & 1000000 & 0 & 1000000 & 0 & 1.3827642 & 1.5953248 &
+0.0899987\tabularnewline
+0.25 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4926700 & 1.7224506 &
+0.0889064\tabularnewline
+1.00 & 24 & 26331 & 973669 & 26478 & 973522 & 0.5561013 & 0.6356655 &
+0.1249651\tabularnewline
+1.00 & 48 & 211199 & 788801 & 214154 & 785846 & 0.7720342 & 0.8881302 &
+0.0991370\tabularnewline
+1.00 & 72 & 613621 & 386379 & 615771 & 384229 & 0.9394912 & 1.0825566 &
+0.0908153\tabularnewline
+1.00 & 96 & 943191 & 56809 & 936396 & 63604 & 1.0812364 & 1.2466510 &
+0.0867047\tabularnewline
+1.00 & 120 & 999157 & 843 & 998396 & 1604 & 1.2065026 & 1.3916458 &
+0.0842561\tabularnewline
+1.00 & 144 & 1000000 & 0 & 999997 & 3 & 1.3199179 & 1.5227509 &
+0.0826325\tabularnewline
+1.00 & 168 & 1000000 & 0 & 1000000 & 0 & 1.4243560 & 1.6434386 &
+0.0814738\tabularnewline
+\bottomrule
+\end{longtable}
+
+In the above, the complexity of \(\mathbf{A}\) and \(\mathbf{M}\), and
+the mean \(C\), are also shown. For similar magnitudes of complexity as
+in random networks of \(\sigma\sqrt{SC} \gtrapprox 1.26\), variation in
+\(\gamma\) typically results in more stable than unstable systems.
+
+Scale-free networks are also not saturated, but are defined by an
+interaction frequency distribution that follows a power law. In other
+words, a small number of components interact with many other components,
+while most components interact with only a small number of other
+components. Scale-free networks can be built by adding new components,
+one by one, to an existing system, with each newly added component
+interacting with a randomly selected subset of \(m\) existing
+components\textsuperscript{\protect\hyperlink{ref-Albert2002}{5}}. The
+network on the left below shows an example of a scale-free network in
+which \(m = 3\). The histogram on the right shows the number of other
+components with which each component interacts.
+
+\begin{center}\includegraphics{unnamed-chunk-26-1} \end{center}
+
+The table below shows how \(\sigma^{2}_\gamma\) affects stability across
+different scale-free networks with different \(S\) and \(m\) values.
+
+\begin{longtable}[]{@{}rrrrrrrrr@{}}
+\toprule
+m & S & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
+complex\_M & C\tabularnewline
+\midrule
+\endhead
+2 & 24 & 152791 & 847209 & 156034 & 843966 & 0.7891257 & 0.9034663 &
+0.1648551\tabularnewline
+3 & 24 & 320481 & 679519 & 326351 & 673649 & 0.9566487 & 1.0967499 &
+0.2409420\tabularnewline
+4 & 24 & 504433 & 495567 & 504826 & 495174 & 1.0922870 & 1.2532761 &
+0.3134058\tabularnewline
+5 & 24 & 670676 & 329324 & 660426 & 339574 & 1.2073054 & 1.3857169 &
+0.3822464\tabularnewline
+6 & 24 & 798637 & 201363 & 779345 & 220655 & 1.3067095 & 1.5004508 &
+0.4474638\tabularnewline
+7 & 24 & 884082 & 115918 & 862215 & 137785 & 1.3942577 & 1.6013368 &
+0.5090580\tabularnewline
+8 & 24 & 936190 & 63810 & 915630 & 84370 & 1.4722315 & 1.6908563 &
+0.5670290\tabularnewline
+9 & 24 & 964868 & 35132 & 948297 & 51703 & 1.5414455 & 1.7707292 &
+0.6213768\tabularnewline
+10 & 24 & 981460 & 18540 & 967911 & 32089 & 1.6030044 & 1.8417459 &
+0.6721014\tabularnewline
+11 & 24 & 989838 & 10162 & 980232 & 19768 & 1.6586511 & 1.9059313 &
+0.7192029\tabularnewline
+12 & 24 & 994393 & 5607 & 987436 & 12564 & 1.7081503 & 1.9628898 &
+0.7626812\tabularnewline
+2 & 48 & 303963 & 696037 & 310053 & 689947 & 0.7946875 & 0.9132519 &
+0.0828901\tabularnewline
+3 & 48 & 577855 & 422145 & 579996 & 420004 & 0.9685494 & 1.1141445 &
+0.1227837\tabularnewline
+4 & 48 & 810001 & 189999 & 799132 & 200868 & 1.1122992 & 1.2799335 &
+0.1617908\tabularnewline
+5 & 48 & 938004 & 61996 & 924613 & 75387 & 1.2369960 & 1.4236817 &
+0.1999113\tabularnewline
+6 & 48 & 984975 & 15025 & 976433 & 23567 & 1.3478291 & 1.5514420 &
+0.2371454\tabularnewline
+7 & 48 & 997160 & 2840 & 994005 & 5995 & 1.4473792 & 1.6663763 &
+0.2734929\tabularnewline
+8 & 48 & 999584 & 416 & 998590 & 1410 & 1.5385445 & 1.7716359 &
+0.3089539\tabularnewline
+9 & 48 & 999955 & 45 & 999707 & 293 & 1.6227742 & 1.8687074 &
+0.3435284\tabularnewline
+10 & 48 & 999992 & 8 & 999939 & 61 & 1.7006157 & 1.9583879 &
+0.3772163\tabularnewline
+11 & 48 & 999999 & 1 & 999990 & 10 & 1.7731759 & 2.0420990 &
+0.4100177\tabularnewline
+12 & 48 & 1000000 & 0 & 999999 & 1 & 1.8410402 & 2.1203112 &
+0.4419326\tabularnewline
+2 & 72 & 427243 & 572757 & 434600 & 565400 & 0.7964226 & 0.9166566 &
+0.0553599\tabularnewline
+3 & 72 & 741345 & 258655 & 739020 & 260980 & 0.9723446 & 1.1195788 &
+0.0823552\tabularnewline
+4 & 72 & 931043 & 68957 & 921145 & 78855 & 1.1188220 & 1.2888100 &
+0.1089593\tabularnewline
+5 & 72 & 989644 & 10356 & 984372 & 15628 & 1.2466268 & 1.4361875 &
+0.1351721\tabularnewline
+6 & 72 & 999131 & 869 & 997914 & 2086 & 1.3604666 & 1.5674966 &
+0.1609937\tabularnewline
+7 & 72 & 999946 & 54 & 999804 & 196 & 1.4642496 & 1.6872501 &
+0.1864241\tabularnewline
+8 & 72 & 999999 & 1 & 999988 & 12 & 1.5596340 & 1.7974044 &
+0.2114632\tabularnewline
+9 & 72 & 1000000 & 0 & 999999 & 1 & 1.6482181 & 1.8994441 &
+0.2361111\tabularnewline
+10 & 72 & 1000000 & 0 & 1000000 & 0 & 1.7307859 & 1.9947150 &
+0.2603678\tabularnewline
+11 & 72 & 1000000 & 0 & 1000000 & 0 & 1.8086766 & 2.0847262 &
+0.2842332\tabularnewline
+12 & 72 & 1000000 & 0 & 1000000 & 0 & 1.8817533 & 2.1689764 &
+0.3077074\tabularnewline
+2 & 96 & 527633 & 472367 & 535188 & 464812 & 0.7974024 & 0.9183557 &
+0.0415570\tabularnewline
+3 & 96 & 842274 & 157726 & 837756 & 162244 & 0.9741293 & 1.1224709 &
+0.0619518\tabularnewline
+4 & 96 & 975834 & 24166 & 969478 & 30522 & 1.1220115 & 1.2931371 &
+0.0821272\tabularnewline
+5 & 96 & 998391 & 1609 & 996991 & 3009 & 1.2511287 & 1.4422331 &
+0.1020833\tabularnewline
+6 & 96 & 999955 & 45 & 999838 & 162 & 1.3669903 & 1.5757699 &
+0.1218202\tabularnewline
+7 & 96 & 999999 & 1 & 999996 & 4 & 1.4725862 & 1.6977057 &
+0.1413377\tabularnewline
+8 & 96 & 1000000 & 0 & 1000000 & 0 & 1.5699145 & 1.8099762 &
+0.1606360\tabularnewline
+9 & 96 & 1000000 & 0 & 1000000 & 0 & 1.6606162 & 1.9146804 &
+0.1797149\tabularnewline
+10 & 96 & 1000000 & 0 & 1000000 & 0 & 1.7457971 & 2.0129344 &
+0.1985746\tabularnewline
+11 & 96 & 1000000 & 0 & 1000000 & 0 & 1.8260368 & 2.1055559 &
+0.2172149\tabularnewline
+12 & 96 & 1000000 & 0 & 1000000 & 0 & 1.9018608 & 2.1929362 &
+0.2356360\tabularnewline
+2 & 120 & 609563 & 390437 & 616036 & 383964 & 0.7979355 & 0.9194404 &
+0.0332633\tabularnewline
+3 & 120 & 904064 & 95936 & 899040 & 100960 & 0.9753815 & 1.1243251 &
+0.0496499\tabularnewline
+4 & 120 & 991710 & 8290 & 988410 & 11590 & 1.1239922 & 1.2957520 &
+0.0658964\tabularnewline
+5 & 120 & 999781 & 219 & 999477 & 523 & 1.2539362 & 1.4458518 &
+0.0820028\tabularnewline
+6 & 120 & 999999 & 1 & 999981 & 19 & 1.3707937 & 1.5806987 &
+0.0979692\tabularnewline
+7 & 120 & 1000000 & 0 & 999999 & 1 & 1.4775366 & 1.7038860 &
+0.1137955\tabularnewline
+8 & 120 & 1000000 & 0 & 1000000 & 0 & 1.5762636 & 1.8177236 &
+0.1294818\tabularnewline
+9 & 120 & 1000000 & 0 & 1000000 & 0 & 1.6680647 & 1.9238257 &
+0.1450280\tabularnewline
+10 & 120 & 1000000 & 0 & 1000000 & 0 & 1.7545110 & 2.0233838 &
+0.1604342\tabularnewline
+11 & 120 & 1000000 & 0 & 1000000 & 0 & 1.8363882 & 2.1178385 &
+0.1757003\tabularnewline
+12 & 120 & 1000000 & 0 & 1000000 & 0 & 1.9135798 & 2.2069806 &
+0.1908263\tabularnewline
+\bottomrule
+\end{longtable}
+
+As in small-world networks, the mean \(C\) is shown, along with the mean
+complexities of \(\mathbf{A}\) and \(\mathbf{M}\). Like all other
+networks, \(\sigma^{2}_\gamma\) increases the stability of scale-free
+networks given sufficiently high complexity.
+
+Cascade food webs are saturated, and similar to predator-prey random
+networks. What distinguishes them from predator-prey networks is that
+cascade food webs are also defined by intactions in which components are
+ranked such that if the rank of \(i > j\), then \(A_{ij} < 0\) and
+\(A_{ji} > 0\)\textsuperscript{\protect\hyperlink{ref-Solow1998}{6},\protect\hyperlink{ref-Williams2000}{7}}.
+In other words, if interpreting components as ecological species,
+species can only feed off of a species of lower rank. The table below
+shows how \(\sigma^{2}_\gamma\) affects stability across system sizes in
+cascade food webs.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & complex\_A &
+complex\_M\tabularnewline
+\midrule
+\endhead
+2 & 0 & 1000000 & 0 & 1000000 & 0.6378839 & 0.6381485\tabularnewline
+3 & 1 & 999999 & 1 & 999999 & 0.7055449 & 0.7525143\tabularnewline
+4 & 2 & 999998 & 2 & 999998 & 0.8060500 & 0.8826100\tabularnewline
+5 & 17 & 999983 & 17 & 999983 & 0.8974749 & 0.9967594\tabularnewline
+6 & 42 & 999958 & 43 & 999957 & 0.9821323 & 1.0999762\tabularnewline
+7 & 124 & 999876 & 124 & 999876 & 1.0600906 & 1.1938910\tabularnewline
+8 & 303 & 999697 & 309 & 999691 & 1.1329713 & 1.2807302\tabularnewline
+9 & 653 & 999347 & 661 & 999339 & 1.2009135 & 1.3616372\tabularnewline
+10 & 1401 & 998599 & 1413 & 998587 & 1.2661142 &
+1.4387567\tabularnewline
+11 & 2534 & 997466 & 2566 & 997434 & 1.3276636 &
+1.5113096\tabularnewline
+12 & 4514 & 995486 & 4597 & 995403 & 1.3865754 &
+1.5804005\tabularnewline
+13 & 7570 & 992430 & 7722 & 992278 & 1.4424479 &
+1.6462780\tabularnewline
+14 & 12223 & 987777 & 12502 & 987498 & 1.4970134 &
+1.7102322\tabularnewline
+15 & 18433 & 981567 & 18879 & 981121 & 1.5498812 &
+1.7719564\tabularnewline
+16 & 26973 & 973027 & 27712 & 972288 & 1.6002970 &
+1.8310447\tabularnewline
+17 & 38272 & 961728 & 39499 & 960501 & 1.6494195 &
+1.8884211\tabularnewline
+18 & 52397 & 947603 & 54099 & 945901 & 1.6975099 &
+1.9443860\tabularnewline
+19 & 69986 & 930014 & 72342 & 927658 & 1.7439233 &
+1.9987398\tabularnewline
+20 & 92851 & 907149 & 95776 & 904224 & 1.7893524 &
+2.0514394\tabularnewline
+21 & 117487 & 882513 & 121095 & 878905 & 1.8335974 &
+2.1030121\tabularnewline
+22 & 147852 & 852148 & 151989 & 848011 & 1.8761874 &
+2.1527108\tabularnewline
+23 & 183501 & 816499 & 187888 & 812112 & 1.9186092 &
+2.2019827\tabularnewline
+24 & 222592 & 777408 & 226021 & 773979 & 1.9591518 &
+2.2491948\tabularnewline
+25 & 267691 & 732309 & 269822 & 730178 & 1.9999089 &
+2.2963949\tabularnewline
+26 & 316090 & 683910 & 316371 & 683629 & 2.0396325 &
+2.3427211\tabularnewline
+27 & 369830 & 630170 & 366550 & 633450 & 2.0785319 &
+2.3879356\tabularnewline
+28 & 426407 & 573593 & 419136 & 580864 & 2.1169703 &
+2.4324407\tabularnewline
+29 & 485068 & 514932 & 473666 & 526334 & 2.1545265 &
+2.4759539\tabularnewline
+30 & 544300 & 455700 & 527568 & 472432 & 2.1912376 &
+2.5187795\tabularnewline
+31 & 605803 & 394197 & 584385 & 415615 & 2.2271037 &
+2.5603818\tabularnewline
+32 & 664689 & 335311 & 638047 & 361953 & 2.2626270 &
+2.6016360\tabularnewline
+33 & 718848 & 281152 & 689172 & 310828 & 2.2979241 &
+2.6424881\tabularnewline
+34 & 770790 & 229210 & 737639 & 262361 & 2.3327303 &
+2.6828460\tabularnewline
+35 & 817531 & 182469 & 783112 & 216888 & 2.3666720 &
+2.7221952\tabularnewline
+36 & 858750 & 141250 & 823548 & 176452 & 2.3998286 &
+2.7608037\tabularnewline
+37 & 893017 & 106983 & 859194 & 140806 & 2.4332806 &
+2.7994470\tabularnewline
+38 & 921268 & 78732 & 890177 & 109823 & 2.4658414 &
+2.8372307\tabularnewline
+39 & 943551 & 56449 & 915655 & 84345 & 2.4974678 &
+2.8741350\tabularnewline
+40 & 961088 & 38912 & 936883 & 63117 & 2.5301278 &
+2.9116114\tabularnewline
+41 & 973664 & 26336 & 953645 & 46355 & 2.5616210 &
+2.9481298\tabularnewline
+42 & 982829 & 17171 & 967044 & 32956 & 2.5925309 &
+2.9841081\tabularnewline
+43 & 989464 & 10536 & 977033 & 22967 & 2.6228949 &
+3.0191690\tabularnewline
+44 & 993622 & 6378 & 984470 & 15530 & 2.6534626 &
+3.0548439\tabularnewline
+45 & 996221 & 3779 & 989678 & 10322 & 2.6832092 &
+3.0890543\tabularnewline
+46 & 997963 & 2037 & 993318 & 6682 & 2.7130588 &
+3.1236201\tabularnewline
+47 & 998818 & 1182 & 995957 & 4043 & 2.7423480 &
+3.1575904\tabularnewline
+48 & 999422 & 578 & 997446 & 2554 & 2.7714223 & 3.1912463\tabularnewline
+49 & 999746 & 254 & 998532 & 1468 & 2.7999596 & 3.2244020\tabularnewline
+50 & 999864 & 136 & 999132 & 868 & 2.8285547 & 3.2574510\tabularnewline
+51 & 999934 & 66 & 999561 & 439 & 2.8566907 & 3.2900943\tabularnewline
+52 & 999970 & 30 & 999761 & 239 & 2.8844703 & 3.3222721\tabularnewline
+53 & 999985 & 15 & 999873 & 127 & 2.9122645 & 3.3544290\tabularnewline
+54 & 999999 & 1 & 999935 & 65 & 2.9395400 & 3.3859103\tabularnewline
+55 & 1000000 & 0 & 999971 & 29 & 2.9665996 & 3.4173273\tabularnewline
+56 & 999999 & 1 & 999988 & 12 & 2.9936263 & 3.4486027\tabularnewline
+57 & 1000000 & 0 & 999989 & 11 & 3.0199283 & 3.4789408\tabularnewline
+58 & 1000000 & 0 & 999998 & 2 & 3.0460952 & 3.5094530\tabularnewline
+59 & 1000000 & 0 & 999999 & 1 & 3.0728115 & 3.5401634\tabularnewline
+60 & 1000000 & 0 & 1000000 & 0 & 3.0983367 & 3.5698067\tabularnewline
+\bottomrule
+\end{longtable}
+
+Cascade food webs are more likely to be stable than small-world or
+scale-free networks at equivalent magnitudes of complexity (note 
+\(C = 1\) for all above rows). A higher number of stable \(\mathbf{M}\)
+than \(\mathbf{A}\) was found given \(S \geq 27\).
 
@@
 
-Hence, in general, \(Var(\gamma)\) does not appear to affect feasibility
-in pure species interaction
-networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{4}}.
+Hence, in general, \(\sigma^{2}_{\gamma}\) does not appear to affect
+feasibility in pure species interaction
+networks\textsuperscript{\protect\hyperlink{ref-Servan2018}{8}}.
+
+\hypertarget{ga}{\section{\texorpdfstring{Stability given targeted
+manipulation of \(\gamma\) (genetic
+algorithm)}{Stability given targeted manipulation of \textbackslash{}gamma (genetic algorithm)}}\label{ga}}
+
+The figure below compares the stability of large complex systems given
+\(\gamma = 1\) versus targeted manipulation of \(\gamma\) elements. For
+each \(S\), 100000 complex systems are randomly generated. Stability of
+each complex system is tested given variation in \(\gamma\) using a
+genetic algorithm to maximise the effect of \(\gamma\) values on
+increasing stability, as compared to stability in an otherwise identical
+system in which \(\gamma\) is the same for all components. Blue bars
+show the number of stable systems in the absence of component response
+rate variation, while red bars show the number of stable systems that
+can be generated if component response rate is varied to maximise system
+stability. The black line shows the proportion of systems that are
+stable when component response rate is targeted to increase stability,
+but would not be stable if \(\sigma^{2}_{\gamma} = 0\). The y-axis shows
+the \(\ln\) number of systems that are stable across
+\(S = \{1, 2, ..., 39, 40\}\) for \(C = 1\), and the proportion of
+systems wherein a targeted search of \(\gamma\) values successfully
+resulted in system stability.
+
+\includegraphics{unnamed-chunk-30-1.pdf}
+
+Stability results are also shown in the table below. Results for
+\texttt{A} indicate systems in which \(\gamma = 1\), while \texttt{M}
+refers to systems in which the genetic algorithm searched for a set of
+\(\gamma\) values that stabilised the system.
+
+\begin{longtable}[]{@{}rrrrrrr@{}}
+\toprule
+S & A\_unstable & A\_stable & M\_unstable & M\_stable & A\_stabilised &
+A\_destabilised\tabularnewline
+\midrule
+\endhead
+2 & 26 & 99974 & 26 & 99974 & 0 & 0\tabularnewline
+3 & 358 & 99642 & 358 & 99642 & 0 & 0\tabularnewline
+4 & 1505 & 98495 & 1505 & 98495 & 0 & 0\tabularnewline
+5 & 3995 & 96005 & 3982 & 96018 & 13 & 0\tabularnewline
+6 & 8060 & 91940 & 7956 & 92044 & 104 & 0\tabularnewline
+7 & 13420 & 86580 & 12953 & 87047 & 468 & 1\tabularnewline
+8 & 20518 & 79482 & 18940 & 81060 & 1578 & 0\tabularnewline
+9 & 28939 & 71061 & 25148 & 74852 & 3793 & 2\tabularnewline
+10 & 38241 & 61759 & 30915 & 69085 & 7327 & 1\tabularnewline
+11 & 48682 & 51318 & 36398 & 63602 & 12286 & 2\tabularnewline
+12 & 58752 & 41248 & 40710 & 59290 & 18043 & 1\tabularnewline
+13 & 68888 & 31112 & 44600 & 55400 & 24289 & 1\tabularnewline
+14 & 77651 & 22349 & 47528 & 52472 & 30124 & 1\tabularnewline
+15 & 84912 & 15088 & 49971 & 50029 & 34942 & 1\tabularnewline
+16 & 90451 & 9549 & 52274 & 47726 & 38178 & 1\tabularnewline
+17 & 94332 & 5668 & 54124 & 45876 & 40209 & 1\tabularnewline
+18 & 96968 & 3032 & 55831 & 44169 & 41139 & 2\tabularnewline
+19 & 98384 & 1616 & 58079 & 41921 & 40305 & 0\tabularnewline
+20 & 99269 & 731 & 60181 & 39819 & 39088 & 0\tabularnewline
+21 & 99677 & 323 & 63338 & 36662 & 36339 & 0\tabularnewline
+22 & 99854 & 146 & 66350 & 33650 & 33504 & 0\tabularnewline
+23 & 99947 & 53 & 70478 & 29522 & 29469 & 0\tabularnewline
+24 & 99983 & 17 & 74121 & 25879 & 25862 & 0\tabularnewline
+25 & 99991 & 9 & 78364 & 21636 & 21627 & 0\tabularnewline
+26 & 99999 & 1 & 82635 & 17365 & 17364 & 0\tabularnewline
+27 & 100000 & 0 & 86433 & 13567 & 13567 & 0\tabularnewline
+28 & 100000 & 0 & 89951 & 10049 & 10049 & 0\tabularnewline
+29 & 100000 & 0 & 92716 & 7284 & 7284 & 0\tabularnewline
+30 & 100000 & 0 & 95171 & 4829 & 4829 & 0\tabularnewline
+31 & 100000 & 0 & 96844 & 3156 & 3156 & 0\tabularnewline
+32 & 100000 & 0 & 98128 & 1872 & 1872 & 0\tabularnewline
+33 & 100000 & 0 & 98941 & 1059 & 1059 & 0\tabularnewline
+34 & 100000 & 0 & 99358 & 642 & 642 & 0\tabularnewline
+35 & 100000 & 0 & 99702 & 298 & 298 & 0\tabularnewline
+36 & 100000 & 0 & 99856 & 144 & 144 & 0\tabularnewline
+37 & 100000 & 0 & 99921 & 79 & 79 & 0\tabularnewline
+38 & 100000 & 0 & 99970 & 30 & 30 & 0\tabularnewline
+39 & 100000 & 0 & 99989 & 11 & 11 & 0\tabularnewline
+40 & 100000 & 0 & 99994 & 6 & 6 & 0\tabularnewline
+\bottomrule
+\end{longtable}
+
+The distributions of nine \(\gamma\) vectors from the highest \(S\)
+values are shown below. This comparison shows the high number of stable
+\(\mathbf{M}\) that can be produced through a targeted search of
+\(\gamma\) values, and suggests that many otherwise unstable systems
+could potentially be stabilised by an informed manipulation of their
+component response times. Such a possibility might conceivably reduce
+the dimensionality of problems involving stability in social-ecological
+or economic systems.
+
+\includegraphics{unnamed-chunk-32-1.pdf}
+
+The distribution of \(\gamma\) values found by the genetic algorithm is
+uniform. A uniform distribution was used to initialise \(\gamma\)
+values, so there is therefore no evidence that a particular distribution
+of \(\gamma\) is likely to be found to stabilise a matrix
+\(\mathbf{M}\).
 
@@
 The question that I address in the main text is distinct from that of
-Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{5}}, which
+Gibbs et al.\textsuperscript{\protect\hyperlink{ref-Gibbs2017}{9}}, who
 focused instead on the effect of a diagonal matrix of biological species
@@
 
-\begin{figure}[h]
-\includegraphics{unnamed-chunk-27-1}
-\end{figure}
+\begin{center}
+\includegraphics{unnamed-chunk-34-1} 
+\end{center}
 
@@
 
-\includegraphics{unnamed-chunk-29-1.pdf}
+\includegraphics{unnamed-chunk-36-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-30-1.pdf}
+\includegraphics{unnamed-chunk-37-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-31-1.pdf}
+\includegraphics{unnamed-chunk-38-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-32-1.pdf}
+\includegraphics{unnamed-chunk-39-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-34-1.pdf}
+\includegraphics{unnamed-chunk-41-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-35-1.pdf}
+\includegraphics{unnamed-chunk-42-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-36-1.pdf}
+\includegraphics{unnamed-chunk-43-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-37-1.pdf}
-
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
+\includegraphics{unnamed-chunk-44-1.pdf}
+
+\hypertarget{repr}{\section{Reproducing simulation results}\label{repr}}
+
+All results in the main text and the literature cited can be reproduced
+using the
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability}
+R package, which can be downloaded as instructed at the beginning of
+this Supplemental Information document. The most relevant R functions
+for reproducing simulations include the following:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \texttt{rand\_gen\_var}: Simulates random complex systems and cascade
+  food webs
+\item
+  \texttt{rand\_rho\_var}: Simulates random complex systems across a
+  fixed correlation of \(\rho = cor(A_{ij}, A_{ji})\)
+\item
+  \texttt{rand\_gen\_swn}: Simulates randomly generated small-world
+  networks
+\item
+  \texttt{rand\_gen\_sfn}: Simulates randomly generated scale-free
+  networks
+\item
+  \texttt{Evo\_rand\_gen\_var}: Use a genetic algorithm to find stable
+  random complex systems
+\end{enumerate}
+
+For the functions 1-4 above, R output will be a table of results. Below
+describes the headers of this table to more clearly explain what is
+being reported.
+
+\footnotesize
+
+\begin{longtable}[]{@{}llll@{}}
+\toprule
+Header & Description & Header\_cont. & Description\_cont.\tabularnewline
+\midrule
+\endhead
+S & The system size & A\_rho & Corr. between elements A{[}ij{]} and
+A{[}ji{]}\tabularnewline
+A\_unstable & No. of A that were unstable & M\_rho & Corr. between
+elements M{[}ij{]} and M{[}ji{]}\tabularnewline
+A\_stable & No. of A that were stable & rho\_diff & Diff. between A and
+M rho values\tabularnewline
+M\_unstable & No. of M that were unstable & rho\_abs & Diff. between A
+and M rho magnitudes\tabularnewline
+M\_stable & No. of M that were stable & complex\_A & Complexity of
+A\tabularnewline
+A\_stabilised & No. of A stabilised by gamma & complex\_M & Complexity
+of M\tabularnewline
+A\_destabilised & No. of A destabilised by gamma & A\_eig & Expected
+real part of leading A eigenvalue\tabularnewline
+A\_infeasible & No. of A that were infeasible & M\_eig & Expected real
+part of leading M eigenvalue\tabularnewline
+A\_feasible & No. of A that were feasible & LR\_A & Lowest obs. real
+part of leading A eigenvalue\tabularnewline
+M\_infeasible & No. of M that were infeasible & UR\_A & Highest obs.
+real part of leading A eigenvalue\tabularnewline
+M\_feasible & No. of M that were feasible & LR\_M & Lowest obs. real
+part of leading M eigenvalue\tabularnewline
+A\_made\_feasible & No. of A made feasible by gamma & UR\_M & Highest
+obs. real part of leading M eigenvalue\tabularnewline
+A\_made\_infeasible & No. of A made infeasible by gamma & C & Obs.
+network connectance\tabularnewline
+\bottomrule
+\end{longtable}
+
+\normalsize
+
+Note that output from \texttt{Evo\_rand\_gen\_var} only includes the
+first seven rows of the table above, and \texttt{rand\_gen\_var} does
+not include \(C\) (which can be defined as an argument). All results
+presented here and in the main text are available in the
+\href{https://github.com/bradduthie/RandomMatrixStability/tree/master/inst/extdata}{inst/extdata}
+folder of the
+\href{https://github.com/bradduthie/RandomMatrixStability}{RandomMatrixStability}
+R package.
+
+\hypertarget{ref}{\section*{Literature cited}\label{ref}}
+\addcontentsline{toc}{section}{Literature cited}
 

### 2020-03-26 notebook/ms.tex
Recompile ms and supplemental information to create TEX files

@@
 \(\gamma\) were sampled from a uniform distribution where
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\) and \(\sigma^{2}_{\gamma} = 1/3\)
-(see Supplementary Information for other \(\gamma\) distributions, which
-gave similar results). In all simulations, diagonal elements were
+\(\gamma \sim \mathcal{U}(0, 2)\) and \(\sigma^{2}_{\gamma} = 1/3\) (see
+Supplementary Information for other \(\gamma\) distributions, which gave
+similar results). In all simulations, diagonal elements were
 standardised to ensure that \(-d\) between individual \(\textbf{A}\) and
@@
 simulated for each \(S\), and the stability of \(\mathbf{A}\) vesus
-\(\mathbf{M}\) was assessed given \(\gamma_{i} \sim \mathcal{U}(0, 2)\)
+\(\mathbf{M}\) was assessed given \(\gamma \sim \mathcal{U}(0, 2)\)
 (\(\sigma^{2}_{\gamma} = 1/3\)). For all \(S > 10\), I found that the
@@
 from a distribution with positive support (typically
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices
-with correlated elements \(A_{ij}\) and \(A_{ji}\) were built using
-Cholesky decomposition. Competitor networks in which off-diagonal
-elements \(A_{ij} \leq 0\) were constructed by first building a random
+\(\gamma \sim \mathcal{U}(0, 2)\)). Random \(\mathbf{A}\) matrices with
+correlated elements \(A_{ij}\) and \(A_{ji}\) were built using Cholesky
+decomposition. Competitor networks in which off-diagonal elements
+\(A_{ij} \leq 0\) were constructed by first building a random
 \(\mathbf{A}\), then flipping the sign of any elements in which
@@
 therefore evaluated \(\mathbf{-M^{-1}r}\) (\(\mathbf{r}\) elements were
-sampled i.i.d. from \(r_{i} \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility
-is satisfied if all of the elements of the resulting vector are
-positive.
+sampled i.i.d. from \(r \sim \mathcal{N}(0, 0.4^{2})\)). Feasibility is
+satisfied if all of the elements of the resulting vector are positive.
 
@@
 1000 different \(diag(\boldsymbol{\gamma})\) vectors with elements
-sampled i.i.d from \(\gamma_{i} \sim \mathcal{U}(0, 2)\). Eigenanalysis
-was performed on the \(\mathbf{M}\) resulting from each
+sampled i.i.d from \(\gamma \sim \mathcal{U}(0, 2)\). Eigenanalysis was
+performed on the \(\mathbf{M}\) resulting from each
 \(\boldsymbol{\gamma}\), and the 20 \(diag(\boldsymbol{\gamma})\)
@@
 the red circle centred at the origin with a radius of
-\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.19.
+\(\sqrt{\sigma^{2}_{A}(1 + \sigma^{2}_{\gamma})S} \approx\) 1.17.
 (\(\textbf{c}\)) A different random system \(\mathbf{A}\) constructed
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{unnamed-chunk-8-1.pdf}
 
@@
 diagonal elements were set to \(-1\). Elements of \(\gamma\) were
-sampled, \(\gamma_{i} \sim \mathcal{U}(0, 2)\).
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
+sampled, \(\gamma \sim \mathcal{U}(0, 2)\).
+
+\includegraphics{unnamed-chunk-11-1.pdf}
 
@@
 set to \(-1\). Elements of \(\gamma\) were sampled such that
-\(\gamma_{i} \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
-
-\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
+\(\gamma \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
+
+\includegraphics{unnamed-chunk-12-1.pdf}
 
@@
 generated. Stability of each complex system is tested given variation in
-\(\gamma\) by randomly sampling \(\gamma_{i} \sim \mathcal{U}(0, 2)\).
+\(\gamma\) by randomly sampling \(\gamma \sim \mathcal{U}(0, 2)\).
 Stability given \(\sigma^{2}_{\gamma}>0\) is then compared to stability
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{unnamed-chunk-13-1.pdf}
 

### 2020-03-26 notebook/SI.Rmd
Recompile SI files with slight adjustment based on arXiv edits

@@
 
-Cascade food webs are more likely to be stable than small-world or scale-free networks at equivalent magnitudes of complexity (note that $C = 1$ for all above rows).
+Cascade food webs are more likely to be stable than small-world or scale-free networks at equivalent magnitudes of complexity (note $C = 1$ for all above rows).
 A higher number of stable $\mathbf{M}$ than $\mathbf{A}$ was found given $S \geq 27$.
@@
 
-\clearpage
-
 Reproducing simulation results {#repr}

### 2020-03-26 notebook/ms.Rmd
Remove numbers on equations that are not referenced in the text

@@
 
-\begin{equation} \label{var_reduced}
-\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\begin{equation} 
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}). \nonumber
 \end{equation}
@@
 
-\begin{equation} \label{rho_stab}
-\Re(\lambda_{max}) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right).
+\begin{equation} 
+\Re(\lambda_{max}) \approx \sigma_{M}\sqrt{SC}\left(1 + \rho\right). \nonumber
 \end{equation}

### 2020-03-27 notebook/arxiv/ms.tex
Remove numbers on equations that are not referenced in the text
Fix error in equation label in arXiv TEX
Some final edits for uploading to arXiv and preparing for resubmission to Scientific Reports

@@
 (\(\sigma^{2}_{\gamma} = 0\)), \(\boldsymbol{\gamma}\) is set to the
-identity matrix (diagonal elements all equal 1) and
+identity matrix (diagonal elements all equal 1), and
 \(\mathbf{M} = \mathbf{A}\). Under these conditions, eigenvalues of
@@
 
-\begin{equation} \labe{var_ref}
+\begin{equation} \label{var_ref}
 \sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.
@@
 
-\begin{equation} \label{var_reduced}
-\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\begin{equation}
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}). \nonumber
 \end{equation}

### 2020-03-28 notebook/ms.tex
Remove numbers on equations that are not referenced in the text
Swith to using the TEX file to prepare for resubmission

@@
 
-\begin{equation} \label{defM}
+\begin{equation} 
 \mathbf{M} = \boldsymbol{\gamma} \mathbf{A}.
@@
 
-Equation \ref{defM} thereby serves as a null model to investigate how
+Equation 1 thereby serves as a null model to investigate how
 variation in component response rate (\(\sigma^{2}_{\gamma}\)) affects
@@
 
-\begin{equation} \label{var_full}
+\begin{equation} 
 \sigma^{2}_{M} = \sigma^{2}_{A}\sigma^{2}_{\gamma} + \sigma^{2}_{A}E[\gamma_{i}]^{2}+\sigma^{2}_{\gamma}E[A_{ij}]^{2}.
@@
 
-Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), Eq. \ref{var_full}
+Given \(E[\gamma_{i}] = 1\) and \(E[A_{ij}] = 0\), Eq. 2
 can be simplified,
 
-\begin{equation} \label{var_reduced}
-\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}).
+\begin{equation} 
+\sigma^{2}_{M} = \sigma^{2}_{A}(1 + \sigma^{2}_{\gamma}). \nonumber
 \end{equation}
@@
 
-\includegraphics{unnamed-chunk-8-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-11-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-12-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
 
@@
 
-\includegraphics{unnamed-chunk-13-1.pdf}
+\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
 

### 2020-03-28 notebook/sci_rep_reviews.Rmd
Fix line numbers in response to reviewers

@@
 I again thank the reviewer for their comments.
-I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (e.g., L48, L88, L209, L234).
+I agree that it is especially relevant that my manuscript focuses on finite S values, and I have made more emphasis on this point in my revision (e.g., L48, L88, L207, L232).
 In particular, in addition to the statement on L48, I now state more explicitly in describing the methods that "the objective is to determine the effect of $\gamma$ as system complexity increases toward the boundary at which stability is realistic for a finite system." (L88)
@@
 I agree with the relevance of Dougoud et al.
-(2018), and have made appropriate links to their results on feasibility (L143, L274).
-Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility (L148).
+(2018), and have made appropriate links to their results on feasibility (L140, L273).
+Nevertheless, one major result of my manuscript is that varying component response rates (i.e., $\gamma$) only affect stability, not feasibility (L147).
 Hence, Reviewer 1 is incorrect that my conclusions concerning varying component response rates should be expected to enhance feasibility in experimentally observed food webs (and therefore that stability should be enhanced for biologically realistic ecological communities in which all population densities are positive).
-Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application (but, as I note in the Discussion L194, empirical food webs do not exist in a vacuum, and instead are parts of larger systems within which component response rate variation might indeed be relevant).
+Checking for validity in empirical food webs is therefore not entirely appropriate because the consequences of such an investigation would have negligible practical application (but, as I note in the Discussion L193, empirical food webs do not exist in a vacuum, and instead are parts of larger systems within which component response rate variation might indeed be relevant).
 
@@
 Overall, the structured networks that I now include in the main text and supporting information encompass a wide range of physical, biological, and social systems.
-Results for thse simulations are presented in a new section entitled, "Simulations of structured M across S" (L123).
-Methods for constructing small-world (L247), scale-free (L260), and cascade food web (L268) networks are now explained, with additional details and figures now in the Supplemental Information.
+Results for thse simulations are presented in a new section entitled, "Simulations of structured M across S" (L122).
+Methods for constructing small-world (L246), scale-free (L259), and cascade food web (L267) networks are now explained, with additional details and figures now in the Supplemental Information.
 The results from these structured networks match those of the random networks; variation in $\gamma$ increases the number of stable systems given sufficiently high system complexity, thereby further showing the generality of my results across complex systems.
 
-Feasibility and how it relates to ecological systems is explained on L141, with discussion starting on L188 for how this relates to ecological systems.
+Feasibility and how it relates to ecological systems is explained on L140, with discussion starting on L187 for how this relates to ecological systems.
 Full details of these results have also been added to the supporting information, and code for building structured networks is publicly available as a documented R package on GitHub (https://github.com/bradduthie/RandomMatrixStability) and accompanying website (https://bradduthie.github.io/RandomMatrixStability).
@@
 
-I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L420).
+I have explained the result shown by the black line of Figures 3 (now 4), and the former figure 4 (now in the SI), now more clearly in my revision (L419).
 
@@
 
-I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by \gamma (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the (co)variation of M (L65-L81).
+I appreciate these comments, and have tried to provide a more analytical explanation of the changes introduced by $\gamma$ (L53-L64), including a new subsection explaining the effect of variation in $\gamma$ on the (co)variation of M (L65-L81).
 I agree that these analytical results improve the manuscript, and I think that they are particularly useful in setting the stage for the numerical simulations that now follow.
@@
 These include food webs with cascade structures (e.g., Williams and Martinez 2000), small-world networks (Watts and Strogatz 1998), and scale-free networks (Albert and Barab\'{a}si 2002).
-Results for these networks are presented in L123-L140, and methods are explained in L247-L273, and in the Supplemental Information.
+Results for these networks are presented in L122-L139, and methods are explained in L246-L272, and in the Supplemental Information.
 Overall, these structured networks were affected by variation in $\gamma$ in the same way as random networks (increasing stability given sufficiently high complexity).
@@
 
-I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (L25, L69, L323, L325, and L367).
+I thank Reviewer 3 for pointing out these articles. I have cited them where appropriate (L25, L69, L322, L324, and L366).
 

### 2020-03-28 notebook/cover_letter.Rmd
New cover letter Rmd for revision to Scientific Reports

---
output:
  pdf_document: default
---

\fontsize{13}{22}
\selectfont

Dear Scientific Reports Editorial Board,

Please find attached my revision of the manuscript, "Component response rate variation underlies the stability of highly complex finite systems", which I hope you will further consider for publication as an Article.

My manuscript addresses a question of broad relevance to the physical, life, and social sciences. I demonstrate that when individual components of complex systems vary in the rate at which they respond to system perturbation, the potential for stability is increased.

In my revision, I have carefully considered the comments of three anonymous reviewers. Addressing these reviewer comments has taken longer than initially anticipated, but I believe that the quality of the manuscript has greatly improved for these additional efforts. I now include new analytical and simulation results that investigate the role of system correlation on stability. I also have performed new simulations for commonly observed network structures, including small-world networks, scale-free networks, and cascade food webs. These new results further clarify the key theory, and further demonstrate its generality across complex systems.

In the interest of transparency, the entire history of this project, including all my efforts in revising and addressing reviewer comments, is accessible on GitHub (https://github.com/bradduthie/RandomMatrixStability). An accompanying R package and website (https://bradduthie.github.io/RandomMatrixStability/) is also available with documentation for easily recreating all simulations and analyses.

I certify that this manuscript is original work and not under review at any other journal or book; a pre-print version of this revised manuscript has been submitted to arXiv (http://arxiv.org/abs/1806.01029).

Sincerely,

A. Bradley Duthie

University of Stirling, Stirling, UK

### 2020-03-28 notebook/SI.Rmd
Change the SI name so that it fits with Scientific Reports

@@
 csl: nature.csl
-subtitle: Supplemental Information
+subtitle: Supplementary Information
 biblio-style: apalike
@@
 
-**This supplemental information supports the manuscript "Component response rate variation underlies the stability of complex systems" with additional analyses to support its conclusions.
+**This supplemental information supports the manuscript "Component response rate variation underlies the stability of highly complex finite systems" with additional analyses to support its conclusions.
 All text, code, and data underlying this manuscript are publicly available on [GitHub](https://github.com/bradduthie/RandomMatrixStability) as part of the RandomMatrixStability R package.**

### 2020-03-28 notebook/SI.tex
Recompile SI files with slight adjustment based on arXiv edits
Change the SI name so that it fits with Scientific Reports

@@
   \posttitle{\par}
-\subtitle{Supplemental Information}
+\subtitle{Supplementary Information}
   \author{A. Bradley Duthie}
@@
 \textbf{This supplemental information supports the manuscript
-``Component response rate variation underlies the stability of complex
-systems'' with additional analyses to support its conclusions. All text,
-code, and data underlying this manuscript are publicly available on
+``Component response rate variation underlies the stability of highly
+complex finite systems'' with additional analyses to support its
+conclusions. All text, code, and data underlying this manuscript are
+publicly available on
 \href{https://github.com/bradduthie/RandomMatrixStability}{GitHub} as
@@
 Cascade food webs are more likely to be stable than small-world or
-scale-free networks at equivalent magnitudes of complexity (note that
+scale-free networks at equivalent magnitudes of complexity (note
 \(C = 1\) for all above rows). A higher number of stable \(\mathbf{M}\)
@@
 
-\clearpage
-
 \hypertarget{repr}{\section{Reproducing simulation results}\label{repr}}

### 2020-03-28 notebook/ms.tex
Add acknowledgement of anonymous reviewers
Minor edits to the manuscript to better incorporate changes for Scientific Reports

@@
 Career Fellowship (ECF-2016-376). Conversations with L. Bussière and N.
-Bunnefeld, and comments from J. J. Cusack and I. L. Jones, improved the
-quality of this work.
+Bunnefeld, and comments from J. J. Cusack, I. L. Jones, and several anonymous 
+reviewers improved the quality of this work.
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{fig1.pdf}
 
@@
 part of the leading eigenvalue is indicated with grey shading. In all
-simulations, system size and connectence were set to \(S = 25\) and
+simulations, system size and connectance were set to \(S = 25\) and
 \(C = 1\), respectively. Off-diagonal elements of \(\textbf{A}\) were
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-11-1.pdf}
+\includegraphics{fig2.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-12-1.pdf}
+\includegraphics{fig3.pdf}
 
@@
 
-\includegraphics{ms_files/figure-latex/unnamed-chunk-13-1.pdf}
+\includegraphics{fig4.pdf}
 

### 2020-03-28 notebook/sci_rep_reviews.Rmd
Add a line of text prior to getting into reviewer responses

@@
 author: "Brad Duthie"
-date: "30 December 2019"
 output:
@@
 ---
+
+**I am grateful for the helpful comments of reviewers on a previous version of this manuscript. I respond to these comments point by point below.**
 

### 2020-04-14 notebook/SI.Rmd
Add changes for minor revision for Sci Rep

@@
 ---
-title: "Component response rate variation underlies the stability of highly complex finite systems"
-author: "A. Bradley Duthie"
+title: "Component response rate variation underlies the stability of highly complex finite systems (Supplementary Information)"
+author: "A. Bradley Duthie (alexander.duthie@stir.ac.uk)"
 date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
-  FK9 4LA alexander.duthie@stir.ac.uk
+  FK9 4LA
 output:
@@
 - \usepackage{float}
-- \linenumbers
 - \bibliographystyle{amnatnat}
@@
 csl: nature.csl
-subtitle: Supplementary Information
 biblio-style: apalike

### 2020-04-14 notebook/SI.tex
Add changes for minor revision for Sci Rep

--- 2020-03-22 notebook/SI.tex
@@
 \IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\usepackage[]{microtype}
 \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
+\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
+\usepackage[unicode=true]{hyperref}
 \PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
-\hypersetup{unicode=true,
-            pdftitle={Component response rate variation underlies the stability of complex systems},
-            pdfauthor={A. Bradley Duthie},
+\hypersetup{
+            pdftitle={Component response rate variation underlies the stability of highly complex finite systems (Supplementary Information)},
+            pdfauthor={A. Bradley Duthie (alexander.duthie@stir.ac.uk)},
             colorlinks=true,
@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage[margin=1in]{geometry}
 \usepackage{color}
@@
 \usepackage{longtable,booktabs}
+% Fix footnotes in tables (requires footnote package)
+\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
 \usepackage{graphicx,grffile}
@@
 
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Component response rate variation underlies the stability of complex
-systems}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-  \subtitle{Supplemental Information}
-  \author{A. Bradley Duthie}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-      \predate{\centering\large\emph}
-  \postdate{\par}
-    \date{Biological and Environmental Sciences, University of Stirling, Stirling,
-UK, FK9 4LA
-\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
@@
 \usepackage{float}
-\linenumbers
 \bibliographystyle{amnatnat}
+
+\title{Component response rate variation underlies the stability of highly
+complex finite systems (Supplementary Information)}
+\author{A. Bradley Duthie
+(\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}})}
+\date{Biological and Environmental Sciences, University of Stirling, Stirling,
+UK, FK9 4LA}
 
@@
 
-\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 

### 2020-04-14 notebook/cover_letter.Rmd
Add changes for minor revision for Sci Rep

@@
 
-\fontsize{13}{22}
+\fontsize{15}{22}
 \selectfont
@@
 
-My manuscript addresses a question of broad relevance to the physical, life, and social sciences.
-I demonstrate that when individual components of complex systems vary in the rate at which they respond to system perturbation, the potential for stability is increased.
-
-In my revision, I have carefully considered the comments of three anonymous reviewers.
-Addressing these reviewer comments has taken longer than initially anticipated, but I believe that the quality of the manuscript has greatly improved for these additional efforts.
-I now include new analytical and simulation results that investigate the role of system correlation on stability.
-I also have performed new simulations for commonly observed network structures, including small-world networks, scale-free networks, and cascade food webs.
-These new results further clarify the key theory, and further demonstrate its generality across complex systems.
+My revision now addresses all remaining concerns from the previous round of review.
 

### 2020-04-14 notebook/ms.tex
Add changes for minor revision for Sci Rep

@@
 a synthesis of eco-evolutionary feedbacks on community stability, Patel
-et al. model a system that includes a vector of potentially changing
+et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}} model a system that includes a vector of potentially changing
 species densities (\(\mathbf{n}\)) and a vector of potentially evolving
 traits
-(\(\mathbf{x}\))\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
+(\(\mathbf{x}\)).
 For any species \(i\) or trait \(j\), change in species density
@@
 effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
+species \(i\) and the value of trait \(j\), respectively. Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}
 were interested in stability when the evolution of traits was relatively
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, and
+densities, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 
-I use the same principle that Patel et al. use to modulate the relative
+I use the same principle that Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}} use to modulate the relative
 rate of evolution to modulate rates of component responses for \(S\)
@@
 
-\textbf{Supplementary Information:} Full tables of stability results for
-simulations across different system size (\(S\)) values, ecological
-community types, connectance (\(C\)) values, interaction strengths
-(\(\sigma\)), and \(\gamma\) distributions are provided as supplementary
-material. An additional table also shows results for how feasibility
-changes across \(S\). All code and simulation outputs are publicly
-available as part of the RandomMatrixStability package on GitHub
-(\url{https://github.com/bradduthie/RandomMatrixStability}).
-
 \textbf{Additional Information:} The author declares no competing
@@
 
-\includegraphics{fig1.pdf}
-
-\clearpage
-
 \textbf{Figure 2: Complex system correlation versus stability with and
@@
 
-\includegraphics{fig2.pdf}
-
-\clearpage
-
 \textbf{Figure 3: System correlation versus stability across different
@@
 \(\gamma \sim \mathcal{U}(0, 2)\), so \(\sigma^{2}_{\gamma} = 1/3\).
-
-\includegraphics{fig3.pdf}
-
-\clearpage
 
@@
 
-\includegraphics{fig4.pdf}
-
 \end{document}

### 2020-04-14 notebook/sci_rep_reviewer_3.Rmd
Add changes for minor revision for Sci Rep

---
title: "Reviewer Comments (Scientific Reports)"
author: "Brad Duthie"
output:
  pdf_document: default
  html_document: default
---

Reviewer comments
================================================================================

Reviewer #3 (Technical Comments to the Author):
--------------------------------------------------------------------------------

The author has followed the reviewers' recommendations and, in my opinion, has greatly improved the manuscript. I recommend accepting to be published.

Author response to Reviewer #3 comment
--------------------------------------------------------------------------------

I thank Reviewer 3 for their kind words, and for reviewing this manuscript a second time.

Reviewer #3 minor comment
--------------------------------------------------------------------------------

Minor comment

in lines 217 and 226 the article has lost el citation of Patel et al.

Author response to Reviewer #3 minor comment
--------------------------------------------------------------------------------

I have re-inserted the citation on these lines.

### 2020-05-04 notebook/arxiv/ms.tex
Minor changes for arXiv that resulted from proofs

@@
 distributed within the blue circle centred at the origin with a radius
-of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as
+of \(\sigma_{A} \sqrt{S} = 1\). (\(\textbf{b}\)) The same system as in 
 \(\textbf{a}\) after including variation in the response rates of \(S\)
@@
 small-world\textsuperscript{\protect\hyperlink{ref-Watts1998}{16}},
-scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}},
-cascade food
+scale-free\textsuperscript{\protect\hyperlink{ref-Albert2002}{17}}, and cascade food
 web\textsuperscript{\protect\hyperlink{ref-Solow1998}{18},\protect\hyperlink{ref-Williams2000}{19}}
@@
 elements of \(\boldsymbol{\gamma}\) are positive, and off-diagonal
-elements are set to zero (i.e. \(\boldsymbol{\gamma}\) is a diagonal
+elements are set to zero (i.e., \(\boldsymbol{\gamma}\) is a diagonal
 matrix with positive support). The distribution of
@@
 density of some tangible quantity, it is relevant to consider the
-feasibility of the system. Feasibilility assumes that values of all
+feasibility of the system. Feasibility assumes that values of all
 components are positive at
@@
 effects of all species densities and trait values on the density of a
-species \(i\) and the value of trait \(j\), respectively. Patel et al.
+species \(i\) and the value of trait \(j\), respectively. 
+Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}
 were interested in stability when the evolution of traits was relatively
 slow or fast in comparison with the change in species
-densities\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}, and
+densities, and
 this is modulated in the above by the scalar \(\epsilon\). The value of
@@
 and evolution, with high \(\epsilon\) modelling relatively fast
-evolution and low \(\epsilon\) modelling relative slow
+evolution and low \(\epsilon\) modelling relatively slow
 evolution\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}}.
 
-I use the same principle that Patel et al. use to modulate the relative
+I use the same principle that Patel et al.\textsuperscript{\protect\hyperlink{ref-Patel2018}{20}} 
+use to modulate the relative
 rate of evolution to modulate rates of component responses for \(S\)
